### 2019-03-15 introduction.Rmd
skeleton of paper first work on introduction

---
title: "introduction"
author: "Jeremy Forbes"
date: "14/03/2019"
output: html_document
---

In the last two decades, political power in Australia has moved back and forth between the two major parties, the Australian Labor Party (Labor) and the Liberal National coalition (Liberal). over the same period Australia has also experienced rising house prices, improvements in education and an ageing population, amongst other changes to its national demographic. These simultaneous changes elicit questions about the relationships between socio-demographics and voting behaviour. Previous work in this area predominantly combine socio-demographic information from the nearest Census, and attach it to the election outcomes at an aggregate level. This poses problems due to temporal differences, and also a spatial mismatch - as the boundaries for aggregate areas (Federal electorates, for example) often change.

This paper outlines an approach to imputing Census information that accounts for both spatial and temporal differences. It then follows with an application of modelling the electoral two-party preferred vote over the past six Australian federal elections.

A key motivation in gathering and imputing this Australian data is to make it easily accessible via the `eechidna` R package, which is available on CRAN.

In the 2007 and 2010 federal elections, the Australian Labor Party (Labor) was victorious, whereas the 2001, 2004, 2013 and 2016 elections were won by the Liberal National coalition (Liberal). A measure of support between these two parties is the two-party preferred vote, which has been as high as 53.5% and as low as 47.3% in favour of the Liberal party over the six elections.

Few studies

The simultaneous changes in characteristics of the Australian population and preference between the two parties lead to our research question: *which socio-demographic characteristics are key determinants of electoral support between the Labor and Liberal parties, and how have their effects changed over time?*

Previous studies of Australian federal elections have shown that socio-demographic characteristics (simply, socio-demographics) are good predictors of party support, at an aggregate level. However, these studies have not ventured beyond the 2007 election. Adding to the existing research, this thesis explores the dynamics of electoral party preference for the six federal elections between 2001 and 2016, using a statistical framework. The two-party preferred vote is the outcome of interest, and is modelled as a function of electoral socio-demographics. An emphasis is placed on understanding how the effects of socio-demographics on voting preferences have changed over time.

The socio-demographic information is obtained from the Australian Census of Population and Housing (Census), a country-wide survey conducted every five years. Federal elections are generally conducted every three years. There are 150 electorates in Australia, with each corresponding to a single seat in the federal House of Representatives. It is electorates that ultimately determine the result of an Australian federal election[^2], which is why electorates are chosen as the level of aggregation for this study. Therefore, the information in this study consists of the electoral two-party preferred vote, and Census data aggregated to Commonwealth Electoral Divisions, which is an approximation of the electoral boundaries.

The two-party preferred vote and the socio-demographic data from the Census possess a temporal component and a spatial component, as both are observed at particular points in time and are aggregate measures for a set of regions in Australia. Censuses and elections often occur in different years, and electoral boundaries change frequently. Therefore, in order to match socio-demographic information from the Census with the electoral two-party preferred vote for each election, a method of spatio-temporal imputation is required. The method used in this study involves estimating an electorate's socio-demographic composition at the time of a Census, using an areal imputation approach, and then linearly interpolating between Census years to approximate the socio-demographics of that electorate, at the time of the election. To adjust for changing variable scales across elections, each socio-demographic variable is centered to have a mean of zero and standardized to have a variance of one, within each election. The resultant cross-sectional data set for each election consists of $n=150$ electorates (observations) and $p=63$ socio-demographic (explanatory) predictors.

Each election is modelled separately with a consistent model specification. This avoids imposing a time-invariant restriction on any regressor, and allows for socio-demographic effects to be compared across elections. In order to obtain the desired model specification, steps are taken to reduce the dimension of the data set, as a large number of predictors $p=63$ and relatively small sample size $n=150$ is problematic.

The first step in dimension reduction involves grouping similar socio-demographic variables together into *factors*. Principal component analysis of socio-demographic variables is employed to determine which variables should be grouped, by identifying variables that have large loadings in one the first four components. Factors are also centered and standardized. The result is a reduced set of $p=30$ predictors. From this reduced set, all possible five variable linear models are fit for each election, with the two-party preferred vote as the response. These models form a set for which Akaike weights are calculated, providing a measure relative variable importance for that election. Using this information-theoretic technique, the five most important variables are identified for each of the six elections. A superset is formed by taking the union of these important socio-demographics, which results in a set of ten variables.

The final specification consists of regressing the two-party preferred vote on the superset of ten predictors and four two-way interactions. Each election model is estimated separately using ordinary least squares. Insights are then found by visualizing the socio-demographic effects within each election. The estimated main effects are described using partial residual plots, and two-way interactions are illustrated with two-dimensional contour plots.

To guide the reader through each step of the analysis, this thesis is structured as follows. Chapter \@ref(ch:litreview) contains a literature review of Australian federal elections modelled using Census information. The data used in this study and the imputation process are described in Chapter \@ref(ch:Data). The steps involved in choosing a statistical model, reducing the dimension of the data, selection of the most important socio-demographic variables, and the final model specification, are explained in Chapter \@ref(ch:Modelling). Chapter \@ref(ch:Insights) contains the insights from the final estimated models. The conclusion and a discussion of possible extensions are detailed in Chapter \@ref(ch:Conclusion).

A key deliverable of this project is the contribution made to the existing `eechidna`[^3] `R` package. This includes the GIS maps, data from all Censuses and elections, and the imputed socio-demographic profiles of each electorate at the time of each election created during this project. When future elections and Censuses come around, the `eechidna` package and the work in this project will provide a resource for anyone wishing to conduct their own socio-political analyses.

[^1]: Socio-demographics are characteristics of a population, such as the distributions of age, gender, ethnicity, education level and income. 
[^2]: The party that wins support from a majority of seats in the House of Representatives forms the government.
[^3]: The name `eechidna` is an acronym for 'Exploring Election and Census Highly Informative Data Nationally for Australia'.

### 2019-03-15 introduction.Rmd
lit review edit

@@
 ---
+
+Australia has changed in many ways over the last two decades.
+Rising house prices, country-wide improvements in education, an ageing population, and a decline in religious affiliation, are just a few facets of the country's evolving socio-demographic characteristics[^1].
+At the same time, political power has moved back and forth between the two major parties.
+In the 2007 and 2010 federal elections, the Australian Labor Party (Labor) was victorious, whereas the 2001, 2004, 2013 and 2016 elections were won by the Liberal National coalition (Liberal).
+The two-party preferred vote, a measure of support between these two parties, fluctuated between 47.3% and 53.5% (in favour of the Liberal party) over this period.
+This study explores the question of what electoral characteristics relate to party preference, and whether their effects have changed over time.
+
+Previous work on modelling voting outcomes as a function of aggregate socio-demographics have found that relationships do exist, and relatively good models can be fit.
+@Forrest01 does this using multiple regression of the Liberal and Labor primary vote for polling booths in the Farrer electorate in 1998, @Stimson06, @Stimson09 and @Stimson12 conduct principal component analysis of polling booths in the 2001, 2004 and 2007 elections respectively, also finding that socio-demographic characteristics of polling booths are linked to their two-party preferred vote.
+@Stimson09 model the polling booth swing vote (change in the two-party preferred vote) in the 2007 election, which has a poor fit.
+
+Previous work on modelling voting outcomes as a function of aggregate socio-demographics is focused on the aggregate level of polling booths - a finer level of aggregation than electorates.
+@Forrest01 use multiple regression to find that polling booth socio-demographic covariates explained a lot of variation in the primary vote for both Liberal and Labor parties in the Farrer electorate during the 1998 election.
+@Stimson09 also apply a multiple regression framework for the swing vote (change in two-party preferred vote) in the 2007 election, but resultant model has a relatively poor fit.
+Principal component analysis has also been used in @Stimson06, @Stimson09 and @Stimson12 to examine the 2001, 2004 and 2007 elections, respectively.
+
+, which find that the Labor party was found to appeal to lower income, younger and more multicultural areas, whereas the Liberal party was associated with higher income, older and more mono-cultural areas in the 2001, 2004 and 2007 elections.
+
+Joining aggregate socio-demographics information from the Census with federal election voting data is problematic, as there in an inherent asynchronicity in the two events - a Census occurs every five years, whereas an election usually occurs every three years.
+The predominant approach in previous studies is to join voting outcomes to the nearest Census, without accounting for any temporal differences (see @DavisStimson98, @Stimson06, @Liao09 and @Stimson09).
+Furthermore, electoral boundaries change regularly, so spatial discrepencies also arise when matching electoral data.
+To combat
+
+Previous studies of aggregate socio-demographics characteristics and voting outcomes commonly use a regression framework and typically use polling booths - rather than electorates - as the level of aggregation, as in @DavisStimson98, @Forrest01, @Stimson06, @Liao09 and @Stimson09.
+Those that use regression to identify significant determinants of party support find that
+
+@Forrest01 use multiple regression to identify significant determinants of primary vote support in the 1998 Farrer electorate, finding that socio-demographic covariates explained a lot of variation for  Liberal and Labor support.
+A similar exercise
+
+@Forrest01 use multiple regression to identify significant determinants of the primary vote for each party in the Farrer electorate during the 1998 election
+
+@Stimson09 model the swing vote in the 2007
+
+To obtain data aggregate socio-demographics characteristics are obtained from the Australian Census of Population and Housing and voting data from the Australian Electoral Commission.
+However, almost all
+
+and voting outcomes involves joining socio-demographic information from the Australian Census of Population and Housing (Census) with voting data from the Australian Electoral Commission.
+The predominant approach to use polling booth level data,
+
+--------
+
+Few existing studies have conducted an analysis of Australian federal elections using socio-demographic information.
+Amongst these, the predominant approach is to merge data from the most recent Census that occurs prior to the election of interest.
+This is done using GIS data obtained at the collection district (Census) and polling booth (election) aggregate levels, as in @DavisStimson98, @Forrest01, @Stimson06, @Liao09 and @Stimson09.
+Collection districts are matched to the nearest polling booth.
+Our study is focused on electorates, which reflects a greater level of aggregation than polling booths.
+Therefore, we adopt a different method of matching Census information with electoral voting outcomes, based on the piece-wise approximation method of areal imputation in @Goodchild1993.
+
+From our revivew of the literature, a regression framework using socio-demographics to understand Australian election outcomes has been applied on only two occasions.
+@Forrest01 use multiple regression to identify significant determinants of the primary vote for each party in the Farrer electorate during the 1998 election, by modelling each party's share of the primary vote as a separate regression, each using ordinary least squares.
+The adjusted R-squared values obtained from these regressions range from 0.158 to 0.785, suggesting that some parties (including the Labor and Liberal parties) are more predictable than others.
+They found that age, education and migrant populations were important in relation to support for both Labor and Liberal.
+@Stimson09 also apply a multiple regression framework to party preference, by modelling the swing vote in the 2007 election.
+The resultant model has relative poor fit (adjusted R-squared of 0.223), and as such no conclusions about socio-demographic effects were made.
+Both of these studies use a stepwise procedure to reduce the variable set, and had limited success in providing rich insights into socio-demographic preferences.
+
+Applying a stepwise procedure to a data set with a low number of observations per predictor is problematic [@Steyerberg2009].
+In our study, each election is modelled as a separate cross-section containing 63 socio-demographic predictors with only 150 observations.
+Therefore, instead of stepwise selection, we adopt an information-theoretic approach that uses Akaike weights to determine relative variable importance [@BurnhamAnderson2002], and use this as a criterion for variable selection.
+Akaike weights are often used in ecological studies (see @Johnson2004 and @Symonds2011).
+
+Another method to differentiate between party preference is to use principal component analysis on the socio-demographic variables, as shown in @Stimson06, @Stimson09 and @Stimson12.
+Electorates are assigned to a particular party based on their primary vote, and each party is then summarized by its mean z-scores in the directions of the first two principal components.
+For the 2001, 2004 and 2007 elections, the Labor party was found to appeal to lower income, younger and more multicultural areas, whereas the Liberal party was associated with higher income, older and more mono-cultural areas [@Stimson12].
+Whilst we do not assign parties to electorates in this way, we do use principal components to understand and group similar predictor variables together.
+
+Other election-related studies that aimed to understand party preference predominantly modelled individual voter behaviour using survey data (see @Schram92, @Wilson2011 and @Gibson2011).
+
+Seemingly absent from previous Australian studies is the use of a regression framework to examine these socio-political relationships over time.
+It also appears that no study has attempted any type of statistical analysis of socio-demographics in conjunction with voter behaviour in Australia since 2007.
+In this thesis, all six elections thus far held in the 21st century are examined.
+
+--------
+
+To understand
+
+Previous studies of Australian federal elections have shown that socio-demographic characteristics (simply, socio-demographics) are good predictors of party support, at an aggregate level.
+However, these studies have not ventured beyond the 2007 election.
+Adding to the existing research, this thesis explores the dynamics of electoral party preference for the six federal elections between 2001 and 2016, using a statistical framework.
+The two-party preferred vote is the outcome of interest, and is modelled as a function of electoral socio-demographics.
+An emphasis is placed on understanding how the effects of socio-demographics on voting preferences have changed over time.
 

### 2019-03-15 introduction.Rmd
structured: motivation, problem, previous work

@@
 The two-party preferred vote, a measure of support between these two parties, fluctuated between 47.3% and 53.5% (in favour of the Liberal party) over this period.
-This study explores the question of what electoral characteristics relate to party preference, and whether their effects have changed over time.
+This study explores the question of how electoral characteristics relate to party preference, and whether their effects have changed over time.
 
-Previous work on modelling voting outcomes as a function of aggregate socio-demographics have found that relationships do exist, and relatively good models can be fit.
-@Forrest01 does this using multiple regression of the Liberal and Labor primary vote for polling booths in the Farrer electorate in 1998, @Stimson06, @Stimson09 and @Stimson12 conduct principal component analysis of polling booths in the 2001, 2004 and 2007 elections respectively, also finding that socio-demographic characteristics of polling booths are linked to their two-party preferred vote.
-@Stimson09 model the polling booth swing vote (change in the two-party preferred vote) in the 2007 election, which has a poor fit.
+Joining aggregate socio-demographics information from the Census with federal election voting data is problematic, as there in an inherent asynchronicity in the two events - a Census occurs every five years, whereas an election usually occurs every three years.
+The predominant approach in previous studies is to join voting outcomes to the nearest Census, without accounting for any temporal differences (see @DavisStimson98, @Stimson06, @Liao09 and @Stimson09).
+Furthermore, electoral boundaries change regularly, so spatial discrepencies also arise when matching electoral data.
+To obtain appropriate Census data for the 2004, 2007, 2010 and 2013 elections, we adopt impute electoral socio-demographics using spatio-temporal imputation that combines areal interpolation [@Goodchild1993] and linear time-interpolation.
+The process of collection and wrangling is detailed in @CHAPTER2 and imputation process is outlined in @CHAPTER3.
+
+Previous work on modelling voting outcomes with aggregate socio-demographics find that relatively good models can be fit.
+@Forrest01 does this using multiple regression of the Liberal and Labor primary vote for polling booths in the Farrer electorate in 1998.
+@Stimson06, @Stimson09 and @Stimson12 conduct principal component analysis of polling booths in the 2001, 2004 and 2007 elections respectively, also finding that socio-demographic characteristics of polling booths are linked to their two-party preferred vote.
+@Stimson09 also model the polling booth swing vote (change in the two-party preferred vote) in the 2007 election, but had little success.
+Instead of analyzing a single election in isolation, this paper employs a consistent model framework across elections so that any temporal changes are observed.
 

### 2019-03-15 introduction.Rmd
added chunk of modelling explanation to intro. needs to be edited.

@@
 The process of collection and wrangling is detailed in @CHAPTER2 and imputation process is outlined in @CHAPTER3.
+All data and documentation involved in this study is reproducible and availabe in the `eechidna` `R` package on `CRAN`.
 
@@
 Instead of analyzing a single election in isolation, this paper employs a consistent model framework across elections so that any temporal changes are observed.
+This process is discussed in @CHAPTER4 and involves standardizing variables across years, using principal components to group variables, and then selecting variables with an information-theoretic approach that uses Akaike weights to determine relative variable importance [@BurnhamAnderson2002].
+Spatial error models are then fit to account for spatial dependence.
+
+---------
 

### 2019-03-18 introduction.Rmd
proof intro 1

--- 2019-03-15 introduction.Rmd
@@
 The two-party preferred vote, a measure of support between these two parties, fluctuated between 47.3% and 53.5% (in favour of the Liberal party) over this period.
-This study explores the question of how electoral characteristics relate to party preference, and whether their effects have changed over time.
+This study explores how electoral characteristics relate to two-party preference, and whether their effects have changed over time.
+Information about electoral socio-demographics are from the Census, and vote counts from the federal election.
 
-Joining aggregate socio-demographics information from the Census with federal election voting data is problematic, as there in an inherent asynchronicity in the two events - a Census occurs every five years, whereas an election usually occurs every three years.
+Joining these two data sources is problematic, as there in an inherent asynchronicity in the two events - a Census occurs every five years, whereas an election usually occurs every three years.
 The predominant approach in previous studies is to join voting outcomes to the nearest Census, without accounting for any temporal differences (see @DavisStimson98, @Stimson06, @Liao09 and @Stimson09).
-Furthermore, electoral boundaries change regularly, so spatial discrepencies also arise when matching electoral data.
+Furthermore, electoral boundaries change regularly, so spatial discrepancies also arise when matching electoral data.
 To obtain appropriate Census data for the 2004, 2007, 2010 and 2013 elections, we adopt impute electoral socio-demographics using spatio-temporal imputation that combines areal interpolation [@Goodchild1993] and linear time-interpolation.
 The process of collection and wrangling is detailed in @CHAPTER2 and imputation process is outlined in @CHAPTER3.
+All data and documentation relating to this procedure is available in the `eechidna` `R` package on `CRAN`.
 
@@
 @Stimson09 also model the polling booth swing vote (change in the two-party preferred vote) in the 2007 election, but had little success.
-Instead of analyzing a single election in isolation, this paper employs a consistent model framework across elections so that any temporal changes are observed.
+Instead of analyzing a single election in isolation, this paper employs a consistent model framework across elections so that any temporal changes are observed, where each of the six federal elections are modelled with a cross-sectional data set.
+This consists of the two-party preferred vote, which is the response variable, and socio-demographic variables which are explanatory variables that characterise each electorate.
+This process is discussed in @CHAPTER4 and involves standardizing variables across years, using principal components to group variables, and then selecting the most influential variables with an information-theoretic approach via Akaike weights to determine relative variable importance [@BurnhamAnderson2002].
+Spatial error models are then fit to account for spatial dependence.
+
+---------
 
@@
 The predominant approach in previous studies is to join voting outcomes to the nearest Census, without accounting for any temporal differences (see @DavisStimson98, @Stimson06, @Liao09 and @Stimson09).
-Furthermore, electoral boundaries change regularly, so spatial discrepencies also arise when matching electoral data.
+Furthermore, electoral boundaries change regularly, so spatial discrepancies also arise when matching electoral data.
 To combat
@@
 
-From our revivew of the literature, a regression framework using socio-demographics to understand Australian election outcomes has been applied on only two occasions.
+From our revive of the literature, a regression framework using socio-demographics to understand Australian election outcomes has been applied on only two occasions.
 @Forrest01 use multiple regression to identify significant determinants of the primary vote for each party in the Farrer electorate during the 1998 election, by modelling each party's share of the primary vote as a separate regression, each using ordinary least squares.
@@
 The resultant model has relative poor fit (adjusted R-squared of 0.223), and as such no conclusions about socio-demographic effects were made.
-Both of these studies use a stepwise procedure to reduce the variable set, and had limited success in providing rich insights into socio-demographic preferences.
+Both of these studies use a step-wise procedure to reduce the variable set, and had limited success in providing rich insights into socio-demographic preferences.
 
-Applying a stepwise procedure to a data set with a low number of observations per predictor is problematic [@Steyerberg2009].
+Applying a step-wise procedure to a data set with a low number of observations per predictor is problematic [@Steyerberg2009].
 In our study, each election is modelled as a separate cross-section containing 63 socio-demographic predictors with only 150 observations.
-Therefore, instead of stepwise selection, we adopt an information-theoretic approach that uses Akaike weights to determine relative variable importance [@BurnhamAnderson2002], and use this as a criterion for variable selection.
+Therefore, instead of step-wise selection, we adopt an information-theoretic approach that uses Akaike weights to determine relative variable importance [@BurnhamAnderson2002], and use this as a criterion for variable selection.
 Akaike weights are often used in ecological studies (see @Johnson2004 and @Symonds2011).

### 2019-03-18 data-collection.Rmd
data collection - initial commit

---
title: "data-collection"
author: "Jeremy Forbes"
date: "18/03/2019"
output: html_document
---

1. Where the data is sourced from

2. Converting the Census raw counts into variables
- Using the same variable definitions as in @Stimson09 
- Tidyverse tools
- Lack of coherence in ABS structures

--------------------------------------------------------------------------------

Data description and collection

--------------------------------------------------------------------------------

## Voting data

The voting outcome of interest is the electoral two-party preferred vote, which is provided by the Australian Electoral Commission (AEC) for the 2001, 2004, 2007, 2010, 2013 and 2016 elections. The AEC divide Australia into 150 regions called electorates, with each corresponding to a single seat in the House of Representatives. Voting is compulsory in Australia, and each voter assigns a numbered preference to each available candidate in their electorate. The two-party preferred vote is determined by a tally of these preferences where, by convention, only the ranks of the Labor and Liberal candidates are considered. This is recorded as a percentage preference in favour of the Liberal party.

## Census data

Socio-demographic variables are derived from the Census of Population and Housing (Census), which is a survey of every household in Australia, recording information such as age, gender, ethnicity, education level and income. There have been four Censuses in the 21st century, being that in 2001, 2006, 2011 and 2016. The Australian Bureau of Statistics (ABS) conducts the Census and publishes aggregated information. The ABS approximation of electorates at the time of the Census is chosen. From this aggregate information, *63* socio-demographic variables are computed for each of the electorates. These variables follow the definitions used in @Stimson09.

## Downloading and wrangling

Raw data is sourced online from the .csv and .xlsx files from the AEC and ABS websites. The format of these files change over the years, meaning that significant time and effort is required to extract the appropriate electoral information. All data wrangling is done using `R`, predominantly using the tools from the `dplyr` and `readxl` packages, providing fast consistent tools for data manipulation and functions to import .xlsx files (respectively). The 2001 and 2006 Census data is formatted such that each electorate has a separate document, making it difficult to use the `dplyr` tools. Instead, cells had to be manually selected from each individual files. This work is entirely reproducible, and all scripts can be found in the github repository for the `eechidna` `R` package. These scripts also provide a resource to help wrangle data for future Censuses and elections.

### 2019-03-18 imputation.Rmd
imputation - added methodology, beginning applied

---
title: "imputation"
author: "Jeremy Forbes"
date: "18/03/2019"
output: html_document
---

## Imputation
Methodology
- Not specific to electorates
- Explain spatial and then temporal

Applied to electorates
- For our models we use electorates
- Visualisations of an example using electorates
- Possible to use finer levels of disaggregation (e.g. SA1). Propose a cascading technique for reduced computation burden (SA4, SA3, SA2, SA1).

-------------------------------------------------------------------------------------

## Differences between Census and electoral data

Between 2001 and 2016 there were six elections and four Censuses (see Figure \@ref(fig:timeline)). Only in 2001 and 2016 did both events occur, and electoral boundaries are also redistributed regularly. Therefore, there are both temporal and spatial differences that need to be accounted for when joining the electoral two-party preferred vote from the four elections between 2004 and 2013 with Census data. For these elections a spatio-temporal imputation method is employed to derive a electoral socio-demographics which uses Census information from both before and after the election of interest.

```{r timeline, fig.cap="Timeline of Australian elections and Censuses. They do not always occur in the same year.", out.width = "90%", fig.align="center"}
library(tidyverse)
timeline <- data.frame(year = c(2001:2016)) %>%
  mutate(Election = ifelse(year %in% c(2001, 2004, 2007, 2010, 2013, 2016),1,0),
         Census = ifelse(year %in% c(2001, 2006, 2011, 2016), 1.4, 0)) %>%
  gather(c(Election,Census), key = event, value = measure) %>%  filter(measure != 0) %>%
  mutate(year_vis = ifelse(event == "Census", year + 0.1, year))

ggplot(timeline, aes(x=year_vis, y=measure)) + geom_point(aes(color=as.factor(event),shape=as.factor(event)),size = 4) + labs(color = "Event", shape = "Event") + geom_hline(yintercept=0, size =1, color = 'grey', show.legend = FALSE) + geom_segment(x=2001,xend=2001, y=1, yend=0, color = '#00BFC4') + geom_segment(x=2004,xend=2004, y=1, yend=0, color = '#00BFC4') + geom_segment(x=2007,xend=2007, y=1, yend=0, color = '#00BFC4') + geom_segment(x=2010,xend=2010, y=1, yend=0, color = '#00BFC4') + geom_segment(x=2013,xend=2013, y=1, yend=0, color = '#00BFC4') + geom_segment(x=2016,xend=2016, y=1, yend=0, color = '#00BFC4') + geom_segment(x=2001.1,xend=2001.1, y=1.4, yend=0, color = '#F8766D') + geom_segment(x=2006.1,xend=2006.1, y=1.4, yend=0, color = '#F8766D') + geom_segment(x=2011.1,xend=2011.1, y=1.4, yend=0, color = '#F8766D') + geom_segment(x=2016.1,xend=2016.1, y=1.4, yend=0, color = '#F8766D') + labs(x="Year") + theme(text=element_text(size=10), axis.text.x=element_text(size=6),axis.text.y=element_blank(), axis.title.y=element_blank(), plot.background=element_blank(), axis.ticks.y = element_blank(), panel.background = element_blank()) + scale_x_discrete(limits = c(2001:2016)) + coord_fixed(ratio = 2)
```

## Spatio-temporal imputation

To account for spatial differences, the piecewise approximation method in @Goodchild1993 is appropriate for our problem. Consider a map of source zones $s = 1,...,S$, for which socio-demographic information is available, and a set of target zones $t = 1,...,T$ for which information is to be imputed. In this context the map of electoral boundaries at the time of a Census would be the source zones, and the boundaries at the time of the election would be the target zones. Denote the area of intersection between source zone $s$ and target zone $t$ as $A_{s,t}$, the population of the source zone $s$ as $U_s$, and the population of intersection between source zone $s$ and target zone $t$ as $P_{s,t}$.

Using tools from the `rgeos` and `rgdal` `R` packages, compute each $A_{s,t}$ and estimate population of the intersection:

$$\hat{P}_{s,t} = U_s*A_{s,t}$$

In order to calculate socio-demographic information for each of the target zones, a weighted average is taken using the estimated population as weights. Denote a given Census variable for the target zone $C_t$, and the same Census variable for the source zone $D_s$:

$$\hat{C}_t = \frac{\sum_{s=1}^{S}{D_s*\hat{P}_{s,t}}}{\sum_{s=1}^{S}{\hat{P}_{s,t}}}$$

Socio-demographic information is now imputed for the target zones, but these are not appropriate for the target year. Denote year $y$, with a Census falling on $y_1$ and $y_3$, and an election on year $y_2$, and add this subscript to the Census variable estimate, $\hat{C}_{t,y}$. To account for temporal changes, linear interpolation is used between Census years to get the final estimate of a Census variable for the target zone in the election year $y_2$.

$$\hat{C}_{t,y_2} = \frac{y_3-y_2}{y_3-y_1}*\hat{C}_{t,y_1} + \frac{y_2-y_1}{y_3-y_1}*\hat{C}_{t,y_3}$$

## Applied

Publically available Census data is aggregated and there are different resolutions accessible, ranging from SA1 (over 50,000 zones) to electoral divisions (150 zones). Any of these resolutions could be used as source zones. For this study, electoral divisions are used and this imputation method is applied to each of the 2004, 2007, 2010 and 2013 elections. To demonstrate its functionality, consider the example of 2013, where the nearest Censuses were in 2011 and 2016.

### 2019-03-18 imputation.Rmd
added example of Hume 2013

@@
 - For our models we use electorates
-- Visualisations of an example using electorates
+- Visualizations of an example using electorates
 - Possible to use finer levels of disaggregation (e.g. SA1). Propose a cascading technique for reduced computation burden (SA4, SA3, SA2, SA1).
@@
 
-```{r timeline, fig.cap="Timeline of Australian elections and Censuses. They do not always occur in the same year.", out.width = "90%", fig.align="center"}
+```{r timeline, messages = F, echo = F, fig.cap="Timeline of Australian elections and Censuses. They do not always occur in the same year.", out.width = "90%", fig.align="center"}
 library(tidyverse)
@@
 
-To account for spatial differences, the piecewise approximation method in @Goodchild1993 is appropriate for our problem.
+To account for spatial differences, the piece-wise approximation method in @Goodchild1993 is appropriate for our problem.
 Consider a map of source zones $s = 1,...,S$, for which socio-demographic information is available, and a set of target zones $t = 1,...,T$ for which information is to be imputed.
@@
 
-Using tools from the `rgeos` and `rgdal` `R` packages, compute each $A_{s,t}$ and estimate population of the intersection:
+Compute each $A_{s,t}$ and estimate population of the intersection:
 
-$$\hat{P}_{s,t} = U_s*A_{s,t}$$
+$$\hat{P}_{s,t} = \frac{U_s*A_{s,t}}{\sum_{t=1}^T A_{s,t}}$$
+This assumes that populations are uniformly distributed within each source zone.
 
@@
 $$\hat{C}_t = \frac{\sum_{s=1}^{S}{D_s*\hat{P}_{s,t}}}{\sum_{s=1}^{S}{\hat{P}_{s,t}}}$$
+This assumes that each individual in the source zones assume the aggregate characteristics of the zone.
 
@@
 To account for temporal changes, linear interpolation is used between Census years to get the final estimate of a Census variable for the target zone in the election year $y_2$.
+This assumes that population evolves in a linear manner over time.
 
@@
 For this study, electoral divisions are used and this imputation method is applied to each of the 2004, 2007, 2010 and 2013 elections.
-To demonstrate its functionality, consider the example of 2013, where the nearest Censuses were in 2011 and 2016.
+To demonstrate its functionality, consider the imputation of the socio-demographic variable $AusCitizen$ for the electorate of Hume in New South Wales (NSW), at the time of the 2013 federal election.
+Figure \@ref(fig:hume13) shows this region amongst other NSW electorates.
+
+```{r hume13, fig.cap = "Some of the electoral boundaries in NSW for 2013, with the electoral boundary for Hume, shown in purple."}
+library(eechidna)
+library(ggplot2)
+library(ggthemes)
+hume_area13 <- nat_map13 %>% 
+  filter(state %in% c("ACT","NSW"), long < 154)
+
+ggplot(data=hume_area13) +
+  geom_polygon(aes(x=long, y=lat, group=group, fill = elect_div == "HUME"),
+               colour="grey50", alpha = 0.4) +
+  scale_fill_manual(name="Boundary", values=c("white", "purple"), labels = c("Other 2013 Electorates", "Hume 2013")) +
+  theme_map() + coord_equal()
+```
+
+The Censuses neighbouring the 2013 election are those in 2011 and 2016, and the Hume boundary is changed, as seen by plotting the Hume boundary (purple) in the 2013 election over the divisions in 2016.
+
+```{r hume16, fig.cap = "Census division boundaries in NSW for 2016, with the 2013 electoral boundary for Hume, shown in purple. The purple region is not contained within a single Census division."}
+hume_area16 <- nat_map16 %>% 
+  filter(state %in% c("ACT","NSW"), long < 154) %>% 
+  mutate(year = "2016") %>% 
+  bind_rows(hume_area13 %>% filter(elect_div == "HUME") %>% mutate(year = "2013"))
+
+ggplot(data=hume_area16) +
+  geom_polygon(aes(x=long, y=lat, group=group, fill = year == "2013", 
+    alpha = year == "2013", colour=year == "2013")) +
+  scale_fill_manual(name="Boundary", values=c("grey95", "purple"), labels = c("2016 Electorates", "Hume 2013")) +
+  scale_alpha_manual(values=c(0, 0.4)) +
+  scale_color_manual(values=c("grey50", NA)) +
+  theme_map() + coord_equal() + guides(alpha = F, color = F)
+```
+
+There are many electorates in 2016 that intersect with the purple region (Hume boundary for 2013), these include the divisions of Riverina, Eden-Monaro and Hume, along with smaller intersecting areas with Fenner, Calare, Gilmore and Whitlam.
+To impute Census information for this purple region, calculate the percentage of each 2016 electorate that intersects with the purple region, which is then used to estimate intersection populations $\hat{P}_{s,t} = \frac{U_s*A_{s,t}}{\sum_{t=1}^T A_{s,t}}$.
+
+Electorate (2016) | Percentage | Population in Electorate | Estimated Population Allocated to Purple Region [$\hat{P}_{s,t}$]
+------------------|------------|--------------------------|------------------------------------------------
+Hume              | 96.54%     | 150643                   | 145427
+Riverina          | 25.11%     | 155793                   | 39117
+Eden-Monaro       | 11.09%     | 147532                   | 16358
+Canberra          | 0.28%      | 196037                   | 548
+Fenner            | 0.23%      | 202955                   | 474
+Whitlam           | 0.06%      | 152280                   | 92
+Gilmore           | 0.06%      | 150436                   | 86
+Calare            | 0.01%      | 161298                   | 21
+
+The socio-demographic of interest is $AusCitizen$, which is the proportion of people in the region who are Australian citizens.
+
+```{r eval = F, include = F}
+library(knitr)
+auscitizen_ex <- abs2016 %>% 
+  filter(DivisionNm %in% c("HUME", "RIVERINA", "EDEN-MONARO", "CANBERRA", "FENNER", "WHITLAM", "GILMORE", "CALARE")) %>%
+  select(DivisionNm, AusCitizen) %>% 
+  mutate(`Estimated Population Allocated to Purple Region` = c(21, 548, 16358, 474, 86, 145427, 39117, 92),
+    AusCitizen = round(AusCitizen, 2)) %>% 
+  arrange(-`Estimated Population Allocated to Purple Region`)
+
+auscitizen_ex %>% 
+  kable()
+
+weighted.mean(x = auscitizen_ex$AusCitizen, w = auscitizen_ex$`Estimated Population Allocated to Purple Region`)
+```
+
+|DivisionNm  | AusCitizen (%): $D_s$| Estimated Population Allocated to Purple Region: $\hat{P}_{s,t}$|
+|------------|---------------|------------------------------------------------|
+|HUME        |          90.02|                                          145427|
+|RIVERINA    |          89.11|                                           39117|
+|EDEN-MONARO |          88.00|                                           16358|
+|CANBERRA    |          85.48|                                             548|
+|FENNER      |          83.64|                                             474|
+|WHITLAM     |          89.52|                                              92|
+|GILMORE     |          89.03|                                              86|
+|CALARE      |          87.56|                                              21|
+
+Then taking a weighted average of $AusCitizen$ using the estimated population as weights yields $\hat{C}_{Hume,2016} = 89.65 \%$.
+Repeating this process using the 2011 Census and electoral boundaries yields $\hat{C}_{Hume,2011} = 91.00 \%$
+
+Finally, linearly interpolate between 2011 and 2016 to arrive at the 2013 estimate:
+\begin{eqnarray*}
+\hat{C}_{Hume,2013} & = &\frac{3}{5} \cdot \hat{C}_{Hume,2011} + \frac{2}{5} \cdot \hat{C}_{Hume,2016} \\ 
+& = & \frac{3}{5} \cdot 91.00 \% + \frac{2}{5} \cdot 89.65 \% \\ 
+& = & 90.46 \%
+\end{eqnarray*}
+
+This is done for each of the socio-demographic variables, and repeated each of the 2013 electorates.

### 2019-03-18 imputation.Rmd
edit tables

@@
 date: "18/03/2019"
-output: html_document
+output:
+  pdf_document: default
+  html_document: default
 ---
 
-## Imputation
-Methodology
-- Not specific to electorates
-- Explain spatial and then temporal
-
-Applied to electorates
-- For our models we use electorates
-- Visualizations of an example using electorates
-- Possible to use finer levels of disaggregation (e.g. SA1). Propose a cascading technique for reduced computation burden (SA4, SA3, SA2, SA1).
-
--------------------------------------------------------------------------------------
+```{r setup, include=FALSE}
+knitr::opts_chunk$set(echo = F)
+library(tidyverse)
+library(eechidna)
+library(ggplot2)
+library(ggthemes)
+library(knitr)
+```
 
@@
 
-```{r timeline, messages = F, echo = F, fig.cap="Timeline of Australian elections and Censuses. They do not always occur in the same year.", out.width = "90%", fig.align="center"}
-library(tidyverse)
+```{r timeline, message = F, echo = F, fig.cap="Timeline of Australian elections and Censuses. They do not always occur in the same year.", out.width = "90%", fig.align="center"}
 timeline <- data.frame(year = c(2001:2016)) %>%
@@
 ```{r hume13, fig.cap = "Some of the electoral boundaries in NSW for 2013, with the electoral boundary for Hume, shown in purple."}
-library(eechidna)
-library(ggplot2)
-library(ggthemes)
 hume_area13 <- nat_map13 %>% 
@@
 
-Electorate (2016) | Percentage | Population in Electorate | Estimated Population Allocated to Purple Region [$\hat{P}_{s,t}$]
-------------------|------------|--------------------------|------------------------------------------------
-Hume              | 96.54%     | 150643                   | 145427
-Riverina          | 25.11%     | 155793                   | 39117
-Eden-Monaro       | 11.09%     | 147532                   | 16358
-Canberra          | 0.28%      | 196037                   | 548
-Fenner            | 0.23%      | 202955                   | 474
-Whitlam           | 0.06%      | 152280                   | 92
-Gilmore           | 0.06%      | 150436                   | 86
-Calare            | 0.01%      | 161298                   | 21
+Electorate (2016) | Percentage | Population in Electorate | Estimated Population Allocated to Purple Region: $\hat{P}_{s,t}$
+---------------------|------------|--------------------------|------------------------------------------------
+HUME              | 96.54%     | 150643                   | 145427
+RIVERINA          | 25.11%     | 155793                   | 39117
+EDEN-MONARO       | 11.09%     | 147532                   | 16358
+CANBERRA          | 0.28%      | 196037                   | 548
+FENNER            | 0.23%      | 202955                   | 474
+WHITLAM           | 0.06%      | 152280                   | 92
+GILMORE           | 0.06%      | 150436                   | 86
+CALARE            | 0.01%      | 161298                   | 21
 
@@
 ```{r eval = F, include = F}
-library(knitr)
 auscitizen_ex <- abs2016 %>% 
@@
 |DivisionNm      | AusCitizen (%): $D_s$| Estimated Population Allocated to Purple Region: $\hat{P}_{s,t}$|
-|------------|---------------|------------------------------------------------|
+|----------------|----------------------|------------------------------------------------|
 |HUME            |                 90.02|                                          145427|

### 2019-03-18 modelling.Rmd
start modelling section

---
title: "modelling"
author: "Jeremy Forbes"
date: "18/03/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Modelling
- Join the data, treat at separate cross-sections
- Address the issue of changing variable scales (rent as example)
- 
- PCA to group variables together
- Akaike weights to derive a variable set for models
- Spatial correlation in response, so model as spatial error model

## Modelling

With socio-demographic information now available for each electorate, each election is joined to the data  corresponding with its two-party preferred vote. Socio-demographic variables within each election year are standardized to have mean zero and variance one, to adjust for changing variable scales. For example, inflation-adjusted median rental prices increased across almost all electorates, with median rent of $200 per week placing an electorate in the 90th percentile in 2001, but only the 30th percentile in 2016.

An identical model specification is used across the six elections, with each election modelled separately. This enables the interpretation of temporal changes of the socio-demographic effects, as the same covariates are used in each election. The response, the two-party preferred vote, lies in the interval $(0,100)$, with $TPP = 70$ representing a 70% preference for Liberal, 30% for Labor. Since the observed values of $TPP$ are never very close to 0 or 100 (minimum $24.05 \%$ and maximum $74.90 \%$), there is no need to impose the constraint of $TPP \in [0,100]$. Furthermore, significant spatial autocorrelation in the response is found using a moran test for each election. Neighbours (and therefore spatial weights) are assigned based on whether electorates share a border (1) or do not (0).

```{r}
# p values for moran test
```

A spatial error model (SEM) is fit to account for the spatial dependence, which is more efficient than OLS estimates. Before any variable selection, since there are only $N = 150$ electorates in each election and over *60* covariates available, the dimension of each cross-sections is reduced by grouping some variables together into factors. This is informed by principal component analysis.

- Need to see if partial residuals are even relevant for spatial error model. If not, an alternative significance measure should be used. Errors are not spherical so doesn't make sense to use lm on partial residuals. Ask Rob perhaps?
- Perhaps test is SEM is even better than LM? If not, then go with LM. No harm done.

As there are only $N = 150$ electorates in each election, the dimension of the cross-sections are reduced by grouping some variables together into factors, before select

An identical model specification is used across the six elections, with each election modelled separately. This enables the interpretation of temporal changes of the socio-demographic effects, as the same covariates are used in each election. As there are only $N = 150$ electorates in each election, the dimension of the cross-sections are reduced by grouping some variables together into factors. This is informed by principal component analysis.

As many of the socio-demographic variables represent similar information (for example, median personal income and median household income) many of the socio-demographic variables are grouped together into factors - informed by principal component analysis. This reduces the

Before selecting what model to use, or which covariates to include, many of the socio-demographic variables are grouped together into factors - informed by principal component analysis. This reduces the

The response, the two-party preferred vote, lies in the interval $(0,100)$, with $TPP = 70$ representing a 70% preference for Liberal, 30% for Labor. Since the observed values of $TPP$ are never very close to 0 or 100 (minimum $24.05 \%$ and maximum $74.90 \%$), there is no need to impose the constraint of $TPP \in [0,100]$. and so linear regression models are estimated using ordinary least squares.

Therefore, for a given election year, we assume $TPP \sim N(X \beta, \sigma^2)$, where $TPP$ is an $n \times 1$ vector containing $n=150$ two-party preferred votes, $X$ is an ($n \times p$) matrix of socio-demographic predictors, $\beta$ is a ($p \times 1$) vector of coefficients to be estimated. By assumption, the regression errors are taken to be independent and identically normally distributed with mean $0$ and variance $\sigma^2$. The estimation is done separately for each election.

Another option would is to combine these cross-sections into a single pooled cross-section with 900 observations, allowing for the construction of a longitudinal model. However, this would require placing some restrictions on which variables are time varying, which is unfavourable in this context. Alternatively it is possible to construct a panel data set for a particular set of electoral boundaries across all elections. This way we could potentially capture any unobserved individual heterogeneity using either fixed or random effects, but since electoral boundaries change, to construct panel data further spatial imputation would be required to generate socio-demographic profiles for each election with a common set of boundaries. Additionally, the two-party preferred vote would have to be imputed for these same boundaries, for all elections, which would introduce greater uncertainty. Finally, the same issues of choosing time-varying effects will arise in the panel data context as for the pooled cross-section.

Alternative without concern for undue bias stemming from an incorrectly imposed restriction on any variable.

Socio-demographic variables for each electorate have now been calculated and matched with the two-party preferred vote, for all six elections. In this chapter, we consider how to best answer the research question; *Which socio-demographics are the key determinants of support between the Labor and Liberal parties, and how have their effects changed over time?* This involves discussing an appropriate statistical model, reducing the dimension of the data by grouping similar variables together, selecting the most important predictors and adding interaction terms. Each of these four steps are described in subsections below, followed by the estimated models.

On this basis, modelling each election as a separate cross-section seems to us as the most appropriate statistical model to pursue. By using identical model specification, effects may be compared across six different elections, enabling the interpretation of temporal changes without concern for undue bias stemming from an incorrectly imposed restriction on any variable.

### Response variable - two-party preferred vote

The two-party preferred vote is a measure of preference between only the two major parties (Labor and Liberal), and is oriented in favour of the Liberal party, sitting in the interval $(0,100)$. For example, $TPP = 70$ represents a 70% preference for Liberal, 30% for Labor. Since the observed values of $TPP$ are never very close to 0 or 100 (minimum $24.05 \%$ and maximum $74.90 \%$), there is no need to impose the constraint of $TPP \in [0,100]$ and so linear regression models are estimated using ordinary least squares.

Therefore, for a given election year, we assume $TPP \sim N(X \beta, \sigma^2)$, where $TPP$ is an $n \times 1$ vector containing $n=150$ two-party preferred votes, $X$ is an ($n \times p$) matrix of socio-demographic predictors, $\beta$ is a ($p \times 1$) vector of coefficients to be estimated. By assumption, the regression errors are taken to be independent and identically normally distributed with mean $0$ and variance $\sigma^2$. The estimation is done separately for each election.

### 2019-03-19 modelling.Rmd
modelling - finished PCA, next Akaike

@@
 ## Modelling
-- Join the data, treat at separate cross-sections
 - Address the issue of changing variable scales (rent as example)
-- 
-- PCA to group variables together
+- Dimension reduction via PCA to group variables together
+- Choosing an appropriate model - discuss the spatial element
 - Akaike weights to derive a variable set for models
-- Spatial correlation in response, so model as spatial error model
+- Final model specification
 
@@
 Socio-demographic variables within each election year are standardized to have mean zero and variance one, to adjust for changing variable scales.
-For example, inflation-adjusted median rental prices increased across almost all electorates, with median rent of $200 per week placing an electorate in the 90th percentile in 2001, but only the 30th percentile in 2016.
+For example, inflation-adjusted median rental prices increased across almost all electorates, with median rent of 200 dollars per week placing an electorate in the 90th percentile in 2001, but only the 30th percentile in 2016.
+The response, the two-party preferred vote in favour of the Liberal party (denoted $Y$), lies in the interval $(0,100)$, where $Y = 70$ would represent a 70% preference for Liberal, 30% for Labor.
+Since the observed values of $Y$ are never very close to 0 or 100 (minimum $24.05 \%$ and maximum $74.90 \%$), there is no need to impose the constraint of $Y \in [0,100]$.
+
+An identical model specification is used across the six elections, with each election modelled separately.
+This allows for the socio-demographic effects to be estimated separately for each year, allowing for interpretation of temporal changes in these effects.
+This is prefereable over a single longitudinal model because it avoids any concerns of undue bias stemming from an incorrectly imposed time-varying restriction on any variable.
+Without such restrictions, a pooled cross-sectional model does not yield any distinct advantage over separate models.
+The panel approach is avoided because of how frequently electoral boundaries change, so any fixed or random effects models would be inappropriate.
+
+For each election a spatial error model [@Anselin99] is to be estimated, as electorates have a spatial structure and it is reasonable to expect errors to have spatial dependence.
+This is a common approach for modelling aggregate spatial units [@Anselin2002].
+Spatial weights are calculated by allowing each electorate to be equally correlated with any electorate that shares any part of its boundary.
+A moran's I test of the two-party preferred vote finds overwhelming evidence significant spatial autocorrelation for each election ($p \le 7\cdot10^{-15}$).
+
+Let $\rho$ be the parameter governing the extent of spatial correlation, $\boldsymbol \varepsilon$ be a spherical error term, ${\bf W}$ be a matrix of spatial weights (containing information about the neighbouring regions), $\bf X$ is a matrix of socio-demographic covariates and $\boldsymbol \beta$ is a vector of regression coefficients.
+
+The spatial error model is specified:
+
+$${\bf y} = {\bf X} {\boldsymbol \beta} + ({\bf I}_n-\rho {\bf W})^{-1}{\boldsymbol \varepsilon}$$
+Where $\varepsilon \sim N(0, \sigma^2)$. Estimation is done via maximum likelihood (using the `spdep` `R` package [@spdep]).
+
+## Dimension reduction
+With only $N = 150$ observations (electorates) in each election and $p = 63$ socio-demographic variables in each cross-section, any model using all variables would face serious problems with multi-collinearity and over-fitting, likely leading to erroneous conclusions regarding variable significance.
+Therefore a form of dimension reduction is adopted that involves a two-stage approach that results in the identification of a reduced predictor set, for which models are fit.
+
+The first step[^2] is to combine socio-demographic variables that represent similar information into "factors" using principal component analysis (PCA).
+The scree plots of the principal components for each election all level off after four components, and the loadings of these four components are similar across the elections.
+Principal components are then computed on the combined set of socio-demographics across all six elections.
+A factor is created by combining several variables all have large loadings in a particular component and when there is an intuitive reason as to why these variables could represent common information.
+A loading with magnitude greater than 0.15 is considered large.
+After computing these sums, each factor is again standardized to have mean zero and variance one, within each election.
+
+Consider the `Incomes` factor as an illustration.
+Independent of principal components, we may suspect that median personal income, median household income and median family income are providing similar information about the financial wellbeing of an electorate.
+Their loadings in the first principal component are large (0.20, 0.21 and 0.22 respectively), which provides the evidence needed to combine these variables into a single factor, which is called `Incomes`.
+
+This process reduces the predictor set to $p = 30$.
+However, it is still relatively large, and an additional variable selection method is required.
+Our target is to end up with a smaller subset of variables in the final stage when we model each election.
+
+[^2]: A preliminary step involved removing all age variables aside from median age.
+
+----------------------------------------------------------------------------------------
+
+Electorates One might suspect an underlying spatial dependence between electorates in their voting behaviour.
+This is  ggregate spatial units are commonly modelled with an ecological regression, which in this study takes the form  [@Anselin2002].
+Electoral voting behaviour  assume the form of the spatial error model
+
+The spatial weight matrix $\bf W$ is therefore specified such that $W_{i,j} = 0$ if electorate $i$ does not share a border with electorate $j$, and $W_{i,j} = \frac{1}{N_i}$ if electorates $i$ and $j$ do share a border, where $N_i$ is the number of electorates that $i$ shares a border with.
+
+Talk about spatial dependence in the structure and decide on SEM.
+
+Address problem of relatively small N large p. PCA for factors. Akaike weights for most important variables.
+
+Interactions. Final model specification.
+
+.
+As there are only $N = 150$ electorates in each election, the dimension of the cross-sections are reduced by grouping some variables together into factors.
+This is informed by principal component analysis.
+
+As there are $N=150$ electorates in each election, but *60*
 

### 2019-03-19 modelling.Rmd
first draft done

@@
 This allows for the socio-demographic effects to be estimated separately for each year, allowing for interpretation of temporal changes in these effects.
-This is prefereable over a single longitudinal model because it avoids any concerns of undue bias stemming from an incorrectly imposed time-varying restriction on any variable.
+This is preferable over a single longitudinal model because it avoids any concerns of undue bias stemming from an incorrectly imposed time-varying restriction on any variable.
 Without such restrictions, a pooled cross-sectional model does not yield any distinct advantage over separate models.
@@
 Spatial weights are calculated by allowing each electorate to be equally correlated with any electorate that shares any part of its boundary.
-A moran's I test of the two-party preferred vote finds overwhelming evidence significant spatial autocorrelation for each election ($p \le 7\cdot10^{-15}$).
+A Moran's I test of the two-party preferred vote finds overwhelming evidence significant spatial autocorrelation for each election ($p \le 7\cdot10^{-15}$).
 
@@
 ## Dimension reduction
-With only $N = 150$ observations (electorates) in each election and $p = 63$ socio-demographic variables in each cross-section, any model using all variables would face serious problems with multi-collinearity and over-fitting, likely leading to erroneous conclusions regarding variable significance.
+With only $N = 150$ observations (electorates) in each election and $p = 59$ socio-demographic variables in each cross-section, any model using all variables would face serious problems with multi-collinearity and over-fitting, likely leading to erroneous conclusions regarding variable significance.
 Therefore a form of dimension reduction is adopted that involves a two-stage approach that results in the identification of a reduced predictor set, for which models are fit.
@@
 
-----------------------------------------------------------------------------------------
+## Selecting a superset
 
-Electorates One might suspect an underlying spatial dependence between electorates in their voting behaviour.
-This is  ggregate spatial units are commonly modelled with an ecological regression, which in this study takes the form  [@Anselin2002].
-Electoral voting behaviour  assume the form of the spatial error model
+The second stage of the data reduction process involves the selection of a superset of predictors that captures the five *most important* variables from each election.
+To determine the five most important variables for each election, Akaike weights are used to measure relative variable importance [@BurnhamAnderson2002].
+This method is commonly used in ecological studies and is fundamentally based on the Akaike information criterion (AIC) [@Akaike73].
+Akaike weights are produced using a method of scoring each model relative to a collection of models $M$ using AIC.
 
-The spatial weight matrix $\bf W$ is therefore specified such that $W_{i,j} = 0$ if electorate $i$ does not share a border with electorate $j$, and $W_{i,j} = \frac{1}{N_i}$ if electorates $i$ and $j$ do share a border, where $N_i$ is the number of electorates that $i$ shares a border with.
+Let $\Delta_m = AIC_m - AIC_{min}$ denote the difference in AIC between models $m$ and that with minimum AIC in the model set $M$ for $m = 1,2,...,R$.
+The Akaike weights $w_m$ for model $m$ is calculated as
 
-Talk about spatial dependence in the structure and decide on SEM.
+$$w_m = \frac {\exp(-\frac{1}{2}\Delta_m)} {\sum_{r=1} ^R \exp(-\frac{1}{2}\Delta_r)},$$
+for $m = 1,...,R$.
 
-Address problem of relatively small N large p. PCA for factors. Akaike weights for most important variables.
+Akaike weights represent the posterior probability that model $m$ is the best model in the set, without imposing any beliefs a priori, as $w_m \in (0,1)$ and $w_m$ sum to 1.
 
-Interactions. Final model specification.
+For each variable $j \in 1, ..., J$, the sum of the Akaike weights is computed over all models that include that variable, and denote this sum by $s_j$. That is, let
 
-.
-As there are only $N = 150$ electorates in each election, the dimension of the cross-sections are reduced by grouping some variables together into factors.
-This is informed by principal component analysis.
+$$s_j = \sum_{m=1}^{R} w_m \cdot \text{I}(\text{variable } j \text{ used in model } m),$$
+where $I(A)$ denotes the indicator function for A.
 
-As there are $N=150$ electorates in each election, but *60*
+Variables are ordered according to $s_j$, which may be used as a measure of variable importance [@BurnhamAnderson2002].
+The variable with largest $s_j$ is deemed to be the most important in the set for the purpose of modelling the response variable.
 
-An identical model specification is used across the six elections, with each election modelled separately.
-This enables the interpretation of temporal changes of the socio-demographic effects, as the same covariates are used in each election.
-The response, the two-party preferred vote, lies in the interval $(0,100)$, with $TPP = 70$ representing a 70% preference for Liberal, 30% for Labor.
-Since the observed values of $TPP$ are never very close to 0 or 100 (minimum $24.05 \%$ and maximum $74.90 \%$), there is no need to impose the constraint of $TPP \in [0,100]$.
-Furthermore, significant spatial autocorrelation in the response is found using a moran test for each election.
-Neighbours (and therefore spatial weights) are assigned based on whether electorates share a border (1) or do not (0).
+For each election, the model set $M$ used, for which Akaike weights are constructed, is the set of all possible five variable linear models.
+This involves fitting $R = {32 \choose 5} = 201,376$ models for each election, with five representing a compromise between an undue computational burden and capturing sufficient information.
+The five most important variables are selected from each election, and used to construct a variable superset obtained by taking the union of these six subsets.
+In doing so, a variable that appears to be important in a particular election (though it may be potentially unimportant in the others) is captured, along with those that are important in multiple elections.
+This superset contains the chosen predictors to be used in all of the final election models.
+Five variables are chosen from each election to ensure that this superset is not too large.
+The resultant superset contains twelve variables.
 
-```{r}
-# p values for moran test
-```
+Before finalizing the common regression model to be fitted for each election year, two-way interactions are added using an iterative procedure.
+First, for each election, a model is fit using the superset of predictors as main effects only.
+Then all models with a single two-way interaction are computed, and the interaction tested using a likelihood-ratio test (1% significance level).
+The number of elections for which a given interaction is significant is tallied, and the interaction that is most frequently significant is selected to remain in the model.
+This process is repeated until no interaction is significant in more than one election.
 
-A spatial error model (SEM) is fit to account for the spatial dependence, which is more efficient than OLS estimates.
-Before any variable selection, since there are only $N = 150$ electorates in each election and over *60* covariates available, the dimension of each cross-sections is reduced by grouping some variables together into factors.
-This is informed by principal component analysis.
+Now that model specification has been determined, the final models are fit using the superset of 12 main effects and the four two-way interactions.
+The equation, estimated separately for each year, is as follows:
 
-- Need to see if partial residuals are even relevant for spatial error model.
-If not, an alternative significance measure should be used.
-Errors are not spherical so doesn't make sense to use lm on partial residuals.
-Ask Rob perhaps?
-- Perhaps test is SEM is even better than LM? If not, then go with LM. No harm done.
+$\small{LNP_i = \beta_0 + \beta_1*\text{Extractive}_i + \beta_2*\text{Born_SE_Europe}_i + \beta_3*\text{ManagerAdminClericalSales}_i + \beta_4*\text{DeFacto}_i + \beta_5*\text{DiffAddress}_i + \beta_6*\text{Married}_i + \beta_7*\text{Born_UK}_i + \beta_8*\text{MedianAge}_i + \beta_9*\text{OneParent_House}_i + \beta_{10}*\text{OtherLanguageHome}_i + \beta_{11}*\text{Transformative}_i + \beta_{12}*\text{Unemployment}_i + \beta_{13}*\Big(\text{Extractive}_i*\text{OtherLanguageHome}_i\Big) + \beta_{14}*\Big(\text{DeFacto}_i*\text{OtherLanguageHome}_i\Big) + \beta_{15}*\Big(\text{Extractive}_i*\text{ManagerAdminClericalSales}_i\Big) + \beta_{16}*\Big(\text{MedianAge}_i*\text{OneParent_House}_i\Big) + \u_i}$
 
-As there are only $N = 150$ electorates in each election, the dimension of the cross-sections are reduced by grouping some variables together into factors, before select
+where ${\bf u} = ({\bf I_n} - \rho \cdot W) \cdot {\boldsymbol \varepsilon}$ and $\varepsilon_i \overset{iid}\sim N(0,\sigma^2)$.
 
-An identical model specification is used across the six elections, with each election modelled separately.
-This enables the interpretation of temporal changes of the socio-demographic effects, as the same covariates are used in each election.
-As there are only $N = 150$ electorates in each election, the dimension of the cross-sections are reduced by grouping some variables together into factors.
-This is informed by principal component analysis.
-
-As many of the socio-demographic variables represent similar information (for example, median personal income and median household income) many of the socio-demographic variables are grouped together into factors - informed by principal component analysis.
-This reduces the
-
-Before selecting what model to use, or which covariates to include, many of the socio-demographic variables are grouped together into factors - informed by principal component analysis.
-This reduces the
-
-The response, the two-party preferred vote, lies in the interval $(0,100)$, with $TPP = 70$ representing a 70% preference for Liberal, 30% for Labor.
-Since the observed values of $TPP$ are never very close to 0 or 100 (minimum $24.05 \%$ and maximum $74.90 \%$), there is no need to impose the constraint of $TPP \in [0,100]$.
-and so linear regression models are estimated using ordinary least squares.
-
-Therefore, for a given election year, we assume $TPP \sim N(X \beta, \sigma^2)$, where $TPP$ is an $n \times 1$ vector containing $n=150$ two-party preferred votes, $X$ is an ($n \times p$) matrix of socio-demographic predictors, $\beta$ is a ($p \times 1$) vector of coefficients to be estimated.
-By assumption, the regression errors are taken to be independent and identically normally distributed with mean $0$ and variance $\sigma^2$.
-The estimation is done separately for each election.
-
-Another option would is to combine these cross-sections into a single pooled cross-section with 900 observations, allowing for the construction of a longitudinal model.
-However, this would require placing some restrictions on which variables are time varying, which is unfavourable in this context.
-Alternatively it is possible to construct a panel data set for a particular set of electoral boundaries across all elections.
-This way we could potentially capture any unobserved individual heterogeneity using either fixed or random effects, but since electoral boundaries change, to construct panel data further spatial imputation would be required to generate socio-demographic profiles for each election with a common set of boundaries.
-Additionally, the two-party preferred vote would have to be imputed for these same boundaries, for all elections, which would introduce greater uncertainty.
-Finally, the same issues of choosing time-varying effects will arise in the panel data context as for the pooled cross-section.
-
-Alternative without concern for undue bias stemming from an incorrectly imposed restriction on any variable.
-
-Socio-demographic variables for each electorate have now been calculated and matched with the two-party preferred vote, for all six elections.
-In this chapter, we consider how to best answer the research question; *Which socio-demographics are the key determinants of support between the Labor and Liberal parties, and how have their effects changed over time?* This involves discussing an appropriate statistical model, reducing the dimension of the data by grouping similar variables together, selecting the most important predictors and adding interaction terms.
-Each of these four steps are described in subsections below, followed by the estimated models.
-
-On this basis, modelling each election as a separate cross-section seems to us as the most appropriate statistical model to pursue.
-By using identical model specification, effects may be compared across six different elections, enabling the interpretation of temporal changes without concern for undue bias stemming from an incorrectly imposed restriction on any variable.
-
-### Response variable - two-party preferred vote
-
-The two-party preferred vote is a measure of preference between only the two major parties (Labor and Liberal), and is oriented in favour of the Liberal party, sitting in the interval $(0,100)$.
-For example, $TPP = 70$ represents a 70% preference for Liberal, 30% for Labor.
-Since the observed values of $TPP$ are never very close to 0 or 100 (minimum $24.05 \%$ and maximum $74.90 \%$), there is no need to impose the constraint of $TPP \in [0,100]$ and so linear regression models are estimated using ordinary least squares.
-
-Therefore, for a given election year, we assume $TPP \sim N(X \beta, \sigma^2)$, where $TPP$ is an $n \times 1$ vector containing $n=150$ two-party preferred votes, $X$ is an ($n \times p$) matrix of socio-demographic predictors, $\beta$ is a ($p \times 1$) vector of coefficients to be estimated.
-By assumption, the regression errors are taken to be independent and identically normally distributed with mean $0$ and variance $\sigma^2$.
-The estimation is done separately for each election.
+Table *XXX* details the estimated model and their standard errors for each of the six elections. These are interpreted in the following section.

### 2019-03-20 data-collection.Rmd
edit

@@
 
-The voting outcome of interest is the electoral two-party preferred vote, which is provided by the Australian Electoral Commission (AEC) for the 2001, 2004, 2007, 2010, 2013 and 2016 elections.
+The voting outcome of interest is the electoral two-party preferred vote, which is provided by the Australian Electoral Commission (AEC) for the 2001, 2004, 2007, 2010, 2013 and 2016 elections via the AEC Tally Room.
 The AEC divide Australia into 150 regions called electorates, with each corresponding to a single seat in the House of Representatives.
@@
 From this aggregate information, *63* socio-demographic variables are computed for each of the electorates.
-These variables follow the definitions used in @Stimson09.
+The choice of variables to construct mimic those used in @Stimson09.
 
@@
 Raw data is sourced online from the .csv and .xlsx files from the AEC and ABS websites.
-The format of these files change over the years, meaning that significant time and effort is required to extract the appropriate electoral information.
-All data wrangling is done using `R`, predominantly using the tools from the `dplyr` and `readxl` packages, providing fast consistent tools for data manipulation and functions to import .xlsx files (respectively).
-The 2001 and 2006 Census data is formatted such that each electorate has a separate document, making it difficult to use the `dplyr` tools.
-Instead, cells had to be manually selected from each individual files.
+The format of these files change over the years, meaning that significant effort is required to extract the appropriate electoral information.
+All data wrangling is done using `R`, predominantly using the tools from the `dplyr` [@dplyr] and `readxl` [@readxl] packages, providing fast consistent tools for data manipulation and functions to import .xlsx files (respectively).
+
+The 2001 and 2006 Census data is published in a format where each electorate has a separate document, making it difficult to use the `dplyr` tools.
+Instead, cells had to be selected from each individual file to construct the desired variables.
+For 2011 and 2016, the available Census data packs contain a separate document for each question and are easily compatible with the `dplyr` verbs, as are the published election vote outcomes.
+
 This work is entirely reproducible, and all scripts can be found in the github repository for the `eechidna` `R` package.

### 2019-03-20 imputation.Rmd
edit

@@
 
+# Joining Census and election data
+
 ## Differences between Census and electoral data
@@
 Between 2001 and 2016 there were six elections and four Censuses (see Figure \@ref(fig:timeline)).
-Only in 2001 and 2016 did both events occur, and electoral boundaries are also redistributed regularly.
-Therefore, there are both temporal and spatial differences that need to be accounted for when joining the electoral two-party preferred vote from the four elections between 2004 and 2013 with Census data.
-For these elections a spatio-temporal imputation method is employed to derive a electoral socio-demographics which uses Census information from both before and after the election of interest.
+Electoral boundaries are redistributed regularly by the AEC, meaning that only in the years where both a Census and election occur will the boundaries match - the case for the 2001 and 2016 election.
+Therefore, for the four elections between 2004 and 2013, both temporal and spatial differences in electorates need to be accounted for when joining the electoral two-party preferred vote with Census data.
+For these elections a spatio-temporal imputation method is employed to obtain electoral socio-demographics.
+This method uses Census information from both before and after the election of interest.
 
@@
 
-To account for spatial differences, the piece-wise approximation method in @Goodchild1993 is appropriate for our problem.
+To account for spatial differences, the piece-wise approximation method in @Goodchild1993 is adopted.
 Consider a map of source zones $s = 1,...,S$, for which socio-demographic information is available, and a set of target zones $t = 1,...,T$ for which information is to be imputed.
@@
 $$\hat{C}_t = \frac{\sum_{s=1}^{S}{D_s*\hat{P}_{s,t}}}{\sum_{s=1}^{S}{\hat{P}_{s,t}}}$$
-This assumes that each individual in the source zones assume the aggregate characteristics of the zone.
+This assumes that each individual in a source zone assumes the aggregate characteristics of the zone.
 
-Socio-demographic information is now imputed for the target zones, but these are not appropriate for the target year.
+Applying this to each of the target zones addresses the spatial component, as it imputes the required socio-demographic for the desired electoral boundaries.
+However these are applicable at the time of the Census (source year) and are not yet appropriate for the election (target year).
+
 Denote year $y$, with a Census falling on $y_1$ and $y_3$, and an election on year $y_2$, and add this subscript to the Census variable estimate, $\hat{C}_{t,y}$.
@@
 For this study, electoral divisions are used and this imputation method is applied to each of the 2004, 2007, 2010 and 2013 elections.
-To demonstrate its functionality, consider the imputation of the socio-demographic variable $AusCitizen$ for the electorate of Hume in New South Wales (NSW), at the time of the 2013 federal election.
+To demonstrate its functionality, consider the imputation of socio-demographic variables for the electorate of Hume in New South Wales (NSW), at the time of the 2013 federal election.
 Figure \@ref(fig:hume13) shows this region amongst other NSW electorates.
@@
 
-The socio-demographic of interest is $AusCitizen$, which is the proportion of people in the region who are Australian citizens.
+Now consider the socio-demographic $AusCitizen$ - the proportion of people in the region who are Australian citizens.
 

### 2019-03-20 introduction.Rmd
edit

@@
 This study explores how electoral characteristics relate to two-party preference, and whether their effects have changed over time.
-Information about electoral socio-demographics are from the Census, and vote counts from the federal election.
+Electoral socio-demographics are derived from the Census, and vote counts are obtained from the federal election.
 
-Joining these two data sources is problematic, as there in an inherent asynchronicity in the two events - a Census occurs every five years, whereas an election usually occurs every three years.
+Joining these two data sources is problematic as there in an inherent asynchronicity in the two events, a Census occurs every five years, whereas an election usually occurs every three years.
+The first problem addressed is that of obtaining appropriate Census data for the 2004, 2007, 2010 and 2013 elections - election years in which a Census does not occur.
 The predominant approach in previous studies is to join voting outcomes to the nearest Census, without accounting for any temporal differences (see @DavisStimson98, @Stimson06, @Liao09 and @Stimson09).
 Furthermore, electoral boundaries change regularly, so spatial discrepancies also arise when matching electoral data.
-To obtain appropriate Census data for the 2004, 2007, 2010 and 2013 elections, we adopt impute electoral socio-demographics using spatio-temporal imputation that combines areal interpolation [@Goodchild1993] and linear time-interpolation.
-The process of collection and wrangling is detailed in @CHAPTER2 and imputation process is outlined in @CHAPTER3.
-All data and documentation relating to this procedure is available in the `eechidna` `R` package on `CRAN`.
+To obtain appropriate Census data for these four elections, electoral socio-demographics are imputed using a spatio-temporal imputation that combines areal interpolation [@Goodchild1993] and linear time-interpolation.
+The process of collection and wrangling the raw data is outlined in @CHAPTER2, and the imputation process is detailed in @CHAPTER3.
+All data and associated documentation relating to this procedure has been made available in the `eechidna` `R` package on `CRAN`, providing a resource for future analysis.
 
@@
 @Stimson06, @Stimson09 and @Stimson12 conduct principal component analysis of polling booths in the 2001, 2004 and 2007 elections respectively, also finding that socio-demographic characteristics of polling booths are linked to their two-party preferred vote.
-@Stimson09 also model the polling booth swing vote (change in the two-party preferred vote) in the 2007 election, but had little success.
-Instead of analyzing a single election in isolation, this paper employs a consistent model framework across elections so that any temporal changes are observed, where each of the six federal elections are modelled with a cross-sectional data set.
-This consists of the two-party preferred vote, which is the response variable, and socio-demographic variables which are explanatory variables that characterise each electorate.
-This process is discussed in @CHAPTER4 and involves standardizing variables across years, using principal components to group variables, and then selecting the most influential variables with an information-theoretic approach via Akaike weights to determine relative variable importance [@BurnhamAnderson2002].
-Spatial error models are then fit to account for spatial dependence.
+@Stimson09 model the polling booth swing vote (change in the two-party preferred vote) in the 2007 election, but had little success.
+Instead of analyzing a single election in isolation, this paper employs a consistent model framework across six elections so that any temporal changes are observed, where each federal elections is modelled with a cross-sectional data set.
+Each cross-section consists of the two-party preferred vote (response variable), and socio-demographic variables (explanatory variables) that characterise each electorate.
+This process involves standardizing variables, using principal components to group variables into "factors" and finally fitting a spatial error model.
+These steps, and the rational behind a spatial model, are discussed in @CHAPTER4.
+Inference is then conducted on the models and insights are discussed in @CHAPTER5.
 
 ---------
+
+, and then selecting the most influential variables with an information-theoretic approach via Akaike weights to determine relative variable importance [@BurnhamAnderson2002].
+Spatial error models are then fit to account for spatial dependence.
 

### 2019-03-20 modelling.Rmd
rearranged so that it is compatible with full model

@@
 
+### Pre-processing
 With socio-demographic information now available for each electorate, each election is joined to the data  corresponding with its two-party preferred vote.
@@
 For example, inflation-adjusted median rental prices increased across almost all electorates, with median rent of 200 dollars per week placing an electorate in the 90th percentile in 2001, but only the 30th percentile in 2016.
-The response, the two-party preferred vote in favour of the Liberal party (denoted $Y$), lies in the interval $(0,100)$, where $Y = 70$ would represent a 70% preference for Liberal, 30% for Labor.
-Since the observed values of $Y$ are never very close to 0 or 100 (minimum $24.05 \%$ and maximum $74.90 \%$), there is no need to impose the constraint of $Y \in [0,100]$.
 
-An identical model specification is used across the six elections, with each election modelled separately.
-This allows for the socio-demographic effects to be estimated separately for each year, allowing for interpretation of temporal changes in these effects.
-This is preferable over a single longitudinal model because it avoids any concerns of undue bias stemming from an incorrectly imposed time-varying restriction on any variable.
-Without such restrictions, a pooled cross-sectional model does not yield any distinct advantage over separate models.
-The panel approach is avoided because of how frequently electoral boundaries change, so any fixed or random effects models would be inappropriate.
-
-For each election a spatial error model [@Anselin99] is to be estimated, as electorates have a spatial structure and it is reasonable to expect errors to have spatial dependence.
-This is a common approach for modelling aggregate spatial units [@Anselin2002].
-Spatial weights are calculated by allowing each electorate to be equally correlated with any electorate that shares any part of its boundary.
-A Moran's I test of the two-party preferred vote finds overwhelming evidence significant spatial autocorrelation for each election ($p \le 7\cdot10^{-15}$).
-
-Let $\rho$ be the parameter governing the extent of spatial correlation, $\boldsymbol \varepsilon$ be a spherical error term, ${\bf W}$ be a matrix of spatial weights (containing information about the neighbouring regions), $\bf X$ is a matrix of socio-demographic covariates and $\boldsymbol \beta$ is a vector of regression coefficients.
-
-The spatial error model is specified:
-
-$${\bf y} = {\bf X} {\boldsymbol \beta} + ({\bf I}_n-\rho {\bf W})^{-1}{\boldsymbol \varepsilon}$$
-Where $\varepsilon \sim N(0, \sigma^2)$. Estimation is done via maximum likelihood (using the `spdep` `R` package [@spdep]).
-
-## Dimension reduction
+#### Dimension reduction
 With only $N = 150$ observations (electorates) in each election and $p = 59$ socio-demographic variables in each cross-section, any model using all variables would face serious problems with multi-collinearity and over-fitting, likely leading to erroneous conclusions regarding variable significance.
@@
 
-[^2]: A preliminary step involved removing all age variables aside from median age.
+[^2]: A preliminary step involved removing all age bands, because age is represented by median age, and to remove some of the other
+
+### Model framework
+An identical model specification is used across the six elections, with each election modelled separately.
+This allows for the socio-demographic effects to be estimated separately for each year, allowing for interpretation of temporal changes in these effects.
+This is preferable over a single longitudinal model because it avoids any concerns of undue bias stemming from an incorrectly imposed time-varying restriction on any variable.
+Without such restrictions, a pooled cross-sectional model does not yield any distinct advantage over separate models.
+The panel approach is avoided because of how frequently electoral boundaries change, so any fixed or random effects models would be inappropriate.
+
+The response, the two-party preferred vote in favour of the Liberal party (denoted $Y$), lies in the interval $(0,100)$, where $Y = 70$ would represent a 70% preference for Liberal, 30% for Labor.
+Since the observed values of $Y$ are never very close to 0 or 100 (minimum $24.05 \%$ and maximum $74.90 \%$), there is no need to impose the constraint of $Y \in [0,100]$.
+Electorates represent aggregate spatial units, and a Moran's I test of the two-party preferred vote finds significant spatial autocorrelation for each election ($p \le 7\cdot10^{-15}$), so the spatial structure of the data must be taken into account.
+The spatial error model [@Anselin99] is chosen because it allows for spatial correlation in the errors.
+This can be thought of as capturing the political climate, where political pressures (unobserved characteristics influencing voter behaviour) of neighbours influence an electorate's voting behaviour.
+Spatial weights are calculated by allowing each electorate to be equally correlated with any electorate that shares any part of its boundary.
+The spatial error model is specified as follows:
+
+Let $\rho$ be the parameter governing the extent of spatial correlation, $\boldsymbol v$ be a spherical error term, ${\bf W}$ be a matrix of spatial weights (containing information about the neighbouring regions), $\bf X$ be a matrix of socio-demographic covariates, $\boldsymbol \beta$ be a vector of regression coefficients and ${\boldsymbol u}$ be a vector of spatially correlated errors.
+
+$${\bf y} = {\bf X} {\boldsymbol \beta} + {\boldsymbol u}$$
+
+with
+
+$${\boldsymbol u} = \rho {\boldsymbol W} {\boldsymbol u} + {\boldsymbol v}$$
+where
+
+$${\boldsymbol v} \sim N({\boldsymbol 0}, \sigma^2 {\boldsymbol I_n})$$.
+
+so it can be written
+
+$${\bf y} = {\bf X} {\boldsymbol \beta} + ({\bf I}_n-\rho {\bf W})^{-1}{\boldsymbol v}$$
+
+Estimation is done via maximum likelihood (using the `spdep` `R` package [@spdep]).
 

### 2019-03-21 modelling.Rmd
rearranged model framework

@@
 ```
-
-## Modelling
-- Address the issue of changing variable scales (rent as example)
-- Dimension reduction via PCA to group variables together
-- Choosing an appropriate model - discuss the spatial element
-- Akaike weights to derive a variable set for models
-- Final model specification
 
@@
 Without such restrictions, a pooled cross-sectional model does not yield any distinct advantage over separate models.
-The panel approach is avoided because of how frequently electoral boundaries change, so any fixed or random effects models would be inappropriate.
+The panel approach is avoided because of how frequently electoral boundaries change - electorates that have the same name across elections are not guaranteed to represent the same geographical region - meaning any fixed or random effects models would be difficult without implementing consistent boundaries requiring further imputation.
 
-The response, the two-party preferred vote in favour of the Liberal party (denoted $Y$), lies in the interval $(0,100)$, where $Y = 70$ would represent a 70% preference for Liberal, 30% for Labor.
-Since the observed values of $Y$ are never very close to 0 or 100 (minimum $24.05 \%$ and maximum $74.90 \%$), there is no need to impose the constraint of $Y \in [0,100]$.
-Electorates represent aggregate spatial units, and a Moran's I test of the two-party preferred vote finds significant spatial autocorrelation for each election ($p \le 7\cdot10^{-15}$), so the spatial structure of the data must be taken into account.
+For each cross-section, let the response variable be the two-party preferred vote in favour of the Liberal party, denoted $Y$, with $Y = 70$ representing a 70% preference for Liberal, 30% for Labor.
+Although $Y$ lies in the interval $(0,100)$, observed values are never very close to 0 or 100 (minimum $24.05 \%$ and maximum $74.90 \%$), so there is no need to impose the constraint of $Y \in [0,100]$.
+Furthermore, the response is found to be spatially correlated in each election (Moran's I test, $p \le 7\cdot10^{-15}$).
+This is expected, as electorates are aggregate spatial units, and hence the spatial structure of the data must modelled appropriately.
+
 The spatial error model [@Anselin99] is chosen because it allows for spatial correlation in the errors.
@@
 Estimation is done via maximum likelihood (using the `spdep` `R` package [@spdep]).
+
+----------------------------------------------------------------------------------------
+
+## Modelling
+- Address the issue of changing variable scales (rent as example)
+- Dimension reduction via PCA to group variables together
+- Choosing an appropriate model - discuss the spatial element
+- Akaike weights to derive a variable set for models
+- Final model specification
 

### 2019-03-21 data-collection.Rmd
made more concise

@@
 
-Raw data is sourced online from the .csv and .xlsx files from the AEC and ABS websites.
-The format of these files change over the years, meaning that significant effort is required to extract the appropriate electoral information.
-All data wrangling is done using `R`, predominantly using the tools from the `dplyr` [@dplyr] and `readxl` [@readxl] packages, providing fast consistent tools for data manipulation and functions to import .xlsx files (respectively).
-
-The 2001 and 2006 Census data is published in a format where each electorate has a separate document, making it difficult to use the `dplyr` tools.
-Instead, cells had to be selected from each individual file to construct the desired variables.
-For 2011 and 2016, the available Census data packs contain a separate document for each question and are easily compatible with the `dplyr` verbs, as are the published election vote outcomes.
-
-This work is entirely reproducible, and all scripts can be found in the github repository for the `eechidna` `R` package.
-These scripts also provide a resource to help wrangle data for future Censuses and elections.
+Raw data is sourced online from the AEC and ABS websites in .csv and .xlsx files.
+The format of these files change over the years, making extracting the appropriate information a big task.
+The functions available in the `dplyr` [@dplyr] and `readxl` [@readxl] `R` packages are very useful, as they provide fast consistent tools for data manipulation and functions to import .xlsx files (respectively).
+The 2001 and 2006 Census data are however published in a format where each electorate has a separate document, making it difficult to use the `dplyr` tools and instead cells have to be selected from each individual file to construct the desired variables.
+All scripts required for the data wrangling process can be found in the github repository for the `eechidna` `R` package, along with the raw data.
+The `eechidna` package makes this study is entirely reproducible and also provides a resource to help wrangle data for future Censuses and elections, when they become available.

### 2019-03-24 01-intro.Rmd
added bookdown files

# Introduction {#intro}

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, fig.pos = 'h', fig.align = 'center')
library(tidyverse)
library(eechidna)
library(ggplot2)
library(ggthemes)
library(knitr)
```

Australia has changed in many ways over the last two decades. Rising house prices, country-wide improvements in education, an ageing population, and a decline in religious affiliation, are just a few facets of the country's evolving socio-demographic characteristics. At the same time, political power has moved back and forth between the two major parties. In the 2007 and 2010 federal elections, the Australian Labor Party (Labor) was victorious, whereas the 2001, 2004, 2013 and 2016 elections were won by the Liberal National coalition (Liberal). The two-party preferred vote, a measure of support between these two parties, fluctuated between 47.3% and 53.5% (in favour of the Liberal party) over this period. This study explores how electoral characteristics relate to two-party preference, and whether their effects have changed over time. Electoral socio-demographics are derived from the Census, and vote counts are obtained from federal elections.

Joining these two data sources is problematic as there in an inherent asynchronicity in the two events. A Census is conducted by the Australian Bureau of Statistics (ABS) every five years, whereas a federal election (conducted by the Australian Electoral Commission (AEC)) usually occurs every three years. The first problem addressed is that of obtaining appropriate Census data for the 2004, 2007, 2010 and 2013 elections - election years in which a Census does not occur. The predominant approach in previous studies is to join voting outcomes to the nearest Census, without accounting for any temporal differences (see @DavisStimson98, @Stimson06, @Liao09 and @Stimson09). Furthermore, electoral boundaries change regularly, so spatial discrepancies also arise when matching electoral data. To obtain appropriate Census data for these four elections, electoral socio-demographics are imputed using a spatio-temporal imputation that combines areal interpolation [@Goodchild1993] and linear time-interpolation. The process of collecting and wrangling the raw data is outlined in section \@ref(data), and the imputation process is detailed in section \@ref(imputation). All data and associated documentation relating to this procedure are available in the `eechidna` `R` package [@eechidna], providing a resource for future analysis.

Previous work on modelling Australian federal elections have found that aggregate socio-demographics are relatively good predictors of voting outcomes. @Forrest01 does this using multiple regression of the Liberal and Labor primary vote for polling booths in the Farrer electorate in 1998. @Stimson06, @Stimson09 and @Stimson12 use principal component analysis of polling booths in the 2001, 2004 and 2007 elections respectively, also finding that socio-demographic characteristics of polling booths are linked to their two-party preferred vote. On the contrary, @Stimson09 models the polling booth swing vote (change in the two-party preferred vote) in the 2007 election, finding that little of swing vote can be explained by Census data. Instead of analyzing a single election in isolation, this paper employs a consistent model framework across six elections so that temporal changes in the effects of socio-demographics can be observed, where each federal elections is modelled with a cross-sectional data set. The use of a regression framework to examine these socio-political relationships over time is seemingly absent from previous Australian studies. It also appears that no study has attempted any type of statistical analysis of socio-demographics in conjunction with voter behaviour in Australia since 2007, making this paper distinctly different from those previous.

The cross-sectional data set for each election consists of the two-party preferred vote (response variable), and socio-demographic variables (explanatory variables) that characterise each electorate. To obtain these cross-sections, socio-demographic variables are first standardized, and then principal components are used to group variables into "factors". To account for the inherent spatial structure of the data, a spatial error model is fit for each election. These steps are discussed in section \@ref(modelling). In section \@ref(results) inference is conducted on the models to see which effects are significant, how effects change over time and which electorates have abnormal voting behaviour.

### 2019-03-24 02-data-collection.Rmd
added bookdown files

--- 2019-03-21 data-collection.Rmd
@@
----
-title: "data-collection"
-author: "Jeremy Forbes"
-date: "18/03/2019"
-output: html_document
----
-
-1. Where the data is sourced from
-
-2. Converting the Census raw counts into variables
-- Using the same variable definitions as in @Stimson09 
-- Tidyverse tools
-- Lack of coherence in ABS structures
-
---------------------------------------------------------------------------------
-
-Data description and collection
-
---------------------------------------------------------------------------------
+# Data collection {#data}
 
@@
 The ABS approximation of electorates at the time of the Census is chosen.
-From this aggregate information, *63* socio-demographic variables are computed for each of the electorates.
-The choice of variables to construct mimic those used in @Stimson09.
+From this aggregate information, 67 socio-demographic variables are computed for each of the electorates.
 
@@
 The 2001 and 2006 Census data are however published in a format where each electorate has a separate document, making it difficult to use the `dplyr` tools and instead cells have to be selected from each individual file to construct the desired variables.
-All scripts required for the data wrangling process can be found in the github repository for the `eechidna` `R` package, along with the raw data.
-The `eechidna` package makes this study is entirely reproducible and also provides a resource to help wrangle data for future Censuses and elections, when they become available.
+All scripts required for the data wrangling process can be found in the github repository for the `eechidna` `R` package [@eechidna], along with the raw data.
+The `eechidna` package makes this study entirely reproducible and provides a resource to help wrangle data for future Censuses and elections, when they become available.

### 2019-03-24 03-imputation.Rmd
added bookdown files

--- 2019-03-20 imputation.Rmd
@@
----
-title: "imputation"
-author: "Jeremy Forbes"
-date: "18/03/2019"
-output:
-  pdf_document: default
-  html_document: default
----
+# Joining Census and election data {#imputation}
 
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = F)
-library(tidyverse)
-library(eechidna)
-library(ggplot2)
-library(ggthemes)
-library(knitr)
-```
-
-# Joining Census and election data
-
-## Differences between Census and electoral data
+## Differences between Census and election data
 
@@
 Publically available Census data is aggregated and there are different resolutions accessible, ranging from SA1 (over 50,000 zones) to electoral divisions (150 zones).
-Any of these resolutions could be used as source zones.
-For this study, electoral divisions are used and this imputation method is applied to each of the 2004, 2007, 2010 and 2013 elections.
+For this study electoral divisions are used as source zones, and this imputation method is applied for each of the 2004, 2007, 2010 and 2013 elections.
 To demonstrate its functionality, consider the imputation of socio-demographic variables for the electorate of Hume in New South Wales (NSW), at the time of the 2013 federal election.
@@
 
-```{r hume13, fig.cap = "Some of the electoral boundaries in NSW for 2013, with the electoral boundary for Hume, shown in purple."}
+```{r hume13, fig.cap = "Some of the electoral boundaries in NSW for 2013, with the electoral boundary for Hume, shown in purple.", out.height="30%"}
 hume_area13 <- nat_map13 %>% 
@@
 
-The Censuses neighbouring the 2013 election are those in 2011 and 2016, and the Hume boundary is changed, as seen by plotting the Hume boundary (purple) in the 2013 election over the divisions in 2016.
+The Censuses neighbouring the 2013 election are those in 2011 and 2016, and the Hume boundary is changed, as seen by plotting the Hume boundary (purple) in the 2013 election over the divisions in 2016 (see Figure \@ref(fig:hume16)).
 
-```{r hume16, fig.cap = "Census division boundaries in NSW for 2016, with the 2013 electoral boundary for Hume, shown in purple. The purple region is not contained within a single Census division."}
+```{r hume16, fig.cap = "Census division boundaries in NSW for 2016, with the 2013 electoral boundary for Hume, shown in purple.
+The purple region is not contained within a single Census division.", out.height="30%"}
 hume_area16 <- nat_map16 %>% 
@@
 There are many electorates in 2016 that intersect with the purple region (Hume boundary for 2013), these include the divisions of Riverina, Eden-Monaro and Hume, along with smaller intersecting areas with Fenner, Calare, Gilmore and Whitlam.
-To impute Census information for this purple region, calculate the percentage of each 2016 electorate that intersects with the purple region, which is then used to estimate intersection populations $\hat{P}_{s,t} = \frac{U_s*A_{s,t}}{\sum_{t=1}^T A_{s,t}}$.
+To impute Census information for this purple region, calculate the percentage of each 2016 electorate that intersects with the purple region, which is then used to estimate intersection populations $\hat{P}_{s,t}$.
 
@@
 |DivisionNm      | AusCitizen (%): $D_s$| Estimated Population Allocated to Purple Region: $\hat{P}_{s,t}$|
-|----------------|----------------------|------------------------------------------------|
+|----------------|-------------------|------------------------------------------------|
 |HUME            |              90.02|                                          145427|

### 2019-03-24 04-modelling.Rmd
added bookdown files

--- 2019-03-21 modelling.Rmd
@@
----
-title: "modelling"
-author: "Jeremy Forbes"
-date: "18/03/2019"
-output: html_document
----
+# Modelling {#modelling}
 
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
-```
-
-## Modelling
-
-### Pre-processing
+## Data pre-processing
 With socio-demographic information now available for each electorate, each election is joined to the data  corresponding with its two-party preferred vote.
@@
 
-#### Dimension reduction
-With only $N = 150$ observations (electorates) in each election and $p = 59$ socio-demographic variables in each cross-section, any model using all variables would face serious problems with multi-collinearity and over-fitting, likely leading to erroneous conclusions regarding variable significance.
-Therefore a form of dimension reduction is adopted that involves a two-stage approach that results in the identification of a reduced predictor set, for which models are fit.
+### Dimension reduction
+With only $N = 150$ observations (electorates) in each election and $p = 65$ socio-demographic variables in each cross-section, any model using all variables would face serious problems with multi-collinearity and over-fitting, likely leading to erroneous conclusions regarding variable significance.
+Therefore a form of dimension reduction is adopted before models are fit.
 
-The first step[^2] is to combine socio-demographic variables that represent similar information into "factors" using principal component analysis (PCA).
+Socio-demographic variables[^2] that represent similar information are combined into "factors" using principal component analysis (PCA).
 The scree plots of the principal components for each election all level off after four components, and the loadings of these four components are similar across the elections.
@@
 This process reduces the predictor set to $p = 30$.
-However, it is still relatively large, and an additional variable selection method is required.
-Our target is to end up with a smaller subset of variables in the final stage when we model each election.
 
-[^2]: A preliminary step involved removing all age bands, because age is represented by median age, and to remove some of the other
+[^2]: A preliminary step involved removing all age bands, because age is represented by median age, and to remove variables relating to particular denominations of Christianity.
 
-### Model framework
+## Model framework
 An identical model specification is used across the six elections, with each election modelled separately.
@@
 This is preferable over a single longitudinal model because it avoids any concerns of undue bias stemming from an incorrectly imposed time-varying restriction on any variable.
-Without such restrictions, a pooled cross-sectional model does not yield any distinct advantage over separate models.
-The panel approach is avoided because of how frequently electoral boundaries change - electorates that have the same name across elections are not guaranteed to represent the same geographical region - meaning any fixed or random effects models would be difficult without implementing consistent boundaries requiring further imputation.
+Without such restrictions, a pooled cross-sectional model does not yield any distinct advantage over separate cross-sections.
+The panel approach is avoided because of how frequently electoral boundaries change, meaning that electorates that have the same name across elections are not guaranteed to represent the same geographical region.
+Therefore any fixed or random effects models would be difficult to estimate without implementing consistent boundaries, which would requiring further imputation.
 
@@
 
-The spatial error model [@Anselin99] is chosen because it allows for spatial correlation in the errors.
-This can be thought of as capturing the political climate, where political pressures (unobserved characteristics influencing voter behaviour) of neighbours influence an electorate's voting behaviour.
-Spatial weights are calculated by allowing each electorate to be equally correlated with any electorate that shares any part of its boundary.
+The spatial error model [@Anselin88] is chosen because captures spatial heterogeneity by incorporating a spatially structured random effect vector [@LeSage2009].
+In this context, the random effect can be thought of as capturing the unobserved political climate in each electorate, where the climate is correlated with the climate in neighbouring electorates.
+This functions under the assumption that the climate is independent of electoral socio-demographics, and that an electorate is equally correlated with any electorate that shares a part of its boundary.
+Spatial weights are calculated in accordance with these assumptions.
 The spatial error model is specified as follows:
 
-Let $\rho$ be the parameter governing the extent of spatial correlation, $\boldsymbol v$ be a spherical error term, ${\bf W}$ be a matrix of spatial weights (containing information about the neighbouring regions), $\bf X$ be a matrix of socio-demographic covariates, $\boldsymbol \beta$ be a vector of regression coefficients and ${\boldsymbol u}$ be a vector of spatially correlated errors.
+Let $\rho$ be spatial autoregressive coefficient, $\boldsymbol v$ be a spherical error term, ${\boldsymbol W}$ be a matrix of spatial weights (containing information about the neighbouring regions), $\boldsymbol X$ be a matrix of socio-demographic covariates, $\boldsymbol \beta$ be a vector of regression coefficients and $\boldsymbol a$ be a spatially structured random effect vector.
 
-$${\bf y} = {\bf X} {\boldsymbol \beta} + {\boldsymbol u}$$
+$${\boldsymbol y} = {\boldsymbol X} {\boldsymbol \beta} + {\boldsymbol a}$$
+and
 
-with
-
-$${\boldsymbol u} = \rho {\boldsymbol W} {\boldsymbol u} + {\boldsymbol v}$$
+$${\boldsymbol a} = \rho {\boldsymbol W} {\boldsymbol a} + {\boldsymbol v}$$
 where
@@
 
-$${\bf y} = {\bf X} {\boldsymbol \beta} + ({\bf I}_n-\rho {\bf W})^{-1}{\boldsymbol v}$$
+$${\boldsymbol y} = {\boldsymbol X} {\boldsymbol \beta} + ({\boldsymbol I}_n-\rho {\boldsymbol W})^{-1}{\boldsymbol v}$$
 
-Estimation is done via maximum likelihood (using the `spdep` `R` package [@spdep]).
+Estimation is done using generalized least squares (using the `spdep` `R` package [@spdep]).
 
-----------------------------------------------------------------------------------------
+Table 4.1 details the resultant estimated model coefficients and their estimated standard errors for each of the six elections. These are interpreted in the next section.
 
-## Modelling
-- Address the issue of changing variable scales (rent as example)
-- Dimension reduction via PCA to group variables together
-- Choosing an appropriate model - discuss the spatial element
-- Akaike weights to derive a variable set for models
-- Final model specification
-
-## Selecting a superset
-
-The second stage of the data reduction process involves the selection of a superset of predictors that captures the five *most important* variables from each election.
-To determine the five most important variables for each election, Akaike weights are used to measure relative variable importance [@BurnhamAnderson2002].
-This method is commonly used in ecological studies and is fundamentally based on the Akaike information criterion (AIC) [@Akaike73].
-Akaike weights are produced using a method of scoring each model relative to a collection of models $M$ using AIC.
-
-Let $\Delta_m = AIC_m - AIC_{min}$ denote the difference in AIC between models $m$ and that with minimum AIC in the model set $M$ for $m = 1,2,...,R$.
-The Akaike weights $w_m$ for model $m$ is calculated as
-
-$$w_m = \frac {\exp(-\frac{1}{2}\Delta_m)} {\sum_{r=1} ^R \exp(-\frac{1}{2}\Delta_r)},$$
-for $m = 1,...,R$.
-
-Akaike weights represent the posterior probability that model $m$ is the best model in the set, without imposing any beliefs a priori, as $w_m \in (0,1)$ and $w_m$ sum to 1.
-
-For each variable $j \in 1, ..., J$, the sum of the Akaike weights is computed over all models that include that variable, and denote this sum by $s_j$. That is, let
-
-$$s_j = \sum_{m=1}^{R} w_m \cdot \text{I}(\text{variable } j \text{ used in model } m),$$
-where $I(A)$ denotes the indicator function for A.
-
-Variables are ordered according to $s_j$, which may be used as a measure of variable importance [@BurnhamAnderson2002].
-The variable with largest $s_j$ is deemed to be the most important in the set for the purpose of modelling the response variable.
-
-For each election, the model set $M$ used, for which Akaike weights are constructed, is the set of all possible five variable linear models.
-This involves fitting $R = {32 \choose 5} = 201,376$ models for each election, with five representing a compromise between an undue computational burden and capturing sufficient information.
-The five most important variables are selected from each election, and used to construct a variable superset obtained by taking the union of these six subsets.
-In doing so, a variable that appears to be important in a particular election (though it may be potentially unimportant in the others) is captured, along with those that are important in multiple elections.
-This superset contains the chosen predictors to be used in all of the final election models.
-Five variables are chosen from each election to ensure that this superset is not too large.
-The resultant superset contains twelve variables.
-
-Before finalizing the common regression model to be fitted for each election year, two-way interactions are added using an iterative procedure.
-First, for each election, a model is fit using the superset of predictors as main effects only.
-Then all models with a single two-way interaction are computed, and the interaction tested using a likelihood-ratio test (1% significance level).
-The number of elections for which a given interaction is significant is tallied, and the interaction that is most frequently significant is selected to remain in the model.
-This process is repeated until no interaction is significant in more than one election.
-
-Now that model specification has been determined, the final models are fit using the superset of 12 main effects and the four two-way interactions.
-The equation, estimated separately for each year, is as follows:
-
-$\small{LNP_i = \beta_0 + \beta_1*\text{Extractive}_i + \beta_2*\text{Born_SE_Europe}_i + \beta_3*\text{ManagerAdminClericalSales}_i + \beta_4*\text{DeFacto}_i + \beta_5*\text{DiffAddress}_i + \beta_6*\text{Married}_i + \beta_7*\text{Born_UK}_i + \beta_8*\text{MedianAge}_i + \beta_9*\text{OneParent_House}_i + \beta_{10}*\text{OtherLanguageHome}_i + \beta_{11}*\text{Transformative}_i + \beta_{12}*\text{Unemployment}_i + \beta_{13}*\Big(\text{Extractive}_i*\text{OtherLanguageHome}_i\Big) + \beta_{14}*\Big(\text{DeFacto}_i*\text{OtherLanguageHome}_i\Big) + \beta_{15}*\Big(\text{Extractive}_i*\text{ManagerAdminClericalSales}_i\Big) + \beta_{16}*\Big(\text{MedianAge}_i*\text{OneParent_House}_i\Big) + \u_i}$
-
-where ${\bf u} = ({\bf I_n} - \rho \cdot W) \cdot {\boldsymbol \varepsilon}$ and $\varepsilon_i \overset{iid}\sim N(0,\sigma^2)$.
-
-Table *XXX* details the estimated model and their standard errors for each of the six elections. These are interpreted in the following section.
+*insert table here*

### 2019-03-24 _book/electoral-modelling-draft.tex
added bookdown files

\documentclass[openany]{book}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Spatial modelling of the electoral two-party preferred vote in Australia. A study of federal elections between 2001 and 2016 via the eechidna R package.},
            pdfauthor={Jeremy Forbes, Di Cook \& Rob Hyndman},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{natbib}
\bibliographystyle{apalike}
\usepackage{longtable,booktabs}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

%%% Change title format to be more compact
\usepackage{titling}

% Create subtitle command for use in maketitle
\newcommand{\subtitle}[1]{
  \posttitle{
    \begin{center}\large#1\end{center}
    }
}

\setlength{\droptitle}{-2em}

  \title{Spatial modelling of the electoral two-party preferred vote in Australia. A study of federal elections between 2001 and 2016 via the \texttt{eechidna} \texttt{R} package.}
    \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
    \author{Jeremy Forbes, Di Cook \& Rob Hyndman}
    \preauthor{\centering\large\emph}
  \postauthor{\par}
      \predate{\centering\large\emph}
  \postdate{\par}
    \date{2019-03-24}

\usepackage{booktabs}
\usepackage{amsthm}
\makeatletter
\def\thm@space@setup{%
  \thm@preskip=8pt plus 2pt minus 4pt
  \thm@postskip=\thm@preskip
}
\makeatother

\begin{document}
\maketitle

{
\setcounter{tocdepth}{1}
\tableofcontents
}
\hypertarget{intro}{%
\chapter{Introduction}\label{intro}}

Australia has changed in many ways over the last two decades. Rising house prices, country-wide improvements in education, an ageing population, and a decline in religious affiliation, are just a few facets of the country's evolving socio-demographic characteristics. At the same time, political power has moved back and forth between the two major parties. In the 2007 and 2010 federal elections, the Australian Labor Party (Labor) was victorious, whereas the 2001, 2004, 2013 and 2016 elections were won by the Liberal National coalition (Liberal). The two-party preferred vote, a measure of support between these two parties, fluctuated between 47.3\% and 53.5\% (in favour of the Liberal party) over this period. This study explores how electoral characteristics relate to two-party preference, and whether their effects have changed over time. Electoral socio-demographics are derived from the Census, and vote counts are obtained from federal elections.

Joining these two data sources is problematic as there in an inherent asynchronicity in the two events. A Census is conducted by the Australian Bureau of Statistics (ABS) every five years, whereas a federal election (conducted by the Australian Electoral Commission (AEC)) usually occurs every three years. The first problem addressed is that of obtaining appropriate Census data for the 2004, 2007, 2010 and 2013 elections - election years in which a Census does not occur. The predominant approach in previous studies is to join voting outcomes to the nearest Census, without accounting for any temporal differences (see \citet{DavisStimson98}, \citet{Stimson06}, \citet{Liao09} and \citet{Stimson09}). Furthermore, electoral boundaries change regularly, so spatial discrepancies also arise when matching electoral data. To obtain appropriate Census data for these four elections, electoral socio-demographics are imputed using a spatio-temporal imputation that combines areal interpolation \citep{Goodchild1993} and linear time-interpolation. The process of collecting and wrangling the raw data is outlined in section \ref{data}, and the imputation process is detailed in section \ref{imputation}. All data and associated documentation relating to this procedure are available in the \texttt{eechidna} \texttt{R} package \citep{eechidna}, providing a resource for future analysis.

Previous work on modelling Australian federal elections have found that aggregate socio-demographics are relatively good predictors of voting outcomes. \citet{Forrest01} does this using multiple regression of the Liberal and Labor primary vote for polling booths in the Farrer electorate in 1998. \citet{Stimson06}, \citet{Stimson09} and \citet{Stimson12} use principal component analysis of polling booths in the 2001, 2004 and 2007 elections respectively, also finding that socio-demographic characteristics of polling booths are linked to their two-party preferred vote. On the contrary, \citet{Stimson09} models the polling booth swing vote (change in the two-party preferred vote) in the 2007 election, finding that little of swing vote can be explained by Census data. Instead of analyzing a single election in isolation, this paper employs a consistent model framework across six elections so that temporal changes in the effects of socio-demographics can be observed, where each federal elections is modelled with a cross-sectional data set. The use of a regression framework to examine these socio-political relationships over time is seemingly absent from previous Australian studies. It also appears that no study has attempted any type of statistical analysis of socio-demographics in conjunction with voter behaviour in Australia since 2007, making this paper distinctly different from those previous.

The cross-sectional data set for each election consists of the two-party preferred vote (response variable), and socio-demographic variables (explanatory variables) that characterise each electorate. To obtain these cross-sections, socio-demographic variables are first standardized, and then principal components are used to group variables into ``factors''. To account for the inherent spatial structure of the data, a spatial error model is fit for each election. These steps are discussed in section \ref{modelling}. In section \ref{results} inference is conducted on the models to see which effects are significant, how effects change over time and which electorates have abnormal voting behaviour.

\hypertarget{data}{%
\chapter{Data collection}\label{data}}

\hypertarget{voting-data}{%
\section{Voting data}\label{voting-data}}

The voting outcome of interest is the electoral two-party preferred vote, which is provided by the Australian Electoral Commission (AEC) for the 2001, 2004, 2007, 2010, 2013 and 2016 elections via the AEC Tally Room. The AEC divide Australia into 150 regions called electorates, with each corresponding to a single seat in the House of Representatives. Voting is compulsory in Australia, and each voter assigns a numbered preference to each available candidate in their electorate. The two-party preferred vote is determined by a tally of these preferences where, by convention, only the ranks of the Labor and Liberal candidates are considered. This is recorded as a percentage preference in favour of the Liberal party.

\hypertarget{census-data}{%
\section{Census data}\label{census-data}}

Socio-demographic variables are derived from the Census of Population and Housing (Census), which is a survey of every household in Australia, recording information such as age, gender, ethnicity, education level and income. There have been four Censuses in the 21st century, being that in 2001, 2006, 2011 and 2016. The Australian Bureau of Statistics (ABS) conducts the Census and publishes aggregated information. The ABS approximation of electorates at the time of the Census is chosen. From this aggregate information, 67 socio-demographic variables are computed for each of the electorates.

\hypertarget{downloading-and-wrangling}{%
\section{Downloading and wrangling}\label{downloading-and-wrangling}}

Raw data is sourced online from the AEC and ABS websites in .csv and .xlsx files. The format of these files change over the years, making extracting the appropriate information a big task. The functions available in the \texttt{dplyr} \citep{dplyr} and \texttt{readxl} \citep{readxl} \texttt{R} packages are very useful, as they provide fast consistent tools for data manipulation and functions to import .xlsx files (respectively). The 2001 and 2006 Census data are however published in a format where each electorate has a separate document, making it difficult to use the \texttt{dplyr} tools and instead cells have to be selected from each individual file to construct the desired variables. All scripts required for the data wrangling process can be found in the github repository for the \texttt{eechidna} \texttt{R} package \citep{eechidna}, along with the raw data. The \texttt{eechidna} package makes this study entirely reproducible and provides a resource to help wrangle data for future Censuses and elections, when they become available.

\hypertarget{imputation}{%
\chapter{Joining Census and election data}\label{imputation}}

\hypertarget{differences-between-census-and-election-data}{%
\section{Differences between Census and election data}\label{differences-between-census-and-election-data}}

Between 2001 and 2016 there were six elections and four Censuses (see Figure \ref{fig:timeline}). Electoral boundaries are redistributed regularly by the AEC, meaning that only in the years where both a Census and election occur will the boundaries match - the case for the 2001 and 2016 election. Therefore, for the four elections between 2004 and 2013, both temporal and spatial differences in electorates need to be accounted for when joining the electoral two-party preferred vote with Census data. For these elections a spatio-temporal imputation method is employed to obtain electoral socio-demographics. This method uses Census information from both before and after the election of interest.

\begin{figure}[h]

{\centering \includegraphics[width=0.9\linewidth]{electoral-modelling-draft_files/figure-latex/timeline-1}

}

\caption{Timeline of Australian elections and Censuses. They do not always occur in the same year.}\label{fig:timeline}
\end{figure}

\hypertarget{spatio-temporal-imputation}{%
\section{Spatio-temporal imputation}\label{spatio-temporal-imputation}}

To account for spatial differences, the piece-wise approximation method in \citet{Goodchild1993} is adopted. Consider a map of source zones \(s = 1,...,S\), for which socio-demographic information is available, and a set of target zones \(t = 1,...,T\) for which information is to be imputed. In this context the map of electoral boundaries at the time of a Census would be the source zones, and the boundaries at the time of the election would be the target zones. Denote the area of intersection between source zone \(s\) and target zone \(t\) as \(A_{s,t}\), the population of the source zone \(s\) as \(U_s\), and the population of intersection between source zone \(s\) and target zone \(t\) as \(P_{s,t}\).

Compute each \(A_{s,t}\) and estimate population of the intersection:

\[\hat{P}_{s,t} = \frac{U_s*A_{s,t}}{\sum_{t=1}^T A_{s,t}}\]
This assumes that populations are uniformly distributed within each source zone.

In order to calculate socio-demographic information for each of the target zones, a weighted average is taken using the estimated population as weights. Denote a given Census variable for the target zone \(C_t\), and the same Census variable for the source zone \(D_s\):

\[\hat{C}_t = \frac{\sum_{s=1}^{S}{D_s*\hat{P}_{s,t}}}{\sum_{s=1}^{S}{\hat{P}_{s,t}}}\]
This assumes that each individual in a source zone assumes the aggregate characteristics of the zone.

Applying this to each of the target zones addresses the spatial component, as it imputes the required socio-demographic for the desired electoral boundaries. However these are applicable at the time of the Census (source year) and are not yet appropriate for the election (target year).

Denote year \(y\), with a Census falling on \(y_1\) and \(y_3\), and an election on year \(y_2\), and add this subscript to the Census variable estimate, \(\hat{C}_{t,y}\). To account for temporal changes, linear interpolation is used between Census years to get the final estimate of a Census variable for the target zone in the election year \(y_2\). This assumes that population evolves in a linear manner over time.

\[\hat{C}_{t,y_2} = \frac{y_3-y_2}{y_3-y_1}*\hat{C}_{t,y_1} + \frac{y_2-y_1}{y_3-y_1}*\hat{C}_{t,y_3}\]

\hypertarget{applied}{%
\section{Applied}\label{applied}}

Publically available Census data is aggregated and there are different resolutions accessible, ranging from SA1 (over 50,000 zones) to electoral divisions (150 zones). For this study electoral divisions are used as source zones, and this imputation method is applied for each of the 2004, 2007, 2010 and 2013 elections. To demonstrate its functionality, consider the imputation of socio-demographic variables for the electorate of Hume in New South Wales (NSW), at the time of the 2013 federal election. Figure \ref{fig:hume13} shows this region amongst other NSW electorates.

\begin{figure}[h]

{\centering \includegraphics[height=0.3\textheight]{electoral-modelling-draft_files/figure-latex/hume13-1}

}

\caption{Some of the electoral boundaries in NSW for 2013, with the electoral boundary for Hume, shown in purple.}\label{fig:hume13}
\end{figure}

The Censuses neighbouring the 2013 election are those in 2011 and 2016, and the Hume boundary is changed, as seen by plotting the Hume boundary (purple) in the 2013 election over the divisions in 2016 (see Figure \ref{fig:hume16}).

\begin{figure}[h]

{\centering \includegraphics[height=0.3\textheight]{electoral-modelling-draft_files/figure-latex/hume16-1}

}

\caption{Census division boundaries in NSW for 2016, with the 2013 electoral boundary for Hume, shown in purple. The purple region is not contained within a single Census division.}\label{fig:hume16}
\end{figure}

There are many electorates in 2016 that intersect with the purple region (Hume boundary for 2013), these include the divisions of Riverina, Eden-Monaro and Hume, along with smaller intersecting areas with Fenner, Calare, Gilmore and Whitlam. To impute Census information for this purple region, calculate the percentage of each 2016 electorate that intersects with the purple region, which is then used to estimate intersection populations \(\hat{P}_{s,t}\).

\begin{longtable}[]{@{}llll@{}}
\toprule
\begin{minipage}[b]{0.17\columnwidth}\raggedright
Electorate (2016)\strut
\end{minipage} & \begin{minipage}[b]{0.10\columnwidth}\raggedright
Percentage\strut
\end{minipage} & \begin{minipage}[b]{0.22\columnwidth}\raggedright
Population in Electorate\strut
\end{minipage} & \begin{minipage}[b]{0.40\columnwidth}\raggedright
Estimated Population Allocated to Purple Region: \(\hat{P}_{s,t}\)\strut
\end{minipage}\tabularnewline
\midrule
\endhead
\begin{minipage}[t]{0.17\columnwidth}\raggedright
HUME\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright
96.54\%\strut
\end{minipage} & \begin{minipage}[t]{0.22\columnwidth}\raggedright
150643\strut
\end{minipage} & \begin{minipage}[t]{0.40\columnwidth}\raggedright
145427\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.17\columnwidth}\raggedright
RIVERINA\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright
25.11\%\strut
\end{minipage} & \begin{minipage}[t]{0.22\columnwidth}\raggedright
155793\strut
\end{minipage} & \begin{minipage}[t]{0.40\columnwidth}\raggedright
39117\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.17\columnwidth}\raggedright
EDEN-MONARO\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright
11.09\%\strut
\end{minipage} & \begin{minipage}[t]{0.22\columnwidth}\raggedright
147532\strut
\end{minipage} & \begin{minipage}[t]{0.40\columnwidth}\raggedright
16358\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.17\columnwidth}\raggedright
CANBERRA\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright
0.28\%\strut
\end{minipage} & \begin{minipage}[t]{0.22\columnwidth}\raggedright
196037\strut
\end{minipage} & \begin{minipage}[t]{0.40\columnwidth}\raggedright
548\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.17\columnwidth}\raggedright
FENNER\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright
0.23\%\strut
\end{minipage} & \begin{minipage}[t]{0.22\columnwidth}\raggedright
202955\strut
\end{minipage} & \begin{minipage}[t]{0.40\columnwidth}\raggedright
474\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.17\columnwidth}\raggedright
WHITLAM\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright
0.06\%\strut
\end{minipage} & \begin{minipage}[t]{0.22\columnwidth}\raggedright
152280\strut
\end{minipage} & \begin{minipage}[t]{0.40\columnwidth}\raggedright
92\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.17\columnwidth}\raggedright
GILMORE\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright
0.06\%\strut
\end{minipage} & \begin{minipage}[t]{0.22\columnwidth}\raggedright
150436\strut
\end{minipage} & \begin{minipage}[t]{0.40\columnwidth}\raggedright
86\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.17\columnwidth}\raggedright
CALARE\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright
0.01\%\strut
\end{minipage} & \begin{minipage}[t]{0.22\columnwidth}\raggedright
161298\strut
\end{minipage} & \begin{minipage}[t]{0.40\columnwidth}\raggedright
21\strut
\end{minipage}\tabularnewline
\bottomrule
\end{longtable}

Now consider the socio-demographic \(AusCitizen\) - the proportion of people in the region who are Australian citizens.

\begin{longtable}[]{@{}lll@{}}
\toprule
\begin{minipage}[b]{0.18\columnwidth}\raggedright
DivisionNm\strut
\end{minipage} & \begin{minipage}[b]{0.21\columnwidth}\raggedright
AusCitizen (\%): \(D_s\)\strut
\end{minipage} & \begin{minipage}[b]{0.53\columnwidth}\raggedright
Estimated Population Allocated to Purple Region: \(\hat{P}_{s,t}\)\strut
\end{minipage}\tabularnewline
\midrule
\endhead
\begin{minipage}[t]{0.18\columnwidth}\raggedright
HUME\strut
\end{minipage} & \begin{minipage}[t]{0.21\columnwidth}\raggedright
90.02\strut
\end{minipage} & \begin{minipage}[t]{0.53\columnwidth}\raggedright
145427\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.18\columnwidth}\raggedright
RIVERINA\strut
\end{minipage} & \begin{minipage}[t]{0.21\columnwidth}\raggedright
89.11\strut
\end{minipage} & \begin{minipage}[t]{0.53\columnwidth}\raggedright
39117\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.18\columnwidth}\raggedright
EDEN-MONARO\strut
\end{minipage} & \begin{minipage}[t]{0.21\columnwidth}\raggedright
88.00\strut
\end{minipage} & \begin{minipage}[t]{0.53\columnwidth}\raggedright
16358\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.18\columnwidth}\raggedright
CANBERRA\strut
\end{minipage} & \begin{minipage}[t]{0.21\columnwidth}\raggedright
85.48\strut
\end{minipage} & \begin{minipage}[t]{0.53\columnwidth}\raggedright
548\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.18\columnwidth}\raggedright
FENNER\strut
\end{minipage} & \begin{minipage}[t]{0.21\columnwidth}\raggedright
83.64\strut
\end{minipage} & \begin{minipage}[t]{0.53\columnwidth}\raggedright
474\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.18\columnwidth}\raggedright
WHITLAM\strut
\end{minipage} & \begin{minipage}[t]{0.21\columnwidth}\raggedright
89.52\strut
\end{minipage} & \begin{minipage}[t]{0.53\columnwidth}\raggedright
92\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.18\columnwidth}\raggedright
GILMORE\strut
\end{minipage} & \begin{minipage}[t]{0.21\columnwidth}\raggedright
89.03\strut
\end{minipage} & \begin{minipage}[t]{0.53\columnwidth}\raggedright
86\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.18\columnwidth}\raggedright
CALARE\strut
\end{minipage} & \begin{minipage}[t]{0.21\columnwidth}\raggedright
87.56\strut
\end{minipage} & \begin{minipage}[t]{0.53\columnwidth}\raggedright
21\strut
\end{minipage}\tabularnewline
\bottomrule
\end{longtable}

Then taking a weighted average of \(AusCitizen\) using the estimated population as weights yields \(\hat{C}_{Hume,2016} = 89.65 \%\). Repeating this process using the 2011 Census and electoral boundaries yields \(\hat{C}_{Hume,2011} = 91.00 \%\)

Finally, linearly interpolate between 2011 and 2016 to arrive at the 2013 estimate:
\begin{eqnarray*}
\hat{C}_{Hume,2013} & = &\frac{3}{5} \cdot \hat{C}_{Hume,2011} + \frac{2}{5} \cdot \hat{C}_{Hume,2016} \\ 
& = & \frac{3}{5} \cdot 91.00 \% + \frac{2}{5} \cdot 89.65 \% \\ 
& = & 90.46 \%
\end{eqnarray*}

This is done for each of the socio-demographic variables, and repeated each of the 2013 electorates.

\hypertarget{modelling}{%
\chapter{Modelling}\label{modelling}}

\hypertarget{data-pre-processing}{%
\section{Data pre-processing}\label{data-pre-processing}}

With socio-demographic information now available for each electorate, each election is joined to the data corresponding with its two-party preferred vote. Socio-demographic variables within each election year are standardized to have mean zero and variance one, to adjust for changing variable scales. For example, inflation-adjusted median rental prices increased across almost all electorates, with median rent of 200 dollars per week placing an electorate in the 90th percentile in 2001, but only the 30th percentile in 2016.

\hypertarget{dimension-reduction}{%
\subsection{Dimension reduction}\label{dimension-reduction}}

With only \(N = 150\) observations (electorates) in each election and \(p = 65\) socio-demographic variables in each cross-section, any model using all variables would face serious problems with multi-collinearity and over-fitting, likely leading to erroneous conclusions regarding variable significance. Therefore a form of dimension reduction is adopted before models are fit.

Socio-demographic variables\footnote{A preliminary step involved removing all age bands, because age is represented by median age, and to remove variables relating to particular denominations of Christianity.} that represent similar information are combined into ``factors'' using principal component analysis (PCA). The scree plots of the principal components for each election all level off after four components, and the loadings of these four components are similar across the elections. Principal components are then computed on the combined set of socio-demographics across all six elections. A factor is created by combining several variables all have large loadings in a particular component and when there is an intuitive reason as to why these variables could represent common information. A loading with magnitude greater than 0.15 is considered large. After computing these sums, each factor is again standardized to have mean zero and variance one, within each election.

Consider the \texttt{Incomes} factor as an illustration. Independent of principal components, we may suspect that median personal income, median household income and median family income are providing similar information about the financial wellbeing of an electorate. Their loadings in the first principal component are large (0.20, 0.21 and 0.22 respectively), which provides the evidence needed to combine these variables into a single factor, which is called \texttt{Incomes}.

This process reduces the predictor set to \(p = 30\).

\hypertarget{model-framework}{%
\section{Model framework}\label{model-framework}}

An identical model specification is used across the six elections, with each election modelled separately. This allows for the socio-demographic effects to be estimated separately for each year, allowing for interpretation of temporal changes in these effects. This is preferable over a single longitudinal model because it avoids any concerns of undue bias stemming from an incorrectly imposed time-varying restriction on any variable. Without such restrictions, a pooled cross-sectional model does not yield any distinct advantage over separate cross-sections. The panel approach is avoided because of how frequently electoral boundaries change, meaning that electorates that have the same name across elections are not guaranteed to represent the same geographical region. Therefore any fixed or random effects models would be difficult to estimate without implementing consistent boundaries, which would requiring further imputation.

For each cross-section, let the response variable be the two-party preferred vote in favour of the Liberal party, denoted \(Y\), with \(Y = 70\) representing a 70\% preference for Liberal, 30\% for Labor. Although \(Y\) lies in the interval \((0,100)\), observed values are never very close to 0 or 100 (minimum \(24.05 \%\) and maximum \(74.90 \%\)), so there is no need to impose the constraint of \(Y \in [0,100]\). Furthermore, the response is found to be spatially correlated in each election (Moran's I test, \(p \le 7\cdot10^{-15}\)). This is expected, as electorates are aggregate spatial units, and hence the spatial structure of the data must modelled appropriately.

The spatial error model \citep{Anselin88} is chosen because captures spatial heterogeneity by incorporating a spatially structured random effect vector \citep{LeSage2009}. In this context, the random effect can be thought of as capturing the unobserved political climate in each electorate, where the climate is correlated with the climate in neighbouring electorates. This functions under the assumption that the climate is independent of electoral socio-demographics, and that an electorate is equally correlated with any electorate that shares a part of its boundary. Spatial weights are calculated in accordance with these assumptions. The spatial error model is specified as follows:

Let \(\rho\) be spatial autoregressive coefficient, \(\boldsymbol v\) be a spherical error term, \({\boldsymbol W}\) be a matrix of spatial weights (containing information about the neighbouring regions), \(\boldsymbol X\) be a matrix of socio-demographic covariates, \(\boldsymbol \beta\) be a vector of regression coefficients and \(\boldsymbol a\) be a spatially structured random effect vector.

\[{\boldsymbol y} = {\boldsymbol X} {\boldsymbol \beta} + {\boldsymbol a}\]
and

\[{\boldsymbol a} = \rho {\boldsymbol W} {\boldsymbol a} + {\boldsymbol v}\]
where

\[{\boldsymbol v} \sim N({\boldsymbol 0}, \sigma^2 {\boldsymbol I_n})\].

so it can be written

\[{\boldsymbol y} = {\boldsymbol X} {\boldsymbol \beta} + ({\boldsymbol I}_n-\rho {\boldsymbol W})^{-1}{\boldsymbol v}\]

Estimation is done using generalized least squares (using the \texttt{spdep} \texttt{R} package \citep{spdep}).

Table 4.1 details the resultant estimated model coefficients and their estimated standard errors for each of the six elections. These are interpreted in the next section.

\emph{insert table here}

\hypertarget{results}{%
\chapter{Results}\label{results}}

\bibliography{book.bib,packages.bib}

\end{document}

### 2019-03-24 index.Rmd
added bookdown files

--- 
title: "Spatial modelling of the electoral two-party preferred vote in Australia. A study of federal elections between 2001 and 2016 via the `eechidna` `R` package."
author: "Jeremy Forbes, Di Cook & Rob Hyndman"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
output: bookdown::gitbook
documentclass: book
classoption: openany
bibliography: [book.bib, packages.bib]
biblio-style: apalike
link-citations: yes
github-repo: rstudio/bookdown-demo
description: "This is an incomplete draft of a paper to submitted to ANZJS."
---

### 2019-03-25 01-intro.Rmd
added data

@@
 
-```{r setup, include=FALSE}
+```{r setup, include=FALSE, message=FALSE}
 knitr::opts_chunk$set(echo = F, fig.pos = 'h', fig.align = 'center')
@@
 library(knitr)
+library(nlme)
+library(rgeos)
+library(spdep)
+
+# Data
+data(tpp01)
+data(tpp04)
+data(tpp07)
+data(tpp10)
+data(tpp13)
+data(tpp16)
+data(abs2001)
+data(abs2004)
+data(abs2007)
+data(abs2010)
+data(abs2013)
+data(abs2016)
+```
+
+```{r wrangle, include = F, message = F}
+# Function to standardize variables
+standardise_vars <- function(df) {
+  hold <- df
+  num_cols <- sapply(df, class) == 'numeric'
+  df[, num_cols] <- lapply(df[, num_cols], scale)
+  names(df) <- names(hold)
+  df$LNP_Percent <- hold$LNP_Percent
+  return(df)
+}
+
+# Combine and standardize
+my_df <- bind_rows(
+  left_join(tpp01, standardise_vars(abs2001) %>% dplyr::select(-c(UniqueID, Area, ends_with("NS"), Population)), by = c("DivisionNm", "StateAb"="State")) %>% mutate(year = "2001"),
+  left_join(tpp04, standardise_vars(abs2004) %>% dplyr::select(-UniqueID), by = c("DivisionNm")) %>% mutate(year = "2004"),
+  left_join(tpp07, standardise_vars(abs2007) %>% dplyr::select(-UniqueID), by = c("DivisionNm")) %>% mutate(year = "2007"),
+  left_join(tpp10, standardise_vars(abs2010) %>% dplyr::select(-UniqueID), by = c("DivisionNm")) %>% mutate(year = "2010"),
+  left_join(tpp13, standardise_vars(abs2013) %>% dplyr::select(-UniqueID), by = c("DivisionNm")) %>% mutate(year = "2013"),
+  left_join(tpp16, standardise_vars(abs2016) %>% dplyr::select(-c(UniqueID, Area, ends_with("NS"), Population)), by = c("DivisionNm", "StateAb"="State")) %>% mutate(year = "2016")
+) %>% 
+  mutate(year = factor(year)) %>% 
+  dplyr::select(-c(starts_with("Age"), StateAb, LNP_Votes, ALP_Votes, ALP_Percent, TotalVotes,
+    Swing, InternetUse, InternetAccess, EnglishOnly, Other_NonChrist, OtherChrist, Volunteer, EmuneratedElsewhere, UniqueID, Catholic, Anglican))
+
+# Create final df for modelling
+
+factors_df <- my_df %>% 
+  mutate(Education = BachelorAbv + HighSchool + Professional + Finance - Laborer - Tradesperson - DipCert,
+    FamHouseSize = FamilyRatio + AverageHouseholdSize + Couple_WChild_House - Couple_NoChild_House -
+      SP_House,
+    PropertyOwned = Owned + Mortgage - Renting - PublicHousing,
+    RentLoanPrice = MedianRent + MedianLoanPay,
+    Incomes = MedianFamilyIncome + MedianHouseholdIncome + MedianPersonalIncome,
+    Unemployment = Unemployed - LFParticipation) %>% 
+  dplyr::select(-c(BachelorAbv, HighSchool, Professional, Finance, Laborer, Tradesperson, DipCert, FamilyRatio,
+    AverageHouseholdSize, Couple_WChild_House, Couple_NoChild_House, SP_House, Owned, Mortgage, Renting,
+    PublicHousing, MedianFamilyIncome, MedianHouseholdIncome, MedianPersonalIncome, MedianRent, 
+    MedianLoanPay, Unemployed, LFParticipation))
+
+# Now standardize factors
+
+small_df <- bind_rows(
+  factors_df %>% filter(year == "2001") %>% standardise_vars(),
+  factors_df %>% filter(year == "2004") %>% standardise_vars(),
+  factors_df %>% filter(year == "2007") %>% standardise_vars(),
+  factors_df %>% filter(year == "2010") %>% standardise_vars(),
+  factors_df %>% filter(year == "2013") %>% standardise_vars(),
+  factors_df %>% filter(year == "2016") %>% standardise_vars()
+)
+
+# Order electorates in alphabetical order to match spatial matrix
+model_df <- small_df %>% 
+  arrange(year, DivisionNm)
+```
+
+```{r spatialweights, eval = F}
+# Compute spatial weights matrix
+sp_weights_matrix <- function(sF) {
+
+dist_matrix <- function(shapefile) {
+  dist_mat <- matrix(NA, nrow = 150, ncol = 150)
+  rownames(dist_mat) <- sort(shapefile$elect_div)
+  colnames(dist_mat) <- sort(shapefile$elect_div)
+
+  for (i in 1:(nrow(dist_mat)-1)) {
+    rname = rownames(dist_mat)[i]
+    row_poly = shapefile %>% subset(elect_div == rname)
+
+    for (j in (i+1):ncol(dist_mat)) {
+
+      cname = rownames(dist_mat)[j]
+      col_poly = shapefile %>% subset(elect_div == cname)
+      dist = gDistance(row_poly, col_poly)
+      dist_mat[i,j] = dist
+
+    }
+    print(i)
+  }
+
+  # Now copy to lower triange
+  for (i in 2:nrow(dist_mat)) {
+    for (j in 1:(i-1)) {
+      dist_mat[i,j] = dist_mat[j,i]
+    }
+  }
+
+  # Check it is symmetric
+  if(!isSymmetric(dist_mat)) {
+    print("Warning! Matrix is not symmetric. Error has occured.")
+  }
+
+  return(dist_mat)
+}
+
+my_dist_mat <- dist_matrix(sF)
+
+# S matrix
+
+s_mat <- function(dist_mat) {
+  s_mat <- dist_mat
+
+  for (i in 1:nrow(dist_mat)) {
+    for (j in 1:nrow(dist_mat)) {
+      a = dist_mat[i,j]
+
+      if (is.na(a)) {
+        b = 0
+      } else {
+        b = ifelse(a == 0, 1, 0)
+      }
+
+      s_mat[i,j] = b
+
+    }
+  }
+
+  return(s_mat)
+}
+
+my_smat <- s_mat(my_dist_mat)
+
+# Turn into W matrix
+
+w_mat <- function(s_mat) {
+
+  w_mat <- s_mat
+  rsums = rowSums(s_mat)
+
+  for (i in 1:nrow(s_mat)) {
+    w_mat[i,] <- s_mat[i,]/rsums[i]
+  }
+
+  return(w_mat)
+}
+
+my_wmat <- w_mat(my_smat)
+
+# Turn into listw
+
+my_listw <- mat2listw(my_wmat)
+
+return(my_listw)
+
+}
+
+# Get spatial weights
+sF_16 <- sF_download(2016)
+sF_13 <- sF_download(2013)
+sF_10 <- sF_download(2010)
+sF_07 <- sF_download(2007)
+sF_04 <- sF_download(2004)
+sF_01 <- sF_download(2001)
+
+sp_weights_16 <- sp_weights_matrix(sF_16)
+sp_weights_13 <- sp_weights_matrix(sF_13)
+sp_weights_10 <- sp_weights_matrix(sF_10)
+sp_weights_07 <- sp_weights_matrix(sF_07)
+sp_weights_04 <- sp_weights_matrix(sF_04)
+sp_weights_01 <- sp_weights_matrix(sF_01)
+```
+
+```{r FGLS, include = F}
+# Instead of running spatial weights
+load("/Users/Jeremy/Documents/R/eechidna-modelling/data/sp_weights_01.rda")
+load("/Users/Jeremy/Documents/R/eechidna-modelling/data/sp_weights_04.rda")
+load("/Users/Jeremy/Documents/R/eechidna-modelling/data/sp_weights_07.rda")
+load("/Users/Jeremy/Documents/R/eechidna-modelling/data/sp_weights_10.rda")
+load("/Users/Jeremy/Documents/R/eechidna-modelling/data/sp_weights_13.rda")
+load("/Users/Jeremy/Documents/R/eechidna-modelling/data/sp_weights_16.rda")
+
+# Function for FGLS
+my_fgls <- function(my_formula, my_data, sp_weights) {
+  # Spatial weights matrix
+  w_mat <- listw2mat(sp_weights)
+
+  # Get OLS residuals
+  ols_model <- lm(my_formula, my_data)
+  my_res <- ols_model$residuals
+
+  # Solve for rho
+  res_model <- lm(my_res ~ w_mat%*%my_res)
+  rho <- res_model$coefficients[2]
+  rho_df <- data.frame(estimate = rho, se = summary(res_model)$coefficients[2,2], p = summary(res_model)$coefficients[2,4])
+
+  # Transform data for GLS
+  trans_mat <- diag(nrow(my_data)) - rho*w_mat
+  gls_data <- data.frame(LNP_Percent = trans_mat %*% my_data$LNP_Percent,
+    Intercept = trans_mat %*% rep(1,nrow(my_data))) %>% 
+    bind_cols(as.data.frame(trans_mat %*% as.matrix(my_data %>% select(-LNP_Percent))))
+
+  # GLS model
+  my_formula <- formula(paste0(my_formula,  " - 1"))
+  gls_model <- gls(my_formula, gls_data)
+  gls_model$rho_df <- rho_df
+  gls_model$gls_data <- gls_data
+  gls_model$my_data <- my_data
+
+  return(gls_model)
+}
+
+## Run full models for each year
+full_formula = "LNP_Percent ~ ."
+
+# 2016
+glsmod16 <- my_fgls(full_formula, 
+  my_data = model_df %>% filter(year == "2016") %>% select(-c(DivisionNm, year)),
+  sp_weights = sp_weights_16)
+
+# 2013
+glsmod13 <- my_fgls(full_formula, 
+  my_data = model_df %>% filter(year == "2013") %>% select(-c(DivisionNm, year)),
+  sp_weights = sp_weights_13)
+
+# 2010
+glsmod10 <- my_fgls(full_formula, 
+  my_data = model_df %>% filter(year == "2010") %>% select(-c(DivisionNm, year)),
+  sp_weights = sp_weights_10)
+
+# 2007
+glsmod07 <- my_fgls(full_formula, 
+  my_data = model_df %>% filter(year == "2007") %>% select(-c(DivisionNm, year)),
+  sp_weights = sp_weights_07)
+
+# 2004
+glsmod04 <- my_fgls(full_formula, 
+  my_data = model_df %>% filter(year == "2004") %>% select(-c(DivisionNm, year)),
+  sp_weights = sp_weights_04)
+
+# 2001
+glsmod01 <- my_fgls(full_formula, 
+  my_data = model_df %>% filter(year == "2001") %>% select(-c(DivisionNm, year)),
+  sp_weights = sp_weights_01)
+
+## Visualise coefficients and significance
+
+coef_df <- bind_rows(
+  data.frame(variable = glsmod16$coefficients %>% names, estimate = glsmod16$coefficients %>% unname, se = summary(glsmod16)$tTable[, "Std.Error"] %>% unname, p = summary(glsmod16)$tTable[, "p-value"] %>% unname, year = 2016),
+  data.frame(variable = glsmod13$coefficients %>% names, estimate = glsmod13$coefficients %>% unname, se = summary(glsmod13)$tTable[, "Std.Error"] %>% unname, p = summary(glsmod13)$tTable[, "p-value"] %>% unname, year = 2013),
+  data.frame(variable = glsmod10$coefficients %>% names, estimate = glsmod10$coefficients %>% unname, se = summary(glsmod10)$tTable[, "Std.Error"] %>% unname, p = summary(glsmod10)$tTable[, "p-value"] %>% unname, year = 2010),
+  data.frame(variable = glsmod07$coefficients %>% names, estimate = glsmod07$coefficients %>% unname, se = summary(glsmod07)$tTable[, "Std.Error"] %>% unname, p = summary(glsmod07)$tTable[, "p-value"] %>% unname, year = 2007),
+  data.frame(variable = glsmod04$coefficients %>% names, estimate = glsmod04$coefficients %>% unname, se = summary(glsmod04)$tTable[, "Std.Error"] %>% unname, p = summary(glsmod04)$tTable[, "p-value"] %>% unname, year = 2004),
+  data.frame(variable = glsmod01$coefficients %>% names, estimate = glsmod01$coefficients %>% unname, se = summary(glsmod01)$tTable[, "Std.Error"] %>% unname, p = summary(glsmod01)$tTable[, "p-value"] %>% unname, year = 2001)
+)
 ```

### 2019-03-25 04-modelling.Rmd
latex table with models

@@
 
-Estimation is done using generalized least squares (using the `spdep` `R` package [@spdep]).
+Estimation is done using feasible generalized least squares.
 
@@
 
-*insert table here*
+```{r printmodels, eval = F}
+# Add rho and make adjustments to this
+temp_fgls <- function(my_formula, my_data, sp_weights) {
+  # Spatial weights matrix
+  w_mat <- listw2mat(sp_weights)
+
+  # Get OLS residuals
+  ols_model <- lm(my_formula, my_data)
+  my_res <- ols_model$residuals
+
+  # Solve for rho
+  res_model <- lm(my_res ~ w_mat%*%my_res)
+  rho <- res_model$coefficients[2]
+  rho_df <- data.frame(estimate = rho, se = summary(res_model)$coefficients[2,2], p = summary(res_model)$coefficients[2,4])
+
+  # Transform data for GLS
+  trans_mat <- diag(nrow(my_data)) - rho*w_mat
+  gls_data <- data.frame(LNP_Percent = trans_mat %*% my_data$LNP_Percent,
+    Intercept = trans_mat %*% rep(1,nrow(my_data))) %>% 
+    bind_cols(as.data.frame(trans_mat %*% as.matrix(my_data %>% select(-LNP_Percent))))
+
+  # GLS model
+  my_formula <- formula(paste0(my_formula,  " - 1"))
+  gls_model <- gls(my_formula, gls_data)
+
+  return(gls_model)
+}
+
+# Models
+full_formula = "LNP_Percent ~ ."
+
+# 2016
+glsmod16 <- temp_fgls(full_formula, 
+  my_data = model_df %>% filter(year == "2016") %>% select(-c(DivisionNm, year)),
+  sp_weights = sp_weights_16)
+
+# 2013
+glsmod13 <- temp_fgls(full_formula, 
+  my_data = model_df %>% filter(year == "2013") %>% select(-c(DivisionNm, year)),
+  sp_weights = sp_weights_13)
+
+# 2010
+glsmod10 <- temp_fgls(full_formula, 
+  my_data = model_df %>% filter(year == "2010") %>% select(-c(DivisionNm, year)),
+  sp_weights = sp_weights_10)
+
+# 2007
+glsmod07 <- temp_fgls(full_formula, 
+  my_data = model_df %>% filter(year == "2007") %>% select(-c(DivisionNm, year)),
+  sp_weights = sp_weights_07)
+
+# 2004
+glsmod04 <- temp_fgls(full_formula, 
+  my_data = model_df %>% filter(year == "2004") %>% select(-c(DivisionNm, year)),
+  sp_weights = sp_weights_04)
+
+# 2001
+glsmod01 <- temp_fgls(full_formula, 
+  my_data = model_df %>% filter(year == "2001") %>% select(-c(DivisionNm, year)),
+  sp_weights = sp_weights_01)
+
+# Stargazer
+stargazer::stargazer(glsmod16, glsmod13, glsmod10, glsmod07, glsmod04, glsmod01, header = FALSE, 
+  column.sep.width = "1pt", 
+  title = "Estimated model using the final set of predictor variables for each of the six elections.",
+  dep.var.labels = "Two-party preferred vote in favor of the Liberal party", 
+  font.size = "scriptsize", df = FALSE, digits = 2, 
+  column.labels = c("2001", "2004", "2007", "2010", "2013", "2016"),
+  notes = c("Estimated coefficients for variable named in column one shown for", "election year indicated by column heading, with estimated standard", "deviation for each coefficient shown below in parenthesis.
+Overall", "summary measures for each regression equation are provided in the", "bottom panel."),
+  notes.align = "l")
+```
+
+\begin{table}[!htbp] \centering 
+  \caption{Estimated model using the final set of predictor variables for each of the six elections.} 
+  \label{} 
+\scriptsize 
+\begin{tabular}{@{\extracolsep{1pt}}lcccccc} 
+\\[-1.8ex]\hline 
+\hline \\[-1.8ex] 
+ & \multicolumn{6}{c}{\textit{Dependent variable:}} \\ 
+\cline{2-7} 
+\\[-1.8ex] & \multicolumn{6}{c}{Two-party preferred vote in favor of the Liberal party} \\ 
+ & 2001 & 2004 & 2007 & 2010 & 2013 & 2016 \\ 
+\\[-1.8ex] & (1) & (2) & (3) & (4) & (5) & (6)\\ 
+\hline \\[-1.8ex] 
+ $\rho$ & 0.48$^{***}$ & 0.29$^{*}$ & 0.15 & $-$0.04 & 0.15 & 0.44$^{***}$ \\ 
+  & (0.16) & (0.16) & (0.17) & (0.19) & (0.17) & (0.15) \\ 
+  & & & & & & \\ 
+ AusCitizen & $-$1.53 & $-$0.87 & 1.00 & 0.67 & $-$2.83 & $-$2.77 \\ 
+  & (2.95) & (2.90) & (2.93) & (2.57) & (2.63) & (2.51) \\ 
+  & & & & & & \\ 
+ Born\_Asia & $-$2.13 & $-$2.56 & $-$7.58$^{***}$ & $-$1.37 & 0.74 & 2.63 \\ 
+  & (2.20) & (2.58) & (2.68) & (1.96) & (2.55) & (3.03) \\ 
+  & & & & & & \\ 
+ Born\_MidEast & $-$0.82 & 1.70 & 1.79 & 2.92$^{*}$ & 3.22$^{*}$ & 1.92 \\ 
+  & (1.56) & (1.54) & (1.54) & (1.51) & (1.78) & (1.53) \\ 
+  & & & & & & \\ 
+ Born\_SE\_Europe & $-$1.63 & $-$1.25 & $-$1.72 & $-$0.20 & 0.04 & 1.02 \\ 
+  & (0.99) & (1.15) & (1.34) & (1.17) & (2.05) & (2.34) \\ 
+  & & & & & & \\ 
+ Born\_UK & $-$1.37 & $-$0.35 & $-$0.29 & 0.29 & $-$0.87 & $-$0.49 \\ 
+  & (0.98) & (1.00) & (1.05) & (0.80) & (0.90) & (1.04) \\ 
+  & & & & & & \\ 
+ BornElsewhere & 1.74 & 4.04 & 9.78$^{**}$ & 5.01 & $-$0.51 & $-$1.08 \\ 
+  & (4.52) & (4.46) & (4.47) & (3.59) & (3.67) & (3.40) \\ 
+  & & & & & & \\ 
+ Buddhism & 1.51 & 1.09 & 2.41$^{*}$ & 0.90 & 2.84$^{*}$ & 1.50 \\ 
+  & (1.24) & (1.25) & (1.40) & (1.25) & (1.61) & (1.94) \\ 
+  & & & & & & \\ 
+ Christianity & 5.17$^{**}$ & 3.08 & $-$0.60 & $-$1.93 & 3.44 & 0.53 \\ 
+  & (2.52) & (2.79) & (3.44) & (3.45) & (4.19) & (4.58) \\ 
+  & & & & & & \\ 
+ CurrentlyStudying & $-$0.09 & 1.73 & 3.28$^{***}$ & 4.03$^{***}$ & 1.69 & $-$1.14 \\ 
+  & (1.13) & (1.17) & (1.21) & (1.12) & (1.09) & (1.01) \\ 
+  & & & & & & \\ 
+ DeFacto & $-$11.02$^{***}$ & $-$9.73$^{***}$ & $-$8.60$^{***}$ & $-$6.55$^{***}$ & $-$5.04$^{**}$ & $-$6.42$^{***}$ \\ 
+  & (2.55) & (2.62) & (2.57) & (1.90) & (2.00) & (1.69) \\ 
+  & & & & & & \\ 
+ DiffAddress & 6.93$^{***}$ & 5.02$^{***}$ & 4.72$^{***}$ & 3.40$^{***}$ & 4.63$^{***}$ & 3.86$^{***}$ \\ 
+  & (1.50) & (1.71) & (1.60) & (0.88) & (1.03) & (0.95) \\ 
+  & & & & & & \\ 
+ Distributive & 1.65 & 2.76$^{**}$ & 2.88$^{**}$ & 2.25$^{**}$ & 2.75$^{**}$ & 1.67 \\ 
+  & (1.17) & (1.25) & (1.30) & (1.02) & (1.13) & (1.08) \\ 
+  & & & & & & \\ 
+ Extractive & 6.81$^{***}$ & 6.09$^{***}$ & 6.97$^{***}$ & 5.81$^{***}$ & 5.31$^{***}$ & 4.69$^{***}$ \\ 
+  & (1.31) & (1.42) & (1.48) & (1.16) & (1.23) & (1.36) \\ 
+  & & & & & & \\ 
+ Indigenous & 4.47$^{***}$ & 6.12$^{***}$ & 5.90$^{***}$ & 6.81$^{***}$ & 5.62$^{***}$ & 3.97$^{***}$ \\ 
+  & (1.38) & (1.49) & (1.49) & (1.17) & (1.35) & (1.26) \\ 
+  & & & & & & \\ 
+ Islam & 2.64 & 1.20 & $-$1.04 & $-$1.21 & 0.61 & 0.15 \\ 
+  & (1.72) & (1.71) & (1.94) & (1.80) & (2.00) & (2.03) \\ 
+  & & & & & & \\ 
+ Judaism & 1.96$^{***}$ & 1.60$^{*}$ & 0.46 & 0.01 & 1.74 & 1.02 \\ 
+  & (0.71) & (0.85) & (1.05) & (1.00) & (1.26) & (1.38) \\ 
+  & & & & & & \\ 
+ ManagerAdminClericalSales & 4.62$^{***}$ & 3.99$^{***}$ & 4.32$^{***}$ & 5.21$^{***}$ & 3.28$^{***}$ & 2.10$^{***}$ \\ 
+  & (1.03) & (0.99) & (1.00) & (0.78) & (0.85) & (0.70) \\ 
+  & & & & & & \\ 
+ Married & $-$1.39 & 1.46 & 5.07$^{*}$ & 5.06$^{*}$ & 5.50$^{**}$ & 2.27 \\ 
+  & (2.64) & (2.71) & (2.91) & (2.59) & (2.79) & (2.26) \\ 
+  & & & & & & \\ 
+ MedianAge & 2.94$^{*}$ & 2.28 & 2.08 & 1.86 & 3.73$^{**}$ & 1.54 \\ 
+  & (1.63) & (1.88) & (2.08) & (1.65) & (1.55) & (1.30) \\ 
+  & & & & & & \\ 
+ NoReligion & 5.76$^{**}$ & 3.53 & 0.25 & $-$0.31 & 3.43 & 1.06 \\ 
+  & (2.59) & (2.58) & (2.98) & (2.90) & (3.45) & (3.85) \\ 
+  & & & & & & \\ 
+ OneParent\_House & $-$0.85 & 0.11 & 0.22 & 1.66 & 1.37 & $-$1.00 \\ 
+  & (1.39) & (1.46) & (1.61) & (1.38) & (1.49) & (1.34) \\ 
+  & & & & & & \\ 
+ OtherLanguageHome & 1.27 & $-$4.69 & $-$4.37 & $-$6.02 & $-$6.76 & $-$10.40$^{**}$ \\ 
+  & (4.95) & (5.50) & (5.70) & (4.44) & (5.61) & (4.97) \\ 
+  & & & & & & \\ 
+ SocialServ & 2.90$^{***}$ & 0.94 & 1.91 & 1.77 & 1.89 & 2.42$^{*}$ \\ 
+  & (1.10) & (1.22) & (1.35) & (1.14) & (1.27) & (1.27) \\ 
+  & & & & & & \\ 
+ Transformative & 4.75$^{***}$ & 3.74$^{**}$ & 3.28$^{*}$ & 5.20$^{***}$ & 5.74$^{***}$ & 2.99$^{**}$ \\ 
+  & (1.41) & (1.57) & (1.74) & (1.47) & (1.59) & (1.41) \\ 
+  & & & & & & \\ 
+ Education & $-$3.04 & $-$0.90 & $-$1.82 & $-$0.77 & 1.94 & 0.62 \\ 
+  & (2.99) & (3.04) & (3.35) & (2.89) & (2.86) & (2.23) \\ 
+  & & & & & & \\ 
+ FamHouseSize & $-$4.35 & $-$6.16$^{*}$ & $-$8.22$^{**}$ & $-$9.67$^{***}$ & $-$5.11$^{*}$ & $-$2.94 \\ 
+  & (2.86) & (3.15) & (3.28) & (2.66) & (2.78) & (2.27) \\ 
+  & & & & & & \\ 
+ PropertyOwned & 1.76 & 0.68 & $-$1.17 & 0.78 & $-$0.92 & $-$0.76 \\ 
+  & (1.38) & (1.43) & (1.49) & (1.09) & (1.21) & (1.19) \\ 
+  & & & & & & \\ 
+ RentLoanPrice & $-$0.70 & 1.01 & 2.12 & $-$0.54 & $-$1.54 & $-$0.31 \\ 
+  & (2.02) & (1.96) & (1.96) & (1.48) & (1.68) & (1.50) \\ 
+  & & & & & & \\ 
+ Incomes & 8.46$^{***}$ & 4.19 & 2.73 & 4.36$^{*}$ & 2.95 & 2.99$^{*}$ \\ 
+  & (2.57) & (2.68) & (3.06) & (2.56) & (2.60) & (1.76) \\ 
+  & & & & & & \\ 
+ Unemployment & 0.36 & $-$1.22 & $-$2.42 & $-$2.88$^{**}$ & $-$5.09$^{***}$ & $-$3.00$^{**}$ \\ 
+  & (1.42) & (1.54) & (1.73) & (1.41) & (1.51) & (1.28) \\ 
+  & & & & & & \\ 
+ Constant & 50.45$^{***}$ & 53.48$^{***}$ & 49.92$^{***}$ & 47.32$^{***}$ & 52.63$^{***}$ & 50.78$^{***}$ \\ 
+  & (0.74) & (0.56) & (0.51) & (0.35) & (0.45) & (0.67) \\ 
+  & & & & & & \\ 
+\hline \\[-1.8ex] 
+Observations & 150 & 150 & 150 & 150 & 150 & 150 \\ 
+Log Likelihood & $-$399.52 & $-$405.41 & $-$418.08 & $-$398.73 & $-$403.34 & $-$398.33 \\ 
+Akaike Inf. Crit. & 863.03 & 874.83 & 900.16 & 861.46 & 870.68 & 860.67 \\ 
+Bayesian Inf. Crit. & 951.96 & 963.76 & 989.10 & 950.40 & 959.61 & 949.60 \\ 
+\hline 
+\hline \\[-1.8ex] 
+\textit{Note:}  & \multicolumn{6}{l}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
+ & \multicolumn{6}{l}{Estimated coefficients for variable named in column one shown for} \\ 
+ & \multicolumn{6}{l}{election year indicated by column heading, with estimated standard} \\ 
+ & \multicolumn{6}{l}{deviation for each coefficient shown below in parenthesis. Overall} \\ 
+ & \multicolumn{6}{l}{summary measures for each regression equation are provided in the} \\ 
+ & \multicolumn{6}{l}{bottom panel.} \\
+\end{tabular} 
+\end{table}

### 2019-03-25 01-intro.Rmd
cache

@@
 ```{r setup, include=FALSE, message=FALSE}
-knitr::opts_chunk$set(echo = F, fig.pos = 'h', fig.align = 'center')
+knitr::opts_chunk$set(echo = F, fig.pos = 'h', fig.align = 'center', cache = T)
 library(tidyverse)

### 2019-03-25 01-intro.Rmd
update gls to function with stargazer,

--- 2019-03-25 01-intro.Rmd
@@
 ```{r setup, include=FALSE, message=FALSE}
-knitr::opts_chunk$set(echo = F, fig.pos = 'h', fig.align = 'center')
+knitr::opts_chunk$set(echo = F, fig.pos = 'h', fig.align = 'center', cache = T, message = F)
 library(tidyverse)
@@
 model_df <- small_df %>% 
-  arrange(year, DivisionNm)
+  arrange(year, DivisionNm) %>% 
+  dplyr::select(order(colnames(.)))
 ```
@@
   gls_model <- gls(my_formula, gls_data)
+
+  # Call to function with stargazer
+  gls_model$call$model <- formula(paste0("LNP_Percent ~ ", paste0(names(gls_data)[-1], collapse = " + ")))
+
+  # Rho and data
   gls_model$rho_df <- rho_df

### 2019-03-25 04-modelling.Rmd
stargazer table fits 1 page

@@
+---
+output:
+  pdf_document: default
+  html_document: default
+---
 # Modelling {#modelling}
@@
 # Add rho and make adjustments to this
-temp_fgls <- function(my_formula, my_data, sp_weights) {
-  # Spatial weights matrix
-  w_mat <- listw2mat(sp_weights)
-
-  # Get OLS residuals
-  ols_model <- lm(my_formula, my_data)
-  my_res <- ols_model$residuals
-
-  # Solve for rho
-  res_model <- lm(my_res ~ w_mat%*%my_res)
-  rho <- res_model$coefficients[2]
-  rho_df <- data.frame(estimate = rho, se = summary(res_model)$coefficients[2,2], p = summary(res_model)$coefficients[2,4])
-
-  # Transform data for GLS
-  trans_mat <- diag(nrow(my_data)) - rho*w_mat
-  gls_data <- data.frame(LNP_Percent = trans_mat %*% my_data$LNP_Percent,
-    Intercept = trans_mat %*% rep(1,nrow(my_data))) %>% 
-    bind_cols(as.data.frame(trans_mat %*% as.matrix(my_data %>% select(-LNP_Percent))))
-
-  # GLS model
-  my_formula <- formula(paste0(my_formula,  " - 1"))
-  gls_model <- gls(my_formula, gls_data)
-
-  return(gls_model)
-}
-
-# Models
-full_formula = "LNP_Percent ~ ."
-
-# 2016
-glsmod16 <- temp_fgls(full_formula, 
-  my_data = model_df %>% filter(year == "2016") %>% select(-c(DivisionNm, year)),
-  sp_weights = sp_weights_16)
-
-# 2013
-glsmod13 <- temp_fgls(full_formula, 
-  my_data = model_df %>% filter(year == "2013") %>% select(-c(DivisionNm, year)),
-  sp_weights = sp_weights_13)
-
-# 2010
-glsmod10 <- temp_fgls(full_formula, 
-  my_data = model_df %>% filter(year == "2010") %>% select(-c(DivisionNm, year)),
-  sp_weights = sp_weights_10)
-
-# 2007
-glsmod07 <- temp_fgls(full_formula, 
-  my_data = model_df %>% filter(year == "2007") %>% select(-c(DivisionNm, year)),
-  sp_weights = sp_weights_07)
-
-# 2004
-glsmod04 <- temp_fgls(full_formula, 
-  my_data = model_df %>% filter(year == "2004") %>% select(-c(DivisionNm, year)),
-  sp_weights = sp_weights_04)
-
-# 2001
-glsmod01 <- temp_fgls(full_formula, 
-  my_data = model_df %>% filter(year == "2001") %>% select(-c(DivisionNm, year)),
-  sp_weights = sp_weights_01)
-
-# Stargazer
 stargazer::stargazer(glsmod16, glsmod13, glsmod10, glsmod07, glsmod04, glsmod01, header = FALSE, 
   column.sep.width = "1pt", 
-  title = "Estimated model using the final set of predictor variables for each of the six elections.",
+  title = "Estimated model for each of the six elections.",
   dep.var.labels = "Two-party preferred vote in favor of the Liberal party", 
-  font.size = "scriptsize", df = FALSE, digits = 2, 
+  font.size = "scriptsize", digits = 2, 
   column.labels = c("2001", "2004", "2007", "2010", "2013", "2016"),
@@
 \begin{table}[!htbp] \centering 
-  \caption{Estimated model using the final set of predictor variables for each of the six elections.} 
+  \caption{Estimated model for each of the six elections.} 
   \label{} 
-\scriptsize 
+\tiny 
 \begin{tabular}{@{\extracolsep{1pt}}lcccccc} 
@@
   & & & & & & \\ 
+ Education & $-$3.04 & $-$0.90 & $-$1.82 & $-$0.77 & 1.94 & 0.62 \\ 
+  & (2.99) & (3.04) & (3.35) & (2.89) & (2.86) & (2.23) \\ 
+  & & & & & & \\ 
  Extractive & 6.81$^{***}$ & 6.09$^{***}$ & 6.97$^{***}$ & 5.81$^{***}$ & 5.31$^{***}$ & 4.69$^{***}$ \\ 
   & (1.31) & (1.42) & (1.48) & (1.16) & (1.23) & (1.36) \\ 
+  & & & & & & \\ 
+ FamHouseSize & $-$4.35 & $-$6.16$^{*}$ & $-$8.22$^{**}$ & $-$9.67$^{***}$ & $-$5.11$^{*}$ & $-$2.94 \\ 
+  & (2.86) & (3.15) & (3.28) & (2.66) & (2.78) & (2.27) \\ 
+  & & & & & & \\ 
+ Incomes & 8.46$^{***}$ & 4.19 & 2.73 & 4.36$^{*}$ & 2.95 & 2.99$^{*}$ \\ 
+  & (2.57) & (2.68) & (3.06) & (2.56) & (2.60) & (1.76) \\ 
   & & & & & & \\ 
@@
   & & & & & & \\ 
- SocialServ & 2.90$^{***}$ & 0.94 & 1.91 & 1.77 & 1.89 & 2.42$^{*}$ \\ 
-  & (1.10) & (1.22) & (1.35) & (1.14) & (1.27) & (1.27) \\ 
-  & & & & & & \\ 
- Transformative & 4.75$^{***}$ & 3.74$^{**}$ & 3.28$^{*}$ & 5.20$^{***}$ & 5.74$^{***}$ & 2.99$^{**}$ \\ 
-  & (1.41) & (1.57) & (1.74) & (1.47) & (1.59) & (1.41) \\ 
-  & & & & & & \\ 
- Education & $-$3.04 & $-$0.90 & $-$1.82 & $-$0.77 & 1.94 & 0.62 \\ 
-  & (2.99) & (3.04) & (3.35) & (2.89) & (2.86) & (2.23) \\ 
-  & & & & & & \\ 
- FamHouseSize & $-$4.35 & $-$6.16$^{*}$ & $-$8.22$^{**}$ & $-$9.67$^{***}$ & $-$5.11$^{*}$ & $-$2.94 \\ 
-  & (2.86) & (3.15) & (3.28) & (2.66) & (2.78) & (2.27) \\ 
-  & & & & & & \\ 
  PropertyOwned & 1.76 & 0.68 & $-$1.17 & 0.78 & $-$0.92 & $-$0.76 \\ 
@@
   & & & & & & \\ 
- Incomes & 8.46$^{***}$ & 4.19 & 2.73 & 4.36$^{*}$ & 2.95 & 2.99$^{*}$ \\ 
-  & (2.57) & (2.68) & (3.06) & (2.56) & (2.60) & (1.76) \\ 
+ SocialServ & 2.90$^{***}$ & 0.94 & 1.91 & 1.77 & 1.89 & 2.42$^{*}$ \\ 
+  & (1.10) & (1.22) & (1.35) & (1.14) & (1.27) & (1.27) \\ 
+  & & & & & & \\ 
+ Transformative & 4.75$^{***}$ & 3.74$^{**}$ & 3.28$^{*}$ & 5.20$^{***}$ & 5.74$^{***}$ & 2.99$^{**}$ \\ 
+  & (1.41) & (1.57) & (1.74) & (1.47) & (1.59) & (1.41) \\ 
   & & & & & & \\ 

### 2019-03-25 05-results.Rmd
added first bits for spatial coefficient, as well as Income and Unemployment effects.

# Results
```{r rhovis}
rho_df <- bind_rows(glsmod16$rho_df %>% mutate(year = "2016"), 
  glsmod13$rho_df %>% mutate(year = "2013"),
  glsmod10$rho_df %>% mutate(year = "2010"), 
  glsmod07$rho_df %>% mutate(year = "2007"),
  glsmod04$rho_df %>% mutate(year = "2004"), 
  glsmod01$rho_df %>% mutate(year = "2001")) %>% 
  mutate(upper95 = estimate + 1.96*se, lower95 = estimate - 1.96*se)

rho_df %>% ggplot() + 
  geom_point(aes(x = year, y = estimate, col = factor(p < 0.05)), size = 3) +
  geom_linerange(aes(x = year, ymin = lower95, ymax = upper95, col = factor(p < 0.05)), size = 1.5) + 
  geom_hline(aes(yintercept = 0), alpha = 0.5, size = 1) +
  scale_color_manual(values = c("grey50", "black")) +
  guides(col = F)
```

The spatial autoregressive coefficient $\rho$ is significant in only the 2001 and 2016 elections, meaning that in these years an electorate's political climate was affected by the attitudes of it's neighbours. Conversely, in the other four elections, it appears that after accounting for electoral socio-demographics, there is no spatial effect, meaning electorates effectively acted independently.

To investigate the socio-demographics that have a strong effect on the two-party preferred vote, partial residual plots are used. These show the direction, size and significance of an estimated effect, as the slope of the prediction line matches the estimated coefficient, and the shaded region represents a 95% confidence band. Typically the Labor party campaigns on more progressive policies, which often include tax reform that adversely affects higher income earners, and more generous social assistance programs. Although topical in the lead up to an election, incomes only significantly affected electoral two-party preference in 2016 (see Figure \@ref(fig:plotincomes)), with higher income areas associated with support for the Liberal party. In the prior elections, the estimated effect is still positive, albeit insignificant. Unemployment on the other hand, was influential in 2001, 2004 and 2007 elections(Figure \@ref(fig:plotunemploy)), with electorates with higher unemployment favouring Labor. This effect has become weaker over time, with unemployment having an estimated effect of virtually zero in 2016.

```{r myvisreg, include = F}
# Function to produce visreg style conditional plots
my_visreg <- function(my_model, sp_weights, varname, 
  nolabs = FALSE, xlimits = NULL, ylimits = NULL, year = "") {

  # Extract fitted parameters
  rho <- my_model$rho_df$estimate
  sigma <- sqrt(sum(my_model$residuals^2)/(my_model$dims$N-my_model$dims$p))

  # Spatial weights
  w_mat <- listw2mat(sp_weights)

  # Q - where u = Qe, Q = (I - pW)^-1
  q_mat <- solve(diag(my_model$dims$N) - rho*w_mat)

  # Omega - QQ'
  omega_mat <- q_mat%*%t(q_mat)

  # X
  x_mat <- my_model$my_data %>% 
    dplyr::select(-LNP_Percent) %>% 
    mutate(Intercept = 1) %>% 
    dplyr::select(Intercept, everything()) %>% 
    as.matrix()

  # Beta
  beta_mat <- my_model$coefficients

  # T value
  t = qt(0.975, nrow(my_model$gls_data)-ncol(my_model$gls_data))

  # Lambda matrix (FGLS)
  x <- round(seq(min(x_mat[, varname]), max(x_mat[, varname]), 0.025), 3)
  lambda_mat <- data.frame(matrix(0, nrow = length(x), ncol = ncol(x_mat)))
  names(lambda_mat) <- dimnames(x_mat)[[2]]
  lambda_mat[, varname] <- x
  lambda_mat$Intercept <- 1
  lambda_mat <- as.matrix(lambda_mat)

  # Confidence interval
  plot_df <- data.frame(variable = x, fitted = lambda_mat%*%beta_mat, variance = 0)

  for (i in 1:nrow(lambda_mat)) {
    lambda <- lambda_mat[i, ]
    plot_df$variance[i] = sigma^2 * t(lambda) %*% 
      solve(t(x_mat) %*% solve(omega_mat) %*% x_mat) %*% 
      lambda
  }

  plot_df <- plot_df %>% 
    mutate(upper95 = fitted + t*sqrt(variance), lower95 = fitted - t*sqrt(variance))

  # Partial residuals
  points_df <- data.frame(
    variable = my_model$my_data[, varname] %>% unname,
    part_res = (my_model$my_data$LNP_Percent - x_mat%*%my_model$coefficients) + my_model$coefficients[varname]*x_mat[, varname] + my_model$coefficients[1]
  )

  # Plot
  myplot <- ggplot(data = plot_df) + 
    geom_ribbon(aes(x = variable, ymin = lower95, ymax = upper95), fill = "grey80") + 
    geom_point(aes(x = variable, y = part_res), data = points_df, size = 0.75, col = "grey50") + 
    geom_line(aes(x = variable, y = fitted), col = "blue", size = 1) +
    #geom_hline(aes(yintercept = min(upper95)), col = "red") +
    #geom_hline(aes(yintercept = max(lower95)), col = "blue") +
    theme_bw() + 
    labs(x = varname, y = "Response") + 
    ggtitle(year) +
    theme(plot.title = element_text(face = "bold", size = 10, hjust = 0.5))

  if (nolabs == TRUE) {
    myplot <- myplot + labs(x = "", y = "")
  }

  if (!is.null(xlimits) & !is.null(ylimits)) {
    myplot <- myplot + coord_cartesian(xlim = xlimits, ylim = ylimits)
  }

  return(myplot)
}

# Grid visreg
library(gridExtra)
library(grid)
grid_visreg <- function(varname, xvar_title, xlimits, ylimits) {
  p16 <- my_visreg(glsmod16, sp_weights_16, varname = varname, nolabs = T, year = "2016", xlimits, ylimits)
  p13 <- my_visreg(glsmod13, sp_weights_13, varname = varname, nolabs = T, year = "2013", xlimits, ylimits)
  p10 <- my_visreg(glsmod10, sp_weights_10, varname = varname, nolabs = T, year = "2010", xlimits, ylimits)
  p07 <- my_visreg(glsmod07, sp_weights_07, varname = varname, nolabs = T, year = "2007", xlimits, ylimits)
  p04 <- my_visreg(glsmod04, sp_weights_04, varname = varname, nolabs = T, year = "2004", xlimits, ylimits)
  p01 <- my_visreg(glsmod01, sp_weights_01, varname = varname, nolabs = T, year = "2001", xlimits, ylimits)

  plots <- grid.arrange(p01, p04, p07, p10, p13, p16, nrow = 2, 
    left = textGrob("Two-party preferred vote (%)", gp = gpar(cex = 0.8), rot = 90), 
    bottom = textGrob(xvar_title, gp = gpar(cex = 0.8)))

  return(plots)
}

```

```{r plotincomes, fig.cap = "Effect of Incomes"}
p <- grid_visreg("Incomes", xvar_title = "Incomes", xlimits = c(-1.5, 3.6), ylimits = c(30, 95))
```

```{r plotunemploy, fig.cap = "Effect of Unemployment"}
p <- grid_visreg("Unemployment", xvar_title = "Unemployment", xlimits = c(-2.1, 3.4), ylimits = c(28, 68))
```

This section summarises the main findings made about the relationships between socio-demographic chracteristics and voting behaviour.

```{r}
coef_df %>% 
  filter(variable != "Intercept") %>% 
  mutate(upper95 = estimate + 1.96*se, lower95 = estimate - 1.96*se) %>% 
  ggplot() +
  geom_point(aes(x = year, y = estimate, col = factor(p < 0.05)), size = 2) +
  geom_linerange(aes(x = year, ymin = lower95, ymax = upper95, col = factor(p < 0.05)), size = 1) + 
  geom_hline(aes(yintercept = 0), alpha = 0.5, size = 1) + 
  facet_wrap(~variable, scales = "free", ncol = 5) +
  scale_color_manual(values = c("grey50", "black")) + 
  theme(axis.text = element_text(size = 6), strip.text = element_text(size = 6)) +
  labs(x = "", y = "Year") +
  scale_x_continuous(breaks = c(2001, 2004, 2007, 2010, 2013, 2016)) +
  guides(col = F)
```

### 2019-03-25 _book/electoral-modelling-draft.tex
bookdown files

@@
   \postdate{\par}
-    \date{2019-03-24}
+    \date{2019-03-25}
 
@@
 \makeatother
+\usepackage{longtable}
 
@@
 
-Estimation is done using generalized least squares (using the \texttt{spdep} \texttt{R} package \citep{spdep}).
+Estimation is done using feasible generalized least squares.
 
@@
 
-\emph{insert table here}
+\begin{table}[!htbp] \centering 
+  \caption{Estimated model for each of the six elections.} 
+  \label{} 
+\tiny 
+\begin{tabular}{@{\extracolsep{1pt}}lcccccc} 
+\\[-1.8ex]\hline 
+\hline \\[-1.8ex] 
+ & \multicolumn{6}{c}{\textit{Dependent variable:}} \\ 
+\cline{2-7} 
+\\[-1.8ex] & \multicolumn{6}{c}{Two-party preferred vote in favor of the Liberal party} \\ 
+ & 2001 & 2004 & 2007 & 2010 & 2013 & 2016 \\ 
+\\[-1.8ex] & (1) & (2) & (3) & (4) & (5) & (6)\\ 
+\hline \\[-1.8ex] 
+ $\rho$ & 0.48$^{***}$ & 0.29$^{*}$ & 0.15 & $-$0.04 & 0.15 & 0.44$^{***}$ \\ 
+  & (0.16) & (0.16) & (0.17) & (0.19) & (0.17) & (0.15) \\ 
+  & & & & & & \\ 
+ AusCitizen & $-$1.53 & $-$0.87 & 1.00 & 0.67 & $-$2.83 & $-$2.77 \\ 
+  & (2.95) & (2.90) & (2.93) & (2.57) & (2.63) & (2.51) \\ 
+  & & & & & & \\ 
+ Born\_Asia & $-$2.13 & $-$2.56 & $-$7.58$^{***}$ & $-$1.37 & 0.74 & 2.63 \\ 
+  & (2.20) & (2.58) & (2.68) & (1.96) & (2.55) & (3.03) \\ 
+  & & & & & & \\ 
+ Born\_MidEast & $-$0.82 & 1.70 & 1.79 & 2.92$^{*}$ & 3.22$^{*}$ & 1.92 \\ 
+  & (1.56) & (1.54) & (1.54) & (1.51) & (1.78) & (1.53) \\ 
+  & & & & & & \\ 
+ Born\_SE\_Europe & $-$1.63 & $-$1.25 & $-$1.72 & $-$0.20 & 0.04 & 1.02 \\ 
+  & (0.99) & (1.15) & (1.34) & (1.17) & (2.05) & (2.34) \\ 
+  & & & & & & \\ 
+ Born\_UK & $-$1.37 & $-$0.35 & $-$0.29 & 0.29 & $-$0.87 & $-$0.49 \\ 
+  & (0.98) & (1.00) & (1.05) & (0.80) & (0.90) & (1.04) \\ 
+  & & & & & & \\ 
+ BornElsewhere & 1.74 & 4.04 & 9.78$^{**}$ & 5.01 & $-$0.51 & $-$1.08 \\ 
+  & (4.52) & (4.46) & (4.47) & (3.59) & (3.67) & (3.40) \\ 
+  & & & & & & \\ 
+ Buddhism & 1.51 & 1.09 & 2.41$^{*}$ & 0.90 & 2.84$^{*}$ & 1.50 \\ 
+  & (1.24) & (1.25) & (1.40) & (1.25) & (1.61) & (1.94) \\ 
+  & & & & & & \\ 
+ Christianity & 5.17$^{**}$ & 3.08 & $-$0.60 & $-$1.93 & 3.44 & 0.53 \\ 
+  & (2.52) & (2.79) & (3.44) & (3.45) & (4.19) & (4.58) \\ 
+  & & & & & & \\ 
+ CurrentlyStudying & $-$0.09 & 1.73 & 3.28$^{***}$ & 4.03$^{***}$ & 1.69 & $-$1.14 \\ 
+  & (1.13) & (1.17) & (1.21) & (1.12) & (1.09) & (1.01) \\ 
+  & & & & & & \\ 
+ DeFacto & $-$11.02$^{***}$ & $-$9.73$^{***}$ & $-$8.60$^{***}$ & $-$6.55$^{***}$ & $-$5.04$^{**}$ & $-$6.42$^{***}$ \\ 
+  & (2.55) & (2.62) & (2.57) & (1.90) & (2.00) & (1.69) \\ 
+  & & & & & & \\ 
+ DiffAddress & 6.93$^{***}$ & 5.02$^{***}$ & 4.72$^{***}$ & 3.40$^{***}$ & 4.63$^{***}$ & 3.86$^{***}$ \\ 
+  & (1.50) & (1.71) & (1.60) & (0.88) & (1.03) & (0.95) \\ 
+  & & & & & & \\ 
+ Distributive & 1.65 & 2.76$^{**}$ & 2.88$^{**}$ & 2.25$^{**}$ & 2.75$^{**}$ & 1.67 \\ 
+  & (1.17) & (1.25) & (1.30) & (1.02) & (1.13) & (1.08) \\ 
+  & & & & & & \\ 
+ Education & $-$3.04 & $-$0.90 & $-$1.82 & $-$0.77 & 1.94 & 0.62 \\ 
+  & (2.99) & (3.04) & (3.35) & (2.89) & (2.86) & (2.23) \\ 
+  & & & & & & \\ 
+ Extractive & 6.81$^{***}$ & 6.09$^{***}$ & 6.97$^{***}$ & 5.81$^{***}$ & 5.31$^{***}$ & 4.69$^{***}$ \\ 
+  & (1.31) & (1.42) & (1.48) & (1.16) & (1.23) & (1.36) \\ 
+  & & & & & & \\ 
+ FamHouseSize & $-$4.35 & $-$6.16$^{*}$ & $-$8.22$^{**}$ & $-$9.67$^{***}$ & $-$5.11$^{*}$ & $-$2.94 \\ 
+  & (2.86) & (3.15) & (3.28) & (2.66) & (2.78) & (2.27) \\ 
+  & & & & & & \\ 
+ Incomes & 8.46$^{***}$ & 4.19 & 2.73 & 4.36$^{*}$ & 2.95 & 2.99$^{*}$ \\ 
+  & (2.57) & (2.68) & (3.06) & (2.56) & (2.60) & (1.76) \\ 
+  & & & & & & \\ 
+ Indigenous & 4.47$^{***}$ & 6.12$^{***}$ & 5.90$^{***}$ & 6.81$^{***}$ & 5.62$^{***}$ & 3.97$^{***}$ \\ 
+  & (1.38) & (1.49) & (1.49) & (1.17) & (1.35) & (1.26) \\ 
+  & & & & & & \\ 
+ Islam & 2.64 & 1.20 & $-$1.04 & $-$1.21 & 0.61 & 0.15 \\ 
+  & (1.72) & (1.71) & (1.94) & (1.80) & (2.00) & (2.03) \\ 
+  & & & & & & \\ 
+ Judaism & 1.96$^{***}$ & 1.60$^{*}$ & 0.46 & 0.01 & 1.74 & 1.02 \\ 
+  & (0.71) & (0.85) & (1.05) & (1.00) & (1.26) & (1.38) \\ 
+  & & & & & & \\ 
+ ManagerAdminClericalSales & 4.62$^{***}$ & 3.99$^{***}$ & 4.32$^{***}$ & 5.21$^{***}$ & 3.28$^{***}$ & 2.10$^{***}$ \\ 
+  & (1.03) & (0.99) & (1.00) & (0.78) & (0.85) & (0.70) \\ 
+  & & & & & & \\ 
+ Married & $-$1.39 & 1.46 & 5.07$^{*}$ & 5.06$^{*}$ & 5.50$^{**}$ & 2.27 \\ 
+  & (2.64) & (2.71) & (2.91) & (2.59) & (2.79) & (2.26) \\ 
+  & & & & & & \\ 
+ MedianAge & 2.94$^{*}$ & 2.28 & 2.08 & 1.86 & 3.73$^{**}$ & 1.54 \\ 
+  & (1.63) & (1.88) & (2.08) & (1.65) & (1.55) & (1.30) \\ 
+  & & & & & & \\ 
+ NoReligion & 5.76$^{**}$ & 3.53 & 0.25 & $-$0.31 & 3.43 & 1.06 \\ 
+  & (2.59) & (2.58) & (2.98) & (2.90) & (3.45) & (3.85) \\ 
+  & & & & & & \\ 
+ OneParent\_House & $-$0.85 & 0.11 & 0.22 & 1.66 & 1.37 & $-$1.00 \\ 
+  & (1.39) & (1.46) & (1.61) & (1.38) & (1.49) & (1.34) \\ 
+  & & & & & & \\ 
+ OtherLanguageHome & 1.27 & $-$4.69 & $-$4.37 & $-$6.02 & $-$6.76 & $-$10.40$^{**}$ \\ 
+  & (4.95) & (5.50) & (5.70) & (4.44) & (5.61) & (4.97) \\ 
+  & & & & & & \\ 
+ PropertyOwned & 1.76 & 0.68 & $-$1.17 & 0.78 & $-$0.92 & $-$0.76 \\ 
+  & (1.38) & (1.43) & (1.49) & (1.09) & (1.21) & (1.19) \\ 
+  & & & & & & \\ 
+ RentLoanPrice & $-$0.70 & 1.01 & 2.12 & $-$0.54 & $-$1.54 & $-$0.31 \\ 
+  & (2.02) & (1.96) & (1.96) & (1.48) & (1.68) & (1.50) \\ 
+  & & & & & & \\ 
+ SocialServ & 2.90$^{***}$ & 0.94 & 1.91 & 1.77 & 1.89 & 2.42$^{*}$ \\ 
+  & (1.10) & (1.22) & (1.35) & (1.14) & (1.27) & (1.27) \\ 
+  & & & & & & \\ 
+ Transformative & 4.75$^{***}$ & 3.74$^{**}$ & 3.28$^{*}$ & 5.20$^{***}$ & 5.74$^{***}$ & 2.99$^{**}$ \\ 
+  & (1.41) & (1.57) & (1.74) & (1.47) & (1.59) & (1.41) \\ 
+  & & & & & & \\ 
+ Unemployment & 0.36 & $-$1.22 & $-$2.42 & $-$2.88$^{**}$ & $-$5.09$^{***}$ & $-$3.00$^{**}$ \\ 
+  & (1.42) & (1.54) & (1.73) & (1.41) & (1.51) & (1.28) \\ 
+  & & & & & & \\ 
+ Constant & 50.45$^{***}$ & 53.48$^{***}$ & 49.92$^{***}$ & 47.32$^{***}$ & 52.63$^{***}$ & 50.78$^{***}$ \\ 
+  & (0.74) & (0.56) & (0.51) & (0.35) & (0.45) & (0.67) \\ 
+  & & & & & & \\ 
+\hline \\[-1.8ex] 
+Observations & 150 & 150 & 150 & 150 & 150 & 150 \\ 
+Log Likelihood & $-$399.52 & $-$405.41 & $-$418.08 & $-$398.73 & $-$403.34 & $-$398.33 \\ 
+Akaike Inf. Crit. & 863.03 & 874.83 & 900.16 & 861.46 & 870.68 & 860.67 \\ 
+Bayesian Inf. Crit. & 951.96 & 963.76 & 989.10 & 950.40 & 959.61 & 949.60 \\ 
+\hline 
+\hline \\[-1.8ex] 
+\textit{Note:}  & \multicolumn{6}{l}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
+ & \multicolumn{6}{l}{Estimated coefficients for variable named in column one shown for} \\ 
+ & \multicolumn{6}{l}{election year indicated by column heading, with estimated standard} \\ 
+ & \multicolumn{6}{l}{deviation for each coefficient shown below in parenthesis. Overall} \\ 
+ & \multicolumn{6}{l}{summary measures for each regression equation are provided in the} \\ 
+ & \multicolumn{6}{l}{bottom panel.} \\ 
+\end{tabular} 
+\end{table}
 
@@
 
+\begin{center}\includegraphics{electoral-modelling-draft_files/figure-latex/rhovis-1} \end{center}
+
+The spatial autoregressive coefficient \(\rho\) is significant in only the 2001 and 2016 elections, meaning that in these years an electorate's political climate was affected by the attitudes of it's neighbours.
+Conversely, in the other four elections, it appears that after accounting for electoral socio-demographics, there is no spatial effect, meaning electorates effectively acted independently.
+
+To investigate the socio-demographics that have a strong effect on the two-party preferred vote, partial residual plots are used.
+These show the direction, size and significance of an estimated effect, as the slope of the prediction line matches the estimated coefficient, and the shaded region represents a 95\% confidence band.
+Typically the Labor party campaigns on more progressive policies, which often include tax reform that adversely affects higher income earners.
+Although extremely topical in the lead up to an election, personal, household and family incomes only affected electoral two-party preference in 2016 (see Figure \ref{fig:}).
+
+\begin{center}\includegraphics{electoral-modelling-draft_files/figure-latex/incomes-1} \end{center}
+
+This section summarises the main findings made about the relationships between socio-demographic chracteristics and voting behaviour.
+
+\begin{center}\includegraphics{electoral-modelling-draft_files/figure-latex/unnamed-chunk-2-1} \end{center}
+
 \bibliography{book.bib,packages.bib}

### 2019-03-25 electoral-modelling-draft.Rmd
bookdown files

--- 
title: "Spatial modelling of the electoral two-party preferred vote in Australia. A study of federal elections between 2001 and 2016 via the `eechidna` `R` package."
author: "Jeremy Forbes, Di Cook & Rob Hyndman"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
output: bookdown::gitbook
documentclass: book
classoption: openany
bibliography: [book.bib, packages.bib]
biblio-style: apalike
link-citations: yes
github-repo: rstudio/bookdown-demo
description: "This is an incomplete draft of a paper to submitted to ANZJS."
---

<!--chapter:end:index.Rmd-->

# Introduction {#intro}

```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = F, fig.pos = 'h', fig.align = 'center', cache = T, message = F)
library(tidyverse)
library(eechidna)
library(ggplot2)
library(ggthemes)
library(knitr)
library(nlme)
library(rgeos)
library(spdep)

# Data
data(tpp01)
data(tpp04)
data(tpp07)
data(tpp10)
data(tpp13)
data(tpp16)
data(abs2001)
data(abs2004)
data(abs2007)
data(abs2010)
data(abs2013)
data(abs2016)
```

```{r wrangle, include = F, message = F}
# Function to standardize variables
standardise_vars <- function(df) {
  hold <- df
  num_cols <- sapply(df, class) == 'numeric'
  df[, num_cols] <- lapply(df[, num_cols], scale)
  names(df) <- names(hold)
  df$LNP_Percent <- hold$LNP_Percent
  return(df)
}

# Combine and standardize
my_df <- bind_rows(
  left_join(tpp01, standardise_vars(abs2001) %>% dplyr::select(-c(UniqueID, Area, ends_with("NS"), Population)), by = c("DivisionNm", "StateAb"="State")) %>% mutate(year = "2001"),
  left_join(tpp04, standardise_vars(abs2004) %>% dplyr::select(-UniqueID), by = c("DivisionNm")) %>% mutate(year = "2004"),
  left_join(tpp07, standardise_vars(abs2007) %>% dplyr::select(-UniqueID), by = c("DivisionNm")) %>% mutate(year = "2007"),
  left_join(tpp10, standardise_vars(abs2010) %>% dplyr::select(-UniqueID), by = c("DivisionNm")) %>% mutate(year = "2010"),
  left_join(tpp13, standardise_vars(abs2013) %>% dplyr::select(-UniqueID), by = c("DivisionNm")) %>% mutate(year = "2013"),
  left_join(tpp16, standardise_vars(abs2016) %>% dplyr::select(-c(UniqueID, Area, ends_with("NS"), Population)), by = c("DivisionNm", "StateAb"="State")) %>% mutate(year = "2016")
) %>% 
  mutate(year = factor(year)) %>% 
  dplyr::select(-c(starts_with("Age"), StateAb, LNP_Votes, ALP_Votes, ALP_Percent, TotalVotes,
    Swing, InternetUse, InternetAccess, EnglishOnly, Other_NonChrist, OtherChrist, Volunteer, EmuneratedElsewhere, UniqueID, Catholic, Anglican))

# Create final df for modelling

factors_df <- my_df %>% 
  mutate(Education = BachelorAbv + HighSchool + Professional + Finance - Laborer - Tradesperson - DipCert,
    FamHouseSize = FamilyRatio + AverageHouseholdSize + Couple_WChild_House - Couple_NoChild_House -
      SP_House,
    PropertyOwned = Owned + Mortgage - Renting - PublicHousing,
    RentLoanPrice = MedianRent + MedianLoanPay,
    Incomes = MedianFamilyIncome + MedianHouseholdIncome + MedianPersonalIncome,
    Unemployment = Unemployed - LFParticipation) %>% 
  dplyr::select(-c(BachelorAbv, HighSchool, Professional, Finance, Laborer, Tradesperson, DipCert, FamilyRatio,
    AverageHouseholdSize, Couple_WChild_House, Couple_NoChild_House, SP_House, Owned, Mortgage, Renting,
    PublicHousing, MedianFamilyIncome, MedianHouseholdIncome, MedianPersonalIncome, MedianRent, 
    MedianLoanPay, Unemployed, LFParticipation))

# Now standardize factors

small_df <- bind_rows(
  factors_df %>% filter(year == "2001") %>% standardise_vars(),
  factors_df %>% filter(year == "2004") %>% standardise_vars(),
  factors_df %>% filter(year == "2007") %>% standardise_vars(),
  factors_df %>% filter(year == "2010") %>% standardise_vars(),
  factors_df %>% filter(year == "2013") %>% standardise_vars(),
  factors_df %>% filter(year == "2016") %>% standardise_vars()
)

# Order electorates in alphabetical order to match spatial matrix
model_df <- small_df %>% 
  arrange(year, DivisionNm) %>% 
  dplyr::select(order(colnames(.)))
```

```{r spatialweights, eval = F}
# Compute spatial weights matrix
sp_weights_matrix <- function(sF) {

dist_matrix <- function(shapefile) {
  dist_mat <- matrix(NA, nrow = 150, ncol = 150)
  rownames(dist_mat) <- sort(shapefile$elect_div)
  colnames(dist_mat) <- sort(shapefile$elect_div)

  for (i in 1:(nrow(dist_mat)-1)) {
    rname = rownames(dist_mat)[i]
    row_poly = shapefile %>% subset(elect_div == rname)

    for (j in (i+1):ncol(dist_mat)) {

      cname = rownames(dist_mat)[j]
      col_poly = shapefile %>% subset(elect_div == cname)
      dist = gDistance(row_poly, col_poly)
      dist_mat[i,j] = dist

    }
    print(i)
  }

  # Now copy to lower triange
  for (i in 2:nrow(dist_mat)) {
    for (j in 1:(i-1)) {
      dist_mat[i,j] = dist_mat[j,i]
    }
  }

  # Check it is symmetric
  if(!isSymmetric(dist_mat)) {
    print("Warning! Matrix is not symmetric. Error has occured.")
  }

  return(dist_mat)
}

my_dist_mat <- dist_matrix(sF)

# S matrix

s_mat <- function(dist_mat) {
  s_mat <- dist_mat

  for (i in 1:nrow(dist_mat)) {
    for (j in 1:nrow(dist_mat)) {
      a = dist_mat[i,j]

      if (is.na(a)) {
        b = 0
      } else {
        b = ifelse(a == 0, 1, 0)
      }

      s_mat[i,j] = b

    }
  }

  return(s_mat)
}

my_smat <- s_mat(my_dist_mat)

# Turn into W matrix

w_mat <- function(s_mat) {

  w_mat <- s_mat
  rsums = rowSums(s_mat)

  for (i in 1:nrow(s_mat)) {
    w_mat[i,] <- s_mat[i,]/rsums[i]
  }

  return(w_mat)
}

my_wmat <- w_mat(my_smat)

# Turn into listw

my_listw <- mat2listw(my_wmat)

return(my_listw)

}

# Get spatial weights
sF_16 <- sF_download(2016)
sF_13 <- sF_download(2013)
sF_10 <- sF_download(2010)
sF_07 <- sF_download(2007)
sF_04 <- sF_download(2004)
sF_01 <- sF_download(2001)

sp_weights_16 <- sp_weights_matrix(sF_16)
sp_weights_13 <- sp_weights_matrix(sF_13)
sp_weights_10 <- sp_weights_matrix(sF_10)
sp_weights_07 <- sp_weights_matrix(sF_07)
sp_weights_04 <- sp_weights_matrix(sF_04)
sp_weights_01 <- sp_weights_matrix(sF_01)
```

```{r FGLS, include = F}
# Instead of running spatial weights
load("/Users/Jeremy/Documents/R/eechidna-modelling/data/sp_weights_01.rda")
load("/Users/Jeremy/Documents/R/eechidna-modelling/data/sp_weights_04.rda")
load("/Users/Jeremy/Documents/R/eechidna-modelling/data/sp_weights_07.rda")
load("/Users/Jeremy/Documents/R/eechidna-modelling/data/sp_weights_10.rda")
load("/Users/Jeremy/Documents/R/eechidna-modelling/data/sp_weights_13.rda")
load("/Users/Jeremy/Documents/R/eechidna-modelling/data/sp_weights_16.rda")

# Function for FGLS
my_fgls <- function(my_formula, my_data, sp_weights) {
  # Spatial weights matrix
  w_mat <- listw2mat(sp_weights)

  # Get OLS residuals
  ols_model <- lm(my_formula, my_data)
  my_res <- ols_model$residuals

  # Solve for rho
  res_model <- lm(my_res ~ w_mat%*%my_res)
  rho <- res_model$coefficients[2]
  rho_df <- data.frame(estimate = rho, se = summary(res_model)$coefficients[2,2], p = summary(res_model)$coefficients[2,4])

  # Transform data for GLS
  trans_mat <- diag(nrow(my_data)) - rho*w_mat
  gls_data <- data.frame(LNP_Percent = trans_mat %*% my_data$LNP_Percent,
    Intercept = trans_mat %*% rep(1,nrow(my_data))) %>% 
    bind_cols(as.data.frame(trans_mat %*% as.matrix(my_data %>% select(-LNP_Percent))))

  # GLS model
  my_formula <- formula(paste0(my_formula,  " - 1"))
  gls_model <- gls(my_formula, gls_data)

  # Call to function with stargazer
  gls_model$call$model <- formula(paste0("LNP_Percent ~ ", paste0(names(gls_data)[-1], collapse = " + ")))

  # Rho and data
  gls_model$rho_df <- rho_df
  gls_model$gls_data <- gls_data
  gls_model$my_data <- my_data

  return(gls_model)
}

## Run full models for each year
full_formula = "LNP_Percent ~ ."

# 2016
glsmod16 <- my_fgls(full_formula, 
  my_data = model_df %>% filter(year == "2016") %>% select(-c(DivisionNm, year)),
  sp_weights = sp_weights_16)

# 2013
glsmod13 <- my_fgls(full_formula, 
  my_data = model_df %>% filter(year == "2013") %>% select(-c(DivisionNm, year)),
  sp_weights = sp_weights_13)

# 2010
glsmod10 <- my_fgls(full_formula, 
  my_data = model_df %>% filter(year == "2010") %>% select(-c(DivisionNm, year)),
  sp_weights = sp_weights_10)

# 2007
glsmod07 <- my_fgls(full_formula, 
  my_data = model_df %>% filter(year == "2007") %>% select(-c(DivisionNm, year)),
  sp_weights = sp_weights_07)

# 2004
glsmod04 <- my_fgls(full_formula, 
  my_data = model_df %>% filter(year == "2004") %>% select(-c(DivisionNm, year)),
  sp_weights = sp_weights_04)

# 2001
glsmod01 <- my_fgls(full_formula, 
  my_data = model_df %>% filter(year == "2001") %>% select(-c(DivisionNm, year)),
  sp_weights = sp_weights_01)

## Visualise coefficients and significance

coef_df <- bind_rows(
  data.frame(variable = glsmod16$coefficients %>% names, estimate = glsmod16$coefficients %>% unname, se = summary(glsmod16)$tTable[, "Std.Error"] %>% unname, p = summary(glsmod16)$tTable[, "p-value"] %>% unname, year = 2016),
  data.frame(variable = glsmod13$coefficients %>% names, estimate = glsmod13$coefficients %>% unname, se = summary(glsmod13)$tTable[, "Std.Error"] %>% unname, p = summary(glsmod13)$tTable[, "p-value"] %>% unname, year = 2013),
  data.frame(variable = glsmod10$coefficients %>% names, estimate = glsmod10$coefficients %>% unname, se = summary(glsmod10)$tTable[, "Std.Error"] %>% unname, p = summary(glsmod10)$tTable[, "p-value"] %>% unname, year = 2010),
  data.frame(variable = glsmod07$coefficients %>% names, estimate = glsmod07$coefficients %>% unname, se = summary(glsmod07)$tTable[, "Std.Error"] %>% unname, p = summary(glsmod07)$tTable[, "p-value"] %>% unname, year = 2007),
  data.frame(variable = glsmod04$coefficients %>% names, estimate = glsmod04$coefficients %>% unname, se = summary(glsmod04)$tTable[, "Std.Error"] %>% unname, p = summary(glsmod04)$tTable[, "p-value"] %>% unname, year = 2004),
  data.frame(variable = glsmod01$coefficients %>% names, estimate = glsmod01$coefficients %>% unname, se = summary(glsmod01)$tTable[, "Std.Error"] %>% unname, p = summary(glsmod01)$tTable[, "p-value"] %>% unname, year = 2001)
)
```

Australia has changed in many ways over the last two decades. Rising house prices, country-wide improvements in education, an ageing population, and a decline in religious affiliation, are just a few facets of the country's evolving socio-demographic characteristics. At the same time, political power has moved back and forth between the two major parties. In the 2007 and 2010 federal elections, the Australian Labor Party (Labor) was victorious, whereas the 2001, 2004, 2013 and 2016 elections were won by the Liberal National coalition (Liberal). The two-party preferred vote, a measure of support between these two parties, fluctuated between 47.3% and 53.5% (in favour of the Liberal party) over this period. This study explores how electoral characteristics relate to two-party preference, and whether their effects have changed over time. Electoral socio-demographics are derived from the Census, and vote counts are obtained from federal elections.

Joining these two data sources is problematic as there in an inherent asynchronicity in the two events. A Census is conducted by the Australian Bureau of Statistics (ABS) every five years, whereas a federal election (conducted by the Australian Electoral Commission (AEC)) usually occurs every three years. The first problem addressed is that of obtaining appropriate Census data for the 2004, 2007, 2010 and 2013 elections - election years in which a Census does not occur. The predominant approach in previous studies is to join voting outcomes to the nearest Census, without accounting for any temporal differences (see @DavisStimson98, @Stimson06, @Liao09 and @Stimson09). Furthermore, electoral boundaries change regularly, so spatial discrepancies also arise when matching electoral data. To obtain appropriate Census data for these four elections, electoral socio-demographics are imputed using a spatio-temporal imputation that combines areal interpolation [@Goodchild1993] and linear time-interpolation. The process of collecting and wrangling the raw data is outlined in section \@ref(data), and the imputation process is detailed in section \@ref(imputation). All data and associated documentation relating to this procedure are available in the `eechidna` `R` package [@eechidna], providing a resource for future analysis.

Previous work on modelling Australian federal elections have found that aggregate socio-demographics are relatively good predictors of voting outcomes. @Forrest01 does this using multiple regression of the Liberal and Labor primary vote for polling booths in the Farrer electorate in 1998. @Stimson06, @Stimson09 and @Stimson12 use principal component analysis of polling booths in the 2001, 2004 and 2007 elections respectively, also finding that socio-demographic characteristics of polling booths are linked to their two-party preferred vote. On the contrary, @Stimson09 models the polling booth swing vote (change in the two-party preferred vote) in the 2007 election, finding that little of swing vote can be explained by Census data. Instead of analyzing a single election in isolation, this paper employs a consistent model framework across six elections so that temporal changes in the effects of socio-demographics can be observed, where each federal elections is modelled with a cross-sectional data set. The use of a regression framework to examine these socio-political relationships over time is seemingly absent from previous Australian studies. It also appears that no study has attempted any type of statistical analysis of socio-demographics in conjunction with voter behaviour in Australia since 2007, making this paper distinctly different from those previous.

The cross-sectional data set for each election consists of the two-party preferred vote (response variable), and socio-demographic variables (explanatory variables) that characterise each electorate. To obtain these cross-sections, socio-demographic variables are first standardized, and then principal components are used to group variables into "factors". To account for the inherent spatial structure of the data, a spatial error model is fit for each election. These steps are discussed in section \@ref(modelling). In section \@ref(results) inference is conducted on the models to see which effects are significant, how effects change over time and which electorates have abnormal voting behaviour.

<!--chapter:end:01-intro.Rmd-->

# Data collection {#data}

## Voting data

The voting outcome of interest is the electoral two-party preferred vote, which is provided by the Australian Electoral Commission (AEC) for the 2001, 2004, 2007, 2010, 2013 and 2016 elections via the AEC Tally Room. The AEC divide Australia into 150 regions called electorates, with each corresponding to a single seat in the House of Representatives. Voting is compulsory in Australia, and each voter assigns a numbered preference to each available candidate in their electorate. The two-party preferred vote is determined by a tally of these preferences where, by convention, only the ranks of the Labor and Liberal candidates are considered. This is recorded as a percentage preference in favour of the Liberal party.

## Census data

Socio-demographic variables are derived from the Census of Population and Housing (Census), which is a survey of every household in Australia, recording information such as age, gender, ethnicity, education level and income. There have been four Censuses in the 21st century, being that in 2001, 2006, 2011 and 2016. The Australian Bureau of Statistics (ABS) conducts the Census and publishes aggregated information. The ABS approximation of electorates at the time of the Census is chosen. From this aggregate information, 67 socio-demographic variables are computed for each of the electorates.

## Downloading and wrangling

Raw data is sourced online from the AEC and ABS websites in .csv and .xlsx files. The format of these files change over the years, making extracting the appropriate information a big task. The functions available in the `dplyr` [@dplyr] and `readxl` [@readxl] `R` packages are very useful, as they provide fast consistent tools for data manipulation and functions to import .xlsx files (respectively). The 2001 and 2006 Census data are however published in a format where each electorate has a separate document, making it difficult to use the `dplyr` tools and instead cells have to be selected from each individual file to construct the desired variables. All scripts required for the data wrangling process can be found in the github repository for the `eechidna` `R` package [@eechidna], along with the raw data. The `eechidna` package makes this study entirely reproducible and provides a resource to help wrangle data for future Censuses and elections, when they become available.

<!--chapter:end:02-data-collection.Rmd-->

# Joining Census and election data {#imputation}

## Differences between Census and election data

Between 2001 and 2016 there were six elections and four Censuses (see Figure \@ref(fig:timeline)). Electoral boundaries are redistributed regularly by the AEC, meaning that only in the years where both a Census and election occur will the boundaries match - the case for the 2001 and 2016 election. Therefore, for the four elections between 2004 and 2013, both temporal and spatial differences in electorates need to be accounted for when joining the electoral two-party preferred vote with Census data. For these elections a spatio-temporal imputation method is employed to obtain electoral socio-demographics. This method uses Census information from both before and after the election of interest.

```{r timeline, message = F, echo = F, fig.cap="Timeline of Australian elections and Censuses. They do not always occur in the same year.", out.width = "90%", fig.align="center"}
timeline <- data.frame(year = c(2001:2016)) %>%
  mutate(Election = ifelse(year %in% c(2001, 2004, 2007, 2010, 2013, 2016),1,0),
         Census = ifelse(year %in% c(2001, 2006, 2011, 2016), 1.4, 0)) %>%
  gather(c(Election,Census), key = event, value = measure) %>%  filter(measure != 0) %>%
  mutate(year_vis = ifelse(event == "Census", year + 0.1, year))

ggplot(timeline, aes(x=year_vis, y=measure)) + geom_point(aes(color=as.factor(event),shape=as.factor(event)),size = 4) + labs(color = "Event", shape = "Event") + geom_hline(yintercept=0, size =1, color = 'grey', show.legend = FALSE) + geom_segment(x=2001,xend=2001, y=1, yend=0, color = '#00BFC4') + geom_segment(x=2004,xend=2004, y=1, yend=0, color = '#00BFC4') + geom_segment(x=2007,xend=2007, y=1, yend=0, color = '#00BFC4') + geom_segment(x=2010,xend=2010, y=1, yend=0, color = '#00BFC4') + geom_segment(x=2013,xend=2013, y=1, yend=0, color = '#00BFC4') + geom_segment(x=2016,xend=2016, y=1, yend=0, color = '#00BFC4') + geom_segment(x=2001.1,xend=2001.1, y=1.4, yend=0, color = '#F8766D') + geom_segment(x=2006.1,xend=2006.1, y=1.4, yend=0, color = '#F8766D') + geom_segment(x=2011.1,xend=2011.1, y=1.4, yend=0, color = '#F8766D') + geom_segment(x=2016.1,xend=2016.1, y=1.4, yend=0, color = '#F8766D') + labs(x="Year") + theme(text=element_text(size=10), axis.text.x=element_text(size=6),axis.text.y=element_blank(), axis.title.y=element_blank(), plot.background=element_blank(), axis.ticks.y = element_blank(), panel.background = element_blank()) + scale_x_discrete(limits = c(2001:2016)) + coord_fixed(ratio = 2)
```

## Spatio-temporal imputation

To account for spatial differences, the piece-wise approximation method in @Goodchild1993 is adopted. Consider a map of source zones $s = 1,...,S$, for which socio-demographic information is available, and a set of target zones $t = 1,...,T$ for which information is to be imputed. In this context the map of electoral boundaries at the time of a Census would be the source zones, and the boundaries at the time of the election would be the target zones. Denote the area of intersection between source zone $s$ and target zone $t$ as $A_{s,t}$, the population of the source zone $s$ as $U_s$, and the population of intersection between source zone $s$ and target zone $t$ as $P_{s,t}$.

Compute each $A_{s,t}$ and estimate population of the intersection:

$$\hat{P}_{s,t} = \frac{U_s*A_{s,t}}{\sum_{t=1}^T A_{s,t}}$$
This assumes that populations are uniformly distributed within each source zone.

In order to calculate socio-demographic information for each of the target zones, a weighted average is taken using the estimated population as weights. Denote a given Census variable for the target zone $C_t$, and the same Census variable for the source zone $D_s$:

$$\hat{C}_t = \frac{\sum_{s=1}^{S}{D_s*\hat{P}_{s,t}}}{\sum_{s=1}^{S}{\hat{P}_{s,t}}}$$
This assumes that each individual in a source zone assumes the aggregate characteristics of the zone.

Applying this to each of the target zones addresses the spatial component, as it imputes the required socio-demographic for the desired electoral boundaries. However these are applicable at the time of the Census (source year) and are not yet appropriate for the election (target year).

Denote year $y$, with a Census falling on $y_1$ and $y_3$, and an election on year $y_2$, and add this subscript to the Census variable estimate, $\hat{C}_{t,y}$. To account for temporal changes, linear interpolation is used between Census years to get the final estimate of a Census variable for the target zone in the election year $y_2$. This assumes that population evolves in a linear manner over time.

$$\hat{C}_{t,y_2} = \frac{y_3-y_2}{y_3-y_1}*\hat{C}_{t,y_1} + \frac{y_2-y_1}{y_3-y_1}*\hat{C}_{t,y_3}$$

## Applied

Publically available Census data is aggregated and there are different resolutions accessible, ranging from SA1 (over 50,000 zones) to electoral divisions (150 zones). For this study electoral divisions are used as source zones, and this imputation method is applied for each of the 2004, 2007, 2010 and 2013 elections. To demonstrate its functionality, consider the imputation of socio-demographic variables for the electorate of Hume in New South Wales (NSW), at the time of the 2013 federal election. Figure \@ref(fig:hume13) shows this region amongst other NSW electorates.

```{r hume13, fig.cap = "Some of the electoral boundaries in NSW for 2013, with the electoral boundary for Hume, shown in purple.", out.height="30%"}
hume_area13 <- nat_map13 %>% 
  filter(state %in% c("ACT","NSW"), long < 154)

ggplot(data=hume_area13) +
  geom_polygon(aes(x=long, y=lat, group=group, fill = elect_div == "HUME"),
               colour="grey50", alpha = 0.4) +
  scale_fill_manual(name="Boundary", values=c("white", "purple"), labels = c("Other 2013 Electorates", "Hume 2013")) +
  theme_map() + coord_equal()
```

The Censuses neighbouring the 2013 election are those in 2011 and 2016, and the Hume boundary is changed, as seen by plotting the Hume boundary (purple) in the 2013 election over the divisions in 2016 (see Figure \@ref(fig:hume16)).

```{r hume16, fig.cap = "Census division boundaries in NSW for 2016, with the 2013 electoral boundary for Hume, shown in purple. The purple region is not contained within a single Census division.", out.height="30%"}
hume_area16 <- nat_map16 %>% 
  filter(state %in% c("ACT","NSW"), long < 154) %>% 
  mutate(year = "2016") %>% 
  bind_rows(hume_area13 %>% filter(elect_div == "HUME") %>% mutate(year = "2013"))

ggplot(data=hume_area16) +
  geom_polygon(aes(x=long, y=lat, group=group, fill = year == "2013", 
    alpha = year == "2013", colour=year == "2013")) +
  scale_fill_manual(name="Boundary", values=c("grey95", "purple"), labels = c("2016 Electorates", "Hume 2013")) +
  scale_alpha_manual(values=c(0, 0.4)) +
  scale_color_manual(values=c("grey50", NA)) +
  theme_map() + coord_equal() + guides(alpha = F, color = F)
```

There are many electorates in 2016 that intersect with the purple region (Hume boundary for 2013), these include the divisions of Riverina, Eden-Monaro and Hume, along with smaller intersecting areas with Fenner, Calare, Gilmore and Whitlam. To impute Census information for this purple region, calculate the percentage of each 2016 electorate that intersects with the purple region, which is then used to estimate intersection populations $\hat{P}_{s,t}$.

Electorate (2016) | Percentage | Population in Electorate | Estimated Population Allocated to Purple Region: $\hat{P}_{s,t}$
---------------------|------------|--------------------------|------------------------------------------------
HUME              | 96.54%     | 150643                   | 145427
RIVERINA          | 25.11%     | 155793                   | 39117
EDEN-MONARO       | 11.09%     | 147532                   | 16358
CANBERRA          | 0.28%      | 196037                   | 548
FENNER            | 0.23%      | 202955                   | 474
WHITLAM           | 0.06%      | 152280                   | 92
GILMORE           | 0.06%      | 150436                   | 86
CALARE            | 0.01%      | 161298                   | 21

Now consider the socio-demographic $AusCitizen$ - the proportion of people in the region who are Australian citizens.

```{r eval = F, include = F}
auscitizen_ex <- abs2016 %>% 
  filter(DivisionNm %in% c("HUME", "RIVERINA", "EDEN-MONARO", "CANBERRA", "FENNER", "WHITLAM", "GILMORE", "CALARE")) %>%
  select(DivisionNm, AusCitizen) %>% 
  mutate(`Estimated Population Allocated to Purple Region` = c(21, 548, 16358, 474, 86, 145427, 39117, 92),
    AusCitizen = round(AusCitizen, 2)) %>% 
  arrange(-`Estimated Population Allocated to Purple Region`)

auscitizen_ex %>% 
  kable()

weighted.mean(x = auscitizen_ex$AusCitizen, w = auscitizen_ex$`Estimated Population Allocated to Purple Region`)
```

|DivisionNm      | AusCitizen (%): $D_s$| Estimated Population Allocated to Purple Region: $\hat{P}_{s,t}$|
|----------------|-------------------|------------------------------------------------|
|HUME            |              90.02|                                          145427|
|RIVERINA        |              89.11|                                           39117|
|EDEN-MONARO     |              88.00|                                           16358|
|CANBERRA        |              85.48|                                             548|
|FENNER          |              83.64|                                             474|
|WHITLAM         |              89.52|                                              92|
|GILMORE         |              89.03|                                              86|
|CALARE          |              87.56|                                              21|

Then taking a weighted average of $AusCitizen$ using the estimated population as weights yields $\hat{C}_{Hume,2016} = 89.65 \%$. Repeating this process using the 2011 Census and electoral boundaries yields $\hat{C}_{Hume,2011} = 91.00 \%$

Finally, linearly interpolate between 2011 and 2016 to arrive at the 2013 estimate:
\begin{eqnarray*}
\hat{C}_{Hume,2013} & = &\frac{3}{5} \cdot \hat{C}_{Hume,2011} + \frac{2}{5} \cdot \hat{C}_{Hume,2016} \\ 
& = & \frac{3}{5} \cdot 91.00 \% + \frac{2}{5} \cdot 89.65 \% \\ 
& = & 90.46 \%
\end{eqnarray*}

This is done for each of the socio-demographic variables, and repeated each of the 2013 electorates.

<!--chapter:end:03-imputation.Rmd-->

---
output:
  pdf_document: default
  html_document: default
---
# Modelling {#modelling}

## Data pre-processing
With socio-demographic information now available for each electorate, each election is joined to the data  corresponding with its two-party preferred vote. Socio-demographic variables within each election year are standardized to have mean zero and variance one, to adjust for changing variable scales. For example, inflation-adjusted median rental prices increased across almost all electorates, with median rent of 200 dollars per week placing an electorate in the 90th percentile in 2001, but only the 30th percentile in 2016.

### Dimension reduction
With only $N = 150$ observations (electorates) in each election and $p = 65$ socio-demographic variables in each cross-section, any model using all variables would face serious problems with multi-collinearity and over-fitting, likely leading to erroneous conclusions regarding variable significance. Therefore a form of dimension reduction is adopted before models are fit.

Socio-demographic variables[^2] that represent similar information are combined into "factors" using principal component analysis (PCA). The scree plots of the principal components for each election all level off after four components, and the loadings of these four components are similar across the elections. Principal components are then computed on the combined set of socio-demographics across all six elections. A factor is created by combining several variables all have large loadings in a particular component and when there is an intuitive reason as to why these variables could represent common information. A loading with magnitude greater than 0.15 is considered large. After computing these sums, each factor is again standardized to have mean zero and variance one, within each election.

Consider the `Incomes` factor as an illustration. Independent of principal components, we may suspect that median personal income, median household income and median family income are providing similar information about the financial wellbeing of an electorate. Their loadings in the first principal component are large (0.20, 0.21 and 0.22 respectively), which provides the evidence needed to combine these variables into a single factor, which is called `Incomes`.

This process reduces the predictor set to $p = 30$.

[^2]: A preliminary step involved removing all age bands, because age is represented by median age, and to remove variables relating to particular denominations of Christianity.

## Model framework
An identical model specification is used across the six elections, with each election modelled separately. This allows for the socio-demographic effects to be estimated separately for each year, allowing for interpretation of temporal changes in these effects. This is preferable over a single longitudinal model because it avoids any concerns of undue bias stemming from an incorrectly imposed time-varying restriction on any variable. Without such restrictions, a pooled cross-sectional model does not yield any distinct advantage over separate cross-sections. The panel approach is avoided because of how frequently electoral boundaries change, meaning that electorates that have the same name across elections are not guaranteed to represent the same geographical region. Therefore any fixed or random effects models would be difficult to estimate without implementing consistent boundaries, which would requiring further imputation.

For each cross-section, let the response variable be the two-party preferred vote in favour of the Liberal party, denoted $Y$, with $Y = 70$ representing a 70% preference for Liberal, 30% for Labor. Although $Y$ lies in the interval $(0,100)$, observed values are never very close to 0 or 100 (minimum $24.05 \%$ and maximum $74.90 \%$), so there is no need to impose the constraint of $Y \in [0,100]$. Furthermore, the response is found to be spatially correlated in each election (Moran's I test, $p \le 7\cdot10^{-15}$). This is expected, as electorates are aggregate spatial units, and hence the spatial structure of the data must modelled appropriately.

The spatial error model [@Anselin88] is chosen because captures spatial heterogeneity by incorporating a spatially structured random effect vector [@LeSage2009]. In this context, the random effect can be thought of as capturing the unobserved political climate in each electorate, where the climate is correlated with the climate in neighbouring electorates. This functions under the assumption that the climate is independent of electoral socio-demographics, and that an electorate is equally correlated with any electorate that shares a part of its boundary. Spatial weights are calculated in accordance with these assumptions. The spatial error model is specified as follows:

Let $\rho$ be spatial autoregressive coefficient, $\boldsymbol v$ be a spherical error term, ${\boldsymbol W}$ be a matrix of spatial weights (containing information about the neighbouring regions), $\boldsymbol X$ be a matrix of socio-demographic covariates, $\boldsymbol \beta$ be a vector of regression coefficients and $\boldsymbol a$ be a spatially structured random effect vector.

$${\boldsymbol y} = {\boldsymbol X} {\boldsymbol \beta} + {\boldsymbol a}$$
and

$${\boldsymbol a} = \rho {\boldsymbol W} {\boldsymbol a} + {\boldsymbol v}$$
where

$${\boldsymbol v} \sim N({\boldsymbol 0}, \sigma^2 {\boldsymbol I_n})$$.

so it can be written

$${\boldsymbol y} = {\boldsymbol X} {\boldsymbol \beta} + ({\boldsymbol I}_n-\rho {\boldsymbol W})^{-1}{\boldsymbol v}$$

Estimation is done using feasible generalized least squares.

Table 4.1 details the resultant estimated model coefficients and their estimated standard errors for each of the six elections. These are interpreted in the next section.

```{r printmodels, eval = F}
# Add rho and make adjustments to this
stargazer::stargazer(glsmod16, glsmod13, glsmod10, glsmod07, glsmod04, glsmod01, header = FALSE, 
  column.sep.width = "1pt", 
  title = "Estimated model for each of the six elections.",
  dep.var.labels = "Two-party preferred vote in favor of the Liberal party", 
  font.size = "scriptsize", digits = 2, 
  column.labels = c("2001", "2004", "2007", "2010", "2013", "2016"),
  notes = c("Estimated coefficients for variable named in column one shown for", "election year indicated by column heading, with estimated standard", "deviation for each coefficient shown below in parenthesis. Overall", "summary measures for each regression equation are provided in the", "bottom panel."),
  notes.align = "l")
```

\begin{table}[!htbp] \centering 
  \caption{Estimated model for each of the six elections.} 
  \label{} 
\tiny 
\begin{tabular}{@{\extracolsep{1pt}}lcccccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{6}{c}{\textit{Dependent variable:}} \\ 
\cline{2-7} 
\\[-1.8ex] & \multicolumn{6}{c}{Two-party preferred vote in favor of the Liberal party} \\ 
 & 2001 & 2004 & 2007 & 2010 & 2013 & 2016 \\ 
\\[-1.8ex] & (1) & (2) & (3) & (4) & (5) & (6)\\ 
\hline \\[-1.8ex] 
 $\rho$ & 0.48$^{***}$ & 0.29$^{*}$ & 0.15 & $-$0.04 & 0.15 & 0.44$^{***}$ \\ 
  & (0.16) & (0.16) & (0.17) & (0.19) & (0.17) & (0.15) \\ 
  & & & & & & \\ 
 AusCitizen & $-$1.53 & $-$0.87 & 1.00 & 0.67 & $-$2.83 & $-$2.77 \\ 
  & (2.95) & (2.90) & (2.93) & (2.57) & (2.63) & (2.51) \\ 
  & & & & & & \\ 
 Born\_Asia & $-$2.13 & $-$2.56 & $-$7.58$^{***}$ & $-$1.37 & 0.74 & 2.63 \\ 
  & (2.20) & (2.58) & (2.68) & (1.96) & (2.55) & (3.03) \\ 
  & & & & & & \\ 
 Born\_MidEast & $-$0.82 & 1.70 & 1.79 & 2.92$^{*}$ & 3.22$^{*}$ & 1.92 \\ 
  & (1.56) & (1.54) & (1.54) & (1.51) & (1.78) & (1.53) \\ 
  & & & & & & \\ 
 Born\_SE\_Europe & $-$1.63 & $-$1.25 & $-$1.72 & $-$0.20 & 0.04 & 1.02 \\ 
  & (0.99) & (1.15) & (1.34) & (1.17) & (2.05) & (2.34) \\ 
  & & & & & & \\ 
 Born\_UK & $-$1.37 & $-$0.35 & $-$0.29 & 0.29 & $-$0.87 & $-$0.49 \\ 
  & (0.98) & (1.00) & (1.05) & (0.80) & (0.90) & (1.04) \\ 
  & & & & & & \\ 
 BornElsewhere & 1.74 & 4.04 & 9.78$^{**}$ & 5.01 & $-$0.51 & $-$1.08 \\ 
  & (4.52) & (4.46) & (4.47) & (3.59) & (3.67) & (3.40) \\ 
  & & & & & & \\ 
 Buddhism & 1.51 & 1.09 & 2.41$^{*}$ & 0.90 & 2.84$^{*}$ & 1.50 \\ 
  & (1.24) & (1.25) & (1.40) & (1.25) & (1.61) & (1.94) \\ 
  & & & & & & \\ 
 Christianity & 5.17$^{**}$ & 3.08 & $-$0.60 & $-$1.93 & 3.44 & 0.53 \\ 
  & (2.52) & (2.79) & (3.44) & (3.45) & (4.19) & (4.58) \\ 
  & & & & & & \\ 
 CurrentlyStudying & $-$0.09 & 1.73 & 3.28$^{***}$ & 4.03$^{***}$ & 1.69 & $-$1.14 \\ 
  & (1.13) & (1.17) & (1.21) & (1.12) & (1.09) & (1.01) \\ 
  & & & & & & \\ 
 DeFacto & $-$11.02$^{***}$ & $-$9.73$^{***}$ & $-$8.60$^{***}$ & $-$6.55$^{***}$ & $-$5.04$^{**}$ & $-$6.42$^{***}$ \\ 
  & (2.55) & (2.62) & (2.57) & (1.90) & (2.00) & (1.69) \\ 
  & & & & & & \\ 
 DiffAddress & 6.93$^{***}$ & 5.02$^{***}$ & 4.72$^{***}$ & 3.40$^{***}$ & 4.63$^{***}$ & 3.86$^{***}$ \\ 
  & (1.50) & (1.71) & (1.60) & (0.88) & (1.03) & (0.95) \\ 
  & & & & & & \\ 
 Distributive & 1.65 & 2.76$^{**}$ & 2.88$^{**}$ & 2.25$^{**}$ & 2.75$^{**}$ & 1.67 \\ 
  & (1.17) & (1.25) & (1.30) & (1.02) & (1.13) & (1.08) \\ 
  & & & & & & \\ 
 Education & $-$3.04 & $-$0.90 & $-$1.82 & $-$0.77 & 1.94 & 0.62 \\ 
  & (2.99) & (3.04) & (3.35) & (2.89) & (2.86) & (2.23) \\ 
  & & & & & & \\ 
 Extractive & 6.81$^{***}$ & 6.09$^{***}$ & 6.97$^{***}$ & 5.81$^{***}$ & 5.31$^{***}$ & 4.69$^{***}$ \\ 
  & (1.31) & (1.42) & (1.48) & (1.16) & (1.23) & (1.36) \\ 
  & & & & & & \\ 
 FamHouseSize & $-$4.35 & $-$6.16$^{*}$ & $-$8.22$^{**}$ & $-$9.67$^{***}$ & $-$5.11$^{*}$ & $-$2.94 \\ 
  & (2.86) & (3.15) & (3.28) & (2.66) & (2.78) & (2.27) \\ 
  & & & & & & \\ 
 Incomes & 8.46$^{***}$ & 4.19 & 2.73 & 4.36$^{*}$ & 2.95 & 2.99$^{*}$ \\ 
  & (2.57) & (2.68) & (3.06) & (2.56) & (2.60) & (1.76) \\ 
  & & & & & & \\ 
 Indigenous & 4.47$^{***}$ & 6.12$^{***}$ & 5.90$^{***}$ & 6.81$^{***}$ & 5.62$^{***}$ & 3.97$^{***}$ \\ 
  & (1.38) & (1.49) & (1.49) & (1.17) & (1.35) & (1.26) \\ 
  & & & & & & \\ 
 Islam & 2.64 & 1.20 & $-$1.04 & $-$1.21 & 0.61 & 0.15 \\ 
  & (1.72) & (1.71) & (1.94) & (1.80) & (2.00) & (2.03) \\ 
  & & & & & & \\ 
 Judaism & 1.96$^{***}$ & 1.60$^{*}$ & 0.46 & 0.01 & 1.74 & 1.02 \\ 
  & (0.71) & (0.85) & (1.05) & (1.00) & (1.26) & (1.38) \\ 
  & & & & & & \\ 
 ManagerAdminClericalSales & 4.62$^{***}$ & 3.99$^{***}$ & 4.32$^{***}$ & 5.21$^{***}$ & 3.28$^{***}$ & 2.10$^{***}$ \\ 
  & (1.03) & (0.99) & (1.00) & (0.78) & (0.85) & (0.70) \\ 
  & & & & & & \\ 
 Married & $-$1.39 & 1.46 & 5.07$^{*}$ & 5.06$^{*}$ & 5.50$^{**}$ & 2.27 \\ 
  & (2.64) & (2.71) & (2.91) & (2.59) & (2.79) & (2.26) \\ 
  & & & & & & \\ 
 MedianAge & 2.94$^{*}$ & 2.28 & 2.08 & 1.86 & 3.73$^{**}$ & 1.54 \\ 
  & (1.63) & (1.88) & (2.08) & (1.65) & (1.55) & (1.30) \\ 
  & & & & & & \\ 
 NoReligion & 5.76$^{**}$ & 3.53 & 0.25 & $-$0.31 & 3.43 & 1.06 \\ 
  & (2.59) & (2.58) & (2.98) & (2.90) & (3.45) & (3.85) \\ 
  & & & & & & \\ 
 OneParent\_House & $-$0.85 & 0.11 & 0.22 & 1.66 & 1.37 & $-$1.00 \\ 
  & (1.39) & (1.46) & (1.61) & (1.38) & (1.49) & (1.34) \\ 
  & & & & & & \\ 
 OtherLanguageHome & 1.27 & $-$4.69 & $-$4.37 & $-$6.02 & $-$6.76 & $-$10.40$^{**}$ \\ 
  & (4.95) & (5.50) & (5.70) & (4.44) & (5.61) & (4.97) \\ 
  & & & & & & \\ 
 PropertyOwned & 1.76 & 0.68 & $-$1.17 & 0.78 & $-$0.92 & $-$0.76 \\ 
  & (1.38) & (1.43) & (1.49) & (1.09) & (1.21) & (1.19) \\ 
  & & & & & & \\ 
 RentLoanPrice & $-$0.70 & 1.01 & 2.12 & $-$0.54 & $-$1.54 & $-$0.31 \\ 
  & (2.02) & (1.96) & (1.96) & (1.48) & (1.68) & (1.50) \\ 
  & & & & & & \\ 
 SocialServ & 2.90$^{***}$ & 0.94 & 1.91 & 1.77 & 1.89 & 2.42$^{*}$ \\ 
  & (1.10) & (1.22) & (1.35) & (1.14) & (1.27) & (1.27) \\ 
  & & & & & & \\ 
 Transformative & 4.75$^{***}$ & 3.74$^{**}$ & 3.28$^{*}$ & 5.20$^{***}$ & 5.74$^{***}$ & 2.99$^{**}$ \\ 
  & (1.41) & (1.57) & (1.74) & (1.47) & (1.59) & (1.41) \\ 
  & & & & & & \\ 
 Unemployment & 0.36 & $-$1.22 & $-$2.42 & $-$2.88$^{**}$ & $-$5.09$^{***}$ & $-$3.00$^{**}$ \\ 
  & (1.42) & (1.54) & (1.73) & (1.41) & (1.51) & (1.28) \\ 
  & & & & & & \\ 
 Constant & 50.45$^{***}$ & 53.48$^{***}$ & 49.92$^{***}$ & 47.32$^{***}$ & 52.63$^{***}$ & 50.78$^{***}$ \\ 
  & (0.74) & (0.56) & (0.51) & (0.35) & (0.45) & (0.67) \\ 
  & & & & & & \\ 
\hline \\[-1.8ex] 
Observations & 150 & 150 & 150 & 150 & 150 & 150 \\ 
Log Likelihood & $-$399.52 & $-$405.41 & $-$418.08 & $-$398.73 & $-$403.34 & $-$398.33 \\ 
Akaike Inf. Crit. & 863.03 & 874.83 & 900.16 & 861.46 & 870.68 & 860.67 \\ 
Bayesian Inf. Crit. & 951.96 & 963.76 & 989.10 & 950.40 & 959.61 & 949.60 \\ 
\hline 
\hline \\[-1.8ex] 
\textit{Note:}  & \multicolumn{6}{l}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
 & \multicolumn{6}{l}{Estimated coefficients for variable named in column one shown for} \\ 
 & \multicolumn{6}{l}{election year indicated by column heading, with estimated standard} \\ 
 & \multicolumn{6}{l}{deviation for each coefficient shown below in parenthesis. Overall} \\ 
 & \multicolumn{6}{l}{summary measures for each regression equation are provided in the} \\ 
 & \multicolumn{6}{l}{bottom panel.} \\ 
\end{tabular} 
\end{table}

<!--chapter:end:04-modelling.Rmd-->

# Results
```{r rhovis}
rho_df <- bind_rows(glsmod16$rho_df %>% mutate(year = "2016"), 
  glsmod13$rho_df %>% mutate(year = "2013"),
  glsmod10$rho_df %>% mutate(year = "2010"), 
  glsmod07$rho_df %>% mutate(year = "2007"),
  glsmod04$rho_df %>% mutate(year = "2004"), 
  glsmod01$rho_df %>% mutate(year = "2001")) %>% 
  mutate(upper95 = estimate + 1.96*se, lower95 = estimate - 1.96*se)

rho_df %>% ggplot() + 
  geom_point(aes(x = year, y = estimate, col = factor(p < 0.05)), size = 3) +
  geom_linerange(aes(x = year, ymin = lower95, ymax = upper95, col = factor(p < 0.05)), size = 1.5) + 
  geom_hline(aes(yintercept = 0), alpha = 0.5, size = 1) +
  scale_color_manual(values = c("grey50", "black")) +
  guides(col = F)
```

The spatial autoregressive coefficient $\rho$ is significant in only the 2001 and 2016 elections, meaning that in these years an electorate's political climate was affected by the attitudes of it's neighbours. Conversely, in the other four elections, it appears that after accounting for electoral socio-demographics, there is no spatial effect, meaning electorates effectively acted independently.

To investigate the socio-demographics that have a strong effect on the two-party preferred vote, partial residual plots are used. These show the direction, size and significance of an estimated effect, as the slope of the prediction line matches the estimated coefficient, and the shaded region represents a 95% confidence band. Typically the Labor party campaigns on more progressive policies, which often include tax reform that adversely affects higher income earners, and more generous social assistance programs. Although topical in the lead up to an election, incomes only significantly affected electoral two-party preference in 2016 (see Figure \@ref(fig:plotincomes)), with higher income areas associated with support for the Liberal party. In the prior elections, the estimated effect is still positive, albeit insignificant. Unemployment on the other hand, was influential in 2001, 2004 and 2007 elections(Figure \@ref(fig:plotunemploy)), with electorates with higher unemployment favouring Labor. This effect has become weaker over time, with unemployment having an estimated effect of virtually zero in 2016.

```{r myvisreg, include = F}
# Function to produce visreg style conditional plots
my_visreg <- function(my_model, sp_weights, varname, 
  nolabs = FALSE, xlimits = NULL, ylimits = NULL, year = "") {

  # Extract fitted parameters
  rho <- my_model$rho_df$estimate
  sigma <- sqrt(sum(my_model$residuals^2)/(my_model$dims$N-my_model$dims$p))

  # Spatial weights
  w_mat <- listw2mat(sp_weights)

  # Q - where u = Qe, Q = (I - pW)^-1
  q_mat <- solve(diag(my_model$dims$N) - rho*w_mat)

  # Omega - QQ'
  omega_mat <- q_mat%*%t(q_mat)

  # X
  x_mat <- my_model$my_data %>% 
    dplyr::select(-LNP_Percent) %>% 
    mutate(Intercept = 1) %>% 
    dplyr::select(Intercept, everything()) %>% 
    as.matrix()

  # Beta
  beta_mat <- my_model$coefficients

  # T value
  t = qt(0.975, nrow(my_model$gls_data)-ncol(my_model$gls_data))

  # Lambda matrix (FGLS)
  x <- round(seq(min(x_mat[, varname]), max(x_mat[, varname]), 0.025), 3)
  lambda_mat <- data.frame(matrix(0, nrow = length(x), ncol = ncol(x_mat)))
  names(lambda_mat) <- dimnames(x_mat)[[2]]
  lambda_mat[, varname] <- x
  lambda_mat$Intercept <- 1
  lambda_mat <- as.matrix(lambda_mat)

  # Confidence interval
  plot_df <- data.frame(variable = x, fitted = lambda_mat%*%beta_mat, variance = 0)

  for (i in 1:nrow(lambda_mat)) {
    lambda <- lambda_mat[i, ]
    plot_df$variance[i] = sigma^2 * t(lambda) %*% 
      solve(t(x_mat) %*% solve(omega_mat) %*% x_mat) %*% 
      lambda
  }

  plot_df <- plot_df %>% 
    mutate(upper95 = fitted + t*sqrt(variance), lower95 = fitted - t*sqrt(variance))

  # Partial residuals
  points_df <- data.frame(
    variable = my_model$my_data[, varname] %>% unname,
    part_res = (my_model$my_data$LNP_Percent - x_mat%*%my_model$coefficients) + my_model$coefficients[varname]*x_mat[, varname] + my_model$coefficients[1]
  )

  # Plot
  myplot <- ggplot(data = plot_df) + 
    geom_ribbon(aes(x = variable, ymin = lower95, ymax = upper95), fill = "grey80") + 
    geom_point(aes(x = variable, y = part_res), data = points_df, size = 0.75, col = "grey50") + 
    geom_line(aes(x = variable, y = fitted), col = "blue", size = 1) +
    #geom_hline(aes(yintercept = min(upper95)), col = "red") +
    #geom_hline(aes(yintercept = max(lower95)), col = "blue") +
    theme_bw() + 
    labs(x = varname, y = "Response") + 
    ggtitle(year) +
    theme(plot.title = element_text(face = "bold", size = 10, hjust = 0.5))

  if (nolabs == TRUE) {
    myplot <- myplot + labs(x = "", y = "")
  }

  if (!is.null(xlimits) & !is.null(ylimits)) {
    myplot <- myplot + coord_cartesian(xlim = xlimits, ylim = ylimits)
  }

  return(myplot)
}

# Grid visreg
library(gridExtra)
library(grid)
grid_visreg <- function(varname, xvar_title, xlimits, ylimits) {
  p16 <- my_visreg(glsmod16, sp_weights_16, varname = varname, nolabs = T, year = "2016", xlimits, ylimits)
  p13 <- my_visreg(glsmod13, sp_weights_13, varname = varname, nolabs = T, year = "2013", xlimits, ylimits)
  p10 <- my_visreg(glsmod10, sp_weights_10, varname = varname, nolabs = T, year = "2010", xlimits, ylimits)
  p07 <- my_visreg(glsmod07, sp_weights_07, varname = varname, nolabs = T, year = "2007", xlimits, ylimits)
  p04 <- my_visreg(glsmod04, sp_weights_04, varname = varname, nolabs = T, year = "2004", xlimits, ylimits)
  p01 <- my_visreg(glsmod01, sp_weights_01, varname = varname, nolabs = T, year = "2001", xlimits, ylimits)

  plots <- grid.arrange(p01, p04, p07, p10, p13, p16, nrow = 2, 
    left = textGrob("Two-party preferred vote (%)", gp = gpar(cex = 0.8), rot = 90), 
    bottom = textGrob(xvar_title, gp = gpar(cex = 0.8)))

  return(plots)
}

```

```{r plotincomes, fig.cap = "Effect of Incomes"}
p <- grid_visreg("Incomes", xvar_title = "Incomes", xlimits = c(-1.5, 3.6), ylimits = c(30, 95))
```

```{r plotunemploy, fig.cap = "Effect of Unemployment"}
p <- grid_visreg("Unemployment", xvar_title = "Unemployment", xlimits = c(-2.1, 3.4), ylimits = c(33, 70))
```

This section summarises the main findings made about the relationships between socio-demographic chracteristics and voting behaviour.

```{r}
coef_df %>% 
  filter(variable != "Intercept") %>% 
  mutate(upper95 = estimate + 1.96*se, lower95 = estimate - 1.96*se) %>% 
  ggplot() +
  geom_point(aes(x = year, y = estimate, col = factor(p < 0.05)), size = 2) +
  geom_linerange(aes(x = year, ymin = lower95, ymax = upper95, col = factor(p < 0.05)), size = 1) + 
  geom_hline(aes(yintercept = 0), alpha = 0.5, size = 1) + 
  facet_wrap(~variable, scales = "free", ncol = 5) +
  scale_color_manual(values = c("grey50", "black")) + 
  theme(axis.text = element_text(size = 6), strip.text = element_text(size = 6)) +
  labs(x = "", y = "Year") +
  scale_x_continuous(breaks = c(2001, 2004, 2007, 2010, 2013, 2016)) +
  guides(col = F)
```

<!--chapter:end:05-results.Rmd-->

`r if (knitr:::is_html_output()) '
# References {-}
'`

<!--chapter:end:06-references.Rmd-->

### 2019-03-26 01-intro.Rmd
added full data to glsmod output

@@
 library(spdep)
+library(gridExtra)
+library(grid)
 
@@
 my_fgls <- function(my_formula, my_data, sp_weights) {
+  # DivisionNm
+  model_data <- my_data %>% dplyr::select(-c(DivisionNm, year))
+
   # Spatial weights matrix
@@
   # Get OLS residuals
-  ols_model <- lm(my_formula, my_data)
+  ols_model <- lm(my_formula, model_data)
   my_res <- ols_model$residuals
@@
   # Transform data for GLS
-  trans_mat <- diag(nrow(my_data)) - rho*w_mat
-  gls_data <- data.frame(LNP_Percent = trans_mat %*% my_data$LNP_Percent,
-    Intercept = trans_mat %*% rep(1,nrow(my_data))) %>% 
-    bind_cols(as.data.frame(trans_mat %*% as.matrix(my_data %>% select(-LNP_Percent))))
+  trans_mat <- diag(nrow(model_data)) - rho*w_mat
+  gls_data <- data.frame(LNP_Percent = trans_mat %*% model_data$LNP_Percent,
+    Intercept = trans_mat %*% rep(1,nrow(model_data))) %>% 
+    bind_cols(as.data.frame(trans_mat %*% as.matrix(model_data %>% dplyr::select(-LNP_Percent))))
 
@@
   gls_model$my_data <- my_data
+  gls_model$actual_residuals <- solve(trans_mat)%*%gls_model$residuals
 
@@
 glsmod16 <- my_fgls(full_formula, 
-  my_data = model_df %>% filter(year == "2016") %>% select(-c(DivisionNm, year)),
+  my_data = model_df %>% filter(year == "2016"),
   sp_weights = sp_weights_16)
@@
 glsmod13 <- my_fgls(full_formula, 
-  my_data = model_df %>% filter(year == "2013") %>% select(-c(DivisionNm, year)),
+  my_data = model_df %>% filter(year == "2013"),
   sp_weights = sp_weights_13)
@@
 glsmod10 <- my_fgls(full_formula, 
-  my_data = model_df %>% filter(year == "2010") %>% select(-c(DivisionNm, year)),
+  my_data = model_df %>% filter(year == "2010"),
   sp_weights = sp_weights_10)
@@
 glsmod07 <- my_fgls(full_formula, 
-  my_data = model_df %>% filter(year == "2007") %>% select(-c(DivisionNm, year)),
+  my_data = model_df %>% filter(year == "2007"),
   sp_weights = sp_weights_07)
@@
 glsmod04 <- my_fgls(full_formula, 
-  my_data = model_df %>% filter(year == "2004") %>% select(-c(DivisionNm, year)),
+  my_data = model_df %>% filter(year == "2004"),
   sp_weights = sp_weights_04)
@@
 glsmod01 <- my_fgls(full_formula, 
-  my_data = model_df %>% filter(year == "2001") %>% select(-c(DivisionNm, year)),
+  my_data = model_df %>% filter(year == "2001"),
   sp_weights = sp_weights_01)

### 2019-03-26 05-results.Rmd
added sections on effects, and residuals

@@
 # Results
+
+## Spatial autoregressive parameter
 ```{r rhovis}
@@
 
-The spatial autoregressive coefficient $\rho$ is significant in only the 2001 and 2016 elections, meaning that in these years an electorate's political climate was affected by the attitudes of it's neighbours.
-Conversely, in the other four elections, it appears that after accounting for electoral socio-demographics, there is no spatial effect, meaning electorates effectively acted independently.
+The spatial autoregressive coefficient $\rho$ is positive and significant in only the 2001 and 2016 elections, meaning that in these elections, an electorate's political climate was affected by the attitudes of it's neighbours.
+Conversely, in the other four elections, it appears that after accounting for electoral socio-demographics, there is no significant spatial effect, meaning electorates effectively acted independently.
 
+## Influential socio-demographics
 To investigate the socio-demographics that have a strong effect on the two-party preferred vote, partial residual plots are used.
-These show the direction, size and significance of an estimated effect, as the slope of the prediction line matches the estimated coefficient, and the shaded region represents a 95% confidence band.
+These show the direction, size and significance of an estimated effect - the slope of the prediction line matches the estimated coefficient, and the shaded region represents a 95% confidence band.
+If a horizontal line can be drawn through the confidence band, then the effect is insignificant.
+Plots for each election are faceted to compare the effects over time.
+
+*It is important here to note the ecological fallacy - insights are being drawn at the electorate level, and cannot be inferred for another disaggregate level (e.g. individual voters).*
+
+### Income and unemployment
 Typically the Labor party campaigns on more progressive policies, which often include tax reform that adversely affects higher income earners, and more generous social assistance programs.
-Although topical in the lead up to an election, incomes only significantly affected electoral two-party preference in 2016 (see Figure \@ref(fig:plotincomes)), with higher income areas associated with support for the Liberal party.
+Despite this, incomes have only significantly affected the electoral two-party preference in 2016 (see Figure \@ref(fig:plotincomes)), with higher income areas associated with support for the Liberal party.
 In the prior elections, the estimated effect is still positive, albeit insignificant.
-Unemployment on the other hand, was influential in 2001, 2004 and 2007 elections(Figure \@ref(fig:plotunemploy)), with electorates with higher unemployment favouring Labor.
-This effect has become weaker over time, with unemployment having an estimated effect of virtually zero in 2016.
+Unemployment on the other hand, was influential in 2001, 2004 and 2007 elections (Figure \@ref(fig:plotunemploy)), with electorates with higher unemployment favouring Labor.
+This estimated effect has become weaker over time, with unemployment having virtually no effect in 2016.
 
@@
 # Grid visreg
-library(gridExtra)
-library(grid)
 grid_visreg <- function(varname, xvar_title, xlimits, ylimits) {
@@
   plots <- grid.arrange(p01, p04, p07, p10, p13, p16, nrow = 2, 
-    left = textGrob("Two-party preferred vote (%)", gp = gpar(cex = 0.8), rot = 90), 
-    bottom = textGrob(xvar_title, gp = gpar(cex = 0.8)))
+    left = grid.text("Two-party preferred vote (%)", gp = gpar(cex = 0.8), rot = 90), 
+    bottom = grid.text(xvar_title, gp = gpar(cex = 0.8)))
 
@@
 
-```{r plotincomes, fig.cap = "Effect of Incomes"}
+```{r plotincomes, fig.cap = "Effect of Incomes", out.width="60%"}
 p <- grid_visreg("Incomes", xvar_title = "Incomes", xlimits = c(-1.5, 3.6), ylimits = c(30, 95))
@@
 
-```{r plotunemploy, fig.cap = "Effect of Unemployment"}
+```{r plotunemploy, fig.cap = "Effect of Unemployment", out.width="60%"}
 p <- grid_visreg("Unemployment", xvar_title = "Unemployment", xlimits = c(-2.1, 3.4), ylimits = c(28, 68))
@@
 
-This section summarises the main findings made about the relationships between socio-demographic chracteristics and voting behaviour.
+### Industry and type of work
+Electorates with higher proportions of workers in mining, gas, water, agriculture, waste and electricity (grouped as `Extractive` industries) are consistently linked with higher support for the Liberal party, with the mangitude of this effect slightly increasing over the years (see Figure \@ref(fig:plotextractive)).
+This is unsurprising, as the Liberal party has close ties with these traditional energy industries, and typically present policies to reduce taxation on energy production.
+Furthermore, electorates with more workers in construction or manufacturing industries (`Transformative`), as well as wholesale trade, retail trade, transport, post or warehousing (`Distributive`), are also linked with the Liberal party.
+However, these effects are not as strong, and `Transformative` was marginally insignificant in the 2010 election, whereas `Distributive` was not a key socio-demographic in 2001 or 2016.
+These effects are shown in Figures \@ref(fig:plottransformative) and \@ref(fig:plotdistributive).
+
+```{r plotextractive, fig.cap = "Effect of Extractive", out.width="60%"}
+p <- grid_visreg("Extractive", xvar_title = "Extractive", xlimits = c(-0.7, 3.4), ylimits = c(35, 81))
+```
+
+```{r plottransformative, fig.cap = "Effect of Transformative", out.width="60%"}
+p <- grid_visreg("Transformative", xvar_title = "Transformative", xlimits = c(-2.25, 2.9), ylimits = c(34, 75))
+```
+```{r plotdistributive, fig.cap = "Effect of Distributive", out.width="60%"}
+p <- grid_visreg("Distributive", xvar_title = "Distributive", xlimits = c(-3.1, 2.5), ylimits = c(34, 65))
+```
+
+Similarly, the proportions of workers in managerial, administrative, clerical and sales roles (`ManagerAdminClericalSales`) is also a significant predictor of two-party preference across all six elections, with higher proportions supporting the Liberal party.
+
+```{r plotmanager, fig.cap = "Effect of ManagerAdminClericalSales", out.width="60%"}
+p <- grid_visreg("ManagerAdminClericalSales", xvar_title = "ManagerAdminClericalSales", xlimits = c(-3.75, 2.8), ylimits = c(32, 70))
+```
+
+### Household mobility
+```{r plotdiffaddress, fig.cap = "Effect of DiffAddress", out.width="60%"}
+p <- grid_visreg("DiffAddress", xvar_title = "DiffAddress", xlimits = c(-2.1, 3.65), ylimits = c(34, 81))
+```
+
+In each of the six elections, electorates with a higher proportion of people that have recently moved house (meaning in last five years) were more likely to support the Liberal party, with this effect becoming stronger since 2010.
+Having controlled for characteristics of house ownership and rental prices (via the factors `PropertyOwned` and `RentLoan` respectively), this is effect is somewhat surprising.
+*Add more here?*
+
+### Indigenous population
+```{r plotindigenous, fig.cap = "Effect of Indigenous population", out.width="60%"}
+p <- grid_visreg("Indigenous", xvar_title = "Indigenous", xlimits = c(-0.6, 9.5), ylimits = c(35, 120))
+```
+
+The indigenous population of an electorate is found to be a significant predictor in every election, with higher indigenous populations supporting the Liberal party.
+There are a few outliers with Lingiari (NT) being the most extreme.
+However, Lingiari itself has been a strong Labor supporter, and this effect appears to be driven by the other electorates with high Indigenous populations (Kalgoorlie (WA), Leichhardt (NSW), Durack (WA), Parkes(NSW) and Kennedy (QLD)), which have almost always elected a Liberal candidate.
+
+### Relationships and families
+In each of the six elections, higher proportions people in de facto marriages are tied to support for the Labor party, whereas the effect of marriages themselves are insignificant.
+Larger family and household size (via the `FamHouseSize` factor) is also linked with Labor support, although only being significant in the 2007 and 2010 elections (see Figure \@ref(fig:plotfamhousesize)).
+
+```{r plotdefacto, fig.cap = "Effect of DeFacto", out.width="60%", eval = F}
+p <- grid_visreg("DeFacto", xvar_title = "DeFacto", xlimits = c(-2.2, 5.1), ylimits = c(12, 83))
+```
+
+```{r plotfamhousesize, fig.cap = "Effect of FamHouseSize", out.width="60%", eval = F}
+p <- grid_visreg("FamHouseSize", xvar_title = "FamHouseSize", xlimits = c(-3.4, 2.7), ylimits = c(12, 83))
+```
+
+### A note on similar variables
+
+Many of the Census variables represent similar information, which is why factors were created and some variables were removed.
+However, there still remain some variables which are closely related.
+For example, `OtherLanguageHome` is the proportion of households that speak a language other than English at home, which is likely to be related to the proportion of people born overseas (a sum of the variables with a `Born` prefix).
+In 2001, the coefficient estiamte for `OtherLanguageHome` is significant and negative, and all `Born` variables are insignificant.
+If `OtherLanguageHome` is removed from the model, populations born in South Eastern Europe (`Born_SE_Europe`) absorb the negative effect ($p = 0.0011$).
+
+```{r removeolh, eval = F}
+no_OLH01 <- my_fgls(full_formula, 
+  my_data = model_df %>% filter(year == "2001") %>% dplyr::select(-OtherLanguageHome),
+  sp_weights = sp_weights_01)
+
+summary(no_OLH01)$tTable
+```
+
+## Residuals
+
+```{r resids, include = F}
+all_resids <- data.frame(
+  Residuals = c(glsmod16$actual_residuals, glsmod13$actual_residuals, glsmod10$actual_residuals, glsmod07$actual_residuals, glsmod04$actual_residuals, glsmod01$actual_residuals),
+  gls_residuals = c(glsmod16$residuals, glsmod13$residuals, glsmod10$residuals, glsmod07$residuals, glsmod04$residuals, glsmod01$residuals),
+  bind_rows(
+    glsmod16$my_data %>% left_join(tpp16 %>% dplyr::select(DivisionNm, UniqueID, StateAb), by = "DivisionNm"), 
+    glsmod13$my_data %>% left_join(tpp13 %>% dplyr::select(DivisionNm, UniqueID, StateAb), by = "DivisionNm"), 
+    glsmod10$my_data %>% left_join(tpp10 %>% dplyr::select(DivisionNm, UniqueID, StateAb), by = "DivisionNm"), 
+    glsmod07$my_data %>% left_join(tpp07 %>% dplyr::select(DivisionNm, UniqueID, StateAb), by = "DivisionNm"), 
+    glsmod04$my_data %>% left_join(tpp04 %>% dplyr::select(DivisionNm, UniqueID, StateAb), by = "DivisionNm"), 
+    glsmod01$my_data %>% left_join(tpp01 %>% dplyr::select(DivisionNm, UniqueID, StateAb), by = "DivisionNm"))
+) %>% dplyr::select(Residuals, DivisionNm, UniqueID, StateAb, year, LNP_Percent, gls_residuals)
+```
+
+## Residuals by state
+
+It is often hypothesized that states have systematic differences that cause their electorates to vote differently.
+Boxplots of residuals grouped by state reveal that only Tasmania and the Australian Capital Territory appear to have a state-specific effect that is not captured by the models.
+These states appear to have a bias towards Labor.
+There are few electorates in these states (five and two, respectively), so this might be due to incumbent effects rather than an actual state-specific bias.
+
+```{r resstate, fig.cap = "Boxplot of residuals by state.", fig.out="75%"}
+all_resids %>% 
+  ggplot(aes(x = StateAb, y = Residuals, col = StateAb)) + 
+  geom_boxplot() + 
+  geom_jitter(aes(label = DivisionNm, shape = year), alpha = 0.4) +
+  #facet_wrap(~year) + 
+  guides(col = F, shape = F, DivisionNm = F)
+```
+
+## Electorates with largest residuals
+
+```{r}
+library(plotly)
+ggplot(all_resids) + geom_jitter(aes(x = year, y = Residuals, label = DivisionNm))
+ggplotly()
+
+all_resids %>% 
+  group_by(UniqueID) %>% 
+  mutate(mean_res = mean(abs(Residuals))) %>% 
+  filter(mean_res > 7) %>%
+  ungroup() %>% View
+  ggplot(aes(x = year, y = Residuals, group = DivisionNm, col = DivisionNm, label = DivisionNm)) + 
+  geom_line() + geom_point() +
+  guides(col = F)
+```
+
+Capricornia, Sydney, Shortland
+
+Residuals by state, by year
 

### 2019-03-26 _book/electoral-modelling-draft.tex
cache

--- 2019-03-24 _book/electoral-modelling-draft.tex
@@
   \postdate{\par}
-    \date{2019-03-24}
+    \date{2019-03-26}
 
@@
 
-\emph{insert table here}
+\begin{table}[!htbp] \centering 
+  \caption{Estimated model for each of the six elections.} 
+  \label{} 
+\scriptsize 
+\begin{tabular}{@{\extracolsep{1pt}}lcccccc} 
+\\[-1.8ex]\hline 
+\hline \\[-1.8ex] 
+ & \multicolumn{6}{c}{\textit{Dependent variable:}} \\ 
+\cline{2-7} 
+\\[-1.8ex] & \multicolumn{6}{c}{Two-party preferred vote in favor of the Liberal party} \\ 
+ & 2001 & 2004 & 2007 & 2010 & 2013 & 2016 \\ 
+\\[-1.8ex] & (1) & (2) & (3) & (4) & (5) & (6)\\ 
+\hline \\[-1.8ex] 
+ $\rho$ & 0.44$^{***}$ & 0.15$^{*}$ & $-$0.04 & 0.15 & 0.29 & 0.48$^{***}$ \\ 
+  & (0.15) & (0.17) & (0.19) & (0.17) & (0.16) & (0.16) \\ 
+  & & & & & & \\ 
+ AusCitizen & $-$2.77 & $-$2.83 & 0.67 & 1.00 & $-$0.87 & $-$1.53 \\ 
+  & (2.51) & (2.63) & (2.57) & (2.93) & (2.90) & (2.95) \\ 
+  & & & & & & \\ 
+ Born\_Asia & 2.63 & 0.74 & $-$1.37 & $-$7.58$^{***}$ & $-$2.56 & $-$2.13 \\ 
+  & (3.03) & (2.55) & (1.96) & (2.68) & (2.58) & (2.20) \\ 
+  & & & & & & \\ 
+ Born\_MidEast & 1.92 & 3.22$^{*}$ & 2.92$^{*}$ & 1.79 & 1.70 & $-$0.82 \\ 
+  & (1.53) & (1.78) & (1.51) & (1.54) & (1.54) & (1.56) \\ 
+  & & & & & & \\ 
+ Born\_SE\_Europe & 1.02 & 0.04 & $-$0.20 & $-$1.72 & $-$1.25 & $-$1.63 \\ 
+  & (2.34) & (2.05) & (1.17) & (1.34) & (1.15) & (0.99) \\ 
+  & & & & & & \\ 
+ Born\_UK & $-$0.49 & $-$0.87 & 0.29 & $-$0.29 & $-$0.35 & $-$1.37 \\ 
+  & (1.04) & (0.90) & (0.80) & (1.05) & (1.00) & (0.98) \\ 
+  & & & & & & \\ 
+ BornElsewhere & $-$1.08 & $-$0.51 & 5.01 & 9.78$^{**}$ & 4.04 & 1.74 \\ 
+  & (3.40) & (3.67) & (3.59) & (4.47) & (4.46) & (4.52) \\ 
+  & & & & & & \\ 
+ Buddhism & 1.50 & 2.84$^{*}$ & 0.90 & 2.41$^{*}$ & 1.09 & 1.51 \\ 
+  & (1.94) & (1.61) & (1.25) & (1.40) & (1.25) & (1.24) \\ 
+  & & & & & & \\ 
+ Christianity & 0.53 & 3.44 & $-$1.93 & $-$0.60 & 3.08 & 5.17$^{**}$ \\ 
+  & (4.58) & (4.19) & (3.45) & (3.44) & (2.79) & (2.52) \\ 
+  & & & & & & \\ 
+ CurrentlyStudying & $-$1.14 & 1.69 & 4.03$^{***}$ & 3.28$^{***}$ & 1.73 & $-$0.09 \\ 
+  & (1.01) & (1.09) & (1.12) & (1.21) & (1.17) & (1.13) \\ 
+  & & & & & & \\ 
+ DeFacto & $-$6.42$^{***}$ & $-$5.04$^{**}$ & $-$6.55$^{***}$ & $-$8.60$^{***}$ & $-$9.73$^{***}$ & $-$11.02$^{***}$ \\ 
+  & (1.69) & (2.00) & (1.90) & (2.57) & (2.62) & (2.55) \\ 
+  & & & & & & \\ 
+ DiffAddress & 3.86$^{***}$ & 4.63$^{***}$ & 3.40$^{***}$ & 4.72$^{***}$ & 5.02$^{***}$ & 6.93$^{***}$ \\ 
+  & (0.95) & (1.03) & (0.88) & (1.60) & (1.71) & (1.50) \\ 
+  & & & & & & \\ 
+ Distributive & 1.67 & 2.75$^{**}$ & 2.25$^{**}$ & 2.88$^{**}$ & 2.76$^{**}$ & 1.65 \\ 
+  & (1.08) & (1.13) & (1.02) & (1.30) & (1.25) & (1.17) \\ 
+  & & & & & & \\ 
+ Education & 0.62 & 1.94 & $-$0.77 & $-$1.82 & $-$0.90 & $-$3.04 \\ 
+  & (2.23) & (2.86) & (2.89) & (3.35) & (3.04) & (2.99) \\ 
+  & & & & & & \\ 
+ Extractive & 4.69$^{***}$ & 5.31$^{***}$ & 5.81$^{***}$ & 6.97$^{***}$ & 6.09$^{***}$ & 6.81$^{***}$ \\ 
+  & (1.36) & (1.23) & (1.16) & (1.48) & (1.42) & (1.31) \\ 
+  & & & & & & \\ 
+ FamHouseSize & $-$2.94 & $-$5.11$^{*}$ & $-$9.67$^{***}$ & $-$8.22$^{**}$ & $-$6.16$^{*}$ & $-$4.35 \\ 
+  & (2.27) & (2.78) & (2.66) & (3.28) & (3.15) & (2.86) \\ 
+  & & & & & & \\ 
+ Incomes & 2.99$^{*}$ & 2.95 & 4.36$^{*}$ & 2.73 & 4.19 & 8.46$^{***}$ \\ 
+  & (1.76) & (2.60) & (2.56) & (3.06) & (2.68) & (2.57) \\ 
+  & & & & & & \\ 
+ Indigenous & 3.97$^{***}$ & 5.62$^{***}$ & 6.81$^{***}$ & 5.90$^{***}$ & 6.12$^{***}$ & 4.47$^{***}$ \\ 
+  & (1.26) & (1.35) & (1.17) & (1.49) & (1.49) & (1.38) \\ 
+  & & & & & & \\ 
+ Islam & 0.15 & 0.61 & $-$1.21 & $-$1.04 & 1.20 & 2.64 \\ 
+  & (2.03) & (2.00) & (1.80) & (1.94) & (1.71) & (1.72) \\ 
+  & & & & & & \\ 
+ Judaism & 1.02 & 1.74 & 0.01 & 0.46 & 1.60$^{*}$ & 1.96$^{***}$ \\ 
+  & (1.38) & (1.26) & (1.00) & (1.05) & (0.85) & (0.71) \\ 
+  & & & & & & \\ 
+ ManagerAdminClericalSales & 2.10$^{***}$ & 3.28$^{***}$ & 5.21$^{***}$ & 4.32$^{***}$ & 3.99$^{***}$ & 4.62$^{***}$ \\ 
+  & (0.70) & (0.85) & (0.78) & (1.00) & (0.99) & (1.03) \\ 
+  & & & & & & \\ 
+ Married & 2.27 & 5.50$^{**}$ & 5.06$^{*}$ & 5.07$^{*}$ & 1.46 & $-$1.39 \\ 
+  & (2.26) & (2.79) & (2.59) & (2.91) & (2.71) & (2.64) \\ 
+  & & & & & & \\ 
+ MedianAge & 1.54 & 3.73$^{**}$ & 1.86 & 2.08 & 2.28 & 2.94$^{*}$ \\ 
+  & (1.30) & (1.55) & (1.65) & (2.08) & (1.88) & (1.63) \\ 
+  & & & & & & \\ 
+ NoReligion & 1.06 & 3.43 & $-$0.31 & 0.25 & 3.53 & 5.76$^{**}$ \\ 
+  & (3.85) & (3.45) & (2.90) & (2.98) & (2.58) & (2.59) \\ 
+  & & & & & & \\ 
+ OneParent\_House & $-$1.00 & 1.37 & 1.66 & 0.22 & 0.11 & $-$0.85 \\ 
+  & (1.34) & (1.49) & (1.38) & (1.61) & (1.46) & (1.39) \\ 
+  & & & & & & \\ 
+ OtherLanguageHome & $-$10.40$^{**}$ & $-$6.76 & $-$6.02 & $-$4.37 & $-$4.69 & 1.27 \\ 
+  & (4.97) & (5.61) & (4.44) & (5.70) & (5.50) & (4.95) \\ 
+  & & & & & & \\ 
+ PropertyOwned & $-$0.76 & $-$0.92 & 0.78 & $-$1.17 & 0.68 & 1.76 \\ 
+  & (1.19) & (1.21) & (1.09) & (1.49) & (1.43) & (1.38) \\ 
+  & & & & & & \\ 
+ RentLoanPrice & $-$0.31 & $-$1.54 & $-$0.54 & 2.12 & 1.01 & $-$0.70 \\ 
+  & (1.50) & (1.68) & (1.48) & (1.96) & (1.96) & (2.02) \\ 
+  & & & & & & \\ 
+ SocialServ & 2.42$^{*}$ & 1.89 & 1.77 & 1.91 & 0.94 & 2.90$^{***}$ \\ 
+  & (1.27) & (1.27) & (1.14) & (1.35) & (1.22) & (1.10) \\ 
+  & & & & & & \\ 
+ Transformative & 2.99$^{**}$ & 5.74$^{***}$ & 5.20$^{***}$ & 3.28$^{*}$ & 3.74$^{**}$ & 4.75$^{***}$ \\ 
+  & (1.41) & (1.59) & (1.47) & (1.74) & (1.57) & (1.41) \\ 
+  & & & & & & \\ 
+ Unemployment & $-$3.00$^{**}$ & $-$5.09$^{***}$ & $-$2.88$^{**}$ & $-$2.42 & $-$1.22 & 0.36 \\ 
+  & (1.28) & (1.51) & (1.41) & (1.73) & (1.54) & (1.42) \\ 
+  & & & & & & \\ 
+ Constant & 50.78$^{***}$ & 52.63$^{***}$ & 47.32$^{***}$ & 49.92$^{***}$ & 53.48$^{***}$ & 50.45$^{***}$ \\ 
+  & (0.67) & (0.45) & (0.35) & (0.51) & (0.56) & (0.74) \\ 
+  & & & & & & \\ 
+\hline \\[-1.8ex] 
+Observations & 150 & 150 & 150 & 150 & 150 & 150 \\ 
+Log Likelihood & $-$398.33 & $-$403.34 & $-$398.73 & $-$418.08 & $-$405.41 & $-$399.52 \\ 
+Akaike Inf. Crit. & 860.67 & 870.68 & 861.46 & 900.16 & 874.83 & 863.03 \\ 
+Bayesian Inf. Crit. & 949.60 & 959.61 & 950.40 & 989.10 & 963.76 & 951.96 \\ 
+\hline 
+\hline \\[-1.8ex] 
+\textit{Note:}  & \multicolumn{6}{l}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
+ & \multicolumn{6}{l}{Estimated coefficients for variable named in column one shown for} \\ 
+ & \multicolumn{6}{l}{election year indicated by column heading, with estimated standard} \\ 
+ & \multicolumn{6}{l}{deviation for each coefficient shown below in parenthesis. Overall} \\ 
+ & \multicolumn{6}{l}{summary measures for each regression equation are provided in the} \\ 
+ & \multicolumn{6}{l}{bottom panel.} \\ 
+\end{tabular} 
+\end{table}
 
@@
 
+\hypertarget{spatial-autoregressive-parameter}{%
+\section{Spatial autoregressive parameter}\label{spatial-autoregressive-parameter}}
+
+\begin{center}\includegraphics{electoral-modelling-draft_files/figure-latex/rhovis-1} \end{center}
+
+The spatial autoregressive coefficient \(\rho\) is positive and significant in only the 2001 and 2016 elections, meaning that in these elections, an electorate's political climate was affected by the attitudes of it's neighbours.
+Conversely, in the other four elections, it appears that after accounting for electoral socio-demographics, there is no spatial effect, meaning electorates effectively acted independently.
+
+\hypertarget{influential-socio-demographics}{%
+\section{Influential socio-demographics}\label{influential-socio-demographics}}
+
+To investigate the socio-demographics that have a strong effect on the two-party preferred vote, partial residual plots are used.
+These show the direction, size and significance of an estimated effect, as the slope of the prediction line matches the estimated coefficient, and the shaded region represents a 95\% confidence band.
+If a horizontal line can be drawn through the confidence band, then the effect is insignificant.
+Plots for each election are faceted to compare the effects over time.
+
+\emph{It is important here to note the ecological fallacy - insights are being drawn at the electorate level, and cannot be inferred for another disaggregate level (e.g.~individual voters).}
+
+\hypertarget{income-and-unemployment}{%
+\subsection{Income and unemployment}\label{income-and-unemployment}}
+
+Typically the Labor party campaigns on more progressive policies, which often include tax reform that adversely affects higher income earners, and more generous social assistance programs.
+Although topical in the lead up to an election, incomes only significantly affected electoral two-party preference in 2016 (see Figure \ref{fig:plotincomes}), with higher income areas associated with support for the Liberal party.
+In the prior elections, the estimated effect is still positive, albeit insignificant.
+Unemployment on the other hand, was influential in 2001, 2004 and 2007 elections(Figure \ref{fig:plotunemploy}), with electorates with higher unemployment favouring Labor.
+This estimated effect has become weaker over time, with unemployment having virtually no effect in 2016.
+
+\begin{figure}[h]
+
+{\centering \includegraphics[width=0.6\linewidth]{electoral-modelling-draft_files/figure-latex/plotincomes-1}
+
+}
+
+\caption{Effect of Incomes}\label{fig:plotincomes}
+\end{figure}
+
+\begin{figure}[h]
+
+{\centering \includegraphics[width=0.6\linewidth]{electoral-modelling-draft_files/figure-latex/plotunemploy-1}
+
+}
+
+\caption{Effect of Unemployment}\label{fig:plotunemploy}
+\end{figure}
+
+\hypertarget{industry-and-type-of-work}{%
+\subsection{Industry and type of work}\label{industry-and-type-of-work}}
+
+Electorates with higher proportions of workers in mining, gas, water, agriculture, waste and electricity (grouped as \texttt{Extractive} industries) are consistently linked with higher support for the Liberal party, with the mangitude of this effect slightly increasing over the years (see Figure \ref{fig:plotextractive}).
+This is unsurprising, as the Liberal party has close ties with these traditional energy industries, and typically present policies to reduce taxation on energy production.
+Furthermore, electorates with more workers in construction or manufacturing industries (\texttt{Transformative}), as well as wholesale trade, retail trade, transport, post or warehousing (\texttt{Distributive}), are also linked with the Liberal party.
+However, these effects are not as strong, and \texttt{Transformative} is marginally insignificant in the 2010 election, whereas \texttt{Distributive} was not a key socio-demographic in 2001 or 2016.
+These effects are shown in Figures \ref{fig:plottransformative} and \ref{fig:plotdistributive}.
+
+\begin{figure}[h]
+
+{\centering \includegraphics[width=0.6\linewidth]{electoral-modelling-draft_files/figure-latex/plotextractive-1}
+
+}
+
+\caption{Effect of Extractive}\label{fig:plotextractive}
+\end{figure}
+
+\begin{figure}[h]
+
+{\centering \includegraphics[width=0.6\linewidth]{electoral-modelling-draft_files/figure-latex/plottransformative-1}
+
+}
+
+\caption{Effect of Transformative}\label{fig:plottransformative}
+\end{figure}
+\begin{figure}[h]
+
+{\centering \includegraphics[width=0.6\linewidth]{electoral-modelling-draft_files/figure-latex/plotdistributive-1}
+
+}
+
+\caption{Effect of Distributive}\label{fig:plotdistributive}
+\end{figure}
+
+Similarly, the proportions of workers in managerial, administrative, clerical and sales roles (\texttt{ManagerAdminClericalSales}) is also a significant predictor of two-party preference across all six elections, with higher proportions supporting the Liberal party.
+
+\begin{figure}[h]
+
+{\centering \includegraphics[width=0.6\linewidth]{electoral-modelling-draft_files/figure-latex/plotmanager-1}
+
+}
+
+\caption{Effect of ManagerAdminClericalSales}\label{fig:plotmanager}
+\end{figure}
+
+\hypertarget{household-mobility}{%
+\subsection{Household mobility}\label{household-mobility}}
+
+\begin{figure}[h]
+
+{\centering \includegraphics[width=0.6\linewidth]{electoral-modelling-draft_files/figure-latex/plotdiffaddress-1}
+
+}
+
+\caption{Effect of DiffAddress}\label{fig:plotdiffaddress}
+\end{figure}
+
+Electorates with a higher proportion of people that have recently moved house (meaning in last five years) are more likely to support the Liberal party, with this effect becoming stronger since the 2010 election.
+Having controlled for characteristics of house ownership and rental prices (via the factors \texttt{PropertyOwned} and \texttt{RentLoan} respectively), this is effect is somewhat surprising.
+\emph{Add more here?}
+
+\hypertarget{indigenous-population}{%
+\subsection{Indigenous population}\label{indigenous-population}}
+
+\begin{figure}[h]
+
+{\centering \includegraphics[width=0.6\linewidth]{electoral-modelling-draft_files/figure-latex/plotindigenous-1}
+
+}
+
+\caption{Effect of Indigenous population}\label{fig:plotindigenous}
+\end{figure}
+
+The indigenous population of an electorate is found to be a significant predictor in every election, with higher indigenous populations supporting the Liberal party.
+There are a few outliers with Lingiari (NT) being the most extreme.
+However, Lingiari itself has been a strong Labor supporter, and this effect appears to be driven by the other electorates with high Indigenous populations (Kalgoorlie (WA), Leichhardt (NSW), Durack (WA), Parkes(NSW) and Kennedy (QLD)), which have almost always elected a Liberal candidate.
+
+\hypertarget{relationships-and-families}{%
+\subsection{Relationships and families}\label{relationships-and-families}}
+
+In each of the six elections, higher proportions people in de facto marriages are tied to support for the Labor party, whereas the effect of marriages themselves are insignificant.
+Larger family and household size (via the \texttt{FamHouseSize} factor) is also linked with Labor support, although only being significant in the 2007 and 2010 elections (see Figure \ref{fig:plotfamhousesize}).
+
+\emph{Potentially remove plots}
+
+\hypertarget{residuals}{%
+\section{Residuals}\label{residuals}}
+
+Largest residuals
+
+Residuals by state
+
+\begin{center}\includegraphics{electoral-modelling-draft_files/figure-latex/unnamed-chunk-2-1} \end{center}
+
 \bibliography{book.bib,packages.bib}

### 2019-03-26 03-modelling.Rmd
merge chapters 2 and 3 into data chapter

--- 2019-03-24 _book/electoral-modelling-draft.tex
@@
-\documentclass[openany]{book}
-\usepackage{lmodern}
-\usepackage{amssymb,amsmath}
-\usepackage{ifxetex,ifluatex}
-\usepackage{fixltx2e} % provides \textsubscript
-\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
-  \usepackage[T1]{fontenc}
-  \usepackage[utf8]{inputenc}
-\else % if luatex or xelatex
-  \ifxetex
-    \usepackage{mathspec}
-  \else
-    \usepackage{fontspec}
-  \fi
-  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
-\fi
-% use upquote if available, for straight quotes in verbatim environments
-\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
-% use microtype if available
-\IfFileExists{microtype.sty}{%
-\usepackage{microtype}
-\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
-}{}
-\usepackage[margin=1in]{geometry}
-\usepackage{hyperref}
-\hypersetup{unicode=true,
-            pdftitle={Spatial modelling of the electoral two-party preferred vote in Australia. A study of federal elections between 2001 and 2016 via the eechidna R package.},
-            pdfauthor={Jeremy Forbes, Di Cook \& Rob Hyndman},
-            pdfborder={0 0 0},
-            breaklinks=true}
-\urlstyle{same}  % don't use monospace font for urls
-\usepackage{natbib}
-\bibliographystyle{apalike}
-\usepackage{longtable,booktabs}
-\usepackage{graphicx,grffile}
-\makeatletter
-\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
-\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
-\makeatother
-% Scale images if necessary, so that they will not overflow the page
-% margins by default, and it is still possible to overwrite the defaults
-% using explicit options in \includegraphics[width, height, ...]{}
-\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
-\IfFileExists{parskip.sty}{%
-\usepackage{parskip}
-}{% else
-\setlength{\parindent}{0pt}
-\setlength{\parskip}{6pt plus 2pt minus 1pt}
-}
-\setlength{\emergencystretch}{3em}  % prevent overfull lines
-\providecommand{\tightlist}{%
-  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
-\setcounter{secnumdepth}{5}
-% Redefines (sub)paragraphs to behave more like sections
-\ifx\paragraph\undefined\else
-\let\oldparagraph\paragraph
-\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
-\fi
-\ifx\subparagraph\undefined\else
-\let\oldsubparagraph\subparagraph
-\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
-\fi
+---
+output:
+  pdf_document: default
+  html_document: default
+---
+# Modelling {#modelling}
 
-%%% Use protect on footnotes to avoid problems with footnotes in titles
-\let\rmarkdownfootnote\footnote%
-\def\footnote{\protect\rmarkdownfootnote}
-
-%%% Change title format to be more compact
-\usepackage{titling}
-
-% Create subtitle command for use in maketitle
-\newcommand{\subtitle}[1]{
-  \posttitle{
-    \begin{center}\large#1\end{center}
-    }
-}
-
-\setlength{\droptitle}{-2em}
-
-  \title{Spatial modelling of the electoral two-party preferred vote in Australia. A study of federal elections between 2001 and 2016 via the \texttt{eechidna} \texttt{R} package.}
-    \pretitle{\vspace{\droptitle}\centering\huge}
-  \posttitle{\par}
-    \author{Jeremy Forbes, Di Cook \& Rob Hyndman}
-    \preauthor{\centering\large\emph}
-  \postauthor{\par}
-      \predate{\centering\large\emph}
-  \postdate{\par}
-    \date{2019-03-24}
-
-\usepackage{booktabs}
-\usepackage{amsthm}
-\makeatletter
-\def\thm@space@setup{%
-  \thm@preskip=8pt plus 2pt minus 4pt
-  \thm@postskip=\thm@preskip
-}
-\makeatother
-
-\begin{document}
-\maketitle
-
-{
-\setcounter{tocdepth}{1}
-\tableofcontents
-}
-\hypertarget{intro}{%
-\chapter{Introduction}\label{intro}}
-
-Australia has changed in many ways over the last two decades.
-Rising house prices, country-wide improvements in education, an ageing population, and a decline in religious affiliation, are just a few facets of the country's evolving socio-demographic characteristics.
-At the same time, political power has moved back and forth between the two major parties.
-In the 2007 and 2010 federal elections, the Australian Labor Party (Labor) was victorious, whereas the 2001, 2004, 2013 and 2016 elections were won by the Liberal National coalition (Liberal).
-The two-party preferred vote, a measure of support between these two parties, fluctuated between 47.3\% and 53.5\% (in favour of the Liberal party) over this period.
-This study explores how electoral characteristics relate to two-party preference, and whether their effects have changed over time.
-Electoral socio-demographics are derived from the Census, and vote counts are obtained from federal elections.
-
-Joining these two data sources is problematic as there in an inherent asynchronicity in the two events.
-A Census is conducted by the Australian Bureau of Statistics (ABS) every five years, whereas a federal election (conducted by the Australian Electoral Commission (AEC)) usually occurs every three years.
-The first problem addressed is that of obtaining appropriate Census data for the 2004, 2007, 2010 and 2013 elections - election years in which a Census does not occur.
-The predominant approach in previous studies is to join voting outcomes to the nearest Census, without accounting for any temporal differences (see \citet{DavisStimson98}, \citet{Stimson06}, \citet{Liao09} and \citet{Stimson09}).
-Furthermore, electoral boundaries change regularly, so spatial discrepancies also arise when matching electoral data.
-To obtain appropriate Census data for these four elections, electoral socio-demographics are imputed using a spatio-temporal imputation that combines areal interpolation \citep{Goodchild1993} and linear time-interpolation.
-The process of collecting and wrangling the raw data is outlined in section \ref{data}, and the imputation process is detailed in section \ref{imputation}.
-All data and associated documentation relating to this procedure are available in the \texttt{eechidna} \texttt{R} package \citep{eechidna}, providing a resource for future analysis.
-
-Previous work on modelling Australian federal elections have found that aggregate socio-demographics are relatively good predictors of voting outcomes.
-\citet{Forrest01} does this using multiple regression of the Liberal and Labor primary vote for polling booths in the Farrer electorate in 1998.
-\citet{Stimson06}, \citet{Stimson09} and \citet{Stimson12} use principal component analysis of polling booths in the 2001, 2004 and 2007 elections respectively, also finding that socio-demographic characteristics of polling booths are linked to their two-party preferred vote.
-On the contrary, \citet{Stimson09} models the polling booth swing vote (change in the two-party preferred vote) in the 2007 election, finding that little of swing vote can be explained by Census data.
-Instead of analyzing a single election in isolation, this paper employs a consistent model framework across six elections so that temporal changes in the effects of socio-demographics can be observed, where each federal elections is modelled with a cross-sectional data set.
-The use of a regression framework to examine these socio-political relationships over time is seemingly absent from previous Australian studies.
-It also appears that no study has attempted any type of statistical analysis of socio-demographics in conjunction with voter behaviour in Australia since 2007, making this paper distinctly different from those previous.
-
-The cross-sectional data set for each election consists of the two-party preferred vote (response variable), and socio-demographic variables (explanatory variables) that characterise each electorate.
-To obtain these cross-sections, socio-demographic variables are first standardized, and then principal components are used to group variables into ``factors''.
-To account for the inherent spatial structure of the data, a spatial error model is fit for each election.
-These steps are discussed in section \ref{modelling}.
-In section \ref{results} inference is conducted on the models to see which effects are significant, how effects change over time and which electorates have abnormal voting behaviour.
-
-\hypertarget{data}{%
-\chapter{Data collection}\label{data}}
-
-\hypertarget{voting-data}{%
-\section{Voting data}\label{voting-data}}
-
-The voting outcome of interest is the electoral two-party preferred vote, which is provided by the Australian Electoral Commission (AEC) for the 2001, 2004, 2007, 2010, 2013 and 2016 elections via the AEC Tally Room.
-The AEC divide Australia into 150 regions called electorates, with each corresponding to a single seat in the House of Representatives.
-Voting is compulsory in Australia, and each voter assigns a numbered preference to each available candidate in their electorate.
-The two-party preferred vote is determined by a tally of these preferences where, by convention, only the ranks of the Labor and Liberal candidates are considered.
-This is recorded as a percentage preference in favour of the Liberal party.
-
-\hypertarget{census-data}{%
-\section{Census data}\label{census-data}}
-
-Socio-demographic variables are derived from the Census of Population and Housing (Census), which is a survey of every household in Australia, recording information such as age, gender, ethnicity, education level and income.
-There have been four Censuses in the 21st century, being that in 2001, 2006, 2011 and 2016.
-The Australian Bureau of Statistics (ABS) conducts the Census and publishes aggregated information.
-The ABS approximation of electorates at the time of the Census is chosen.
-From this aggregate information, 67 socio-demographic variables are computed for each of the electorates.
-
-\hypertarget{downloading-and-wrangling}{%
-\section{Downloading and wrangling}\label{downloading-and-wrangling}}
-
-Raw data is sourced online from the AEC and ABS websites in .csv and .xlsx files.
-The format of these files change over the years, making extracting the appropriate information a big task.
-The functions available in the \texttt{dplyr} \citep{dplyr} and \texttt{readxl} \citep{readxl} \texttt{R} packages are very useful, as they provide fast consistent tools for data manipulation and functions to import .xlsx files (respectively).
-The 2001 and 2006 Census data are however published in a format where each electorate has a separate document, making it difficult to use the \texttt{dplyr} tools and instead cells have to be selected from each individual file to construct the desired variables.
-All scripts required for the data wrangling process can be found in the github repository for the \texttt{eechidna} \texttt{R} package \citep{eechidna}, along with the raw data.
-The \texttt{eechidna} package makes this study entirely reproducible and provides a resource to help wrangle data for future Censuses and elections, when they become available.
-
-\hypertarget{imputation}{%
-\chapter{Joining Census and election data}\label{imputation}}
-
-\hypertarget{differences-between-census-and-election-data}{%
-\section{Differences between Census and election data}\label{differences-between-census-and-election-data}}
-
-Between 2001 and 2016 there were six elections and four Censuses (see Figure \ref{fig:timeline}).
-Electoral boundaries are redistributed regularly by the AEC, meaning that only in the years where both a Census and election occur will the boundaries match - the case for the 2001 and 2016 election.
-Therefore, for the four elections between 2004 and 2013, both temporal and spatial differences in electorates need to be accounted for when joining the electoral two-party preferred vote with Census data.
-For these elections a spatio-temporal imputation method is employed to obtain electoral socio-demographics.
-This method uses Census information from both before and after the election of interest.
-
-\begin{figure}[h]
-
-{\centering \includegraphics[width=0.9\linewidth]{electoral-modelling-draft_files/figure-latex/timeline-1}
-
-}
-
-\caption{Timeline of Australian elections and Censuses. They do not always occur in the same year.}\label{fig:timeline}
-\end{figure}
-
-\hypertarget{spatio-temporal-imputation}{%
-\section{Spatio-temporal imputation}\label{spatio-temporal-imputation}}
-
-To account for spatial differences, the piece-wise approximation method in \citet{Goodchild1993} is adopted.
-Consider a map of source zones \(s = 1,...,S\), for which socio-demographic information is available, and a set of target zones \(t = 1,...,T\) for which information is to be imputed.
-In this context the map of electoral boundaries at the time of a Census would be the source zones, and the boundaries at the time of the election would be the target zones.
-Denote the area of intersection between source zone \(s\) and target zone \(t\) as \(A_{s,t}\), the population of the source zone \(s\) as \(U_s\), and the population of intersection between source zone \(s\) and target zone \(t\) as \(P_{s,t}\).
-
-Compute each \(A_{s,t}\) and estimate population of the intersection:
-
-\[\hat{P}_{s,t} = \frac{U_s*A_{s,t}}{\sum_{t=1}^T A_{s,t}}\]
-This assumes that populations are uniformly distributed within each source zone.
-
-In order to calculate socio-demographic information for each of the target zones, a weighted average is taken using the estimated population as weights.
-Denote a given Census variable for the target zone \(C_t\), and the same Census variable for the source zone \(D_s\):
-
-\[\hat{C}_t = \frac{\sum_{s=1}^{S}{D_s*\hat{P}_{s,t}}}{\sum_{s=1}^{S}{\hat{P}_{s,t}}}\]
-This assumes that each individual in a source zone assumes the aggregate characteristics of the zone.
-
-Applying this to each of the target zones addresses the spatial component, as it imputes the required socio-demographic for the desired electoral boundaries.
-However these are applicable at the time of the Census (source year) and are not yet appropriate for the election (target year).
-
-Denote year \(y\), with a Census falling on \(y_1\) and \(y_3\), and an election on year \(y_2\), and add this subscript to the Census variable estimate, \(\hat{C}_{t,y}\).
-To account for temporal changes, linear interpolation is used between Census years to get the final estimate of a Census variable for the target zone in the election year \(y_2\).
-This assumes that population evolves in a linear manner over time.
-
-\[\hat{C}_{t,y_2} = \frac{y_3-y_2}{y_3-y_1}*\hat{C}_{t,y_1} + \frac{y_2-y_1}{y_3-y_1}*\hat{C}_{t,y_3}\]
-
-\hypertarget{applied}{%
-\section{Applied}\label{applied}}
-
-Publically available Census data is aggregated and there are different resolutions accessible, ranging from SA1 (over 50,000 zones) to electoral divisions (150 zones).
-For this study electoral divisions are used as source zones, and this imputation method is applied for each of the 2004, 2007, 2010 and 2013 elections.
-To demonstrate its functionality, consider the imputation of socio-demographic variables for the electorate of Hume in New South Wales (NSW), at the time of the 2013 federal election.
-Figure \ref{fig:hume13} shows this region amongst other NSW electorates.
-
-\begin{figure}[h]
-
-{\centering \includegraphics[height=0.3\textheight]{electoral-modelling-draft_files/figure-latex/hume13-1}
-
-}
-
-\caption{Some of the electoral boundaries in NSW for 2013, with the electoral boundary for Hume, shown in purple.}\label{fig:hume13}
-\end{figure}
-
-The Censuses neighbouring the 2013 election are those in 2011 and 2016, and the Hume boundary is changed, as seen by plotting the Hume boundary (purple) in the 2013 election over the divisions in 2016 (see Figure \ref{fig:hume16}).
-
-\begin{figure}[h]
-
-{\centering \includegraphics[height=0.3\textheight]{electoral-modelling-draft_files/figure-latex/hume16-1}
-
-}
-
-\caption{Census division boundaries in NSW for 2016, with the 2013 electoral boundary for Hume, shown in purple. The purple region is not contained within a single Census division.}\label{fig:hume16}
-\end{figure}
-
-There are many electorates in 2016 that intersect with the purple region (Hume boundary for 2013), these include the divisions of Riverina, Eden-Monaro and Hume, along with smaller intersecting areas with Fenner, Calare, Gilmore and Whitlam.
-To impute Census information for this purple region, calculate the percentage of each 2016 electorate that intersects with the purple region, which is then used to estimate intersection populations \(\hat{P}_{s,t}\).
-
-\begin{longtable}[]{@{}llll@{}}
-\toprule
-\begin{minipage}[b]{0.17\columnwidth}\raggedright
-Electorate (2016)\strut
-\end{minipage} & \begin{minipage}[b]{0.10\columnwidth}\raggedright
-Percentage\strut
-\end{minipage} & \begin{minipage}[b]{0.22\columnwidth}\raggedright
-Population in Electorate\strut
-\end{minipage} & \begin{minipage}[b]{0.40\columnwidth}\raggedright
-Estimated Population Allocated to Purple Region: \(\hat{P}_{s,t}\)\strut
-\end{minipage}\tabularnewline
-\midrule
-\endhead
-\begin{minipage}[t]{0.17\columnwidth}\raggedright
-HUME\strut
-\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright
-96.54\%\strut
-\end{minipage} & \begin{minipage}[t]{0.22\columnwidth}\raggedright
-150643\strut
-\end{minipage} & \begin{minipage}[t]{0.40\columnwidth}\raggedright
-145427\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.17\columnwidth}\raggedright
-RIVERINA\strut
-\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright
-25.11\%\strut
-\end{minipage} & \begin{minipage}[t]{0.22\columnwidth}\raggedright
-155793\strut
-\end{minipage} & \begin{minipage}[t]{0.40\columnwidth}\raggedright
-39117\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.17\columnwidth}\raggedright
-EDEN-MONARO\strut
-\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright
-11.09\%\strut
-\end{minipage} & \begin{minipage}[t]{0.22\columnwidth}\raggedright
-147532\strut
-\end{minipage} & \begin{minipage}[t]{0.40\columnwidth}\raggedright
-16358\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.17\columnwidth}\raggedright
-CANBERRA\strut
-\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright
-0.28\%\strut
-\end{minipage} & \begin{minipage}[t]{0.22\columnwidth}\raggedright
-196037\strut
-\end{minipage} & \begin{minipage}[t]{0.40\columnwidth}\raggedright
-548\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.17\columnwidth}\raggedright
-FENNER\strut
-\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright
-0.23\%\strut
-\end{minipage} & \begin{minipage}[t]{0.22\columnwidth}\raggedright
-202955\strut
-\end{minipage} & \begin{minipage}[t]{0.40\columnwidth}\raggedright
-474\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.17\columnwidth}\raggedright
-WHITLAM\strut
-\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright
-0.06\%\strut
-\end{minipage} & \begin{minipage}[t]{0.22\columnwidth}\raggedright
-152280\strut
-\end{minipage} & \begin{minipage}[t]{0.40\columnwidth}\raggedright
-92\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.17\columnwidth}\raggedright
-GILMORE\strut
-\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright
-0.06\%\strut
-\end{minipage} & \begin{minipage}[t]{0.22\columnwidth}\raggedright
-150436\strut
-\end{minipage} & \begin{minipage}[t]{0.40\columnwidth}\raggedright
-86\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.17\columnwidth}\raggedright
-CALARE\strut
-\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright
-0.01\%\strut
-\end{minipage} & \begin{minipage}[t]{0.22\columnwidth}\raggedright
-161298\strut
-\end{minipage} & \begin{minipage}[t]{0.40\columnwidth}\raggedright
-21\strut
-\end{minipage}\tabularnewline
-\bottomrule
-\end{longtable}
-
-Now consider the socio-demographic \(AusCitizen\) - the proportion of people in the region who are Australian citizens.
-
-\begin{longtable}[]{@{}lll@{}}
-\toprule
-\begin{minipage}[b]{0.18\columnwidth}\raggedright
-DivisionNm\strut
-\end{minipage} & \begin{minipage}[b]{0.21\columnwidth}\raggedright
-AusCitizen (\%): \(D_s\)\strut
-\end{minipage} & \begin{minipage}[b]{0.53\columnwidth}\raggedright
-Estimated Population Allocated to Purple Region: \(\hat{P}_{s,t}\)\strut
-\end{minipage}\tabularnewline
-\midrule
-\endhead
-\begin{minipage}[t]{0.18\columnwidth}\raggedright
-HUME\strut
-\end{minipage} & \begin{minipage}[t]{0.21\columnwidth}\raggedright
-90.02\strut
-\end{minipage} & \begin{minipage}[t]{0.53\columnwidth}\raggedright
-145427\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.18\columnwidth}\raggedright
-RIVERINA\strut
-\end{minipage} & \begin{minipage}[t]{0.21\columnwidth}\raggedright
-89.11\strut
-\end{minipage} & \begin{minipage}[t]{0.53\columnwidth}\raggedright
-39117\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.18\columnwidth}\raggedright
-EDEN-MONARO\strut
-\end{minipage} & \begin{minipage}[t]{0.21\columnwidth}\raggedright
-88.00\strut
-\end{minipage} & \begin{minipage}[t]{0.53\columnwidth}\raggedright
-16358\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.18\columnwidth}\raggedright
-CANBERRA\strut
-\end{minipage} & \begin{minipage}[t]{0.21\columnwidth}\raggedright
-85.48\strut
-\end{minipage} & \begin{minipage}[t]{0.53\columnwidth}\raggedright
-548\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.18\columnwidth}\raggedright
-FENNER\strut
-\end{minipage} & \begin{minipage}[t]{0.21\columnwidth}\raggedright
-83.64\strut
-\end{minipage} & \begin{minipage}[t]{0.53\columnwidth}\raggedright
-474\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.18\columnwidth}\raggedright
-WHITLAM\strut
-\end{minipage} & \begin{minipage}[t]{0.21\columnwidth}\raggedright
-89.52\strut
-\end{minipage} & \begin{minipage}[t]{0.53\columnwidth}\raggedright
-92\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.18\columnwidth}\raggedright
-GILMORE\strut
-\end{minipage} & \begin{minipage}[t]{0.21\columnwidth}\raggedright
-89.03\strut
-\end{minipage} & \begin{minipage}[t]{0.53\columnwidth}\raggedright
-86\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.18\columnwidth}\raggedright
-CALARE\strut
-\end{minipage} & \begin{minipage}[t]{0.21\columnwidth}\raggedright
-87.56\strut
-\end{minipage} & \begin{minipage}[t]{0.53\columnwidth}\raggedright
-21\strut
-\end{minipage}\tabularnewline
-\bottomrule
-\end{longtable}
-
-Then taking a weighted average of \(AusCitizen\) using the estimated population as weights yields \(\hat{C}_{Hume,2016} = 89.65 \%\).
-Repeating this process using the 2011 Census and electoral boundaries yields \(\hat{C}_{Hume,2011} = 91.00 \%\)
-
-Finally, linearly interpolate between 2011 and 2016 to arrive at the 2013 estimate:
-\begin{eqnarray*}
-\hat{C}_{Hume,2013} & = &\frac{3}{5} \cdot \hat{C}_{Hume,2011} + \frac{2}{5} \cdot \hat{C}_{Hume,2016} \\ 
-& = & \frac{3}{5} \cdot 91.00 \% + \frac{2}{5} \cdot 89.65 \% \\ 
-& = & 90.46 \%
-\end{eqnarray*}
-
-This is done for each of the socio-demographic variables, and repeated each of the 2013 electorates.
-
-\hypertarget{modelling}{%
-\chapter{Modelling}\label{modelling}}
-
-\hypertarget{data-pre-processing}{%
-\section{Data pre-processing}\label{data-pre-processing}}
-
+## Data pre-processing
 With socio-demographic information now available for each electorate, each election is joined to the data  corresponding with its two-party preferred vote.
@@
 
-\hypertarget{dimension-reduction}{%
-\subsection{Dimension reduction}\label{dimension-reduction}}
-
-With only \(N = 150\) observations (electorates) in each election and \(p = 65\) socio-demographic variables in each cross-section, any model using all variables would face serious problems with multi-collinearity and over-fitting, likely leading to erroneous conclusions regarding variable significance.
+### Dimension reduction
+With only $N = 150$ observations (electorates) in each election and $p = 65$ socio-demographic variables in each cross-section, any model using all variables would face serious problems with multi-collinearity and over-fitting, likely leading to erroneous conclusions regarding variable significance.
 Therefore a form of dimension reduction is adopted before models are fit.
 
-Socio-demographic variables\footnote{A preliminary step involved removing all age bands, because age is represented by median age, and to remove variables relating to particular denominations of Christianity.} that represent similar information are combined into ``factors'' using principal component analysis (PCA).
+Socio-demographic variables[^2] that represent similar information are combined into "factors" using principal component analysis (PCA).
 The scree plots of the principal components for each election all level off after four components, and the loadings of these four components are similar across the elections.
@@
 
-Consider the \texttt{Incomes} factor as an illustration.
+Consider the `Incomes` factor as an illustration.
 Independent of principal components, we may suspect that median personal income, median household income and median family income are providing similar information about the financial wellbeing of an electorate.
-Their loadings in the first principal component are large (0.20, 0.21 and 0.22 respectively), which provides the evidence needed to combine these variables into a single factor, which is called \texttt{Incomes}.
+Their loadings in the first principal component are large (0.20, 0.21 and 0.22 respectively), which provides the evidence needed to combine these variables into a single factor, which is called `Incomes`.
 
-This process reduces the predictor set to \(p = 30\).
+This process reduces the predictor set to $p = 30$.
 
-\hypertarget{model-framework}{%
-\section{Model framework}\label{model-framework}}
+[^2]: A preliminary step involved removing all age bands, because age is represented by median age, and to remove variables relating to particular denominations of Christianity.
 
+## Model framework
 An identical model specification is used across the six elections, with each election modelled separately.
@@
 
-For each cross-section, let the response variable be the two-party preferred vote in favour of the Liberal party, denoted \(Y\), with \(Y = 70\) representing a 70\% preference for Liberal, 30\% for Labor.
-Although \(Y\) lies in the interval \((0,100)\), observed values are never very close to 0 or 100 (minimum \(24.05 \%\) and maximum \(74.90 \%\)), so there is no need to impose the constraint of \(Y \in [0,100]\).
-Furthermore, the response is found to be spatially correlated in each election (Moran's I test, \(p \le 7\cdot10^{-15}\)).
+For each cross-section, let the response variable be the two-party preferred vote in favour of the Liberal party, denoted $Y$, with $Y = 70$ representing a 70% preference for Liberal, 30% for Labor.
+Although $Y$ lies in the interval $(0,100)$, observed values are never very close to 0 or 100 (minimum $24.05 \%$ and maximum $74.90 \%$), so there is no need to impose the constraint of $Y \in [0,100]$.
+Furthermore, the response is found to be spatially correlated in each election (Moran's I test, $p \le 7\cdot10^{-15}$).
 This is expected, as electorates are aggregate spatial units, and hence the spatial structure of the data must modelled appropriately.
 
-The spatial error model \citep{Anselin88} is chosen because captures spatial heterogeneity by incorporating a spatially structured random effect vector \citep{LeSage2009}.
+The spatial error model [@Anselin88] is chosen because captures spatial heterogeneity by incorporating a spatially structured random effect vector [@LeSage2009].
 In this context, the random effect can be thought of as capturing the unobserved political climate in each electorate, where the climate is correlated with the climate in neighbouring electorates.
@@
 
-Let \(\rho\) be spatial autoregressive coefficient, \(\boldsymbol v\) be a spherical error term, \({\boldsymbol W}\) be a matrix of spatial weights (containing information about the neighbouring regions), \(\boldsymbol X\) be a matrix of socio-demographic covariates, \(\boldsymbol \beta\) be a vector of regression coefficients and \(\boldsymbol a\) be a spatially structured random effect vector.
+Let $\rho$ be spatial autoregressive coefficient, $\boldsymbol v$ be a spherical error term, ${\boldsymbol W}$ be a matrix of spatial weights (containing information about the neighbouring regions), $\boldsymbol X$ be a matrix of socio-demographic covariates, $\boldsymbol \beta$ be a vector of regression coefficients and $\boldsymbol a$ be a spatially structured random effect vector.
 
-\[{\boldsymbol y} = {\boldsymbol X} {\boldsymbol \beta} + {\boldsymbol a}\]
+$${\boldsymbol y} = {\boldsymbol X} {\boldsymbol \beta} + {\boldsymbol a}$$
 and
 
-\[{\boldsymbol a} = \rho {\boldsymbol W} {\boldsymbol a} + {\boldsymbol v}\]
+$${\boldsymbol a} = \rho {\boldsymbol W} {\boldsymbol a} + {\boldsymbol v}$$
 where
 
-\[{\boldsymbol v} \sim N({\boldsymbol 0}, \sigma^2 {\boldsymbol I_n})\].
+$${\boldsymbol v} \sim N({\boldsymbol 0}, \sigma^2 {\boldsymbol I_n})$$.
 
@@
 
-\[{\boldsymbol y} = {\boldsymbol X} {\boldsymbol \beta} + ({\boldsymbol I}_n-\rho {\boldsymbol W})^{-1}{\boldsymbol v}\]
+$${\boldsymbol y} = {\boldsymbol X} {\boldsymbol \beta} + ({\boldsymbol I}_n-\rho {\boldsymbol W})^{-1}{\boldsymbol v}$$
 
-Estimation is done using generalized least squares (using the \texttt{spdep} \texttt{R} package \citep{spdep}).
+Estimation is done using feasible generalized least squares.
 
@@
 
-\emph{insert table here}
+```{r printmodels, eval = F}
+# Add rho and make adjustments to this
+stargazer::stargazer(glsmod01, glsmod04, glsmod07, glsmod10, glsmod13, glsmod16, header = FALSE, 
+  column.sep.width = "1pt", 
+  title = "Estimated model for each of the six elections.",
+  dep.var.labels = "Two-party preferred vote in favor of the Liberal party", 
+  font.size = "scriptsize", digits = 2, 
+  column.labels = c("2001", "2004", "2007", "2010", "2013", "2016"),
+  notes = c("Estimated coefficients for variable named in column one shown for", "election year indicated by column heading, with estimated standard", "deviation for each coefficient shown below in parenthesis.
+Overall", "summary measures for each regression equation are provided in the", "bottom panel."),
+  notes.align = "l")
+```
 
-\hypertarget{results}{%
-\chapter{Results}\label{results}}
-
-\bibliography{book.bib,packages.bib}
-
-\end{document}
+\begin{table}[!htbp] \centering 
+  \caption{Estimated model for each of the six elections.} 
+  \label{} 
+\scriptsize 
+\begin{tabular}{@{\extracolsep{1pt}}lcccccc} 
+\\[-1.8ex]\hline 
+\hline \\[-1.8ex] 
+ & \multicolumn{6}{c}{\textit{Dependent variable:}} \\ 
+\cline{2-7} 
+\\[-1.8ex] & \multicolumn{6}{c}{Two-party preferred vote in favor of the Liberal party} \\ 
+ & 2001 & 2004 & 2007 & 2010 & 2013 & 2016 \\ 
+\\[-1.8ex] & (1) & (2) & (3) & (4) & (5) & (6)\\ 
+\hline \\[-1.8ex] 
+ $\rho$ & 0.44$^{***}$ & 0.15$^{*}$ & $-$0.04 & 0.15 & 0.29 & 0.48$^{***}$ \\ 
+  & (0.15) & (0.17) & (0.19) & (0.17) & (0.16) & (0.16) \\ 
+  & & & & & & \\ 
+ AusCitizen & $-$2.77 & $-$2.83 & 0.67 & 1.00 & $-$0.87 & $-$1.53 \\ 
+  & (2.51) & (2.63) & (2.57) & (2.93) & (2.90) & (2.95) \\ 
+  & & & & & & \\ 
+ Born\_Asia & 2.63 & 0.74 & $-$1.37 & $-$7.58$^{***}$ & $-$2.56 & $-$2.13 \\ 
+  & (3.03) & (2.55) & (1.96) & (2.68) & (2.58) & (2.20) \\ 
+  & & & & & & \\ 
+ Born\_MidEast & 1.92 & 3.22$^{*}$ & 2.92$^{*}$ & 1.79 & 1.70 & $-$0.82 \\ 
+  & (1.53) & (1.78) & (1.51) & (1.54) & (1.54) & (1.56) \\ 
+  & & & & & & \\ 
+ Born\_SE\_Europe & 1.02 & 0.04 & $-$0.20 & $-$1.72 & $-$1.25 & $-$1.63 \\ 
+  & (2.34) & (2.05) & (1.17) & (1.34) & (1.15) & (0.99) \\ 
+  & & & & & & \\ 
+ Born\_UK & $-$0.49 & $-$0.87 & 0.29 & $-$0.29 & $-$0.35 & $-$1.37 \\ 
+  & (1.04) & (0.90) & (0.80) & (1.05) & (1.00) & (0.98) \\ 
+  & & & & & & \\ 
+ BornElsewhere & $-$1.08 & $-$0.51 & 5.01 & 9.78$^{**}$ & 4.04 & 1.74 \\ 
+  & (3.40) & (3.67) & (3.59) & (4.47) & (4.46) & (4.52) \\ 
+  & & & & & & \\ 
+ Buddhism & 1.50 & 2.84$^{*}$ & 0.90 & 2.41$^{*}$ & 1.09 & 1.51 \\ 
+  & (1.94) & (1.61) & (1.25) & (1.40) & (1.25) & (1.24) \\ 
+  & & & & & & \\ 
+ Christianity & 0.53 & 3.44 & $-$1.93 & $-$0.60 & 3.08 & 5.17$^{**}$ \\ 
+  & (4.58) & (4.19) & (3.45) & (3.44) & (2.79) & (2.52) \\ 
+  & & & & & & \\ 
+ CurrentlyStudying & $-$1.14 & 1.69 & 4.03$^{***}$ & 3.28$^{***}$ & 1.73 & $-$0.09 \\ 
+  & (1.01) & (1.09) & (1.12) & (1.21) & (1.17) & (1.13) \\ 
+  & & & & & & \\ 
+ DeFacto & $-$6.42$^{***}$ & $-$5.04$^{**}$ & $-$6.55$^{***}$ & $-$8.60$^{***}$ & $-$9.73$^{***}$ & $-$11.02$^{***}$ \\ 
+  & (1.69) & (2.00) & (1.90) & (2.57) & (2.62) & (2.55) \\ 
+  & & & & & & \\ 
+ DiffAddress & 3.86$^{***}$ & 4.63$^{***}$ & 3.40$^{***}$ & 4.72$^{***}$ & 5.02$^{***}$ & 6.93$^{***}$ \\ 
+  & (0.95) & (1.03) & (0.88) & (1.60) & (1.71) & (1.50) \\ 
+  & & & & & & \\ 
+ Distributive & 1.67 & 2.75$^{**}$ & 2.25$^{**}$ & 2.88$^{**}$ & 2.76$^{**}$ & 1.65 \\ 
+  & (1.08) & (1.13) & (1.02) & (1.30) & (1.25) & (1.17) \\ 
+  & & & & & & \\ 
+ Education & 0.62 & 1.94 & $-$0.77 & $-$1.82 & $-$0.90 & $-$3.04 \\ 
+  & (2.23) & (2.86) & (2.89) & (3.35) & (3.04) & (2.99) \\ 
+  & & & & & & \\ 
+ Extractive & 4.69$^{***}$ & 5.31$^{***}$ & 5.81$^{***}$ & 6.97$^{***}$ & 6.09$^{***}$ & 6.81$^{***}$ \\ 
+  & (1.36) & (1.23) & (1.16) & (1.48) & (1.42) & (1.31) \\ 
+  & & & & & & \\ 
+ FamHouseSize & $-$2.94 & $-$5.11$^{*}$ & $-$9.67$^{***}$ & $-$8.22$^{**}$ & $-$6.16$^{*}$ & $-$4.35 \\ 
+  & (2.27) & (2.78) & (2.66) & (3.28) & (3.15) & (2.86) \\ 
+  & & & & & & \\ 
+ Incomes & 2.99$^{*}$ & 2.95 & 4.36$^{*}$ & 2.73 & 4.19 & 8.46$^{***}$ \\ 
+  & (1.76) & (2.60) & (2.56) & (3.06) & (2.68) & (2.57) \\ 
+  & & & & & & \\ 
+ Indigenous & 3.97$^{***}$ & 5.62$^{***}$ & 6.81$^{***}$ & 5.90$^{***}$ & 6.12$^{***}$ & 4.47$^{***}$ \\ 
+  & (1.26) & (1.35) & (1.17) & (1.49) & (1.49) & (1.38) \\ 
+  & & & & & & \\ 
+ Islam & 0.15 & 0.61 & $-$1.21 & $-$1.04 & 1.20 & 2.64 \\ 
+  & (2.03) & (2.00) & (1.80) & (1.94) & (1.71) & (1.72) \\ 
+  & & & & & & \\ 
+ Judaism & 1.02 & 1.74 & 0.01 & 0.46 & 1.60$^{*}$ & 1.96$^{***}$ \\ 
+  & (1.38) & (1.26) & (1.00) & (1.05) & (0.85) & (0.71) \\ 
+  & & & & & & \\ 
+ ManagerAdminClericalSales & 2.10$^{***}$ & 3.28$^{***}$ & 5.21$^{***}$ & 4.32$^{***}$ & 3.99$^{***}$ & 4.62$^{***}$ \\ 
+  & (0.70) & (0.85) & (0.78) & (1.00) & (0.99) & (1.03) \\ 
+  & & & & & & \\ 
+ Married & 2.27 & 5.50$^{**}$ & 5.06$^{*}$ & 5.07$^{*}$ & 1.46 & $-$1.39 \\ 
+  & (2.26) & (2.79) & (2.59) & (2.91) & (2.71) & (2.64) \\ 
+  & & & & & & \\ 
+ MedianAge & 1.54 & 3.73$^{**}$ & 1.86 & 2.08 & 2.28 & 2.94$^{*}$ \\ 
+  & (1.30) & (1.55) & (1.65) & (2.08) & (1.88) & (1.63) \\ 
+  & & & & & & \\ 
+ NoReligion & 1.06 & 3.43 & $-$0.31 & 0.25 & 3.53 & 5.76$^{**}$ \\ 
+  & (3.85) & (3.45) & (2.90) & (2.98) & (2.58) & (2.59) \\ 
+  & & & & & & \\ 
+ OneParent\_House & $-$1.00 & 1.37 & 1.66 & 0.22 & 0.11 & $-$0.85 \\ 
+  & (1.34) & (1.49) & (1.38) & (1.61) & (1.46) & (1.39) \\ 
+  & & & & & & \\ 
+ OtherLanguageHome & $-$10.40$^{**}$ & $-$6.76 & $-$6.02 & $-$4.37 & $-$4.69 & 1.27 \\ 
+  & (4.97) & (5.61) & (4.44) & (5.70) & (5.50) & (4.95) \\ 
+  & & & & & & \\ 
+ PropertyOwned & $-$0.76 & $-$0.92 & 0.78 & $-$1.17 & 0.68 & 1.76 \\ 
+  & (1.19) & (1.21) & (1.09) & (1.49) & (1.43) & (1.38) \\ 
+  & & & & & & \\ 
+ RentLoanPrice & $-$0.31 & $-$1.54 & $-$0.54 & 2.12 & 1.01 & $-$0.70 \\ 
+  & (1.50) & (1.68) & (1.48) & (1.96) & (1.96) & (2.02) \\ 
+  & & & & & & \\ 
+ SocialServ & 2.42$^{*}$ & 1.89 & 1.77 & 1.91 & 0.94 & 2.90$^{***}$ \\ 
+  & (1.27) & (1.27) & (1.14) & (1.35) & (1.22) & (1.10) \\ 
+  & & & & & & \\ 
+ Transformative & 2.99$^{**}$ & 5.74$^{***}$ & 5.20$^{***}$ & 3.28$^{*}$ & 3.74$^{**}$ & 4.75$^{***}$ \\ 
+  & (1.41) & (1.59) & (1.47) & (1.74) & (1.57) & (1.41) \\ 
+  & & & & & & \\ 
+ Unemployment & $-$3.00$^{**}$ & $-$5.09$^{***}$ & $-$2.88$^{**}$ & $-$2.42 & $-$1.22 & 0.36 \\ 
+  & (1.28) & (1.51) & (1.41) & (1.73) & (1.54) & (1.42) \\ 
+  & & & & & & \\ 
+ Constant & 50.78$^{***}$ & 52.63$^{***}$ & 47.32$^{***}$ & 49.92$^{***}$ & 53.48$^{***}$ & 50.45$^{***}$ \\ 
+  & (0.67) & (0.45) & (0.35) & (0.51) & (0.56) & (0.74) \\ 
+  & & & & & & \\ 
+\hline \\[-1.8ex] 
+Observations & 150 & 150 & 150 & 150 & 150 & 150 \\ 
+Log Likelihood & $-$398.33 & $-$403.34 & $-$398.73 & $-$418.08 & $-$405.41 & $-$399.52 \\ 
+Akaike Inf. Crit. & 860.67 & 870.68 & 861.46 & 900.16 & 874.83 & 863.03 \\ 
+Bayesian Inf. Crit. & 949.60 & 959.61 & 950.40 & 989.10 & 963.76 & 951.96 \\ 
+\hline 
+\hline \\[-1.8ex] 
+\textit{Note:}  & \multicolumn{6}{l}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
+ & \multicolumn{6}{l}{Estimated coefficients for variable named in column one shown for} \\ 
+ & \multicolumn{6}{l}{election year indicated by column heading, with estimated standard} \\ 
+ & \multicolumn{6}{l}{deviation for each coefficient shown below in parenthesis. Overall} \\ 
+ & \multicolumn{6}{l}{summary measures for each regression equation are provided in the} \\ 
+ & \multicolumn{6}{l}{bottom panel.} \\ 
+\end{tabular} 
+\end{table}

### 2019-03-26 04-results.Rmd
merge chapters 2 and 3 into data chapter

--- 2019-03-26 05-results.Rmd
@@
 
-## Residuals
+## A closer look at the residuals
 
@@
     glsmod01$my_data %>% left_join(tpp01 %>% dplyr::select(DivisionNm, UniqueID, StateAb), by = "DivisionNm"))
-) %>% dplyr::select(Residuals, DivisionNm, UniqueID, StateAb, year, LNP_Percent, gls_residuals)
+) 
+#%>% dplyr::select(Residuals, DivisionNm, UniqueID, StateAb, year, LNP_Percent, gls_residuals)
 ```
 
-## Residuals by state
+### Residuals by state
 
@@
 
-```{r resstate, fig.cap = "Boxplot of residuals by state.", fig.out="75%"}
+```{r resstate, fig.cap = "Boxplot of residuals by state.", out.width="75%"}
 all_resids %>% 
@@
   geom_boxplot() + 
-  geom_jitter(aes(label = DivisionNm, shape = year), alpha = 0.4) +
+  geom_jitter(alpha = 0.4) +
+  #geom_jitter(aes(label = DivisionNm, shape = year), alpha = 0.4) +
   #facet_wrap(~year) + 
@@
 
-## Electorates with largest residuals
+### Electorates with largest residuals
 
-```{r}
+The electorates Capricornia (QLD), Leichhardt (QLD) and Riverina (NSW) have the three largest maximum absolute residuals across the six elections.
+Capricornia is a swing electorate, which strongly supported Labor between 2001 and 2010, before being narrowly won by Liberal in the last two elections.
+It has many people working in extractive jobs, an above-average Christian population, and above-average proportion of de facto relationships.
+The model tends to overpredict its two-party preferred vote (in favour of the Liberal party).
+Leichhardt is an area encompassing the northernmost tip of Queensland and consists of mainly small indigenous communities and uninhabited rural areas, which was won by the Liberal party in all but the 2007 election.
+It too has an above-average proportion of de facto relationships, along with an above-average proportion of people moving houses and a high proportion of single parent households.
+The model consistently underpredicts its two-party vote.
+Riverina is a rural electorate of NSW, and has been a very safe Liberal (National) seat since 1980.
+It has a relatively monocultural population consisting of a high proportion of Australian citizens and Christians, an above-average amount of extractive jobs and below average education levels.
+Its Liberal preference is also underpredicted.
+
+```{r eval = F}
+all_resids %>% 
+  group_by(UniqueID) %>% 
+  mutate(max_res = max(abs(Residuals))) %>% 
+  arrange(-max_res) %>% 
+  filter(max_res > 13.4) %>% 
+  dplyr::select(DivisionNm, LNP_Percent, year, everything())
+```
+
+```{r eval = F}
 library(plotly)
@@
   group_by(UniqueID) %>% 
-  mutate(mean_res = mean(abs(Residuals))) %>% 
-  filter(mean_res > 7) %>%
-  ungroup() %>% View
+  mutate(mean_res = mean(abs(Residuals)), min_res = min(abs(Residuals))) %>%
+  filter(min_res > 5) %>%
+  ungroup() %>% dplyr::select(Residuals, DivisionNm, year, min_res, mean_res, StateAb) %>% View
   ggplot(aes(x = year, y = Residuals, group = DivisionNm, col = DivisionNm, label = DivisionNm)) + 
@@
 ```
-
-Capricornia, Sydney, Shortland
-
-Residuals by state, by year
-
-```{r}
-coef_df %>% 
-  filter(variable != "Intercept") %>% 
-  mutate(upper95 = estimate + 1.96*se, lower95 = estimate - 1.96*se) %>% 
-  ggplot() +
-  geom_point(aes(x = year, y = estimate, col = factor(p < 0.05)), size = 2) +
-  geom_linerange(aes(x = year, ymin = lower95, ymax = upper95, col = factor(p < 0.05)), size = 1) + 
-  geom_hline(aes(yintercept = 0), alpha = 0.5, size = 1) + 
-  facet_wrap(~variable, scales = "free", ncol = 5) +
-  scale_color_manual(values = c("grey50", "black")) + 
-  theme(axis.text = element_text(size = 6), strip.text = element_text(size = 6)) +
-  labs(x = "", y = "Year") +
-  scale_x_continuous(breaks = c(2001, 2004, 2007, 2010, 2013, 2016)) +
-  guides(col = F)
-```

### 2019-03-26 05-acknowledgements.Rmd
add small conclusion and reference to eechidna - maybe remove software section because its just doubling up

# Concluding comments

Across the past six elections, the spatial error models reveal that the electoral socio-demographics driving the two-party preferred vote have remained relatively steady, with industry and type of work being particularly influential, along with the type of relationships and household mobility, among others. The neighbourhood (spatial) effects were only found to be significant in the 2001 and 2016 elections. Surprisingly, unemployment, incomes and family/household sizes were significant in only some elections, and other characteristics like education were not significant at all.

The cross-sectional data set of Census and election data were obtained by collecting information from the ABS and AEC, and then applying a spatio-temporal imputation process, as detailed in section \@ref(data). This data is available in the `eechidna` `R` package, which can be downloaded from CRAN. All associated code to produce the data in the package from raw files can be found in the github repository (https://github.com/ropenscilabs/eechidna), and vignettes on how to use the data are located on the github page (https://ropenscilabs.github.io/eechidna/).

Furthermore, this paper is generated using the `knitr` [@knitr] package, so code to reproduce the results are contained in the `RMarkdown` file.

### 2019-03-26 06-software.Rmd
add small conclusion and reference to eechidna - maybe remove software section because its just doubling up

# Software

`eechidna` (Exploring Election and Census Highly Informative Data Nationally for Australia) is an R package that makes it easy to look at the data from the Australian Federal elections and Censuses that occurred between 2001 and 2016. The data in this package includes voting results for each polling booth and electoral division (electorate), Census information for each electorate and maps of the electorates that were in place at the time of each event. All data in this package is obtained from the Australian Electoral Commission, the Australian Bureau of Statistics and the Australian Government. Census data is imputed for the election years in which a Census does not occur via a spatio-temporal imputation method.

The Census and election datasets in the package can be linked by electorate using the unique identifier variable `UniqueID`. Vignettes describing how to use the data and how the data was obtained, can also be found on https://ropenscilabs.github.io/eechidna/.

### 2019-03-26 01-intro.Rmd
change data ref

@@
 To obtain appropriate Census data for these four elections, electoral socio-demographics are imputed using a spatio-temporal imputation that combines areal interpolation [@Goodchild1993] and linear time-interpolation.
-The process of collecting and wrangling the raw data is outlined in section \@ref(data), and the imputation process is detailed in section \@ref(imputation).
-All data and associated documentation relating to this procedure are available in the `eechidna` `R` package [@eechidna], providing a resource for future analysis.
+Collecting and wrangling the raw data, along with the imputation process, are detailed in section \@ref(data).
+All data and associated documentation relating to this procedure are available in the `eechidna` `R` package [@eechidna], providing a resource for future analysis .
 

### 2019-03-26 03-modelling.Rmd
add residual standard error GLS to table

@@
 
-Table 4.1 details the resultant estimated model coefficients and their estimated standard errors for each of the six elections. These are interpreted in the next section.
+Table 3.1 details the resultant estimated model coefficients and their estimated standard errors for each of the six elections. These are interpreted in the next section.
 
 ```{r printmodels, eval = F}
-# Add rho and make adjustments to this
+# Add rho and GLS residual standard error
 stargazer::stargazer(glsmod01, glsmod04, glsmod07, glsmod10, glsmod13, glsmod16, header = FALSE, 
@@
 Observations & 150 & 150 & 150 & 150 & 150 & 150 \\ 
+Residual Standard Error (GLS) & 4.55 & 4.69 & 4.44 & 5.31 & 4.85 & 4.65 \\ 
 Log Likelihood & $-$398.33 & $-$403.34 & $-$398.73 & $-$418.08 & $-$405.41 & $-$399.52 \\ 

### 2019-03-26 04-results.Rmd
make rho visualization look tidy

@@
 ## Spatial autoregressive parameter
-```{r rhovis}
+```{r rhovis, out.height= "15%", fig.cap = "Estimates of the spatial autoregressive parameter for each of the six elections, with a 95% confidence interval.
+Only in 2001 and 2016 is there a significant spatial component."}
 rho_df <- bind_rows(glsmod16$rho_df %>% mutate(year = "2016"), 
@@
   scale_color_manual(values = c("grey50", "black")) +
-  guides(col = F)
+  labs(x = "Election year", y = "Estimate of spatial autoregressive parameter") +
+  lims(y = c(-1,1)) +
+  guides(col = F) +
+  coord_fixed(ratio = 1)
 ```
@@
   x_mat <- my_model$my_data %>% 
-    dplyr::select(-LNP_Percent) %>% 
+    dplyr::select(-c(LNP_Percent, year, DivisionNm)) %>% 
     mutate(Intercept = 1) %>% 
@@
   # Lambda matrix (FGLS)
-  x <- round(seq(min(x_mat[, varname]), max(x_mat[, varname]), 0.025), 3)
+  x <- round(seq(min(as.numeric(x_mat[, varname])), max(as.numeric(x_mat[, varname])), 0.025), 3)
   lambda_mat <- data.frame(matrix(0, nrow = length(x), ncol = ncol(x_mat)))
@@
 
-```{r plotincomes, fig.cap = "Effect of Incomes", out.width="60%"}
+```{r plotincomes, fig.cap = "Effect of Incomes", out.width="100%"}
 p <- grid_visreg("Incomes", xvar_title = "Incomes", xlimits = c(-1.5, 3.6), ylimits = c(30, 95))

### 2019-03-26 _book/electoral-modelling-draft.tex
update cache

@@
 To obtain appropriate Census data for these four elections, electoral socio-demographics are imputed using a spatio-temporal imputation that combines areal interpolation \citep{Goodchild1993} and linear time-interpolation.
-The process of collecting and wrangling the raw data is outlined in section \ref{data}, and the imputation process is detailed in section \ref{imputation}.
-All data and associated documentation relating to this procedure are available in the \texttt{eechidna} \texttt{R} package \citep{eechidna}, providing a resource for future analysis.
+Collecting and wrangling the raw data, along with the imputation process, are detailed in section \ref{data}.
+All data and associated documentation relating to this procedure are available in the \texttt{eechidna} \texttt{R} package \citep{eechidna}, providing a resource for future analysis .
 
@@
 \hypertarget{data}{%
-\chapter{Data collection}\label{data}}
+\chapter{Data collection, wrangling and imputation}\label{data}}
 
-\hypertarget{voting-data}{%
-\section{Voting data}\label{voting-data}}
+\hypertarget{collecting-the-data}{%
+\section{Collecting the data}\label{collecting-the-data}}
 
@@
 
-\hypertarget{census-data}{%
-\section{Census data}\label{census-data}}
-
 Socio-demographic variables are derived from the Census of Population and Housing (Census), which is a survey of every household in Australia, recording information such as age, gender, ethnicity, education level and income.
@@
 From this aggregate information, 67 socio-demographic variables are computed for each of the electorates.
-
-\hypertarget{downloading-and-wrangling}{%
-\section{Downloading and wrangling}\label{downloading-and-wrangling}}
 
@@
 
-\hypertarget{imputation}{%
-\chapter{Joining Census and election data}\label{imputation}}
+\hypertarget{joining-census-and-election-data}{%
+\section{Joining Census and election data}\label{joining-census-and-election-data}}
 
 \hypertarget{differences-between-census-and-election-data}{%
-\section{Differences between Census and election data}\label{differences-between-census-and-election-data}}
+\subsection{Differences between Census and election data}\label{differences-between-census-and-election-data}}
 
@@
 \hypertarget{spatio-temporal-imputation}{%
-\section{Spatio-temporal imputation}\label{spatio-temporal-imputation}}
+\subsection{Spatio-temporal imputation}\label{spatio-temporal-imputation}}
 
@@
 \hypertarget{applied}{%
-\section{Applied}\label{applied}}
+\subsection{Applied}\label{applied}}
 
@@
 
-Table 4.1 details the resultant estimated model coefficients and their estimated standard errors for each of the six elections. These are interpreted in the next section.
+Table 3.1 details the resultant estimated model coefficients and their estimated standard errors for each of the six elections. These are interpreted in the next section.
 
@@
 
-\begin{center}\includegraphics{electoral-modelling-draft_files/figure-latex/rhovis-1} \end{center}
+\textbackslash begin\{figure\}{[}h{]}
+
+\{\centering \includegraphics[height=0.15\textheight]{electoral-modelling-draft_files/figure-latex/rhovis-1}
+
+\}
+
+\textbackslash caption\{Estimates of the spatial autoregressive parameter for each of the six elections, with a 95\% confidence interval.
+Only in 2001 and 2016 is there a significant spatial component.\}\label{fig:rhovis}
+\textbackslash end\{figure\}
 
 The spatial autoregressive coefficient \(\rho\) is positive and significant in only the 2001 and 2016 elections, meaning that in these elections, an electorate's political climate was affected by the attitudes of it's neighbours.
-Conversely, in the other four elections, it appears that after accounting for electoral socio-demographics, there is no spatial effect, meaning electorates effectively acted independently.
+Conversely, in the other four elections, it appears that after accounting for electoral socio-demographics, there is no significant spatial effect, meaning electorates effectively acted independently.
 
@@
 To investigate the socio-demographics that have a strong effect on the two-party preferred vote, partial residual plots are used.
-These show the direction, size and significance of an estimated effect, as the slope of the prediction line matches the estimated coefficient, and the shaded region represents a 95\% confidence band.
+These show the direction, size and significance of an estimated effect - the slope of the prediction line matches the estimated coefficient, and the shaded region represents a 95\% confidence band.
 If a horizontal line can be drawn through the confidence band, then the effect is insignificant.
@@
 Typically the Labor party campaigns on more progressive policies, which often include tax reform that adversely affects higher income earners, and more generous social assistance programs.
-Although topical in the lead up to an election, incomes only significantly affected electoral two-party preference in 2016 (see Figure \ref{fig:plotincomes}), with higher income areas associated with support for the Liberal party.
+Despite this, incomes have only significantly affected the electoral two-party preference in 2016 (see Figure \ref{fig:plotincomes}), with higher income areas associated with support for the Liberal party.
 In the prior elections, the estimated effect is still positive, albeit insignificant.
-Unemployment on the other hand, was influential in 2001, 2004 and 2007 elections(Figure \ref{fig:plotunemploy}), with electorates with higher unemployment favouring Labor.
+Unemployment on the other hand, was influential in 2001, 2004 and 2007 elections (Figure \ref{fig:plotunemploy}), with electorates with higher unemployment favouring Labor.
 This estimated effect has become weaker over time, with unemployment having virtually no effect in 2016.
@@
 
-{\centering \includegraphics[width=0.6\linewidth]{electoral-modelling-draft_files/figure-latex/plotincomes-1}
+{\centering \includegraphics[width=1\linewidth]{electoral-modelling-draft_files/figure-latex/plotincomes-1}
 
@@
 Furthermore, electorates with more workers in construction or manufacturing industries (\texttt{Transformative}), as well as wholesale trade, retail trade, transport, post or warehousing (\texttt{Distributive}), are also linked with the Liberal party.
-However, these effects are not as strong, and \texttt{Transformative} is marginally insignificant in the 2010 election, whereas \texttt{Distributive} was not a key socio-demographic in 2001 or 2016.
+However, these effects are not as strong, and \texttt{Transformative} was marginally insignificant in the 2010 election, whereas \texttt{Distributive} was not a key socio-demographic in 2001 or 2016.
 These effects are shown in Figures \ref{fig:plottransformative} and \ref{fig:plotdistributive}.
@@
 
-Electorates with a higher proportion of people that have recently moved house (meaning in last five years) are more likely to support the Liberal party, with this effect becoming stronger since the 2010 election.
+In each of the six elections, electorates with a higher proportion of people that have recently moved house (meaning in last five years) were more likely to support the Liberal party, with this effect becoming stronger since 2010.
 Having controlled for characteristics of house ownership and rental prices (via the factors \texttt{PropertyOwned} and \texttt{RentLoan} respectively), this is effect is somewhat surprising.
@@
 
-\emph{Potentially remove plots}
+\hypertarget{a-note-on-similar-variables}{%
+\subsection{A note on similar variables}\label{a-note-on-similar-variables}}
 
-\hypertarget{residuals}{%
-\section{Residuals}\label{residuals}}
+Many of the Census variables represent similar information, which is why factors were created and some variables were removed.
+However, there still remain some variables which are closely related.
+For example, \texttt{OtherLanguageHome} is the proportion of households that speak a language other than English at home, which is likely to be related to the proportion of people born overseas (a sum of the variables with a \texttt{Born} prefix).
+In 2001, the coefficient estiamte for \texttt{OtherLanguageHome} is significant and negative, and all \texttt{Born} variables are insignificant.
+If \texttt{OtherLanguageHome} is removed from the model, populations born in South Eastern Europe (\texttt{Born\_SE\_Europe}) absorb the negative effect (\(p = 0.0011\)).
 
-Largest residuals
+\hypertarget{a-closer-look-at-the-residuals}{%
+\section{A closer look at the residuals}\label{a-closer-look-at-the-residuals}}
 
-Residuals by state
+\hypertarget{residuals-by-state}{%
+\subsection{Residuals by state}\label{residuals-by-state}}
 
-\begin{center}\includegraphics{electoral-modelling-draft_files/figure-latex/unnamed-chunk-2-1} \end{center}
+It is often hypothesized that states have systematic differences that cause their electorates to vote differently.
+Boxplots of residuals grouped by state reveal that only Tasmania and the Australian Capital Territory appear to have a state-specific effect that is not captured by the models.
+These states appear to have a bias towards Labor.
+There are few electorates in these states (five and two, respectively), so this might be due to incumbent effects rather than an actual state-specific bias.
+
+\begin{figure}[h]
+
+{\centering \includegraphics[width=0.75\linewidth]{electoral-modelling-draft_files/figure-latex/resstate-1}
+
+}
+
+\caption{Boxplot of residuals by state.}\label{fig:resstate}
+\end{figure}
+
+\hypertarget{electorates-with-largest-residuals}{%
+\subsection{Electorates with largest residuals}\label{electorates-with-largest-residuals}}
+
+The electorates Capricornia (QLD), Leichhardt (QLD) and Riverina (NSW) have the three largest maximum absolute residuals across the six elections.
+Capricornia is a swing electorate, which strongly supported Labor between 2001 and 2010, before being narrowly won by Liberal in the last two elections.
+It has many people working in extractive jobs, an above-average Christian population, and above-average proportion of de facto relationships.
+The model tends to overpredict its two-party preferred vote (in favour of the Liberal party).
+Leichhardt is an area encompassing the northernmost tip of Queensland and consists of mainly small indigenous communities and uninhabited rural areas, which was won by the Liberal party in all but the 2007 election.
+It too has an above-average proportion of de facto relationships, along with an above-average proportion of people moving houses and a high proportion of single parent households.
+The model consistently underpredicts its two-party vote.
+Riverina is a rural electorate of NSW, and has been a very safe Liberal (National) seat since 1980.
+It has a relatively monocultural population consisting of a high proportion of Australian citizens and Christians, an above-average amount of extractive jobs and below average education levels.
+Its Liberal preference is also underpredicted.
+
+\hypertarget{concluding-comments}{%
+\chapter{Concluding comments}\label{concluding-comments}}
+
+Across the past six elections, the spatial error models reveal that the electoral socio-demographics driving the two-party preferred vote have remained relatively steady, with industry and type of work being particularly influential, along with the type of relationships and household mobility, among others.
+The neighbourhood (spatial) effects were only found to be significant in the 2001 and 2016 elections.
+Surprisingly, unemployment, incomes and family/household sizes were significant in only some elections, and other characteristics like education were not significant at all.
+
+The cross-sectional data set of Census and election data were obtained by collecting information from the ABS and AEC, and then applying a spatio-temporal imputation process, as detailed in section \ref{data}.
+This data is available in the \texttt{eechidna} \texttt{R} package, which can be downloaded from CRAN.
+All associated code to produce the data in the package from raw files can be found in the github repository (\url{https://github.com/ropenscilabs/eechidna}), and vignettes on how to use the data are located on the github page (\url{https://ropenscilabs.github.io/eechidna/}).
+
+Furthermore, this paper is generated using the \texttt{knitr} \citep{knitr} package, so code to reproduce the results are contained in the \texttt{RMarkdown} file.
+
+\hypertarget{software}{%
+\chapter{Software}\label{software}}
+
+\texttt{eechidna} (Exploring Election and Census Highly Informative Data Nationally for Australia) is an R package that makes it easy to look at the data from the Australian Federal elections and Censuses that occurred between 2001 and 2016.
+The data in this package includes voting results for each polling booth and electoral division (electorate), Census information for each electorate and maps of the electorates that were in place at the time of each event.
+All data in this package is obtained from the Australian Electoral Commission, the Australian Bureau of Statistics and the Australian Government.
+Census data is imputed for the election years in which a Census does not occur via a spatio-temporal imputation method.
+
+The Census and election datasets in the package can be linked by electorate using the unique identifier variable \texttt{UniqueID}.
+Vignettes describing how to use the data and how the data was obtained, can also be found on \url{https://ropenscilabs.github.io/eechidna/}.
 

### 2019-03-26 04-results.Rmd
fix gitignore and % sign in fig cap

@@
 ## Spatial autoregressive parameter
-```{r rhovis, out.height= "15%", fig.cap = "Estimates of the spatial autoregressive parameter for each of the six elections, with a 95% confidence interval.
-Only in 2001 and 2016 is there a significant spatial component."}
+
+The spatial autoregressive coefficient $\rho$ is positive and significant in only the 2001 and 2016 elections (Figure \@ref(fig:rhovis)), meaning that in these elections, an electorate's political climate was affected by the attitudes of it's neighbours.
+Conversely, in the other four elections, it appears that after accounting for electoral socio-demographics, there is no significant spatial effect, meaning electorates effectively acted independently.
+
+```{r rhovis, fig.cap="Estimates of the spatial autoregressive parameter for each of the six elections, with a 95\\% confidence interval.
+Only in 2001 and 2016 is there a significant spatial component", out.height='20%'}
 rho_df <- bind_rows(glsmod16$rho_df %>% mutate(year = "2016"), 
@@
 ```
-
-The spatial autoregressive coefficient $\rho$ is positive and significant in only the 2001 and 2016 elections, meaning that in these elections, an electorate's political climate was affected by the attitudes of it's neighbours.
-Conversely, in the other four elections, it appears that after accounting for electoral socio-demographics, there is no significant spatial effect, meaning electorates effectively acted independently.
 

### 2019-03-26 _book/electoral-modelling-draft.tex
fix gitignore and % sign in fig cap

@@
   \label{} 
-\scriptsize 
+\tiny 
 \begin{tabular}{@{\extracolsep{1pt}}lcccccc} 
@@
 
-\textbackslash begin\{figure\}{[}h{]}
+The spatial autoregressive coefficient \(\rho\) is positive and significant in only the 2001 and 2016 elections (Figure \ref{fig:rhovis}), meaning that in these elections, an electorate's political climate was affected by the attitudes of it's neighbours.
+Conversely, in the other four elections, it appears that after accounting for electoral socio-demographics, there is no significant spatial effect, meaning electorates effectively acted independently.
 
-\{\centering \includegraphics[height=0.15\textheight]{electoral-modelling-draft_files/figure-latex/rhovis-1}
+\begin{figure}[h]
 
-\}
+{\centering \includegraphics[height=0.2\textheight]{electoral-modelling-draft_files/figure-latex/rhovis-1}
 
-\textbackslash caption\{Estimates of the spatial autoregressive parameter for each of the six elections, with a 95\% confidence interval.
-Only in 2001 and 2016 is there a significant spatial component.\}\label{fig:rhovis}
-\textbackslash end\{figure\}
+}
 
-The spatial autoregressive coefficient \(\rho\) is positive and significant in only the 2001 and 2016 elections, meaning that in these elections, an electorate's political climate was affected by the attitudes of it's neighbours.
-Conversely, in the other four elections, it appears that after accounting for electoral socio-demographics, there is no significant spatial effect, meaning electorates effectively acted independently.
+\caption{Estimates of the spatial autoregressive parameter for each of the six elections, with a 95\% confidence interval.
+Only in 2001 and 2016 is there a significant spatial component}\label{fig:rhovis}
+\end{figure}
 

### 2019-03-27 04-results.Rmd
add country wide trend, diversity, cooks distance,

@@
 The spatial autoregressive coefficient $\rho$ is positive and significant in only the 2001 and 2016 elections (Figure \@ref(fig:rhovis)), meaning that in these elections, an electorate's political climate was affected by the attitudes of it's neighbours.
-Conversely, in the other four elections, it appears that after accounting for electoral socio-demographics, there is no significant spatial effect, meaning electorates effectively acted independently.
+Conversely, in the other four elections, the spatial effect weakens to become marginally insignficant.
+In these years, it appears that the spatial component does not explain anything over and above the electoral socio-demographics, meaning electorates effectively acted more independently.
 
@@
 rho_df %>% ggplot() + 
-  geom_point(aes(x = year, y = estimate, col = factor(p < 0.05)), size = 3) +
-  geom_linerange(aes(x = year, ymin = lower95, ymax = upper95, col = factor(p < 0.05)), size = 1.5) + 
+  geom_point(aes(x = year, y = estimate), size = 3) +
+  geom_linerange(aes(x = year, ymin = lower95, ymax = upper95), size = 1.5) + 
   geom_hline(aes(yintercept = 0), alpha = 0.5, size = 1) +
@@
   coord_fixed(ratio = 1)
+```
+
+## Country-wide trend
+Since all socio-demographics have been standardized to have a mean of zero and a variance of one, the intercept in each model can be interpreted as the estimated two-party preferred vote for an electorate with mean characteristics[^3].
+The baseline of party preference has varied over the elections, with the biggest swing occuring in the 2007 election where the mean electorate shifted more than five percentage points in favour of the Labor party.
+
+[^3]: Mean of all variables aside from Judaism, Indigenous, Islam and Buddhism, where it assumes the mean of the log value.
+
+```{r plotintercept, fig.cap = "Estimated intercept for each election, which represents the two-party preferred vote for an electorate with mean characteristics.", out.height="20%"}
+intercept_df <- bind_rows(data.frame(summary(glsmod16)$tTable)[1, ]  %>% mutate(year = "2016"), 
+  data.frame(summary(glsmod13)$tTable)[1, ] %>% mutate(year = "2013"),
+  data.frame(summary(glsmod10)$tTable)[1, ] %>% mutate(year = "2010"), 
+  data.frame(summary(glsmod07)$tTable)[1, ] %>% mutate(year = "2007"),
+  data.frame(summary(glsmod04)$tTable)[1, ] %>% mutate(year = "2004"), 
+  data.frame(summary(glsmod01)$tTable)[1, ] %>% mutate(year = "2001")) %>% 
+  rename(se = `Std.Error`, estimate = Value, p = `p.value`) %>% 
+  mutate(upper95 = estimate + 1.96*se, lower95 = estimate - 1.96*se)
+
+intercept_df %>% ggplot() + 
+  geom_point(aes(x = year, y = estimate), size = 2) +
+  geom_linerange(aes(x = year, ymin = lower95, ymax = upper95), size = 1.5) + 
+  geom_hline(aes(yintercept = 50), alpha = 0.5, size = 1) +
+  labs(x = "Election year", y = "Estimate of intercept") +
+  guides(col = F) +
+  coord_fixed(ratio = 0.2) +
+  lims(y = c(45, 55))
 ```
@@
 
+### Diversity
+In the lead up to the 2001 election, the Liberal government (in power) refused permission for the MV Tampa, a ship carrying 433 rescued refugees and 5 crew to enter Australian waters.
+This, along with the 9/11 attacks, were two key issues that are widely believed to have pushed the Liberal party to victory.
+In this 2001 election, electorates with a higher proportion of people speaking languages other than english at home (`OtherLanguageHome`) were significantly more likely to support Labor.
+This remains negative up until the 2016 election albeit insignificant.
+Higher proportions of people born in Asia (`Born_Asia`) is also significantly linked with Labor support in 2010, which may be due to refugee policy again arising as a key election issue in lead up to the election.
+
+```{r plototherlanguagehome, fig.cap = "Effect of OtherLanguageHome", out.width="60%", eval = F}
+p <- grid_visreg("OtherLanguageHome", xvar_title = "OtherLanguageHome", xlimits = c(-1.1, 3.5), ylimits = c(5, 80))
+```
+
 ### A note on similar variables
@@
 For example, `OtherLanguageHome` is the proportion of households that speak a language other than English at home, which is likely to be related to the proportion of people born overseas (a sum of the variables with a `Born` prefix).
-In 2001, the coefficient estiamte for `OtherLanguageHome` is significant and negative, and all `Born` variables are insignificant.
-If `OtherLanguageHome` is removed from the model, populations born in South Eastern Europe (`Born_SE_Europe`) absorb the negative effect ($p = 0.0011$).
+In 2004, the coefficient estiamte for `OtherLanguageHome` negative but not significant, and all `Born` variables are insignificant.
+If all `Born` variables are removed from the model, `OtherLanguageHome` absorbs the negative effect, becoming significant ($p = 0.01$).
 
 ```{r removeolh, eval = F}
-no_OLH01 <- my_fgls(full_formula, 
-  my_data = model_df %>% filter(year == "2001") %>% dplyr::select(-OtherLanguageHome),
-  sp_weights = sp_weights_01)
+OLH04 <- my_fgls(full_formula, 
+  my_data = model_df %>% filter(year == "2004") %>% dplyr::select(-starts_with("Born")),
+  sp_weights = sp_weights_04)
 
-summary(no_OLH01)$tTable
+summary(OLH04)$tTable
 ```
@@
 
+### Outlier electorates
+
+In
+
+```{r readcooksd, include = F, echo=FALSE}
+read_chunk('cooksd.R')
+```
+
+```{r getcooksd}
+<<GetCookDistance>>
+```
+
+```{r cooksd, include = F}
+cooksd_df <- data.frame(
+  Residuals = c(glsmod16$actual_residuals, glsmod13$actual_residuals, glsmod10$actual_residuals, glsmod07$actual_residuals, glsmod04$actual_residuals, glsmod01$actual_residuals),
+  gls_residuals = c(glsmod16$residuals, glsmod13$residuals, glsmod10$residuals, glsmod07$residuals, glsmod04$residuals, glsmod01$residuals),
+  cooksd = c(glsmod16$cooksd, glsmod13$cooksd, glsmod10$cooksd, glsmod07$cooksd, glsmod04$cooksd, glsmod01$cooksd),
+  bind_rows(
+    glsmod16$my_data %>% left_join(tpp16 %>% dplyr::select(DivisionNm, UniqueID, StateAb), by = "DivisionNm"), 
+    glsmod13$my_data %>% left_join(tpp13 %>% dplyr::select(DivisionNm, UniqueID, StateAb), by = "DivisionNm"), 
+    glsmod10$my_data %>% left_join(tpp10 %>% dplyr::select(DivisionNm, UniqueID, StateAb), by = "DivisionNm"), 
+    glsmod07$my_data %>% left_join(tpp07 %>% dplyr::select(DivisionNm, UniqueID, StateAb), by = "DivisionNm"), 
+    glsmod04$my_data %>% left_join(tpp04 %>% dplyr::select(DivisionNm, UniqueID, StateAb), by = "DivisionNm"), 
+    glsmod01$my_data %>% left_join(tpp01 %>% dplyr::select(DivisionNm, UniqueID, StateAb), by = "DivisionNm"))
+) 
+```
+
+```{r plotcooksd, fig.cap = "Electorates' Cook's Distance values in each election."}
+cooksd_df %>% 
+  ggplot(aes(x = cooksd)) + 
+  geom_dotplot() + 
+  facet_wrap(~year, nrow = 1) + 
+  coord_fixed(ratio = .5) +
+  labs(x = "Cook's Distance", y = "")
+```
+
+```{r}
+cooksd_df %>% 
+  filter(cooksd > 0.13) %>% 
+  select(DivisionNm, StateAb, year, cooksd, Residuals, LNP_Percent, everything()) %>% View
+```
+
+The dotplots in Figure \@ref(fig:plotcooksd) show that there are a few outliers in each election, using cook's distance as a measure of influence.
+The electorate of Sydney (NSW) has the largest cook's distance in 2001, 2004, 2007 and 2010, due to its diverse population (language, birthplace and religion), high number of defacto relationships,  high income, high household mobility and small amount of workers in extractive, distributive and transformative jobs.
+It has remained a strong supporter of the Labor party and is severely underpredicted by the model.
+Nearby in metropolitan NSW, the electorate of Wentworth is found to be an outlier in the 2010, 2013 and 2016 elections - coinciding with its Liberal member, Malcolm Turnbull, being the leader of the Liberal party.
+Wentworth was historically Liberal but increased its two-party vote by over 10 percentage points in 2010 with Turnbull as Liberal leader, whilst experiencing minimal change in its socio-demographics, causing the models to heavily underpredict its Liberal preference.
+
+Melbourne Ports, an upper-middle class electorate just south of Melbourne's the central business district,
+
 ### Electorates with largest residuals
@@
   dplyr::select(DivisionNm, LNP_Percent, year, everything())
+
+all_resids %>% 
+  group_by(year) %>% 
+  mutate(yr_max = max(abs(Residuals))) %>% 
+  filter(abs(Residuals) == yr_max) %>% 
+  select(DivisionNm, StateAb, year, Residuals, LNP_Percent)
+
+cooksd_df %>% 
+  mutate(LNP_win = ifelse(LNP_Percent > 50, 1, 0),
+    LNP_predicted = ifelse(LNP_Percent - Residuals > 50, 1, 0),
+    correct = ifelse(LNP_win == LNP_predicted, 1, 0)) %>% 
+  filter(correct == 0) %>% dim
+  group_by(year) %>% 
+  mutate(yr_max = max(abs(Residuals))) %>% 
+  filter(abs(Residuals) == yr_max) %>% 
+  select(DivisionNm, StateAb, year, Residuals, LNP_Percent)
 ```

### 2019-03-27 01-intro.Rmd
add logs, cooksd

@@
 library(grid)
+library(predictmeans)
 
@@
 }
+
+# Take log of indigneous, judaism, islam, buddhism
+abs2016 <- abs2016 %>% mutate(Indigenous = log(Indigenous), Judaism = log(Judaism), Islam = log(Islam), Buddhism = log(Buddhism))
+abs2013 <- abs2013 %>% mutate(Indigenous = log(Indigenous), Judaism = log(Judaism), Islam = log(Islam), Buddhism = log(Buddhism))
+abs2010 <- abs2010 %>% mutate(Indigenous = log(Indigenous), Judaism = log(Judaism), Islam = log(Islam), Buddhism = log(Buddhism))
+abs2007 <- abs2007 %>% mutate(Indigenous = log(Indigenous), Judaism = log(Judaism), Islam = log(Islam), Buddhism = log(Buddhism))
+abs2004 <- abs2004 %>% mutate(Indigenous = log(Indigenous), Judaism = log(Judaism), Islam = log(Islam), Buddhism = log(Buddhism))
+abs2001 <- abs2001 %>% mutate(Indigenous = log(Indigenous), Judaism = log(Judaism), Islam = log(Islam), Buddhism = log(Buddhism))
 
@@
 
+  # Cooks distance
+  #gls_model$cooksd <- unname(predictmeans::CookD(gls_model, plot = FALSE))
+
   # Call to function with stargazer

### 2019-03-27 cooksd.R
script to get cooksd via predictmeans::CookD

@@ -1,0 +1,306 @@
+## @knitr GetCookDistance
+
+# 2016
+
+my_formula = "LNP_Percent ~ ."
+my_data = model_df %>% filter(year == "2016")
+sp_weights = sp_weights_16
+
+# DivisionNm
+model_data <- my_data %>% dplyr::select(-c(DivisionNm, year))
+
+# Spatial weights matrix
+w_mat <- listw2mat(sp_weights)
+
+# Get OLS residuals
+ols_model <- lm(my_formula, model_data)
+my_res <- ols_model$residuals
+
+# Solve for rho
+res_model <- lm(my_res ~ w_mat%*%my_res)
+rho <- res_model$coefficients[2]
+rho_df <- data.frame(estimate = rho, se = summary(res_model)$coefficients[2,2], p = summary(res_model)$coefficients[2,4])
+
+# Transform data for GLS
+trans_mat <- diag(nrow(model_data)) - rho*w_mat
+gls_data <- data.frame(LNP_Percent = trans_mat %*% model_data$LNP_Percent,
+  Intercept = trans_mat %*% rep(1,nrow(model_data))) %>% 
+  bind_cols(as.data.frame(trans_mat %*% as.matrix(model_data %>% dplyr::select(-LNP_Percent))))
+
+# GLS model
+my_formula <- formula(paste0(my_formula,  " - 1"))
+gls_model <- gls(my_formula, gls_data)
+
+# Cooks distance
+glsmod16$cooksd <- unname(predictmeans::CookD(gls_model, plot = FALSE))
+
+# REMOVE
+
+remove(gls_model)
+remove(my_formula)
+remove(gls_data)
+remove(trans_mat)
+remove(rho)
+remove(rho_df)
+remove(res_model)
+remove(my_res)
+remove(ols_model)
+remove(w_mat)
+remove(model_data)
+remove(sp_weights)
+remove(my_data)
+
+
+# 2013
+
+my_formula = "LNP_Percent ~ ."
+my_data = model_df %>% filter(year == "2013")
+sp_weights = sp_weights_13
+
+# DivisionNm
+model_data <- my_data %>% dplyr::select(-c(DivisionNm, year))
+
+# Spatial weights matrix
+w_mat <- listw2mat(sp_weights)
+
+# Get OLS residuals
+ols_model <- lm(my_formula, model_data)
+my_res <- ols_model$residuals
+
+# Solve for rho
[…]

### 2019-03-27 04-results.Rmd
update visreg functions to return data (plot = F), and begin new (updated) results sections. Incomes and unemployment complete.

@@
 The spatial autoregressive coefficient $\rho$ is positive and significant in only the 2001 and 2016 elections (Figure \@ref(fig:rhovis)), meaning that in these elections, an electorate's political climate was affected by the attitudes of it's neighbours.
-Conversely, in the other four elections, the spatial effect weakens to become marginally insignficant.
+Conversely, in the other four elections, the spatial effect weakens to become insignficant.
 In these years, it appears that the spatial component does not explain anything over and above the electoral socio-demographics, meaning electorates effectively acted more independently.
@@
 
-### Income and unemployment
-Typically the Labor party campaigns on more progressive policies, which often include tax reform that adversely affects higher income earners, and more generous social assistance programs.
-Despite this, incomes have only significantly affected the electoral two-party preference in 2016 (see Figure \@ref(fig:plotincomes)), with higher income areas associated with support for the Liberal party.
-In the prior elections, the estimated effect is still positive, albeit insignificant.
-Unemployment on the other hand, was influential in 2001, 2004 and 2007 elections (Figure \@ref(fig:plotunemploy)), with electorates with higher unemployment favouring Labor.
-This estimated effect has become weaker over time, with unemployment having virtually no effect in 2016.
-
 ```{r myvisreg, include = F}
@@
 my_visreg <- function(my_model, sp_weights, varname, 
-  nolabs = FALSE, xlimits = NULL, ylimits = NULL, year = "") {
+  plot = FALSE, nolabs = FALSE, xlimits = NULL, ylimits = NULL, year = "") {
 
@@
   # Plot
+  if (plot == TRUE) {
       myplot <- ggplot(data = plot_df) + 
@@
   }
-
       return(myplot)
@@
 
+  # Points
+  if (plot == FALSE) {
+    return_ls <- list(bands = plot_df, points = points_df)
+
+    return(return_ls)
+  }
+
+}
+
 # Grid visreg
-grid_visreg <- function(varname, xvar_title, xlimits, ylimits) {
-  p16 <- my_visreg(glsmod16, sp_weights_16, varname = varname, nolabs = T, year = "2016", xlimits, ylimits)
-  p13 <- my_visreg(glsmod13, sp_weights_13, varname = varname, nolabs = T, year = "2013", xlimits, ylimits)
-  p10 <- my_visreg(glsmod10, sp_weights_10, varname = varname, nolabs = T, year = "2010", xlimits, ylimits)
-  p07 <- my_visreg(glsmod07, sp_weights_07, varname = varname, nolabs = T, year = "2007", xlimits, ylimits)
-  p04 <- my_visreg(glsmod04, sp_weights_04, varname = varname, nolabs = T, year = "2004", xlimits, ylimits)
-  p01 <- my_visreg(glsmod01, sp_weights_01, varname = varname, nolabs = T, year = "2001", xlimits, ylimits)
+grid_visreg <- function(varname, xvar_title = "", xlimits = NULL, ylimits = NULL, plot = FALSE) {
+  if (plot == TRUE) {
+    p16 <- my_visreg(glsmod16, sp_weights_16, varname = varname, plot = T, nolabs = T, 
+      year = "2016", xlimits, ylimits)
+    p13 <- my_visreg(glsmod13, sp_weights_13, varname = varname, plot = T, nolabs = T, 
+      year = "2013", xlimits, ylimits)
+    p10 <- my_visreg(glsmod10, sp_weights_10, varname = varname, plot = T, nolabs = T, 
+      year = "2010", xlimits, ylimits)
+    p07 <- my_visreg(glsmod07, sp_weights_07, varname = varname, plot = T, nolabs = T, 
+      year = "2007", xlimits, ylimits)
+    p04 <- my_visreg(glsmod04, sp_weights_04, varname = varname, plot = T, nolabs = T, 
+      year = "2004", xlimits, ylimits)
+    p01 <- my_visreg(glsmod01, sp_weights_01, varname = varname, plot = T, nolabs = T, 
+      year = "2001", xlimits, ylimits)
 
@@
 
+  if (plot == FALSE) {
+    p16 <- my_visreg(glsmod16, sp_weights_16, varname = varname, plot = F, year = "2016")
+    p13 <- my_visreg(glsmod13, sp_weights_13, varname = varname, plot = F, year = "2013")
+    p10 <- my_visreg(glsmod10, sp_weights_10, varname = varname, plot = F, year = "2010")
+    p07 <- my_visreg(glsmod07, sp_weights_07, varname = varname, plot = F, year = "2007")
+    p04 <- my_visreg(glsmod04, sp_weights_04, varname = varname, plot = F, year = "2004")
+    p01 <- my_visreg(glsmod01, sp_weights_01, varname = varname, plot = F, year = "2001")
+
+    bands_df <- bind_rows(
+      p16$bands %>% mutate(year = "2016"),
+      p13$bands %>% mutate(year = "2013"),
+      p10$bands %>% mutate(year = "2010"),
+      p07$bands %>% mutate(year = "2007"),
+      p04$bands %>% mutate(year = "2004"),
+      p01$bands %>% mutate(year = "2001")
+    )
+
+    points_df <- bind_rows(
+      p16$points %>% mutate(year = "2016"),
+      p13$points %>% mutate(year = "2013"),
+      p10$points %>% mutate(year = "2010"),
+      p07$points %>% mutate(year = "2007"),
+      p04$points %>% mutate(year = "2004"),
+      p01$points %>% mutate(year = "2001")
+    )
+
+    return_df <- list(bands = bands_df, points = points_df)
+
+    return(return_df)
+  }
+
+}
+
 ```
+
+### Income and unemployment
+
+Typically the Labor party campaigns on more progressive policies, which often include tax reform that adversely affects higher income earners, and more generous social assistance programs.
+Perhaps due to these policies, higher income electorates have been more likely to support the Liberal party, as the `Incomes` factor has a positive effect on Liberal preference (see Figure \@ref(fig:plotincomes)).
+This effect is signficant in every election aside from 2004, in which it is only marginally insignificant ($p = 0.0613$).
+Unemployment however, has not been as influential.
+In 2001 and 2004, electorates with higher unemployment were more likely to support Labor, but over time this has shifted towards support for the Liberal party, culminating in a significantly positive effect in 2016.
+
+```{r plotincomes, fig.cap = "Effect of Incomes", out.width="100%"}
+p <- grid_visreg("Incomes", xvar_title = "Incomes", plot = T, xlimits = c(-1.5, 3.6), ylimits = c(30, 95))
+```
+
+```{r plotunemploy, fig.cap = "Effect of Unemployment", out.width="60%"}
+p <- grid_visreg("Unemployment", xvar_title = "Unemployment", plot = T, xlimits = c(-2.1, 3.4), ylimits = c(28, 68))
+```
+`
+
+### Income and unemployment
+Typically the Labor party campaigns on more progressive policies, which often include tax reform that adversely affects higher income earners, and more generous social assistance programs.
+Despite this, incomes have only significantly affected the electoral two-party preference in 2016 (see Figure \@ref(fig:plotincomes)), with higher income areas associated with support for the Liberal party.
+In the prior elections, the estimated effect is still positive, albeit insignificant.
+Unemployment on the other hand, was influential in 2001, 2004 and 2007 elections (Figure \@ref(fig:plotunemploy)), with electorates with higher unemployment favouring Labor.
+This estimated effect has become weaker over time, with unemployment having virtually no effect in 2016.
 

### 2019-03-27 04-results.Rmd
first attempt at all but diversity part of results

@@
 ```
-`
-
-### Income and unemployment
-Typically the Labor party campaigns on more progressive policies, which often include tax reform that adversely affects higher income earners, and more generous social assistance programs.
-Despite this, incomes have only significantly affected the electoral two-party preference in 2016 (see Figure \@ref(fig:plotincomes)), with higher income areas associated with support for the Liberal party.
-In the prior elections, the estimated effect is still positive, albeit insignificant.
-Unemployment on the other hand, was influential in 2001, 2004 and 2007 elections (Figure \@ref(fig:plotunemploy)), with electorates with higher unemployment favouring Labor.
-This estimated effect has become weaker over time, with unemployment having virtually no effect in 2016.
-
-```{r plotincomes, fig.cap = "Effect of Incomes", out.width="100%"}
-p <- grid_visreg("Incomes", xvar_title = "Incomes", xlimits = c(-1.5, 3.6), ylimits = c(30, 95))
-```
-
-```{r plotunemploy, fig.cap = "Effect of Unemployment", out.width="60%"}
-p <- grid_visreg("Unemployment", xvar_title = "Unemployment", xlimits = c(-2.1, 3.4), ylimits = c(28, 68))
-```
 
@@
 This is unsurprising, as the Liberal party has close ties with these traditional energy industries, and typically present policies to reduce taxation on energy production.
-Furthermore, electorates with more workers in construction or manufacturing industries (`Transformative`), as well as wholesale trade, retail trade, transport, post or warehousing (`Distributive`), are also linked with the Liberal party.
-However, these effects are not as strong, and `Transformative` was marginally insignificant in the 2010 election, whereas `Distributive` was not a key socio-demographic in 2001 or 2016.
-These effects are shown in Figures \@ref(fig:plottransformative) and \@ref(fig:plotdistributive).
+Furthermore, electorates with more workers in construction or manufacturing industries (`Transformative`) are also more likely to support the Liberal party (see Figure \@ref(fig:plottransformative)).
 
 ```{r plotextractive, fig.cap = "Effect of Extractive", out.width="60%"}
-p <- grid_visreg("Extractive", xvar_title = "Extractive", xlimits = c(-0.7, 3.4), ylimits = c(35, 81))
+p <- grid_visreg("Extractive", xvar_title = "Extractive", xlimits = c(-0.7, 3.4), ylimits = c(35, 81), plot = T)
 ```
@@
 ```{r plottransformative, fig.cap = "Effect of Transformative", out.width="60%"}
-p <- grid_visreg("Transformative", xvar_title = "Transformative", xlimits = c(-2.25, 2.9), ylimits = c(34, 75))
-```
-```{r plotdistributive, fig.cap = "Effect of Distributive", out.width="60%"}
-p <- grid_visreg("Distributive", xvar_title = "Distributive", xlimits = c(-3.1, 2.5), ylimits = c(34, 65))
+p <- grid_visreg("Transformative", xvar_title = "Transformative", xlimits = c(-2.25, 2.9), ylimits = c(34, 75), plot = T)
 ```
 
-Similarly, the proportions of workers in managerial, administrative, clerical and sales roles (`ManagerAdminClericalSales`) is also a significant predictor of two-party preference across all six elections, with higher proportions supporting the Liberal party.
+Similarly, workers in managerial, administrative, clerical and sales roles (`ManagerAdminClericalSales`) is also a significant predictor of two-party preference vote across all six elections, with a higher proportion of people working these jobs increasing Liberal support.
+The magnitude of this effect has also increased over the years.
 
 ```{r plotmanager, fig.cap = "Effect of ManagerAdminClericalSales", out.width="60%"}
-p <- grid_visreg("ManagerAdminClericalSales", xvar_title = "ManagerAdminClericalSales", xlimits = c(-3.75, 2.8), ylimits = c(32, 70))
+p <- grid_visreg("ManagerAdminClericalSales", xvar_title = "ManagerAdminClericalSales", xlimits = c(-3.75, 2.8), ylimits = c(32, 70), plot = T)
 ```
@@
 ```{r plotdiffaddress, fig.cap = "Effect of DiffAddress", out.width="60%"}
-p <- grid_visreg("DiffAddress", xvar_title = "DiffAddress", xlimits = c(-2.1, 3.65), ylimits = c(34, 81))
+p <- grid_visreg("DiffAddress", xvar_title = "DiffAddress", xlimits = c(-2.1, 3.65), ylimits = c(34, 81), plot = T)
 ```
 
-In each of the six elections, electorates with a higher proportion of people that have recently moved house (meaning in last five years) were more likely to support the Liberal party, with this effect becoming stronger since 2010.
+In each of the six elections, electorates with a higher proportion of people that have recently moved house (meaning in last five years) were more likely to support the Liberal party, although this effect was marginally insignificant in 2013.
 Having controlled for characteristics of house ownership and rental prices (via the factors `PropertyOwned` and `RentLoan` respectively), this is effect is somewhat surprising.
-*Add more here?*
 
-### Indigenous population
-```{r plotindigenous, fig.cap = "Effect of Indigenous population", out.width="60%"}
-p <- grid_visreg("Indigenous", xvar_title = "Indigenous", xlimits = c(-0.6, 9.5), ylimits = c(35, 120))
+### Relationships and age
+De facto relationships, but not marriages, is found to be an important (and significant) predictor of the two-party preferred vote in all six elections, with more de facto relationships associated with higher support for the Labor party.
+
+```{r plotdefacto, fig.cap = "Effect of DeFacto", out.width="60%", eval = F}
+p <- grid_visreg("DeFacto", xvar_title = "DeFacto", xlimits = c(-2.2, 5.1), ylimits = c(12, 83), plot = T)
 ```
 
-The indigenous population of an electorate is found to be a significant predictor in every election, with higher indigenous populations supporting the Liberal party.
-There are a few outliers with Lingiari (NT) being the most extreme.
-However, Lingiari itself has been a strong Labor supporter, and this effect appears to be driven by the other electorates with high Indigenous populations (Kalgoorlie (WA), Leichhardt (NSW), Durack (WA), Parkes(NSW) and Kennedy (QLD)), which have almost always elected a Liberal candidate.
+Although only significant in 2007 and 2010, older electorates (higher median age) are consistently more likely to support the Liberal party.
 
-### Relationships and families
-In each of the six elections, higher proportions people in de facto marriages are tied to support for the Labor party, whereas the effect of marriages themselves are insignificant.
-Larger family and household size (via the `FamHouseSize` factor) is also linked with Labor support, although only being significant in the 2007 and 2010 elections (see Figure \@ref(fig:plotfamhousesize)).
-
-```{r plotdefacto, fig.cap = "Effect of DeFacto", out.width="60%", eval = F}
-p <- grid_visreg("DeFacto", xvar_title = "DeFacto", xlimits = c(-2.2, 5.1), ylimits = c(12, 83))
+```{r plotage, fig.cap = "Effect of MedianAge", out.width="60%", eval = F}
+p <- grid_visreg("MedianAge", xvar_title = "MedianAge", xlimits = c(-2.5, 2.7), ylimits = c(31, 70), plot = T)
 ```
 
-```{r plotfamhousesize, fig.cap = "Effect of FamHouseSize", out.width="60%", eval = F}
-p <- grid_visreg("FamHouseSize", xvar_title = "FamHouseSize", xlimits = c(-3.4, 2.7), ylimits = c(12, 83))
+### Education
+Since 2007, electorates with higher education levels have been associated with supporting the Labor party, although this effect is only significant in 2016.
+Prior to 2007, education had no material effect.
+
+```{r ploteducation, fig.cap = "Effect of Education", out.width="60%", eval = F}
+p <- grid_visreg("Education", xvar_title = "Education", xlimits = c(-2.5, 2.7), ylimits = c(31, 70), plot = T)
 ```
@@
 ```{r plototherlanguagehome, fig.cap = "Effect of OtherLanguageHome", out.width="60%", eval = F}
-p <- grid_visreg("OtherLanguageHome", xvar_title = "OtherLanguageHome", xlimits = c(-1.1, 3.5), ylimits = c(5, 80))
+p <- grid_visreg("OtherLanguageHome", xvar_title = "OtherLanguageHome", xlimits = c(-1.1, 3.5), ylimits = c(5, 80), plot = T)
+```
+
+```{r plotbornse, fig.cap = "Effect of Born_SE_Europe", out.width="60%", eval = F}
+p <- grid_visreg("Born_SE_Europe", xvar_title = "Born_SE_Europe", xlimits = c(-0.9, 5), ylimits = c(20, 70), plot = T)
 ```

### 2019-03-27 03-modelling.Rmd
fix gitignore and % sign in fig cap
updated model table

@@
 Independent of principal components, we may suspect that median personal income, median household income and median family income are providing similar information about the financial wellbeing of an electorate.
-Their loadings in the first principal component are large (0.20, 0.21 and 0.22 respectively), which provides the evidence needed to combine these variables into a single factor, which is called `Incomes`.
+Their loadings in the first principal component are large (0.19, 0.21 and 0.22 respectively), which provides the evidence needed to combine these variables into a single factor, which is called `Incomes`.
 
@@
 ```
-
 \begin{table}[!htbp] \centering 
@@
 \hline \\[-1.8ex] 
- $\rho$ & 0.44$^{***}$ & 0.15$^{*}$ & $-$0.04 & 0.15 & 0.29 & 0.48$^{***}$ \\ 
-  & (0.15) & (0.17) & (0.19) & (0.17) & (0.16) & (0.16) \\ 
+ $\rho$ & 0.46$^{***}$ & 0.29$^{*}$ & 0.24 & 0.19 & 0.27$^{*}$ & 0.50$^{***}$ \\ 
+  & (0.15) & (0.17) & (0.17) & (0.16) & (0.16) & (0.17) \\ 
   & & & & & & \\ 
- AusCitizen & $-$2.77 & $-$2.83 & 0.67 & 1.00 & $-$0.87 & $-$1.53 \\ 
-  & (2.51) & (2.63) & (2.57) & (2.93) & (2.90) & (2.95) \\ 
+ AusCitizen & $-$3.13 & $-$2.64 & $-$2.53 & $-$0.08 & $-$3.40 & $-$1.80 \\ 
+  & (2.26) & (2.43) & (2.34) & (2.79) & (2.76) & (2.71) \\ 
   & & & & & & \\ 
- Born\_Asia & 2.63 & 0.74 & $-$1.37 & $-$7.58$^{***}$ & $-$2.56 & $-$2.13 \\ 
-  & (3.03) & (2.55) & (1.96) & (2.68) & (2.58) & (2.20) \\ 
+ Born\_Asia & 2.22 & $-$0.95 & $-$1.60 & $-$6.83$^{**}$ & $-$3.03 & $-$0.55 \\ 
+  & (2.18) & (2.44) & (2.19) & (2.73) & (2.71) & (2.17) \\ 
   & & & & & & \\ 
- Born\_MidEast & 1.92 & 3.22$^{*}$ & 2.92$^{*}$ & 1.79 & 1.70 & $-$0.82 \\ 
-  & (1.53) & (1.78) & (1.51) & (1.54) & (1.54) & (1.56) \\ 
+ Born\_MidEast & $-$1.15 & $-$1.59 & $-$2.01$^{*}$ & $-$2.03 & $-$0.92 & $-$1.44 \\ 
+  & (1.07) & (1.20) & (1.11) & (1.27) & (1.24) & (1.13) \\ 
   & & & & & & \\ 
- Born\_SE\_Europe & 1.02 & 0.04 & $-$0.20 & $-$1.72 & $-$1.25 & $-$1.63 \\ 
-  & (2.34) & (2.05) & (1.17) & (1.34) & (1.15) & (0.99) \\ 
+ Born\_SE\_Europe & $-$3.21$^{**}$ & $-$4.24$^{***}$ & $-$3.61$^{***}$ & $-$4.14$^{***}$ & $-$3.69$^{***}$ & $-$2.72$^{***}$ \\ 
+  & (1.42) & (1.46) & (1.02) & (1.19) & (1.07) & (0.97) \\ 
   & & & & & & \\ 
- Born\_UK & $-$0.49 & $-$0.87 & 0.29 & $-$0.29 & $-$0.35 & $-$1.37 \\ 
-  & (1.04) & (0.90) & (0.80) & (1.05) & (1.00) & (0.98) \\ 
+ Born\_UK & 0.25 & $-$0.07 & 0.34 & 0.56 & $-$0.09 & $-$1.32 \\ 
+  & (1.00) & (0.98) & (0.90) & (1.07) & (1.04) & (1.04) \\ 
   & & & & & & \\ 
- BornElsewhere & $-$1.08 & $-$0.51 & 5.01 & 9.78$^{**}$ & 4.04 & 1.74 \\ 
-  & (3.40) & (3.67) & (3.59) & (4.47) & (4.46) & (4.52) \\ 
+ BornElsewhere & $-$5.04 & $-$4.91 & $-$4.13 & 2.35 & $-$5.23 & $-$4.14 \\ 
+  & (3.30) & (3.68) & (3.38) & (4.23) & (4.15) & (3.97) \\ 
   & & & & & & \\ 
- Buddhism & 1.50 & 2.84$^{*}$ & 0.90 & 2.41$^{*}$ & 1.09 & 1.51 \\ 
-  & (1.94) & (1.61) & (1.25) & (1.40) & (1.25) & (1.24) \\ 
+ Buddhism & $-$0.49 & $-$0.17 & $-$1.37 & $-$0.83 & $-$0.12 & $-$1.60 \\ 
+  & (1.39) & (1.61) & (1.61) & (1.80) & (1.68) & (1.56) \\ 
   & & & & & & \\ 
- Christianity & 0.53 & 3.44 & $-$1.93 & $-$0.60 & 3.08 & 5.17$^{**}$ \\ 
-  & (4.58) & (4.19) & (3.45) & (3.44) & (2.79) & (2.52) \\ 
+ Christianity & $-$2.48 & $-$1.23 & 0.38 & 0.50 & 2.41 & 1.68 \\ 
+  & (1.73) & (1.85) & (1.83) & (1.99) & (1.85) & (1.78) \\ 
   & & & & & & \\ 
- CurrentlyStudying & $-$1.14 & 1.69 & 4.03$^{***}$ & 3.28$^{***}$ & 1.73 & $-$0.09 \\ 
-  & (1.01) & (1.09) & (1.12) & (1.21) & (1.17) & (1.13) \\ 
+ CurrentlyStudying & $-$2.19$^{**}$ & $-$0.13 & 2.06$^{*}$ & 2.12$^{*}$ & 1.15 & $-$0.16 \\ 
+  & (0.99) & (1.13) & (1.17) & (1.25) & (1.26) & (1.18) \\ 
   & & & & & & \\ 
- DeFacto & $-$6.42$^{***}$ & $-$5.04$^{**}$ & $-$6.55$^{***}$ & $-$8.60$^{***}$ & $-$9.73$^{***}$ & $-$11.02$^{***}$ \\ 
-  & (1.69) & (2.00) & (1.90) & (2.57) & (2.62) & (2.55) \\ 
+ DeFacto & $-$6.44$^{***}$ & $-$5.37$^{**}$ & $-$6.43$^{***}$ & $-$8.07$^{***}$ & $-$6.56$^{**}$ & $-$8.53$^{***}$ \\ 
+  & (1.87) & (2.48) & (2.31) & (3.06) & (3.11) & (2.83) \\ 
   & & & & & & \\ 
- DiffAddress & 3.86$^{***}$ & 4.63$^{***}$ & 3.40$^{***}$ & 4.72$^{***}$ & 5.02$^{***}$ & 6.93$^{***}$ \\ 
-  & (0.95) & (1.03) & (0.88) & (1.60) & (1.71) & (1.50) \\ 
+ DiffAddress & 3.88$^{***}$ & 5.06$^{***}$ & 4.22$^{***}$ & 5.57$^{***}$ & 3.53$^{*}$ & 5.67$^{***}$ \\ 
+  & (0.94) & (1.12) & (0.99) & (1.76) & (1.91) & (1.60) \\ 
   & & & & & & \\ 
- Distributive & 1.67 & 2.75$^{**}$ & 2.25$^{**}$ & 2.88$^{**}$ & 2.76$^{**}$ & 1.65 \\ 
-  & (1.08) & (1.13) & (1.02) & (1.30) & (1.25) & (1.17) \\ 
+ Distributive & 1.27 & 2.01$^{*}$ & 1.36 & 1.57 & 2.10$^{*}$ & 1.20 \\ 
+  & (1.12) & (1.21) & (1.13) & (1.34) & (1.27) & (1.21) \\ 
   & & & & & & \\ 
- Education & 0.62 & 1.94 & $-$0.77 & $-$1.82 & $-$0.90 & $-$3.04 \\ 
-  & (2.23) & (2.86) & (2.89) & (3.35) & (3.04) & (2.99) \\ 
+ Education & 1.08 & 0.52 & $-$5.52$^{*}$ & $-$4.08 & $-$4.44 & $-$7.07$^{**}$ \\ 
+  & (2.38) & (3.12) & (3.27) & (3.95) & (3.78) & (3.55) \\ 
   & & & & & & \\ 
- Extractive & 4.69$^{***}$ & 5.31$^{***}$ & 5.81$^{***}$ & 6.97$^{***}$ & 6.09$^{***}$ & 6.81$^{***}$ \\ 
-  & (1.36) & (1.23) & (1.16) & (1.48) & (1.42) & (1.31) \\ 
+ Extractive & 4.83$^{***}$ & 5.45$^{***}$ & 5.37$^{***}$ & 7.31$^{***}$ & 6.71$^{***}$ & 7.43$^{***}$ \\ 
+  & (1.48) & (1.42) & (1.36) & (1.56) & (1.47) & (1.39) \\ 
   & & & & & & \\ 
- FamHouseSize & $-$2.94 & $-$5.11$^{*}$ & $-$9.67$^{***}$ & $-$8.22$^{**}$ & $-$6.16$^{*}$ & $-$4.35 \\ 
-  & (2.27) & (2.78) & (2.66) & (3.28) & (3.15) & (2.86) \\ 
+ FamHouseSize & $-$0.16 & 0.87 & $-$2.40 & $-$2.53 & $-$3.26 & $-$2.91 \\ 
+  & (2.19) & (2.72) & (2.69) & (3.25) & (3.28) & (2.90) \\ 
   & & & & & & \\ 
- Incomes & 2.99$^{*}$ & 2.95 & 4.36$^{*}$ & 2.73 & 4.19 & 8.46$^{***}$ \\ 
-  & (1.76) & (2.60) & (2.56) & (3.06) & (2.68) & (2.57) \\ 
+ Incomes & 4.36$^{**}$ & 5.03$^{*}$ & 9.45$^{***}$ & 7.09$^{**}$ & 7.97$^{***}$ & 12.20$^{***}$ \\ 
+  & (1.77) & (2.66) & (2.75) & (3.25) & (2.92) & (2.75) \\ 
   & & & & & & \\ 
- Indigenous & 3.97$^{***}$ & 5.62$^{***}$ & 6.81$^{***}$ & 5.90$^{***}$ & 6.12$^{***}$ & 4.47$^{***}$ \\ 
-  & (1.26) & (1.35) & (1.17) & (1.49) & (1.49) & (1.38) \\ 
+ Indigenous & 2.91$^{*}$ & 1.97 & 2.48 & 2.84 & 0.67 & $-$0.05 \\ 
+  & (1.68) & (1.95) & (1.75) & (2.16) & (2.14) & (2.00) \\ 
   & & & & & & \\ 
- Islam & 0.15 & 0.61 & $-$1.21 & $-$1.04 & 1.20 & 2.64 \\ 
-  & (2.03) & (2.00) & (1.80) & (1.94) & (1.71) & (1.72) \\ 
+ Islam & $-$0.92 & $-$0.97 & $-$0.54 & $-$2.50 & $-$0.82 & $-$0.95 \\ 
+  & (1.22) & (1.36) & (1.27) & (1.52) & (1.42) & (1.34) \\ 
   & & & & & & \\ 
- Judaism & 1.02 & 1.74 & 0.01 & 0.46 & 1.60$^{*}$ & 1.96$^{***}$ \\ 
-  & (1.38) & (1.26) & (1.00) & (1.05) & (0.85) & (0.71) \\ 
+ Judaism & 1.88$^{*}$ & 1.78 & 2.66$^{***}$ & 1.97$^{*}$ & 2.74$^{**}$ & 1.65$^{*}$ \\ 
+  & (1.05) & (1.13) & (1.01) & (1.15) & (1.10) & (1.00) \\ 
   & & & & & & \\ 
- ManagerAdminClericalSales & 2.10$^{***}$ & 3.28$^{***}$ & 5.21$^{***}$ & 4.32$^{***}$ & 3.99$^{***}$ & 4.62$^{***}$ \\ 
-  & (0.70) & (0.85) & (0.78) & (1.00) & (0.99) & (1.03) \\ 
+ ManagerAdminClericalSales & 2.06$^{***}$ & 3.32$^{***}$ & 6.00$^{***}$ & 5.47$^{***}$ & 5.04$^{***}$ & 5.78$^{***}$ \\ 
+  & (0.71) & (0.93) & (0.90) & (1.08) & (1.03) & (1.06) \\ 
   & & & & & & \\ 
- Married & 2.27 & 5.50$^{**}$ & 5.06$^{*}$ & 5.07$^{*}$ & 1.46 & $-$1.39 \\ 
-  & (2.26) & (2.79) & (2.59) & (2.91) & (2.71) & (2.64) \\ 
+ Married & 0.44 & 0.11 & $-$1.22 & $-$0.22 & 0.91 & $-$2.34 \\ 
+  & (2.31) & (2.96) & (2.83) & (3.15) & (3.03) & (2.81) \\ 
   & & & & & & \\ 
- MedianAge & 1.54 & 3.73$^{**}$ & 1.86 & 2.08 & 2.28 & 2.94$^{*}$ \\ 
-  & (1.30) & (1.55) & (1.65) & (2.08) & (1.88) & (1.63) \\ 
+ MedianAge & 2.32$^{*}$ & 4.96$^{***}$ & 3.66$^{**}$ & 4.00$^{*}$ & 2.30 & 2.87 \\ 
+  & (1.32) & (1.65) & (1.81) & (2.26) & (2.08) & (1.79) \\ 
   & & & & & & \\ 
- NoReligion & 1.06 & 3.43 & $-$0.31 & 0.25 & 3.53 & 5.76$^{**}$ \\ 
-  & (3.85) & (3.45) & (2.90) & (2.98) & (2.58) & (2.59) \\ 
+ NoReligion & $-$1.57 & $-$0.92 & 0.56 & $-$0.30 & 1.02 & 1.31 \\ 
+  & (1.59) & (1.71) & (1.73) & (1.92) & (1.94) & (2.04) \\ 
   & & & & & & \\ 
- OneParent\_House & $-$1.00 & 1.37 & 1.66 & 0.22 & 0.11 & $-$0.85 \\ 
-  & (1.34) & (1.49) & (1.38) & (1.61) & (1.46) & (1.39) \\ 
+ OneParent\_House & $-$1.73 & $-$0.45 & $-$0.75 & $-$1.46 & $-$0.77 & $-$0.74 \\ 
+  & (1.36) & (1.59) & (1.49) & (1.69) & (1.57) & (1.47) \\ 
   & & & & & & \\ 
- OtherLanguageHome & $-$10.40$^{**}$ & $-$6.76 & $-$6.02 & $-$4.37 & $-$4.69 & 1.27 \\ 
-  & (4.97) & (5.61) & (4.44) & (5.70) & (5.50) & (4.95) \\ 
+ OtherLanguageHome & $-$0.44 & 5.92 & 9.98$^{**}$ & 11.24$^{**}$ & 9.00$^{*}$ & 9.84$^{**}$ \\ 
+  & (3.22) & (4.16) & (3.91) & (4.76) & (4.66) & (4.44) \\ 
   & & & & & & \\ 
- PropertyOwned & $-$0.76 & $-$0.92 & 0.78 & $-$1.17 & 0.68 & 1.76 \\ 
-  & (1.19) & (1.21) & (1.09) & (1.49) & (1.43) & (1.38) \\ 
+ PropertyOwned & $-$0.46 & $-$0.53 & 0.67 & $-$0.94 & $-$0.48 & 1.41 \\ 
+  & (1.37) & (1.50) & (1.43) & (1.76) & (1.67) & (1.50) \\ 
   & & & & & & \\ 
- RentLoanPrice & $-$0.31 & $-$1.54 & $-$0.54 & 2.12 & 1.01 & $-$0.70 \\ 
-  & (1.50) & (1.68) & (1.48) & (1.96) & (1.96) & (2.02) \\ 
+ RentLoanPrice & $-$1.57 & $-$3.32$^{*}$ & $-$4.01$^{**}$ & $-$0.97 & $-$0.70 & $-$0.89 \\ 
+  & (1.49) & (1.76) & (1.67) & (2.07) & (2.07) & (2.16) \\ 
   & & & & & & \\ 
- SocialServ & 2.42$^{*}$ & 1.89 & 1.77 & 1.91 & 0.94 & 2.90$^{***}$ \\ 
-  & (1.27) & (1.27) & (1.14) & (1.35) & (1.22) & (1.10) \\ 
+ SocialServ & 2.51$^{*}$ & 1.65 & 2.47$^{*}$ & 2.53$^{*}$ & 2.35$^{*}$ & 4.45$^{***}$ \\ 
+  & (1.33) & (1.41) & (1.29) & (1.47) & (1.32) & (1.19) \\ 
   & & & & & & \\ 
- Transformative & 2.99$^{**}$ & 5.74$^{***}$ & 5.20$^{***}$ & 3.28$^{*}$ & 3.74$^{**}$ & 4.75$^{***}$ \\ 
-  & (1.41) & (1.59) & (1.47) & (1.74) & (1.57) & (1.41) \\ 
+ Transformative & 3.24$^{**}$ & 4.73$^{***}$ & 4.84$^{***}$ & 4.46$^{**}$ & 3.56$^{**}$ & 4.58$^{***}$ \\ 
+  & (1.55) & (1.78) & (1.74) & (1.98) & (1.78) & (1.53) \\ 
   & & & & & & \\ 
- Unemployment & $-$3.00$^{**}$ & $-$5.09$^{***}$ & $-$2.88$^{**}$ & $-$2.42 & $-$1.22 & 0.36 \\ 
-  & (1.28) & (1.51) & (1.41) & (1.73) & (1.54) & (1.42) \\ 
+ Unemployment & $-$2.45$^{*}$ & $-$3.07$^{*}$ & 0.29 & 0.08 & 1.67 & 2.79$^{**}$ \\ 
+  & (1.40) & (1.63) & (1.51) & (1.76) & (1.51) & (1.37) \\ 
   & & & & & & \\ 
- Constant & 50.78$^{***}$ & 52.63$^{***}$ & 47.32$^{***}$ & 49.92$^{***}$ & 53.48$^{***}$ & 50.45$^{***}$ \\ 
-  & (0.67) & (0.45) & (0.35) & (0.51) & (0.56) & (0.74) \\ 
+ Constant & 50.81$^{***}$ & 52.60$^{***}$ & 47.31$^{***}$ & 49.93$^{***}$ & 53.51$^{***}$ & 50.49$^{***}$ \\ 
+  & (0.71) & (0.58) & (0.52) & (0.57) & (0.58) & (0.80) \\ 
   & & & & & & \\ 
@@
 Observations & 150 & 150 & 150 & 150 & 150 & 150 \\ 
-Residual Standard Error (GLS) & 4.55 & 4.69 & 4.44 & 5.31 & 4.85 & 4.65 \\ 
-Log Likelihood & $-$398.33 & $-$403.34 & $-$398.73 & $-$418.08 & $-$405.41 & $-$399.52 \\ 
-Akaike Inf. Crit. & 860.67 & 870.68 & 861.46 & 900.16 & 874.83 & 863.03 \\ 
-Bayesian Inf. Crit. & 949.60 & 959.61 & 950.40 & 989.10 & 963.76 & 951.96 \\ 
+Residual Standard Error (GLS) & 4.69 & 5.04 & 4.79 & 5.63 & 5.18 & 4.88 \\ 
+Log Likelihood & $-$402.11 & $-$410.52 & $-$404.94 & $-$423.96 & $-$412.62 & $-$404.34 \\ 
+Akaike Inf. Crit. & 868.21 & 885.05 & 873.88 & 911.92 & 889.24 & 872.68 \\ 
+Bayesian Inf. Crit. & 957.14 & 973.98 & 962.81 & 1,000.85 & 978.17 & 961.62 \\ 
 \hline 

### 2019-03-27 04-results.Rmd
created the faceting of effects - 1 row for each variable. still have to get the limits to work (fixed for each variable)

@@
 ### Diversity
-In the lead up to the 2001 election, the Liberal government (in power) refused permission for the MV Tampa, a ship carrying 433 rescued refugees and 5 crew to enter Australian waters.
-This, along with the 9/11 attacks, were two key issues that are widely believed to have pushed the Liberal party to victory.
-In this 2001 election, electorates with a higher proportion of people speaking languages other than english at home (`OtherLanguageHome`) were significantly more likely to support Labor.
-This remains negative up until the 2016 election albeit insignificant.
-Higher proportions of people born in Asia (`Born_Asia`) is also significantly linked with Labor support in 2010, which may be due to refugee policy again arising as a key election issue in lead up to the election.
+Larger migrant populations from Asia, the Middle East, South-Eastern Europe, the United Kingdom and elsewhere, are assocated with Labor support, or had no effect.
+Of these areas, only South-Eastern European populations were signficant in each election and Asian migrants were a significant in 2010.
+However, the effect of other languages (aside from English) is a far stronger effect, as observed through `OtherLanguageHome`.
+Surprisingly, electorates with more diverse speech were more likely to support the Liberal party from 2004 onwards, with this effect being significant in 2007, 2010 and 2016.
+
+Of the variables relating to religion, only Judaism shows a consistent effect, with electorates with larger Jewish populations more likely to vote Liberal.
 
@@
 p <- grid_visreg("Born_SE_Europe", xvar_title = "Born_SE_Europe", xlimits = c(-0.9, 5), ylimits = c(20, 70), plot = T)
+```
+
+```{r plotbornasia, fig.cap = "Effect of Born_Asia", out.width="60%", eval = F}
+p <- grid_visreg("Born_Asia", xvar_title = "Born_Asia", xlimits = c(-0.9, 5), ylimits = c(20, 70), plot = T)
+```
+
+```{r plotjudaism, fig.cap = "Effect of Judaism", out.width="60%", eval = F}
+p <- grid_visreg("Judaism", xvar_title = "Judaism", xlimits = c(-1.5, 3.8), ylimits = c(40, 70), plot = T)
+```
+
+### Partial residual plots
+
+```{r}
+vars_to_plot <- c("Judaism", "Born_Asia", "Born_SE_Europe", "OtherLanguageHome", "Education", "MedianAge", "DeFacto", "DiffAddress", "ManagerAdminClericalSales", "Transformative", "Extractive", "Unemployment", "Incomes")
+
+ls_visreg <- list(Judaism = grid_visreg("Judaism"), Born_Asia = grid_visreg("Born_Asia"), Born_SE_Europe = grid_visreg("Born_SE_Europe"), OtherLanguageHome = grid_visreg("OtherLanguageHome"), Education = grid_visreg("Education"), MedianAge = grid_visreg("MedianAge"), DeFacto = grid_visreg("DeFacto"), DiffAddress = grid_visreg("DiffAddress"), ManagerAdminClericalSales = grid_visreg("ManagerAdminClericalSales"), Transformative = grid_visreg("Transformative"), Extractive = grid_visreg("Extractive"), Unemployment = grid_visreg("Unemployment"), Incomes = grid_visreg("Incomes"))
+
+all_bands_df <- bind_rows(
+  ls_visreg$Judaism$bands %>% mutate(varname = "Judaism", xmin = -1.5, xmax = 3.8),
+  ls_visreg$Born_Asia$bands %>% mutate(varname = "Born_Asia", xmin = -0.9, xmax = 5.15),
+  ls_visreg$Born_SE_Europe$bands %>% mutate(varname = "Born_SE_Europe", xmin = -0.85, xmax = 5.05),
+  ls_visreg$OtherLanguageHome$bands %>% mutate(varname = "OtherLanguageHome", xmin = -1, xmax = 3.4),
+  ls_visreg$Education$bands %>% mutate(varname = "Education", xmin = -1.3, xmax = 3),
+  ls_visreg$MedianAge$bands %>% mutate(varname = "MedianAge", xmin = -2.55, xmax = 2.8),
+  ls_visreg$DeFacto$bands %>% mutate(varname = "DeFacto", xmin = -2.2, xmax = 5.05),
+  ls_visreg$DiffAddress$bands %>% mutate(varname = "DiffAddress", xmin = -2.2, xmax = 3.75),
+  ls_visreg$ManagerAdminClericalSales$bands %>% mutate(varname = "ManagerAdminClericalSales", xmin = -3.75, xmax = 2.85),
+  ls_visreg$Transformative$bands %>% mutate(varname = "Transformative", xmin = -2.3, xmax = 2.92),
+  ls_visreg$Extractive$bands %>% mutate(varname = "Extractive", xmin = -0.75, xmax = 3.45),
+  ls_visreg$Unemployment$bands %>% mutate(varname = "Unemployment", xmin = -2.08, xmax = 3.37),
+  ls_visreg$Incomes$bands %>% mutate(varname = "Incomes", xmin = -1.5, xmax = 3.65)
+)
+
+all_points_df <- bind_rows(
+  ls_visreg$Judaism$points %>% mutate(varname = "Judaism"),
+  ls_visreg$Born_Asia$points %>% mutate(varname = "Born_Asia"),
+  ls_visreg$Born_SE_Europe$points %>% mutate(varname = "Born_SE_Europe"),
+  ls_visreg$OtherLanguageHome$points %>% mutate(varname = "OtherLanguageHome"),
+  ls_visreg$Education$points %>% mutate(varname = "Education"),
+  ls_visreg$MedianAge$points %>% mutate(varname = "MedianAge"),
+  ls_visreg$DeFacto$points %>% mutate(varname = "DeFacto"),
+  ls_visreg$DiffAddress$points %>% mutate(varname = "DiffAddress"),
+  ls_visreg$ManagerAdminClericalSales$points %>% mutate(varname = "ManagerAdminClericalSales"),
+  ls_visreg$Transformative$points %>% mutate(varname = "Transformative"),
+  ls_visreg$Extractive$points %>% mutate(varname = "Extractive"),
+  ls_visreg$Unemployment$points %>% mutate(varname = "Unemployment"),
+  ls_visreg$Incomes$points %>% mutate(varname = "Incomes")
+)
+
+ggplot(data = all_bands_df %>% filter(varname %in% vars_to_plot[8:13])) +
+  geom_ribbon(aes(x = variable, ymin = lower95, ymax = upper95), fill = "grey80") + 
+  geom_point(aes(x = variable, y = part_res), data = all_points_df %>% filter(varname %in% vars_to_plot[8:13]), size = 0.75, col = "grey50") + 
+  geom_line(aes(x = variable, y = fitted), col = "blue", size = 1) +
+  theme_bw()+ 
+  labs(x = varname, y = "Response") + 
+  facet_wrap(varname ~ year, ncol = 6, scales = "free")
+
+my_x_ranges <- range(range(model_df$Judaism), range(model_df$Born_Asia), range(model_df$Born_SE_Europe), range(model_df$OtherLanguageHome), range(model_df$Education), range(model_df$MedianAge), range(model_df$DeFacto), range(model_df$DiffAddress), range(model_df$ManagerAdminClericalSales), range(model_df$Transformative), range(model_df$Extractive), range(model_df$Unemployment), range(model_df$Incomes))
+
+my_y_ranges <- range(
+  range(c(ls_visreg$Judaism$bands$upper95, ls_visreg$Judaism$bands$lower95, ls_visreg$Judaism$points$part_res)),
+  range(c(ls_visreg$Born_Asia$bands$upper95, ls_visreg$Born_Asia$bands$lower95, ls_visreg$Born_Asia$points$part_res)),
+  range(c(ls_visreg$Born_SE_Europe$bands$upper95, ls_visreg$Born_SE_Europe$bands$lower95, ls_visreg$Born_SE_Europe$points$part_res)),
+  range(c(ls_visreg$OtherLanguageHome$bands$upper95, ls_visreg$OtherLanguageHome$bands$lower95, ls_visreg$OtherLanguageHome$points$part_res)),
+  range(c(ls_visreg$Education$bands$upper95, ls_visreg$Education$bands$lower95, ls_visreg$Education$points$part_res)),
+  range(c(ls_visreg$MedianAge$bands$upper95, ls_visreg$MedianAge$bands$lower95, ls_visreg$MedianAge$points$part_res)),
+  range(c(ls_visreg$DeFacto$bands$upper95, ls_visreg$DeFacto$bands$lower95, ls_visreg$DeFacto$points$part_res)),
+  range(c(ls_visreg$DiffAddress$bands$upper95, ls_visreg$DiffAddress$bands$lower95, ls_visreg$DiffAddress$points$part_res)),
+  range(c(ls_visreg$ManagerAdminClericalSales$bands$upper95, ls_visreg$ManagerAdminClericalSales$bands$lower95, ls_visreg$ManagerAdminClericalSales$points$part_res)),
+  range(c(ls_visreg$Transformative$bands$upper95, ls_visreg$Transformative$bands$lower95, ls_visreg$Transformative$points$part_res)),
+  range(c(ls_visreg$Extractive$bands$upper95, ls_visreg$Extractive$bands$lower95, ls_visreg$Extractive$points$part_res)),
+  range(c(ls_visreg$Unemployment$bands$upper95, ls_visreg$Unemployment$bands$lower95, ls_visreg$Unemployment$points$part_res)),
+  range(c(ls_visreg$Incomes$bands$upper95, ls_visreg$Incomes$bands$lower95, ls_visreg$Incomes$points$part_res))
+  )
+
+ggplot(data = all_bands_df %>% filter(varname %in% vars_to_plot[12:13])) +
+  geom_ribbon(aes(x = variable, ymin = lower95, ymax = upper95), fill = "grey80") + 
+  geom_point(aes(x = variable, y = part_res), data = all_points_df %>% filter(varname %in% vars_to_plot[12:13]), size = 0.75, col = "grey50") + 
+  geom_line(aes(x = variable, y = fitted), col = "blue", size = 1) +
+  theme_bw() + 
+  labs(x = varname, y = "Response") + 
+  facet_wrap(varname ~ year, ncol = 6, scales = "free") +
+  expand_limits(x = my_x_ranges[23:26], y = my_y_ranges[23:26])
+
+# https://stackoverflow.com/questions/18046051/setting-individual-axis-limits-with-facet-wrap-and-scales-free-in-ggplot2
+
 ```

### 2019-03-28 04-results.Rmd
more playing with plots - still need to fix it to facet properly.
added section on outliers, and briefly proofread other results.

@@
 # Grid visreg
-grid_visreg <- function(varname, xvar_title = "", xlimits = NULL, ylimits = NULL, plot = FALSE) {
+grid_visreg <- function(varname, xvar_title = "", xlimits = NULL, ylimits = NULL, plot = FALSE, table = TRUE, years = FALSE) {
   if (plot == TRUE) {
+
+    if (years == TRUE) {
     p16 <- my_visreg(glsmod16, sp_weights_16, varname = varname, plot = T, nolabs = T, 
@@
       year = "2001", xlimits, ylimits)
+    } else {
+      p16 <- my_visreg(glsmod16, sp_weights_16, varname = varname, plot = T, nolabs = T, 
+      xlimits, ylimits)
+    p13 <- my_visreg(glsmod13, sp_weights_13, varname = varname, plot = T, nolabs = T, 
+      xlimits, ylimits)
+    p10 <- my_visreg(glsmod10, sp_weights_10, varname = varname, plot = T, nolabs = T, 
+      xlimits, ylimits)
+    p07 <- my_visreg(glsmod07, sp_weights_07, varname = varname, plot = T, nolabs = T, 
+      xlimits, ylimits)
+    p04 <- my_visreg(glsmod04, sp_weights_04, varname = varname, plot = T, nolabs = T, 
+      xlimits, ylimits)
+    p01 <- my_visreg(glsmod01, sp_weights_01, varname = varname, plot = T, nolabs = T, 
+      xlimits, ylimits)
+    }
 
-    plots <- grid.arrange(p01, p04, p07, p10, p13, p16, nrow = 2, 
+    if (table == TRUE) {
+      plots <- grid.arrange(p01, p04, p07, p10, p13, p16, nrow = 1, 
+        bottom = grid.text(xvar_title, gp = gpar(fontface = "bold", cex = 0.6)))
+    } else {
+      plots <- grid.arrange(p01, p04, p07, p10, p13, p16, nrow = 1, 
         left = grid.text("Two-party preferred vote (%)", gp = gpar(cex = 0.8), rot = 90), 
         bottom = grid.text(xvar_title, gp = gpar(cex = 0.8)))
+    }
 
@@
 }
-
 ```
@@
 ```{r plotincomes, fig.cap = "Effect of Incomes", out.width="100%"}
-p <- grid_visreg("Incomes", xvar_title = "Incomes", plot = T, xlimits = c(-1.5, 3.6), ylimits = c(30, 95))
+p <- grid_visreg("Incomes", xvar_title = "Incomes", plot = T, xlimits = c(-1.5, 3.6), ylimits = c(30, 95), years = FALSE)
 ```
 
-```{r plotunemploy, fig.cap = "Effect of Unemployment", out.width="60%"}
+```{r plotunemploy, fig.cap = "Effect of Unemployment", out.width="100%"}
 p <- grid_visreg("Unemployment", xvar_title = "Unemployment", plot = T, xlimits = c(-2.1, 3.4), ylimits = c(28, 68))
@@
 ### Relationships and age
-De facto relationships, but not marriages, is found to be an important (and significant) predictor of the two-party preferred vote in all six elections, with more de facto relationships associated with higher support for the Labor party.
+De facto relationships, but not marriages, are found to be an important (and significant) predictor of the two-party preferred vote in all six elections, with more de facto relationships associated with higher support for the Labor party.
+Marriages however, are insignficiant.
 
@@
 
-Although only significant in 2007 and 2010, older electorates (higher median age) are consistently more likely to support the Liberal party.
+Older regions are often believed to be more conservative, and it can be seen that electorates with a higher median age have been more likely to support the Liberal party - although this effect is only significant in 2007 and 2010.
 
@@
 Since 2007, electorates with higher education levels have been associated with supporting the Labor party, although this effect is only significant in 2016.
-Prior to 2007, education had no material effect.
+Prior to 2007, education had an almost zero effect.
 
@@
 ### Diversity
-Larger migrant populations from Asia, the Middle East, South-Eastern Europe, the United Kingdom and elsewhere, are assocated with Labor support, or had no effect.
+Larger migrant populations from Asia, the Middle East, South-Eastern Europe, the United Kingdom and elsewhere, are either assocated with Labor support, or have no effect.
 Of these areas, only South-Eastern European populations were signficant in each election and Asian migrants were a significant in 2010.
-However, the effect of other languages (aside from English) is a far stronger effect, as observed through `OtherLanguageHome`.
-Surprisingly, electorates with more diverse speech were more likely to support the Liberal party from 2004 onwards, with this effect being significant in 2007, 2010 and 2016.
-
-Of the variables relating to religion, only Judaism shows a consistent effect, with electorates with larger Jewish populations more likely to vote Liberal.
+Speaking other language (aside from English) however, have a far stronger effect, as observed through `OtherLanguageHome`.
+Electorates with more diverse speech were more likely to support the Liberal party from 2004 onwards, with this effect being significant in 2007, 2010 and 2016.
+Furthermore, of the variables relating to religion, only Judaism shows a consistent effect, with electorates with larger Jewish populations more likely to vote Liberal.
 
@@
 
-```{r}
+*Move the individual plots in each section to a global grid here*
+
+```{r, eval = F}
 vars_to_plot <- c("Judaism", "Born_Asia", "Born_SE_Europe", "OtherLanguageHome", "Education", "MedianAge", "DeFacto", "DiffAddress", "ManagerAdminClericalSales", "Transformative", "Extractive", "Unemployment", "Incomes")
@@
 
-ggplot(data = all_bands_df %>% filter(varname %in% vars_to_plot[12:13])) +
+ggplot(data = all_bands_df %>% filter(varname %in% vars_to_plot[13])) +
   geom_ribbon(aes(x = variable, ymin = lower95, ymax = upper95), fill = "grey80") + 
-  geom_point(aes(x = variable, y = part_res), data = all_points_df %>% filter(varname %in% vars_to_plot[12:13]), size = 0.75, col = "grey50") + 
+  geom_point(aes(x = variable, y = part_res), data = all_points_df %>% filter(varname %in% vars_to_plot[13]), size = 0.75, col = "grey50") + 
   geom_line(aes(x = variable, y = fitted), col = "blue", size = 1) +
@@
   facet_wrap(varname ~ year, ncol = 6, scales = "free") +
-  expand_limits(x = my_x_ranges[23:26], y = my_y_ranges[23:26])
+  expand_limits(x = c(-2, 2))
+  expand_limits(x = my_x_ranges[25:26], y = my_y_ranges[25:26])
 
 # https://stackoverflow.com/questions/18046051/setting-individual-axis-limits-with-facet-wrap-and-scales-free-in-ggplot2
+
+```
+
+```{r tablegrids, eval = F}
+vars_to_plot <- c("Judaism", "Born_Asia", "Born_SE_Europe", "OtherLanguageHome", "Education", "MedianAge", "DeFacto", "DiffAddress", "ManagerAdminClericalSales", "Transformative", "Extractive", "Unemployment", "Incomes")
+
+p_Judaism <- grid_visreg("Judaism", xvar_title = "Judaism", xlimits = c(-1.5, 3.8), ylimits = c(40, 70), plot = T, years = T)
+p_Born_Asia <- grid_visreg("Born_Asia", xvar_title = "Born_Asia", xlimits = c(-0.9, 5), ylimits = c(20, 70), plot = T)
+p_Born_SE_Europe <- grid_visreg("Born_SE_Europe", xvar_title = "Born_SE_Europe", xlimits = c(-0.9, 5), ylimits = c(20, 70), plot = T)
+p_OtherLanguageHome <- grid_visreg("OtherLanguageHome", xvar_title = "OtherLanguageHome", xlimits = c(-1.1, 3.5), ylimits = c(5, 80), plot = T)
+p_Education <- grid_visreg("Education", xvar_title = "Education", xlimits = c(-2.5, 2.7), ylimits = c(31, 70), plot = T)
+
+grid.arrange(p_Judaism, p_Born_Asia, ncol = 1, left = grid.text("Two-party preferred vote (%)", gp = gpar(cex = 0.8), rot = 90))
 
@@
 However, there still remain some variables which are closely related.
-For example, `OtherLanguageHome` is the proportion of households that speak a language other than English at home, which is likely to be related to the proportion of people born overseas (a sum of the variables with a `Born` prefix).
-In 2004, the coefficient estiamte for `OtherLanguageHome` negative but not significant, and all `Born` variables are insignificant.
-If all `Born` variables are removed from the model, `OtherLanguageHome` absorbs the negative effect, becoming significant ($p = 0.01$).
+For example, electorate income levels (via `Incomes`) is likely to be related to electoral unemployment and labor force participation (via `Unemployment`).
+In 2001, the coefficient estiamte for `Unemployment` negative but not significant, and the `Incomes` variables  is significant.
+If the `Incomes` variable is removed from the model, `Unemployment` absorbs the negative effect, becoming significant ($p = 0.0056$).
 
 ```{r removeolh, eval = F}
-OLH04 <- my_fgls(full_formula, 
-  my_data = model_df %>% filter(year == "2004") %>% dplyr::select(-starts_with("Born")),
-  sp_weights = sp_weights_04)
+unemp01 <- my_fgls(full_formula, 
+  my_data = model_df %>% filter(year == "2001") %>% dplyr::select(-Incomes),
+  sp_weights = sp_weights_01)
 
-summary(OLH04)$tTable
+summary(unemp01)$tTable["Unemployment",]
 ```
@@
 It is often hypothesized that states have systematic differences that cause their electorates to vote differently.
-Boxplots of residuals grouped by state reveal that only Tasmania and the Australian Capital Territory appear to have a state-specific effect that is not captured by the models.
-These states appear to have a bias towards Labor.
-There are few electorates in these states (five and two, respectively), so this might be due to incumbent effects rather than an actual state-specific bias.
+Boxplots of residuals grouped by state reveal that only Tasmania, the Australian Capital Territory and the Northern Territory appear to have a state-specific effect that is not captured by the models.
+Tasmania and the Australian Capital Territory appear to have a bias towards Labor, whereas the Northern Territory has one towards Liberal.
+There are few electorates in these states (five, two and two respectively), so this might be due to incumbent effects rather than an actual state-specific bias.
 
@@
 ### Outlier electorates
-
-In
 
@@
   coord_fixed(ratio = .5) +
-  labs(x = "Cook's Distance", y = "")
+  labs(x = "Cook's Distance", y = "") + 
+  theme(axis.text.x = element_text(angle = 60, hjust = 1, size = 8), axis.text.y = element_text(size = 8))
 ```
@@
 cooksd_df %>% 
-  filter(cooksd > 0.13) %>% 
+  filter(cooksd > 0.1) %>% 
   select(DivisionNm, StateAb, year, cooksd, Residuals, LNP_Percent, everything()) %>% View
@@
 The dotplots in Figure \@ref(fig:plotcooksd) show that there are a few outliers in each election, using cook's distance as a measure of influence.
-The electorate of Sydney (NSW) has the largest cook's distance in 2001, 2004, 2007 and 2010, due to its diverse population (language, birthplace and religion), high number of defacto relationships,  high income, high household mobility and small amount of workers in extractive, distributive and transformative jobs.
-It has remained a strong supporter of the Labor party and is severely underpredicted by the model.
-Nearby in metropolitan NSW, the electorate of Wentworth is found to be an outlier in the 2010, 2013 and 2016 elections - coinciding with its Liberal member, Malcolm Turnbull, being the leader of the Liberal party.
-Wentworth was historically Liberal but increased its two-party vote by over 10 percentage points in 2010 with Turnbull as Liberal leader, whilst experiencing minimal change in its socio-demographics, causing the models to heavily underpredict its Liberal preference.
+Based on the distribution of the cook's distance values, a cook's distance greater than $0.1$ is considered to be an outlier.
+The electorate of Sydney (NSW) has a large cook's distance from 2001 to 2013, due to its diverse population (language, birthplace and religion), high number of defacto relationships, high income, high household mobility and small amount of workers in extractive, distributive and transformative jobs.
+It has remained a strong supporter of the Labor party and Liberal vote is severely overpredicted by the model.
+Nearby in metropolitan NSW, the electorate of Wentworth is found to be an outlier in all but the 2007 election.
+Although historically Liberal, its two-party vote jumped by over 10 percentage points in 2010 without experiencing any notable changes in its socio-demographic makeup - implying that this may be the direct effect of its Liberal member, Malcolm Turnbull, becoming the leader of the Liberal party.
+Liberal support in Wentworth is underpredicted by the model in each year, and more so with Turnbull as Liberal leader.
 
-Melbourne Ports, an upper-middle class electorate just south of Melbourne's the central business district,
+Lingiari, an electorate taking up almost all of the Northern Territory, is an outlier in the 2001-2007 elections.
+This is likely due to its large Indigenous population, young age profile and low rates of property ownership.
+Other electorates with large cook's distance are Barton (NSW, 2016) and Leichhardt (QLD) in 2016, and Canberra (ACT) in 2007.
 
-### Electorates with largest residuals
+```{r, eval = F}
+x <- model_df %>% filter(year == "2016") %>% select(-c(year, DivisionNm)) %>% as.matrix
+myhat <- hat(x)
 
-The electorates Capricornia (QLD), Leichhardt (QLD) and Riverina (NSW) have the three largest maximum absolute residuals across the six elections.
-Capricornia is a swing electorate, which strongly supported Labor between 2001 and 2010, before being narrowly won by Liberal in the last two elections.
-It has many people working in extractive jobs, an above-average Christian population, and above-average proportion of de facto relationships.
-The model tends to overpredict its two-party preferred vote (in favour of the Liberal party).
-Leichhardt is an area encompassing the northernmost tip of Queensland and consists of mainly small indigenous communities and uninhabited rural areas, which was won by the Liberal party in all but the 2007 election.
-It too has an above-average proportion of de facto relationships, along with an above-average proportion of people moving houses and a high proportion of single parent households.
-The model consistently underpredicts its two-party vote.
-Riverina is a rural electorate of NSW, and has been a very safe Liberal (National) seat since 1980.
-It has a relatively monocultural population consisting of a high proportion of Australian citizens and Christians, an above-average amount of extractive jobs and below average education levels.
-Its Liberal preference is also underpredicted.
+plot(glsmod10$cooksd, hat(model_df %>% filter(year == "2010") %>% select(-c(year, DivisionNm)) %>% as.matrix))
+```
 

### 2019-03-29 04-results.Rmd
updated gridvisreg

@@
 # Grid visreg
-grid_visreg <- function(varname, xvar_title = "", xlimits = NULL, ylimits = NULL, plot = FALSE, table = TRUE, years = FALSE) {
-  if (plot == TRUE) {
+grid_visreg <- function(varname, plot = TRUE, myscale = "free") {
 
-    if (years == TRUE) {
-    p16 <- my_visreg(glsmod16, sp_weights_16, varname = varname, plot = T, nolabs = T, 
-      year = "2016", xlimits, ylimits)
-    p13 <- my_visreg(glsmod13, sp_weights_13, varname = varname, plot = T, nolabs = T, 
-      year = "2013", xlimits, ylimits)
-    p10 <- my_visreg(glsmod10, sp_weights_10, varname = varname, plot = T, nolabs = T, 
-      year = "2010", xlimits, ylimits)
-    p07 <- my_visreg(glsmod07, sp_weights_07, varname = varname, plot = T, nolabs = T, 
-      year = "2007", xlimits, ylimits)
-    p04 <- my_visreg(glsmod04, sp_weights_04, varname = varname, plot = T, nolabs = T, 
-      year = "2004", xlimits, ylimits)
-    p01 <- my_visreg(glsmod01, sp_weights_01, varname = varname, plot = T, nolabs = T, 
-      year = "2001", xlimits, ylimits)
-    } else {
-      p16 <- my_visreg(glsmod16, sp_weights_16, varname = varname, plot = T, nolabs = T, 
-      xlimits, ylimits)
-    p13 <- my_visreg(glsmod13, sp_weights_13, varname = varname, plot = T, nolabs = T, 
-      xlimits, ylimits)
-    p10 <- my_visreg(glsmod10, sp_weights_10, varname = varname, plot = T, nolabs = T, 
-      xlimits, ylimits)
-    p07 <- my_visreg(glsmod07, sp_weights_07, varname = varname, plot = T, nolabs = T, 
-      xlimits, ylimits)
-    p04 <- my_visreg(glsmod04, sp_weights_04, varname = varname, plot = T, nolabs = T, 
-      xlimits, ylimits)
-    p01 <- my_visreg(glsmod01, sp_weights_01, varname = varname, plot = T, nolabs = T, 
-      xlimits, ylimits)
-    }
-
-    if (table == TRUE) {
-      plots <- grid.arrange(p01, p04, p07, p10, p13, p16, nrow = 1, 
-        bottom = grid.text(xvar_title, gp = gpar(fontface = "bold", cex = 0.6)))
-    } else {
-      plots <- grid.arrange(p01, p04, p07, p10, p13, p16, nrow = 1, 
-        left = grid.text("Two-party preferred vote (%)", gp = gpar(cex = 0.8), rot = 90), 
-        bottom = grid.text(xvar_title, gp = gpar(cex = 0.8)))
-    }
-
-  return(plots)
-  }
-
-  if (plot == FALSE) {
     p16 <- my_visreg(glsmod16, sp_weights_16, varname = varname, plot = F, year = "2016")
@@
       p01$bands %>% mutate(year = "2001")
-    )
+    ) %>% mutate(varname = varname)
 
@@
       p01$points %>% mutate(year = "2001")
-    )
+    ) %>% mutate(varname = varname)
 
-    return_df <- list(bands = bands_df, points = points_df)
+    if (plot == T) {
+      return_object <- ggplot(data = bands_df) +
+        geom_ribbon(aes(x = variable, ymin = lower95, ymax = upper95), fill = "grey80") + 
+        geom_point(aes(x = variable, y = part_res), data = points_df, size = 0.5, shape = 1, alpha = 0.5) +
+        geom_line(aes(x = variable, y = fitted), col = "blue", size = 1) +
+        theme_bw() + 
+        labs(x = "", y = "") + 
+        facet_grid(varname ~ year, scales = "free") + 
+        theme(strip.text.x = element_blank())
 
-    return(return_df)
+      if (myscale == "free") {
+        return_object <- return_object + lims(x = range(model_df %>% select(varname)), 
+          y = range(c(bands_df$upper95, bands_df$lower95, points_df$part_res)))
       }
 
+      if (myscale == "free_x") {
+        return_object <- return_object + lims(x = range(model_df %>% select(varname)))
       }
+
+      if (myscale == "free_y") {
+        return_object <- return_object + lims(y = range(c(bands_df$upper95, bands_df$lower95, points_df$part_res)))
+      }
+
+      if (myscale == "none") {
+        return_object <- return_object
+      }
+    }
+
+    if (plot == F) {
+      return_object <- list(bands = bands_df, points = points_df)
+    }
+
+    return(return_object)
+
+}
+```
+
+```{r}
+grid_visreg(varname = "Incomes", plot = T)
 ```
@@
 
+### Industry and type of work
+Electorates with higher proportions of workers in mining, gas, water, agriculture, waste and electricity (grouped as `Extractive` industries) are consistently linked with higher support for the Liberal party, with the mangitude of this effect slightly increasing over the years (see Figure \@ref(fig:plotextractive)).
+This is unsurprising, as the Liberal party has close ties with these traditional energy industries, and typically present policies to reduce taxation on energy production.
+Furthermore, electorates with more workers in construction or manufacturing industries (`Transformative`) are also more likely to support the Liberal party (see Figure \@ref(fig:plottransformative)).
+
+Similarly, workers in managerial, administrative, clerical and sales roles (`ManagerAdminClericalSales`) is also a significant predictor of two-party preference vote across all six elections, with a higher proportion of people working these jobs increasing Liberal support.
+The magnitude of this effect has also increased over the years.
+
+### Household mobility
+In each of the six elections, electorates with a higher proportion of people that have recently moved house (meaning in last five years) were more likely to support the Liberal party, although this effect was marginally insignificant in 2013.
+Having controlled for characteristics of house ownership and rental prices (via the factors `PropertyOwned` and `RentLoan` respectively), this is effect is somewhat surprising.
+
+### Relationships and age
+De facto relationships, but not marriages, are found to be an important (and significant) predictor of the two-party preferred vote in all six elections, with more de facto relationships associated with higher support for the Labor party.
+Marriages however, are insignficiant.
+
+Older regions are often believed to be more conservative, and it can be seen that electorates with a higher median age have been more likely to support the Liberal party - although this effect is only significant in 2007 and 2010.
+
+### Education
+Since 2007, electorates with higher education levels have been associated with supporting the Labor party, although this effect is only significant in 2016.
+Prior to 2007, education had an almost zero effect.
+
+### Diversity
+Larger migrant populations from Asia, the Middle East, South-Eastern Europe, the United Kingdom and elsewhere, are either assocated with Labor support, or have no effect.
+Of these areas, only South-Eastern European populations were signficant in each election and Asian migrants were a significant in 2010.
+Speaking other language (aside from English) however, have a far stronger effect, as observed through `OtherLanguageHome`.
+Electorates with more diverse speech were more likely to support the Liberal party from 2004 onwards, with this effect being significant in 2007, 2010 and 2016.
+Furthermore, of the variables relating to religion, only Judaism shows a consistent effect, with electorates with larger Jewish populations more likely to vote Liberal.
+
+### Partial residual plots
+
+#### Individual*
+
 ```{r plotincomes, fig.cap = "Effect of Incomes", out.width="100%"}
@@
 ```
-
-### Industry and type of work
-Electorates with higher proportions of workers in mining, gas, water, agriculture, waste and electricity (grouped as `Extractive` industries) are consistently linked with higher support for the Liberal party, with the mangitude of this effect slightly increasing over the years (see Figure \@ref(fig:plotextractive)).
-This is unsurprising, as the Liberal party has close ties with these traditional energy industries, and typically present policies to reduce taxation on energy production.
-Furthermore, electorates with more workers in construction or manufacturing industries (`Transformative`) are also more likely to support the Liberal party (see Figure \@ref(fig:plottransformative)).
 
@@
 
-Similarly, workers in managerial, administrative, clerical and sales roles (`ManagerAdminClericalSales`) is also a significant predictor of two-party preference vote across all six elections, with a higher proportion of people working these jobs increasing Liberal support.
-The magnitude of this effect has also increased over the years.
-
 ```{r plotmanager, fig.cap = "Effect of ManagerAdminClericalSales", out.width="60%"}
@@
 
-### Household mobility
 ```{r plotdiffaddress, fig.cap = "Effect of DiffAddress", out.width="60%"}
@@
 ```
-
-In each of the six elections, electorates with a higher proportion of people that have recently moved house (meaning in last five years) were more likely to support the Liberal party, although this effect was marginally insignificant in 2013.
-Having controlled for characteristics of house ownership and rental prices (via the factors `PropertyOwned` and `RentLoan` respectively), this is effect is somewhat surprising.
-
-### Relationships and age
-De facto relationships, but not marriages, are found to be an important (and significant) predictor of the two-party preferred vote in all six elections, with more de facto relationships associated with higher support for the Labor party.
-Marriages however, are insignficiant.
 
@@
 
-Older regions are often believed to be more conservative, and it can be seen that electorates with a higher median age have been more likely to support the Liberal party - although this effect is only significant in 2007 and 2010.
-
 ```{r plotage, fig.cap = "Effect of MedianAge", out.width="60%", eval = F}
@@
 
-### Education
-Since 2007, electorates with higher education levels have been associated with supporting the Labor party, although this effect is only significant in 2016.
-Prior to 2007, education had an almost zero effect.
-
 ```{r ploteducation, fig.cap = "Effect of Education", out.width="60%", eval = F}
@@
 ```
-
-### Diversity
-Larger migrant populations from Asia, the Middle East, South-Eastern Europe, the United Kingdom and elsewhere, are either assocated with Labor support, or have no effect.
-Of these areas, only South-Eastern European populations were signficant in each election and Asian migrants were a significant in 2010.
-Speaking other language (aside from English) however, have a far stronger effect, as observed through `OtherLanguageHome`.
-Electorates with more diverse speech were more likely to support the Liberal party from 2004 onwards, with this effect being significant in 2007, 2010 and 2016.
-Furthermore, of the variables relating to religion, only Judaism shows a consistent effect, with electorates with larger Jewish populations more likely to vote Liberal.
 
@@
 
-### Partial residual plots
-
-*Move the individual plots in each section to a global grid here*
+#### *Move the individual plots in each section to a global grid here*
 
@@
 
-ggplot(data = all_bands_df %>% filter(varname %in% vars_to_plot[8:13])) +
+ggplot(data = all_bands_df %>% filter(varname %in% vars_to_plot[1:6])) +
   geom_ribbon(aes(x = variable, ymin = lower95, ymax = upper95), fill = "grey80") + 
-  geom_point(aes(x = variable, y = part_res), data = all_points_df %>% filter(varname %in% vars_to_plot[8:13]), size = 0.75, col = "grey50") + 
+  geom_point(aes(x = variable, y = part_res), data = all_points_df %>% filter(varname %in% vars_to_plot[1:6]), size = 0.5, shape = 1, alpha = 0.5) + 
   geom_line(aes(x = variable, y = fitted), col = "blue", size = 1) +
   theme_bw()+ 
-  labs(x = varname, y = "Response") + 
-  facet_wrap(varname ~ year, ncol = 6, scales = "free")
+  labs(x = "", y = "Two-party preferred vote (%)") + 
+  facet_grid(varname ~ year, scales = "free_x") + theme(strip.text.x = element_blank())
 
-my_x_ranges <- range(range(model_df$Judaism), range(model_df$Born_Asia), range(model_df$Born_SE_Europe), range(model_df$OtherLanguageHome), range(model_df$Education), range(model_df$MedianAge), range(model_df$DeFacto), range(model_df$DiffAddress), range(model_df$ManagerAdminClericalSales), range(model_df$Transformative), range(model_df$Extractive), range(model_df$Unemployment), range(model_df$Incomes))
+my_x_ranges <- c(range(model_df$Judaism), range(model_df$Born_Asia), range(model_df$Born_SE_Europe), range(model_df$OtherLanguageHome), range(model_df$Education), range(model_df$MedianAge), range(model_df$DeFacto), range(model_df$DiffAddress), range(model_df$ManagerAdminClericalSales), range(model_df$Transformative), range(model_df$Extractive), range(model_df$Unemployment), range(model_df$Incomes))
 
-my_y_ranges <- range(
+my_y_ranges <- c(
   range(c(ls_visreg$Judaism$bands$upper95, ls_visreg$Judaism$bands$lower95, ls_visreg$Judaism$points$part_res)),
@@
 
-ggplot(data = all_bands_df %>% filter(varname %in% vars_to_plot[13])) +
+ggplot(data = all_bands_df %>% filter(varname %in% vars_to_plot[1])) +
   geom_ribbon(aes(x = variable, ymin = lower95, ymax = upper95), fill = "grey80") + 
-  geom_point(aes(x = variable, y = part_res), data = all_points_df %>% filter(varname %in% vars_to_plot[13]), size = 0.75, col = "grey50") + 
+  geom_point(aes(x = variable, y = part_res), data = all_points_df %>% filter(varname %in% vars_to_plot[1]), size = 0.5, shape = 1, alpha = 0.5) + 
   geom_line(aes(x = variable, y = fitted), col = "blue", size = 1) +
-  theme_bw() + 
-  labs(x = varname, y = "Response") + 
-  facet_wrap(varname ~ year, ncol = 6, scales = "free") +
-  expand_limits(x = c(-2, 2))
-  expand_limits(x = my_x_ranges[25:26], y = my_y_ranges[25:26])
+  theme_bw()+ 
+  labs(x = "", y = "") + 
+  facet_grid(varname ~ year, scales = "free") + theme(strip.text.x = element_blank()) +
+  expand_limits(x = my_x_ranges[1:2], y = my_y_ranges[1:2])
+
+, y = my_y_ranges[1:12])
+"Two-party preferred vote (%)"
 
@@
 
-```{r}
+```{r, eval = F}
 cooksd_df %>% 

### 2019-03-29 04-results.Rmd
removed old attempts at faceting

@@
 
-#### *Move the individual plots in each section to a global grid here*
-
-```{r, eval = F}
-vars_to_plot <- c("Judaism", "Born_Asia", "Born_SE_Europe", "OtherLanguageHome", "Education", "MedianAge", "DeFacto", "DiffAddress", "ManagerAdminClericalSales", "Transformative", "Extractive", "Unemployment", "Incomes")
-
-ls_visreg <- list(Judaism = grid_visreg("Judaism"), Born_Asia = grid_visreg("Born_Asia"), Born_SE_Europe = grid_visreg("Born_SE_Europe"), OtherLanguageHome = grid_visreg("OtherLanguageHome"), Education = grid_visreg("Education"), MedianAge = grid_visreg("MedianAge"), DeFacto = grid_visreg("DeFacto"), DiffAddress = grid_visreg("DiffAddress"), ManagerAdminClericalSales = grid_visreg("ManagerAdminClericalSales"), Transformative = grid_visreg("Transformative"), Extractive = grid_visreg("Extractive"), Unemployment = grid_visreg("Unemployment"), Incomes = grid_visreg("Incomes"))
-
-all_bands_df <- bind_rows(
-  ls_visreg$Judaism$bands %>% mutate(varname = "Judaism", xmin = -1.5, xmax = 3.8),
-  ls_visreg$Born_Asia$bands %>% mutate(varname = "Born_Asia", xmin = -0.9, xmax = 5.15),
-  ls_visreg$Born_SE_Europe$bands %>% mutate(varname = "Born_SE_Europe", xmin = -0.85, xmax = 5.05),
-  ls_visreg$OtherLanguageHome$bands %>% mutate(varname = "OtherLanguageHome", xmin = -1, xmax = 3.4),
-  ls_visreg$Education$bands %>% mutate(varname = "Education", xmin = -1.3, xmax = 3),
-  ls_visreg$MedianAge$bands %>% mutate(varname = "MedianAge", xmin = -2.55, xmax = 2.8),
-  ls_visreg$DeFacto$bands %>% mutate(varname = "DeFacto", xmin = -2.2, xmax = 5.05),
-  ls_visreg$DiffAddress$bands %>% mutate(varname = "DiffAddress", xmin = -2.2, xmax = 3.75),
-  ls_visreg$ManagerAdminClericalSales$bands %>% mutate(varname = "ManagerAdminClericalSales", xmin = -3.75, xmax = 2.85),
-  ls_visreg$Transformative$bands %>% mutate(varname = "Transformative", xmin = -2.3, xmax = 2.92),
-  ls_visreg$Extractive$bands %>% mutate(varname = "Extractive", xmin = -0.75, xmax = 3.45),
-  ls_visreg$Unemployment$bands %>% mutate(varname = "Unemployment", xmin = -2.08, xmax = 3.37),
-  ls_visreg$Incomes$bands %>% mutate(varname = "Incomes", xmin = -1.5, xmax = 3.65)
-)
-
-all_points_df <- bind_rows(
-  ls_visreg$Judaism$points %>% mutate(varname = "Judaism"),
-  ls_visreg$Born_Asia$points %>% mutate(varname = "Born_Asia"),
-  ls_visreg$Born_SE_Europe$points %>% mutate(varname = "Born_SE_Europe"),
-  ls_visreg$OtherLanguageHome$points %>% mutate(varname = "OtherLanguageHome"),
-  ls_visreg$Education$points %>% mutate(varname = "Education"),
-  ls_visreg$MedianAge$points %>% mutate(varname = "MedianAge"),
-  ls_visreg$DeFacto$points %>% mutate(varname = "DeFacto"),
-  ls_visreg$DiffAddress$points %>% mutate(varname = "DiffAddress"),
-  ls_visreg$ManagerAdminClericalSales$points %>% mutate(varname = "ManagerAdminClericalSales"),
-  ls_visreg$Transformative$points %>% mutate(varname = "Transformative"),
-  ls_visreg$Extractive$points %>% mutate(varname = "Extractive"),
-  ls_visreg$Unemployment$points %>% mutate(varname = "Unemployment"),
-  ls_visreg$Incomes$points %>% mutate(varname = "Incomes")
-)
-
-ggplot(data = all_bands_df %>% filter(varname %in% vars_to_plot[1:6])) +
-  geom_ribbon(aes(x = variable, ymin = lower95, ymax = upper95), fill = "grey80") + 
-  geom_point(aes(x = variable, y = part_res), data = all_points_df %>% filter(varname %in% vars_to_plot[1:6]), size = 0.5, shape = 1, alpha = 0.5) + 
-  geom_line(aes(x = variable, y = fitted), col = "blue", size = 1) +
-  theme_bw()+ 
-  labs(x = "", y = "Two-party preferred vote (%)") + 
-  facet_grid(varname ~ year, scales = "free_x") + theme(strip.text.x = element_blank())
-
-my_x_ranges <- c(range(model_df$Judaism), range(model_df$Born_Asia), range(model_df$Born_SE_Europe), range(model_df$OtherLanguageHome), range(model_df$Education), range(model_df$MedianAge), range(model_df$DeFacto), range(model_df$DiffAddress), range(model_df$ManagerAdminClericalSales), range(model_df$Transformative), range(model_df$Extractive), range(model_df$Unemployment), range(model_df$Incomes))
-
-my_y_ranges <- c(
-  range(c(ls_visreg$Judaism$bands$upper95, ls_visreg$Judaism$bands$lower95, ls_visreg$Judaism$points$part_res)),
-  range(c(ls_visreg$Born_Asia$bands$upper95, ls_visreg$Born_Asia$bands$lower95, ls_visreg$Born_Asia$points$part_res)),
-  range(c(ls_visreg$Born_SE_Europe$bands$upper95, ls_visreg$Born_SE_Europe$bands$lower95, ls_visreg$Born_SE_Europe$points$part_res)),
-  range(c(ls_visreg$OtherLanguageHome$bands$upper95, ls_visreg$OtherLanguageHome$bands$lower95, ls_visreg$OtherLanguageHome$points$part_res)),
-  range(c(ls_visreg$Education$bands$upper95, ls_visreg$Education$bands$lower95, ls_visreg$Education$points$part_res)),
-  range(c(ls_visreg$MedianAge$bands$upper95, ls_visreg$MedianAge$bands$lower95, ls_visreg$MedianAge$points$part_res)),
-  range(c(ls_visreg$DeFacto$bands$upper95, ls_visreg$DeFacto$bands$lower95, ls_visreg$DeFacto$points$part_res)),
-  range(c(ls_visreg$DiffAddress$bands$upper95, ls_visreg$DiffAddress$bands$lower95, ls_visreg$DiffAddress$points$part_res)),
-  range(c(ls_visreg$ManagerAdminClericalSales$bands$upper95, ls_visreg$ManagerAdminClericalSales$bands$lower95, ls_visreg$ManagerAdminClericalSales$points$part_res)),
-  range(c(ls_visreg$Transformative$bands$upper95, ls_visreg$Transformative$bands$lower95, ls_visreg$Transformative$points$part_res)),
-  range(c(ls_visreg$Extractive$bands$upper95, ls_visreg$Extractive$bands$lower95, ls_visreg$Extractive$points$part_res)),
-  range(c(ls_visreg$Unemployment$bands$upper95, ls_visreg$Unemployment$bands$lower95, ls_visreg$Unemployment$points$part_res)),
-  range(c(ls_visreg$Incomes$bands$upper95, ls_visreg$Incomes$bands$lower95, ls_visreg$Incomes$points$part_res))
-  )
-
-ggplot(data = all_bands_df %>% filter(varname %in% vars_to_plot[1])) +
-  geom_ribbon(aes(x = variable, ymin = lower95, ymax = upper95), fill = "grey80") + 
-  geom_point(aes(x = variable, y = part_res), data = all_points_df %>% filter(varname %in% vars_to_plot[1]), size = 0.5, shape = 1, alpha = 0.5) + 
-  geom_line(aes(x = variable, y = fitted), col = "blue", size = 1) +
-  theme_bw()+ 
-  labs(x = "", y = "") + 
-  facet_grid(varname ~ year, scales = "free") + theme(strip.text.x = element_blank()) +
-  expand_limits(x = my_x_ranges[1:2], y = my_y_ranges[1:2])
-
-, y = my_y_ranges[1:12])
-"Two-party preferred vote (%)"
-
-# https://stackoverflow.com/questions/18046051/setting-individual-axis-limits-with-facet-wrap-and-scales-free-in-ggplot2
-
-```
-
-```{r tablegrids, eval = F}
-vars_to_plot <- c("Judaism", "Born_Asia", "Born_SE_Europe", "OtherLanguageHome", "Education", "MedianAge", "DeFacto", "DiffAddress", "ManagerAdminClericalSales", "Transformative", "Extractive", "Unemployment", "Incomes")
-
-p_Judaism <- grid_visreg("Judaism", xvar_title = "Judaism", xlimits = c(-1.5, 3.8), ylimits = c(40, 70), plot = T, years = T)
-p_Born_Asia <- grid_visreg("Born_Asia", xvar_title = "Born_Asia", xlimits = c(-0.9, 5), ylimits = c(20, 70), plot = T)
-p_Born_SE_Europe <- grid_visreg("Born_SE_Europe", xvar_title = "Born_SE_Europe", xlimits = c(-0.9, 5), ylimits = c(20, 70), plot = T)
-p_OtherLanguageHome <- grid_visreg("OtherLanguageHome", xvar_title = "OtherLanguageHome", xlimits = c(-1.1, 3.5), ylimits = c(5, 80), plot = T)
-p_Education <- grid_visreg("Education", xvar_title = "Education", xlimits = c(-2.5, 2.7), ylimits = c(31, 70), plot = T)
-
-grid.arrange(p_Judaism, p_Born_Asia, ncol = 1, left = grid.text("Two-party preferred vote (%)", gp = gpar(cex = 0.8), rot = 90))
-
-```
-
 ### A note on similar variables

### 2019-03-29 old_book_files/02-data.Rmd
moved all old bookdown files to old_book_files

--- 2019-03-25 electoral-modelling-draft.Rmd
@@
---- 
-title: "Spatial modelling of the electoral two-party preferred vote in Australia. A study of federal elections between 2001 and 2016 via the `eechidna` `R` package."
-author: "Jeremy Forbes, Di Cook & Rob Hyndman"
-date: "`r Sys.Date()`"
-site: bookdown::bookdown_site
-output: bookdown::gitbook
-documentclass: book
-classoption: openany
-bibliography: [book.bib, packages.bib]
-biblio-style: apalike
-link-citations: yes
-github-repo: rstudio/bookdown-demo
-description: "This is an incomplete draft of a paper to submitted to ANZJS."
----
+# Data collection, wrangling and imputation {#data}
 
-<!--chapter:end:index.Rmd-->
-
-# Introduction {#intro}
-
-```{r setup, include=FALSE, message=FALSE}
-knitr::opts_chunk$set(echo = F, fig.pos = 'h', fig.align = 'center', cache = T, message = F)
-library(tidyverse)
-library(eechidna)
-library(ggplot2)
-library(ggthemes)
-library(knitr)
-library(nlme)
-library(rgeos)
-library(spdep)
-
-# Data
-data(tpp01)
-data(tpp04)
-data(tpp07)
-data(tpp10)
-data(tpp13)
-data(tpp16)
-data(abs2001)
-data(abs2004)
-data(abs2007)
-data(abs2010)
-data(abs2013)
-data(abs2016)
-```
-
-```{r wrangle, include = F, message = F}
-# Function to standardize variables
-standardise_vars <- function(df) {
-  hold <- df
-  num_cols <- sapply(df, class) == 'numeric'
-  df[, num_cols] <- lapply(df[, num_cols], scale)
-  names(df) <- names(hold)
-  df$LNP_Percent <- hold$LNP_Percent
-  return(df)
-}
-
-# Combine and standardize
-my_df <- bind_rows(
-  left_join(tpp01, standardise_vars(abs2001) %>% dplyr::select(-c(UniqueID, Area, ends_with("NS"), Population)), by = c("DivisionNm", "StateAb"="State")) %>% mutate(year = "2001"),
-  left_join(tpp04, standardise_vars(abs2004) %>% dplyr::select(-UniqueID), by = c("DivisionNm")) %>% mutate(year = "2004"),
-  left_join(tpp07, standardise_vars(abs2007) %>% dplyr::select(-UniqueID), by = c("DivisionNm")) %>% mutate(year = "2007"),
-  left_join(tpp10, standardise_vars(abs2010) %>% dplyr::select(-UniqueID), by = c("DivisionNm")) %>% mutate(year = "2010"),
-  left_join(tpp13, standardise_vars(abs2013) %>% dplyr::select(-UniqueID), by = c("DivisionNm")) %>% mutate(year = "2013"),
-  left_join(tpp16, standardise_vars(abs2016) %>% dplyr::select(-c(UniqueID, Area, ends_with("NS"), Population)), by = c("DivisionNm", "StateAb"="State")) %>% mutate(year = "2016")
-) %>% 
-  mutate(year = factor(year)) %>% 
-  dplyr::select(-c(starts_with("Age"), StateAb, LNP_Votes, ALP_Votes, ALP_Percent, TotalVotes,
-    Swing, InternetUse, InternetAccess, EnglishOnly, Other_NonChrist, OtherChrist, Volunteer, EmuneratedElsewhere, UniqueID, Catholic, Anglican))
-
-# Create final df for modelling
-
-factors_df <- my_df %>% 
-  mutate(Education = BachelorAbv + HighSchool + Professional + Finance - Laborer - Tradesperson - DipCert,
-    FamHouseSize = FamilyRatio + AverageHouseholdSize + Couple_WChild_House - Couple_NoChild_House -
-      SP_House,
-    PropertyOwned = Owned + Mortgage - Renting - PublicHousing,
-    RentLoanPrice = MedianRent + MedianLoanPay,
-    Incomes = MedianFamilyIncome + MedianHouseholdIncome + MedianPersonalIncome,
-    Unemployment = Unemployed - LFParticipation) %>% 
-  dplyr::select(-c(BachelorAbv, HighSchool, Professional, Finance, Laborer, Tradesperson, DipCert, FamilyRatio,
-    AverageHouseholdSize, Couple_WChild_House, Couple_NoChild_House, SP_House, Owned, Mortgage, Renting,
-    PublicHousing, MedianFamilyIncome, MedianHouseholdIncome, MedianPersonalIncome, MedianRent, 
-    MedianLoanPay, Unemployed, LFParticipation))
-
-# Now standardize factors
-
-small_df <- bind_rows(
-  factors_df %>% filter(year == "2001") %>% standardise_vars(),
-  factors_df %>% filter(year == "2004") %>% standardise_vars(),
-  factors_df %>% filter(year == "2007") %>% standardise_vars(),
-  factors_df %>% filter(year == "2010") %>% standardise_vars(),
-  factors_df %>% filter(year == "2013") %>% standardise_vars(),
-  factors_df %>% filter(year == "2016") %>% standardise_vars()
-)
-
-# Order electorates in alphabetical order to match spatial matrix
-model_df <- small_df %>% 
-  arrange(year, DivisionNm) %>% 
-  dplyr::select(order(colnames(.)))
-```
-
-```{r spatialweights, eval = F}
-# Compute spatial weights matrix
-sp_weights_matrix <- function(sF) {
-
-dist_matrix <- function(shapefile) {
-  dist_mat <- matrix(NA, nrow = 150, ncol = 150)
-  rownames(dist_mat) <- sort(shapefile$elect_div)
-  colnames(dist_mat) <- sort(shapefile$elect_div)
-
-  for (i in 1:(nrow(dist_mat)-1)) {
-    rname = rownames(dist_mat)[i]
-    row_poly = shapefile %>% subset(elect_div == rname)
-
-    for (j in (i+1):ncol(dist_mat)) {
-
-      cname = rownames(dist_mat)[j]
-      col_poly = shapefile %>% subset(elect_div == cname)
-      dist = gDistance(row_poly, col_poly)
-      dist_mat[i,j] = dist
-
-    }
-    print(i)
-  }
-
-  # Now copy to lower triange
-  for (i in 2:nrow(dist_mat)) {
-    for (j in 1:(i-1)) {
-      dist_mat[i,j] = dist_mat[j,i]
-    }
-  }
-
-  # Check it is symmetric
-  if(!isSymmetric(dist_mat)) {
-    print("Warning! Matrix is not symmetric. Error has occured.")
-  }
-
-  return(dist_mat)
-}
-
-my_dist_mat <- dist_matrix(sF)
-
-# S matrix
-
-s_mat <- function(dist_mat) {
-  s_mat <- dist_mat
-
-  for (i in 1:nrow(dist_mat)) {
-    for (j in 1:nrow(dist_mat)) {
-      a = dist_mat[i,j]
-
-      if (is.na(a)) {
-        b = 0
-      } else {
-        b = ifelse(a == 0, 1, 0)
-      }
-
-      s_mat[i,j] = b
-
-    }
-  }
-
-  return(s_mat)
-}
-
-my_smat <- s_mat(my_dist_mat)
-
-# Turn into W matrix
-
-w_mat <- function(s_mat) {
-
-  w_mat <- s_mat
-  rsums = rowSums(s_mat)
-
-  for (i in 1:nrow(s_mat)) {
-    w_mat[i,] <- s_mat[i,]/rsums[i]
-  }
-
-  return(w_mat)
-}
-
-my_wmat <- w_mat(my_smat)
-
-# Turn into listw
-
-my_listw <- mat2listw(my_wmat)
-
-return(my_listw)
-
-}
-
-# Get spatial weights
-sF_16 <- sF_download(2016)
-sF_13 <- sF_download(2013)
-sF_10 <- sF_download(2010)
-sF_07 <- sF_download(2007)
-sF_04 <- sF_download(2004)
-sF_01 <- sF_download(2001)
-
-sp_weights_16 <- sp_weights_matrix(sF_16)
-sp_weights_13 <- sp_weights_matrix(sF_13)
-sp_weights_10 <- sp_weights_matrix(sF_10)
-sp_weights_07 <- sp_weights_matrix(sF_07)
-sp_weights_04 <- sp_weights_matrix(sF_04)
-sp_weights_01 <- sp_weights_matrix(sF_01)
-```
-
-```{r FGLS, include = F}
-# Instead of running spatial weights
-load("/Users/Jeremy/Documents/R/eechidna-modelling/data/sp_weights_01.rda")
-load("/Users/Jeremy/Documents/R/eechidna-modelling/data/sp_weights_04.rda")
-load("/Users/Jeremy/Documents/R/eechidna-modelling/data/sp_weights_07.rda")
-load("/Users/Jeremy/Documents/R/eechidna-modelling/data/sp_weights_10.rda")
-load("/Users/Jeremy/Documents/R/eechidna-modelling/data/sp_weights_13.rda")
-load("/Users/Jeremy/Documents/R/eechidna-modelling/data/sp_weights_16.rda")
-
-# Function for FGLS
-my_fgls <- function(my_formula, my_data, sp_weights) {
-  # Spatial weights matrix
-  w_mat <- listw2mat(sp_weights)
-
-  # Get OLS residuals
-  ols_model <- lm(my_formula, my_data)
-  my_res <- ols_model$residuals
-
-  # Solve for rho
-  res_model <- lm(my_res ~ w_mat%*%my_res)
-  rho <- res_model$coefficients[2]
-  rho_df <- data.frame(estimate = rho, se = summary(res_model)$coefficients[2,2], p = summary(res_model)$coefficients[2,4])
-
-  # Transform data for GLS
-  trans_mat <- diag(nrow(my_data)) - rho*w_mat
-  gls_data <- data.frame(LNP_Percent = trans_mat %*% my_data$LNP_Percent,
-    Intercept = trans_mat %*% rep(1,nrow(my_data))) %>% 
-    bind_cols(as.data.frame(trans_mat %*% as.matrix(my_data %>% select(-LNP_Percent))))
-
-  # GLS model
-  my_formula <- formula(paste0(my_formula,  " - 1"))
-  gls_model <- gls(my_formula, gls_data)
-
-  # Call to function with stargazer
-  gls_model$call$model <- formula(paste0("LNP_Percent ~ ", paste0(names(gls_data)[-1], collapse = " + ")))
-
-  # Rho and data
-  gls_model$rho_df <- rho_df
-  gls_model$gls_data <- gls_data
-  gls_model$my_data <- my_data
-
-  return(gls_model)
-}
-
-## Run full models for each year
-full_formula = "LNP_Percent ~ ."
-
-# 2016
-glsmod16 <- my_fgls(full_formula, 
-  my_data = model_df %>% filter(year == "2016") %>% select(-c(DivisionNm, year)),
-  sp_weights = sp_weights_16)
-
-# 2013
-glsmod13 <- my_fgls(full_formula, 
-  my_data = model_df %>% filter(year == "2013") %>% select(-c(DivisionNm, year)),
-  sp_weights = sp_weights_13)
-
-# 2010
-glsmod10 <- my_fgls(full_formula, 
-  my_data = model_df %>% filter(year == "2010") %>% select(-c(DivisionNm, year)),
-  sp_weights = sp_weights_10)
-
-# 2007
-glsmod07 <- my_fgls(full_formula, 
-  my_data = model_df %>% filter(year == "2007") %>% select(-c(DivisionNm, year)),
-  sp_weights = sp_weights_07)
-
-# 2004
-glsmod04 <- my_fgls(full_formula, 
-  my_data = model_df %>% filter(year == "2004") %>% select(-c(DivisionNm, year)),
-  sp_weights = sp_weights_04)
-
-# 2001
-glsmod01 <- my_fgls(full_formula, 
-  my_data = model_df %>% filter(year == "2001") %>% select(-c(DivisionNm, year)),
-  sp_weights = sp_weights_01)
-
-## Visualise coefficients and significance
-
-coef_df <- bind_rows(
-  data.frame(variable = glsmod16$coefficients %>% names, estimate = glsmod16$coefficients %>% unname, se = summary(glsmod16)$tTable[, "Std.Error"] %>% unname, p = summary(glsmod16)$tTable[, "p-value"] %>% unname, year = 2016),
-  data.frame(variable = glsmod13$coefficients %>% names, estimate = glsmod13$coefficients %>% unname, se = summary(glsmod13)$tTable[, "Std.Error"] %>% unname, p = summary(glsmod13)$tTable[, "p-value"] %>% unname, year = 2013),
-  data.frame(variable = glsmod10$coefficients %>% names, estimate = glsmod10$coefficients %>% unname, se = summary(glsmod10)$tTable[, "Std.Error"] %>% unname, p = summary(glsmod10)$tTable[, "p-value"] %>% unname, year = 2010),
-  data.frame(variable = glsmod07$coefficients %>% names, estimate = glsmod07$coefficients %>% unname, se = summary(glsmod07)$tTable[, "Std.Error"] %>% unname, p = summary(glsmod07)$tTable[, "p-value"] %>% unname, year = 2007),
-  data.frame(variable = glsmod04$coefficients %>% names, estimate = glsmod04$coefficients %>% unname, se = summary(glsmod04)$tTable[, "Std.Error"] %>% unname, p = summary(glsmod04)$tTable[, "p-value"] %>% unname, year = 2004),
-  data.frame(variable = glsmod01$coefficients %>% names, estimate = glsmod01$coefficients %>% unname, se = summary(glsmod01)$tTable[, "Std.Error"] %>% unname, p = summary(glsmod01)$tTable[, "p-value"] %>% unname, year = 2001)
-)
-```
-
-Australia has changed in many ways over the last two decades.
-Rising house prices, country-wide improvements in education, an ageing population, and a decline in religious affiliation, are just a few facets of the country's evolving socio-demographic characteristics.
-At the same time, political power has moved back and forth between the two major parties.
-In the 2007 and 2010 federal elections, the Australian Labor Party (Labor) was victorious, whereas the 2001, 2004, 2013 and 2016 elections were won by the Liberal National coalition (Liberal).
-The two-party preferred vote, a measure of support between these two parties, fluctuated between 47.3% and 53.5% (in favour of the Liberal party) over this period.
-This study explores how electoral characteristics relate to two-party preference, and whether their effects have changed over time.
-Electoral socio-demographics are derived from the Census, and vote counts are obtained from federal elections.
-
-Joining these two data sources is problematic as there in an inherent asynchronicity in the two events.
-A Census is conducted by the Australian Bureau of Statistics (ABS) every five years, whereas a federal election (conducted by the Australian Electoral Commission (AEC)) usually occurs every three years.
-The first problem addressed is that of obtaining appropriate Census data for the 2004, 2007, 2010 and 2013 elections - election years in which a Census does not occur.
-The predominant approach in previous studies is to join voting outcomes to the nearest Census, without accounting for any temporal differences (see @DavisStimson98, @Stimson06, @Liao09 and @Stimson09).
-Furthermore, electoral boundaries change regularly, so spatial discrepancies also arise when matching electoral data.
-To obtain appropriate Census data for these four elections, electoral socio-demographics are imputed using a spatio-temporal imputation that combines areal interpolation [@Goodchild1993] and linear time-interpolation.
-The process of collecting and wrangling the raw data is outlined in section \@ref(data), and the imputation process is detailed in section \@ref(imputation).
-All data and associated documentation relating to this procedure are available in the `eechidna` `R` package [@eechidna], providing a resource for future analysis.
-
-Previous work on modelling Australian federal elections have found that aggregate socio-demographics are relatively good predictors of voting outcomes.
-@Forrest01 does this using multiple regression of the Liberal and Labor primary vote for polling booths in the Farrer electorate in 1998.
-@Stimson06, @Stimson09 and @Stimson12 use principal component analysis of polling booths in the 2001, 2004 and 2007 elections respectively, also finding that socio-demographic characteristics of polling booths are linked to their two-party preferred vote.
-On the contrary, @Stimson09 models the polling booth swing vote (change in the two-party preferred vote) in the 2007 election, finding that little of swing vote can be explained by Census data.
-Instead of analyzing a single election in isolation, this paper employs a consistent model framework across six elections so that temporal changes in the effects of socio-demographics can be observed, where each federal elections is modelled with a cross-sectional data set.
-The use of a regression framework to examine these socio-political relationships over time is seemingly absent from previous Australian studies.
-It also appears that no study has attempted any type of statistical analysis of socio-demographics in conjunction with voter behaviour in Australia since 2007, making this paper distinctly different from those previous.
-
-The cross-sectional data set for each election consists of the two-party preferred vote (response variable), and socio-demographic variables (explanatory variables) that characterise each electorate.
-To obtain these cross-sections, socio-demographic variables are first standardized, and then principal components are used to group variables into "factors".
-To account for the inherent spatial structure of the data, a spatial error model is fit for each election.
-These steps are discussed in section \@ref(modelling).
-In section \@ref(results) inference is conducted on the models to see which effects are significant, how effects change over time and which electorates have abnormal voting behaviour.
-
-<!--chapter:end:01-intro.Rmd-->
-
-# Data collection {#data}
-
-## Voting data
+## Collecting the data
 
@@
 
-## Census data
-
 Socio-demographic variables are derived from the Census of Population and Housing (Census), which is a survey of every household in Australia, recording information such as age, gender, ethnicity, education level and income.
@@
 From this aggregate information, 67 socio-demographic variables are computed for each of the electorates.
-
-## Downloading and wrangling
 
@@
 
-<!--chapter:end:02-data-collection.Rmd-->
+## Joining Census and election data
 
-# Joining Census and election data {#imputation}
-
-## Differences between Census and election data
+### Differences between Census and election data
 
@@
 
-## Spatio-temporal imputation
+### Spatio-temporal imputation
 
@@
 
-## Applied
+### Applied
 
@@
 This is done for each of the socio-demographic variables, and repeated each of the 2013 electorates.
-
-<!--chapter:end:03-imputation.Rmd-->
-
----
-output:
-  pdf_document: default
-  html_document: default
----
-# Modelling {#modelling}
-
-## Data pre-processing
-With socio-demographic information now available for each electorate, each election is joined to the data  corresponding with its two-party preferred vote.
-Socio-demographic variables within each election year are standardized to have mean zero and variance one, to adjust for changing variable scales.
-For example, inflation-adjusted median rental prices increased across almost all electorates, with median rent of 200 dollars per week placing an electorate in the 90th percentile in 2001, but only the 30th percentile in 2016.
-
-### Dimension reduction
-With only $N = 150$ observations (electorates) in each election and $p = 65$ socio-demographic variables in each cross-section, any model using all variables would face serious problems with multi-collinearity and over-fitting, likely leading to erroneous conclusions regarding variable significance.
-Therefore a form of dimension reduction is adopted before models are fit.
-
-Socio-demographic variables[^2] that represent similar information are combined into "factors" using principal component analysis (PCA).
-The scree plots of the principal components for each election all level off after four components, and the loadings of these four components are similar across the elections.
-Principal components are then computed on the combined set of socio-demographics across all six elections.
-A factor is created by combining several variables all have large loadings in a particular component and when there is an intuitive reason as to why these variables could represent common information.
-A loading with magnitude greater than 0.15 is considered large.
-After computing these sums, each factor is again standardized to have mean zero and variance one, within each election.
-
-Consider the `Incomes` factor as an illustration.
-Independent of principal components, we may suspect that median personal income, median household income and median family income are providing similar information about the financial wellbeing of an electorate.
-Their loadings in the first principal component are large (0.20, 0.21 and 0.22 respectively), which provides the evidence needed to combine these variables into a single factor, which is called `Incomes`.
-
-This process reduces the predictor set to $p = 30$.
-
-[^2]: A preliminary step involved removing all age bands, because age is represented by median age, and to remove variables relating to particular denominations of Christianity.
-
-## Model framework
-An identical model specification is used across the six elections, with each election modelled separately.
-This allows for the socio-demographic effects to be estimated separately for each year, allowing for interpretation of temporal changes in these effects.
-This is preferable over a single longitudinal model because it avoids any concerns of undue bias stemming from an incorrectly imposed time-varying restriction on any variable.
-Without such restrictions, a pooled cross-sectional model does not yield any distinct advantage over separate cross-sections.
-The panel approach is avoided because of how frequently electoral boundaries change, meaning that electorates that have the same name across elections are not guaranteed to represent the same geographical region.
-Therefore any fixed or random effects models would be difficult to estimate without implementing consistent boundaries, which would requiring further imputation.
-
-For each cross-section, let the response variable be the two-party preferred vote in favour of the Liberal party, denoted $Y$, with $Y = 70$ representing a 70% preference for Liberal, 30% for Labor.
-Although $Y$ lies in the interval $(0,100)$, observed values are never very close to 0 or 100 (minimum $24.05 \%$ and maximum $74.90 \%$), so there is no need to impose the constraint of $Y \in [0,100]$.
-Furthermore, the response is found to be spatially correlated in each election (Moran's I test, $p \le 7\cdot10^{-15}$).
-This is expected, as electorates are aggregate spatial units, and hence the spatial structure of the data must modelled appropriately.
-
-The spatial error model [@Anselin88] is chosen because captures spatial heterogeneity by incorporating a spatially structured random effect vector [@LeSage2009].
-In this context, the random effect can be thought of as capturing the unobserved political climate in each electorate, where the climate is correlated with the climate in neighbouring electorates.
-This functions under the assumption that the climate is independent of electoral socio-demographics, and that an electorate is equally correlated with any electorate that shares a part of its boundary.
-Spatial weights are calculated in accordance with these assumptions.
-The spatial error model is specified as follows:
-
-Let $\rho$ be spatial autoregressive coefficient, $\boldsymbol v$ be a spherical error term, ${\boldsymbol W}$ be a matrix of spatial weights (containing information about the neighbouring regions), $\boldsymbol X$ be a matrix of socio-demographic covariates, $\boldsymbol \beta$ be a vector of regression coefficients and $\boldsymbol a$ be a spatially structured random effect vector.
-
-$${\boldsymbol y} = {\boldsymbol X} {\boldsymbol \beta} + {\boldsymbol a}$$
-and
-
-$${\boldsymbol a} = \rho {\boldsymbol W} {\boldsymbol a} + {\boldsymbol v}$$
-where
-
-$${\boldsymbol v} \sim N({\boldsymbol 0}, \sigma^2 {\boldsymbol I_n})$$.
-
-so it can be written
-
-$${\boldsymbol y} = {\boldsymbol X} {\boldsymbol \beta} + ({\boldsymbol I}_n-\rho {\boldsymbol W})^{-1}{\boldsymbol v}$$
-
-Estimation is done using feasible generalized least squares.
-
-Table 4.1 details the resultant estimated model coefficients and their estimated standard errors for each of the six elections. These are interpreted in the next section.
-
-```{r printmodels, eval = F}
-# Add rho and make adjustments to this
-stargazer::stargazer(glsmod16, glsmod13, glsmod10, glsmod07, glsmod04, glsmod01, header = FALSE, 
-  column.sep.width = "1pt", 
-  title = "Estimated model for each of the six elections.",
-  dep.var.labels = "Two-party preferred vote in favor of the Liberal party", 
-  font.size = "scriptsize", digits = 2, 
-  column.labels = c("2001", "2004", "2007", "2010", "2013", "2016"),
-  notes = c("Estimated coefficients for variable named in column one shown for", "election year indicated by column heading, with estimated standard", "deviation for each coefficient shown below in parenthesis.
-Overall", "summary measures for each regression equation are provided in the", "bottom panel."),
-  notes.align = "l")
-```
-
-\begin{table}[!htbp] \centering 
-  \caption{Estimated model for each of the six elections.} 
-  \label{} 
-\tiny 
-\begin{tabular}{@{\extracolsep{1pt}}lcccccc} 
-\\[-1.8ex]\hline 
-\hline \\[-1.8ex] 
- & \multicolumn{6}{c}{\textit{Dependent variable:}} \\ 
-\cline{2-7} 
-\\[-1.8ex] & \multicolumn{6}{c}{Two-party preferred vote in favor of the Liberal party} \\ 
- & 2001 & 2004 & 2007 & 2010 & 2013 & 2016 \\ 
-\\[-1.8ex] & (1) & (2) & (3) & (4) & (5) & (6)\\ 
-\hline \\[-1.8ex] 
- $\rho$ & 0.48$^{***}$ & 0.29$^{*}$ & 0.15 & $-$0.04 & 0.15 & 0.44$^{***}$ \\ 
-  & (0.16) & (0.16) & (0.17) & (0.19) & (0.17) & (0.15) \\ 
-  & & & & & & \\ 
- AusCitizen & $-$1.53 & $-$0.87 & 1.00 & 0.67 & $-$2.83 & $-$2.77 \\ 
-  & (2.95) & (2.90) & (2.93) & (2.57) & (2.63) & (2.51) \\ 
-  & & & & & & \\ 
- Born\_Asia & $-$2.13 & $-$2.56 & $-$7.58$^{***}$ & $-$1.37 & 0.74 & 2.63 \\ 
-  & (2.20) & (2.58) & (2.68) & (1.96) & (2.55) & (3.03) \\ 
-  & & & & & & \\ 
- Born\_MidEast & $-$0.82 & 1.70 & 1.79 & 2.92$^{*}$ & 3.22$^{*}$ & 1.92 \\ 
-  & (1.56) & (1.54) & (1.54) & (1.51) & (1.78) & (1.53) \\ 
-  & & & & & & \\ 
- Born\_SE\_Europe & $-$1.63 & $-$1.25 & $-$1.72 & $-$0.20 & 0.04 & 1.02 \\ 
-  & (0.99) & (1.15) & (1.34) & (1.17) & (2.05) & (2.34) \\ 
-  & & & & & & \\ 
- Born\_UK & $-$1.37 & $-$0.35 & $-$0.29 & 0.29 & $-$0.87 & $-$0.49 \\ 
-  & (0.98) & (1.00) & (1.05) & (0.80) & (0.90) & (1.04) \\ 
-  & & & & & & \\ 
- BornElsewhere & 1.74 & 4.04 & 9.78$^{**}$ & 5.01 & $-$0.51 & $-$1.08 \\ 
-  & (4.52) & (4.46) & (4.47) & (3.59) & (3.67) & (3.40) \\ 
-  & & & & & & \\ 
- Buddhism & 1.51 & 1.09 & 2.41$^{*}$ & 0.90 & 2.84$^{*}$ & 1.50 \\ 
-  & (1.24) & (1.25) & (1.40) & (1.25) & (1.61) & (1.94) \\ 
-  & & & & & & \\ 
- Christianity & 5.17$^{**}$ & 3.08 & $-$0.60 & $-$1.93 & 3.44 & 0.53 \\ 
-  & (2.52) & (2.79) & (3.44) & (3.45) & (4.19) & (4.58) \\ 
-  & & & & & & \\ 
- CurrentlyStudying & $-$0.09 & 1.73 & 3.28$^{***}$ & 4.03$^{***}$ & 1.69 & $-$1.14 \\ 
-  & (1.13) & (1.17) & (1.21) & (1.12) & (1.09) & (1.01) \\ 
-  & & & & & & \\ 
- DeFacto & $-$11.02$^{***}$ & $-$9.73$^{***}$ & $-$8.60$^{***}$ & $-$6.55$^{***}$ & $-$5.04$^{**}$ & $-$6.42$^{***}$ \\ 
-  & (2.55) & (2.62) & (2.57) & (1.90) & (2.00) & (1.69) \\ 
-  & & & & & & \\ 
- DiffAddress & 6.93$^{***}$ & 5.02$^{***}$ & 4.72$^{***}$ & 3.40$^{***}$ & 4.63$^{***}$ & 3.86$^{***}$ \\ 
-  & (1.50) & (1.71) & (1.60) & (0.88) & (1.03) & (0.95) \\ 
-  & & & & & & \\ 
- Distributive & 1.65 & 2.76$^{**}$ & 2.88$^{**}$ & 2.25$^{**}$ & 2.75$^{**}$ & 1.67 \\ 
-  & (1.17) & (1.25) & (1.30) & (1.02) & (1.13) & (1.08) \\ 
-  & & & & & & \\ 
- Education & $-$3.04 & $-$0.90 & $-$1.82 & $-$0.77 & 1.94 & 0.62 \\ 
-  & (2.99) & (3.04) & (3.35) & (2.89) & (2.86) & (2.23) \\ 
-  & & & & & & \\ 
- Extractive & 6.81$^{***}$ & 6.09$^{***}$ & 6.97$^{***}$ & 5.81$^{***}$ & 5.31$^{***}$ & 4.69$^{***}$ \\ 
-  & (1.31) & (1.42) & (1.48) & (1.16) & (1.23) & (1.36) \\ 
-  & & & & & & \\ 
- FamHouseSize & $-$4.35 & $-$6.16$^{*}$ & $-$8.22$^{**}$ & $-$9.67$^{***}$ & $-$5.11$^{*}$ & $-$2.94 \\ 
-  & (2.86) & (3.15) & (3.28) & (2.66) & (2.78) & (2.27) \\ 
-  & & & & & & \\ 
- Incomes & 8.46$^{***}$ & 4.19 & 2.73 & 4.36$^{*}$ & 2.95 & 2.99$^{*}$ \\ 
-  & (2.57) & (2.68) & (3.06) & (2.56) & (2.60) & (1.76) \\ 
-  & & & & & & \\ 
- Indigenous & 4.47$^{***}$ & 6.12$^{***}$ & 5.90$^{***}$ & 6.81$^{***}$ & 5.62$^{***}$ & 3.97$^{***}$ \\ 
-  & (1.38) & (1.49) & (1.49) & (1.17) & (1.35) & (1.26) \\ 
-  & & & & & & \\ 
- Islam & 2.64 & 1.20 & $-$1.04 & $-$1.21 & 0.61 & 0.15 \\ 
-  & (1.72) & (1.71) & (1.94) & (1.80) & (2.00) & (2.03) \\ 
-  & & & & & & \\ 
- Judaism & 1.96$^{***}$ & 1.60$^{*}$ & 0.46 & 0.01 & 1.74 & 1.02 \\ 
-  & (0.71) & (0.85) & (1.05) & (1.00) & (1.26) & (1.38) \\ 
-  & & & & & & \\ 
- ManagerAdminClericalSales & 4.62$^{***}$ & 3.99$^{***}$ & 4.32$^{***}$ & 5.21$^{***}$ & 3.28$^{***}$ & 2.10$^{***}$ \\ 
-  & (1.03) & (0.99) & (1.00) & (0.78) & (0.85) & (0.70) \\ 
-  & & & & & & \\ 
- Married & $-$1.39 & 1.46 & 5.07$^{*}$ & 5.06$^{*}$ & 5.50$^{**}$ & 2.27 \\ 
-  & (2.64) & (2.71) & (2.91) & (2.59) & (2.79) & (2.26) \\ 
-  & & & & & & \\ 
- MedianAge & 2.94$^{*}$ & 2.28 & 2.08 & 1.86 & 3.73$^{**}$ & 1.54 \\ 
-  & (1.63) & (1.88) & (2.08) & (1.65) & (1.55) & (1.30) \\ 
-  & & & & & & \\ 
- NoReligion & 5.76$^{**}$ & 3.53 & 0.25 & $-$0.31 & 3.43 & 1.06 \\ 
-  & (2.59) & (2.58) & (2.98) & (2.90) & (3.45) & (3.85) \\ 
-  & & & & & & \\ 
- OneParent\_House & $-$0.85 & 0.11 & 0.22 & 1.66 & 1.37 & $-$1.00 \\ 
-  & (1.39) & (1.46) & (1.61) & (1.38) & (1.49) & (1.34) \\ 
-  & & & & & & \\ 
- OtherLanguageHome & 1.27 & $-$4.69 & $-$4.37 & $-$6.02 & $-$6.76 & $-$10.40$^{**}$ \\ 
-  & (4.95) & (5.50) & (5.70) & (4.44) & (5.61) & (4.97) \\ 
-  & & & & & & \\ 
- PropertyOwned & 1.76 & 0.68 & $-$1.17 & 0.78 & $-$0.92 & $-$0.76 \\ 
-  & (1.38) & (1.43) & (1.49) & (1.09) & (1.21) & (1.19) \\ 
-  & & & & & & \\ 
- RentLoanPrice & $-$0.70 & 1.01 & 2.12 & $-$0.54 & $-$1.54 & $-$0.31 \\ 
-  & (2.02) & (1.96) & (1.96) & (1.48) & (1.68) & (1.50) \\ 
-  & & & & & & \\ 
- SocialServ & 2.90$^{***}$ & 0.94 & 1.91 & 1.77 & 1.89 & 2.42$^{*}$ \\ 
-  & (1.10) & (1.22) & (1.35) & (1.14) & (1.27) & (1.27) \\ 
-  & & & & & & \\ 
- Transformative & 4.75$^{***}$ & 3.74$^{**}$ & 3.28$^{*}$ & 5.20$^{***}$ & 5.74$^{***}$ & 2.99$^{**}$ \\ 
-  & (1.41) & (1.57) & (1.74) & (1.47) & (1.59) & (1.41) \\ 
-  & & & & & & \\ 
- Unemployment & 0.36 & $-$1.22 & $-$2.42 & $-$2.88$^{**}$ & $-$5.09$^{***}$ & $-$3.00$^{**}$ \\ 
-  & (1.42) & (1.54) & (1.73) & (1.41) & (1.51) & (1.28) \\ 
-  & & & & & & \\ 
- Constant & 50.45$^{***}$ & 53.48$^{***}$ & 49.92$^{***}$ & 47.32$^{***}$ & 52.63$^{***}$ & 50.78$^{***}$ \\ 
-  & (0.74) & (0.56) & (0.51) & (0.35) & (0.45) & (0.67) \\ 
-  & & & & & & \\ 
-\hline \\[-1.8ex] 
-Observations & 150 & 150 & 150 & 150 & 150 & 150 \\ 
-Log Likelihood & $-$399.52 & $-$405.41 & $-$418.08 & $-$398.73 & $-$403.34 & $-$398.33 \\ 
-Akaike Inf. Crit. & 863.03 & 874.83 & 900.16 & 861.46 & 870.68 & 860.67 \\ 
-Bayesian Inf. Crit. & 951.96 & 963.76 & 989.10 & 950.40 & 959.61 & 949.60 \\ 
-\hline 
-\hline \\[-1.8ex] 
-\textit{Note:}  & \multicolumn{6}{l}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
- & \multicolumn{6}{l}{Estimated coefficients for variable named in column one shown for} \\ 
- & \multicolumn{6}{l}{election year indicated by column heading, with estimated standard} \\ 
- & \multicolumn{6}{l}{deviation for each coefficient shown below in parenthesis. Overall} \\ 
- & \multicolumn{6}{l}{summary measures for each regression equation are provided in the} \\ 
- & \multicolumn{6}{l}{bottom panel.} \\ 
-\end{tabular} 
-\end{table}
-
-<!--chapter:end:04-modelling.Rmd-->
-
-# Results
-```{r rhovis}
-rho_df <- bind_rows(glsmod16$rho_df %>% mutate(year = "2016"), 
-  glsmod13$rho_df %>% mutate(year = "2013"),
-  glsmod10$rho_df %>% mutate(year = "2010"), 
-  glsmod07$rho_df %>% mutate(year = "2007"),
-  glsmod04$rho_df %>% mutate(year = "2004"), 
-  glsmod01$rho_df %>% mutate(year = "2001")) %>% 
-  mutate(upper95 = estimate + 1.96*se, lower95 = estimate - 1.96*se)
-
-rho_df %>% ggplot() + 
-  geom_point(aes(x = year, y = estimate, col = factor(p < 0.05)), size = 3) +
-  geom_linerange(aes(x = year, ymin = lower95, ymax = upper95, col = factor(p < 0.05)), size = 1.5) + 
-  geom_hline(aes(yintercept = 0), alpha = 0.5, size = 1) +
-  scale_color_manual(values = c("grey50", "black")) +
-  guides(col = F)
-```
-
-The spatial autoregressive coefficient $\rho$ is significant in only the 2001 and 2016 elections, meaning that in these years an electorate's political climate was affected by the attitudes of it's neighbours.
-Conversely, in the other four elections, it appears that after accounting for electoral socio-demographics, there is no spatial effect, meaning electorates effectively acted independently.
-
-To investigate the socio-demographics that have a strong effect on the two-party preferred vote, partial residual plots are used.
-These show the direction, size and significance of an estimated effect, as the slope of the prediction line matches the estimated coefficient, and the shaded region represents a 95% confidence band.
-Typically the Labor party campaigns on more progressive policies, which often include tax reform that adversely affects higher income earners, and more generous social assistance programs.
-Although topical in the lead up to an election, incomes only significantly affected electoral two-party preference in 2016 (see Figure \@ref(fig:plotincomes)), with higher income areas associated with support for the Liberal party.
-In the prior elections, the estimated effect is still positive, albeit insignificant.
-Unemployment on the other hand, was influential in 2001, 2004 and 2007 elections(Figure \@ref(fig:plotunemploy)), with electorates with higher unemployment favouring Labor.
-This effect has become weaker over time, with unemployment having an estimated effect of virtually zero in 2016.
-
-```{r myvisreg, include = F}
-# Function to produce visreg style conditional plots
-my_visreg <- function(my_model, sp_weights, varname, 
-  nolabs = FALSE, xlimits = NULL, ylimits = NULL, year = "") {
-
-  # Extract fitted parameters
-  rho <- my_model$rho_df$estimate
-  sigma <- sqrt(sum(my_model$residuals^2)/(my_model$dims$N-my_model$dims$p))
-
-  # Spatial weights
-  w_mat <- listw2mat(sp_weights)
-
-  # Q - where u = Qe, Q = (I - pW)^-1
-  q_mat <- solve(diag(my_model$dims$N) - rho*w_mat)
-
-  # Omega - QQ'
-  omega_mat <- q_mat%*%t(q_mat)
-
-  # X
-  x_mat <- my_model$my_data %>% 
-    dplyr::select(-LNP_Percent) %>% 
-    mutate(Intercept = 1) %>% 
-    dplyr::select(Intercept, everything()) %>% 
-    as.matrix()
-
-  # Beta
-  beta_mat <- my_model$coefficients
-
-  # T value
-  t = qt(0.975, nrow(my_model$gls_data)-ncol(my_model$gls_data))
-
-  # Lambda matrix (FGLS)
-  x <- round(seq(min(x_mat[, varname]), max(x_mat[, varname]), 0.025), 3)
-  lambda_mat <- data.frame(matrix(0, nrow = length(x), ncol = ncol(x_mat)))
-  names(lambda_mat) <- dimnames(x_mat)[[2]]
-  lambda_mat[, varname] <- x
-  lambda_mat$Intercept <- 1
-  lambda_mat <- as.matrix(lambda_mat)
-
-  # Confidence interval
-  plot_df <- data.frame(variable = x, fitted = lambda_mat%*%beta_mat, variance = 0)
-
-  for (i in 1:nrow(lambda_mat)) {
-    lambda <- lambda_mat[i, ]
-    plot_df$variance[i] = sigma^2 * t(lambda) %*% 
-      solve(t(x_mat) %*% solve(omega_mat) %*% x_mat) %*% 
-      lambda
-  }
-
-  plot_df <- plot_df %>% 
-    mutate(upper95 = fitted + t*sqrt(variance), lower95 = fitted - t*sqrt(variance))
-
-  # Partial residuals
-  points_df <- data.frame(
-    variable = my_model$my_data[, varname] %>% unname,
-    part_res = (my_model$my_data$LNP_Percent - x_mat%*%my_model$coefficients) + my_model$coefficients[varname]*x_mat[, varname] + my_model$coefficients[1]
-  )
-
-  # Plot
-  myplot <- ggplot(data = plot_df) + 
-    geom_ribbon(aes(x = variable, ymin = lower95, ymax = upper95), fill = "grey80") + 
-    geom_point(aes(x = variable, y = part_res), data = points_df, size = 0.75, col = "grey50") + 
-    geom_line(aes(x = variable, y = fitted), col = "blue", size = 1) +
-    #geom_hline(aes(yintercept = min(upper95)), col = "red") +
-    #geom_hline(aes(yintercept = max(lower95)), col = "blue") +
-    theme_bw() + 
-    labs(x = varname, y = "Response") + 
-    ggtitle(year) +
-    theme(plot.title = element_text(face = "bold", size = 10, hjust = 0.5))
-
-  if (nolabs == TRUE) {
-    myplot <- myplot + labs(x = "", y = "")
-  }
-
-  if (!is.null(xlimits) & !is.null(ylimits)) {
-    myplot <- myplot + coord_cartesian(xlim = xlimits, ylim = ylimits)
-  }
-
-  return(myplot)
-}
-
-# Grid visreg
-library(gridExtra)
-library(grid)
-grid_visreg <- function(varname, xvar_title, xlimits, ylimits) {
-  p16 <- my_visreg(glsmod16, sp_weights_16, varname = varname, nolabs = T, year = "2016", xlimits, ylimits)
-  p13 <- my_visreg(glsmod13, sp_weights_13, varname = varname, nolabs = T, year = "2013", xlimits, ylimits)
-  p10 <- my_visreg(glsmod10, sp_weights_10, varname = varname, nolabs = T, year = "2010", xlimits, ylimits)
-  p07 <- my_visreg(glsmod07, sp_weights_07, varname = varname, nolabs = T, year = "2007", xlimits, ylimits)
-  p04 <- my_visreg(glsmod04, sp_weights_04, varname = varname, nolabs = T, year = "2004", xlimits, ylimits)
-  p01 <- my_visreg(glsmod01, sp_weights_01, varname = varname, nolabs = T, year = "2001", xlimits, ylimits)
-
-  plots <- grid.arrange(p01, p04, p07, p10, p13, p16, nrow = 2, 
-    left = textGrob("Two-party preferred vote (%)", gp = gpar(cex = 0.8), rot = 90), 
-    bottom = textGrob(xvar_title, gp = gpar(cex = 0.8)))
-
-  return(plots)
-}
-
-```
-
-```{r plotincomes, fig.cap = "Effect of Incomes"}
-p <- grid_visreg("Incomes", xvar_title = "Incomes", xlimits = c(-1.5, 3.6), ylimits = c(30, 95))
-```
-
-```{r plotunemploy, fig.cap = "Effect of Unemployment"}
-p <- grid_visreg("Unemployment", xvar_title = "Unemployment", xlimits = c(-2.1, 3.4), ylimits = c(33, 70))
-```
-
-This section summarises the main findings made about the relationships between socio-demographic chracteristics and voting behaviour.
-
-```{r}
-coef_df %>% 
-  filter(variable != "Intercept") %>% 
-  mutate(upper95 = estimate + 1.96*se, lower95 = estimate - 1.96*se) %>% 
-  ggplot() +
-  geom_point(aes(x = year, y = estimate, col = factor(p < 0.05)), size = 2) +
-  geom_linerange(aes(x = year, ymin = lower95, ymax = upper95, col = factor(p < 0.05)), size = 1) + 
-  geom_hline(aes(yintercept = 0), alpha = 0.5, size = 1) + 
-  facet_wrap(~variable, scales = "free", ncol = 5) +
-  scale_color_manual(values = c("grey50", "black")) + 
-  theme(axis.text = element_text(size = 6), strip.text = element_text(size = 6)) +
-  labs(x = "", y = "Year") +
-  scale_x_continuous(breaks = c(2001, 2004, 2007, 2010, 2013, 2016)) +
-  guides(col = F)
-```
-
-<!--chapter:end:05-results.Rmd-->
-
-`r if (knitr:::is_html_output()) '
-# References {-}
-'`
-
-<!--chapter:end:06-references.Rmd-->

### 2019-03-29 old_book_files/04-results.Rmd
moved all old bookdown files to old_book_files

--- 2019-03-29 04-results.Rmd
@@
 Conversely, in the other four elections, the spatial effect weakens to become insignficant.
-In these years, it appears that the spatial component does not explain anything over and above the electoral socio-demographics, meaning electorates effectively acted more independently.
+In these years, it appears that the spatial component does not explain anything over and above the electoral socio-demographics, meaning electorates voted effectively independently.
 
 ```{r rhovis, fig.cap="Estimates of the spatial autoregressive parameter for each of the six elections, with a 95\\% confidence interval.
-Only in 2001 and 2016 is there a significant spatial component", out.height='20%'}
+Only in 2001 and 2016 is there a significant spatial component", out.height='20%', out.extra = ''}
 rho_df <- bind_rows(glsmod16$rho_df %>% mutate(year = "2016"), 
@@
   guides(col = F) +
-  coord_fixed(ratio = 1)
+  coord_fixed(ratio = 1) +
+  theme_bw()
 ```
@@
 
-```{r plotintercept, fig.cap = "Estimated intercept for each election, which represents the two-party preferred vote for an electorate with mean characteristics.", out.height="20%"}
+```{r plotintercept, fig.cap = "Estimated intercept for each election, which represents the two-party preferred vote for an electorate with mean characteristics.", out.height="20%", out.extra = ''}
 intercept_df <- bind_rows(data.frame(summary(glsmod16)$tTable)[1, ]  %>% mutate(year = "2016"), 
@@
   coord_fixed(ratio = 0.2) +
-  lims(y = c(45, 55))
+  lims(y = c(45, 55)) +
+  theme_bw()
 ```
@@
 If a horizontal line can be drawn through the confidence band, then the effect is insignificant.
-Plots for each election are faceted to compare the effects over time.
+Plots for each election are faceted to compare the effects over time in Figure \@ref(fig:partresplot1) and Figure \@ref(fig:partresplot2).
 
-*It is important here to note the ecological fallacy - insights are being drawn at the electorate level, and cannot be inferred for another disaggregate level (e.g. individual voters).*
+It is important here to note the ecological fallacy - insights are being drawn at the electorate level, and cannot be inferred for another disaggregate level (e.g. individual voters).
 
@@
 # Grid visreg
-grid_visreg <- function(varname, plot = TRUE, myscale = "free") {
+grid_visreg <- function(varname, plot = TRUE, myscale = "free", top = FALSE) {
 
@@
         facet_grid(varname ~ year, scales = "free") +
-        theme(strip.text.x = element_blank())
+        theme(plot.margin=unit(c(b = 0.05, l = 0, t = -0.2, r = 0),"cm"))
+
+      if (top == FALSE) {
+        return_object <- return_object + theme(strip.text.x = element_blank()) 
+      }
 
@@
 
-```{r}
-grid_visreg(varname = "Incomes", plot = T)
-```
-
 ### Income and unemployment
@@
 Typically the Labor party campaigns on more progressive policies, which often include tax reform that adversely affects higher income earners, and more generous social assistance programs.
-Perhaps due to these policies, higher income electorates have been more likely to support the Liberal party, as the `Incomes` factor has a positive effect on Liberal preference (see Figure \@ref(fig:plotincomes)).
+Perhaps due to these policies, higher income electorates have been more likely to support the Liberal party, as the `Incomes` factor has a positive effect on Liberal preference (see row 1 in Figure \@ref(fig:partresplot1)).
 This effect is signficant in every election aside from 2004, in which it is only marginally insignificant ($p = 0.0613$).
@@
 ### Industry and type of work
-Electorates with higher proportions of workers in mining, gas, water, agriculture, waste and electricity (grouped as `Extractive` industries) are consistently linked with higher support for the Liberal party, with the mangitude of this effect slightly increasing over the years (see Figure \@ref(fig:plotextractive)).
+Electorates with higher proportions of workers in mining, gas, water, agriculture, waste and electricity (grouped as `Extractive` industries) are consistently linked with higher support for the Liberal party, with the mangitude of this effect slightly increasing over the years (see row 3 in Figure \@ref(fig:partresplot1)).
 This is unsurprising, as the Liberal party has close ties with these traditional energy industries, and typically present policies to reduce taxation on energy production.
-Furthermore, electorates with more workers in construction or manufacturing industries (`Transformative`) are also more likely to support the Liberal party (see Figure \@ref(fig:plottransformative)).
+Furthermore, electorates with more workers in construction or manufacturing industries (`Transformative`) are also more likely to support the Liberal party (see row 4 in Figure \@ref(fig:partresplot1)).
 
@@
 ### Household mobility
-In each of the six elections, electorates with a higher proportion of people that have recently moved house (meaning in last five years) were more likely to support the Liberal party, although this effect was marginally insignificant in 2013.
+In each of the six elections, electorates with a higher proportion of people that have recently moved house (meaning in last five years) were more likely to support the Liberal party, although this effect was marginally insignificant in 2013 (see row 6 in Figure \@ref(fig:partresplot1).
 Having controlled for characteristics of house ownership and rental prices (via the factors `PropertyOwned` and `RentLoan` respectively), this is effect is somewhat surprising.
 
-### Relationships and age
+### Relationships
 De facto relationships, but not marriages, are found to be an important (and significant) predictor of the two-party preferred vote in all six elections, with more de facto relationships associated with higher support for the Labor party.
@@
 
-Older regions are often believed to be more conservative, and it can be seen that electorates with a higher median age have been more likely to support the Liberal party - although this effect is only significant in 2007 and 2010.
+### Age
+Older regions are often believed to be more conservative, and it can be seen that electorates with a higher median age have been more likely to support the Liberal party - although this effect is only significant in 2007 and 2010 (see row 2 in Figure \@ref(fig:partresplot2)).
 
@@
 Since 2007, electorates with higher education levels have been associated with supporting the Labor party, although this effect is only significant in 2016.
-Prior to 2007, education had an almost zero effect.
+Prior to 2007, education had an almost zero effect (see row 3 in \@ref(fig:partresplot1)).
 
@@
 
-#### Individual*
-
-```{r plotincomes, fig.cap = "Effect of Incomes", out.width="100%"}
-p <- grid_visreg("Incomes", xvar_title = "Incomes", plot = T, xlimits = c(-1.5, 3.6), ylimits = c(30, 95), years = FALSE)
+```{r prepplot, include = F}
+p1 <- grid_visreg("Incomes", top = T)
+p2 <- grid_visreg("Unemployment")
+p3 <- grid_visreg("Extractive")
+p4 <- grid_visreg("Transformative")
+p5 <- grid_visreg("ManagerAdminClericalSales")
+p6 <- grid_visreg("DiffAddress")
+p7 <- grid_visreg("DeFacto", top = T)
+p8 <- grid_visreg("MedianAge")
+p9 <- grid_visreg("Education")
+p10 <- grid_visreg("OtherLanguageHome")
+p11 <- grid_visreg("Born_SE_Europe")
+p12 <- grid_visreg("Born_Asia")
+p13 <- grid_visreg("Judaism")
 ```
 
-```{r plotunemploy, fig.cap = "Effect of Unemployment", out.width="100%"}
-p <- grid_visreg("Unemployment", xvar_title = "Unemployment", plot = T, xlimits = c(-2.1, 3.4), ylimits = c(28, 68))
+```{r partresplot1, fig.cap = "Partial residual plots for: income and unemployment, industry and type of work, and household mobility.", fig.height = 12,  fig.width = 9, warning = F}
+grid.arrange(p1, p2, p3, p4, p5, p6, nrow = 6,
+  heights = c(1.2, 1, 1, 1, 1, 1),
+  left = grid.text("Two-party preferred vote (%)", gp = gpar(cex = 0.8), rot = 90))
 ```
 
-```{r plotextractive, fig.cap = "Effect of Extractive", out.width="60%"}
-p <- grid_visreg("Extractive", xvar_title = "Extractive", xlimits = c(-0.7, 3.4), ylimits = c(35, 81), plot = T)
-```
-
-```{r plottransformative, fig.cap = "Effect of Transformative", out.width="60%"}
-p <- grid_visreg("Transformative", xvar_title = "Transformative", xlimits = c(-2.25, 2.9), ylimits = c(34, 75), plot = T)
-```
-
-```{r plotmanager, fig.cap = "Effect of ManagerAdminClericalSales", out.width="60%"}
-p <- grid_visreg("ManagerAdminClericalSales", xvar_title = "ManagerAdminClericalSales", xlimits = c(-3.75, 2.8), ylimits = c(32, 70), plot = T)
-```
-
-```{r plotdiffaddress, fig.cap = "Effect of DiffAddress", out.width="60%"}
-p <- grid_visreg("DiffAddress", xvar_title = "DiffAddress", xlimits = c(-2.1, 3.65), ylimits = c(34, 81), plot = T)
-```
-
-```{r plotdefacto, fig.cap = "Effect of DeFacto", out.width="60%", eval = F}
-p <- grid_visreg("DeFacto", xvar_title = "DeFacto", xlimits = c(-2.2, 5.1), ylimits = c(12, 83), plot = T)
-```
-
-```{r plotage, fig.cap = "Effect of MedianAge", out.width="60%", eval = F}
-p <- grid_visreg("MedianAge", xvar_title = "MedianAge", xlimits = c(-2.5, 2.7), ylimits = c(31, 70), plot = T)
-```
-
-```{r ploteducation, fig.cap = "Effect of Education", out.width="60%", eval = F}
-p <- grid_visreg("Education", xvar_title = "Education", xlimits = c(-2.5, 2.7), ylimits = c(31, 70), plot = T)
-```
-
-```{r plototherlanguagehome, fig.cap = "Effect of OtherLanguageHome", out.width="60%", eval = F}
-p <- grid_visreg("OtherLanguageHome", xvar_title = "OtherLanguageHome", xlimits = c(-1.1, 3.5), ylimits = c(5, 80), plot = T)
-```
-
-```{r plotbornse, fig.cap = "Effect of Born_SE_Europe", out.width="60%", eval = F}
-p <- grid_visreg("Born_SE_Europe", xvar_title = "Born_SE_Europe", xlimits = c(-0.9, 5), ylimits = c(20, 70), plot = T)
-```
-
-```{r plotbornasia, fig.cap = "Effect of Born_Asia", out.width="60%", eval = F}
-p <- grid_visreg("Born_Asia", xvar_title = "Born_Asia", xlimits = c(-0.9, 5), ylimits = c(20, 70), plot = T)
-```
-
-```{r plotjudaism, fig.cap = "Effect of Judaism", out.width="60%", eval = F}
-p <- grid_visreg("Judaism", xvar_title = "Judaism", xlimits = c(-1.5, 3.8), ylimits = c(40, 70), plot = T)
+```{r partresplot2, fig.cap = "Partial residual plots for: relationships and age, education and diversity.", fig.height = 12,  fig.width = 9, warning = F}
+grid.arrange(p7, p8, p9, p10, p11, p12, p13, nrow = 7,
+  heights = c(1.2, 1, 1, 1, 1, 1, 1),
+  left = grid.text("Two-party preferred vote (%)", gp = gpar(cex = 0.8), rot = 90))
 ```
@@
 
-```{r removeolh, eval = F}
+```{r removeunemp, eval = F}
 unemp01 <- my_fgls(full_formula, 
@@
 
-```{r resstate, fig.cap = "Boxplot of residuals by state.", out.width="75%"}
+```{r resstate, fig.cap = "Boxplot of residuals by state with jittered points.", fig.asp=0.6, out.width="80%", out.extra = ''}
 all_resids %>% 
-  ggplot(aes(x = StateAb, y = Residuals, col = StateAb)) + 
+  ggplot(aes(x = StateAb, y = Residuals)) + 
+  geom_hline(aes(yintercept = 0), col = "grey50") +
   geom_boxplot() + 
   geom_jitter(alpha = 0.4) +
-  #geom_jitter(aes(label = DivisionNm, shape = year), alpha = 0.4) +
-  #facet_wrap(~year) + 
-  guides(col = F, shape = F, DivisionNm = F)
+  guides(col = F, shape = F, DivisionNm = F) +
+  theme_bw()
 ```
@@
 
-```{r readcooksd, include = F, echo=FALSE}
+```{r readcooksd, eval = F}
 read_chunk('cooksd.R')
@@
 
-```{r getcooksd}
+```{r getcooksd, eval = F}
 <<GetCookDistance>>
@@
 
-```{r cooksd, include = F}
+```{r cooksd, eval = F}
 cooksd_df <- data.frame(
@@
 
-```{r plotcooksd, fig.cap = "Electorates' Cook's Distance values in each election."}
+```{r plotcooksd, fig.cap = "Electorates' Cook's Distance values in each election.", out.height = "15%", eval = F}
 cooksd_df %>% 
@@
   facet_wrap(~year, nrow = 1) + 
-  coord_fixed(ratio = .5) +
+  #coord_fixed(ratio = .5) +
   labs(x = "Cook's Distance", y = "") + 
@@
 
-The dotplots in Figure \@ref(fig:plotcooksd) show that there are a few outliers in each election, using cook's distance as a measure of influence.
-Based on the distribution of the cook's distance values, a cook's distance greater than $0.1$ is considered to be an outlier.
-The electorate of Sydney (NSW) has a large cook's distance from 2001 to 2013, due to its diverse population (language, birthplace and religion), high number of defacto relationships, high income, high household mobility and small amount of workers in extractive, distributive and transformative jobs.
-It has remained a strong supporter of the Labor party and Liberal vote is severely overpredicted by the model.
+Based on the distribution of the cook's distance values, a cook's distance greater than $0.1$ is considered to be influential and a potential outlier.
+The electorate of Sydney (NSW) has a large cook's distance from 2001 to 2013, due to its diverse population (language, birthplace and religion), high number of defacto relationships, high income, high household mobility and small amount of workers in extractive and transformative jobs.
+It has remained a strong supporter of the Labor party and Liberal vote is severely overpredicted by the model, making it an outlier.
 Nearby in metropolitan NSW, the electorate of Wentworth is found to be an outlier in all but the 2007 election.
@@
 
-Lingiari, an electorate taking up almost all of the Northern Territory, is an outlier in the 2001-2007 elections.
-This is likely due to its large Indigenous population, young age profile and low rates of property ownership.
-Other electorates with large cook's distance are Barton (NSW, 2016) and Leichhardt (QLD) in 2016, and Canberra (ACT) in 2007.
+Lingiari, an electorate taking up almost all of the Northern Territory, is an outlier in the 2001-2007 elections due to its large Indigenous population, young age profile and low rates of property ownership.
+Fowler (NSW) has a diverse population with a high proportion of migrants, many Buddhists and Muslims, and has very strong Labor support, making it influential in 2001, 2004 and 2010.
+Other electorates with large cook's distance are Barton (NSW) and Leichhardt (QLD) in 2016, and Canberra (ACT) in 2007.
 
@@
 
-plot(glsmod10$cooksd, hat(model_df %>% filter(year == "2010") %>% select(-c(year, DivisionNm)) %>% as.matrix))
+plot(glsmod07$cooksd, hat(model_df %>% filter(year == "2007") %>% select(-c(year, DivisionNm)) %>% as.matrix))
 ```
-
-```{r eval = F}
-all_resids %>% 
-  group_by(UniqueID) %>% 
-  mutate(max_res = max(abs(Residuals))) %>% 
-  arrange(-max_res) %>% 
-  filter(max_res > 13.4) %>% 
-  dplyr::select(DivisionNm, LNP_Percent, year, everything())
-
-all_resids %>% 
-  group_by(year) %>% 
-  mutate(yr_max = max(abs(Residuals))) %>% 
-  filter(abs(Residuals) == yr_max) %>% 
-  select(DivisionNm, StateAb, year, Residuals, LNP_Percent)
-
-cooksd_df %>% 
-  mutate(LNP_win = ifelse(LNP_Percent > 50, 1, 0),
-    LNP_predicted = ifelse(LNP_Percent - Residuals > 50, 1, 0),
-    correct = ifelse(LNP_win == LNP_predicted, 1, 0)) %>% 
-  filter(correct == 0) %>% dim
-  group_by(year) %>% 
-  mutate(yr_max = max(abs(Residuals))) %>% 
-  filter(abs(Residuals) == yr_max) %>% 
-  select(DivisionNm, StateAb, year, Residuals, LNP_Percent)
-```
-
-```{r eval = F}
-library(plotly)
-ggplot(all_resids) + geom_jitter(aes(x = year, y = Residuals, label = DivisionNm))
-ggplotly()
-
-all_resids %>% 
-  group_by(UniqueID) %>% 
-  mutate(mean_res = mean(abs(Residuals)), min_res = min(abs(Residuals))) %>%
-  filter(min_res > 5) %>%
-  ungroup() %>% dplyr::select(Residuals, DivisionNm, year, min_res, mean_res, StateAb) %>% View
-  ggplot(aes(x = year, y = Residuals, group = DivisionNm, col = DivisionNm, label = DivisionNm)) + 
-  geom_line() + geom_point() +
-  guides(col = F)
-```

### 2019-03-29 old_book_files/05-conclusion.Rmd
moved all old bookdown files to old_book_files

# Conclusion

Across the past six elections, most of the socio-demographics that drive the electoral two-party preferred vote have remained steady, whilst a few (typically weaker) effects have varied over time. Industry and type of work are particularly influential, with energy-related and manufacturing/construction jobs, as well as administrative roles being strongly linked with the Liberal party in all elections. Incomes have had a similarly consistent effect, with higher income areas supporting Liberal. Higher levels of unemployment have shifted from weak association with Labor to a significant Liberal effect over the years, and higher education levels have been associated with Labor since 2007 (although only significant in 2016). It is also found that electorates with higher household mobility support Liberal, birthplace diversity favours Labor and more de facto relationships align with Labor preference - although marriages, family and household sizes have no material influence. Furthermore, the neighbourhood (spatial) effects were found to be positive in all elections, although only significant in 2001 and 2016, meaning that in the 2004-2013 elections, electorates effectively voted independently.

The findings in this paper complement the existing literature by modelling temporal trends, which as far as the authors are aware, has not been done previously for Australian elections using a regression framework. It is also the first study to model any Australian election since 2010 using Census information. Additionally, a key contribution of this research is the spatio-temporal imputation, which is applied to obtain Census data for the 2004, 2007, 2010 and 2013 elections. All data sets used in this study have been contributed to the `eechidna` `R` package, which provides a rich, accessible data resource for future Australian electoral analysis.

### 2019-03-29 old_book_files/07-software.Rmd
moved all old bookdown files to old_book_files

# Software

All election and Census data sets, along with electoral maps and more, are available in the `eechidna` `R` package, which can be downloaded from CRAN. `eechidna` (Exploring Election and Census Highly Informative Data Nationally for Australia) makes it easy to look at the data from the Australian Federal elections and Censuses that occurred between 2001 and 2016. This study contributed a large revision to the `eechidna` package, which included the addition of election and Census data for 2001-2010, vote outcomes for polling booths and imputed Census data for election years. For more details on using `eechidna`, please see the articles (vignettes) on the github page (https://ropenscilabs.github.io/eechidna/).

The authors would like to give a sincere thanks to Anthony Ebert, Heike Hofmann, Thomas Lumley, Ben Marwick, Carson Sievert, Mingzhu Sun, Dilini Talagala, Nicholas Tierney, Nathaniel Tomasetti, Earo Wang and Fang Zhou, all of whom have contributed to the `eechidna` package.

### 2019-03-29 old_book_files/_book/electoral-modelling-draft.tex
moved all old bookdown files to old_book_files

--- 2019-03-26 _book/electoral-modelling-draft.tex
@@
   \postdate{\par}
-    \date{2019-03-26}
+    \date{2019-03-29}
 
@@
 Independent of principal components, we may suspect that median personal income, median household income and median family income are providing similar information about the financial wellbeing of an electorate.
-Their loadings in the first principal component are large (0.20, 0.21 and 0.22 respectively), which provides the evidence needed to combine these variables into a single factor, which is called \texttt{Incomes}.
+Their loadings in the first principal component are large (0.19, 0.21 and 0.22 respectively), which provides the evidence needed to combine these variables into a single factor, which is called \texttt{Incomes}.
 
@@
 Observations & 150 & 150 & 150 & 150 & 150 & 150 \\ 
-Log Likelihood & $-$398.33 & $-$403.34 & $-$398.73 & $-$418.08 & $-$405.41 & $-$399.52 \\ 
-Akaike Inf. Crit. & 860.67 & 870.68 & 861.46 & 900.16 & 874.83 & 863.03 \\ 
-Bayesian Inf. Crit. & 949.60 & 959.61 & 950.40 & 989.10 & 963.76 & 951.96 \\ 
+Residual Standard Error (GLS) & 4.69 & 5.04 & 4.79 & 5.63 & 5.18 & 4.88 \\ 
+Log Likelihood & $-$402.11 & $-$410.52 & $-$404.94 & $-$423.96 & $-$412.62 & $-$404.34 \\ 
+Akaike Inf. Crit. & 868.21 & 885.05 & 873.88 & 911.92 & 889.24 & 872.68 \\ 
+Bayesian Inf. Crit. & 957.14 & 973.98 & 962.81 & 1,000.85 & 978.17 & 961.62 \\ 
 \hline 
@@
 
-\begin{center}\includegraphics{electoral-modelling-draft_files/figure-latex/rhovis-1} \end{center}
+The spatial autoregressive coefficient \(\rho\) is positive and significant in only the 2001 and 2016 elections (Figure \ref{fig:rhovis}), meaning that in these elections, an electorate's political climate was affected by the attitudes of it's neighbours.
+Conversely, in the other four elections, the spatial effect weakens to become insignficant.
+In these years, it appears that the spatial component does not explain anything over and above the electoral socio-demographics, meaning electorates voted effectively independently.
 
-The spatial autoregressive coefficient \(\rho\) is positive and significant in only the 2001 and 2016 elections, meaning that in these elections, an electorate's political climate was affected by the attitudes of it's neighbours.
-Conversely, in the other four elections, it appears that after accounting for electoral socio-demographics, there is no spatial effect, meaning electorates effectively acted independently.
+\begin{figure}[h]
+
+{\centering \includegraphics[height=0.2\textheight,]{electoral-modelling-draft_files/figure-latex/rhovis-1}
+
+}
+
+\caption{Estimates of the spatial autoregressive parameter for each of the six elections, with a 95\% confidence interval.
+Only in 2001 and 2016 is there a significant spatial component}\label{fig:rhovis}
+\end{figure}
+
+\hypertarget{country-wide-trend}{%
+\section{Country-wide trend}\label{country-wide-trend}}
+
+Since all socio-demographics have been standardized to have a mean of zero and a variance of one, the intercept in each model can be interpreted as the estimated two-party preferred vote for an electorate with mean characteristics\footnote{Mean of all variables aside from Judaism, Indigenous, Islam and Buddhism, where it assumes the mean of the log value.}.
+The baseline of party preference has varied over the elections, with the biggest swing occuring in the 2007 election where the mean electorate shifted more than five percentage points in favour of the Labor party.
+
+\begin{figure}[h]
+
+{\centering \includegraphics[height=0.2\textheight,]{electoral-modelling-draft_files/figure-latex/plotintercept-1}
+
+}
+
+\caption{Estimated intercept for each election, which represents the two-party preferred vote for an electorate with mean characteristics.}\label{fig:plotintercept}
+\end{figure}
 
@@
 To investigate the socio-demographics that have a strong effect on the two-party preferred vote, partial residual plots are used.
-These show the direction, size and significance of an estimated effect, as the slope of the prediction line matches the estimated coefficient, and the shaded region represents a 95\% confidence band.
+These show the direction, size and significance of an estimated effect - the slope of the prediction line matches the estimated coefficient, and the shaded region represents a 95\% confidence band.
 If a horizontal line can be drawn through the confidence band, then the effect is insignificant.
-Plots for each election are faceted to compare the effects over time.
+Plots for each election are faceted to compare the effects over time in Figure \ref{fig:partresplot1} and Figure \ref{fig:partresplot2}.
 
-\emph{It is important here to note the ecological fallacy - insights are being drawn at the electorate level, and cannot be inferred for another disaggregate level (e.g.~individual voters).}
+It is important here to note the ecological fallacy - insights are being drawn at the electorate level, and cannot be inferred for another disaggregate level (e.g.~individual voters).
 
@@
 Typically the Labor party campaigns on more progressive policies, which often include tax reform that adversely affects higher income earners, and more generous social assistance programs.
-Although topical in the lead up to an election, incomes only significantly affected electoral two-party preference in 2016 (see Figure \ref{fig:plotincomes}), with higher income areas associated with support for the Liberal party.
-In the prior elections, the estimated effect is still positive, albeit insignificant.
-Unemployment on the other hand, was influential in 2001, 2004 and 2007 elections(Figure \ref{fig:plotunemploy}), with electorates with higher unemployment favouring Labor.
-This estimated effect has become weaker over time, with unemployment having virtually no effect in 2016.
+Perhaps due to these policies, higher income electorates have been more likely to support the Liberal party, as the \texttt{Incomes} factor has a positive effect on Liberal preference (see row 1 in Figure \ref{fig:partresplot1}).
+This effect is signficant in every election aside from 2004, in which it is only marginally insignificant (\(p = 0.0613\)).
+Unemployment however, has not been as influential.
+In 2001 and 2004, electorates with higher unemployment were more likely to support Labor, but over time this has shifted towards support for the Liberal party, culminating in a significantly positive effect in 2016.
+
+\hypertarget{industry-and-type-of-work}{%
+\subsection{Industry and type of work}\label{industry-and-type-of-work}}
+
+Electorates with higher proportions of workers in mining, gas, water, agriculture, waste and electricity (grouped as \texttt{Extractive} industries) are consistently linked with higher support for the Liberal party, with the mangitude of this effect slightly increasing over the years (see row 3 in Figure \ref{fig:partresplot1}).
+This is unsurprising, as the Liberal party has close ties with these traditional energy industries, and typically present policies to reduce taxation on energy production.
+Furthermore, electorates with more workers in construction or manufacturing industries (\texttt{Transformative}) are also more likely to support the Liberal party (see row 4 in Figure \ref{fig:partresplot1}).
+
+Similarly, workers in managerial, administrative, clerical and sales roles (\texttt{ManagerAdminClericalSales}) is also a significant predictor of two-party preference vote across all six elections, with a higher proportion of people working these jobs increasing Liberal support.
+The magnitude of this effect has also increased over the years.
+
+\hypertarget{household-mobility}{%
+\subsection{Household mobility}\label{household-mobility}}
+
+In each of the six elections, electorates with a higher proportion of people that have recently moved house (meaning in last five years) were more likely to support the Liberal party, although this effect was marginally insignificant in 2013 (see row 6 in Figure \ref{fig:partresplot1}.
+Having controlled for characteristics of house ownership and rental prices (via the factors \texttt{PropertyOwned} and \texttt{RentLoan} respectively), this is effect is somewhat surprising.
+
+\hypertarget{relationships}{%
+\subsection{Relationships}\label{relationships}}
+
+De facto relationships, but not marriages, are found to be an important (and significant) predictor of the two-party preferred vote in all six elections, with more de facto relationships associated with higher support for the Labor party.
+Marriages however, are insignficiant.
+
+\hypertarget{age}{%
+\subsection{Age}\label{age}}
+
+Older regions are often believed to be more conservative, and it can be seen that electorates with a higher median age have been more likely to support the Liberal party - although this effect is only significant in 2007 and 2010 (see row 2 in Figure \ref{fig:partresplot2}).
+
+\hypertarget{education}{%
+\subsection{Education}\label{education}}
+
+Since 2007, electorates with higher education levels have been associated with supporting the Labor party, although this effect is only significant in 2016.
+Prior to 2007, education had an almost zero effect (see row 3 in \ref{fig:partresplot1}).
+
+\hypertarget{diversity}{%
+\subsection{Diversity}\label{diversity}}
+
+Larger migrant populations from Asia, the Middle East, South-Eastern Europe, the United Kingdom and elsewhere, are either assocated with Labor support, or have no effect.
+Of these areas, only South-Eastern European populations were signficant in each election and Asian migrants were a significant in 2010.
+Speaking other language (aside from English) however, have a far stronger effect, as observed through \texttt{OtherLanguageHome}.
+Electorates with more diverse speech were more likely to support the Liberal party from 2004 onwards, with this effect being significant in 2007, 2010 and 2016.
+Furthermore, of the variables relating to religion, only Judaism shows a consistent effect, with electorates with larger Jewish populations more likely to vote Liberal.
+
+\hypertarget{partial-residual-plots}{%
+\subsection{Partial residual plots}\label{partial-residual-plots}}
 
@@
 
-{\centering \includegraphics[width=0.6\linewidth]{electoral-modelling-draft_files/figure-latex/plotincomes-1}
+{\centering \includegraphics{electoral-modelling-draft_files/figure-latex/partresplot1-1}
 
@@
 
-\caption{Effect of Incomes}\label{fig:plotincomes}
+\caption{Partial residual plots for: income and unemployment, industry and type of work, and household mobility.}\label{fig:partresplot1}
 \end{figure}
@@
 
-{\centering \includegraphics[width=0.6\linewidth]{electoral-modelling-draft_files/figure-latex/plotunemploy-1}
+{\centering \includegraphics{electoral-modelling-draft_files/figure-latex/partresplot2-1}
 
@@
 
-\caption{Effect of Unemployment}\label{fig:plotunemploy}
+\caption{Partial residual plots for: relationships and age, education and diversity.}\label{fig:partresplot2}
 \end{figure}
 
-\hypertarget{industry-and-type-of-work}{%
-\subsection{Industry and type of work}\label{industry-and-type-of-work}}
+\hypertarget{a-note-on-similar-variables}{%
+\subsection{A note on similar variables}\label{a-note-on-similar-variables}}
 
-Electorates with higher proportions of workers in mining, gas, water, agriculture, waste and electricity (grouped as \texttt{Extractive} industries) are consistently linked with higher support for the Liberal party, with the mangitude of this effect slightly increasing over the years (see Figure \ref{fig:plotextractive}).
-This is unsurprising, as the Liberal party has close ties with these traditional energy industries, and typically present policies to reduce taxation on energy production.
-Furthermore, electorates with more workers in construction or manufacturing industries (\texttt{Transformative}), as well as wholesale trade, retail trade, transport, post or warehousing (\texttt{Distributive}), are also linked with the Liberal party.
-However, these effects are not as strong, and \texttt{Transformative} is marginally insignificant in the 2010 election, whereas \texttt{Distributive} was not a key socio-demographic in 2001 or 2016.
-These effects are shown in Figures \ref{fig:plottransformative} and \ref{fig:plotdistributive}.
+Many of the Census variables represent similar information, which is why factors were created and some variables were removed.
+However, there still remain some variables which are closely related.
+For example, electorate income levels (via \texttt{Incomes}) is likely to be related to electoral unemployment and labor force participation (via \texttt{Unemployment}).
+In 2001, the coefficient estiamte for \texttt{Unemployment} negative but not significant, and the \texttt{Incomes} variables is significant.
+If the \texttt{Incomes} variable is removed from the model, \texttt{Unemployment} absorbs the negative effect, becoming significant (\(p = 0.0056\)).
+
+\hypertarget{a-closer-look-at-the-residuals}{%
+\section{A closer look at the residuals}\label{a-closer-look-at-the-residuals}}
+
+\hypertarget{residuals-by-state}{%
+\subsection{Residuals by state}\label{residuals-by-state}}
+
+It is often hypothesized that states have systematic differences that cause their electorates to vote differently.
+Boxplots of residuals grouped by state reveal that only Tasmania, the Australian Capital Territory and the Northern Territory appear to have a state-specific effect that is not captured by the models.
+Tasmania and the Australian Capital Territory appear to have a bias towards Labor, whereas the Northern Territory has one towards Liberal.
+There are few electorates in these states (five, two and two respectively), so this might be due to incumbent effects rather than an actual state-specific bias.
 
@@
 
-{\centering \includegraphics[width=0.6\linewidth]{electoral-modelling-draft_files/figure-latex/plotextractive-1}
+{\centering \includegraphics[width=0.8\linewidth,]{electoral-modelling-draft_files/figure-latex/resstate-1}
 
@@
 
-\caption{Effect of Extractive}\label{fig:plotextractive}
+\caption{Boxplot of residuals by state with jittered points.}\label{fig:resstate}
 \end{figure}
 
-\begin{figure}[h]
+\hypertarget{outlier-electorates}{%
+\subsection{Outlier electorates}\label{outlier-electorates}}
 
-{\centering \includegraphics[width=0.6\linewidth]{electoral-modelling-draft_files/figure-latex/plottransformative-1}
+Based on the distribution of the cook's distance values, a cook's distance greater than \(0.1\) is considered to be influential and a potential outlier.
+The electorate of Sydney (NSW) has a large cook's distance from 2001 to 2013, due to its diverse population (language, birthplace and religion), high number of defacto relationships, high income, high household mobility and small amount of workers in extractive and transformative jobs.
+It has remained a strong supporter of the Labor party and Liberal vote is severely overpredicted by the model, making it an outlier.
+Nearby in metropolitan NSW, the electorate of Wentworth is found to be an outlier in all but the 2007 election.
+Although historically Liberal, its two-party vote jumped by over 10 percentage points in 2010 without experiencing any notable changes in its socio-demographic makeup - implying that this may be the direct effect of its Liberal member, Malcolm Turnbull, becoming the leader of the Liberal party.
+Liberal support in Wentworth is underpredicted by the model in each year, and more so with Turnbull as Liberal leader.
 
-}
+Lingiari, an electorate taking up almost all of the Northern Territory, is an outlier in the 2001-2007 elections due to its large Indigenous population, young age profile and low rates of property ownership.
+Fowler (NSW) has a diverse population with a high proportion of migrants, many Buddhists and Muslims, and has very strong Labor support, making it influential in 2001, 2004 and 2010.
+Other electorates with large cook's distance are Barton (NSW) and Leichhardt (QLD) in 2016, and Canberra (ACT) in 2007.
 
-\caption{Effect of Transformative}\label{fig:plottransformative}
-\end{figure}
-\begin{figure}[h]
+\hypertarget{conclusion}{%
+\chapter{Conclusion}\label{conclusion}}
 
-{\centering \includegraphics[width=0.6\linewidth]{electoral-modelling-draft_files/figure-latex/plotdistributive-1}
+Across the past six elections, most of the socio-demographics that drive the electoral two-party preferred vote have remained steady, whilst a few (typically weaker) effects have varied over time.
+Industry and type of work are particularly influential, with energy-related and manufacturing/construction jobs, as well as administrative roles being strongly linked with the Liberal party in all elections.
+Incomes have had a similarly consistent effect, with higher income areas supporting Liberal.
+Higher levels of unemployment have shifted from weak association with Labor to a significant Liberal effect over the years, and higher education levels have been associated with Labor since 2007 (although only significant in 2016).
+It is also found that electorates with higher household mobility support Liberal, birthplace diversity favours Labor and more de facto relationships align with Labor preference - although marriages, family and household sizes have no material influence.
+Furthermore, the neighbourhood (spatial) effects were found to be positive in all elections, although only significant in 2001 and 2016, meaning that in the 2004-2013 elections, electorates effectively voted independently.
 
-}
+The findings in this paper complement the existing literature by modelling temporal trends, which as far as the authors are aware, has not been done previously for Australian elections using a regression framework.
+It is also the first study to model any Australian election since 2010 using Census information.
+Additionally, a key contribution of this research is the spatio-temporal imputation, which is applied to obtain Census data for the 2004, 2007, 2010 and 2013 elections.
+All data sets used in this study have been contributed to the \texttt{eechidna} \texttt{R} package, which provides a rich, accessible data resource for future Australian electoral analysis.
 
-\caption{Effect of Distributive}\label{fig:plotdistributive}
-\end{figure}
+\hypertarget{acknowledgements}{%
+\chapter{Acknowledgements}\label{acknowledgements}}
 
-Similarly, the proportions of workers in managerial, administrative, clerical and sales roles (\texttt{ManagerAdminClericalSales}) is also a significant predictor of two-party preference across all six elections, with higher proportions supporting the Liberal party.
+This paper was produced using \texttt{RMarkdown}\citep{rmarkdown} and \texttt{knitr} \citep{knitr}.
+All corresponding code for this paper can be found in the github repository (\url{https://github.com/jforbes14/journal-writing}), and the data used is available in the \texttt{eechidna} package \citep{eechidna}.
+All raw data was obtained from the Australian Electoral Commission, the Australian Bureau of Statistics and the Australian Government.
 
-\begin{figure}[h]
+\hypertarget{software}{%
+\chapter{Software}\label{software}}
 
-{\centering \includegraphics[width=0.6\linewidth]{electoral-modelling-draft_files/figure-latex/plotmanager-1}
+All election and Census data sets, along with electoral maps and more, are available in the \texttt{eechidna} \texttt{R} package, which can be downloaded from CRAN.
+\texttt{eechidna} (Exploring Election and Census Highly Informative Data Nationally for Australia) makes it easy to look at the data from the Australian Federal elections and Censuses that occurred between 2001 and 2016.
+This study contributed a large revision to the \texttt{eechidna} package, which included the addition of election and Census data for 2001-2010, vote outcomes for polling booths and imputed Census data for election years.
+For more details on using \texttt{eechidna}, please see the articles (vignettes) on the github page (\url{https://ropenscilabs.github.io/eechidna/}).
 
-}
-
-\caption{Effect of ManagerAdminClericalSales}\label{fig:plotmanager}
-\end{figure}
-
-\hypertarget{household-mobility}{%
-\subsection{Household mobility}\label{household-mobility}}
-
-\begin{figure}[h]
-
-{\centering \includegraphics[width=0.6\linewidth]{electoral-modelling-draft_files/figure-latex/plotdiffaddress-1}
-
-}
-
-\caption{Effect of DiffAddress}\label{fig:plotdiffaddress}
-\end{figure}
-
-Electorates with a higher proportion of people that have recently moved house (meaning in last five years) are more likely to support the Liberal party, with this effect becoming stronger since the 2010 election.
-Having controlled for characteristics of house ownership and rental prices (via the factors \texttt{PropertyOwned} and \texttt{RentLoan} respectively), this is effect is somewhat surprising.
-\emph{Add more here?}
-
-\hypertarget{indigenous-population}{%
-\subsection{Indigenous population}\label{indigenous-population}}
-
-\begin{figure}[h]
-
-{\centering \includegraphics[width=0.6\linewidth]{electoral-modelling-draft_files/figure-latex/plotindigenous-1}
-
-}
-
-\caption{Effect of Indigenous population}\label{fig:plotindigenous}
-\end{figure}
-
-The indigenous population of an electorate is found to be a significant predictor in every election, with higher indigenous populations supporting the Liberal party.
-There are a few outliers with Lingiari (NT) being the most extreme.
-However, Lingiari itself has been a strong Labor supporter, and this effect appears to be driven by the other electorates with high Indigenous populations (Kalgoorlie (WA), Leichhardt (NSW), Durack (WA), Parkes(NSW) and Kennedy (QLD)), which have almost always elected a Liberal candidate.
-
-\hypertarget{relationships-and-families}{%
-\subsection{Relationships and families}\label{relationships-and-families}}
-
-In each of the six elections, higher proportions people in de facto marriages are tied to support for the Labor party, whereas the effect of marriages themselves are insignificant.
-Larger family and household size (via the \texttt{FamHouseSize} factor) is also linked with Labor support, although only being significant in the 2007 and 2010 elections (see Figure \ref{fig:plotfamhousesize}).
-
-\emph{Potentially remove plots}
-
-\hypertarget{residuals}{%
-\section{Residuals}\label{residuals}}
-
-Largest residuals
-
-Residuals by state
-
-\begin{center}\includegraphics{electoral-modelling-draft_files/figure-latex/unnamed-chunk-2-1} \end{center}
+The authors would like to give a sincere thanks to Anthony Ebert, Heike Hofmann, Thomas Lumley, Ben Marwick, Carson Sievert, Mingzhu Sun, Dilini Talagala, Nicholas Tierney, Nathaniel Tomasetti, Earo Wang and Fang Zhou, all of whom have contributed to the \texttt{eechidna} package.
 

### 2019-03-29 old_book_files/raw-files/introduction.Rmd
moved all old bookdown files to old_book_files

--- 2019-03-20 introduction.Rmd
@@
 
-Joining these two data sources is problematic as there in an inherent asynchronicity in the two events, a Census occurs every five years, whereas an election usually occurs every three years.
+Joining these two data sources is problematic as there in an inherent asynchronicity in the two events, a Census is conducted by the Australian Bureau of Statistics (ABS) every five years, whereas a federal election (conducted by the Australian Electoral Commission (AEC)) usually occurs every three years.
 The first problem addressed is that of obtaining appropriate Census data for the 2004, 2007, 2010 and 2013 elections - election years in which a Census does not occur.
@@
 
-Previous work on modelling voting outcomes with aggregate socio-demographics find that relatively good models can be fit.
+Previous work on modelling Australian federal elections have found that aggregate socio-demographics are relatively good predictors of voting outcomes.
 @Forrest01 does this using multiple regression of the Liberal and Labor primary vote for polling booths in the Farrer electorate in 1998.
-@Stimson06, @Stimson09 and @Stimson12 conduct principal component analysis of polling booths in the 2001, 2004 and 2007 elections respectively, also finding that socio-demographic characteristics of polling booths are linked to their two-party preferred vote.
-@Stimson09 model the polling booth swing vote (change in the two-party preferred vote) in the 2007 election, but had little success.
-Instead of analyzing a single election in isolation, this paper employs a consistent model framework across six elections so that any temporal changes are observed, where each federal elections is modelled with a cross-sectional data set.
-Each cross-section consists of the two-party preferred vote (response variable), and socio-demographic variables (explanatory variables) that characterise each electorate.
-This process involves standardizing variables, using principal components to group variables into "factors" and finally fitting a spatial error model.
-These steps, and the rational behind a spatial model, are discussed in @CHAPTER4.
-Inference is then conducted on the models and insights are discussed in @CHAPTER5.
+@Stimson06, @Stimson09 and @Stimson12 use principal component analysis of polling booths in the 2001, 2004 and 2007 elections respectively, also finding that socio-demographic characteristics of polling booths are linked to their two-party preferred vote.
+On the contrary, @Stimson09 models the polling booth swing vote (change in the two-party preferred vote) in the 2007 election, finding that little of swing vote can be explained by Census data.
+Instead of analyzing a single election in isolation, this paper employs a consistent model framework across six elections so that temporal changes in the effects of socio-demographics can be observed, where each federal elections is modelled with a cross-sectional data set.
+The use of a regression framework to examine these socio-political relationships over time is seemingly absent from previous Australian studies.
+It also appears that no study has attempted any type of statistical analysis of socio-demographics in conjunction with voter behaviour in Australia since 2007, making this paper distinctly different from those previous.
+
+The cross-sectional data set for each election consists of the two-party preferred vote (response variable), and socio-demographic variables (explanatory variables) that characterise each electorate.
+To obtain these cross-sections, socio-demographic variables are first standardized, and then principal components are used to group variables into "factors".
+To account for the inherent spatial structure of the data, a spatial error model is fit for each election.
+These steps are discussed in @CHAPTER4.
+In @CHAPTER5 inference and diagnostics are conducted on the models to see which effects are significant, how effects change over time and which electorates have abnormal voting behaviour.
 

### 2019-03-29 old_book_files/raw-files/modelling.Rmd
moved all old bookdown files to old_book_files

--- 2019-03-21 modelling.Rmd
@@
 With only $N = 150$ observations (electorates) in each election and $p = 59$ socio-demographic variables in each cross-section, any model using all variables would face serious problems with multi-collinearity and over-fitting, likely leading to erroneous conclusions regarding variable significance.
-Therefore a form of dimension reduction is adopted that involves a two-stage approach that results in the identification of a reduced predictor set, for which models are fit.
+Therefore a form of dimension reduction is adopted that involves a *two-stage approach* that results in the identification of a reduced predictor set, for which models are fit.
 
@@
 
-The spatial error model [@Anselin99] is chosen because it allows for spatial correlation in the errors.
-This can be thought of as capturing the political climate, where political pressures (unobserved characteristics influencing voter behaviour) of neighbours influence an electorate's voting behaviour.
-Spatial weights are calculated by allowing each electorate to be equally correlated with any electorate that shares any part of its boundary.
+The spatial error model [@Anselin99] is chosen because captures spatial heterogeneity by incorporating a spatially structured random effect vector [@LeSage2009].
+In this context, the random effect thought of as capturing the political climate in each electorate, where the climate is correlated with the climate in neighbouring electorates.
+This functions under the assumption that the climate is independent of electoral socio-demographics, and that an electorate is equally correlated with any electorate that shares a part of its boundary.
+Spatial weights are calculated in accordance with these assumptions.
 The spatial error model is specified as follows:
 
-Let $\rho$ be the parameter governing the extent of spatial correlation, $\boldsymbol v$ be a spherical error term, ${\bf W}$ be a matrix of spatial weights (containing information about the neighbouring regions), $\bf X$ be a matrix of socio-demographic covariates, $\boldsymbol \beta$ be a vector of regression coefficients and ${\boldsymbol u}$ be a vector of spatially correlated errors.
+Let $\rho$ be spatial autoregressive coefficient, $\boldsymbol v$ be a spherical error term, ${\bf W}$ be a matrix of spatial weights (containing information about the neighbouring regions), $\bf X$ be a matrix of socio-demographic covariates, $\boldsymbol \beta$ be a vector of regression coefficients and $\boldsymbol a$ be a spatially structured random effect vector.
 
-$${\bf y} = {\bf X} {\boldsymbol \beta} + {\boldsymbol u}$$
+$${\bf y} = {\bf X} {\boldsymbol \beta} + {\boldsymbol a}$$
+and
 
-with
-
-$${\boldsymbol u} = \rho {\boldsymbol W} {\boldsymbol u} + {\boldsymbol v}$$
+$${\boldsymbol a} = \rho {\boldsymbol W} {\boldsymbol a} + {\boldsymbol v}$$
 where
@@
 
-Estimation is done via maximum likelihood (using the `spdep` `R` package [@spdep]).
+Estimation is done using generalized least squares (using the `spdep` `R` package [@spdep]).
+
+Table 4.1 details the resultant estimated model coefficients and their estimated standard errors for each of the six elections. These are interpreted in the next section.
+
+*Model diagnostics are shown in Appendix \@ref(a:diagnostics), including the outcomes of tests for heteroskedasticity, residual non-normality, non-linear patterns in the residuals against included predictors, patterns in the residuals against missing predictors and influential points.
+These assessments did not indicate a major violation of any of the assumptions.*
 

### 2019-03-29 old_book_files/raw-files/results.Rmd
moved all old bookdown files to old_book_files

---
title: "results"
author: "Jeremy Forbes"
date: "19/03/2019"
output: html_document
---

```{r}
library(spdep)
library(tidyverse)
load("/Users/Jeremy/Documents/R/eechidna-modelling/data/model_df.rda")
load("/Users/Jeremy/Documents/R/eechidna-modelling/data/sp_weights_16.rda")
load("/Users/Jeremy/Documents/R/eechidna-modelling/data/sp_weights_13.rda")
load("/Users/Jeremy/Documents/R/eechidna-modelling/data/sp_weights_10.rda")
load("/Users/Jeremy/Documents/R/eechidna-modelling/data/sp_weights_07.rda")
load("/Users/Jeremy/Documents/R/eechidna-modelling/data/sp_weights_04.rda")
load("/Users/Jeremy/Documents/R/eechidna-modelling/data/sp_weights_01.rda")

fmod16 <- errorsarlm(LNP_Percent ~ ., 
  data=(model_df %>% filter(year == "2016") %>% dplyr::select(-c(year, DivisionNm))),
  sp_weights_16, etype="error", method="eigen", interval=c(-1,0.999))

fmod13 <- errorsarlm(LNP_Percent ~ ., 
  data=(model_df %>% filter(year == "2013") %>% dplyr::select(-c(year, DivisionNm))),
  sp_weights_13, etype="error", method="eigen", interval=c(-1,0.999))

fmod10 <- errorsarlm(LNP_Percent ~ ., 
  data=(model_df %>% filter(year == "2010") %>% dplyr::select(-c(year, DivisionNm))),
  sp_weights_10, etype="error", method="eigen", interval=c(-1,0.999))

fmod07 <- errorsarlm(LNP_Percent ~ ., 
  data=(model_df %>% filter(year == "2007") %>% dplyr::select(-c(year, DivisionNm))),
  sp_weights_07, etype="error", method="eigen", interval=c(-1,0.999))

fmod04 <- errorsarlm(LNP_Percent ~ ., 
  data=(model_df %>% filter(year == "2004") %>% dplyr::select(-c(year, DivisionNm))),
  sp_weights_04, etype="error", method="eigen", interval=c(-1,0.999))

fmod01 <- errorsarlm(LNP_Percent ~ ., 
  data=(model_df %>% filter(year == "2001") %>% dplyr::select(-c(year, DivisionNm))),
  sp_weights_01, etype="error", method="eigen", interval=c(-1,0.999))

# Coefficients and confidence bands
coef_df <- bind_rows(
  data.frame(variable = fmod16$coefficients %>% names, estimate = fmod16$coefficients %>% unname, se = fmod16$rest.se %>% unname, p = summary(fmod16)$Coef[,4] %>% unname, year = 2016),
  data.frame(variable = fmod13$coefficients %>% names, estimate = fmod13$coefficients %>% unname, se = fmod13$rest.se %>% unname, p = summary(fmod13)$Coef[,4] %>% unname, year = 2013),
  data.frame(variable = fmod10$coefficients %>% names, estimate = fmod10$coefficients %>% unname, se = fmod10$rest.se %>% unname, p = summary(fmod10)$Coef[,4] %>% unname, year = 2010),
  data.frame(variable = fmod07$coefficients %>% names, estimate = fmod07$coefficients %>% unname, se = fmod07$rest.se %>% unname, p = summary(fmod07)$Coef[,4] %>% unname, year = 2007),
  data.frame(variable = fmod04$coefficients %>% names, estimate = fmod04$coefficients %>% unname, se = fmod04$rest.se %>% unname, p = summary(fmod04)$Coef[,4] %>% unname, year = 2004),
  data.frame(variable = fmod01$coefficients %>% names, estimate = fmod01$coefficients %>% unname, se = fmod01$rest.se %>% unname, p = summary(fmod01)$Coef[,4] %>% unname, year = 2001)
)
```

## Results

```{r}
#stargazer::stargazer(fmod16)
```

### Spatial correlation

The spatial autoregressive coefficient represents the extent to which an electorate is affected by the surrounding political climate. In the 2001, 2013 and 2016 elections $\rho$ is positive and significant, indicating that electoral voting outcomes were influenced by attitudes of neighbouring electorates, in the sense that two-party preference. However, in the 2004, 2007 and 2010 elections, $\rho$ is not significantly different from zero, meaning that neighbours did not influence the two-party preferred vote over and above the effect of electoral socio-demographics.

```{r}
coef_df %>% 
  filter(variable != "(Intercept)") %>% 
  mutate(upper95 = estimate + 1.96*se, lower95 = estimate - 1.96*se) %>% 
  ggplot() +
  geom_point(aes(x = year, y = estimate, col = factor(p < 0.05)), size = 2) +
  geom_linerange(aes(x = year, ymin = lower95, ymax = upper95, col = factor(p < 0.05)), size = 1.5) + 
  geom_hline(aes(yintercept = 0), alpha = 0.5, size = 1) + 
  facet_wrap(~variable, scales = "free") +
  scale_color_manual(values = c("grey50", "black")) +
  guides(col = F) +
  theme(axis.text = element_text(size = 8), axis.text.x = element_text(angle = 60, hjust=1)) +
  scale_x_continuous(breaks = c(2001, 2004, 2007, 2010, 2013, 2016))
```

### Persistent effects
Defacto - Labor
DiffAddress - Liberal
Extractive - Liberal
Indigenous - Liberal
Admin - Liberal

```{r, include = F}
coef_df %>% 
  filter(variable %in% c("DeFacto", "DiffAddress", "Extractive", "Indigenous", "ManagerAdminClericalSales")) %>% 
  mutate(upper95 = estimate + 1.96*se, lower95 = estimate - 1.96*se) %>% 
  ggplot() +
  geom_point(aes(x = year, y = estimate), size = 3) +
  geom_linerange(aes(x = year, ymin = lower95, ymax = upper95), size = 1.5) + 
  geom_hline(aes(yintercept = 0), alpha = 0.5, size = 1) + 
  facet_wrap(~variable) +
  scale_color_manual(values = c("grey50", "black"))
```

### Changing effects
Unemployment - Labor then none
FamilyHouseSize - Labor 2004-2010
Incomes - Liberal in 2016
MedianAge - Liberal 2004 and 2016

```{r, include = F}
coef_df %>% 
  filter(variable %in% c("Unemployment", "FamHouseSize", "Incomes", "MedianAge")) %>% 
  mutate(upper95 = estimate + 1.96*se, lower95 = estimate - 1.96*se) %>% 
  ggplot() +
  geom_point(aes(x = year, y = estimate, col = factor(p < 0.05)), size = 3) +
  geom_linerange(aes(x = year, ymin = lower95, ymax = upper95, col = factor(p < 0.05)), size = 1.5) + 
  geom_hline(aes(yintercept = 0), alpha = 0.5, size = 1) + 
  facet_wrap(~variable) +
  scale_color_manual(values = c("grey50", "black")) +
  guides(col = F)
```

### Alternative: discuss each election individually/over time

Over the six elections, it is apparent that five socio-demographic characterstics have been strongly associated with support for one of the parties in all elections, whilst others are influential in some, and others not at all. The liberal party appeals to XXX. Labor party XXX. In 2001, ....

### Diagnostics
Goodness of fit?

Residuals: GLS residuals should be normal (QQ-plot)
- fmod16$residuals

Outliers:
-- Leverage: Hat values (X)
-- Influence: Cook's distance

#### Electorates that are very different to their neighbours
- Residuals vs weighted neighbours residuals
- Socio-demographics that are quite different?

$${\bf r_j} = {\bf \hat{u}} + \hat\beta_j{\boldsymbol X_j}$$
where

$${\bf \hat{u}} = ({\bf I_n}-\hat\rho \cdot {\bf W}) {\bf  \hat v}$$

where

$${\bf  \hat v} \sim N({\bf 0}, \sigma^2 \cdot {\bf I_n}) $$

### 2019-03-29 old_book_files/raw-files/some-workings.Rmd
moved all old bookdown files to old_book_files

---
title: "Some Workings"
author: "Jeremy Forbes"
date: "14/03/2019"
output: html_document
---

Structure of the article

## Introduction
Motivation:
- Few existing studies on relationship between aggregate socio-demographics and voting behaviour in Australia.
- Common approach is to borrow socio-demographic information from the nearest Census - but there's both a spatial and temporal problem with this. Motivates interpolation to obtain Census data for election years.
- This analysis - Create an easily accessible resource for exploration of Census and election data
- Gain insights into electoral voting determinants over time

Intro:
- Data wrangling from raw ABS and AEC structures, navigating their webpages
- Problem: Asynchronicity with boundaries and time
- Spatial and temporal imputation
- Model characteristics of the Australian electorate over time, to see which socio-demographics influence electoral two-party preference

## Data collection
Consistency of data formats
- Raw files in excel and all over the place
- Construct dataframes using tidyverse tools
Reproducibility
- Scripts to deal with upcoming elections and Censuses
- All data can be regenerated for transparecy

## Imputation
Methodology
- Not specific to electorates
- Explain spatial and then temporal

Applied to electorates
- For our models we use electorates
- Visualisations of an example using electorates
- Possible to use finer levels of disaggregation (e.g. SA1). Propose a cascading technique for reduced computation burden (SA4, SA3, SA2, SA1).

## Modelling
- Join the data, treat at separate cross-sections
- Address the issue of changing variable scales (rent as example)
- PCA to group variables together
- Akaike weights to derive a variable set for models
- Spatial correlation in response, so model as spatial error model

## Results
Visualise interesting and significant effects
- Discuss what this means
- What changes over time
Mention what else is significant
Discuss significance of spatial dependence
Discuss which are apparently not important
- Reference articles on what is believed to be important if possible

## Conclusion
Summarise findings and make broader statement about electorate characteristics
Discuss the eechidna package and encourage its use
Further extensions - model polling booths, revise imputation using cascading approach

## Areal interpolation

Following the piecewise approximation method in Goodchild, Anselin and Deichmann (1993), is appropriate for our problem. Publically available Census data is aggregated, of which there are different resolutions available, ranging from SA1 (over 50,000 zones) to electoral divisions (150 zones). The boundaries for these regions change over time. Therefore, we denote the map of electoral divisions at the time of an election to be the target zones $t = 1,...,T$, and the Census map to be the source zones $s = 1,...,S$.

Denote the population of intersection source zone $s$ and target zone $t$ be $P_{s,t}$, population of the source zone $s$ $U_s$, and area of intersection between source zone $s$ and target zone $t$ be denoted $A_{s,t}$. Therefore the estimated population of the intersection is given by:
$$\hat{P}_{s,t} = U_s*A_{s,t}$$

In order to calculate Census information for the target zone, denote a given Census variable for the target zone $C_t$, and the same Census variable for the source zone $D_s$.

$$\hat{C}_t = \frac{\sum_{s=1}^{S}{D_s*\hat{P}_{s,t}}}{\sum_{s=1}^{S}{\hat{P}_{s,t}}}$$

Estimated Census information is now imputed for the target zones, but these are not appropriate for the target year. Denote year $y$, with a Census falling on $y_1$ and $y_3$, and an election on year $Y=y_2$, and add this subscript to the Census variable estimate, $\hat{C}_{t,y}$. Linear interpolation is then used between Census years to get the final estimate of a Census variable for the target zone in the election year $y_2$.

$$\hat{C}_{t,y_2} = \frac{y_3-y_2}{y_3-y_1}*\hat{C}_{t,y_1} + \frac{y_2-y_1}{y_3-y_1}*\hat{C}_{t,y_3}$$

## Uncertainty in the estimates

Think of a way to get some kind of uncertainty estimate? If not, some kind of sensitivity? Say introduce some random error to the population intersection estimates and see how that manifests. Also what about different time interpolation - using say quadratic distance instead of linear? See how that changes the models/conclusions?

Instead of linear interpolation, could use electorate specific weights based on electoral enrolments as a proxy for population.

### 2019-03-29 draft-paper.Rmd
new EBS working paper files

--- 2019-03-25 electoral-modelling-draft.Rmd
@@
 title: "Spatial modelling of the electoral two-party preferred vote in Australia. A study of federal elections between 2001 and 2016 via the `eechidna` `R` package."
-author: "Jeremy Forbes, Di Cook & Rob Hyndman"
-date: "`r Sys.Date()`"
-site: bookdown::bookdown_site
-output: bookdown::gitbook
-documentclass: book
-classoption: openany
-bibliography: [book.bib, packages.bib]
-biblio-style: apalike
-link-citations: yes
-github-repo: rstudio/bookdown-demo
-description: "This is an incomplete draft of a paper to submitted to ANZJS."
+author:
+- familyname: Forbes
+  othernames: Jeremy
+  address: Monash University\newline Melbourne, Australia
+  email: jeremyforbes1995@gmail.com
+  correspondingauthor: true
+- familyname: Cook
+  othernames: Di
+  address: Monash University\newline Melbourne, Australia
+- familyname: Hyndman
+  othernames: Rob
+  address: Monash University\newline Melbourne, Australia
+abstract: "A brief summary of our ideas"
+keywords: "blah, blah"
+wpnumber: no/yr
+jelcodes: C10,C14,C22
+blind: false
+cover: true
+toc: false
+bibliography: references.bib
+biblio-style: authoryear-comp
+header-includes: \usepackage{float}
+output:
+  MonashEBSTemplates::workingpaper:
+    fig_caption: yes
+    fig_height: 5
+    fig_width: 8
+    includes:
+      in_header: preamble.tex
+    keep_tex: yes
+    number_sections: yes
+    citation_package: biblatex
 ---
-
-<!--chapter:end:index.Rmd-->
 
@@
 library(spdep)
+library(gridExtra)
+library(grid)
+library(predictmeans)
 
@@
 
-```{r wrangle, include = F, message = F}
+```{r wrangle, include = F}
 # Function to standardize variables
@@
   gls_model <- gls(my_formula, gls_data)
+
+  # Cooks distance
+  #gls_model$cooksd <- unname(predictmeans::CookD(gls_model, plot = FALSE))
 
@@
 
-<!--chapter:end:01-intro.Rmd-->
+# Data collection, wrangling and imputation {#data}
 
-# Data collection {#data}
-
-## Voting data
+## Collecting the data
 
@@
 
-```{r timeline, message = F, echo = F, fig.cap="Timeline of Australian elections and Censuses. They do not always occur in the same year.", out.width = "90%", fig.align="center"}
+```{r timeline, echo = F, fig.cap="Timeline of Australian elections and Censuses. They do not always occur in the same year.", out.width = "90%", fig.align="center"}
 timeline <- data.frame(year = c(2001:2016)) %>%
@@
 
-<!--chapter:end:03-imputation.Rmd-->
-
----
-output:
-  pdf_document: default
-  html_document: default
----
 # Modelling {#modelling}
@@
 
-Table 4.1 details the resultant estimated model coefficients and their estimated standard errors for each of the six elections. These are interpreted in the next section.
+Table 3 details the resultant estimated model coefficients and their estimated standard errors for each of the six elections. These are interpreted in the next section.
 
 ```{r printmodels, eval = F}
-# Add rho and make adjustments to this
-stargazer::stargazer(glsmod16, glsmod13, glsmod10, glsmod07, glsmod04, glsmod01, header = FALSE, 
+# Add rho and GLS residual standard error
+stargazer::stargazer(glsmod01, glsmod04, glsmod07, glsmod10, glsmod13, glsmod16, header = FALSE, 
   column.sep.width = "1pt", 
@@
   \label{} 
-\tiny 
+\scriptsize 
 \begin{tabular}{@{\extracolsep{1pt}}lcccccc} 
@@
  & 2001 & 2004 & 2007 & 2010 & 2013 & 2016 \\ 
-\\[-1.8ex] & (1) & (2) & (3) & (4) & (5) & (6)\\ 
 \hline \\[-1.8ex] 
- $\rho$ & 0.48$^{***}$ & 0.29$^{*}$ & 0.15 & $-$0.04 & 0.15 & 0.44$^{***}$ \\ 
-  & (0.16) & (0.16) & (0.17) & (0.19) & (0.17) & (0.15) \\ 
-  & & & & & & \\ 
- AusCitizen & $-$1.53 & $-$0.87 & 1.00 & 0.67 & $-$2.83 & $-$2.77 \\ 
-  & (2.95) & (2.90) & (2.93) & (2.57) & (2.63) & (2.51) \\ 
-  & & & & & & \\ 
- Born\_Asia & $-$2.13 & $-$2.56 & $-$7.58$^{***}$ & $-$1.37 & 0.74 & 2.63 \\ 
-  & (2.20) & (2.58) & (2.68) & (1.96) & (2.55) & (3.03) \\ 
-  & & & & & & \\ 
- Born\_MidEast & $-$0.82 & 1.70 & 1.79 & 2.92$^{*}$ & 3.22$^{*}$ & 1.92 \\ 
-  & (1.56) & (1.54) & (1.54) & (1.51) & (1.78) & (1.53) \\ 
-  & & & & & & \\ 
- Born\_SE\_Europe & $-$1.63 & $-$1.25 & $-$1.72 & $-$0.20 & 0.04 & 1.02 \\ 
-  & (0.99) & (1.15) & (1.34) & (1.17) & (2.05) & (2.34) \\ 
-  & & & & & & \\ 
- Born\_UK & $-$1.37 & $-$0.35 & $-$0.29 & 0.29 & $-$0.87 & $-$0.49 \\ 
-  & (0.98) & (1.00) & (1.05) & (0.80) & (0.90) & (1.04) \\ 
-  & & & & & & \\ 
- BornElsewhere & 1.74 & 4.04 & 9.78$^{**}$ & 5.01 & $-$0.51 & $-$1.08 \\ 
-  & (4.52) & (4.46) & (4.47) & (3.59) & (3.67) & (3.40) \\ 
-  & & & & & & \\ 
- Buddhism & 1.51 & 1.09 & 2.41$^{*}$ & 0.90 & 2.84$^{*}$ & 1.50 \\ 
-  & (1.24) & (1.25) & (1.40) & (1.25) & (1.61) & (1.94) \\ 
-  & & & & & & \\ 
- Christianity & 5.17$^{**}$ & 3.08 & $-$0.60 & $-$1.93 & 3.44 & 0.53 \\ 
-  & (2.52) & (2.79) & (3.44) & (3.45) & (4.19) & (4.58) \\ 
-  & & & & & & \\ 
- CurrentlyStudying & $-$0.09 & 1.73 & 3.28$^{***}$ & 4.03$^{***}$ & 1.69 & $-$1.14 \\ 
-  & (1.13) & (1.17) & (1.21) & (1.12) & (1.09) & (1.01) \\ 
-  & & & & & & \\ 
- DeFacto & $-$11.02$^{***}$ & $-$9.73$^{***}$ & $-$8.60$^{***}$ & $-$6.55$^{***}$ & $-$5.04$^{**}$ & $-$6.42$^{***}$ \\ 
-  & (2.55) & (2.62) & (2.57) & (1.90) & (2.00) & (1.69) \\ 
-  & & & & & & \\ 
- DiffAddress & 6.93$^{***}$ & 5.02$^{***}$ & 4.72$^{***}$ & 3.40$^{***}$ & 4.63$^{***}$ & 3.86$^{***}$ \\ 
-  & (1.50) & (1.71) & (1.60) & (0.88) & (1.03) & (0.95) \\ 
-  & & & & & & \\ 
- Distributive & 1.65 & 2.76$^{**}$ & 2.88$^{**}$ & 2.25$^{**}$ & 2.75$^{**}$ & 1.67 \\ 
-  & (1.17) & (1.25) & (1.30) & (1.02) & (1.13) & (1.08) \\ 
-  & & & & & & \\ 
- Education & $-$3.04 & $-$0.90 & $-$1.82 & $-$0.77 & 1.94 & 0.62 \\ 
-  & (2.99) & (3.04) & (3.35) & (2.89) & (2.86) & (2.23) \\ 
-  & & & & & & \\ 
- Extractive & 6.81$^{***}$ & 6.09$^{***}$ & 6.97$^{***}$ & 5.81$^{***}$ & 5.31$^{***}$ & 4.69$^{***}$ \\ 
-  & (1.31) & (1.42) & (1.48) & (1.16) & (1.23) & (1.36) \\ 
-  & & & & & & \\ 
- FamHouseSize & $-$4.35 & $-$6.16$^{*}$ & $-$8.22$^{**}$ & $-$9.67$^{***}$ & $-$5.11$^{*}$ & $-$2.94 \\ 
-  & (2.86) & (3.15) & (3.28) & (2.66) & (2.78) & (2.27) \\ 
-  & & & & & & \\ 
- Incomes & 8.46$^{***}$ & 4.19 & 2.73 & 4.36$^{*}$ & 2.95 & 2.99$^{*}$ \\ 
-  & (2.57) & (2.68) & (3.06) & (2.56) & (2.60) & (1.76) \\ 
-  & & & & & & \\ 
- Indigenous & 4.47$^{***}$ & 6.12$^{***}$ & 5.90$^{***}$ & 6.81$^{***}$ & 5.62$^{***}$ & 3.97$^{***}$ \\ 
-  & (1.38) & (1.49) & (1.49) & (1.17) & (1.35) & (1.26) \\ 
-  & & & & & & \\ 
- Islam & 2.64 & 1.20 & $-$1.04 & $-$1.21 & 0.61 & 0.15 \\ 
-  & (1.72) & (1.71) & (1.94) & (1.80) & (2.00) & (2.03) \\ 
-  & & & & & & \\ 
- Judaism & 1.96$^{***}$ & 1.60$^{*}$ & 0.46 & 0.01 & 1.74 & 1.02 \\ 
-  & (0.71) & (0.85) & (1.05) & (1.00) & (1.26) & (1.38) \\ 
-  & & & & & & \\ 
- ManagerAdminClericalSales & 4.62$^{***}$ & 3.99$^{***}$ & 4.32$^{***}$ & 5.21$^{***}$ & 3.28$^{***}$ & 2.10$^{***}$ \\ 
-  & (1.03) & (0.99) & (1.00) & (0.78) & (0.85) & (0.70) \\ 
-  & & & & & & \\ 
- Married & $-$1.39 & 1.46 & 5.07$^{*}$ & 5.06$^{*}$ & 5.50$^{**}$ & 2.27 \\ 
-  & (2.64) & (2.71) & (2.91) & (2.59) & (2.79) & (2.26) \\ 
-  & & & & & & \\ 
- MedianAge & 2.94$^{*}$ & 2.28 & 2.08 & 1.86 & 3.73$^{**}$ & 1.54 \\ 
-  & (1.63) & (1.88) & (2.08) & (1.65) & (1.55) & (1.30) \\ 
-  & & & & & & \\ 
- NoReligion & 5.76$^{**}$ & 3.53 & 0.25 & $-$0.31 & 3.43 & 1.06 \\ 
-  & (2.59) & (2.58) & (2.98) & (2.90) & (3.45) & (3.85) \\ 
-  & & & & & & \\ 
- OneParent\_House & $-$0.85 & 0.11 & 0.22 & 1.66 & 1.37 & $-$1.00 \\ 
-  & (1.39) & (1.46) & (1.61) & (1.38) & (1.49) & (1.34) \\ 
-  & & & & & & \\ 
- OtherLanguageHome & 1.27 & $-$4.69 & $-$4.37 & $-$6.02 & $-$6.76 & $-$10.40$^{**}$ \\ 
-  & (4.95) & (5.50) & (5.70) & (4.44) & (5.61) & (4.97) \\ 
-  & & & & & & \\ 
- PropertyOwned & 1.76 & 0.68 & $-$1.17 & 0.78 & $-$0.92 & $-$0.76 \\ 
-  & (1.38) & (1.43) & (1.49) & (1.09) & (1.21) & (1.19) \\ 
-  & & & & & & \\ 
- RentLoanPrice & $-$0.70 & 1.01 & 2.12 & $-$0.54 & $-$1.54 & $-$0.31 \\ 
-  & (2.02) & (1.96) & (1.96) & (1.48) & (1.68) & (1.50) \\ 
-  & & & & & & \\ 
- SocialServ & 2.90$^{***}$ & 0.94 & 1.91 & 1.77 & 1.89 & 2.42$^{*}$ \\ 
-  & (1.10) & (1.22) & (1.35) & (1.14) & (1.27) & (1.27) \\ 
-  & & & & & & \\ 
- Transformative & 4.75$^{***}$ & 3.74$^{**}$ & 3.28$^{*}$ & 5.20$^{***}$ & 5.74$^{***}$ & 2.99$^{**}$ \\ 
-  & (1.41) & (1.57) & (1.74) & (1.47) & (1.59) & (1.41) \\ 
-  & & & & & & \\ 
- Unemployment & 0.36 & $-$1.22 & $-$2.42 & $-$2.88$^{**}$ & $-$5.09$^{***}$ & $-$3.00$^{**}$ \\ 
-  & (1.42) & (1.54) & (1.73) & (1.41) & (1.51) & (1.28) \\ 
-  & & & & & & \\ 
- Constant & 50.45$^{***}$ & 53.48$^{***}$ & 49.92$^{***}$ & 47.32$^{***}$ & 52.63$^{***}$ & 50.78$^{***}$ \\ 
-  & (0.74) & (0.56) & (0.51) & (0.35) & (0.45) & (0.67) \\ 
-  & & & & & & \\ 
+ $\rho$ & 0.46$^{***}$ & 0.29$^{*}$ & 0.24 & 0.19 & 0.27$^{*}$ & 0.50$^{***}$ \\ 
+  & (0.15) & (0.17) & (0.17) & (0.16) & (0.16) & (0.17) \\
+
+ AusCitizen & $-$3.13 & $-$2.64 & $-$2.53 & $-$0.08 & $-$3.40 & $-$1.80 \\ 
+  & (2.26) & (2.43) & (2.34) & (2.79) & (2.76) & (2.71) \\
+
+ Born\_Asia & 2.22 & $-$0.95 & $-$1.60 & $-$6.83$^{**}$ & $-$3.03 & $-$0.55 \\ 
+  & (2.18) & (2.44) & (2.19) & (2.73) & (2.71) & (2.17) \\
+
+ Born\_MidEast & $-$1.15 & $-$1.59 & $-$2.01$^{*}$ & $-$2.03 & $-$0.92 & $-$1.44 \\ 
+  & (1.07) & (1.20) & (1.11) & (1.27) & (1.24) & (1.13) \\
+
+ Born\_SE\_Europe & $-$3.21$^{**}$ & $-$4.24$^{***}$ & $-$3.61$^{***}$ & $-$4.14$^{***}$ & $-$3.69$^{***}$ & $-$2.72$^{***}$ \\ 
+  & (1.42) & (1.46) & (1.02) & (1.19) & (1.07) & (0.97) \\
+
+ Born\_UK & 0.25 & $-$0.07 & 0.34 & 0.56 & $-$0.09 & $-$1.32 \\ 
+  & (1.00) & (0.98) & (0.90) & (1.07) & (1.04) & (1.04) \\
+
+ BornElsewhere & $-$5.04 & $-$4.91 & $-$4.13 & 2.35 & $-$5.23 & $-$4.14 \\ 
+  & (3.30) & (3.68) & (3.38) & (4.23) & (4.15) & (3.97) \\
+
+ Buddhism & $-$0.49 & $-$0.17 & $-$1.37 & $-$0.83 & $-$0.12 & $-$1.60 \\ 
+  & (1.39) & (1.61) & (1.61) & (1.80) & (1.68) & (1.56) \\
+
+ Christianity & $-$2.48 & $-$1.23 & 0.38 & 0.50 & 2.41 & 1.68 \\ 
+  & (1.73) & (1.85) & (1.83) & (1.99) & (1.85) & (1.78) \\
+
+ CurrentlyStudying & $-$2.19$^{**}$ & $-$0.13 & 2.06$^{*}$ & 2.12$^{*}$ & 1.15 & $-$0.16 \\ 
+  & (0.99) & (1.13) & (1.17) & (1.25) & (1.26) & (1.18) \\
+
+ DeFacto & $-$6.44$^{***}$ & $-$5.37$^{**}$ & $-$6.43$^{***}$ & $-$8.07$^{***}$ & $-$6.56$^{**}$ & $-$8.53$^{***}$ \\ 
+  & (1.87) & (2.48) & (2.31) & (3.06) & (3.11) & (2.83) \\
+
+ DiffAddress & 3.88$^{***}$ & 5.06$^{***}$ & 4.22$^{***}$ & 5.57$^{***}$ & 3.53$^{*}$ & 5.67$^{***}$ \\ 
+  & (0.94) & (1.12) & (0.99) & (1.76) & (1.91) & (1.60) \\
+
+ Distributive & 1.27 & 2.01$^{*}$ & 1.36 & 1.57 & 2.10$^{*}$ & 1.20 \\ 
+  & (1.12) & (1.21) & (1.13) & (1.34) & (1.27) & (1.21) \\
+
+ Education & 1.08 & 0.52 & $-$5.52$^{*}$ & $-$4.08 & $-$4.44 & $-$7.07$^{**}$ \\ 
+  & (2.38) & (3.12) & (3.27) & (3.95) & (3.78) & (3.55) \\
+
+ Extractive & 4.83$^{***}$ & 5.45$^{***}$ & 5.37$^{***}$ & 7.31$^{***}$ & 6.71$^{***}$ & 7.43$^{***}$ \\ 
+  & (1.48) & (1.42) & (1.36) & (1.56) & (1.47) & (1.39) \\
+
+ FamHouseSize & $-$0.16 & 0.87 & $-$2.40 & $-$2.53 & $-$3.26 & $-$2.91 \\ 
+  & (2.19) & (2.72) & (2.69) & (3.25) & (3.28) & (2.90) \\
+
+ Incomes & 4.36$^{**}$ & 5.03$^{*}$ & 9.45$^{***}$ & 7.09$^{**}$ & 7.97$^{***}$ & 12.20$^{***}$ \\ 
+  & (1.77) & (2.66) & (2.75) & (3.25) & (2.92) & (2.75) \\
+
+ Indigenous & 2.91$^{*}$ & 1.97 & 2.48 & 2.84 & 0.67 & $-$0.05 \\ 
+  & (1.68) & (1.95) & (1.75) & (2.16) & (2.14) & (2.00) \\
+
+ Islam & $-$0.92 & $-$0.97 & $-$0.54 & $-$2.50 & $-$0.82 & $-$0.95 \\ 
+  & (1.22) & (1.36) & (1.27) & (1.52) & (1.42) & (1.34) \\
+
+ Judaism & 1.88$^{*}$ & 1.78 & 2.66$^{***}$ & 1.97$^{*}$ & 2.74$^{**}$ & 1.65$^{*}$ \\ 
+  & (1.05) & (1.13) & (1.01) & (1.15) & (1.10) & (1.00) \\
+
+ ManagerAdminClericalSales & 2.06$^{***}$ & 3.32$^{***}$ & 6.00$^{***}$ & 5.47$^{***}$ & 5.04$^{***}$ & 5.78$^{***}$ \\ 
+  & (0.71) & (0.93) & (0.90) & (1.08) & (1.03) & (1.06) \\
+
+ Married & 0.44 & 0.11 & $-$1.22 & $-$0.22 & 0.91 & $-$2.34 \\ 
+  & (2.31) & (2.96) & (2.83) & (3.15) & (3.03) & (2.81) \\
+
+ MedianAge & 2.32$^{*}$ & 4.96$^{***}$ & 3.66$^{**}$ & 4.00$^{*}$ & 2.30 & 2.87 \\ 
+  & (1.32) & (1.65) & (1.81) & (2.26) & (2.08) & (1.79) \\
+
+ NoReligion & $-$1.57 & $-$0.92 & 0.56 & $-$0.30 & 1.02 & 1.31 \\ 
+  & (1.59) & (1.71) & (1.73) & (1.92) & (1.94) & (2.04) \\
+
+ OneParent\_House & $-$1.73 & $-$0.45 & $-$0.75 & $-$1.46 & $-$0.77 & $-$0.74 \\ 
+  & (1.36) & (1.59) & (1.49) & (1.69) & (1.57) & (1.47) \\
+
+ OtherLanguageHome & $-$0.44 & 5.92 & 9.98$^{**}$ & 11.24$^{**}$ & 9.00$^{*}$ & 9.84$^{**}$ \\ 
+  & (3.22) & (4.16) & (3.91) & (4.76) & (4.66) & (4.44) \\
+
+ PropertyOwned & $-$0.46 & $-$0.53 & 0.67 & $-$0.94 & $-$0.48 & 1.41 \\ 
+  & (1.37) & (1.50) & (1.43) & (1.76) & (1.67) & (1.50) \\
+
+ RentLoanPrice & $-$1.57 & $-$3.32$^{*}$ & $-$4.01$^{**}$ & $-$0.97 & $-$0.70 & $-$0.89 \\ 
+  & (1.49) & (1.76) & (1.67) & (2.07) & (2.07) & (2.16) \\
+
+ SocialServ & 2.51$^{*}$ & 1.65 & 2.47$^{*}$ & 2.53$^{*}$ & 2.35$^{*}$ & 4.45$^{***}$ \\ 
+  & (1.33) & (1.41) & (1.29) & (1.47) & (1.32) & (1.19) \\
+
+ Transformative & 3.24$^{**}$ & 4.73$^{***}$ & 4.84$^{***}$ & 4.46$^{**}$ & 3.56$^{**}$ & 4.58$^{***}$ \\ 
+  & (1.55) & (1.78) & (1.74) & (1.98) & (1.78) & (1.53) \\
+
+ Unemployment & $-$2.45$^{*}$ & $-$3.07$^{*}$ & 0.29 & 0.08 & 1.67 & 2.79$^{**}$ \\ 
+  & (1.40) & (1.63) & (1.51) & (1.76) & (1.51) & (1.37) \\
+
+ Constant & 50.81$^{***}$ & 52.60$^{***}$ & 47.31$^{***}$ & 49.93$^{***}$ & 53.51$^{***}$ & 50.49$^{***}$ \\ 
+  & (0.71) & (0.58) & (0.52) & (0.57) & (0.58) & (0.80) \\
+
 \hline \\[-1.8ex] 
 Observations & 150 & 150 & 150 & 150 & 150 & 150 \\ 
-Log Likelihood & $-$399.52 & $-$405.41 & $-$418.08 & $-$398.73 & $-$403.34 & $-$398.33 \\ 
-Akaike Inf. Crit. & 863.03 & 874.83 & 900.16 & 861.46 & 870.68 & 860.67 \\ 
-Bayesian Inf. Crit. & 951.96 & 963.76 & 989.10 & 950.40 & 959.61 & 949.60 \\ 
+Residual Standard Error (GLS) & 4.69 & 5.04 & 4.79 & 5.63 & 5.18 & 4.88 \\ 
 \hline 
@@
 \textit{Note:}  & \multicolumn{6}{l}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
- & \multicolumn{6}{l}{Estimated coefficients for variable named in column one shown for} \\ 
- & \multicolumn{6}{l}{election year indicated by column heading, with estimated standard} \\ 
- & \multicolumn{6}{l}{deviation for each coefficient shown below in parenthesis. Overall} \\ 
- & \multicolumn{6}{l}{summary measures for each regression equation are provided in the} \\ 
- & \multicolumn{6}{l}{bottom panel.} \\ 
 \end{tabular} 
@@
 
-<!--chapter:end:04-modelling.Rmd-->
+\newpage
 
 # Results
-```{r rhovis}
+
+## Spatial autoregressive parameter
+
+The spatial autoregressive coefficient $\rho$ is positive and significant in only the 2001 and 2016 elections (Figure \@ref(fig:rhovis)), meaning that in these elections, an electorate's political climate was affected by the attitudes of it's neighbours.
+Conversely, in the other four elections, the spatial effect weakens to become insignficant.
+In these years, it appears that the spatial component does not explain anything over and above the electoral socio-demographics, meaning electorates voted effectively independently.
+
+```{r rhovis, fig.cap="Estimates of the spatial autoregressive parameter for each of the six elections, with a 95\\% confidence interval.
+Only in 2001 and 2016 is there a significant spatial component", out.height='15%', out.extra = ''}
 rho_df <- bind_rows(glsmod16$rho_df %>% mutate(year = "2016"), 
@@
 rho_df %>% ggplot() + 
-  geom_point(aes(x = year, y = estimate, col = factor(p < 0.05)), size = 3) +
-  geom_linerange(aes(x = year, ymin = lower95, ymax = upper95, col = factor(p < 0.05)), size = 1.5) + 
+  geom_point(aes(x = year, y = estimate), size = 3) +
+  geom_linerange(aes(x = year, ymin = lower95, ymax = upper95), size = 1.2) + 
   geom_hline(aes(yintercept = 0), alpha = 0.5, size = 1) +
   scale_color_manual(values = c("grey50", "black")) +
-  guides(col = F)
+  labs(x = "Election year", y = "Estimate of spatial autoregressive parameter") +
+  lims(y = c(-1,1)) +
+  guides(col = F) +
+  coord_fixed(ratio = 1) +
+  theme_bw()
 ```
 
-The spatial autoregressive coefficient $\rho$ is significant in only the 2001 and 2016 elections, meaning that in these years an electorate's political climate was affected by the attitudes of it's neighbours.
-Conversely, in the other four elections, it appears that after accounting for electoral socio-demographics, there is no spatial effect, meaning electorates effectively acted independently.
+## Country-wide trend
+Since all socio-demographics have been standardized to have a mean of zero and a variance of one, the intercept in each model can be interpreted as the estimated two-party preferred vote for an electorate with mean characteristics[^3].
+The baseline of party preference has varied over the elections, with the biggest swing occuring in the 2007 election where the mean electorate shifted more than five percentage points in favour of the Labor party.
 
+[^3]: Mean of all variables aside from Judaism, Indigenous, Islam and Buddhism, where it assumes the mean of the log value.
+
+```{r plotintercept, fig.cap = "Estimated intercept for each election, which represents the two-party preferred vote for an electorate with mean characteristics.", out.height="15%", out.extra = ''}
+intercept_df <- bind_rows(data.frame(summary(glsmod16)$tTable)[1, ]  %>% mutate(year = "2016"), 
+  data.frame(summary(glsmod13)$tTable)[1, ] %>% mutate(year = "2013"),
+  data.frame(summary(glsmod10)$tTable)[1, ] %>% mutate(year = "2010"), 
+  data.frame(summary(glsmod07)$tTable)[1, ] %>% mutate(year = "2007"),
+  data.frame(summary(glsmod04)$tTable)[1, ] %>% mutate(year = "2004"), 
+  data.frame(summary(glsmod01)$tTable)[1, ] %>% mutate(year = "2001")) %>% 
+  rename(se = `Std.Error`, estimate = Value, p = `p.value`) %>% 
+  mutate(upper95 = estimate + 1.96*se, lower95 = estimate - 1.96*se)
+
+intercept_df %>% ggplot() + 
+  geom_point(aes(x = year, y = estimate), size = 3) +
+  geom_linerange(aes(x = year, ymin = lower95, ymax = upper95), size = 1.2) + 
+  geom_hline(aes(yintercept = 50), alpha = 0.5, size = 1) +
+  labs(x = "Election year", y = "Estimate of intercept") +
+  guides(col = F) +
+  coord_fixed(ratio = 0.2) +
+  lims(y = c(45, 55)) +
+  theme_bw()
+```
+
+## Influential socio-demographics
 To investigate the socio-demographics that have a strong effect on the two-party preferred vote, partial residual plots are used.
-These show the direction, size and significance of an estimated effect, as the slope of the prediction line matches the estimated coefficient, and the shaded region represents a 95% confidence band.
-Typically the Labor party campaigns on more progressive policies, which often include tax reform that adversely affects higher income earners, and more generous social assistance programs.
-Although topical in the lead up to an election, incomes only significantly affected electoral two-party preference in 2016 (see Figure \@ref(fig:plotincomes)), with higher income areas associated with support for the Liberal party.
-In the prior elections, the estimated effect is still positive, albeit insignificant.
-Unemployment on the other hand, was influential in 2001, 2004 and 2007 elections(Figure \@ref(fig:plotunemploy)), with electorates with higher unemployment favouring Labor.
-This effect has become weaker over time, with unemployment having an estimated effect of virtually zero in 2016.
+These show the direction, size and significance of an estimated effect - the slope of the prediction line matches the estimated coefficient, and the shaded region represents a 95% confidence band.
+If a horizontal line can be drawn through the confidence band, then the effect is insignificant.
+Plots for each election are faceted to compare the effects over time in Figure \@ref(fig:partresplot1) and Figure \@ref(fig:partresplot2).
+
+It is important here to note the ecological fallacy - insights are being drawn at the electorate level, and cannot be inferred for another disaggregate level (e.g. individual voters).
 
@@
 
-# Grid visreg
-library(gridExtra)
-library(grid)
-grid_visreg <- function(varname, xvar_title, xlimits, ylimits) {
-  p16 <- my_visreg(glsmod16, sp_weights_16, varname = varname, nolabs = T, year = "2016", xlimits, ylimits)
-  p13 <- my_visreg(glsmod13, sp_weights_13, varname = varname, nolabs = T, year = "2013", xlimits, ylimits)
-  p10 <- my_visreg(glsmod10, sp_weights_10, varname = varname, nolabs = T, year = "2010", xlimits, ylimits)
-  p07 <- my_visreg(glsmod07, sp_weights_07, varname = varname, nolabs = T, year = "2007", xlimits, ylimits)
-  p04 <- my_visreg(glsmod04, sp_weights_04, varname = varname, nolabs = T, year = "2004", xlimits, ylimits)
-  p01 <- my_visreg(glsmod01, sp_weights_01, varname = varname, nolabs = T, year = "2001", xlimits, ylimits)
+  # Points
+  if (plot == FALSE) {
+    return_ls <- list(bands = plot_df, points = points_df)
 
-  plots <- grid.arrange(p01, p04, p07, p10, p13, p16, nrow = 2, 
-    left = textGrob("Two-party preferred vote (%)", gp = gpar(cex = 0.8), rot = 90), 
-    bottom = textGrob(xvar_title, gp = gpar(cex = 0.8)))
-
-  return(plots)
+    return(return_ls)
   }
 
+}
+
+# Grid visreg
+grid_visreg <- function(varname, plot = TRUE, myscale = "free", top = FALSE) {
+
+    p16 <- my_visreg(glsmod16, sp_weights_16, varname = varname, plot = F, year = "2016")
+    p13 <- my_visreg(glsmod13, sp_weights_13, varname = varname, plot = F, year = "2013")
+    p10 <- my_visreg(glsmod10, sp_weights_10, varname = varname, plot = F, year = "2010")
+    p07 <- my_visreg(glsmod07, sp_weights_07, varname = varname, plot = F, year = "2007")
+    p04 <- my_visreg(glsmod04, sp_weights_04, varname = varname, plot = F, year = "2004")
+    p01 <- my_visreg(glsmod01, sp_weights_01, varname = varname, plot = F, year = "2001")
+
+    bands_df <- bind_rows(
+      p16$bands %>% mutate(year = "2016"),
+      p13$bands %>% mutate(year = "2013"),
+      p10$bands %>% mutate(year = "2010"),
+      p07$bands %>% mutate(year = "2007"),
+      p04$bands %>% mutate(year = "2004"),
+      p01$bands %>% mutate(year = "2001")
+    ) %>% mutate(varname = varname)
+
+    points_df <- bind_rows(
+      p16$points %>% mutate(year = "2016"),
+      p13$points %>% mutate(year = "2013"),
+      p10$points %>% mutate(year = "2010"),
+      p07$points %>% mutate(year = "2007"),
+      p04$points %>% mutate(year = "2004"),
+      p01$points %>% mutate(year = "2001")
+    ) %>% mutate(varname = varname)
+
+    if (plot == T) {
+      return_object <- ggplot(data = bands_df) +
+        geom_ribbon(aes(x = variable, ymin = lower95, ymax = upper95), fill = "grey80") + 
+        geom_point(aes(x = variable, y = part_res), data = points_df, size = 0.5, shape = 1, alpha = 0.5) +
+        geom_line(aes(x = variable, y = fitted), col = "blue", size = 1) +
+        theme_bw() + 
+        labs(x = "", y = "") + 
+        facet_grid(varname ~ year, scales = "free") +
+        theme(plot.margin=unit(c(b = 0.05, l = 0, t = -0.2, r = 0),"cm"))
+
+      if (top == FALSE) {
+        return_object <- return_object + theme(strip.text.x = element_blank()) 
+      }
+
+      if (myscale == "free") {
+        return_object <- return_object + lims(x = range(model_df %>% select(varname)), 
+          y = range(c(bands_df$upper95, bands_df$lower95, points_df$part_res)))
+      }
+
+      if (myscale == "free_x") {
+        return_object <- return_object + lims(x = range(model_df %>% select(varname)))
+      }
+
+      if (myscale == "free_y") {
+        return_object <- return_object + lims(y = range(c(bands_df$upper95, bands_df$lower95, points_df$part_res)))
+      }
+
+      if (myscale == "none") {
+        return_object <- return_object
+      }
+    }
+
+    if (plot == F) {
+      return_object <- list(bands = bands_df, points = points_df)
+    }
+
+    return(return_object)
+
+}
 ```
 
-```{r plotincomes, fig.cap = "Effect of Incomes"}
-p <- grid_visreg("Incomes", xvar_title = "Incomes", xlimits = c(-1.5, 3.6), ylimits = c(30, 95))
+### Income and unemployment
+
+Typically the Labor party campaigns on more progressive policies, which often include tax reform that adversely affects higher income earners, and more generous social assistance programs.
+Perhaps due to these policies, higher income electorates have been more likely to support the Liberal party, as the `Incomes` factor has a positive effect on Liberal preference (see row 1 in Figure \@ref(fig:partresplot1)).
+This effect is signficant in every election aside from 2004, in which it is only marginally insignificant ($p = 0.0613$).
+Unemployment however, has not been as influential.
+In 2001 and 2004, electorates with higher unemployment were more likely to support Labor, but over time this has shifted towards support for the Liberal party, culminating in a significantly positive effect in 2016.
+
+### Industry and type of work
+Electorates with higher proportions of workers in mining, gas, water, agriculture, waste and electricity (grouped as `Extractive` industries) are consistently linked with higher support for the Liberal party, with the mangitude of this effect slightly increasing over the years (see row 3 in Figure \@ref(fig:partresplot1)).
+This is unsurprising, as the Liberal party has close ties with these traditional energy industries, and typically present policies to reduce taxation on energy production.
+Furthermore, electorates with more workers in construction or manufacturing industries (`Transformative`) are also more likely to support the Liberal party (see row 4 in Figure \@ref(fig:partresplot1)).
+
+Similarly, workers in managerial, administrative, clerical and sales roles (`ManagerAdminClericalSales`) is also a significant predictor of two-party preference vote across all six elections, with a higher proportion of people working these jobs increasing Liberal support.
+The magnitude of this effect has also increased over the years.
+
+### Household mobility
+In each of the six elections, electorates with a higher proportion of people that have recently moved house (meaning in last five years) were more likely to support the Liberal party, although this effect was marginally insignificant in 2013 (see row 6 in Figure \@ref(fig:partresplot1).
+Having controlled for characteristics of house ownership and rental prices (via the factors `PropertyOwned` and `RentLoan` respectively), this is effect is somewhat surprising.
+
+### Relationships
+De facto relationships, but not marriages, are found to be an important (and significant) predictor of the two-party preferred vote in all six elections, with more de facto relationships associated with higher support for the Labor party.
+Marriages however, are insignficiant.
+
+### Age
+Older regions are often believed to be more conservative, and it can be seen that electorates with a higher median age have been more likely to support the Liberal party - although this effect is only significant in 2007 and 2010 (see row 2 in Figure \@ref(fig:partresplot2)).
+
+### Education
+Since 2007, electorates with higher education levels have been associated with supporting the Labor party, although this effect is only significant in 2016.
+Prior to 2007, education had an almost zero effect (see row 3 in \@ref(fig:partresplot1)).
+
+### Diversity
+Larger migrant populations from Asia, the Middle East, South-Eastern Europe, the United Kingdom and elsewhere, are either assocated with Labor support, or have no effect.
+Of these areas, only South-Eastern European populations were signficant in each election and Asian migrants were a significant in 2010.
+Speaking other language (aside from English) however, have a far stronger effect, as observed through `OtherLanguageHome`.
+Electorates with more diverse speech were more likely to support the Liberal party from 2004 onwards, with this effect being significant in 2007, 2010 and 2016.
+Furthermore, of the variables relating to religion, only Judaism shows a consistent effect, with electorates with larger Jewish populations more likely to vote Liberal.
+
+```{r prepplot, include = F}
+p1 <- grid_visreg("Incomes", top = T)
+p2 <- grid_visreg("Unemployment")
+p3 <- grid_visreg("Extractive")
+p4 <- grid_visreg("Transformative")
+p5 <- grid_visreg("ManagerAdminClericalSales")
+p6 <- grid_visreg("DiffAddress")
+p7 <- grid_visreg("DeFacto", top = T)
+p8 <- grid_visreg("MedianAge")
+p9 <- grid_visreg("Education")
+p10 <- grid_visreg("OtherLanguageHome")
+p11 <- grid_visreg("Born_SE_Europe")
+p12 <- grid_visreg("Born_Asia")
+p13 <- grid_visreg("Judaism")
 ```
 
-```{r plotunemploy, fig.cap = "Effect of Unemployment"}
-p <- grid_visreg("Unemployment", xvar_title = "Unemployment", xlimits = c(-2.1, 3.4), ylimits = c(33, 70))
+### A note on similar variables
+
+Many of the Census variables represent similar information, which is why factors were created and some variables were removed.
+However, there still remain some variables which are closely related.
+For example, electorate income levels (via `Incomes`) is likely to be related to electoral unemployment and labor force participation (via `Unemployment`).
+In 2001, the coefficient estiamte for `Unemployment` negative but not significant, and the `Incomes` variables  is significant.
+If the `Incomes` variable is removed from the model, `Unemployment` absorbs the negative effect, becoming significant ($p = 0.0056$).
+
+```{r removeunemp, eval = F}
+unemp01 <- my_fgls(full_formula, 
+  my_data = model_df %>% filter(year == "2001") %>% dplyr::select(-Incomes),
+  sp_weights = sp_weights_01)
+
+summary(unemp01)$tTable["Unemployment",]
 ```
 
-This section summarises the main findings made about the relationships between socio-demographic chracteristics and voting behaviour.
+## A closer look at the residuals
 
-```{r}
-coef_df %>% 
-  filter(variable != "Intercept") %>% 
-  mutate(upper95 = estimate + 1.96*se, lower95 = estimate - 1.96*se) %>% 
-  ggplot() +
-  geom_point(aes(x = year, y = estimate, col = factor(p < 0.05)), size = 2) +
-  geom_linerange(aes(x = year, ymin = lower95, ymax = upper95, col = factor(p < 0.05)), size = 1) + 
-  geom_hline(aes(yintercept = 0), alpha = 0.5, size = 1) + 
-  facet_wrap(~variable, scales = "free", ncol = 5) +
-  scale_color_manual(values = c("grey50", "black")) + 
-  theme(axis.text = element_text(size = 6), strip.text = element_text(size = 6)) +
-  labs(x = "", y = "Year") +
-  scale_x_continuous(breaks = c(2001, 2004, 2007, 2010, 2013, 2016)) +
-  guides(col = F)
+```{r resids, include = F}
+all_resids <- data.frame(
+  Residuals = c(glsmod16$actual_residuals, glsmod13$actual_residuals, glsmod10$actual_residuals, glsmod07$actual_residuals, glsmod04$actual_residuals, glsmod01$actual_residuals),
+  gls_residuals = c(glsmod16$residuals, glsmod13$residuals, glsmod10$residuals, glsmod07$residuals, glsmod04$residuals, glsmod01$residuals),
+  bind_rows(
+    glsmod16$my_data %>% left_join(tpp16 %>% dplyr::select(DivisionNm, UniqueID, StateAb), by = "DivisionNm"), 
+    glsmod13$my_data %>% left_join(tpp13 %>% dplyr::select(DivisionNm, UniqueID, StateAb), by = "DivisionNm"), 
+    glsmod10$my_data %>% left_join(tpp10 %>% dplyr::select(DivisionNm, UniqueID, StateAb), by = "DivisionNm"), 
+    glsmod07$my_data %>% left_join(tpp07 %>% dplyr::select(DivisionNm, UniqueID, StateAb), by = "DivisionNm"), 
+    glsmod04$my_data %>% left_join(tpp04 %>% dplyr::select(DivisionNm, UniqueID, StateAb), by = "DivisionNm"), 
+    glsmod01$my_data %>% left_join(tpp01 %>% dplyr::select(DivisionNm, UniqueID, StateAb), by = "DivisionNm"))
+) 
+#%>% dplyr::select(Residuals, DivisionNm, UniqueID, StateAb, year, LNP_Percent, gls_residuals)
 ```
 
-<!--chapter:end:05-results.Rmd-->
+### Residuals by state
 
-`r if (knitr:::is_html_output()) '
-# References {-}
-'`
+It is often hypothesized that states have systematic differences that cause their electorates to vote differently.
+Boxplots of residuals grouped by state reveal that only Tasmania, the Australian Capital Territory and the Northern Territory appear to have a state-specific effect that is not captured by the models.
+Tasmania and the Australian Capital Territory appear to have a bias towards Labor, whereas the Northern Territory has one towards Liberal.
+There are few electorates in these states (five, two and two respectively), so this might be due to incumbent effects rather than an actual state-specific bias.
 
-<!--chapter:end:06-references.Rmd-->
+```{r resstate, fig.cap = "Boxplot of residuals by state with jittered points.", fig.asp=0.6, out.width="80%", fig.pos = 'h', out.extra = ''}
+all_resids %>% 
+  ggplot(aes(x = StateAb, y = Residuals)) + 
+  geom_hline(aes(yintercept = 0), col = "grey50") +
+  geom_boxplot() + 
+  geom_jitter(alpha = 0.5, shape = 1) +
+  guides(col = F, shape = F, DivisionNm = F) +
+  theme_bw()
+```
+
+### Outlier electorates
+
+```{r readcooksd, eval = F}
+read_chunk('cooksd.R')
+```
+
+```{r getcooksd, eval = F}
+<<GetCookDistance>>
+```
+
+```{r cooksd, eval = F}
+cooksd_df <- data.frame(
+  Residuals = c(glsmod16$actual_residuals, glsmod13$actual_residuals, glsmod10$actual_residuals, glsmod07$actual_residuals, glsmod04$actual_residuals, glsmod01$actual_residuals),
+  gls_residuals = c(glsmod16$residuals, glsmod13$residuals, glsmod10$residuals, glsmod07$residuals, glsmod04$residuals, glsmod01$residuals),
+  cooksd = c(glsmod16$cooksd, glsmod13$cooksd, glsmod10$cooksd, glsmod07$cooksd, glsmod04$cooksd, glsmod01$cooksd),
+  bind_rows(
+    glsmod16$my_data %>% left_join(tpp16 %>% dplyr::select(DivisionNm, UniqueID, StateAb), by = "DivisionNm"), 
+    glsmod13$my_data %>% left_join(tpp13 %>% dplyr::select(DivisionNm, UniqueID, StateAb), by = "DivisionNm"), 
+    glsmod10$my_data %>% left_join(tpp10 %>% dplyr::select(DivisionNm, UniqueID, StateAb), by = "DivisionNm"), 
+    glsmod07$my_data %>% left_join(tpp07 %>% dplyr::select(DivisionNm, UniqueID, StateAb), by = "DivisionNm"), 
+    glsmod04$my_data %>% left_join(tpp04 %>% dplyr::select(DivisionNm, UniqueID, StateAb), by = "DivisionNm"), 
+    glsmod01$my_data %>% left_join(tpp01 %>% dplyr::select(DivisionNm, UniqueID, StateAb), by = "DivisionNm"))
+) 
+```
+
+```{r plotcooksd, fig.cap = "Electorates' Cook's Distance values in each election.", out.height = "15%", eval = F}
+cooksd_df %>% 
+  ggplot(aes(x = cooksd)) + 
+  geom_dotplot() + 
+  facet_wrap(~year, nrow = 1) + 
+  #coord_fixed(ratio = .5) +
+  labs(x = "Cook's Distance", y = "") + 
+  theme(axis.text.x = element_text(angle = 60, hjust = 1, size = 8), axis.text.y = element_text(size = 8))
+```
+
+```{r, eval = F}
+cooksd_df %>% 
+  filter(cooksd > 0.1) %>% 
+  select(DivisionNm, StateAb, year, cooksd, Residuals, LNP_Percent, everything()) %>% View
+```
+
+Based on the distribution of the cook's distance values, a cook's distance greater than $0.1$ is considered to be influential and a potential outlier.
+The electorate of Sydney (NSW) has a large cook's distance from 2001 to 2013, due to its diverse population (language, birthplace and religion), high number of defacto relationships, high income, high household mobility and small amount of workers in extractive and transformative jobs.
+It has remained a strong supporter of the Labor party and Liberal vote is severely overpredicted by the model, making it an outlier.
+Nearby in metropolitan NSW, the electorate of Wentworth is found to be an outlier in all but the 2007 election.
+Although historically Liberal, its two-party vote jumped by over 10 percentage points in 2010 without experiencing any notable changes in its socio-demographic makeup - implying that this may be the direct effect of its Liberal member, Malcolm Turnbull, becoming the leader of the Liberal party.
+Liberal support in Wentworth is underpredicted by the model in each year, and more so with Turnbull as Liberal leader.
+
+Lingiari, an electorate taking up almost all of the Northern Territory, is an outlier in the 2001-2007 elections due to its large Indigenous population, young age profile and low rates of property ownership.
+Fowler (NSW) has a diverse population with a high proportion of migrants, many Buddhists and Muslims, and has very strong Labor support, making it influential in 2001, 2004 and 2010.
+Other electorates with large cook's distance are Barton (NSW) and Leichhardt (QLD) in 2016, and Canberra (ACT) in 2007.
+
+```{r, eval = F}
+x <- model_df %>% filter(year == "2016") %>% select(-c(year, DivisionNm)) %>% as.matrix
+myhat <- hat(x)
+
+plot(glsmod07$cooksd, hat(model_df %>% filter(year == "2007") %>% select(-c(year, DivisionNm)) %>% as.matrix))
+```
+
+# Conclusion
+
+Across the past six elections, most of the socio-demographics that drive the electoral two-party preferred vote have remained steady, whilst a few (typically weaker) effects have varied over time.
+Industry and type of work are particularly influential, with energy-related and manufacturing/construction jobs, as well as administrative roles being strongly linked with the Liberal party in all elections.
+Incomes have had a similarly consistent effect, with higher income areas supporting Liberal.
+Higher levels of unemployment have shifted from weak association with Labor to a significant Liberal effect over the years, and higher education levels have been associated with Labor since 2007 (although only significant in 2016).
+It is also found that electorates with higher household mobility support Liberal, birthplace diversity favours Labor and more de facto relationships align with Labor preference - although marriages, family and household sizes have no material influence.
+Furthermore, the neighbourhood (spatial) effects were found to be positive in all elections, although only significant in 2001 and 2016, meaning that in the 2004-2013 elections, electorates effectively voted independently.
+
+The findings in this paper complement the existing literature by modelling temporal trends, which as far as the authors are aware, has not been done previously for Australian elections using a regression framework.
+It is also the first study to model any Australian election since 2010 using Census information.
+Additionally, a key contribution of this research is the spatio-temporal imputation, which is applied to obtain Census data for the 2004, 2007, 2010 and 2013 elections.
+All data sets used in this study have been contributed to the `eechidna` `R` package, which provides a rich, accessible data resource for future Australian electoral analysis.
+
+# Acknowledgements
+
+This paper was produced using `RMarkdown`[@rmarkdown] and `knitr` [@knitr].
+All corresponding code for this paper can be found in the github repository (https://github.com/jforbes14/journal-writing), and the data used is available in the `eechidna` package [@eechidna].
+All raw data was obtained from the Australian Electoral Commission, the Australian Bureau of Statistics and the Australian Government.
+
+# Software
+
+All election and Census data sets, along with electoral maps and more, are available in the `eechidna` `R` package, which can be downloaded from CRAN.
+`eechidna` (Exploring Election and Census Highly Informative Data Nationally for Australia) makes it easy to look at the data from the Australian Federal elections and Censuses that occurred between 2001 and 2016.
+This study contributed a large revision to the `eechidna` package, which included the addition of election and Census data for 2001-2010, vote outcomes for polling booths and imputed Census data for election years.
+For more details on using `eechidna`, please see the articles (vignettes) on the github page (https://ropenscilabs.github.io/eechidna/).
+
+The authors would like to give a sincere thanks to Anthony Ebert, Heike Hofmann, Thomas Lumley, Ben Marwick, Carson Sievert, Mingzhu Sun, Dilini Talagala, Nicholas Tierney, Nathaniel Tomasetti, Earo Wang and Fang Zhou, all of whom have contributed to the `eechidna` package.
+
+```{r partresplot1, fig.cap = "Partial residual plots for: income and unemployment, industry and type of work, and household mobility.", fig.height = 9,  fig.width = 6, warning = F, fig.pos = 'h'}
+grid.arrange(p1, p2, p3, p4, p5, p6, nrow = 6,
+  heights = c(1.2, 1, 1, 1, 1, 1),
+  left = grid.text("Two-party preferred vote (%)", gp = gpar(cex = 0.8), rot = 90))
+```
+
+```{r partresplot2, fig.cap = "Partial residual plots for: relationships and age, education and diversity.", fig.height = 9,  fig.width = 6, warning = F, fig.pos = 'h'}
+grid.arrange(p7, p8, p9, p10, p11, p12, p13, nrow = 7,
+  heights = c(1.2, 1, 1, 1, 1, 1, 1),
+  left = grid.text("Two-party preferred vote (%)", gp = gpar(cex = 0.8), rot = 90))
+```

### 2019-03-29 draft-paper.tex
new EBS working paper files

--- 2019-03-29 old_book_files/_book/electoral-modelling-draft.tex
@@
-\documentclass[openany]{book}
+\documentclass[11pt,a4paper,]{article}
 \usepackage{lmodern}
+
 \usepackage{amssymb,amsmath}
@@
 \else % if luatex or xelatex
-  \ifxetex
-    \usepackage{mathspec}
-  \else
-    \usepackage{fontspec}
-  \fi
+  \usepackage{unicode-math}
   \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
@@
 \IfFileExists{microtype.sty}{%
-\usepackage{microtype}
+\usepackage[]{microtype}
 \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
 }{}
-\usepackage[margin=1in]{geometry}
-\usepackage{hyperref}
-\hypersetup{unicode=true,
+\PassOptionsToPackage{hyphens}{url} % url is loaded by hyperref
+\usepackage[unicode=true]{hyperref}
+\hypersetup{
             pdftitle={Spatial modelling of the electoral two-party preferred vote in Australia. A study of federal elections between 2001 and 2016 via the eechidna R package.},
-            pdfauthor={Jeremy Forbes, Di Cook \& Rob Hyndman},
+            pdfkeywords={blah, blah},
             pdfborder={0 0 0},
@@
 \urlstyle{same}  % don't use monospace font for urls
-\usepackage{natbib}
-\bibliographystyle{apalike}
+\usepackage{geometry}
+\geometry{left=2.5cm,right=2.5cm,top=2.5cm,bottom=2.5cm}
+\usepackage[style=authoryear-comp,]{biblatex}
+\addbibresource{references.bib}
 \usepackage{longtable,booktabs}
-\usepackage{graphicx,grffile}
-\makeatletter
-\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
-\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
-\makeatother
-% Scale images if necessary, so that they will not overflow the page
-% margins by default, and it is still possible to overwrite the defaults
-% using explicit options in \includegraphics[width, height, ...]{}
-\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
+% Fix footnotes in tables (requires footnote package)
+\IfFileExists{footnote.sty}{\usepackage{footnote}\makesavenoteenv{long table}}{}
 \IfFileExists{parskip.sty}{%
@@
 \setcounter{secnumdepth}{5}
-% Redefines (sub)paragraphs to behave more like sections
-\ifx\paragraph\undefined\else
-\let\oldparagraph\paragraph
-\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
-\fi
-\ifx\subparagraph\undefined\else
-\let\oldsubparagraph\subparagraph
-\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
-\fi
 
-%%% Use protect on footnotes to avoid problems with footnotes in titles
-\let\rmarkdownfootnote\footnote%
-\def\footnote{\protect\rmarkdownfootnote}
+% set default figure placement to htbp
+\makeatletter
+\def\fps@figure{htbp}
+\makeatother
 
-%%% Change title format to be more compact
-\usepackage{titling}
+\title{Spatial modelling of the electoral two-party preferred vote in Australia. A study of federal elections between 2001 and 2016 via the \texttt{eechidna} \texttt{R} package.}
 
-% Create subtitle command for use in maketitle
-\newcommand{\subtitle}[1]{
-  \posttitle{
-    \begin{center}\large#1\end{center}
-    }
+%% MONASH STUFF
+
+%% CAPTIONS
+\RequirePackage{caption}
+\DeclareCaptionStyle{italic}[justification=centering]
+ {labelfont={bf},textfont={it},labelsep=colon}
+\captionsetup[figure]{style=italic,format=hang,singlelinecheck=true}
+\captionsetup[table]{style=italic,format=hang,singlelinecheck=true}
+
+%% FONT
+\RequirePackage{bera}
+\RequirePackage{mathpazo}
+
+%% HEADERS AND FOOTERS
+\RequirePackage{fancyhdr}
+\pagestyle{fancy}
+\rfoot{\Large\sffamily\raisebox{-0.1cm}{\textbf{\thepage}}}
+\makeatletter
+\lhead{\textsf{\expandafter{\@title}}}
+\makeatother
+\rhead{}
+\cfoot{}
+\setlength{\headheight}{15pt}
+\renewcommand{\headrulewidth}{0.4pt}
+\renewcommand{\footrulewidth}{0.4pt}
+\fancypagestyle{plain}{%
+\fancyhf{} % clear all header and footer fields
+\fancyfoot[C]{\sffamily\thepage} % except the center
+\renewcommand{\headrulewidth}{0pt}
+\renewcommand{\footrulewidth}{0pt}}
+
+%% MATHS
+\RequirePackage{bm,amsmath}
+\allowdisplaybreaks
+
+%% GRAPHICS
+\RequirePackage{graphicx}
+\setcounter{topnumber}{2}
+\setcounter{bottomnumber}{2}
+\setcounter{totalnumber}{4}
+\renewcommand{\topfraction}{0.85}
+\renewcommand{\bottomfraction}{0.85}
+\renewcommand{\textfraction}{0.15}
+\renewcommand{\floatpagefraction}{0.8}
+
+%\RequirePackage[section]{placeins}
+
+%% SECTION TITLES
+\RequirePackage[compact,sf,bf]{titlesec}
+\titleformat{\section}[block]
+  {\fontsize{15}{17}\bfseries\sffamily}
+  {\thesection}
+  {0.4em}{}
+\titleformat{\subsection}[block]
+  {\fontsize{12}{14}\bfseries\sffamily}
+  {\thesubsection}
+  {0.4em}{}
+\titlespacing{\section}{0pt}{*5}{*1}
+\titlespacing{\subsection}{0pt}{*2}{*0.2}
+
+%% TITLE PAGE
+\def\Date{\number\day}
+\def\Month{\ifcase\month\or
+ January\or February\or March\or April\or May\or June\or
+ July\or August\or September\or October\or November\or December\fi}
+\def\Year{\number\year}
+
+\makeatletter
+\def\wp#1{\gdef\@wp{#1}}\def\@wp{??/??}
+\def\jel#1{\gdef\@jel{#1}}\def\@jel{??}
+\def\showjel{{\large\textsf{\textbf{JEL classification:}}~\@jel}}
+\def\nojel{\def\showjel{}}
+\def\addresses#1{\gdef\@addresses{#1}}\def\@addresses{??}
+\def\cover{{\sffamily\setcounter{page}{0}
+        \thispagestyle{empty}
+        \placefig{2}{1.5}{width=5cm}{monash2}
+        \placefig{16.9}{1.5}{width=2.1cm}{MBusSchool}
+        \begin{textblock}{4}(16.9,4)ISSN 1440-771X\end{textblock}
+        \begin{textblock}{7}(12.7,27.9)\hfill
+        \includegraphics[height=0.7cm]{AACSB}~~~
+        \includegraphics[height=0.7cm]{EQUIS}~~~
+        \includegraphics[height=0.7cm]{AMBA}
+        \end{textblock}
+        \vspace*{2cm}
+        \begin{center}\Large
+        Department of Econometrics and Business Statistics\\[.5cm]
+        \footnotesize http://monash.edu/business/ebs/research/publications
+        \end{center}\vspace{2cm}
+        \begin{center}
+        \fbox{\parbox{14cm}{\begin{onehalfspace}\centering\Huge\vspace*{0.3cm}
+                \textsf{\textbf{\expandafter{\@title}}}\vspace{1cm}\par
+                \LARGE\@author\end{onehalfspace}
+        }}
+        \end{center}
+        \vfill
+                \begin{center}\Large
+                \Month~\Year\\[1cm]
+                Working Paper \@wp
+        \end{center}\vspace*{2cm}}}
+\def\pageone{{\sffamily\setstretch{1}%
+        \thispagestyle{empty}%
+        \vbox to \textheight{%
+        \raggedright\baselineskip=1.2cm
+     {\fontsize{24.88}{30}\sffamily\textbf{\expandafter{\@title}}}
+        \vspace{2cm}\par
+        \hspace{1cm}\parbox{14cm}{\sffamily\large\@addresses}\vspace{1cm}\vfill
+        \hspace{1cm}{\large\Date~\Month~\Year}\\[1cm]
+        \hspace{1cm}\showjel\vss}}}
+\def\blindtitle{{\sffamily
+     \thispagestyle{plain}\raggedright\baselineskip=1.2cm
+     {\fontsize{24.88}{30}\sffamily\textbf{\expandafter{\@title}}}\vspace{1cm}\par
+        }}
+\def\titlepage{{\cover\newpage\pageone\newpage\blindtitle}}
+
+\def\blind{\def\titlepage{{\blindtitle}}\let\maketitle\blindtitle}
+\def\titlepageonly{\def\titlepage{{\pageone\end{document}}}}
+\def\nocover{\def\titlepage{{\pageone\newpage\blindtitle}}\let\maketitle\titlepage}
+\let\maketitle\titlepage
+\makeatother
+
+%% SPACING
+\RequirePackage{setspace}
+\spacing{1.5}
+
+%% LINE AND PAGE BREAKING
+\sloppy
+\clubpenalty = 10000
+\widowpenalty = 10000
+\brokenpenalty = 10000
+\RequirePackage{microtype}
+
+%% PARAGRAPH BREAKS
+\setlength{\parskip}{1.4ex}
+\setlength{\parindent}{0em}
+
+%% HYPERLINKS
+\RequirePackage{xcolor} % Needed for links
+\definecolor{darkblue}{rgb}{0,0,.6}
+\RequirePackage{url}
+
+\makeatletter
+\@ifpackageloaded{hyperref}{}{\RequirePackage{hyperref}}
+\makeatother
+\hypersetup{
+     citecolor=0 0 0,
+     breaklinks=true,
+     bookmarksopen=true,
+     bookmarksnumbered=true,
+     linkcolor=darkblue,
+     urlcolor=blue,
+     citecolor=darkblue,
+     colorlinks=true}
+
+%% KEYWORDS
+\newenvironment{keywords}{\par\vspace{0.5cm}\noindent{\sffamily\textbf{Keywords:}}}{\vspace{0.25cm}\par\hrule\vspace{0.5cm}\par}
+
+%% ABSTRACT
+\renewenvironment{abstract}{\begin{minipage}{\textwidth}\parskip=1.4ex\noindent
+\hrule\vspace{0.1cm}\par{\sffamily\textbf{\abstractname}}\newline}
+  {\end{minipage}}
+
+\usepackage[T1]{fontenc}
+\usepackage[utf8]{inputenc}
+
+\usepackage[showonlyrefs]{mathtools}
+\usepackage[no-weekday]{eukdate}
+
+%% BIBLIOGRAPHY
+
+\makeatletter
+\@ifpackageloaded{biblatex}{}{\usepackage[style=authoryear-comp, backend=biber, natbib=true]{biblatex}}
+\makeatother
+\ExecuteBibliographyOptions{bibencoding=utf8,minnames=1,maxnames=3, maxbibnames=99,dashed=false,terseinits=true,giveninits=true,uniquename=false,uniquelist=false,doi=false, isbn=false,url=true,sortcites=false}
+
+\DeclareFieldFormat{url}{\texttt{\url{#1}}}
+\DeclareFieldFormat[article]{pages}{#1}
+\DeclareFieldFormat[inproceedings]{pages}{\lowercase{pp.}#1}
+\DeclareFieldFormat[incollection]{pages}{\lowercase{pp.}#1}
+\DeclareFieldFormat[article]{volume}{\mkbibbold{#1}}
+\DeclareFieldFormat[article]{number}{\mkbibparens{#1}}
+\DeclareFieldFormat[article]{title}{\MakeCapital{#1}}
+\DeclareFieldFormat[inproceedings]{title}{#1}
+\DeclareFieldFormat{shorthandwidth}{#1}
+% No dot before number of articles
+\usepackage{xpatch}
+\xpatchbibmacro{volume+number+eid}{\setunit*{\adddot}}{}{}{}
+% Remove In: for an article.
+\renewbibmacro{in:}{%
+  \ifentrytype{article}{}{%
+  \printtext{\bibstring{in}\intitlepunct}}}
+
+\makeatletter
+\DeclareDelimFormat[cbx@textcite]{nameyeardelim}{\addspace}
+\makeatother
+\renewcommand*{\finalnamedelim}{%
+  %\ifnumgreater{\value{liststop}}{2}{\finalandcomma}{}% there really should be no funny Oxford comma business here
+  \addspace\&\space}
+
+\wp{no/yr}
+\jel{C10,C14,C22}
+
+\RequirePackage[absolute,overlay]{textpos}
+\setlength{\TPHorizModule}{1cm}
+\setlength{\TPVertModule}{1cm}
+\def\placefig#1#2#3#4{\begin{textblock}{.1}(#1,#2)\rlap{\includegraphics[#3]{#4}}\end{textblock}}
+
+\author{Jeremy~Forbes, Di~Cook, Rob~Hyndman}
+\addresses{\textbf{Jeremy Forbes}\newline
+Monash University\newline Melbourne, Australia
+\newline{Email: \href{mailto:jeremyforbes1995@gmail.com}{\nolinkurl{jeremyforbes1995@gmail.com}}}\newline Corresponding author\\[1cm]
+\textbf{Di Cook}\newline
+Monash University\newline Melbourne, Australia
+\\[1cm]
+\textbf{Rob Hyndman}\newline
+Monash University\newline Melbourne, Australia
+\\[1cm]
 }
 
-\setlength{\droptitle}{-2em}
+\date{\sf\Date~\Month~\Year}
+\makeatletter
+ \lfoot{\sf Forbes, Cook, Hyndman: \@date}
+\makeatother
 
-  \title{Spatial modelling of the electoral two-party preferred vote in Australia. A study of federal elections between 2001 and 2016 via the \texttt{eechidna} \texttt{R} package.}
-    \pretitle{\vspace{\droptitle}\centering\huge}
-  \posttitle{\par}
-    \author{Jeremy Forbes, Di Cook \& Rob Hyndman}
-    \preauthor{\centering\large\emph}
-  \postauthor{\par}
-      \predate{\centering\large\emph}
-  \postdate{\par}
-    \date{2019-03-29}
-
-\usepackage{booktabs}
-\usepackage{amsthm}
-\makeatletter
-\def\thm@space@setup{%
-  \thm@preskip=8pt plus 2pt minus 4pt
-  \thm@postskip=\thm@preskip
+%% Any special functions or other packages can be loaded here.
+\usepackage{longtable}
+\usepackage{float}
+\let\origfigure\figure
+\let\endorigfigure\endfigure
+\renewenvironment{figure}[1][2] {
+    \expandafter\origfigure\expandafter[H]
+} {
+    \endorigfigure
 }
-\makeatother
-\usepackage{longtable}
 
@@
 \maketitle
+\begin{abstract}
+A brief summary of our ideas
+\end{abstract}
+\begin{keywords}
+blah, blah
+\end{keywords}
 
-{
-\setcounter{tocdepth}{1}
-\tableofcontents
-}
 \hypertarget{intro}{%
-\chapter{Introduction}\label{intro}}
+\section{Introduction}\label{intro}}
 
@@
 The first problem addressed is that of obtaining appropriate Census data for the 2004, 2007, 2010 and 2013 elections - election years in which a Census does not occur.
-The predominant approach in previous studies is to join voting outcomes to the nearest Census, without accounting for any temporal differences (see \citet{DavisStimson98}, \citet{Stimson06}, \citet{Liao09} and \citet{Stimson09}).
+The predominant approach in previous studies is to join voting outcomes to the nearest Census, without accounting for any temporal differences (see \textcite{DavisStimson98}, \textcite{Stimson06}, \textcite{Liao09} and \textcite{Stimson09}).
 Furthermore, electoral boundaries change regularly, so spatial discrepancies also arise when matching electoral data.
-To obtain appropriate Census data for these four elections, electoral socio-demographics are imputed using a spatio-temporal imputation that combines areal interpolation \citep{Goodchild1993} and linear time-interpolation.
+To obtain appropriate Census data for these four elections, electoral socio-demographics are imputed using a spatio-temporal imputation that combines areal interpolation \autocite{Goodchild1993} and linear time-interpolation.
 Collecting and wrangling the raw data, along with the imputation process, are detailed in section \ref{data}.
-All data and associated documentation relating to this procedure are available in the \texttt{eechidna} \texttt{R} package \citep{eechidna}, providing a resource for future analysis .
+All data and associated documentation relating to this procedure are available in the \texttt{eechidna} \texttt{R} package \autocite{eechidna}, providing a resource for future analysis .
 
 Previous work on modelling Australian federal elections have found that aggregate socio-demographics are relatively good predictors of voting outcomes.
-\citet{Forrest01} does this using multiple regression of the Liberal and Labor primary vote for polling booths in the Farrer electorate in 1998.
-\citet{Stimson06}, \citet{Stimson09} and \citet{Stimson12} use principal component analysis of polling booths in the 2001, 2004 and 2007 elections respectively, also finding that socio-demographic characteristics of polling booths are linked to their two-party preferred vote.
-On the contrary, \citet{Stimson09} models the polling booth swing vote (change in the two-party preferred vote) in the 2007 election, finding that little of swing vote can be explained by Census data.
+\textcite{Forrest01} does this using multiple regression of the Liberal and Labor primary vote for polling booths in the Farrer electorate in 1998.
+\textcite{Stimson06}, \textcite{Stimson09} and \textcite{Stimson12} use principal component analysis of polling booths in the 2001, 2004 and 2007 elections respectively, also finding that socio-demographic characteristics of polling booths are linked to their two-party preferred vote.
+On the contrary, \textcite{Stimson09} models the polling booth swing vote (change in the two-party preferred vote) in the 2007 election, finding that little of swing vote can be explained by Census data.
 Instead of analyzing a single election in isolation, this paper employs a consistent model framework across six elections so that temporal changes in the effects of socio-demographics can be observed, where each federal elections is modelled with a cross-sectional data set.
@@
 \hypertarget{data}{%
-\chapter{Data collection, wrangling and imputation}\label{data}}
+\section{Data collection, wrangling and imputation}\label{data}}
 
 \hypertarget{collecting-the-data}{%
-\section{Collecting the data}\label{collecting-the-data}}
+\subsection{Collecting the data}\label{collecting-the-data}}
 
@@
 The format of these files change over the years, making extracting the appropriate information a big task.
-The functions available in the \texttt{dplyr} \citep{dplyr} and \texttt{readxl} \citep{readxl} \texttt{R} packages are very useful, as they provide fast consistent tools for data manipulation and functions to import .xlsx files (respectively).
+The functions available in the \texttt{dplyr} \autocite{dplyr} and \texttt{readxl} \autocite{readxl} \texttt{R} packages are very useful, as they provide fast consistent tools for data manipulation and functions to import .xlsx files (respectively).
 The 2001 and 2006 Census data are however published in a format where each electorate has a separate document, making it difficult to use the \texttt{dplyr} tools and instead cells have to be selected from each individual file to construct the desired variables.
-All scripts required for the data wrangling process can be found in the github repository for the \texttt{eechidna} \texttt{R} package \citep{eechidna}, along with the raw data.
+All scripts required for the data wrangling process can be found in the github repository for the \texttt{eechidna} \texttt{R} package \autocite{eechidna}, along with the raw data.
 The \texttt{eechidna} package makes this study entirely reproducible and provides a resource to help wrangle data for future Censuses and elections, when they become available.
@@
 \hypertarget{joining-census-and-election-data}{%
-\section{Joining Census and election data}\label{joining-census-and-election-data}}
+\subsection{Joining Census and election data}\label{joining-census-and-election-data}}
 
 \hypertarget{differences-between-census-and-election-data}{%
-\subsection{Differences between Census and election data}\label{differences-between-census-and-election-data}}
+\subsubsection{Differences between Census and election data}\label{differences-between-census-and-election-data}}
 
@@
 
-{\centering \includegraphics[width=0.9\linewidth]{electoral-modelling-draft_files/figure-latex/timeline-1}
+{\centering \includegraphics[width=0.9\linewidth]{skeleton_files/figure-latex/timeline-1}
 
@@
 \hypertarget{spatio-temporal-imputation}{%
-\subsection{Spatio-temporal imputation}\label{spatio-temporal-imputation}}
+\subsubsection{Spatio-temporal imputation}\label{spatio-temporal-imputation}}
 
-To account for spatial differences, the piece-wise approximation method in \citet{Goodchild1993} is adopted.
+To account for spatial differences, the piece-wise approximation method in \textcite{Goodchild1993} is adopted.
 Consider a map of source zones \(s = 1,...,S\), for which socio-demographic information is available, and a set of target zones \(t = 1,...,T\) for which information is to be imputed.
@@
 \hypertarget{applied}{%
-\subsection{Applied}\label{applied}}
+\subsubsection{Applied}\label{applied}}
 
@@
 
-{\centering \includegraphics[height=0.3\textheight]{electoral-modelling-draft_files/figure-latex/hume13-1}
+{\centering \includegraphics[height=0.3\textheight]{skeleton_files/figure-latex/hume13-1}
 
@@
 
-{\centering \includegraphics[height=0.3\textheight]{electoral-modelling-draft_files/figure-latex/hume16-1}
+{\centering \includegraphics[height=0.3\textheight]{skeleton_files/figure-latex/hume16-1}
 
@@
 \hypertarget{modelling}{%
-\chapter{Modelling}\label{modelling}}
+\section{Modelling}\label{modelling}}
 
 \hypertarget{data-pre-processing}{%
-\section{Data pre-processing}\label{data-pre-processing}}
+\subsection{Data pre-processing}\label{data-pre-processing}}
 
@@
 \hypertarget{dimension-reduction}{%
-\subsection{Dimension reduction}\label{dimension-reduction}}
+\subsubsection{Dimension reduction}\label{dimension-reduction}}
 
@@
 \hypertarget{model-framework}{%
-\section{Model framework}\label{model-framework}}
+\subsection{Model framework}\label{model-framework}}
 
@@
 
-The spatial error model \citep{Anselin88} is chosen because captures spatial heterogeneity by incorporating a spatially structured random effect vector \citep{LeSage2009}.
+The spatial error model \autocite{Anselin88} is chosen because captures spatial heterogeneity by incorporating a spatially structured random effect vector \autocite{LeSage2009}.
 In this context, the random effect can be thought of as capturing the unobserved political climate in each electorate, where the climate is correlated with the climate in neighbouring electorates.
@@
 
-Table 3.1 details the resultant estimated model coefficients and their estimated standard errors for each of the six elections. These are interpreted in the next section.
+Table 3 details the resultant estimated model coefficients and their estimated standard errors for each of the six elections. These are interpreted in the next section.
 
@@
  & 2001 & 2004 & 2007 & 2010 & 2013 & 2016 \\ 
-\\[-1.8ex] & (1) & (2) & (3) & (4) & (5) & (6)\\ 
 \hline \\[-1.8ex] 
@@
   & (0.15) & (0.17) & (0.17) & (0.16) & (0.16) & (0.17) \\
-  & & & & & & \\ 
+
  AusCitizen & $-$3.13 & $-$2.64 & $-$2.53 & $-$0.08 & $-$3.40 & $-$1.80 \\ 
   & (2.26) & (2.43) & (2.34) & (2.79) & (2.76) & (2.71) \\
-  & & & & & & \\ 
+
  Born\_Asia & 2.22 & $-$0.95 & $-$1.60 & $-$6.83$^{**}$ & $-$3.03 & $-$0.55 \\ 
   & (2.18) & (2.44) & (2.19) & (2.73) & (2.71) & (2.17) \\
-  & & & & & & \\ 
+
  Born\_MidEast & $-$1.15 & $-$1.59 & $-$2.01$^{*}$ & $-$2.03 & $-$0.92 & $-$1.44 \\ 
   & (1.07) & (1.20) & (1.11) & (1.27) & (1.24) & (1.13) \\
-  & & & & & & \\ 
+
  Born\_SE\_Europe & $-$3.21$^{**}$ & $-$4.24$^{***}$ & $-$3.61$^{***}$ & $-$4.14$^{***}$ & $-$3.69$^{***}$ & $-$2.72$^{***}$ \\ 
   & (1.42) & (1.46) & (1.02) & (1.19) & (1.07) & (0.97) \\
-  & & & & & & \\ 
+
  Born\_UK & 0.25 & $-$0.07 & 0.34 & 0.56 & $-$0.09 & $-$1.32 \\ 
   & (1.00) & (0.98) & (0.90) & (1.07) & (1.04) & (1.04) \\
-  & & & & & & \\ 
+
  BornElsewhere & $-$5.04 & $-$4.91 & $-$4.13 & 2.35 & $-$5.23 & $-$4.14 \\ 
   & (3.30) & (3.68) & (3.38) & (4.23) & (4.15) & (3.97) \\
-  & & & & & & \\ 
+
  Buddhism & $-$0.49 & $-$0.17 & $-$1.37 & $-$0.83 & $-$0.12 & $-$1.60 \\ 
   & (1.39) & (1.61) & (1.61) & (1.80) & (1.68) & (1.56) \\
-  & & & & & & \\ 
+
  Christianity & $-$2.48 & $-$1.23 & 0.38 & 0.50 & 2.41 & 1.68 \\ 
   & (1.73) & (1.85) & (1.83) & (1.99) & (1.85) & (1.78) \\
-  & & & & & & \\ 
+
  CurrentlyStudying & $-$2.19$^{**}$ & $-$0.13 & 2.06$^{*}$ & 2.12$^{*}$ & 1.15 & $-$0.16 \\ 
   & (0.99) & (1.13) & (1.17) & (1.25) & (1.26) & (1.18) \\
-  & & & & & & \\ 
+
  DeFacto & $-$6.44$^{***}$ & $-$5.37$^{**}$ & $-$6.43$^{***}$ & $-$8.07$^{***}$ & $-$6.56$^{**}$ & $-$8.53$^{***}$ \\ 
   & (1.87) & (2.48) & (2.31) & (3.06) & (3.11) & (2.83) \\
-  & & & & & & \\ 
+
  DiffAddress & 3.88$^{***}$ & 5.06$^{***}$ & 4.22$^{***}$ & 5.57$^{***}$ & 3.53$^{*}$ & 5.67$^{***}$ \\ 
   & (0.94) & (1.12) & (0.99) & (1.76) & (1.91) & (1.60) \\
-  & & & & & & \\ 
+
  Distributive & 1.27 & 2.01$^{*}$ & 1.36 & 1.57 & 2.10$^{*}$ & 1.20 \\ 
   & (1.12) & (1.21) & (1.13) & (1.34) & (1.27) & (1.21) \\
-  & & & & & & \\ 
+
  Education & 1.08 & 0.52 & $-$5.52$^{*}$ & $-$4.08 & $-$4.44 & $-$7.07$^{**}$ \\ 
   & (2.38) & (3.12) & (3.27) & (3.95) & (3.78) & (3.55) \\
-  & & & & & & \\ 
+
  Extractive & 4.83$^{***}$ & 5.45$^{***}$ & 5.37$^{***}$ & 7.31$^{***}$ & 6.71$^{***}$ & 7.43$^{***}$ \\ 
   & (1.48) & (1.42) & (1.36) & (1.56) & (1.47) & (1.39) \\
-  & & & & & & \\ 
+
  FamHouseSize & $-$0.16 & 0.87 & $-$2.40 & $-$2.53 & $-$3.26 & $-$2.91 \\ 
   & (2.19) & (2.72) & (2.69) & (3.25) & (3.28) & (2.90) \\
-  & & & & & & \\ 
+
  Incomes & 4.36$^{**}$ & 5.03$^{*}$ & 9.45$^{***}$ & 7.09$^{**}$ & 7.97$^{***}$ & 12.20$^{***}$ \\ 
   & (1.77) & (2.66) & (2.75) & (3.25) & (2.92) & (2.75) \\
-  & & & & & & \\ 
+
  Indigenous & 2.91$^{*}$ & 1.97 & 2.48 & 2.84 & 0.67 & $-$0.05 \\ 
   & (1.68) & (1.95) & (1.75) & (2.16) & (2.14) & (2.00) \\
-  & & & & & & \\ 
+
  Islam & $-$0.92 & $-$0.97 & $-$0.54 & $-$2.50 & $-$0.82 & $-$0.95 \\ 
   & (1.22) & (1.36) & (1.27) & (1.52) & (1.42) & (1.34) \\
-  & & & & & & \\ 
+
  Judaism & 1.88$^{*}$ & 1.78 & 2.66$^{***}$ & 1.97$^{*}$ & 2.74$^{**}$ & 1.65$^{*}$ \\ 
   & (1.05) & (1.13) & (1.01) & (1.15) & (1.10) & (1.00) \\
-  & & & & & & \\ 
+
  ManagerAdminClericalSales & 2.06$^{***}$ & 3.32$^{***}$ & 6.00$^{***}$ & 5.47$^{***}$ & 5.04$^{***}$ & 5.78$^{***}$ \\ 
   & (0.71) & (0.93) & (0.90) & (1.08) & (1.03) & (1.06) \\
-  & & & & & & \\ 
+
  Married & 0.44 & 0.11 & $-$1.22 & $-$0.22 & 0.91 & $-$2.34 \\ 
   & (2.31) & (2.96) & (2.83) & (3.15) & (3.03) & (2.81) \\
-  & & & & & & \\ 
+
  MedianAge & 2.32$^{*}$ & 4.96$^{***}$ & 3.66$^{**}$ & 4.00$^{*}$ & 2.30 & 2.87 \\ 
   & (1.32) & (1.65) & (1.81) & (2.26) & (2.08) & (1.79) \\
-  & & & & & & \\ 
+
  NoReligion & $-$1.57 & $-$0.92 & 0.56 & $-$0.30 & 1.02 & 1.31 \\ 
   & (1.59) & (1.71) & (1.73) & (1.92) & (1.94) & (2.04) \\
-  & & & & & & \\ 
+
  OneParent\_House & $-$1.73 & $-$0.45 & $-$0.75 & $-$1.46 & $-$0.77 & $-$0.74 \\ 
   & (1.36) & (1.59) & (1.49) & (1.69) & (1.57) & (1.47) \\
-  & & & & & & \\ 
+
  OtherLanguageHome & $-$0.44 & 5.92 & 9.98$^{**}$ & 11.24$^{**}$ & 9.00$^{*}$ & 9.84$^{**}$ \\ 
   & (3.22) & (4.16) & (3.91) & (4.76) & (4.66) & (4.44) \\
-  & & & & & & \\ 
+
  PropertyOwned & $-$0.46 & $-$0.53 & 0.67 & $-$0.94 & $-$0.48 & 1.41 \\ 
   & (1.37) & (1.50) & (1.43) & (1.76) & (1.67) & (1.50) \\
-  & & & & & & \\ 
+
  RentLoanPrice & $-$1.57 & $-$3.32$^{*}$ & $-$4.01$^{**}$ & $-$0.97 & $-$0.70 & $-$0.89 \\ 
   & (1.49) & (1.76) & (1.67) & (2.07) & (2.07) & (2.16) \\
-  & & & & & & \\ 
+
  SocialServ & 2.51$^{*}$ & 1.65 & 2.47$^{*}$ & 2.53$^{*}$ & 2.35$^{*}$ & 4.45$^{***}$ \\ 
   & (1.33) & (1.41) & (1.29) & (1.47) & (1.32) & (1.19) \\
-  & & & & & & \\ 
+
  Transformative & 3.24$^{**}$ & 4.73$^{***}$ & 4.84$^{***}$ & 4.46$^{**}$ & 3.56$^{**}$ & 4.58$^{***}$ \\ 
   & (1.55) & (1.78) & (1.74) & (1.98) & (1.78) & (1.53) \\
-  & & & & & & \\ 
+
  Unemployment & $-$2.45$^{*}$ & $-$3.07$^{*}$ & 0.29 & 0.08 & 1.67 & 2.79$^{**}$ \\ 
   & (1.40) & (1.63) & (1.51) & (1.76) & (1.51) & (1.37) \\
-  & & & & & & \\ 
+
  Constant & 50.81$^{***}$ & 52.60$^{***}$ & 47.31$^{***}$ & 49.93$^{***}$ & 53.51$^{***}$ & 50.49$^{***}$ \\ 
   & (0.71) & (0.58) & (0.52) & (0.57) & (0.58) & (0.80) \\
-  & & & & & & \\ 
+
 \hline \\[-1.8ex] 
@@
 Residual Standard Error (GLS) & 4.69 & 5.04 & 4.79 & 5.63 & 5.18 & 4.88 \\ 
-Log Likelihood & $-$402.11 & $-$410.52 & $-$404.94 & $-$423.96 & $-$412.62 & $-$404.34 \\ 
-Akaike Inf. Crit. & 868.21 & 885.05 & 873.88 & 911.92 & 889.24 & 872.68 \\ 
-Bayesian Inf. Crit. & 957.14 & 973.98 & 962.81 & 1,000.85 & 978.17 & 961.62 \\ 
 \hline 
@@
 
+\newpage
+
 \hypertarget{results}{%
-\chapter{Results}\label{results}}
+\section{Results}\label{results}}
 
 \hypertarget{spatial-autoregressive-parameter}{%
-\section{Spatial autoregressive parameter}\label{spatial-autoregressive-parameter}}
+\subsection{Spatial autoregressive parameter}\label{spatial-autoregressive-parameter}}
 
@@
 
-{\centering \includegraphics[height=0.2\textheight,]{electoral-modelling-draft_files/figure-latex/rhovis-1}
+{\centering \includegraphics[height=0.15\textheight,]{skeleton_files/figure-latex/rhovis-1}
 
@@
 \hypertarget{country-wide-trend}{%
-\section{Country-wide trend}\label{country-wide-trend}}
+\subsection{Country-wide trend}\label{country-wide-trend}}
 
@@
 
-{\centering \includegraphics[height=0.2\textheight,]{electoral-modelling-draft_files/figure-latex/plotintercept-1}
+{\centering \includegraphics[height=0.15\textheight,]{skeleton_files/figure-latex/plotintercept-1}
 
@@
 \hypertarget{influential-socio-demographics}{%
-\section{Influential socio-demographics}\label{influential-socio-demographics}}
+\subsection{Influential socio-demographics}\label{influential-socio-demographics}}
 
@@
 \hypertarget{income-and-unemployment}{%
-\subsection{Income and unemployment}\label{income-and-unemployment}}
+\subsubsection{Income and unemployment}\label{income-and-unemployment}}
 
@@
 \hypertarget{industry-and-type-of-work}{%
-\subsection{Industry and type of work}\label{industry-and-type-of-work}}
+\subsubsection{Industry and type of work}\label{industry-and-type-of-work}}
 
@@
 \hypertarget{household-mobility}{%
-\subsection{Household mobility}\label{household-mobility}}
+\subsubsection{Household mobility}\label{household-mobility}}
 
@@
 \hypertarget{relationships}{%
-\subsection{Relationships}\label{relationships}}
+\subsubsection{Relationships}\label{relationships}}
 
@@
 \hypertarget{age}{%
-\subsection{Age}\label{age}}
+\subsubsection{Age}\label{age}}
 
@@
 \hypertarget{education}{%
-\subsection{Education}\label{education}}
+\subsubsection{Education}\label{education}}
 
@@
 \hypertarget{diversity}{%
-\subsection{Diversity}\label{diversity}}
+\subsubsection{Diversity}\label{diversity}}
 
@@
 
-\hypertarget{partial-residual-plots}{%
-\subsection{Partial residual plots}\label{partial-residual-plots}}
-
-\begin{figure}[h]
-
-{\centering \includegraphics{electoral-modelling-draft_files/figure-latex/partresplot1-1}
-
-}
-
-\caption{Partial residual plots for: income and unemployment, industry and type of work, and household mobility.}\label{fig:partresplot1}
-\end{figure}
-
-\begin{figure}[h]
-
-{\centering \includegraphics{electoral-modelling-draft_files/figure-latex/partresplot2-1}
-
-}
-
-\caption{Partial residual plots for: relationships and age, education and diversity.}\label{fig:partresplot2}
-\end{figure}
-
 \hypertarget{a-note-on-similar-variables}{%
-\subsection{A note on similar variables}\label{a-note-on-similar-variables}}
+\subsubsection{A note on similar variables}\label{a-note-on-similar-variables}}
 
@@
 \hypertarget{a-closer-look-at-the-residuals}{%
-\section{A closer look at the residuals}\label{a-closer-look-at-the-residuals}}
+\subsection{A closer look at the residuals}\label{a-closer-look-at-the-residuals}}
 
 \hypertarget{residuals-by-state}{%
-\subsection{Residuals by state}\label{residuals-by-state}}
+\subsubsection{Residuals by state}\label{residuals-by-state}}
 
@@
 
-{\centering \includegraphics[width=0.8\linewidth,]{electoral-modelling-draft_files/figure-latex/resstate-1}
+{\centering \includegraphics[width=0.8\linewidth,]{skeleton_files/figure-latex/resstate-1}
 
@@
 \hypertarget{outlier-electorates}{%
-\subsection{Outlier electorates}\label{outlier-electorates}}
+\subsubsection{Outlier electorates}\label{outlier-electorates}}
 
@@
 \hypertarget{conclusion}{%
-\chapter{Conclusion}\label{conclusion}}
+\section{Conclusion}\label{conclusion}}
 
@@
 \hypertarget{acknowledgements}{%
-\chapter{Acknowledgements}\label{acknowledgements}}
+\section{Acknowledgements}\label{acknowledgements}}
 
-This paper was produced using \texttt{RMarkdown}\citep{rmarkdown} and \texttt{knitr} \citep{knitr}.
-All corresponding code for this paper can be found in the github repository (\url{https://github.com/jforbes14/journal-writing}), and the data used is available in the \texttt{eechidna} package \citep{eechidna}.
+This paper was produced using \texttt{RMarkdown}\autocite{rmarkdown} and \texttt{knitr} \autocite{knitr}.
+All corresponding code for this paper can be found in the github repository (\url{https://github.com/jforbes14/journal-writing}), and the data used is available in the \texttt{eechidna} package \autocite{eechidna}.
 All raw data was obtained from the Australian Electoral Commission, the Australian Bureau of Statistics and the Australian Government.
@@
 \hypertarget{software}{%
-\chapter{Software}\label{software}}
+\section{Software}\label{software}}
 
@@
 
-\bibliography{book.bib,packages.bib}
+\begin{figure}[h]
+
+{\centering \includegraphics{skeleton_files/figure-latex/partresplot1-1}
+
+}
+
+\caption{Partial residual plots for: income and unemployment, industry and type of work, and household mobility.}\label{fig:partresplot1}
+\end{figure}
+
+\begin{figure}[h]
+
+{\centering \includegraphics{skeleton_files/figure-latex/partresplot2-1}
+
+}
+
+\caption{Partial residual plots for: relationships and age, education and diversity.}\label{fig:partresplot2}
+\end{figure}
+
+\printbibliography
 

### 2019-03-29 draft-paper.Rmd
spellcheck

@@
 
-The cross-sectional data set for each election consists of the two-party preferred vote (response variable), and socio-demographic variables (explanatory variables) that characterise each electorate.
+The cross-sectional data set for each election consists of the two-party preferred vote (response variable), and socio-demographic variables (explanatory variables) that characterize each electorate.
 To obtain these cross-sections, socio-demographic variables are first standardized, and then principal components are used to group variables into "factors".
@@
 
-Publically available Census data is aggregated and there are different resolutions accessible, ranging from SA1 (over 50,000 zones) to electoral divisions (150 zones).
+Publicly available Census data is aggregated and there are different resolutions accessible, ranging from SA1 (over 50,000 zones) to electoral divisions (150 zones).
 For this study electoral divisions are used as source zones, and this imputation method is applied for each of the 2004, 2007, 2010 and 2013 elections.
@@
 
-Electorate (2016) | Percentage | Population in Electorate | Estimated Population Allocated to Purple Region: $\hat{P}_{s,t}$
----------------------|------------|--------------------------|------------------------------------------------
+\footnotesize
+
+Electorate (2016) | Percentage | Population | Estimated Population Allocated to Purple Region: $\hat{P}_{s,t}$
+--------------------|----------|-----------|---------------------------------------------
 HUME              | 96.54%     | 150643                   | 145427
@@
 
+\normalsize
+
 Now consider the socio-demographic $AusCitizen$ - the proportion of people in the region who are Australian citizens.
@@
 
+\footnotesize
+
 |DivisionNm      | AusCitizen (%): $D_s$| Estimated Population Allocated to Purple Region: $\hat{P}_{s,t}$|
-|----------------|-------------------|------------------------------------------------|
+|--------------------------|-------------------|------------------------------------------------|
 |HUME            |              90.02|                                          145427|
@@
 |CALARE          |              87.56|                                              21|
+
+\normalsize
 
@@
 The spatial autoregressive coefficient $\rho$ is positive and significant in only the 2001 and 2016 elections (Figure \@ref(fig:rhovis)), meaning that in these elections, an electorate's political climate was affected by the attitudes of it's neighbours.
-Conversely, in the other four elections, the spatial effect weakens to become insignficant.
+Conversely, in the other four elections, the spatial effect weakens to become insignificant.
 In these years, it appears that the spatial component does not explain anything over and above the electoral socio-demographics, meaning electorates voted effectively independently.
@@
 Since all socio-demographics have been standardized to have a mean of zero and a variance of one, the intercept in each model can be interpreted as the estimated two-party preferred vote for an electorate with mean characteristics[^3].
-The baseline of party preference has varied over the elections, with the biggest swing occuring in the 2007 election where the mean electorate shifted more than five percentage points in favour of the Labor party.
+The baseline of party preference has varied over the elections, with the biggest swing occurring in the 2007 election where the mean electorate shifted more than five percentage points in favour of the Labor party.
 
@@
 Perhaps due to these policies, higher income electorates have been more likely to support the Liberal party, as the `Incomes` factor has a positive effect on Liberal preference (see row 1 in Figure \@ref(fig:partresplot1)).
-This effect is signficant in every election aside from 2004, in which it is only marginally insignificant ($p = 0.0613$).
+This effect is significant in every election aside from 2004, in which it is only marginally insignificant ($p = 0.0613$).
 Unemployment however, has not been as influential.
@@
 ### Industry and type of work
-Electorates with higher proportions of workers in mining, gas, water, agriculture, waste and electricity (grouped as `Extractive` industries) are consistently linked with higher support for the Liberal party, with the mangitude of this effect slightly increasing over the years (see row 3 in Figure \@ref(fig:partresplot1)).
+Electorates with higher proportions of workers in mining, gas, water, agriculture, waste and electricity (grouped as `Extractive` industries) are consistently linked with higher support for the Liberal party, with the magnitude of this effect slightly increasing over the years (see row 3 in Figure \@ref(fig:partresplot1)).
 This is unsurprising, as the Liberal party has close ties with these traditional energy industries, and typically present policies to reduce taxation on energy production.
@@
 De facto relationships, but not marriages, are found to be an important (and significant) predictor of the two-party preferred vote in all six elections, with more de facto relationships associated with higher support for the Labor party.
-Marriages however, are insignficiant.
+Marriages however, are insignificant.
 
@@
 ### Diversity
-Larger migrant populations from Asia, the Middle East, South-Eastern Europe, the United Kingdom and elsewhere, are either assocated with Labor support, or have no effect.
-Of these areas, only South-Eastern European populations were signficant in each election and Asian migrants were a significant in 2010.
+Larger migrant populations from Asia, the Middle East, South-Eastern Europe, the United Kingdom and elsewhere, are either associated with Labor support, or have no effect.
+Of these areas, only South-Eastern European populations were significant in each election and Asian migrants were a significant in 2010.
 Speaking other language (aside from English) however, have a far stronger effect, as observed through `OtherLanguageHome`.
@@
 For example, electorate income levels (via `Incomes`) is likely to be related to electoral unemployment and labor force participation (via `Unemployment`).
-In 2001, the coefficient estiamte for `Unemployment` negative but not significant, and the `Incomes` variables  is significant.
+In 2001, the coefficient estimate for `Unemployment` negative but not significant, and the `Incomes` variables  is significant.
 If the `Incomes` variable is removed from the model, `Unemployment` absorbs the negative effect, becoming significant ($p = 0.0056$).

### 2019-03-29 draft-paper.Rmd
add abstract. small proofread

@@
   address: Monash University\newline Melbourne, Australia
-abstract: "A brief summary of our ideas"
-keywords: "blah, blah"
+abstract: "This research examines the relationships between electoral socio-demographic characteristics and two-party preference in Australian federal elections, with the six elections from 2001 to 2016 considered.
+Socio-demographic information is derived from the Australian Census, which occurs every five years.
+Since a Census is not directly available for each election, a method of spatio-temporal imputation is employed to impute Census data for the electorates in each election.
+This accounts for both spatial and temporal changes in electoral characteristics between Censuses.
+To capture any spatial heterogeneity, a spatial error model is estimated for each election, which incorporates a spatially structured random effect vector that can thought of as the unobserved political climate in each electorate.
+Over time, most socio-demographic characteristics that affect electoral two-party preference do not vary, with industry of work, incomes, household mobility and de facto relationships having strong effects in the six elections.
+Education and unemployment are amongst those that have varying effects.
+It is also found that between 2004 and 2013, the spatial effect is insignificant, meaning that electorates voted effectively independently.
+All data featured in this study has been contributed to the `eechidna` `R` package (available on CRAN)."
+keywords: "federal election, Census, Australia, spatial modelling, imputation, socio-demographics, electorates, R, eechidna"
 wpnumber: no/yr
@@
 
-It is important here to note the ecological fallacy - insights are being drawn at the electorate level, and cannot be inferred for another disaggregate level (e.g. individual voters).
+It is important here to note the ecological fallacy: insights are being drawn at the electorate level, and cannot be inferred for another disaggregate level (e.g. individual voters).
 
@@
 In 2001, the coefficient estimate for `Unemployment` negative but not significant, and the `Incomes` variables is significant.
-If the `Incomes` variable is removed from the model, `Unemployment` absorbs the negative effect, becoming significant ($p = 0.0056$).
+If the `Incomes` variable is removed from the model in 2016, `Unemployment` absorbs the negative effect, becoming significant ($p = 0.0056$).
 
@@
 It is also the first study to model any Australian election since 2010 using Census information.
-Additionally, a key contribution of this research is the spatio-temporal imputation, which is applied to obtain Census data for the 2004, 2007, 2010 and 2013 elections.
+Additionally, a key contribution of this research is the wrangling of the raw data and spatio-temporal imputation, which is applied to obtain Census data for the 2004, 2007, 2010 and 2013 elections.
 All data sets used in this study have been contributed to the `eechidna` `R` package, which provides a rich, accessible data resource for future Australian electoral analysis.
@@
 `eechidna` (Exploring Election and Census Highly Informative Data Nationally for Australia) makes it easy to look at the data from the Australian Federal elections and Censuses that occurred between 2001 and 2016.
-This study contributed a large revision to the `eechidna` package, which included the addition of election and Census data for 2001-2010, vote outcomes for polling booths and imputed Census data for election years.
+This study contributed a large revision to the `eechidna` package, which included the addition of election and Census data for 2001-2010, voting outcomes for polling booths and imputed Census data for election years.
 For more details on using `eechidna`, please see the articles (vignettes) on the github page (https://ropenscilabs.github.io/eechidna/).

### 2019-03-29 draft-paper.Rmd
mentioning: add intercept for interpretability
updated with mum edits

@@
 The two-party preferred vote, a measure of support between these two parties, fluctuated between 47.3% and 53.5% (in favour of the Liberal party) over this period.
-This study explores how electoral characteristics relate to two-party preference, and whether their effects have changed over time.
-Electoral socio-demographics are derived from the Census, and vote counts are obtained from federal elections.
-
-Joining these two data sources is problematic as there in an inherent asynchronicity in the two events.
+This study explores how electoral socio-demographic characteristics relate to two-party preference, and whether their effects have changed over time.
+Electoral socio-demographics are derived from the Australian Census, and vote counts are obtained from federal elections.
+
+Joining these two data sources is problematic as there in an inherent asynchronicity in the two types of events.
 A Census is conducted by the Australian Bureau of Statistics (ABS) every five years, whereas a federal election (conducted by the Australian Electoral Commission (AEC)) usually occurs every three years.
-The first problem addressed is that of obtaining appropriate Census data for the 2004, 2007, 2010 and 2013 elections - election years in which a Census does not occur.
+The first problem addressed is that of constructing appropriate Census data for the 2004, 2007, 2010 and 2013 elections - election years in which a Census does not occur.
 The predominant approach in previous studies is to join voting outcomes to the nearest Census, without accounting for any temporal differences (see @DavisStimson98, @Stimson06, @Liao09 and @Stimson09).
 Furthermore, electoral boundaries change regularly, so spatial discrepancies also arise when matching electoral data.
-To obtain appropriate Census data for these four elections, electoral socio-demographics are imputed using a spatio-temporal imputation that combines areal interpolation [@Goodchild1993] and linear time-interpolation.
-Collecting and wrangling the raw data, along with the imputation process, are detailed in section \@ref(data).
+To obtain appropriate Census data for these four elections, electoral socio-demographics are constructed using a spatio-temporal imputation that combines areal interpolation [@Goodchild1993] and linear time-interpolation.
+Collecting and wrangling the raw data, along with the imputation process, are detailed in Section \@ref(data).
 All data and associated documentation relating to this procedure are available in the `eechidna` `R` package [@eechidna], providing a resource for future analysis .
@@
 Previous work on modelling Australian federal elections have found that aggregate socio-demographics are relatively good predictors of voting outcomes.
-@Forrest01 does this using multiple regression of the Liberal and Labor primary vote for polling booths in the Farrer electorate in 1998.
+@Forrest01 use multiple regression to model the Liberal and Labor primary vote for polling booths in the Farrer electorate in 1998 as a function of Census variables from 1996.
 @Stimson06, @Stimson09 and @Stimson12 use principal component analysis of polling booths in the 2001, 2004 and 2007 elections respectively, also finding that socio-demographic characteristics of polling booths are linked to their two-party preferred vote.
 On the contrary, @Stimson09 models the polling booth swing vote (change in the two-party preferred vote) in the 2007 election, finding that little of swing vote can be explained by Census data.
-Instead of analyzing a single election in isolation, this paper employs a consistent model framework across six elections so that temporal changes in the effects of socio-demographics can be observed, where each federal elections is modelled with a cross-sectional data set.
+Instead of analyzing a single election in isolation, this paper employs a consistent model framework across six elections so that temporal changes in the effects of socio-demographics can be observed, where each federal elections is modelled with a cross-sectional dataset.
 The use of a regression framework to examine these socio-political relationships over time is seemingly absent from previous Australian studies.
@@
 
-The cross-sectional data set for each election consists of the two-party preferred vote (response variable), and socio-demographic variables (explanatory variables) that characterize each electorate.
-To obtain these cross-sections, socio-demographic variables are first standardized, and then principal components are used to group variables into "factors".
-To account for the inherent spatial structure of the data, a spatial error model is fit for each election.
-These steps are discussed in section \@ref(modelling).
-In section \@ref(results) inference is conducted on the models to see which effects are significant, how effects change over time and which electorates have abnormal voting behaviour.
+The cross-sectional dataset for each election used here consists of the two-party preferred vote (response variable), and a set of common socio-demographic variables (explanatory variables) that characterize each electorate.
+To obtain these datasets, socio-demographic variables are first standardized, and then a principal component analysis is used to group variables into "factors".
+To account for the inherent spatial structure of the data, a spatial error model is then fit for each election.
+These steps are discussed in Section \@ref(modelling).
+In Section \@ref(results) inference is conducted to see which effects are significant and how these effects change over time.
+
+Section \@ref(conclusion) concludes, followed by two supplementary sections.
 
@@
 The voting outcome of interest is the electoral two-party preferred vote, which is provided by the Australian Electoral Commission (AEC) for the 2001, 2004, 2007, 2010, 2013 and 2016 elections via the AEC Tally Room.
-The AEC divide Australia into 150 regions called electorates, with each corresponding to a single seat in the House of Representatives.
+The AEC divides Australia into 150 regions, called electorates, with each corresponding to a single seat in the House of Representatives.
 Voting is compulsory in Australia, and each voter assigns a numbered preference to each available candidate in their electorate.
@@
 Socio-demographic variables are derived from the Census of Population and Housing (Census), which is a survey of every household in Australia, recording information such as age, gender, ethnicity, education level and income.
-There have been four Censuses in the 21st century, being that in 2001, 2006, 2011 and 2016.
+There have been four Censuses so far in the 21st century, conducted in 2001, 2006, 2011 and 2016.
 The Australian Bureau of Statistics (ABS) conducts the Census and publishes aggregated information.
-The ABS approximation of electorates at the time of the Census is chosen.
-From this aggregate information, 67 socio-demographic variables are computed for each of the electorates.
+The ABS uses electoral boundaries as defined by the AEC at the time of each Census, which may not match those in place at the subsequent election.
+From the available Census information aggregated at the electorate level, 65 socio-demographic variables are defined for each of the electorates to be used in the analysis.
 
 Raw data is sourced online from the AEC and ABS websites in .csv and .xlsx files.
-The format of these files change over the years, making extracting the appropriate information a big task.
+The formats of these files differ over the years, making extracting the appropriate information a big task.
 The functions available in the `dplyr` [@dplyr] and `readxl` [@readxl] `R` packages are very useful, as they provide fast consistent tools for data manipulation and functions to import .xlsx files (respectively).
-The 2001 and 2006 Census data are however published in a format where each electorate has a separate document, making it difficult to use the `dplyr` tools and instead cells have to be selected from each individual file to construct the desired variables.
+The 2001 and 2006 Census data are published in a format where the information for each electorate is held in a separate document making it difficult to use the `dplyr` tools.
+Instead, cells have to be selected from each individual file to construct the desired variables.
 All scripts required for the data wrangling process can be found in the github repository for the `eechidna` `R` package [@eechidna], along with the raw data.
@@
 
+For each election, neighbouring Census information has to be combined in some way so that it represents the boundaries in place at the time of the election.
+This is done by taking the electoral boundaries and imputing the corresponding socio-demographic characteristics for each of neighbouring Censuses, thereby addressing the spatial aspect.
+Next, to deal with the temporal component, characteristcs at the time of the election are constructed using linear interpolation between the spatially imputed neighbouring Census variables.
+
 To account for spatial differences, the piece-wise approximation method in @Goodchild1993 is adopted.
 Consider a map of source zones $s = 1,...,S$, for which socio-demographic information is available, and a set of target zones $t = 1,...,T$ for which information is to be imputed.
-In this context the map of electoral boundaries at the time of a Census would be the source zones, and the boundaries at the time of the election would be the target zones.
-Denote the area of intersection between source zone $s$ and target zone $t$ as $A_{s,t}$, the population of the source zone $s$ as $U_s$, and the population of intersection between source zone $s$ and target zone $t$ as $P_{s,t}$.
-
-Compute each $A_{s,t}$ and estimate population of the intersection:
-
-$$\hat{P}_{s,t} = \frac{U_s*A_{s,t}}{\sum_{t=1}^T A_{s,t}}$$
-This assumes that populations are uniformly distributed within each source zone.
-
-In order to calculate socio-demographic information for each of the target zones, a weighted average is taken using the estimated population as weights.
-Denote a given Census variable for the target zone $C_t$, and the same Census variable for the source zone $D_s$:
-
-$$\hat{C}_t = \frac{\sum_{s=1}^{S}{D_s*\hat{P}_{s,t}}}{\sum_{s=1}^{S}{\hat{P}_{s,t}}}$$
-This assumes that each individual in a source zone assumes the aggregate characteristics of the zone.
-
-Applying this to each of the target zones addresses the spatial component, as it imputes the required socio-demographic for the desired electoral boundaries.
-However these are applicable at the time of the Census (source year) and are not yet appropriate for the election (target year).
-
-Denote year $y$, with a Census falling on $y_1$ and $y_3$, and an election on year $y_2$, and add this subscript to the Census variable estimate, $\hat{C}_{t,y}$.
-To account for temporal changes, linear interpolation is used between Census years to get the final estimate of a Census variable for the target zone in the election year $y_2$.
-This assumes that population evolves in a linear manner over time.
-
-$$\hat{C}_{t,y_2} = \frac{y_3-y_2}{y_3-y_1}*\hat{C}_{t,y_1} + \frac{y_2-y_1}{y_3-y_1}*\hat{C}_{t,y_3}$$
-
-### Applied
-
-Publicly available Census data is aggregated and there are different resolutions accessible, ranging from SA1 (over 50,000 zones) to electoral divisions (150 zones).
-For this study electoral divisions are used as source zones, and this imputation method is applied for each of the 2004, 2007, 2010 and 2013 elections.
-To demonstrate its functionality, consider the imputation of socio-demographic variables for the electorate of Hume in New South Wales (NSW), at the time of the 2013 federal election.
-Figure \@ref(fig:hume13) shows this region amongst other NSW electorates.
-
-```{r hume13, fig.cap = "Some of the electoral boundaries in NSW for 2013, with the electoral boundary for Hume, shown in purple.", out.height="30%"}
+This is described in the context of a single election, and a single neighbouring Census.
+
+Let the map of electoral boundaries at the time of a Census define the source zones, and let the boundaries at the time of the election be the target zones.
+Denote the area of intersection between a source zone $s$ and a target zone $t$ as $A_{s,t}$.
+Additionally, let the population of the source zone $s$ be $U_s$ and the population of intersection between source zone $s$ and target zone $t$ be $P_{s,t}$.
+The estimated population of intersection is given by:
+
+$$\hat{P}_{s,t} = \frac{U_s*A_{s,t}}{\sum_{t=1}^T A_{s,t}}, \mbox{ for all } s=1,...,S \mbox{ and } t=1,...,T.$$
+Note that this estimator implicitly assumes that populations are uniformly distributed within each source zone.
+
+In order to calculate socio-demographic information for each of the target zones, a weighted average is taken using the estimated intersection populations as weights.
+Denote a given Census variable for the target zone by $C_t$, and the same Census variable for the source zone as $D_s$.
+Then, estimate $C_t$ using:
+
+$$\hat{C}_t = \frac{\sum_{s=1}^{S}{D_s*\hat{P}_{s,t}}}{\sum_{s=1}^{S}{\hat{P}_{s,t}}}, \mbox{ for each } t=1,...,T.$$
+This concludes the spatial imputation of the socio-demographic characteristics for one target zone (a single electoral boundary), at time of only one of the neighbouring Censuses.
+This process is repeated for all of the target zones, and then for the other neighbouring Census.
+
+To account for temporal changes, linear interpolation is used between Census years to get the final estimate of a Census variable for the target zone in the election year.
+Let $y_1$ be the year of the Census preceeding an election, let $y_2$ be the year of the election and $y_3$ be the year of the Census that follows.
+Add this year subscript to the Census variable estimate $\hat{C}_t$, resulting in $\hat{C}_{t,y}$.
+Linear interpolating between these Census years results an imputed value for the election year, given by
+
+$$\hat{C}_{t,y_2} = \frac{y_3-y_2}{y_3-y_1}*\hat{C}_{t,y_1} + \frac{y_2-y_1}{y_3-y_1}*\hat{C}_{t,y_3}.$$
+Implicitly this assumes that population characteristics change in a linear manner over time.
+
+### An illustration of the spatio-temporal imputation
+
+Census data is publicly available that different levels of aggregation, ranging from SA1 (over 50,000 zones) to electoral divisions (150 zones).
+For this study electoral divisions are used as source zones, and the imputation method is applied to produce socio-demographic variables for each of the 2004, 2007, 2010 and 2013 elections.
+As mentioned earlier, there is no need to impute socio-demographic variables for the 2001 and 2016 elections.
+To illustrate the method, consider the imputation of socio-demographic variables for the electorate of Hume in New South Wales (NSW) at the time of the 2013 federal election.
+The boundaries shown in Figure \@ref(fig:hume13) define all  target zones in NSW for 2013, with the target zone of interest (Hume) shaded purple.
+
+```{r hume13, fig.cap = "Some of the electoral boundaries in NSW for 2013, with the electoral boundary for Hume shown in purple.", out.height="30%"}
 hume_area13 <- nat_map13 %>% 
@@
 
-The Censuses neighbouring the 2013 election are those in 2011 and 2016, and the Hume boundary is changed, as seen by plotting the Hume boundary (purple) in the 2013 election over the divisions in 2016 (see Figure \@ref(fig:hume16)).
+The corresponding source zones from the 2016 Census are shown in Figure \@ref(fig:hume16).
+As can be seen, the Hume boundary from the 2013 election (shaded purple) does not perfectly match any of the source zones.
 
@@
 
-There are many electorates in 2016 that intersect with the purple region (Hume boundary for 2013), these include the divisions of Riverina, Eden-Monaro and Hume, along with smaller intersecting areas with Fenner, Calare, Gilmore and Whitlam.
-To impute Census information for this purple region, calculate the percentage of each 2016 electorate that intersects with the purple region, which is then used to estimate intersection populations $\hat{P}_{s,t}$.
+There are many source zones from the 2016 Census that intersect with this purple region, including the divisions of Riverina, Eden-Monaro and Hume, along with smaller intersecting areas with Fenner, Calare, Gilmore and Whitlam.
+The proportion of each source zone that overlaps with the purple region is calculated, and used to obtain the intersecting populations $\hat{P}_{s,t}$.
 
@@
 
-Electorate (2016) | Percentage | Population | Estimated Population Allocated to Purple Region: $\hat{P}_{s,t}$
+Source Zone (2016) | Proportion | Source Zone Population | Population Allocated to Purple Region: $\hat{P}_{s,t}$
 --------------------|----------|-----------|---------------------------------------------
-HUME              | 96.54%     | 150643                   | 145427
-RIVERINA          | 25.11%     | 155793                   | 39117
-EDEN-MONARO       | 11.09%     | 147532                   | 16358
-CANBERRA          | 0.28%      | 196037                   | 548
-FENNER            | 0.23%      | 202955                   | 474
-WHITLAM           | 0.06%      | 152280                   | 92
-GILMORE           | 0.06%      | 150436                   | 86
-CALARE            | 0.01%      | 161298                   | 21
+HUME              | 0.9654     | 150643                   | 145427
+RIVERINA          | 0.2511     | 155793                   | 39117
+EDEN-MONARO       | 0.1109     | 147532                   | 16358
+CANBERRA          | 0.028      | 196037                   | 548
+FENNER            | 0.0023     | 202955                   | 474
+WHITLAM           | 0.0006     | 152280                   | 92
+GILMORE           | 0.0006     | 150436                   | 86
+CALARE            | 0.0001     | 161298                   | 21
 
@@
 
-Now consider the socio-demographic $AusCitizen$ - the proportion of people in the region who are Australian citizens.
+Now consider the socio-demographic variable $AusCitizen$ - the proportion of people in the region who are Australian citizens.
 
@@
     AusCitizen = round(AusCitizen, 2)) %>% 
-  arrange(-`Estimated Population Allocated to Purple Region`)
+  arrange(-`Estimated Population Allocated to Purple Region`) %>% 
+  rename(`Source Zone (2016)` = DivisionNm)
 
@@
 
-|DivisionNm      | AusCitizen (%): $D_s$| Estimated Population Allocated to Purple Region: $\hat{P}_{s,t}$|
+|Source Zone (2016)   | AusCitizen (%): $D_s$| Estimated Population Allocated to Purple Region: $\hat{P}_{s,t}$|
 |--------------------------|-------------------|------------------------------------------------|
@@
 
-Then taking a weighted average of $AusCitizen$ using the estimated population as weights yields $\hat{C}_{Hume,2016} = 89.65 \%$.
-Repeating this process using the 2011 Census and electoral boundaries yields $\hat{C}_{Hume,2011} = 91.00 \%$
-
-Finally, linearly interpolate between 2011 and 2016 to arrive at the 2013 estimate:
+A weighted average of $AusCitizen$, with the allocated population from each source zone as weights, yields $\hat{C}_{Hume,2016} = 89.65 \%$.
+Repeating this process using the 2011 Census yields $\hat{C}_{Hume,2011} = 91.00 \%$
+
+Finally, linear interpolation between 2011 and 2016 yields the 2013 estimate:
 \begin{eqnarray*}
@@
 
-This is done for each of the socio-demographic variables, and repeated each of the 2013 electorates.
+This is done for each of the 65 socio-demographic variables, and is repeated for each of the 149 remaining target zones corresponding with 2013 electorates.
 
@@
 
+*1: HERE Introductory paragraph*
+
 ## Data pre-processing
-With socio-demographic information now available for each electorate, each election is joined to the data  corresponding with its two-party preferred vote.
+Having used the imputation method detailed in Section \@ref(data), all socio-demographic variables are not for each electorate in each election, and are joined to with the corresponding two-party preferred vote data.
 Socio-demographic variables within each election year are standardized to have mean zero and variance one, to adjust for changing variable scales.
 For example, inflation-adjusted median rental prices increased across almost all electorates, with median rent of 200 dollars per week placing an electorate in the 90th percentile in 2001, but only the 30th percentile in 2016.
+By standardizing, the median rental price of an electorate is reported relative to all other electorates in the same year.
 
@@
 With only $N = 150$ observations (electorates) in each election and $p = 65$ socio-demographic variables in each cross-section, any model using all variables would face serious problems with multi-collinearity and over-fitting, likely leading to erroneous conclusions regarding variable significance.
-Therefore a form of dimension reduction is adopted before models are fit.
-
-Socio-demographic variables[^2] that represent similar information are combined into "factors" using principal component analysis (PCA).
+Therefore a form of dimension reduction is adopted before any models are fitted.
+
+Socio-demographic variables[^2] that represent similar information are combined into "factors", informed by principal component analysis (PCA) on a combined dataset of socio-demographic variables across the six elections.
+A factor is created by combining several variables that have large loadings in a particular component and an intuitive justification for their grouping.
+Only the first four principal components are considered (noting that the scree plot levels off after the fourth component).
+A loading with magnitude greater than 0.15 is considered large.
+After computing these *sums*, each factor is again standardized to have mean zero and variance one, within each election.
+*2: HERE Explain the different factors and how they are constructed - also justify combining all cross-sections*
+
 The scree plots of the principal components for each election all level off after four components, and the loadings of these four components are similar across the elections.
@@
 Consider the `Incomes` factor as an illustration.
-Independent of principal components, we may suspect that median personal income, median household income and median family income are providing similar information about the financial wellbeing of an electorate.
-Their loadings in the first principal component are large (0.19, 0.21 and 0.22 respectively), which provides the evidence needed to combine these variables into a single factor, which is called `Incomes`.
-
-This process reduces the predictor set to $p = 30$.
+Independent of principal components, we may suspect that median personal income, median household income and median family income all provide similar information about the financial wellbeing of an electorate.
+Their loadings on the first principal component are large (0.19, 0.21 and 0.22 respectively), suggesting that these variables should be combined into a single factor.
+
+Having produced these *Seven* factors, the predictor set is reduced to $p = 30$. All of these variables are used in the regression for each election.
 
@@
 
-## Model framework
-An identical model specification is used across the six elections, with each election modelled separately.
-This allows for the socio-demographic effects to be estimated separately for each year, allowing for interpretation of temporal changes in these effects.
-This is preferable over a single longitudinal model because it avoids any concerns of undue bias stemming from an incorrectly imposed time-varying restriction on any variable.
+## Regression incorporating spatially dependent errors
+An identical model specification is used for each of the six elections, with each election modelled separately.
+This allows for the socio-demographic effects to be estimated separately for each election year, facilitating analysis of temporal changes in variable effects.
+This approach is preferable to using a single longitudinal model because it avoids any concerns of undue bias stemming from incorrectly imposed time-varying restrictions on any variable.
 Without such restrictions, a pooled cross-sectional model does not yield any distinct advantage over separate cross-sections.
-The panel approach is avoided because of how frequently electoral boundaries change, meaning that electorates that have the same name across elections are not guaranteed to represent the same geographical region.
-Therefore any fixed or random effects models would be difficult to estimate without implementing consistent boundaries, which would requiring further imputation.
-
-For each cross-section, let the response variable be the two-party preferred vote in favour of the Liberal party, denoted $Y$, with $Y = 70$ representing a 70% preference for Liberal, 30% for Labor.
-Although $Y$ lies in the interval $(0,100)$, observed values are never very close to 0 or 100 (minimum $24.05 \%$ and maximum $74.90 \%$), so there is no need to impose the constraint of $Y \in [0,100]$.
-Furthermore, the response is found to be spatially correlated in each election (Moran's I test, $p \le 7\cdot10^{-15}$).
-This is expected, as electorates are aggregate spatial units, and hence the spatial structure of the data must modelled appropriately.
+The panel approach is avoided because of how frequently electoral boundaries change, noting that electorates with same name across elections are not guaranteed to represent the same geographical region.
+Therefore any fixed or random effects models would be difficult to estimate without implementing consistent boundaries, which would require further imputation.
+
+For each cross-section, let the response $\boldsymbol y$ be the vector two-party preferred vote in favour of the Liberal party, with $y_i = 70$ representing a 70% preference for Liberal, 30% for Labor, in electorate $i$.
+Although $y_i$ lies in the interval $(0,100)$, observed values are never very close to 0 or 100 (minimum $24.05 \%$ and maximum $74.90 \%$), so there is no need to formally impose the constraint of $y_i \in [0,100]$.
+Furthermore, the responses are found to be spatially correlated in each election (Moran's I test, $p \le 7\cdot10^{-15}$).
+This not surprising as electorates are aggregate spatial units, and hence the spatial structure of the data must modelled appropriately.
 
 The spatial error model [@Anselin88] is chosen because captures spatial heterogeneity by incorporating a spatially structured random effect vector [@LeSage2009].
-In this context, the random effect can be thought of as capturing the unobserved political climate in each electorate, where the climate is correlated with the climate in neighbouring electorates.
-This functions under the assumption that the climate is independent of electoral socio-demographics, and that an electorate is equally correlated with any electorate that shares a part of its boundary.
-Spatial weights are calculated in accordance with these assumptions.
-The spatial error model is specified as follows:
-
+In this context, the random effect can be thought of as capturing the unobserved political climate in each electorate, where this climate is correlated with the climate in neighbouring electorates, under the assumption that the climate is independent of electoral socio-demographics.
+
+Spatial weights are calculated in accordance with the assumption that an electorate is equally correlated with any electorate that shares a part of its boundary.
 Let $\rho$ be spatial autoregressive coefficient, $\boldsymbol v$ be a spherical error term, ${\boldsymbol W}$ be a matrix of spatial weights (containing information about the neighbouring regions), $\boldsymbol X$ be a matrix of socio-demographic covariates, $\boldsymbol \beta$ be a vector of regression coefficients and $\boldsymbol a$ be a spatially structured random effect vector.
 
-$${\boldsymbol y} = {\boldsymbol X} {\boldsymbol \beta} + {\boldsymbol a}$$
+$${\boldsymbol y} = {\boldsymbol X} {\boldsymbol \beta} + {\boldsymbol a},$$
 and
 
-$${\boldsymbol a} = \rho {\boldsymbol W} {\boldsymbol a} + {\boldsymbol v}$$
+$${\boldsymbol a} = \rho {\boldsymbol W} {\boldsymbol a} + {\boldsymbol v},$$
 where
 
-$${\boldsymbol v} \sim N({\boldsymbol 0}, \sigma^2 {\boldsymbol I_n})$$.
-
-so it can be written
-
-$${\boldsymbol y} = {\boldsymbol X} {\boldsymbol \beta} + ({\boldsymbol I}_n-\rho {\boldsymbol W})^{-1}{\boldsymbol v}$$
-
-Estimation is done using feasible generalized least squares.
-
-Table 3 details the resultant estimated model coefficients and their estimated standard errors for each of the six elections. These are interpreted in the next section.
+$${\boldsymbol v} \sim N({\boldsymbol 0}, \sigma^2 {\boldsymbol I_n}),$$.
+
+and hence
+
+$${\boldsymbol y} = {\boldsymbol X} {\boldsymbol \beta} + ({\boldsymbol I}_n-\rho {\boldsymbol W})^{-1}{\boldsymbol v}.$$
+
+Estimation of the above spatial error model is undertaken using feasible generalized least squares.
+
+Table 3 details the estimated model coefficients and their estimated standard errors, for each of the six elections. An interpretation of these estimated values is provided in the next section.
 
@@
   column.sep.width = "1pt", 
-  title = "Estimated model for each of the six elections.",
+  title = "Estimated spatial error model parameters (standard errors) for each of the six election years.",
   dep.var.labels = "Two-party preferred vote in favor of the Liberal party", 
@@
 \begin{table}[!htbp] \centering 
-  \caption{Estimated model for each of the six elections.} 
+  \caption{Estimated spatial error model parameters (standard errors) for each of the six election years.} 
   \label{} 
@@
 \\[-1.8ex]\hline 
-\hline \\[-1.8ex] 
- & \multicolumn{6}{c}{\textit{Dependent variable:}} \\ 
-\cline{2-7} 
-\\[-1.8ex] & \multicolumn{6}{c}{Two-party preferred vote in favor of the Liberal party} \\ 
  & 2001 & 2004 & 2007 & 2010 & 2013 & 2016 \\ 
@@
 
-The spatial autoregressive coefficient $\rho$ is positive and significant in only the 2001 and 2016 elections (Figure \@ref(fig:rhovis)), meaning that in these elections, an electorate's political climate was affected by the attitudes of it's neighbours.
+The spatial autoregressive coefficient $\rho$ is positive and significant in only the 2001 and 2016 elections (Figure \@ref(fig:rhovis)), meaning that, on average, in these elections, the political climate of an electorate appears to be affected by the attitudes of it's neighbours.
 Conversely, in the other four elections, the spatial effect weakens to become insignificant.
-In these years, it appears that the spatial component does not explain anything over and above the electoral socio-demographics, meaning electorates voted effectively independently.
-
-```{r rhovis, fig.cap="Estimates of the spatial autoregressive parameter for each of the six elections, with a 95\\% confidence interval.
-Only in 2001 and 2016 is there a significant spatial component", out.height='15%', out.extra = ''}
+In these years, it appears that the spatial component does not explain anything over and above the electoral socio-demographics, meaning electorates effectively voted independently.
+
+```{r rhovis, fig.cap="Estimates of the spatial autoregressive parameter for each of the six elections, reported with their individual 95\\% confidence intervals.
+Only in 2001 and 2016 is there a significant spatial component.", out.height='15%', out.extra = ''}
 rho_df <- bind_rows(glsmod16$rho_df %>% mutate(year = "2016"), 
@@
 ## Influential socio-demographics
-To investigate the socio-demographics that have a strong effect on the two-party preferred vote, partial residual plots are used.
+To investigate the socio-demographics that have a strong effect on the two-party preferred vote, partial residual plots are used and shown in Figures \@ref(fig:partresplot1) and \@ref(fig:partresplot2).
+The partial residuals are the residuals from the fitted model with the estimated effect an individual variable added.
+The estimated intercept is also added to the partial residuals for interpretability.
 These show the direction, size and significance of an estimated effect - the slope of the prediction line matches the estimated coefficient, and the shaded region represents a 95% confidence band.
@@
 Typically the Labor party campaigns on more progressive policies, which often include tax reform that adversely affects higher income earners, and more generous social assistance programs.
-Perhaps due to these policies, higher income electorates have been more likely to support the Liberal party, as the `Incomes` factor has a positive effect on Liberal preference (see row 1 in Figure \@ref(fig:partresplot1)).
-This effect is significant in every election aside from 2004, in which it is only marginally insignificant ($p = 0.0613$).
+Perhaps it is due to these policies, that higher income electorates appear more likely to support the Liberal party, as the `Incomes` factor has a positive effect on Liberal preference (see row 1 in Figure \@ref(fig:partresplot1)).
+This effect is significant in every election aside from 2004, where it is only marginally insignificant ($p = 0.0613$).
 Unemployment however, has not been as influential.
@@
 Similarly, workers in managerial, administrative, clerical and sales roles (`ManagerAdminClericalSales`) is also a significant predictor of two-party preference vote across all six elections, with a higher proportion of people working these jobs increasing Liberal support.
-The magnitude of this effect has also increased over the years.
+The magnitude of this effect also seems to increase over the years.
 
 ### Household mobility
-In each of the six elections, electorates with a higher proportion of people that have recently moved house (meaning in last five years) were more likely to support the Liberal party, although this effect was marginally insignificant in 2013 (see row 6 in Figure \@ref(fig:partresplot1).
+In each of the six elections, electorates with a higher proportion of people that have recently (past five years) moved house were more likely to support the Liberal party, although this effect was marginally insignificant in 2013 (see row 6 in Figure \@ref(fig:partresplot1).
 Having controlled for characteristics of house ownership and rental prices (via the factors `PropertyOwned` and `RentLoan` respectively), this is effect is somewhat surprising.
@@
 De facto relationships, but not marriages, are found to be an important (and significant) predictor of the two-party preferred vote in all six elections, with more de facto relationships associated with higher support for the Labor party.
-Marriages however, are insignificant.
+The proportion of individuals who are marriaged however, are insignificant (not shown).
 
 ### Age
-Older regions are often believed to be more conservative, and it can be seen that electorates with a higher median age have been more likely to support the Liberal party - although this effect is only significant in 2007 and 2010 (see row 2 in Figure \@ref(fig:partresplot2)).
+Regions comprised of older people are often believed to be more conservative, and indeed it found that electorates with a higher median age are more likely to support the Liberal party - although this effect is only significant in 2007 and 2010 (see row 2 in Figure \@ref(fig:partresplot2)).
 
 ### Education
-Since 2007, electorates with higher education levels have been associated with supporting the Labor party, although this effect is only significant in 2016.
-Prior to 2007, education had an almost zero effect (see row 3 in \@ref(fig:partresplot1)).
+Since 2007, electorates with higher education levels are associated with supporting the Labor party, although this effect is only significant in 2016.
+Before 2007, education has an almost zero effect (see row 3 in Figure \@ref(fig:partresplot1)).
 
@@
 Larger migrant populations from Asia, the Middle East, South-Eastern Europe, the United Kingdom and elsewhere, are either associated with Labor support, or have no effect.
-Of these areas, only South-Eastern European populations were significant in each election and Asian migrants were a significant in 2010.
-Speaking other language (aside from English) however, have a far stronger effect, as observed through `OtherLanguageHome`.
-Electorates with more diverse speech were more likely to support the Liberal party from 2004 onwards, with this effect being significant in 2007, 2010 and 2016.
-Furthermore, of the variables relating to religion, only Judaism shows a consistent effect, with electorates with larger Jewish populations more likely to vote Liberal.
+Of these areas, only South-Eastern European populations appear significant in each election, with the proportion of Asian migrants also being significant in 2010.
+Speaking other languages (aside from English) however, appears to have a far stronger effect, as observed through the `OtherLanguageHome` variable.
+Electorates with more diverse speech are associated with higher support for the Liberal party from 2004 onwards, with this effect being significant in 2007, 2010 and 2016.
+Furthermore, of the variables relating to religion, only Judaism shows a consistent effect, with electorates with relatively large Jewish populations more likely to vote Liberal.
 
@@
 However, there still remain some variables which are closely related.
-For example, electorate income levels (via `Incomes`) is likely to be related to electoral unemployment and labor force participation (via `Unemployment`).
-In 2001, the coefficient estimate for `Unemployment` negative but not significant, and the `Incomes` variables is significant.
-If the `Incomes` variable is removed from the model in 2016, `Unemployment` absorbs the negative effect, becoming significant ($p = 0.0056$).
+For example, an electorate's income level (via `Incomes`) is likely to be related to electoral unemployment and labor force participation (via `Unemployment`).
+In 2001, the coefficient estimate for `Unemployment` is negative but not significant, whilst the `Incomes` variables is significant.
+If the `Incomes` variable is removed from the model in 2001, `Unemployment` absorbs the negative effect, becoming significant ($p = 0.0056$).
 
@@
 It is often hypothesized that states have systematic differences that cause their electorates to vote differently.
-Boxplots of residuals grouped by state reveal that only Tasmania, the Australian Capital Territory and the Northern Territory appear to have a state-specific effect that is not captured by the models.
-Tasmania and the Australian Capital Territory appear to have a bias towards Labor, whereas the Northern Territory has one towards Liberal.
-There are few electorates in these states (five, two and two respectively), so this might be due to incumbent effects rather than an actual state-specific bias.
+Boxplots of residuals grouped by state reveal that only Tasmania, the Australian Capital Territory and the Northern Territory appear to have a state-specific effect not captured by the models.
+Tasmania and the Australian Capital Territory appear to have a bias towards Labor, whereas the Northern Territory tends towards voting Liberal.
+However, there are relatively few electorates in each of these states (five, two and two respectively), so this apparent result may be due to incumbent effects rather than an actual state-specific bias.
 
@@
 
+*3: HERE Add paragraph on the overall strategy and contributions - can restate stuff from intro*
+
 Across the past six elections, most of the socio-demographics that drive the electoral two-party preferred vote have remained steady, whilst a few (typically weaker) effects have varied over time.
@@
 Additionally, a key contribution of this research is the wrangling of the raw data and spatio-temporal imputation, which is applied to obtain Census data for the 2004, 2007, 2010 and 2013 elections.
-All data sets used in this study have been contributed to the `eechidna` `R` package, which provides a rich, accessible data resource for future Australian electoral analysis.
+All datasets used in this study have been contributed to the `eechidna` `R` package, which provides a rich, accessible data resource for future Australian electoral analysis.
 
@@
 
-This paper was produced using `RMarkdown`[@rmarkdown] and `knitr` [@knitr].
+This paper was produced using `RMarkdown` [@rmarkdown] and `knitr` [@knitr].
 All corresponding code for this paper can be found in the github repository (https://github.com/jforbes14/journal-writing), and the data used is available in the `eechidna` package [@eechidna].
@@
 
-All election and Census data sets, along with electoral maps and more, are available in the `eechidna` `R` package, which can be downloaded from CRAN.
+All election and Census datasets, along with electoral maps and more, are available in the `eechidna` `R` package, which can be downloaded from CRAN.
 `eechidna` (Exploring Election and Census Highly Informative Data Nationally for Australia) makes it easy to look at the data from the Australian Federal elections and Censuses that occurred between 2001 and 2016.

### 2019-03-31 draft-paper.Rmd
add modelling intro bit

@@
 *1: HERE Introductory paragraph*
+Electoral socio-demographic variables are now available for each of the six elections and are joined with their corresponding two-party preferred votes.
+Before choosing an appropriate model, two issues with the socio-demographic variables are addressed.
+Firstly, variable scales change over the years and as such the data is standardized accordingly.
+Secondly, many variables represent similar information and where appropriate, should be combined in some way.
+Principal component analysis is used to determine which intuitive groups of variables covary and should be combined into a single variable.
+This also reduces the dimension of the data.
+After these steps, a model specification is chosen.
 
 ## Data pre-processing
-Having used the imputation method detailed in Section \@ref(data), all socio-demographic variables are not for each electorate in each election, and are joined to with the corresponding two-party preferred vote data.
-Socio-demographic variables within each election year are standardized to have mean zero and variance one, to adjust for changing variable scales.
+Many of the socio-demographic variables have changing scales over the years.
 For example, inflation-adjusted median rental prices increased across almost all electorates, with median rent of 200 dollars per week placing an electorate in the 90th percentile in 2001, but only the 30th percentile in 2016.
-By standardizing, the median rental price of an electorate is reported relative to all other electorates in the same year.
+In order for socio-demographic effects to be comparable across years, the explanatory variables are standardized to have mean zero and variance one within each election year.
+By standardizing, each variable is reported as a relative measure compared to all other electorates in the same year.
 
@@
 After computing these *sums*, each factor is again standardized to have mean zero and variance one, within each election.
+
 *2: HERE Explain the different factors and how they are constructed - also justify combining all cross-sections*

### 2019-03-31 draft-paper.Rmd
added explanation of factors created + justification of doing single PCA, and edited general dimension reduction section

@@
 Secondly, many variables represent similar information and where appropriate, should be combined in some way.
-Principal component analysis is used to determine which intuitive groups of variables covary and should be combined into a single variable.
+Principal component analysis is used to identify variables covary, from which intuitive groupings are selected to be combined into a single variable.
 This also reduces the dimension of the data.
@@
 ### Dimension reduction
-With only $N = 150$ observations (electorates) in each election and $p = 65$ socio-demographic variables in each cross-section, any model using all variables would face serious problems with multi-collinearity and over-fitting, likely leading to erroneous conclusions regarding variable significance.
-Therefore a form of dimension reduction is adopted before any models are fitted.
-
-Socio-demographic variables[^2] that represent similar information are combined into "factors", informed by principal component analysis (PCA) on a combined dataset of socio-demographic variables across the six elections.
-A factor is created by combining several variables that have large loadings in a particular component and an intuitive justification for their grouping.
-Only the first four principal components are considered (noting that the scree plot levels off after the fourth component).
-A loading with magnitude greater than 0.15 is considered large.
-After computing these *sums*, each factor is again standardized to have mean zero and variance one, within each election.
-
-*2: HERE Explain the different factors and how they are constructed - also justify combining all cross-sections*
-
-The scree plots of the principal components for each election all level off after four components, and the loadings of these four components are similar across the elections.
-Principal components are then computed on the combined set of socio-demographics across all six elections.
-A factor is created by combining several variables all have large loadings in a particular component and when there is an intuitive reason as to why these variables could represent common information.
-A loading with magnitude greater than 0.15 is considered large.
-After computing these sums, each factor is again standardized to have mean zero and variance one, within each election.
-
-Consider the `Incomes` factor as an illustration.
-Independent of principal components, we may suspect that median personal income, median household income and median family income all provide similar information about the financial wellbeing of an electorate.
-Their loadings on the first principal component are large (0.19, 0.21 and 0.22 respectively), suggesting that these variables should be combined into a single factor.
-
-Having produced these *Seven* factors, the predictor set is reduced to $p = 30$. All of these variables are used in the regression for each election.
+There are only $N = 150$ observations (electorates) in each election and $p = 65$ socio-demographic variables in each cross-section, with many variables represent similar information about an electorate.
+Any model that uses all variables would face serious problems with multi-collinearity and over-fitting, which would likely lead to erroneous conclusions regarding variable significance.
+To address this, groups of variables that represent similar information are combined into "factors"[^2].
+
+A factor is created from a group of variables if there is an intuitive reason as to why they should represent similar information and if there is evidence to suggest that they covary.
+For example, a potential group would be variables relating to electoral incomes - median family, household and personal incomes.
+To determine which variables covary, principal component analysis is used on a combined dataset of socio-demographic variables from all six elections.
+It is appropriate to compute principal components in this way because when computed separately for each election, scree plots level off after four components and the loadings of the first four components are similar across the elections.
+
+*2: HERE - ADD JUSTIFICATION FOR DOING SINGLE PCA AND EXPLAIN FACTORS CREATED*
+
+Only the first four principal components from the combined dataset are considered, as the scree plot levels off after the fourth component.
+Variables that have a large loading in a particular component are deemed to covary, with a loading with magnitude greater than 0.15 being considered large.
+Six factors are created using this criteria.
+These are; `Incomes` - median personal, household and family incomes, `Unemployment` - unemployment and labor force participation rates, `ProportyOwned` - rates of housing ownership, mortgages, renting and government housing, `RentLoanPrice` - median rental and loan repayments, `FamHouseSize` - average household size, ratio of people to families and household makeup (single person, couple with kids and couple without kids) and `Education` - high school and university qualifications, jobs requiring higher levels of education as well as vocational course completions and jobs that do not require higher education levels (e.g.
+laborer, tradesperson).
+For each of these groups, variables with positive loadings are added and those with negative loadings are substracted to create a factor.
+After computing these sums, each factor is standardized to have mean zero and variance one, within each election.
+
+There are $p=30$ variables in the resultant predictor set, with all of these used in the regression for each election.
 

### 2019-03-31 draft-paper.Rmd
added first para conclusion

@@
 
-## Data pre-processing
+## Standardizing variables
 Many of the socio-demographic variables have changing scales over the years.
@@
 
-### Dimension reduction
+## Creating factors
 There are only $N = 150$ observations (electorates) in each election and $p = 65$ socio-demographic variables in each cross-section, with many variables represent similar information about an electorate.
@@
 
-Across the past six elections, most of the socio-demographics that drive the electoral two-party preferred vote have remained steady, whilst a few (typically weaker) effects have varied over time.
+This paper explores the effects of electoral socio-demographic characteristics on the two-party preferred vote in the 2001-2016 elections, using information from the corresponding Australian federal elections and Censuses.
+As a Census does not always occur in the same year as an election, Census data for the 2004-2013 elections are generated by employing a method of spatio-temporal imputation.
+This imputes electoral socio-demographics for the electoral boundaries in place at the time of the election - an approach that is distinctly different from previous work on modelling election outcomes, where Census and election data are typically joined without addressing their temporal differences.
+Before estimating a model, these socio-demographic variables are standardized (to adjust for changing variable scales) and many variables (representing similar information) are combined into factors, resulting in a reduced predictor set.
+A spatial error model is then fit for each election, accounting for the inherent spatial structure of the data.
+
+Across the past six elections, most of the socio-demographics that drive the electoral two-party preferred vote are found to remain steady, whilst a few (typically weaker) effects vary over time.
 Industry and type of work are particularly influential, with energy-related and manufacturing/construction jobs, as well as administrative roles being strongly linked with the Liberal party in all elections.
-Incomes have had a similarly consistent effect, with higher income areas supporting Liberal.
-Higher levels of unemployment have shifted from weak association with Labor to a significant Liberal effect over the years, and higher education levels have been associated with Labor since 2007 (although only significant in 2016).
+Incomes have a similarly consistent effect, with higher income areas supporting Liberal.
+Higher levels of unemployment shift from weak association with Labor to a significant Liberal effect over the years, and higher education levels are associated with Labor from 2007 (although only significant in 2016).
 It is also found that electorates with higher household mobility support Liberal, birthplace diversity favours Labor and more de facto relationships align with Labor preference - although marriages, family and household sizes have no material influence.
-Furthermore, the neighbourhood (spatial) effects were found to be positive in all elections, although only significant in 2001 and 2016, meaning that in the 2004-2013 elections, electorates effectively voted independently.
+Furthermore, the neighbourhood (spatial) effects are found to be positive in all elections, although only significant in 2001 and 2016, meaning that in the 2004-2013 elections, electorates effectively voted independently.
 
@@
 It is also the first study to model any Australian election since 2010 using Census information.
-Additionally, a key contribution of this research is the wrangling of the raw data and spatio-temporal imputation, which is applied to obtain Census data for the 2004, 2007, 2010 and 2013 elections.
-All datasets used in this study have been contributed to the `eechidna` `R` package, which provides a rich, accessible data resource for future Australian electoral analysis.
+Additionally, a key contribution of this research is the wrangling of the raw data and imputed data sets for the 2004, 2007, 2010 and 2013 elections, which have been contributed to the `eechidna` `R` package - providing a rich, accessible data resource for future Australian electoral analysis.
 

### 2019-03-31 draft-paper.Rmd
present tense

@@
 This study explores how electoral socio-demographic characteristics relate to two-party preference, and whether their effects have changed over time.
-Electoral socio-demographics are derived from the Australian Census, and vote counts are obtained from federal elections.
+Electoral socio-demographics are derived from the Australian Census, and vote counts are obtained from Australian federal elections.
 
@@
 The predominant approach in previous studies is to join voting outcomes to the nearest Census, without accounting for any temporal differences (see @DavisStimson98, @Stimson06, @Liao09 and @Stimson09).
-Furthermore, electoral boundaries change regularly, so spatial discrepancies also arise when matching electoral data.
+Furthermore, electoral boundaries change regularly, so spatial discrepancies also arise when matching with electoral data.
 To obtain appropriate Census data for these four elections, electoral socio-demographics are constructed using a spatio-temporal imputation that combines areal interpolation [@Goodchild1993] and linear time-interpolation.
@@
 Between 2001 and 2016 there were six elections and four Censuses (see Figure \@ref(fig:timeline)).
-Electoral boundaries are redistributed regularly by the AEC, meaning that only in the years where both a Census and election occur will the boundaries match - the case for the 2001 and 2016 election.
+Electoral boundaries are redistributed regularly by the AEC, meaning that only in the years where both a Census and election occur will the boundaries match - the case for the 2001 and 2016 elections.
 Therefore, for the four elections between 2004 and 2013, both temporal and spatial differences in electorates need to be accounted for when joining the electoral two-party preferred vote with Census data.
@@
                colour="grey50", alpha = 0.4) +
-  scale_fill_manual(name="Boundary", values=c("white", "purple"), labels = c("Other 2013 Electorates", "Hume 2013")) +
+  scale_fill_manual(name="Electoral boundaries", values=c("white", "purple"), labels = c("Other 2013 Electorates", "Hume 2013")) +
   theme_map() + coord_equal()
@@
     alpha = year == "2013", colour=year == "2013")) +
-  scale_fill_manual(name="Boundary", values=c("grey95", "purple"), labels = c("2016 Electorates", "Hume 2013")) +
+  scale_fill_manual(name="Electoral boundaries", values=c("white", "purple"), labels = c("2016 Electorates", "Hume 2013")) +
   scale_alpha_manual(values=c(0, 0.4)) +
   scale_color_manual(values=c("grey50", NA)) +
-  theme_map() + coord_equal() + guides(alpha = F, color = F)
+  theme_map() + coord_equal() + guides(alpha = F, col = F)
 ```
@@
 The partial residuals are the residuals from the fitted model with the estimated effect an individual variable added.
-The estimated intercept is also added to the partial residuals for interpretability.
 These show the direction, size and significance of an estimated effect - the slope of the prediction line matches the estimated coefficient, and the shaded region represents a 95% confidence band.
 If a horizontal line can be drawn through the confidence band, then the effect is insignificant.
+The estimated intercept is also added to the partial residuals for interpretability.
 Plots for each election are faceted to compare the effects over time in Figure \@ref(fig:partresplot1) and Figure \@ref(fig:partresplot2).
@@
 This effect is significant in every election aside from 2004, where it is only marginally insignificant ($p = 0.0613$).
-Unemployment however, has not been as influential.
-In 2001 and 2004, electorates with higher unemployment were more likely to support Labor, but over time this has shifted towards support for the Liberal party, culminating in a significantly positive effect in 2016.
+Unemployment however, is not as influential.
+In 2001 and 2004, electorates with higher unemployment align with Labor, but over time this shifts towards support for the Liberal party, culminating in a significantly positive effect in 2016.
 
@@
 ### Household mobility
-In each of the six elections, electorates with a higher proportion of people that have recently (past five years) moved house were more likely to support the Liberal party, although this effect was marginally insignificant in 2013 (see row 6 in Figure \@ref(fig:partresplot1).
+In each of the six elections, electorates with a higher proportion of people that have recently (past five years) moved house are more likely to support the Liberal party, although this effect was marginally insignificant in 2013 (see row 6 in Figure \@ref(fig:partresplot1).
 Having controlled for characteristics of house ownership and rental prices (via the factors `PropertyOwned` and `RentLoan` respectively), this is effect is somewhat surprising.

### 2019-03-31 draft-paper.Rmd
small change to abstract

@@
 Socio-demographic information is derived from the Australian Census, which occurs every five years.
-Since a Census is not directly available for each election, a method of spatio-temporal imputation is employed to impute Census data for the electorates in each election.
+Since a Census is not directly available for each election, a method of spatio-temporal imputation is employed to impute Census data for the electorates in 2004-2013 elections.
 This accounts for both spatial and temporal changes in electoral characteristics between Censuses.

### 2019-03-31 draft-paper.tex
pdf and tex files

@@
             pdftitle={Spatial modelling of the electoral two-party preferred vote in Australia. A study of federal elections between 2001 and 2016 via the eechidna R package.},
-            pdfkeywords={blah, blah},
+            pdfkeywords={federal election, Census, Australia, spatial modelling, imputation, socio-demographics, electorates, R, eechidna},
             pdfborder={0 0 0},
@@
 \begin{abstract}
-A brief summary of our ideas
+This research examines the relationships between electoral socio-demographic characteristics and two-party preference in Australian federal elections, with the six elections from 2001 to 2016 considered.
+Socio-demographic information is derived from the Australian Census, which occurs every five years.
+Since a Census is not directly available for each election, a method of spatio-temporal imputation is employed to impute Census data for the electorates in 2004-2013 elections.
+This accounts for both spatial and temporal changes in electoral characteristics between Censuses.
+To capture any spatial heterogeneity, a spatial error model is estimated for each election, which incorporates a spatially structured random effect vector that can thought of as the unobserved political climate in each electorate.
+Over time, most socio-demographic characteristics that affect electoral two-party preference do not vary, with industry of work, incomes, household mobility and de facto relationships having strong effects in the six elections.
+Education and unemployment are amongst those that have varying effects.
+It is also found that between 2004 and 2013, the spatial effect is insignificant, meaning that electorates voted effectively independently.
+All data featured in this study has been contributed to the \texttt{eechidna} \texttt{R} package (available on CRAN).
 \end{abstract}
 \begin{keywords}
-blah, blah
+federal election, Census, Australia, spatial modelling, imputation, socio-demographics, electorates, R, eechidna
 \end{keywords}
@@
 The two-party preferred vote, a measure of support between these two parties, fluctuated between 47.3\% and 53.5\% (in favour of the Liberal party) over this period.
-This study explores how electoral characteristics relate to two-party preference, and whether their effects have changed over time.
-Electoral socio-demographics are derived from the Census, and vote counts are obtained from federal elections.
-
-Joining these two data sources is problematic as there in an inherent asynchronicity in the two events.
+This study explores how electoral socio-demographic characteristics relate to two-party preference, and whether their effects have changed over time.
+Electoral socio-demographics are derived from the Australian Census, and vote counts are obtained from Australian federal elections.
+
+Joining these two data sources is problematic as there in an inherent asynchronicity in the two types of events.
 A Census is conducted by the Australian Bureau of Statistics (ABS) every five years, whereas a federal election (conducted by the Australian Electoral Commission (AEC)) usually occurs every three years.
-The first problem addressed is that of obtaining appropriate Census data for the 2004, 2007, 2010 and 2013 elections - election years in which a Census does not occur.
+The first problem addressed is that of constructing appropriate Census data for the 2004, 2007, 2010 and 2013 elections - election years in which a Census does not occur.
 The predominant approach in previous studies is to join voting outcomes to the nearest Census, without accounting for any temporal differences (see \textcite{DavisStimson98}, \textcite{Stimson06}, \textcite{Liao09} and \textcite{Stimson09}).
-Furthermore, electoral boundaries change regularly, so spatial discrepancies also arise when matching electoral data.
-To obtain appropriate Census data for these four elections, electoral socio-demographics are imputed using a spatio-temporal imputation that combines areal interpolation \autocite{Goodchild1993} and linear time-interpolation.
-Collecting and wrangling the raw data, along with the imputation process, are detailed in section \ref{data}.
+Furthermore, electoral boundaries change regularly, so spatial discrepancies also arise when matching with electoral data.
+To obtain appropriate Census data for these four elections, electoral socio-demographics are constructed using a spatio-temporal imputation that combines areal interpolation \autocite{Goodchild1993} and linear time-interpolation.
+Collecting and wrangling the raw data, along with the imputation process, are detailed in Section \ref{data}.
 All data and associated documentation relating to this procedure are available in the \texttt{eechidna} \texttt{R} package \autocite{eechidna}, providing a resource for future analysis .
@@
 Previous work on modelling Australian federal elections have found that aggregate socio-demographics are relatively good predictors of voting outcomes.
-\textcite{Forrest01} does this using multiple regression of the Liberal and Labor primary vote for polling booths in the Farrer electorate in 1998.
+\textcite{Forrest01} use multiple regression to model the Liberal and Labor primary vote for polling booths in the Farrer electorate in 1998 as a function of Census variables from 1996.
 \textcite{Stimson06}, \textcite{Stimson09} and \textcite{Stimson12} use principal component analysis of polling booths in the 2001, 2004 and 2007 elections respectively, also finding that socio-demographic characteristics of polling booths are linked to their two-party preferred vote.
 On the contrary, \textcite{Stimson09} models the polling booth swing vote (change in the two-party preferred vote) in the 2007 election, finding that little of swing vote can be explained by Census data.
-Instead of analyzing a single election in isolation, this paper employs a consistent model framework across six elections so that temporal changes in the effects of socio-demographics can be observed, where each federal elections is modelled with a cross-sectional data set.
+Instead of analyzing a single election in isolation, this paper employs a consistent model framework across six elections so that temporal changes in the effects of socio-demographics can be observed, where each federal elections is modelled with a cross-sectional dataset.
 The use of a regression framework to examine these socio-political relationships over time is seemingly absent from previous Australian studies.
@@
 
-The cross-sectional data set for each election consists of the two-party preferred vote (response variable), and socio-demographic variables (explanatory variables) that characterise each electorate.
-To obtain these cross-sections, socio-demographic variables are first standardized, and then principal components are used to group variables into ``factors''.
-To account for the inherent spatial structure of the data, a spatial error model is fit for each election.
-These steps are discussed in section \ref{modelling}.
-In section \ref{results} inference is conducted on the models to see which effects are significant, how effects change over time and which electorates have abnormal voting behaviour.
+The cross-sectional dataset for each election used here consists of the two-party preferred vote (response variable), and a set of common socio-demographic variables (explanatory variables) that characterize each electorate.
+To obtain these datasets, socio-demographic variables are first standardized, and then a principal component analysis is used to group variables into ``factors''.
+To account for the inherent spatial structure of the data, a spatial error model is then fit for each election.
+These steps are discussed in Section \ref{modelling}.
+In Section \ref{results} inference is conducted to see which effects are significant and how these effects change over time.
+
+Section \ref{conclusion} concludes, followed by two supplementary sections.
 
@@
 Socio-demographic variables are derived from the Census of Population and Housing (Census), which is a survey of every household in Australia, recording information such as age, gender, ethnicity, education level and income.
-There have been four Censuses in the 21st century, being that in 2001, 2006, 2011 and 2016.
+There have been four Censuses so far in the 21st century, conducted in 2001, 2006, 2011 and 2016.
 The Australian Bureau of Statistics (ABS) conducts the Census and publishes aggregated information.
-The ABS approximation of electorates at the time of the Census is chosen.
-From this aggregate information, 67 socio-demographic variables are computed for each of the electorates.
+The ABS uses electoral boundaries as defined by the AEC at the time of each Census, which may not match those in place at the subsequent election.
+From the available Census information aggregated at the electorate level, 65 socio-demographic variables are defined for each of the electorates to be used in the analysis.
 
 Raw data is sourced online from the AEC and ABS websites in .csv and .xlsx files.
-The format of these files change over the years, making extracting the appropriate information a big task.
+The formats of these files differ over the years, making extracting the appropriate information a big task.
 The functions available in the \texttt{dplyr} \autocite{dplyr} and \texttt{readxl} \autocite{readxl} \texttt{R} packages are very useful, as they provide fast consistent tools for data manipulation and functions to import .xlsx files (respectively).
-The 2001 and 2006 Census data are however published in a format where each electorate has a separate document, making it difficult to use the \texttt{dplyr} tools and instead cells have to be selected from each individual file to construct the desired variables.
+The 2001 and 2006 Census data are published in a format where the information for each electorate is held in a separate document making it difficult to use the \texttt{dplyr} tools.
+Instead, cells have to be selected from each individual file to construct the desired variables.
 All scripts required for the data wrangling process can be found in the github repository for the \texttt{eechidna} \texttt{R} package \autocite{eechidna}, along with the raw data.
@@
 Between 2001 and 2016 there were six elections and four Censuses (see Figure \ref{fig:timeline}).
-Electoral boundaries are redistributed regularly by the AEC, meaning that only in the years where both a Census and election occur will the boundaries match - the case for the 2001 and 2016 election.
+Electoral boundaries are redistributed regularly by the AEC, meaning that only in the years where both a Census and election occur will the boundaries match - the case for the 2001 and 2016 elections.
 Therefore, for the four elections between 2004 and 2013, both temporal and spatial differences in electorates need to be accounted for when joining the electoral two-party preferred vote with Census data.
@@
 
-{\centering \includegraphics[width=0.9\linewidth]{skeleton_files/figure-latex/timeline-1}
+{\centering \includegraphics[width=0.9\linewidth]{draft-paper_files/figure-latex/timeline-1}
 
@@
 
+For each election, neighbouring Census information has to be combined in some way so that it represents the boundaries in place at the time of the election.
+This is done by taking the electoral boundaries and imputing the corresponding socio-demographic characteristics for each of neighbouring Censuses, thereby addressing the spatial aspect.
+Next, to deal with the temporal component, characteristcs at the time of the election are constructed using linear interpolation between the spatially imputed neighbouring Census variables.
+
 To account for spatial differences, the piece-wise approximation method in \textcite{Goodchild1993} is adopted.
 Consider a map of source zones \(s = 1,...,S\), for which socio-demographic information is available, and a set of target zones \(t = 1,...,T\) for which information is to be imputed.
-In this context the map of electoral boundaries at the time of a Census would be the source zones, and the boundaries at the time of the election would be the target zones.
-Denote the area of intersection between source zone \(s\) and target zone \(t\) as \(A_{s,t}\), the population of the source zone \(s\) as \(U_s\), and the population of intersection between source zone \(s\) and target zone \(t\) as \(P_{s,t}\).
-
-Compute each \(A_{s,t}\) and estimate population of the intersection:
-
-\[\hat{P}_{s,t} = \frac{U_s*A_{s,t}}{\sum_{t=1}^T A_{s,t}}\]
-This assumes that populations are uniformly distributed within each source zone.
-
-In order to calculate socio-demographic information for each of the target zones, a weighted average is taken using the estimated population as weights.
-Denote a given Census variable for the target zone \(C_t\), and the same Census variable for the source zone \(D_s\):
-
-\[\hat{C}_t = \frac{\sum_{s=1}^{S}{D_s*\hat{P}_{s,t}}}{\sum_{s=1}^{S}{\hat{P}_{s,t}}}\]
-This assumes that each individual in a source zone assumes the aggregate characteristics of the zone.
-
-Applying this to each of the target zones addresses the spatial component, as it imputes the required socio-demographic for the desired electoral boundaries.
-However these are applicable at the time of the Census (source year) and are not yet appropriate for the election (target year).
-
-Denote year \(y\), with a Census falling on \(y_1\) and \(y_3\), and an election on year \(y_2\), and add this subscript to the Census variable estimate, \(\hat{C}_{t,y}\).
-To account for temporal changes, linear interpolation is used between Census years to get the final estimate of a Census variable for the target zone in the election year \(y_2\).
-This assumes that population evolves in a linear manner over time.
-
-\[\hat{C}_{t,y_2} = \frac{y_3-y_2}{y_3-y_1}*\hat{C}_{t,y_1} + \frac{y_2-y_1}{y_3-y_1}*\hat{C}_{t,y_3}\]
-
-\hypertarget{applied}{%
-\subsubsection{Applied}\label{applied}}
-
-Publically available Census data is aggregated and there are different resolutions accessible, ranging from SA1 (over 50,000 zones) to electoral divisions (150 zones).
-For this study electoral divisions are used as source zones, and this imputation method is applied for each of the 2004, 2007, 2010 and 2013 elections.
-To demonstrate its functionality, consider the imputation of socio-demographic variables for the electorate of Hume in New South Wales (NSW), at the time of the 2013 federal election.
-Figure \ref{fig:hume13} shows this region amongst other NSW electorates.
+This is described in the context of a single election, and a single neighbouring Census.
+
+Let the map of electoral boundaries at the time of a Census define the source zones, and let the boundaries at the time of the election be the target zones.
+Denote the area of intersection between a source zone \(s\) and a target zone \(t\) as \(A_{s,t}\).
+Additionally, let the population of the source zone \(s\) be \(U_s\) and the population of intersection between source zone \(s\) and target zone \(t\) be \(P_{s,t}\).
+The estimated population of intersection is given by:
+
+\[\hat{P}_{s,t} = \frac{U_s*A_{s,t}}{\sum_{t=1}^T A_{s,t}}, \mbox{ for all } s=1,...,S \mbox{ and } t=1,...,T.\]
+Note that this estimator implicitly assumes that populations are uniformly distributed within each source zone.
+
+In order to calculate socio-demographic information for each of the target zones, a weighted average is taken using the estimated intersection populations as weights.
+Denote a given Census variable for the target zone by \(C_t\), and the same Census variable for the source zone as \(D_s\).
+Then, estimate \(C_t\) using:
+
+\[\hat{C}_t = \frac{\sum_{s=1}^{S}{D_s*\hat{P}_{s,t}}}{\sum_{s=1}^{S}{\hat{P}_{s,t}}}, \mbox{ for each } t=1,...,T.\]
+This concludes the spatial imputation of the socio-demographic characteristics for one target zone (a single electoral boundary), at time of only one of the neighbouring Censuses.
+This process is repeated for all of the target zones, and then for the other neighbouring Census.
+
+To account for temporal changes, linear interpolation is used between Census years to get the final estimate of a Census variable for the target zone in the election year.
+Let \(y_1\) be the year of the Census preceeding an election, let \(y_2\) be the year of the election and \(y_3\) be the year of the Census that follows.
+Add this year subscript to the Census variable estimate \(\hat{C}_t\), resulting in \(\hat{C}_{t,y}\).
+Linear interpolating between these Census years results an imputed value for the election year, given by
+
+\[\hat{C}_{t,y_2} = \frac{y_3-y_2}{y_3-y_1}*\hat{C}_{t,y_1} + \frac{y_2-y_1}{y_3-y_1}*\hat{C}_{t,y_3}.\]
+Implicitly this assumes that population characteristics change in a linear manner over time.
+
+\hypertarget{an-illustration-of-the-spatio-temporal-imputation}{%
+\subsubsection{An illustration of the spatio-temporal imputation}\label{an-illustration-of-the-spatio-temporal-imputation}}
+
+Census data is publicly available that different levels of aggregation, ranging from SA1 (over 50,000 zones) to electoral divisions (150 zones).
+For this study electoral divisions are used as source zones, and the imputation method is applied to produce socio-demographic variables for each of the 2004, 2007, 2010 and 2013 elections.
+As mentioned earlier, there is no need to impute socio-demographic variables for the 2001 and 2016 elections.
+To illustrate the method, consider the imputation of socio-demographic variables for the electorate of Hume in New South Wales (NSW) at the time of the 2013 federal election.
+The boundaries shown in Figure \ref{fig:hume13} define all target zones in NSW for 2013, with the target zone of interest (Hume) shaded purple.
 
@@
 
-{\centering \includegraphics[height=0.3\textheight]{skeleton_files/figure-latex/hume13-1}
+{\centering \includegraphics[height=0.3\textheight]{draft-paper_files/figure-latex/hume13-1}
 
@@
 
-\caption{Some of the electoral boundaries in NSW for 2013, with the electoral boundary for Hume, shown in purple.}\label{fig:hume13}
+\caption{Some of the electoral boundaries in NSW for 2013, with the electoral boundary for Hume shown in purple.}\label{fig:hume13}
 \end{figure}
 
-The Censuses neighbouring the 2013 election are those in 2011 and 2016, and the Hume boundary is changed, as seen by plotting the Hume boundary (purple) in the 2013 election over the divisions in 2016 (see Figure \ref{fig:hume16}).
+The corresponding source zones from the 2016 Census are shown in Figure \ref{fig:hume16}.
+As can be seen, the Hume boundary from the 2013 election (shaded purple) does not perfectly match any of the source zones.
 
@@
 
-{\centering \includegraphics[height=0.3\textheight]{skeleton_files/figure-latex/hume16-1}
+{\centering \includegraphics[height=0.3\textheight]{draft-paper_files/figure-latex/hume16-1}
 
@@
 
-There are many electorates in 2016 that intersect with the purple region (Hume boundary for 2013), these include the divisions of Riverina, Eden-Monaro and Hume, along with smaller intersecting areas with Fenner, Calare, Gilmore and Whitlam.
-To impute Census information for this purple region, calculate the percentage of each 2016 electorate that intersects with the purple region, which is then used to estimate intersection populations \(\hat{P}_{s,t}\).
+There are many source zones from the 2016 Census that intersect with this purple region, including the divisions of Riverina, Eden-Monaro and Hume, along with smaller intersecting areas with Fenner, Calare, Gilmore and Whitlam.
+The proportion of each source zone that overlaps with the purple region is calculated, and used to obtain the intersecting populations \(\hat{P}_{s,t}\).
+
+\footnotesize
 
@@
 \toprule
-\begin{minipage}[b]{0.17\columnwidth}\raggedright
-Electorate (2016)\strut
+\begin{minipage}[b]{0.21\columnwidth}\raggedright
+Source Zone (2016)\strut
 \end{minipage} & \begin{minipage}[b]{0.10\columnwidth}\raggedright
-Percentage\strut
-\end{minipage} & \begin{minipage}[b]{0.22\columnwidth}\raggedright
-Population in Electorate\strut
-\end{minipage} & \begin{minipage}[b]{0.40\columnwidth}\raggedright
-Estimated Population Allocated to Purple Region: \(\hat{P}_{s,t}\)\strut
+Proportion\strut
+\end{minipage} & \begin{minipage}[b]{0.11\columnwidth}\raggedright
+Source Zone Population\strut
+\end{minipage} & \begin{minipage}[b]{0.46\columnwidth}\raggedright
+Population Allocated to Purple Region: \(\hat{P}_{s,t}\)\strut
 \end{minipage}\tabularnewline
@@
 \endhead
-\begin{minipage}[t]{0.17\columnwidth}\raggedright
+\begin{minipage}[t]{0.21\columnwidth}\raggedright
 HUME\strut
 \end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright
-96.54\%\strut
-\end{minipage} & \begin{minipage}[t]{0.22\columnwidth}\raggedright
+0.9654\strut
+\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright
 150643\strut
-\end{minipage} & \begin{minipage}[t]{0.40\columnwidth}\raggedright
+\end{minipage} & \begin{minipage}[t]{0.46\columnwidth}\raggedright
 145427\strut
 \end{minipage}\tabularnewline
-\begin{minipage}[t]{0.17\columnwidth}\raggedright
+\begin{minipage}[t]{0.21\columnwidth}\raggedright
 RIVERINA\strut
 \end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright
-25.11\%\strut
-\end{minipage} & \begin{minipage}[t]{0.22\columnwidth}\raggedright
+0.2511\strut
+\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright
 155793\strut
-\end{minipage} & \begin{minipage}[t]{0.40\columnwidth}\raggedright
+\end{minipage} & \begin{minipage}[t]{0.46\columnwidth}\raggedright
 39117\strut
 \end{minipage}\tabularnewline
-\begin{minipage}[t]{0.17\columnwidth}\raggedright
+\begin{minipage}[t]{0.21\columnwidth}\raggedright
 EDEN-MONARO\strut
 \end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright
-11.09\%\strut
-\end{minipage} & \begin{minipage}[t]{0.22\columnwidth}\raggedright
+0.1109\strut
+\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright
 147532\strut
-\end{minipage} & \begin{minipage}[t]{0.40\columnwidth}\raggedright
+\end{minipage} & \begin{minipage}[t]{0.46\columnwidth}\raggedright
 16358\strut
 \end{minipage}\tabularnewline
-\begin{minipage}[t]{0.17\columnwidth}\raggedright
+\begin{minipage}[t]{0.21\columnwidth}\raggedright
 CANBERRA\strut
 \end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright
-0.28\%\strut
-\end{minipage} & \begin{minipage}[t]{0.22\columnwidth}\raggedright
+0.028\strut
+\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright
 196037\strut
-\end{minipage} & \begin{minipage}[t]{0.40\columnwidth}\raggedright
+\end{minipage} & \begin{minipage}[t]{0.46\columnwidth}\raggedright
 548\strut
 \end{minipage}\tabularnewline
-\begin{minipage}[t]{0.17\columnwidth}\raggedright
+\begin{minipage}[t]{0.21\columnwidth}\raggedright
 FENNER\strut
 \end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright
-0.23\%\strut
-\end{minipage} & \begin{minipage}[t]{0.22\columnwidth}\raggedright
+0.0023\strut
+\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright
 202955\strut
-\end{minipage} & \begin{minipage}[t]{0.40\columnwidth}\raggedright
+\end{minipage} & \begin{minipage}[t]{0.46\columnwidth}\raggedright
 474\strut
 \end{minipage}\tabularnewline
-\begin{minipage}[t]{0.17\columnwidth}\raggedright
+\begin{minipage}[t]{0.21\columnwidth}\raggedright
 WHITLAM\strut
 \end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright
-0.06\%\strut
-\end{minipage} & \begin{minipage}[t]{0.22\columnwidth}\raggedright
+0.0006\strut
+\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright
 152280\strut
-\end{minipage} & \begin{minipage}[t]{0.40\columnwidth}\raggedright
+\end{minipage} & \begin{minipage}[t]{0.46\columnwidth}\raggedright
 92\strut
 \end{minipage}\tabularnewline
-\begin{minipage}[t]{0.17\columnwidth}\raggedright
+\begin{minipage}[t]{0.21\columnwidth}\raggedright
 GILMORE\strut
 \end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright
-0.06\%\strut
-\end{minipage} & \begin{minipage}[t]{0.22\columnwidth}\raggedright
+0.0006\strut
+\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright
 150436\strut
-\end{minipage} & \begin{minipage}[t]{0.40\columnwidth}\raggedright
+\end{minipage} & \begin{minipage}[t]{0.46\columnwidth}\raggedright
 86\strut
 \end{minipage}\tabularnewline
-\begin{minipage}[t]{0.17\columnwidth}\raggedright
+\begin{minipage}[t]{0.21\columnwidth}\raggedright
 CALARE\strut
 \end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright
-0.01\%\strut
-\end{minipage} & \begin{minipage}[t]{0.22\columnwidth}\raggedright
+0.0001\strut
+\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright
 161298\strut
-\end{minipage} & \begin{minipage}[t]{0.40\columnwidth}\raggedright
+\end{minipage} & \begin{minipage}[t]{0.46\columnwidth}\raggedright
 21\strut
@@
 
-Now consider the socio-demographic \(AusCitizen\) - the proportion of people in the region who are Australian citizens.
+\normalsize
+
+Now consider the socio-demographic variable \(AusCitizen\) - the proportion of people in the region who are Australian citizens.
+
+\footnotesize
 
@@
 \toprule
-\begin{minipage}[b]{0.18\columnwidth}\raggedright
-DivisionNm\strut
-\end{minipage} & \begin{minipage}[b]{0.21\columnwidth}\raggedright
+\begin{minipage}[b]{0.26\columnwidth}\raggedright
+Source Zone (2016)\strut
+\end{minipage} & \begin{minipage}[b]{0.19\columnwidth}\raggedright
 AusCitizen (\%): \(D_s\)\strut
-\end{minipage} & \begin{minipage}[b]{0.53\columnwidth}\raggedright
+\end{minipage} & \begin{minipage}[b]{0.47\columnwidth}\raggedright
 Estimated Population Allocated to Purple Region: \(\hat{P}_{s,t}\)\strut
@@
 \endhead
-\begin{minipage}[t]{0.18\columnwidth}\raggedright
+\begin{minipage}[t]{0.26\columnwidth}\raggedright
 HUME\strut
-\end{minipage} & \begin{minipage}[t]{0.21\columnwidth}\raggedright
+\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedright
 90.02\strut
-\end{minipage} & \begin{minipage}[t]{0.53\columnwidth}\raggedright
+\end{minipage} & \begin{minipage}[t]{0.47\columnwidth}\raggedright
 145427\strut
 \end{minipage}\tabularnewline
-\begin{minipage}[t]{0.18\columnwidth}\raggedright
+\begin{minipage}[t]{0.26\columnwidth}\raggedright
 RIVERINA\strut
-\end{minipage} & \begin{minipage}[t]{0.21\columnwidth}\raggedright
+\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedright
 89.11\strut
-\end{minipage} & \begin{minipage}[t]{0.53\columnwidth}\raggedright
+\end{minipage} & \begin{minipage}[t]{0.47\columnwidth}\raggedright
 39117\strut
 \end{minipage}\tabularnewline
-\begin{minipage}[t]{0.18\columnwidth}\raggedright
+\begin{minipage}[t]{0.26\columnwidth}\raggedright
 EDEN-MONARO\strut
-\end{minipage} & \begin{minipage}[t]{0.21\columnwidth}\raggedright
+\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedright
 88.00\strut
-\end{minipage} & \begin{minipage}[t]{0.53\columnwidth}\raggedright
+\end{minipage} & \begin{minipage}[t]{0.47\columnwidth}\raggedright
 16358\strut
 \end{minipage}\tabularnewline
-\begin{minipage}[t]{0.18\columnwidth}\raggedright
+\begin{minipage}[t]{0.26\columnwidth}\raggedright
 CANBERRA\strut
-\end{minipage} & \begin{minipage}[t]{0.21\columnwidth}\raggedright
+\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedright
 85.48\strut
-\end{minipage} & \begin{minipage}[t]{0.53\columnwidth}\raggedright
+\end{minipage} & \begin{minipage}[t]{0.47\columnwidth}\raggedright
 548\strut
 \end{minipage}\tabularnewline
-\begin{minipage}[t]{0.18\columnwidth}\raggedright
+\begin{minipage}[t]{0.26\columnwidth}\raggedright
 FENNER\strut
-\end{minipage} & \begin{minipage}[t]{0.21\columnwidth}\raggedright
+\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedright
 83.64\strut
-\end{minipage} & \begin{minipage}[t]{0.53\columnwidth}\raggedright
+\end{minipage} & \begin{minipage}[t]{0.47\columnwidth}\raggedright
 474\strut
 \end{minipage}\tabularnewline
-\begin{minipage}[t]{0.18\columnwidth}\raggedright
+\begin{minipage}[t]{0.26\columnwidth}\raggedright
 WHITLAM\strut
-\end{minipage} & \begin{minipage}[t]{0.21\columnwidth}\raggedright
+\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedright
 89.52\strut
-\end{minipage} & \begin{minipage}[t]{0.53\columnwidth}\raggedright
+\end{minipage} & \begin{minipage}[t]{0.47\columnwidth}\raggedright
 92\strut
 \end{minipage}\tabularnewline
-\begin{minipage}[t]{0.18\columnwidth}\raggedright
+\begin{minipage}[t]{0.26\columnwidth}\raggedright
 GILMORE\strut
-\end{minipage} & \begin{minipage}[t]{0.21\columnwidth}\raggedright
+\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedright
 89.03\strut
-\end{minipage} & \begin{minipage}[t]{0.53\columnwidth}\raggedright
+\end{minipage} & \begin{minipage}[t]{0.47\columnwidth}\raggedright
 86\strut
 \end{minipage}\tabularnewline
-\begin{minipage}[t]{0.18\columnwidth}\raggedright
+\begin{minipage}[t]{0.26\columnwidth}\raggedright
 CALARE\strut
-\end{minipage} & \begin{minipage}[t]{0.21\columnwidth}\raggedright
+\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedright
 87.56\strut
-\end{minipage} & \begin{minipage}[t]{0.53\columnwidth}\raggedright
+\end{minipage} & \begin{minipage}[t]{0.47\columnwidth}\raggedright
 21\strut
@@
 
-Then taking a weighted average of \(AusCitizen\) using the estimated population as weights yields \(\hat{C}_{Hume,2016} = 89.65 \%\).
-Repeating this process using the 2011 Census and electoral boundaries yields \(\hat{C}_{Hume,2011} = 91.00 \%\)
-
-Finally, linearly interpolate between 2011 and 2016 to arrive at the 2013 estimate:
+\normalsize
+
+A weighted average of \(AusCitizen\), with the allocated population from each source zone as weights, yields \(\hat{C}_{Hume,2016} = 89.65 \%\).
+Repeating this process using the 2011 Census yields \(\hat{C}_{Hume,2011} = 91.00 \%\)
+
+Finally, linear interpolation between 2011 and 2016 yields the 2013 estimate:
 \begin{eqnarray*}
@@
 
-\hypertarget{data-pre-processing}{%
-\subsection{Data pre-processing}\label{data-pre-processing}}
-
-With socio-demographic information now available for each electorate, each election is joined to the data corresponding with its two-party preferred vote.
-Socio-demographic variables within each election year are standardized to have mean zero and variance one, to adjust for changing variable scales.
+\emph{1: HERE Introductory paragraph}
+Electoral socio-demographic variables are now available for each of the six elections and are joined with their corresponding two-party preferred votes.
+Before choosing an appropriate model, two issues with the socio-demographic variables are addressed.
+Firstly, variable scales change over the years and as such the data is standardized accordingly.
+Secondly, many variables represent similar information and where appropriate, should be combined in some way.
+Principal component analysis is used to identify variables covary, from which intuitive groupings are selected to be combined into a single variable.
+This also reduces the dimension of the data.
+After these steps, a model specification is chosen.
+
+\hypertarget{standardizing-variables}{%
+\subsection{Standardizing variables}\label{standardizing-variables}}
+
+Many of the socio-demographic variables have changing scales over the years.
 For example, inflation-adjusted median rental prices increased across almost all electorates, with median rent of 200 dollars per week placing an electorate in the 90th percentile in 2001, but only the 30th percentile in 2016.
-
-\hypertarget{dimension-reduction}{%
-\subsubsection{Dimension reduction}\label{dimension-reduction}}
-
-With only \(N = 150\) observations (electorates) in each election and \(p = 65\) socio-demographic variables in each cross-section, any model using all variables would face serious problems with multi-collinearity and over-fitting, likely leading to erroneous conclusions regarding variable significance.
-Therefore a form of dimension reduction is adopted before models are fit.
-
-Socio-demographic variables\footnote{A preliminary step involved removing all age bands, because age is represented by median age, and to remove variables relating to particular denominations of Christianity.} that represent similar information are combined into ``factors'' using principal component analysis (PCA).
-The scree plots of the principal components for each election all level off after four components, and the loadings of these four components are similar across the elections.
-Principal components are then computed on the combined set of socio-demographics across all six elections.
-A factor is created by combining several variables all have large loadings in a particular component and when there is an intuitive reason as to why these variables could represent common information.
-A loading with magnitude greater than 0.15 is considered large.
-After computing these sums, each factor is again standardized to have mean zero and variance one, within each election.
-
-Consider the \texttt{Incomes} factor as an illustration.
-Independent of principal components, we may suspect that median personal income, median household income and median family income are providing similar information about the financial wellbeing of an electorate.
-Their loadings in the first principal component are large (0.19, 0.21 and 0.22 respectively), which provides the evidence needed to combine these variables into a single factor, which is called \texttt{Incomes}.
-
-This process reduces the predictor set to \(p = 30\).
-
-\hypertarget{model-framework}{%
-\subsection{Model framework}\label{model-framework}}
-
-An identical model specification is used across the six elections, with each election modelled separately.
-This allows for the socio-demographic effects to be estimated separately for each year, allowing for interpretation of temporal changes in these effects.
-This is preferable over a single longitudinal model because it avoids any concerns of undue bias stemming from an incorrectly imposed time-varying restriction on any variable.
+In order for socio-demographic effects to be comparable across years, the explanatory variables are standardized to have mean zero and variance one within each election year.
+By standardizing, each variable is reported as a relative measure compared to all other electorates in the same year.
+
+\hypertarget{creating-factors}{%
+\subsection{Creating factors}\label{creating-factors}}
+
+There are only \(N = 150\) observations (electorates) in each election and \(p = 65\) socio-demographic variables in each cross-section, with many variables represent similar information about an electorate.
+Any model that uses all variables would face serious problems with multi-collinearity and over-fitting, which would likely lead to erroneous conclusions regarding variable significance.
+To address this, groups of variables that represent similar information are combined into ``factors''\footnote{A preliminary step involved removing all age bands, because age is represented by median age, and to remove variables relating to particular denominations of Christianity.}.
+
+A factor is created from a group of variables if there is an intuitive reason as to why they should represent similar information and if there is evidence to suggest that they covary.
+For example, a potential group would be variables relating to electoral incomes - median family, household and personal incomes.
+To determine which variables covary, principal component analysis is used on a combined dataset of socio-demographic variables from all six elections.
+It is appropriate to compute principal components in this way because when computed separately for each election, scree plots level off after four components and the loadings of the first four components are similar across the elections.
+
+\emph{2: HERE - ADD JUSTIFICATION FOR DOING SINGLE PCA AND EXPLAIN FACTORS CREATED}
+
+Only the first four principal components from the combined dataset are considered, as the scree plot levels off after the fourth component.
+Variables that have a large loading in a particular component are deemed to covary, with a loading with magnitude greater than 0.15 being considered large.
+Six factors are created using this criteria.
+These are; \texttt{Incomes} - median personal, household and family incomes, \texttt{Unemployment} - unemployment and labor force participation rates, \texttt{ProportyOwned} - rates of housing ownership, mortgages, renting and government housing, \texttt{RentLoanPrice} - median rental and loan repayments, \texttt{FamHouseSize} - average household size, ratio of people to families and household makeup (single person, couple with kids and couple without kids) and \texttt{Education} - high school and university qualifications, jobs requiring higher levels of education as well as vocational course completions and jobs that do not require higher education levels (e.g.~laborer, tradesperson).
+For each of these groups, variables with positive loadings are added and those with negative loadings are substracted to create a factor.
+After computing these sums, each factor is standardized to have mean zero and variance one, within each election.
+
+There are \(p=30\) variables in the resultant predictor set, with all of these used in the regression for each election.
+
+\hypertarget{regression-incorporating-spatially-dependent-errors}{%
+\subsection{Regression incorporating spatially dependent errors}\label{regression-incorporating-spatially-dependent-errors}}
+
+An identical model specification is used for each of the six elections, with each election modelled separately.
+This allows for the socio-demographic effects to be estimated separately for each election year, facilitating analysis of temporal changes in variable effects.
+This approach is preferable to using a single longitudinal model because it avoids any concerns of undue bias stemming from incorrectly imposed time-varying restrictions on any variable.
 Without such restrictions, a pooled cross-sectional model does not yield any distinct advantage over separate cross-sections.
-The panel approach is avoided because of how frequently electoral boundaries change, meaning that electorates that have the same name across elections are not guaranteed to represent the same geographical region.
-Therefore any fixed or random effects models would be difficult to estimate without implementing consistent boundaries, which would requiring further imputation.
-
-For each cross-section, let the response variable be the two-party preferred vote in favour of the Liberal party, denoted \(Y\), with \(Y = 70\) representing a 70\% preference for Liberal, 30\% for Labor.
-Although \(Y\) lies in the interval \((0,100)\), observed values are never very close to 0 or 100 (minimum \(24.05 \%\) and maximum \(74.90 \%\)), so there is no need to impose the constraint of \(Y \in [0,100]\).
-Furthermore, the response is found to be spatially correlated in each election (Moran's I test, \(p \le 7\cdot10^{-15}\)).
-This is expected, as electorates are aggregate spatial units, and hence the spatial structure of the data must modelled appropriately.
+The panel approach is avoided because of how frequently electoral boundaries change, noting that electorates with same name across elections are not guaranteed to represent the same geographical region.
+Therefore any fixed or random effects models would be difficult to estimate without implementing consistent boundaries, which would require further imputation.
+
+For each cross-section, let the response \(\boldsymbol y\) be the vector two-party preferred vote in favour of the Liberal party, with \(y_i = 70\) representing a 70\% preference for Liberal, 30\% for Labor, in electorate \(i\).
+Although \(y_i\) lies in the interval \((0,100)\), observed values are never very close to 0 or 100 (minimum \(24.05 \%\) and maximum \(74.90 \%\)), so there is no need to formally impose the constraint of \(y_i \in [0,100]\).
+Furthermore, the responses are found to be spatially correlated in each election (Moran's I test, \(p \le 7\cdot10^{-15}\)).
+This not surprising as electorates are aggregate spatial units, and hence the spatial structure of the data must modelled appropriately.
 
 The spatial error model \autocite{Anselin88} is chosen because captures spatial heterogeneity by incorporating a spatially structured random effect vector \autocite{LeSage2009}.
-In this context, the random effect can be thought of as capturing the unobserved political climate in each electorate, where the climate is correlated with the climate in neighbouring electorates.
-This functions under the assumption that the climate is independent of electoral socio-demographics, and that an electorate is equally correlated with any electorate that shares a part of its boundary.
-Spatial weights are calculated in accordance with these assumptions.
-The spatial error model is specified as follows:
-
+In this context, the random effect can be thought of as capturing the unobserved political climate in each electorate, where this climate is correlated with the climate in neighbouring electorates, under the assumption that the climate is independent of electoral socio-demographics.
+
+Spatial weights are calculated in accordance with the assumption that an electorate is equally correlated with any electorate that shares a part of its boundary.
 Let \(\rho\) be spatial autoregressive coefficient, \(\boldsymbol v\) be a spherical error term, \({\boldsymbol W}\) be a matrix of spatial weights (containing information about the neighbouring regions), \(\boldsymbol X\) be a matrix of socio-demographic covariates, \(\boldsymbol \beta\) be a vector of regression coefficients and \(\boldsymbol a\) be a spatially structured random effect vector.
 
-\[{\boldsymbol y} = {\boldsymbol X} {\boldsymbol \beta} + {\boldsymbol a}\]
+\[{\boldsymbol y} = {\boldsymbol X} {\boldsymbol \beta} + {\boldsymbol a},\]
 and
 
-\[{\boldsymbol a} = \rho {\boldsymbol W} {\boldsymbol a} + {\boldsymbol v}\]
+\[{\boldsymbol a} = \rho {\boldsymbol W} {\boldsymbol a} + {\boldsymbol v},\]
 where
 
-\[{\boldsymbol v} \sim N({\boldsymbol 0}, \sigma^2 {\boldsymbol I_n})\].
-
-so it can be written
-
-\[{\boldsymbol y} = {\boldsymbol X} {\boldsymbol \beta} + ({\boldsymbol I}_n-\rho {\boldsymbol W})^{-1}{\boldsymbol v}\]
-
-Estimation is done using feasible generalized least squares.
-
-Table 3 details the resultant estimated model coefficients and their estimated standard errors for each of the six elections. These are interpreted in the next section.
+\[{\boldsymbol v} \sim N({\boldsymbol 0}, \sigma^2 {\boldsymbol I_n}),\].
+
+and hence
+
+\[{\boldsymbol y} = {\boldsymbol X} {\boldsymbol \beta} + ({\boldsymbol I}_n-\rho {\boldsymbol W})^{-1}{\boldsymbol v}.\]
+
+Estimation of the above spatial error model is undertaken using feasible generalized least squares.
+
+Table 3 details the estimated model coefficients and their estimated standard errors, for each of the six elections. An interpretation of these estimated values is provided in the next section.
 
 \begin{table}[!htbp] \centering 
-  \caption{Estimated model for each of the six elections.} 
+  \caption{Estimated spatial error model parameters (standard errors) for each of the six election years.} 
   \label{} 
@@
 
-The spatial autoregressive coefficient \(\rho\) is positive and significant in only the 2001 and 2016 elections (Figure \ref{fig:rhovis}), meaning that in these elections, an electorate's political climate was affected by the attitudes of it's neighbours.
-Conversely, in the other four elections, the spatial effect weakens to become insignficant.
-In these years, it appears that the spatial component does not explain anything over and above the electoral socio-demographics, meaning electorates voted effectively independently.
+The spatial autoregressive coefficient \(\rho\) is positive and significant in only the 2001 and 2016 elections (Figure \ref{fig:rhovis}), meaning that, on average, in these elections, the political climate of an electorate appears to be affected by the attitudes of it's neighbours.
+Conversely, in the other four elections, the spatial effect weakens to become insignificant.
+In these years, it appears that the spatial component does not explain anything over and above the electoral socio-demographics, meaning electorates effectively voted independently.
 
@@
 
-{\centering \includegraphics[height=0.15\textheight,]{skeleton_files/figure-latex/rhovis-1}
+{\centering \includegraphics[height=0.15\textheight,]{draft-paper_files/figure-latex/rhovis-1}
 
@@
 
-\caption{Estimates of the spatial autoregressive parameter for each of the six elections, with a 95\% confidence interval.
-Only in 2001 and 2016 is there a significant spatial component}\label{fig:rhovis}
+\caption{Estimates of the spatial autoregressive parameter for each of the six elections, reported with their individual 95\% confidence intervals.
+Only in 2001 and 2016 is there a significant spatial component.}\label{fig:rhovis}
 \end{figure}
@@
 Since all socio-demographics have been standardized to have a mean of zero and a variance of one, the intercept in each model can be interpreted as the estimated two-party preferred vote for an electorate with mean characteristics\footnote{Mean of all variables aside from Judaism, Indigenous, Islam and Buddhism, where it assumes the mean of the log value.}.
-The baseline of party preference has varied over the elections, with the biggest swing occuring in the 2007 election where the mean electorate shifted more than five percentage points in favour of the Labor party.
+The baseline of party preference has varied over the elections, with the biggest swing occurring in the 2007 election where the mean electorate shifted more than five percentage points in favour of the Labor party.
 
@@
 
-{\centering \includegraphics[height=0.15\textheight,]{skeleton_files/figure-latex/plotintercept-1}
+{\centering \includegraphics[height=0.15\textheight,]{draft-paper_files/figure-latex/plotintercept-1}
 
@@
 
-To investigate the socio-demographics that have a strong effect on the two-party preferred vote, partial residual plots are used.
+To investigate the socio-demographics that have a strong effect on the two-party preferred vote, partial residual plots are used and shown in Figures \ref{fig:partresplot1} and \ref{fig:partresplot2}.
+The partial residuals are the residuals from the fitted model with the estimated effect an individual variable added.
 These show the direction, size and significance of an estimated effect - the slope of the prediction line matches the estimated coefficient, and the shaded region represents a 95\% confidence band.
 If a horizontal line can be drawn through the confidence band, then the effect is insignificant.
+The estimated intercept is also added to the partial residuals for interpretability.
 Plots for each election are faceted to compare the effects over time in Figure \ref{fig:partresplot1} and Figure \ref{fig:partresplot2}.
 
-It is important here to note the ecological fallacy - insights are being drawn at the electorate level, and cannot be inferred for another disaggregate level (e.g.~individual voters).
+It is important here to note the ecological fallacy: insights are being drawn at the electorate level, and cannot be inferred for another disaggregate level (e.g.~individual voters).
 
@@
 Typically the Labor party campaigns on more progressive policies, which often include tax reform that adversely affects higher income earners, and more generous social assistance programs.
-Perhaps due to these policies, higher income electorates have been more likely to support the Liberal party, as the \texttt{Incomes} factor has a positive effect on Liberal preference (see row 1 in Figure \ref{fig:partresplot1}).
-This effect is signficant in every election aside from 2004, in which it is only marginally insignificant (\(p = 0.0613\)).
-Unemployment however, has not been as influential.
-In 2001 and 2004, electorates with higher unemployment were more likely to support Labor, but over time this has shifted towards support for the Liberal party, culminating in a significantly positive effect in 2016.
+Perhaps it is due to these policies, that higher income electorates appear more likely to support the Liberal party, as the \texttt{Incomes} factor has a positive effect on Liberal preference (see row 1 in Figure \ref{fig:partresplot1}).
+This effect is significant in every election aside from 2004, where it is only marginally insignificant (\(p = 0.0613\)).
+Unemployment however, is not as influential.
+In 2001 and 2004, electorates with higher unemployment align with Labor, but over time this shifts towards support for the Liberal party, culminating in a significantly positive effect in 2016.
 
@@
 
-Electorates with higher proportions of workers in mining, gas, water, agriculture, waste and electricity (grouped as \texttt{Extractive} industries) are consistently linked with higher support for the Liberal party, with the mangitude of this effect slightly increasing over the years (see row 3 in Figure \ref{fig:partresplot1}).
+Electorates with higher proportions of workers in mining, gas, water, agriculture, waste and electricity (grouped as \texttt{Extractive} industries) are consistently linked with higher support for the Liberal party, with the magnitude of this effect slightly increasing over the years (see row 3 in Figure \ref{fig:partresplot1}).
 This is unsurprising, as the Liberal party has close ties with these traditional energy industries, and typically present policies to reduce taxation on energy production.
@@
 Similarly, workers in managerial, administrative, clerical and sales roles (\texttt{ManagerAdminClericalSales}) is also a significant predictor of two-party preference vote across all six elections, with a higher proportion of people working these jobs increasing Liberal support.
-The magnitude of this effect has also increased over the years.
+The magnitude of this effect also seems to increase over the years.
 
@@
 
-In each of the six elections, electorates with a higher proportion of people that have recently moved house (meaning in last five years) were more likely to support the Liberal party, although this effect was marginally insignificant in 2013 (see row 6 in Figure \ref{fig:partresplot1}.
+In each of the six elections, electorates with a higher proportion of people that have recently (past five years) moved house are more likely to support the Liberal party, although this effect was marginally insignificant in 2013 (see row 6 in Figure \ref{fig:partresplot1}.
 Having controlled for characteristics of house ownership and rental prices (via the factors \texttt{PropertyOwned} and \texttt{RentLoan} respectively), this is effect is somewhat surprising.
@@
 De facto relationships, but not marriages, are found to be an important (and significant) predictor of the two-party preferred vote in all six elections, with more de facto relationships associated with higher support for the Labor party.
-Marriages however, are insignficiant.
+The proportion of individuals who are marriaged however, are insignificant (not shown).
 
@@
 
-Older regions are often believed to be more conservative, and it can be seen that electorates with a higher median age have been more likely to support the Liberal party - although this effect is only significant in 2007 and 2010 (see row 2 in Figure \ref{fig:partresplot2}).
+Regions comprised of older people are often believed to be more conservative, and indeed it found that electorates with a higher median age are more likely to support the Liberal party - although this effect is only significant in 2007 and 2010 (see row 2 in Figure \ref{fig:partresplot2}).
 
@@
 
-Since 2007, electorates with higher education levels have been associated with supporting the Labor party, although this effect is only significant in 2016.
-Prior to 2007, education had an almost zero effect (see row 3 in \ref{fig:partresplot1}).
+Since 2007, electorates with higher education levels are associated with supporting the Labor party, although this effect is only significant in 2016.
+Before 2007, education has an almost zero effect (see row 3 in Figure \ref{fig:partresplot1}).
 
@@
 
-Larger migrant populations from Asia, the Middle East, South-Eastern Europe, the United Kingdom and elsewhere, are either assocated with Labor support, or have no effect.
-Of these areas, only South-Eastern European populations were signficant in each election and Asian migrants were a significant in 2010.
-Speaking other language (aside from English) however, have a far stronger effect, as observed through \texttt{OtherLanguageHome}.
-Electorates with more diverse speech were more likely to support the Liberal party from 2004 onwards, with this effect being significant in 2007, 2010 and 2016.
-Furthermore, of the variables relating to religion, only Judaism shows a consistent effect, with electorates with larger Jewish populations more likely to vote Liberal.
+Larger migrant populations from Asia, the Middle East, South-Eastern Europe, the United Kingdom and elsewhere, are either associated with Labor support, or have no effect.
+Of these areas, only South-Eastern European populations appear significant in each election, with the proportion of Asian migrants also being significant in 2010.
+Speaking other languages (aside from English) however, appears to have a far stronger effect, as observed through the \texttt{OtherLanguageHome} variable.
+Electorates with more diverse speech are associated with higher support for the Liberal party from 2004 onwards, with this effect being significant in 2007, 2010 and 2016.
+Furthermore, of the variables relating to religion, only Judaism shows a consistent effect, with electorates with relatively large Jewish populations more likely to vote Liberal.
 
@@
 However, there still remain some variables which are closely related.
-For example, electorate income levels (via \texttt{Incomes}) is likely to be related to electoral unemployment and labor force participation (via \texttt{Unemployment}).
-In 2001, the coefficient estiamte for \texttt{Unemployment} negative but not significant, and the \texttt{Incomes} variables is significant.
-If the \texttt{Incomes} variable is removed from the model, \texttt{Unemployment} absorbs the negative effect, becoming significant (\(p = 0.0056\)).
+For example, an electorate's income level (via \texttt{Incomes}) is likely to be related to electoral unemployment and labor force participation (via \texttt{Unemployment}).
+In 2001, the coefficient estimate for \texttt{Unemployment} is negative but not significant, whilst the \texttt{Incomes} variables is significant.
+If the \texttt{Incomes} variable is removed from the model in 2001, \texttt{Unemployment} absorbs the negative effect, becoming significant (\(p = 0.0056\)).
 
@@
 
-{\centering \includegraphics[width=0.8\linewidth,]{skeleton_files/figure-latex/resstate-1}
+{\centering \includegraphics[width=0.8\linewidth,]{draft-paper_files/figure-latex/resstate-1}
 
@@
 
-Across the past six elections, most of the socio-demographics that drive the electoral two-party preferred vote have remained steady, whilst a few (typically weaker) effects have varied over time.
+\emph{3: HERE Add paragraph on the overall strategy and contributions - can restate stuff from intro}
+
+This paper explores the effects of electoral socio-demographic characteristics on the two-party preferred vote in the 2001-2016 elections, using information from the corresponding Australian federal elections and Censuses.
+As a Census does not always occur in the same year as an election, Census data for the 2004-2013 elections are generated by employing a method of spatio-temporal imputation.
+This imputes electoral socio-demographics for the electoral boundaries in place at the time of the election - an approach that is distinctly different from previous work on modelling election outcomes, where Census and election data are typically joined without addressing their temporal differences.
+Before estimating a model, these socio-demographic variables are standardized (to adjust for changing variable scales) and many variables (representing similar information) are combined into factors, resulting in a reduced predictor set.
+A spatial error model is then fit for each election, accounting for the inherent spatial structure of the data.
+
+Across the past six elections, most of the socio-demographics that drive the electoral two-party preferred vote are found to remain steady, whilst a few (typically weaker) effects vary over time.
 Industry and type of work are particularly influential, with energy-related and manufacturing/construction jobs, as well as administrative roles being strongly linked with the Liberal party in all elections.
-Incomes have had a similarly consistent effect, with higher income areas supporting Liberal.
-Higher levels of unemployment have shifted from weak association with Labor to a significant Liberal effect over the years, and higher education levels have been associated with Labor since 2007 (although only significant in 2016).
+Incomes have a similarly consistent effect, with higher income areas supporting Liberal.
+Higher levels of unemployment shift from weak association with Labor to a significant Liberal effect over the years, and higher education levels are associated with Labor from 2007 (although only significant in 2016).
 It is also found that electorates with higher household mobility support Liberal, birthplace diversity favours Labor and more de facto relationships align with Labor preference - although marriages, family and household sizes have no material influence.
-Furthermore, the neighbourhood (spatial) effects were found to be positive in all elections, although only significant in 2001 and 2016, meaning that in the 2004-2013 elections, electorates effectively voted independently.
+Furthermore, the neighbourhood (spatial) effects are found to be positive in all elections, although only significant in 2001 and 2016, meaning that in the 2004-2013 elections, electorates effectively voted independently.
 
@@
 It is also the first study to model any Australian election since 2010 using Census information.
-Additionally, a key contribution of this research is the spatio-temporal imputation, which is applied to obtain Census data for the 2004, 2007, 2010 and 2013 elections.
-All data sets used in this study have been contributed to the \texttt{eechidna} \texttt{R} package, which provides a rich, accessible data resource for future Australian electoral analysis.
+Additionally, a key contribution of this research is the wrangling of the raw data and imputed data sets for the 2004, 2007, 2010 and 2013 elections, which have been contributed to the \texttt{eechidna} \texttt{R} package - providing a rich, accessible data resource for future Australian electoral analysis.
 
@@
 
-This paper was produced using \texttt{RMarkdown}\autocite{rmarkdown} and \texttt{knitr} \autocite{knitr}.
+This paper was produced using \texttt{RMarkdown} \autocite{rmarkdown} and \texttt{knitr} \autocite{knitr}.
 All corresponding code for this paper can be found in the github repository (\url{https://github.com/jforbes14/journal-writing}), and the data used is available in the \texttt{eechidna} package \autocite{eechidna}.
@@
 
-All election and Census data sets, along with electoral maps and more, are available in the \texttt{eechidna} \texttt{R} package, which can be downloaded from CRAN.
+All election and Census datasets, along with electoral maps and more, are available in the \texttt{eechidna} \texttt{R} package, which can be downloaded from CRAN.
 \texttt{eechidna} (Exploring Election and Census Highly Informative Data Nationally for Australia) makes it easy to look at the data from the Australian Federal elections and Censuses that occurred between 2001 and 2016.
-This study contributed a large revision to the \texttt{eechidna} package, which included the addition of election and Census data for 2001-2010, vote outcomes for polling booths and imputed Census data for election years.
+This study contributed a large revision to the \texttt{eechidna} package, which included the addition of election and Census data for 2001-2010, voting outcomes for polling booths and imputed Census data for election years.
 For more details on using \texttt{eechidna}, please see the articles (vignettes) on the github page (\url{https://ropenscilabs.github.io/eechidna/}).
@@
 
-{\centering \includegraphics{skeleton_files/figure-latex/partresplot1-1}
+{\centering \includegraphics{draft-paper_files/figure-latex/partresplot1-1}
 
@@
 
-{\centering \includegraphics{skeleton_files/figure-latex/partresplot2-1}
+{\centering \includegraphics{draft-paper_files/figure-latex/partresplot2-1}
 

### 2019-03-31 draft-paper.Rmd
post proofread edit - complete first draft

--- 2019-03-31 draft-paper.Rmd
@@
   address: Monash University\newline Melbourne, Australia
-  email: jeremyforbes1995@gmail.com
+  email: jeremy.forbes@monash.edu
   correspondingauthor: true
@@
 Socio-demographic information is derived from the Australian Census, which occurs every five years.
-Since a Census is not directly available for each election, a method of spatio-temporal imputation is employed to impute Census data for the electorates in each election.
+Since a Census is not directly available for each election, a method of spatio-temporal imputation is employed to construct Census data for the electorates in 2004-2013 elections.
 This accounts for both spatial and temporal changes in electoral characteristics between Censuses.
 To capture any spatial heterogeneity, a spatial error model is estimated for each election, which incorporates a spatially structured random effect vector that can thought of as the unobserved political climate in each electorate.
-Over time, most socio-demographic characteristics that affect electoral two-party preference do not vary, with industry of work, incomes, household mobility and de facto relationships having strong effects in the six elections.
+Over time, the impact of most socio-demographic characteristics that affect electoral two-party preference do not vary, with industry of work, incomes, household mobility and de facto relationships having strong effects in each of the six elections.
 Education and unemployment are amongst those that have varying effects.
@@
 keywords: "federal election, Census, Australia, spatial modelling, imputation, socio-demographics, electorates, R, eechidna"
-wpnumber: no/yr
-jelcodes: C10,C14,C22
+wpnumber: 
+jelcodes: C31, C33, D72
 blind: false
-cover: true
+cover: false
 toc: false
@@
   arrange(year, DivisionNm) %>% 
-  dplyr::select(order(colnames(.)))
+  dplyr::select(order(colnames(.))) %>% 
+  rename(ManagerAdmin = ManagerAdminClericalSales)
 ```
@@
 This is done by taking the electoral boundaries and imputing the corresponding socio-demographic characteristics for each of neighbouring Censuses, thereby addressing the spatial aspect.
-Next, to deal with the temporal component, characteristcs at the time of the election are constructed using linear interpolation between the spatially imputed neighbouring Census variables.
+Next, to deal with the temporal component, characteristics at the time of the election are constructed using linear interpolation between the spatially imputed neighbouring Census variables.
 
@@
 To account for temporal changes, linear interpolation is used between Census years to get the final estimate of a Census variable for the target zone in the election year.
-Let $y_1$ be the year of the Census preceeding an election, let $y_2$ be the year of the election and $y_3$ be the year of the Census that follows.
+Let $y_1$ be the year of the Census preceding an election, let $y_2$ be the year of the election and $y_3$ be the year of the Census that follows.
 Add this year subscript to the Census variable estimate $\hat{C}_t$, resulting in $\hat{C}_{t,y}$.
@@
 Before choosing an appropriate model, two issues with the socio-demographic variables are addressed.
-Firstly, variable scales change over the years and as such the data is standardized accordingly.
+Firstly, variable scales change over the years, making it important to standardize variables.
 Secondly, many variables represent similar information and where appropriate, should be combined in some way.
@@
 For example, inflation-adjusted median rental prices increased across almost all electorates, with median rent of 200 dollars per week placing an electorate in the 90th percentile in 2001, but only the 30th percentile in 2016.
-In order for socio-demographic effects to be comparable across years, the explanatory variables are standardized to have mean zero and variance one within each election year.
+In order for socio-demographic effects to be comparable across years, all explanatory variables are standardized to have mean zero and variance one within each election year.
 By standardizing, each variable is reported as a relative measure compared to all other electorates in the same year.
@@
 
-Only the first four principal components from the combined dataset are considered, as the scree plot levels off after the fourth component.
+Only the first four principal components from the combined dataset are considered, as the scree plot corresponding to the combined dataset levels off after the fourth component.
 Variables that have a large loading in a particular component are deemed to covary, with a loading with magnitude greater than 0.15 being considered large.
@@
 laborer, tradesperson).
-For each of these groups, variables with positive loadings are added and those with negative loadings are substracted to create a factor.
+For each of these groups, variables with positive loadings are added and those with negative loadings are subtracted to create a factor.
 After computing these sums, each factor is standardized to have mean zero and variance one, within each election.
@@
 
- ManagerAdminClericalSales & 2.06$^{***}$ & 3.32$^{***}$ & 6.00$^{***}$ & 5.47$^{***}$ & 5.04$^{***}$ & 5.78$^{***}$ \\ 
+ ManagerAdmin & 2.06$^{***}$ & 3.32$^{***}$ & 6.00$^{***}$ & 5.47$^{***}$ & 5.04$^{***}$ & 5.78$^{***}$ \\ 
   & (0.71) & (0.93) & (0.90) & (1.08) & (1.03) & (1.06) \\
@@
 
-Similarly, workers in managerial, administrative, clerical and sales roles (`ManagerAdminClericalSales`) is also a significant predictor of two-party preference vote across all six elections, with a higher proportion of people working these jobs increasing Liberal support.
+Similarly, workers in managerial, administrative, clerical and sales roles (`ManagerAdmin`) is also a significant predictor of two-party preference vote across all six elections, with a higher proportion of people working these jobs increasing Liberal support.
 The magnitude of this effect also seems to increase over the years.
@@
 De facto relationships, but not marriages, are found to be an important (and significant) predictor of the two-party preferred vote in all six elections, with more de facto relationships associated with higher support for the Labor party.
-The proportion of individuals who are marriaged however, are insignificant (not shown).
+The proportion of individuals who are marriages however, are insignificant (not shown).
 
@@
 p4 <- grid_visreg("Transformative")
-p5 <- grid_visreg("ManagerAdminClericalSales")
+p5 <- grid_visreg("ManagerAdmin")
 p6 <- grid_visreg("DiffAddress")

### 2019-03-31 draft-paper.Rmd
change variable names to fit in plots

@@
   dplyr::select(order(colnames(.))) %>% 
-  rename(ManagerAdmin = ManagerAdminClericalSales)
+  rename(ManagerAdmin = ManagerAdminClericalSales, 
+    OtherLanguage = OtherLanguageHome,
+    Born_SE_EU = Born_SE_Europe)
 ```
@@
 # Instead of running spatial weights
-load("/Users/Jeremy/Documents/R/eechidna-modelling/data/sp_weights_01.rda")
-load("/Users/Jeremy/Documents/R/eechidna-modelling/data/sp_weights_04.rda")
-load("/Users/Jeremy/Documents/R/eechidna-modelling/data/sp_weights_07.rda")
-load("/Users/Jeremy/Documents/R/eechidna-modelling/data/sp_weights_10.rda")
-load("/Users/Jeremy/Documents/R/eechidna-modelling/data/sp_weights_13.rda")
-load("/Users/Jeremy/Documents/R/eechidna-modelling/data/sp_weights_16.rda")
+load("data/sp_weights_01.rda")
+load("data/sp_weights_04.rda")
+load("data/sp_weights_07.rda")
+load("data/sp_weights_10.rda")
+load("data/sp_weights_13.rda")
+load("data/sp_weights_16.rda")
 
@@
 
- OtherLanguageHome & $-$0.44 & 5.92 & 9.98$^{**}$ & 11.24$^{**}$ & 9.00$^{*}$ & 9.84$^{**}$ \\ 
+ OtherLanguage & $-$0.44 & 5.92 & 9.98$^{**}$ & 11.24$^{**}$ & 9.00$^{*}$ & 9.84$^{**}$ \\ 
   & (3.22) & (4.16) & (3.91) & (4.76) & (4.66) & (4.44) \\
@@
 Of these areas, only South-Eastern European populations appear significant in each election, with the proportion of Asian migrants also being significant in 2010.
-Speaking other languages (aside from English) however, appears to have a far stronger effect, as observed through the `OtherLanguageHome` variable.
+Speaking other languages (aside from English) however, appears to have a far stronger effect, as observed through the `OtherLanguage` variable.
 Electorates with more diverse speech are associated with higher support for the Liberal party from 2004 onwards, with this effect being significant in 2007, 2010 and 2016.
@@
 p9 <- grid_visreg("Education")
-p10 <- grid_visreg("OtherLanguageHome")
-p11 <- grid_visreg("Born_SE_Europe")
+p10 <- grid_visreg("OtherLanguage")
+p11 <- grid_visreg("Born_SE_EU")
 p12 <- grid_visreg("Born_Asia")

### 2019-03-31 draft-paper.Rmd
fixed variable names. draft 1 complete and ready for review.

@@
     OtherLanguage = OtherLanguageHome,
-    Born_SE_EU = Born_SE_Europe)
+    BornSEEuro = BornSEEurope,
+    BornAsia = Born_Asia)
 ```
@@
 p10 <- grid_visreg("OtherLanguage")
-p11 <- grid_visreg("Born_SE_EU")
-p12 <- grid_visreg("Born_Asia")
+p11 <- grid_visreg("BornSEEuro")
+p12 <- grid_visreg("BornAsia")
 p13 <- grid_visreg("Judaism")

### 2019-03-31 draft-paper.Rmd
fixed variable names. draft 1 complete and ready for review.

@@
     OtherLanguage = OtherLanguageHome,
-    BornSEEuro = BornSEEurope,
+    BornSEEuro = BornSEEurorope,
     BornAsia = Born_Asia)

### 2019-03-31 draft-paper.Rmd

--- 2019-03-31 draft-paper.Rmd
@@
     BornAsia = Born_Asia)
+
 ```

### 2019-03-31 draft-paper.Rmd
removed underscores from all variables - final first draft.

@@
     OtherLanguage = OtherLanguageHome,
-    BornSEEuro = BornSEEurope,
-    BornAsia = Born_Asia)
+    BornSEEuro = Born_SE_Europe,
+    BornAsia = Born_Asia,
+    OneParentHouse = OneParent_House)
 
@@
 
- Born\_Asia & 2.22 & $-$0.95 & $-$1.60 & $-$6.83$^{**}$ & $-$3.03 & $-$0.55 \\ 
+ BornAsia & 2.22 & $-$0.95 & $-$1.60 & $-$6.83$^{**}$ & $-$3.03 & $-$0.55 \\ 
   & (2.18) & (2.44) & (2.19) & (2.73) & (2.71) & (2.17) \\
 
- Born\_MidEast & $-$1.15 & $-$1.59 & $-$2.01$^{*}$ & $-$2.03 & $-$0.92 & $-$1.44 \\ 
+ BornMidEast & $-$1.15 & $-$1.59 & $-$2.01$^{*}$ & $-$2.03 & $-$0.92 & $-$1.44 \\ 
   & (1.07) & (1.20) & (1.11) & (1.27) & (1.24) & (1.13) \\
 
- Born\_SE\_Europe & $-$3.21$^{**}$ & $-$4.24$^{***}$ & $-$3.61$^{***}$ & $-$4.14$^{***}$ & $-$3.69$^{***}$ & $-$2.72$^{***}$ \\ 
+ BornSEEuro & $-$3.21$^{**}$ & $-$4.24$^{***}$ & $-$3.61$^{***}$ & $-$4.14$^{***}$ & $-$3.69$^{***}$ & $-$2.72$^{***}$ \\ 
   & (1.42) & (1.46) & (1.02) & (1.19) & (1.07) & (0.97) \\
 
- Born\_UK & 0.25 & $-$0.07 & 0.34 & 0.56 & $-$0.09 & $-$1.32 \\ 
+ BornUK & 0.25 & $-$0.07 & 0.34 & 0.56 & $-$0.09 & $-$1.32 \\ 
   & (1.00) & (0.98) & (0.90) & (1.07) & (1.04) & (1.04) \\
@@
 
- OneParent\_House & $-$1.73 & $-$0.45 & $-$0.75 & $-$1.46 & $-$0.77 & $-$0.74 \\ 
+ OneParentHouse & $-$1.73 & $-$0.45 & $-$0.75 & $-$1.46 & $-$0.77 & $-$0.74 \\ 
   & (1.36) & (1.59) & (1.49) & (1.69) & (1.57) & (1.47) \\

### 2019-03-31 draft-paper.tex
first draft

@@
 
-\wp{no/yr}
-\jel{C10,C14,C22}
+\jel{C31, C33, D72}
 
@@
 
+\nocover
+
 \author{Jeremy~Forbes, Di~Cook, Rob~Hyndman}
@@
 Monash University\newline Melbourne, Australia
-\newline{Email: \href{mailto:jeremyforbes1995@gmail.com}{\nolinkurl{jeremyforbes1995@gmail.com}}}\newline Corresponding author\\[1cm]
+\newline{Email: \href{mailto:jeremy.forbes@monash.edu}{\nolinkurl{jeremy.forbes@monash.edu}}}\newline Corresponding author\\[1cm]
 \textbf{Di Cook}\newline
@@
 Socio-demographic information is derived from the Australian Census, which occurs every five years.
-Since a Census is not directly available for each election, a method of spatio-temporal imputation is employed to impute Census data for the electorates in 2004-2013 elections.
+Since a Census is not directly available for each election, a method of spatio-temporal imputation is employed to construct Census data for the electorates in 2004-2013 elections.
 This accounts for both spatial and temporal changes in electoral characteristics between Censuses.
 To capture any spatial heterogeneity, a spatial error model is estimated for each election, which incorporates a spatially structured random effect vector that can thought of as the unobserved political climate in each electorate.
-Over time, most socio-demographic characteristics that affect electoral two-party preference do not vary, with industry of work, incomes, household mobility and de facto relationships having strong effects in the six elections.
+Over time, the impact of most socio-demographic characteristics that affect electoral two-party preference do not vary, with industry of work, incomes, household mobility and de facto relationships having strong effects in each of the six elections.
 Education and unemployment are amongst those that have varying effects.
@@
 To account for temporal changes, linear interpolation is used between Census years to get the final estimate of a Census variable for the target zone in the election year.
-Let \(y_1\) be the year of the Census preceeding an election, let \(y_2\) be the year of the election and \(y_3\) be the year of the Census that follows.
+Let \(y_1\) be the year of the Census preceding an election, let \(y_2\) be the year of the election and \(y_3\) be the year of the Census that follows.
 Add this year subscript to the Census variable estimate \(\hat{C}_t\), resulting in \(\hat{C}_{t,y}\).
@@
 These are; \texttt{Incomes} - median personal, household and family incomes, \texttt{Unemployment} - unemployment and labor force participation rates, \texttt{ProportyOwned} - rates of housing ownership, mortgages, renting and government housing, \texttt{RentLoanPrice} - median rental and loan repayments, \texttt{FamHouseSize} - average household size, ratio of people to families and household makeup (single person, couple with kids and couple without kids) and \texttt{Education} - high school and university qualifications, jobs requiring higher levels of education as well as vocational course completions and jobs that do not require higher education levels (e.g.~laborer, tradesperson).
-For each of these groups, variables with positive loadings are added and those with negative loadings are substracted to create a factor.
+For each of these groups, variables with positive loadings are added and those with negative loadings are subtracted to create a factor.
 After computing these sums, each factor is standardized to have mean zero and variance one, within each election.
@@
 
- OneParent\_House & $-$1.73 & $-$0.45 & $-$0.75 & $-$1.46 & $-$0.77 & $-$0.74 \\ 
+ OneParentHouse & $-$1.73 & $-$0.45 & $-$0.75 & $-$1.46 & $-$0.77 & $-$0.74 \\ 
   & (1.36) & (1.59) & (1.49) & (1.69) & (1.57) & (1.47) \\
 
- OtherLanguageHome & $-$0.44 & 5.92 & 9.98$^{**}$ & 11.24$^{**}$ & 9.00$^{*}$ & 9.84$^{**}$ \\ 
+ OtherLanguage & $-$0.44 & 5.92 & 9.98$^{**}$ & 11.24$^{**}$ & 9.00$^{*}$ & 9.84$^{**}$ \\ 
   & (3.22) & (4.16) & (3.91) & (4.76) & (4.66) & (4.44) \\
@@
 
-Similarly, workers in managerial, administrative, clerical and sales roles (\texttt{ManagerAdminClericalSales}) is also a significant predictor of two-party preference vote across all six elections, with a higher proportion of people working these jobs increasing Liberal support.
+Similarly, workers in managerial, administrative, clerical and sales roles (\texttt{ManagerAdmin}) is also a significant predictor of two-party preference vote across all six elections, with a higher proportion of people working these jobs increasing Liberal support.
 The magnitude of this effect also seems to increase over the years.
@@
 Of these areas, only South-Eastern European populations appear significant in each election, with the proportion of Asian migrants also being significant in 2010.
-Speaking other languages (aside from English) however, appears to have a far stronger effect, as observed through the \texttt{OtherLanguageHome} variable.
+Speaking other languages (aside from English) however, appears to have a far stronger effect, as observed through the \texttt{OtherLanguage} variable.
 Electorates with more diverse speech are associated with higher support for the Liberal party from 2004 onwards, with this effect being significant in 2007, 2010 and 2016.

### 2019-03-31 draft-paper.Rmd
fixed partial residual plots figure caption

@@
 
-*1: HERE Introductory paragraph*
 Electoral socio-demographic variables are now available for each of the six elections and are joined with their corresponding two-party preferred votes.
@@
 It is appropriate to compute principal components in this way because when computed separately for each election, scree plots level off after four components and the loadings of the first four components are similar across the elections.
-
-*2: HERE - ADD JUSTIFICATION FOR DOING SINGLE PCA AND EXPLAIN FACTORS CREATED*
 
@@
 
-```{r resstate, fig.cap = "Boxplot of residuals by state with jittered points.", fig.asp=0.6, out.width="80%", fig.pos = 'h', out.extra = ''}
+```{r resstate, fig.cap = "Boxplot of residuals by state with jittered points.", fig.asp=0.6, out.width="70%", fig.pos = 'h', out.extra = ''}
 all_resids %>% 
@@
 
-*3: HERE Add paragraph on the overall strategy and contributions - can restate stuff from intro*
-
 This paper explores the effects of electoral socio-demographic characteristics on the two-party preferred vote in the 2001-2016 elections, using information from the corresponding Australian federal elections and Censuses.
@@
 
-```{r partresplot1, fig.cap = "Partial residual plots for: income and unemployment, industry and type of work, and household mobility.", fig.height = 9,  fig.width = 6, warning = F, fig.pos = 'h'}
+```{r partresplot1, fig.cap = "Partial residual plots for: Incomes, Unemployment, Extractive, Transformative, ManagerAdmin and DiffAddress", fig.height = 9,  fig.width = 6, warning = F, fig.pos = 'h'}
 grid.arrange(p1, p2, p3, p4, p5, p6, nrow = 6,
@@
 
-```{r partresplot2, fig.cap = "Partial residual plots for: relationships and age, education and diversity.", fig.height = 9,  fig.width = 6, warning = F, fig.pos = 'h'}
+```{r partresplot2, fig.cap = "Partial residual plots for: DeFacto, MedianAge, Education, OtherLanguage, BornSEEuro, BornAsia and Judaism.", fig.height = 9,  fig.width = 6, warning = F, fig.pos = 'h'}
 grid.arrange(p7, p8, p9, p10, p11, p12, p13, nrow = 7,

### 2019-03-31 draft-paper.tex
updated first draft

@@
 
-\emph{1: HERE Introductory paragraph}
 Electoral socio-demographic variables are now available for each of the six elections and are joined with their corresponding two-party preferred votes.
@@
 It is appropriate to compute principal components in this way because when computed separately for each election, scree plots level off after four components and the loadings of the first four components are similar across the elections.
-
-\emph{2: HERE - ADD JUSTIFICATION FOR DOING SINGLE PCA AND EXPLAIN FACTORS CREATED}
 
@@
 
-{\centering \includegraphics[width=0.8\linewidth,]{draft-paper_files/figure-latex/resstate-1}
+{\centering \includegraphics[width=0.7\linewidth,]{draft-paper_files/figure-latex/resstate-1}
 
@@
 
-\emph{3: HERE Add paragraph on the overall strategy and contributions - can restate stuff from intro}
-
 This paper explores the effects of electoral socio-demographic characteristics on the two-party preferred vote in the 2001-2016 elections, using information from the corresponding Australian federal elections and Censuses.
@@
 
-\caption{Partial residual plots for: income and unemployment, industry and type of work, and household mobility.}\label{fig:partresplot1}
+\caption{Partial residual plots for: Incomes, Unemployment, Extractive, Transformative, ManagerAdmin and DiffAddress}\label{fig:partresplot1}
 \end{figure}
@@
 
-\caption{Partial residual plots for: relationships and age, education and diversity.}\label{fig:partresplot2}
+\caption{Partial residual plots for: DeFacto, MedianAge, Education, OtherLanguage, BornSEEuro, BornAsia and Judaism.}\label{fig:partresplot2}
 \end{figure}

### 2019-04-01 draft-paper.Rmd
Made title shorter

@@
 ---
-title: "Spatial modelling of the electoral two-party preferred vote in Australia. A study of federal elections between 2001 and 2016 via the `eechidna` `R` package."
+title: "Spatial modelling of the two-party preferred vote in Australian federal elections: 2001--2016"
 author:
@@
 - familyname: Hyndman
-  othernames: Rob
+  othernames: Rob J
   address: Monash University\newline Melbourne, Australia

### 2019-04-01 draft-paper.Rmd
Moved functions to separate file and styled all code

@@
 ```{r setup, include=FALSE, message=FALSE}
-knitr::opts_chunk$set(echo = F, fig.pos = 'h', fig.align = 'center', cache = T, message = F)
+knitr::opts_chunk$set(
+  echo = FALSE,
+  fig.pos = "h",
+  fig.align = "center",
+  cache = TRUE,
+  message = FALSE
+)
 library(tidyverse)
@@
 library(predictmeans)
-
-# Data
-data(tpp01)
-data(tpp04)
-data(tpp07)
-data(tpp10)
-data(tpp13)
-data(tpp16)
-data(abs2001)
-data(abs2004)
-data(abs2007)
-data(abs2010)
-data(abs2013)
-data(abs2016)
-```
-
-```{r wrangle, include = F}
-# Function to standardize variables
-standardise_vars <- function(df) {
-  hold <- df
-  num_cols <- sapply(df, class) == 'numeric'
-  df[, num_cols] <- lapply(df[, num_cols], scale)
-  names(df) <- names(hold)
-  df$LNP_Percent <- hold$LNP_Percent
-  return(df)
-}
-
+source("functions.R")
+```
+
+```{r wrangle, include = FALSE}
 # Take log of indigneous, judaism, islam, buddhism
-abs2016 <- abs2016 %>% mutate(Indigenous = log(Indigenous), Judaism = log(Judaism), Islam = log(Islam), Buddhism = log(Buddhism))
-abs2013 <- abs2013 %>% mutate(Indigenous = log(Indigenous), Judaism = log(Judaism), Islam = log(Islam), Buddhism = log(Buddhism))
-abs2010 <- abs2010 %>% mutate(Indigenous = log(Indigenous), Judaism = log(Judaism), Islam = log(Islam), Buddhism = log(Buddhism))
-abs2007 <- abs2007 %>% mutate(Indigenous = log(Indigenous), Judaism = log(Judaism), Islam = log(Islam), Buddhism = log(Buddhism))
-abs2004 <- abs2004 %>% mutate(Indigenous = log(Indigenous), Judaism = log(Judaism), Islam = log(Islam), Buddhism = log(Buddhism))
-abs2001 <- abs2001 %>% mutate(Indigenous = log(Indigenous), Judaism = log(Judaism), Islam = log(Islam), Buddhism = log(Buddhism))
+abs2016 <- eechidna::abs2016 %>%
+  mutate(
+    Indigenous = log(Indigenous),
+    Judaism = log(Judaism),
+    Islam = log(Islam),
+    Buddhism = log(Buddhism)
+  )
+abs2013 <- eechidna::abs2013 %>%
+  mutate(
+    Indigenous = log(Indigenous),
+    Judaism = log(Judaism),
+    Islam = log(Islam),
+    Buddhism = log(Buddhism)
+  )
+abs2010 <- eechidna::abs2010 %>%
+  mutate(
+    Indigenous = log(Indigenous),
+    Judaism = log(Judaism),
+    Islam = log(Islam),
+    Buddhism = log(Buddhism)
+  )
+abs2007 <- eechidna::abs2007 %>%
+  mutate(
+    Indigenous = log(Indigenous),
+    Judaism = log(Judaism),
+    Islam = log(Islam),
+    Buddhism = log(Buddhism)
+  )
+abs2004 <- eechidna::abs2004 %>%
+  mutate(
+    Indigenous = log(Indigenous),
+    Judaism = log(Judaism),
+    Islam = log(Islam),
+    Buddhism = log(Buddhism)
+  )
+abs2001 <- eechidna::abs2001 %>%
+  mutate(
+    Indigenous = log(Indigenous),
+    Judaism = log(Judaism),
+    Islam = log(Islam),
+    Buddhism = log(Buddhism)
+  )
 
@@
 my_df <- bind_rows(
-  left_join(tpp01, standardise_vars(abs2001) %>% dplyr::select(-c(UniqueID, Area, ends_with("NS"), Population)), by = c("DivisionNm", "StateAb"="State")) %>% mutate(year = "2001"),
-  left_join(tpp04, standardise_vars(abs2004) %>% dplyr::select(-UniqueID), by = c("DivisionNm")) %>% mutate(year = "2004"),
-  left_join(tpp07, standardise_vars(abs2007) %>% dplyr::select(-UniqueID), by = c("DivisionNm")) %>% mutate(year = "2007"),
-  left_join(tpp10, standardise_vars(abs2010) %>% dplyr::select(-UniqueID), by = c("DivisionNm")) %>% mutate(year = "2010"),
-  left_join(tpp13, standardise_vars(abs2013) %>% dplyr::select(-UniqueID), by = c("DivisionNm")) %>% mutate(year = "2013"),
-  left_join(tpp16, standardise_vars(abs2016) %>% dplyr::select(-c(UniqueID, Area, ends_with("NS"), Population)), by = c("DivisionNm", "StateAb"="State")) %>% mutate(year = "2016")
+  left_join(eechidna::tpp01,
+    standardise_vars(abs2001) %>%
+      dplyr::select(-c(UniqueID, Area, ends_with("NS"), Population)),
+    by = c("DivisionNm", "StateAb" = "State")
+  ) %>% mutate(year = "2001"),
+  left_join(eechidna::tpp04,
+    standardise_vars(abs2004) %>%
+      dplyr::select(-UniqueID),
+    by = c("DivisionNm")
+  ) %>% mutate(year = "2004"),
+  left_join(eechidna::tpp07,
+    standardise_vars(abs2007) %>%
+      dplyr::select(-UniqueID),
+    by = c("DivisionNm")
+  ) %>% mutate(year = "2007"),
+  left_join(eechidna::tpp10,
+    standardise_vars(abs2010) %>%
+      dplyr::select(-UniqueID),
+    by = c("DivisionNm")
+  ) %>% mutate(year = "2010"),
+  left_join(eechidna::tpp13,
+    standardise_vars(abs2013) %>%
+      dplyr::select(-UniqueID),
+    by = c("DivisionNm")
+  ) %>% mutate(year = "2013"),
+  left_join(eechidna::tpp16,
+    standardise_vars(abs2016) %>%
+      dplyr::select(-c(UniqueID, Area, ends_with("NS"), Population)),
+    by = c("DivisionNm", "StateAb" = "State")
+  ) %>% mutate(year = "2016")
 ) %>%
   mutate(year = factor(year)) %>%
-  dplyr::select(-c(starts_with("Age"), StateAb, LNP_Votes, ALP_Votes, ALP_Percent, TotalVotes,
-    Swing, InternetUse, InternetAccess, EnglishOnly, Other_NonChrist, OtherChrist, Volunteer, EmuneratedElsewhere, UniqueID, Catholic, Anglican))
+  dplyr::select(-c(
+    starts_with("Age"), StateAb,
+    LNP_Votes, ALP_Votes, ALP_Percent, TotalVotes, Swing,
+    InternetUse, InternetAccess, EnglishOnly,
+    Other_NonChrist, OtherChrist, Catholic, Anglican,
+    Volunteer, EmuneratedElsewhere, UniqueID
+  ))
 
@@
 factors_df <- my_df %>%
-  mutate(Education = BachelorAbv + HighSchool + Professional + Finance - Laborer - Tradesperson - DipCert,
-    FamHouseSize = FamilyRatio + AverageHouseholdSize + Couple_WChild_House - Couple_NoChild_House -
-      SP_House,
+  mutate(
+    Education = BachelorAbv + HighSchool + Professional + Finance - Laborer - Tradesperson - DipCert,
+    FamHouseSize = FamilyRatio + AverageHouseholdSize + Couple_WChild_House - Couple_NoChild_House - SP_House,
     PropertyOwned = Owned + Mortgage - Renting - PublicHousing,
@@
     Incomes = MedianFamilyIncome + MedianHouseholdIncome + MedianPersonalIncome,
-    Unemployment = Unemployed - LFParticipation) %>% 
-  dplyr::select(-c(BachelorAbv, HighSchool, Professional, Finance, Laborer, Tradesperson, DipCert, FamilyRatio,
+    Unemployment = Unemployed - LFParticipation
+  ) %>%
+  dplyr::select(-c(
+    BachelorAbv, HighSchool, Professional, Finance, Laborer, Tradesperson, DipCert, FamilyRatio,
     AverageHouseholdSize, Couple_WChild_House, Couple_NoChild_House, SP_House, Owned, Mortgage, Renting,
     PublicHousing, MedianFamilyIncome, MedianHouseholdIncome, MedianPersonalIncome, MedianRent,
-    MedianLoanPay, Unemployed, LFParticipation))
+    MedianLoanPay, Unemployed, LFParticipation
+  ))
 
@@
   dplyr::select(order(colnames(.))) %>%
-  rename(ManagerAdmin = ManagerAdminClericalSales, 
+  rename(
+    ManagerAdmin = ManagerAdminClericalSales,
     OtherLanguage = OtherLanguageHome,
@@
     BornAsia = Born_Asia,
-    OneParentHouse = OneParent_House)
-
-```
-
-```{r spatialweights, eval = F}
-# Compute spatial weights matrix
-sp_weights_matrix <- function(sF) {
-
-dist_matrix <- function(shapefile) {
-  dist_mat <- matrix(NA, nrow = 150, ncol = 150)
-  rownames(dist_mat) <- sort(shapefile$elect_div)
-  colnames(dist_mat) <- sort(shapefile$elect_div)
-
-  for (i in 1:(nrow(dist_mat)-1)) {
-    rname = rownames(dist_mat)[i]
-    row_poly = shapefile %>% subset(elect_div == rname)
-
-    for (j in (i+1):ncol(dist_mat)) {
-
-      cname = rownames(dist_mat)[j]
-      col_poly = shapefile %>% subset(elect_div == cname)
-      dist = gDistance(row_poly, col_poly)
-      dist_mat[i,j] = dist
-
-    }
-    print(i)
-  }
-
-  # Now copy to lower triange
-  for (i in 2:nrow(dist_mat)) {
-    for (j in 1:(i-1)) {
-      dist_mat[i,j] = dist_mat[j,i]
-    }
-  }
-
-  # Check it is symmetric
-  if(!isSymmetric(dist_mat)) {
-    print("Warning! Matrix is not symmetric. Error has occured.")
-  }
-
-  return(dist_mat)
-}
-
-my_dist_mat <- dist_matrix(sF)
-
-# S matrix
-
-s_mat <- function(dist_mat) {
-  s_mat <- dist_mat
-
-  for (i in 1:nrow(dist_mat)) {
-    for (j in 1:nrow(dist_mat)) {
-      a = dist_mat[i,j]
-
-      if (is.na(a)) {
-        b = 0
-      } else {
-        b = ifelse(a == 0, 1, 0)
-      }
-
-      s_mat[i,j] = b
-
-    }
-  }
-
-  return(s_mat)
-}
-
-my_smat <- s_mat(my_dist_mat)
-
-# Turn into W matrix
-
-w_mat <- function(s_mat) {
-
-  w_mat <- s_mat
-  rsums = rowSums(s_mat)
-
-  for (i in 1:nrow(s_mat)) {
-    w_mat[i,] <- s_mat[i,]/rsums[i]
-  }
-
-  return(w_mat)
-}
-
-my_wmat <- w_mat(my_smat)
-
-# Turn into listw
-
-my_listw <- mat2listw(my_wmat)
-
-return(my_listw)
-
-}
-
+    OneParentHouse = OneParent_House
+  )
+```
+
+```{r spatialweights, eval = FALSE}
 # Get spatial weights
@@
 
-```{r FGLS, include = F}
+```{r FGLS, include = FALSE}
 # Instead of running spatial weights
@@
 
-# Function for FGLS
-my_fgls <- function(my_formula, my_data, sp_weights) {
-  # DivisionNm
-  model_data <- my_data %>% dplyr::select(-c(DivisionNm, year))
-
-  # Spatial weights matrix
-  w_mat <- listw2mat(sp_weights)
-
-  # Get OLS residuals
-  ols_model <- lm(my_formula, model_data)
-  my_res <- ols_model$residuals
-
-  # Solve for rho
-  res_model <- lm(my_res ~ w_mat%*%my_res)
-  rho <- res_model$coefficients[2]
-  rho_df <- data.frame(estimate = rho, se = summary(res_model)$coefficients[2,2], p = summary(res_model)$coefficients[2,4])
-
-  # Transform data for GLS
-  trans_mat <- diag(nrow(model_data)) - rho*w_mat
-  gls_data <- data.frame(LNP_Percent = trans_mat %*% model_data$LNP_Percent,
-    Intercept = trans_mat %*% rep(1,nrow(model_data))) %>% 
-    bind_cols(as.data.frame(trans_mat %*% as.matrix(model_data %>% dplyr::select(-LNP_Percent))))
-
-  # GLS model
-  my_formula <- formula(paste0(my_formula,  " - 1"))
-  gls_model <- gls(my_formula, gls_data)
-
-  # Cooks distance
-  #gls_model$cooksd <- unname(predictmeans::CookD(gls_model, plot = FALSE))
-
-  # Call to function with stargazer
-  gls_model$call$model <- formula(paste0("LNP_Percent ~ ", paste0(names(gls_data)[-1], collapse = " + ")))
-
-  # Rho and data
-  gls_model$rho_df <- rho_df
-  gls_model$gls_data <- gls_data
-  gls_model$my_data <- my_data
-  gls_model$actual_residuals <- solve(trans_mat)%*%gls_model$residuals
-
-  return(gls_model)
-}
-
 ## Run full models for each year
-full_formula = "LNP_Percent ~ ."
+full_formula <- "LNP_Percent ~ ."
 
@@
   my_data = model_df %>% filter(year == "2016"),
-  sp_weights = sp_weights_16)
+  sp_weights = sp_weights_16
+)
 
@@
   my_data = model_df %>% filter(year == "2013"),
-  sp_weights = sp_weights_13)
+  sp_weights = sp_weights_13
+)
 
@@
   my_data = model_df %>% filter(year == "2010"),
-  sp_weights = sp_weights_10)
+  sp_weights = sp_weights_10
+)
 
@@
   my_data = model_df %>% filter(year == "2007"),
-  sp_weights = sp_weights_07)
+  sp_weights = sp_weights_07
+)
 
@@
   my_data = model_df %>% filter(year == "2004"),
-  sp_weights = sp_weights_04)
+  sp_weights = sp_weights_04
+)
 
@@
   my_data = model_df %>% filter(year == "2001"),
-  sp_weights = sp_weights_01)
+  sp_weights = sp_weights_01
+)
 
@@
 coef_df <- bind_rows(
-  data.frame(variable = glsmod16$coefficients %>% names, estimate = glsmod16$coefficients %>% unname, se = summary(glsmod16)$tTable[, "Std.Error"] %>% unname, p = summary(glsmod16)$tTable[, "p-value"] %>% unname, year = 2016),
-  data.frame(variable = glsmod13$coefficients %>% names, estimate = glsmod13$coefficients %>% unname, se = summary(glsmod13)$tTable[, "Std.Error"] %>% unname, p = summary(glsmod13)$tTable[, "p-value"] %>% unname, year = 2013),
-  data.frame(variable = glsmod10$coefficients %>% names, estimate = glsmod10$coefficients %>% unname, se = summary(glsmod10)$tTable[, "Std.Error"] %>% unname, p = summary(glsmod10)$tTable[, "p-value"] %>% unname, year = 2010),
-  data.frame(variable = glsmod07$coefficients %>% names, estimate = glsmod07$coefficients %>% unname, se = summary(glsmod07)$tTable[, "Std.Error"] %>% unname, p = summary(glsmod07)$tTable[, "p-value"] %>% unname, year = 2007),
-  data.frame(variable = glsmod04$coefficients %>% names, estimate = glsmod04$coefficients %>% unname, se = summary(glsmod04)$tTable[, "Std.Error"] %>% unname, p = summary(glsmod04)$tTable[, "p-value"] %>% unname, year = 2004),
-  data.frame(variable = glsmod01$coefficients %>% names, estimate = glsmod01$coefficients %>% unname, se = summary(glsmod01)$tTable[, "Std.Error"] %>% unname, p = summary(glsmod01)$tTable[, "p-value"] %>% unname, year = 2001)
+  data.frame(
+    variable = glsmod16$coefficients %>% names(),
+    estimate = glsmod16$coefficients %>% unname(),
+    se = summary(glsmod16)$tTable[, "Std.Error"] %>% unname(),
+    p = summary(glsmod16)$tTable[, "p-value"] %>% unname(),
+    year = 2016),
+  data.frame(
+    variable = glsmod13$coefficients %>% names(),
+    estimate = glsmod13$coefficients %>% unname(),
+    se = summary(glsmod13)$tTable[, "Std.Error"] %>% unname(),
+    p = summary(glsmod13)$tTable[, "p-value"] %>% unname(),
+    year = 2013),
+  data.frame(
+    variable = glsmod10$coefficients %>% names(),
+    estimate = glsmod10$coefficients %>% unname(),
+    se = summary(glsmod10)$tTable[, "Std.Error"] %>% unname(),
+    p = summary(glsmod10)$tTable[, "p-value"] %>% unname(),
+    year = 2010),
+  data.frame(
+    variable = glsmod07$coefficients %>% names(),
+    estimate = glsmod07$coefficients %>% unname(),
+    se = summary(glsmod07)$tTable[, "Std.Error"] %>% unname(),
+    p = summary(glsmod07)$tTable[, "p-value"] %>% unname(),
+    year = 2007),
+  data.frame(
+    variable = glsmod04$coefficients %>% names(),
+    estimate = glsmod04$coefficients %>% unname(),
+    se = summary(glsmod04)$tTable[, "Std.Error"] %>% unname(),
+    p = summary(glsmod04)$tTable[, "p-value"] %>% unname(),
+    year = 2004),
+  data.frame(
+    variable = glsmod01$coefficients %>% names(),
+    estimate = glsmod01$coefficients %>% unname(),
+    se = summary(glsmod01)$tTable[, "Std.Error"] %>% unname(),
+    p = summary(glsmod01)$tTable[, "p-value"] %>% unname(),
+    year = 2001)
 )
@@
 
-```{r timeline, echo = F, fig.cap="Timeline of Australian elections and Censuses. They do not always occur in the same year.", out.width = "90%", fig.align="center"}
+```{r timeline, echo = FALSE, fig.cap="Timeline of Australian elections and Censuses. They do not always occur in the same year.", out.width = "90%", fig.align="center"}
 timeline <- data.frame(year = c(2001:2016)) %>%
-  mutate(Election = ifelse(year %in% c(2001, 2004, 2007, 2010, 2013, 2016),1,0),
-         Census = ifelse(year %in% c(2001, 2006, 2011, 2016), 1.4, 0)) %>%
-  gather(c(Election,Census), key = event, value = measure) %>%  filter(measure != 0) %>%
+  mutate(
+    Election = ifelse(year %in% c(2001, 2004, 2007, 2010, 2013, 2016), 1, 0),
+    Census = ifelse(year %in% c(2001, 2006, 2011, 2016), 1.4, 0)
+  ) %>%
+  gather(c(Election, Census), key = event, value = measure) %>%
+  filter(measure != 0) %>%
   mutate(year_vis = ifelse(event == "Census", year + 0.1, year))
 
-ggplot(timeline, aes(x=year_vis, y=measure)) + geom_point(aes(color=as.factor(event),shape=as.factor(event)),size = 4) + labs(color = "Event", shape = "Event") + geom_hline(yintercept=0, size =1, color = 'grey', show.legend = FALSE) + geom_segment(x=2001,xend=2001, y=1, yend=0, color = '#00BFC4') + geom_segment(x=2004,xend=2004, y=1, yend=0, color = '#00BFC4') + geom_segment(x=2007,xend=2007, y=1, yend=0, color = '#00BFC4') + geom_segment(x=2010,xend=2010, y=1, yend=0, color = '#00BFC4') + geom_segment(x=2013,xend=2013, y=1, yend=0, color = '#00BFC4') + geom_segment(x=2016,xend=2016, y=1, yend=0, color = '#00BFC4') + geom_segment(x=2001.1,xend=2001.1, y=1.4, yend=0, color = '#F8766D') + geom_segment(x=2006.1,xend=2006.1, y=1.4, yend=0, color = '#F8766D') + geom_segment(x=2011.1,xend=2011.1, y=1.4, yend=0, color = '#F8766D') + geom_segment(x=2016.1,xend=2016.1, y=1.4, yend=0, color = '#F8766D') + labs(x="Year") + theme(text=element_text(size=10), axis.text.x=element_text(size=6),axis.text.y=element_blank(), axis.title.y=element_blank(), plot.background=element_blank(), axis.ticks.y = element_blank(), panel.background = element_blank()) + scale_x_discrete(limits = c(2001:2016)) + coord_fixed(ratio = 2)
+ggplot(timeline, aes(x = year_vis, y = measure)) +
+  geom_point(aes(color = as.factor(event), shape = as.factor(event)), size = 4) +
+  labs(color = "Event", shape = "Event", x = "Year") +
+  geom_hline(yintercept = 0, size = 1, color = "grey", show.legend = FALSE) +
+  geom_segment(x = 2001, xend = 2001, y = 1, yend = 0, color = "#00BFC4") +
+  geom_segment(x = 2004, xend = 2004, y = 1, yend = 0, color = "#00BFC4") +
+  geom_segment(x = 2007, xend = 2007, y = 1, yend = 0, color = "#00BFC4") +
+  geom_segment(x = 2010, xend = 2010, y = 1, yend = 0, color = "#00BFC4") +
+  geom_segment(x = 2013, xend = 2013, y = 1, yend = 0, color = "#00BFC4") +
+  geom_segment(x = 2016, xend = 2016, y = 1, yend = 0, color = "#00BFC4") +
+  geom_segment(x = 2001.1, xend = 2001.1, y = 1.4, yend = 0, color = "#F8766D") +
+  geom_segment(x = 2006.1, xend = 2006.1, y = 1.4, yend = 0, color = "#F8766D") +
+  geom_segment(x = 2011.1, xend = 2011.1, y = 1.4, yend = 0, color = "#F8766D") +
+  geom_segment(x = 2016.1, xend = 2016.1, y = 1.4, yend = 0, color = "#F8766D") +
+  theme(
+    text = element_text(size = 10),
+    axis.text.x = element_text(size = 6),
+    axis.text.y = element_blank(),
+    axis.title.y = element_blank(),
+    plot.background = element_blank(),
+    axis.ticks.y = element_blank(),
+    panel.background = element_blank()
+  ) +
+  scale_x_discrete(limits = c(2001:2016)) +
+  coord_fixed(ratio = 2)
 ```
@@
 hume_area13 <- nat_map13 %>%
-  filter(state %in% c("ACT","NSW"), long < 154)
-
-ggplot(data=hume_area13) +
-  geom_polygon(aes(x=long, y=lat, group=group, fill = elect_div == "HUME"),
-               colour="grey50", alpha = 0.4) +
-  scale_fill_manual(name="Electoral boundaries", values=c("white", "purple"), labels = c("Other 2013 Electorates", "Hume 2013")) +
-  theme_map() + coord_equal()
+  filter(state %in% c("ACT", "NSW"), long < 154)
+
+ggplot(data = hume_area13) +
+  geom_polygon(aes(x = long, y = lat, group = group, fill = elect_div == "HUME"),
+    colour = "grey50", alpha = 0.4
+  ) +
+  scale_fill_manual(
+    name = "Electoral boundaries",
+    values = c("white", "purple"),
+    labels = c("Other 2013 Electorates", "Hume 2013")
+  ) +
+  theme_map() +
+  coord_equal()
 ```
@@
 hume_area16 <- nat_map16 %>%
-  filter(state %in% c("ACT","NSW"), long < 154) %>% 
+  filter(state %in% c("ACT", "NSW"), long < 154) %>%
   mutate(year = "2016") %>%
-  bind_rows(hume_area13 %>% filter(elect_div == "HUME") %>% mutate(year = "2013"))
-
-ggplot(data=hume_area16) +
-  geom_polygon(aes(x=long, y=lat, group=group, fill = year == "2013", 
-    alpha = year == "2013", colour=year == "2013")) +
-  scale_fill_manual(name="Electoral boundaries", values=c("white", "purple"), labels = c("2016 Electorates", "Hume 2013")) +
-  scale_alpha_manual(values=c(0, 0.4)) +
-  scale_color_manual(values=c("grey50", NA)) +
-  theme_map() + coord_equal() + guides(alpha = F, col = F)
+  bind_rows(
+    hume_area13 %>%
+      filter(elect_div == "HUME") %>%
+      mutate(year = "2013")
+  )
+
+ggplot(data = hume_area16) +
+  geom_polygon(aes(
+    x = long, y = lat, group = group, fill = year == "2013",
+    alpha = year == "2013", colour = year == "2013"
+  )) +
+  scale_fill_manual(
+    name = "Electoral boundaries",
+    values = c("white", "purple"),
+    labels = c("2016 Electorates", "Hume 2013")
+  ) +
+  scale_alpha_manual(values = c(0, 0.4)) +
+  scale_color_manual(values = c("grey50", NA)) +
+  theme_map() +
+  coord_equal() +
+  guides(alpha = FALSE, col = FALSE)
 ```
@@
 
-```{r eval = F, include = F}
+```{r eval = FALSE, include = FALSE}
 auscitizen_ex <- abs2016 %>%
@@
   select(DivisionNm, AusCitizen) %>%
-  mutate(`Estimated Population Allocated to Purple Region` = c(21, 548, 16358, 474, 86, 145427, 39117, 92),
-    AusCitizen = round(AusCitizen, 2)) %>% 
+  mutate(
+    `Estimated Population Allocated to Purple Region` = c(21, 548, 16358, 474, 86, 145427, 39117, 92),
+    AusCitizen = round(AusCitizen, 2)
+  ) %>%
   arrange(-`Estimated Population Allocated to Purple Region`) %>%
@@
 
-```{r printmodels, eval = F}
+```{r printmodels, eval = FALSE}
 # Add rho and GLS residual standard error
-stargazer::stargazer(glsmod01, glsmod04, glsmod07, glsmod10, glsmod13, glsmod16, header = FALSE, 
+stargazer::stargazer(glsmod01, glsmod04, glsmod07, glsmod10, glsmod13, glsmod16,
+  header = FALSE,
   column.sep.width = "1pt",
@@
   dep.var.labels = "Two-party preferred vote in favor of the Liberal party",
-  font.size = "scriptsize", digits = 2, 
+  font.size = "scriptsize",
+  digits = 2,
   column.labels = c("2001", "2004", "2007", "2010", "2013", "2016"),
@@
 Overall", "summary measures for each regression equation are provided in the", "bottom panel."),
-  notes.align = "l")
-```
+  notes.align = "l"
+)
+```
+
 \begin{table}[!htbp] \centering
@@
 Only in 2001 and 2016 is there a significant spatial component.", out.height='15%', out.extra = ''}
-rho_df <- bind_rows(glsmod16$rho_df %>% mutate(year = "2016"), 
+rho_df <- bind_rows(
+  glsmod16$rho_df %>% mutate(year = "2016"),
   glsmod13$rho_df %>% mutate(year = "2013"),
@@
   glsmod04$rho_df %>% mutate(year = "2004"),
-  glsmod01$rho_df %>% mutate(year = "2001")) %>% 
-  mutate(upper95 = estimate + 1.96*se, lower95 = estimate - 1.96*se)
-
-rho_df %>% ggplot() + 
+  glsmod01$rho_df %>% mutate(year = "2001")
+) %>%
+  mutate(
+    upper95 = estimate + 1.96 * se,
+    lower95 = estimate - 1.96 * se
+  )
+
+rho_df %>%
+  ggplot() +
   geom_point(aes(x = year, y = estimate), size = 3) +
@@
   labs(x = "Election year", y = "Estimate of spatial autoregressive parameter") +
-  lims(y = c(-1,1)) +
-  guides(col = F) +
+  lims(y = c(-1, 1)) +
+  guides(col = FALSE) +
   coord_fixed(ratio = 1) +
@@
 ```{r plotintercept, fig.cap = "Estimated intercept for each election, which represents the two-party preferred vote for an electorate with mean characteristics.", out.height="15%", out.extra = ''}
-intercept_df <- bind_rows(data.frame(summary(glsmod16)$tTable)[1, ]  %>% mutate(year = "2016"), 
+intercept_df <- bind_rows(
+  data.frame(summary(glsmod16)$tTable)[1, ] %>% mutate(year = "2016"),
   data.frame(summary(glsmod13)$tTable)[1, ] %>% mutate(year = "2013"),
@@
   data.frame(summary(glsmod04)$tTable)[1, ] %>% mutate(year = "2004"),
-  data.frame(summary(glsmod01)$tTable)[1, ] %>% mutate(year = "2001")) %>% 
+  data.frame(summary(glsmod01)$tTable)[1, ] %>% mutate(year = "2001")
+) %>%
   rename(se = `Std.Error`, estimate = Value, p = `p.value`) %>%
-  mutate(upper95 = estimate + 1.96*se, lower95 = estimate - 1.96*se)
-
-intercept_df %>% ggplot() + 
+  mutate(upper95 = estimate + 1.96 * se, lower95 = estimate - 1.96 * se)
+
+intercept_df %>%
+  ggplot() +
   geom_point(aes(x = year, y = estimate), size = 3) +
@@
   labs(x = "Election year", y = "Estimate of intercept") +
-  guides(col = F) +
+  guides(col = FALSE) +
   coord_fixed(ratio = 0.2) +
@@
 
-```{r myvisreg, include = F}
-# Function to produce visreg style conditional plots
-my_visreg <- function(my_model, sp_weights, varname, 
-  plot = FALSE, nolabs = FALSE, xlimits = NULL, ylimits = NULL, year = "") {
-
-  # Extract fitted parameters
-  rho <- my_model$rho_df$estimate
-  sigma <- sqrt(sum(my_model$residuals^2)/(my_model$dims$N-my_model$dims$p))
-
-  # Spatial weights
-  w_mat <- listw2mat(sp_weights)
-
-  # Q - where u = Qe, Q = (I - pW)^-1
-  q_mat <- solve(diag(my_model$dims$N) - rho*w_mat)
-
-  # Omega - QQ'
-  omega_mat <- q_mat%*%t(q_mat)
-
-  # X
-  x_mat <- my_model$my_data %>% 
-    dplyr::select(-c(LNP_Percent, year, DivisionNm)) %>% 
-    mutate(Intercept = 1) %>% 
-    dplyr::select(Intercept, everything()) %>% 
-    as.matrix()
-
-  # Beta
-  beta_mat <- my_model$coefficients
-
-  # T value
-  t = qt(0.975, nrow(my_model$gls_data)-ncol(my_model$gls_data))
-
-  # Lambda matrix (FGLS)
-  x <- round(seq(min(as.numeric(x_mat[, varname])), max(as.numeric(x_mat[, varname])), 0.025), 3)
-  lambda_mat <- data.frame(matrix(0, nrow = length(x), ncol = ncol(x_mat)))
-  names(lambda_mat) <- dimnames(x_mat)[[2]]
-  lambda_mat[, varname] <- x
-  lambda_mat$Intercept <- 1
-  lambda_mat <- as.matrix(lambda_mat)
-
-  # Confidence interval
-  plot_df <- data.frame(variable = x, fitted = lambda_mat%*%beta_mat, variance = 0)
-
-  for (i in 1:nrow(lambda_mat)) {
-    lambda <- lambda_mat[i, ]
-    plot_df$variance[i] = sigma^2 * t(lambda) %*% 
-      solve(t(x_mat) %*% solve(omega_mat) %*% x_mat) %*% 
-      lambda
-  }
-
-  plot_df <- plot_df %>% 
-    mutate(upper95 = fitted + t*sqrt(variance), lower95 = fitted - t*sqrt(variance))
-
-  # Partial residuals
-  points_df <- data.frame(
-    variable = my_model$my_data[, varname] %>% unname,
-    part_res = (my_model$my_data$LNP_Percent - x_mat%*%my_model$coefficients) + my_model$coefficients[varname]*x_mat[, varname] + my_model$coefficients[1]
-  )
-
-  # Plot
-  if (plot == TRUE) {
-      myplot <- ggplot(data = plot_df) + 
-    geom_ribbon(aes(x = variable, ymin = lower95, ymax = upper95), fill = "grey80") + 
-    geom_point(aes(x = variable, y = part_res), data = points_df, size = 0.75, col = "grey50") + 
-    geom_line(aes(x = variable, y = fitted), col = "blue", size = 1) +
-    #geom_hline(aes(yintercept = min(upper95)), col = "red") +
-    #geom_hline(aes(yintercept = max(lower95)), col = "blue") +
-    theme_bw() + 
-    labs(x = varname, y = "Response") + 
-    ggtitle(year) +
-    theme(plot.title = element_text(face = "bold", size = 10, hjust = 0.5))
-
-  if (nolabs == TRUE) {
-    myplot <- myplot + labs(x = "", y = "")
-  }
-
-  if (!is.null(xlimits) & !is.null(ylimits)) {
-    myplot <- myplot + coord_cartesian(xlim = xlimits, ylim = ylimits)
-  }
-      return(myplot)
-  }
-
-  # Points
-  if (plot == FALSE) {
-    return_ls <- list(bands = plot_df, points = points_df)
-
-    return(return_ls)
-  }
-
-}
-
-# Grid visreg
-grid_visreg <- function(varname, plot = TRUE, myscale = "free", top = FALSE) {
-
-    p16 <- my_visreg(glsmod16, sp_weights_16, varname = varname, plot = F, year = "2016")
-    p13 <- my_visreg(glsmod13, sp_weights_13, varname = varname, plot = F, year = "2013")
-    p10 <- my_visreg(glsmod10, sp_weights_10, varname = varname, plot = F, year = "2010")
-    p07 <- my_visreg(glsmod07, sp_weights_07, varname = varname, plot = F, year = "2007")
-    p04 <- my_visreg(glsmod04, sp_weights_04, varname = varname, plot = F, year = "2004")
-    p01 <- my_visreg(glsmod01, sp_weights_01, varname = varname, plot = F, year = "2001")
-
-    bands_df <- bind_rows(
-      p16$bands %>% mutate(year = "2016"),
-      p13$bands %>% mutate(year = "2013"),
-      p10$bands %>% mutate(year = "2010"),
-      p07$bands %>% mutate(year = "2007"),
-      p04$bands %>% mutate(year = "2004"),
-      p01$bands %>% mutate(year = "2001")
-    ) %>% mutate(varname = varname)
-
-    points_df <- bind_rows(
-      p16$points %>% mutate(year = "2016"),
-      p13$points %>% mutate(year = "2013"),
-      p10$points %>% mutate(year = "2010"),
-      p07$points %>% mutate(year = "2007"),
-      p04$points %>% mutate(year = "2004"),
-      p01$points %>% mutate(year = "2001")
-    ) %>% mutate(varname = varname)
-
-    if (plot == T) {
-      return_object <- ggplot(data = bands_df) +
-        geom_ribbon(aes(x = variable, ymin = lower95, ymax = upper95), fill = "grey80") + 
-        geom_point(aes(x = variable, y = part_res), data = points_df, size = 0.5, shape = 1, alpha = 0.5) +
-        geom_line(aes(x = variable, y = fitted), col = "blue", size = 1) +
-        theme_bw() + 
-        labs(x = "", y = "") + 
-        facet_grid(varname ~ year, scales = "free") +
-        theme(plot.margin=unit(c(b = 0.05, l = 0, t = -0.2, r = 0),"cm"))
-
-      if (top == FALSE) {
-        return_object <- return_object + theme(strip.text.x = element_blank()) 
-      }
-
-      if (myscale == "free") {
-        return_object <- return_object + lims(x = range(model_df %>% select(varname)), 
-          y = range(c(bands_df$upper95, bands_df$lower95, points_df$part_res)))
-      }
-
-      if (myscale == "free_x") {
-        return_object <- return_object + lims(x = range(model_df %>% select(varname)))
-      }
-
-      if (myscale == "free_y") {
-        return_object <- return_object + lims(y = range(c(bands_df$upper95, bands_df$lower95, points_df$part_res)))
-      }
-
-      if (myscale == "none") {
-        return_object <- return_object
-      }
-    }
-
-    if (plot == F) {
-      return_object <- list(bands = bands_df, points = points_df)
-    }
-
-    return(return_object)
-
-}
-```
-
 ### Income and unemployment
@@
 
-```{r prepplot, include = F}
-p1 <- grid_visreg("Incomes", top = T)
+```{r prepplot, include = FALSE}
+p1 <- grid_visreg("Incomes", top = TRUE)
 p2 <- grid_visreg("Unemployment")
@@
 p6 <- grid_visreg("DiffAddress")
-p7 <- grid_visreg("DeFacto", top = T)
+p7 <- grid_visreg("DeFacto", top = TRUE)
 p8 <- grid_visreg("MedianAge")
@@
 
-```{r removeunemp, eval = F}
+```{r removeunemp, eval = FALSE}
 unemp01 <- my_fgls(full_formula,
   my_data = model_df %>% filter(year == "2001") %>% dplyr::select(-Incomes),
-  sp_weights = sp_weights_01)
+  sp_weights = sp_weights_01
+)
 
@@
 
-```{r resids, include = F}
+```{r resids, include = FALSE}
 all_resids <- data.frame(
@@
     glsmod04$my_data %>% left_join(tpp04 %>% dplyr::select(DivisionNm, UniqueID, StateAb), by = "DivisionNm"),
-    glsmod01$my_data %>% left_join(tpp01 %>% dplyr::select(DivisionNm, UniqueID, StateAb), by = "DivisionNm"))
-) 
-#%>% dplyr::select(Residuals, DivisionNm, UniqueID, StateAb, year, LNP_Percent, gls_residuals)
+    glsmod01$my_data %>% left_join(tpp01 %>% dplyr::select(DivisionNm, UniqueID, StateAb), by = "DivisionNm")
+  )
+)
+# %>% dplyr::select(Residuals, DivisionNm, UniqueID, StateAb, year, LNP_Percent, gls_residuals)
 ```
@@
   geom_jitter(alpha = 0.5, shape = 1) +
-  guides(col = F, shape = F, DivisionNm = F) +
+  guides(col = FALSE, shape = FALSE, DivisionNm = FALSE) +
   theme_bw()
@@
 
-```{r readcooksd, eval = F}
+```{r readcooksd, eval = FALSE}
 read_chunk('cooksd.R')
@@
 
-```{r getcooksd, eval = F}
+```{r getcooksd, eval = FALSE}
 <<GetCookDistance>>
@@
 
-```{r cooksd, eval = F}
+```{r cooksd, eval = FALSE}
 cooksd_df <- data.frame(
@@
 
-```{r plotcooksd, fig.cap = "Electorates' Cook's Distance values in each election.", out.height = "15%", eval = F}
+```{r plotcooksd, fig.cap = "Electorates' Cook's Distance values in each election.", out.height = "15%", eval = FALSE}
 cooksd_df %>%
@@
   facet_wrap(~year, nrow = 1) +
-  #coord_fixed(ratio = .5) +
+  # coord_fixed(ratio = .5) +
   labs(x = "Cook's Distance", y = "") +
@@
 
-```{r, eval = F}
+```{r, eval = FALSE}
 cooksd_df %>%
   filter(cooksd > 0.1) %>%
-  select(DivisionNm, StateAb, year, cooksd, Residuals, LNP_Percent, everything()) %>% View
+  select(DivisionNm, StateAb, year, cooksd, Residuals, LNP_Percent, everything()) %>%
+  View()
 ```
@@
 
-```{r, eval = F}
-x <- model_df %>% filter(year == "2016") %>% select(-c(year, DivisionNm)) %>% as.matrix
+```{r, eval = FALSE}
+x <- model_df %>%
+  filter(year == "2016") %>%
+  select(-c(year, DivisionNm)) %>%
+  as.matrix()
 myhat <- hat(x)
 
-plot(glsmod07$cooksd, hat(model_df %>% filter(year == "2007") %>% select(-c(year, DivisionNm)) %>% as.matrix))
+plot(
+  glsmod07$cooksd,
+  hat(
+    model_df %>%
+      filter(year == "2007") %>%
+      select(-c(year, DivisionNm)) %>%
+      as.matrix()
+  )
+)
 ```
@@
 
-```{r partresplot1, fig.cap = "Partial residual plots for: Incomes, Unemployment, Extractive, Transformative, ManagerAdmin and DiffAddress", fig.height = 9,  fig.width = 6, warning = F, fig.pos = 'h'}
-grid.arrange(p1, p2, p3, p4, p5, p6, nrow = 6,
+```{r partresplot1, fig.cap = "Partial residual plots for: Incomes, Unemployment, Extractive, Transformative, ManagerAdmin and DiffAddress", fig.height = 9,  fig.width = 6, warning = FALSE, fig.pos = 'h'}
+grid.arrange(p1, p2, p3, p4, p5, p6,
+  nrow = 6,
   heights = c(1.2, 1, 1, 1, 1, 1),
-  left = grid.text("Two-party preferred vote (%)", gp = gpar(cex = 0.8), rot = 90))
-```
-
-```{r partresplot2, fig.cap = "Partial residual plots for: DeFacto, MedianAge, Education, OtherLanguage, BornSEEuro, BornAsia and Judaism.", fig.height = 9,  fig.width = 6, warning = F, fig.pos = 'h'}
-grid.arrange(p7, p8, p9, p10, p11, p12, p13, nrow = 7,
+  left = grid.text("Two-party preferred vote (%)", gp = gpar(cex = 0.8), rot = 90)
+)
+```
+
+```{r partresplot2, fig.cap = "Partial residual plots for: DeFacto, MedianAge, Education, OtherLanguage, BornSEEuro, BornAsia and Judaism.", fig.height = 9,  fig.width = 6, warning = FALSE, fig.pos = 'h'}
+grid.arrange(p7, p8, p9, p10, p11, p12, p13,
+  nrow = 7,
   heights = c(1.2, 1, 1, 1, 1, 1, 1),
-  left = grid.text("Two-party preferred vote (%)", gp = gpar(cex = 0.8), rot = 90))
-```
+  left = grid.text("Two-party preferred vote (%)", gp = gpar(cex = 0.8), rot = 90)
+)
+```

### 2019-04-01 draft-paper.tex
Moved functions to separate file and styled all code

@@
 \hypersetup{
-            pdftitle={Spatial modelling of the electoral two-party preferred vote in Australia. A study of federal elections between 2001 and 2016 via the eechidna R package.},
+            pdftitle={Spatial modelling of the two-party preferred vote in Australian federal elections: 2001--2016},
             pdfkeywords={federal election, Census, Australia, spatial modelling, imputation, socio-demographics, electorates, R, eechidna},
@@
 
-\title{Spatial modelling of the electoral two-party preferred vote in Australia. A study of federal elections between 2001 and 2016 via the \texttt{eechidna} \texttt{R} package.}
+\title{Spatial modelling of the two-party preferred vote in Australian federal elections: 2001--2016}
 
@@
 
-\author{Jeremy~Forbes, Di~Cook, Rob~Hyndman}
+\author{Jeremy~Forbes, Di~Cook, Rob J~Hyndman}
 \addresses{\textbf{Jeremy Forbes}\newline
@@
 \\[1cm]
-\textbf{Rob Hyndman}\newline
+\textbf{Rob J Hyndman}\newline
 Monash University\newline Melbourne, Australia

### 2019-04-01 functions.R
functions code

@@ -1,0 +1,294 @@
+# Function to standardize variables
+standardise_vars <- function(df) {
+  hold <- df
+  num_cols <- sapply(df, class) == "numeric"
+  df[, num_cols] <- lapply(df[, num_cols], scale)
+  names(df) <- names(hold)
+  df$LNP_Percent <- hold$LNP_Percent
+  return(df)
+}
+
+# Compute spatial weights matrix
+sp_weights_matrix <- function(sF) {
+  dist_matrix <- function(shapefile) {
+    dist_mat <- matrix(NA, nrow = 150, ncol = 150)
+    rownames(dist_mat) <- sort(shapefile$elect_div)
+    colnames(dist_mat) <- sort(shapefile$elect_div)
+
+    for (i in 1:(nrow(dist_mat) - 1)) {
+      rname <- rownames(dist_mat)[i]
+      row_poly <- shapefile %>% subset(elect_div == rname)
+
+      for (j in (i + 1):ncol(dist_mat)) {
+        cname <- rownames(dist_mat)[j]
+        col_poly <- shapefile %>% subset(elect_div == cname)
+        dist <- gDistance(row_poly, col_poly)
+        dist_mat[i, j] <- dist
+      }
+      print(i)
+    }
+
+    # Now copy to lower triange
+    for (i in 2:nrow(dist_mat)) {
+      for (j in 1:(i - 1)) {
+        dist_mat[i, j] <- dist_mat[j, i]
+      }
+    }
+
+    # Check it is symmetric
+    if (!isSymmetric(dist_mat)) {
+      print("Warning! Matrix is not symmetric. Error has occured.")
+    }
+
+    return(dist_mat)
+  }
+
+  my_dist_mat <- dist_matrix(sF)
+
+  # S matrix
+
+  s_mat <- function(dist_mat) {
+    s_mat <- dist_mat
+
+    for (i in 1:nrow(dist_mat)) {
+      for (j in 1:nrow(dist_mat)) {
+        a <- dist_mat[i, j]
+
+        if (is.na(a)) {
+          b <- 0
+        } else {
+          b <- ifelse(a == 0, 1, 0)
+        }
+
+        s_mat[i, j] <- b
+      }
+    }
+
+    return(s_mat)
+  }
+
+  my_smat <- s_mat(my_dist_mat)
+
+  # Turn into W matrix
+
+  w_mat <- function(s_mat) {
+    w_mat <- s_mat
[…]

### 2019-04-01 draft-paper.Rmd
Minor edits

@@
 The formats of these files differ over the years, making extracting the appropriate information a big task.
-The functions available in the `dplyr` [@dplyr] and `readxl` [@readxl] `R` packages are very useful, as they provide fast consistent tools for data manipulation and functions to import .xlsx files (respectively).
+The functions available in the `dplyr` [@dplyr] and `readxl` [@readxl] `R` packages are particularly useful, as they provide fast consistent tools for data manipulation and functions to import .xlsx files (respectively).
 The 2001 and 2006 Census data are published in a format where the information for each electorate is held in a separate document making it difficult to use the `dplyr` tools.
@@
 For each cross-section, let the response $\boldsymbol y$ be the vector two-party preferred vote in favour of the Liberal party, with $y_i = 70$ representing a 70% preference for Liberal, 30% for Labor, in electorate $i$.
-Although $y_i$ lies in the interval $(0,100)$, observed values are never very close to 0 or 100 (minimum $24.05 \%$ and maximum $74.90 \%$), so there is no need to formally impose the constraint of $y_i \in [0,100]$.
+Although $y_i$ lies in the interval $(0,100)$, observed values are never close to 0 or 100 (minimum $24.05 \%$ and maximum $74.90 \%$), so there is no need to formally impose the constraint of $y_i \in [0,100]$.
 Furthermore, the responses are found to be spatially correlated in each election (Moran's I test, $p \le 7\cdot10^{-15}$).
@@
 ## Country-wide trend
+
 Since all socio-demographics have been standardized to have a mean of zero and a variance of one, the intercept in each model can be interpreted as the estimated two-party preferred vote for an electorate with mean characteristics[^3].
@@
 ## Influential socio-demographics
+
 To investigate the socio-demographics that have a strong effect on the two-party preferred vote, partial residual plots are used and shown in Figures \@ref(fig:partresplot1) and \@ref(fig:partresplot2).
@@
 
-Based on the distribution of the cook's distance values, a cook's distance greater than $0.1$ is considered to be influential and a potential outlier.
-The electorate of Sydney (NSW) has a large cook's distance from 2001 to 2013, due to its diverse population (language, birthplace and religion), high number of defacto relationships, high income, high household mobility and small amount of workers in extractive and transformative jobs.
+Based on the distribution of the Cook's distance values, a Cook's distance greater than $0.1$ is considered to be influential and a potential outlier.
+The electorate of Sydney (NSW) has a large Cook's distance from 2001 to 2013, due to its diverse population (language, birthplace and religion), high number of defacto relationships, high income, high household mobility and small amount of workers in extractive and transformative jobs.
 It has remained a strong supporter of the Labor party and Liberal vote is severely overpredicted by the model, making it an outlier.
@@
 Lingiari, an electorate taking up almost all of the Northern Territory, is an outlier in the 2001-2007 elections due to its large Indigenous population, young age profile and low rates of property ownership.
-Fowler (NSW) has a diverse population with a high proportion of migrants, many Buddhists and Muslims, and has very strong Labor support, making it influential in 2001, 2004 and 2010.
-Other electorates with large cook's distance are Barton (NSW) and Leichhardt (QLD) in 2016, and Canberra (ACT) in 2007.
+Fowler (NSW) has a diverse population with a high proportion of migrants, many Buddhists and Muslims, and has strong Labor support, making it influential in 2001, 2004 and 2010.
+Other electorates with large Cook's distance are Barton (NSW) and Leichhardt (QLD) in 2016, and Canberra (ACT) in 2007.
 

### 2019-04-01 draft-paper.tex
Minor edits

@@
 The formats of these files differ over the years, making extracting the appropriate information a big task.
-The functions available in the \texttt{dplyr} \autocite{dplyr} and \texttt{readxl} \autocite{readxl} \texttt{R} packages are very useful, as they provide fast consistent tools for data manipulation and functions to import .xlsx files (respectively).
+The functions available in the \texttt{dplyr} \autocite{dplyr} and \texttt{readxl} \autocite{readxl} \texttt{R} packages are particularly useful, as they provide fast consistent tools for data manipulation and functions to import .xlsx files (respectively).
 The 2001 and 2006 Census data are published in a format where the information for each electorate is held in a separate document making it difficult to use the \texttt{dplyr} tools.
@@
 For each cross-section, let the response \(\boldsymbol y\) be the vector two-party preferred vote in favour of the Liberal party, with \(y_i = 70\) representing a 70\% preference for Liberal, 30\% for Labor, in electorate \(i\).
-Although \(y_i\) lies in the interval \((0,100)\), observed values are never very close to 0 or 100 (minimum \(24.05 \%\) and maximum \(74.90 \%\)), so there is no need to formally impose the constraint of \(y_i \in [0,100]\).
+Although \(y_i\) lies in the interval \((0,100)\), observed values are never close to 0 or 100 (minimum \(24.05 \%\) and maximum \(74.90 \%\)), so there is no need to formally impose the constraint of \(y_i \in [0,100]\).
 Furthermore, the responses are found to be spatially correlated in each election (Moran's I test, \(p \le 7\cdot10^{-15}\)).
@@
 
-Based on the distribution of the cook's distance values, a cook's distance greater than \(0.1\) is considered to be influential and a potential outlier.
-The electorate of Sydney (NSW) has a large cook's distance from 2001 to 2013, due to its diverse population (language, birthplace and religion), high number of defacto relationships, high income, high household mobility and small amount of workers in extractive and transformative jobs.
+Based on the distribution of the Cook's distance values, a Cook's distance greater than \(0.1\) is considered to be influential and a potential outlier.
+The electorate of Sydney (NSW) has a large Cook's distance from 2001 to 2013, due to its diverse population (language, birthplace and religion), high number of defacto relationships, high income, high household mobility and small amount of workers in extractive and transformative jobs.
 It has remained a strong supporter of the Labor party and Liberal vote is severely overpredicted by the model, making it an outlier.

### 2019-04-03 draft-paper.Rmd
Minor wording changes only

@@
   address: Monash University\newline Melbourne, Australia
-abstract: "This research examines the relationships between electoral socio-demographic characteristics and two-party preference in Australian federal elections, with the six elections from 2001 to 2016 considered.
-Socio-demographic information is derived from the Australian Census, which occurs every five years.
-Since a Census is not directly available for each election, a method of spatio-temporal imputation is employed to construct Census data for the electorates in 2004-2013 elections.
+abstract: "We examine the relationships between electoral socio-demographic characteristics and two-party preference in the six Australian federal elections held between 2001 to 2016.
+Socio-demographic information was derived from the Australian Census, which occurs every five years.
+Since a Census is not directly available for each election, spatio-temporal imputation was employed to estimate Census data for the electorates at the time of each election.
 This accounts for both spatial and temporal changes in electoral characteristics between Censuses.
-To capture any spatial heterogeneity, a spatial error model is estimated for each election, which incorporates a spatially structured random effect vector that can thought of as the unobserved political climate in each electorate.
+To capture any spatial heterogeneity, a spatial error model was estimated for each election, which incorporates a spatially structured random effect vector that can be thought of as the unobserved political climate in each electorate.
 Over time, the impact of most socio-demographic characteristics that affect electoral two-party preference do not vary, with industry of work, incomes, household mobility and de facto relationships having strong effects in each of the six elections.
 Education and unemployment are amongst those that have varying effects.
-It is also found that between 2004 and 2013, the spatial effect is insignificant, meaning that electorates voted effectively independently.
+It is also found that between 2004 and 2013, the spatial effect was insignificant, meaning that electorates voted effectively independently.
 All data featured in this study has been contributed to the `eechidna` `R` package (available on CRAN)."
@@
 Joining these two data sources is problematic as there in an inherent asynchronicity in the two types of events.
-A Census is conducted by the Australian Bureau of Statistics (ABS) every five years, whereas a federal election (conducted by the Australian Electoral Commission (AEC)) usually occurs every three years.
-The first problem addressed is that of constructing appropriate Census data for the 2004, 2007, 2010 and 2013 elections - election years in which a Census does not occur.
-The predominant approach in previous studies is to join voting outcomes to the nearest Census, without accounting for any temporal differences (see @DavisStimson98, @Stimson06, @Liao09 and @Stimson09).
+A Census is conducted by the Australian Bureau of Statistics (ABS) every five years, whereas federal elections, conducted by the Australian Electoral Commission (AEC), usually occur every three years or so.
+The first problem addressed is that of constructing appropriate Census data for the 2004, 2007, 2010 and 2013 elections --- election years in which a Census does not occur.
+The predominant approach in previous studies was to join voting outcomes to the nearest Census, without accounting for any temporal differences (see @DavisStimson98, @Stimson06, @Liao09 and @Stimson09).
 Furthermore, electoral boundaries change regularly, so spatial discrepancies also arise when matching with electoral data.
-To obtain appropriate Census data for these four elections, electoral socio-demographics are constructed using a spatio-temporal imputation that combines areal interpolation [@Goodchild1993] and linear time-interpolation.
+To obtain appropriate "Census-like" data for these four elections, electoral socio-demographics are constructed using a spatio-temporal imputation that combines areal interpolation [@Goodchild1993] and linear time-interpolation.
 Collecting and wrangling the raw data, along with the imputation process, are detailed in Section \@ref(data).
-All data and associated documentation relating to this procedure are available in the `eechidna` `R` package [@eechidna], providing a resource for future analysis .
-
-Previous work on modelling Australian federal elections have found that aggregate socio-demographics are relatively good predictors of voting outcomes.
-@Forrest01 use multiple regression to model the Liberal and Labor primary vote for polling booths in the Farrer electorate in 1998 as a function of Census variables from 1996.
-@Stimson06, @Stimson09 and @Stimson12 use principal component analysis of polling booths in the 2001, 2004 and 2007 elections respectively, also finding that socio-demographic characteristics of polling booths are linked to their two-party preferred vote.
-On the contrary, @Stimson09 models the polling booth swing vote (change in the two-party preferred vote) in the 2007 election, finding that little of swing vote can be explained by Census data.
+All data and associated documentation relating to this procedure are available in the `eechidna` `R` package [@eechidna], providing a resource for future analysis.
+
+Previous work on modelling Australian federal elections has found that aggregate socio-demographics are relatively good predictors of voting outcomes.
+@Forrest01 used multiple regression to model the Liberal and Labor primary vote for polling booths in the Farrer electorate in 1998 as a function of Census variables from 1996.
+@Stimson06, @Stimson09 and @Stimson12 used principal component analysis of polling booths in the 2001, 2004 and 2007 elections respectively, also finding that socio-demographic characteristics of polling booths are linked to their two-party preferred vote.
+In contrast, @Stimson09 models the polling booth swing vote (change in the two-party preferred vote) in the 2007 election, finding that little of the swing vote can be explained by Census data.
 Instead of analyzing a single election in isolation, this paper employs a consistent model framework across six elections so that temporal changes in the effects of socio-demographics can be observed, where each federal elections is modelled with a cross-sectional dataset.
 The use of a regression framework to examine these socio-political relationships over time is seemingly absent from previous Australian studies.
-It also appears that no study has attempted any type of statistical analysis of socio-demographics in conjunction with voter behaviour in Australia since 2007, making this paper distinctly different from those previous.
-
-The cross-sectional dataset for each election used here consists of the two-party preferred vote (response variable), and a set of common socio-demographic variables (explanatory variables) that characterize each electorate.
-To obtain these datasets, socio-demographic variables are first standardized, and then a principal component analysis is used to group variables into "factors".
-To account for the inherent spatial structure of the data, a spatial error model is then fit for each election.
+It also appears that no study has attempted any type of statistical analysis of socio-demographics in conjunction with voter behaviour in Australia since 2007, making this paper distinctly different from previous work.
+
+The cross-sectional dataset for each election used here consists of the two-party preferred vote (as the response variable), and a set of common socio-demographic variables (as the explanatory variables) that characterize each electorate.
+To prepare these datasets, socio-demographic variables are first standardized, and then a principal component analysis is used to group variables into "factors".
+To account for the inherent spatial structure of the data, a spatial error model is then estimated for each election.
 These steps are discussed in Section \@ref(modelling).
@@
 
-Socio-demographic variables are derived from the Census of Population and Housing (Census), which is a survey of every household in Australia, recording information such as age, gender, ethnicity, education level and income.
+Socio-demographic variables were derived from the Australian Census of Population and Housing (Census), which is a survey of every household in Australia, recording information such as age, gender, ethnicity, education level and income.
 There have been four Censuses so far in the 21st century, conducted in 2001, 2006, 2011 and 2016.
 The Australian Bureau of Statistics (ABS) conducts the Census and publishes aggregated information.
-The ABS uses electoral boundaries as defined by the AEC at the time of each Census, which may not match those in place at the subsequent election.
-From the available Census information aggregated at the electorate level, 65 socio-demographic variables are defined for each of the electorates to be used in the analysis.
-
-Raw data is sourced online from the AEC and ABS websites in .csv and .xlsx files.
+The ABS uses electoral boundaries as defined by the AEC at the time of each Census, which may not match those in place at the subsequent and previous elections.
+From the available Census information aggregated at the electorate level, 65 socio-demographic variables were defined for each of the electorates to be used in the analysis.
+
+Raw data was sourced online from the AEC and ABS websites in `.csv` and `.xlsx` files.
 The formats of these files differ over the years, making extracting the appropriate information a big task.
-The functions available in the `dplyr` [@dplyr] and `readxl` [@readxl] `R` packages are particularly useful, as they provide fast consistent tools for data manipulation and functions to import .xlsx files (respectively).
+The functions available in the `dplyr` [@dplyr] and `readxl` [@readxl] `R` packages are particularly useful, as they provide fast consistent tools for data manipulation and functions to import `.xlsx` files (respectively).
 The 2001 and 2006 Census data are published in a format where the information for each electorate is held in a separate document making it difficult to use the `dplyr` tools.
@@
 Between 2001 and 2016 there were six elections and four Censuses (see Figure \@ref(fig:timeline)).
-Electoral boundaries are redistributed regularly by the AEC, meaning that only in the years where both a Census and election occur will the boundaries match - the case for the 2001 and 2016 elections.
+Electoral boundaries are redistributed regularly by the AEC, meaning that only in the years where both a Census and election occur are all boundaries likely to match --- the case for the 2001 and 2016 elections.
 Therefore, for the four elections between 2004 and 2013, both temporal and spatial differences in electorates need to be accounted for when joining the electoral two-party preferred vote with Census data.
-For these elections a spatio-temporal imputation method is employed to obtain electoral socio-demographics.
+For these elections a spatio-temporal imputation method was employed to obtain electoral socio-demographics.
 This method uses Census information from both before and after the election of interest.
@@
 To account for spatial differences, the piece-wise approximation method in @Goodchild1993 is adopted.
-Consider a map of source zones $s = 1,...,S$, for which socio-demographic information is available, and a set of target zones $t = 1,...,T$ for which information is to be imputed.
+Consider a map of source zones $s = 1,\dots,S$, for which socio-demographic information is available, and a set of target zones $t = 1,\dots,T$ for which information is to be imputed.
 This is described in the context of a single election, and a single neighbouring Census.
@@
 Additionally, let the population of the source zone $s$ be $U_s$ and the population of intersection between source zone $s$ and target zone $t$ be $P_{s,t}$.
-The estimated population of intersection is given by:
-
-$$\hat{P}_{s,t} = \frac{U_s*A_{s,t}}{\sum_{t=1}^T A_{s,t}}, \mbox{ for all } s=1,...,S \mbox{ and } t=1,...,T.$$
+The estimated population of intersection is given by
+$$
+  \hat{P}_{s,t} = \frac{U_s*A_{s,t}}{\sum_{t=1}^T A_{s,t}},
+  \quad\text{for all $s=1,\dots,S$ and $t=1,\dots,T$}.
+$$
 Note that this estimator implicitly assumes that populations are uniformly distributed within each source zone.
@@
 Denote a given Census variable for the target zone by $C_t$, and the same Census variable for the source zone as $D_s$.
-Then, estimate $C_t$ using:
-
-$$\hat{C}_t = \frac{\sum_{s=1}^{S}{D_s*\hat{P}_{s,t}}}{\sum_{s=1}^{S}{\hat{P}_{s,t}}}, \mbox{ for each } t=1,...,T.$$
-This concludes the spatial imputation of the socio-demographic characteristics for one target zone (a single electoral boundary), at time of only one of the neighbouring Censuses.
+Then, estimate $C_t$ using
+$$
+  \hat{C}_t = \frac{\sum_{s=1}^{S}{D_s*\hat{P}_{s,t}}}%
+  {\sum_{s=1}^{S}{\hat{P}_{s,t}}},
+  \quad\text{for each $t=1,\dots,T$}.
+$$
+This concludes the spatial imputation of the socio-demographic characteristics for one target zone (a single electoral boundary), at the time of only one of the neighbouring Censuses.
 This process is repeated for all of the target zones, and then for the other neighbouring Census.
@@
 To account for temporal changes, linear interpolation is used between Census years to get the final estimate of a Census variable for the target zone in the election year.
-Let $y_1$ be the year of the Census preceding an election, let $y_2$ be the year of the election and $y_3$ be the year of the Census that follows.
+Let $y_1$ be the year of the Census preceding an election, let $y_2$ be the year of the election, and $y_3$ be the year of the Census that follows.
 Add this year subscript to the Census variable estimate $\hat{C}_t$, resulting in $\hat{C}_{t,y}$.
 Linear interpolating between these Census years results an imputed value for the election year, given by
-
-$$\hat{C}_{t,y_2} = \frac{y_3-y_2}{y_3-y_1}*\hat{C}_{t,y_1} + \frac{y_2-y_1}{y_3-y_1}*\hat{C}_{t,y_3}.$$
+$$
+  \hat{C}_{t,y_2} = \frac{y_3-y_2}{y_3-y_1} * \hat{C}_{t,y_1} +
+    \frac{y_2-y_1}{y_3-y_1} * \hat{C}_{t,y_3}.
+$$
 Implicitly this assumes that population characteristics change in a linear manner over time.
@@
 
-Census data is publicly available that different levels of aggregation, ranging from SA1 (over 50,000 zones) to electoral divisions (150 zones).
-For this study electoral divisions are used as source zones, and the imputation method is applied to produce socio-demographic variables for each of the 2004, 2007, 2010 and 2013 elections.
+Census data is publicly available at different levels of aggregation, ranging from SA1 (over 50,000 zones) to electoral divisions (150 zones).
+For this study, electoral divisions are used as source zones, and the imputation method is applied to produce socio-demographic variables for each of the 2004, 2007, 2010 and 2013 elections.
 As mentioned earlier, there is no need to impute socio-demographic variables for the 2001 and 2016 elections.
@@
 Source Zone (2016) | Proportion | Source Zone Population | Population Allocated to Purple Region: $\hat{P}_{s,t}$
---------------------|----------|-----------|---------------------------------------------
+--------------------|----------|-----------|--------------------------------------------:
 HUME              | 0.9654     | 150643                   | 145427
@@
 EDEN-MONARO       | 0.1109     | 147532                   | 16358
-CANBERRA          | 0.028      | 196037                   | 548
+CANBERRA          | 0.0280     | 196037                   | 548
 FENNER            | 0.0023     | 202955                   | 474
@@
 
-Now consider the socio-demographic variable $AusCitizen$ - the proportion of people in the region who are Australian citizens.
+Now consider the socio-demographic variable *AusCitizen*, the proportion of people in the region who are Australian citizens.
 
@@
 |Source Zone (2016)   | AusCitizen (%): $D_s$| Estimated Population Allocated to Purple Region: $\hat{P}_{s,t}$|
-|--------------------------|-------------------|------------------------------------------------|
+|--------------------------|-------------------|-----------------------------------------------:|
 |HUME            |              90.02|                                          145427|
@@
 ### Industry and type of work
+
 Electorates with higher proportions of workers in mining, gas, water, agriculture, waste and electricity (grouped as `Extractive` industries) are consistently linked with higher support for the Liberal party, with the magnitude of this effect slightly increasing over the years (see row 3 in Figure \@ref(fig:partresplot1)).

### 2019-04-03 draft-paper.Rmd
Reverted abstract to present tense

@@
 abstract: "We examine the relationships between electoral socio-demographic characteristics and two-party preference in the six Australian federal elections held between 2001 to 2016.
-Socio-demographic information was derived from the Australian Census, which occurs every five years.
-Since a Census is not directly available for each election, spatio-temporal imputation was employed to estimate Census data for the electorates at the time of each election.
+Socio-demographic information is derived from the Australian Census, which occurs every five years.
+Since a Census is not directly available for each election, spatio-temporal imputation is employed to estimate Census data for the electorates at the time of each election.
 This accounts for both spatial and temporal changes in electoral characteristics between Censuses.
-To capture any spatial heterogeneity, a spatial error model was estimated for each election, which incorporates a spatially structured random effect vector that can be thought of as the unobserved political climate in each electorate.
+To capture any spatial heterogeneity, a spatial error model is estimated for each election, which incorporates a spatially structured random effect vector that can be thought of as the unobserved political climate in each electorate.
 Over time, the impact of most socio-demographic characteristics that affect electoral two-party preference do not vary, with industry of work, incomes, household mobility and de facto relationships having strong effects in each of the six elections.

### 2019-04-03 draft-paper.Rmd
Use patchwork rather than grid/gridExtra

@@
 library(spdep)
-library(gridExtra)
-library(grid)
+library(patchwork)
+#library(gridExtra)
+#library(grid)
 library(predictmeans)
@@
 ```{r partresplot1, fig.cap = "Partial residual plots for: Incomes, Unemployment, Extractive, Transformative, ManagerAdmin and DiffAddress", fig.height = 9, fig.width = 6, warning = FALSE, fig.pos = 'h'}
-grid.arrange(p1, p2, p3, p4, p5, p6,
-  nrow = 6,
-  heights = c(1.2, 1, 1, 1, 1, 1),
-  left = grid.text("Two-party preferred vote (%)", gp = gpar(cex = 0.8), rot = 90)
-)
+(p1 / p2 / (p3 + ylab("Two-party preferred vote (%)                     ")) / p4 / p5 / p6)
 ```
@@
 ```{r partresplot2, fig.cap = "Partial residual plots for: DeFacto, MedianAge, Education, OtherLanguage, BornSEEuro, BornAsia and Judaism.", fig.height = 9,  fig.width = 6, warning = FALSE, fig.pos = 'h'}
-grid.arrange(p7, p8, p9, p10, p11, p12, p13,
-  nrow = 7,
-  heights = c(1.2, 1, 1, 1, 1, 1, 1),
-  left = grid.text("Two-party preferred vote (%)", gp = gpar(cex = 0.8), rot = 90)
-)
-```
+(p7 / p8 / (p9 + ylab("Two-party preferred vote (%)                     ")) / p10 / p11 / p12)
+```

### 2019-04-03 draft-paper.Rmd
Mostly work on tables in Section 2 and 3

@@
 The first problem addressed is that of constructing appropriate Census data for the 2004, 2007, 2010 and 2013 elections --- election years in which a Census does not occur.
-The predominant approach in previous studies was to join voting outcomes to the nearest Census, without accounting for any temporal differences (see @DavisStimson98, @Stimson06, @Liao09 and @Stimson09).
+The predominant approach in previous studies was to join voting outcomes to the nearest Census, without accounting for any temporal differences [see @DavisStimson98;@Stimson06;@Liao09;@Stimson09].
 Furthermore, electoral boundaries change regularly, so spatial discrepancies also arise when matching with electoral data.
@@
 
+\enlargethispage*{0.5cm}
+
 To account for temporal changes, linear interpolation is used between Census years to get the final estimate of a Census variable for the target zone in the election year.
@@
 
-### An illustration of the spatio-temporal imputation
+### An illustration of the spatio-temporal imputation {-}
 
@@
 ```
-
-The corresponding source zones from the 2016 Census are shown in Figure \@ref(fig:hume16).
-As can be seen, the Hume boundary from the 2013 election (shaded purple) does not perfectly match any of the source zones.
 
@@
 
+The corresponding source zones from the 2016 Census are shown in Figure \@ref(fig:hume16).
+As can be seen, the Hume boundary from the 2013 election (shaded purple) does not perfectly match any of the source zones.
+
 There are many source zones from the 2016 Census that intersect with this purple region, including the divisions of Riverina, Eden-Monaro and Hume, along with smaller intersecting areas with Fenner, Calare, Gilmore and Whitlam.
@@
 
-\footnotesize
-
-Source Zone (2016) | Proportion | Source Zone Population | Population Allocated to Purple Region: $\hat{P}_{s,t}$
---------------------|----------|-----------|--------------------------------------------:
-HUME              | 0.9654     | 150643                   | 145427
-RIVERINA          | 0.2511     | 155793                   | 39117
-EDEN-MONARO       | 0.1109     | 147532                   | 16358
-CANBERRA          | 0.0280     | 196037                   | 548
-FENNER            | 0.0023     | 202955                   | 474
-WHITLAM           | 0.0006     | 152280                   | 92
-GILMORE           | 0.0006     | 150436                   | 86
-CALARE            | 0.0001     | 161298                   | 21
-
-\normalsize
+```{r hume1}
+hume <- abs2016 %>%
+  filter(DivisionNm %in%
+           c("HUME", "RIVERINA", "EDEN-MONARO", "CANBERRA",
+             "FENNER", "WHITLAM", "GILMORE", "CALARE")) %>%
+  mutate(
+    Pst = c(21, 548, 16358, 474, 86, 145427, 39117, 92),
+    Proportion = Pst/Population,
+    DivisionNm = tools::toTitleCase(tolower(DivisionNm))
+  ) %>%
+  select(DivisionNm, Proportion, Population, Pst, AusCitizen) %>%
+  arrange(desc(Proportion))
+
+hume %>%
+  select(-AusCitizen) %>%
+  kable(longtable=FALSE, booktabs=TRUE, digits=4,
+        format='latex',escape=FALSE,
+        col.names=c(
+          "Source zone (2016)",
+          "Proportion",
+          "Source zone population",
+          "Population allocated to purple region: $\\hat{P}_{s,t}$")) %>%
+  kableExtra::kable_styling(font_size=10, latex_options='scale_down')
+```
 
@@
 
-```{r eval = FALSE, include = FALSE}
-auscitizen_ex <- abs2016 %>%
-  filter(DivisionNm %in% c("HUME", "RIVERINA", "EDEN-MONARO", "CANBERRA", "FENNER", "WHITLAM", "GILMORE", "CALARE")) %>%
-  select(DivisionNm, AusCitizen) %>%
+```{r hume2, dependson='hume1'}
+hume %>%
+  select(DivisionNm, AusCitizen, Pst) %>%
+  kable(longtable=FALSE, booktabs=TRUE, digits=2,
+        format='latex',escape=FALSE,
+        col.names=c(
+          "Source zone (2016)",
+          "AusCitizen (\\%): $D_s$",
+          "Population allocated to purple region: $\\hat{P}_{s,t}$")) %>%
+  kableExtra::kable_styling(font_size=10)
+```
+
+```{r wave, include=FALSE}
+wave2016 <- weighted.mean(x = hume$AusCitizen, w = hume$Pst)
+hume11 <- abs2011 %>%
+  filter(DivisionNm %in%
+           c("HUME", "RIVERINA", "EDEN-MONARO", "CANBERRA",
+             "FENNER", "WHITLAM", "GILMORE", "CALARE")) %>%
   mutate(
-    `Estimated Population Allocated to Purple Region` = c(21, 548, 16358, 474, 86, 145427, 39117, 92),
-    AusCitizen = round(AusCitizen, 2)
+    Pst = c(21, 548, 16358, 474, 86, 145427), # Needs adjusting
+    Proportion = Pst/Population,
+    DivisionNm = tools::toTitleCase(tolower(DivisionNm))
   ) %>%
-  arrange(-`Estimated Population Allocated to Purple Region`) %>%
-  rename(`Source Zone (2016)` = DivisionNm)
-
-auscitizen_ex %>%
-  kable()
-
-weighted.mean(x = auscitizen_ex$AusCitizen, w = auscitizen_ex$`Estimated Population Allocated to Purple Region`)
-```
-
-\footnotesize
-
-|Source Zone (2016)   | AusCitizen (%): $D_s$| Estimated Population Allocated to Purple Region: $\hat{P}_{s,t}$|
-|--------------------------|-------------------|-----------------------------------------------:|
-|HUME            |              90.02|                                          145427|
-|RIVERINA        |              89.11|                                           39117|
-|EDEN-MONARO     |              88.00|                                           16358|
-|CANBERRA        |              85.48|                                             548|
-|FENNER          |              83.64|                                             474|
-|WHITLAM         |              89.52|                                              92|
-|GILMORE         |              89.03|                                              86|
-|CALARE          |              87.56|                                              21|
-
-\normalsize
-
-A weighted average of $AusCitizen$, with the allocated population from each source zone as weights, yields $\hat{C}_{Hume,2016} = 89.65 \%$.
-Repeating this process using the 2011 Census yields $\hat{C}_{Hume,2011} = 91.00 \%$
-
+  select(DivisionNm, Proportion, Population, Pst, AusCitizen) %>%
+  arrange(desc(Proportion))
+wave2011 <- 91.00 #weighted.mean(x = hume11$AusCitizen, w = hume11$Pst)
+```
+
+A weighted average of *AusCitizen*, with the allocated population from each source zone as weights, yields $\hat{C}_{\text{Hume},2016} = `r sprintf("%4.2f",wave2016)`\%$.
+Repeating this process using the 2011 Census yields $\hat{C}_{\text{Hume},2011} = `r sprintf("%4.2f",wave2011)`\%$.
 Finally, linear interpolation between 2011 and 2016 yields the 2013 estimate:
-\begin{eqnarray*}
-\hat{C}_{Hume,2013} & = &\frac{3}{5} \cdot \hat{C}_{Hume,2011} + \frac{2}{5} \cdot \hat{C}_{Hume,2016} \\
-& = & \frac{3}{5} \cdot 91.00 \% + \frac{2}{5} \cdot 89.65 \% \\
-& = & 90.46 \%
-\end{eqnarray*}
-
+$$
+  \hat{C}_{\text{Hume},2013}
+    = \frac{3}{5} \hat{C}_{\text{Hume},2011} + \frac{2}{5} \hat{C}_{\text{Hume},2016}
+    % = \frac{3}{5} \cdot 91.00 \% + \frac{2}{5} \cdot 89.65 \% \\
+    =  90.46\%.
+$$
 This is done for each of the 65 socio-demographic variables, and is repeated for each of the 149 remaining target zones corresponding with 2013 electorates.
@@
 In this context, the random effect can be thought of as capturing the unobserved political climate in each electorate, where this climate is correlated with the climate in neighbouring electorates, under the assumption that the climate is independent of electoral socio-demographics.
-
-Spatial weights are calculated in accordance with the assumption that an electorate is equally correlated with any electorate that shares a part of its boundary.
-Let $\rho$ be spatial autoregressive coefficient, $\boldsymbol v$ be a spherical error term, ${\boldsymbol W}$ be a matrix of spatial weights (containing information about the neighbouring regions), $\boldsymbol X$ be a matrix of socio-demographic covariates, $\boldsymbol \beta$ be a vector of regression coefficients and $\boldsymbol a$ be a spatially structured random effect vector.
-
-$${\boldsymbol y} = {\boldsymbol X} {\boldsymbol \beta} + {\boldsymbol a},$$
-and
-
-$${\boldsymbol a} = \rho {\boldsymbol W} {\boldsymbol a} + {\boldsymbol v},$$
-where
-
-$${\boldsymbol v} \sim N({\boldsymbol 0}, \sigma^2 {\boldsymbol I_n}),$$.
-
-and hence
-
-$${\boldsymbol y} = {\boldsymbol X} {\boldsymbol \beta} + ({\boldsymbol I}_n-\rho {\boldsymbol W})^{-1}{\boldsymbol v}.$$
-
-Estimation of the above spatial error model is undertaken using feasible generalized least squares.
-
-Table 3 details the estimated model coefficients and their estimated standard errors, for each of the six elections. An interpretation of these estimated values is provided in the next section.
 
@@
 
-\begin{table}[!htbp] \centering
+\begin{table}[!p] \centering
   \caption{Estimated spatial error model parameters (standard errors) for each of the six election years.}
@@
 
-\newpage
+Spatial weights are calculated in accordance with the assumption that an electorate is equally correlated with any electorate that shares a part of its boundary.
+Let $\rho$ be spatial autoregressive coefficient, $\boldsymbol v$ be a spherical error term, $\bm{W}$ be a matrix of spatial weights (containing information about the neighbouring regions), $\boldsymbol X$ be a matrix of socio-demographic covariates, $\boldsymbol \beta$ be a vector of regression coefficients and $\boldsymbol a$ be a spatially structured random effect vector.
+$$
+  \bm{y} = \bm{X} \bm{\beta} + \bm{a},
+$$
+and
+$$
+  \bm{a} = \rho \bm{W} \bm{a} + \bm{v},
+$$
+where
+$$
+  \bm{v} \sim N(\bm{0}, \sigma^2 \bm{I_n}),
+$$.
+and hence
+$$
+  \bm{y} = \bm{X} \bm{\beta} + (\bm{I}_n-\rho \bm{W})^{-1}\bm{v}.
+$$
+Estimation of the above spatial error model is undertaken using feasible generalized least squares.
+
+Table 3 details the estimated model coefficients and their estimated standard errors, for each of the six elections. An interpretation of these estimated values is provided in the next section.
 

### 2019-04-03 draft-paper.tex
Mostly work on tables in Section 2 and 3

@@
 \renewenvironment{figure}[1][2] {
-    \expandafter\origfigure\expandafter[H]
+    \expandafter\origfigure\expandafter[!htbp]
 } {
@@
 \begin{abstract}
-This research examines the relationships between electoral socio-demographic characteristics and two-party preference in Australian federal elections, with the six elections from 2001 to 2016 considered.
+We examine the relationships between electoral socio-demographic characteristics and two-party preference in the six Australian federal elections held between 2001 to 2016.
 Socio-demographic information is derived from the Australian Census, which occurs every five years.
-Since a Census is not directly available for each election, a method of spatio-temporal imputation is employed to construct Census data for the electorates in 2004-2013 elections.
+Since a Census is not directly available for each election, spatio-temporal imputation is employed to estimate Census data for the electorates at the time of each election.
 This accounts for both spatial and temporal changes in electoral characteristics between Censuses.
-To capture any spatial heterogeneity, a spatial error model is estimated for each election, which incorporates a spatially structured random effect vector that can thought of as the unobserved political climate in each electorate.
+To capture any spatial heterogeneity, a spatial error model is estimated for each election, which incorporates a spatially structured random effect vector that can be thought of as the unobserved political climate in each electorate.
 Over time, the impact of most socio-demographic characteristics that affect electoral two-party preference do not vary, with industry of work, incomes, household mobility and de facto relationships having strong effects in each of the six elections.
 Education and unemployment are amongst those that have varying effects.
-It is also found that between 2004 and 2013, the spatial effect is insignificant, meaning that electorates voted effectively independently.
+It is also found that between 2004 and 2013, the spatial effect was insignificant, meaning that electorates voted effectively independently.
 All data featured in this study has been contributed to the \texttt{eechidna} \texttt{R} package (available on CRAN).
@@
 Joining these two data sources is problematic as there in an inherent asynchronicity in the two types of events.
-A Census is conducted by the Australian Bureau of Statistics (ABS) every five years, whereas a federal election (conducted by the Australian Electoral Commission (AEC)) usually occurs every three years.
-The first problem addressed is that of constructing appropriate Census data for the 2004, 2007, 2010 and 2013 elections - election years in which a Census does not occur.
-The predominant approach in previous studies is to join voting outcomes to the nearest Census, without accounting for any temporal differences (see \textcite{DavisStimson98}, \textcite{Stimson06}, \textcite{Liao09} and \textcite{Stimson09}).
+A Census is conducted by the Australian Bureau of Statistics (ABS) every five years, whereas federal elections, conducted by the Australian Electoral Commission (AEC), usually occur every three years or so.
+The first problem addressed is that of constructing appropriate Census data for the 2004, 2007, 2010 and 2013 elections --- election years in which a Census does not occur.
+The predominant approach in previous studies was to join voting outcomes to the nearest Census, without accounting for any temporal differences \autocites[see][]{DavisStimson98}{Stimson06}{Liao09}{Stimson09}.
 Furthermore, electoral boundaries change regularly, so spatial discrepancies also arise when matching with electoral data.
-To obtain appropriate Census data for these four elections, electoral socio-demographics are constructed using a spatio-temporal imputation that combines areal interpolation \autocite{Goodchild1993} and linear time-interpolation.
+To obtain appropriate ``Census-like'' data for these four elections, electoral socio-demographics are constructed using a spatio-temporal imputation that combines areal interpolation \autocite{Goodchild1993} and linear time-interpolation.
 Collecting and wrangling the raw data, along with the imputation process, are detailed in Section \ref{data}.
-All data and associated documentation relating to this procedure are available in the \texttt{eechidna} \texttt{R} package \autocite{eechidna}, providing a resource for future analysis .
+All data and associated documentation relating to this procedure are available in the \texttt{eechidna} \texttt{R} package \autocite{eechidna}, providing a resource for future analysis.
 
-Previous work on modelling Australian federal elections have found that aggregate socio-demographics are relatively good predictors of voting outcomes.
-\textcite{Forrest01} use multiple regression to model the Liberal and Labor primary vote for polling booths in the Farrer electorate in 1998 as a function of Census variables from 1996.
-\textcite{Stimson06}, \textcite{Stimson09} and \textcite{Stimson12} use principal component analysis of polling booths in the 2001, 2004 and 2007 elections respectively, also finding that socio-demographic characteristics of polling booths are linked to their two-party preferred vote.
-On the contrary, \textcite{Stimson09} models the polling booth swing vote (change in the two-party preferred vote) in the 2007 election, finding that little of swing vote can be explained by Census data.
+Previous work on modelling Australian federal elections has found that aggregate socio-demographics are relatively good predictors of voting outcomes.
+\textcite{Forrest01} used multiple regression to model the Liberal and Labor primary vote for polling booths in the Farrer electorate in 1998 as a function of Census variables from 1996.
+\textcite{Stimson06}, \textcite{Stimson09} and \textcite{Stimson12} used principal component analysis of polling booths in the 2001, 2004 and 2007 elections respectively, also finding that socio-demographic characteristics of polling booths are linked to their two-party preferred vote.
+In contrast, \textcite{Stimson09} models the polling booth swing vote (change in the two-party preferred vote) in the 2007 election, finding that little of the swing vote can be explained by Census data.
 Instead of analyzing a single election in isolation, this paper employs a consistent model framework across six elections so that temporal changes in the effects of socio-demographics can be observed, where each federal elections is modelled with a cross-sectional dataset.
 The use of a regression framework to examine these socio-political relationships over time is seemingly absent from previous Australian studies.
-It also appears that no study has attempted any type of statistical analysis of socio-demographics in conjunction with voter behaviour in Australia since 2007, making this paper distinctly different from those previous.
+It also appears that no study has attempted any type of statistical analysis of socio-demographics in conjunction with voter behaviour in Australia since 2007, making this paper distinctly different from previous work.
 
-The cross-sectional dataset for each election used here consists of the two-party preferred vote (response variable), and a set of common socio-demographic variables (explanatory variables) that characterize each electorate.
-To obtain these datasets, socio-demographic variables are first standardized, and then a principal component analysis is used to group variables into ``factors''.
-To account for the inherent spatial structure of the data, a spatial error model is then fit for each election.
+The cross-sectional dataset for each election used here consists of the two-party preferred vote (as the response variable), and a set of common socio-demographic variables (as the explanatory variables) that characterize each electorate.
+To prepare these datasets, socio-demographic variables are first standardized, and then a principal component analysis is used to group variables into ``factors''.
+To account for the inherent spatial structure of the data, a spatial error model is then estimated for each election.
 These steps are discussed in Section \ref{modelling}.
@@
 
-Socio-demographic variables are derived from the Census of Population and Housing (Census), which is a survey of every household in Australia, recording information such as age, gender, ethnicity, education level and income.
+Socio-demographic variables were derived from the Australian Census of Population and Housing (Census), which is a survey of every household in Australia, recording information such as age, gender, ethnicity, education level and income.
 There have been four Censuses so far in the 21st century, conducted in 2001, 2006, 2011 and 2016.
 The Australian Bureau of Statistics (ABS) conducts the Census and publishes aggregated information.
-The ABS uses electoral boundaries as defined by the AEC at the time of each Census, which may not match those in place at the subsequent election.
-From the available Census information aggregated at the electorate level, 65 socio-demographic variables are defined for each of the electorates to be used in the analysis.
+The ABS uses electoral boundaries as defined by the AEC at the time of each Census, which may not match those in place at the subsequent and previous elections.
+From the available Census information aggregated at the electorate level, 65 socio-demographic variables were defined for each of the electorates to be used in the analysis.
 
-Raw data is sourced online from the AEC and ABS websites in .csv and .xlsx files.
+Raw data was sourced online from the AEC and ABS websites in \texttt{.csv} and \texttt{.xlsx} files.
 The formats of these files differ over the years, making extracting the appropriate information a big task.
-The functions available in the \texttt{dplyr} \autocite{dplyr} and \texttt{readxl} \autocite{readxl} \texttt{R} packages are particularly useful, as they provide fast consistent tools for data manipulation and functions to import .xlsx files (respectively).
+The functions available in the \texttt{dplyr} \autocite{dplyr} and \texttt{readxl} \autocite{readxl} \texttt{R} packages are particularly useful, as they provide fast consistent tools for data manipulation and functions to import \texttt{.xlsx} files (respectively).
 The 2001 and 2006 Census data are published in a format where the information for each electorate is held in a separate document making it difficult to use the \texttt{dplyr} tools.
@@
 Between 2001 and 2016 there were six elections and four Censuses (see Figure \ref{fig:timeline}).
-Electoral boundaries are redistributed regularly by the AEC, meaning that only in the years where both a Census and election occur will the boundaries match - the case for the 2001 and 2016 elections.
+Electoral boundaries are redistributed regularly by the AEC, meaning that only in the years where both a Census and election occur are all boundaries likely to match --- the case for the 2001 and 2016 elections.
 Therefore, for the four elections between 2004 and 2013, both temporal and spatial differences in electorates need to be accounted for when joining the electoral two-party preferred vote with Census data.
-For these elections a spatio-temporal imputation method is employed to obtain electoral socio-demographics.
+For these elections a spatio-temporal imputation method was employed to obtain electoral socio-demographics.
 This method uses Census information from both before and after the election of interest.
@@
 To account for spatial differences, the piece-wise approximation method in \textcite{Goodchild1993} is adopted.
-Consider a map of source zones \(s = 1,...,S\), for which socio-demographic information is available, and a set of target zones \(t = 1,...,T\) for which information is to be imputed.
+Consider a map of source zones \(s = 1,\dots,S\), for which socio-demographic information is available, and a set of target zones \(t = 1,\dots,T\) for which information is to be imputed.
 This is described in the context of a single election, and a single neighbouring Census.
@@
 Additionally, let the population of the source zone \(s\) be \(U_s\) and the population of intersection between source zone \(s\) and target zone \(t\) be \(P_{s,t}\).
-The estimated population of intersection is given by:
-
-\[\hat{P}_{s,t} = \frac{U_s*A_{s,t}}{\sum_{t=1}^T A_{s,t}}, \mbox{ for all } s=1,...,S \mbox{ and } t=1,...,T.\]
+The estimated population of intersection is given by
+\[
+  \hat{P}_{s,t} = \frac{U_s*A_{s,t}}{\sum_{t=1}^T A_{s,t}},
+  \quad\text{for all $s=1,\dots,S$ and $t=1,\dots,T$}.
+\]
 Note that this estimator implicitly assumes that populations are uniformly distributed within each source zone.
@@
 Denote a given Census variable for the target zone by \(C_t\), and the same Census variable for the source zone as \(D_s\).
-Then, estimate \(C_t\) using:
-
-\[\hat{C}_t = \frac{\sum_{s=1}^{S}{D_s*\hat{P}_{s,t}}}{\sum_{s=1}^{S}{\hat{P}_{s,t}}}, \mbox{ for each } t=1,...,T.\]
-This concludes the spatial imputation of the socio-demographic characteristics for one target zone (a single electoral boundary), at time of only one of the neighbouring Censuses.
+Then, estimate \(C_t\) using
+\[
+  \hat{C}_t = \frac{\sum_{s=1}^{S}{D_s*\hat{P}_{s,t}}}%
+  {\sum_{s=1}^{S}{\hat{P}_{s,t}}},
+  \quad\text{for each $t=1,\dots,T$}.
+\]
+This concludes the spatial imputation of the socio-demographic characteristics for one target zone (a single electoral boundary), at the time of only one of the neighbouring Censuses.
 This process is repeated for all of the target zones, and then for the other neighbouring Census.
 
+\enlargethispage*{0.5cm}
+
 To account for temporal changes, linear interpolation is used between Census years to get the final estimate of a Census variable for the target zone in the election year.
-Let \(y_1\) be the year of the Census preceding an election, let \(y_2\) be the year of the election and \(y_3\) be the year of the Census that follows.
+Let \(y_1\) be the year of the Census preceding an election, let \(y_2\) be the year of the election, and \(y_3\) be the year of the Census that follows.
 Add this year subscript to the Census variable estimate \(\hat{C}_t\), resulting in \(\hat{C}_{t,y}\).
 Linear interpolating between these Census years results an imputed value for the election year, given by
-
-\[\hat{C}_{t,y_2} = \frac{y_3-y_2}{y_3-y_1}*\hat{C}_{t,y_1} + \frac{y_2-y_1}{y_3-y_1}*\hat{C}_{t,y_3}.\]
+\[
+  \hat{C}_{t,y_2} = \frac{y_3-y_2}{y_3-y_1} * \hat{C}_{t,y_1} +
+    \frac{y_2-y_1}{y_3-y_1} * \hat{C}_{t,y_3}.
+\]
 Implicitly this assumes that population characteristics change in a linear manner over time.
@@
 \hypertarget{an-illustration-of-the-spatio-temporal-imputation}{%
-\subsubsection{An illustration of the spatio-temporal imputation}\label{an-illustration-of-the-spatio-temporal-imputation}}
+\subsubsection*{An illustration of the spatio-temporal imputation}\label{an-illustration-of-the-spatio-temporal-imputation}}
+\addcontentsline{toc}{subsubsection}{An illustration of the spatio-temporal imputation}
 
-Census data is publicly available that different levels of aggregation, ranging from SA1 (over 50,000 zones) to electoral divisions (150 zones).
-For this study electoral divisions are used as source zones, and the imputation method is applied to produce socio-demographic variables for each of the 2004, 2007, 2010 and 2013 elections.
+Census data is publicly available at different levels of aggregation, ranging from SA1 (over 50,000 zones) to electoral divisions (150 zones).
+For this study, electoral divisions are used as source zones, and the imputation method is applied to produce socio-demographic variables for each of the 2004, 2007, 2010 and 2013 elections.
 As mentioned earlier, there is no need to impute socio-demographic variables for the 2001 and 2016 elections.
@@
 
-The corresponding source zones from the 2016 Census are shown in Figure \ref{fig:hume16}.
-As can be seen, the Hume boundary from the 2013 election (shaded purple) does not perfectly match any of the source zones.
-
 \begin{figure}[h]
@@
 
+The corresponding source zones from the 2016 Census are shown in Figure \ref{fig:hume16}.
+As can be seen, the Hume boundary from the 2013 election (shaded purple) does not perfectly match any of the source zones.
+
 There are many source zones from the 2016 Census that intersect with this purple region, including the divisions of Riverina, Eden-Monaro and Hume, along with smaller intersecting areas with Fenner, Calare, Gilmore and Whitlam.
@@
 
-\footnotesize
+\begin{table}[H]
+\centering\begingroup\fontsize{10}{12}\selectfont
 
-\begin{longtable}[]{@{}llll@{}}
+\resizebox{\linewidth}{!}{
+\begin{tabular}{lrrr}
 \toprule
-\begin{minipage}[b]{0.21\columnwidth}\raggedright
-Source Zone (2016)\strut
-\end{minipage} & \begin{minipage}[b]{0.10\columnwidth}\raggedright
-Proportion\strut
-\end{minipage} & \begin{minipage}[b]{0.11\columnwidth}\raggedright
-Source Zone Population\strut
-\end{minipage} & \begin{minipage}[b]{0.46\columnwidth}\raggedright
-Population Allocated to Purple Region: \(\hat{P}_{s,t}\)\strut
-\end{minipage}\tabularnewline
+Source zone (2016) & Proportion & Source zone population & Population allocated to purple region: $\hat{P}_{s,t}$\\
 \midrule
-\endhead
-\begin{minipage}[t]{0.21\columnwidth}\raggedright
-HUME\strut
-\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright
-0.9654\strut
-\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright
-150643\strut
-\end{minipage} & \begin{minipage}[t]{0.46\columnwidth}\raggedright
-145427\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.21\columnwidth}\raggedright
-RIVERINA\strut
-\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright
-0.2511\strut
-\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright
-155793\strut
-\end{minipage} & \begin{minipage}[t]{0.46\columnwidth}\raggedright
-39117\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.21\columnwidth}\raggedright
-EDEN-MONARO\strut
-\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright
-0.1109\strut
-\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright
-147532\strut
-\end{minipage} & \begin{minipage}[t]{0.46\columnwidth}\raggedright
-16358\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.21\columnwidth}\raggedright
-CANBERRA\strut
-\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright
-0.028\strut
-\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright
-196037\strut
-\end{minipage} & \begin{minipage}[t]{0.46\columnwidth}\raggedright
-548\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.21\columnwidth}\raggedright
-FENNER\strut
-\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright
-0.0023\strut
-\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright
-202955\strut
-\end{minipage} & \begin{minipage}[t]{0.46\columnwidth}\raggedright
-474\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.21\columnwidth}\raggedright
-WHITLAM\strut
-\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright
-0.0006\strut
-\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright
-152280\strut
-\end{minipage} & \begin{minipage}[t]{0.46\columnwidth}\raggedright
-92\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.21\columnwidth}\raggedright
-GILMORE\strut
-\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright
-0.0006\strut
-\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright
-150436\strut
-\end{minipage} & \begin{minipage}[t]{0.46\columnwidth}\raggedright
-86\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.21\columnwidth}\raggedright
-CALARE\strut
-\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright
-0.0001\strut
-\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright
-161298\strut
-\end{minipage} & \begin{minipage}[t]{0.46\columnwidth}\raggedright
-21\strut
-\end{minipage}\tabularnewline
+Hume & 0.9654 & 150643 & 145427\\
+Riverina & 0.2511 & 155793 & 39117\\
+Eden-Monaro & 0.1109 & 147532 & 16358\\
+Canberra & 0.0028 & 196037 & 548\\
+Fenner & 0.0023 & 202955 & 474\\
+\addlinespace
+Whitlam & 0.0006 & 152280 & 92\\
+Gilmore & 0.0006 & 150436 & 86\\
+Calare & 0.0001 & 161298 & 21\\
 \bottomrule
-\end{longtable}
+\end{tabular}}
+\endgroup{}
+\end{table}
 
-\normalsize
+Now consider the socio-demographic variable \emph{AusCitizen}, the proportion of people in the region who are Australian citizens.
 
-Now consider the socio-demographic variable \(AusCitizen\) - the proportion of people in the region who are Australian citizens.
+\begin{table}[H]
+\centering\begingroup\fontsize{10}{12}\selectfont
 
-\footnotesize
+\begin{tabular}{lrr}
+\toprule
+Source zone (2016) & AusCitizen (\%): $D_s$ & Population allocated to purple region: $\hat{P}_{s,t}$\\
+\midrule
+Hume & 90.02 & 145427\\
+Riverina & 89.11 & 39117\\
+Eden-Monaro & 88.00 & 16358\\
+Canberra & 85.48 & 548\\
+Fenner & 83.64 & 474\\
+\addlinespace
+Whitlam & 89.52 & 92\\
+Gilmore & 89.03 & 86\\
+Calare & 87.56 & 21\\
+\bottomrule
+\end{tabular}
+\endgroup{}
+\end{table}
 
-\begin{longtable}[]{@{}lll@{}}
-\toprule
-\begin{minipage}[b]{0.26\columnwidth}\raggedright
-Source Zone (2016)\strut
-\end{minipage} & \begin{minipage}[b]{0.19\columnwidth}\raggedright
-AusCitizen (\%): \(D_s\)\strut
-\end{minipage} & \begin{minipage}[b]{0.47\columnwidth}\raggedright
-Estimated Population Allocated to Purple Region: \(\hat{P}_{s,t}\)\strut
-\end{minipage}\tabularnewline
-\midrule
-\endhead
-\begin{minipage}[t]{0.26\columnwidth}\raggedright
-HUME\strut
-\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedright
-90.02\strut
-\end{minipage} & \begin{minipage}[t]{0.47\columnwidth}\raggedright
-145427\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.26\columnwidth}\raggedright
-RIVERINA\strut
-\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedright
-89.11\strut
-\end{minipage} & \begin{minipage}[t]{0.47\columnwidth}\raggedright
-39117\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.26\columnwidth}\raggedright
-EDEN-MONARO\strut
-\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedright
-88.00\strut
-\end{minipage} & \begin{minipage}[t]{0.47\columnwidth}\raggedright
-16358\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.26\columnwidth}\raggedright
-CANBERRA\strut
-\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedright
-85.48\strut
-\end{minipage} & \begin{minipage}[t]{0.47\columnwidth}\raggedright
-548\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.26\columnwidth}\raggedright
-FENNER\strut
-\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedright
-83.64\strut
-\end{minipage} & \begin{minipage}[t]{0.47\columnwidth}\raggedright
-474\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.26\columnwidth}\raggedright
-WHITLAM\strut
-\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedright
-89.52\strut
-\end{minipage} & \begin{minipage}[t]{0.47\columnwidth}\raggedright
-92\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.26\columnwidth}\raggedright
-GILMORE\strut
-\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedright
-89.03\strut
-\end{minipage} & \begin{minipage}[t]{0.47\columnwidth}\raggedright
-86\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.26\columnwidth}\raggedright
-CALARE\strut
-\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedright
-87.56\strut
-\end{minipage} & \begin{minipage}[t]{0.47\columnwidth}\raggedright
-21\strut
-\end{minipage}\tabularnewline
-\bottomrule
-\end{longtable}
-
-\normalsize
-
-A weighted average of \(AusCitizen\), with the allocated population from each source zone as weights, yields \(\hat{C}_{Hume,2016} = 89.65 \%\).
-Repeating this process using the 2011 Census yields \(\hat{C}_{Hume,2011} = 91.00 \%\)
-
+A weighted average of \emph{AusCitizen}, with the allocated population from each source zone as weights, yields \(\hat{C}_{\text{Hume},2016} = 89.65\%\).
+Repeating this process using the 2011 Census yields \(\hat{C}_{\text{Hume},2011} = 91.00\%\).
 Finally, linear interpolation between 2011 and 2016 yields the 2013 estimate:
-\begin{eqnarray*}
-\hat{C}_{Hume,2013} & = &\frac{3}{5} \cdot \hat{C}_{Hume,2011} + \frac{2}{5} \cdot \hat{C}_{Hume,2016} \\
-& = & \frac{3}{5} \cdot 91.00 \% + \frac{2}{5} \cdot 89.65 \% \\
-& = & 90.46 \%
-\end{eqnarray*}
-
+\[
+  \hat{C}_{\text{Hume},2013}
+    = \frac{3}{5} \hat{C}_{\text{Hume},2011} + \frac{2}{5} \hat{C}_{\text{Hume},2016}
+    % = \frac{3}{5} \cdot 91.00 \% + \frac{2}{5} \cdot 89.65 \% \\
+    =  90.46\%.
+\]
 This is done for each of the 65 socio-demographic variables, and is repeated for each of the 149 remaining target zones corresponding with 2013 electorates.
@@
 
-Spatial weights are calculated in accordance with the assumption that an electorate is equally correlated with any electorate that shares a part of its boundary.
-Let \(\rho\) be spatial autoregressive coefficient, \(\boldsymbol v\) be a spherical error term, \({\boldsymbol W}\) be a matrix of spatial weights (containing information about the neighbouring regions), \(\boldsymbol X\) be a matrix of socio-demographic covariates, \(\boldsymbol \beta\) be a vector of regression coefficients and \(\boldsymbol a\) be a spatially structured random effect vector.
-
-\[{\boldsymbol y} = {\boldsymbol X} {\boldsymbol \beta} + {\boldsymbol a},\]
-and
-
-\[{\boldsymbol a} = \rho {\boldsymbol W} {\boldsymbol a} + {\boldsymbol v},\]
-where
-
-\[{\boldsymbol v} \sim N({\boldsymbol 0}, \sigma^2 {\boldsymbol I_n}),\].
-
-and hence
-
-\[{\boldsymbol y} = {\boldsymbol X} {\boldsymbol \beta} + ({\boldsymbol I}_n-\rho {\boldsymbol W})^{-1}{\boldsymbol v}.\]
-
-Estimation of the above spatial error model is undertaken using feasible generalized least squares.
-
-Table 3 details the estimated model coefficients and their estimated standard errors, for each of the six elections. An interpretation of these estimated values is provided in the next section.
-
-\begin{table}[!htbp] \centering
+\begin{table}[!p] \centering
   \caption{Estimated spatial error model parameters (standard errors) for each of the six election years.}
@@
 
-\newpage
+Spatial weights are calculated in accordance with the assumption that an electorate is equally correlated with any electorate that shares a part of its boundary.
+Let \(\rho\) be spatial autoregressive coefficient, \(\boldsymbol v\) be a spherical error term, \(\bm{W}\) be a matrix of spatial weights (containing information about the neighbouring regions), \(\boldsymbol X\) be a matrix of socio-demographic covariates, \(\boldsymbol \beta\) be a vector of regression coefficients and \(\boldsymbol a\) be a spatially structured random effect vector.
+\[
+  \bm{y} = \bm{X} \bm{\beta} + \bm{a},
+\]
+and
+\[
+  \bm{a} = \rho \bm{W} \bm{a} + \bm{v},
+\]
+where
+\[
+  \bm{v} \sim N(\bm{0}, \sigma^2 \bm{I_n}),
+\].
+and hence
+\[
+  \bm{y} = \bm{X} \bm{\beta} + (\bm{I}_n-\rho \bm{W})^{-1}\bm{v}.
+\]
+Estimation of the above spatial error model is undertaken using feasible generalized least squares.
+
+Table 3 details the estimated model coefficients and their estimated standard errors, for each of the six elections. An interpretation of these estimated values is provided in the next section.
 

### 2019-04-04 draft-paper.Rmd
Making stargazer table reproducible

@@
 hume %>%
-  select(-AusCitizen) %>%
   kable(longtable=FALSE, booktabs=TRUE, digits=4,
@@
           "Source zone population",
-          "Population allocated to purple region: $\\hat{P}_{s,t}$")) %>%
+          "Population allocated to purple region: $\\hat{P}_{s,t}$",
+          "AusCitizen (\\%): $D_s$")) %>%
   kableExtra::kable_styling(font_size=10, latex_options='scale_down')
-```
-
-Now consider the socio-demographic variable *AusCitizen*, the proportion of people in the region who are Australian citizens.
-
-```{r hume2, dependson='hume1'}
-hume %>%
-  select(DivisionNm, AusCitizen, Pst) %>%
-  kable(longtable=FALSE, booktabs=TRUE, digits=2,
-        format='latex',escape=FALSE,
-        col.names=c(
-          "Source zone (2016)",
-          "AusCitizen (\\%): $D_s$",
-          "Population allocated to purple region: $\\hat{P}_{s,t}$")) %>%
-  kableExtra::kable_styling(font_size=10)
 ```
@@
 
+Now consider the socio-demographic variable *AusCitizen*, the proportion of people in the region who are Australian citizens.
 A weighted average of *AusCitizen*, with the allocated population from each source zone as weights, yields $\hat{C}_{\text{Hume},2016} = `r sprintf("%4.2f",wave2016)`\%$.
@@
 
-Electoral socio-demographic variables are now available for each of the six elections and are joined with their corresponding two-party preferred votes.
-Before choosing an appropriate model, two issues with the socio-demographic variables are addressed.
-Firstly, variable scales change over the years, making it important to standardize variables.
-Secondly, many variables represent similar information and where appropriate, should be combined in some way.
+Following this process, electoral socio-demographic variables are available for each of the six elections and can be joined with their corresponding two-party preferred votes.
+Before choosing an appropriate model, two issues with the socio-demographic variables need to be addressed.
+First, variable scales change over the years, making it important to standardize variables.
+Second, many variables represent similar information and where appropriate, should be combined in some way.
 Principal component analysis is used to identify variables covary, from which intuitive groupings are selected to be combined into a single variable.
@@
 ## Standardizing variables
+
 Many of the socio-demographic variables have changing scales over the years.
@@
 ## Creating factors
+
 There are only $N = 150$ observations (electorates) in each election and $p = 65$ socio-demographic variables in each cross-section, with many variables represent similar information about an electorate.
@@
 ## Regression incorporating spatially dependent errors
+
 An identical model specification is used for each of the six elections, with each election modelled separately.
@@
 
-```{r printmodels, eval = FALSE}
+```{r printmodels, warning=FALSE, message=FALSE, results='asis'}
 # Add rho and GLS residual standard error
-stargazer::stargazer(glsmod01, glsmod04, glsmod07, glsmod10, glsmod13, glsmod16,
+regtable <- stargazer::stargazer(glsmod01, glsmod04, glsmod07, glsmod10, glsmod13, glsmod16,
   header = FALSE,
   column.sep.width = "1pt",
+  no.space=TRUE,
+  model.numbers=FALSE,
+  keep.stat = "n",
   title = "Estimated spatial error model parameters (standard errors) for each of the six election years.",
-  dep.var.labels = "Two-party preferred vote in favor of the Liberal party",
+  #dep.var.labels = "Two-party preferred vote in favor of the Liberal party",
   font.size = "scriptsize",
   digits = 2,
-  column.labels = c("2001", "2004", "2007", "2010", "2013", "2016"),
-  notes = c("Estimated coefficients for variable named in column one shown for", "election year indicated by column heading, with estimated standard", "deviation for each coefficient shown below in parenthesis.
+  column.labels = c("2001", "2004", "2007", "2010", "2013", "2016")
+  #notes = c("Estimated coefficients for variable named in column one shown for", "election year indicated by column heading, with estimated standard", "deviation for each coefficient shown below in parenthesis.
 Overall", "summary measures for each regression equation are provided in the", "bottom panel."),
-  notes.align = "l"
+  #notes.align = "l"
 )

### 2019-04-04 draft-paper.tex
Making stargazer table reproducible

@@
 \resizebox{\linewidth}{!}{
-\begin{tabular}{lrrr}
+\begin{tabular}{lrrrr}
 \toprule
-Source zone (2016) & Proportion & Source zone population & Population allocated to purple region: $\hat{P}_{s,t}$\\
+Source zone (2016) & Proportion & Source zone population & Population allocated to purple region: $\hat{P}_{s,t}$ & AusCitizen (\%): $D_s$\\
 \midrule
-Hume & 0.9654 & 150643 & 145427\\
-Riverina & 0.2511 & 155793 & 39117\\
-Eden-Monaro & 0.1109 & 147532 & 16358\\
-Canberra & 0.0028 & 196037 & 548\\
-Fenner & 0.0023 & 202955 & 474\\
+Hume & 0.9654 & 150643 & 145427 & 90.0168\\
+Riverina & 0.2511 & 155793 & 39117 & 89.1144\\
+Eden-Monaro & 0.1109 & 147532 & 16358 & 87.9999\\
+Canberra & 0.0028 & 196037 & 548 & 85.4793\\
+Fenner & 0.0023 & 202955 & 474 & 83.6432\\
 \addlinespace
-Whitlam & 0.0006 & 152280 & 92\\
-Gilmore & 0.0006 & 150436 & 86\\
-Calare & 0.0001 & 161298 & 21\\
+Whitlam & 0.0006 & 152280 & 92 & 89.5173\\
+Gilmore & 0.0006 & 150436 & 86 & 89.0266\\
+Calare & 0.0001 & 161298 & 21 & 87.5603\\
 \bottomrule
@@
 Now consider the socio-demographic variable \emph{AusCitizen}, the proportion of people in the region who are Australian citizens.
-
-\begin{table}[H]
-\centering\begingroup\fontsize{10}{12}\selectfont
-
-\begin{tabular}{lrr}
-\toprule
-Source zone (2016) & AusCitizen (\%): $D_s$ & Population allocated to purple region: $\hat{P}_{s,t}$\\
-\midrule
-Hume & 90.02 & 145427\\
-Riverina & 89.11 & 39117\\
-Eden-Monaro & 88.00 & 16358\\
-Canberra & 85.48 & 548\\
-Fenner & 83.64 & 474\\
-\addlinespace
-Whitlam & 89.52 & 92\\
-Gilmore & 89.03 & 86\\
-Calare & 87.56 & 21\\
-\bottomrule
-\end{tabular}
-\endgroup{}
-\end{table}
-
 A weighted average of \emph{AusCitizen}, with the allocated population from each source zone as weights, yields \(\hat{C}_{\text{Hume},2016} = 89.65\%\).
@@
 In this context, the random effect can be thought of as capturing the unobserved political climate in each electorate, where this climate is correlated with the climate in neighbouring electorates, under the assumption that the climate is independent of electoral socio-demographics.
+
+\begin{table}[!htbp] \centering 
+  \caption{Estimated spatial error model parameters (standard errors) for each of the six election years.} 
+  \label{} 
+\scriptsize 
+\begin{tabular}{@{\extracolsep{1pt}}lcccccc} 
+\\[-1.8ex]\hline 
+\hline \\[-1.8ex] 
+ & \multicolumn{6}{c}{\textit{Dependent variable:}} \\ 
+\cline{2-7} 
+\\[-1.8ex] & \multicolumn{6}{c}{LNP\_Percent} \\ 
+ & 2001 & 2004 & 2007 & 2010 & 2013 & 2016 \\ 
+\hline \\[-1.8ex] 
+ AusCitizen & $-$3.13 & $-$2.64 & $-$2.53 & $-$0.08 & $-$3.40 & $-$1.80 \\ 
+  & (2.26) & (2.43) & (2.34) & (2.79) & (2.76) & (2.71) \\ 
+  BornAsia & 2.22 & $-$0.95 & $-$1.60 & $-$6.83$^{**}$ & $-$3.03 & $-$0.55 \\ 
+  & (2.18) & (2.44) & (2.19) & (2.73) & (2.71) & (2.17) \\ 
+  Born\_MidEast & $-$1.15 & $-$1.59 & $-$2.01$^{*}$ & $-$2.03 & $-$0.92 & $-$1.44 \\ 
+  & (1.07) & (1.20) & (1.11) & (1.27) & (1.24) & (1.13) \\ 
+  BornSEEuro & $-$3.21$^{**}$ & $-$4.24$^{***}$ & $-$3.61$^{***}$ & $-$4.14$^{***}$ & $-$3.69$^{***}$ & $-$2.72$^{***}$ \\ 
+  & (1.42) & (1.46) & (1.02) & (1.19) & (1.07) & (0.97) \\ 
+  Born\_UK & 0.25 & $-$0.07 & 0.34 & 0.56 & $-$0.09 & $-$1.32 \\ 
+  & (1.00) & (0.98) & (0.90) & (1.07) & (1.04) & (1.04) \\ 
+  BornElsewhere & $-$5.04 & $-$4.91 & $-$4.13 & 2.35 & $-$5.23 & $-$4.14 \\ 
+  & (3.30) & (3.68) & (3.38) & (4.23) & (4.15) & (3.97) \\ 
+  Buddhism & $-$0.49 & $-$0.17 & $-$1.37 & $-$0.83 & $-$0.12 & $-$1.60 \\ 
+  & (1.39) & (1.61) & (1.61) & (1.80) & (1.68) & (1.56) \\ 
+  Christianity & $-$2.48 & $-$1.23 & 0.38 & 0.50 & 2.41 & 1.68 \\ 
+  & (1.73) & (1.85) & (1.83) & (1.99) & (1.85) & (1.78) \\ 
+  CurrentlyStudying & $-$2.19$^{**}$ & $-$0.13 & 2.06$^{*}$ & 2.12$^{*}$ & 1.15 & $-$0.16 \\ 
+  & (0.99) & (1.13) & (1.17) & (1.25) & (1.26) & (1.18) \\ 
+  DeFacto & $-$6.44$^{***}$ & $-$5.37$^{**}$ & $-$6.43$^{***}$ & $-$8.07$^{***}$ & $-$6.56$^{**}$ & $-$8.53$^{***}$ \\ 
+  & (1.87) & (2.48) & (2.31) & (3.06) & (3.11) & (2.83) \\ 
+  DiffAddress & 3.88$^{***}$ & 5.06$^{***}$ & 4.22$^{***}$ & 5.57$^{***}$ & 3.53$^{*}$ & 5.67$^{***}$ \\ 
+  & (0.94) & (1.12) & (0.99) & (1.76) & (1.91) & (1.60) \\ 
+  Distributive & 1.27 & 2.01$^{*}$ & 1.36 & 1.57 & 2.10$^{*}$ & 1.20 \\ 
+  & (1.12) & (1.21) & (1.13) & (1.34) & (1.27) & (1.21) \\ 
+  Education & 1.08 & 0.52 & $-$5.52$^{*}$ & $-$4.08 & $-$4.44 & $-$7.07$^{**}$ \\ 
+  & (2.38) & (3.12) & (3.27) & (3.95) & (3.78) & (3.55) \\ 
+  Extractive & 4.83$^{***}$ & 5.45$^{***}$ & 5.37$^{***}$ & 7.31$^{***}$ & 6.71$^{***}$ & 7.43$^{***}$ \\ 
+  & (1.48) & (1.42) & (1.36) & (1.56) & (1.47) & (1.39) \\ 
+  FamHouseSize & $-$0.16 & 0.87 & $-$2.40 & $-$2.53 & $-$3.26 & $-$2.91 \\ 
+  & (2.19) & (2.72) & (2.69) & (3.25) & (3.28) & (2.90) \\ 
+  Incomes & 4.36$^{**}$ & 5.03$^{*}$ & 9.45$^{***}$ & 7.09$^{**}$ & 7.97$^{***}$ & 12.20$^{***}$ \\ 
+  & (1.77) & (2.66) & (2.75) & (3.25) & (2.92) & (2.75) \\ 
+  Indigenous & 2.91$^{*}$ & 1.97 & 2.48 & 2.84 & 0.67 & $-$0.05 \\ 
+  & (1.68) & (1.95) & (1.75) & (2.16) & (2.14) & (2.00) \\ 
+  Islam & $-$0.92 & $-$0.97 & $-$0.54 & $-$2.50 & $-$0.82 & $-$0.95 \\ 
+  & (1.22) & (1.36) & (1.27) & (1.52) & (1.42) & (1.34) \\ 
+  Judaism & 1.88$^{*}$ & 1.78 & 2.66$^{***}$ & 1.97$^{*}$ & 2.74$^{**}$ & 1.65$^{*}$ \\ 
+  & (1.05) & (1.13) & (1.01) & (1.15) & (1.10) & (1.00) \\ 
+  ManagerAdmin & 2.06$^{***}$ & 3.32$^{***}$ & 6.00$^{***}$ & 5.47$^{***}$ & 5.04$^{***}$ & 5.78$^{***}$ \\ 
+  & (0.71) & (0.93) & (0.90) & (1.08) & (1.03) & (1.06) \\ 
+  Married & 0.44 & 0.11 & $-$1.22 & $-$0.22 & 0.91 & $-$2.34 \\ 
+  & (2.31) & (2.96) & (2.83) & (3.15) & (3.03) & (2.81) \\ 
+  MedianAge & 2.32$^{*}$ & 4.96$^{***}$ & 3.66$^{**}$ & 4.00$^{*}$ & 2.30 & 2.87 \\ 
+  & (1.32) & (1.65) & (1.81) & (2.26) & (2.08) & (1.79) \\ 
+  NoReligion & $-$1.57 & $-$0.92 & 0.56 & $-$0.30 & 1.02 & 1.31 \\ 
+  & (1.59) & (1.71) & (1.73) & (1.92) & (1.94) & (2.04) \\ 
+  OneParentHouse & $-$1.73 & $-$0.45 & $-$0.75 & $-$1.46 & $-$0.77 & $-$0.74 \\ 
+  & (1.36) & (1.59) & (1.49) & (1.69) & (1.57) & (1.47) \\ 
+  OtherLanguage & $-$0.44 & 5.92 & 9.98$^{**}$ & 11.24$^{**}$ & 9.00$^{*}$ & 9.84$^{**}$ \\ 
+  & (3.22) & (4.16) & (3.91) & (4.76) & (4.66) & (4.44) \\ 
+  PropertyOwned & $-$0.46 & $-$0.53 & 0.67 & $-$0.94 & $-$0.48 & 1.41 \\ 
+  & (1.37) & (1.50) & (1.43) & (1.76) & (1.67) & (1.50) \\ 
+  RentLoanPrice & $-$1.57 & $-$3.32$^{*}$ & $-$4.01$^{**}$ & $-$0.97 & $-$0.70 & $-$0.89 \\ 
+  & (1.49) & (1.76) & (1.67) & (2.07) & (2.07) & (2.16) \\ 
+  SocialServ & 2.51$^{*}$ & 1.65 & 2.47$^{*}$ & 2.53$^{*}$ & 2.35$^{*}$ & 4.45$^{***}$ \\ 
+  & (1.33) & (1.41) & (1.29) & (1.47) & (1.32) & (1.19) \\ 
+  Transformative & 3.24$^{**}$ & 4.73$^{***}$ & 4.84$^{***}$ & 4.46$^{**}$ & 3.56$^{**}$ & 4.58$^{***}$ \\ 
+  & (1.55) & (1.78) & (1.74) & (1.98) & (1.78) & (1.53) \\ 
+  Unemployment & $-$2.45$^{*}$ & $-$3.07$^{*}$ & 0.29 & 0.08 & 1.67 & 2.79$^{**}$ \\ 
+  & (1.40) & (1.63) & (1.51) & (1.76) & (1.51) & (1.37) \\ 
+  Constant & 50.81$^{***}$ & 52.60$^{***}$ & 47.31$^{***}$ & 49.93$^{***}$ & 53.51$^{***}$ & 50.49$^{***}$ \\ 
+  & (0.71) & (0.58) & (0.52) & (0.57) & (0.58) & (0.80) \\ 
+ \hline \\[-1.8ex] 
+Observations & 150 & 150 & 150 & 150 & 150 & 150 \\ 
+\hline 
+\hline \\[-1.8ex] 
+\textit{Note:}  & \multicolumn{6}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
+\end{tabular} 
+\end{table}
 

### 2019-04-10 draft-paper.Rmd
Table 1 now reproducible

@@
 
-```{r printmodels, warning=FALSE, message=FALSE, results='asis'}
+```{r printmodels, warning=FALSE, message=FALSE, include=FALSE}
 # Add rho and GLS residual standard error
@@
   model.numbers=FALSE,
-  keep.stat = "n",
+  keep.stat = c("n","ser"),
+  dep.var.caption="",
+  dep.var.labels="\\vspace{-0.5cm}",
   title = "Estimated spatial error model parameters (standard errors) for each of the six election years.",
@@
 )
-```
-
-\begin{table}[!p] \centering
-  \caption{Estimated spatial error model parameters (standard errors) for each of the six election years.}
-  \label{}
-\scriptsize
-\begin{tabular}{@{\extracolsep{1pt}}lcccccc}
-\\[-1.8ex]\hline
- & 2001 & 2004 & 2007 & 2010 & 2013 & 2016 \\
-\hline \\[-1.8ex]
- $\rho$ & 0.46$^{***}$ & 0.29$^{*}$ & 0.24 & 0.19 & 0.27$^{*}$ & 0.50$^{***}$ \\
-  & (0.15) & (0.17) & (0.17) & (0.16) & (0.16) & (0.17) \\
-
- AusCitizen & $-$3.13 & $-$2.64 & $-$2.53 & $-$0.08 & $-$3.40 & $-$1.80 \\
-  & (2.26) & (2.43) & (2.34) & (2.79) & (2.76) & (2.71) \\
-
- BornAsia & 2.22 & $-$0.95 & $-$1.60 & $-$6.83$^{**}$ & $-$3.03 & $-$0.55 \\
-  & (2.18) & (2.44) & (2.19) & (2.73) & (2.71) & (2.17) \\
-
- BornMidEast & $-$1.15 & $-$1.59 & $-$2.01$^{*}$ & $-$2.03 & $-$0.92 & $-$1.44 \\
-  & (1.07) & (1.20) & (1.11) & (1.27) & (1.24) & (1.13) \\
-
- BornSEEuro & $-$3.21$^{**}$ & $-$4.24$^{***}$ & $-$3.61$^{***}$ & $-$4.14$^{***}$ & $-$3.69$^{***}$ & $-$2.72$^{***}$ \\
-  & (1.42) & (1.46) & (1.02) & (1.19) & (1.07) & (0.97) \\
-
- BornUK & 0.25 & $-$0.07 & 0.34 & 0.56 & $-$0.09 & $-$1.32 \\
-  & (1.00) & (0.98) & (0.90) & (1.07) & (1.04) & (1.04) \\
-
- BornElsewhere & $-$5.04 & $-$4.91 & $-$4.13 & 2.35 & $-$5.23 & $-$4.14 \\
-  & (3.30) & (3.68) & (3.38) & (4.23) & (4.15) & (3.97) \\
-
- Buddhism & $-$0.49 & $-$0.17 & $-$1.37 & $-$0.83 & $-$0.12 & $-$1.60 \\
-  & (1.39) & (1.61) & (1.61) & (1.80) & (1.68) & (1.56) \\
-
- Christianity & $-$2.48 & $-$1.23 & 0.38 & 0.50 & 2.41 & 1.68 \\
-  & (1.73) & (1.85) & (1.83) & (1.99) & (1.85) & (1.78) \\
-
- CurrentlyStudying & $-$2.19$^{**}$ & $-$0.13 & 2.06$^{*}$ & 2.12$^{*}$ & 1.15 & $-$0.16 \\
-  & (0.99) & (1.13) & (1.17) & (1.25) & (1.26) & (1.18) \\
-
- DeFacto & $-$6.44$^{***}$ & $-$5.37$^{**}$ & $-$6.43$^{***}$ & $-$8.07$^{***}$ & $-$6.56$^{**}$ & $-$8.53$^{***}$ \\
-  & (1.87) & (2.48) & (2.31) & (3.06) & (3.11) & (2.83) \\
-
- DiffAddress & 3.88$^{***}$ & 5.06$^{***}$ & 4.22$^{***}$ & 5.57$^{***}$ & 3.53$^{*}$ & 5.67$^{***}$ \\
-  & (0.94) & (1.12) & (0.99) & (1.76) & (1.91) & (1.60) \\
-
- Distributive & 1.27 & 2.01$^{*}$ & 1.36 & 1.57 & 2.10$^{*}$ & 1.20 \\
-  & (1.12) & (1.21) & (1.13) & (1.34) & (1.27) & (1.21) \\
-
- Education & 1.08 & 0.52 & $-$5.52$^{*}$ & $-$4.08 & $-$4.44 & $-$7.07$^{**}$ \\
-  & (2.38) & (3.12) & (3.27) & (3.95) & (3.78) & (3.55) \\
-
- Extractive & 4.83$^{***}$ & 5.45$^{***}$ & 5.37$^{***}$ & 7.31$^{***}$ & 6.71$^{***}$ & 7.43$^{***}$ \\
-  & (1.48) & (1.42) & (1.36) & (1.56) & (1.47) & (1.39) \\
-
- FamHouseSize & $-$0.16 & 0.87 & $-$2.40 & $-$2.53 & $-$3.26 & $-$2.91 \\
-  & (2.19) & (2.72) & (2.69) & (3.25) & (3.28) & (2.90) \\
-
- Incomes & 4.36$^{**}$ & 5.03$^{*}$ & 9.45$^{***}$ & 7.09$^{**}$ & 7.97$^{***}$ & 12.20$^{***}$ \\
-  & (1.77) & (2.66) & (2.75) & (3.25) & (2.92) & (2.75) \\
-
- Indigenous & 2.91$^{*}$ & 1.97 & 2.48 & 2.84 & 0.67 & $-$0.05 \\
-  & (1.68) & (1.95) & (1.75) & (2.16) & (2.14) & (2.00) \\
-
- Islam & $-$0.92 & $-$0.97 & $-$0.54 & $-$2.50 & $-$0.82 & $-$0.95 \\
-  & (1.22) & (1.36) & (1.27) & (1.52) & (1.42) & (1.34) \\
-
- Judaism & 1.88$^{*}$ & 1.78 & 2.66$^{***}$ & 1.97$^{*}$ & 2.74$^{**}$ & 1.65$^{*}$ \\
-  & (1.05) & (1.13) & (1.01) & (1.15) & (1.10) & (1.00) \\
-
- ManagerAdmin & 2.06$^{***}$ & 3.32$^{***}$ & 6.00$^{***}$ & 5.47$^{***}$ & 5.04$^{***}$ & 5.78$^{***}$ \\
-  & (0.71) & (0.93) & (0.90) & (1.08) & (1.03) & (1.06) \\
-
- Married & 0.44 & 0.11 & $-$1.22 & $-$0.22 & 0.91 & $-$2.34 \\
-  & (2.31) & (2.96) & (2.83) & (3.15) & (3.03) & (2.81) \\
-
- MedianAge & 2.32$^{*}$ & 4.96$^{***}$ & 3.66$^{**}$ & 4.00$^{*}$ & 2.30 & 2.87 \\
-  & (1.32) & (1.65) & (1.81) & (2.26) & (2.08) & (1.79) \\
-
- NoReligion & $-$1.57 & $-$0.92 & 0.56 & $-$0.30 & 1.02 & 1.31 \\
-  & (1.59) & (1.71) & (1.73) & (1.92) & (1.94) & (2.04) \\
-
- OneParentHouse & $-$1.73 & $-$0.45 & $-$0.75 & $-$1.46 & $-$0.77 & $-$0.74 \\
-  & (1.36) & (1.59) & (1.49) & (1.69) & (1.57) & (1.47) \\
-
- OtherLanguage & $-$0.44 & 5.92 & 9.98$^{**}$ & 11.24$^{**}$ & 9.00$^{*}$ & 9.84$^{**}$ \\
-  & (3.22) & (4.16) & (3.91) & (4.76) & (4.66) & (4.44) \\
-
- PropertyOwned & $-$0.46 & $-$0.53 & 0.67 & $-$0.94 & $-$0.48 & 1.41 \\
-  & (1.37) & (1.50) & (1.43) & (1.76) & (1.67) & (1.50) \\
-
- RentLoanPrice & $-$1.57 & $-$3.32$^{*}$ & $-$4.01$^{**}$ & $-$0.97 & $-$0.70 & $-$0.89 \\
-  & (1.49) & (1.76) & (1.67) & (2.07) & (2.07) & (2.16) \\
-
- SocialServ & 2.51$^{*}$ & 1.65 & 2.47$^{*}$ & 2.53$^{*}$ & 2.35$^{*}$ & 4.45$^{***}$ \\
-  & (1.33) & (1.41) & (1.29) & (1.47) & (1.32) & (1.19) \\
-
- Transformative & 3.24$^{**}$ & 4.73$^{***}$ & 4.84$^{***}$ & 4.46$^{**}$ & 3.56$^{**}$ & 4.58$^{***}$ \\
-  & (1.55) & (1.78) & (1.74) & (1.98) & (1.78) & (1.53) \\
-
- Unemployment & $-$2.45$^{*}$ & $-$3.07$^{*}$ & 0.29 & 0.08 & 1.67 & 2.79$^{**}$ \\
-  & (1.40) & (1.63) & (1.51) & (1.76) & (1.51) & (1.37) \\
-
- Constant & 50.81$^{***}$ & 52.60$^{***}$ & 47.31$^{***}$ & 49.93$^{***}$ & 53.51$^{***}$ & 50.49$^{***}$ \\
-  & (0.71) & (0.58) & (0.52) & (0.57) & (0.58) & (0.80) \\
-
-\hline \\[-1.8ex]
-Observations & 150 & 150 & 150 & 150 & 150 & 150 \\
-Residual Standard Error (GLS) & 4.69 & 5.04 & 4.79 & 5.63 & 5.18 & 4.88 \\
-\hline
-\hline \\[-1.8ex]
-\textit{Note:}  & \multicolumn{6}{l}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\
-\end{tabular}
-\end{table}
+# Add correlations
+getrho <- function(x, element) {x$rho_df[[element]]}
+rho <- map_dbl(list(glsmod01,glsmod04,glsmod07,glsmod10,glsmod13,glsmod16),
+           getrho, element="estimate")
+rhop <- map_dbl(list(glsmod01,glsmod04,glsmod07,glsmod10,glsmod13,glsmod16),
+           getrho, element="p")
+rhose <- map_dbl(list(glsmod01,glsmod04,glsmod07,glsmod10,glsmod13,glsmod16),
+           getrho, element="se")
+rholine1 <- paste("$\\rho$",
+            paste(" & ",
+                  sprintf("%3.2f",rho), 
+                  ifelse(rhop < 0.01, "$^{***}$", 
+                  ifelse(rhop < 0.05, "$^{**}$",
+                  ifelse(rhop < 0.10, "$^{*}$",""))), 
+              sep="", collapse=""),
+            "\\\\", sep="", collapse="")
+rholine2 <- paste("",
+            paste(" & (", 
+                  sprintf("%3.2f",rhose), ")", 
+              sep="", collapse=""),
+            "\\\\", sep="", collapse="")
+regtable <- c(regtable[1:11], rholine1, rholine2, regtable[12:length(regtable)])
+# Add residual standard error (GLS)
+getsigma <- function(x, element) {x$sigma}
+sigma <- map_dbl(list(glsmod01,glsmod04,glsmod07,glsmod10,glsmod13,glsmod16),
+           getsigma)
+sigmaline <- paste("Residual Standard Error (GLS)",
+             paste(" & ",
+                   sprintf("%4.2f", sigma),
+                   sep="", collapse=""),
+             "\\\\", sep="", collapse="")
+regtable <- c(regtable[1:77],sigmaline, regtable[78:length(regtable)])
+```
+
+```{r printmodels2, results='asis', dependson='printmodels'}
+cat(regtable)
+```
 

### 2019-04-10 draft-paper.tex
Table 1 now reproducible

--- 2019-04-03 draft-paper.tex
@@
 
-\begin{table}[!p] \centering
-  \caption{Estimated spatial error model parameters (standard errors) for each of the six election years.}
-  \label{}
-\scriptsize
-\begin{tabular}{@{\extracolsep{1pt}}lcccccc}
-\\[-1.8ex]\hline
- & 2001 & 2004 & 2007 & 2010 & 2013 & 2016 \\
-\hline \\[-1.8ex]
- $\rho$ & 0.46$^{***}$ & 0.29$^{*}$ & 0.24 & 0.19 & 0.27$^{*}$ & 0.50$^{***}$ \\
-  & (0.15) & (0.17) & (0.17) & (0.16) & (0.16) & (0.17) \\
-
- AusCitizen & $-$3.13 & $-$2.64 & $-$2.53 & $-$0.08 & $-$3.40 & $-$1.80 \\
-  & (2.26) & (2.43) & (2.34) & (2.79) & (2.76) & (2.71) \\
-
- BornAsia & 2.22 & $-$0.95 & $-$1.60 & $-$6.83$^{**}$ & $-$3.03 & $-$0.55 \\
-  & (2.18) & (2.44) & (2.19) & (2.73) & (2.71) & (2.17) \\
-
- BornMidEast & $-$1.15 & $-$1.59 & $-$2.01$^{*}$ & $-$2.03 & $-$0.92 & $-$1.44 \\
-  & (1.07) & (1.20) & (1.11) & (1.27) & (1.24) & (1.13) \\
-
- BornSEEuro & $-$3.21$^{**}$ & $-$4.24$^{***}$ & $-$3.61$^{***}$ & $-$4.14$^{***}$ & $-$3.69$^{***}$ & $-$2.72$^{***}$ \\
-  & (1.42) & (1.46) & (1.02) & (1.19) & (1.07) & (0.97) \\
-
- BornUK & 0.25 & $-$0.07 & 0.34 & 0.56 & $-$0.09 & $-$1.32 \\
-  & (1.00) & (0.98) & (0.90) & (1.07) & (1.04) & (1.04) \\
-
- BornElsewhere & $-$5.04 & $-$4.91 & $-$4.13 & 2.35 & $-$5.23 & $-$4.14 \\
-  & (3.30) & (3.68) & (3.38) & (4.23) & (4.15) & (3.97) \\
-
- Buddhism & $-$0.49 & $-$0.17 & $-$1.37 & $-$0.83 & $-$0.12 & $-$1.60 \\
-  & (1.39) & (1.61) & (1.61) & (1.80) & (1.68) & (1.56) \\
-
- Christianity & $-$2.48 & $-$1.23 & 0.38 & 0.50 & 2.41 & 1.68 \\
-  & (1.73) & (1.85) & (1.83) & (1.99) & (1.85) & (1.78) \\
-
- CurrentlyStudying & $-$2.19$^{**}$ & $-$0.13 & 2.06$^{*}$ & 2.12$^{*}$ & 1.15 & $-$0.16 \\
-  & (0.99) & (1.13) & (1.17) & (1.25) & (1.26) & (1.18) \\
-
- DeFacto & $-$6.44$^{***}$ & $-$5.37$^{**}$ & $-$6.43$^{***}$ & $-$8.07$^{***}$ & $-$6.56$^{**}$ & $-$8.53$^{***}$ \\
-  & (1.87) & (2.48) & (2.31) & (3.06) & (3.11) & (2.83) \\
-
- DiffAddress & 3.88$^{***}$ & 5.06$^{***}$ & 4.22$^{***}$ & 5.57$^{***}$ & 3.53$^{*}$ & 5.67$^{***}$ \\
-  & (0.94) & (1.12) & (0.99) & (1.76) & (1.91) & (1.60) \\
-
- Distributive & 1.27 & 2.01$^{*}$ & 1.36 & 1.57 & 2.10$^{*}$ & 1.20 \\
-  & (1.12) & (1.21) & (1.13) & (1.34) & (1.27) & (1.21) \\
-
- Education & 1.08 & 0.52 & $-$5.52$^{*}$ & $-$4.08 & $-$4.44 & $-$7.07$^{**}$ \\
-  & (2.38) & (3.12) & (3.27) & (3.95) & (3.78) & (3.55) \\
-
- Extractive & 4.83$^{***}$ & 5.45$^{***}$ & 5.37$^{***}$ & 7.31$^{***}$ & 6.71$^{***}$ & 7.43$^{***}$ \\
-  & (1.48) & (1.42) & (1.36) & (1.56) & (1.47) & (1.39) \\
-
- FamHouseSize & $-$0.16 & 0.87 & $-$2.40 & $-$2.53 & $-$3.26 & $-$2.91 \\
-  & (2.19) & (2.72) & (2.69) & (3.25) & (3.28) & (2.90) \\
-
- Incomes & 4.36$^{**}$ & 5.03$^{*}$ & 9.45$^{***}$ & 7.09$^{**}$ & 7.97$^{***}$ & 12.20$^{***}$ \\
-  & (1.77) & (2.66) & (2.75) & (3.25) & (2.92) & (2.75) \\
-
- Indigenous & 2.91$^{*}$ & 1.97 & 2.48 & 2.84 & 0.67 & $-$0.05 \\
-  & (1.68) & (1.95) & (1.75) & (2.16) & (2.14) & (2.00) \\
-
- Islam & $-$0.92 & $-$0.97 & $-$0.54 & $-$2.50 & $-$0.82 & $-$0.95 \\
-  & (1.22) & (1.36) & (1.27) & (1.52) & (1.42) & (1.34) \\
-
- Judaism & 1.88$^{*}$ & 1.78 & 2.66$^{***}$ & 1.97$^{*}$ & 2.74$^{**}$ & 1.65$^{*}$ \\
-  & (1.05) & (1.13) & (1.01) & (1.15) & (1.10) & (1.00) \\
-
- ManagerAdmin & 2.06$^{***}$ & 3.32$^{***}$ & 6.00$^{***}$ & 5.47$^{***}$ & 5.04$^{***}$ & 5.78$^{***}$ \\
-  & (0.71) & (0.93) & (0.90) & (1.08) & (1.03) & (1.06) \\
-
- Married & 0.44 & 0.11 & $-$1.22 & $-$0.22 & 0.91 & $-$2.34 \\
-  & (2.31) & (2.96) & (2.83) & (3.15) & (3.03) & (2.81) \\
-
- MedianAge & 2.32$^{*}$ & 4.96$^{***}$ & 3.66$^{**}$ & 4.00$^{*}$ & 2.30 & 2.87 \\
-  & (1.32) & (1.65) & (1.81) & (2.26) & (2.08) & (1.79) \\
-
- NoReligion & $-$1.57 & $-$0.92 & 0.56 & $-$0.30 & 1.02 & 1.31 \\
-  & (1.59) & (1.71) & (1.73) & (1.92) & (1.94) & (2.04) \\
-
- OneParentHouse & $-$1.73 & $-$0.45 & $-$0.75 & $-$1.46 & $-$0.77 & $-$0.74 \\
-  & (1.36) & (1.59) & (1.49) & (1.69) & (1.57) & (1.47) \\
-
- OtherLanguage & $-$0.44 & 5.92 & 9.98$^{**}$ & 11.24$^{**}$ & 9.00$^{*}$ & 9.84$^{**}$ \\
-  & (3.22) & (4.16) & (3.91) & (4.76) & (4.66) & (4.44) \\
-
- PropertyOwned & $-$0.46 & $-$0.53 & 0.67 & $-$0.94 & $-$0.48 & 1.41 \\
-  & (1.37) & (1.50) & (1.43) & (1.76) & (1.67) & (1.50) \\
-
- RentLoanPrice & $-$1.57 & $-$3.32$^{*}$ & $-$4.01$^{**}$ & $-$0.97 & $-$0.70 & $-$0.89 \\
-  & (1.49) & (1.76) & (1.67) & (2.07) & (2.07) & (2.16) \\
-
- SocialServ & 2.51$^{*}$ & 1.65 & 2.47$^{*}$ & 2.53$^{*}$ & 2.35$^{*}$ & 4.45$^{***}$ \\
-  & (1.33) & (1.41) & (1.29) & (1.47) & (1.32) & (1.19) \\
-
- Transformative & 3.24$^{**}$ & 4.73$^{***}$ & 4.84$^{***}$ & 4.46$^{**}$ & 3.56$^{**}$ & 4.58$^{***}$ \\
-  & (1.55) & (1.78) & (1.74) & (1.98) & (1.78) & (1.53) \\
-
- Unemployment & $-$2.45$^{*}$ & $-$3.07$^{*}$ & 0.29 & 0.08 & 1.67 & 2.79$^{**}$ \\
-  & (1.40) & (1.63) & (1.51) & (1.76) & (1.51) & (1.37) \\
-
- Constant & 50.81$^{***}$ & 52.60$^{***}$ & 47.31$^{***}$ & 49.93$^{***}$ & 53.51$^{***}$ & 50.49$^{***}$ \\
-  & (0.71) & (0.58) & (0.52) & (0.57) & (0.58) & (0.80) \\
-
-\hline \\[-1.8ex]
-Observations & 150 & 150 & 150 & 150 & 150 & 150 \\
-Residual Standard Error (GLS) & 4.69 & 5.04 & 4.79 & 5.63 & 5.18 & 4.88 \\
-\hline
-\hline \\[-1.8ex]
-\textit{Note:}  & \multicolumn{6}{l}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\
-\end{tabular}
-\end{table}
+\begin{table}[!htbp] \centering    \caption{Estimated spatial error model parameters (standard errors) for each of the six election years.}    \label{}  \scriptsize  \begin{tabular}{@{\extracolsep{1pt}}lcccccc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]  \\[-1.8ex] & \multicolumn{6}{c}{\vspace{-0.5cm}} \\   & 2001 & 2004 & 2007 & 2010 & 2013 & 2016 \\  \hline \\[-1.8ex]  $\rho$ & 0.46$^{***}$ & 0.29$^{*}$ & 0.24 & 0.19 & 0.27$^{*}$ & 0.50$^{***}$\\  & (0.15) & (0.17) & (0.17) & (0.16) & (0.16) & (0.17)\\  AusCitizen & $-$3.13 & $-$2.64 & $-$2.53 & $-$0.08 & $-$3.40 & $-$1.80 \\    & (2.26) & (2.43) & (2.34) & (2.79) & (2.76) & (2.71) \\    BornAsia & 2.22 & $-$0.95 & $-$1.60 & $-$6.83$^{**}$ & $-$3.03 & $-$0.55 \\    & (2.18) & (2.44) & (2.19) & (2.73) & (2.71) & (2.17) \\    Born\_MidEast & $-$1.15 & $-$1.59 & $-$2.01$^{*}$ & $-$2.03 & $-$0.92 & $-$1.44 \\    & (1.07) & (1.20) & (1.11) & (1.27) & (1.24) & (1.13) \\    BornSEEuro & $-$3.21$^{**}$ & $-$4.24$^{***}$ & $-$3.61$^{***}$ & $-$4.14$^{***}$ & $-$3.69$^{***}$ & $-$2.72$^{***}$ \\    & (1.42) & (1.46) & (1.02) & (1.19) & (1.07) & (0.97) \\    Born\_UK & 0.25 & $-$0.07 & 0.34 & 0.56 & $-$0.09 & $-$1.32 \\    & (1.00) & (0.98) & (0.90) & (1.07) & (1.04) & (1.04) \\    BornElsewhere & $-$5.04 & $-$4.91 & $-$4.13 & 2.35 & $-$5.23 & $-$4.14 \\    & (3.30) & (3.68) & (3.38) & (4.23) & (4.15) & (3.97) \\    Buddhism & $-$0.49 & $-$0.17 & $-$1.37 & $-$0.83 & $-$0.12 & $-$1.60 \\    & (1.39) & (1.61) & (1.61) & (1.80) & (1.68) & (1.56) \\    Christianity & $-$2.48 & $-$1.23 & 0.38 & 0.50 & 2.41 & 1.68 \\    & (1.73) & (1.85) & (1.83) & (1.99) & (1.85) & (1.78) \\    CurrentlyStudying & $-$2.19$^{**}$ & $-$0.13 & 2.06$^{*}$ & 2.12$^{*}$ & 1.15 & $-$0.16 \\    & (0.99) & (1.13) & (1.17) & (1.25) & (1.26) & (1.18) \\    DeFacto & $-$6.44$^{***}$ & $-$5.37$^{**}$ & $-$6.43$^{***}$ & $-$8.07$^{***}$ & $-$6.56$^{**}$ & $-$8.53$^{***}$ \\    & (1.87) & (2.48) & (2.31) & (3.06) & (3.11) & (2.83) \\    DiffAddress & 3.88$^{***}$ & 5.06$^{***}$ & 4.22$^{***}$ & 5.57$^{***}$ & 3.53$^{*}$ & 5.67$^{***}$ \\    & (0.94) & (1.12) & (0.99) & (1.76) & (1.91) & (1.60) \\    Distributive & 1.27 & 2.01$^{*}$ & 1.36 & 1.57 & 2.10$^{*}$ & 1.20 \\    & (1.12) & (1.21) & (1.13) & (1.34) & (1.27) & (1.21) \\    Education & 1.08 & 0.52 & $-$5.52$^{*}$ & $-$4.08 & $-$4.44 & $-$7.07$^{**}$ \\    & (2.38) & (3.12) & (3.27) & (3.95) & (3.78) & (3.55) \\    Extractive & 4.83$^{***}$ & 5.45$^{***}$ & 5.37$^{***}$ & 7.31$^{***}$ & 6.71$^{***}$ & 7.43$^{***}$ \\    & (1.48) & (1.42) & (1.36) & (1.56) & (1.47) & (1.39) \\    FamHouseSize & $-$0.16 & 0.87 & $-$2.40 & $-$2.53 & $-$3.26 & $-$2.91 \\    & (2.19) & (2.72) & (2.69) & (3.25) & (3.28) & (2.90) \\    Incomes & 4.36$^{**}$ & 5.03$^{*}$ & 9.45$^{***}$ & 7.09$^{**}$ & 7.97$^{***}$ & 12.20$^{***}$ \\    & (1.77) & (2.66) & (2.75) & (3.25) & (2.92) & (2.75) \\    Indigenous & 2.91$^{*}$ & 1.97 & 2.48 & 2.84 & 0.67 & $-$0.05 \\    & (1.68) & (1.95) & (1.75) & (2.16) & (2.14) & (2.00) \\    Islam & $-$0.92 & $-$0.97 & $-$0.54 & $-$2.50 & $-$0.82 & $-$0.95 \\    & (1.22) & (1.36) & (1.27) & (1.52) & (1.42) & (1.34) \\    Judaism & 1.88$^{*}$ & 1.78 & 2.66$^{***}$ & 1.97$^{*}$ & 2.74$^{**}$ & 1.65$^{*}$ \\    & (1.05) & (1.13) & (1.01) & (1.15) & (1.10) & (1.00) \\    ManagerAdmin & 2.06$^{***}$ & 3.32$^{***}$ & 6.00$^{***}$ & 5.47$^{***}$ & 5.04$^{***}$ & 5.78$^{***}$ \\    & (0.71) & (0.93) & (0.90) & (1.08) & (1.03) & (1.06) \\    Married & 0.44 & 0.11 & $-$1.22 & $-$0.22 & 0.91 & $-$2.34 \\    & (2.31) & (2.96) & (2.83) & (3.15) & (3.03) & (2.81) \\    MedianAge & 2.32$^{*}$ & 4.96$^{***}$ & 3.66$^{**}$ & 4.00$^{*}$ & 2.30 & 2.87 \\    & (1.32) & (1.65) & (1.81) & (2.26) & (2.08) & (1.79) \\    NoReligion & $-$1.57 & $-$0.92 & 0.56 & $-$0.30 & 1.02 & 1.31 \\    & (1.59) & (1.71) & (1.73) & (1.92) & (1.94) & (2.04) \\    OneParentHouse & $-$1.73 & $-$0.45 & $-$0.75 & $-$1.46 & $-$0.77 & $-$0.74 \\    & (1.36) & (1.59) & (1.49) & (1.69) & (1.57) & (1.47) \\    OtherLanguage & $-$0.44 & 5.92 & 9.98$^{**}$ & 11.24$^{**}$ & 9.00$^{*}$ & 9.84$^{**}$ \\    & (3.22) & (4.16) & (3.91) & (4.76) & (4.66) & (4.44) \\    PropertyOwned & $-$0.46 & $-$0.53 & 0.67 & $-$0.94 & $-$0.48 & 1.41 \\    & (1.37) & (1.50) & (1.43) & (1.76) & (1.67) & (1.50) \\    RentLoanPrice & $-$1.57 & $-$3.32$^{*}$ & $-$4.01$^{**}$ & $-$0.97 & $-$0.70 & $-$0.89 \\    & (1.49) & (1.76) & (1.67) & (2.07) & (2.07) & (2.16) \\    SocialServ & 2.51$^{*}$ & 1.65 & 2.47$^{*}$ & 2.53$^{*}$ & 2.35$^{*}$ & 4.45$^{***}$ \\    & (1.33) & (1.41) & (1.29) & (1.47) & (1.32) & (1.19) \\    Transformative & 3.24$^{**}$ & 4.73$^{***}$ & 4.84$^{***}$ & 4.46$^{**}$ & 3.56$^{**}$ & 4.58$^{***}$ \\    & (1.55) & (1.78) & (1.74) & (1.98) & (1.78) & (1.53) \\    Unemployment & $-$2.45$^{*}$ & $-$3.07$^{*}$ & 0.29 & 0.08 & 1.67 & 2.79$^{**}$ \\    & (1.40) & (1.63) & (1.51) & (1.76) & (1.51) & (1.37) \\    Constant & 50.81$^{***}$ & 52.60$^{***}$ & 47.31$^{***}$ & 49.93$^{***}$ & 53.51$^{***}$ & 50.49$^{***}$ \\    & (0.71) & (0.58) & (0.52) & (0.57) & (0.58) & (0.80) \\   \hline \\[-1.8ex]  Observations & 150 & 150 & 150 & 150 & 150 & 150 \\  Residual Standard Error (GLS) & 4.69 & 5.04 & 4.79 & 5.63 & 5.18 & 4.88\\ \hline  \hline \\[-1.8ex]  \textit{Note:}  & \multicolumn{6}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\  \end{tabular}  \end{table}
 

### 2019-04-10 draft-paper.Rmd
Nearly ready to submit

@@
 ```{r wrangle, include = FALSE}
-# Take log of indigneous, judaism, islam, buddhism
+# Take log of indigenous, judaism, islam, buddhism
 abs2016 <- eechidna::abs2016 %>%
@@
 
-To account for spatial differences, the piece-wise approximation method in @Goodchild1993 is adopted.
+To account for spatial differences, the piecewise approximation method in @Goodchild1993 is adopted.
 Consider a map of source zones $s = 1,\dots,S$, for which socio-demographic information is available, and a set of target zones $t = 1,\dots,T$ for which information is to be imputed.
@@
 
-```{r hume13, fig.cap = "Some of the electoral boundaries in NSW for 2013, with the electoral boundary for Hume shown in purple.", out.height="30%"}
+```{r hume13, fig.cap = "Some of the electoral boundaries in NSW for 2013, with the electoral boundary for Hume shown in purple.", out.height="28%"}
 hume_area13 <- nat_map13 %>%
@@
 ```{r hume16, fig.cap = "Census division boundaries in NSW for 2016, with the 2013 electoral boundary for Hume, shown in purple.
-The purple region is not contained within a single Census division.", out.height="30%"}
+The purple region is not contained within a single Census division.", out.height="28%"}
 hume_area16 <- nat_map16 %>%
@@
 
+[^2]: A preliminary step involved removing all age bands, because age is represented by median age, and to remove variables relating to particular denominations of Christianity.
+
 A factor is created from a group of variables if there is an intuitive reason as to why they should represent similar information and if there is evidence to suggest that they covary.
-For example, a potential group would be variables relating to electoral incomes - median family, household and personal incomes.
+For example, a potential group would be variables relating to electoral incomes --- median family, household and personal incomes.
 To determine which variables covary, principal component analysis is used on a combined dataset of socio-demographic variables from all six elections.
@@
 Six factors are created using this criteria.
-These are; `Incomes` - median personal, household and family incomes, `Unemployment` - unemployment and labor force participation rates, `ProportyOwned` - rates of housing ownership, mortgages, renting and government housing, `RentLoanPrice` - median rental and loan repayments, `FamHouseSize` - average household size, ratio of people to families and household makeup (single person, couple with kids and couple without kids) and `Education` - high school and university qualifications, jobs requiring higher levels of education as well as vocational course completions and jobs that do not require higher education levels (e.g.
-laborer, tradesperson).
+These are:
+`Incomes` (median personal, household and family incomes);
+`Unemployment` (unemployment and labor force participation rates);
+`ProportyOwned` (rates of housing ownership, mortgages, renting and government housing);
+`RentLoanPrice` (median rental and loan repayments);
+`FamHouseSize` (average household size, ratio of people to families and household makeup (single person, couple with kids and couple without kids); and
+`Education` (high school and university qualifications, jobs requiring higher levels of education as well as vocational course completions and jobs that do not require higher education levels, such as laborer or tradesperson).
 For each of these groups, variables with positive loadings are added and those with negative loadings are subtracted to create a factor.
@@
 
-[^2]: A preliminary step involved removing all age bands, because age is represented by median age, and to remove variables relating to particular denominations of Christianity.
-
 ## Regression incorporating spatially dependent errors
@@
 An identical model specification is used for each of the six elections, with each election modelled separately.
-This allows for the socio-demographic effects to be estimated separately for each election year, facilitating analysis of temporal changes in variable effects.
+This allows the socio-demographic effects to be estimated separately for each election year, facilitating analysis of temporal changes in variable effects.
 This approach is preferable to using a single longitudinal model because it avoids any concerns of undue bias stemming from incorrectly imposed time-varying restrictions on any variable.
 Without such restrictions, a pooled cross-sectional model does not yield any distinct advantage over separate cross-sections.
-The panel approach is avoided because of how frequently electoral boundaries change, noting that electorates with same name across elections are not guaranteed to represent the same geographical region.
+The panel approach is avoided because of how frequently electoral boundaries change, noting that electorates with the same name across elections are not guaranteed to represent the same geographical region.
 Therefore any fixed or random effects models would be difficult to estimate without implementing consistent boundaries, which would require further imputation.
 
-For each cross-section, let the response $\boldsymbol y$ be the vector two-party preferred vote in favour of the Liberal party, with $y_i = 70$ representing a 70% preference for Liberal, 30% for Labor, in electorate $i$.
+For each cross-section, let the response $\bm{y}$ be the vector two-party preferred vote in favour of the Liberal party; for example,  $y_i = 70$ represents a 70% preference for Liberal, 30% for Labor, in electorate $i$.
 Although $y_i$ lies in the interval $(0,100)$, observed values are never close to 0 or 100 (minimum $24.05 \%$ and maximum $74.90 \%$), so there is no need to formally impose the constraint of $y_i \in [0,100]$.
 Furthermore, the responses are found to be spatially correlated in each election (Moran's I test, $p \le 7\cdot10^{-15}$).
-This not surprising as electorates are aggregate spatial units, and hence the spatial structure of the data must modelled appropriately.
-
-The spatial error model [@Anselin88] is chosen because captures spatial heterogeneity by incorporating a spatially structured random effect vector [@LeSage2009].
+This is not surprising as electorates are aggregate spatial units, and hence the spatial structure of the data must modelled appropriately.
+
+The spatial error model [@Anselin88] is chosen because it captures spatial heterogeneity by incorporating a spatially structured random effect vector [@LeSage2009].
 In this context, the random effect can be thought of as capturing the unobserved political climate in each electorate, where this climate is correlated with the climate in neighbouring electorates, under the assumption that the climate is independent of electoral socio-demographics.
@@
   digits = 2,
-  column.labels = c("2001", "2004", "2007", "2010", "2013", "2016")
+  column.labels = c("2001", "2004", "2007", "2010", "2013", "2016"),
+  label="tab:printmodels"
   #notes = c("Estimated coefficients for variable named in column one shown for", "election year indicated by column heading, with estimated standard", "deviation for each coefficient shown below in parenthesis.
@@
 Spatial weights are calculated in accordance with the assumption that an electorate is equally correlated with any electorate that shares a part of its boundary.
-Let $\rho$ be spatial autoregressive coefficient, $\boldsymbol v$ be a spherical error term, $\bm{W}$ be a matrix of spatial weights (containing information about the neighbouring regions), $\boldsymbol X$ be a matrix of socio-demographic covariates, $\boldsymbol \beta$ be a vector of regression coefficients and $\boldsymbol a$ be a spatially structured random effect vector.
+Let $\rho$ be the spatial autoregressive coefficient, $\bm{v}$ be a spherical error term, $\bm{W}$ be a matrix of spatial weights (containing information about the neighbouring regions), $\bm{X}$ be a matrix of socio-demographic covariates, $\bm{\beta}$ be a vector of regression coefficients and $\bm{a}$ be a spatially structured random effect vector.
 $$
@@
 $$
-where
-$$
-  \bm{v} \sim N(\bm{0}, \sigma^2 \bm{I_n}),
-$$.
-and hence
+where $\bm{v} \sim N(\bm{0}, \sigma^2 \bm{I_n})$, and hence
 $$
@@
 
-Table 3 details the estimated model coefficients and their estimated standard errors, for each of the six elections. An interpretation of these estimated values is provided in the next section.
+Table \@ref(tab:printmodels) details the estimated model coefficients and their estimated standard errors, for each of the six elections.
+An interpretation of these estimated values is provided in the next section.
 
@@
 
-The spatial autoregressive coefficient $\rho$ is positive and significant in only the 2001 and 2016 elections (Figure \@ref(fig:rhovis)), meaning that, on average, in these elections, the political climate of an electorate appears to be affected by the attitudes of it's neighbours.
+The spatial autoregressive coefficient $\rho$ is positive and significant in only the 2001 and 2016 elections (Figure \@ref(fig:rhovis)), meaning that in these elections, the political climate of an electorate appears to be affected by the attitudes of its neighbours.
 Conversely, in the other four elections, the spatial effect weakens to become insignificant.
-In these years, it appears that the spatial component does not explain anything over and above the electoral socio-demographics, meaning electorates effectively voted independently.
+In these years, it appears that the spatial component does not explain anything not already explained by the electoral socio-demographics, meaning electorates effectively voted independently.
 
@@
 
-Since all socio-demographics have been standardized to have a mean of zero and a variance of one, the intercept in each model can be interpreted as the estimated two-party preferred vote for an electorate with mean characteristics[^3].
-The baseline of party preference has varied over the elections, with the biggest swing occurring in the 2007 election where the mean electorate shifted more than five percentage points in favour of the Labor party.
-
-[^3]: Mean of all variables aside from Judaism, Indigenous, Islam and Buddhism, where it assumes the mean of the log value.
-
 ```{r plotintercept, fig.cap = "Estimated intercept for each election, which represents the two-party preferred vote for an electorate with mean characteristics.", out.height="15%", out.extra = ''}
@@
 
+Since all socio-demographics have been standardized to have a mean of zero and a variance of one, the intercept in each model can be interpreted as the estimated two-party preferred vote for an electorate with mean characteristics[^3].
+Figure \@ref(fig:plotintercept) shows that the baseline of party preference has varied over the elections, with the biggest swing occurring in the 2007 election where the mean electorate shifted more than five percentage points in favour of the Labor party.
+
+[^3]: Mean of all variables aside from Judaism, Indigenous, Islam and Buddhism, where it assumes the mean of the log value.
+
 ## Influential socio-demographics
+
+```{r partresplot1, fig.cap = "Partial residual plots for: Incomes, Unemployment, Extractive, Transformative, ManagerAdmin and DiffAddress", fig.height = 9, fig.width = 6, warning = FALSE, fig.pos = 'h'}
+(p1 / p2 / (p3 + ylab("Two-party preferred vote (%)                     ")) / p4 / p5 / p6)
+```
+
+```{r partresplot2, fig.cap = "Partial residual plots for: DeFacto, MedianAge, Education, OtherLanguage, BornSEEuro, BornAsia and Judaism.", fig.height = 9,  fig.width = 6, warning = FALSE, fig.pos = 'h'}
+(p7 / p8 / (p9 + ylab("Two-party preferred vote (%)                     ")) / p10 / p11 / p12)
+```
 
@@
 The partial residuals are the residuals from the fitted model with the estimated effect an individual variable added.
-These show the direction, size and significance of an estimated effect - the slope of the prediction line matches the estimated coefficient, and the shaded region represents a 95% confidence band.
+These show the direction, size and significance of an estimated effect --- the slope of the prediction line matches the estimated coefficient, and the shaded region represents a 95% confidence band.
 If a horizontal line can be drawn through the confidence band, then the effect is insignificant.
 The estimated intercept is also added to the partial residuals for interpretability.
-Plots for each election are faceted to compare the effects over time in Figure \@ref(fig:partresplot1) and Figure \@ref(fig:partresplot2).
-
-It is important here to note the ecological fallacy: insights are being drawn at the electorate level, and cannot be inferred for another disaggregate level (e.g. individual voters).
+Plots for each election are faceted to compare the effects over time in Figures \@ref(fig:partresplot1) and \@ref(fig:partresplot2).
+
+It is important here to note the ecological fallacy: insights are being drawn at the electorate level, and cannot be inferred for another disaggregate level (e.g., individual voters).
 
@@
 Typically the Labor party campaigns on more progressive policies, which often include tax reform that adversely affects higher income earners, and more generous social assistance programs.
-Perhaps it is due to these policies, that higher income electorates appear more likely to support the Liberal party, as the `Incomes` factor has a positive effect on Liberal preference (see row 1 in Figure \@ref(fig:partresplot1)).
+Perhaps it is due to these policies that higher income electorates appear more likely to support the Liberal party, as the `Incomes` factor has a positive effect on Liberal preference (see row 1 in Figure \@ref(fig:partresplot1)).
 This effect is significant in every election aside from 2004, where it is only marginally insignificant ($p = 0.0613$).
@@
 
-Similarly, workers in managerial, administrative, clerical and sales roles (`ManagerAdmin`) is also a significant predictor of two-party preference vote across all six elections, with a higher proportion of people working these jobs increasing Liberal support.
+Similarly, the proportion of workers in managerial, administrative, clerical and sales roles (`ManagerAdmin`) is also a significant predictor of two-party preference vote across all six elections, with a higher proportion of people working these jobs increasing Liberal support.
 The magnitude of this effect also seems to increase over the years.
@@
 ### Household mobility
-In each of the six elections, electorates with a higher proportion of people that have recently (past five years) moved house are more likely to support the Liberal party, although this effect was marginally insignificant in 2013 (see row 6 in Figure \@ref(fig:partresplot1).
-Having controlled for characteristics of house ownership and rental prices (via the factors `PropertyOwned` and `RentLoan` respectively), this is effect is somewhat surprising.
+
+In each of the six elections, electorates with a higher proportion of people that have recently (in the past five years) moved house are more likely to support the Liberal party, although this effect was marginally insignificant in 2013 (see row 6 in Figure \@ref(fig:partresplot1).
+Having controlled for characteristics of house ownership and rental prices (via the factors `PropertyOwned` and `RentLoan` respectively), this effect is somewhat surprising.
 
 ### Relationships
+
 De facto relationships, but not marriages, are found to be an important (and significant) predictor of the two-party preferred vote in all six elections, with more de facto relationships associated with higher support for the Labor party.
-The proportion of individuals who are marriages however, are insignificant (not shown).
+The proportion of individuals who are married however, is insignificant (not shown).
 
 ### Age
-Regions comprised of older people are often believed to be more conservative, and indeed it found that electorates with a higher median age are more likely to support the Liberal party - although this effect is only significant in 2007 and 2010 (see row 2 in Figure \@ref(fig:partresplot2)).
+
+Regions comprising more older people are often believed to be more conservative, and indeed it found that electorates with a higher median age are more likely to support the Liberal party --- although this effect is only significant in 2007 and 2010 (see row 2 in Figure \@ref(fig:partresplot2)).
 
 ### Education
+
 Since 2007, electorates with higher education levels are associated with supporting the Labor party, although this effect is only significant in 2016.
@@
 ### Diversity
+
 Larger migrant populations from Asia, the Middle East, South-Eastern Europe, the United Kingdom and elsewhere, are either associated with Labor support, or have no effect.
@@
 Many of the Census variables represent similar information, which is why factors were created and some variables were removed.
-However, there still remain some variables which are closely related.
+However, some variables remain which are closely related.
 For example, an electorate's income level (via `Incomes`) is likely to be related to electoral unemployment and labor force participation (via `Unemployment`).
@@
 It is often hypothesized that states have systematic differences that cause their electorates to vote differently.
-Boxplots of residuals grouped by state reveal that only Tasmania, the Australian Capital Territory and the Northern Territory appear to have a state-specific effect not captured by the models.
+Boxplots of residuals grouped by state (Figure \@ref(fig:resstate)) reveal that only Tasmania, the Australian Capital Territory and the Northern Territory appear to have a state-specific effect not captured by the models.
 Tasmania and the Australian Capital Territory appear to have a bias towards Labor, whereas the Northern Territory tends towards voting Liberal.
@@
 The electorate of Sydney (NSW) has a large Cook's distance from 2001 to 2013, due to its diverse population (language, birthplace and religion), high number of defacto relationships, high income, high household mobility and small amount of workers in extractive and transformative jobs.
-It has remained a strong supporter of the Labor party and Liberal vote is severely overpredicted by the model, making it an outlier.
+It has remained a strong supporter of the Labor party and the Liberal vote is severely overpredicted by the model, making it an outlier.
 Nearby in metropolitan NSW, the electorate of Wentworth is found to be an outlier in all but the 2007 election.
-Although historically Liberal, its two-party vote jumped by over 10 percentage points in 2010 without experiencing any notable changes in its socio-demographic makeup - implying that this may be the direct effect of its Liberal member, Malcolm Turnbull, becoming the leader of the Liberal party.
+Although historically Liberal, its two-party vote jumped by over 10 percentage points in 2010 without experiencing any notable changes in its socio-demographic makeup --- implying that this may be the direct effect of its Liberal member, Malcolm Turnbull, becoming the leader of the Liberal party.
 Liberal support in Wentworth is underpredicted by the model in each year, and more so with Turnbull as Liberal leader.
 
-Lingiari, an electorate taking up almost all of the Northern Territory, is an outlier in the 2001-2007 elections due to its large Indigenous population, young age profile and low rates of property ownership.
+Lingiari, an electorate taking up almost all of the Northern Territory, is an outlier in the 2001--2007 elections due to its large Indigenous population, young age profile and low rates of property ownership.
 Fowler (NSW) has a diverse population with a high proportion of migrants, many Buddhists and Muslims, and has strong Labor support, making it influential in 2001, 2004 and 2010.
@@
 
-This paper explores the effects of electoral socio-demographic characteristics on the two-party preferred vote in the 2001-2016 elections, using information from the corresponding Australian federal elections and Censuses.
-As a Census does not always occur in the same year as an election, Census data for the 2004-2013 elections are generated by employing a method of spatio-temporal imputation.
-This imputes electoral socio-demographics for the electoral boundaries in place at the time of the election - an approach that is distinctly different from previous work on modelling election outcomes, where Census and election data are typically joined without addressing their temporal differences.
+This paper explores the effects of electoral socio-demographic characteristics on the two-party preferred vote in the 2001--2016 elections, using information from the corresponding Australian federal elections and Censuses.
+As a Census does not always occur in the same year as an election, Census data for the 2004--2013 elections are generated by employing a method of spatio-temporal imputation.
+This imputes electoral socio-demographics for the electoral boundaries in place at the time of the election --- an approach that is distinctly different from previous work on modelling election outcomes, where Census and election data are typically joined without addressing their temporal differences.
 Before estimating a model, these socio-demographic variables are standardized (to adjust for changing variable scales) and many variables (representing similar information) are combined into factors, resulting in a reduced predictor set.
-A spatial error model is then fit for each election, accounting for the inherent spatial structure of the data.
+A spatial error model is then estimated for each election, accounting for the inherent spatial structure of the data.
 
@@
 Higher levels of unemployment shift from weak association with Labor to a significant Liberal effect over the years, and higher education levels are associated with Labor from 2007 (although only significant in 2016).
-It is also found that electorates with higher household mobility support Liberal, birthplace diversity favours Labor and more de facto relationships align with Labor preference - although marriages, family and household sizes have no material influence.
-Furthermore, the neighbourhood (spatial) effects are found to be positive in all elections, although only significant in 2001 and 2016, meaning that in the 2004-2013 elections, electorates effectively voted independently.
+It is also found that electorates with higher household mobility support Liberal, birthplace diversity favours Labor and more de facto relationships align with Labor preference --- although marriages, family and household sizes have no material influence.
+Furthermore, the neighbourhood (spatial) effects are found to be positive in all elections, although only significant in 2001 and 2016, meaning that in the 2004--2013 elections, electorates effectively voted independently.
 
@@
 It is also the first study to model any Australian election since 2010 using Census information.
-Additionally, a key contribution of this research is the wrangling of the raw data and imputed data sets for the 2004, 2007, 2010 and 2013 elections, which have been contributed to the `eechidna` `R` package - providing a rich, accessible data resource for future Australian electoral analysis.
+
+Additionally, a key contribution of this research is the wrangling of the raw data and imputed data sets for the 2004, 2007, 2010 and 2013 elections, which have been contributed to the `eechidna` `R` package --- providing a rich, accessible data resource for future Australian electoral analysis.
 
@@
 This paper was produced using `RMarkdown` [@rmarkdown] and `knitr` [@knitr].
-All corresponding code for this paper can be found in the github repository (https://github.com/jforbes14/journal-writing), and the data used is available in the `eechidna` package [@eechidna].
+All corresponding code for this paper can be found in the github repository [github.com/jforbes14/journal-writing](https://github.com/jforbes14/journal-writing), and the data used is available in the `eechidna` package [@eechidna].
 All raw data was obtained from the Australian Electoral Commission, the Australian Bureau of Statistics and the Australian Government.
@@
 
-All election and Census datasets, along with electoral maps and more, are available in the `eechidna` `R` package, which can be downloaded from CRAN.
-`eechidna` (Exploring Election and Census Highly Informative Data Nationally for Australia) makes it easy to look at the data from the Australian Federal elections and Censuses that occurred between 2001 and 2016.
-This study contributed a large revision to the `eechidna` package, which included the addition of election and Census data for 2001-2010, voting outcomes for polling booths and imputed Census data for election years.
-For more details on using `eechidna`, please see the articles (vignettes) on the github page (https://ropenscilabs.github.io/eechidna/).
-
-The authors would like to give a sincere thanks to Anthony Ebert, Heike Hofmann, Thomas Lumley, Ben Marwick, Carson Sievert, Mingzhu Sun, Dilini Talagala, Nicholas Tierney, Nathaniel Tomasetti, Earo Wang and Fang Zhou, all of whom have contributed to the `eechidna` package.
-
-```{r partresplot1, fig.cap = "Partial residual plots for: Incomes, Unemployment, Extractive, Transformative, ManagerAdmin and DiffAddress", fig.height = 9, fig.width = 6, warning = FALSE, fig.pos = 'h'}
-(p1 / p2 / (p3 + ylab("Two-party preferred vote (%)                     ")) / p4 / p5 / p6)
-```
-
-```{r partresplot2, fig.cap = "Partial residual plots for: DeFacto, MedianAge, Education, OtherLanguage, BornSEEuro, BornAsia and Judaism.", fig.height = 9,  fig.width = 6, warning = FALSE, fig.pos = 'h'}
-(p7 / p8 / (p9 + ylab("Two-party preferred vote (%)                     ")) / p10 / p11 / p12)
-```
+All election and Census datasets, along with electoral maps and more, are available in the `eechidna` (Exploring Election and Census Highly Informative Data Nationally for Australia) `R` package, which can be downloaded from CRAN.
+The `eechidna`  package makes it easy to look at the data from the Australian Federal elections and Censuses that occurred between 2001 and 2016.
+This study contributed a large revision to the `eechidna` package, which included the addition of election and Census data for 2001--2010, voting outcomes for polling booths and imputed Census data for election years.
+For more details on using `eechidna`, please see the articles (vignettes) on the github page [ropenscilabs.github.io/eechidna/](https://ropenscilabs.github.io/eechidna/).
+
+The authors would like to sincerely thank Anthony Ebert, Heike Hofmann, Thomas Lumley, Ben Marwick, Carson Sievert, Mingzhu Sun, Dilini Talagala, Nicholas Tierney, Nathaniel Tomasetti, Earo Wang and Fang Zhou, all of whom have contributed to the `eechidna` package.

### 2019-04-10 draft-paper.tex
Nearly ready to submit

@@
 %% Any special functions or other packages can be loaded here.
-\usepackage{longtable}
 \usepackage{float}
@@
 \renewenvironment{figure}[1][2] {
-    \expandafter\origfigure\expandafter[!htbp]
+    \expandafter\origfigure\expandafter[htbp]
 } {
@@
 }
+
+\let\origtable\table
+\let\endorigtable\endtable
+\renewenvironment{table}[1][2] {
+    \expandafter\origtable\expandafter[htbp]
+} {
+    \endorigtable
+}
+
+% Don't number subsubsections
+
+\titleformat{\subsubsection}
+  {\normalfont\bfseries\sffamily}   % The style of the section title
+  {}                             % a prefix
+  {0pt}                          % How much space exists between the prefix and the title
+  {}    % How the section is represented
 
@@
 
-To account for spatial differences, the piece-wise approximation method in \textcite{Goodchild1993} is adopted.
+To account for spatial differences, the piecewise approximation method in \textcite{Goodchild1993} is adopted.
 Consider a map of source zones \(s = 1,\dots,S\), for which socio-demographic information is available, and a set of target zones \(t = 1,\dots,T\) for which information is to be imputed.
@@
 
-{\centering \includegraphics[height=0.3\textheight]{draft-paper_files/figure-latex/hume13-1}
+{\centering \includegraphics[height=0.28\textheight]{draft-paper_files/figure-latex/hume13-1}
 
@@
 
-{\centering \includegraphics[height=0.3\textheight]{draft-paper_files/figure-latex/hume16-1}
+{\centering \includegraphics[height=0.28\textheight]{draft-paper_files/figure-latex/hume16-1}
 
@@
 A factor is created from a group of variables if there is an intuitive reason as to why they should represent similar information and if there is evidence to suggest that they covary.
-For example, a potential group would be variables relating to electoral incomes - median family, household and personal incomes.
+For example, a potential group would be variables relating to electoral incomes --- median family, household and personal incomes.
 To determine which variables covary, principal component analysis is used on a combined dataset of socio-demographic variables from all six elections.
@@
 Six factors are created using this criteria.
-These are; \texttt{Incomes} - median personal, household and family incomes, \texttt{Unemployment} - unemployment and labor force participation rates, \texttt{ProportyOwned} - rates of housing ownership, mortgages, renting and government housing, \texttt{RentLoanPrice} - median rental and loan repayments, \texttt{FamHouseSize} - average household size, ratio of people to families and household makeup (single person, couple with kids and couple without kids) and \texttt{Education} - high school and university qualifications, jobs requiring higher levels of education as well as vocational course completions and jobs that do not require higher education levels (e.g.~laborer, tradesperson).
+These are:
+\texttt{Incomes} (median personal, household and family incomes);
+\texttt{Unemployment} (unemployment and labor force participation rates);
+\texttt{ProportyOwned} (rates of housing ownership, mortgages, renting and government housing);
+\texttt{RentLoanPrice} (median rental and loan repayments);
+\texttt{FamHouseSize} (average household size, ratio of people to families and household makeup (single person, couple with kids and couple without kids); and
+\texttt{Education} (high school and university qualifications, jobs requiring higher levels of education as well as vocational course completions and jobs that do not require higher education levels, such as laborer or tradesperson).
 For each of these groups, variables with positive loadings are added and those with negative loadings are subtracted to create a factor.
@@
 An identical model specification is used for each of the six elections, with each election modelled separately.
-This allows for the socio-demographic effects to be estimated separately for each election year, facilitating analysis of temporal changes in variable effects.
+This allows the socio-demographic effects to be estimated separately for each election year, facilitating analysis of temporal changes in variable effects.
 This approach is preferable to using a single longitudinal model because it avoids any concerns of undue bias stemming from incorrectly imposed time-varying restrictions on any variable.
 Without such restrictions, a pooled cross-sectional model does not yield any distinct advantage over separate cross-sections.
-The panel approach is avoided because of how frequently electoral boundaries change, noting that electorates with same name across elections are not guaranteed to represent the same geographical region.
+The panel approach is avoided because of how frequently electoral boundaries change, noting that electorates with the same name across elections are not guaranteed to represent the same geographical region.
 Therefore any fixed or random effects models would be difficult to estimate without implementing consistent boundaries, which would require further imputation.
 
-For each cross-section, let the response \(\boldsymbol y\) be the vector two-party preferred vote in favour of the Liberal party, with \(y_i = 70\) representing a 70\% preference for Liberal, 30\% for Labor, in electorate \(i\).
+For each cross-section, let the response \(\bm{y}\) be the vector two-party preferred vote in favour of the Liberal party; for example, \(y_i = 70\) represents a 70\% preference for Liberal, 30\% for Labor, in electorate \(i\).
 Although \(y_i\) lies in the interval \((0,100)\), observed values are never close to 0 or 100 (minimum \(24.05 \%\) and maximum \(74.90 \%\)), so there is no need to formally impose the constraint of \(y_i \in [0,100]\).
 Furthermore, the responses are found to be spatially correlated in each election (Moran's I test, \(p \le 7\cdot10^{-15}\)).
-This not surprising as electorates are aggregate spatial units, and hence the spatial structure of the data must modelled appropriately.
+This is not surprising as electorates are aggregate spatial units, and hence the spatial structure of the data must modelled appropriately.
 
-The spatial error model \autocite{Anselin88} is chosen because captures spatial heterogeneity by incorporating a spatially structured random effect vector \autocite{LeSage2009}.
+The spatial error model \autocite{Anselin88} is chosen because it captures spatial heterogeneity by incorporating a spatially structured random effect vector \autocite{LeSage2009}.
 In this context, the random effect can be thought of as capturing the unobserved political climate in each electorate, where this climate is correlated with the climate in neighbouring electorates, under the assumption that the climate is independent of electoral socio-demographics.
 
-\begin{table}[!htbp] \centering    \caption{Estimated spatial error model parameters (standard errors) for each of the six election years.}    \label{}  \scriptsize  \begin{tabular}{@{\extracolsep{1pt}}lcccccc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]  \\[-1.8ex] & \multicolumn{6}{c}{\vspace{-0.5cm}} \\   & 2001 & 2004 & 2007 & 2010 & 2013 & 2016 \\  \hline \\[-1.8ex]  $\rho$ & 0.46$^{***}$ & 0.29$^{*}$ & 0.24 & 0.19 & 0.27$^{*}$ & 0.50$^{***}$\\  & (0.15) & (0.17) & (0.17) & (0.16) & (0.16) & (0.17)\\  AusCitizen & $-$3.13 & $-$2.64 & $-$2.53 & $-$0.08 & $-$3.40 & $-$1.80 \\    & (2.26) & (2.43) & (2.34) & (2.79) & (2.76) & (2.71) \\    BornAsia & 2.22 & $-$0.95 & $-$1.60 & $-$6.83$^{**}$ & $-$3.03 & $-$0.55 \\    & (2.18) & (2.44) & (2.19) & (2.73) & (2.71) & (2.17) \\    Born\_MidEast & $-$1.15 & $-$1.59 & $-$2.01$^{*}$ & $-$2.03 & $-$0.92 & $-$1.44 \\    & (1.07) & (1.20) & (1.11) & (1.27) & (1.24) & (1.13) \\    BornSEEuro & $-$3.21$^{**}$ & $-$4.24$^{***}$ & $-$3.61$^{***}$ & $-$4.14$^{***}$ & $-$3.69$^{***}$ & $-$2.72$^{***}$ \\    & (1.42) & (1.46) & (1.02) & (1.19) & (1.07) & (0.97) \\    Born\_UK & 0.25 & $-$0.07 & 0.34 & 0.56 & $-$0.09 & $-$1.32 \\    & (1.00) & (0.98) & (0.90) & (1.07) & (1.04) & (1.04) \\    BornElsewhere & $-$5.04 & $-$4.91 & $-$4.13 & 2.35 & $-$5.23 & $-$4.14 \\    & (3.30) & (3.68) & (3.38) & (4.23) & (4.15) & (3.97) \\    Buddhism & $-$0.49 & $-$0.17 & $-$1.37 & $-$0.83 & $-$0.12 & $-$1.60 \\    & (1.39) & (1.61) & (1.61) & (1.80) & (1.68) & (1.56) \\    Christianity & $-$2.48 & $-$1.23 & 0.38 & 0.50 & 2.41 & 1.68 \\    & (1.73) & (1.85) & (1.83) & (1.99) & (1.85) & (1.78) \\    CurrentlyStudying & $-$2.19$^{**}$ & $-$0.13 & 2.06$^{*}$ & 2.12$^{*}$ & 1.15 & $-$0.16 \\    & (0.99) & (1.13) & (1.17) & (1.25) & (1.26) & (1.18) \\    DeFacto & $-$6.44$^{***}$ & $-$5.37$^{**}$ & $-$6.43$^{***}$ & $-$8.07$^{***}$ & $-$6.56$^{**}$ & $-$8.53$^{***}$ \\    & (1.87) & (2.48) & (2.31) & (3.06) & (3.11) & (2.83) \\    DiffAddress & 3.88$^{***}$ & 5.06$^{***}$ & 4.22$^{***}$ & 5.57$^{***}$ & 3.53$^{*}$ & 5.67$^{***}$ \\    & (0.94) & (1.12) & (0.99) & (1.76) & (1.91) & (1.60) \\    Distributive & 1.27 & 2.01$^{*}$ & 1.36 & 1.57 & 2.10$^{*}$ & 1.20 \\    & (1.12) & (1.21) & (1.13) & (1.34) & (1.27) & (1.21) \\    Education & 1.08 & 0.52 & $-$5.52$^{*}$ & $-$4.08 & $-$4.44 & $-$7.07$^{**}$ \\    & (2.38) & (3.12) & (3.27) & (3.95) & (3.78) & (3.55) \\    Extractive & 4.83$^{***}$ & 5.45$^{***}$ & 5.37$^{***}$ & 7.31$^{***}$ & 6.71$^{***}$ & 7.43$^{***}$ \\    & (1.48) & (1.42) & (1.36) & (1.56) & (1.47) & (1.39) \\    FamHouseSize & $-$0.16 & 0.87 & $-$2.40 & $-$2.53 & $-$3.26 & $-$2.91 \\    & (2.19) & (2.72) & (2.69) & (3.25) & (3.28) & (2.90) \\    Incomes & 4.36$^{**}$ & 5.03$^{*}$ & 9.45$^{***}$ & 7.09$^{**}$ & 7.97$^{***}$ & 12.20$^{***}$ \\    & (1.77) & (2.66) & (2.75) & (3.25) & (2.92) & (2.75) \\    Indigenous & 2.91$^{*}$ & 1.97 & 2.48 & 2.84 & 0.67 & $-$0.05 \\    & (1.68) & (1.95) & (1.75) & (2.16) & (2.14) & (2.00) \\    Islam & $-$0.92 & $-$0.97 & $-$0.54 & $-$2.50 & $-$0.82 & $-$0.95 \\    & (1.22) & (1.36) & (1.27) & (1.52) & (1.42) & (1.34) \\    Judaism & 1.88$^{*}$ & 1.78 & 2.66$^{***}$ & 1.97$^{*}$ & 2.74$^{**}$ & 1.65$^{*}$ \\    & (1.05) & (1.13) & (1.01) & (1.15) & (1.10) & (1.00) \\    ManagerAdmin & 2.06$^{***}$ & 3.32$^{***}$ & 6.00$^{***}$ & 5.47$^{***}$ & 5.04$^{***}$ & 5.78$^{***}$ \\    & (0.71) & (0.93) & (0.90) & (1.08) & (1.03) & (1.06) \\    Married & 0.44 & 0.11 & $-$1.22 & $-$0.22 & 0.91 & $-$2.34 \\    & (2.31) & (2.96) & (2.83) & (3.15) & (3.03) & (2.81) \\    MedianAge & 2.32$^{*}$ & 4.96$^{***}$ & 3.66$^{**}$ & 4.00$^{*}$ & 2.30 & 2.87 \\    & (1.32) & (1.65) & (1.81) & (2.26) & (2.08) & (1.79) \\    NoReligion & $-$1.57 & $-$0.92 & 0.56 & $-$0.30 & 1.02 & 1.31 \\    & (1.59) & (1.71) & (1.73) & (1.92) & (1.94) & (2.04) \\    OneParentHouse & $-$1.73 & $-$0.45 & $-$0.75 & $-$1.46 & $-$0.77 & $-$0.74 \\    & (1.36) & (1.59) & (1.49) & (1.69) & (1.57) & (1.47) \\    OtherLanguage & $-$0.44 & 5.92 & 9.98$^{**}$ & 11.24$^{**}$ & 9.00$^{*}$ & 9.84$^{**}$ \\    & (3.22) & (4.16) & (3.91) & (4.76) & (4.66) & (4.44) \\    PropertyOwned & $-$0.46 & $-$0.53 & 0.67 & $-$0.94 & $-$0.48 & 1.41 \\    & (1.37) & (1.50) & (1.43) & (1.76) & (1.67) & (1.50) \\    RentLoanPrice & $-$1.57 & $-$3.32$^{*}$ & $-$4.01$^{**}$ & $-$0.97 & $-$0.70 & $-$0.89 \\    & (1.49) & (1.76) & (1.67) & (2.07) & (2.07) & (2.16) \\    SocialServ & 2.51$^{*}$ & 1.65 & 2.47$^{*}$ & 2.53$^{*}$ & 2.35$^{*}$ & 4.45$^{***}$ \\    & (1.33) & (1.41) & (1.29) & (1.47) & (1.32) & (1.19) \\    Transformative & 3.24$^{**}$ & 4.73$^{***}$ & 4.84$^{***}$ & 4.46$^{**}$ & 3.56$^{**}$ & 4.58$^{***}$ \\    & (1.55) & (1.78) & (1.74) & (1.98) & (1.78) & (1.53) \\    Unemployment & $-$2.45$^{*}$ & $-$3.07$^{*}$ & 0.29 & 0.08 & 1.67 & 2.79$^{**}$ \\    & (1.40) & (1.63) & (1.51) & (1.76) & (1.51) & (1.37) \\    Constant & 50.81$^{***}$ & 52.60$^{***}$ & 47.31$^{***}$ & 49.93$^{***}$ & 53.51$^{***}$ & 50.49$^{***}$ \\    & (0.71) & (0.58) & (0.52) & (0.57) & (0.58) & (0.80) \\   \hline \\[-1.8ex]  Observations & 150 & 150 & 150 & 150 & 150 & 150 \\  Residual Standard Error (GLS) & 4.69 & 5.04 & 4.79 & 5.63 & 5.18 & 4.88\\ \hline  \hline \\[-1.8ex]  \textit{Note:}  & \multicolumn{6}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\  \end{tabular}  \end{table}
+\begin{table}[!htbp] \centering    \caption{Estimated spatial error model parameters (standard errors) for each of the six election years.}    \label{tab:printmodels}  \scriptsize  \begin{tabular}{@{\extracolsep{1pt}}lcccccc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]  \\[-1.8ex] & \multicolumn{6}{c}{\vspace{-0.5cm}} \\   & 2001 & 2004 & 2007 & 2010 & 2013 & 2016 \\  \hline \\[-1.8ex]  $\rho$ & 0.46$^{***}$ & 0.29$^{*}$ & 0.24 & 0.19 & 0.27$^{*}$ & 0.50$^{***}$\\  & (0.15) & (0.17) & (0.17) & (0.16) & (0.16) & (0.17)\\  AusCitizen & $-$3.13 & $-$2.64 & $-$2.53 & $-$0.08 & $-$3.40 & $-$1.80 \\    & (2.26) & (2.43) & (2.34) & (2.79) & (2.76) & (2.71) \\    BornAsia & 2.22 & $-$0.95 & $-$1.60 & $-$6.83$^{**}$ & $-$3.03 & $-$0.55 \\    & (2.18) & (2.44) & (2.19) & (2.73) & (2.71) & (2.17) \\    Born\_MidEast & $-$1.15 & $-$1.59 & $-$2.01$^{*}$ & $-$2.03 & $-$0.92 & $-$1.44 \\    & (1.07) & (1.20) & (1.11) & (1.27) & (1.24) & (1.13) \\    BornSEEuro & $-$3.21$^{**}$ & $-$4.24$^{***}$ & $-$3.61$^{***}$ & $-$4.14$^{***}$ & $-$3.69$^{***}$ & $-$2.72$^{***}$ \\    & (1.42) & (1.46) & (1.02) & (1.19) & (1.07) & (0.97) \\    Born\_UK & 0.25 & $-$0.07 & 0.34 & 0.56 & $-$0.09 & $-$1.32 \\    & (1.00) & (0.98) & (0.90) & (1.07) & (1.04) & (1.04) \\    BornElsewhere & $-$5.04 & $-$4.91 & $-$4.13 & 2.35 & $-$5.23 & $-$4.14 \\    & (3.30) & (3.68) & (3.38) & (4.23) & (4.15) & (3.97) \\    Buddhism & $-$0.49 & $-$0.17 & $-$1.37 & $-$0.83 & $-$0.12 & $-$1.60 \\    & (1.39) & (1.61) & (1.61) & (1.80) & (1.68) & (1.56) \\    Christianity & $-$2.48 & $-$1.23 & 0.38 & 0.50 & 2.41 & 1.68 \\    & (1.73) & (1.85) & (1.83) & (1.99) & (1.85) & (1.78) \\    CurrentlyStudying & $-$2.19$^{**}$ & $-$0.13 & 2.06$^{*}$ & 2.12$^{*}$ & 1.15 & $-$0.16 \\    & (0.99) & (1.13) & (1.17) & (1.25) & (1.26) & (1.18) \\    DeFacto & $-$6.44$^{***}$ & $-$5.37$^{**}$ & $-$6.43$^{***}$ & $-$8.07$^{***}$ & $-$6.56$^{**}$ & $-$8.53$^{***}$ \\    & (1.87) & (2.48) & (2.31) & (3.06) & (3.11) & (2.83) \\    DiffAddress & 3.88$^{***}$ & 5.06$^{***}$ & 4.22$^{***}$ & 5.57$^{***}$ & 3.53$^{*}$ & 5.67$^{***}$ \\    & (0.94) & (1.12) & (0.99) & (1.76) & (1.91) & (1.60) \\    Distributive & 1.27 & 2.01$^{*}$ & 1.36 & 1.57 & 2.10$^{*}$ & 1.20 \\    & (1.12) & (1.21) & (1.13) & (1.34) & (1.27) & (1.21) \\    Education & 1.08 & 0.52 & $-$5.52$^{*}$ & $-$4.08 & $-$4.44 & $-$7.07$^{**}$ \\    & (2.38) & (3.12) & (3.27) & (3.95) & (3.78) & (3.55) \\    Extractive & 4.83$^{***}$ & 5.45$^{***}$ & 5.37$^{***}$ & 7.31$^{***}$ & 6.71$^{***}$ & 7.43$^{***}$ \\    & (1.48) & (1.42) & (1.36) & (1.56) & (1.47) & (1.39) \\    FamHouseSize & $-$0.16 & 0.87 & $-$2.40 & $-$2.53 & $-$3.26 & $-$2.91 \\    & (2.19) & (2.72) & (2.69) & (3.25) & (3.28) & (2.90) \\    Incomes & 4.36$^{**}$ & 5.03$^{*}$ & 9.45$^{***}$ & 7.09$^{**}$ & 7.97$^{***}$ & 12.20$^{***}$ \\    & (1.77) & (2.66) & (2.75) & (3.25) & (2.92) & (2.75) \\    Indigenous & 2.91$^{*}$ & 1.97 & 2.48 & 2.84 & 0.67 & $-$0.05 \\    & (1.68) & (1.95) & (1.75) & (2.16) & (2.14) & (2.00) \\    Islam & $-$0.92 & $-$0.97 & $-$0.54 & $-$2.50 & $-$0.82 & $-$0.95 \\    & (1.22) & (1.36) & (1.27) & (1.52) & (1.42) & (1.34) \\    Judaism & 1.88$^{*}$ & 1.78 & 2.66$^{***}$ & 1.97$^{*}$ & 2.74$^{**}$ & 1.65$^{*}$ \\    & (1.05) & (1.13) & (1.01) & (1.15) & (1.10) & (1.00) \\    ManagerAdmin & 2.06$^{***}$ & 3.32$^{***}$ & 6.00$^{***}$ & 5.47$^{***}$ & 5.04$^{***}$ & 5.78$^{***}$ \\    & (0.71) & (0.93) & (0.90) & (1.08) & (1.03) & (1.06) \\    Married & 0.44 & 0.11 & $-$1.22 & $-$0.22 & 0.91 & $-$2.34 \\    & (2.31) & (2.96) & (2.83) & (3.15) & (3.03) & (2.81) \\    MedianAge & 2.32$^{*}$ & 4.96$^{***}$ & 3.66$^{**}$ & 4.00$^{*}$ & 2.30 & 2.87 \\    & (1.32) & (1.65) & (1.81) & (2.26) & (2.08) & (1.79) \\    NoReligion & $-$1.57 & $-$0.92 & 0.56 & $-$0.30 & 1.02 & 1.31 \\    & (1.59) & (1.71) & (1.73) & (1.92) & (1.94) & (2.04) \\    OneParentHouse & $-$1.73 & $-$0.45 & $-$0.75 & $-$1.46 & $-$0.77 & $-$0.74 \\    & (1.36) & (1.59) & (1.49) & (1.69) & (1.57) & (1.47) \\    OtherLanguage & $-$0.44 & 5.92 & 9.98$^{**}$ & 11.24$^{**}$ & 9.00$^{*}$ & 9.84$^{**}$ \\    & (3.22) & (4.16) & (3.91) & (4.76) & (4.66) & (4.44) \\    PropertyOwned & $-$0.46 & $-$0.53 & 0.67 & $-$0.94 & $-$0.48 & 1.41 \\    & (1.37) & (1.50) & (1.43) & (1.76) & (1.67) & (1.50) \\    RentLoanPrice & $-$1.57 & $-$3.32$^{*}$ & $-$4.01$^{**}$ & $-$0.97 & $-$0.70 & $-$0.89 \\    & (1.49) & (1.76) & (1.67) & (2.07) & (2.07) & (2.16) \\    SocialServ & 2.51$^{*}$ & 1.65 & 2.47$^{*}$ & 2.53$^{*}$ & 2.35$^{*}$ & 4.45$^{***}$ \\    & (1.33) & (1.41) & (1.29) & (1.47) & (1.32) & (1.19) \\    Transformative & 3.24$^{**}$ & 4.73$^{***}$ & 4.84$^{***}$ & 4.46$^{**}$ & 3.56$^{**}$ & 4.58$^{***}$ \\    & (1.55) & (1.78) & (1.74) & (1.98) & (1.78) & (1.53) \\    Unemployment & $-$2.45$^{*}$ & $-$3.07$^{*}$ & 0.29 & 0.08 & 1.67 & 2.79$^{**}$ \\    & (1.40) & (1.63) & (1.51) & (1.76) & (1.51) & (1.37) \\    Constant & 50.81$^{***}$ & 52.60$^{***}$ & 47.31$^{***}$ & 49.93$^{***}$ & 53.51$^{***}$ & 50.49$^{***}$ \\    & (0.71) & (0.58) & (0.52) & (0.57) & (0.58) & (0.80) \\   \hline \\[-1.8ex]  Observations & 150 & 150 & 150 & 150 & 150 & 150 \\  Residual Standard Error (GLS) & 4.69 & 5.04 & 4.79 & 5.63 & 5.18 & 4.88\\ \hline  \hline \\[-1.8ex]  \textit{Note:}  & \multicolumn{6}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\  \end{tabular}  \end{table}
 
 Spatial weights are calculated in accordance with the assumption that an electorate is equally correlated with any electorate that shares a part of its boundary.
-Let \(\rho\) be spatial autoregressive coefficient, \(\boldsymbol v\) be a spherical error term, \(\bm{W}\) be a matrix of spatial weights (containing information about the neighbouring regions), \(\boldsymbol X\) be a matrix of socio-demographic covariates, \(\boldsymbol \beta\) be a vector of regression coefficients and \(\boldsymbol a\) be a spatially structured random effect vector.
+Let \(\rho\) be the spatial autoregressive coefficient, \(\bm{v}\) be a spherical error term, \(\bm{W}\) be a matrix of spatial weights (containing information about the neighbouring regions), \(\bm{X}\) be a matrix of socio-demographic covariates, \(\bm{\beta}\) be a vector of regression coefficients and \(\bm{a}\) be a spatially structured random effect vector.
 \[
@@
 \]
-where
-\[
-  \bm{v} \sim N(\bm{0}, \sigma^2 \bm{I_n}),
-\].
-and hence
+where \(\bm{v} \sim N(\bm{0}, \sigma^2 \bm{I_n})\), and hence
 \[
@@
 
-Table 3 details the estimated model coefficients and their estimated standard errors, for each of the six elections. An interpretation of these estimated values is provided in the next section.
+Table \ref{tab:printmodels} details the estimated model coefficients and their estimated standard errors, for each of the six elections.
+An interpretation of these estimated values is provided in the next section.
 
@@
 
-The spatial autoregressive coefficient \(\rho\) is positive and significant in only the 2001 and 2016 elections (Figure \ref{fig:rhovis}), meaning that, on average, in these elections, the political climate of an electorate appears to be affected by the attitudes of it's neighbours.
+The spatial autoregressive coefficient \(\rho\) is positive and significant in only the 2001 and 2016 elections (Figure \ref{fig:rhovis}), meaning that in these elections, the political climate of an electorate appears to be affected by the attitudes of its neighbours.
 Conversely, in the other four elections, the spatial effect weakens to become insignificant.
-In these years, it appears that the spatial component does not explain anything over and above the electoral socio-demographics, meaning electorates effectively voted independently.
+In these years, it appears that the spatial component does not explain anything not already explained by the electoral socio-demographics, meaning electorates effectively voted independently.
 
@@
 
-Since all socio-demographics have been standardized to have a mean of zero and a variance of one, the intercept in each model can be interpreted as the estimated two-party preferred vote for an electorate with mean characteristics\footnote{Mean of all variables aside from Judaism, Indigenous, Islam and Buddhism, where it assumes the mean of the log value.}.
-The baseline of party preference has varied over the elections, with the biggest swing occurring in the 2007 election where the mean electorate shifted more than five percentage points in favour of the Labor party.
-
 \begin{figure}[h]
@@
 
+Since all socio-demographics have been standardized to have a mean of zero and a variance of one, the intercept in each model can be interpreted as the estimated two-party preferred vote for an electorate with mean characteristics\footnote{Mean of all variables aside from Judaism, Indigenous, Islam and Buddhism, where it assumes the mean of the log value.}.
+Figure \ref{fig:plotintercept} shows that the baseline of party preference has varied over the elections, with the biggest swing occurring in the 2007 election where the mean electorate shifted more than five percentage points in favour of the Labor party.
+
 \hypertarget{influential-socio-demographics}{%
@@
 
+\begin{figure}[h]
+
+{\centering \includegraphics{draft-paper_files/figure-latex/partresplot1-1}
+
+}
+
+\caption{Partial residual plots for: Incomes, Unemployment, Extractive, Transformative, ManagerAdmin and DiffAddress}\label{fig:partresplot1}
+\end{figure}
+
+\begin{figure}[h]
+
+{\centering \includegraphics{draft-paper_files/figure-latex/partresplot2-1}
+
+}
+
+\caption{Partial residual plots for: DeFacto, MedianAge, Education, OtherLanguage, BornSEEuro, BornAsia and Judaism.}\label{fig:partresplot2}
+\end{figure}
+
 To investigate the socio-demographics that have a strong effect on the two-party preferred vote, partial residual plots are used and shown in Figures \ref{fig:partresplot1} and \ref{fig:partresplot2}.
 The partial residuals are the residuals from the fitted model with the estimated effect an individual variable added.
-These show the direction, size and significance of an estimated effect - the slope of the prediction line matches the estimated coefficient, and the shaded region represents a 95\% confidence band.
+These show the direction, size and significance of an estimated effect --- the slope of the prediction line matches the estimated coefficient, and the shaded region represents a 95\% confidence band.
 If a horizontal line can be drawn through the confidence band, then the effect is insignificant.
 The estimated intercept is also added to the partial residuals for interpretability.
-Plots for each election are faceted to compare the effects over time in Figure \ref{fig:partresplot1} and Figure \ref{fig:partresplot2}.
+Plots for each election are faceted to compare the effects over time in Figures \ref{fig:partresplot1} and \ref{fig:partresplot2}.
 
-It is important here to note the ecological fallacy: insights are being drawn at the electorate level, and cannot be inferred for another disaggregate level (e.g.~individual voters).
+It is important here to note the ecological fallacy: insights are being drawn at the electorate level, and cannot be inferred for another disaggregate level (e.g., individual voters).
 
@@
 Typically the Labor party campaigns on more progressive policies, which often include tax reform that adversely affects higher income earners, and more generous social assistance programs.
-Perhaps it is due to these policies, that higher income electorates appear more likely to support the Liberal party, as the \texttt{Incomes} factor has a positive effect on Liberal preference (see row 1 in Figure \ref{fig:partresplot1}).
+Perhaps it is due to these policies that higher income electorates appear more likely to support the Liberal party, as the \texttt{Incomes} factor has a positive effect on Liberal preference (see row 1 in Figure \ref{fig:partresplot1}).
 This effect is significant in every election aside from 2004, where it is only marginally insignificant (\(p = 0.0613\)).
@@
 
-Similarly, workers in managerial, administrative, clerical and sales roles (\texttt{ManagerAdmin}) is also a significant predictor of two-party preference vote across all six elections, with a higher proportion of people working these jobs increasing Liberal support.
+Similarly, the proportion of workers in managerial, administrative, clerical and sales roles (\texttt{ManagerAdmin}) is also a significant predictor of two-party preference vote across all six elections, with a higher proportion of people working these jobs increasing Liberal support.
 The magnitude of this effect also seems to increase over the years.
@@
 
-In each of the six elections, electorates with a higher proportion of people that have recently (past five years) moved house are more likely to support the Liberal party, although this effect was marginally insignificant in 2013 (see row 6 in Figure \ref{fig:partresplot1}.
-Having controlled for characteristics of house ownership and rental prices (via the factors \texttt{PropertyOwned} and \texttt{RentLoan} respectively), this is effect is somewhat surprising.
+In each of the six elections, electorates with a higher proportion of people that have recently (in the past five years) moved house are more likely to support the Liberal party, although this effect was marginally insignificant in 2013 (see row 6 in Figure \ref{fig:partresplot1}.
+Having controlled for characteristics of house ownership and rental prices (via the factors \texttt{PropertyOwned} and \texttt{RentLoan} respectively), this effect is somewhat surprising.
 
@@
 De facto relationships, but not marriages, are found to be an important (and significant) predictor of the two-party preferred vote in all six elections, with more de facto relationships associated with higher support for the Labor party.
-The proportion of individuals who are marriages however, are insignificant (not shown).
+The proportion of individuals who are married however, is insignificant (not shown).
 
@@
 
-Regions comprised of older people are often believed to be more conservative, and indeed it found that electorates with a higher median age are more likely to support the Liberal party - although this effect is only significant in 2007 and 2010 (see row 2 in Figure \ref{fig:partresplot2}).
+Regions comprising more older people are often believed to be more conservative, and indeed it found that electorates with a higher median age are more likely to support the Liberal party --- although this effect is only significant in 2007 and 2010 (see row 2 in Figure \ref{fig:partresplot2}).
 
@@
 Many of the Census variables represent similar information, which is why factors were created and some variables were removed.
-However, there still remain some variables which are closely related.
+However, some variables remain which are closely related.
 For example, an electorate's income level (via \texttt{Incomes}) is likely to be related to electoral unemployment and labor force participation (via \texttt{Unemployment}).
@@
 It is often hypothesized that states have systematic differences that cause their electorates to vote differently.
-Boxplots of residuals grouped by state reveal that only Tasmania, the Australian Capital Territory and the Northern Territory appear to have a state-specific effect not captured by the models.
+Boxplots of residuals grouped by state (Figure \ref{fig:resstate}) reveal that only Tasmania, the Australian Capital Territory and the Northern Territory appear to have a state-specific effect not captured by the models.
 Tasmania and the Australian Capital Territory appear to have a bias towards Labor, whereas the Northern Territory tends towards voting Liberal.
@@
 It is also the first study to model any Australian election since 2010 using Census information.
-Additionally, a key contribution of this research is the wrangling of the raw data and imputed data sets for the 2004, 2007, 2010 and 2013 elections, which have been contributed to the \texttt{eechidna} \texttt{R} package - providing a rich, accessible data resource for future Australian electoral analysis.
+
+Additionally, a key contribution of this research is the wrangling of the raw data and imputed data sets for the 2004, 2007, 2010 and 2013 elections, which have been contributed to the \texttt{eechidna} \texttt{R} package --- providing a rich, accessible data resource for future Australian electoral analysis.
 
@@
 This paper was produced using \texttt{RMarkdown} \autocite{rmarkdown} and \texttt{knitr} \autocite{knitr}.
-All corresponding code for this paper can be found in the github repository (\url{https://github.com/jforbes14/journal-writing}), and the data used is available in the \texttt{eechidna} package \autocite{eechidna}.
+All corresponding code for this paper can be found in the github repository \href{https://github.com/jforbes14/journal-writing}{github.com/jforbes14/journal-writing}, and the data used is available in the \texttt{eechidna} package \autocite{eechidna}.
 All raw data was obtained from the Australian Electoral Commission, the Australian Bureau of Statistics and the Australian Government.
@@
 
-All election and Census datasets, along with electoral maps and more, are available in the \texttt{eechidna} \texttt{R} package, which can be downloaded from CRAN.
-\texttt{eechidna} (Exploring Election and Census Highly Informative Data Nationally for Australia) makes it easy to look at the data from the Australian Federal elections and Censuses that occurred between 2001 and 2016.
-This study contributed a large revision to the \texttt{eechidna} package, which included the addition of election and Census data for 2001-2010, voting outcomes for polling booths and imputed Census data for election years.
-For more details on using \texttt{eechidna}, please see the articles (vignettes) on the github page (\url{https://ropenscilabs.github.io/eechidna/}).
+All election and Census datasets, along with electoral maps and more, are available in the \texttt{eechidna} (Exploring Election and Census Highly Informative Data Nationally for Australia) \texttt{R} package, which can be downloaded from CRAN.
+The \texttt{eechidna} package makes it easy to look at the data from the Australian Federal elections and Censuses that occurred between 2001 and 2016.
+This study contributed a large revision to the \texttt{eechidna} package, which included the addition of election and Census data for 2001--2010, voting outcomes for polling booths and imputed Census data for election years.
+For more details on using \texttt{eechidna}, please see the articles (vignettes) on the github page \href{https://ropenscilabs.github.io/eechidna/}{ropenscilabs.github.io/eechidna/}.
 
-The authors would like to give a sincere thanks to Anthony Ebert, Heike Hofmann, Thomas Lumley, Ben Marwick, Carson Sievert, Mingzhu Sun, Dilini Talagala, Nicholas Tierney, Nathaniel Tomasetti, Earo Wang and Fang Zhou, all of whom have contributed to the \texttt{eechidna} package.
-
-\begin{figure}[h]
-
-{\centering \includegraphics{draft-paper_files/figure-latex/partresplot1-1}
-
-}
-
-\caption{Partial residual plots for: Incomes, Unemployment, Extractive, Transformative, ManagerAdmin and DiffAddress}\label{fig:partresplot1}
-\end{figure}
-
-\begin{figure}[h]
-
-{\centering \includegraphics{draft-paper_files/figure-latex/partresplot2-1}
-
-}
-
-\caption{Partial residual plots for: DeFacto, MedianAge, Education, OtherLanguage, BornSEEuro, BornAsia and Judaism.}\label{fig:partresplot2}
-\end{figure}
+The authors would like to sincerely thank Anthony Ebert, Heike Hofmann, Thomas Lumley, Ben Marwick, Carson Sievert, Mingzhu Sun, Dilini Talagala, Nicholas Tierney, Nathaniel Tomasetti, Earo Wang and Fang Zhou, all of whom have contributed to the \texttt{eechidna} package.
 

### 2019-04-13 draft-paper.Rmd
tweaked a few figures and language in the results section

@@
 - familyname: Cook
-  othernames: Di
+  othernames: Dianne
   address: Monash University\newline Melbourne, Australia
@@
 library(eechidna)
-library(ggplot2)
+library(forcats)
 library(ggthemes)
@@
 ```{r rhovis, fig.cap="Estimates of the spatial autoregressive parameter for each of the six elections, reported with their individual 95\\% confidence intervals.
-Only in 2001 and 2016 is there a significant spatial component.", out.height='15%', out.extra = ''}
+Only in 2001 and 2016 is there a significant spatial component.", out.height='20%', out.extra = ''}
 rho_df <- bind_rows(
@@
 
-```{r plotintercept, fig.cap = "Estimated intercept for each election, which represents the two-party preferred vote for an electorate with mean characteristics.", out.height="15%", out.extra = ''}
+```{r plotintercept, fig.cap = "Estimated intercept for each election, which represents the two-party preferred vote for an electorate with mean characteristics.", out.height="20%", out.extra = ''}
 intercept_df <- bind_rows(
@@
 ## Influential socio-demographics
-
-```{r partresplot1, fig.cap = "Partial residual plots for: Incomes, Unemployment, Extractive, Transformative, ManagerAdmin and DiffAddress", fig.height = 9, fig.width = 6, warning = FALSE, fig.pos = 'h'}
-(p1 / p2 / (p3 + ylab("Two-party preferred vote (%)                     ")) / p4 / p5 / p6)
-```
-
-```{r partresplot2, fig.cap = "Partial residual plots for: DeFacto, MedianAge, Education, OtherLanguage, BornSEEuro, BornAsia and Judaism.", fig.height = 9,  fig.width = 6, warning = FALSE, fig.pos = 'h'}
-(p7 / p8 / (p9 + ylab("Two-party preferred vote (%)                     ")) / p10 / p11 / p12)
-```
-
-To investigate the socio-demographics that have a strong effect on the two-party preferred vote, partial residual plots are used and shown in Figures \@ref(fig:partresplot1) and \@ref(fig:partresplot2).
-The partial residuals are the residuals from the fitted model with the estimated effect an individual variable added.
-These show the direction, size and significance of an estimated effect --- the slope of the prediction line matches the estimated coefficient, and the shaded region represents a 95% confidence band.
-If a horizontal line can be drawn through the confidence band, then the effect is insignificant.
-The estimated intercept is also added to the partial residuals for interpretability.
-Plots for each election are faceted to compare the effects over time in Figures \@ref(fig:partresplot1) and \@ref(fig:partresplot2).
-
-It is important here to note the ecological fallacy: insights are being drawn at the electorate level, and cannot be inferred for another disaggregate level (e.g., individual voters).
-
-### Income and unemployment
-
-Typically the Labor party campaigns on more progressive policies, which often include tax reform that adversely affects higher income earners, and more generous social assistance programs.
-Perhaps it is due to these policies that higher income electorates appear more likely to support the Liberal party, as the `Incomes` factor has a positive effect on Liberal preference (see row 1 in Figure \@ref(fig:partresplot1)).
-This effect is significant in every election aside from 2004, where it is only marginally insignificant ($p = 0.0613$).
-Unemployment however, is not as influential.
-In 2001 and 2004, electorates with higher unemployment align with Labor, but over time this shifts towards support for the Liberal party, culminating in a significantly positive effect in 2016.
-
-### Industry and type of work
-
-Electorates with higher proportions of workers in mining, gas, water, agriculture, waste and electricity (grouped as `Extractive` industries) are consistently linked with higher support for the Liberal party, with the magnitude of this effect slightly increasing over the years (see row 3 in Figure \@ref(fig:partresplot1)).
-This is unsurprising, as the Liberal party has close ties with these traditional energy industries, and typically present policies to reduce taxation on energy production.
-Furthermore, electorates with more workers in construction or manufacturing industries (`Transformative`) are also more likely to support the Liberal party (see row 4 in Figure \@ref(fig:partresplot1)).
-
-Similarly, the proportion of workers in managerial, administrative, clerical and sales roles (`ManagerAdmin`) is also a significant predictor of two-party preference vote across all six elections, with a higher proportion of people working these jobs increasing Liberal support.
-The magnitude of this effect also seems to increase over the years.
-
-### Household mobility
-
-In each of the six elections, electorates with a higher proportion of people that have recently (in the past five years) moved house are more likely to support the Liberal party, although this effect was marginally insignificant in 2013 (see row 6 in Figure \@ref(fig:partresplot1).
-Having controlled for characteristics of house ownership and rental prices (via the factors `PropertyOwned` and `RentLoan` respectively), this effect is somewhat surprising.
-
-### Relationships
-
-De facto relationships, but not marriages, are found to be an important (and significant) predictor of the two-party preferred vote in all six elections, with more de facto relationships associated with higher support for the Labor party.
-The proportion of individuals who are married however, is insignificant (not shown).
-
-### Age
-
-Regions comprising more older people are often believed to be more conservative, and indeed it found that electorates with a higher median age are more likely to support the Liberal party --- although this effect is only significant in 2007 and 2010 (see row 2 in Figure \@ref(fig:partresplot2)).
-
-### Education
-
-Since 2007, electorates with higher education levels are associated with supporting the Labor party, although this effect is only significant in 2016.
-Before 2007, education has an almost zero effect (see row 3 in Figure \@ref(fig:partresplot1)).
-
-### Diversity
-
-Larger migrant populations from Asia, the Middle East, South-Eastern Europe, the United Kingdom and elsewhere, are either associated with Labor support, or have no effect.
-Of these areas, only South-Eastern European populations appear significant in each election, with the proportion of Asian migrants also being significant in 2010.
-Speaking other languages (aside from English) however, appears to have a far stronger effect, as observed through the `OtherLanguage` variable.
-Electorates with more diverse speech are associated with higher support for the Liberal party from 2004 onwards, with this effect being significant in 2007, 2010 and 2016.
-Furthermore, of the variables relating to religion, only Judaism shows a consistent effect, with electorates with relatively large Jewish populations more likely to vote Liberal.
 
@@
 
+```{r partresplot1, fig.cap = "Partial residual plots by election year for a selection of predictors.
+Linear model with 95\\% confidence bands is overlaid.
+Most predictors have a positive relationship: the larger the value the more likely the electorate preferences the Coalition.
+The relationship is relatively robust over time, with the exception of Unemployment.", fig.height = 8.5, fig.width = 6, warning = FALSE, fig.pos = 'h'}
+(p1 / p2 / (p3 + ylab("Two-party preferred vote (%)                     ")) / p4 / p5 / p6)
+```
+
+```{r partresplot2, fig.cap = "Partial residual plots by election year for a selection of predictors.
+Linear model with 95\\% confidence bands is overlaid.
+Several predictors have a negative relationship: with larger values indicating the electorate more likely preferences Labor.
+Education exhibits a strong temporal change.", fig.height = 8.5,  fig.width = 6, warning = FALSE, fig.pos = 'h'}
+(p7 / p8 / (p9 + ylab("Two-party preferred vote (%)                     ")) / p10 / p11 / p12)
+```
+
+To investigate the socio-demographics that have a strong effect on the two-party preferred vote, partial residual plots are used and shown in Figures \@ref(fig:partresplot1) and \@ref(fig:partresplot2).
+The partial residuals are the residuals from the fitted model with the estimated effect an individual variable added.
+These show the direction, size and significance of an estimated effect --- the slope of the prediction line matches the estimated coefficient, and the shaded region represents a 95% confidence band.
+If a horizontal line can be drawn through the confidence band, then the effect is insignificant.
+The estimated intercept is also added to the partial residuals for interpretability.
+Plots for each election are faceted to compare the effects over time in Figures \@ref(fig:partresplot1) and \@ref(fig:partresplot2).
+
+It is important here to note the ecological fallacy: insights are being drawn at the electorate level, and cannot be inferred for another disaggregate level (e.g., individual voters).
+
+### Income and unemployment
+
+Typically the Labor party campaigns on more progressive policies, which often include tax reform that adversely affects higher income earners, and more generous social assistance programs.
+Perhaps it is due to these policies that higher income electorates appear more likely to support the Liberal party, as the `Incomes` factor has a positive effect on Liberal preference (see row 1 in Figure \@ref(fig:partresplot1)).
+This effect is significant in every election aside from 2004, where it is only marginally insignificant ($p = 0.0613$).
+Unemployment however, is not as influential.
+In 2001 and 2004, electorates with higher unemployment align with Labor, but over time this shifts towards support for the Liberal party, culminating in a significantly positive effect in 2016.
+
+### Industry and type of work
+
+Electorates with higher proportions of workers in mining, gas, water, agriculture, waste and electricity (grouped as `Extractive` industries) are consistently linked with higher support for the Liberal party, with the magnitude of this effect slightly increasing over the years (see row 3 in Figure \@ref(fig:partresplot1)).
+This is unsurprising, as the Liberal party has close ties with these traditional energy industries, and typically present policies to reduce taxation on energy production.
+Furthermore, electorates with more workers in construction or manufacturing industries (`Transformative`) are also more likely to support the Liberal party (see row 4 in Figure \@ref(fig:partresplot1)).
+
+Similarly, the proportion of workers in managerial, administrative, clerical and sales roles (`ManagerAdmin`) is also a significant predictor of two-party preference vote across all six elections, with a higher proportion of people working these jobs increasing Liberal support.
+The magnitude of this effect also seems to increase over the years.
+
+### Household mobility
+
+In each of the six elections, electorates with a higher proportion of people that have recently (in the past five years) moved house (`DiffAddress`) are more likely to support the Liberal party, although this effect was marginally insignificant in 2013 (see row 6 in Figure \@ref(fig:partresplot1).
+Having controlled for characteristics of house ownership and rental prices (via the factors `PropertyOwned` and `RentLoan` respectively), this effect is somewhat surprising.
+
+### Relationships
+
+De facto relationships, but not marriages, are found to be an important (and significant) predictor of the two-party preferred vote in all six elections, with more de facto relationships associated with higher support for the Labor party.
+The proportion of individuals who are married however, is insignificant (not shown).
+
+### Age
+
+Regions comprising more older people are often believed to be more conservative, and indeed it found that electorates with a higher median age are more likely to support the Liberal party --- although this effect is only significant in 2007 and 2010 (see row 2 in Figure \@ref(fig:partresplot2)).
+
+### Education
+
+Since 2007, electorates with higher education levels are associated with supporting the Labor party, although this effect is only significant in 2016.
+Before 2007, education has an almost zero effect (see row 3 in Figure \@ref(fig:partresplot1)).
+
+### Diversity
+
+Larger migrant populations from Asia, the Middle East, South-Eastern Europe, the United Kingdom and elsewhere, are either associated with Labor support, or have no effect.
+Of these areas, only South-Eastern European populations appear significant in each election, with the proportion of Asian migrants also being significant in 2010.
+Speaking other languages (aside from English) however, appears to have a far stronger effect, as observed through the `OtherLanguage` variable.
+Electorates with more diverse speech are associated with higher support for the Liberal party from 2004 onwards, with this effect being significant in 2007, 2010 and 2016.
+Furthermore, of the variables relating to religion, only Judaism shows a consistent effect, with electorates with relatively large Jewish populations more likely to vote Liberal.
+
 ### A note on similar variables
@@
 It is often hypothesized that states have systematic differences that cause their electorates to vote differently.
-Boxplots of residuals grouped by state (Figure \@ref(fig:resstate)) reveal that only Tasmania, the Australian Capital Territory and the Northern Territory appear to have a state-specific effect not captured by the models.
+Boxplots of residuals grouped by state (Figure \@ref(fig:resstate)) reveal that the data reflects this -- there appears to be a state-specific effect not captured by the models.
 Tasmania and the Australian Capital Territory appear to have a bias towards Labor, whereas the Northern Territory tends towards voting Liberal.
@@
 
-```{r resstate, fig.cap = "Boxplot of residuals by state with jittered points.", fig.asp=0.6, out.width="70%", fig.pos = 'h', out.extra = ''}
+```{r resstate, fig.cap = "Boxplot of residuals by state with jittered points.
+States ordered by median residual.
+A state-specific bias not captured by the model is evident.", fig.asp=0.6, out.width="70%", fig.pos = 'h', out.extra = ''}
 all_resids %>%
-  ggplot(aes(x = StateAb, y = Residuals)) +
+  ggplot(aes(x = fct_reorder(StateAb, Residuals), y = Residuals)) +
   geom_hline(aes(yintercept = 0), col = "grey50") +

### 2019-04-13 draft-paper.tex
tweaked a few figures and language in the results section

@@
             pdftitle={Spatial modelling of the two-party preferred vote in Australian federal elections: 2001--2016},
-            pdfkeywords={federal election, Census, Australia, spatial modelling, imputation, socio-demographics, electorates, R, eechidna},
+            pdfkeywords={federal election, Census, Australia, spatial modelling, imputation,
+socio-demographics, electorates, R, eechidna},
             pdfborder={0 0 0},
@@
 
-\title{Spatial modelling of the two-party preferred vote in Australian federal elections: 2001--2016}
+\title{Spatial modelling of the two-party preferred vote in Australian federal
+elections: 2001--2016}
 
@@
 
-\author{Jeremy~Forbes, Di~Cook, Rob J~Hyndman}
+\author{Jeremy~Forbes, Dianne~Cook, Rob J~Hyndman}
 \addresses{\textbf{Jeremy Forbes}\newline
@@
 \newline{Email: \href{mailto:jeremy.forbes@monash.edu}{\nolinkurl{jeremy.forbes@monash.edu}}}\newline Corresponding author\\[1cm]
-\textbf{Di Cook}\newline
+\textbf{Dianne Cook}\newline
 Monash University\newline Melbourne, Australia
@@
 \begin{abstract}
-We examine the relationships between electoral socio-demographic characteristics and two-party preference in the six Australian federal elections held between 2001 to 2016.
-Socio-demographic information is derived from the Australian Census, which occurs every five years.
-Since a Census is not directly available for each election, spatio-temporal imputation is employed to estimate Census data for the electorates at the time of each election.
-This accounts for both spatial and temporal changes in electoral characteristics between Censuses.
-To capture any spatial heterogeneity, a spatial error model is estimated for each election, which incorporates a spatially structured random effect vector that can be thought of as the unobserved political climate in each electorate.
-Over time, the impact of most socio-demographic characteristics that affect electoral two-party preference do not vary, with industry of work, incomes, household mobility and de facto relationships having strong effects in each of the six elections.
+We examine the relationships between electoral socio-demographic
+characteristics and two-party preference in the six Australian federal
+elections held between 2001 to 2016. Socio-demographic information is
+derived from the Australian Census, which occurs every five years. Since
+a Census is not directly available for each election, spatio-temporal
+imputation is employed to estimate Census data for the electorates at
+the time of each election. This accounts for both spatial and temporal
+changes in electoral characteristics between Censuses. To capture any
+spatial heterogeneity, a spatial error model is estimated for each
+election, which incorporates a spatially structured random effect vector
+that can be thought of as the unobserved political climate in each
+electorate. Over time, the impact of most socio-demographic
+characteristics that affect electoral two-party preference do not vary,
+with industry of work, incomes, household mobility and de facto
+relationships having strong effects in each of the six elections.
 Education and unemployment are amongst those that have varying effects.
-It is also found that between 2004 and 2013, the spatial effect was insignificant, meaning that electorates voted effectively independently.
-All data featured in this study has been contributed to the \texttt{eechidna} \texttt{R} package (available on CRAN).
+It is also found that between 2004 and 2013, the spatial effect was
+insignificant, meaning that electorates voted effectively independently.
+All data featured in this study has been contributed to the
+\texttt{eechidna} \texttt{R} package (available on CRAN).
 \end{abstract}
 \begin{keywords}
-federal election, Census, Australia, spatial modelling, imputation, socio-demographics, electorates, R, eechidna
+federal election, Census, Australia, spatial modelling, imputation,
+socio-demographics, electorates, R, eechidna
 \end{keywords}
 
-\hypertarget{intro}{%
-\section{Introduction}\label{intro}}
+\section{Introduction}\label{intro}
 
-Australia has changed in many ways over the last two decades.
-Rising house prices, country-wide improvements in education, an ageing population, and a decline in religious affiliation, are just a few facets of the country's evolving socio-demographic characteristics.
-At the same time, political power has moved back and forth between the two major parties.
-In the 2007 and 2010 federal elections, the Australian Labor Party (Labor) was victorious, whereas the 2001, 2004, 2013 and 2016 elections were won by the Liberal National coalition (Liberal).
-The two-party preferred vote, a measure of support between these two parties, fluctuated between 47.3\% and 53.5\% (in favour of the Liberal party) over this period.
-This study explores how electoral socio-demographic characteristics relate to two-party preference, and whether their effects have changed over time.
-Electoral socio-demographics are derived from the Australian Census, and vote counts are obtained from Australian federal elections.
+Australia has changed in many ways over the last two decades. Rising
+house prices, country-wide improvements in education, an ageing
+population, and a decline in religious affiliation, are just a few
+facets of the country's evolving socio-demographic characteristics. At
+the same time, political power has moved back and forth between the two
+major parties. In the 2007 and 2010 federal elections, the Australian
+Labor Party (Labor) was victorious, whereas the 2001, 2004, 2013 and
+2016 elections were won by the Liberal National coalition (Liberal). The
+two-party preferred vote, a measure of support between these two
+parties, fluctuated between 47.3\% and 53.5\% (in favour of the Liberal
+party) over this period. This study explores how electoral
+socio-demographic characteristics relate to two-party preference, and
+whether their effects have changed over time. Electoral
+socio-demographics are derived from the Australian Census, and vote
+counts are obtained from Australian federal elections.
 
-Joining these two data sources is problematic as there in an inherent asynchronicity in the two types of events.
-A Census is conducted by the Australian Bureau of Statistics (ABS) every five years, whereas federal elections, conducted by the Australian Electoral Commission (AEC), usually occur every three years or so.
-The first problem addressed is that of constructing appropriate Census data for the 2004, 2007, 2010 and 2013 elections --- election years in which a Census does not occur.
-The predominant approach in previous studies was to join voting outcomes to the nearest Census, without accounting for any temporal differences \autocites[see][]{DavisStimson98}{Stimson06}{Liao09}{Stimson09}.
-Furthermore, electoral boundaries change regularly, so spatial discrepancies also arise when matching with electoral data.
-To obtain appropriate ``Census-like'' data for these four elections, electoral socio-demographics are constructed using a spatio-temporal imputation that combines areal interpolation \autocite{Goodchild1993} and linear time-interpolation.
-Collecting and wrangling the raw data, along with the imputation process, are detailed in Section \ref{data}.
-All data and associated documentation relating to this procedure are available in the \texttt{eechidna} \texttt{R} package \autocite{eechidna}, providing a resource for future analysis.
+Joining these two data sources is problematic as there in an inherent
+asynchronicity in the two types of events. A Census is conducted by the
+Australian Bureau of Statistics (ABS) every five years, whereas federal
+elections, conducted by the Australian Electoral Commission (AEC),
+usually occur every three years or so. The first problem addressed is
+that of constructing appropriate Census data for the 2004, 2007, 2010
+and 2013 elections --- election years in which a Census does not occur.
+The predominant approach in previous studies was to join voting outcomes
+to the nearest Census, without accounting for any temporal differences
+\autocites[see][]{DavisStimson98}{Stimson06}{Liao09}{Stimson09}.
+Furthermore, electoral boundaries change regularly, so spatial
+discrepancies also arise when matching with electoral data. To obtain
+appropriate ``Census-like'' data for these four elections, electoral
+socio-demographics are constructed using a spatio-temporal imputation
+that combines areal interpolation \autocite{Goodchild1993} and linear
+time-interpolation. Collecting and wrangling the raw data, along with
+the imputation process, are detailed in Section \ref{data}. All data and
+associated documentation relating to this procedure are available in the
+\texttt{eechidna} \texttt{R} package \autocite{eechidna}, providing a
+resource for future analysis.
 
-Previous work on modelling Australian federal elections has found that aggregate socio-demographics are relatively good predictors of voting outcomes.
-\textcite{Forrest01} used multiple regression to model the Liberal and Labor primary vote for polling booths in the Farrer electorate in 1998 as a function of Census variables from 1996.
-\textcite{Stimson06}, \textcite{Stimson09} and \textcite{Stimson12} used principal component analysis of polling booths in the 2001, 2004 and 2007 elections respectively, also finding that socio-demographic characteristics of polling booths are linked to their two-party preferred vote.
-In contrast, \textcite{Stimson09} models the polling booth swing vote (change in the two-party preferred vote) in the 2007 election, finding that little of the swing vote can be explained by Census data.
-Instead of analyzing a single election in isolation, this paper employs a consistent model framework across six elections so that temporal changes in the effects of socio-demographics can be observed, where each federal elections is modelled with a cross-sectional dataset.
-The use of a regression framework to examine these socio-political relationships over time is seemingly absent from previous Australian studies.
-It also appears that no study has attempted any type of statistical analysis of socio-demographics in conjunction with voter behaviour in Australia since 2007, making this paper distinctly different from previous work.
+Previous work on modelling Australian federal elections has found that
+aggregate socio-demographics are relatively good predictors of voting
+outcomes. \textcite{Forrest01} used multiple regression to model the
+Liberal and Labor primary vote for polling booths in the Farrer
+electorate in 1998 as a function of Census variables from 1996.
+\textcite{Stimson06}, \textcite{Stimson09} and \textcite{Stimson12} used
+principal component analysis of polling booths in the 2001, 2004 and
+2007 elections respectively, also finding that socio-demographic
+characteristics of polling booths are linked to their two-party
+preferred vote. In contrast, \textcite{Stimson09} models the polling
+booth swing vote (change in the two-party preferred vote) in the 2007
+election, finding that little of the swing vote can be explained by
+Census data. Instead of analyzing a single election in isolation, this
+paper employs a consistent model framework across six elections so that
+temporal changes in the effects of socio-demographics can be observed,
+where each federal elections is modelled with a cross-sectional dataset.
+The use of a regression framework to examine these socio-political
+relationships over time is seemingly absent from previous Australian
+studies. It also appears that no study has attempted any type of
+statistical analysis of socio-demographics in conjunction with voter
+behaviour in Australia since 2007, making this paper distinctly
+different from previous work.
 
-The cross-sectional dataset for each election used here consists of the two-party preferred vote (as the response variable), and a set of common socio-demographic variables (as the explanatory variables) that characterize each electorate.
-To prepare these datasets, socio-demographic variables are first standardized, and then a principal component analysis is used to group variables into ``factors''.
-To account for the inherent spatial structure of the data, a spatial error model is then estimated for each election.
-These steps are discussed in Section \ref{modelling}.
-In Section \ref{results} inference is conducted to see which effects are significant and how these effects change over time.
+The cross-sectional dataset for each election used here consists of the
+two-party preferred vote (as the response variable), and a set of common
+socio-demographic variables (as the explanatory variables) that
+characterize each electorate. To prepare these datasets,
+socio-demographic variables are first standardized, and then a principal
+component analysis is used to group variables into ``factors''. To
+account for the inherent spatial structure of the data, a spatial error
+model is then estimated for each election. These steps are discussed in
+Section \ref{modelling}. In Section \ref{results} inference is conducted
+to see which effects are significant and how these effects change over
+time.
 
-Section \ref{conclusion} concludes, followed by two supplementary sections.
+Section \ref{conclusion} concludes, followed by two supplementary
+sections.
 
-\hypertarget{data}{%
-\section{Data collection, wrangling and imputation}\label{data}}
+\section{Data collection, wrangling and imputation}\label{data}
 
-\hypertarget{collecting-the-data}{%
-\subsection{Collecting the data}\label{collecting-the-data}}
+\subsection{Collecting the data}\label{collecting-the-data}
 
-The voting outcome of interest is the electoral two-party preferred vote, which is provided by the Australian Electoral Commission (AEC) for the 2001, 2004, 2007, 2010, 2013 and 2016 elections via the AEC Tally Room.
-The AEC divides Australia into 150 regions, called electorates, with each corresponding to a single seat in the House of Representatives.
-Voting is compulsory in Australia, and each voter assigns a numbered preference to each available candidate in their electorate.
-The two-party preferred vote is determined by a tally of these preferences where, by convention, only the ranks of the Labor and Liberal candidates are considered.
-This is recorded as a percentage preference in favour of the Liberal party.
+The voting outcome of interest is the electoral two-party preferred
+vote, which is provided by the Australian Electoral Commission (AEC) for
+the 2001, 2004, 2007, 2010, 2013 and 2016 elections via the AEC Tally
+Room. The AEC divides Australia into 150 regions, called electorates,
+with each corresponding to a single seat in the House of
+Representatives. Voting is compulsory in Australia, and each voter
+assigns a numbered preference to each available candidate in their
+electorate. The two-party preferred vote is determined by a tally of
+these preferences where, by convention, only the ranks of the Labor and
+Liberal candidates are considered. This is recorded as a percentage
+preference in favour of the Liberal party.
 
-Socio-demographic variables were derived from the Australian Census of Population and Housing (Census), which is a survey of every household in Australia, recording information such as age, gender, ethnicity, education level and income.
-There have been four Censuses so far in the 21st century, conducted in 2001, 2006, 2011 and 2016.
-The Australian Bureau of Statistics (ABS) conducts the Census and publishes aggregated information.
-The ABS uses electoral boundaries as defined by the AEC at the time of each Census, which may not match those in place at the subsequent and previous elections.
-From the available Census information aggregated at the electorate level, 65 socio-demographic variables were defined for each of the electorates to be used in the analysis.
+Socio-demographic variables were derived from the Australian Census of
+Population and Housing (Census), which is a survey of every household in
+Australia, recording information such as age, gender, ethnicity,
+education level and income. There have been four Censuses so far in the
+21st century, conducted in 2001, 2006, 2011 and 2016. The Australian
+Bureau of Statistics (ABS) conducts the Census and publishes aggregated
+information. The ABS uses electoral boundaries as defined by the AEC at
+the time of each Census, which may not match those in place at the
+subsequent and previous elections. From the available Census information
+aggregated at the electorate level, 65 socio-demographic variables were
+defined for each of the electorates to be used in the analysis.
 
-Raw data was sourced online from the AEC and ABS websites in \texttt{.csv} and \texttt{.xlsx} files.
-The formats of these files differ over the years, making extracting the appropriate information a big task.
-The functions available in the \texttt{dplyr} \autocite{dplyr} and \texttt{readxl} \autocite{readxl} \texttt{R} packages are particularly useful, as they provide fast consistent tools for data manipulation and functions to import \texttt{.xlsx} files (respectively).
-The 2001 and 2006 Census data are published in a format where the information for each electorate is held in a separate document making it difficult to use the \texttt{dplyr} tools.
-Instead, cells have to be selected from each individual file to construct the desired variables.
-All scripts required for the data wrangling process can be found in the github repository for the \texttt{eechidna} \texttt{R} package \autocite{eechidna}, along with the raw data.
-The \texttt{eechidna} package makes this study entirely reproducible and provides a resource to help wrangle data for future Censuses and elections, when they become available.
+Raw data was sourced online from the AEC and ABS websites in
+\texttt{.csv} and \texttt{.xlsx} files. The formats of these files
+differ over the years, making extracting the appropriate information a
+big task. The functions available in the \texttt{dplyr} \autocite{dplyr}
+and \texttt{readxl} \autocite{readxl} \texttt{R} packages are
+particularly useful, as they provide fast consistent tools for data
+manipulation and functions to import \texttt{.xlsx} files
+(respectively). The 2001 and 2006 Census data are published in a format
+where the information for each electorate is held in a separate document
+making it difficult to use the \texttt{dplyr} tools. Instead, cells have
+to be selected from each individual file to construct the desired
+variables. All scripts required for the data wrangling process can be
+found in the github repository for the \texttt{eechidna} \texttt{R}
+package \autocite{eechidna}, along with the raw data. The
+\texttt{eechidna} package makes this study entirely reproducible and
+provides a resource to help wrangle data for future Censuses and
+elections, when they become available.
 
-\hypertarget{joining-census-and-election-data}{%
-\subsection{Joining Census and election data}\label{joining-census-and-election-data}}
+\subsection{Joining Census and election
+data}\label{joining-census-and-election-data}
 
-\hypertarget{differences-between-census-and-election-data}{%
-\subsubsection{Differences between Census and election data}\label{differences-between-census-and-election-data}}
+\subsubsection{Differences between Census and election
+data}\label{differences-between-census-and-election-data}
 
-Between 2001 and 2016 there were six elections and four Censuses (see Figure \ref{fig:timeline}).
-Electoral boundaries are redistributed regularly by the AEC, meaning that only in the years where both a Census and election occur are all boundaries likely to match --- the case for the 2001 and 2016 elections.
-Therefore, for the four elections between 2004 and 2013, both temporal and spatial differences in electorates need to be accounted for when joining the electoral two-party preferred vote with Census data.
-For these elections a spatio-temporal imputation method was employed to obtain electoral socio-demographics.
-This method uses Census information from both before and after the election of interest.
+Between 2001 and 2016 there were six elections and four Censuses (see
+Figure \ref{fig:timeline}). Electoral boundaries are redistributed
+regularly by the AEC, meaning that only in the years where both a Census
+and election occur are all boundaries likely to match --- the case for
+the 2001 and 2016 elections. Therefore, for the four elections between
+2004 and 2013, both temporal and spatial differences in electorates need
+to be accounted for when joining the electoral two-party preferred vote
+with Census data. For these elections a spatio-temporal imputation
+method was employed to obtain electoral socio-demographics. This method
+uses Census information from both before and after the election of
+interest.
 
@@
 
-\hypertarget{spatio-temporal-imputation}{%
-\subsubsection{Spatio-temporal imputation}\label{spatio-temporal-imputation}}
+\subsubsection{Spatio-temporal
+imputation}\label{spatio-temporal-imputation}
 
-For each election, neighbouring Census information has to be combined in some way so that it represents the boundaries in place at the time of the election.
-This is done by taking the electoral boundaries and imputing the corresponding socio-demographic characteristics for each of neighbouring Censuses, thereby addressing the spatial aspect.
-Next, to deal with the temporal component, characteristics at the time of the election are constructed using linear interpolation between the spatially imputed neighbouring Census variables.
+For each election, neighbouring Census information has to be combined in
+some way so that it represents the boundaries in place at the time of
+the election. This is done by taking the electoral boundaries and
+imputing the corresponding socio-demographic characteristics for each of
+neighbouring Censuses, thereby addressing the spatial aspect. Next, to
+deal with the temporal component, characteristics at the time of the
+election are constructed using linear interpolation between the
+spatially imputed neighbouring Census variables.
 
-To account for spatial differences, the piecewise approximation method in \textcite{Goodchild1993} is adopted.
-Consider a map of source zones \(s = 1,\dots,S\), for which socio-demographic information is available, and a set of target zones \(t = 1,\dots,T\) for which information is to be imputed.
-This is described in the context of a single election, and a single neighbouring Census.
+To account for spatial differences, the piecewise approximation method
+in \textcite{Goodchild1993} is adopted. Consider a map of source zones
+\(s = 1,\dots,S\), for which socio-demographic information is available,
+and a set of target zones \(t = 1,\dots,T\) for which information is to
+be imputed. This is described in the context of a single election, and a
+single neighbouring Census.
 
-Let the map of electoral boundaries at the time of a Census define the source zones, and let the boundaries at the time of the election be the target zones.
-Denote the area of intersection between a source zone \(s\) and a target zone \(t\) as \(A_{s,t}\).
-Additionally, let the population of the source zone \(s\) be \(U_s\) and the population of intersection between source zone \(s\) and target zone \(t\) be \(P_{s,t}\).
-The estimated population of intersection is given by
-\[
+Let the map of electoral boundaries at the time of a Census define the
+source zones, and let the boundaries at the time of the election be the
+target zones. Denote the area of intersection between a source zone
+\(s\) and a target zone \(t\) as \(A_{s,t}\). Additionally, let the
+population of the source zone \(s\) be \(U_s\) and the population of
+intersection between source zone \(s\) and target zone \(t\) be
+\(P_{s,t}\). The estimated population of intersection is given by \[
   \hat{P}_{s,t} = \frac{U_s*A_{s,t}}{\sum_{t=1}^T A_{s,t}},
   \quad\text{for all $s=1,\dots,S$ and $t=1,\dots,T$}.
-\]
-Note that this estimator implicitly assumes that populations are uniformly distributed within each source zone.
+\] Note that this estimator implicitly assumes that populations are
+uniformly distributed within each source zone.
 
-In order to calculate socio-demographic information for each of the target zones, a weighted average is taken using the estimated intersection populations as weights.
-Denote a given Census variable for the target zone by \(C_t\), and the same Census variable for the source zone as \(D_s\).
-Then, estimate \(C_t\) using
-\[
+In order to calculate socio-demographic information for each of the
+target zones, a weighted average is taken using the estimated
+intersection populations as weights. Denote a given Census variable for
+the target zone by \(C_t\), and the same Census variable for the source
+zone as \(D_s\). Then, estimate \(C_t\) using \[
   \hat{C}_t = \frac{\sum_{s=1}^{S}{D_s*\hat{P}_{s,t}}}%
@@
   \quad\text{for each $t=1,\dots,T$}.
-\]
-This concludes the spatial imputation of the socio-demographic characteristics for one target zone (a single electoral boundary), at the time of only one of the neighbouring Censuses.
-This process is repeated for all of the target zones, and then for the other neighbouring Census.
+\] This concludes the spatial imputation of the socio-demographic
+characteristics for one target zone (a single electoral boundary), at
+the time of only one of the neighbouring Censuses. This process is
+repeated for all of the target zones, and then for the other
+neighbouring Census.
 
@@
 
-To account for temporal changes, linear interpolation is used between Census years to get the final estimate of a Census variable for the target zone in the election year.
-Let \(y_1\) be the year of the Census preceding an election, let \(y_2\) be the year of the election, and \(y_3\) be the year of the Census that follows.
-Add this year subscript to the Census variable estimate \(\hat{C}_t\), resulting in \(\hat{C}_{t,y}\).
-Linear interpolating between these Census years results an imputed value for the election year, given by
-\[
+To account for temporal changes, linear interpolation is used between
+Census years to get the final estimate of a Census variable for the
+target zone in the election year. Let \(y_1\) be the year of the Census
+preceding an election, let \(y_2\) be the year of the election, and
+\(y_3\) be the year of the Census that follows. Add this year subscript
+to the Census variable estimate \(\hat{C}_t\), resulting in
+\(\hat{C}_{t,y}\). Linear interpolating between these Census years
+results an imputed value for the election year, given by \[
   \hat{C}_{t,y_2} = \frac{y_3-y_2}{y_3-y_1} * \hat{C}_{t,y_1} +
     \frac{y_2-y_1}{y_3-y_1} * \hat{C}_{t,y_3}.
-\]
-Implicitly this assumes that population characteristics change in a linear manner over time.
+\] Implicitly this assumes that population characteristics change in a
+linear manner over time.
 
-\hypertarget{an-illustration-of-the-spatio-temporal-imputation}{%
-\subsubsection*{An illustration of the spatio-temporal imputation}\label{an-illustration-of-the-spatio-temporal-imputation}}
-\addcontentsline{toc}{subsubsection}{An illustration of the spatio-temporal imputation}
+\subsubsection*{An illustration of the spatio-temporal
+imputation}\label{an-illustration-of-the-spatio-temporal-imputation}
+\addcontentsline{toc}{subsubsection}{An illustration of the
+spatio-temporal imputation}
 
-Census data is publicly available at different levels of aggregation, ranging from SA1 (over 50,000 zones) to electoral divisions (150 zones).
-For this study, electoral divisions are used as source zones, and the imputation method is applied to produce socio-demographic variables for each of the 2004, 2007, 2010 and 2013 elections.
-As mentioned earlier, there is no need to impute socio-demographic variables for the 2001 and 2016 elections.
-To illustrate the method, consider the imputation of socio-demographic variables for the electorate of Hume in New South Wales (NSW) at the time of the 2013 federal election.
-The boundaries shown in Figure \ref{fig:hume13} define all target zones in NSW for 2013, with the target zone of interest (Hume) shaded purple.
+Census data is publicly available at different levels of aggregation,
+ranging from SA1 (over 50,000 zones) to electoral divisions (150 zones).
+For this study, electoral divisions are used as source zones, and the
+imputation method is applied to produce socio-demographic variables for
+each of the 2004, 2007, 2010 and 2013 elections. As mentioned earlier,
+there is no need to impute socio-demographic variables for the 2001 and
+2016 elections. To illustrate the method, consider the imputation of
+socio-demographic variables for the electorate of Hume in New South
+Wales (NSW) at the time of the 2013 federal election. The boundaries
+shown in Figure \ref{fig:hume13} define all target zones in NSW for
+2013, with the target zone of interest (Hume) shaded purple.
 
@@
 
-The corresponding source zones from the 2016 Census are shown in Figure \ref{fig:hume16}.
-As can be seen, the Hume boundary from the 2013 election (shaded purple) does not perfectly match any of the source zones.
+The corresponding source zones from the 2016 Census are shown in Figure
+\ref{fig:hume16}. As can be seen, the Hume boundary from the 2013
+election (shaded purple) does not perfectly match any of the source
+zones.
 
-There are many source zones from the 2016 Census that intersect with this purple region, including the divisions of Riverina, Eden-Monaro and Hume, along with smaller intersecting areas with Fenner, Calare, Gilmore and Whitlam.
-The proportion of each source zone that overlaps with the purple region is calculated, and used to obtain the intersecting populations \(\hat{P}_{s,t}\).
+There are many source zones from the 2016 Census that intersect with
+this purple region, including the divisions of Riverina, Eden-Monaro and
+Hume, along with smaller intersecting areas with Fenner, Calare, Gilmore
+and Whitlam. The proportion of each source zone that overlaps with the
+purple region is calculated, and used to obtain the intersecting
+populations \(\hat{P}_{s,t}\).
 
@@
 
-Now consider the socio-demographic variable \emph{AusCitizen}, the proportion of people in the region who are Australian citizens.
-A weighted average of \emph{AusCitizen}, with the allocated population from each source zone as weights, yields \(\hat{C}_{\text{Hume},2016} = 89.65\%\).
-Repeating this process using the 2011 Census yields \(\hat{C}_{\text{Hume},2011} = 91.00\%\).
-Finally, linear interpolation between 2011 and 2016 yields the 2013 estimate:
-\[
+Now consider the socio-demographic variable \emph{AusCitizen}, the
+proportion of people in the region who are Australian citizens. A
+weighted average of \emph{AusCitizen}, with the allocated population
+from each source zone as weights, yields
+\(\hat{C}_{\text{Hume},2016} = 89.65\%\). Repeating this process using
+the 2011 Census yields \(\hat{C}_{\text{Hume},2011} = 91.00\%\).
+Finally, linear interpolation between 2011 and 2016 yields the 2013
+estimate: \[
   \hat{C}_{\text{Hume},2013}
@@
     =  90.46\%.
-\]
-This is done for each of the 65 socio-demographic variables, and is repeated for each of the 149 remaining target zones corresponding with 2013 electorates.
+\] This is done for each of the 65 socio-demographic variables, and is
+repeated for each of the 149 remaining target zones corresponding with
+2013 electorates.
 
-\hypertarget{modelling}{%
-\section{Modelling}\label{modelling}}
+\section{Modelling}\label{modelling}
 
-Following this process, electoral socio-demographic variables are available for each of the six elections and can be joined with their corresponding two-party preferred votes.
-Before choosing an appropriate model, two issues with the socio-demographic variables need to be addressed.
-First, variable scales change over the years, making it important to standardize variables.
-Second, many variables represent similar information and where appropriate, should be combined in some way.
-Principal component analysis is used to identify variables covary, from which intuitive groupings are selected to be combined into a single variable.
-This also reduces the dimension of the data.
-After these steps, a model specification is chosen.
+Following this process, electoral socio-demographic variables are
+available for each of the six elections and can be joined with their
+corresponding two-party preferred votes. Before choosing an appropriate
+model, two issues with the socio-demographic variables need to be
+addressed. First, variable scales change over the years, making it
+important to standardize variables. Second, many variables represent
+similar information and where appropriate, should be combined in some
+way. Principal component analysis is used to identify variables covary,
+from which intuitive groupings are selected to be combined into a single
+variable. This also reduces the dimension of the data. After these
+steps, a model specification is chosen.
 
-\hypertarget{standardizing-variables}{%
-\subsection{Standardizing variables}\label{standardizing-variables}}
+\subsection{Standardizing variables}\label{standardizing-variables}
 
-Many of the socio-demographic variables have changing scales over the years.
-For example, inflation-adjusted median rental prices increased across almost all electorates, with median rent of 200 dollars per week placing an electorate in the 90th percentile in 2001, but only the 30th percentile in 2016.
-In order for socio-demographic effects to be comparable across years, all explanatory variables are standardized to have mean zero and variance one within each election year.
-By standardizing, each variable is reported as a relative measure compared to all other electorates in the same year.
+Many of the socio-demographic variables have changing scales over the
+years. For example, inflation-adjusted median rental prices increased
+across almost all electorates, with median rent of 200 dollars per week
+placing an electorate in the 90th percentile in 2001, but only the 30th
+percentile in 2016. In order for socio-demographic effects to be
+comparable across years, all explanatory variables are standardized to
+have mean zero and variance one within each election year. By
+standardizing, each variable is reported as a relative measure compared
+to all other electorates in the same year.
 
-\hypertarget{creating-factors}{%
-\subsection{Creating factors}\label{creating-factors}}
+\subsection{Creating factors}\label{creating-factors}
 
-There are only \(N = 150\) observations (electorates) in each election and \(p = 65\) socio-demographic variables in each cross-section, with many variables represent similar information about an electorate.
-Any model that uses all variables would face serious problems with multi-collinearity and over-fitting, which would likely lead to erroneous conclusions regarding variable significance.
-To address this, groups of variables that represent similar information are combined into ``factors''\footnote{A preliminary step involved removing all age bands, because age is represented by median age, and to remove variables relating to particular denominations of Christianity.}.
+There are only \(N = 150\) observations (electorates) in each election
+and \(p = 65\) socio-demographic variables in each cross-section, with
+many variables represent similar information about an electorate. Any
+model that uses all variables would face serious problems with
+multi-collinearity and over-fitting, which would likely lead to
+erroneous conclusions regarding variable significance. To address this,
+groups of variables that represent similar information are combined into
+``factors''\footnote{A preliminary step involved removing all age bands,
+  because age is represented by median age, and to remove variables
+  relating to particular denominations of Christianity.}.
 
-A factor is created from a group of variables if there is an intuitive reason as to why they should represent similar information and if there is evidence to suggest that they covary.
-For example, a potential group would be variables relating to electoral incomes --- median family, household and personal incomes.
-To determine which variables covary, principal component analysis is used on a combined dataset of socio-demographic variables from all six elections.
-It is appropriate to compute principal components in this way because when computed separately for each election, scree plots level off after four components and the loadings of the first four components are similar across the elections.
+A factor is created from a group of variables if there is an intuitive
+reason as to why they should represent similar information and if there
+is evidence to suggest that they covary. For example, a potential group
+would be variables relating to electoral incomes --- median family,
+household and personal incomes. To determine which variables covary,
+principal component analysis is used on a combined dataset of
+socio-demographic variables from all six elections. It is appropriate to
+compute principal components in this way because when computed
+separately for each election, scree plots level off after four
+components and the loadings of the first four components are similar
+across the elections.
 
-Only the first four principal components from the combined dataset are considered, as the scree plot corresponding to the combined dataset levels off after the fourth component.
-Variables that have a large loading in a particular component are deemed to covary, with a loading with magnitude greater than 0.15 being considered large.
-Six factors are created using this criteria.
-These are:
-\texttt{Incomes} (median personal, household and family incomes);
-\texttt{Unemployment} (unemployment and labor force participation rates);
-\texttt{ProportyOwned} (rates of housing ownership, mortgages, renting and government housing);
-\texttt{RentLoanPrice} (median rental and loan repayments);
-\texttt{FamHouseSize} (average household size, ratio of people to families and household makeup (single person, couple with kids and couple without kids); and
-\texttt{Education} (high school and university qualifications, jobs requiring higher levels of education as well as vocational course completions and jobs that do not require higher education levels, such as laborer or tradesperson).
-For each of these groups, variables with positive loadings are added and those with negative loadings are subtracted to create a factor.
-After computing these sums, each factor is standardized to have mean zero and variance one, within each election.
+Only the first four principal components from the combined dataset are
+considered, as the scree plot corresponding to the combined dataset
+levels off after the fourth component. Variables that have a large
+loading in a particular component are deemed to covary, with a loading
+with magnitude greater than 0.15 being considered large. Six factors are
+created using this criteria. These are: \texttt{Incomes} (median
+personal, household and family incomes); \texttt{Unemployment}
+(unemployment and labor force participation rates);
+\texttt{ProportyOwned} (rates of housing ownership, mortgages, renting
+and government housing); \texttt{RentLoanPrice} (median rental and loan
+repayments); \texttt{FamHouseSize} (average household size, ratio of
+people to families and household makeup (single person, couple with kids
+and couple without kids); and \texttt{Education} (high school and
+university qualifications, jobs requiring higher levels of education as
+well as vocational course completions and jobs that do not require
+higher education levels, such as laborer or tradesperson). For each of
+these groups, variables with positive loadings are added and those with
+negative loadings are subtracted to create a factor. After computing
+these sums, each factor is standardized to have mean zero and variance
+one, within each election.
 
-There are \(p=30\) variables in the resultant predictor set, with all of these used in the regression for each election.
+There are \(p=30\) variables in the resultant predictor set, with all of
+these used in the regression for each election.
 
-\hypertarget{regression-incorporating-spatially-dependent-errors}{%
-\subsection{Regression incorporating spatially dependent errors}\label{regression-incorporating-spatially-dependent-errors}}
+\subsection{Regression incorporating spatially dependent
+errors}\label{regression-incorporating-spatially-dependent-errors}
 
-An identical model specification is used for each of the six elections, with each election modelled separately.
-This allows the socio-demographic effects to be estimated separately for each election year, facilitating analysis of temporal changes in variable effects.
-This approach is preferable to using a single longitudinal model because it avoids any concerns of undue bias stemming from incorrectly imposed time-varying restrictions on any variable.
-Without such restrictions, a pooled cross-sectional model does not yield any distinct advantage over separate cross-sections.
-The panel approach is avoided because of how frequently electoral boundaries change, noting that electorates with the same name across elections are not guaranteed to represent the same geographical region.
-Therefore any fixed or random effects models would be difficult to estimate without implementing consistent boundaries, which would require further imputation.
+An identical model specification is used for each of the six elections,
+with each election modelled separately. This allows the
+socio-demographic effects to be estimated separately for each election
+year, facilitating analysis of temporal changes in variable effects.
+This approach is preferable to using a single longitudinal model because
+it avoids any concerns of undue bias stemming from incorrectly imposed
+time-varying restrictions on any variable. Without such restrictions, a
+pooled cross-sectional model does not yield any distinct advantage over
+separate cross-sections. The panel approach is avoided because of how
+frequently electoral boundaries change, noting that electorates with the
+same name across elections are not guaranteed to represent the same
+geographical region. Therefore any fixed or random effects models would
+be difficult to estimate without implementing consistent boundaries,
+which would require further imputation.
 
-For each cross-section, let the response \(\bm{y}\) be the vector two-party preferred vote in favour of the Liberal party; for example, \(y_i = 70\) represents a 70\% preference for Liberal, 30\% for Labor, in electorate \(i\).
-Although \(y_i\) lies in the interval \((0,100)\), observed values are never close to 0 or 100 (minimum \(24.05 \%\) and maximum \(74.90 \%\)), so there is no need to formally impose the constraint of \(y_i \in [0,100]\).
-Furthermore, the responses are found to be spatially correlated in each election (Moran's I test, \(p \le 7\cdot10^{-15}\)).
-This is not surprising as electorates are aggregate spatial units, and hence the spatial structure of the data must modelled appropriately.
+For each cross-section, let the response \(\bm{y}\) be the vector
+two-party preferred vote in favour of the Liberal party; for example,
+\(y_i = 70\) represents a 70\% preference for Liberal, 30\% for Labor,
+in electorate \(i\). Although \(y_i\) lies in the interval \((0,100)\),
+observed values are never close to 0 or 100 (minimum \(24.05 \%\) and
+maximum \(74.90 \%\)), so there is no need to formally impose the
+constraint of \(y_i \in [0,100]\). Furthermore, the responses are found
+to be spatially correlated in each election (Moran's I test,
+\(p \le 7\cdot10^{-15}\)). This is not surprising as electorates are
+aggregate spatial units, and hence the spatial structure of the data
+must modelled appropriately.
 
-The spatial error model \autocite{Anselin88} is chosen because it captures spatial heterogeneity by incorporating a spatially structured random effect vector \autocite{LeSage2009}.
-In this context, the random effect can be thought of as capturing the unobserved political climate in each electorate, where this climate is correlated with the climate in neighbouring electorates, under the assumption that the climate is independent of electoral socio-demographics.
+The spatial error model \autocite{Anselin88} is chosen because it
+captures spatial heterogeneity by incorporating a spatially structured
+random effect vector \autocite{LeSage2009}. In this context, the random
+effect can be thought of as capturing the unobserved political climate
+in each electorate, where this climate is correlated with the climate in
+neighbouring electorates, under the assumption that the climate is
+independent of electoral socio-demographics.
 
@@
 
-Spatial weights are calculated in accordance with the assumption that an electorate is equally correlated with any electorate that shares a part of its boundary.
-Let \(\rho\) be the spatial autoregressive coefficient, \(\bm{v}\) be a spherical error term, \(\bm{W}\) be a matrix of spatial weights (containing information about the neighbouring regions), \(\bm{X}\) be a matrix of socio-demographic covariates, \(\bm{\beta}\) be a vector of regression coefficients and \(\bm{a}\) be a spatially structured random effect vector.
-\[
+Spatial weights are calculated in accordance with the assumption that an
+electorate is equally correlated with any electorate that shares a part
+of its boundary. Let \(\rho\) be the spatial autoregressive coefficient,
+\(\bm{v}\) be a spherical error term, \(\bm{W}\) be a matrix of spatial
+weights (containing information about the neighbouring regions),
+\(\bm{X}\) be a matrix of socio-demographic covariates, \(\bm{\beta}\)
+be a vector of regression coefficients and \(\bm{a}\) be a spatially
+structured random effect vector. \[
   \bm{y} = \bm{X} \bm{\beta} + \bm{a},
-\]
-and
-\[
+\] and \[
   \bm{a} = \rho \bm{W} \bm{a} + \bm{v},
-\]
-where \(\bm{v} \sim N(\bm{0}, \sigma^2 \bm{I_n})\), and hence
-\[
+\] where \(\bm{v} \sim N(\bm{0}, \sigma^2 \bm{I_n})\), and hence \[
   \bm{y} = \bm{X} \bm{\beta} + (\bm{I}_n-\rho \bm{W})^{-1}\bm{v}.
-\]
-Estimation of the above spatial error model is undertaken using feasible generalized least squares.
+\] Estimation of the above spatial error model is undertaken using
+feasible generalized least squares.
 
-Table \ref{tab:printmodels} details the estimated model coefficients and their estimated standard errors, for each of the six elections.
-An interpretation of these estimated values is provided in the next section.
+Table \ref{tab:printmodels} details the estimated model coefficients and
+their estimated standard errors, for each of the six elections. An
+interpretation of these estimated values is provided in the next
+section.
 
-\hypertarget{results}{%
-\section{Results}\label{results}}
+\section{Results}\label{results}
 
-\hypertarget{spatial-autoregressive-parameter}{%
-\subsection{Spatial autoregressive parameter}\label{spatial-autoregressive-parameter}}
+\subsection{Spatial autoregressive
+parameter}\label{spatial-autoregressive-parameter}
 
-The spatial autoregressive coefficient \(\rho\) is positive and significant in only the 2001 and 2016 elections (Figure \ref{fig:rhovis}), meaning that in these elections, the political climate of an electorate appears to be affected by the attitudes of its neighbours.
-Conversely, in the other four elections, the spatial effect weakens to become insignificant.
-In these years, it appears that the spatial component does not explain anything not already explained by the electoral socio-demographics, meaning electorates effectively voted independently.
+The spatial autoregressive coefficient \(\rho\) is positive and
+significant in only the 2001 and 2016 elections (Figure
+\ref{fig:rhovis}), meaning that in these elections, the political
+climate of an electorate appears to be affected by the attitudes of its
+neighbours. Conversely, in the other four elections, the spatial effect
+weakens to become insignificant. In these years, it appears that the
+spatial component does not explain anything not already explained by the
+electoral socio-demographics, meaning electorates effectively voted
+independently.
 
@@
 
-{\centering \includegraphics[height=0.15\textheight,]{draft-paper_files/figure-latex/rhovis-1}
+{\centering \includegraphics[height=0.2\textheight,]{draft-paper_files/figure-latex/rhovis-1}
 
@@
 
-\hypertarget{country-wide-trend}{%
-\subsection{Country-wide trend}\label{country-wide-trend}}
+\subsection{Country-wide trend}\label{country-wide-trend}
 
@@
 
-{\centering \includegraphics[height=0.15\textheight,]{draft-paper_files/figure-latex/plotintercept-1}
+{\centering \includegraphics[height=0.2\textheight,]{draft-paper_files/figure-latex/plotintercept-1}
 
@@
 
-Since all socio-demographics have been standardized to have a mean of zero and a variance of one, the intercept in each model can be interpreted as the estimated two-party preferred vote for an electorate with mean characteristics\footnote{Mean of all variables aside from Judaism, Indigenous, Islam and Buddhism, where it assumes the mean of the log value.}.
-Figure \ref{fig:plotintercept} shows that the baseline of party preference has varied over the elections, with the biggest swing occurring in the 2007 election where the mean electorate shifted more than five percentage points in favour of the Labor party.
+Since all socio-demographics have been standardized to have a mean of
+zero and a variance of one, the intercept in each model can be
+interpreted as the estimated two-party preferred vote for an electorate
+with mean characteristics\footnote{Mean of all variables aside from
+  Judaism, Indigenous, Islam and Buddhism, where it assumes the mean of
+  the log value.}. Figure \ref{fig:plotintercept} shows that the
+baseline of party preference has varied over the elections, with the
+biggest swing occurring in the 2007 election where the mean electorate
+shifted more than five percentage points in favour of the Labor party.
 
-\hypertarget{influential-socio-demographics}{%
-\subsection{Influential socio-demographics}\label{influential-socio-demographics}}
+\subsection{Influential
+socio-demographics}\label{influential-socio-demographics}
 
@@
 
-\caption{Partial residual plots for: Incomes, Unemployment, Extractive, Transformative, ManagerAdmin and DiffAddress}\label{fig:partresplot1}
+\caption{Partial residual plots by election year for a selection of predictors.
+Linear model with 95\% confidence bands is overlaid.
+Most predictors have a positive relationship: the larger the value the more likely the electorate preferences the Coalition.
+The relationship is relatively robust over time, with the exception of Unemployment.}\label{fig:partresplot1}
 \end{figure}
@@
 
-\caption{Partial residual plots for: DeFacto, MedianAge, Education, OtherLanguage, BornSEEuro, BornAsia and Judaism.}\label{fig:partresplot2}
+\caption{Partial residual plots by election year for a selection of predictors.
+Linear model with 95\% confidence bands is overlaid.
+Several predictors have a negative relationship: with larger values indicating the electorate more likely preferences Labor.
+Education exhibits a strong temporal change.}\label{fig:partresplot2}
 \end{figure}
 
-To investigate the socio-demographics that have a strong effect on the two-party preferred vote, partial residual plots are used and shown in Figures \ref{fig:partresplot1} and \ref{fig:partresplot2}.
-The partial residuals are the residuals from the fitted model with the estimated effect an individual variable added.
-These show the direction, size and significance of an estimated effect --- the slope of the prediction line matches the estimated coefficient, and the shaded region represents a 95\% confidence band.
-If a horizontal line can be drawn through the confidence band, then the effect is insignificant.
-The estimated intercept is also added to the partial residuals for interpretability.
-Plots for each election are faceted to compare the effects over time in Figures \ref{fig:partresplot1} and \ref{fig:partresplot2}.
+To investigate the socio-demographics that have a strong effect on the
+two-party preferred vote, partial residual plots are used and shown in
+Figures \ref{fig:partresplot1} and \ref{fig:partresplot2}. The partial
+residuals are the residuals from the fitted model with the estimated
+effect an individual variable added. These show the direction, size and
+significance of an estimated effect --- the slope of the prediction line
+matches the estimated coefficient, and the shaded region represents a
+95\% confidence band. If a horizontal line can be drawn through the
+confidence band, then the effect is insignificant. The estimated
+intercept is also added to the partial residuals for interpretability.
+Plots for each election are faceted to compare the effects over time in
+Figures \ref{fig:partresplot1} and \ref{fig:partresplot2}.
 
-It is important here to note the ecological fallacy: insights are being drawn at the electorate level, and cannot be inferred for another disaggregate level (e.g., individual voters).
+It is important here to note the ecological fallacy: insights are being
+drawn at the electorate level, and cannot be inferred for another
+disaggregate level (e.g., individual voters).
 
-\hypertarget{income-and-unemployment}{%
-\subsubsection{Income and unemployment}\label{income-and-unemployment}}
+\subsubsection{Income and unemployment}\label{income-and-unemployment}
 
-Typically the Labor party campaigns on more progressive policies, which often include tax reform that adversely affects higher income earners, and more generous social assistance programs.
-Perhaps it is due to these policies that higher income electorates appear more likely to support the Liberal party, as the \texttt{Incomes} factor has a positive effect on Liberal preference (see row 1 in Figure \ref{fig:partresplot1}).
-This effect is significant in every election aside from 2004, where it is only marginally insignificant (\(p = 0.0613\)).
-Unemployment however, is not as influential.
-In 2001 and 2004, electorates with higher unemployment align with Labor, but over time this shifts towards support for the Liberal party, culminating in a significantly positive effect in 2016.
+Typically the Labor party campaigns on more progressive policies, which
+often include tax reform that adversely affects higher income earners,
+and more generous social assistance programs. Perhaps it is due to these
+policies that higher income electorates appear more likely to support
+the Liberal party, as the \texttt{Incomes} factor has a positive effect
+on Liberal preference (see row 1 in Figure \ref{fig:partresplot1}). This
+effect is significant in every election aside from 2004, where it is
+only marginally insignificant (\(p = 0.0613\)). Unemployment however, is
+not as influential. In 2001 and 2004, electorates with higher
+unemployment align with Labor, but over time this shifts towards support
+for the Liberal party, culminating in a significantly positive effect in
+2016.
 
-\hypertarget{industry-and-type-of-work}{%
-\subsubsection{Industry and type of work}\label{industry-and-type-of-work}}
+\subsubsection{Industry and type of
+work}\label{industry-and-type-of-work}
 
-Electorates with higher proportions of workers in mining, gas, water, agriculture, waste and electricity (grouped as \texttt{Extractive} industries) are consistently linked with higher support for the Liberal party, with the magnitude of this effect slightly increasing over the years (see row 3 in Figure \ref{fig:partresplot1}).
-This is unsurprising, as the Liberal party has close ties with these traditional energy industries, and typically present policies to reduce taxation on energy production.
-Furthermore, electorates with more workers in construction or manufacturing industries (\texttt{Transformative}) are also more likely to support the Liberal party (see row 4 in Figure \ref{fig:partresplot1}).
+Electorates with higher proportions of workers in mining, gas, water,
+agriculture, waste and electricity (grouped as \texttt{Extractive}
+industries) are consistently linked with higher support for the Liberal
+party, with the magnitude of this effect slightly increasing over the
+years (see row 3 in Figure \ref{fig:partresplot1}). This is
+unsurprising, as the Liberal party has close ties with these traditional
+energy industries, and typically present policies to reduce taxation on
+energy production. Furthermore, electorates with more workers in
+construction or manufacturing industries (\texttt{Transformative}) are
+also more likely to support the Liberal party (see row 4 in Figure
+\ref{fig:partresplot1}).
 
-Similarly, the proportion of workers in managerial, administrative, clerical and sales roles (\texttt{ManagerAdmin}) is also a significant predictor of two-party preference vote across all six elections, with a higher proportion of people working these jobs increasing Liberal support.
-The magnitude of this effect also seems to increase over the years.
+Similarly, the proportion of workers in managerial, administrative,
+clerical and sales roles (\texttt{ManagerAdmin}) is also a significant
+predictor of two-party preference vote across all six elections, with a
+higher proportion of people working these jobs increasing Liberal
+support. The magnitude of this effect also seems to increase over the
+years.
 
-\hypertarget{household-mobility}{%
-\subsubsection{Household mobility}\label{household-mobility}}
+\subsubsection{Household mobility}\label{household-mobility}
 
-In each of the six elections, electorates with a higher proportion of people that have recently (in the past five years) moved house are more likely to support the Liberal party, although this effect was marginally insignificant in 2013 (see row 6 in Figure \ref{fig:partresplot1}.
-Having controlled for characteristics of house ownership and rental prices (via the factors \texttt{PropertyOwned} and \texttt{RentLoan} respectively), this effect is somewhat surprising.
+In each of the six elections, electorates with a higher proportion of
+people that have recently (in the past five years) moved house
+(\texttt{DiffAddress}) are more likely to support the Liberal party,
+although this effect was marginally insignificant in 2013 (see row 6 in
+Figure \ref{fig:partresplot1}. Having controlled for characteristics of
+house ownership and rental prices (via the factors
+\texttt{PropertyOwned} and \texttt{RentLoan} respectively), this effect
+is somewhat surprising.
 
-\hypertarget{relationships}{%
-\subsubsection{Relationships}\label{relationships}}
+\subsubsection{Relationships}\label{relationships}
 
-De facto relationships, but not marriages, are found to be an important (and significant) predictor of the two-party preferred vote in all six elections, with more de facto relationships associated with higher support for the Labor party.
-The proportion of individuals who are married however, is insignificant (not shown).
+De facto relationships, but not marriages, are found to be an important
+(and significant) predictor of the two-party preferred vote in all six
+elections, with more de facto relationships associated with higher
+support for the Labor party. The proportion of individuals who are
+married however, is insignificant (not shown).
 
-\hypertarget{age}{%
-\subsubsection{Age}\label{age}}
+\subsubsection{Age}\label{age}
 
-Regions comprising more older people are often believed to be more conservative, and indeed it found that electorates with a higher median age are more likely to support the Liberal party --- although this effect is only significant in 2007 and 2010 (see row 2 in Figure \ref{fig:partresplot2}).
+Regions comprising more older people are often believed to be more
+conservative, and indeed it found that electorates with a higher median
+age are more likely to support the Liberal party --- although this
+effect is only significant in 2007 and 2010 (see row 2 in Figure
+\ref{fig:partresplot2}).
 
-\hypertarget{education}{%
-\subsubsection{Education}\label{education}}
+\subsubsection{Education}\label{education}
 
-Since 2007, electorates with higher education levels are associated with supporting the Labor party, although this effect is only significant in 2016.
-Before 2007, education has an almost zero effect (see row 3 in Figure \ref{fig:partresplot1}).
+Since 2007, electorates with higher education levels are associated with
+supporting the Labor party, although this effect is only significant in
+2016. Before 2007, education has an almost zero effect (see row 3 in
+Figure \ref{fig:partresplot1}).
 
-\hypertarget{diversity}{%
-\subsubsection{Diversity}\label{diversity}}
+\subsubsection{Diversity}\label{diversity}
 
-Larger migrant populations from Asia, the Middle East, South-Eastern Europe, the United Kingdom and elsewhere, are either associated with Labor support, or have no effect.
-Of these areas, only South-Eastern European populations appear significant in each election, with the proportion of Asian migrants also being significant in 2010.
-Speaking other languages (aside from English) however, appears to have a far stronger effect, as observed through the \texttt{OtherLanguage} variable.
-Electorates with more diverse speech are associated with higher support for the Liberal party from 2004 onwards, with this effect being significant in 2007, 2010 and 2016.
-Furthermore, of the variables relating to religion, only Judaism shows a consistent effect, with electorates with relatively large Jewish populations more likely to vote Liberal.
+Larger migrant populations from Asia, the Middle East, South-Eastern
+Europe, the United Kingdom and elsewhere, are either associated with
+Labor support, or have no effect. Of these areas, only South-Eastern
+European populations appear significant in each election, with the
+proportion of Asian migrants also being significant in 2010. Speaking
+other languages (aside from English) however, appears to have a far
+stronger effect, as observed through the \texttt{OtherLanguage}
+variable. Electorates with more diverse speech are associated with
+higher support for the Liberal party from 2004 onwards, with this effect
+being significant in 2007, 2010 and 2016. Furthermore, of the variables
+relating to religion, only Judaism shows a consistent effect, with
+electorates with relatively large Jewish populations more likely to vote
+Liberal.
 
-\hypertarget{a-note-on-similar-variables}{%
-\subsubsection{A note on similar variables}\label{a-note-on-similar-variables}}
+\subsubsection{A note on similar
+variables}\label{a-note-on-similar-variables}
 
-Many of the Census variables represent similar information, which is why factors were created and some variables were removed.
-However, some variables remain which are closely related.
-For example, an electorate's income level (via \texttt{Incomes}) is likely to be related to electoral unemployment and labor force participation (via \texttt{Unemployment}).
-In 2001, the coefficient estimate for \texttt{Unemployment} is negative but not significant, whilst the \texttt{Incomes} variables is significant.
-If the \texttt{Incomes} variable is removed from the model in 2001, \texttt{Unemployment} absorbs the negative effect, becoming significant (\(p = 0.0056\)).
+Many of the Census variables represent similar information, which is why
+factors were created and some variables were removed. However, some
+variables remain which are closely related. For example, an electorate's
+income level (via \texttt{Incomes}) is likely to be related to electoral
+unemployment and labor force participation (via \texttt{Unemployment}).
+In 2001, the coefficient estimate for \texttt{Unemployment} is negative
+but not significant, whilst the \texttt{Incomes} variables is
+significant. If the \texttt{Incomes} variable is removed from the model
+in 2001, \texttt{Unemployment} absorbs the negative effect, becoming
+significant (\(p = 0.0056\)).
 
-\hypertarget{a-closer-look-at-the-residuals}{%
-\subsection{A closer look at the residuals}\label{a-closer-look-at-the-residuals}}
+\subsection{A closer look at the
+residuals}\label{a-closer-look-at-the-residuals}
 
-\hypertarget{residuals-by-state}{%
-\subsubsection{Residuals by state}\label{residuals-by-state}}
+\subsubsection{Residuals by state}\label{residuals-by-state}
 
-It is often hypothesized that states have systematic differences that cause their electorates to vote differently.
-Boxplots of residuals grouped by state (Figure \ref{fig:resstate}) reveal that only Tasmania, the Australian Capital Territory and the Northern Territory appear to have a state-specific effect not captured by the models.
-Tasmania and the Australian Capital Territory appear to have a bias towards Labor, whereas the Northern Territory tends towards voting Liberal.
-However, there are relatively few electorates in each of these states (five, two and two respectively), so this apparent result may be due to incumbent effects rather than an actual state-specific bias.
+It is often hypothesized that states have systematic differences that
+cause their electorates to vote differently. Boxplots of residuals
+grouped by state (Figure \ref{fig:resstate}) reveal that the data
+reflects this -- there appears to be a state-specific effect not
+captured by the models. Tasmania and the Australian Capital Territory
+appear to have a bias towards Labor, whereas the Northern Territory
+tends towards voting Liberal. However, there are relatively few
+electorates in each of these states (five, two and two respectively), so
+this apparent result may be due to incumbent effects rather than an
+actual state-specific bias.
 
@@
 
-\caption{Boxplot of residuals by state with jittered points.}\label{fig:resstate}
+\caption{Boxplot of residuals by state with jittered points. States ordered by median residual. A state-specific bias not captured by the model is evident.}\label{fig:resstate}
 \end{figure}
 
-\hypertarget{outlier-electorates}{%
-\subsubsection{Outlier electorates}\label{outlier-electorates}}
+\subsubsection{Outlier electorates}\label{outlier-electorates}
 
-Based on the distribution of the Cook's distance values, a Cook's distance greater than \(0.1\) is considered to be influential and a potential outlier.
-The electorate of Sydney (NSW) has a large Cook's distance from 2001 to 2013, due to its diverse population (language, birthplace and religion), high number of defacto relationships, high income, high household mobility and small amount of workers in extractive and transformative jobs.
-It has remained a strong supporter of the Labor party and the Liberal vote is severely overpredicted by the model, making it an outlier.
-Nearby in metropolitan NSW, the electorate of Wentworth is found to be an outlier in all but the 2007 election.
-Although historically Liberal, its two-party vote jumped by over 10 percentage points in 2010 without experiencing any notable changes in its socio-demographic makeup --- implying that this may be the direct effect of its Liberal member, Malcolm Turnbull, becoming the leader of the Liberal party.
-Liberal support in Wentworth is underpredicted by the model in each year, and more so with Turnbull as Liberal leader.
+Based on the distribution of the Cook's distance values, a Cook's
+distance greater than \(0.1\) is considered to be influential and a
+potential outlier. The electorate of Sydney (NSW) has a large Cook's
+distance from 2001 to 2013, due to its diverse population (language,
+birthplace and religion), high number of defacto relationships, high
+income, high household mobility and small amount of workers in
+extractive and transformative jobs. It has remained a strong supporter
+of the Labor party and the Liberal vote is severely overpredicted by the
+model, making it an outlier. Nearby in metropolitan NSW, the electorate
+of Wentworth is found to be an outlier in all but the 2007 election.
+Although historically Liberal, its two-party vote jumped by over 10
+percentage points in 2010 without experiencing any notable changes in
+its socio-demographic makeup --- implying that this may be the direct
+effect of its Liberal member, Malcolm Turnbull, becoming the leader of
+the Liberal party. Liberal support in Wentworth is underpredicted by the
+model in each year, and more so with Turnbull as Liberal leader.
 
-Lingiari, an electorate taking up almost all of the Northern Territory, is an outlier in the 2001--2007 elections due to its large Indigenous population, young age profile and low rates of property ownership.
-Fowler (NSW) has a diverse population with a high proportion of migrants, many Buddhists and Muslims, and has strong Labor support, making it influential in 2001, 2004 and 2010.
-Other electorates with large Cook's distance are Barton (NSW) and Leichhardt (QLD) in 2016, and Canberra (ACT) in 2007.
+Lingiari, an electorate taking up almost all of the Northern Territory,
+is an outlier in the 2001--2007 elections due to its large Indigenous
+population, young age profile and low rates of property ownership.
+Fowler (NSW) has a diverse population with a high proportion of
+migrants, many Buddhists and Muslims, and has strong Labor support,
+making it influential in 2001, 2004 and 2010. Other electorates with
+large Cook's distance are Barton (NSW) and Leichhardt (QLD) in 2016, and
+Canberra (ACT) in 2007.
 
-\hypertarget{conclusion}{%
-\section{Conclusion}\label{conclusion}}
+\section{Conclusion}\label{conclusion}
 
-This paper explores the effects of electoral socio-demographic characteristics on the two-party preferred vote in the 2001--2016 elections, using information from the corresponding Australian federal elections and Censuses.
-As a Census does not always occur in the same year as an election, Census data for the 2004--2013 elections are generated by employing a method of spatio-temporal imputation.
-This imputes electoral socio-demographics for the electoral boundaries in place at the time of the election --- an approach that is distinctly different from previous work on modelling election outcomes, where Census and election data are typically joined without addressing their temporal differences.
-Before estimating a model, these socio-demographic variables are standardized (to adjust for changing variable scales) and many variables (representing similar information) are combined into factors, resulting in a reduced predictor set.
-A spatial error model is then estimated for each election, accounting for the inherent spatial structure of the data.
+This paper explores the effects of electoral socio-demographic
+characteristics on the two-party preferred vote in the 2001--2016
+elections, using information from the corresponding Australian federal
+elections and Censuses. As a Census does not always occur in the same
+year as an election, Census data for the 2004--2013 elections are
+generated by employing a method of spatio-temporal imputation. This
+imputes electoral socio-demographics for the electoral boundaries in
+place at the time of the election --- an approach that is distinctly
+different from previous work on modelling election outcomes, where
+Census and election data are typically joined without addressing their
+temporal differences. Before estimating a model, these socio-demographic
+variables are standardized (to adjust for changing variable scales) and
+many variables (representing similar information) are combined into
+factors, resulting in a reduced predictor set. A spatial error model is
+then estimated for each election, accounting for the inherent spatial
+structure of the data.
 
-Across the past six elections, most of the socio-demographics that drive the electoral two-party preferred vote are found to remain steady, whilst a few (typically weaker) effects vary over time.
-Industry and type of work are particularly influential, with energy-related and manufacturing/construction jobs, as well as administrative roles being strongly linked with the Liberal party in all elections.
-Incomes have a similarly consistent effect, with higher income areas supporting Liberal.
-Higher levels of unemployment shift from weak association with Labor to a significant Liberal effect over the years, and higher education levels are associated with Labor from 2007 (although only significant in 2016).
-It is also found that electorates with higher household mobility support Liberal, birthplace diversity favours Labor and more de facto relationships align with Labor preference --- although marriages, family and household sizes have no material influence.
-Furthermore, the neighbourhood (spatial) effects are found to be positive in all elections, although only significant in 2001 and 2016, meaning that in the 2004--2013 elections, electorates effectively voted independently.
+Across the past six elections, most of the socio-demographics that drive
+the electoral two-party preferred vote are found to remain steady,
+whilst a few (typically weaker) effects vary over time. Industry and
+type of work are particularly influential, with energy-related and
+manufacturing/construction jobs, as well as administrative roles being
+strongly linked with the Liberal party in all elections. Incomes have a
+similarly consistent effect, with higher income areas supporting
+Liberal. Higher levels of unemployment shift from weak association with
+Labor to a significant Liberal effect over the years, and higher
+education levels are associated with Labor from 2007 (although only
+significant in 2016). It is also found that electorates with higher
+household mobility support Liberal, birthplace diversity favours Labor
+and more de facto relationships align with Labor preference --- although
+marriages, family and household sizes have no material influence.
+Furthermore, the neighbourhood (spatial) effects are found to be
+positive in all elections, although only significant in 2001 and 2016,
+meaning that in the 2004--2013 elections, electorates effectively voted
+independently.
 
-The findings in this paper complement the existing literature by modelling temporal trends, which as far as the authors are aware, has not been done previously for Australian elections using a regression framework.
-It is also the first study to model any Australian election since 2010 using Census information.
+The findings in this paper complement the existing literature by
+modelling temporal trends, which as far as the authors are aware, has
+not been done previously for Australian elections using a regression
+framework. It is also the first study to model any Australian election
+since 2010 using Census information.
 
-Additionally, a key contribution of this research is the wrangling of the raw data and imputed data sets for the 2004, 2007, 2010 and 2013 elections, which have been contributed to the \texttt{eechidna} \texttt{R} package --- providing a rich, accessible data resource for future Australian electoral analysis.
+Additionally, a key contribution of this research is the wrangling of
+the raw data and imputed data sets for the 2004, 2007, 2010 and 2013
+elections, which have been contributed to the \texttt{eechidna}
+\texttt{R} package --- providing a rich, accessible data resource for
+future Australian electoral analysis.
 
-\hypertarget{acknowledgements}{%
-\section{Acknowledgements}\label{acknowledgements}}
+\section{Acknowledgements}\label{acknowledgements}
 
-This paper was produced using \texttt{RMarkdown} \autocite{rmarkdown} and \texttt{knitr} \autocite{knitr}.
-All corresponding code for this paper can be found in the github repository \href{https://github.com/jforbes14/journal-writing}{github.com/jforbes14/journal-writing}, and the data used is available in the \texttt{eechidna} package \autocite{eechidna}.
-All raw data was obtained from the Australian Electoral Commission, the Australian Bureau of Statistics and the Australian Government.
+This paper was produced using \texttt{RMarkdown} \autocite{rmarkdown}
+and \texttt{knitr} \autocite{knitr}. All corresponding code for this
+paper can be found in the github repository
+\href{https://github.com/jforbes14/journal-writing}{github.com/jforbes14/journal-writing},
+and the data used is available in the \texttt{eechidna} package
+\autocite{eechidna}. All raw data was obtained from the Australian
+Electoral Commission, the Australian Bureau of Statistics and the
+Australian Government.
 
-\hypertarget{software}{%
-\section{Software}\label{software}}
+\section{Software}\label{software}
 
-All election and Census datasets, along with electoral maps and more, are available in the \texttt{eechidna} (Exploring Election and Census Highly Informative Data Nationally for Australia) \texttt{R} package, which can be downloaded from CRAN.
-The \texttt{eechidna} package makes it easy to look at the data from the Australian Federal elections and Censuses that occurred between 2001 and 2016.
-This study contributed a large revision to the \texttt{eechidna} package, which included the addition of election and Census data for 2001--2010, voting outcomes for polling booths and imputed Census data for election years.
-For more details on using \texttt{eechidna}, please see the articles (vignettes) on the github page \href{https://ropenscilabs.github.io/eechidna/}{ropenscilabs.github.io/eechidna/}.
+All election and Census datasets, along with electoral maps and more,
+are available in the \texttt{eechidna} (Exploring Election and Census
+Highly Informative Data Nationally for Australia) \texttt{R} package,
+which can be downloaded from CRAN. The \texttt{eechidna} package makes
+it easy to look at the data from the Australian Federal elections and
+Censuses that occurred between 2001 and 2016. This study contributed a
+large revision to the \texttt{eechidna} package, which included the
+addition of election and Census data for 2001--2010, voting outcomes for
+polling booths and imputed Census data for election years. For more
+details on using \texttt{eechidna}, please see the articles (vignettes)
+on the github page
+\href{https://ropenscilabs.github.io/eechidna/}{ropenscilabs.github.io/eechidna/}.
 
-The authors would like to sincerely thank Anthony Ebert, Heike Hofmann, Thomas Lumley, Ben Marwick, Carson Sievert, Mingzhu Sun, Dilini Talagala, Nicholas Tierney, Nathaniel Tomasetti, Earo Wang and Fang Zhou, all of whom have contributed to the \texttt{eechidna} package.
+The authors would like to sincerely thank Anthony Ebert, Heike Hofmann,
+Thomas Lumley, Ben Marwick, Carson Sievert, Mingzhu Sun, Dilini
+Talagala, Nicholas Tierney, Nathaniel Tomasetti, Earo Wang and Fang
+Zhou, all of whom have contributed to the \texttt{eechidna} package.
 

### 2019-04-13 functions.R

@@ -226,7 +226,7 @@ my_visreg <- function(my_model, sp_weights, varname,
 }
 
 # Grid visreg
-grid_visreg <- function(varname, plot = TRUE, myscale = "free", top = FALSE) {
+grid_visreg <- function(varname, plot = TRUE, myscale = "free_y", top = FALSE) {
 
     p16 <- my_visreg(glsmod16, sp_weights_16, varname = varname, plot = FALSE, year = "2016")
     p13 <- my_visreg(glsmod13, sp_weights_13, varname = varname, plot = FALSE, year = "2013")
@@ -260,7 +260,7 @@ grid_visreg <- function(varname, plot = TRUE, myscale = "free", top = FALSE) {
         geom_line(aes(x = variable, y = fitted), col = "blue", size = 1) +
         theme_bw() +
         labs(x = "", y = "") +
-        facet_grid(varname ~ year, scales = "free") +
+        facet_grid(varname ~ year, scales = "free_y") +
         theme(plot.margin=unit(c(b = 0.05, l = 0, t = -0.2, r = 0),"cm"))
 
       if (top == FALSE) {

### 2019-04-14 draft-paper.Rmd
Fixed table to be aligned correctly

@@
 # Add rho and GLS residual standard error
-  regtable <- stargazer::stargazer(glsmod01, glsmod04, glsmod07, glsmod10, glsmod13, glsmod16,
+regtable <- stargazer::stargazer(
+  glsmod01, glsmod04, glsmod07, glsmod10, glsmod13, glsmod16,
   header = FALSE,
@@
 regtable <- c(regtable[1:77],sigmaline, regtable[78:length(regtable)])
+# Allow alignment on decimal point
+regtable[6] <- "\\begin{tabular}{@{\\extracolsep{1pt}}lrrrrrr}" 
+regtable[12:75] <- gsub(")","\\\\rlap{)}",regtable[12:75])
+regtable[12:75] <- gsub("$^{*}$","\\rlap{$^{*}$}", regtable[12:75], fixed=TRUE)
+regtable[12:75] <- gsub("$^{**}$","\\rlap{$^{**}$}", regtable[12:75], fixed=TRUE)
+regtable[12:75] <- gsub("$^{***}$","\\rlap{$^{***}$}", regtable[12:75], fixed=TRUE)
+regtable[5] <- "\\scriptsize\\tabcolsep=0.35cm"
 ```

### 2019-04-14 draft-paper.tex
Fixed table to be aligned correctly

--- 2019-04-10 draft-paper.tex
@@
   {}    % How the section is represented
+
+% Align tabular column on decimal point
+\usepackage{dcolumn}
+\newcolumntype{d}[1]{D{.}{\cdot}{#1}}
 
@@
 
-\begin{table}[!htbp] \centering    \caption{Estimated spatial error model parameters (standard errors) for each of the six election years.}    \label{tab:printmodels}  \scriptsize  \begin{tabular}{@{\extracolsep{1pt}}lcccccc}  \\[-1.8ex]\hline  \hline \\[-1.8ex]  \\[-1.8ex] & \multicolumn{6}{c}{\vspace{-0.5cm}} \\   & 2001 & 2004 & 2007 & 2010 & 2013 & 2016 \\  \hline \\[-1.8ex]  $\rho$ & 0.46$^{***}$ & 0.29$^{*}$ & 0.24 & 0.19 & 0.27$^{*}$ & 0.50$^{***}$\\  & (0.15) & (0.17) & (0.17) & (0.16) & (0.16) & (0.17)\\  AusCitizen & $-$3.13 & $-$2.64 & $-$2.53 & $-$0.08 & $-$3.40 & $-$1.80 \\    & (2.26) & (2.43) & (2.34) & (2.79) & (2.76) & (2.71) \\    BornAsia & 2.22 & $-$0.95 & $-$1.60 & $-$6.83$^{**}$ & $-$3.03 & $-$0.55 \\    & (2.18) & (2.44) & (2.19) & (2.73) & (2.71) & (2.17) \\    Born\_MidEast & $-$1.15 & $-$1.59 & $-$2.01$^{*}$ & $-$2.03 & $-$0.92 & $-$1.44 \\    & (1.07) & (1.20) & (1.11) & (1.27) & (1.24) & (1.13) \\    BornSEEuro & $-$3.21$^{**}$ & $-$4.24$^{***}$ & $-$3.61$^{***}$ & $-$4.14$^{***}$ & $-$3.69$^{***}$ & $-$2.72$^{***}$ \\    & (1.42) & (1.46) & (1.02) & (1.19) & (1.07) & (0.97) \\    Born\_UK & 0.25 & $-$0.07 & 0.34 & 0.56 & $-$0.09 & $-$1.32 \\    & (1.00) & (0.98) & (0.90) & (1.07) & (1.04) & (1.04) \\    BornElsewhere & $-$5.04 & $-$4.91 & $-$4.13 & 2.35 & $-$5.23 & $-$4.14 \\    & (3.30) & (3.68) & (3.38) & (4.23) & (4.15) & (3.97) \\    Buddhism & $-$0.49 & $-$0.17 & $-$1.37 & $-$0.83 & $-$0.12 & $-$1.60 \\    & (1.39) & (1.61) & (1.61) & (1.80) & (1.68) & (1.56) \\    Christianity & $-$2.48 & $-$1.23 & 0.38 & 0.50 & 2.41 & 1.68 \\    & (1.73) & (1.85) & (1.83) & (1.99) & (1.85) & (1.78) \\    CurrentlyStudying & $-$2.19$^{**}$ & $-$0.13 & 2.06$^{*}$ & 2.12$^{*}$ & 1.15 & $-$0.16 \\    & (0.99) & (1.13) & (1.17) & (1.25) & (1.26) & (1.18) \\    DeFacto & $-$6.44$^{***}$ & $-$5.37$^{**}$ & $-$6.43$^{***}$ & $-$8.07$^{***}$ & $-$6.56$^{**}$ & $-$8.53$^{***}$ \\    & (1.87) & (2.48) & (2.31) & (3.06) & (3.11) & (2.83) \\    DiffAddress & 3.88$^{***}$ & 5.06$^{***}$ & 4.22$^{***}$ & 5.57$^{***}$ & 3.53$^{*}$ & 5.67$^{***}$ \\    & (0.94) & (1.12) & (0.99) & (1.76) & (1.91) & (1.60) \\    Distributive & 1.27 & 2.01$^{*}$ & 1.36 & 1.57 & 2.10$^{*}$ & 1.20 \\    & (1.12) & (1.21) & (1.13) & (1.34) & (1.27) & (1.21) \\    Education & 1.08 & 0.52 & $-$5.52$^{*}$ & $-$4.08 & $-$4.44 & $-$7.07$^{**}$ \\    & (2.38) & (3.12) & (3.27) & (3.95) & (3.78) & (3.55) \\    Extractive & 4.83$^{***}$ & 5.45$^{***}$ & 5.37$^{***}$ & 7.31$^{***}$ & 6.71$^{***}$ & 7.43$^{***}$ \\    & (1.48) & (1.42) & (1.36) & (1.56) & (1.47) & (1.39) \\    FamHouseSize & $-$0.16 & 0.87 & $-$2.40 & $-$2.53 & $-$3.26 & $-$2.91 \\    & (2.19) & (2.72) & (2.69) & (3.25) & (3.28) & (2.90) \\    Incomes & 4.36$^{**}$ & 5.03$^{*}$ & 9.45$^{***}$ & 7.09$^{**}$ & 7.97$^{***}$ & 12.20$^{***}$ \\    & (1.77) & (2.66) & (2.75) & (3.25) & (2.92) & (2.75) \\    Indigenous & 2.91$^{*}$ & 1.97 & 2.48 & 2.84 & 0.67 & $-$0.05 \\    & (1.68) & (1.95) & (1.75) & (2.16) & (2.14) & (2.00) \\    Islam & $-$0.92 & $-$0.97 & $-$0.54 & $-$2.50 & $-$0.82 & $-$0.95 \\    & (1.22) & (1.36) & (1.27) & (1.52) & (1.42) & (1.34) \\    Judaism & 1.88$^{*}$ & 1.78 & 2.66$^{***}$ & 1.97$^{*}$ & 2.74$^{**}$ & 1.65$^{*}$ \\    & (1.05) & (1.13) & (1.01) & (1.15) & (1.10) & (1.00) \\    ManagerAdmin & 2.06$^{***}$ & 3.32$^{***}$ & 6.00$^{***}$ & 5.47$^{***}$ & 5.04$^{***}$ & 5.78$^{***}$ \\    & (0.71) & (0.93) & (0.90) & (1.08) & (1.03) & (1.06) \\    Married & 0.44 & 0.11 & $-$1.22 & $-$0.22 & 0.91 & $-$2.34 \\    & (2.31) & (2.96) & (2.83) & (3.15) & (3.03) & (2.81) \\    MedianAge & 2.32$^{*}$ & 4.96$^{***}$ & 3.66$^{**}$ & 4.00$^{*}$ & 2.30 & 2.87 \\    & (1.32) & (1.65) & (1.81) & (2.26) & (2.08) & (1.79) \\    NoReligion & $-$1.57 & $-$0.92 & 0.56 & $-$0.30 & 1.02 & 1.31 \\    & (1.59) & (1.71) & (1.73) & (1.92) & (1.94) & (2.04) \\    OneParentHouse & $-$1.73 & $-$0.45 & $-$0.75 & $-$1.46 & $-$0.77 & $-$0.74 \\    & (1.36) & (1.59) & (1.49) & (1.69) & (1.57) & (1.47) \\    OtherLanguage & $-$0.44 & 5.92 & 9.98$^{**}$ & 11.24$^{**}$ & 9.00$^{*}$ & 9.84$^{**}$ \\    & (3.22) & (4.16) & (3.91) & (4.76) & (4.66) & (4.44) \\    PropertyOwned & $-$0.46 & $-$0.53 & 0.67 & $-$0.94 & $-$0.48 & 1.41 \\    & (1.37) & (1.50) & (1.43) & (1.76) & (1.67) & (1.50) \\    RentLoanPrice & $-$1.57 & $-$3.32$^{*}$ & $-$4.01$^{**}$ & $-$0.97 & $-$0.70 & $-$0.89 \\    & (1.49) & (1.76) & (1.67) & (2.07) & (2.07) & (2.16) \\    SocialServ & 2.51$^{*}$ & 1.65 & 2.47$^{*}$ & 2.53$^{*}$ & 2.35$^{*}$ & 4.45$^{***}$ \\    & (1.33) & (1.41) & (1.29) & (1.47) & (1.32) & (1.19) \\    Transformative & 3.24$^{**}$ & 4.73$^{***}$ & 4.84$^{***}$ & 4.46$^{**}$ & 3.56$^{**}$ & 4.58$^{***}$ \\    & (1.55) & (1.78) & (1.74) & (1.98) & (1.78) & (1.53) \\    Unemployment & $-$2.45$^{*}$ & $-$3.07$^{*}$ & 0.29 & 0.08 & 1.67 & 2.79$^{**}$ \\    & (1.40) & (1.63) & (1.51) & (1.76) & (1.51) & (1.37) \\    Constant & 50.81$^{***}$ & 52.60$^{***}$ & 47.31$^{***}$ & 49.93$^{***}$ & 53.51$^{***}$ & 50.49$^{***}$ \\    & (0.71) & (0.58) & (0.52) & (0.57) & (0.58) & (0.80) \\   \hline \\[-1.8ex]  Observations & 150 & 150 & 150 & 150 & 150 & 150 \\  Residual Standard Error (GLS) & 4.69 & 5.04 & 4.79 & 5.63 & 5.18 & 4.88\\ \hline  \hline \\[-1.8ex]  \textit{Note:}  & \multicolumn{6}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\  \end{tabular}  \end{table}
+\begin{table}[!htbp] \centering    \caption{Estimated spatial error model parameters (standard errors) for each of the six election years.}    \label{tab:printmodels}  \scriptsize\tabcolsep=0.35cm \begin{tabular}{@{\extracolsep{1pt}}lrrrrrr} \\[-1.8ex]\hline  \hline \\[-1.8ex]  \\[-1.8ex] & \multicolumn{6}{c}{\vspace{-0.5cm}} \\   & 2001 & 2004 & 2007 & 2010 & 2013 & 2016 \\  \hline \\[-1.8ex]  $\rho$ & 0.46\rlap{$^{***}$} & 0.29\rlap{$^{*}$} & 0.24 & 0.19 & 0.27\rlap{$^{*}$} & 0.50\rlap{$^{***}$}\\  & (0.15\rlap{)} & (0.17\rlap{)} & (0.17\rlap{)} & (0.16\rlap{)} & (0.16\rlap{)} & (0.17\rlap{)}\\  AusCitizen & $-$3.13 & $-$2.64 & $-$2.53 & $-$0.08 & $-$3.40 & $-$1.80 \\    & (2.26\rlap{)} & (2.43\rlap{)} & (2.34\rlap{)} & (2.79\rlap{)} & (2.76\rlap{)} & (2.71\rlap{)} \\    BornAsia & 2.22 & $-$0.95 & $-$1.60 & $-$6.83\rlap{$^{**}$} & $-$3.03 & $-$0.55 \\    & (2.18\rlap{)} & (2.44\rlap{)} & (2.19\rlap{)} & (2.73\rlap{)} & (2.71\rlap{)} & (2.17\rlap{)} \\    Born\_MidEast & $-$1.15 & $-$1.59 & $-$2.01\rlap{$^{*}$} & $-$2.03 & $-$0.92 & $-$1.44 \\    & (1.07\rlap{)} & (1.20\rlap{)} & (1.11\rlap{)} & (1.27\rlap{)} & (1.24\rlap{)} & (1.13\rlap{)} \\    BornSEEuro & $-$3.21\rlap{$^{**}$} & $-$4.24\rlap{$^{***}$} & $-$3.61\rlap{$^{***}$} & $-$4.14\rlap{$^{***}$} & $-$3.69\rlap{$^{***}$} & $-$2.72\rlap{$^{***}$} \\    & (1.42\rlap{)} & (1.46\rlap{)} & (1.02\rlap{)} & (1.19\rlap{)} & (1.07\rlap{)} & (0.97\rlap{)} \\    Born\_UK & 0.25 & $-$0.07 & 0.34 & 0.56 & $-$0.09 & $-$1.32 \\    & (1.00\rlap{)} & (0.98\rlap{)} & (0.90\rlap{)} & (1.07\rlap{)} & (1.04\rlap{)} & (1.04\rlap{)} \\    BornElsewhere & $-$5.04 & $-$4.91 & $-$4.13 & 2.35 & $-$5.23 & $-$4.14 \\    & (3.30\rlap{)} & (3.68\rlap{)} & (3.38\rlap{)} & (4.23\rlap{)} & (4.15\rlap{)} & (3.97\rlap{)} \\    Buddhism & $-$0.49 & $-$0.17 & $-$1.37 & $-$0.83 & $-$0.12 & $-$1.60 \\    & (1.39\rlap{)} & (1.61\rlap{)} & (1.61\rlap{)} & (1.80\rlap{)} & (1.68\rlap{)} & (1.56\rlap{)} \\    Christianity & $-$2.48 & $-$1.23 & 0.38 & 0.50 & 2.41 & 1.68 \\    & (1.73\rlap{)} & (1.85\rlap{)} & (1.83\rlap{)} & (1.99\rlap{)} & (1.85\rlap{)} & (1.78\rlap{)} \\    CurrentlyStudying & $-$2.19\rlap{$^{**}$} & $-$0.13 & 2.06\rlap{$^{*}$} & 2.12\rlap{$^{*}$} & 1.15 & $-$0.16 \\    & (0.99\rlap{)} & (1.13\rlap{)} & (1.17\rlap{)} & (1.25\rlap{)} & (1.26\rlap{)} & (1.18\rlap{)} \\    DeFacto & $-$6.44\rlap{$^{***}$} & $-$5.37\rlap{$^{**}$} & $-$6.43\rlap{$^{***}$} & $-$8.07\rlap{$^{***}$} & $-$6.56\rlap{$^{**}$} & $-$8.53\rlap{$^{***}$} \\    & (1.87\rlap{)} & (2.48\rlap{)} & (2.31\rlap{)} & (3.06\rlap{)} & (3.11\rlap{)} & (2.83\rlap{)} \\    DiffAddress & 3.88\rlap{$^{***}$} & 5.06\rlap{$^{***}$} & 4.22\rlap{$^{***}$} & 5.57\rlap{$^{***}$} & 3.53\rlap{$^{*}$} & 5.67\rlap{$^{***}$} \\    & (0.94\rlap{)} & (1.12\rlap{)} & (0.99\rlap{)} & (1.76\rlap{)} & (1.91\rlap{)} & (1.60\rlap{)} \\    Distributive & 1.27 & 2.01\rlap{$^{*}$} & 1.36 & 1.57 & 2.10\rlap{$^{*}$} & 1.20 \\    & (1.12\rlap{)} & (1.21\rlap{)} & (1.13\rlap{)} & (1.34\rlap{)} & (1.27\rlap{)} & (1.21\rlap{)} \\    Education & 1.08 & 0.52 & $-$5.52\rlap{$^{*}$} & $-$4.08 & $-$4.44 & $-$7.07\rlap{$^{**}$} \\    & (2.38\rlap{)} & (3.12\rlap{)} & (3.27\rlap{)} & (3.95\rlap{)} & (3.78\rlap{)} & (3.55\rlap{)} \\    Extractive & 4.83\rlap{$^{***}$} & 5.45\rlap{$^{***}$} & 5.37\rlap{$^{***}$} & 7.31\rlap{$^{***}$} & 6.71\rlap{$^{***}$} & 7.43\rlap{$^{***}$} \\    & (1.48\rlap{)} & (1.42\rlap{)} & (1.36\rlap{)} & (1.56\rlap{)} & (1.47\rlap{)} & (1.39\rlap{)} \\    FamHouseSize & $-$0.16 & 0.87 & $-$2.40 & $-$2.53 & $-$3.26 & $-$2.91 \\    & (2.19\rlap{)} & (2.72\rlap{)} & (2.69\rlap{)} & (3.25\rlap{)} & (3.28\rlap{)} & (2.90\rlap{)} \\    Incomes & 4.36\rlap{$^{**}$} & 5.03\rlap{$^{*}$} & 9.45\rlap{$^{***}$} & 7.09\rlap{$^{**}$} & 7.97\rlap{$^{***}$} & 12.20\rlap{$^{***}$} \\    & (1.77\rlap{)} & (2.66\rlap{)} & (2.75\rlap{)} & (3.25\rlap{)} & (2.92\rlap{)} & (2.75\rlap{)} \\    Indigenous & 2.91\rlap{$^{*}$} & 1.97 & 2.48 & 2.84 & 0.67 & $-$0.05 \\    & (1.68\rlap{)} & (1.95\rlap{)} & (1.75\rlap{)} & (2.16\rlap{)} & (2.14\rlap{)} & (2.00\rlap{)} \\    Islam & $-$0.92 & $-$0.97 & $-$0.54 & $-$2.50 & $-$0.82 & $-$0.95 \\    & (1.22\rlap{)} & (1.36\rlap{)} & (1.27\rlap{)} & (1.52\rlap{)} & (1.42\rlap{)} & (1.34\rlap{)} \\    Judaism & 1.88\rlap{$^{*}$} & 1.78 & 2.66\rlap{$^{***}$} & 1.97\rlap{$^{*}$} & 2.74\rlap{$^{**}$} & 1.65\rlap{$^{*}$} \\    & (1.05\rlap{)} & (1.13\rlap{)} & (1.01\rlap{)} & (1.15\rlap{)} & (1.10\rlap{)} & (1.00\rlap{)} \\    ManagerAdmin & 2.06\rlap{$^{***}$} & 3.32\rlap{$^{***}$} & 6.00\rlap{$^{***}$} & 5.47\rlap{$^{***}$} & 5.04\rlap{$^{***}$} & 5.78\rlap{$^{***}$} \\    & (0.71\rlap{)} & (0.93\rlap{)} & (0.90\rlap{)} & (1.08\rlap{)} & (1.03\rlap{)} & (1.06\rlap{)} \\    Married & 0.44 & 0.11 & $-$1.22 & $-$0.22 & 0.91 & $-$2.34 \\    & (2.31\rlap{)} & (2.96\rlap{)} & (2.83\rlap{)} & (3.15\rlap{)} & (3.03\rlap{)} & (2.81\rlap{)} \\    MedianAge & 2.32\rlap{$^{*}$} & 4.96\rlap{$^{***}$} & 3.66\rlap{$^{**}$} & 4.00\rlap{$^{*}$} & 2.30 & 2.87 \\    & (1.32\rlap{)} & (1.65\rlap{)} & (1.81\rlap{)} & (2.26\rlap{)} & (2.08\rlap{)} & (1.79\rlap{)} \\    NoReligion & $-$1.57 & $-$0.92 & 0.56 & $-$0.30 & 1.02 & 1.31 \\    & (1.59\rlap{)} & (1.71\rlap{)} & (1.73\rlap{)} & (1.92\rlap{)} & (1.94\rlap{)} & (2.04\rlap{)} \\    OneParentHouse & $-$1.73 & $-$0.45 & $-$0.75 & $-$1.46 & $-$0.77 & $-$0.74 \\    & (1.36\rlap{)} & (1.59\rlap{)} & (1.49\rlap{)} & (1.69\rlap{)} & (1.57\rlap{)} & (1.47\rlap{)} \\    OtherLanguage & $-$0.44 & 5.92 & 9.98\rlap{$^{**}$} & 11.24\rlap{$^{**}$} & 9.00\rlap{$^{*}$} & 9.84\rlap{$^{**}$} \\    & (3.22\rlap{)} & (4.16\rlap{)} & (3.91\rlap{)} & (4.76\rlap{)} & (4.66\rlap{)} & (4.44\rlap{)} \\    PropertyOwned & $-$0.46 & $-$0.53 & 0.67 & $-$0.94 & $-$0.48 & 1.41 \\    & (1.37\rlap{)} & (1.50\rlap{)} & (1.43\rlap{)} & (1.76\rlap{)} & (1.67\rlap{)} & (1.50\rlap{)} \\    RentLoanPrice & $-$1.57 & $-$3.32\rlap{$^{*}$} & $-$4.01\rlap{$^{**}$} & $-$0.97 & $-$0.70 & $-$0.89 \\    & (1.49\rlap{)} & (1.76\rlap{)} & (1.67\rlap{)} & (2.07\rlap{)} & (2.07\rlap{)} & (2.16\rlap{)} \\    SocialServ & 2.51\rlap{$^{*}$} & 1.65 & 2.47\rlap{$^{*}$} & 2.53\rlap{$^{*}$} & 2.35\rlap{$^{*}$} & 4.45\rlap{$^{***}$} \\    & (1.33\rlap{)} & (1.41\rlap{)} & (1.29\rlap{)} & (1.47\rlap{)} & (1.32\rlap{)} & (1.19\rlap{)} \\    Transformative & 3.24\rlap{$^{**}$} & 4.73\rlap{$^{***}$} & 4.84\rlap{$^{***}$} & 4.46\rlap{$^{**}$} & 3.56\rlap{$^{**}$} & 4.58\rlap{$^{***}$} \\    & (1.55\rlap{)} & (1.78\rlap{)} & (1.74\rlap{)} & (1.98\rlap{)} & (1.78\rlap{)} & (1.53\rlap{)} \\    Unemployment & $-$2.45\rlap{$^{*}$} & $-$3.07\rlap{$^{*}$} & 0.29 & 0.08 & 1.67 & 2.79\rlap{$^{**}$} \\    & (1.40\rlap{)} & (1.63\rlap{)} & (1.51\rlap{)} & (1.76\rlap{)} & (1.51\rlap{)} & (1.37\rlap{)} \\    Constant & 50.81\rlap{$^{***}$} & 52.60\rlap{$^{***}$} & 47.31\rlap{$^{***}$} & 49.93\rlap{$^{***}$} & 53.51\rlap{$^{***}$} & 50.49\rlap{$^{***}$} \\    & (0.71\rlap{)} & (0.58\rlap{)} & (0.52\rlap{)} & (0.57\rlap{)} & (0.58\rlap{)} & (0.80\rlap{)} \\   \hline \\[-1.8ex]  Observations & 150 & 150 & 150 & 150 & 150 & 150 \\  Residual Standard Error (GLS) & 4.69 & 5.04 & 4.79 & 5.63 & 5.18 & 4.88\\ \hline  \hline \\[-1.8ex]  \textit{Note:}  & \multicolumn{6}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\  \end{tabular}  \end{table}
 
@@
 
-\caption{Partial residual plots for: DeFacto, MedianAge, Education, OtherLanguage, BornSEEuro, BornAsia and Judaism.}\label{fig:partresplot2}
+\caption{Partial residual plots by election year for a selection of predictors.
+Linear model with 95\% confidence bands is overlaid.
+Several predictors have a negative relationship: with larger values indicating the electorate more likely preferences Labor.
+Education exhibits a strong temporal change.}\label{fig:partresplot2}
 \end{figure}
@@
 
-In each of the six elections, electorates with a higher proportion of people that have recently (in the past five years) moved house are more likely to support the Liberal party, although this effect was marginally insignificant in 2013 (see row 6 in Figure \ref{fig:partresplot1}.
+In each of the six elections, electorates with a higher proportion of people that have recently (in the past five years) moved house (\texttt{DiffAddress}) are more likely to support the Liberal party, although this effect was marginally insignificant in 2013 (see row 6 in Figure \ref{fig:partresplot1}.
 Having controlled for characteristics of house ownership and rental prices (via the factors \texttt{PropertyOwned} and \texttt{RentLoan} respectively), this effect is somewhat surprising.
@@
 It is often hypothesized that states have systematic differences that cause their electorates to vote differently.
-Boxplots of residuals grouped by state (Figure \ref{fig:resstate}) reveal that only Tasmania, the Australian Capital Territory and the Northern Territory appear to have a state-specific effect not captured by the models.
+Boxplots of residuals grouped by state (Figure \ref{fig:resstate}) reveal that the data reflects this -- there appears to be a state-specific effect not captured by the models.
 Tasmania and the Australian Capital Territory appear to have a bias towards Labor, whereas the Northern Territory tends towards voting Liberal.
@@
 
-\caption{Boxplot of residuals by state with jittered points.}\label{fig:resstate}
+\caption{Boxplot of residuals by state with jittered points. States ordered by median residual. A state-specific bias not captured by the model is evident.}\label{fig:resstate}
 \end{figure}

### 2019-04-19 functions.R
make strip text smaller for partial residual plots

@@ -261,7 +261,9 @@ grid_visreg <- function(varname, plot = TRUE, myscale = "free_y", top = FALSE) {
         theme_bw() +
         labs(x = "", y = "") +
         facet_grid(varname ~ year, scales = "free_y") +
-        theme(plot.margin=unit(c(b = 0.05, l = 0, t = -0.2, r = 0),"cm"))
+        theme(plot.margin=unit(c(b = 0.05, l = 0, t = -0.2, r = 0),"cm"),
+          strip.text = element_text(size = 8),
+          axis.text = element_text(size = 8))
 
       if (top == FALSE) {
         return_object <- return_object + theme(strip.text.x = element_blank())

### 2019-04-19 draft-paper.Rmd
partial residual plot confidence band reference (visreg)

@@
 # Get spatial weights
+# This is not run, as weights are pulled from github, but this can be run to get them from scratch.
 sF_16 <- sF_download(2016)
@@
 ```{r FGLS, include = FALSE}
-# Instead of running spatial weights
+# Instead of running spatial weights, load from github
 load("data/sp_weights_01.rda")
@@
           "Population allocated to purple region: $\\hat{P}_{s,t}$",
-          "AusCitizen (\\%): $D_s$")) %>%
+          "AusCitizen (\\%): $D_s$"),
+    caption = "Population from each intersecting Census source zone (2016) that is allocated to the target zone (purple region - Hume electoral boundary in 2013), and the corresponding proportion of Australian citizens in each of these source zones.") %>%
   kableExtra::kable_styling(font_size=10, latex_options='scale_down')
@@
 ```{r prepplot, include = FALSE}
-p1 <- grid_visreg("Incomes", top = TRUE)
-p2 <- grid_visreg("Unemployment")
-p3 <- grid_visreg("Extractive")
+p1 <- grid_visreg("Extractive", top = TRUE)
+p2 <- grid_visreg("DiffAddress")
+p3 <- grid_visreg("ManagerAdmin")
 p4 <- grid_visreg("Transformative")
-p5 <- grid_visreg("ManagerAdmin")
-p6 <- grid_visreg("DiffAddress")
-p7 <- grid_visreg("DeFacto", top = TRUE)
-p8 <- grid_visreg("MedianAge")
-p9 <- grid_visreg("Education")
-p10 <- grid_visreg("OtherLanguage")
-p11 <- grid_visreg("BornSEEuro")
-p12 <- grid_visreg("BornAsia")
-p13 <- grid_visreg("Judaism")
+p5 <- grid_visreg("Incomes")
+p6 <- grid_visreg("Unemployment")
+
+p7 <- grid_visreg("MedianAge", top = TRUE)
+p8 <- grid_visreg("Judaism")
+p9 <- grid_visreg("OtherLanguage")
+temp <- grid_visreg("OtherLanguage")
+p10 <- grid_visreg("BornAsia")
+p11 <- grid_visreg("Education")
+p12 <- grid_visreg("BornSEEuro")
+p13 <- grid_visreg("DeFacto")
 ```
@@
 Education exhibits a strong temporal change.", fig.height = 8.5,  fig.width = 6, warning = FALSE, fig.pos = 'h'}
-(p7 / p8 / (p9 + ylab("Two-party preferred vote (%)                     ")) / p10 / p11 / p12)
+(p7 / p8 / (p9 + ylab("Two-party preferred vote (%)                     ")) / p10 / p11 / p12 / p13)
 ```
@@
 These show the direction, size and significance of an estimated effect --- the slope of the prediction line matches the estimated coefficient, and the shaded region represents a 95% confidence band.
+The confidence bands around each prediction line are constructed in the usual way, as described in @visreg.
 If a horizontal line can be drawn through the confidence band, then the effect is insignificant.
@@
 Plots for each election are faceted to compare the effects over time in Figures \@ref(fig:partresplot1) and \@ref(fig:partresplot2).
-
-It is important here to note the ecological fallacy: insights are being drawn at the electorate level, and cannot be inferred for another disaggregate level (e.g., individual voters).
+Only socio-demographics that have a significant effect in at least one election are displayed in Figures \@ref(fig:partresplot1) and \@ref(fig:partresplot2).
+
+It is important here to note the ecological fallacy: insights are being drawn at the electorate level, and cannot be inferred for another disaggregate level (e.g. individual voters).
 

### 2019-04-19 draft-paper.tex
partial residual plot confidence band reference (visreg)

@@
 
-% Align tabular column on decimal point
-\usepackage{dcolumn}
-\newcolumntype{d}[1]{D{.}{\cdot}{#1}}
-
 \begin{document}
@@
 
-\begin{table}[H]
-\centering\begingroup\fontsize{10}{12}\selectfont
+\begin{table}[t]
 
+\caption{\label{tab:hume1}Population from each intersecting Census source zone (2016) that is allocated to the target zone (purple region - Hume electoral boundary in 2013), and the corresponding proportion of Australian citizens in each of these source zones.}
+\centering
 \resizebox{\linewidth}{!}{
+\fontsize{10}{12}\selectfont
 \begin{tabular}{lrrrr}
@@
 \end{tabular}}
-\endgroup{}
 \end{table}
@@
 These show the direction, size and significance of an estimated effect --- the slope of the prediction line matches the estimated coefficient, and the shaded region represents a 95\% confidence band.
+The confidence bands around each prediction line are constructed in the usual way, as described in \textcite{visreg}.
 If a horizontal line can be drawn through the confidence band, then the effect is insignificant.
@@
 Plots for each election are faceted to compare the effects over time in Figures \ref{fig:partresplot1} and \ref{fig:partresplot2}.
+Only socio-demographics that have a significant effect in at least one election are displayed in Figures \ref{fig:partresplot1} and \ref{fig:partresplot2}.
 
-It is important here to note the ecological fallacy: insights are being drawn at the electorate level, and cannot be inferred for another disaggregate level (e.g., individual voters).
+It is important here to note the ecological fallacy: insights are being drawn at the electorate level, and cannot be inferred for another disaggregate level (e.g.~individual voters).
 

### 2019-04-19 draft-paper.Rmd
change repo name

@@
 This paper was produced using `RMarkdown` [@rmarkdown] and `knitr` [@knitr].
-All corresponding code for this paper can be found in the github repository [github.com/jforbes14/journal-writing](https://github.com/jforbes14/journal-writing), and the data used is available in the `eechidna` package [@eechidna].
+All corresponding code for this paper can be found in the github repository [github.com/jforbes14/eechidna-paper](https://github.com/jforbes14/eechidna-paper), and the data used is available in the `eechidna` package [@eechidna].
 All raw data was obtained from the Australian Electoral Commission, the Australian Bureau of Statistics and the Australian Government.

### 2019-04-19 draft-paper.tex
change repo name

@@
 This paper was produced using \texttt{RMarkdown} \autocite{rmarkdown} and \texttt{knitr} \autocite{knitr}.
-All corresponding code for this paper can be found in the github repository \href{https://github.com/jforbes14/journal-writing}{github.com/jforbes14/journal-writing}, and the data used is available in the \texttt{eechidna} package \autocite{eechidna}.
+All corresponding code for this paper can be found in the github repository \href{https://github.com/jforbes14/eechidna-paper}{github.com/jforbes14/eechidna-paper}, and the data used is available in the \texttt{eechidna} package \autocite{eechidna}.
 All raw data was obtained from the Australian Electoral Commission, the Australian Bureau of Statistics and the Australian Government.

### 2019-04-20 draft-paper.Rmd
tweak on language in paper, and submission letter

@@
 ```{r partresplot1, fig.cap = "Partial residual plots by election year for a selection of predictors.
-Linear model with 95\\% confidence bands is overlaid.
+Linear model with 95\\% confidence bands overlaid.
 Most predictors have a positive relationship: the larger the value the more likely the electorate preferences the Coalition.
@@
 ```{r partresplot2, fig.cap = "Partial residual plots by election year for a selection of predictors.
-Linear model with 95\\% confidence bands is overlaid.
+Linear model with 95\\% confidence bands overlaid.
 Several predictors have a negative relationship: with larger values indicating the electorate more likely preferences Labor.
-Education exhibits a strong temporal change.", fig.height = 8.5,  fig.width = 6, warning = FALSE, fig.pos = 'h'}
+Most relationships are relatively stable over elections, except OtherLanguage and Education.", fig.height = 8.5,  fig.width = 6, warning = FALSE, fig.pos = 'h'}
 (p7 / p8 / (p9 + ylab("Two-party preferred vote (%)                     ")) / p10 / p11 / p12 / p13)
@@
 The partial residuals are the residuals from the fitted model with the estimated effect an individual variable added.
-These show the direction, size and significance of an estimated effect --- the slope of the prediction line matches the estimated coefficient, and the shaded region represents a 95% confidence band.
-The confidence bands around each prediction line are constructed in the usual way, as described in @visreg.
+These show the direction, size and significance of an estimated effect --- the slope of the prediction line matches the estimated coefficient, and the shaded region represents a 95% confidence band, computed using the method in @visreg.
 If a horizontal line can be drawn through the confidence band, then the effect is insignificant.

### 2019-04-20 draft-paper.tex
tweak on language in paper, and submission letter

@@
 \caption{Partial residual plots by election year for a selection of predictors.
-Linear model with 95\% confidence bands is overlaid.
+Linear model with 95\% confidence bands overlaid.
 Most predictors have a positive relationship: the larger the value the more likely the electorate preferences the Coalition.
@@
 \caption{Partial residual plots by election year for a selection of predictors.
-Linear model with 95\% confidence bands is overlaid.
+Linear model with 95\% confidence bands overlaid.
 Several predictors have a negative relationship: with larger values indicating the electorate more likely preferences Labor.
-Education exhibits a strong temporal change.}\label{fig:partresplot2}
+Most relationships are relatively stable over elections, except OtherLanguage and Education.}\label{fig:partresplot2}
 \end{figure}
 
-To investigate the socio-demographics that have a strong effect on the two-party preferred vote, partial residual plots are used and shown in Figures \ref{fig:partresplot1} and \ref{fig:partresplot2}.
-The partial residuals are the residuals from the fitted model with the estimated effect an individual variable added.
-These show the direction, size and significance of an estimated effect --- the slope of the prediction line matches the estimated coefficient, and the shaded region represents a 95\% confidence band.
-The confidence bands around each prediction line are constructed in the usual way, as described in \textcite{visreg}.
-If a horizontal line can be drawn through the confidence band, then the effect is insignificant.
-The estimated intercept is also added to the partial residuals for interpretability.
-Plots for each election are faceted to compare the effects over time in Figures \ref{fig:partresplot1} and \ref{fig:partresplot2}.
-Only socio-demographics that have a significant effect in at least one election are displayed in Figures \ref{fig:partresplot1} and \ref{fig:partresplot2}.
+To investigate the socio-demographics that have a strong effect on the
+two-party preferred vote, partial residual plots are used and shown in
+Figures \ref{fig:partresplot1} and \ref{fig:partresplot2}. The partial
+residuals are the residuals from the fitted model with the estimated
+effect an individual variable added. These show the direction, size and
+significance of an estimated effect --- the slope of the prediction line
+matches the estimated coefficient, and the shaded region represents a
+95\% confidence band, computed using the method in \textcite{visreg}. If
+a horizontal line can be drawn through the confidence band, then the
+effect is insignificant. The estimated intercept is also added to the
+partial residuals for interpretability. Plots for each election are
+faceted to compare the effects over time in Figures
+\ref{fig:partresplot1} and \ref{fig:partresplot2}. Only
+socio-demographics that have a significant effect in at least one
+election are displayed in Figures \ref{fig:partresplot1} and
+\ref{fig:partresplot2}.
 
-It is important here to note the ecological fallacy: insights are being drawn at the electorate level, and cannot be inferred for another disaggregate level (e.g.~individual voters).
+It is important here to note the ecological fallacy: insights are being
+drawn at the electorate level, and cannot be inferred for another
+disaggregate level (e.g.~individual voters).
 
-\hypertarget{income-and-unemployment}{%
-\subsubsection{Income and unemployment}\label{income-and-unemployment}}
+\subsubsection{Income and unemployment}\label{income-and-unemployment}
 
-Typically the Labor party campaigns on more progressive policies, which often include tax reform that adversely affects higher income earners, and more generous social assistance programs.
-Perhaps it is due to these policies that higher income electorates appear more likely to support the Liberal party, as the \texttt{Incomes} factor has a positive effect on Liberal preference (see row 1 in Figure \ref{fig:partresplot1}).
-This effect is significant in every election aside from 2004, where it is only marginally insignificant (\(p = 0.0613\)).
-Unemployment however, is not as influential.
-In 2001 and 2004, electorates with higher unemployment align with Labor, but over time this shifts towards support for the Liberal party, culminating in a significantly positive effect in 2016.
+Typically the Labor party campaigns on more progressive policies, which
+often include tax reform that adversely affects higher income earners,
+and more generous social assistance programs. Perhaps it is due to these
+policies that higher income electorates appear more likely to support
+the Liberal party, as the \texttt{Incomes} factor has a positive effect
+on Liberal preference (see row 1 in Figure \ref{fig:partresplot1}). This
+effect is significant in every election aside from 2004, where it is
+only marginally insignificant (\(p = 0.0613\)). Unemployment however, is
+not as influential. In 2001 and 2004, electorates with higher
+unemployment align with Labor, but over time this shifts towards support
+for the Liberal party, culminating in a significantly positive effect in
+2016.
 
-\hypertarget{industry-and-type-of-work}{%
-\subsubsection{Industry and type of work}\label{industry-and-type-of-work}}
+\subsubsection{Industry and type of
+work}\label{industry-and-type-of-work}
 
-Electorates with higher proportions of workers in mining, gas, water, agriculture, waste and electricity (grouped as \texttt{Extractive} industries) are consistently linked with higher support for the Liberal party, with the magnitude of this effect slightly increasing over the years (see row 3 in Figure \ref{fig:partresplot1}).
-This is unsurprising, as the Liberal party has close ties with these traditional energy industries, and typically present policies to reduce taxation on energy production.
-Furthermore, electorates with more workers in construction or manufacturing industries (\texttt{Transformative}) are also more likely to support the Liberal party (see row 4 in Figure \ref{fig:partresplot1}).
+Electorates with higher proportions of workers in mining, gas, water,
+agriculture, waste and electricity (grouped as \texttt{Extractive}
+industries) are consistently linked with higher support for the Liberal
+party, with the magnitude of this effect slightly increasing over the
+years (see row 3 in Figure \ref{fig:partresplot1}). This is
+unsurprising, as the Liberal party has close ties with these traditional
+energy industries, and typically present policies to reduce taxation on
+energy production. Furthermore, electorates with more workers in
+construction or manufacturing industries (\texttt{Transformative}) are
+also more likely to support the Liberal party (see row 4 in Figure
+\ref{fig:partresplot1}).
 
-Similarly, the proportion of workers in managerial, administrative, clerical and sales roles (\texttt{ManagerAdmin}) is also a significant predictor of two-party preference vote across all six elections, with a higher proportion of people working these jobs increasing Liberal support.
-The magnitude of this effect also seems to increase over the years.
+Similarly, the proportion of workers in managerial, administrative,
+clerical and sales roles (\texttt{ManagerAdmin}) is also a significant
+predictor of two-party preference vote across all six elections, with a
+higher proportion of people working these jobs increasing Liberal
+support. The magnitude of this effect also seems to increase over the
+years.
 
-\hypertarget{household-mobility}{%
-\subsubsection{Household mobility}\label{household-mobility}}
+\subsubsection{Household mobility}\label{household-mobility}
 
-In each of the six elections, electorates with a higher proportion of people that have recently (in the past five years) moved house (\texttt{DiffAddress}) are more likely to support the Liberal party, although this effect was marginally insignificant in 2013 (see row 6 in Figure \ref{fig:partresplot1}.
-Having controlled for characteristics of house ownership and rental prices (via the factors \texttt{PropertyOwned} and \texttt{RentLoan} respectively), this effect is somewhat surprising.
+In each of the six elections, electorates with a higher proportion of
+people that have recently (in the past five years) moved house
+(\texttt{DiffAddress}) are more likely to support the Liberal party,
+although this effect was marginally insignificant in 2013 (see row 6 in
+Figure \ref{fig:partresplot1}. Having controlled for characteristics of
+house ownership and rental prices (via the factors
+\texttt{PropertyOwned} and \texttt{RentLoan} respectively), this effect
+is somewhat surprising.
 
-\hypertarget{relationships}{%
-\subsubsection{Relationships}\label{relationships}}
+\subsubsection{Relationships}\label{relationships}
 
-De facto relationships, but not marriages, are found to be an important (and significant) predictor of the two-party preferred vote in all six elections, with more de facto relationships associated with higher support for the Labor party.
-The proportion of individuals who are married however, is insignificant (not shown).
+De facto relationships, but not marriages, are found to be an important
+(and significant) predictor of the two-party preferred vote in all six
+elections, with more de facto relationships associated with higher
+support for the Labor party. The proportion of individuals who are
+married however, is insignificant (not shown).
 
-\hypertarget{age}{%
-\subsubsection{Age}\label{age}}
+\subsubsection{Age}\label{age}
 
-Regions comprising more older people are often believed to be more conservative, and indeed it found that electorates with a higher median age are more likely to support the Liberal party --- although this effect is only significant in 2007 and 2010 (see row 2 in Figure \ref{fig:partresplot2}).
+Regions comprising more older people are often believed to be more
+conservative, and indeed it found that electorates with a higher median
+age are more likely to support the Liberal party --- although this
+effect is only significant in 2007 and 2010 (see row 2 in Figure
+\ref{fig:partresplot2}).
 
-\hypertarget{education}{%
-\subsubsection{Education}\label{education}}
+\subsubsection{Education}\label{education}
 
-Since 2007, electorates with higher education levels are associated with supporting the Labor party, although this effect is only significant in 2016.
-Before 2007, education has an almost zero effect (see row 3 in Figure \ref{fig:partresplot1}).
+Since 2007, electorates with higher education levels are associated with
+supporting the Labor party, although this effect is only significant in
+2016. Before 2007, education has an almost zero effect (see row 3 in
+Figure \ref{fig:partresplot1}).
 
-\hypertarget{diversity}{%
-\subsubsection{Diversity}\label{diversity}}
+\subsubsection{Diversity}\label{diversity}
 
-Larger migrant populations from Asia, the Middle East, South-Eastern Europe, the United Kingdom and elsewhere, are either associated with Labor support, or have no effect.
-Of these areas, only South-Eastern European populations appear significant in each election, with the proportion of Asian migrants also being significant in 2010.
-Speaking other languages (aside from English) however, appears to have a far stronger effect, as observed through the \texttt{OtherLanguage} variable.
-Electorates with more diverse speech are associated with higher support for the Liberal party from 2004 onwards, with this effect being significant in 2007, 2010 and 2016.
-Furthermore, of the variables relating to religion, only Judaism shows a consistent effect, with electorates with relatively large Jewish populations more likely to vote Liberal.
+Larger migrant populations from Asia, the Middle East, South-Eastern
+Europe, the United Kingdom and elsewhere, are either associated with
+Labor support, or have no effect. Of these areas, only South-Eastern
+European populations appear significant in each election, with the
+proportion of Asian migrants also being significant in 2010. Speaking
+other languages (aside from English) however, appears to have a far
+stronger effect, as observed through the \texttt{OtherLanguage}
+variable. Electorates with more diverse speech are associated with
+higher support for the Liberal party from 2004 onwards, with this effect
+being significant in 2007, 2010 and 2016. Furthermore, of the variables
+relating to religion, only Judaism shows a consistent effect, with
+electorates with relatively large Jewish populations more likely to vote
+Liberal.
 
-\hypertarget{a-note-on-similar-variables}{%
-\subsubsection{A note on similar variables}\label{a-note-on-similar-variables}}
+\subsubsection{A note on similar
+variables}\label{a-note-on-similar-variables}
 
-Many of the Census variables represent similar information, which is why factors were created and some variables were removed.
-However, some variables remain which are closely related.
-For example, an electorate's income level (via \texttt{Incomes}) is likely to be related to electoral unemployment and labor force participation (via \texttt{Unemployment}).
-In 2001, the coefficient estimate for \texttt{Unemployment} is negative but not significant, whilst the \texttt{Incomes} variables is significant.
-If the \texttt{Incomes} variable is removed from the model in 2001, \texttt{Unemployment} absorbs the negative effect, becoming significant (\(p = 0.0056\)).
+Many of the Census variables represent similar information, which is why
+factors were created and some variables were removed. However, some
+variables remain which are closely related. For example, an electorate's
+income level (via \texttt{Incomes}) is likely to be related to electoral
+unemployment and labor force participation (via \texttt{Unemployment}).
+In 2001, the coefficient estimate for \texttt{Unemployment} is negative
+but not significant, whilst the \texttt{Incomes} variables is
+significant. If the \texttt{Incomes} variable is removed from the model
+in 2001, \texttt{Unemployment} absorbs the negative effect, becoming
+significant (\(p = 0.0056\)).
 
-\hypertarget{a-closer-look-at-the-residuals}{%
-\subsection{A closer look at the residuals}\label{a-closer-look-at-the-residuals}}
+\subsection{A closer look at the
+residuals}\label{a-closer-look-at-the-residuals}
 
-\hypertarget{residuals-by-state}{%
-\subsubsection{Residuals by state}\label{residuals-by-state}}
+\subsubsection{Residuals by state}\label{residuals-by-state}
 
-It is often hypothesized that states have systematic differences that cause their electorates to vote differently.
-Boxplots of residuals grouped by state (Figure \ref{fig:resstate}) reveal that the data reflects this -- there appears to be a state-specific effect not captured by the models.
-Tasmania and the Australian Capital Territory appear to have a bias towards Labor, whereas the Northern Territory tends towards voting Liberal.
-However, there are relatively few electorates in each of these states (five, two and two respectively), so this apparent result may be due to incumbent effects rather than an actual state-specific bias.
+It is often hypothesized that states have systematic differences that
+cause their electorates to vote differently. Boxplots of residuals
+grouped by state (Figure \ref{fig:resstate}) reveal that the data
+reflects this -- there appears to be a state-specific effect not
+captured by the models. Tasmania and the Australian Capital Territory
+appear to have a bias towards Labor, whereas the Northern Territory
+tends towards voting Liberal. However, there are relatively few
+electorates in each of these states (five, two and two respectively), so
+this apparent result may be due to incumbent effects rather than an
+actual state-specific bias.
 
@@
 
-\hypertarget{outlier-electorates}{%
-\subsubsection{Outlier electorates}\label{outlier-electorates}}
+\subsubsection{Outlier electorates}\label{outlier-electorates}
 
-Based on the distribution of the Cook's distance values, a Cook's distance greater than \(0.1\) is considered to be influential and a potential outlier.
-The electorate of Sydney (NSW) has a large Cook's distance from 2001 to 2013, due to its diverse population (language, birthplace and religion), high number of defacto relationships, high income, high household mobility and small amount of workers in extractive and transformative jobs.
-It has remained a strong supporter of the Labor party and the Liberal vote is severely overpredicted by the model, making it an outlier.
-Nearby in metropolitan NSW, the electorate of Wentworth is found to be an outlier in all but the 2007 election.
-Although historically Liberal, its two-party vote jumped by over 10 percentage points in 2010 without experiencing any notable changes in its socio-demographic makeup --- implying that this may be the direct effect of its Liberal member, Malcolm Turnbull, becoming the leader of the Liberal party.
-Liberal support in Wentworth is underpredicted by the model in each year, and more so with Turnbull as Liberal leader.
+Based on the distribution of the Cook's distance values, a Cook's
+distance greater than \(0.1\) is considered to be influential and a
+potential outlier. The electorate of Sydney (NSW) has a large Cook's
+distance from 2001 to 2013, due to its diverse population (language,
+birthplace and religion), high number of defacto relationships, high
+income, high household mobility and small amount of workers in
+extractive and transformative jobs. It has remained a strong supporter
+of the Labor party and the Liberal vote is severely overpredicted by the
+model, making it an outlier. Nearby in metropolitan NSW, the electorate
+of Wentworth is found to be an outlier in all but the 2007 election.
+Although historically Liberal, its two-party vote jumped by over 10
+percentage points in 2010 without experiencing any notable changes in
+its socio-demographic makeup --- implying that this may be the direct
+effect of its Liberal member, Malcolm Turnbull, becoming the leader of
+the Liberal party. Liberal support in Wentworth is underpredicted by the
+model in each year, and more so with Turnbull as Liberal leader.
 
-Lingiari, an electorate taking up almost all of the Northern Territory, is an outlier in the 2001--2007 elections due to its large Indigenous population, young age profile and low rates of property ownership.
-Fowler (NSW) has a diverse population with a high proportion of migrants, many Buddhists and Muslims, and has strong Labor support, making it influential in 2001, 2004 and 2010.
-Other electorates with large Cook's distance are Barton (NSW) and Leichhardt (QLD) in 2016, and Canberra (ACT) in 2007.
+Lingiari, an electorate taking up almost all of the Northern Territory,
+is an outlier in the 2001--2007 elections due to its large Indigenous
+population, young age profile and low rates of property ownership.
+Fowler (NSW) has a diverse population with a high proportion of
+migrants, many Buddhists and Muslims, and has strong Labor support,
+making it influential in 2001, 2004 and 2010. Other electorates with
+large Cook's distance are Barton (NSW) and Leichhardt (QLD) in 2016, and
+Canberra (ACT) in 2007.
 
-\hypertarget{conclusion}{%
-\section{Conclusion}\label{conclusion}}
+\section{Conclusion}\label{conclusion}
 
-This paper explores the effects of electoral socio-demographic characteristics on the two-party preferred vote in the 2001--2016 elections, using information from the corresponding Australian federal elections and Censuses.
-As a Census does not always occur in the same year as an election, Census data for the 2004--2013 elections are generated by employing a method of spatio-temporal imputation.
-This imputes electoral socio-demographics for the electoral boundaries in place at the time of the election --- an approach that is distinctly different from previous work on modelling election outcomes, where Census and election data are typically joined without addressing their temporal differences.
-Before estimating a model, these socio-demographic variables are standardized (to adjust for changing variable scales) and many variables (representing similar information) are combined into factors, resulting in a reduced predictor set.
-A spatial error model is then estimated for each election, accounting for the inherent spatial structure of the data.
+This paper explores the effects of electoral socio-demographic
+characteristics on the two-party preferred vote in the 2001--2016
+elections, using information from the corresponding Australian federal
+elections and Censuses. As a Census does not always occur in the same
+year as an election, Census data for the 2004--2013 elections are
+generated by employing a method of spatio-temporal imputation. This
+imputes electoral socio-demographics for the electoral boundaries in
+place at the time of the election --- an approach that is distinctly
+different from previous work on modelling election outcomes, where
+Census and election data are typically joined without addressing their
+temporal differences. Before estimating a model, these socio-demographic
+variables are standardized (to adjust for changing variable scales) and
+many variables (representing similar information) are combined into
+factors, resulting in a reduced predictor set. A spatial error model is
+then estimated for each election, accounting for the inherent spatial
+structure of the data.
 
-Across the past six elections, most of the socio-demographics that drive the electoral two-party preferred vote are found to remain steady, whilst a few (typically weaker) effects vary over time.
-Industry and type of work are particularly influential, with energy-related and manufacturing/construction jobs, as well as administrative roles being strongly linked with the Liberal party in all elections.
-Incomes have a similarly consistent effect, with higher income areas supporting Liberal.
-Higher levels of unemployment shift from weak association with Labor to a significant Liberal effect over the years, and higher education levels are associated with Labor from 2007 (although only significant in 2016).
-It is also found that electorates with higher household mobility support Liberal, birthplace diversity favours Labor and more de facto relationships align with Labor preference --- although marriages, family and household sizes have no material influence.
-Furthermore, the neighbourhood (spatial) effects are found to be positive in all elections, although only significant in 2001 and 2016, meaning that in the 2004--2013 elections, electorates effectively voted independently.
+Across the past six elections, most of the socio-demographics that drive
+the electoral two-party preferred vote are found to remain steady,
+whilst a few (typically weaker) effects vary over time. Industry and
+type of work are particularly influential, with energy-related and
+manufacturing/construction jobs, as well as administrative roles being
+strongly linked with the Liberal party in all elections. Incomes have a
+similarly consistent effect, with higher income areas supporting
+Liberal. Higher levels of unemployment shift from weak association with
+Labor to a significant Liberal effect over the years, and higher
+education levels are associated with Labor from 2007 (although only
+significant in 2016). It is also found that electorates with higher
+household mobility support Liberal, birthplace diversity favours Labor
+and more de facto relationships align with Labor preference --- although
+marriages, family and household sizes have no material influence.
+Furthermore, the neighbourhood (spatial) effects are found to be
+positive in all elections, although only significant in 2001 and 2016,
+meaning that in the 2004--2013 elections, electorates effectively voted
+independently.
 
-The findings in this paper complement the existing literature by modelling temporal trends, which as far as the authors are aware, has not been done previously for Australian elections using a regression framework.
-It is also the first study to model any Australian election since 2010 using Census information.
+The findings in this paper complement the existing literature by
+modelling temporal trends, which as far as the authors are aware, has
+not been done previously for Australian elections using a regression
+framework. It is also the first study to model any Australian election
+since 2010 using Census information.
 
-Additionally, a key contribution of this research is the wrangling of the raw data and imputed data sets for the 2004, 2007, 2010 and 2013 elections, which have been contributed to the \texttt{eechidna} \texttt{R} package --- providing a rich, accessible data resource for future Australian electoral analysis.
+Additionally, a key contribution of this research is the wrangling of
+the raw data and imputed data sets for the 2004, 2007, 2010 and 2013
+elections, which have been contributed to the \texttt{eechidna}
+\texttt{R} package --- providing a rich, accessible data resource for
+future Australian electoral analysis.
 
-\hypertarget{acknowledgements}{%
-\section{Acknowledgements}\label{acknowledgements}}
+\section{Acknowledgements}\label{acknowledgements}
 
-This paper was produced using \texttt{RMarkdown} \autocite{rmarkdown} and \texttt{knitr} \autocite{knitr}.
-All corresponding code for this paper can be found in the github repository \href{https://github.com/jforbes14/eechidna-paper}{github.com/jforbes14/eechidna-paper}, and the data used is available in the \texttt{eechidna} package \autocite{eechidna}.
-All raw data was obtained from the Australian Electoral Commission, the Australian Bureau of Statistics and the Australian Government.
+This paper was produced using \texttt{RMarkdown} \autocite{rmarkdown}
+and \texttt{knitr} \autocite{knitr}. All corresponding code for this
+paper can be found in the github repository
+\href{https://github.com/jforbes14/eechidna-paper}{github.com/jforbes14/eechidna-paper},
+and the data used is available in the \texttt{eechidna} package
+\autocite{eechidna}. All raw data was obtained from the Australian
+Electoral Commission, the Australian Bureau of Statistics and the
+Australian Government.
 
-\hypertarget{software}{%
-\section{Software}\label{software}}
+\section{Software}\label{software}
 
-All election and Census datasets, along with electoral maps and more, are available in the \texttt{eechidna} (Exploring Election and Census Highly Informative Data Nationally for Australia) \texttt{R} package, which can be downloaded from CRAN.
-The \texttt{eechidna} package makes it easy to look at the data from the Australian Federal elections and Censuses that occurred between 2001 and 2016.
-This study contributed a large revision to the \texttt{eechidna} package, which included the addition of election and Census data for 2001--2010, voting outcomes for polling booths and imputed Census data for election years.
-For more details on using \texttt{eechidna}, please see the articles (vignettes) on the github page \href{https://ropenscilabs.github.io/eechidna/}{ropenscilabs.github.io/eechidna/}.
+All election and Census datasets, along with electoral maps and more,
+are available in the \texttt{eechidna} (Exploring Election and Census
+Highly Informative Data Nationally for Australia) \texttt{R} package,
+which can be downloaded from CRAN. The \texttt{eechidna} package makes
+it easy to look at the data from the Australian Federal elections and
+Censuses that occurred between 2001 and 2016. This study contributed a
+large revision to the \texttt{eechidna} package, which included the
+addition of election and Census data for 2001--2010, voting outcomes for
+polling booths and imputed Census data for election years. For more
+details on using \texttt{eechidna}, please see the articles (vignettes)
+on the github page
+\href{https://ropenscilabs.github.io/eechidna/}{ropenscilabs.github.io/eechidna/}.
 
-The authors would like to sincerely thank Anthony Ebert, Heike Hofmann, Thomas Lumley, Ben Marwick, Carson Sievert, Mingzhu Sun, Dilini Talagala, Nicholas Tierney, Nathaniel Tomasetti, Earo Wang and Fang Zhou, all of whom have contributed to the \texttt{eechidna} package.
+The authors would like to sincerely thank Anthony Ebert, Heike Hofmann,
+Thomas Lumley, Ben Marwick, Carson Sievert, Mingzhu Sun, Dilini
+Talagala, Nicholas Tierney, Nathaniel Tomasetti, Earo Wang and Fang
+Zhou, all of whom have contributed to the \texttt{eechidna} package.
 

### 2019-04-20 paper.Rmd
almost have format for journal together

--- 2019-04-20 draft-paper.Rmd
@@
 title: "Spatial modelling of the two-party preferred vote in Australian federal elections: 2001--2016"
-author:
-- familyname: Forbes
-  othernames: Jeremy
-  address: Monash University\newline Melbourne, Australia
-  email: jeremy.forbes@monash.edu
-  correspondingauthor: true
-- familyname: Cook
-  othernames: Dianne
-  address: Monash University\newline Melbourne, Australia
-- familyname: Hyndman
-  othernames: Rob J
-  address: Monash University\newline Melbourne, Australia
-abstract: "We examine the relationships between electoral socio-demographic characteristics and two-party preference in the six Australian federal elections held between 2001 to 2016.
+author: "Jeremy Forbes*, Dianne Cook, Rob J Hyndman Department of Econometrics and Business Statistics, Monash University"
+wpnumber:
+jelcodes: C31, C33, D72
+blind: false
+cover: false
+toc: false
+documentclass: anzsauth
+bibliography: references.bib
+biblio-style: anzsj
+classoption: times,doublespace
+header-includes: \usepackage{float, bm}
+output: 
+  pdf_document:
+    latex_engine: xelatex
+    fig_caption: yes
+    keep_tex: yes
+    number_sections: yes
+    fig_height: 5
+    fig_width: 8
+---
+
+# Summary{-}
+
+We examine the relationships between electoral socio-demographic characteristics and two-party preference in the six Australian federal elections held between 2001 to 2016.
 Socio-demographic information is derived from the Australian Census, which occurs every five years.
@@
 It is also found that between 2004 and 2013, the spatial effect was insignificant, meaning that electorates voted effectively independently.
-All data featured in this study has been contributed to the `eechidna` `R` package (available on CRAN)."
-keywords: "federal election, Census, Australia, spatial modelling, imputation, socio-demographics, electorates, R, eechidna"
-wpnumber:
-jelcodes: C31, C33, D72
-blind: false
-cover: false
-toc: false
-bibliography: references.bib
-biblio-style: authoryear-comp
-header-includes: \usepackage{float}
-output:
-  MonashEBSTemplates::workingpaper:
-    fig_caption: yes
-    fig_height: 5
-    fig_width: 8
-    includes:
-      in_header: preamble.tex
-    keep_tex: yes
-    number_sections: yes
-    citation_package: biblatex
----
+All data featured in this study has been contributed to the `eechidna` `R` package (available on CRAN).
+
+**Keywords:** federal election, Census, Australia, spatial modelling, imputation, socio-demographics, electorates, R, eechidna
 
@@
 The partial residuals are the residuals from the fitted model with the estimated effect an individual variable added.
-These show the direction, size and significance of an estimated effect --- the slope of the prediction line matches the estimated coefficient, and the shaded region represents a 95% confidence band, computed using the method in @visreg.
+These show the direction, size and significance of an estimated effect --- the slope of the prediction line matches the estimated coefficient, and the shaded region represents a 95% confidence band, computed using the method  @visreg.
 If a horizontal line can be drawn through the confidence band, then the effect is insignificant.

### 2019-04-21 draft-paper.Rmd
Updated Jeremy's email address

@@
   address: Monash University\newline Melbourne, Australia
-  email: jeremy.forbes@monash.edu
+  email: jeremyforbes1995@gmail.com
   correspondingauthor: true

### 2019-04-21 draft-paper.tex
Removed duplicate bib entry

--- 2019-04-19 draft-paper.tex
@@
 Monash University\newline Melbourne, Australia
-\newline{Email: \href{mailto:jeremy.forbes@monash.edu}{\nolinkurl{jeremy.forbes@monash.edu}}}\newline Corresponding author\\[1cm]
+\newline{Email: \href{mailto:jeremyforbes1995@gmail.com}{\nolinkurl{jeremyforbes1995@gmail.com}}}\newline Corresponding author\\[1cm]
 \textbf{Dianne Cook}\newline
@@
 \caption{Partial residual plots by election year for a selection of predictors.
-Linear model with 95\% confidence bands is overlaid.
+Linear model with 95\% confidence bands overlaid.
 Several predictors have a negative relationship: with larger values indicating the electorate more likely preferences Labor.
-Education exhibits a strong temporal change.}\label{fig:partresplot2}
+Most relationships are relatively stable over elections, except OtherLanguage and Education.}\label{fig:partresplot2}
 \end{figure}
@@
 The partial residuals are the residuals from the fitted model with the estimated effect an individual variable added.
-These show the direction, size and significance of an estimated effect --- the slope of the prediction line matches the estimated coefficient, and the shaded region represents a 95\% confidence band.
-The confidence bands around each prediction line are constructed in the usual way, as described in \textcite{visreg}.
+These show the direction, size and significance of an estimated effect --- the slope of the prediction line matches the estimated coefficient, and the shaded region represents a 95\% confidence band, computed using the method in \textcite{visreg}.
 If a horizontal line can be drawn through the confidence band, then the effect is insignificant.

### 2019-04-21 paper.Rmd
Updated ANZJS paper

@@
 title: "Spatial modelling of the two-party preferred vote in Australian federal elections: 2001--2016"
-author: "Jeremy Forbes*, Dianne Cook, Rob J Hyndman Department of Econometrics and Business Statistics, Monash University"
-wpnumber:
-jelcodes: C31, C33, D72
-blind: false
-cover: false
+author: Jeremy Forbes$^*$, Dianne Cook, Rob J Hyndman
 toc: false
 documentclass: anzsauth
+classoption: times, doublespace
+fontfamily: times
 bibliography: references.bib
 biblio-style: anzsj
-classoption: times,doublespace
-header-includes: \usepackage{float, bm}
 output: 
-  pdf_document:
-    latex_engine: xelatex
+  bookdown::pdf_document2:
     fig_caption: yes
@@
     fig_width: 8
+    includes:
+      in_header: anzjs_preamble.tex
 ---

### 2019-04-21 paper.tex
Updated ANZJS paper

--- 2019-04-21 draft-paper.tex
@@
-\documentclass[11pt,a4paper,]{article}
-\usepackage{lmodern}
-
+\documentclass[times, doublespace]{anzsauth}
+\usepackage[]{times}
 \usepackage{amssymb,amsmath}
@@
 \else % if luatex or xelatex
-  \usepackage{unicode-math}
+  \ifxetex
+    \usepackage{mathspec}
+  \else
+    \usepackage{fontspec}
+  \fi
   \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
@@
 \IfFileExists{microtype.sty}{%
-\usepackage[]{microtype}
+\usepackage{microtype}
 \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
 }{}
-\PassOptionsToPackage{hyphens}{url} % url is loaded by hyperref
-\usepackage[unicode=true]{hyperref}
-\hypersetup{
+\usepackage{hyperref}
+\hypersetup{unicode=true,
             pdftitle={Spatial modelling of the two-party preferred vote in Australian federal elections: 2001--2016},
-            pdfkeywords={federal election, Census, Australia, spatial modelling, imputation, socio-demographics, electorates, R, eechidna},
+            pdfauthor={Jeremy Forbes\^{}*, Dianne Cook, Rob J Hyndman},
             pdfborder={0 0 0},
@@
 \urlstyle{same}  % don't use monospace font for urls
-\usepackage{geometry}
-\geometry{left=2.5cm,right=2.5cm,top=2.5cm,bottom=2.5cm}
-\usepackage[style=authoryear-comp,]{biblatex}
-\addbibresource{references.bib}
 \usepackage{longtable,booktabs}
-% Fix footnotes in tables (requires footnote package)
-\IfFileExists{footnote.sty}{\usepackage{footnote}\makesavenoteenv{long table}}{}
+\usepackage{graphicx,grffile}
+\makeatletter
+\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
+\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
+\makeatother
+% Scale images if necessary, so that they will not overflow the page
+% margins by default, and it is still possible to overwrite the defaults
+% using explicit options in \includegraphics[width, height, ...]{}
+\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
 \IfFileExists{parskip.sty}{%
@@
 \setcounter{secnumdepth}{5}
+% Redefines (sub)paragraphs to behave more like sections
+\ifx\paragraph\undefined\else
+\let\oldparagraph\paragraph
+\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
+\fi
+\ifx\subparagraph\undefined\else
+\let\oldsubparagraph\subparagraph
+\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
+\fi
 
-% set default figure placement to htbp
-\makeatletter
-\def\fps@figure{htbp}
-\makeatother
+%%% Use protect on footnotes to avoid problems with footnotes in titles
+\let\rmarkdownfootnote\footnote%
+\def\footnote{\protect\rmarkdownfootnote}
+
+%%% Change title format to be more compact
+\usepackage{titling}
+
+% Create subtitle command for use in maketitle
+\providecommand{\subtitle}[1]{
+  \posttitle{
+    \begin{center}\large#1\end{center}
+    }
+}
+
+\setlength{\droptitle}{-2em}
 
   \title{Spatial modelling of the two-party preferred vote in Australian federal elections: 2001--2016}
+    \pretitle{\vspace{\droptitle}\centering\huge}
+  \posttitle{\par}
+    \author{Jeremy Forbes\(^*\), Dianne Cook, Rob J Hyndman}
+    \preauthor{\centering\large\emph}
+  \postauthor{\par}
+    \date{}
+    \predate{}\postdate{}
 
-%% MONASH STUFF
-
-%% CAPTIONS
-\RequirePackage{caption}
-\DeclareCaptionStyle{italic}[justification=centering]
- {labelfont={bf},textfont={it},labelsep=colon}
-\captionsetup[figure]{style=italic,format=hang,singlelinecheck=true}
-\captionsetup[table]{style=italic,format=hang,singlelinecheck=true}
-
-%% FONT
-\RequirePackage{bera}
-\RequirePackage{mathpazo}
-
-%% HEADERS AND FOOTERS
-\RequirePackage{fancyhdr}
-\pagestyle{fancy}
-\rfoot{\Large\sffamily\raisebox{-0.1cm}{\textbf{\thepage}}}
-\makeatletter
-\lhead{\textsf{\expandafter{\@title}}}
-\makeatother
-\rhead{}
-\cfoot{}
-\setlength{\headheight}{15pt}
-\renewcommand{\headrulewidth}{0.4pt}
-\renewcommand{\footrulewidth}{0.4pt}
-\fancypagestyle{plain}{%
-\fancyhf{} % clear all header and footer fields
-\fancyfoot[C]{\sffamily\thepage} % except the center
-\renewcommand{\headrulewidth}{0pt}
-\renewcommand{\footrulewidth}{0pt}}
-
-%% MATHS
-\RequirePackage{bm,amsmath}
-\allowdisplaybreaks
-
-%% GRAPHICS
-\RequirePackage{graphicx}
-\setcounter{topnumber}{2}
-\setcounter{bottomnumber}{2}
-\setcounter{totalnumber}{4}
-\renewcommand{\topfraction}{0.85}
-\renewcommand{\bottomfraction}{0.85}
-\renewcommand{\textfraction}{0.15}
-\renewcommand{\floatpagefraction}{0.8}
-
-%\RequirePackage[section]{placeins}
-
-%% SECTION TITLES
-\RequirePackage[compact,sf,bf]{titlesec}
-\titleformat{\section}[block]
-  {\fontsize{15}{17}\bfseries\sffamily}
-  {\thesection}
-  {0.4em}{}
-\titleformat{\subsection}[block]
-  {\fontsize{12}{14}\bfseries\sffamily}
-  {\thesubsection}
-  {0.4em}{}
-\titlespacing{\section}{0pt}{*5}{*1}
-\titlespacing{\subsection}{0pt}{*2}{*0.2}
-
-%% TITLE PAGE
-\def\Date{\number\day}
-\def\Month{\ifcase\month\or
- January\or February\or March\or April\or May\or June\or
- July\or August\or September\or October\or November\or December\fi}
-\def\Year{\number\year}
-
-\makeatletter
-\def\wp#1{\gdef\@wp{#1}}\def\@wp{??/??}
-\def\jel#1{\gdef\@jel{#1}}\def\@jel{??}
-\def\showjel{{\large\textsf{\textbf{JEL classification:}}~\@jel}}
-\def\nojel{\def\showjel{}}
-\def\addresses#1{\gdef\@addresses{#1}}\def\@addresses{??}
-\def\cover{{\sffamily\setcounter{page}{0}
-        \thispagestyle{empty}
-        \placefig{2}{1.5}{width=5cm}{monash2}
-        \placefig{16.9}{1.5}{width=2.1cm}{MBusSchool}
-        \begin{textblock}{4}(16.9,4)ISSN 1440-771X\end{textblock}
-        \begin{textblock}{7}(12.7,27.9)\hfill
-        \includegraphics[height=0.7cm]{AACSB}~~~
-        \includegraphics[height=0.7cm]{EQUIS}~~~
-        \includegraphics[height=0.7cm]{AMBA}
-        \end{textblock}
-        \vspace*{2cm}
-        \begin{center}\Large
-        Department of Econometrics and Business Statistics\\[.5cm]
-        \footnotesize http://monash.edu/business/ebs/research/publications
-        \end{center}\vspace{2cm}
-        \begin{center}
-        \fbox{\parbox{14cm}{\begin{onehalfspace}\centering\Huge\vspace*{0.3cm}
-                \textsf{\textbf{\expandafter{\@title}}}\vspace{1cm}\par
-                \LARGE\@author\end{onehalfspace}
-        }}
-        \end{center}
-        \vfill
-                \begin{center}\Large
-                \Month~\Year\\[1cm]
-                Working Paper \@wp
-        \end{center}\vspace*{2cm}}}
-\def\pageone{{\sffamily\setstretch{1}%
-        \thispagestyle{empty}%
-        \vbox to \textheight{%
-        \raggedright\baselineskip=1.2cm
-     {\fontsize{24.88}{30}\sffamily\textbf{\expandafter{\@title}}}
-        \vspace{2cm}\par
-        \hspace{1cm}\parbox{14cm}{\sffamily\large\@addresses}\vspace{1cm}\vfill
-        \hspace{1cm}{\large\Date~\Month~\Year}\\[1cm]
-        \hspace{1cm}\showjel\vss}}}
-\def\blindtitle{{\sffamily
-     \thispagestyle{plain}\raggedright\baselineskip=1.2cm
-     {\fontsize{24.88}{30}\sffamily\textbf{\expandafter{\@title}}}\vspace{1cm}\par
-        }}
-\def\titlepage{{\cover\newpage\pageone\newpage\blindtitle}}
-
-\def\blind{\def\titlepage{{\blindtitle}}\let\maketitle\blindtitle}
-\def\titlepageonly{\def\titlepage{{\pageone\end{document}}}}
-\def\nocover{\def\titlepage{{\pageone\newpage\blindtitle}}\let\maketitle\titlepage}
-\let\maketitle\titlepage
-\makeatother
-
-%% SPACING
-\RequirePackage{setspace}
-\spacing{1.5}
-
-%% LINE AND PAGE BREAKING
-\sloppy
-\clubpenalty = 10000
-\widowpenalty = 10000
-\brokenpenalty = 10000
-\RequirePackage{microtype}
-
-%% PARAGRAPH BREAKS
-\setlength{\parskip}{1.4ex}
-\setlength{\parindent}{0em}
-
-%% HYPERLINKS
-\RequirePackage{xcolor} % Needed for links
-\definecolor{darkblue}{rgb}{0,0,.6}
-\RequirePackage{url}
-
-\makeatletter
-\@ifpackageloaded{hyperref}{}{\RequirePackage{hyperref}}
-\makeatother
-\hypersetup{
-     citecolor=0 0 0,
-     breaklinks=true,
-     bookmarksopen=true,
-     bookmarksnumbered=true,
-     linkcolor=darkblue,
-     urlcolor=blue,
-     citecolor=darkblue,
-     colorlinks=true}
-
-%% KEYWORDS
-\newenvironment{keywords}{\par\vspace{0.5cm}\noindent{\sffamily\textbf{Keywords:}}}{\vspace{0.25cm}\par\hrule\vspace{0.5cm}\par}
-
-%% ABSTRACT
-\renewenvironment{abstract}{\begin{minipage}{\textwidth}\parskip=1.4ex\noindent
-\hrule\vspace{0.1cm}\par{\sffamily\textbf{\abstractname}}\newline}
-  {\end{minipage}}
-
+%% Any special functions or other packages can be loaded here.
 \usepackage[T1]{fontenc}
@@
 
-\usepackage[showonlyrefs]{mathtools}
-\usepackage[no-weekday]{eukdate}
+\usepackage{float, bm}
+\usepackage{moreverb}
+\usepackage{url}
+\usepackage{grffile}
+\usepackage{lineno}
+\usepackage[UKenglish]{isodate}
 
-%% BIBLIOGRAPHY
-
-\makeatletter
-\@ifpackageloaded{biblatex}{}{\usepackage[style=authoryear-comp, backend=biber, natbib=true]{biblatex}}
-\makeatother
-\ExecuteBibliographyOptions{bibencoding=utf8,minnames=1,maxnames=3, maxbibnames=99,dashed=false,terseinits=true,giveninits=true,uniquename=false,uniquelist=false,doi=false, isbn=false,url=true,sortcites=false}
-
-\DeclareFieldFormat{url}{\texttt{\url{#1}}}
-\DeclareFieldFormat[article]{pages}{#1}
-\DeclareFieldFormat[inproceedings]{pages}{\lowercase{pp.}#1}
-\DeclareFieldFormat[incollection]{pages}{\lowercase{pp.}#1}
-\DeclareFieldFormat[article]{volume}{\mkbibbold{#1}}
-\DeclareFieldFormat[article]{number}{\mkbibparens{#1}}
-\DeclareFieldFormat[article]{title}{\MakeCapital{#1}}
-\DeclareFieldFormat[inproceedings]{title}{#1}
-\DeclareFieldFormat{shorthandwidth}{#1}
-% No dot before number of articles
-\usepackage{xpatch}
-\xpatchbibmacro{volume+number+eid}{\setunit*{\adddot}}{}{}{}
-% Remove In: for an article.
-\renewbibmacro{in:}{%
-  \ifentrytype{article}{}{%
-  \printtext{\bibstring{in}\intitlepunct}}}
-
-\makeatletter
-\DeclareDelimFormat[cbx@textcite]{nameyeardelim}{\addspace}
-\makeatother
-\renewcommand*{\finalnamedelim}{%
-  %\ifnumgreater{\value{liststop}}{2}{\finalandcomma}{}% there really should be no funny Oxford comma business here
-  \addspace\&\space}
-
-\jel{C31, C33, D72}
-
-\RequirePackage[absolute,overlay]{textpos}
-\setlength{\TPHorizModule}{1cm}
-\setlength{\TPVertModule}{1cm}
-\def\placefig#1#2#3#4{\begin{textblock}{.1}(#1,#2)\rlap{\includegraphics[#3]{#4}}\end{textblock}}
-
-\nocover
-
-\author{Jeremy~Forbes, Dianne~Cook, Rob J~Hyndman}
-\addresses{\textbf{Jeremy Forbes}\newline
-Monash University\newline Melbourne, Australia
-\newline{Email: \href{mailto:jeremyforbes1995@gmail.com}{\nolinkurl{jeremyforbes1995@gmail.com}}}\newline Corresponding author\\[1cm]
-\textbf{Dianne Cook}\newline
-Monash University\newline Melbourne, Australia
-\\[1cm]
-\textbf{Rob J Hyndman}\newline
-Monash University\newline Melbourne, Australia
-\\[1cm]
-}
-
-\date{\sf\Date~\Month~\Year}
-\makeatletter
- \lfoot{\sf Forbes, Cook, Hyndman: \@date}
-\makeatother
-
-%% Any special functions or other packages can be loaded here.
-\usepackage{float}
 \let\origfigure\figure
@@
 
-% Don't number subsubsections
+\runningheads{Spatial modelling of Australian federal elections: 2001--2016}{FORBES, COOK and HYNDMAN}
 
-\titleformat{\subsubsection}
-  {\normalfont\bfseries\sffamily}   % The style of the section title
-  {}                             % a prefix
-  {0pt}                          % How much space exists between the prefix and the title
-  {}    % How the section is represented
+\def\volumeyear{2019}
+\let\oldmaketitle\maketitle
+\def\maketitle{\cleanlookdateon\oldmaketitle}
+\linenumbers
+
+\pretitle{\begin{center}\Large\bfseries}
+\posttitle{\end{center}\vspace{0.5cm}}
+\preauthor{\begin{center}\large}
+\postauthor{\end{center}\begin{center}\emph{Department of Econometrics \& Business Statistics, Monash University, Australia}\footnotetext{$^*$Corresponding author.
+Email: \texttt{jeremyforbes1995@gmail.com}}\end{center}}
 
@@
 \maketitle
-\begin{abstract}
+
+\hypertarget{summary}{%
+\section*{Summary}\label{summary}}
+\addcontentsline{toc}{section}{Summary}
+
 We examine the relationships between electoral socio-demographic characteristics and two-party preference in the six Australian federal elections held between 2001 to 2016.
@@
 All data featured in this study has been contributed to the \texttt{eechidna} \texttt{R} package (available on CRAN).
-\end{abstract}
-\begin{keywords}
-federal election, Census, Australia, spatial modelling, imputation, socio-demographics, electorates, R, eechidna
-\end{keywords}
+
+\textbf{Keywords:} federal election, Census, Australia, spatial modelling, imputation, socio-demographics, electorates, R, eechidna
 
@@
 The first problem addressed is that of constructing appropriate Census data for the 2004, 2007, 2010 and 2013 elections --- election years in which a Census does not occur.
-The predominant approach in previous studies was to join voting outcomes to the nearest Census, without accounting for any temporal differences \autocites[see][]{DavisStimson98}{Stimson06}{Liao09}{Stimson09}.
+The predominant approach in previous studies was to join voting outcomes to the nearest Census, without accounting for any temporal differences (see Davis and Stimson 1998; Stimson, McCrea, and Shyy 2006; Liao, Shyy, and Stimson 2009; Stimson and Shyy 2009).
 Furthermore, electoral boundaries change regularly, so spatial discrepancies also arise when matching with electoral data.
-To obtain appropriate ``Census-like'' data for these four elections, electoral socio-demographics are constructed using a spatio-temporal imputation that combines areal interpolation \autocite{Goodchild1993} and linear time-interpolation.
+To obtain appropriate ``Census-like'' data for these four elections, electoral socio-demographics are constructed using a spatio-temporal imputation that combines areal interpolation (Goodchild, Anselin, and Deichmann 1993) and linear time-interpolation.
 Collecting and wrangling the raw data, along with the imputation process, are detailed in Section \ref{data}.
-All data and associated documentation relating to this procedure are available in the \texttt{eechidna} \texttt{R} package \autocite{eechidna}, providing a resource for future analysis.
+All data and associated documentation relating to this procedure are available in the \texttt{eechidna} \texttt{R} package (Forbes et al.
+2019), providing a resource for future analysis.
 
 Previous work on modelling Australian federal elections has found that aggregate socio-demographics are relatively good predictors of voting outcomes.
-\textcite{Forrest01} used multiple regression to model the Liberal and Labor primary vote for polling booths in the Farrer electorate in 1998 as a function of Census variables from 1996.
-\textcite{Stimson06}, \textcite{Stimson09} and \textcite{Stimson12} used principal component analysis of polling booths in the 2001, 2004 and 2007 elections respectively, also finding that socio-demographic characteristics of polling booths are linked to their two-party preferred vote.
-In contrast, \textcite{Stimson09} models the polling booth swing vote (change in the two-party preferred vote) in the 2007 election, finding that little of the swing vote can be explained by Census data.
+Forrest et al.
+(2001) used multiple regression to model the Liberal and Labor primary vote for polling booths in the Farrer electorate in 1998 as a function of Census variables from 1996.
+Stimson, McCrea, and Shyy (2006), Stimson and Shyy (2009) and Stimson and Shyy (2012) used principal component analysis of polling booths in the 2001, 2004 and 2007 elections respectively, also finding that socio-demographic characteristics of polling booths are linked to their two-party preferred vote.
+In contrast, Stimson and Shyy (2009) models the polling booth swing vote (change in the two-party preferred vote) in the 2007 election, finding that little of the swing vote can be explained by Census data.
 Instead of analyzing a single election in isolation, this paper employs a consistent model framework across six elections so that temporal changes in the effects of socio-demographics can be observed, where each federal elections is modelled with a cross-sectional dataset.
@@
 The formats of these files differ over the years, making extracting the appropriate information a big task.
-The functions available in the \texttt{dplyr} \autocite{dplyr} and \texttt{readxl} \autocite{readxl} \texttt{R} packages are particularly useful, as they provide fast consistent tools for data manipulation and functions to import \texttt{.xlsx} files (respectively).
+The functions available in the \texttt{dplyr} (Wickham, François, et al.
+2019) and \texttt{readxl} (Wickham, Bryan, et al.
+2019) \texttt{R} packages are particularly useful, as they provide fast consistent tools for data manipulation and functions to import \texttt{.xlsx} files (respectively).
 The 2001 and 2006 Census data are published in a format where the information for each electorate is held in a separate document making it difficult to use the \texttt{dplyr} tools.
 Instead, cells have to be selected from each individual file to construct the desired variables.
-All scripts required for the data wrangling process can be found in the github repository for the \texttt{eechidna} \texttt{R} package \autocite{eechidna}, along with the raw data.
+All scripts required for the data wrangling process can be found in the github repository for the \texttt{eechidna} \texttt{R} package (Forbes et al.
+2019), along with the raw data.
 The \texttt{eechidna} package makes this study entirely reproducible and provides a resource to help wrangle data for future Censuses and elections, when they become available.
@@
 
-{\centering \includegraphics[width=0.9\linewidth]{draft-paper_files/figure-latex/timeline-1}
+{\centering \includegraphics[width=0.9\linewidth]{paper_files/figure-latex/timeline-1}
 
@@
 
-To account for spatial differences, the piecewise approximation method in \textcite{Goodchild1993} is adopted.
+To account for spatial differences, the piecewise approximation method in Goodchild, Anselin, and Deichmann (1993) is adopted.
 Consider a map of source zones \(s = 1,\dots,S\), for which socio-demographic information is available, and a set of target zones \(t = 1,\dots,T\) for which information is to be imputed.
@@
 
-{\centering \includegraphics[height=0.28\textheight]{draft-paper_files/figure-latex/hume13-1}
+{\centering \includegraphics[height=0.28\textheight]{paper_files/figure-latex/hume13-1}
 
@@
 
-{\centering \includegraphics[height=0.28\textheight]{draft-paper_files/figure-latex/hume16-1}
+{\centering \includegraphics[height=0.28\textheight]{paper_files/figure-latex/hume16-1}
 
@@
 
-The spatial error model \autocite{Anselin88} is chosen because it captures spatial heterogeneity by incorporating a spatially structured random effect vector \autocite{LeSage2009}.
+The spatial error model (Anselin 1988) is chosen because it captures spatial heterogeneity by incorporating a spatially structured random effect vector (LeSage, Kelley Pace, and Pace 2009).
 In this context, the random effect can be thought of as capturing the unobserved political climate in each electorate, where this climate is correlated with the climate in neighbouring electorates, under the assumption that the climate is independent of electoral socio-demographics.
@@
 
-{\centering \includegraphics[height=0.2\textheight,]{draft-paper_files/figure-latex/rhovis-1}
+{\centering \includegraphics[height=0.2\textheight,]{paper_files/figure-latex/rhovis-1}
 
@@
 
-{\centering \includegraphics[height=0.2\textheight,]{draft-paper_files/figure-latex/plotintercept-1}
+{\centering \includegraphics[height=0.2\textheight,]{paper_files/figure-latex/plotintercept-1}
 
@@
 
-{\centering \includegraphics{draft-paper_files/figure-latex/partresplot1-1}
+{\centering \includegraphics{paper_files/figure-latex/partresplot1-1}
 
@@
 
-{\centering \includegraphics{draft-paper_files/figure-latex/partresplot2-1}
+{\centering \includegraphics{paper_files/figure-latex/partresplot2-1}
 
@@
 The partial residuals are the residuals from the fitted model with the estimated effect an individual variable added.
-These show the direction, size and significance of an estimated effect --- the slope of the prediction line matches the estimated coefficient, and the shaded region represents a 95\% confidence band, computed using the method in \textcite{visreg}.
+These show the direction, size and significance of an estimated effect --- the slope of the prediction line matches the estimated coefficient, and the shaded region represents a 95\% confidence band, computed using the method Breheny and Burchett (2017).
 If a horizontal line can be drawn through the confidence band, then the effect is insignificant.
@@
 
-{\centering \includegraphics[width=0.7\linewidth,]{draft-paper_files/figure-latex/resstate-1}
+{\centering \includegraphics[width=0.7\linewidth,]{paper_files/figure-latex/resstate-1}
 
@@
 
-This paper was produced using \texttt{RMarkdown} \autocite{rmarkdown} and \texttt{knitr} \autocite{knitr}.
-All corresponding code for this paper can be found in the github repository \href{https://github.com/jforbes14/eechidna-paper}{github.com/jforbes14/eechidna-paper}, and the data used is available in the \texttt{eechidna} package \autocite{eechidna}.
+This paper was produced using \texttt{RMarkdown} (Allaire et al.
+2019) and \texttt{knitr} (Xie 2015).
+All corresponding code for this paper can be found in the github repository \href{https://github.com/jforbes14/eechidna-paper}{github.com/jforbes14/eechidna-paper}, and the data used is available in the \texttt{eechidna} package (Forbes et al.
+2019).
 All raw data was obtained from the Australian Electoral Commission, the Australian Bureau of Statistics and the Australian Government.
@@
 
-\printbibliography
+\hypertarget{refs}{}
+\leavevmode\hypertarget{ref-rmarkdown}{}%
+Allaire, JJ, Yihui Xie, Jonathan McPherson, Javier Luraschi, Kevin Ushey, Aron Atkins, Hadley Wickham, Joe Cheng, Winston Chang, and Richard Iannone.
+2019.
+\emph{Rmarkdown: Dynamic Documents for R}.
+\url{https://rmarkdown.rstudio.com}.
+
+\leavevmode\hypertarget{ref-Anselin88}{}%
+Anselin, Luc. 1988. \emph{Spatial Econometrics: Methods and Models}. Vol. 4. Springer Science \& Business Media.
+
+\leavevmode\hypertarget{ref-visreg}{}%
+Breheny, Patrick, and Woodrow Burchett.
+2017.
+``Visualization of Regression Models Using Visreg.'' \emph{The R Journal} 9 (2): 56--71.
+\url{https://journal.r-project.org/archive/2017/RJ-2017-046/index.html}.
+
+\leavevmode\hypertarget{ref-DavisStimson98}{}%
+Davis, R., and R.
+Stimson.
+1998.
+``Disillusionment and Disenchantment at the Fringe: Explaining the Geography of the One Nation Party Vote at the Queensland Election.'' \emph{People and Place} 6 (3): 69--82.
+
+\leavevmode\hypertarget{ref-eechidna}{}%
+Forbes, Jeremy, Di Cook, Anthony Ebert, Heike Hofmann, Rob J Hyndman, Thomas Lumley, Ben Marwick, et al.
+2019.
+\emph{Eechidna: Exploring Election and Census Highly Informative Data Nationally for Australia}.
+\url{https://CRAN.R-project.org/package=eechidna}.
+
+\leavevmode\hypertarget{ref-Forrest01}{}%
+Forrest, James, Margaret Alston, Chris Medlin, and Siti Amri.
+2001.
+``Voter Behaviour in Rural Areas: A Study of the Farrer Electoral Division in Southern New South Wales at the 1998 Federal Election.'' \emph{Australian Geographical Studies} 39 (2).
+Wiley/Blackwell: 167--82.
+
+\leavevmode\hypertarget{ref-Goodchild1993}{}%
+Goodchild, M F, L Anselin, and U Deichmann.
+1993.
+``A Framework for the Areal Interpolation of Socioeconomic Data.'' \emph{Environment and Planning A} 25 (3).
+London, England: SAGE Publications: 383--97.
+
+\leavevmode\hypertarget{ref-LeSage2009}{}%
+LeSage, James, R Kelley Pace, and Robert Kelley Pace. 2009. \emph{Introduction to Spatial Econometrics}. Chapman; Hall/CRC.
+
+\leavevmode\hypertarget{ref-Liao09}{}%
+Liao, E., T.
+Shyy, and R.
+Stimson.
+2009.
+``Developing a Web‐based E‐research Facility for Socio‐spatial Analysis to Investigate Relationships Between Voting Patterns and Local Population Characteristics.'' \emph{Journal of Spatial Science} 54 (2): 63--88.
+
+\leavevmode\hypertarget{ref-Stimson06}{}%
+Stimson, R., R.
+McCrea, and T.
+Shyy.
+2006.
+``Spatially Disaggregated Modelling of Voting Outcomes and Socio-Economic Characteristics at the 2001 Australian Federal Election.'' \emph{Geographical Research} 44 (3): 242--54.
+
+\leavevmode\hypertarget{ref-Stimson09}{}%
+Stimson, Robert, and Tung-Kai Shyy.
+2009.
+``A Socio-Spatial Analysis of Voting for Political Parties at the 2007 Federal Election.'' \emph{People and Place} 17 (1).
+Monash University, Centre for Population; Urban Research: 39--54.
+
+\leavevmode\hypertarget{ref-Stimson12}{}%
+Stimson, R., and T. Shyy. 2012. ``And Now for Something Different: Modelling Socio-Political Landscapes.'' \emph{Annals of Regional Science} 50: 623--43.
+
+\leavevmode\hypertarget{ref-readxl}{}%
+Wickham, Hadley, Jennifer Bryan, Marcin Kalicinski, Komarov Valery, Christophe Leitenne, Bob Colbert, David Hoerl, and Evan Miller.
+2019.
+\emph{Readxl: Read Excel Files}.
+\url{https://CRAN.R-project.org/package=readxl}.
+
+\leavevmode\hypertarget{ref-dplyr}{}%
+Wickham, Hadley, Romain François, Lionel Henry, and Kirill Müller. 2019. \emph{Dplyr: A Grammar of Data Manipulation}. \url{https://CRAN.R-project.org/package=dplyr}.
+
+\leavevmode\hypertarget{ref-knitr}{}%
+Xie, Yihui. 2015. \emph{Dynamic Documents with R and Knitr}. 2nd ed. Boca Raton, Florida: Chapman; Hall/CRC. \url{http://yihui.name/knitr/}.
 

### 2019-04-21 draft-paper.Rmd
fixed the margins for the paper and revised introduction

@@
 All data featured in this study has been contributed to the `eechidna` `R` package (available on CRAN)."
-keywords: "federal election, Census, Australia, spatial modelling, imputation, socio-demographics, electorates, R, eechidna"
+keywords: "federal election, Census, Australia, spatial modelling, imputation, data science, socio-demographics, electorates, R, eechidna"
 wpnumber:
@@
 This study explores how electoral socio-demographic characteristics relate to two-party preference, and whether their effects have changed over time.
+
 Electoral socio-demographics are derived from the Australian Census, and vote counts are obtained from Australian federal elections.
-
-Joining these two data sources is problematic as there in an inherent asynchronicity in the two types of events.
+Joining these two data sources is problematic as there is an inherent asynchronicity in the two types of events.
 A Census is conducted by the Australian Bureau of Statistics (ABS) every five years, whereas federal elections, conducted by the Australian Electoral Commission (AEC), usually occur every three years or so.
@@
 In contrast, @Stimson09 models the polling booth swing vote (change in the two-party preferred vote) in the 2007 election, finding that little of the swing vote can be explained by Census data.
-Instead of analyzing a single election in isolation, this paper employs a consistent model framework across six elections so that temporal changes in the effects of socio-demographics can be observed, where each federal elections is modelled with a cross-sectional dataset.
-The use of a regression framework to examine these socio-political relationships over time is seemingly absent from previous Australian studies.
-It also appears that no study has attempted any type of statistical analysis of socio-demographics in conjunction with voter behaviour in Australia since 2007, making this paper distinctly different from previous work.
-
+
+Instead of analyzing a single election in isolation, this paper employs a consistent model framework across six elections so that temporal changes in the effects of socio-demographics can be observed.
+Each federal election is modelled with a cross-sectional dataset.
+<!--The use of a regression framework to examine these socio-political relationships over time is seemingly absent from previous Australian studies.
+It also appears that no study has attempted any type of statistical analysis of socio-demographics in conjunction with voter behaviour in Australia since 2007, making this paper distinctly different from previous work.-->
 The cross-sectional dataset for each election used here consists of the two-party preferred vote (as the response variable), and a set of common socio-demographic variables (as the explanatory variables) that characterize each electorate.
@@
 To account for the inherent spatial structure of the data, a spatial error model is then estimated for each election.
-These steps are discussed in Section \@ref(modelling).
-In Section \@ref(results) inference is conducted to see which effects are significant and how these effects change over time.
-
-Section \@ref(conclusion) concludes, followed by two supplementary sections.
+
+The paper is organised as follows.
+Section \@ref(data) describes the data collection, joining and cleaning.
+These pre-processing steps and model details are discussed in Section \@ref(modelling).
+Section \@ref(results) describes the inference conducted to determine signficance of effects and how these change over time.
+Section \@ref(conclusion) summarises the work.
+Two supplementary sections document the contributions of others to this work and the software.
 

### 2019-04-21 paper.Rmd
fixed the margins for the paper and revised introduction

@@
 biblio-style: anzsj
+geometry: "left=5cm,top=5cm"
 output: 
@@
 
-**Keywords:** federal election, Census, Australia, spatial modelling, imputation, socio-demographics, electorates, R, eechidna
+**Keywords:** federal election, Census, Australia, spatial modelling, imputation, data science, socio-demographics, electorates, R, eechidna
 
@@
     caption = "Population from each intersecting Census source zone (2016) that is allocated to the target zone (purple region - Hume electoral boundary in 2013), and the corresponding proportion of Australian citizens in each of these source zones.") %>%
-  kableExtra::kable_styling(font_size=10, latex_options='scale_down')
+  kableExtra::kable_styling(font_size=12, latex_options='scale_down')
 ```

### 2019-04-21 paper.tex
fixed the margins for the paper and revised introduction

@@
 }{}
+\usepackage[left=5cm,top=5cm]{geometry}
 \usepackage{hyperref}
@@
 Email: \texttt{jeremyforbes1995@gmail.com}}\end{center}}
+\usepackage{booktabs}
+\usepackage{longtable}
+\usepackage{array}
+\usepackage{multirow}
+\usepackage{wrapfig}
+\usepackage{float}
+\usepackage{colortbl}
+\usepackage{pdflscape}
+\usepackage{tabu}
+\usepackage{threeparttable}
+\usepackage{threeparttablex}
+\usepackage[normalem]{ulem}
+\usepackage{makecell}
+\usepackage{xcolor}
 
@@
 
-\textbf{Keywords:} federal election, Census, Australia, spatial modelling, imputation, socio-demographics, electorates, R, eechidna
+\textbf{Keywords:} federal election, Census, Australia, spatial modelling, imputation, data science, socio-demographics, electorates, R, eechidna
 
@@
 In contrast, Stimson and Shyy (2009) models the polling booth swing vote (change in the two-party preferred vote) in the 2007 election, finding that little of the swing vote can be explained by Census data.
-Instead of analyzing a single election in isolation, this paper employs a consistent model framework across six elections so that temporal changes in the effects of socio-demographics can be observed, where each federal elections is modelled with a cross-sectional dataset.
-The use of a regression framework to examine these socio-political relationships over time is seemingly absent from previous Australian studies.
-It also appears that no study has attempted any type of statistical analysis of socio-demographics in conjunction with voter behaviour in Australia since 2007, making this paper distinctly different from previous work.
 
+Instead of analyzing a single election in isolation, this paper employs a consistent model framework across six elections so that temporal changes in the effects of socio-demographics can be observed.
+Each federal election is modelled with a cross-sectional dataset.
 The cross-sectional dataset for each election used here consists of the two-party preferred vote (as the response variable), and a set of common socio-demographic variables (as the explanatory variables) that characterize each electorate.
@@
 To account for the inherent spatial structure of the data, a spatial error model is then estimated for each election.
-These steps are discussed in Section \ref{modelling}.
-In Section \ref{results} inference is conducted to see which effects are significant and how these effects change over time.
 
-Section \ref{conclusion} concludes, followed by two supplementary sections.
+The paper is organised as follows.
+Section \ref{data} describes the data collection, joining and cleaning.
+These pre-processing steps and model details are discussed in Section \ref{modelling}.
+Section \ref{results} describes the inference conducted to determine signficance of effects and how these change over time.
+Section \ref{conclusion} summarises the work.
+Two supplementary sections document the contributions of others to this work and the software.
 
@@
 \resizebox{\linewidth}{!}{
-\fontsize{10}{12}\selectfont
+\fontsize{12}{14}\selectfont
 \begin{tabular}{lrrrr}
@@
 2001.
-``Voter Behaviour in Rural Areas: A Study of the Farrer Electoral Division in Southern New South Wales at the 1998 Federal Election.'' \emph{Australian Geographical Studies} 39 (2).
-Wiley/Blackwell: 167--82.
+``Voter Behaviour in Rural Areas: A Study of the Farrer Electoral Division in Southern New South Wales at the 1998 Federal Election.'' \emph{Australian Geographical Studies} 39 (2): 167--82.
 
 \leavevmode\hypertarget{ref-Goodchild1993}{}%
-Goodchild, M F, L Anselin, and U Deichmann.
-1993.
-``A Framework for the Areal Interpolation of Socioeconomic Data.'' \emph{Environment and Planning A} 25 (3).
-London, England: SAGE Publications: 383--97.
+Goodchild, M F, L Anselin, and U Deichmann. 1993. ``A Framework for the Areal Interpolation of Socioeconomic Data.'' \emph{Environment and Planning A} 25 (3): 383--97.
 
@@
 \leavevmode\hypertarget{ref-Stimson09}{}%
-Stimson, Robert, and Tung-Kai Shyy.
-2009.
-``A Socio-Spatial Analysis of Voting for Political Parties at the 2007 Federal Election.'' \emph{People and Place} 17 (1).
-Monash University, Centre for Population; Urban Research: 39--54.
+Stimson, Robert, and Tung-Kai Shyy. 2009. ``A Socio-Spatial Analysis of Voting for Political Parties at the 2007 Federal Election.'' \emph{People and Place} 17 (1): 39--54.
 

### 2019-04-21 paper.Rmd
No need to modify geometry

@@
 biblio-style: anzsj
-geometry: "left=5cm,top=5cm"
 output: 

### 2019-04-21 paper.Rmd
Use natbib

@@
     fig_caption: yes
+    citation_package: natbib
     keep_tex: yes

### 2019-04-21 paper.Rmd
Squeeze the table onto one page

@@
   header = FALSE,
-  column.sep.width = "1pt",
+  column.sep.width = "2pt",
   no.space=TRUE,
@@
   dep.var.caption="",
-  dep.var.labels="\\vspace{-0.5cm}",
-  title = "Estimated spatial error model parameters (standard errors) for each of the six election years.",
+  dep.var.labels="\\vspace{-0.4cm}",
+  title = "Estimated spatial error model parameters (standard errors) for each of the six election years.\\vspace*{-0.4cm}",
   #dep.var.labels = "Two-party preferred vote in favor of the Liberal party",
@@
 regtable <- c(regtable[1:77],sigmaline, regtable[78:length(regtable)])
+# Squeeze some more space
+regtable[2] <- "\\begin{table}[!htbp]\\vspace*{-0.4cm}\\centering "     
 # Allow alignment on decimal point
@@
 regtable[12:75] <- gsub("$^{***}$","\\rlap{$^{***}$}", regtable[12:75], fixed=TRUE)
+# Column spacing
 regtable[5] <- "\\scriptsize\\tabcolsep=0.35cm"
+# Remove "Note"
+regtable[81] <- "& \\multicolumn{6}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\\\"
 ```

### 2019-04-21 paper.tex
Remove numbering of subsubsections

@@
 }{}
-\usepackage[left=5cm,top=5cm]{geometry}
 \usepackage{hyperref}
@@
 \urlstyle{same}  % don't use monospace font for urls
+\usepackage{natbib}
+\bibliographystyle{anzsj}
 \usepackage{longtable,booktabs}
@@
 Email: \texttt{jeremyforbes1995@gmail.com}}\end{center}}
-\usepackage{booktabs}
-\usepackage{longtable}
-\usepackage{array}
-\usepackage{multirow}
-\usepackage{wrapfig}
-\usepackage{float}
-\usepackage{colortbl}
-\usepackage{pdflscape}
-\usepackage{tabu}
-\usepackage{threeparttable}
-\usepackage{threeparttablex}
-\usepackage[normalem]{ulem}
-\usepackage{makecell}
-\usepackage{xcolor}
 
@@
 The first problem addressed is that of constructing appropriate Census data for the 2004, 2007, 2010 and 2013 elections --- election years in which a Census does not occur.
-The predominant approach in previous studies was to join voting outcomes to the nearest Census, without accounting for any temporal differences (see Davis and Stimson 1998; Stimson, McCrea, and Shyy 2006; Liao, Shyy, and Stimson 2009; Stimson and Shyy 2009).
+The predominant approach in previous studies was to join voting outcomes to the nearest Census, without accounting for any temporal differences \citep[see][]{DavisStimson98, Stimson06, Liao09, Stimson09}.
 Furthermore, electoral boundaries change regularly, so spatial discrepancies also arise when matching with electoral data.
-To obtain appropriate ``Census-like'' data for these four elections, electoral socio-demographics are constructed using a spatio-temporal imputation that combines areal interpolation (Goodchild, Anselin, and Deichmann 1993) and linear time-interpolation.
+To obtain appropriate ``Census-like'' data for these four elections, electoral socio-demographics are constructed using a spatio-temporal imputation that combines areal interpolation \citep{Goodchild1993} and linear time-interpolation.
 Collecting and wrangling the raw data, along with the imputation process, are detailed in Section \ref{data}.
-All data and associated documentation relating to this procedure are available in the \texttt{eechidna} \texttt{R} package (Forbes et al.
-2019), providing a resource for future analysis.
+All data and associated documentation relating to this procedure are available in the \texttt{eechidna} \texttt{R} package \citep{eechidna}, providing a resource for future analysis.
 
 Previous work on modelling Australian federal elections has found that aggregate socio-demographics are relatively good predictors of voting outcomes.
-Forrest et al.
-(2001) used multiple regression to model the Liberal and Labor primary vote for polling booths in the Farrer electorate in 1998 as a function of Census variables from 1996.
-Stimson, McCrea, and Shyy (2006), Stimson and Shyy (2009) and Stimson and Shyy (2012) used principal component analysis of polling booths in the 2001, 2004 and 2007 elections respectively, also finding that socio-demographic characteristics of polling booths are linked to their two-party preferred vote.
-In contrast, Stimson and Shyy (2009) models the polling booth swing vote (change in the two-party preferred vote) in the 2007 election, finding that little of the swing vote can be explained by Census data.
+\citet{Forrest01} used multiple regression to model the Liberal and Labor primary vote for polling booths in the Farrer electorate in 1998 as a function of Census variables from 1996.
+\citet{Stimson06}, \citet{Stimson09} and \citet{Stimson12} used principal component analysis of polling booths in the 2001, 2004 and 2007 elections respectively, also finding that socio-demographic characteristics of polling booths are linked to their two-party preferred vote.
+In contrast, \citet{Stimson09} models the polling booth swing vote (change in the two-party preferred vote) in the 2007 election, finding that little of the swing vote can be explained by Census data.
 
@@
 The formats of these files differ over the years, making extracting the appropriate information a big task.
-The functions available in the \texttt{dplyr} (Wickham, François, et al.
-2019) and \texttt{readxl} (Wickham, Bryan, et al.
-2019) \texttt{R} packages are particularly useful, as they provide fast consistent tools for data manipulation and functions to import \texttt{.xlsx} files (respectively).
+The functions available in the \texttt{dplyr} \citep{dplyr} and \texttt{readxl} \citep{readxl} \texttt{R} packages are particularly useful, as they provide fast consistent tools for data manipulation and functions to import \texttt{.xlsx} files (respectively).
 The 2001 and 2006 Census data are published in a format where the information for each electorate is held in a separate document making it difficult to use the \texttt{dplyr} tools.
 Instead, cells have to be selected from each individual file to construct the desired variables.
-All scripts required for the data wrangling process can be found in the github repository for the \texttt{eechidna} \texttt{R} package (Forbes et al.
-2019), along with the raw data.
+All scripts required for the data wrangling process can be found in the github repository for the \texttt{eechidna} \texttt{R} package \citep{eechidna}, along with the raw data.
 The \texttt{eechidna} package makes this study entirely reproducible and provides a resource to help wrangle data for future Censuses and elections, when they become available.
@@
 \hypertarget{differences-between-census-and-election-data}{%
-\subsubsection{Differences between Census and election data}\label{differences-between-census-and-election-data}}
+\subsubsection*{Differences between Census and election data}\label{differences-between-census-and-election-data}}
+\addcontentsline{toc}{subsubsection}{Differences between Census and election data}
 
@@
 \hypertarget{spatio-temporal-imputation}{%
-\subsubsection{Spatio-temporal imputation}\label{spatio-temporal-imputation}}
+\subsubsection*{Spatio-temporal imputation}\label{spatio-temporal-imputation}}
+\addcontentsline{toc}{subsubsection}{Spatio-temporal imputation}
 
@@
 
-To account for spatial differences, the piecewise approximation method in Goodchild, Anselin, and Deichmann (1993) is adopted.
+To account for spatial differences, the piecewise approximation method in \citet{Goodchild1993} is adopted.
 Consider a map of source zones \(s = 1,\dots,S\), for which socio-demographic information is available, and a set of target zones \(t = 1,\dots,T\) for which information is to be imputed.
@@
 
-The spatial error model (Anselin 1988) is chosen because it captures spatial heterogeneity by incorporating a spatially structured random effect vector (LeSage, Kelley Pace, and Pace 2009).
+The spatial error model \citep{Anselin88} is chosen because it captures spatial heterogeneity by incorporating a spatially structured random effect vector \citep{LeSage2009}.
 In this context, the random effect can be thought of as capturing the unobserved political climate in each electorate, where this climate is correlated with the climate in neighbouring electorates, under the assumption that the climate is independent of electoral socio-demographics.
 
-\begin{table}[!htbp] \centering    \caption{Estimated spatial error model parameters (standard errors) for each of the six election years.}    \label{tab:printmodels}  \scriptsize\tabcolsep=0.35cm \begin{tabular}{@{\extracolsep{1pt}}lrrrrrr} \\[-1.8ex]\hline  \hline \\[-1.8ex]  \\[-1.8ex] & \multicolumn{6}{c}{\vspace{-0.5cm}} \\   & 2001 & 2004 & 2007 & 2010 & 2013 & 2016 \\  \hline \\[-1.8ex]  $\rho$ & 0.46\rlap{$^{***}$} & 0.29\rlap{$^{*}$} & 0.24 & 0.19 & 0.27\rlap{$^{*}$} & 0.50\rlap{$^{***}$}\\  & (0.15\rlap{)} & (0.17\rlap{)} & (0.17\rlap{)} & (0.16\rlap{)} & (0.16\rlap{)} & (0.17\rlap{)}\\  AusCitizen & $-$3.13 & $-$2.64 & $-$2.53 & $-$0.08 & $-$3.40 & $-$1.80 \\    & (2.26\rlap{)} & (2.43\rlap{)} & (2.34\rlap{)} & (2.79\rlap{)} & (2.76\rlap{)} & (2.71\rlap{)} \\    BornAsia & 2.22 & $-$0.95 & $-$1.60 & $-$6.83\rlap{$^{**}$} & $-$3.03 & $-$0.55 \\    & (2.18\rlap{)} & (2.44\rlap{)} & (2.19\rlap{)} & (2.73\rlap{)} & (2.71\rlap{)} & (2.17\rlap{)} \\    Born\_MidEast & $-$1.15 & $-$1.59 & $-$2.01\rlap{$^{*}$} & $-$2.03 & $-$0.92 & $-$1.44 \\    & (1.07\rlap{)} & (1.20\rlap{)} & (1.11\rlap{)} & (1.27\rlap{)} & (1.24\rlap{)} & (1.13\rlap{)} \\    BornSEEuro & $-$3.21\rlap{$^{**}$} & $-$4.24\rlap{$^{***}$} & $-$3.61\rlap{$^{***}$} & $-$4.14\rlap{$^{***}$} & $-$3.69\rlap{$^{***}$} & $-$2.72\rlap{$^{***}$} \\    & (1.42\rlap{)} & (1.46\rlap{)} & (1.02\rlap{)} & (1.19\rlap{)} & (1.07\rlap{)} & (0.97\rlap{)} \\    Born\_UK & 0.25 & $-$0.07 & 0.34 & 0.56 & $-$0.09 & $-$1.32 \\    & (1.00\rlap{)} & (0.98\rlap{)} & (0.90\rlap{)} & (1.07\rlap{)} & (1.04\rlap{)} & (1.04\rlap{)} \\    BornElsewhere & $-$5.04 & $-$4.91 & $-$4.13 & 2.35 & $-$5.23 & $-$4.14 \\    & (3.30\rlap{)} & (3.68\rlap{)} & (3.38\rlap{)} & (4.23\rlap{)} & (4.15\rlap{)} & (3.97\rlap{)} \\    Buddhism & $-$0.49 & $-$0.17 & $-$1.37 & $-$0.83 & $-$0.12 & $-$1.60 \\    & (1.39\rlap{)} & (1.61\rlap{)} & (1.61\rlap{)} & (1.80\rlap{)} & (1.68\rlap{)} & (1.56\rlap{)} \\    Christianity & $-$2.48 & $-$1.23 & 0.38 & 0.50 & 2.41 & 1.68 \\    & (1.73\rlap{)} & (1.85\rlap{)} & (1.83\rlap{)} & (1.99\rlap{)} & (1.85\rlap{)} & (1.78\rlap{)} \\    CurrentlyStudying & $-$2.19\rlap{$^{**}$} & $-$0.13 & 2.06\rlap{$^{*}$} & 2.12\rlap{$^{*}$} & 1.15 & $-$0.16 \\    & (0.99\rlap{)} & (1.13\rlap{)} & (1.17\rlap{)} & (1.25\rlap{)} & (1.26\rlap{)} & (1.18\rlap{)} \\    DeFacto & $-$6.44\rlap{$^{***}$} & $-$5.37\rlap{$^{**}$} & $-$6.43\rlap{$^{***}$} & $-$8.07\rlap{$^{***}$} & $-$6.56\rlap{$^{**}$} & $-$8.53\rlap{$^{***}$} \\    & (1.87\rlap{)} & (2.48\rlap{)} & (2.31\rlap{)} & (3.06\rlap{)} & (3.11\rlap{)} & (2.83\rlap{)} \\    DiffAddress & 3.88\rlap{$^{***}$} & 5.06\rlap{$^{***}$} & 4.22\rlap{$^{***}$} & 5.57\rlap{$^{***}$} & 3.53\rlap{$^{*}$} & 5.67\rlap{$^{***}$} \\    & (0.94\rlap{)} & (1.12\rlap{)} & (0.99\rlap{)} & (1.76\rlap{)} & (1.91\rlap{)} & (1.60\rlap{)} \\    Distributive & 1.27 & 2.01\rlap{$^{*}$} & 1.36 & 1.57 & 2.10\rlap{$^{*}$} & 1.20 \\    & (1.12\rlap{)} & (1.21\rlap{)} & (1.13\rlap{)} & (1.34\rlap{)} & (1.27\rlap{)} & (1.21\rlap{)} \\    Education & 1.08 & 0.52 & $-$5.52\rlap{$^{*}$} & $-$4.08 & $-$4.44 & $-$7.07\rlap{$^{**}$} \\    & (2.38\rlap{)} & (3.12\rlap{)} & (3.27\rlap{)} & (3.95\rlap{)} & (3.78\rlap{)} & (3.55\rlap{)} \\    Extractive & 4.83\rlap{$^{***}$} & 5.45\rlap{$^{***}$} & 5.37\rlap{$^{***}$} & 7.31\rlap{$^{***}$} & 6.71\rlap{$^{***}$} & 7.43\rlap{$^{***}$} \\    & (1.48\rlap{)} & (1.42\rlap{)} & (1.36\rlap{)} & (1.56\rlap{)} & (1.47\rlap{)} & (1.39\rlap{)} \\    FamHouseSize & $-$0.16 & 0.87 & $-$2.40 & $-$2.53 & $-$3.26 & $-$2.91 \\    & (2.19\rlap{)} & (2.72\rlap{)} & (2.69\rlap{)} & (3.25\rlap{)} & (3.28\rlap{)} & (2.90\rlap{)} \\    Incomes & 4.36\rlap{$^{**}$} & 5.03\rlap{$^{*}$} & 9.45\rlap{$^{***}$} & 7.09\rlap{$^{**}$} & 7.97\rlap{$^{***}$} & 12.20\rlap{$^{***}$} \\    & (1.77\rlap{)} & (2.66\rlap{)} & (2.75\rlap{)} & (3.25\rlap{)} & (2.92\rlap{)} & (2.75\rlap{)} \\    Indigenous & 2.91\rlap{$^{*}$} & 1.97 & 2.48 & 2.84 & 0.67 & $-$0.05 \\    & (1.68\rlap{)} & (1.95\rlap{)} & (1.75\rlap{)} & (2.16\rlap{)} & (2.14\rlap{)} & (2.00\rlap{)} \\    Islam & $-$0.92 & $-$0.97 & $-$0.54 & $-$2.50 & $-$0.82 & $-$0.95 \\    & (1.22\rlap{)} & (1.36\rlap{)} & (1.27\rlap{)} & (1.52\rlap{)} & (1.42\rlap{)} & (1.34\rlap{)} \\    Judaism & 1.88\rlap{$^{*}$} & 1.78 & 2.66\rlap{$^{***}$} & 1.97\rlap{$^{*}$} & 2.74\rlap{$^{**}$} & 1.65\rlap{$^{*}$} \\    & (1.05\rlap{)} & (1.13\rlap{)} & (1.01\rlap{)} & (1.15\rlap{)} & (1.10\rlap{)} & (1.00\rlap{)} \\    ManagerAdmin & 2.06\rlap{$^{***}$} & 3.32\rlap{$^{***}$} & 6.00\rlap{$^{***}$} & 5.47\rlap{$^{***}$} & 5.04\rlap{$^{***}$} & 5.78\rlap{$^{***}$} \\    & (0.71\rlap{)} & (0.93\rlap{)} & (0.90\rlap{)} & (1.08\rlap{)} & (1.03\rlap{)} & (1.06\rlap{)} \\    Married & 0.44 & 0.11 & $-$1.22 & $-$0.22 & 0.91 & $-$2.34 \\    & (2.31\rlap{)} & (2.96\rlap{)} & (2.83\rlap{)} & (3.15\rlap{)} & (3.03\rlap{)} & (2.81\rlap{)} \\    MedianAge & 2.32\rlap{$^{*}$} & 4.96\rlap{$^{***}$} & 3.66\rlap{$^{**}$} & 4.00\rlap{$^{*}$} & 2.30 & 2.87 \\    & (1.32\rlap{)} & (1.65\rlap{)} & (1.81\rlap{)} & (2.26\rlap{)} & (2.08\rlap{)} & (1.79\rlap{)} \\    NoReligion & $-$1.57 & $-$0.92 & 0.56 & $-$0.30 & 1.02 & 1.31 \\    & (1.59\rlap{)} & (1.71\rlap{)} & (1.73\rlap{)} & (1.92\rlap{)} & (1.94\rlap{)} & (2.04\rlap{)} \\    OneParentHouse & $-$1.73 & $-$0.45 & $-$0.75 & $-$1.46 & $-$0.77 & $-$0.74 \\    & (1.36\rlap{)} & (1.59\rlap{)} & (1.49\rlap{)} & (1.69\rlap{)} & (1.57\rlap{)} & (1.47\rlap{)} \\    OtherLanguage & $-$0.44 & 5.92 & 9.98\rlap{$^{**}$} & 11.24\rlap{$^{**}$} & 9.00\rlap{$^{*}$} & 9.84\rlap{$^{**}$} \\    & (3.22\rlap{)} & (4.16\rlap{)} & (3.91\rlap{)} & (4.76\rlap{)} & (4.66\rlap{)} & (4.44\rlap{)} \\    PropertyOwned & $-$0.46 & $-$0.53 & 0.67 & $-$0.94 & $-$0.48 & 1.41 \\    & (1.37\rlap{)} & (1.50\rlap{)} & (1.43\rlap{)} & (1.76\rlap{)} & (1.67\rlap{)} & (1.50\rlap{)} \\    RentLoanPrice & $-$1.57 & $-$3.32\rlap{$^{*}$} & $-$4.01\rlap{$^{**}$} & $-$0.97 & $-$0.70 & $-$0.89 \\    & (1.49\rlap{)} & (1.76\rlap{)} & (1.67\rlap{)} & (2.07\rlap{)} & (2.07\rlap{)} & (2.16\rlap{)} \\    SocialServ & 2.51\rlap{$^{*}$} & 1.65 & 2.47\rlap{$^{*}$} & 2.53\rlap{$^{*}$} & 2.35\rlap{$^{*}$} & 4.45\rlap{$^{***}$} \\    & (1.33\rlap{)} & (1.41\rlap{)} & (1.29\rlap{)} & (1.47\rlap{)} & (1.32\rlap{)} & (1.19\rlap{)} \\    Transformative & 3.24\rlap{$^{**}$} & 4.73\rlap{$^{***}$} & 4.84\rlap{$^{***}$} & 4.46\rlap{$^{**}$} & 3.56\rlap{$^{**}$} & 4.58\rlap{$^{***}$} \\    & (1.55\rlap{)} & (1.78\rlap{)} & (1.74\rlap{)} & (1.98\rlap{)} & (1.78\rlap{)} & (1.53\rlap{)} \\    Unemployment & $-$2.45\rlap{$^{*}$} & $-$3.07\rlap{$^{*}$} & 0.29 & 0.08 & 1.67 & 2.79\rlap{$^{**}$} \\    & (1.40\rlap{)} & (1.63\rlap{)} & (1.51\rlap{)} & (1.76\rlap{)} & (1.51\rlap{)} & (1.37\rlap{)} \\    Constant & 50.81\rlap{$^{***}$} & 52.60\rlap{$^{***}$} & 47.31\rlap{$^{***}$} & 49.93\rlap{$^{***}$} & 53.51\rlap{$^{***}$} & 50.49\rlap{$^{***}$} \\    & (0.71\rlap{)} & (0.58\rlap{)} & (0.52\rlap{)} & (0.57\rlap{)} & (0.58\rlap{)} & (0.80\rlap{)} \\   \hline \\[-1.8ex]  Observations & 150 & 150 & 150 & 150 & 150 & 150 \\  Residual Standard Error (GLS) & 4.69 & 5.04 & 4.79 & 5.63 & 5.18 & 4.88\\ \hline  \hline \\[-1.8ex]  \textit{Note:}  & \multicolumn{6}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\  \end{tabular}  \end{table}
+\begin{table}[!htbp]\vspace*{-0.4cm}\centering    \caption{Estimated spatial error model parameters (standard errors) for each of the six election years.\vspace*{-0.4cm}}    \label{tab:printmodels}  \scriptsize\tabcolsep=0.35cm \begin{tabular}{@{\extracolsep{1pt}}lrrrrrr} \\[-1.8ex]\hline  \hline \\[-1.8ex]  \\[-1.8ex] & \multicolumn{6}{c}{\vspace{-0.4cm}} \\   & 2001 & 2004 & 2007 & 2010 & 2013 & 2016 \\  \hline \\[-1.8ex]  $\rho$ & 0.46\rlap{$^{***}$} & 0.29\rlap{$^{*}$} & 0.24 & 0.19 & 0.27\rlap{$^{*}$} & 0.50\rlap{$^{***}$}\\  & (0.15\rlap{)} & (0.17\rlap{)} & (0.17\rlap{)} & (0.16\rlap{)} & (0.16\rlap{)} & (0.17\rlap{)}\\  AusCitizen & $-$3.13 & $-$2.64 & $-$2.53 & $-$0.08 & $-$3.40 & $-$1.80 \\    & (2.26\rlap{)} & (2.43\rlap{)} & (2.34\rlap{)} & (2.79\rlap{)} & (2.76\rlap{)} & (2.71\rlap{)} \\    BornAsia & 2.22 & $-$0.95 & $-$1.60 & $-$6.83\rlap{$^{**}$} & $-$3.03 & $-$0.55 \\    & (2.18\rlap{)} & (2.44\rlap{)} & (2.19\rlap{)} & (2.73\rlap{)} & (2.71\rlap{)} & (2.17\rlap{)} \\    Born\_MidEast & $-$1.15 & $-$1.59 & $-$2.01\rlap{$^{*}$} & $-$2.03 & $-$0.92 & $-$1.44 \\    & (1.07\rlap{)} & (1.20\rlap{)} & (1.11\rlap{)} & (1.27\rlap{)} & (1.24\rlap{)} & (1.13\rlap{)} \\    BornSEEuro & $-$3.21\rlap{$^{**}$} & $-$4.24\rlap{$^{***}$} & $-$3.61\rlap{$^{***}$} & $-$4.14\rlap{$^{***}$} & $-$3.69\rlap{$^{***}$} & $-$2.72\rlap{$^{***}$} \\    & (1.42\rlap{)} & (1.46\rlap{)} & (1.02\rlap{)} & (1.19\rlap{)} & (1.07\rlap{)} & (0.97\rlap{)} \\    Born\_UK & 0.25 & $-$0.07 & 0.34 & 0.56 & $-$0.09 & $-$1.32 \\    & (1.00\rlap{)} & (0.98\rlap{)} & (0.90\rlap{)} & (1.07\rlap{)} & (1.04\rlap{)} & (1.04\rlap{)} \\    BornElsewhere & $-$5.04 & $-$4.91 & $-$4.13 & 2.35 & $-$5.23 & $-$4.14 \\    & (3.30\rlap{)} & (3.68\rlap{)} & (3.38\rlap{)} & (4.23\rlap{)} & (4.15\rlap{)} & (3.97\rlap{)} \\    Buddhism & $-$0.49 & $-$0.17 & $-$1.37 & $-$0.83 & $-$0.12 & $-$1.60 \\    & (1.39\rlap{)} & (1.61\rlap{)} & (1.61\rlap{)} & (1.80\rlap{)} & (1.68\rlap{)} & (1.56\rlap{)} \\    Christianity & $-$2.48 & $-$1.23 & 0.38 & 0.50 & 2.41 & 1.68 \\    & (1.73\rlap{)} & (1.85\rlap{)} & (1.83\rlap{)} & (1.99\rlap{)} & (1.85\rlap{)} & (1.78\rlap{)} \\    CurrentlyStudying & $-$2.19\rlap{$^{**}$} & $-$0.13 & 2.06\rlap{$^{*}$} & 2.12\rlap{$^{*}$} & 1.15 & $-$0.16 \\    & (0.99\rlap{)} & (1.13\rlap{)} & (1.17\rlap{)} & (1.25\rlap{)} & (1.26\rlap{)} & (1.18\rlap{)} \\    DeFacto & $-$6.44\rlap{$^{***}$} & $-$5.37\rlap{$^{**}$} & $-$6.43\rlap{$^{***}$} & $-$8.07\rlap{$^{***}$} & $-$6.56\rlap{$^{**}$} & $-$8.53\rlap{$^{***}$} \\    & (1.87\rlap{)} & (2.48\rlap{)} & (2.31\rlap{)} & (3.06\rlap{)} & (3.11\rlap{)} & (2.83\rlap{)} \\    DiffAddress & 3.88\rlap{$^{***}$} & 5.06\rlap{$^{***}$} & 4.22\rlap{$^{***}$} & 5.57\rlap{$^{***}$} & 3.53\rlap{$^{*}$} & 5.67\rlap{$^{***}$} \\    & (0.94\rlap{)} & (1.12\rlap{)} & (0.99\rlap{)} & (1.76\rlap{)} & (1.91\rlap{)} & (1.60\rlap{)} \\    Distributive & 1.27 & 2.01\rlap{$^{*}$} & 1.36 & 1.57 & 2.10\rlap{$^{*}$} & 1.20 \\    & (1.12\rlap{)} & (1.21\rlap{)} & (1.13\rlap{)} & (1.34\rlap{)} & (1.27\rlap{)} & (1.21\rlap{)} \\    Education & 1.08 & 0.52 & $-$5.52\rlap{$^{*}$} & $-$4.08 & $-$4.44 & $-$7.07\rlap{$^{**}$} \\    & (2.38\rlap{)} & (3.12\rlap{)} & (3.27\rlap{)} & (3.95\rlap{)} & (3.78\rlap{)} & (3.55\rlap{)} \\    Extractive & 4.83\rlap{$^{***}$} & 5.45\rlap{$^{***}$} & 5.37\rlap{$^{***}$} & 7.31\rlap{$^{***}$} & 6.71\rlap{$^{***}$} & 7.43\rlap{$^{***}$} \\    & (1.48\rlap{)} & (1.42\rlap{)} & (1.36\rlap{)} & (1.56\rlap{)} & (1.47\rlap{)} & (1.39\rlap{)} \\    FamHouseSize & $-$0.16 & 0.87 & $-$2.40 & $-$2.53 & $-$3.26 & $-$2.91 \\    & (2.19\rlap{)} & (2.72\rlap{)} & (2.69\rlap{)} & (3.25\rlap{)} & (3.28\rlap{)} & (2.90\rlap{)} \\    Incomes & 4.36\rlap{$^{**}$} & 5.03\rlap{$^{*}$} & 9.45\rlap{$^{***}$} & 7.09\rlap{$^{**}$} & 7.97\rlap{$^{***}$} & 12.20\rlap{$^{***}$} \\    & (1.77\rlap{)} & (2.66\rlap{)} & (2.75\rlap{)} & (3.25\rlap{)} & (2.92\rlap{)} & (2.75\rlap{)} \\    Indigenous & 2.91\rlap{$^{*}$} & 1.97 & 2.48 & 2.84 & 0.67 & $-$0.05 \\    & (1.68\rlap{)} & (1.95\rlap{)} & (1.75\rlap{)} & (2.16\rlap{)} & (2.14\rlap{)} & (2.00\rlap{)} \\    Islam & $-$0.92 & $-$0.97 & $-$0.54 & $-$2.50 & $-$0.82 & $-$0.95 \\    & (1.22\rlap{)} & (1.36\rlap{)} & (1.27\rlap{)} & (1.52\rlap{)} & (1.42\rlap{)} & (1.34\rlap{)} \\    Judaism & 1.88\rlap{$^{*}$} & 1.78 & 2.66\rlap{$^{***}$} & 1.97\rlap{$^{*}$} & 2.74\rlap{$^{**}$} & 1.65\rlap{$^{*}$} \\    & (1.05\rlap{)} & (1.13\rlap{)} & (1.01\rlap{)} & (1.15\rlap{)} & (1.10\rlap{)} & (1.00\rlap{)} \\    ManagerAdmin & 2.06\rlap{$^{***}$} & 3.32\rlap{$^{***}$} & 6.00\rlap{$^{***}$} & 5.47\rlap{$^{***}$} & 5.04\rlap{$^{***}$} & 5.78\rlap{$^{***}$} \\    & (0.71\rlap{)} & (0.93\rlap{)} & (0.90\rlap{)} & (1.08\rlap{)} & (1.03\rlap{)} & (1.06\rlap{)} \\    Married & 0.44 & 0.11 & $-$1.22 & $-$0.22 & 0.91 & $-$2.34 \\    & (2.31\rlap{)} & (2.96\rlap{)} & (2.83\rlap{)} & (3.15\rlap{)} & (3.03\rlap{)} & (2.81\rlap{)} \\    MedianAge & 2.32\rlap{$^{*}$} & 4.96\rlap{$^{***}$} & 3.66\rlap{$^{**}$} & 4.00\rlap{$^{*}$} & 2.30 & 2.87 \\    & (1.32\rlap{)} & (1.65\rlap{)} & (1.81\rlap{)} & (2.26\rlap{)} & (2.08\rlap{)} & (1.79\rlap{)} \\    NoReligion & $-$1.57 & $-$0.92 & 0.56 & $-$0.30 & 1.02 & 1.31 \\    & (1.59\rlap{)} & (1.71\rlap{)} & (1.73\rlap{)} & (1.92\rlap{)} & (1.94\rlap{)} & (2.04\rlap{)} \\    OneParentHouse & $-$1.73 & $-$0.45 & $-$0.75 & $-$1.46 & $-$0.77 & $-$0.74 \\    & (1.36\rlap{)} & (1.59\rlap{)} & (1.49\rlap{)} & (1.69\rlap{)} & (1.57\rlap{)} & (1.47\rlap{)} \\    OtherLanguage & $-$0.44 & 5.92 & 9.98\rlap{$^{**}$} & 11.24\rlap{$^{**}$} & 9.00\rlap{$^{*}$} & 9.84\rlap{$^{**}$} \\    & (3.22\rlap{)} & (4.16\rlap{)} & (3.91\rlap{)} & (4.76\rlap{)} & (4.66\rlap{)} & (4.44\rlap{)} \\    PropertyOwned & $-$0.46 & $-$0.53 & 0.67 & $-$0.94 & $-$0.48 & 1.41 \\    & (1.37\rlap{)} & (1.50\rlap{)} & (1.43\rlap{)} & (1.76\rlap{)} & (1.67\rlap{)} & (1.50\rlap{)} \\    RentLoanPrice & $-$1.57 & $-$3.32\rlap{$^{*}$} & $-$4.01\rlap{$^{**}$} & $-$0.97 & $-$0.70 & $-$0.89 \\    & (1.49\rlap{)} & (1.76\rlap{)} & (1.67\rlap{)} & (2.07\rlap{)} & (2.07\rlap{)} & (2.16\rlap{)} \\    SocialServ & 2.51\rlap{$^{*}$} & 1.65 & 2.47\rlap{$^{*}$} & 2.53\rlap{$^{*}$} & 2.35\rlap{$^{*}$} & 4.45\rlap{$^{***}$} \\    & (1.33\rlap{)} & (1.41\rlap{)} & (1.29\rlap{)} & (1.47\rlap{)} & (1.32\rlap{)} & (1.19\rlap{)} \\    Transformative & 3.24\rlap{$^{**}$} & 4.73\rlap{$^{***}$} & 4.84\rlap{$^{***}$} & 4.46\rlap{$^{**}$} & 3.56\rlap{$^{**}$} & 4.58\rlap{$^{***}$} \\    & (1.55\rlap{)} & (1.78\rlap{)} & (1.74\rlap{)} & (1.98\rlap{)} & (1.78\rlap{)} & (1.53\rlap{)} \\    Unemployment & $-$2.45\rlap{$^{*}$} & $-$3.07\rlap{$^{*}$} & 0.29 & 0.08 & 1.67 & 2.79\rlap{$^{**}$} \\    & (1.40\rlap{)} & (1.63\rlap{)} & (1.51\rlap{)} & (1.76\rlap{)} & (1.51\rlap{)} & (1.37\rlap{)} \\    Constant & 50.81\rlap{$^{***}$} & 52.60\rlap{$^{***}$} & 47.31\rlap{$^{***}$} & 49.93\rlap{$^{***}$} & 53.51\rlap{$^{***}$} & 50.49\rlap{$^{***}$} \\    & (0.71\rlap{)} & (0.58\rlap{)} & (0.52\rlap{)} & (0.57\rlap{)} & (0.58\rlap{)} & (0.80\rlap{)} \\   \hline \\[-1.8ex]  Observations & 150 & 150 & 150 & 150 & 150 & 150 \\  Residual Standard Error (GLS) & 4.69 & 5.04 & 4.79 & 5.63 & 5.18 & 4.88\\ \hline  \hline \\[-1.8ex]  & \multicolumn{6}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ \end{tabular}  \end{table}
 
@@
 The partial residuals are the residuals from the fitted model with the estimated effect an individual variable added.
-These show the direction, size and significance of an estimated effect --- the slope of the prediction line matches the estimated coefficient, and the shaded region represents a 95\% confidence band, computed using the method Breheny and Burchett (2017).
+These show the direction, size and significance of an estimated effect --- the slope of the prediction line matches the estimated coefficient, and the shaded region represents a 95\% confidence band, computed using the method \citet{visreg}.
 If a horizontal line can be drawn through the confidence band, then the effect is insignificant.
@@
 \hypertarget{income-and-unemployment}{%
-\subsubsection{Income and unemployment}\label{income-and-unemployment}}
+\subsubsection*{Income and unemployment}\label{income-and-unemployment}}
+\addcontentsline{toc}{subsubsection}{Income and unemployment}
 
@@
 \hypertarget{industry-and-type-of-work}{%
-\subsubsection{Industry and type of work}\label{industry-and-type-of-work}}
+\subsubsection*{Industry and type of work}\label{industry-and-type-of-work}}
+\addcontentsline{toc}{subsubsection}{Industry and type of work}
 
@@
 \hypertarget{household-mobility}{%
-\subsubsection{Household mobility}\label{household-mobility}}
+\subsubsection*{Household mobility}\label{household-mobility}}
+\addcontentsline{toc}{subsubsection}{Household mobility}
 
@@
 \hypertarget{relationships}{%
-\subsubsection{Relationships}\label{relationships}}
+\subsubsection*{Relationships}\label{relationships}}
+\addcontentsline{toc}{subsubsection}{Relationships}
 
@@
 \hypertarget{age}{%
-\subsubsection{Age}\label{age}}
+\subsubsection*{Age}\label{age}}
+\addcontentsline{toc}{subsubsection}{Age}
 
@@
 \hypertarget{education}{%
-\subsubsection{Education}\label{education}}
+\subsubsection*{Education}\label{education}}
+\addcontentsline{toc}{subsubsection}{Education}
 
@@
 \hypertarget{diversity}{%
-\subsubsection{Diversity}\label{diversity}}
+\subsubsection*{Diversity}\label{diversity}}
+\addcontentsline{toc}{subsubsection}{Diversity}
 
@@
 \hypertarget{a-note-on-similar-variables}{%
-\subsubsection{A note on similar variables}\label{a-note-on-similar-variables}}
+\subsubsection*{A note on similar variables}\label{a-note-on-similar-variables}}
+\addcontentsline{toc}{subsubsection}{A note on similar variables}
 
@@
 \hypertarget{residuals-by-state}{%
-\subsubsection{Residuals by state}\label{residuals-by-state}}
+\subsubsection*{Residuals by state}\label{residuals-by-state}}
+\addcontentsline{toc}{subsubsection}{Residuals by state}
 
@@
 \hypertarget{outlier-electorates}{%
-\subsubsection{Outlier electorates}\label{outlier-electorates}}
+\subsubsection*{Outlier electorates}\label{outlier-electorates}}
+\addcontentsline{toc}{subsubsection}{Outlier electorates}
 
@@
 
-This paper was produced using \texttt{RMarkdown} (Allaire et al.
-2019) and \texttt{knitr} (Xie 2015).
-All corresponding code for this paper can be found in the github repository \href{https://github.com/jforbes14/eechidna-paper}{github.com/jforbes14/eechidna-paper}, and the data used is available in the \texttt{eechidna} package (Forbes et al.
-2019).
+This paper was produced using \texttt{RMarkdown} \citep{rmarkdown} and \texttt{knitr} \citep{knitr}.
+All corresponding code for this paper can be found in the github repository \href{https://github.com/jforbes14/eechidna-paper}{github.com/jforbes14/eechidna-paper}, and the data used is available in the \texttt{eechidna} package \citep{eechidna}.
 All raw data was obtained from the Australian Electoral Commission, the Australian Bureau of Statistics and the Australian Government.
@@
 
-\hypertarget{refs}{}
-\leavevmode\hypertarget{ref-rmarkdown}{}%
-Allaire, JJ, Yihui Xie, Jonathan McPherson, Javier Luraschi, Kevin Ushey, Aron Atkins, Hadley Wickham, Joe Cheng, Winston Chang, and Richard Iannone.
-2019.
-\emph{Rmarkdown: Dynamic Documents for R}.
-\url{https://rmarkdown.rstudio.com}.
-
-\leavevmode\hypertarget{ref-Anselin88}{}%
-Anselin, Luc. 1988. \emph{Spatial Econometrics: Methods and Models}. Vol. 4. Springer Science \& Business Media.
-
-\leavevmode\hypertarget{ref-visreg}{}%
-Breheny, Patrick, and Woodrow Burchett.
-2017.
-``Visualization of Regression Models Using Visreg.'' \emph{The R Journal} 9 (2): 56--71.
-\url{https://journal.r-project.org/archive/2017/RJ-2017-046/index.html}.
-
-\leavevmode\hypertarget{ref-DavisStimson98}{}%
-Davis, R., and R.
-Stimson.
-1998.
-``Disillusionment and Disenchantment at the Fringe: Explaining the Geography of the One Nation Party Vote at the Queensland Election.'' \emph{People and Place} 6 (3): 69--82.
-
-\leavevmode\hypertarget{ref-eechidna}{}%
-Forbes, Jeremy, Di Cook, Anthony Ebert, Heike Hofmann, Rob J Hyndman, Thomas Lumley, Ben Marwick, et al.
-2019.
-\emph{Eechidna: Exploring Election and Census Highly Informative Data Nationally for Australia}.
-\url{https://CRAN.R-project.org/package=eechidna}.
-
-\leavevmode\hypertarget{ref-Forrest01}{}%
-Forrest, James, Margaret Alston, Chris Medlin, and Siti Amri.
-2001.
-``Voter Behaviour in Rural Areas: A Study of the Farrer Electoral Division in Southern New South Wales at the 1998 Federal Election.'' \emph{Australian Geographical Studies} 39 (2): 167--82.
-
-\leavevmode\hypertarget{ref-Goodchild1993}{}%
-Goodchild, M F, L Anselin, and U Deichmann. 1993. ``A Framework for the Areal Interpolation of Socioeconomic Data.'' \emph{Environment and Planning A} 25 (3): 383--97.
-
-\leavevmode\hypertarget{ref-LeSage2009}{}%
-LeSage, James, R Kelley Pace, and Robert Kelley Pace. 2009. \emph{Introduction to Spatial Econometrics}. Chapman; Hall/CRC.
-
-\leavevmode\hypertarget{ref-Liao09}{}%
-Liao, E., T.
-Shyy, and R.
-Stimson.
-2009.
-``Developing a Web‐based E‐research Facility for Socio‐spatial Analysis to Investigate Relationships Between Voting Patterns and Local Population Characteristics.'' \emph{Journal of Spatial Science} 54 (2): 63--88.
-
-\leavevmode\hypertarget{ref-Stimson06}{}%
-Stimson, R., R.
-McCrea, and T.
-Shyy.
-2006.
-``Spatially Disaggregated Modelling of Voting Outcomes and Socio-Economic Characteristics at the 2001 Australian Federal Election.'' \emph{Geographical Research} 44 (3): 242--54.
-
-\leavevmode\hypertarget{ref-Stimson09}{}%
-Stimson, Robert, and Tung-Kai Shyy. 2009. ``A Socio-Spatial Analysis of Voting for Political Parties at the 2007 Federal Election.'' \emph{People and Place} 17 (1): 39--54.
-
-\leavevmode\hypertarget{ref-Stimson12}{}%
-Stimson, R., and T. Shyy. 2012. ``And Now for Something Different: Modelling Socio-Political Landscapes.'' \emph{Annals of Regional Science} 50: 623--43.
-
-\leavevmode\hypertarget{ref-readxl}{}%
-Wickham, Hadley, Jennifer Bryan, Marcin Kalicinski, Komarov Valery, Christophe Leitenne, Bob Colbert, David Hoerl, and Evan Miller.
-2019.
-\emph{Readxl: Read Excel Files}.
-\url{https://CRAN.R-project.org/package=readxl}.
-
-\leavevmode\hypertarget{ref-dplyr}{}%
-Wickham, Hadley, Romain François, Lionel Henry, and Kirill Müller. 2019. \emph{Dplyr: A Grammar of Data Manipulation}. \url{https://CRAN.R-project.org/package=dplyr}.
-
-\leavevmode\hypertarget{ref-knitr}{}%
-Xie, Yihui. 2015. \emph{Dynamic Documents with R and Knitr}. 2nd ed. Boca Raton, Florida: Chapman; Hall/CRC. \url{http://yihui.name/knitr/}.
+\renewcommand\refname{References}
+\bibliography{references.bib}
 

### 2019-04-21 paper.Rmd
Add references subheading
Remove numbering of subsubsections
digits in table 1 fixed

@@
 
-### Differences between Census and election data
+### Differences between Census and election data {-}
 
@@
 
-### Spatio-temporal imputation
+### Spatio-temporal imputation {-}
 
@@
 hume %>%
-  kable(longtable=FALSE, booktabs=TRUE, digits=4,
+  kable(longtable=FALSE, booktabs=TRUE, digits=c(0,4,0,0,1),
         format='latex',escape=FALSE,
@@
 
-### Income and unemployment
+### Income and unemployment {-}
 
@@
 
-### Industry and type of work
+### Industry and type of work {-}
 
@@
 
-### Household mobility
+### Household mobility {-}
 
@@
 
-### Relationships
+### Relationships {-}
 
@@
 
-### Age
+### Age {-}
 
@@
 
-### Education
+### Education {-}
 
@@
 
-### Diversity
+### Diversity {-}
 
@@
 
-### A note on similar variables
+### A note on similar variables {-}
 
@@
 
-### Residuals by state
+### Residuals by state {-}
 
@@
 
-### Outlier electorates
+### Outlier electorates {-}
 
@@
 The authors would like to sincerely thank Anthony Ebert, Heike Hofmann, Thomas Lumley, Ben Marwick, Carson Sievert, Mingzhu Sun, Dilini Talagala, Nicholas Tierney, Nathaniel Tomasetti, Earo Wang and Fang Zhou, all of whom have contributed to the `eechidna` package.
+
+# References

### 2019-04-21 paper.tex
digits in table 1 fixed

@@
 }{}
+\usepackage[margin=1in]{geometry}
 \usepackage{hyperref}
@@
 \midrule
-Hume & 0.9654 & 150643 & 145427 & 90.0168\\
-Riverina & 0.2511 & 155793 & 39117 & 89.1144\\
-Eden-Monaro & 0.1109 & 147532 & 16358 & 87.9999\\
-Canberra & 0.0028 & 196037 & 548 & 85.4793\\
-Fenner & 0.0023 & 202955 & 474 & 83.6432\\
+Hume & 0.9654 & 150643 & 145427 & 90.0\\
+Riverina & 0.2511 & 155793 & 39117 & 89.1\\
+Eden-Monaro & 0.1109 & 147532 & 16358 & 88.0\\
+Canberra & 0.0028 & 196037 & 548 & 85.5\\
+Fenner & 0.0023 & 202955 & 474 & 83.6\\
 \addlinespace
-Whitlam & 0.0006 & 152280 & 92 & 89.5173\\
-Gilmore & 0.0006 & 150436 & 86 & 89.0266\\
-Calare & 0.0001 & 161298 & 21 & 87.5603\\
+Whitlam & 0.0006 & 152280 & 92 & 89.5\\
+Gilmore & 0.0006 & 150436 & 86 & 89.0\\
+Calare & 0.0001 & 161298 & 21 & 87.6\\
 \bottomrule

### 2019-04-25 paper.Rmd
Fixed axis labels of residual plots and legend for hume boundary plot.

@@
 ```
-
 ```{r hume16, fig.cap = "Census division boundaries in NSW for 2016, with the 2013 electoral boundary for Hume, shown in purple.
@@
   coord_equal() +
-  guides(alpha = FALSE, col = FALSE)
+  guides(alpha = FALSE, col = FALSE, 
+    fill = guide_legend(override.aes = list(col = c("grey50", NA), alpha = 0.4)))
 ```
@@
   guides(col = FALSE, shape = FALSE, DivisionNm = FALSE) +
-  theme_bw()
+  theme_bw() +
+  labs(x = "State", y = "Residuals")
 ```

### 2019-04-25 anzjs_elections.tex
More meaningful filenames

--- 2019-04-21 paper.tex
@@
 }{}
-\usepackage[margin=1in]{geometry}
 \usepackage{hyperref}
@@
 \end{figure}
-
 \begin{figure}[h]
@@
 Based on the distribution of the Cook's distance values, a Cook's distance greater than \(0.1\) is considered to be influential and a potential outlier.
-The electorate of Sydney (NSW) has a large Cook's distance from 2001 to 2013, due to its diverse population (language, birthplace and religion), high number of defacto relationships, high income, high household mobility and small amount of workers in extractive and transformative jobs.
+The electorate of Sydney (NSW) has a large Cook's distance from 2001 to 2013, due to its diverse population (language, birthplace and religion), high number of de facto relationships, high income, high household mobility and small amount of workers in extractive and transformative jobs.
 It has remained a strong supporter of the Labor party and the Liberal vote is severely overpredicted by the model, making it an outlier.

### 2019-04-25 anzjs_preamble.tex
Rob's re-compiled files. Ready for submission.

--- 2019-04-25 anzjs_elections.tex
@@
-\documentclass[times, doublespace]{anzsauth}
-\usepackage[]{times}
-\usepackage{amssymb,amsmath}
-\usepackage{ifxetex,ifluatex}
-\usepackage{fixltx2e} % provides \textsubscript
-\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
-  \usepackage[T1]{fontenc}
-  \usepackage[utf8]{inputenc}
-\else % if luatex or xelatex
-  \ifxetex
-    \usepackage{mathspec}
-  \else
-    \usepackage{fontspec}
-  \fi
-  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
-\fi
-% use upquote if available, for straight quotes in verbatim environments
-\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
-% use microtype if available
-\IfFileExists{microtype.sty}{%
-\usepackage{microtype}
-\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
-}{}
-\usepackage{hyperref}
-\hypersetup{unicode=true,
-            pdftitle={Spatial modelling of the two-party preferred vote in Australian federal elections: 2001--2016},
-            pdfauthor={Jeremy Forbes\^{}*, Dianne Cook, Rob J Hyndman},
-            pdfborder={0 0 0},
-            breaklinks=true}
-\urlstyle{same}  % don't use monospace font for urls
-\usepackage{natbib}
-\bibliographystyle{anzsj}
-\usepackage{longtable,booktabs}
-\usepackage{graphicx,grffile}
-\makeatletter
-\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
-\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
-\makeatother
-% Scale images if necessary, so that they will not overflow the page
-% margins by default, and it is still possible to overwrite the defaults
-% using explicit options in \includegraphics[width, height, ...]{}
-\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
-\IfFileExists{parskip.sty}{%
-\usepackage{parskip}
-}{% else
-\setlength{\parindent}{0pt}
-\setlength{\parskip}{6pt plus 2pt minus 1pt}
-}
-\setlength{\emergencystretch}{3em}  % prevent overfull lines
-\providecommand{\tightlist}{%
-  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
-\setcounter{secnumdepth}{5}
-% Redefines (sub)paragraphs to behave more like sections
-\ifx\paragraph\undefined\else
-\let\oldparagraph\paragraph
-\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
-\fi
-\ifx\subparagraph\undefined\else
-\let\oldsubparagraph\subparagraph
-\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
-\fi
-
-%%% Use protect on footnotes to avoid problems with footnotes in titles
-\let\rmarkdownfootnote\footnote%
-\def\footnote{\protect\rmarkdownfootnote}
-
-%%% Change title format to be more compact
-\usepackage{titling}
-
-% Create subtitle command for use in maketitle
-\providecommand{\subtitle}[1]{
-  \posttitle{
-    \begin{center}\large#1\end{center}
-    }
-}
-
-\setlength{\droptitle}{-2em}
-
-  \title{Spatial modelling of the two-party preferred vote in Australian federal elections: 2001--2016}
-    \pretitle{\vspace{\droptitle}\centering\huge}
-  \posttitle{\par}
-    \author{Jeremy Forbes\(^*\), Dianne Cook, Rob J Hyndman}
-    \preauthor{\centering\large\emph}
-  \postauthor{\par}
-    \date{}
-    \predate{}\postdate{}
-
 %% Any special functions or other packages can be loaded here.
@@
 Email: \texttt{jeremyforbes1995@gmail.com}}\end{center}}
-
-\begin{document}
-\maketitle
-
-\hypertarget{summary}{%
-\section*{Summary}\label{summary}}
-\addcontentsline{toc}{section}{Summary}
-
-We examine the relationships between electoral socio-demographic characteristics and two-party preference in the six Australian federal elections held between 2001 to 2016.
-Socio-demographic information is derived from the Australian Census, which occurs every five years.
-Since a Census is not directly available for each election, spatio-temporal imputation is employed to estimate Census data for the electorates at the time of each election.
-This accounts for both spatial and temporal changes in electoral characteristics between Censuses.
-To capture any spatial heterogeneity, a spatial error model is estimated for each election, which incorporates a spatially structured random effect vector that can be thought of as the unobserved political climate in each electorate.
-Over time, the impact of most socio-demographic characteristics that affect electoral two-party preference do not vary, with industry of work, incomes, household mobility and de facto relationships having strong effects in each of the six elections.
-Education and unemployment are amongst those that have varying effects.
-It is also found that between 2004 and 2013, the spatial effect was insignificant, meaning that electorates voted effectively independently.
-All data featured in this study has been contributed to the \texttt{eechidna} \texttt{R} package (available on CRAN).
-
-\textbf{Keywords:} federal election, Census, Australia, spatial modelling, imputation, data science, socio-demographics, electorates, R, eechidna
-
-\hypertarget{intro}{%
-\section{Introduction}\label{intro}}
-
-Australia has changed in many ways over the last two decades.
-Rising house prices, country-wide improvements in education, an ageing population, and a decline in religious affiliation, are just a few facets of the country's evolving socio-demographic characteristics.
-At the same time, political power has moved back and forth between the two major parties.
-In the 2007 and 2010 federal elections, the Australian Labor Party (Labor) was victorious, whereas the 2001, 2004, 2013 and 2016 elections were won by the Liberal National coalition (Liberal).
-The two-party preferred vote, a measure of support between these two parties, fluctuated between 47.3\% and 53.5\% (in favour of the Liberal party) over this period.
-This study explores how electoral socio-demographic characteristics relate to two-party preference, and whether their effects have changed over time.
-
-Electoral socio-demographics are derived from the Australian Census, and vote counts are obtained from Australian federal elections.
-Joining these two data sources is problematic as there is an inherent asynchronicity in the two types of events.
-A Census is conducted by the Australian Bureau of Statistics (ABS) every five years, whereas federal elections, conducted by the Australian Electoral Commission (AEC), usually occur every three years or so.
-The first problem addressed is that of constructing appropriate Census data for the 2004, 2007, 2010 and 2013 elections --- election years in which a Census does not occur.
-The predominant approach in previous studies was to join voting outcomes to the nearest Census, without accounting for any temporal differences \citep[see][]{DavisStimson98, Stimson06, Liao09, Stimson09}.
-Furthermore, electoral boundaries change regularly, so spatial discrepancies also arise when matching with electoral data.
-To obtain appropriate ``Census-like'' data for these four elections, electoral socio-demographics are constructed using a spatio-temporal imputation that combines areal interpolation \citep{Goodchild1993} and linear time-interpolation.
-Collecting and wrangling the raw data, along with the imputation process, are detailed in Section \ref{data}.
-All data and associated documentation relating to this procedure are available in the \texttt{eechidna} \texttt{R} package \citep{eechidna}, providing a resource for future analysis.
-
-Previous work on modelling Australian federal elections has found that aggregate socio-demographics are relatively good predictors of voting outcomes.
-\citet{Forrest01} used multiple regression to model the Liberal and Labor primary vote for polling booths in the Farrer electorate in 1998 as a function of Census variables from 1996.
-\citet{Stimson06}, \citet{Stimson09} and \citet{Stimson12} used principal component analysis of polling booths in the 2001, 2004 and 2007 elections respectively, also finding that socio-demographic characteristics of polling booths are linked to their two-party preferred vote.
-In contrast, \citet{Stimson09} models the polling booth swing vote (change in the two-party preferred vote) in the 2007 election, finding that little of the swing vote can be explained by Census data.
-
-Instead of analyzing a single election in isolation, this paper employs a consistent model framework across six elections so that temporal changes in the effects of socio-demographics can be observed.
-Each federal election is modelled with a cross-sectional dataset.
-The cross-sectional dataset for each election used here consists of the two-party preferred vote (as the response variable), and a set of common socio-demographic variables (as the explanatory variables) that characterize each electorate.
-To prepare these datasets, socio-demographic variables are first standardized, and then a principal component analysis is used to group variables into ``factors''.
-To account for the inherent spatial structure of the data, a spatial error model is then estimated for each election.
-
-The paper is organised as follows.
-Section \ref{data} describes the data collection, joining and cleaning.
-These pre-processing steps and model details are discussed in Section \ref{modelling}.
-Section \ref{results} describes the inference conducted to determine signficance of effects and how these change over time.
-Section \ref{conclusion} summarises the work.
-Two supplementary sections document the contributions of others to this work and the software.
-
-\hypertarget{data}{%
-\section{Data collection, wrangling and imputation}\label{data}}
-
-\hypertarget{collecting-the-data}{%
-\subsection{Collecting the data}\label{collecting-the-data}}
-
-The voting outcome of interest is the electoral two-party preferred vote, which is provided by the Australian Electoral Commission (AEC) for the 2001, 2004, 2007, 2010, 2013 and 2016 elections via the AEC Tally Room.
-The AEC divides Australia into 150 regions, called electorates, with each corresponding to a single seat in the House of Representatives.
-Voting is compulsory in Australia, and each voter assigns a numbered preference to each available candidate in their electorate.
-The two-party preferred vote is determined by a tally of these preferences where, by convention, only the ranks of the Labor and Liberal candidates are considered.
-This is recorded as a percentage preference in favour of the Liberal party.
-
-Socio-demographic variables were derived from the Australian Census of Population and Housing (Census), which is a survey of every household in Australia, recording information such as age, gender, ethnicity, education level and income.
-There have been four Censuses so far in the 21st century, conducted in 2001, 2006, 2011 and 2016.
-The Australian Bureau of Statistics (ABS) conducts the Census and publishes aggregated information.
-The ABS uses electoral boundaries as defined by the AEC at the time of each Census, which may not match those in place at the subsequent and previous elections.
-From the available Census information aggregated at the electorate level, 65 socio-demographic variables were defined for each of the electorates to be used in the analysis.
-
-Raw data was sourced online from the AEC and ABS websites in \texttt{.csv} and \texttt{.xlsx} files.
-The formats of these files differ over the years, making extracting the appropriate information a big task.
-The functions available in the \texttt{dplyr} \citep{dplyr} and \texttt{readxl} \citep{readxl} \texttt{R} packages are particularly useful, as they provide fast consistent tools for data manipulation and functions to import \texttt{.xlsx} files (respectively).
-The 2001 and 2006 Census data are published in a format where the information for each electorate is held in a separate document making it difficult to use the \texttt{dplyr} tools.
-Instead, cells have to be selected from each individual file to construct the desired variables.
-All scripts required for the data wrangling process can be found in the github repository for the \texttt{eechidna} \texttt{R} package \citep{eechidna}, along with the raw data.
-The \texttt{eechidna} package makes this study entirely reproducible and provides a resource to help wrangle data for future Censuses and elections, when they become available.
-
-\hypertarget{joining-census-and-election-data}{%
-\subsection{Joining Census and election data}\label{joining-census-and-election-data}}
-
-\hypertarget{differences-between-census-and-election-data}{%
-\subsubsection*{Differences between Census and election data}\label{differences-between-census-and-election-data}}
-\addcontentsline{toc}{subsubsection}{Differences between Census and election data}
-
-Between 2001 and 2016 there were six elections and four Censuses (see Figure \ref{fig:timeline}).
-Electoral boundaries are redistributed regularly by the AEC, meaning that only in the years where both a Census and election occur are all boundaries likely to match --- the case for the 2001 and 2016 elections.
-Therefore, for the four elections between 2004 and 2013, both temporal and spatial differences in electorates need to be accounted for when joining the electoral two-party preferred vote with Census data.
-For these elections a spatio-temporal imputation method was employed to obtain electoral socio-demographics.
-This method uses Census information from both before and after the election of interest.
-
-\begin{figure}[h]
-
-{\centering \includegraphics[width=0.9\linewidth]{paper_files/figure-latex/timeline-1}
-
-}
-
-\caption{Timeline of Australian elections and Censuses. They do not always occur in the same year.}\label{fig:timeline}
-\end{figure}
-
-\hypertarget{spatio-temporal-imputation}{%
-\subsubsection*{Spatio-temporal imputation}\label{spatio-temporal-imputation}}
-\addcontentsline{toc}{subsubsection}{Spatio-temporal imputation}
-
-For each election, neighbouring Census information has to be combined in some way so that it represents the boundaries in place at the time of the election.
-This is done by taking the electoral boundaries and imputing the corresponding socio-demographic characteristics for each of neighbouring Censuses, thereby addressing the spatial aspect.
-Next, to deal with the temporal component, characteristics at the time of the election are constructed using linear interpolation between the spatially imputed neighbouring Census variables.
-
-To account for spatial differences, the piecewise approximation method in \citet{Goodchild1993} is adopted.
-Consider a map of source zones \(s = 1,\dots,S\), for which socio-demographic information is available, and a set of target zones \(t = 1,\dots,T\) for which information is to be imputed.
-This is described in the context of a single election, and a single neighbouring Census.
-
-Let the map of electoral boundaries at the time of a Census define the source zones, and let the boundaries at the time of the election be the target zones.
-Denote the area of intersection between a source zone \(s\) and a target zone \(t\) as \(A_{s,t}\).
-Additionally, let the population of the source zone \(s\) be \(U_s\) and the population of intersection between source zone \(s\) and target zone \(t\) be \(P_{s,t}\).
-The estimated population of intersection is given by
-\[
-  \hat{P}_{s,t} = \frac{U_s*A_{s,t}}{\sum_{t=1}^T A_{s,t}},
-  \quad\text{for all $s=1,\dots,S$ and $t=1,\dots,T$}.
-\]
-Note that this estimator implicitly assumes that populations are uniformly distributed within each source zone.
-
-In order to calculate socio-demographic information for each of the target zones, a weighted average is taken using the estimated intersection populations as weights.
-Denote a given Census variable for the target zone by \(C_t\), and the same Census variable for the source zone as \(D_s\).
-Then, estimate \(C_t\) using
-\[
-  \hat{C}_t = \frac{\sum_{s=1}^{S}{D_s*\hat{P}_{s,t}}}%
-  {\sum_{s=1}^{S}{\hat{P}_{s,t}}},
-  \quad\text{for each $t=1,\dots,T$}.
-\]
-This concludes the spatial imputation of the socio-demographic characteristics for one target zone (a single electoral boundary), at the time of only one of the neighbouring Censuses.
-This process is repeated for all of the target zones, and then for the other neighbouring Census.
-
-\enlargethispage*{0.5cm}
-
-To account for temporal changes, linear interpolation is used between Census years to get the final estimate of a Census variable for the target zone in the election year.
-Let \(y_1\) be the year of the Census preceding an election, let \(y_2\) be the year of the election, and \(y_3\) be the year of the Census that follows.
-Add this year subscript to the Census variable estimate \(\hat{C}_t\), resulting in \(\hat{C}_{t,y}\).
-Linear interpolating between these Census years results an imputed value for the election year, given by
-\[
-  \hat{C}_{t,y_2} = \frac{y_3-y_2}{y_3-y_1} * \hat{C}_{t,y_1} +
-    \frac{y_2-y_1}{y_3-y_1} * \hat{C}_{t,y_3}.
-\]
-Implicitly this assumes that population characteristics change in a linear manner over time.
-
-\hypertarget{an-illustration-of-the-spatio-temporal-imputation}{%
-\subsubsection*{An illustration of the spatio-temporal imputation}\label{an-illustration-of-the-spatio-temporal-imputation}}
-\addcontentsline{toc}{subsubsection}{An illustration of the spatio-temporal imputation}
-
-Census data is publicly available at different levels of aggregation, ranging from SA1 (over 50,000 zones) to electoral divisions (150 zones).
-For this study, electoral divisions are used as source zones, and the imputation method is applied to produce socio-demographic variables for each of the 2004, 2007, 2010 and 2013 elections.
-As mentioned earlier, there is no need to impute socio-demographic variables for the 2001 and 2016 elections.
-To illustrate the method, consider the imputation of socio-demographic variables for the electorate of Hume in New South Wales (NSW) at the time of the 2013 federal election.
-The boundaries shown in Figure \ref{fig:hume13} define all target zones in NSW for 2013, with the target zone of interest (Hume) shaded purple.
-
-\begin{figure}[h]
-
-{\centering \includegraphics[height=0.28\textheight]{paper_files/figure-latex/hume13-1}
-
-}
-
-\caption{Some of the electoral boundaries in NSW for 2013, with the electoral boundary for Hume shown in purple.}\label{fig:hume13}
-\end{figure}
-\begin{figure}[h]
-
-{\centering \includegraphics[height=0.28\textheight]{paper_files/figure-latex/hume16-1}
-
-}
-
-\caption{Census division boundaries in NSW for 2016, with the 2013 electoral boundary for Hume, shown in purple. The purple region is not contained within a single Census division.}\label{fig:hume16}
-\end{figure}
-
-The corresponding source zones from the 2016 Census are shown in Figure \ref{fig:hume16}.
-As can be seen, the Hume boundary from the 2013 election (shaded purple) does not perfectly match any of the source zones.
-
-There are many source zones from the 2016 Census that intersect with this purple region, including the divisions of Riverina, Eden-Monaro and Hume, along with smaller intersecting areas with Fenner, Calare, Gilmore and Whitlam.
-The proportion of each source zone that overlaps with the purple region is calculated, and used to obtain the intersecting populations \(\hat{P}_{s,t}\).
-
-\begin{table}[t]
-
-\caption{\label{tab:hume1}Population from each intersecting Census source zone (2016) that is allocated to the target zone (purple region - Hume electoral boundary in 2013), and the corresponding proportion of Australian citizens in each of these source zones.}
-\centering
-\resizebox{\linewidth}{!}{
-\fontsize{12}{14}\selectfont
-\begin{tabular}{lrrrr}
-\toprule
-Source zone (2016) & Proportion & Source zone population & Population allocated to purple region: $\hat{P}_{s,t}$ & AusCitizen (\%): $D_s$\\
-\midrule
-Hume & 0.9654 & 150643 & 145427 & 90.0\\
-Riverina & 0.2511 & 155793 & 39117 & 89.1\\
-Eden-Monaro & 0.1109 & 147532 & 16358 & 88.0\\
-Canberra & 0.0028 & 196037 & 548 & 85.5\\
-Fenner & 0.0023 & 202955 & 474 & 83.6\\
-\addlinespace
-Whitlam & 0.0006 & 152280 & 92 & 89.5\\
-Gilmore & 0.0006 & 150436 & 86 & 89.0\\
-Calare & 0.0001 & 161298 & 21 & 87.6\\
-\bottomrule
-\end{tabular}}
-\end{table}
-
-Now consider the socio-demographic variable \emph{AusCitizen}, the proportion of people in the region who are Australian citizens.
-A weighted average of \emph{AusCitizen}, with the allocated population from each source zone as weights, yields \(\hat{C}_{\text{Hume},2016} = 89.65\%\).
-Repeating this process using the 2011 Census yields \(\hat{C}_{\text{Hume},2011} = 91.00\%\).
-Finally, linear interpolation between 2011 and 2016 yields the 2013 estimate:
-\[
-  \hat{C}_{\text{Hume},2013}
-    = \frac{3}{5} \hat{C}_{\text{Hume},2011} + \frac{2}{5} \hat{C}_{\text{Hume},2016}
-    % = \frac{3}{5} \cdot 91.00 \% + \frac{2}{5} \cdot 89.65 \% \\
-    =  90.46\%.
-\]
-This is done for each of the 65 socio-demographic variables, and is repeated for each of the 149 remaining target zones corresponding with 2013 electorates.
-
-\hypertarget{modelling}{%
-\section{Modelling}\label{modelling}}
-
-Following this process, electoral socio-demographic variables are available for each of the six elections and can be joined with their corresponding two-party preferred votes.
-Before choosing an appropriate model, two issues with the socio-demographic variables need to be addressed.
-First, variable scales change over the years, making it important to standardize variables.
-Second, many variables represent similar information and where appropriate, should be combined in some way.
-Principal component analysis is used to identify variables covary, from which intuitive groupings are selected to be combined into a single variable.
-This also reduces the dimension of the data.
-After these steps, a model specification is chosen.
-
-\hypertarget{standardizing-variables}{%
-\subsection{Standardizing variables}\label{standardizing-variables}}
-
-Many of the socio-demographic variables have changing scales over the years.
-For example, inflation-adjusted median rental prices increased across almost all electorates, with median rent of 200 dollars per week placing an electorate in the 90th percentile in 2001, but only the 30th percentile in 2016.
-In order for socio-demographic effects to be comparable across years, all explanatory variables are standardized to have mean zero and variance one within each election year.
-By standardizing, each variable is reported as a relative measure compared to all other electorates in the same year.
-
-\hypertarget{creating-factors}{%
-\subsection{Creating factors}\label{creating-factors}}
-
-There are only \(N = 150\) observations (electorates) in each election and \(p = 65\) socio-demographic variables in each cross-section, with many variables represent similar information about an electorate.
-Any model that uses all variables would face serious problems with multi-collinearity and over-fitting, which would likely lead to erroneous conclusions regarding variable significance.
-To address this, groups of variables that represent similar information are combined into ``factors''\footnote{A preliminary step involved removing all age bands, because age is represented by median age, and to remove variables relating to particular denominations of Christianity.}.
-
-A factor is created from a group of variables if there is an intuitive reason as to why they should represent similar information and if there is evidence to suggest that they covary.
-For example, a potential group would be variables relating to electoral incomes --- median family, household and personal incomes.
-To determine which variables covary, principal component analysis is used on a combined dataset of socio-demographic variables from all six elections.
-It is appropriate to compute principal components in this way because when computed separately for each election, scree plots level off after four components and the loadings of the first four components are similar across the elections.
-
-Only the first four principal components from the combined dataset are considered, as the scree plot corresponding to the combined dataset levels off after the fourth component.
-Variables that have a large loading in a particular component are deemed to covary, with a loading with magnitude greater than 0.15 being considered large.
-Six factors are created using this criteria.
-These are:
-\texttt{Incomes} (median personal, household and family incomes);
-\texttt{Unemployment} (unemployment and labor force participation rates);
-\texttt{ProportyOwned} (rates of housing ownership, mortgages, renting and government housing);
-\texttt{RentLoanPrice} (median rental and loan repayments);
-\texttt{FamHouseSize} (average household size, ratio of people to families and household makeup (single person, couple with kids and couple without kids); and
-\texttt{Education} (high school and university qualifications, jobs requiring higher levels of education as well as vocational course completions and jobs that do not require higher education levels, such as laborer or tradesperson).
-For each of these groups, variables with positive loadings are added and those with negative loadings are subtracted to create a factor.
-After computing these sums, each factor is standardized to have mean zero and variance one, within each election.
-
-There are \(p=30\) variables in the resultant predictor set, with all of these used in the regression for each election.
-
-\hypertarget{regression-incorporating-spatially-dependent-errors}{%
-\subsection{Regression incorporating spatially dependent errors}\label{regression-incorporating-spatially-dependent-errors}}
-
-An identical model specification is used for each of the six elections, with each election modelled separately.
-This allows the socio-demographic effects to be estimated separately for each election year, facilitating analysis of temporal changes in variable effects.
-This approach is preferable to using a single longitudinal model because it avoids any concerns of undue bias stemming from incorrectly imposed time-varying restrictions on any variable.
-Without such restrictions, a pooled cross-sectional model does not yield any distinct advantage over separate cross-sections.
-The panel approach is avoided because of how frequently electoral boundaries change, noting that electorates with the same name across elections are not guaranteed to represent the same geographical region.
-Therefore any fixed or random effects models would be difficult to estimate without implementing consistent boundaries, which would require further imputation.
-
-For each cross-section, let the response \(\bm{y}\) be the vector two-party preferred vote in favour of the Liberal party; for example, \(y_i = 70\) represents a 70\% preference for Liberal, 30\% for Labor, in electorate \(i\).
-Although \(y_i\) lies in the interval \((0,100)\), observed values are never close to 0 or 100 (minimum \(24.05 \%\) and maximum \(74.90 \%\)), so there is no need to formally impose the constraint of \(y_i \in [0,100]\).
-Furthermore, the responses are found to be spatially correlated in each election (Moran's I test, \(p \le 7\cdot10^{-15}\)).
-This is not surprising as electorates are aggregate spatial units, and hence the spatial structure of the data must modelled appropriately.
-
-The spatial error model \citep{Anselin88} is chosen because it captures spatial heterogeneity by incorporating a spatially structured random effect vector \citep{LeSage2009}.
-In this context, the random effect can be thought of as capturing the unobserved political climate in each electorate, where this climate is correlated with the climate in neighbouring electorates, under the assumption that the climate is independent of electoral socio-demographics.
-
-\begin{table}[!htbp]\vspace*{-0.4cm}\centering    \caption{Estimated spatial error model parameters (standard errors) for each of the six election years.\vspace*{-0.4cm}}    \label{tab:printmodels}  \scriptsize\tabcolsep=0.35cm \begin{tabular}{@{\extracolsep{1pt}}lrrrrrr} \\[-1.8ex]\hline  \hline \\[-1.8ex]  \\[-1.8ex] & \multicolumn{6}{c}{\vspace{-0.4cm}} \\   & 2001 & 2004 & 2007 & 2010 & 2013 & 2016 \\  \hline \\[-1.8ex]  $\rho$ & 0.46\rlap{$^{***}$} & 0.29\rlap{$^{*}$} & 0.24 & 0.19 & 0.27\rlap{$^{*}$} & 0.50\rlap{$^{***}$}\\  & (0.15\rlap{)} & (0.17\rlap{)} & (0.17\rlap{)} & (0.16\rlap{)} & (0.16\rlap{)} & (0.17\rlap{)}\\  AusCitizen & $-$3.13 & $-$2.64 & $-$2.53 & $-$0.08 & $-$3.40 & $-$1.80 \\    & (2.26\rlap{)} & (2.43\rlap{)} & (2.34\rlap{)} & (2.79\rlap{)} & (2.76\rlap{)} & (2.71\rlap{)} \\    BornAsia & 2.22 & $-$0.95 & $-$1.60 & $-$6.83\rlap{$^{**}$} & $-$3.03 & $-$0.55 \\    & (2.18\rlap{)} & (2.44\rlap{)} & (2.19\rlap{)} & (2.73\rlap{)} & (2.71\rlap{)} & (2.17\rlap{)} \\    Born\_MidEast & $-$1.15 & $-$1.59 & $-$2.01\rlap{$^{*}$} & $-$2.03 & $-$0.92 & $-$1.44 \\    & (1.07\rlap{)} & (1.20\rlap{)} & (1.11\rlap{)} & (1.27\rlap{)} & (1.24\rlap{)} & (1.13\rlap{)} \\    BornSEEuro & $-$3.21\rlap{$^{**}$} & $-$4.24\rlap{$^{***}$} & $-$3.61\rlap{$^{***}$} & $-$4.14\rlap{$^{***}$} & $-$3.69\rlap{$^{***}$} & $-$2.72\rlap{$^{***}$} \\    & (1.42\rlap{)} & (1.46\rlap{)} & (1.02\rlap{)} & (1.19\rlap{)} & (1.07\rlap{)} & (0.97\rlap{)} \\    Born\_UK & 0.25 & $-$0.07 & 0.34 & 0.56 & $-$0.09 & $-$1.32 \\    & (1.00\rlap{)} & (0.98\rlap{)} & (0.90\rlap{)} & (1.07\rlap{)} & (1.04\rlap{)} & (1.04\rlap{)} \\    BornElsewhere & $-$5.04 & $-$4.91 & $-$4.13 & 2.35 & $-$5.23 & $-$4.14 \\    & (3.30\rlap{)} & (3.68\rlap{)} & (3.38\rlap{)} & (4.23\rlap{)} & (4.15\rlap{)} & (3.97\rlap{)} \\    Buddhism & $-$0.49 & $-$0.17 & $-$1.37 & $-$0.83 & $-$0.12 & $-$1.60 \\    & (1.39\rlap{)} & (1.61\rlap{)} & (1.61\rlap{)} & (1.80\rlap{)} & (1.68\rlap{)} & (1.56\rlap{)} \\    Christianity & $-$2.48 & $-$1.23 & 0.38 & 0.50 & 2.41 & 1.68 \\    & (1.73\rlap{)} & (1.85\rlap{)} & (1.83\rlap{)} & (1.99\rlap{)} & (1.85\rlap{)} & (1.78\rlap{)} \\    CurrentlyStudying & $-$2.19\rlap{$^{**}$} & $-$0.13 & 2.06\rlap{$^{*}$} & 2.12\rlap{$^{*}$} & 1.15 & $-$0.16 \\    & (0.99\rlap{)} & (1.13\rlap{)} & (1.17\rlap{)} & (1.25\rlap{)} & (1.26\rlap{)} & (1.18\rlap{)} \\    DeFacto & $-$6.44\rlap{$^{***}$} & $-$5.37\rlap{$^{**}$} & $-$6.43\rlap{$^{***}$} & $-$8.07\rlap{$^{***}$} & $-$6.56\rlap{$^{**}$} & $-$8.53\rlap{$^{***}$} \\    & (1.87\rlap{)} & (2.48\rlap{)} & (2.31\rlap{)} & (3.06\rlap{)} & (3.11\rlap{)} & (2.83\rlap{)} \\    DiffAddress & 3.88\rlap{$^{***}$} & 5.06\rlap{$^{***}$} & 4.22\rlap{$^{***}$} & 5.57\rlap{$^{***}$} & 3.53\rlap{$^{*}$} & 5.67\rlap{$^{***}$} \\    & (0.94\rlap{)} & (1.12\rlap{)} & (0.99\rlap{)} & (1.76\rlap{)} & (1.91\rlap{)} & (1.60\rlap{)} \\    Distributive & 1.27 & 2.01\rlap{$^{*}$} & 1.36 & 1.57 & 2.10\rlap{$^{*}$} & 1.20 \\    & (1.12\rlap{)} & (1.21\rlap{)} & (1.13\rlap{)} & (1.34\rlap{)} & (1.27\rlap{)} & (1.21\rlap{)} \\    Education & 1.08 & 0.52 & $-$5.52\rlap{$^{*}$} & $-$4.08 & $-$4.44 & $-$7.07\rlap{$^{**}$} \\    & (2.38\rlap{)} & (3.12\rlap{)} & (3.27\rlap{)} & (3.95\rlap{)} & (3.78\rlap{)} & (3.55\rlap{)} \\    Extractive & 4.83\rlap{$^{***}$} & 5.45\rlap{$^{***}$} & 5.37\rlap{$^{***}$} & 7.31\rlap{$^{***}$} & 6.71\rlap{$^{***}$} & 7.43\rlap{$^{***}$} \\    & (1.48\rlap{)} & (1.42\rlap{)} & (1.36\rlap{)} & (1.56\rlap{)} & (1.47\rlap{)} & (1.39\rlap{)} \\    FamHouseSize & $-$0.16 & 0.87 & $-$2.40 & $-$2.53 & $-$3.26 & $-$2.91 \\    & (2.19\rlap{)} & (2.72\rlap{)} & (2.69\rlap{)} & (3.25\rlap{)} & (3.28\rlap{)} & (2.90\rlap{)} \\    Incomes & 4.36\rlap{$^{**}$} & 5.03\rlap{$^{*}$} & 9.45\rlap{$^{***}$} & 7.09\rlap{$^{**}$} & 7.97\rlap{$^{***}$} & 12.20\rlap{$^{***}$} \\    & (1.77\rlap{)} & (2.66\rlap{)} & (2.75\rlap{)} & (3.25\rlap{)} & (2.92\rlap{)} & (2.75\rlap{)} \\    Indigenous & 2.91\rlap{$^{*}$} & 1.97 & 2.48 & 2.84 & 0.67 & $-$0.05 \\    & (1.68\rlap{)} & (1.95\rlap{)} & (1.75\rlap{)} & (2.16\rlap{)} & (2.14\rlap{)} & (2.00\rlap{)} \\    Islam & $-$0.92 & $-$0.97 & $-$0.54 & $-$2.50 & $-$0.82 & $-$0.95 \\    & (1.22\rlap{)} & (1.36\rlap{)} & (1.27\rlap{)} & (1.52\rlap{)} & (1.42\rlap{)} & (1.34\rlap{)} \\    Judaism & 1.88\rlap{$^{*}$} & 1.78 & 2.66\rlap{$^{***}$} & 1.97\rlap{$^{*}$} & 2.74\rlap{$^{**}$} & 1.65\rlap{$^{*}$} \\    & (1.05\rlap{)} & (1.13\rlap{)} & (1.01\rlap{)} & (1.15\rlap{)} & (1.10\rlap{)} & (1.00\rlap{)} \\    ManagerAdmin & 2.06\rlap{$^{***}$} & 3.32\rlap{$^{***}$} & 6.00\rlap{$^{***}$} & 5.47\rlap{$^{***}$} & 5.04\rlap{$^{***}$} & 5.78\rlap{$^{***}$} \\    & (0.71\rlap{)} & (0.93\rlap{)} & (0.90\rlap{)} & (1.08\rlap{)} & (1.03\rlap{)} & (1.06\rlap{)} \\    Married & 0.44 & 0.11 & $-$1.22 & $-$0.22 & 0.91 & $-$2.34 \\    & (2.31\rlap{)} & (2.96\rlap{)} & (2.83\rlap{)} & (3.15\rlap{)} & (3.03\rlap{)} & (2.81\rlap{)} \\    MedianAge & 2.32\rlap{$^{*}$} & 4.96\rlap{$^{***}$} & 3.66\rlap{$^{**}$} & 4.00\rlap{$^{*}$} & 2.30 & 2.87 \\    & (1.32\rlap{)} & (1.65\rlap{)} & (1.81\rlap{)} & (2.26\rlap{)} & (2.08\rlap{)} & (1.79\rlap{)} \\    NoReligion & $-$1.57 & $-$0.92 & 0.56 & $-$0.30 & 1.02 & 1.31 \\    & (1.59\rlap{)} & (1.71\rlap{)} & (1.73\rlap{)} & (1.92\rlap{)} & (1.94\rlap{)} & (2.04\rlap{)} \\    OneParentHouse & $-$1.73 & $-$0.45 & $-$0.75 & $-$1.46 & $-$0.77 & $-$0.74 \\    & (1.36\rlap{)} & (1.59\rlap{)} & (1.49\rlap{)} & (1.69\rlap{)} & (1.57\rlap{)} & (1.47\rlap{)} \\    OtherLanguage & $-$0.44 & 5.92 & 9.98\rlap{$^{**}$} & 11.24\rlap{$^{**}$} & 9.00\rlap{$^{*}$} & 9.84\rlap{$^{**}$} \\    & (3.22\rlap{)} & (4.16\rlap{)} & (3.91\rlap{)} & (4.76\rlap{)} & (4.66\rlap{)} & (4.44\rlap{)} \\    PropertyOwned & $-$0.46 & $-$0.53 & 0.67 & $-$0.94 & $-$0.48 & 1.41 \\    & (1.37\rlap{)} & (1.50\rlap{)} & (1.43\rlap{)} & (1.76\rlap{)} & (1.67\rlap{)} & (1.50\rlap{)} \\    RentLoanPrice & $-$1.57 & $-$3.32\rlap{$^{*}$} & $-$4.01\rlap{$^{**}$} & $-$0.97 & $-$0.70 & $-$0.89 \\    & (1.49\rlap{)} & (1.76\rlap{)} & (1.67\rlap{)} & (2.07\rlap{)} & (2.07\rlap{)} & (2.16\rlap{)} \\    SocialServ & 2.51\rlap{$^{*}$} & 1.65 & 2.47\rlap{$^{*}$} & 2.53\rlap{$^{*}$} & 2.35\rlap{$^{*}$} & 4.45\rlap{$^{***}$} \\    & (1.33\rlap{)} & (1.41\rlap{)} & (1.29\rlap{)} & (1.47\rlap{)} & (1.32\rlap{)} & (1.19\rlap{)} \\    Transformative & 3.24\rlap{$^{**}$} & 4.73\rlap{$^{***}$} & 4.84\rlap{$^{***}$} & 4.46\rlap{$^{**}$} & 3.56\rlap{$^{**}$} & 4.58\rlap{$^{***}$} \\    & (1.55\rlap{)} & (1.78\rlap{)} & (1.74\rlap{)} & (1.98\rlap{)} & (1.78\rlap{)} & (1.53\rlap{)} \\    Unemployment & $-$2.45\rlap{$^{*}$} & $-$3.07\rlap{$^{*}$} & 0.29 & 0.08 & 1.67 & 2.79\rlap{$^{**}$} \\    & (1.40\rlap{)} & (1.63\rlap{)} & (1.51\rlap{)} & (1.76\rlap{)} & (1.51\rlap{)} & (1.37\rlap{)} \\    Constant & 50.81\rlap{$^{***}$} & 52.60\rlap{$^{***}$} & 47.31\rlap{$^{***}$} & 49.93\rlap{$^{***}$} & 53.51\rlap{$^{***}$} & 50.49\rlap{$^{***}$} \\    & (0.71\rlap{)} & (0.58\rlap{)} & (0.52\rlap{)} & (0.57\rlap{)} & (0.58\rlap{)} & (0.80\rlap{)} \\   \hline \\[-1.8ex]  Observations & 150 & 150 & 150 & 150 & 150 & 150 \\  Residual Standard Error (GLS) & 4.69 & 5.04 & 4.79 & 5.63 & 5.18 & 4.88\\ \hline  \hline \\[-1.8ex]  & \multicolumn{6}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ \end{tabular}  \end{table}
-
-Spatial weights are calculated in accordance with the assumption that an electorate is equally correlated with any electorate that shares a part of its boundary.
-Let \(\rho\) be the spatial autoregressive coefficient, \(\bm{v}\) be a spherical error term, \(\bm{W}\) be a matrix of spatial weights (containing information about the neighbouring regions), \(\bm{X}\) be a matrix of socio-demographic covariates, \(\bm{\beta}\) be a vector of regression coefficients and \(\bm{a}\) be a spatially structured random effect vector.
-\[
-  \bm{y} = \bm{X} \bm{\beta} + \bm{a},
-\]
-and
-\[
-  \bm{a} = \rho \bm{W} \bm{a} + \bm{v},
-\]
-where \(\bm{v} \sim N(\bm{0}, \sigma^2 \bm{I_n})\), and hence
-\[
-  \bm{y} = \bm{X} \bm{\beta} + (\bm{I}_n-\rho \bm{W})^{-1}\bm{v}.
-\]
-Estimation of the above spatial error model is undertaken using feasible generalized least squares.
-
-Table \ref{tab:printmodels} details the estimated model coefficients and their estimated standard errors, for each of the six elections.
-An interpretation of these estimated values is provided in the next section.
-
-\hypertarget{results}{%
-\section{Results}\label{results}}
-
-\hypertarget{spatial-autoregressive-parameter}{%
-\subsection{Spatial autoregressive parameter}\label{spatial-autoregressive-parameter}}
-
-The spatial autoregressive coefficient \(\rho\) is positive and significant in only the 2001 and 2016 elections (Figure \ref{fig:rhovis}), meaning that in these elections, the political climate of an electorate appears to be affected by the attitudes of its neighbours.
-Conversely, in the other four elections, the spatial effect weakens to become insignificant.
-In these years, it appears that the spatial component does not explain anything not already explained by the electoral socio-demographics, meaning electorates effectively voted independently.
-
-\begin{figure}[h]
-
-{\centering \includegraphics[height=0.2\textheight,]{paper_files/figure-latex/rhovis-1}
-
-}
-
-\caption{Estimates of the spatial autoregressive parameter for each of the six elections, reported with their individual 95\% confidence intervals.
-Only in 2001 and 2016 is there a significant spatial component.}\label{fig:rhovis}
-\end{figure}
-
-\hypertarget{country-wide-trend}{%
-\subsection{Country-wide trend}\label{country-wide-trend}}
-
-\begin{figure}[h]
-
-{\centering \includegraphics[height=0.2\textheight,]{paper_files/figure-latex/plotintercept-1}
-
-}
-
-\caption{Estimated intercept for each election, which represents the two-party preferred vote for an electorate with mean characteristics.}\label{fig:plotintercept}
-\end{figure}
-
-Since all socio-demographics have been standardized to have a mean of zero and a variance of one, the intercept in each model can be interpreted as the estimated two-party preferred vote for an electorate with mean characteristics\footnote{Mean of all variables aside from Judaism, Indigenous, Islam and Buddhism, where it assumes the mean of the log value.}.
-Figure \ref{fig:plotintercept} shows that the baseline of party preference has varied over the elections, with the biggest swing occurring in the 2007 election where the mean electorate shifted more than five percentage points in favour of the Labor party.
-
-\hypertarget{influential-socio-demographics}{%
-\subsection{Influential socio-demographics}\label{influential-socio-demographics}}
-
-\begin{figure}[h]
-
-{\centering \includegraphics{paper_files/figure-latex/partresplot1-1}
-
-}
-
-\caption{Partial residual plots by election year for a selection of predictors.
-Linear model with 95\% confidence bands overlaid.
-Most predictors have a positive relationship: the larger the value the more likely the electorate preferences the Coalition.
-The relationship is relatively robust over time, with the exception of Unemployment.}\label{fig:partresplot1}
-\end{figure}
-
-\begin{figure}[h]
-
-{\centering \includegraphics{paper_files/figure-latex/partresplot2-1}
-
-}
-
-\caption{Partial residual plots by election year for a selection of predictors.
-Linear model with 95\% confidence bands overlaid.
-Several predictors have a negative relationship: with larger values indicating the electorate more likely preferences Labor.
-Most relationships are relatively stable over elections, except OtherLanguage and Education.}\label{fig:partresplot2}
-\end{figure}
-
-To investigate the socio-demographics that have a strong effect on the two-party preferred vote, partial residual plots are used and shown in Figures \ref{fig:partresplot1} and \ref{fig:partresplot2}.
-The partial residuals are the residuals from the fitted model with the estimated effect an individual variable added.
-These show the direction, size and significance of an estimated effect --- the slope of the prediction line matches the estimated coefficient, and the shaded region represents a 95\% confidence band, computed using the method \citet{visreg}.
-If a horizontal line can be drawn through the confidence band, then the effect is insignificant.
-The estimated intercept is also added to the partial residuals for interpretability.
-Plots for each election are faceted to compare the effects over time in Figures \ref{fig:partresplot1} and \ref{fig:partresplot2}.
-Only socio-demographics that have a significant effect in at least one election are displayed in Figures \ref{fig:partresplot1} and \ref{fig:partresplot2}.
-
-It is important here to note the ecological fallacy: insights are being drawn at the electorate level, and cannot be inferred for another disaggregate level (e.g.~individual voters).
-
-\hypertarget{income-and-unemployment}{%
-\subsubsection*{Income and unemployment}\label{income-and-unemployment}}
-\addcontentsline{toc}{subsubsection}{Income and unemployment}
-
-Typically the Labor party campaigns on more progressive policies, which often include tax reform that adversely affects higher income earners, and more generous social assistance programs.
-Perhaps it is due to these policies that higher income electorates appear more likely to support the Liberal party, as the \texttt{Incomes} factor has a positive effect on Liberal preference (see row 1 in Figure \ref{fig:partresplot1}).
-This effect is significant in every election aside from 2004, where it is only marginally insignificant (\(p = 0.0613\)).
-Unemployment however, is not as influential.
-In 2001 and 2004, electorates with higher unemployment align with Labor, but over time this shifts towards support for the Liberal party, culminating in a significantly positive effect in 2016.
-
-\hypertarget{industry-and-type-of-work}{%
-\subsubsection*{Industry and type of work}\label{industry-and-type-of-work}}
-\addcontentsline{toc}{subsubsection}{Industry and type of work}
-
-Electorates with higher proportions of workers in mining, gas, water, agriculture, waste and electricity (grouped as \texttt{Extractive} industries) are consistently linked with higher support for the Liberal party, with the magnitude of this effect slightly increasing over the years (see row 3 in Figure \ref{fig:partresplot1}).
-This is unsurprising, as the Liberal party has close ties with these traditional energy industries, and typically present policies to reduce taxation on energy production.
-Furthermore, electorates with more workers in construction or manufacturing industries (\texttt{Transformative}) are also more likely to support the Liberal party (see row 4 in Figure \ref{fig:partresplot1}).
-
-Similarly, the proportion of workers in managerial, administrative, clerical and sales roles (\texttt{ManagerAdmin}) is also a significant predictor of two-party preference vote across all six elections, with a higher proportion of people working these jobs increasing Liberal support.
-The magnitude of this effect also seems to increase over the years.
-
-\hypertarget{household-mobility}{%
-\subsubsection*{Household mobility}\label{household-mobility}}
-\addcontentsline{toc}{subsubsection}{Household mobility}
-
-In each of the six elections, electorates with a higher proportion of people that have recently (in the past five years) moved house (\texttt{DiffAddress}) are more likely to support the Liberal party, although this effect was marginally insignificant in 2013 (see row 6 in Figure \ref{fig:partresplot1}.
-Having controlled for characteristics of house ownership and rental prices (via the factors \texttt{PropertyOwned} and \texttt{RentLoan} respectively), this effect is somewhat surprising.
-
-\hypertarget{relationships}{%
-\subsubsection*{Relationships}\label{relationships}}
-\addcontentsline{toc}{subsubsection}{Relationships}
-
-De facto relationships, but not marriages, are found to be an important (and significant) predictor of the two-party preferred vote in all six elections, with more de facto relationships associated with higher support for the Labor party.
-The proportion of individuals who are married however, is insignificant (not shown).
-
-\hypertarget{age}{%
-\subsubsection*{Age}\label{age}}
-\addcontentsline{toc}{subsubsection}{Age}
-
-Regions comprising more older people are often believed to be more conservative, and indeed it found that electorates with a higher median age are more likely to support the Liberal party --- although this effect is only significant in 2007 and 2010 (see row 2 in Figure \ref{fig:partresplot2}).
-
-\hypertarget{education}{%
-\subsubsection*{Education}\label{education}}
-\addcontentsline{toc}{subsubsection}{Education}
-
-Since 2007, electorates with higher education levels are associated with supporting the Labor party, although this effect is only significant in 2016.
-Before 2007, education has an almost zero effect (see row 3 in Figure \ref{fig:partresplot1}).
-
-\hypertarget{diversity}{%
-\subsubsection*{Diversity}\label{diversity}}
-\addcontentsline{toc}{subsubsection}{Diversity}
-
-Larger migrant populations from Asia, the Middle East, South-Eastern Europe, the United Kingdom and elsewhere, are either associated with Labor support, or have no effect.
-Of these areas, only South-Eastern European populations appear significant in each election, with the proportion of Asian migrants also being significant in 2010.
-Speaking other languages (aside from English) however, appears to have a far stronger effect, as observed through the \texttt{OtherLanguage} variable.
-Electorates with more diverse speech are associated with higher support for the Liberal party from 2004 onwards, with this effect being significant in 2007, 2010 and 2016.
-Furthermore, of the variables relating to religion, only Judaism shows a consistent effect, with electorates with relatively large Jewish populations more likely to vote Liberal.
-
-\hypertarget{a-note-on-similar-variables}{%
-\subsubsection*{A note on similar variables}\label{a-note-on-similar-variables}}
-\addcontentsline{toc}{subsubsection}{A note on similar variables}
-
-Many of the Census variables represent similar information, which is why factors were created and some variables were removed.
-However, some variables remain which are closely related.
-For example, an electorate's income level (via \texttt{Incomes}) is likely to be related to electoral unemployment and labor force participation (via \texttt{Unemployment}).
-In 2001, the coefficient estimate for \texttt{Unemployment} is negative but not significant, whilst the \texttt{Incomes} variables is significant.
-If the \texttt{Incomes} variable is removed from the model in 2001, \texttt{Unemployment} absorbs the negative effect, becoming significant (\(p = 0.0056\)).
-
-\hypertarget{a-closer-look-at-the-residuals}{%
-\subsection{A closer look at the residuals}\label{a-closer-look-at-the-residuals}}
-
-\hypertarget{residuals-by-state}{%
-\subsubsection*{Residuals by state}\label{residuals-by-state}}
-\addcontentsline{toc}{subsubsection}{Residuals by state}
-
-It is often hypothesized that states have systematic differences that cause their electorates to vote differently.
-Boxplots of residuals grouped by state (Figure \ref{fig:resstate}) reveal that the data reflects this -- there appears to be a state-specific effect not captured by the models.
-Tasmania and the Australian Capital Territory appear to have a bias towards Labor, whereas the Northern Territory tends towards voting Liberal.
-However, there are relatively few electorates in each of these states (five, two and two respectively), so this apparent result may be due to incumbent effects rather than an actual state-specific bias.
-
-\begin{figure}[h]
-
-{\centering \includegraphics[width=0.7\linewidth,]{paper_files/figure-latex/resstate-1}
-
-}
-
-\caption{Boxplot of residuals by state with jittered points. States ordered by median residual. A state-specific bias not captured by the model is evident.}\label{fig:resstate}
-\end{figure}
-
-\hypertarget{outlier-electorates}{%
-\subsubsection*{Outlier electorates}\label{outlier-electorates}}
-\addcontentsline{toc}{subsubsection}{Outlier electorates}
-
-Based on the distribution of the Cook's distance values, a Cook's distance greater than \(0.1\) is considered to be influential and a potential outlier.
-The electorate of Sydney (NSW) has a large Cook's distance from 2001 to 2013, due to its diverse population (language, birthplace and religion), high number of de facto relationships, high income, high household mobility and small amount of workers in extractive and transformative jobs.
-It has remained a strong supporter of the Labor party and the Liberal vote is severely overpredicted by the model, making it an outlier.
-Nearby in metropolitan NSW, the electorate of Wentworth is found to be an outlier in all but the 2007 election.
-Although historically Liberal, its two-party vote jumped by over 10 percentage points in 2010 without experiencing any notable changes in its socio-demographic makeup --- implying that this may be the direct effect of its Liberal member, Malcolm Turnbull, becoming the leader of the Liberal party.
-Liberal support in Wentworth is underpredicted by the model in each year, and more so with Turnbull as Liberal leader.
-
-Lingiari, an electorate taking up almost all of the Northern Territory, is an outlier in the 2001--2007 elections due to its large Indigenous population, young age profile and low rates of property ownership.
-Fowler (NSW) has a diverse population with a high proportion of migrants, many Buddhists and Muslims, and has strong Labor support, making it influential in 2001, 2004 and 2010.
-Other electorates with large Cook's distance are Barton (NSW) and Leichhardt (QLD) in 2016, and Canberra (ACT) in 2007.
-
-\hypertarget{conclusion}{%
-\section{Conclusion}\label{conclusion}}
-
-This paper explores the effects of electoral socio-demographic characteristics on the two-party preferred vote in the 2001--2016 elections, using information from the corresponding Australian federal elections and Censuses.
-As a Census does not always occur in the same year as an election, Census data for the 2004--2013 elections are generated by employing a method of spatio-temporal imputation.
-This imputes electoral socio-demographics for the electoral boundaries in place at the time of the election --- an approach that is distinctly different from previous work on modelling election outcomes, where Census and election data are typically joined without addressing their temporal differences.
-Before estimating a model, these socio-demographic variables are standardized (to adjust for changing variable scales) and many variables (representing similar information) are combined into factors, resulting in a reduced predictor set.
-A spatial error model is then estimated for each election, accounting for the inherent spatial structure of the data.
-
-Across the past six elections, most of the socio-demographics that drive the electoral two-party preferred vote are found to remain steady, whilst a few (typically weaker) effects vary over time.
-Industry and type of work are particularly influential, with energy-related and manufacturing/construction jobs, as well as administrative roles being strongly linked with the Liberal party in all elections.
-Incomes have a similarly consistent effect, with higher income areas supporting Liberal.
-Higher levels of unemployment shift from weak association with Labor to a significant Liberal effect over the years, and higher education levels are associated with Labor from 2007 (although only significant in 2016).
-It is also found that electorates with higher household mobility support Liberal, birthplace diversity favours Labor and more de facto relationships align with Labor preference --- although marriages, family and household sizes have no material influence.
-Furthermore, the neighbourhood (spatial) effects are found to be positive in all elections, although only significant in 2001 and 2016, meaning that in the 2004--2013 elections, electorates effectively voted independently.
-
-The findings in this paper complement the existing literature by modelling temporal trends, which as far as the authors are aware, has not been done previously for Australian elections using a regression framework.
-It is also the first study to model any Australian election since 2010 using Census information.
-
-Additionally, a key contribution of this research is the wrangling of the raw data and imputed data sets for the 2004, 2007, 2010 and 2013 elections, which have been contributed to the \texttt{eechidna} \texttt{R} package --- providing a rich, accessible data resource for future Australian electoral analysis.
-
-\hypertarget{acknowledgements}{%
-\section{Acknowledgements}\label{acknowledgements}}
-
-This paper was produced using \texttt{RMarkdown} \citep{rmarkdown} and \texttt{knitr} \citep{knitr}.
-All corresponding code for this paper can be found in the github repository \href{https://github.com/jforbes14/eechidna-paper}{github.com/jforbes14/eechidna-paper}, and the data used is available in the \texttt{eechidna} package \citep{eechidna}.
-All raw data was obtained from the Australian Electoral Commission, the Australian Bureau of Statistics and the Australian Government.
-
-\hypertarget{software}{%
-\section{Software}\label{software}}
-
-All election and Census datasets, along with electoral maps and more, are available in the \texttt{eechidna} (Exploring Election and Census Highly Informative Data Nationally for Australia) \texttt{R} package, which can be downloaded from CRAN.
-The \texttt{eechidna} package makes it easy to look at the data from the Australian Federal elections and Censuses that occurred between 2001 and 2016.
-This study contributed a large revision to the \texttt{eechidna} package, which included the addition of election and Census data for 2001--2010, voting outcomes for polling booths and imputed Census data for election years.
-For more details on using \texttt{eechidna}, please see the articles (vignettes) on the github page \href{https://ropenscilabs.github.io/eechidna/}{ropenscilabs.github.io/eechidna/}.
-
-The authors would like to sincerely thank Anthony Ebert, Heike Hofmann, Thomas Lumley, Ben Marwick, Carson Sievert, Mingzhu Sun, Dilini Talagala, Nicholas Tierney, Nathaniel Tomasetti, Earo Wang and Fang Zhou, all of whom have contributed to the \texttt{eechidna} package.
-
-\renewcommand\refname{References}
-\bibliography{references.bib}
-
-\end{document}

### 2019-04-25 elections.tex
Rob's re-compiled files. Ready for submission.

--- 2019-04-21 draft-paper.tex
@@
             pdftitle={Spatial modelling of the two-party preferred vote in Australian federal elections: 2001--2016},
-            pdfkeywords={federal election, Census, Australia, spatial modelling, imputation, socio-demographics, electorates, R, eechidna},
+            pdfkeywords={federal election, Census, Australia, spatial modelling, imputation, data science, socio-demographics, electorates, R, eechidna},
             pdfborder={0 0 0},
@@
 \begin{keywords}
-federal election, Census, Australia, spatial modelling, imputation, socio-demographics, electorates, R, eechidna
+federal election, Census, Australia, spatial modelling, imputation, data science, socio-demographics, electorates, R, eechidna
 \end{keywords}
@@
 In contrast, \textcite{Stimson09} models the polling booth swing vote (change in the two-party preferred vote) in the 2007 election, finding that little of the swing vote can be explained by Census data.
-Instead of analyzing a single election in isolation, this paper employs a consistent model framework across six elections so that temporal changes in the effects of socio-demographics can be observed, where each federal elections is modelled with a cross-sectional dataset.
-The use of a regression framework to examine these socio-political relationships over time is seemingly absent from previous Australian studies.
-It also appears that no study has attempted any type of statistical analysis of socio-demographics in conjunction with voter behaviour in Australia since 2007, making this paper distinctly different from previous work.
 
+Instead of analyzing a single election in isolation, this paper employs a consistent model framework across six elections so that temporal changes in the effects of socio-demographics can be observed.
+Each federal election is modelled with a cross-sectional dataset.
 The cross-sectional dataset for each election used here consists of the two-party preferred vote (as the response variable), and a set of common socio-demographic variables (as the explanatory variables) that characterize each electorate.

### 2019-04-30 elections.Rmd
Changes as per Mervyn's comments

@@
 
-Electoral socio-demographics are derived from the Australian Census, and vote counts are obtained from Australian federal elections.
+Data on electoral socio-demographics are derived from the Australian Census, and vote counts are obtained from Australian federal elections.
 Joining these two data sources is problematic as there is an inherent asynchronicity in the two types of events.
@@
 Furthermore, the responses are found to be spatially correlated in each election (Moran's I test, $p \le 7\cdot10^{-15}$).
-This is not surprising as electorates are aggregate spatial units, and hence the spatial structure of the data must modelled appropriately.
+This is not surprising as electorates are aggregate spatial units, and hence the spatial structure of the data must be modelled appropriately.
 
@@
 
-Regions comprising more older people are often believed to be more conservative, and indeed it found that electorates with a higher median age are more likely to support the Liberal party --- although this effect is only significant in 2007 and 2010 (see row 2 in Figure \@ref(fig:partresplot2)).
+Regions comprising more older people are often believed to be more conservative, and indeed it found that electorates with a higher median age are more likely to support the Liberal party --- although this effect is significant only in 2007 and 2010 (see row 2 in Figure \@ref(fig:partresplot2)).
 
@@
 
-Since 2007, electorates with higher education levels are associated with supporting the Labor party, although this effect is only significant in 2016.
+Since 2007, electorates with higher education levels are associated with supporting the Labor party, although this effect is significant only in 2016.
 Before 2007, education has an almost zero effect (see row 3 in Figure \@ref(fig:partresplot1)).
@@
 Incomes have a similarly consistent effect, with higher income areas supporting Liberal.
-Higher levels of unemployment shift from weak association with Labor to a significant Liberal effect over the years, and higher education levels are associated with Labor from 2007 (although only significant in 2016).
+Higher levels of unemployment shift from weak association with Labor to a significant Liberal effect over the years, and higher education levels are associated with Labor from 2007 (although significant only in 2016).
 It is also found that electorates with higher household mobility support Liberal, birthplace diversity favours Labor and more de facto relationships align with Labor preference --- although marriages, family and household sizes have no material influence.
-Furthermore, the neighbourhood (spatial) effects are found to be positive in all elections, although only significant in 2001 and 2016, meaning that in the 2004--2013 elections, electorates effectively voted independently.
+Furthermore, the neighbourhood (spatial) effects are found to be positive in all elections, although significant only in 2001 and 2016, meaning that in the 2004--2013 elections, electorates effectively voted independently.
 

### 2019-05-01 elections.Rmd
Updated to run with latest tidyverse

@@
 
+# Change abs2016 to data frame to prevent error below
+abs2016 <- as.data.frame(abs2016)
+
 # Combine and standardize
@@
   mutate(year = factor(year)) %>%
-  dplyr::select(-c(
+  select(-c(
     starts_with("Age"), StateAb,
@@
     Other_NonChrist, OtherChrist, Catholic, Anglican,
-    Volunteer, EmuneratedElsewhere, UniqueID
+    Volunteer, EmuneratedElsewhere, DivisionID
   ))
@@
   bind_rows(
-    glsmod16$my_data %>% left_join(tpp16 %>% dplyr::select(DivisionNm, UniqueID, StateAb), by = "DivisionNm"),
-    glsmod13$my_data %>% left_join(tpp13 %>% dplyr::select(DivisionNm, UniqueID, StateAb), by = "DivisionNm"),
-    glsmod10$my_data %>% left_join(tpp10 %>% dplyr::select(DivisionNm, UniqueID, StateAb), by = "DivisionNm"),
-    glsmod07$my_data %>% left_join(tpp07 %>% dplyr::select(DivisionNm, UniqueID, StateAb), by = "DivisionNm"),
-    glsmod04$my_data %>% left_join(tpp04 %>% dplyr::select(DivisionNm, UniqueID, StateAb), by = "DivisionNm"),
-    glsmod01$my_data %>% left_join(tpp01 %>% dplyr::select(DivisionNm, UniqueID, StateAb), by = "DivisionNm")
-  )
-)
-# %>% dplyr::select(Residuals, DivisionNm, UniqueID, StateAb, year, LNP_Percent, gls_residuals)
+    glsmod16$my_data %>% left_join(tpp16 %>% dplyr::select(DivisionNm, DivisionID, StateAb), by = "DivisionNm"),
+    glsmod13$my_data %>% left_join(tpp13 %>% dplyr::select(DivisionNm, DivisionID, StateAb), by = "DivisionNm"),
+    glsmod10$my_data %>% left_join(tpp10 %>% dplyr::select(DivisionNm, DivisionID, StateAb), by = "DivisionNm"),
+    glsmod07$my_data %>% left_join(tpp07 %>% dplyr::select(DivisionNm, DivisionID, StateAb), by = "DivisionNm"),
+    glsmod04$my_data %>% left_join(tpp04 %>% dplyr::select(DivisionNm, DivisionID, StateAb), by = "DivisionNm"),
+    glsmod01$my_data %>% left_join(tpp01 %>% dplyr::select(DivisionNm, StateAb), by = "DivisionNm")
+  )
+)
+# %>% dplyr::select(Residuals, DivisionNm, DivisionID, StateAb, year, LNP_Percent, gls_residuals)
 ```
@@
   bind_rows(
-    glsmod16$my_data %>% left_join(tpp16 %>% dplyr::select(DivisionNm, UniqueID, StateAb), by = "DivisionNm"),
-    glsmod13$my_data %>% left_join(tpp13 %>% dplyr::select(DivisionNm, UniqueID, StateAb), by = "DivisionNm"),
-    glsmod10$my_data %>% left_join(tpp10 %>% dplyr::select(DivisionNm, UniqueID, StateAb), by = "DivisionNm"),
-    glsmod07$my_data %>% left_join(tpp07 %>% dplyr::select(DivisionNm, UniqueID, StateAb), by = "DivisionNm"),
-    glsmod04$my_data %>% left_join(tpp04 %>% dplyr::select(DivisionNm, UniqueID, StateAb), by = "DivisionNm"),
-    glsmod01$my_data %>% left_join(tpp01 %>% dplyr::select(DivisionNm, UniqueID, StateAb), by = "DivisionNm"))
+    glsmod16$my_data %>% left_join(tpp16 %>% dplyr::select(DivisionNm, DivisionID, StateAb), by = "DivisionNm"),
+    glsmod13$my_data %>% left_join(tpp13 %>% dplyr::select(DivisionNm, DivisionID, StateAb), by = "DivisionNm"),
+    glsmod10$my_data %>% left_join(tpp10 %>% dplyr::select(DivisionNm, DivisionID, StateAb), by = "DivisionNm"),
+    glsmod07$my_data %>% left_join(tpp07 %>% dplyr::select(DivisionNm, DivisionID, StateAb), by = "DivisionNm"),
+    glsmod04$my_data %>% left_join(tpp04 %>% dplyr::select(DivisionNm, DivisionID, StateAb), by = "DivisionNm"),
+    glsmod01$my_data %>% left_join(tpp01 %>% dplyr::select(DivisionNm, DivisionID, StateAb), by = "DivisionNm"))
 )

### 2019-05-01 elections.tex
Updated to run with latest tidyverse

@@
 
-{\centering \includegraphics[width=0.9\linewidth]{draft-paper_files/figure-latex/timeline-1}
+{\centering \includegraphics[width=0.9\linewidth]{elections_files/figure-latex/timeline-1}
 
@@
 
-{\centering \includegraphics[height=0.28\textheight]{draft-paper_files/figure-latex/hume13-1}
+{\centering \includegraphics[height=0.28\textheight]{elections_files/figure-latex/hume13-1}
 
@@
 
-{\centering \includegraphics[height=0.28\textheight]{draft-paper_files/figure-latex/hume16-1}
+{\centering \includegraphics[height=0.28\textheight]{elections_files/figure-latex/hume16-1}
 
@@
 Furthermore, the responses are found to be spatially correlated in each election (Moran's I test, \(p \le 7\cdot10^{-15}\)).
-This is not surprising as electorates are aggregate spatial units, and hence the spatial structure of the data must modelled appropriately.
+This is not surprising as electorates are aggregate spatial units, and hence the spatial structure of the data must be modelled appropriately.
 
@@
 
-{\centering \includegraphics[height=0.2\textheight,]{draft-paper_files/figure-latex/rhovis-1}
+{\centering \includegraphics[height=0.2\textheight,]{elections_files/figure-latex/rhovis-1}
 
@@
 
-{\centering \includegraphics[height=0.2\textheight,]{draft-paper_files/figure-latex/plotintercept-1}
+{\centering \includegraphics[height=0.2\textheight,]{elections_files/figure-latex/plotintercept-1}
 
@@
 
-{\centering \includegraphics{draft-paper_files/figure-latex/partresplot1-1}
+{\centering \includegraphics{elections_files/figure-latex/partresplot1-1}
 
@@
 
-{\centering \includegraphics{draft-paper_files/figure-latex/partresplot2-1}
+{\centering \includegraphics{elections_files/figure-latex/partresplot2-1}
 
@@
 
-Regions comprising more older people are often believed to be more conservative, and indeed it found that electorates with a higher median age are more likely to support the Liberal party --- although this effect is only significant in 2007 and 2010 (see row 2 in Figure \ref{fig:partresplot2}).
+Regions comprising more older people are often believed to be more conservative, and indeed it found that electorates with a higher median age are more likely to support the Liberal party --- although this effect is significant only in 2007 and 2010 (see row 2 in Figure \ref{fig:partresplot2}).
 
@@
 
-Since 2007, electorates with higher education levels are associated with supporting the Labor party, although this effect is only significant in 2016.
+Since 2007, electorates with higher education levels are associated with supporting the Labor party, although this effect is significant only in 2016.
 Before 2007, education has an almost zero effect (see row 3 in Figure \ref{fig:partresplot1}).
@@
 
-{\centering \includegraphics[width=0.7\linewidth,]{draft-paper_files/figure-latex/resstate-1}
+{\centering \includegraphics[width=0.7\linewidth,]{elections_files/figure-latex/resstate-1}
 

### 2019-05-01 functions.R

@@ -4,7 +4,8 @@ standardise_vars <- function(df) {
   num_cols <- sapply(df, class) == "numeric"
   df[, num_cols] <- lapply(df[, num_cols], scale)
   names(df) <- names(hold)
-  df$LNP_Percent <- hold$LNP_Percent
+  if("LNP_Percent" %in% names(hold))
+    df$LNP_Percent <- hold$LNP_Percent
   return(df)
 }
 

### 2019-05-02 elections_conversation.Rmd
Started conversation version of article

--- 2019-05-01 elections.Rmd
@@
 title: "Spatial modelling of the two-party preferred vote in Australian federal elections: 2001--2016"
-author:
-- familyname: Forbes
-  othernames: Jeremy
-  address: Monash University\newline Melbourne, Australia
-  email: jeremyforbes1995@gmail.com
-  correspondingauthor: true
-- familyname: Cook
-  othernames: Dianne
-  address: Monash University\newline Melbourne, Australia
-- familyname: Hyndman
-  othernames: Rob J
-  address: Monash University\newline Melbourne, Australia
-abstract: "We examine the relationships between electoral socio-demographic characteristics and two-party preference in the six Australian federal elections held between 2001 to 2016.
-Socio-demographic information is derived from the Australian Census, which occurs every five years.
-Since a Census is not directly available for each election, spatio-temporal imputation is employed to estimate Census data for the electorates at the time of each election.
-This accounts for both spatial and temporal changes in electoral characteristics between Censuses.
-To capture any spatial heterogeneity, a spatial error model is estimated for each election, which incorporates a spatially structured random effect vector that can be thought of as the unobserved political climate in each electorate.
-Over time, the impact of most socio-demographic characteristics that affect electoral two-party preference do not vary, with industry of work, incomes, household mobility and de facto relationships having strong effects in each of the six elections.
-Education and unemployment are amongst those that have varying effects.
-It is also found that between 2004 and 2013, the spatial effect was insignificant, meaning that electorates voted effectively independently.
-All data featured in this study has been contributed to the `eechidna` `R` package (available on CRAN)."
-keywords: "federal election, Census, Australia, spatial modelling, imputation, data science, socio-demographics, electorates, R, eechidna"
-wpnumber:
-jelcodes: C31, C33, D72
-blind: false
-cover: false
+author: Rob J Hyndman and Di Cook
 toc: false
@@
 output:
-  MonashEBSTemplates::workingpaper:
+  bookdown::pdf_document2:
     fig_caption: yes
@@
     includes:
-      in_header: preamble.tex
+      in_header: conversation_preamble.tex
     keep_tex: yes
@@
 The relationship is relatively robust over time, with the exception of Unemployment.", fig.height = 8.5, fig.width = 6, warning = FALSE, fig.pos = 'h'}
-(p1 / p2 / (p3 + ylab("Two-party preferred vote (%)                     ")) / p4 / p5 / p6)
+#(p1 / p2 / (p3 + ylab("Two-party preferred vote (%)                     ")) / p4 / p5 / p6)
 ```
@@
 Most relationships are relatively stable over elections, except OtherLanguage and Education.", fig.height = 8.5,  fig.width = 6, warning = FALSE, fig.pos = 'h'}
-(p7 / p8 / (p9 + ylab("Two-party preferred vote (%)                     ")) / p10 / p11 / p12 / p13)
+#(p7 / p8 / (p9 + ylab("Two-party preferred vote (%)                     ")) / p10 / p11 / p12 / p13)
 ```

### 2019-05-02 elections_conversation.Rmd
Made a start on conversation article

@@
 ---
-title: "Spatial modelling of the two-party preferred vote in Australian federal elections: 2001--2016"
+title: "How who we are affects how we vote"
 author: Rob J Hyndman and Di Cook
@@
 
+We often hear stereotypes about voting patterns --- people in their 20s without kids are more likely to be left wing, migrants are more conservative, wealthier people tend to favour the conservative parties, and so on.
+So we thought we would test these ideas by matching census data to election data, to see if we can identify what are the socio-demographic characteristics of an electorate that are most closely related to how they vote.
+
+# Previous intro
+
 Australia has changed in many ways over the last two decades.

### 2019-05-02 elections_conversation.Rmd
Added discussion of data

@@
 ---
-title: "How who we are affects how we vote"
+title: "Who we are affects how we vote"
 author: Rob J Hyndman and Di Cook
@@
 biblio-style: authoryear-comp
-header-includes: \usepackage{float}
+fontsize: 12pt
+papersize: a4
 output:
   bookdown::pdf_document2:
+    latex_engine: xelatex
     fig_caption: yes
@@
 ---
-
-# Introduction {#intro}
 
@@
 
+# Introduction {#intro}
+
 We often hear stereotypes about voting patterns --- people in their 20s without kids are more likely to be left wing, migrants are more conservative, wealthier people tend to favour the conservative parties, and so on.
@@
 
-# Previous intro
+It is also interesing to see how this might have changed over time.
+For example, if wealth was a good predictor of voting patterns in the 2001 election, is it still a good predictor in 2019?
+Australia has changed in many ways over the last two decades.
+Rising house prices, country-wide improvements in education, an ageing population, and a decline in religious affiliation, are just some of the ways we have changed.
+At the same time, political power has moved back and forth between the two major parties.
+How much can we attribute changes in political power to changes in who we are?
 
-Australia has changed in many ways over the last two decades.
-Rising house prices, country-wide improvements in education, an ageing population, and a decline in religious affiliation, are just a few facets of the country's evolving socio-demographic characteristics.
-At the same time, political power has moved back and forth between the two major parties.
-In the 2007 and 2010 federal elections, the Australian Labor Party (Labor) was victorious, whereas the 2001, 2004, 2013 and 2016 elections were won by the Liberal National coalition (Liberal).
-The two-party preferred vote, a measure of support between these two parties, fluctuated between 47.3% and 53.5% (in favour of the Liberal party) over this period.
-This study explores how electoral socio-demographic characteristics relate to two-party preference, and whether their effects have changed over time.
+# Census and electoral data
 
-Data on electoral socio-demographics are derived from the Australian Census, and vote counts are obtained from Australian federal elections.
-Joining these two data sources is problematic as there is an inherent asynchronicity in the two types of events.
-A Census is conducted by the Australian Bureau of Statistics (ABS) every five years, whereas federal elections, conducted by the Australian Electoral Commission (AEC), usually occur every three years or so.
-The first problem addressed is that of constructing appropriate Census data for the 2004, 2007, 2010 and 2013 elections --- election years in which a Census does not occur.
-The predominant approach in previous studies was to join voting outcomes to the nearest Census, without accounting for any temporal differences [see @DavisStimson98;@Stimson06;@Liao09;@Stimson09].
-Furthermore, electoral boundaries change regularly, so spatial discrepancies also arise when matching with electoral data.
-To obtain appropriate "Census-like" data for these four elections, electoral socio-demographics are constructed using a spatio-temporal imputation that combines areal interpolation [@Goodchild1993] and linear time-interpolation.
-Collecting and wrangling the raw data, along with the imputation process, are detailed in Section \@ref(data).
-All data and associated documentation relating to this procedure are available in the `eechidna` `R` package [@eechidna], providing a resource for future analysis.
-
-Previous work on modelling Australian federal elections has found that aggregate socio-demographics are relatively good predictors of voting outcomes.
-@Forrest01 used multiple regression to model the Liberal and Labor primary vote for polling booths in the Farrer electorate in 1998 as a function of Census variables from 1996.
-@Stimson06, @Stimson09 and @Stimson12 used principal component analysis of polling booths in the 2001, 2004 and 2007 elections respectively, also finding that socio-demographic characteristics of polling booths are linked to their two-party preferred vote.
-In contrast, @Stimson09 models the polling booth swing vote (change in the two-party preferred vote) in the 2007 election, finding that little of the swing vote can be explained by Census data.
-
-Instead of analyzing a single election in isolation, this paper employs a consistent model framework across six elections so that temporal changes in the effects of socio-demographics can be observed.
-Each federal election is modelled with a cross-sectional dataset.
-<!--The use of a regression framework to examine these socio-political relationships over time is seemingly absent from previous Australian studies.
-It also appears that no study has attempted any type of statistical analysis of socio-demographics in conjunction with voter behaviour in Australia since 2007, making this paper distinctly different from previous work.-->
-The cross-sectional dataset for each election used here consists of the two-party preferred vote (as the response variable), and a set of common socio-demographic variables (as the explanatory variables) that characterize each electorate.
-To prepare these datasets, socio-demographic variables are first standardized, and then a principal component analysis is used to group variables into "factors".
-To account for the inherent spatial structure of the data, a spatial error model is then estimated for each election.
-
-The paper is organised as follows.
-Section \@ref(data) describes the data collection, joining and cleaning.
-These pre-processing steps and model details are discussed in Section \@ref(modelling).
-Section \@ref(results) describes the inference conducted to determine signficance of effects and how these change over time.
-Section \@ref(conclusion) summarises the work.
-Two supplementary sections document the contributions of others to this work and the software.
-
-# Data collection, wrangling and imputation {#data}
-
-## Collecting the data
-
-The voting outcome of interest is the electoral two-party preferred vote, which is provided by the Australian Electoral Commission (AEC) for the 2001, 2004, 2007, 2010, 2013 and 2016 elections via the AEC Tally Room.
-The AEC divides Australia into 150 regions, called electorates, with each corresponding to a single seat in the House of Representatives.
-Voting is compulsory in Australia, and each voter assigns a numbered preference to each available candidate in their electorate.
-The two-party preferred vote is determined by a tally of these preferences where, by convention, only the ranks of the Labor and Liberal candidates are considered.
-This is recorded as a percentage preference in favour of the Liberal party.
-
-Socio-demographic variables were derived from the Australian Census of Population and Housing (Census), which is a survey of every household in Australia, recording information such as age, gender, ethnicity, education level and income.
-There have been four Censuses so far in the 21st century, conducted in 2001, 2006, 2011 and 2016.
-The Australian Bureau of Statistics (ABS) conducts the Census and publishes aggregated information.
-The ABS uses electoral boundaries as defined by the AEC at the time of each Census, which may not match those in place at the subsequent and previous elections.
-From the available Census information aggregated at the electorate level, 65 socio-demographic variables were defined for each of the electorates to be used in the analysis.
-
-Raw data was sourced online from the AEC and ABS websites in `.csv` and `.xlsx` files.
-The formats of these files differ over the years, making extracting the appropriate information a big task.
-The functions available in the `dplyr` [@dplyr] and `readxl` [@readxl] `R` packages are particularly useful, as they provide fast consistent tools for data manipulation and functions to import `.xlsx` files (respectively).
-The 2001 and 2006 Census data are published in a format where the information for each electorate is held in a separate document making it difficult to use the `dplyr` tools.
-Instead, cells have to be selected from each individual file to construct the desired variables.
-All scripts required for the data wrangling process can be found in the github repository for the `eechidna` `R` package [@eechidna], along with the raw data.
-The `eechidna` package makes this study entirely reproducible and provides a resource to help wrangle data for future Censuses and elections, when they become available.
-
-## Joining Census and election data
-
-### Differences between Census and election data
-
-Between 2001 and 2016 there were six elections and four Censuses (see Figure \@ref(fig:timeline)).
-Electoral boundaries are redistributed regularly by the AEC, meaning that only in the years where both a Census and election occur are all boundaries likely to match --- the case for the 2001 and 2016 elections.
-Therefore, for the four elections between 2004 and 2013, both temporal and spatial differences in electorates need to be accounted for when joining the electoral two-party preferred vote with Census data.
-For these elections a spatio-temporal imputation method was employed to obtain electoral socio-demographics.
-This method uses Census information from both before and after the election of interest.
+The Census provides data on electoral socio-demographics, and vote counts in each electorate can be obtained from Australian federal elections.
+However, joining these two data sources is difficult because the Censuses are not held at the same time as the elections.
+Between 2001 and 2016 there were six elections and four Censuses, as shown in the timeline below.
 
@@
 
-### Spatio-temporal imputation
+Not only can an electorate change between the last Census and an election, but even the electorate boundaries can change.
+Some electorates can disappear altogether and new electorates can arise.
+Electoral boundaries are redistributed regularly by the AEC, meaning that only in the years where both a Census and election occur are all boundaries likely to match --- the case for the 2001 and 2016 elections.
+So we first had to estimate what the socio-demographic characteristics of an electorate would have been at the time of each election using a complicated method of interpolation over time and geography.
+This method uses Census information from both before and after the election of interest, and information from neighbouring electorates when boundaries have changed.
 
-For each election, neighbouring Census information has to be combined in some way so that it represents the boundaries in place at the time of the election.
-This is done by taking the electoral boundaries and imputing the corresponding socio-demographic characteristics for each of neighbouring Censuses, thereby addressing the spatial aspect.
-Next, to deal with the temporal component, characteristics at the time of the election are constructed using linear interpolation between the spatially imputed neighbouring Census variables.
+A simple way to measure voting patterns is to consider the two-party preferred (2PP) vote, which is based on the tally of preferences for the Labour and Liberal candidates, ignoring all other candidates.
+By convention, this is recorded as a percentage preference in favour of the Liberal party -- so a 2PP value of 45% indicates that 45% of voters preferred a Liberal candidate and 55% of voters preferred a Labour candidate.
 
-To account for spatial differences, the piecewise approximation method in @Goodchild1993 is adopted.
-Consider a map of source zones $s = 1,\dots,S$, for which socio-demographic information is available, and a set of target zones $t = 1,\dots,T$ for which information is to be imputed.
-This is described in the context of a single election, and a single neighbouring Census.
-
-Let the map of electoral boundaries at the time of a Census define the source zones, and let the boundaries at the time of the election be the target zones.
-Denote the area of intersection between a source zone $s$ and a target zone $t$ as $A_{s,t}$.
-Additionally, let the population of the source zone $s$ be $U_s$ and the population of intersection between source zone $s$ and target zone $t$ be $P_{s,t}$.
-The estimated population of intersection is given by
-$$
-  \hat{P}_{s,t} = \frac{U_s*A_{s,t}}{\sum_{t=1}^T A_{s,t}},
-  \quad\text{for all $s=1,\dots,S$ and $t=1,\dots,T$}.
-$$
-Note that this estimator implicitly assumes that populations are uniformly distributed within each source zone.
-
-In order to calculate socio-demographic information for each of the target zones, a weighted average is taken using the estimated intersection populations as weights.
-Denote a given Census variable for the target zone by $C_t$, and the same Census variable for the source zone as $D_s$.
-Then, estimate $C_t$ using
-$$
-  \hat{C}_t = \frac{\sum_{s=1}^{S}{D_s*\hat{P}_{s,t}}}%
-  {\sum_{s=1}^{S}{\hat{P}_{s,t}}},
-  \quad\text{for each $t=1,\dots,T$}.
-$$
-This concludes the spatial imputation of the socio-demographic characteristics for one target zone (a single electoral boundary), at the time of only one of the neighbouring Censuses.
-This process is repeated for all of the target zones, and then for the other neighbouring Census.
-
-\enlargethispage*{0.5cm}
-
-To account for temporal changes, linear interpolation is used between Census years to get the final estimate of a Census variable for the target zone in the election year.
-Let $y_1$ be the year of the Census preceding an election, let $y_2$ be the year of the election, and $y_3$ be the year of the Census that follows.
-Add this year subscript to the Census variable estimate $\hat{C}_t$, resulting in $\hat{C}_{t,y}$.
-Linear interpolating between these Census years results an imputed value for the election year, given by
-$$
-  \hat{C}_{t,y_2} = \frac{y_3-y_2}{y_3-y_1} * \hat{C}_{t,y_1} +
-    \frac{y_2-y_1}{y_3-y_1} * \hat{C}_{t,y_3}.
-$$
-Implicitly this assumes that population characteristics change in a linear manner over time.
-
-### An illustration of the spatio-temporal imputation {-}
-
-Census data is publicly available at different levels of aggregation, ranging from SA1 (over 50,000 zones) to electoral divisions (150 zones).
-For this study, electoral divisions are used as source zones, and the imputation method is applied to produce socio-demographic variables for each of the 2004, 2007, 2010 and 2013 elections.
-As mentioned earlier, there is no need to impute socio-demographic variables for the 2001 and 2016 elections.
-To illustrate the method, consider the imputation of socio-demographic variables for the electorate of Hume in New South Wales (NSW) at the time of the 2013 federal election.
-The boundaries shown in Figure \@ref(fig:hume13) define all target zones in NSW for 2013, with the target zone of interest (Hume) shaded purple.
-
-```{r hume13, fig.cap = "Some of the electoral boundaries in NSW for 2013, with the electoral boundary for Hume shown in purple.", out.height="28%"}
-hume_area13 <- nat_map13 %>%
-  filter(state %in% c("ACT", "NSW"), long < 154)
-
-ggplot(data = hume_area13) +
-  geom_polygon(aes(x = long, y = lat, group = group, fill = elect_div == "HUME"),
-    colour = "grey50", alpha = 0.4
-  ) +
-  scale_fill_manual(
-    name = "Electoral boundaries",
-    values = c("white", "purple"),
-    labels = c("Other 2013 Electorates", "Hume 2013")
-  ) +
-  theme_map() +
-  coord_equal()
-```
-
-```{r hume16, fig.cap = "Census division boundaries in NSW for 2016, with the 2013 electoral boundary for Hume, shown in purple.
-The purple region is not contained within a single Census division.", out.height="28%"}
-hume_area16 <- nat_map16 %>%
-  filter(state %in% c("ACT", "NSW"), long < 154) %>%
-  mutate(year = "2016") %>%
-  bind_rows(
-    hume_area13 %>%
-      filter(elect_div == "HUME") %>%
-      mutate(year = "2013")
-  )
-
-ggplot(data = hume_area16) +
-  geom_polygon(aes(
-    x = long, y = lat, group = group, fill = year == "2013",
-    alpha = year == "2013", colour = year == "2013"
-  )) +
-  scale_fill_manual(
-    name = "Electoral boundaries",
-    values = c("white", "purple"),
-    labels = c("2016 Electorates", "Hume 2013")
-  ) +
-  scale_alpha_manual(values = c(0, 0.4)) +
-  scale_color_manual(values = c("grey50", NA)) +
-  theme_map() +
-  coord_equal() +
-  guides(alpha = FALSE, col = FALSE,
-    fill = guide_legend(override.aes = list(col = c("grey50", NA), alpha = 0.4)))
-```
-
-The corresponding source zones from the 2016 Census are shown in Figure \@ref(fig:hume16).
-As can be seen, the Hume boundary from the 2013 election (shaded purple) does not perfectly match any of the source zones.
-
-There are many source zones from the 2016 Census that intersect with this purple region, including the divisions of Riverina, Eden-Monaro and Hume, along with smaller intersecting areas with Fenner, Calare, Gilmore and Whitlam.
-The proportion of each source zone that overlaps with the purple region is calculated, and used to obtain the intersecting populations $\hat{P}_{s,t}$.
-
-```{r hume1}
-hume <- abs2016 %>%
-  filter(DivisionNm %in%
-           c("HUME", "RIVERINA", "EDEN-MONARO", "CANBERRA",
-             "FENNER", "WHITLAM", "GILMORE", "CALARE")) %>%
-  mutate(
-    Pst = c(21, 548, 16358, 474, 86, 145427, 39117, 92),
-    Proportion = Pst/Population,
-    DivisionNm = tools::toTitleCase(tolower(DivisionNm))
-  ) %>%
-  select(DivisionNm, Proportion, Population, Pst, AusCitizen) %>%
-  arrange(desc(Proportion))
-
-hume %>%
-  kable(longtable=FALSE, booktabs=TRUE, digits=4,
-        format='latex',escape=FALSE,
-        col.names=c(
-          "Source zone (2016)",
-          "Proportion",
-          "Source zone population",
-          "Population allocated to purple region: $\\hat{P}_{s,t}$",
-          "AusCitizen (\\%): $D_s$"),
-    caption = "Population from each intersecting Census source zone (2016) that is allocated to the target zone (purple region - Hume electoral boundary in 2013), and the corresponding proportion of Australian citizens in each of these source zones.") %>%
-  kableExtra::kable_styling(font_size=10, latex_options='scale_down')
-```
-
-```{r wave, include=FALSE}
-wave2016 <- weighted.mean(x = hume$AusCitizen, w = hume$Pst)
-hume11 <- abs2011 %>%
-  filter(DivisionNm %in%
-           c("HUME", "RIVERINA", "EDEN-MONARO", "CANBERRA",
-             "FENNER", "WHITLAM", "GILMORE", "CALARE")) %>%
-  mutate(
-    Pst = c(21, 548, 16358, 474, 86, 145427), # Needs adjusting
-    Proportion = Pst/Population,
-    DivisionNm = tools::toTitleCase(tolower(DivisionNm))
-  ) %>%
-  select(DivisionNm, Proportion, Population, Pst, AusCitizen) %>%
-  arrange(desc(Proportion))
-wave2011 <- 91.00 #weighted.mean(x = hume11$AusCitizen, w = hume11$Pst)
-```
-
-Now consider the socio-demographic variable *AusCitizen*, the proportion of people in the region who are Australian citizens.
-A weighted average of *AusCitizen*, with the allocated population from each source zone as weights, yields $\hat{C}_{\text{Hume},2016} = `r sprintf("%4.2f",wave2016)`\%$.
-Repeating this process using the 2011 Census yields $\hat{C}_{\text{Hume},2011} = `r sprintf("%4.2f",wave2011)`\%$.
-Finally, linear interpolation between 2011 and 2016 yields the 2013 estimate:
-$$
-  \hat{C}_{\text{Hume},2013}
-    = \frac{3}{5} \hat{C}_{\text{Hume},2011} + \frac{2}{5} \hat{C}_{\text{Hume},2016}
-    % = \frac{3}{5} \cdot 91.00 \% + \frac{2}{5} \cdot 89.65 \% \\
-    =  90.46\%.
-$$
-This is done for each of the 65 socio-demographic variables, and is repeated for each of the 149 remaining target zones corresponding with 2013 electorates.
+We consider how various socio-demographic variables obtained from Census data can be used to explain the 2PP values for each of the 150 electorates in each of the federal elections between 2001 and 2016.
 

### 2019-05-02 elections_conversation.tex
Added discussion of data

--- 2019-05-01 elections.tex
@@
-\documentclass[11pt,a4paper,]{article}
+\documentclass[12pt,a4paper,]{article}
 \usepackage{lmodern}
-
 \usepackage{amssymb,amsmath}
@@
 \IfFileExists{microtype.sty}{%
-\usepackage[]{microtype}
+\usepackage{microtype}
 \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
 }{}
-\PassOptionsToPackage{hyphens}{url} % url is loaded by hyperref
-\usepackage[unicode=true]{hyperref}
-\hypersetup{
-            pdftitle={Spatial modelling of the two-party preferred vote in Australian federal elections: 2001--2016},
-            pdfkeywords={federal election, Census, Australia, spatial modelling, imputation, data science, socio-demographics, electorates, R, eechidna},
+\usepackage[margin=1in]{geometry}
+\usepackage{hyperref}
+\hypersetup{unicode=true,
+            pdftitle={Who we are affects how we vote},
+            pdfauthor={Rob J Hyndman and Di Cook},
             pdfborder={0 0 0},
@@
 \urlstyle{same}  % don't use monospace font for urls
-\usepackage{geometry}
-\geometry{left=2.5cm,right=2.5cm,top=2.5cm,bottom=2.5cm}
-\usepackage[style=authoryear-comp,]{biblatex}
+\usepackage[style=authoryear-comp]{biblatex}
+
 \addbibresource{references.bib}
 \usepackage{longtable,booktabs}
-% Fix footnotes in tables (requires footnote package)
-\IfFileExists{footnote.sty}{\usepackage{footnote}\makesavenoteenv{long table}}{}
+\usepackage{graphicx,grffile}
+\makeatletter
+\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
+\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
+\makeatother
+% Scale images if necessary, so that they will not overflow the page
+% margins by default, and it is still possible to overwrite the defaults
+% using explicit options in \includegraphics[width, height, ...]{}
+\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
 \IfFileExists{parskip.sty}{%
@@
 \setcounter{secnumdepth}{5}
+% Redefines (sub)paragraphs to behave more like sections
+\ifx\paragraph\undefined\else
+\let\oldparagraph\paragraph
+\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
+\fi
+\ifx\subparagraph\undefined\else
+\let\oldsubparagraph\subparagraph
+\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
+\fi
 
-% set default figure placement to htbp
-\makeatletter
-\def\fps@figure{htbp}
-\makeatother
+%%% Use protect on footnotes to avoid problems with footnotes in titles
+\let\rmarkdownfootnote\footnote%
+\def\footnote{\protect\rmarkdownfootnote}
 
-\title{Spatial modelling of the two-party preferred vote in Australian federal elections: 2001--2016}
+%%% Change title format to be more compact
+\usepackage{titling}
 
-%% MONASH STUFF
+% Create subtitle command for use in maketitle
+\providecommand{\subtitle}[1]{
+  \posttitle{
+    \begin{center}\large#1\end{center}
+    }
+}
+
+\setlength{\droptitle}{-2em}
+
+  \title{Who we are affects how we vote}
+    \pretitle{\vspace{\droptitle}\centering\huge}
+  \posttitle{\par}
+    \author{Rob J Hyndman and Di Cook}
+    \preauthor{\centering\large\emph}
+  \postauthor{\par}
+    \date{}
+    \predate{}\postdate{}
+
+%% Any special functions or other packages can be loaded here.
+
+\usepackage[no-weekday]{eukdate}
+\usepackage{sourcecodepro,sourceserifpro}
+\usepackage{float,bm}
+\usepackage{setspace}
+\setstretch{1.2}
 
@@
 
-%% FONT
-\RequirePackage{bera}
-\RequirePackage{mathpazo}
+\let\origfigure\figure
+\let\endorigfigure\endfigure
+\renewenvironment{figure}[1][2] {
+    \expandafter\origfigure\expandafter[htbp]
+} {
+    \endorigfigure
+}
 
-%% HEADERS AND FOOTERS
-\RequirePackage{fancyhdr}
-\pagestyle{fancy}
-\rfoot{\Large\sffamily\raisebox{-0.1cm}{\textbf{\thepage}}}
-\makeatletter
-\lhead{\textsf{\expandafter{\@title}}}
-\makeatother
-\rhead{}
-\cfoot{}
-\setlength{\headheight}{15pt}
-\renewcommand{\headrulewidth}{0.4pt}
-\renewcommand{\footrulewidth}{0.4pt}
-\fancypagestyle{plain}{%
-\fancyhf{} % clear all header and footer fields
-\fancyfoot[C]{\sffamily\thepage} % except the center
-\renewcommand{\headrulewidth}{0pt}
-\renewcommand{\footrulewidth}{0pt}}
-
-%% MATHS
-\RequirePackage{bm,amsmath}
-\allowdisplaybreaks
+\let\origtable\table
+\let\endorigtable\endtable
+\renewenvironment{table}[1][2] {
+    \expandafter\origtable\expandafter[htbp]
+} {
+    \endorigtable
+}
 
@@
 \renewcommand{\floatpagefraction}{0.8}
-
-%\RequirePackage[section]{placeins}
-
-%% SECTION TITLES
-\RequirePackage[compact,sf,bf]{titlesec}
-\titleformat{\section}[block]
-  {\fontsize{15}{17}\bfseries\sffamily}
-  {\thesection}
-  {0.4em}{}
-\titleformat{\subsection}[block]
-  {\fontsize{12}{14}\bfseries\sffamily}
-  {\thesubsection}
-  {0.4em}{}
-\titlespacing{\section}{0pt}{*5}{*1}
-\titlespacing{\subsection}{0pt}{*2}{*0.2}
-
-%% TITLE PAGE
-\def\Date{\number\day}
-\def\Month{\ifcase\month\or
- January\or February\or March\or April\or May\or June\or
- July\or August\or September\or October\or November\or December\fi}
-\def\Year{\number\year}
-
-\makeatletter
-\def\wp#1{\gdef\@wp{#1}}\def\@wp{??/??}
-\def\jel#1{\gdef\@jel{#1}}\def\@jel{??}
-\def\showjel{{\large\textsf{\textbf{JEL classification:}}~\@jel}}
-\def\nojel{\def\showjel{}}
-\def\addresses#1{\gdef\@addresses{#1}}\def\@addresses{??}
-\def\cover{{\sffamily\setcounter{page}{0}
-        \thispagestyle{empty}
-        \placefig{2}{1.5}{width=5cm}{monash2}
-        \placefig{16.9}{1.5}{width=2.1cm}{MBusSchool}
-        \begin{textblock}{4}(16.9,4)ISSN 1440-771X\end{textblock}
-        \begin{textblock}{7}(12.7,27.9)\hfill
-        \includegraphics[height=0.7cm]{AACSB}~~~
-        \includegraphics[height=0.7cm]{EQUIS}~~~
-        \includegraphics[height=0.7cm]{AMBA}
-        \end{textblock}
-        \vspace*{2cm}
-        \begin{center}\Large
-        Department of Econometrics and Business Statistics\\[.5cm]
-        \footnotesize http://monash.edu/business/ebs/research/publications
-        \end{center}\vspace{2cm}
-        \begin{center}
-        \fbox{\parbox{14cm}{\begin{onehalfspace}\centering\Huge\vspace*{0.3cm}
-                \textsf{\textbf{\expandafter{\@title}}}\vspace{1cm}\par
-                \LARGE\@author\end{onehalfspace}
-        }}
-        \end{center}
-        \vfill
-                \begin{center}\Large
-                \Month~\Year\\[1cm]
-                Working Paper \@wp
-        \end{center}\vspace*{2cm}}}
-\def\pageone{{\sffamily\setstretch{1}%
-        \thispagestyle{empty}%
-        \vbox to \textheight{%
-        \raggedright\baselineskip=1.2cm
-     {\fontsize{24.88}{30}\sffamily\textbf{\expandafter{\@title}}}
-        \vspace{2cm}\par
-        \hspace{1cm}\parbox{14cm}{\sffamily\large\@addresses}\vspace{1cm}\vfill
-        \hspace{1cm}{\large\Date~\Month~\Year}\\[1cm]
-        \hspace{1cm}\showjel\vss}}}
-\def\blindtitle{{\sffamily
-     \thispagestyle{plain}\raggedright\baselineskip=1.2cm
-     {\fontsize{24.88}{30}\sffamily\textbf{\expandafter{\@title}}}\vspace{1cm}\par
-        }}
-\def\titlepage{{\cover\newpage\pageone\newpage\blindtitle}}
-
-\def\blind{\def\titlepage{{\blindtitle}}\let\maketitle\blindtitle}
-\def\titlepageonly{\def\titlepage{{\pageone\end{document}}}}
-\def\nocover{\def\titlepage{{\pageone\newpage\blindtitle}}\let\maketitle\titlepage}
-\let\maketitle\titlepage
-\makeatother
-
-%% SPACING
-\RequirePackage{setspace}
-\spacing{1.5}
-
-%% LINE AND PAGE BREAKING
-\sloppy
-\clubpenalty = 10000
-\widowpenalty = 10000
-\brokenpenalty = 10000
-\RequirePackage{microtype}
-
-%% PARAGRAPH BREAKS
-\setlength{\parskip}{1.4ex}
-\setlength{\parindent}{0em}
-
-%% HYPERLINKS
-\RequirePackage{xcolor} % Needed for links
-\definecolor{darkblue}{rgb}{0,0,.6}
-\RequirePackage{url}
-
-\makeatletter
-\@ifpackageloaded{hyperref}{}{\RequirePackage{hyperref}}
-\makeatother
-\hypersetup{
-     citecolor=0 0 0,
-     breaklinks=true,
-     bookmarksopen=true,
-     bookmarksnumbered=true,
-     linkcolor=darkblue,
-     urlcolor=blue,
-     citecolor=darkblue,
-     colorlinks=true}
-
-%% KEYWORDS
-\newenvironment{keywords}{\par\vspace{0.5cm}\noindent{\sffamily\textbf{Keywords:}}}{\vspace{0.25cm}\par\hrule\vspace{0.5cm}\par}
-
-%% ABSTRACT
-\renewenvironment{abstract}{\begin{minipage}{\textwidth}\parskip=1.4ex\noindent
-\hrule\vspace{0.1cm}\par{\sffamily\textbf{\abstractname}}\newline}
-  {\end{minipage}}
-
-\usepackage[T1]{fontenc}
-\usepackage[utf8]{inputenc}
-
-\usepackage[showonlyrefs]{mathtools}
-\usepackage[no-weekday]{eukdate}
 
@@
 
-\jel{C31, C33, D72}
-
-\RequirePackage[absolute,overlay]{textpos}
-\setlength{\TPHorizModule}{1cm}
-\setlength{\TPVertModule}{1cm}
-\def\placefig#1#2#3#4{\begin{textblock}{.1}(#1,#2)\rlap{\includegraphics[#3]{#4}}\end{textblock}}
-
-\nocover
-
-\author{Jeremy~Forbes, Dianne~Cook, Rob J~Hyndman}
-\addresses{\textbf{Jeremy Forbes}\newline
-Monash University\newline Melbourne, Australia
-\newline{Email: \href{mailto:jeremyforbes1995@gmail.com}{\nolinkurl{jeremyforbes1995@gmail.com}}}\newline Corresponding author\\[1cm]
-\textbf{Dianne Cook}\newline
-Monash University\newline Melbourne, Australia
-\\[1cm]
-\textbf{Rob J Hyndman}\newline
-Monash University\newline Melbourne, Australia
-\\[1cm]
-}
-
-\date{\sf\Date~\Month~\Year}
-\makeatletter
- \lfoot{\sf Forbes, Cook, Hyndman: \@date}
-\makeatother
-
-%% Any special functions or other packages can be loaded here.
-\usepackage{float}
-\let\origfigure\figure
-\let\endorigfigure\endfigure
-\renewenvironment{figure}[1][2] {
-    \expandafter\origfigure\expandafter[htbp]
-} {
-    \endorigfigure
-}
-
-\let\origtable\table
-\let\endorigtable\endtable
-\renewenvironment{table}[1][2] {
-    \expandafter\origtable\expandafter[htbp]
-} {
-    \endorigtable
-}
-
-% Don't number subsubsections
-
-\titleformat{\subsubsection}
-  {\normalfont\bfseries\sffamily}   % The style of the section title
-  {}                             % a prefix
-  {0pt}                          % How much space exists between the prefix and the title
-  {}    % How the section is represented
-
 \begin{document}
 \maketitle
-\begin{abstract}
-We examine the relationships between electoral socio-demographic characteristics and two-party preference in the six Australian federal elections held between 2001 to 2016.
-Socio-demographic information is derived from the Australian Census, which occurs every five years.
-Since a Census is not directly available for each election, spatio-temporal imputation is employed to estimate Census data for the electorates at the time of each election.
-This accounts for both spatial and temporal changes in electoral characteristics between Censuses.
-To capture any spatial heterogeneity, a spatial error model is estimated for each election, which incorporates a spatially structured random effect vector that can be thought of as the unobserved political climate in each electorate.
-Over time, the impact of most socio-demographic characteristics that affect electoral two-party preference do not vary, with industry of work, incomes, household mobility and de facto relationships having strong effects in each of the six elections.
-Education and unemployment are amongst those that have varying effects.
-It is also found that between 2004 and 2013, the spatial effect was insignificant, meaning that electorates voted effectively independently.
-All data featured in this study has been contributed to the \texttt{eechidna} \texttt{R} package (available on CRAN).
-\end{abstract}
-\begin{keywords}
-federal election, Census, Australia, spatial modelling, imputation, data science, socio-demographics, electorates, R, eechidna
-\end{keywords}
 
@@
 
+We often hear stereotypes about voting patterns --- people in their 20s without kids are more likely to be left wing, migrants are more conservative, wealthier people tend to favour the conservative parties, and so on.
+So we thought we would test these ideas by matching census data to election data, to see if we can identify what are the socio-demographic characteristics of an electorate that are most closely related to how they vote.
+
+It is also interesing to see how this might have changed over time.
+For example, if wealth was a good predictor of voting patterns in the 2001 election, is it still a good predictor in 2019?
 Australia has changed in many ways over the last two decades.
-Rising house prices, country-wide improvements in education, an ageing population, and a decline in religious affiliation, are just a few facets of the country's evolving socio-demographic characteristics.
+Rising house prices, country-wide improvements in education, an ageing population, and a decline in religious affiliation, are just some of the ways we have changed.
 At the same time, political power has moved back and forth between the two major parties.
-In the 2007 and 2010 federal elections, the Australian Labor Party (Labor) was victorious, whereas the 2001, 2004, 2013 and 2016 elections were won by the Liberal National coalition (Liberal).
-The two-party preferred vote, a measure of support between these two parties, fluctuated between 47.3\% and 53.5\% (in favour of the Liberal party) over this period.
-This study explores how electoral socio-demographic characteristics relate to two-party preference, and whether their effects have changed over time.
+How much can we attribute changes in political power to changes in who we are?
 
-Data on electoral socio-demographics are derived from the Australian Census, and vote counts are obtained from Australian federal elections.
-Joining these two data sources is problematic as there is an inherent asynchronicity in the two types of events.
-A Census is conducted by the Australian Bureau of Statistics (ABS) every five years, whereas federal elections, conducted by the Australian Electoral Commission (AEC), usually occur every three years or so.
-The first problem addressed is that of constructing appropriate Census data for the 2004, 2007, 2010 and 2013 elections --- election years in which a Census does not occur.
-The predominant approach in previous studies was to join voting outcomes to the nearest Census, without accounting for any temporal differences \autocites[see][]{DavisStimson98}{Stimson06}{Liao09}{Stimson09}.
-Furthermore, electoral boundaries change regularly, so spatial discrepancies also arise when matching with electoral data.
-To obtain appropriate ``Census-like'' data for these four elections, electoral socio-demographics are constructed using a spatio-temporal imputation that combines areal interpolation \autocite{Goodchild1993} and linear time-interpolation.
-Collecting and wrangling the raw data, along with the imputation process, are detailed in Section \ref{data}.
-All data and associated documentation relating to this procedure are available in the \texttt{eechidna} \texttt{R} package \autocite{eechidna}, providing a resource for future analysis.
+\hypertarget{census-and-electoral-data}{%
+\section{Census and electoral data}\label{census-and-electoral-data}}
 
-Previous work on modelling Australian federal elections has found that aggregate socio-demographics are relatively good predictors of voting outcomes.
-\textcite{Forrest01} used multiple regression to model the Liberal and Labor primary vote for polling booths in the Farrer electorate in 1998 as a function of Census variables from 1996.
-\textcite{Stimson06}, \textcite{Stimson09} and \textcite{Stimson12} used principal component analysis of polling booths in the 2001, 2004 and 2007 elections respectively, also finding that socio-demographic characteristics of polling booths are linked to their two-party preferred vote.
-In contrast, \textcite{Stimson09} models the polling booth swing vote (change in the two-party preferred vote) in the 2007 election, finding that little of the swing vote can be explained by Census data.
-
-Instead of analyzing a single election in isolation, this paper employs a consistent model framework across six elections so that temporal changes in the effects of socio-demographics can be observed.
-Each federal election is modelled with a cross-sectional dataset.
-The cross-sectional dataset for each election used here consists of the two-party preferred vote (as the response variable), and a set of common socio-demographic variables (as the explanatory variables) that characterize each electorate.
-To prepare these datasets, socio-demographic variables are first standardized, and then a principal component analysis is used to group variables into ``factors''.
-To account for the inherent spatial structure of the data, a spatial error model is then estimated for each election.
-
-The paper is organised as follows.
-Section \ref{data} describes the data collection, joining and cleaning.
-These pre-processing steps and model details are discussed in Section \ref{modelling}.
-Section \ref{results} describes the inference conducted to determine signficance of effects and how these change over time.
-Section \ref{conclusion} summarises the work.
-Two supplementary sections document the contributions of others to this work and the software.
-
-\hypertarget{data}{%
-\section{Data collection, wrangling and imputation}\label{data}}
-
-\hypertarget{collecting-the-data}{%
-\subsection{Collecting the data}\label{collecting-the-data}}
-
-The voting outcome of interest is the electoral two-party preferred vote, which is provided by the Australian Electoral Commission (AEC) for the 2001, 2004, 2007, 2010, 2013 and 2016 elections via the AEC Tally Room.
-The AEC divides Australia into 150 regions, called electorates, with each corresponding to a single seat in the House of Representatives.
-Voting is compulsory in Australia, and each voter assigns a numbered preference to each available candidate in their electorate.
-The two-party preferred vote is determined by a tally of these preferences where, by convention, only the ranks of the Labor and Liberal candidates are considered.
-This is recorded as a percentage preference in favour of the Liberal party.
-
-Socio-demographic variables were derived from the Australian Census of Population and Housing (Census), which is a survey of every household in Australia, recording information such as age, gender, ethnicity, education level and income.
-There have been four Censuses so far in the 21st century, conducted in 2001, 2006, 2011 and 2016.
-The Australian Bureau of Statistics (ABS) conducts the Census and publishes aggregated information.
-The ABS uses electoral boundaries as defined by the AEC at the time of each Census, which may not match those in place at the subsequent and previous elections.
-From the available Census information aggregated at the electorate level, 65 socio-demographic variables were defined for each of the electorates to be used in the analysis.
-
-Raw data was sourced online from the AEC and ABS websites in \texttt{.csv} and \texttt{.xlsx} files.
-The formats of these files differ over the years, making extracting the appropriate information a big task.
-The functions available in the \texttt{dplyr} \autocite{dplyr} and \texttt{readxl} \autocite{readxl} \texttt{R} packages are particularly useful, as they provide fast consistent tools for data manipulation and functions to import \texttt{.xlsx} files (respectively).
-The 2001 and 2006 Census data are published in a format where the information for each electorate is held in a separate document making it difficult to use the \texttt{dplyr} tools.
-Instead, cells have to be selected from each individual file to construct the desired variables.
-All scripts required for the data wrangling process can be found in the github repository for the \texttt{eechidna} \texttt{R} package \autocite{eechidna}, along with the raw data.
-The \texttt{eechidna} package makes this study entirely reproducible and provides a resource to help wrangle data for future Censuses and elections, when they become available.
-
-\hypertarget{joining-census-and-election-data}{%
-\subsection{Joining Census and election data}\label{joining-census-and-election-data}}
-
-\hypertarget{differences-between-census-and-election-data}{%
-\subsubsection{Differences between Census and election data}\label{differences-between-census-and-election-data}}
-
-Between 2001 and 2016 there were six elections and four Censuses (see Figure \ref{fig:timeline}).
-Electoral boundaries are redistributed regularly by the AEC, meaning that only in the years where both a Census and election occur are all boundaries likely to match --- the case for the 2001 and 2016 elections.
-Therefore, for the four elections between 2004 and 2013, both temporal and spatial differences in electorates need to be accounted for when joining the electoral two-party preferred vote with Census data.
-For these elections a spatio-temporal imputation method was employed to obtain electoral socio-demographics.
-This method uses Census information from both before and after the election of interest.
+The Census provides data on electoral socio-demographics, and vote counts in each electorate can be obtained from Australian federal elections.
+However, joining these two data sources is difficult because the Censuses are not held at the same time as the elections.
+Between 2001 and 2016 there were six elections and four Censuses, as shown in the timeline below.
 
@@
 
-{\centering \includegraphics[width=0.9\linewidth]{elections_files/figure-latex/timeline-1}
+{\centering \includegraphics[width=0.9\linewidth]{elections_conversation_files/figure-latex/timeline-1}
 
@@
 
-\hypertarget{spatio-temporal-imputation}{%
-\subsubsection{Spatio-temporal imputation}\label{spatio-temporal-imputation}}
+Not only can an electorate change between the last Census and an election, but even the electorate boundaries can change.
+Some electorates can disappear altogether and new electorates can arise.
+Electoral boundaries are redistributed regularly by the AEC, meaning that only in the years where both a Census and election occur are all boundaries likely to match --- the case for the 2001 and 2016 elections.
+So we first had to estimate what the socio-demographic characteristics of an electorate would have been at the time of each election using a complicated method of interpolation over time and geography.
+This method uses Census information from both before and after the election of interest, and information from neighbouring electorates when boundaries have changed.
 
-For each election, neighbouring Census information has to be combined in some way so that it represents the boundaries in place at the time of the election.
-This is done by taking the electoral boundaries and imputing the corresponding socio-demographic characteristics for each of neighbouring Censuses, thereby addressing the spatial aspect.
-Next, to deal with the temporal component, characteristics at the time of the election are constructed using linear interpolation between the spatially imputed neighbouring Census variables.
+A simple way to measure voting patterns is to consider the two-party preferred (2PP) vote, which is based on the tally of preferences for the Labour and Liberal candidates, ignoring all other candidates.
+By convention, this is recorded as a percentage preference in favour of the Liberal party -- so a 2PP value of 45\% indicates that 45\% of voters preferred a Liberal candidate and 55\% of voters preferred a Labour candidate.
 
-To account for spatial differences, the piecewise approximation method in \textcite{Goodchild1993} is adopted.
-Consider a map of source zones \(s = 1,\dots,S\), for which socio-demographic information is available, and a set of target zones \(t = 1,\dots,T\) for which information is to be imputed.
-This is described in the context of a single election, and a single neighbouring Census.
-
-Let the map of electoral boundaries at the time of a Census define the source zones, and let the boundaries at the time of the election be the target zones.
-Denote the area of intersection between a source zone \(s\) and a target zone \(t\) as \(A_{s,t}\).
-Additionally, let the population of the source zone \(s\) be \(U_s\) and the population of intersection between source zone \(s\) and target zone \(t\) be \(P_{s,t}\).
-The estimated population of intersection is given by
-\[
-  \hat{P}_{s,t} = \frac{U_s*A_{s,t}}{\sum_{t=1}^T A_{s,t}},
-  \quad\text{for all $s=1,\dots,S$ and $t=1,\dots,T$}.
-\]
-Note that this estimator implicitly assumes that populations are uniformly distributed within each source zone.
-
-In order to calculate socio-demographic information for each of the target zones, a weighted average is taken using the estimated intersection populations as weights.
-Denote a given Census variable for the target zone by \(C_t\), and the same Census variable for the source zone as \(D_s\).
-Then, estimate \(C_t\) using
-\[
-  \hat{C}_t = \frac{\sum_{s=1}^{S}{D_s*\hat{P}_{s,t}}}%
-  {\sum_{s=1}^{S}{\hat{P}_{s,t}}},
-  \quad\text{for each $t=1,\dots,T$}.
-\]
-This concludes the spatial imputation of the socio-demographic characteristics for one target zone (a single electoral boundary), at the time of only one of the neighbouring Censuses.
-This process is repeated for all of the target zones, and then for the other neighbouring Census.
-
-\enlargethispage*{0.5cm}
-
-To account for temporal changes, linear interpolation is used between Census years to get the final estimate of a Census variable for the target zone in the election year.
-Let \(y_1\) be the year of the Census preceding an election, let \(y_2\) be the year of the election, and \(y_3\) be the year of the Census that follows.
-Add this year subscript to the Census variable estimate \(\hat{C}_t\), resulting in \(\hat{C}_{t,y}\).
-Linear interpolating between these Census years results an imputed value for the election year, given by
-\[
-  \hat{C}_{t,y_2} = \frac{y_3-y_2}{y_3-y_1} * \hat{C}_{t,y_1} +
-    \frac{y_2-y_1}{y_3-y_1} * \hat{C}_{t,y_3}.
-\]
-Implicitly this assumes that population characteristics change in a linear manner over time.
-
-\hypertarget{an-illustration-of-the-spatio-temporal-imputation}{%
-\subsubsection*{An illustration of the spatio-temporal imputation}\label{an-illustration-of-the-spatio-temporal-imputation}}
-\addcontentsline{toc}{subsubsection}{An illustration of the spatio-temporal imputation}
-
-Census data is publicly available at different levels of aggregation, ranging from SA1 (over 50,000 zones) to electoral divisions (150 zones).
-For this study, electoral divisions are used as source zones, and the imputation method is applied to produce socio-demographic variables for each of the 2004, 2007, 2010 and 2013 elections.
-As mentioned earlier, there is no need to impute socio-demographic variables for the 2001 and 2016 elections.
-To illustrate the method, consider the imputation of socio-demographic variables for the electorate of Hume in New South Wales (NSW) at the time of the 2013 federal election.
-The boundaries shown in Figure \ref{fig:hume13} define all target zones in NSW for 2013, with the target zone of interest (Hume) shaded purple.
-
-\begin{figure}[h]
-
-{\centering \includegraphics[height=0.28\textheight]{elections_files/figure-latex/hume13-1}
-
-}
-
-\caption{Some of the electoral boundaries in NSW for 2013, with the electoral boundary for Hume shown in purple.}\label{fig:hume13}
-\end{figure}
-
-\begin{figure}[h]
-
-{\centering \includegraphics[height=0.28\textheight]{elections_files/figure-latex/hume16-1}
-
-}
-
-\caption{Census division boundaries in NSW for 2016, with the 2013 electoral boundary for Hume, shown in purple. The purple region is not contained within a single Census division.}\label{fig:hume16}
-\end{figure}
-
-The corresponding source zones from the 2016 Census are shown in Figure \ref{fig:hume16}.
-As can be seen, the Hume boundary from the 2013 election (shaded purple) does not perfectly match any of the source zones.
-
-There are many source zones from the 2016 Census that intersect with this purple region, including the divisions of Riverina, Eden-Monaro and Hume, along with smaller intersecting areas with Fenner, Calare, Gilmore and Whitlam.
-The proportion of each source zone that overlaps with the purple region is calculated, and used to obtain the intersecting populations \(\hat{P}_{s,t}\).
-
-\begin{table}[t]
-
-\caption{\label{tab:hume1}Population from each intersecting Census source zone (2016) that is allocated to the target zone (purple region - Hume electoral boundary in 2013), and the corresponding proportion of Australian citizens in each of these source zones.}
-\centering
-\resizebox{\linewidth}{!}{
-\fontsize{10}{12}\selectfont
-\begin{tabular}{lrrrr}
-\toprule
-Source zone (2016) & Proportion & Source zone population & Population allocated to purple region: $\hat{P}_{s,t}$ & AusCitizen (\%): $D_s$\\
-\midrule
-Hume & 0.9654 & 150643 & 145427 & 90.0168\\
-Riverina & 0.2511 & 155793 & 39117 & 89.1144\\
-Eden-Monaro & 0.1109 & 147532 & 16358 & 87.9999\\
-Canberra & 0.0028 & 196037 & 548 & 85.4793\\
-Fenner & 0.0023 & 202955 & 474 & 83.6432\\
-\addlinespace
-Whitlam & 0.0006 & 152280 & 92 & 89.5173\\
-Gilmore & 0.0006 & 150436 & 86 & 89.0266\\
-Calare & 0.0001 & 161298 & 21 & 87.5603\\
-\bottomrule
-\end{tabular}}
-\end{table}
-
-Now consider the socio-demographic variable \emph{AusCitizen}, the proportion of people in the region who are Australian citizens.
-A weighted average of \emph{AusCitizen}, with the allocated population from each source zone as weights, yields \(\hat{C}_{\text{Hume},2016} = 89.65\%\).
-Repeating this process using the 2011 Census yields \(\hat{C}_{\text{Hume},2011} = 91.00\%\).
-Finally, linear interpolation between 2011 and 2016 yields the 2013 estimate:
-\[
-  \hat{C}_{\text{Hume},2013}
-    = \frac{3}{5} \hat{C}_{\text{Hume},2011} + \frac{2}{5} \hat{C}_{\text{Hume},2016}
-    % = \frac{3}{5} \cdot 91.00 \% + \frac{2}{5} \cdot 89.65 \% \\
-    =  90.46\%.
-\]
-This is done for each of the 65 socio-demographic variables, and is repeated for each of the 149 remaining target zones corresponding with 2013 electorates.
+We consider how various socio-demographic variables obtained from Census data can be used to explain the 2PP values for each of the 150 electorates in each of the federal elections between 2001 and 2016.
 
@@
 
-{\centering \includegraphics[height=0.2\textheight,]{elections_files/figure-latex/rhovis-1}
+{\centering \includegraphics[height=0.2\textheight,]{elections_conversation_files/figure-latex/rhovis-1}
 
@@
 
-{\centering \includegraphics[height=0.2\textheight,]{elections_files/figure-latex/plotintercept-1}
+{\centering \includegraphics[height=0.2\textheight,]{elections_conversation_files/figure-latex/plotintercept-1}
 
@@
 \subsection{Influential socio-demographics}\label{influential-socio-demographics}}
-
-\begin{figure}[h]
-
-{\centering \includegraphics{elections_files/figure-latex/partresplot1-1}
-
-}
-
-\caption{Partial residual plots by election year for a selection of predictors.
-Linear model with 95\% confidence bands overlaid.
-Most predictors have a positive relationship: the larger the value the more likely the electorate preferences the Coalition.
-The relationship is relatively robust over time, with the exception of Unemployment.}\label{fig:partresplot1}
-\end{figure}
-
-\begin{figure}[h]
-
-{\centering \includegraphics{elections_files/figure-latex/partresplot2-1}
-
-}
-
-\caption{Partial residual plots by election year for a selection of predictors.
-Linear model with 95\% confidence bands overlaid.
-Several predictors have a negative relationship: with larger values indicating the electorate more likely preferences Labor.
-Most relationships are relatively stable over elections, except OtherLanguage and Education.}\label{fig:partresplot2}
-\end{figure}
 
@@
 
-{\centering \includegraphics[width=0.7\linewidth,]{elections_files/figure-latex/resstate-1}
+{\centering \includegraphics[width=0.7\linewidth,]{elections_conversation_files/figure-latex/resstate-1}
 

### 2019-05-05 elections_conversation.Rmd
Working on modelling

@@
 
-# Modelling {#modelling}
-
-Following this process, electoral socio-demographic variables are available for each of the six elections and can be joined with their corresponding two-party preferred votes.
-Before choosing an appropriate model, two issues with the socio-demographic variables need to be addressed.
-First, variable scales change over the years, making it important to standardize variables.
-Second, many variables represent similar information and where appropriate, should be combined in some way.
-Principal component analysis is used to identify variables covary, from which intuitive groupings are selected to be combined into a single variable.
-This also reduces the dimension of the data.
-After these steps, a model specification is chosen.
-
-## Standardizing variables
-
 Many of the socio-demographic variables have changing scales over the years.
@@
 
-## Creating factors
+# Modelling {#modelling}
 
-There are only $N = 150$ observations (electorates) in each election and $p = 65$ socio-demographic variables in each cross-section, with many variables represent similar information about an electorate.
-Any model that uses all variables would face serious problems with multi-collinearity and over-fitting, which would likely lead to erroneous conclusions regarding variable significance.
-To address this, groups of variables that represent similar information are combined into "factors"[^2].
-
-[^2]: A preliminary step involved removing all age bands, because age is represented by median age, and to remove variables relating to particular denominations of Christianity.
-
-A factor is created from a group of variables if there is an intuitive reason as to why they should represent similar information and if there is evidence to suggest that they covary.
+There are dozens of socio-demographic variables available in each Census, with many variables representing similar information about an electorate.
+To reduce the numbers of variables, we combine variables with similar information.
 For example, a potential group would be variables relating to electoral incomes --- median family, household and personal incomes.

### 2019-05-05 elections_conversation.tex
Working on modelling

@@
 
-\hypertarget{modelling}{%
-\section{Modelling}\label{modelling}}
-
-Following this process, electoral socio-demographic variables are available for each of the six elections and can be joined with their corresponding two-party preferred votes.
-Before choosing an appropriate model, two issues with the socio-demographic variables need to be addressed.
-First, variable scales change over the years, making it important to standardize variables.
-Second, many variables represent similar information and where appropriate, should be combined in some way.
-Principal component analysis is used to identify variables covary, from which intuitive groupings are selected to be combined into a single variable.
-This also reduces the dimension of the data.
-After these steps, a model specification is chosen.
-
-\hypertarget{standardizing-variables}{%
-\subsection{Standardizing variables}\label{standardizing-variables}}
-
 Many of the socio-demographic variables have changing scales over the years.
@@
 
-\hypertarget{creating-factors}{%
-\subsection{Creating factors}\label{creating-factors}}
+\hypertarget{modelling}{%
+\section{Modelling}\label{modelling}}
 
-There are only \(N = 150\) observations (electorates) in each election and \(p = 65\) socio-demographic variables in each cross-section, with many variables represent similar information about an electorate.
-Any model that uses all variables would face serious problems with multi-collinearity and over-fitting, which would likely lead to erroneous conclusions regarding variable significance.
-To address this, groups of variables that represent similar information are combined into ``factors''\footnote{A preliminary step involved removing all age bands, because age is represented by median age, and to remove variables relating to particular denominations of Christianity.}.
-
-A factor is created from a group of variables if there is an intuitive reason as to why they should represent similar information and if there is evidence to suggest that they covary.
+There are dozens of socio-demographic variables available in each Census, with many variables representing similar information about an electorate.
+To reduce the numbers of variables, we combine variables with similar information.
 For example, a potential group would be variables relating to electoral incomes --- median family, household and personal incomes.

### 2019-05-05 elections_conversation.Rmd
Moved some code to wrangle.R

@@
 source("functions.R")
-```
-
-```{r wrangle, include = FALSE}
-# Take log of indigenous, judaism, islam, buddhism
-abs2016 <- eechidna::abs2016 %>%
-  mutate(
-    Indigenous = log(Indigenous),
-    Judaism = log(Judaism),
-    Islam = log(Islam),
-    Buddhism = log(Buddhism)
-  )
-abs2013 <- eechidna::abs2013 %>%
-  mutate(
-    Indigenous = log(Indigenous),
-    Judaism = log(Judaism),
-    Islam = log(Islam),
-    Buddhism = log(Buddhism)
-  )
-abs2010 <- eechidna::abs2010 %>%
-  mutate(
-    Indigenous = log(Indigenous),
-    Judaism = log(Judaism),
-    Islam = log(Islam),
-    Buddhism = log(Buddhism)
-  )
-abs2007 <- eechidna::abs2007 %>%
-  mutate(
-    Indigenous = log(Indigenous),
-    Judaism = log(Judaism),
-    Islam = log(Islam),
-    Buddhism = log(Buddhism)
-  )
-abs2004 <- eechidna::abs2004 %>%
-  mutate(
-    Indigenous = log(Indigenous),
-    Judaism = log(Judaism),
-    Islam = log(Islam),
-    Buddhism = log(Buddhism)
-  )
-abs2001 <- eechidna::abs2001 %>%
-  mutate(
-    Indigenous = log(Indigenous),
-    Judaism = log(Judaism),
-    Islam = log(Islam),
-    Buddhism = log(Buddhism)
-  )
-
-# Change abs2016 to data frame to prevent error below
-abs2016 <- as.data.frame(abs2016)
-
-# Combine and standardize
-my_df <- bind_rows(
-  left_join(eechidna::tpp01,
-    standardise_vars(abs2001) %>%
-      dplyr::select(-c(UniqueID, Area, ends_with("NS"), Population)),
-    by = c("DivisionNm", "StateAb" = "State")
-  ) %>% mutate(year = "2001"),
-  left_join(eechidna::tpp04,
-    standardise_vars(abs2004) %>%
-      dplyr::select(-UniqueID),
-    by = c("DivisionNm")
-  ) %>% mutate(year = "2004"),
-  left_join(eechidna::tpp07,
-    standardise_vars(abs2007) %>%
-      dplyr::select(-UniqueID),
-    by = c("DivisionNm")
-  ) %>% mutate(year = "2007"),
-  left_join(eechidna::tpp10,
-    standardise_vars(abs2010) %>%
-      dplyr::select(-UniqueID),
-    by = c("DivisionNm")
-  ) %>% mutate(year = "2010"),
-  left_join(eechidna::tpp13,
-    standardise_vars(abs2013) %>%
-      dplyr::select(-UniqueID),
-    by = c("DivisionNm")
-  ) %>% mutate(year = "2013"),
-  left_join(eechidna::tpp16,
-    standardise_vars(abs2016) %>%
-      dplyr::select(-c(UniqueID, Area, ends_with("NS"), Population)),
-    by = c("DivisionNm", "StateAb" = "State")
-  ) %>% mutate(year = "2016")
-) %>%
-  mutate(year = factor(year)) %>%
-  select(-c(
-    starts_with("Age"), StateAb,
-    LNP_Votes, ALP_Votes, ALP_Percent, TotalVotes, Swing,
-    InternetUse, InternetAccess, EnglishOnly,
-    Other_NonChrist, OtherChrist, Catholic, Anglican,
-    Volunteer, EmuneratedElsewhere, DivisionID
-  ))
-
-# Create final df for modelling
-
-factors_df <- my_df %>%
-  mutate(
-    Education = BachelorAbv + HighSchool + Professional + Finance - Laborer - Tradesperson - DipCert,
-    FamHouseSize = FamilyRatio + AverageHouseholdSize + Couple_WChild_House - Couple_NoChild_House - SP_House,
-    PropertyOwned = Owned + Mortgage - Renting - PublicHousing,
-    RentLoanPrice = MedianRent + MedianLoanPay,
-    Incomes = MedianFamilyIncome + MedianHouseholdIncome + MedianPersonalIncome,
-    Unemployment = Unemployed - LFParticipation
-  ) %>%
-  dplyr::select(-c(
-    BachelorAbv, HighSchool, Professional, Finance, Laborer, Tradesperson, DipCert, FamilyRatio,
-    AverageHouseholdSize, Couple_WChild_House, Couple_NoChild_House, SP_House, Owned, Mortgage, Renting,
-    PublicHousing, MedianFamilyIncome, MedianHouseholdIncome, MedianPersonalIncome, MedianRent,
-    MedianLoanPay, Unemployed, LFParticipation
-  ))
-
-# Now standardize factors
-
-small_df <- bind_rows(
-  factors_df %>% filter(year == "2001") %>% standardise_vars(),
-  factors_df %>% filter(year == "2004") %>% standardise_vars(),
-  factors_df %>% filter(year == "2007") %>% standardise_vars(),
-  factors_df %>% filter(year == "2010") %>% standardise_vars(),
-  factors_df %>% filter(year == "2013") %>% standardise_vars(),
-  factors_df %>% filter(year == "2016") %>% standardise_vars()
-)
-
-# Order electorates in alphabetical order to match spatial matrix
-model_df <- small_df %>%
-  arrange(year, DivisionNm) %>%
-  dplyr::select(order(colnames(.))) %>%
-  rename(
-    ManagerAdmin = ManagerAdminClericalSales,
-    OtherLanguage = OtherLanguageHome,
-    BornSEEuro = Born_SE_Europe,
-    BornAsia = Born_Asia,
-    OneParentHouse = OneParent_House
-  )
-```
-
-```{r spatialweights, eval = FALSE}
-# Get spatial weights
-# This is not run, as weights are pulled from github, but this can be run to get them from scratch.
-sF_16 <- sF_download(2016)
-sF_13 <- sF_download(2013)
-sF_10 <- sF_download(2010)
-sF_07 <- sF_download(2007)
-sF_04 <- sF_download(2004)
-sF_01 <- sF_download(2001)
-
-sp_weights_16 <- sp_weights_matrix(sF_16)
-sp_weights_13 <- sp_weights_matrix(sF_13)
-sp_weights_10 <- sp_weights_matrix(sF_10)
-sp_weights_07 <- sp_weights_matrix(sF_07)
-sp_weights_04 <- sp_weights_matrix(sF_04)
-sp_weights_01 <- sp_weights_matrix(sF_01)
-```
-
-```{r FGLS, include = FALSE}
-# Instead of running spatial weights, load from github
-load("data/sp_weights_01.rda")
-load("data/sp_weights_04.rda")
-load("data/sp_weights_07.rda")
-load("data/sp_weights_10.rda")
-load("data/sp_weights_13.rda")
-load("data/sp_weights_16.rda")
-
-## Run full models for each year
-full_formula <- "LNP_Percent ~ ."
-
-# 2016
-glsmod16 <- my_fgls(full_formula,
-  my_data = model_df %>% filter(year == "2016"),
-  sp_weights = sp_weights_16
-)
-
-# 2013
-glsmod13 <- my_fgls(full_formula,
-  my_data = model_df %>% filter(year == "2013"),
-  sp_weights = sp_weights_13
-)
-
-# 2010
-glsmod10 <- my_fgls(full_formula,
-  my_data = model_df %>% filter(year == "2010"),
-  sp_weights = sp_weights_10
-)
-
-# 2007
-glsmod07 <- my_fgls(full_formula,
-  my_data = model_df %>% filter(year == "2007"),
-  sp_weights = sp_weights_07
-)
-
-# 2004
-glsmod04 <- my_fgls(full_formula,
-  my_data = model_df %>% filter(year == "2004"),
-  sp_weights = sp_weights_04
-)
-
-# 2001
-glsmod01 <- my_fgls(full_formula,
-  my_data = model_df %>% filter(year == "2001"),
-  sp_weights = sp_weights_01
-)
-
-## Visualise coefficients and significance
-
-coef_df <- bind_rows(
-  data.frame(
-    variable = glsmod16$coefficients %>% names(),
-    estimate = glsmod16$coefficients %>% unname(),
-    se = summary(glsmod16)$tTable[, "Std.Error"] %>% unname(),
-    p = summary(glsmod16)$tTable[, "p-value"] %>% unname(),
-    year = 2016),
-  data.frame(
-    variable = glsmod13$coefficients %>% names(),
-    estimate = glsmod13$coefficients %>% unname(),
-    se = summary(glsmod13)$tTable[, "Std.Error"] %>% unname(),
-    p = summary(glsmod13)$tTable[, "p-value"] %>% unname(),
-    year = 2013),
-  data.frame(
-    variable = glsmod10$coefficients %>% names(),
-    estimate = glsmod10$coefficients %>% unname(),
-    se = summary(glsmod10)$tTable[, "Std.Error"] %>% unname(),
-    p = summary(glsmod10)$tTable[, "p-value"] %>% unname(),
-    year = 2010),
-  data.frame(
-    variable = glsmod07$coefficients %>% names(),
-    estimate = glsmod07$coefficients %>% unname(),
-    se = summary(glsmod07)$tTable[, "Std.Error"] %>% unname(),
-    p = summary(glsmod07)$tTable[, "p-value"] %>% unname(),
-    year = 2007),
-  data.frame(
-    variable = glsmod04$coefficients %>% names(),
-    estimate = glsmod04$coefficients %>% unname(),
-    se = summary(glsmod04)$tTable[, "Std.Error"] %>% unname(),
-    p = summary(glsmod04)$tTable[, "p-value"] %>% unname(),
-    year = 2004),
-  data.frame(
-    variable = glsmod01$coefficients %>% names(),
-    estimate = glsmod01$coefficients %>% unname(),
-    se = summary(glsmod01)$tTable[, "Std.Error"] %>% unname(),
-    p = summary(glsmod01)$tTable[, "p-value"] %>% unname(),
-    year = 2001)
-)
+source("wrangle.R")
 ```
@@
 
-A simple way to measure voting patterns is to consider the two-party preferred (2PP) vote, which is based on the tally of preferences for the Labour and Liberal candidates, ignoring all other candidates.
-By convention, this is recorded as a percentage preference in favour of the Liberal party -- so a 2PP value of 45% indicates that 45% of voters preferred a Liberal candidate and 55% of voters preferred a Labour candidate.
+A simple way to measure voting patterns is to consider the two-party preferred (2PP) vote, which is based on the tally of preferences for the Labor and Liberal candidates, ignoring all other candidates.
+By convention, this is recorded as a percentage preference in favour of the Liberal party --- for example, a 2PP value of 45% indicates that 45% of voters ranked the Liberal candidate higher than the Labor candidate, while the remaining 55% ranked them in the reverse order.
 
@@
 
+# 2PP Modelling {#modelling}
+
 Many of the socio-demographic variables have changing scales over the years.
 For example, inflation-adjusted median rental prices increased across almost all electorates, with median rent of 200 dollars per week placing an electorate in the 90th percentile in 2001, but only the 30th percentile in 2016.
-In order for socio-demographic effects to be comparable across years, all explanatory variables are standardized to have mean zero and variance one within each election year.
+In order for socio-demographic effects to be comparable across years, all explanatory variables are standardized so they are comparable across years.
 By standardizing, each variable is reported as a relative measure compared to all other electorates in the same year.
 
-# Modelling {#modelling}
+There are dozens of socio-demographic variables available in the Censuses, with many variables representing similar information about an electorate.
+So we combined some variables to avoid redundant information.
+For example, our "Incomes" variable is a combination of median personal income, household income and family income.
+In this way, we reduced the data set to include 30 socio-demographic variables, listed below.
 
-There are dozens of socio-demographic variables available in each Census, with many variables representing similar information about an electorate.
-To reduce the numbers of variables, we combine variables with similar information.
-For example, a potential group would be variables relating to electoral incomes --- median family, household and personal incomes.
-To determine which variables covary, principal component analysis is used on a combined dataset of socio-demographic variables from all six elections.
-It is appropriate to compute principal components in this way because when computed separately for each election, scree plots level off after four components and the loadings of the first four components are similar across the elections.
-
-Only the first four principal components from the combined dataset are considered, as the scree plot corresponding to the combined dataset levels off after the fourth component.
-Variables that have a large loading in a particular component are deemed to covary, with a loading with magnitude greater than 0.15 being considered large.
-Six factors are created using this criteria.
-These are:
-`Incomes` (median personal, household and family incomes);
-`Unemployment` (unemployment and labor force participation rates);
-`ProportyOwned` (rates of housing ownership, mortgages, renting and government housing);
-`RentLoanPrice` (median rental and loan repayments);
-`FamHouseSize` (average household size, ratio of people to families and household makeup (single person, couple with kids and couple without kids); and
-`Education` (high school and university qualifications, jobs requiring higher levels of education as well as vocational course completions and jobs that do not require higher education levels, such as laborer or tradesperson).
-For each of these groups, variables with positive loadings are added and those with negative loadings are subtracted to create a factor.
-After computing these sums, each factor is standardized to have mean zero and variance one, within each election.
-
-There are $p=30$ variables in the resultant predictor set, with all of these used in the regression for each election.
-
-## Regression incorporating spatially dependent errors
-
-An identical model specification is used for each of the six elections, with each election modelled separately.
-This allows the socio-demographic effects to be estimated separately for each election year, facilitating analysis of temporal changes in variable effects.
-This approach is preferable to using a single longitudinal model because it avoids any concerns of undue bias stemming from incorrectly imposed time-varying restrictions on any variable.
-Without such restrictions, a pooled cross-sectional model does not yield any distinct advantage over separate cross-sections.
-The panel approach is avoided because of how frequently electoral boundaries change, noting that electorates with the same name across elections are not guaranteed to represent the same geographical region.
-Therefore any fixed or random effects models would be difficult to estimate without implementing consistent boundaries, which would require further imputation.
-
-For each cross-section, let the response $\bm{y}$ be the vector two-party preferred vote in favour of the Liberal party; for example,  $y_i = 70$ represents a 70% preference for Liberal, 30% for Labor, in electorate $i$.
-Although $y_i$ lies in the interval $(0,100)$, observed values are never close to 0 or 100 (minimum $24.05 \%$ and maximum $74.90 \%$), so there is no need to formally impose the constraint of $y_i \in [0,100]$.
-Furthermore, the responses are found to be spatially correlated in each election (Moran's I test, $p \le 7\cdot10^{-15}$).
-This is not surprising as electorates are aggregate spatial units, and hence the spatial structure of the data must be modelled appropriately.
-
-The spatial error model [@Anselin88] is chosen because it captures spatial heterogeneity by incorporating a spatially structured random effect vector [@LeSage2009].
-In this context, the random effect can be thought of as capturing the unobserved political climate in each electorate, where this climate is correlated with the climate in neighbouring electorates, under the assumption that the climate is independent of electoral socio-demographics.
-
-```{r printmodels, warning=FALSE, message=FALSE, include=FALSE}
+```{r models, warning=FALSE, message=FALSE, include=FALSE}
 # Add rho and GLS residual standard error
@@
 ```
+
+## Regression incorporating spatially dependent errors
+
+An identical model specification is used for each of the six elections, with each election modelled separately.
+This allows the socio-demographic effects to be estimated separately for each election year, facilitating analysis of temporal changes in variable effects.
+This approach is preferable to using a single longitudinal model because it avoids any concerns of undue bias stemming from incorrectly imposed time-varying restrictions on any variable.
+Without such restrictions, a pooled cross-sectional model does not yield any distinct advantage over separate cross-sections.
+The panel approach is avoided because of how frequently electoral boundaries change, noting that electorates with the same name across elections are not guaranteed to represent the same geographical region.
+Therefore any fixed or random effects models would be difficult to estimate without implementing consistent boundaries, which would require further imputation.
+
+For each cross-section, let the response $\bm{y}$ be the vector two-party preferred vote in favour of the Liberal party; for example,  $y_i = 70$ represents a 70% preference for Liberal, 30% for Labor, in electorate $i$.
+Although $y_i$ lies in the interval $(0,100)$, observed values are never close to 0 or 100 (minimum $24.05 \%$ and maximum $74.90 \%$), so there is no need to formally impose the constraint of $y_i \in [0,100]$.
+Furthermore, the responses are found to be spatially correlated in each election (Moran's I test, $p \le 7\cdot10^{-15}$).
+This is not surprising as electorates are aggregate spatial units, and hence the spatial structure of the data must be modelled appropriately.
+
+The spatial error model [@Anselin88] is chosen because it captures spatial heterogeneity by incorporating a spatially structured random effect vector [@LeSage2009].
+In this context, the random effect can be thought of as capturing the unobserved political climate in each electorate, where this climate is correlated with the climate in neighbouring electorates, under the assumption that the climate is independent of electoral socio-demographics.
 

### 2019-05-05 elections_conversation.Rmd
Updated to the start of the graphics

@@
 
-```{r models, warning=FALSE, message=FALSE, include=FALSE}
-# Add rho and GLS residual standard error
-regtable <- stargazer::stargazer(
-  glsmod01, glsmod04, glsmod07, glsmod10, glsmod13, glsmod16,
-  header = FALSE,
-  column.sep.width = "1pt",
-  no.space=TRUE,
-  model.numbers=FALSE,
-  keep.stat = c("n","ser"),
-  dep.var.caption="",
-  dep.var.labels="\\vspace{-0.5cm}",
-  title = "Estimated spatial error model parameters (standard errors) for each of the six election years.",
-  #dep.var.labels = "Two-party preferred vote in favor of the Liberal party",
-  font.size = "scriptsize",
-  digits = 2,
-  column.labels = c("2001", "2004", "2007", "2010", "2013", "2016"),
-  label="tab:printmodels"
-  #notes = c("Estimated coefficients for variable named in column one shown for", "election year indicated by column heading, with estimated standard", "deviation for each coefficient shown below in parenthesis.
-Overall", "summary measures for each regression equation are provided in the", "bottom panel."),
-  #notes.align = "l"
-)
-# Add correlations
-getrho <- function(x, element) {x$rho_df[[element]]}
-rho <- map_dbl(list(glsmod01,glsmod04,glsmod07,glsmod10,glsmod13,glsmod16),
-           getrho, element="estimate")
-rhop <- map_dbl(list(glsmod01,glsmod04,glsmod07,glsmod10,glsmod13,glsmod16),
-           getrho, element="p")
-rhose <- map_dbl(list(glsmod01,glsmod04,glsmod07,glsmod10,glsmod13,glsmod16),
-           getrho, element="se")
-rholine1 <- paste("$\\rho$",
-            paste(" & ",
-                  sprintf("%3.2f",rho),
-                  ifelse(rhop < 0.01, "$^{***}$",
-                  ifelse(rhop < 0.05, "$^{**}$",
-                  ifelse(rhop < 0.10, "$^{*}$",""))),
-              sep="", collapse=""),
-            "\\\\", sep="", collapse="")
-rholine2 <- paste("",
-            paste(" & (",
-                  sprintf("%3.2f",rhose), ")",
-              sep="", collapse=""),
-            "\\\\", sep="", collapse="")
-regtable <- c(regtable[1:11], rholine1, rholine2, regtable[12:length(regtable)])
-# Add residual standard error (GLS)
-getsigma <- function(x, element) {x$sigma}
-sigma <- map_dbl(list(glsmod01,glsmod04,glsmod07,glsmod10,glsmod13,glsmod16),
-           getsigma)
-sigmaline <- paste("Residual Standard Error (GLS)",
-             paste(" & ",
-                   sprintf("%4.2f", sigma),
-                   sep="", collapse=""),
-             "\\\\", sep="", collapse="")
-regtable <- c(regtable[1:77],sigmaline, regtable[78:length(regtable)])
-# Allow alignment on decimal point
-regtable[6] <- "\\begin{tabular}{@{\\extracolsep{1pt}}lrrrrrr}"
-regtable[12:75] <- gsub(")","\\\\rlap{)}",regtable[12:75])
-regtable[12:75] <- gsub("$^{*}$","\\rlap{$^{*}$}", regtable[12:75], fixed=TRUE)
-regtable[12:75] <- gsub("$^{**}$","\\rlap{$^{**}$}", regtable[12:75], fixed=TRUE)
-regtable[12:75] <- gsub("$^{***}$","\\rlap{$^{***}$}", regtable[12:75], fixed=TRUE)
-regtable[5] <- "\\scriptsize\\tabcolsep=0.35cm"
-```
+ * "AusCitizen"
+ * "BornAsia"
+ * "Born_MidEast"
+ * "BornSEEuro"
+ * "Born_UK"
+ * "BornElsewhere"
+ * "Buddhism"
+ * "Christianity"
+ * "CurrentlyStudying"
+ * "DeFacto"
+ * "DiffAddress"
+ * "Distributive"
+ * "Education"
+ * "Extractive"
+ * "FamHouseSize"
+ * "Incomes"
+ * "Indigenous"
+ * "Islam"
+ * "Judaism"
+ * "ManagerAdmin"
+ * "Married"
+ * "MedianAge"
+ * "NoReligion"
+ * "OneParentHouse"
+ * "OtherLanguage"
+ * "PropertyOwned"
+ * "RentLoanPrice"
+ * "SocialServ"
+ * "Transformative"
+ * "Unemployment"
 
-```{r printmodels2, results='asis', dependson='printmodels'}
-cat(regtable)
-```
-
-## Regression incorporating spatially dependent errors
-
-An identical model specification is used for each of the six elections, with each election modelled separately.
-This allows the socio-demographic effects to be estimated separately for each election year, facilitating analysis of temporal changes in variable effects.
-This approach is preferable to using a single longitudinal model because it avoids any concerns of undue bias stemming from incorrectly imposed time-varying restrictions on any variable.
-Without such restrictions, a pooled cross-sectional model does not yield any distinct advantage over separate cross-sections.
-The panel approach is avoided because of how frequently electoral boundaries change, noting that electorates with the same name across elections are not guaranteed to represent the same geographical region.
-Therefore any fixed or random effects models would be difficult to estimate without implementing consistent boundaries, which would require further imputation.
-
-For each cross-section, let the response $\bm{y}$ be the vector two-party preferred vote in favour of the Liberal party; for example,  $y_i = 70$ represents a 70% preference for Liberal, 30% for Labor, in electorate $i$.
-Although $y_i$ lies in the interval $(0,100)$, observed values are never close to 0 or 100 (minimum $24.05 \%$ and maximum $74.90 \%$), so there is no need to formally impose the constraint of $y_i \in [0,100]$.
-Furthermore, the responses are found to be spatially correlated in each election (Moran's I test, $p \le 7\cdot10^{-15}$).
-This is not surprising as electorates are aggregate spatial units, and hence the spatial structure of the data must be modelled appropriately.
-
-The spatial error model [@Anselin88] is chosen because it captures spatial heterogeneity by incorporating a spatially structured random effect vector [@LeSage2009].
-In this context, the random effect can be thought of as capturing the unobserved political climate in each electorate, where this climate is correlated with the climate in neighbouring electorates, under the assumption that the climate is independent of electoral socio-demographics.
-
-Spatial weights are calculated in accordance with the assumption that an electorate is equally correlated with any electorate that shares a part of its boundary.
-Let $\rho$ be the spatial autoregressive coefficient, $\bm{v}$ be a spherical error term, $\bm{W}$ be a matrix of spatial weights (containing information about the neighbouring regions), $\bm{X}$ be a matrix of socio-demographic covariates, $\bm{\beta}$ be a vector of regression coefficients and $\bm{a}$ be a spatially structured random effect vector.
-$$
-  \bm{y} = \bm{X} \bm{\beta} + \bm{a},
-$$
-and
-$$
-  \bm{a} = \rho \bm{W} \bm{a} + \bm{v},
-$$
-where $\bm{v} \sim N(\bm{0}, \sigma^2 \bm{I_n})$, and hence
-$$
-  \bm{y} = \bm{X} \bm{\beta} + (\bm{I}_n-\rho \bm{W})^{-1}\bm{v}.
-$$
-Estimation of the above spatial error model is undertaken using feasible generalized least squares.
-
-Table \@ref(tab:printmodels) details the estimated model coefficients and their estimated standard errors, for each of the six elections.
-An interpretation of these estimated values is provided in the next section.
-
-# Results
-
-## Spatial autoregressive parameter
-
-The spatial autoregressive coefficient $\rho$ is positive and significant in only the 2001 and 2016 elections (Figure \@ref(fig:rhovis)), meaning that in these elections, the political climate of an electorate appears to be affected by the attitudes of its neighbours.
-Conversely, in the other four elections, the spatial effect weakens to become insignificant.
-In these years, it appears that the spatial component does not explain anything not already explained by the electoral socio-demographics, meaning electorates effectively voted independently.
-
-```{r rhovis, fig.cap="Estimates of the spatial autoregressive parameter for each of the six elections, reported with their individual 95\\% confidence intervals.
-Only in 2001 and 2016 is there a significant spatial component.", out.height='20%', out.extra = ''}
-rho_df <- bind_rows(
-  glsmod16$rho_df %>% mutate(year = "2016"),
-  glsmod13$rho_df %>% mutate(year = "2013"),
-  glsmod10$rho_df %>% mutate(year = "2010"),
-  glsmod07$rho_df %>% mutate(year = "2007"),
-  glsmod04$rho_df %>% mutate(year = "2004"),
-  glsmod01$rho_df %>% mutate(year = "2001")
-) %>%
-  mutate(
-    upper95 = estimate + 1.96 * se,
-    lower95 = estimate - 1.96 * se
-  )
-
-rho_df %>%
-  ggplot() +
-  geom_point(aes(x = year, y = estimate), size = 3) +
-  geom_linerange(aes(x = year, ymin = lower95, ymax = upper95), size = 1.2) +
-  geom_hline(aes(yintercept = 0), alpha = 0.5, size = 1) +
-  scale_color_manual(values = c("grey50", "black")) +
-  labs(x = "Election year", y = "Estimate of spatial autoregressive parameter") +
-  lims(y = c(-1, 1)) +
-  guides(col = FALSE) +
-  coord_fixed(ratio = 1) +
-  theme_bw()
-```
+Each election was modelled separately, to allow us to see any changes over time, and to account for changing electorate boundaries.
 
@@
 
-```{r prepplot, include = FALSE}
+```{r prepplot, echo = FALSE}
 p1 <- grid_visreg("Extractive", top = TRUE)
@@
 The relationship is relatively robust over time, with the exception of Unemployment.", fig.height = 8.5, fig.width = 6, warning = FALSE, fig.pos = 'h'}
-#(p1 / p2 / (p3 + ylab("Two-party preferred vote (%)                     ")) / p4 / p5 / p6)
+(p1 / p2 / (p3 + ylab("Two-party preferred vote (%)                     ")) / p4 / p5 / p6)
 ```
@@
 Most relationships are relatively stable over elections, except OtherLanguage and Education.", fig.height = 8.5,  fig.width = 6, warning = FALSE, fig.pos = 'h'}
-#(p7 / p8 / (p9 + ylab("Two-party preferred vote (%)                     ")) / p10 / p11 / p12 / p13)
+(p7 / p8 / (p9 + ylab("Two-party preferred vote (%)                     ")) / p10 / p11 / p12 / p13)
 ```

### 2019-05-05 elections_conversation.tex
Updated to the start of the graphics

@@
 \usepackage{sourcecodepro,sourceserifpro}
-\usepackage{float,bm}
-\usepackage{setspace}
+\usepackage{float,bm,setspace}
 \setstretch{1.2}
@@
 
-A simple way to measure voting patterns is to consider the two-party preferred (2PP) vote, which is based on the tally of preferences for the Labour and Liberal candidates, ignoring all other candidates.
-By convention, this is recorded as a percentage preference in favour of the Liberal party -- so a 2PP value of 45\% indicates that 45\% of voters preferred a Liberal candidate and 55\% of voters preferred a Labour candidate.
+A simple way to measure voting patterns is to consider the two-party preferred (2PP) vote, which is based on the tally of preferences for the Labor and Liberal candidates, ignoring all other candidates.
+By convention, this is recorded as a percentage preference in favour of the Liberal party --- for example, a 2PP value of 45\% indicates that 45\% of voters ranked the Liberal candidate higher than the Labor candidate, while the remaining 55\% ranked them in the reverse order.
 
@@
 
+\hypertarget{modelling}{%
+\section{2PP Modelling}\label{modelling}}
+
 Many of the socio-demographic variables have changing scales over the years.
 For example, inflation-adjusted median rental prices increased across almost all electorates, with median rent of 200 dollars per week placing an electorate in the 90th percentile in 2001, but only the 30th percentile in 2016.
-In order for socio-demographic effects to be comparable across years, all explanatory variables are standardized to have mean zero and variance one within each election year.
+In order for socio-demographic effects to be comparable across years, all explanatory variables are standardized so they are comparable across years.
 By standardizing, each variable is reported as a relative measure compared to all other electorates in the same year.
 
-\hypertarget{modelling}{%
-\section{Modelling}\label{modelling}}
+There are dozens of socio-demographic variables available in the Censuses, with many variables representing similar information about an electorate.
+So we combined some variables to avoid redundant information.
+For example, our ``Incomes'' variable is a combination of median personal income, household income and family income.
+In this way, we reduced the data set to include 30 socio-demographic variables, listed below.
 
-There are dozens of socio-demographic variables available in each Census, with many variables representing similar information about an electorate.
-To reduce the numbers of variables, we combine variables with similar information.
-For example, a potential group would be variables relating to electoral incomes --- median family, household and personal incomes.
-To determine which variables covary, principal component analysis is used on a combined dataset of socio-demographic variables from all six elections.
-It is appropriate to compute principal components in this way because when computed separately for each election, scree plots level off after four components and the loadings of the first four components are similar across the elections.
+\begin{itemize}
+\tightlist
+\item
+  ``AusCitizen''
+\item
+  ``BornAsia''
+\item
+  ``Born\_MidEast''
+\item
+  ``BornSEEuro''
+\item
+  ``Born\_UK''
+\item
+  ``BornElsewhere''
+\item
+  ``Buddhism''
+\item
+  ``Christianity''
+\item
+  ``CurrentlyStudying''
+\item
+  ``DeFacto''
+\item
+  ``DiffAddress''
+\item
+  ``Distributive''
+\item
+  ``Education''
+\item
+  ``Extractive''
+\item
+  ``FamHouseSize''
+\item
+  ``Incomes''
+\item
+  ``Indigenous''
+\item
+  ``Islam''
+\item
+  ``Judaism''
+\item
+  ``ManagerAdmin''
+\item
+  ``Married''
+\item
+  ``MedianAge''
+\item
+  ``NoReligion''
+\item
+  ``OneParentHouse''
+\item
+  ``OtherLanguage''
+\item
+  ``PropertyOwned''
+\item
+  ``RentLoanPrice''
+\item
+  ``SocialServ''
+\item
+  ``Transformative''
+\item
+  ``Unemployment''
+\end{itemize}
 
-Only the first four principal components from the combined dataset are considered, as the scree plot corresponding to the combined dataset levels off after the fourth component.
-Variables that have a large loading in a particular component are deemed to covary, with a loading with magnitude greater than 0.15 being considered large.
-Six factors are created using this criteria.
-These are:
-\texttt{Incomes} (median personal, household and family incomes);
-\texttt{Unemployment} (unemployment and labor force participation rates);
-\texttt{ProportyOwned} (rates of housing ownership, mortgages, renting and government housing);
-\texttt{RentLoanPrice} (median rental and loan repayments);
-\texttt{FamHouseSize} (average household size, ratio of people to families and household makeup (single person, couple with kids and couple without kids); and
-\texttt{Education} (high school and university qualifications, jobs requiring higher levels of education as well as vocational course completions and jobs that do not require higher education levels, such as laborer or tradesperson).
-For each of these groups, variables with positive loadings are added and those with negative loadings are subtracted to create a factor.
-After computing these sums, each factor is standardized to have mean zero and variance one, within each election.
-
-There are \(p=30\) variables in the resultant predictor set, with all of these used in the regression for each election.
-
-\hypertarget{regression-incorporating-spatially-dependent-errors}{%
-\subsection{Regression incorporating spatially dependent errors}\label{regression-incorporating-spatially-dependent-errors}}
-
-An identical model specification is used for each of the six elections, with each election modelled separately.
-This allows the socio-demographic effects to be estimated separately for each election year, facilitating analysis of temporal changes in variable effects.
-This approach is preferable to using a single longitudinal model because it avoids any concerns of undue bias stemming from incorrectly imposed time-varying restrictions on any variable.
-Without such restrictions, a pooled cross-sectional model does not yield any distinct advantage over separate cross-sections.
-The panel approach is avoided because of how frequently electoral boundaries change, noting that electorates with the same name across elections are not guaranteed to represent the same geographical region.
-Therefore any fixed or random effects models would be difficult to estimate without implementing consistent boundaries, which would require further imputation.
-
-For each cross-section, let the response \(\bm{y}\) be the vector two-party preferred vote in favour of the Liberal party; for example, \(y_i = 70\) represents a 70\% preference for Liberal, 30\% for Labor, in electorate \(i\).
-Although \(y_i\) lies in the interval \((0,100)\), observed values are never close to 0 or 100 (minimum \(24.05 \%\) and maximum \(74.90 \%\)), so there is no need to formally impose the constraint of \(y_i \in [0,100]\).
-Furthermore, the responses are found to be spatially correlated in each election (Moran's I test, \(p \le 7\cdot10^{-15}\)).
-This is not surprising as electorates are aggregate spatial units, and hence the spatial structure of the data must be modelled appropriately.
-
-The spatial error model \autocite{Anselin88} is chosen because it captures spatial heterogeneity by incorporating a spatially structured random effect vector \autocite{LeSage2009}.
-In this context, the random effect can be thought of as capturing the unobserved political climate in each electorate, where this climate is correlated with the climate in neighbouring electorates, under the assumption that the climate is independent of electoral socio-demographics.
-
-\begin{table}[!htbp] \centering    \caption{Estimated spatial error model parameters (standard errors) for each of the six election years.}    \label{tab:printmodels}  \scriptsize\tabcolsep=0.35cm \begin{tabular}{@{\extracolsep{1pt}}lrrrrrr} \\[-1.8ex]\hline  \hline \\[-1.8ex]  \\[-1.8ex] & \multicolumn{6}{c}{\vspace{-0.5cm}} \\   & 2001 & 2004 & 2007 & 2010 & 2013 & 2016 \\  \hline \\[-1.8ex]  $\rho$ & 0.46\rlap{$^{***}$} & 0.29\rlap{$^{*}$} & 0.24 & 0.19 & 0.27\rlap{$^{*}$} & 0.50\rlap{$^{***}$}\\  & (0.15\rlap{)} & (0.17\rlap{)} & (0.17\rlap{)} & (0.16\rlap{)} & (0.16\rlap{)} & (0.17\rlap{)}\\  AusCitizen & $-$3.13 & $-$2.64 & $-$2.53 & $-$0.08 & $-$3.40 & $-$1.80 \\    & (2.26\rlap{)} & (2.43\rlap{)} & (2.34\rlap{)} & (2.79\rlap{)} & (2.76\rlap{)} & (2.71\rlap{)} \\    BornAsia & 2.22 & $-$0.95 & $-$1.60 & $-$6.83\rlap{$^{**}$} & $-$3.03 & $-$0.55 \\    & (2.18\rlap{)} & (2.44\rlap{)} & (2.19\rlap{)} & (2.73\rlap{)} & (2.71\rlap{)} & (2.17\rlap{)} \\    Born\_MidEast & $-$1.15 & $-$1.59 & $-$2.01\rlap{$^{*}$} & $-$2.03 & $-$0.92 & $-$1.44 \\    & (1.07\rlap{)} & (1.20\rlap{)} & (1.11\rlap{)} & (1.27\rlap{)} & (1.24\rlap{)} & (1.13\rlap{)} \\    BornSEEuro & $-$3.21\rlap{$^{**}$} & $-$4.24\rlap{$^{***}$} & $-$3.61\rlap{$^{***}$} & $-$4.14\rlap{$^{***}$} & $-$3.69\rlap{$^{***}$} & $-$2.72\rlap{$^{***}$} \\    & (1.42\rlap{)} & (1.46\rlap{)} & (1.02\rlap{)} & (1.19\rlap{)} & (1.07\rlap{)} & (0.97\rlap{)} \\    Born\_UK & 0.25 & $-$0.07 & 0.34 & 0.56 & $-$0.09 & $-$1.32 \\    & (1.00\rlap{)} & (0.98\rlap{)} & (0.90\rlap{)} & (1.07\rlap{)} & (1.04\rlap{)} & (1.04\rlap{)} \\    BornElsewhere & $-$5.04 & $-$4.91 & $-$4.13 & 2.35 & $-$5.23 & $-$4.14 \\    & (3.30\rlap{)} & (3.68\rlap{)} & (3.38\rlap{)} & (4.23\rlap{)} & (4.15\rlap{)} & (3.97\rlap{)} \\    Buddhism & $-$0.49 & $-$0.17 & $-$1.37 & $-$0.83 & $-$0.12 & $-$1.60 \\    & (1.39\rlap{)} & (1.61\rlap{)} & (1.61\rlap{)} & (1.80\rlap{)} & (1.68\rlap{)} & (1.56\rlap{)} \\    Christianity & $-$2.48 & $-$1.23 & 0.38 & 0.50 & 2.41 & 1.68 \\    & (1.73\rlap{)} & (1.85\rlap{)} & (1.83\rlap{)} & (1.99\rlap{)} & (1.85\rlap{)} & (1.78\rlap{)} \\    CurrentlyStudying & $-$2.19\rlap{$^{**}$} & $-$0.13 & 2.06\rlap{$^{*}$} & 2.12\rlap{$^{*}$} & 1.15 & $-$0.16 \\    & (0.99\rlap{)} & (1.13\rlap{)} & (1.17\rlap{)} & (1.25\rlap{)} & (1.26\rlap{)} & (1.18\rlap{)} \\    DeFacto & $-$6.44\rlap{$^{***}$} & $-$5.37\rlap{$^{**}$} & $-$6.43\rlap{$^{***}$} & $-$8.07\rlap{$^{***}$} & $-$6.56\rlap{$^{**}$} & $-$8.53\rlap{$^{***}$} \\    & (1.87\rlap{)} & (2.48\rlap{)} & (2.31\rlap{)} & (3.06\rlap{)} & (3.11\rlap{)} & (2.83\rlap{)} \\    DiffAddress & 3.88\rlap{$^{***}$} & 5.06\rlap{$^{***}$} & 4.22\rlap{$^{***}$} & 5.57\rlap{$^{***}$} & 3.53\rlap{$^{*}$} & 5.67\rlap{$^{***}$} \\    & (0.94\rlap{)} & (1.12\rlap{)} & (0.99\rlap{)} & (1.76\rlap{)} & (1.91\rlap{)} & (1.60\rlap{)} \\    Distributive & 1.27 & 2.01\rlap{$^{*}$} & 1.36 & 1.57 & 2.10\rlap{$^{*}$} & 1.20 \\    & (1.12\rlap{)} & (1.21\rlap{)} & (1.13\rlap{)} & (1.34\rlap{)} & (1.27\rlap{)} & (1.21\rlap{)} \\    Education & 1.08 & 0.52 & $-$5.52\rlap{$^{*}$} & $-$4.08 & $-$4.44 & $-$7.07\rlap{$^{**}$} \\    & (2.38\rlap{)} & (3.12\rlap{)} & (3.27\rlap{)} & (3.95\rlap{)} & (3.78\rlap{)} & (3.55\rlap{)} \\    Extractive & 4.83\rlap{$^{***}$} & 5.45\rlap{$^{***}$} & 5.37\rlap{$^{***}$} & 7.31\rlap{$^{***}$} & 6.71\rlap{$^{***}$} & 7.43\rlap{$^{***}$} \\    & (1.48\rlap{)} & (1.42\rlap{)} & (1.36\rlap{)} & (1.56\rlap{)} & (1.47\rlap{)} & (1.39\rlap{)} \\    FamHouseSize & $-$0.16 & 0.87 & $-$2.40 & $-$2.53 & $-$3.26 & $-$2.91 \\    & (2.19\rlap{)} & (2.72\rlap{)} & (2.69\rlap{)} & (3.25\rlap{)} & (3.28\rlap{)} & (2.90\rlap{)} \\    Incomes & 4.36\rlap{$^{**}$} & 5.03\rlap{$^{*}$} & 9.45\rlap{$^{***}$} & 7.09\rlap{$^{**}$} & 7.97\rlap{$^{***}$} & 12.20\rlap{$^{***}$} \\    & (1.77\rlap{)} & (2.66\rlap{)} & (2.75\rlap{)} & (3.25\rlap{)} & (2.92\rlap{)} & (2.75\rlap{)} \\    Indigenous & 2.91\rlap{$^{*}$} & 1.97 & 2.48 & 2.84 & 0.67 & $-$0.05 \\    & (1.68\rlap{)} & (1.95\rlap{)} & (1.75\rlap{)} & (2.16\rlap{)} & (2.14\rlap{)} & (2.00\rlap{)} \\    Islam & $-$0.92 & $-$0.97 & $-$0.54 & $-$2.50 & $-$0.82 & $-$0.95 \\    & (1.22\rlap{)} & (1.36\rlap{)} & (1.27\rlap{)} & (1.52\rlap{)} & (1.42\rlap{)} & (1.34\rlap{)} \\    Judaism & 1.88\rlap{$^{*}$} & 1.78 & 2.66\rlap{$^{***}$} & 1.97\rlap{$^{*}$} & 2.74\rlap{$^{**}$} & 1.65\rlap{$^{*}$} \\    & (1.05\rlap{)} & (1.13\rlap{)} & (1.01\rlap{)} & (1.15\rlap{)} & (1.10\rlap{)} & (1.00\rlap{)} \\    ManagerAdmin & 2.06\rlap{$^{***}$} & 3.32\rlap{$^{***}$} & 6.00\rlap{$^{***}$} & 5.47\rlap{$^{***}$} & 5.04\rlap{$^{***}$} & 5.78\rlap{$^{***}$} \\    & (0.71\rlap{)} & (0.93\rlap{)} & (0.90\rlap{)} & (1.08\rlap{)} & (1.03\rlap{)} & (1.06\rlap{)} \\    Married & 0.44 & 0.11 & $-$1.22 & $-$0.22 & 0.91 & $-$2.34 \\    & (2.31\rlap{)} & (2.96\rlap{)} & (2.83\rlap{)} & (3.15\rlap{)} & (3.03\rlap{)} & (2.81\rlap{)} \\    MedianAge & 2.32\rlap{$^{*}$} & 4.96\rlap{$^{***}$} & 3.66\rlap{$^{**}$} & 4.00\rlap{$^{*}$} & 2.30 & 2.87 \\    & (1.32\rlap{)} & (1.65\rlap{)} & (1.81\rlap{)} & (2.26\rlap{)} & (2.08\rlap{)} & (1.79\rlap{)} \\    NoReligion & $-$1.57 & $-$0.92 & 0.56 & $-$0.30 & 1.02 & 1.31 \\    & (1.59\rlap{)} & (1.71\rlap{)} & (1.73\rlap{)} & (1.92\rlap{)} & (1.94\rlap{)} & (2.04\rlap{)} \\    OneParentHouse & $-$1.73 & $-$0.45 & $-$0.75 & $-$1.46 & $-$0.77 & $-$0.74 \\    & (1.36\rlap{)} & (1.59\rlap{)} & (1.49\rlap{)} & (1.69\rlap{)} & (1.57\rlap{)} & (1.47\rlap{)} \\    OtherLanguage & $-$0.44 & 5.92 & 9.98\rlap{$^{**}$} & 11.24\rlap{$^{**}$} & 9.00\rlap{$^{*}$} & 9.84\rlap{$^{**}$} \\    & (3.22\rlap{)} & (4.16\rlap{)} & (3.91\rlap{)} & (4.76\rlap{)} & (4.66\rlap{)} & (4.44\rlap{)} \\    PropertyOwned & $-$0.46 & $-$0.53 & 0.67 & $-$0.94 & $-$0.48 & 1.41 \\    & (1.37\rlap{)} & (1.50\rlap{)} & (1.43\rlap{)} & (1.76\rlap{)} & (1.67\rlap{)} & (1.50\rlap{)} \\    RentLoanPrice & $-$1.57 & $-$3.32\rlap{$^{*}$} & $-$4.01\rlap{$^{**}$} & $-$0.97 & $-$0.70 & $-$0.89 \\    & (1.49\rlap{)} & (1.76\rlap{)} & (1.67\rlap{)} & (2.07\rlap{)} & (2.07\rlap{)} & (2.16\rlap{)} \\    SocialServ & 2.51\rlap{$^{*}$} & 1.65 & 2.47\rlap{$^{*}$} & 2.53\rlap{$^{*}$} & 2.35\rlap{$^{*}$} & 4.45\rlap{$^{***}$} \\    & (1.33\rlap{)} & (1.41\rlap{)} & (1.29\rlap{)} & (1.47\rlap{)} & (1.32\rlap{)} & (1.19\rlap{)} \\    Transformative & 3.24\rlap{$^{**}$} & 4.73\rlap{$^{***}$} & 4.84\rlap{$^{***}$} & 4.46\rlap{$^{**}$} & 3.56\rlap{$^{**}$} & 4.58\rlap{$^{***}$} \\    & (1.55\rlap{)} & (1.78\rlap{)} & (1.74\rlap{)} & (1.98\rlap{)} & (1.78\rlap{)} & (1.53\rlap{)} \\    Unemployment & $-$2.45\rlap{$^{*}$} & $-$3.07\rlap{$^{*}$} & 0.29 & 0.08 & 1.67 & 2.79\rlap{$^{**}$} \\    & (1.40\rlap{)} & (1.63\rlap{)} & (1.51\rlap{)} & (1.76\rlap{)} & (1.51\rlap{)} & (1.37\rlap{)} \\    Constant & 50.81\rlap{$^{***}$} & 52.60\rlap{$^{***}$} & 47.31\rlap{$^{***}$} & 49.93\rlap{$^{***}$} & 53.51\rlap{$^{***}$} & 50.49\rlap{$^{***}$} \\    & (0.71\rlap{)} & (0.58\rlap{)} & (0.52\rlap{)} & (0.57\rlap{)} & (0.58\rlap{)} & (0.80\rlap{)} \\   \hline \\[-1.8ex]  Observations & 150 & 150 & 150 & 150 & 150 & 150 \\  Residual Standard Error (GLS) & 4.69 & 5.04 & 4.79 & 5.63 & 5.18 & 4.88\\ \hline  \hline \\[-1.8ex]  \textit{Note:}  & \multicolumn{6}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\  \end{tabular}  \end{table}
-
-Spatial weights are calculated in accordance with the assumption that an electorate is equally correlated with any electorate that shares a part of its boundary.
-Let \(\rho\) be the spatial autoregressive coefficient, \(\bm{v}\) be a spherical error term, \(\bm{W}\) be a matrix of spatial weights (containing information about the neighbouring regions), \(\bm{X}\) be a matrix of socio-demographic covariates, \(\bm{\beta}\) be a vector of regression coefficients and \(\bm{a}\) be a spatially structured random effect vector.
-\[
-  \bm{y} = \bm{X} \bm{\beta} + \bm{a},
-\]
-and
-\[
-  \bm{a} = \rho \bm{W} \bm{a} + \bm{v},
-\]
-where \(\bm{v} \sim N(\bm{0}, \sigma^2 \bm{I_n})\), and hence
-\[
-  \bm{y} = \bm{X} \bm{\beta} + (\bm{I}_n-\rho \bm{W})^{-1}\bm{v}.
-\]
-Estimation of the above spatial error model is undertaken using feasible generalized least squares.
-
-Table \ref{tab:printmodels} details the estimated model coefficients and their estimated standard errors, for each of the six elections.
-An interpretation of these estimated values is provided in the next section.
-
-\hypertarget{results}{%
-\section{Results}\label{results}}
-
-\hypertarget{spatial-autoregressive-parameter}{%
-\subsection{Spatial autoregressive parameter}\label{spatial-autoregressive-parameter}}
-
-The spatial autoregressive coefficient \(\rho\) is positive and significant in only the 2001 and 2016 elections (Figure \ref{fig:rhovis}), meaning that in these elections, the political climate of an electorate appears to be affected by the attitudes of its neighbours.
-Conversely, in the other four elections, the spatial effect weakens to become insignificant.
-In these years, it appears that the spatial component does not explain anything not already explained by the electoral socio-demographics, meaning electorates effectively voted independently.
-
-\begin{figure}[h]
-
-{\centering \includegraphics[height=0.2\textheight,]{elections_conversation_files/figure-latex/rhovis-1}
-
-}
-
-\caption{Estimates of the spatial autoregressive parameter for each of the six elections, reported with their individual 95\% confidence intervals.
-Only in 2001 and 2016 is there a significant spatial component.}\label{fig:rhovis}
-\end{figure}
+Each election was modelled separately, to allow us to see any changes over time, and to account for changing electorate boundaries.
 
@@
 \subsection{Influential socio-demographics}\label{influential-socio-demographics}}
+
+\begin{figure}[h]
+
+{\centering \includegraphics{elections_conversation_files/figure-latex/partresplot1-1}
+
+}
+
+\caption{Partial residual plots by election year for a selection of predictors.
+Linear model with 95\% confidence bands overlaid.
+Most predictors have a positive relationship: the larger the value the more likely the electorate preferences the Coalition.
+The relationship is relatively robust over time, with the exception of Unemployment.}\label{fig:partresplot1}
+\end{figure}
+
+\begin{figure}[h]
+
+{\centering \includegraphics{elections_conversation_files/figure-latex/partresplot2-1}
+
+}
+
+\caption{Partial residual plots by election year for a selection of predictors.
+Linear model with 95\% confidence bands overlaid.
+Several predictors have a negative relationship: with larger values indicating the electorate more likely preferences Labor.
+Most relationships are relatively stable over elections, except OtherLanguage and Education.}\label{fig:partresplot2}
+\end{figure}
 

### 2019-05-06 elections_conversation.Rmd
Updating on train

@@
 biblio-style: authoryear-comp
-fontsize: 12pt
+fontsize: 11pt
 papersize: a4
@@
   bookdown::pdf_document2:
-    latex_engine: xelatex
     fig_caption: yes
@@
   echo = FALSE,
-  fig.pos = "h",
+  fig.pos = "H",
   fig.align = "center",
@@
 library(forcats)
-library(ggthemes)
+#library(ggthemes)
 library(knitr)
@@
 library(spdep)
-library(patchwork)
+#library(patchwork)
 #library(gridExtra)
@@
 
-It is also interesing to see how this might have changed over time.
+It is also interesting to see how this might have changed over time.
 For example, if wealth was a good predictor of voting patterns in the 2001 election, is it still a good predictor in 2019?
@@
   geom_segment(x = 2016.1, xend = 2016.1, y = 1.4, yend = 0, color = "#F8766D") +
-  theme(
-    text = element_text(size = 10),
-    axis.text.x = element_text(size = 6),
-    axis.text.y = element_blank(),
-    axis.title.y = element_blank(),
-    plot.background = element_blank(),
-    axis.ticks.y = element_blank(),
-    panel.background = element_blank()
-  ) +
   scale_x_discrete(limits = c(2001:2016)) +
@@
 
+# 2PP Modelling {#modelling}
+
 A simple way to measure voting patterns is to consider the two-party preferred (2PP) vote, which is based on the tally of preferences for the Labor and Liberal candidates, ignoring all other candidates.
@@
 
-# 2PP Modelling {#modelling}
-
 Many of the socio-demographic variables have changing scales over the years.
 For example, inflation-adjusted median rental prices increased across almost all electorates, with median rent of 200 dollars per week placing an electorate in the 90th percentile in 2001, but only the 30th percentile in 2016.
-In order for socio-demographic effects to be comparable across years, all explanatory variables are standardized so they are comparable across years.
-By standardizing, each variable is reported as a relative measure compared to all other electorates in the same year.
+In order for socio-demographic effects to be comparable across years, all socio-demographic variables were standardized.
 
@@
 For example, our "Incomes" variable is a combination of median personal income, household income and family income.
-In this way, we reduced the data set to include 30 socio-demographic variables, listed below.
-
- * "AusCitizen"
- * "BornAsia"
- * "Born_MidEast"
- * "BornSEEuro"
- * "Born_UK"
- * "BornElsewhere"
- * "Buddhism"
- * "Christianity"
- * "CurrentlyStudying"
- * "DeFacto"
- * "DiffAddress"
- * "Distributive"
- * "Education"
- * "Extractive"
- * "FamHouseSize"
- * "Incomes"
- * "Indigenous"
- * "Islam"
- * "Judaism"
- * "ManagerAdmin"
- * "Married"
- * "MedianAge"
- * "NoReligion"
- * "OneParentHouse"
- * "OtherLanguage"
- * "PropertyOwned"
- * "RentLoanPrice"
- * "SocialServ"
- * "Transformative"
- * "Unemployment"
 
 Each election was modelled separately, to allow us to see any changes over time, and to account for changing electorate boundaries.
+In this article, we highlight the variables with the strongest relationship to the two-party preferred vote, or which have had substantial changes over time.
+The full analysis is available at https://robjhyndman.com/publications/elections/.
 
@@
 
-```{r plotintercept, fig.cap = "Estimated intercept for each election, which represents the two-party preferred vote for an electorate with mean characteristics.", out.height="20%", out.extra = ''}
+```{r plotintercept, fig.cap = "Estimated intercept for each election, which represents the two-party preferred vote for an electorate with average characteristics.", fig.pos="H"}
 intercept_df <- bind_rows(
@@
   geom_hline(aes(yintercept = 50), alpha = 0.5, size = 1) +
-  labs(x = "Election year", y = "Estimate of intercept") +
-  guides(col = FALSE) +
+  labs(x = "Election year", y = "Estimate of 2PP vote in average electorate") +
   coord_fixed(ratio = 0.2) +
-  lims(y = c(45, 55)) +
-  theme_bw()
+  ylim(45.3, 54.7)
 ```
 
-Since all socio-demographics have been standardized to have a mean of zero and a variance of one, the intercept in each model can be interpreted as the estimated two-party preferred vote for an electorate with mean characteristics[^3].
+Since all socio-demographics have been standardized, the intercept in each model can be interpreted as the estimated two-party preferred vote for an "average" electorate.
 Figure \@ref(fig:plotintercept) shows that the baseline of party preference has varied over the elections, with the biggest swing occurring in the 2007 election where the mean electorate shifted more than five percentage points in favour of the Labor party.
-
-[^3]: Mean of all variables aside from Judaism, Indigenous, Islam and Buddhism, where it assumes the mean of the log value.
+The dots represent the estimated 2PP value, and the lines indicate a 95% confidence interval providing a guide to the uncertainty in the estimate.
 
@@
 ```{r prepplot, echo = FALSE}
-p1 <- grid_visreg("Extractive", top = TRUE)
-p2 <- grid_visreg("DiffAddress")
-p3 <- grid_visreg("ManagerAdmin")
-p4 <- grid_visreg("Transformative")
-p5 <- grid_visreg("Incomes")
-p6 <- grid_visreg("Unemployment")
-
-p7 <- grid_visreg("MedianAge", top = TRUE)
-p8 <- grid_visreg("Judaism")
-p9 <- grid_visreg("OtherLanguage")
-temp <- grid_visreg("OtherLanguage")
-p10 <- grid_visreg("BornAsia")
-p11 <- grid_visreg("Education")
-p12 <- grid_visreg("BornSEEuro")
-p13 <- grid_visreg("DeFacto")
+grid_visreg("Extractive", top = TRUE)
+grid_visreg("DiffAddress")
+grid_visreg("ManagerAdmin")
+grid_visreg("Transformative")
+grid_visreg("Incomes")
+grid_visreg("Unemployment")
+grid_visreg("MedianAge", top = TRUE)
+grid_visreg("Judaism")
+grid_visreg("OtherLanguage")
+grid_visreg("OtherLanguage")
+grid_visreg("BornAsia")
+grid_visreg("Education")
+grid_visreg("BornSEEuro")
+grid_visreg("DeFacto")
 ```
 
-```{r partresplot1, fig.cap = "Partial residual plots by election year for a selection of predictors.
+Partial residual plots by election year for a selection of predictors.
 Linear model with 95\\% confidence bands overlaid.
 Most predictors have a positive relationship: the larger the value the more likely the electorate preferences the Coalition.
-The relationship is relatively robust over time, with the exception of Unemployment.", fig.height = 8.5, fig.width = 6, warning = FALSE, fig.pos = 'h'}
-(p1 / p2 / (p3 + ylab("Two-party preferred vote (%)                     ")) / p4 / p5 / p6)
-```
+The relationship is relatively robust over time, with the exception of Unemployment.
 
-```{r partresplot2, fig.cap = "Partial residual plots by election year for a selection of predictors.
-Linear model with 95\\% confidence bands overlaid.
 Several predictors have a negative relationship: with larger values indicating the electorate more likely preferences Labor.
-Most relationships are relatively stable over elections, except OtherLanguage and Education.", fig.height = 8.5,  fig.width = 6, warning = FALSE, fig.pos = 'h'}
-(p7 / p8 / (p9 + ylab("Two-party preferred vote (%)                     ")) / p10 / p11 / p12 / p13)
-```
+Most relationships are relatively stable over elections, except OtherLanguage and Education.
 
@@
 Furthermore, of the variables relating to religion, only Judaism shows a consistent effect, with electorates with relatively large Jewish populations more likely to vote Liberal.
-
-### A note on similar variables
-
-Many of the Census variables represent similar information, which is why factors were created and some variables were removed.
-However, some variables remain which are closely related.
-For example, an electorate's income level (via `Incomes`) is likely to be related to electoral unemployment and labor force participation (via `Unemployment`).
-In 2001, the coefficient estimate for `Unemployment` is negative but not significant, whilst the `Incomes` variables is significant.
-If the `Incomes` variable is removed from the model in 2001, `Unemployment` absorbs the negative effect, becoming significant ($p = 0.0056$).
-
-```{r removeunemp, eval = FALSE}
-unemp01 <- my_fgls(full_formula,
-  my_data = model_df %>% filter(year == "2001") %>% dplyr::select(-Incomes),
-  sp_weights = sp_weights_01
-)
-
-summary(unemp01)$tTable["Unemployment",]
-```
 
@@
 The authors would like to sincerely thank Anthony Ebert, Heike Hofmann, Thomas Lumley, Ben Marwick, Carson Sievert, Mingzhu Sun, Dilini Talagala, Nicholas Tierney, Nathaniel Tomasetti, Earo Wang and Fang Zhou, all of whom have contributed to the `eechidna` package.
+
+# From Emil
+
+In terms of structure, start with a brief paragraph or two on what the research is, and what the reader can expect to find below.
+
+Then, break the article into small sections for each of the factors that have an effect (a short paragraph or two for each factor).
+And it would be great to include examples of specific seats where the relationship was strong, or interesting for other reasons.
+
+So the sections will be:
+
+- Age
+- Income
+- Unemployment
+- Industry and type of work
+- Education
+- Diversity
+- Household mobility
+- Relationships
+
+An example of this structure can be seen here:
+https://theconversation.com/confused-about-aged-care-in-the-home-these-10-charts-explain-how-it-works-113923
+
+Questions on particular factors:
+
+Incomes:
+- Was this correlation stronger in 2016? (and increasingly influential?)
+
+Household mobility
+- In the report, you say: "“Having controlled for characteristics of house ownership and rental prices”, whis effect is somewhat surprising”.
+Why is this surprising to you?
+(I would have expected this to support Labor more)
+
+Age
+- Would it be possible to show how the correlation changes for different age brackets (ie. younger voters = more Labor, and this swings more to Liberal as the age brackets gets older)?
+- If not, can you go into more detail specific age brackets? Eg. Is it the over-65s that skew Liberal?
+
+Diversity
+- Migrant population: Would it possible to break this out into particular groups?
+eg.
+Seats with higher Chinese population correlates with Liberal, seats with higher SEAsia pop correlates with Labor, etc?
+- The same for languages?

### 2019-05-06 elections_conversation.tex
Updating on train

@@
-\documentclass[12pt,a4paper,]{article}
+\documentclass[11pt,a4paper,]{article}
 \usepackage{lmodern}
@@
 \usepackage[no-weekday]{eukdate}
-\usepackage{sourcecodepro,sourceserifpro}
+%\usepackage{sourceserifpro}
+%\usepackage[scaled=0.86]{DejaVuSansMono}
 \usepackage{float,bm,setspace}
@@
 \captionsetup[table]{style=italic,format=hang,singlelinecheck=true}
-
-\let\origfigure\figure
-\let\endorigfigure\endfigure
-\renewenvironment{figure}[1][2] {
-    \expandafter\origfigure\expandafter[htbp]
-} {
-    \endorigfigure
-}
-
-\let\origtable\table
-\let\endorigtable\endtable
-\renewenvironment{table}[1][2] {
-    \expandafter\origtable\expandafter[htbp]
-} {
-    \endorigtable
-}
 
@@
 
-\begin{figure}[h]
+\begin{figure}[H]
 
@@
 
+\hypertarget{modelling}{%
+\section{2PP Modelling}\label{modelling}}
+
 A simple way to measure voting patterns is to consider the two-party preferred (2PP) vote, which is based on the tally of preferences for the Labor and Liberal candidates, ignoring all other candidates.
@@
 
-\hypertarget{modelling}{%
-\section{2PP Modelling}\label{modelling}}
-
 Many of the socio-demographic variables have changing scales over the years.
 For example, inflation-adjusted median rental prices increased across almost all electorates, with median rent of 200 dollars per week placing an electorate in the 90th percentile in 2001, but only the 30th percentile in 2016.
-In order for socio-demographic effects to be comparable across years, all explanatory variables are standardized so they are comparable across years.
-By standardizing, each variable is reported as a relative measure compared to all other electorates in the same year.
+In order for socio-demographic effects to be comparable across years, all socio-demographic variables were standardized.
 
@@
 For example, our ``Incomes'' variable is a combination of median personal income, household income and family income.
-In this way, we reduced the data set to include 30 socio-demographic variables, listed below.
-
-\begin{itemize}
-\tightlist
-\item
-  ``AusCitizen''
-\item
-  ``BornAsia''
-\item
-  ``Born\_MidEast''
-\item
-  ``BornSEEuro''
-\item
-  ``Born\_UK''
-\item
-  ``BornElsewhere''
-\item
-  ``Buddhism''
-\item
-  ``Christianity''
-\item
-  ``CurrentlyStudying''
-\item
-  ``DeFacto''
-\item
-  ``DiffAddress''
-\item
-  ``Distributive''
-\item
-  ``Education''
-\item
-  ``Extractive''
-\item
-  ``FamHouseSize''
-\item
-  ``Incomes''
-\item
-  ``Indigenous''
-\item
-  ``Islam''
-\item
-  ``Judaism''
-\item
-  ``ManagerAdmin''
-\item
-  ``Married''
-\item
-  ``MedianAge''
-\item
-  ``NoReligion''
-\item
-  ``OneParentHouse''
-\item
-  ``OtherLanguage''
-\item
-  ``PropertyOwned''
-\item
-  ``RentLoanPrice''
-\item
-  ``SocialServ''
-\item
-  ``Transformative''
-\item
-  ``Unemployment''
-\end{itemize}
 
 Each election was modelled separately, to allow us to see any changes over time, and to account for changing electorate boundaries.
+In this article, we highlight the variables with the strongest relationship to the two-party preferred vote, or which have had substantial changes over time.
+The full analysis is available at \url{https://robjhyndman.com/publications/elections/}.
 
@@
 
-\begin{figure}[h]
+\begin{figure}[H]
 
-{\centering \includegraphics[height=0.2\textheight,]{elections_conversation_files/figure-latex/plotintercept-1}
+{\centering \includegraphics{elections_conversation_files/figure-latex/plotintercept-1}
 
@@
 
-\caption{Estimated intercept for each election, which represents the two-party preferred vote for an electorate with mean characteristics.}\label{fig:plotintercept}
+\caption{Estimated intercept for each election, which represents the two-party preferred vote for an electorate with average characteristics.}\label{fig:plotintercept}
 \end{figure}
 
-Since all socio-demographics have been standardized to have a mean of zero and a variance of one, the intercept in each model can be interpreted as the estimated two-party preferred vote for an electorate with mean characteristics\footnote{Mean of all variables aside from Judaism, Indigenous, Islam and Buddhism, where it assumes the mean of the log value.}.
+Since all socio-demographics have been standardized, the intercept in each model can be interpreted as the estimated two-party preferred vote for an ``average'' electorate.
 Figure \ref{fig:plotintercept} shows that the baseline of party preference has varied over the elections, with the biggest swing occurring in the 2007 election where the mean electorate shifted more than five percentage points in favour of the Labor party.
+The dots represent the estimated 2PP value, and the lines indicate a 95\% confidence interval providing a guide to the uncertainty in the estimate.
 
@@
 
-\begin{figure}[h]
+\begin{center}\includegraphics{elections_conversation_files/figure-latex/prepplot-1} \end{center}
 
-{\centering \includegraphics{elections_conversation_files/figure-latex/partresplot1-1}
+\begin{center}\includegraphics{elections_conversation_files/figure-latex/prepplot-2} \end{center}
 
-}
+\begin{center}\includegraphics{elections_conversation_files/figure-latex/prepplot-3} \end{center}
 
-\caption{Partial residual plots by election year for a selection of predictors.
-Linear model with 95\% confidence bands overlaid.
+\begin{center}\includegraphics{elections_conversation_files/figure-latex/prepplot-4} \end{center}
+
+\begin{center}\includegraphics{elections_conversation_files/figure-latex/prepplot-5} \end{center}
+
+\begin{center}\includegraphics{elections_conversation_files/figure-latex/prepplot-6} \end{center}
+
+\begin{center}\includegraphics{elections_conversation_files/figure-latex/prepplot-7} \end{center}
+
+\begin{center}\includegraphics{elections_conversation_files/figure-latex/prepplot-8} \end{center}
+
+\begin{center}\includegraphics{elections_conversation_files/figure-latex/prepplot-9} \end{center}
+
+\begin{center}\includegraphics{elections_conversation_files/figure-latex/prepplot-10} \end{center}
+
+\begin{center}\includegraphics{elections_conversation_files/figure-latex/prepplot-11} \end{center}
+
+\begin{center}\includegraphics{elections_conversation_files/figure-latex/prepplot-12} \end{center}
+
+\begin{center}\includegraphics{elections_conversation_files/figure-latex/prepplot-13} \end{center}
+
+\begin{center}\includegraphics{elections_conversation_files/figure-latex/prepplot-14} \end{center}
+
+Partial residual plots by election year for a selection of predictors.
+Linear model with 95\textbackslash{}\% confidence bands overlaid.
 Most predictors have a positive relationship: the larger the value the more likely the electorate preferences the Coalition.
-The relationship is relatively robust over time, with the exception of Unemployment.}\label{fig:partresplot1}
-\end{figure}
+The relationship is relatively robust over time, with the exception of Unemployment.
 
-\begin{figure}[h]
-
-{\centering \includegraphics{elections_conversation_files/figure-latex/partresplot2-1}
-
-}
-
-\caption{Partial residual plots by election year for a selection of predictors.
-Linear model with 95\% confidence bands overlaid.
 Several predictors have a negative relationship: with larger values indicating the electorate more likely preferences Labor.
-Most relationships are relatively stable over elections, except OtherLanguage and Education.}\label{fig:partresplot2}
-\end{figure}
+Most relationships are relatively stable over elections, except OtherLanguage and Education.
 
@@
 Furthermore, of the variables relating to religion, only Judaism shows a consistent effect, with electorates with relatively large Jewish populations more likely to vote Liberal.
-
-\hypertarget{a-note-on-similar-variables}{%
-\subsubsection{A note on similar variables}\label{a-note-on-similar-variables}}
-
-Many of the Census variables represent similar information, which is why factors were created and some variables were removed.
-However, some variables remain which are closely related.
-For example, an electorate's income level (via \texttt{Incomes}) is likely to be related to electoral unemployment and labor force participation (via \texttt{Unemployment}).
-In 2001, the coefficient estimate for \texttt{Unemployment} is negative but not significant, whilst the \texttt{Incomes} variables is significant.
-If the \texttt{Incomes} variable is removed from the model in 2001, \texttt{Unemployment} absorbs the negative effect, becoming significant (\(p = 0.0056\)).
 
@@
 
+\hypertarget{from-emil}{%
+\section{From Emil}\label{from-emil}}
+
+In terms of structure, start with a brief paragraph or two on what the research is, and what the reader can expect to find below.
+
+Then, break the article into small sections for each of the factors that have an effect (a short paragraph or two for each factor).
+And it would be great to include examples of specific seats where the relationship was strong, or interesting for other reasons.
+
+So the sections will be:
+
+\begin{itemize}
+\tightlist
+\item
+  Age
+\item
+  Income
+\item
+  Unemployment
+\item
+  Industry and type of work
+\item
+  Education
+\item
+  Diversity
+\item
+  Household mobility
+\item
+  Relationships
+\end{itemize}
+
+An example of this structure can be seen here:
+\url{https://theconversation.com/confused-about-aged-care-in-the-home-these-10-charts-explain-how-it-works-113923}
+
+Questions on particular factors:
+
+Incomes:
+- Was this correlation stronger in 2016? (and increasingly influential?)
+
+Household mobility
+- In the report, you say: ````Having controlled for characteristics of house ownership and rental prices'', whis effect is somewhat surprising''.
+Why is this surprising to you?
+(I would have expected this to support Labor more)
+
+Age
+- Would it be possible to show how the correlation changes for different age brackets (ie. younger voters = more Labor, and this swings more to Liberal as the age brackets gets older)?
+- If not, can you go into more detail specific age brackets? Eg. Is it the over-65s that skew Liberal?
+
+Diversity
+- Migrant population: Would it possible to break this out into particular groups?
+eg.
+Seats with higher Chinese population correlates with Liberal, seats with higher SEAsia pop correlates with Labor, etc?
+- The same for languages?
+
 \printbibliography

### 2019-05-06 elections_conversation.Rmd
Working on graphs

@@
 
+```{r coefficients}
+coefs <- bind_rows(
+    data.frame(summary(glsmod16)$tTable) %>% 
+      rownames_to_column("variable") %>%
+      mutate(year = 2016),
+    data.frame(summary(glsmod13)$tTable) %>% 
+      rownames_to_column("variable") %>%
+      mutate(year = 2013),
+    data.frame(summary(glsmod10)$tTable) %>% 
+      rownames_to_column("variable") %>%
+      mutate(year = 2010),
+    data.frame(summary(glsmod07)$tTable) %>% 
+      rownames_to_column("variable") %>%
+      mutate(year = 2007),
+    data.frame(summary(glsmod04)$tTable) %>% 
+      rownames_to_column("variable") %>%
+      mutate(year = 2004),
+    data.frame(summary(glsmod01)$tTable) %>% 
+      rownames_to_column("variable") %>%
+      mutate(year = 2001)
+  ) %>%
+  as_tibble() %>%
+  rename(se = `Std.Error`, estimate = Value, p = `p.value`) %>%
+  mutate(upper95 = estimate + 1.96 * se, lower95 = estimate - 1.96 * se)
+```
+
 ## Country-wide trend
@@
 ```{r plotintercept, fig.cap = "Estimated intercept for each election, which represents the two-party preferred vote for an electorate with average characteristics.", fig.pos="H"}
-intercept_df <- bind_rows(
-  data.frame(summary(glsmod16)$tTable)[1, ] %>% mutate(year = "2016"),
-  data.frame(summary(glsmod13)$tTable)[1, ] %>% mutate(year = "2013"),
-  data.frame(summary(glsmod10)$tTable)[1, ] %>% mutate(year = "2010"),
-  data.frame(summary(glsmod07)$tTable)[1, ] %>% mutate(year = "2007"),
-  data.frame(summary(glsmod04)$tTable)[1, ] %>% mutate(year = "2004"),
-  data.frame(summary(glsmod01)$tTable)[1, ] %>% mutate(year = "2001")
-) %>%
-  rename(se = `Std.Error`, estimate = Value, p = `p.value`) %>%
-  mutate(upper95 = estimate + 1.96 * se, lower95 = estimate - 1.96 * se)
-
-intercept_df %>%
+coefs %>% filter(variable=="Intercept") %>%
   ggplot() +
-  geom_point(aes(x = year, y = estimate), size = 3) +
+  geom_point(aes(x = year, y = estimate), size = 2) +
   geom_linerange(aes(x = year, ymin = lower95, ymax = upper95), size = 1.2) +
   geom_hline(aes(yintercept = 50), alpha = 0.5, size = 1) +
-  labs(x = "Election year", y = "Estimate of 2PP vote in average electorate") +
-  coord_fixed(ratio = 0.2) +
-  ylim(45.3, 54.7)
+  labs(x = "Election year", y = "2PP percentage") +
+  coord_fixed(ratio = 0.25) +
+  scale_x_continuous(breaks=seq(2001,2016,by=3), minor_breaks=NULL) +
+  scale_y_continuous(breaks=seq(46,54,by=2), minor_breaks=45:55)
 ```
@@
 
-## Influential socio-demographics
+It is important here to note the ecological fallacy: insights are being drawn at the electorate level, and cannot be inferred for another disaggregate level (e.g. individual voters).
 
-```{r prepplot, echo = FALSE}
-grid_visreg("Extractive", top = TRUE)
-grid_visreg("DiffAddress")
-grid_visreg("ManagerAdmin")
-grid_visreg("Transformative")
-grid_visreg("Incomes")
-grid_visreg("Unemployment")
-grid_visreg("MedianAge", top = TRUE)
-grid_visreg("Judaism")
-grid_visreg("OtherLanguage")
-grid_visreg("OtherLanguage")
-grid_visreg("BornAsia")
-grid_visreg("Education")
-grid_visreg("BornSEEuro")
-grid_visreg("DeFacto")
+## Age
+
+- Would it be possible to show how the correlation changes for different age brackets (ie. younger voters = more Labor, and this swings more to Liberal as the age brackets gets older)?
+- If not, can you go into more detail specific age brackets? Eg. Is it the over-65s that skew Liberal?
+
+```{r age, fig.cap = "Estimated age effect for each election, which represents the impact of electorate median age on the two-party preferred vote. Higher values indicate a ", fig.pos="H"}
+coefs %>% filter(variable=="MedianAge") %>%
+  ggplot(aes(x=year)) +
+  geom_hline(aes(yintercept = 0), col='gray', size = 1) +
+  geom_point(aes(x = year, y = estimate), size = 2) +
+  geom_linerange(aes(x = year, ymin = lower95, ymax = upper95), size = 1.2) +
+  labs(x = "Election year", y = "2PP impact") +
+  coord_fixed(ratio = 0.25) + 
+  scale_x_continuous(breaks=seq(2001,2016,by=3), minor_breaks=NULL)
 ```
 
-Partial residual plots by election year for a selection of predictors.
-Linear model with 95\\% confidence bands overlaid.
-Most predictors have a positive relationship: the larger the value the more likely the electorate preferences the Coalition.
-The relationship is relatively robust over time, with the exception of Unemployment.
+### Income and unemployment
 
-Several predictors have a negative relationship: with larger values indicating the electorate more likely preferences Labor.
-Most relationships are relatively stable over elections, except OtherLanguage and Education.
-
-To investigate the socio-demographics that have a strong effect on the two-party preferred vote, partial residual plots are used and shown in Figures \@ref(fig:partresplot1) and \@ref(fig:partresplot2).
-The partial residuals are the residuals from the fitted model with the estimated effect an individual variable added.
-These show the direction, size and significance of an estimated effect --- the slope of the prediction line matches the estimated coefficient, and the shaded region represents a 95% confidence band, computed using the method in @visreg.
-If a horizontal line can be drawn through the confidence band, then the effect is insignificant.
-The estimated intercept is also added to the partial residuals for interpretability.
-Plots for each election are faceted to compare the effects over time in Figures \@ref(fig:partresplot1) and \@ref(fig:partresplot2).
-Only socio-demographics that have a significant effect in at least one election are displayed in Figures \@ref(fig:partresplot1) and \@ref(fig:partresplot2).
-
-It is important here to note the ecological fallacy: insights are being drawn at the electorate level, and cannot be inferred for another disaggregate level (e.g. individual voters).
-
-### Income and unemployment
+```{r age, fig.cap = "Estimated age effect for each election, which represents the impact of electorate median age on the two-party preferred vote. Higher values indicate a ", fig.pos="H"}
+coefs %>% filter(variable=="MedianAge") %>%
+  ggplot(aes(x=year)) +
+  geom_hline(aes(yintercept = 0), col='gray', size = 1) +
+  geom_point(aes(x = year, y = estimate), size = 2) +
+  geom_linerange(aes(x = year, ymin = lower95, ymax = upper95), size = 1.2) +
+  labs(x = "Election year", y = "2PP impact") +
+  coord_fixed(ratio = 0.25) + 
+  scale_x_continuous(breaks=seq(2001,2016,by=3), minor_breaks=NULL)
+```
 

### 2019-05-06 elections_conversation.tex
Working on graphs

@@
 \usepackage[no-weekday]{eukdate}
-%\usepackage{sourceserifpro}
-%\usepackage[scaled=0.86]{DejaVuSansMono}
+\usepackage{sourceserifpro}
+\usepackage[scaled=0.86]{DejaVuSansMono}
 \usepackage{float,bm,setspace}
@@
 
+\begin{figure}[H]
+
+{\centering \includegraphics{elections_conversation_files/figure-latex/intercept2-1}
+
+}
+
+\caption{Estimated intercept for each election, which represents the two-party preferred vote for an electorate with average characteristics.}\label{fig:intercept2}
+\end{figure}
+
 Since all socio-demographics have been standardized, the intercept in each model can be interpreted as the estimated two-party preferred vote for an ``average'' electorate.
@@
 The dots represent the estimated 2PP value, and the lines indicate a 95\% confidence interval providing a guide to the uncertainty in the estimate.
+
+\hypertarget{age}{%
+\subsection{Age}\label{age}}
+
+\begin{itemize}
+\tightlist
+\item
+  Would it be possible to show how the correlation changes for different age brackets (ie. younger voters = more Labor, and this swings more to Liberal as the age brackets gets older)?
+\item
+  If not, can you go into more detail specific age brackets? Eg. Is it the over-65s that skew Liberal?
+\end{itemize}
+
+\begin{center}\includegraphics{elections_conversation_files/figure-latex/age-1} \end{center}
 
@@
 
-\hypertarget{age}{%
-\subsubsection{Age}\label{age}}
+\hypertarget{age-1}{%
+\subsubsection{Age}\label{age-1}}
 

### 2019-05-06 elections_conversation.Rmd
Working on graphs

@@
 
-```{r age, fig.cap = "Estimated age effect for each election, which represents the impact of electorate median age on the two-party preferred vote. Higher values indicate a ", fig.pos="H"}
-coefs %>% filter(variable=="MedianAge") %>%
+```{r income, fig.cap = "Estimated age effect for each election, which represents the impact of electorate median age on the two-party preferred vote. Higher values indicate a ", fig.pos="H"}
+coefs %>% filter(variable=="Income") %>%
   ggplot(aes(x=year)) +

### 2019-05-07 elections_conversation.Rmd
substit uniqueID for DivisionID to make code work, and revised first section

@@
 
-We often hear stereotypes about voting patterns --- people in their 20s without kids are more likely to be left wing, migrants are more conservative, wealthier people tend to favour the conservative parties, and so on.
-So we thought we would test these ideas by matching census data to election data, to see if we can identify what are the socio-demographic characteristics of an electorate that are most closely related to how they vote.
+We often hear commentary about voting patterns --- people in their 20s without kids are more likely to be left wing, migrants are more conservative, wealthier people tend to favour the conservative parties, and so on.
+These might be built on myth and stereotypes.
+There is open data available from each election and from regular national census that can be used to examine, just what are the indicators for voting tendency.
+<!-- So we thought we would test these ideas by matching census data to election data, to see if we can identify what are the socio-demographic characteristics of an electorate that are most closely related to how they vote.-->
 
-It is also interesting to see how this might have changed over time.
-For example, if wealth was a good predictor of voting patterns in the 2001 election, is it still a good predictor in 2019?
+Voting tendencies can chang over time.
+For example, if wealth was a good predictor of conservative voting in the 2001 election, is it still a good predictor in 2019?
 Australia has changed in many ways over the last two decades.
@@
   bind_rows(
-    glsmod16$my_data %>% left_join(tpp16 %>% dplyr::select(DivisionNm, DivisionID, StateAb), by = "DivisionNm"),
-    glsmod13$my_data %>% left_join(tpp13 %>% dplyr::select(DivisionNm, DivisionID, StateAb), by = "DivisionNm"),
-    glsmod10$my_data %>% left_join(tpp10 %>% dplyr::select(DivisionNm, DivisionID, StateAb), by = "DivisionNm"),
-    glsmod07$my_data %>% left_join(tpp07 %>% dplyr::select(DivisionNm, DivisionID, StateAb), by = "DivisionNm"),
-    glsmod04$my_data %>% left_join(tpp04 %>% dplyr::select(DivisionNm, DivisionID, StateAb), by = "DivisionNm"),
+    glsmod16$my_data %>% left_join(tpp16 %>% dplyr::select(DivisionNm, UniqueID, StateAb), by = "DivisionNm"),
+    glsmod13$my_data %>% left_join(tpp13 %>% dplyr::select(DivisionNm, UniqueID, StateAb), by = "DivisionNm"),
+    glsmod10$my_data %>% left_join(tpp10 %>% dplyr::select(DivisionNm, UniqueID, StateAb), by = "DivisionNm"),
+    glsmod07$my_data %>% left_join(tpp07 %>% dplyr::select(DivisionNm, UniqueID, StateAb), by = "DivisionNm"),
+    glsmod04$my_data %>% left_join(tpp04 %>% dplyr::select(DivisionNm, UniqueID, StateAb), by = "DivisionNm"),
     glsmod01$my_data %>% left_join(tpp01 %>% dplyr::select(DivisionNm, StateAb), by = "DivisionNm")
@@
   bind_rows(
-    glsmod16$my_data %>% left_join(tpp16 %>% dplyr::select(DivisionNm, DivisionID, StateAb), by = "DivisionNm"),
-    glsmod13$my_data %>% left_join(tpp13 %>% dplyr::select(DivisionNm, DivisionID, StateAb), by = "DivisionNm"),
-    glsmod10$my_data %>% left_join(tpp10 %>% dplyr::select(DivisionNm, DivisionID, StateAb), by = "DivisionNm"),
-    glsmod07$my_data %>% left_join(tpp07 %>% dplyr::select(DivisionNm, DivisionID, StateAb), by = "DivisionNm"),
-    glsmod04$my_data %>% left_join(tpp04 %>% dplyr::select(DivisionNm, DivisionID, StateAb), by = "DivisionNm"),
-    glsmod01$my_data %>% left_join(tpp01 %>% dplyr::select(DivisionNm, DivisionID, StateAb), by = "DivisionNm"))
+    glsmod16$my_data %>% left_join(tpp16 %>% dplyr::select(DivisionNm, UniqueID, StateAb), by = "DivisionNm"),
+    glsmod13$my_data %>% left_join(tpp13 %>% dplyr::select(DivisionNm, UniqueID, StateAb), by = "DivisionNm"),
+    glsmod10$my_data %>% left_join(tpp10 %>% dplyr::select(DivisionNm, UniqueID, StateAb), by = "DivisionNm"),
+    glsmod07$my_data %>% left_join(tpp07 %>% dplyr::select(DivisionNm, UniqueID, StateAb), by = "DivisionNm"),
+    glsmod04$my_data %>% left_join(tpp04 %>% dplyr::select(DivisionNm, UniqueID, StateAb), by = "DivisionNm"),
+    glsmod01$my_data %>% left_join(tpp01 %>% dplyr::select(DivisionNm, UniqueID, StateAb), by = "DivisionNm"))
 )

### 2019-05-07 elections_conversation.tex
substit uniqueID for DivisionID to make code work, and revised first section

@@
 
-We often hear stereotypes about voting patterns --- people in their 20s without kids are more likely to be left wing, migrants are more conservative, wealthier people tend to favour the conservative parties, and so on.
-So we thought we would test these ideas by matching census data to election data, to see if we can identify what are the socio-demographic characteristics of an electorate that are most closely related to how they vote.
+We often hear commentary about voting patterns --- people in their 20s without kids are more likely to be left wing, migrants are more conservative, wealthier people tend to favour the conservative parties, and so on.
+These might be built on myth and stereotypes.
+There is open data available from each election and from regular national census that can be used to examine, just what are the indicators for voting tendency.
 
-It is also interesting to see how this might have changed over time.
-For example, if wealth was a good predictor of voting patterns in the 2001 election, is it still a good predictor in 2019?
+Voting tendencies can chang over time.
+For example, if wealth was a good predictor of conservative voting in the 2001 election, is it still a good predictor in 2019?
 Australia has changed in many ways over the last two decades.
@@
 
-\begin{figure}[H]
-
-{\centering \includegraphics{elections_conversation_files/figure-latex/intercept2-1}
-
-}
-
-\caption{Estimated intercept for each election, which represents the two-party preferred vote for an electorate with average characteristics.}\label{fig:intercept2}
-\end{figure}
-
 Since all socio-demographics have been standardized, the intercept in each model can be interpreted as the estimated two-party preferred vote for an ``average'' electorate.
@@
 The dots represent the estimated 2PP value, and the lines indicate a 95\% confidence interval providing a guide to the uncertainty in the estimate.
+
+It is important here to note the ecological fallacy: insights are being drawn at the electorate level, and cannot be inferred for another disaggregate level (e.g.~individual voters).
 
@@
 
-\begin{center}\includegraphics{elections_conversation_files/figure-latex/age-1} \end{center}
+\begin{figure}[H]
 
-\hypertarget{influential-socio-demographics}{%
-\subsection{Influential socio-demographics}\label{influential-socio-demographics}}
+{\centering \includegraphics{elections_conversation_files/figure-latex/age-1}
 
-\begin{center}\includegraphics{elections_conversation_files/figure-latex/prepplot-1} \end{center}
+}
 
-\begin{center}\includegraphics{elections_conversation_files/figure-latex/prepplot-2} \end{center}
-
-\begin{center}\includegraphics{elections_conversation_files/figure-latex/prepplot-3} \end{center}
-
-\begin{center}\includegraphics{elections_conversation_files/figure-latex/prepplot-4} \end{center}
-
-\begin{center}\includegraphics{elections_conversation_files/figure-latex/prepplot-5} \end{center}
-
-\begin{center}\includegraphics{elections_conversation_files/figure-latex/prepplot-6} \end{center}
-
-\begin{center}\includegraphics{elections_conversation_files/figure-latex/prepplot-7} \end{center}
-
-\begin{center}\includegraphics{elections_conversation_files/figure-latex/prepplot-8} \end{center}
-
-\begin{center}\includegraphics{elections_conversation_files/figure-latex/prepplot-9} \end{center}
-
-\begin{center}\includegraphics{elections_conversation_files/figure-latex/prepplot-10} \end{center}
-
-\begin{center}\includegraphics{elections_conversation_files/figure-latex/prepplot-11} \end{center}
-
-\begin{center}\includegraphics{elections_conversation_files/figure-latex/prepplot-12} \end{center}
-
-\begin{center}\includegraphics{elections_conversation_files/figure-latex/prepplot-13} \end{center}
-
-\begin{center}\includegraphics{elections_conversation_files/figure-latex/prepplot-14} \end{center}
-
-Partial residual plots by election year for a selection of predictors.
-Linear model with 95\textbackslash{}\% confidence bands overlaid.
-Most predictors have a positive relationship: the larger the value the more likely the electorate preferences the Coalition.
-The relationship is relatively robust over time, with the exception of Unemployment.
-
-Several predictors have a negative relationship: with larger values indicating the electorate more likely preferences Labor.
-Most relationships are relatively stable over elections, except OtherLanguage and Education.
-
-To investigate the socio-demographics that have a strong effect on the two-party preferred vote, partial residual plots are used and shown in Figures \ref{fig:partresplot1} and \ref{fig:partresplot2}.
-The partial residuals are the residuals from the fitted model with the estimated effect an individual variable added.
-These show the direction, size and significance of an estimated effect --- the slope of the prediction line matches the estimated coefficient, and the shaded region represents a 95\% confidence band, computed using the method in \textcite{visreg}.
-If a horizontal line can be drawn through the confidence band, then the effect is insignificant.
-The estimated intercept is also added to the partial residuals for interpretability.
-Plots for each election are faceted to compare the effects over time in Figures \ref{fig:partresplot1} and \ref{fig:partresplot2}.
-Only socio-demographics that have a significant effect in at least one election are displayed in Figures \ref{fig:partresplot1} and \ref{fig:partresplot2}.
-
-It is important here to note the ecological fallacy: insights are being drawn at the electorate level, and cannot be inferred for another disaggregate level (e.g.~individual voters).
+\caption{Estimated age effect for each election, which represents the impact of electorate median age on the two-party preferred vote. Higher values indicate a }\label{fig:age}
+\end{figure}
 
@@
 \subsubsection{Income and unemployment}\label{income-and-unemployment}}
+
+\begin{figure}[H]
+
+{\centering \includegraphics{elections_conversation_files/figure-latex/income-1}
+
+}
+
+\caption{Estimated age effect for each election, which represents the impact of electorate median age on the two-party preferred vote. Higher values indicate a }\label{fig:income}
+\end{figure}
 

### 2019-05-07 wrangle.R

@@ -84,7 +84,7 @@ my_df <- bind_rows(
     LNP_Votes, ALP_Votes, ALP_Percent, TotalVotes, Swing,
     InternetUse, InternetAccess, EnglishOnly,
     Other_NonChrist, OtherChrist, Catholic, Anglican,
-    Volunteer, EmuneratedElsewhere, DivisionID
+    Volunteer, EmuneratedElsewhere #, DivisionID
   ))
 
 # Create final df for modelling

### 2019-05-07 elections_conversation.Rmd
Adding some interpretation

--- 2019-05-06 elections_conversation.Rmd
@@
 
-We often hear stereotypes about voting patterns --- people in their 20s without kids are more likely to be left wing, migrants are more conservative, wealthier people tend to favour the conservative parties, and so on.
-So we thought we would test these ideas by matching census data to election data, to see if we can identify what are the socio-demographic characteristics of an electorate that are most closely related to how they vote.
+We often hear commentary about voting patterns --- people in their 20s without kids are more likely to be left wing, migrants are more conservative, wealthier people tend to favour the conservative parties, and so on.
+These might be built on myth and stereotypes.
+There is open data available from each election and from regular national censuses; this data can be used to examine just what are the indicators for voting tendency.
 
-It is also interesting to see how this might have changed over time.
-For example, if wealth was a good predictor of voting patterns in the 2001 election, is it still a good predictor in 2019?
+Voting tendencies can change over time.
+For example, if wealth was a good predictor of conservative voting in the 2001 election, is it still a good predictor in 2019?
 Australia has changed in many ways over the last two decades.
@@
 Some electorates can disappear altogether and new electorates can arise.
-Electoral boundaries are redistributed regularly by the AEC, meaning that only in the years where both a Census and election occur are all boundaries likely to match --- the case for the 2001 and 2016 elections.
+Electoral boundaries are redistributed regularly by the AEC, meaning that only in the years where both a Census and an election occur are all boundaries likely to match --- the case for the 2001 and 2016 elections.
 So we first had to estimate what the socio-demographic characteristics of an electorate would have been at the time of each election using a complicated method of interpolation over time and geography.
@@
 For example, our "Incomes" variable is a combination of median personal income, household income and family income.
+
+We are using data at the electorate level. The results do not directly address the voting intentions of individuals.
 
@@
 
-## Country-wide trend
+## Country-wide trend {-}
 
@@
 
-Since all socio-demographics have been standardized, the intercept in each model can be interpreted as the estimated two-party preferred vote for an "average" electorate.
+First, we show the estimated two-party preferred vote for an "average" electorate.
 Figure \@ref(fig:plotintercept) shows that the baseline of party preference has varied over the elections, with the biggest swing occurring in the 2007 election where the mean electorate shifted more than five percentage points in favour of the Labor party.
@@
 
-It is important here to note the ecological fallacy: insights are being drawn at the electorate level, and cannot be inferred for another disaggregate level (e.g. individual voters).
-
-## Age
-
-- Would it be possible to show how the correlation changes for different age brackets (ie. younger voters = more Labor, and this swings more to Liberal as the age brackets gets older)?
-- If not, can you go into more detail specific age brackets? Eg. Is it the over-65s that skew Liberal?
+## Age {-}
 
@@
 
-### Income and unemployment
+Regions comprising more older people are often believed to be more conservative, and indeed we found that electorates with a higher median age are more likely to support the Liberal party --- although this effect is significant only in 2007 and 2010.
+
+Figure \@ref(fig:age) shows the effect of median age on the 2PP vote.
+The vertical scale can be interpreted as a score measuring how much the 2PP vote differs between an average electorate, and an electorate with median age in the top third of all electorates.
+
+- Would it be possible to show how the correlation changes for different age brackets (ie. younger voters = more Labor, and this swings more to Liberal as the age brackets gets older)?
+- If not, can you go into more detail specific age brackets? Eg. Is it the over-65s that skew Liberal?
+
+## Income and unemployment {-}
 
@@
 
-### Industry and type of work
+## Industry and type of work {-}
 
@@
 
-### Household mobility
+## Household mobility {-}
 
@@
 
-### Relationships
+## Relationships {-}
 
@@
 
-### Age
-
-Regions comprising more older people are often believed to be more conservative, and indeed it found that electorates with a higher median age are more likely to support the Liberal party --- although this effect is significant only in 2007 and 2010 (see row 2 in Figure \@ref(fig:partresplot2)).
-
-### Education
+## Education {-}
 
@@
 
-### Diversity
+## Diversity {-}
 
@@
 
-## A closer look at the residuals
+# A closer look at the residuals
 
@@
 
-### Residuals by state
+## Residuals by state
 
@@
 
-### Outlier electorates
+## Outlier electorates
 

### 2019-05-07 elections_conversation.tex
Adding some interpretation

@@
 These might be built on myth and stereotypes.
-There is open data available from each election and from regular national census that can be used to examine, just what are the indicators for voting tendency.
+There is open data available from each election and from regular national censuses; this data can be used to examine just what are the indicators for voting tendency.
 
-Voting tendencies can chang over time.
+Voting tendencies can change over time.
 For example, if wealth was a good predictor of conservative voting in the 2001 election, is it still a good predictor in 2019?
@@
 
+We are using data at the electorate level. The results do not directly address the voting intentions of individuals.
+
 Each election was modelled separately, to allow us to see any changes over time, and to account for changing electorate boundaries.
@@
 \hypertarget{country-wide-trend}{%
-\subsection{Country-wide trend}\label{country-wide-trend}}
+\subsection*{Country-wide trend}\label{country-wide-trend}}
+\addcontentsline{toc}{subsection}{Country-wide trend}
 
@@
 
-Since all socio-demographics have been standardized, the intercept in each model can be interpreted as the estimated two-party preferred vote for an ``average'' electorate.
+First, we show the estimated two-party preferred vote for an ``average'' electorate.
 Figure \ref{fig:plotintercept} shows that the baseline of party preference has varied over the elections, with the biggest swing occurring in the 2007 election where the mean electorate shifted more than five percentage points in favour of the Labor party.
@@
 
-It is important here to note the ecological fallacy: insights are being drawn at the electorate level, and cannot be inferred for another disaggregate level (e.g.~individual voters).
-
 \hypertarget{age}{%
-\subsection{Age}\label{age}}
-
-\begin{itemize}
-\tightlist
-\item
-  Would it be possible to show how the correlation changes for different age brackets (ie. younger voters = more Labor, and this swings more to Liberal as the age brackets gets older)?
-\item
-  If not, can you go into more detail specific age brackets? Eg. Is it the over-65s that skew Liberal?
-\end{itemize}
+\subsection*{Age}\label{age}}
+\addcontentsline{toc}{subsection}{Age}
 
@@
 
+Regions comprising more older people are often believed to be more conservative, and indeed we found that electorates with a higher median age are more likely to support the Liberal party --- although this effect is significant only in 2007 and 2010.
+
+Figure \ref{fig:age} shows the effect of median age on the 2PP vote.
+The vertical scale can be interpreted as a score measuring how much the 2PP vote differs between an average electorate, and an electorate with median age in the top third of all electorates.
+
+\begin{itemize}
+\tightlist
+\item
+  Would it be possible to show how the correlation changes for different age brackets (ie. younger voters = more Labor, and this swings more to Liberal as the age brackets gets older)?
+\item
+  If not, can you go into more detail specific age brackets? Eg. Is it the over-65s that skew Liberal?
+\end{itemize}
+
 \hypertarget{income-and-unemployment}{%
-\subsubsection{Income and unemployment}\label{income-and-unemployment}}
+\subsection*{Income and unemployment}\label{income-and-unemployment}}
+\addcontentsline{toc}{subsection}{Income and unemployment}
 
@@
 \hypertarget{industry-and-type-of-work}{%
-\subsubsection{Industry and type of work}\label{industry-and-type-of-work}}
+\subsection*{Industry and type of work}\label{industry-and-type-of-work}}
+\addcontentsline{toc}{subsection}{Industry and type of work}
 
@@
 \hypertarget{household-mobility}{%
-\subsubsection{Household mobility}\label{household-mobility}}
+\subsection*{Household mobility}\label{household-mobility}}
+\addcontentsline{toc}{subsection}{Household mobility}
 
@@
 \hypertarget{relationships}{%
-\subsubsection{Relationships}\label{relationships}}
+\subsection*{Relationships}\label{relationships}}
+\addcontentsline{toc}{subsection}{Relationships}
 
@@
 
-\hypertarget{age-1}{%
-\subsubsection{Age}\label{age-1}}
-
-Regions comprising more older people are often believed to be more conservative, and indeed it found that electorates with a higher median age are more likely to support the Liberal party --- although this effect is significant only in 2007 and 2010 (see row 2 in Figure \ref{fig:partresplot2}).
-
 \hypertarget{education}{%
-\subsubsection{Education}\label{education}}
+\subsection*{Education}\label{education}}
+\addcontentsline{toc}{subsection}{Education}
 
@@
 \hypertarget{diversity}{%
-\subsubsection{Diversity}\label{diversity}}
+\subsection*{Diversity}\label{diversity}}
+\addcontentsline{toc}{subsection}{Diversity}
 
@@
 \hypertarget{a-closer-look-at-the-residuals}{%
-\subsection{A closer look at the residuals}\label{a-closer-look-at-the-residuals}}
+\section{A closer look at the residuals}\label{a-closer-look-at-the-residuals}}
 
 \hypertarget{residuals-by-state}{%
-\subsubsection{Residuals by state}\label{residuals-by-state}}
+\subsection{Residuals by state}\label{residuals-by-state}}
 
@@
 \hypertarget{outlier-electorates}{%
-\subsubsection{Outlier electorates}\label{outlier-electorates}}
+\subsection{Outlier electorates}\label{outlier-electorates}}
 

### 2019-05-07 wrangle.R

@@ -84,7 +84,7 @@ my_df <- bind_rows(
     LNP_Votes, ALP_Votes, ALP_Percent, TotalVotes, Swing,
     InternetUse, InternetAccess, EnglishOnly,
     Other_NonChrist, OtherChrist, Catholic, Anglican,
-    Volunteer, EmuneratedElsewhere #, DivisionID
+    Volunteer, EmuneratedElsewhere, DivisionID
   ))
 
 # Create final df for modelling

### 2019-05-07 elections_conversation.Rmd
color and caption for fig 1

@@
 
-```{r timeline, echo = FALSE, fig.cap="Timeline of Australian elections and Censuses. They do not always occur in the same year.", out.width = "90%", fig.align="center"}
+```{r timeline, echo = FALSE, fig.cap="Timeline of Australian elections and Censuses.
+Censuses happen regularly, elections not quite regularly, which means data needs to be interpolated by time.", out.width = "90%", fig.align="center"}
 timeline <- data.frame(year = c(2001:2016)) %>%
@@
   geom_hline(yintercept = 0, size = 1, color = "grey", show.legend = FALSE) +
-  geom_segment(x = 2001, xend = 2001, y = 1, yend = 0, color = "#00BFC4") +
-  geom_segment(x = 2004, xend = 2004, y = 1, yend = 0, color = "#00BFC4") +
-  geom_segment(x = 2007, xend = 2007, y = 1, yend = 0, color = "#00BFC4") +
-  geom_segment(x = 2010, xend = 2010, y = 1, yend = 0, color = "#00BFC4") +
-  geom_segment(x = 2013, xend = 2013, y = 1, yend = 0, color = "#00BFC4") +
-  geom_segment(x = 2016, xend = 2016, y = 1, yend = 0, color = "#00BFC4") +
-  geom_segment(x = 2001.1, xend = 2001.1, y = 1.4, yend = 0, color = "#F8766D") +
-  geom_segment(x = 2006.1, xend = 2006.1, y = 1.4, yend = 0, color = "#F8766D") +
-  geom_segment(x = 2011.1, xend = 2011.1, y = 1.4, yend = 0, color = "#F8766D") +
-  geom_segment(x = 2016.1, xend = 2016.1, y = 1.4, yend = 0, color = "#F8766D") +
+  geom_segment(x = 2001, xend = 2001, y = 1, yend = 0, color = "#1B9E77") +
+  geom_segment(x = 2004, xend = 2004, y = 1, yend = 0, color = "#1B9E77") +
+  geom_segment(x = 2007, xend = 2007, y = 1, yend = 0, color = "#1B9E77") +
+  geom_segment(x = 2010, xend = 2010, y = 1, yend = 0, color = "#1B9E77") +
+  geom_segment(x = 2013, xend = 2013, y = 1, yend = 0, color = "#1B9E77") +
+  geom_segment(x = 2016, xend = 2016, y = 1, yend = 0, color = "#1B9E77") +
+  geom_segment(x = 2001.1, xend = 2001.1, y = 1.4, yend = 0, color = "#D95F02") +
+  geom_segment(x = 2006.1, xend = 2006.1, y = 1.4, yend = 0, color = "#D95F02") +
+  geom_segment(x = 2011.1, xend = 2011.1, y = 1.4, yend = 0, color = "#D95F02") +
+  geom_segment(x = 2016.1, xend = 2016.1, y = 1.4, yend = 0, color = "#D95F02") +
   scale_x_discrete(limits = c(2001:2016)) +

### 2019-05-07 wrangle.R

@@ -84,7 +84,7 @@ my_df <- bind_rows(
     LNP_Votes, ALP_Votes, ALP_Percent, TotalVotes, Swing,
     InternetUse, InternetAccess, EnglishOnly,
     Other_NonChrist, OtherChrist, Catholic, Anglican,
-    Volunteer, EmuneratedElsewhere, DivisionID
+    Volunteer, EmuneratedElsewhere
   ))
 
 # Create final df for modelling

### 2019-05-08 elections_conversation.Rmd
Coefficient analysis

--- 2019-05-07 elections_conversation.Rmd
@@
     fig_caption: yes
-    fig_height: 5
+    fig_height: 3
     fig_width: 8
@@
   fig.align = "center",
-  cache = TRUE,
+  cache = FALSE,
   message = FALSE
@@
 
-```{r timeline, echo = FALSE, fig.cap="Timeline of Australian elections and Censuses. They do not always occur in the same year.", out.width = "90%", fig.align="center"}
-timeline <- data.frame(year = c(2001:2016)) %>%
+```{r timeline, echo = FALSE, fig.cap="Timeline of Australian elections and Censuses. Censuses happen regularly, elections not quite regularly, which means data needs to be interpolated by time."}
+data.frame(year = c(2001:2016)) %>%
   mutate(
@@
   filter(measure != 0) %>%
-  mutate(year_vis = ifelse(event == "Census", year + 0.1, year))
-
-ggplot(timeline, aes(x = year_vis, y = measure)) +
+  mutate(year_vis = ifelse(event == "Census", year + 0.1, year)) %>%
+  ggplot(aes(x = year_vis, y = measure)) +
     geom_point(aes(color = as.factor(event), shape = as.factor(event)), size = 4) +
@@
     geom_hline(yintercept = 0, size = 1, color = "grey", show.legend = FALSE) +
-  geom_segment(x = 2001, xend = 2001, y = 1, yend = 0, color = "#00BFC4") +
-  geom_segment(x = 2004, xend = 2004, y = 1, yend = 0, color = "#00BFC4") +
-  geom_segment(x = 2007, xend = 2007, y = 1, yend = 0, color = "#00BFC4") +
-  geom_segment(x = 2010, xend = 2010, y = 1, yend = 0, color = "#00BFC4") +
-  geom_segment(x = 2013, xend = 2013, y = 1, yend = 0, color = "#00BFC4") +
-  geom_segment(x = 2016, xend = 2016, y = 1, yend = 0, color = "#00BFC4") +
-  geom_segment(x = 2001.1, xend = 2001.1, y = 1.4, yend = 0, color = "#F8766D") +
-  geom_segment(x = 2006.1, xend = 2006.1, y = 1.4, yend = 0, color = "#F8766D") +
-  geom_segment(x = 2011.1, xend = 2011.1, y = 1.4, yend = 0, color = "#F8766D") +
-  geom_segment(x = 2016.1, xend = 2016.1, y = 1.4, yend = 0, color = "#F8766D") +
+    geom_segment(x = 2001, xend = 2001, y = 1, yend = 0, color = "#1B9E77") +
+    geom_segment(x = 2004, xend = 2004, y = 1, yend = 0, color = "#1B9E77") +
+    geom_segment(x = 2007, xend = 2007, y = 1, yend = 0, color = "#1B9E77") +
+    geom_segment(x = 2010, xend = 2010, y = 1, yend = 0, color = "#1B9E77") +
+    geom_segment(x = 2013, xend = 2013, y = 1, yend = 0, color = "#1B9E77") +
+    geom_segment(x = 2016, xend = 2016, y = 1, yend = 0, color = "#1B9E77") +
+    geom_segment(x = 2001.1, xend = 2001.1, y = 1.4, yend = 0, color = "#D95F02") +
+    geom_segment(x = 2006.1, xend = 2006.1, y = 1.4, yend = 0, color = "#D95F02") +
+    geom_segment(x = 2011.1, xend = 2011.1, y = 1.4, yend = 0, color = "#D95F02") +
+    geom_segment(x = 2016.1, xend = 2016.1, y = 1.4, yend = 0, color = "#D95F02") +
     scale_x_discrete(limits = c(2001:2016)) +
@@
 
-```{r plotintercept, fig.cap = "Estimated intercept for each election, which represents the two-party preferred vote for an electorate with average characteristics.", fig.pos="H"}
-coefs %>% filter(variable=="Intercept") %>%
+First, we show the estimated two-party preferred vote for an "average" electorate.
+
+```{r plotintercept, fig.cap = "Estimated two-party preferred vote for an electorate with average characteristics at each election."}
+coefs %>%
+  filter(variable=="Intercept") %>%
   ggplot() +
+    geom_hline(aes(yintercept = 50), col='gray', size = 1) +
     geom_point(aes(x = year, y = estimate), size = 2) +
-  geom_linerange(aes(x = year, ymin = lower95, ymax = upper95), size = 1.2) +
-  geom_hline(aes(yintercept = 50), alpha = 0.5, size = 1) +
+    geom_linerange(aes(x = year, ymin = lower95, ymax = upper95), size = .7) +
     labs(x = "Election year", y = "2PP percentage") +
-  coord_fixed(ratio = 0.25) +
     scale_x_continuous(breaks=seq(2001,2016,by=3), minor_breaks=NULL) +
-  scale_y_continuous(breaks=seq(46,54,by=2), minor_breaks=45:55)
+    scale_y_continuous(breaks=seq(44,56,by=2), minor_breaks=NULL, limits=50 + c(-1,1)*4.66)
 ```
 
-First, we show the estimated two-party preferred vote for an "average" electorate.
-Figure \@ref(fig:plotintercept) shows that the baseline of party preference has varied over the elections, with the biggest swing occurring in the 2007 election where the mean electorate shifted more than five percentage points in favour of the Labor party.
+This shows that the baseline of party preference has varied over the elections, with the biggest swing occurring in the 2007 election where the mean electorate shifted more than five percentage points in favour of the Labor party.
 The dots represent the estimated 2PP value, and the lines indicate a 95% confidence interval providing a guide to the uncertainty in the estimate.
+Where the vertical lines corresponding to each election cross the 50% horizontal line, the 2PP vote for an average electorate is not statistically distinguishable from 50% for that election.
+
+```{r scaling}
+# For the remaining graphics, we will rescale the coefficients to allow easier interpretation
+coefs <- coefs %>%
+  mutate(
+    estimate = estimate * qnorm(.9),
+    lower95 = lower95 * qnorm(.9),
+    upper95 = upper95 * qnorm(.9),
+  )
+
+yrange <- coefs %>%
+  filter(variable %in% c(
+      "MedianAge",
+      "Incomes",
+      "Unemployment",
+      "Education",
+      "Extractive",
+      "Transformative",
+      "ManagerAdmin",
+      "DiffAddress",
+      "DeFacto",
+      "Education",
+      "OtherLanguage"
+    )) %>%
+  summarise(
+    upper = max(upper95),
+    lower = min(lower95)
+  )
+yscale <- scale_y_continuous(
+  breaks=seq(-50,50,by=5),
+  minor_breaks=NULL,
+  limits=c(yrange$lower,yrange$upper))
+```
 
@@
 
-```{r age, fig.cap = "Estimated age effect for each election, which represents the impact of electorate median age on the two-party preferred vote. Higher values indicate a ", fig.pos="H"}
-coefs %>% filter(variable=="MedianAge") %>%
+Regions comprising more older people are often believed to be more conservative, and indeed we found that electorates with a higher median age are more likely to support the Liberal party --- although this effect is significant only in 2004 and 2007.
+
+```{r age, fig.cap = "Estimated impact of electorate median age on the two-party preferred vote."}
+coefs %>%
+  filter(variable=="MedianAge") %>%
   ggplot(aes(x=year)) +
@@
     geom_point(aes(x = year, y = estimate), size = 2) +
-  geom_linerange(aes(x = year, ymin = lower95, ymax = upper95), size = 1.2) +
+    geom_linerange(aes(x = year, ymin = lower95, ymax = upper95), size = .7) +
     labs(x = "Election year", y = "2PP impact") +
-  coord_fixed(ratio = 0.25) +
-  scale_x_continuous(breaks=seq(2001,2016,by=3), minor_breaks=NULL)
+    scale_x_continuous(breaks=seq(2001,2016,by=3), minor_breaks=NULL) + yscale
 ```
 
-Regions comprising more older people are often believed to be more conservative, and indeed we found that electorates with a higher median age are more likely to support the Liberal party --- although this effect is significant only in 2007 and 2010.
+This graph shows the effect of median age on the 2PP vote.
+The vertical scale can be interpreted as a score measuring how much the 2PP vote differs between an average electorate, and an electorate with median age in the top 10% of all electorates (with all other socio-demographic variables unchanged).
+The reverse is true for young electorates --- the score can be interpreted as how much the 2PP vote would be reduced if an average electorate kept all socio-demographic variables fixed but median age was reduced to be in the bottom 10% of all eletorates.
 
-Figure \@ref(fig:age) shows the effect of median age on the 2PP vote.
-The vertical scale can be interpreted as a score measuring how much the 2PP vote differs between an average electorate, and an electorate with median age in the top third of all electorates.
+## Income {-}
 
-- Would it be possible to show how the correlation changes for different age brackets (ie. younger voters = more Labor, and this swings more to Liberal as the age brackets gets older)?
-- If not, can you go into more detail specific age brackets? Eg. Is it the over-65s that skew Liberal?
+Typically the Labor party campaigns on more progressive policies, which often include tax reform that adversely affects higher income earners, and more generous social assistance programs.
+Perhaps it is due to these policies that higher income electorates appear more likely to support the Liberal party, as the `Incomes` factor has a positive effect on Liberal preference.
+The effect also seems to have been increasing over time.
 
-## Income and unemployment {-}
-
-```{r income, fig.cap = "Estimated age effect for each election, which represents the impact of electorate median age on the two-party preferred vote. Higher values indicate a ", fig.pos="H"}
-coefs %>% filter(variable=="Income") %>%
+```{r income, fig.cap = "Estimated impact of electorate income on the two-party preferred vote."}
+coefs %>%
+  filter(variable=="Incomes") %>%
   ggplot(aes(x=year)) +
@@
     geom_point(aes(x = year, y = estimate), size = 2) +
-  geom_linerange(aes(x = year, ymin = lower95, ymax = upper95), size = 1.2) +
+    geom_linerange(aes(x = year, ymin = lower95, ymax = upper95), size = .7) +
     labs(x = "Election year", y = "2PP impact") +
-  coord_fixed(ratio = 0.25) +
-  scale_x_continuous(breaks=seq(2001,2016,by=3), minor_breaks=NULL)
+    scale_x_continuous(breaks=seq(2001,2016,by=3), minor_breaks=NULL) + yscale
 ```
 
-Typically the Labor party campaigns on more progressive policies, which often include tax reform that adversely affects higher income earners, and more generous social assistance programs.
-Perhaps it is due to these policies that higher income electorates appear more likely to support the Liberal party, as the `Incomes` factor has a positive effect on Liberal preference (see row 1 in Figure \@ref(fig:partresplot1)).
-This effect is significant in every election aside from 2004, where it is only marginally insignificant ($p = 0.0613$).
+As before, the vertical scale can be interpreted as a score measuring how much the 2PP vote differs between an average electorate, and an electorate with income in the top 10% of all electorates (with all other socio-demographic variables unchanged).
+
+## Unemployment {-}
+
 Unemployment however, is not as influential.
-In 2001 and 2004, electorates with higher unemployment align with Labor, but over time this shifts towards support for the Liberal party, culminating in a significantly positive effect in 2016.
+In 2001 and 2004, electorates with higher unemployment align with Labor, but over time this shifts towards support for the Liberal party, culminating in a significantly positive (but small) effect in 2016.
+
+```{r unemployment, fig.cap = "Estimated impact of electorate unemployment on the two-party preferred vote."}
+coefs %>%
+  filter(variable=="Unemployment") %>%
+  ggplot(aes(x=year)) +
+    geom_hline(aes(yintercept = 0), col='gray', size = 1) +
+    geom_point(aes(x = year, y = estimate), size = 2) +
+    geom_linerange(aes(x = year, ymin = lower95, ymax = upper95), size = .7) +
+    labs(x = "Election year", y = "2PP impact") +
+    scale_x_continuous(breaks=seq(2001,2016,by=3), minor_breaks=NULL) + yscale
+```
+
+The vertical scale can be interpreted as a score measuring how much the 2PP vote differs between an average electorate, and an electorate with unemployment in the top 10% of all electorates (with all other socio-demographic variables unchanged)
 
@@
 
-Electorates with higher proportions of workers in mining, gas, water, agriculture, waste and electricity (grouped as `Extractive` industries) are consistently linked with higher support for the Liberal party, with the magnitude of this effect slightly increasing over the years (see row 3 in Figure \@ref(fig:partresplot1)).
+Electorates with higher proportions of workers in extractive industries (mining, gas, water, agriculture, waste and electricity) are consistently linked with higher support for the Liberal party, with the magnitude of this effect slightly increasing over the years.
+
+```{r extractive, fig.cap = "Estimated impact of employment in the extractive industries on the two-party preferred vote."}
+coefs %>%
+  filter(variable=="Extractive") %>%
+  ggplot(aes(x=year)) +
+    geom_hline(aes(yintercept = 0), col='gray', size = 1) +
+    geom_point(aes(x = year, y = estimate), size = 2) +
+    geom_linerange(aes(x = year, ymin = lower95, ymax = upper95), size = .7) +
+    labs(x = "Election year", y = "2PP impact") +
+    scale_x_continuous(breaks=seq(2001,2016,by=3), minor_breaks=NULL) + yscale
+```
+
 This is unsurprising, as the Liberal party has close ties with these traditional energy industries, and typically present policies to reduce taxation on energy production.
-Furthermore, electorates with more workers in construction or manufacturing industries (`Transformative`) are also more likely to support the Liberal party (see row 4 in Figure \@ref(fig:partresplot1)).
 
-Similarly, the proportion of workers in managerial, administrative, clerical and sales roles (`ManagerAdmin`) is also a significant predictor of two-party preference vote across all six elections, with a higher proportion of people working these jobs increasing Liberal support.
-The magnitude of this effect also seems to increase over the years.
+Electorates with more workers in transformative industries (construction or manufacturing) are also more likely to support the Liberal party.
+
+```{r transformative, fig.cap = "Estimated impact of employment in the transformative industries on the two-party preferred vote."}
+coefs %>%
+  filter(variable=="Transformative") %>%
+  ggplot(aes(x=year)) +
+    geom_hline(aes(yintercept = 0), col='gray', size = 1) +
+    geom_point(aes(x = year, y = estimate), size = 2) +
+    geom_linerange(aes(x = year, ymin = lower95, ymax = upper95), size = .7) +
+    labs(x = "Election year", y = "2PP impact") +
+    scale_x_continuous(breaks=seq(2001,2016,by=3), minor_breaks=NULL) + yscale
+```
+
+The proportion of workers in managerial, administrative, clerical and sales roles is also a significant predictor of two-party preference vote across all six elections, with a higher proportion of people working these jobs increasing Liberal support.
+
+```{r manager, fig.cap = "Estimated impact of employment in management or administration on the two-party preferred vote."}
+coefs %>%
+  filter(variable=="ManagerAdmin") %>%
+  ggplot(aes(x=year)) +
+    geom_hline(aes(yintercept = 0), col='gray', size = 1) +
+    geom_point(aes(x = year, y = estimate), size = 2) +
+    geom_linerange(aes(x = year, ymin = lower95, ymax = upper95), size = .7) +
+    labs(x = "Election year", y = "2PP impact") +
+    scale_x_continuous(breaks=seq(2001,2016,by=3), minor_breaks=NULL) + yscale
+```
 
@@
 
-In each of the six elections, electorates with a higher proportion of people that have recently (in the past five years) moved house (`DiffAddress`) are more likely to support the Liberal party, although this effect was marginally insignificant in 2013 (see row 6 in Figure \@ref(fig:partresplot1).
-Having controlled for characteristics of house ownership and rental prices (via the factors `PropertyOwned` and `RentLoan` respectively), this effect is somewhat surprising.
+In each of the six elections, electorates with a higher proportion of people that have recently (in the past five years) moved house are more likely to support the Liberal party.
+
+```{r mobility, fig.cap = "Estimated impact of household mobility on the two-party preferred vote."}
+coefs %>%
+  filter(variable=="DiffAddress") %>%
+  ggplot(aes(x=year)) +
+    geom_hline(aes(yintercept = 0), col='gray', size = 1) +
+    geom_point(aes(x = year, y = estimate), size = 2) +
+    geom_linerange(aes(x = year, ymin = lower95, ymax = upper95), size = .7) +
+    labs(x = "Election year", y = "2PP impact") +
+    scale_x_continuous(breaks=seq(2001,2016,by=3), minor_breaks=NULL) + yscale
+```
+
+Our analysis controls for characteristics of home ownership and rental prices, so this effect is not simply due to electorates having low rates of home ownership, or due to electorates having high rental prices.
+Instead, it suggests that people who are more transient are also more likely to be conservative voters, regardless of their home ownership or rental status.
+(This would need further study, as we do not have individual level voting data.)
 
@@
 
+```{r defacto, fig.cap = "Estimated impact of de facto relationships on the two-party preferred vote."}
+coefs %>%
+  filter(variable=="DeFacto") %>%
+  ggplot(aes(x=year)) +
+    geom_hline(aes(yintercept = 0), col='gray', size = 1) +
+    geom_point(aes(x = year, y = estimate), size = 2) +
+    geom_linerange(aes(x = year, ymin = lower95, ymax = upper95), size = .7) +
+    labs(x = "Election year", y = "2PP impact") +
+    scale_x_continuous(breaks=seq(2001,2016,by=3), minor_breaks=NULL) + yscale
+```
+
 ## Education {-}
 
-Since 2007, electorates with higher education levels are associated with supporting the Labor party, although this effect is significant only in 2016.
-Before 2007, education has an almost zero effect (see row 3 in Figure \@ref(fig:partresplot1)).
+Since 2007, electorates with higher education levels are associated with supporting the Labor party, although this effect is significant only in 2016. Before 2007, education had a negligible effect.
+
+```{r education, fig.cap = "Estimated impact of education on the two-party preferred vote."}
+coefs %>%
+  filter(variable=="Education") %>%
+  ggplot(aes(x=year)) +
+    geom_hline(aes(yintercept = 0), col='gray', size = 1) +
+    geom_point(aes(x = year, y = estimate), size = 2) +
+    geom_linerange(aes(x = year, ymin = lower95, ymax = upper95), size = .7) +
+    labs(x = "Election year", y = "2PP impact") +
+    scale_x_continuous(breaks=seq(2001,2016,by=3), minor_breaks=NULL) + yscale
+```
 
@@
 Larger migrant populations from Asia, the Middle East, South-Eastern Europe, the United Kingdom and elsewhere, are either associated with Labor support, or have no effect.
-Of these areas, only South-Eastern European populations appear significant in each election, with the proportion of Asian migrants also being significant in 2010.
-Speaking other languages (aside from English) however, appears to have a far stronger effect, as observed through the `OtherLanguage` variable.
-Electorates with more diverse speech are associated with higher support for the Liberal party from 2004 onwards, with this effect being significant in 2007, 2010 and 2016.
-Furthermore, of the variables relating to religion, only Judaism shows a consistent effect, with electorates with relatively large Jewish populations more likely to vote Liberal.
+Of these areas, only differences in South-Eastern European populations appear to have had a significant impact in each election, with the proportion of Asian migrants also being significant in 2010.
 
-# A closer look at the residuals
+Speaking languages other than English however, appears to have a far stronger effect.
+Electorates with more diverse languages are associated with higher support for the Liberal party from 2004 onwards, with this effect being significant in 2007, 2010 and 2016.
+
+```{r otherlanguage, fig.cap = "Estimated impact of ethnic diversity on the two-party preferred vote."}
+coefs %>%
+  filter(variable=="OtherLanguage") %>%
+  ggplot(aes(x=year)) +
+    geom_hline(aes(yintercept = 0), col='gray', size = 1) +
+    geom_point(aes(x = year, y = estimate), size = 2) +
+    geom_linerange(aes(x = year, ymin = lower95, ymax = upper95), size = .7) +
+    labs(x = "Election year", y = "2PP impact") +
+    scale_x_continuous(breaks=seq(2001,2016,by=3), minor_breaks=NULL) + yscale
+```
+
+Of the variables relating to religion, only Judaism shows a consistent effect, with electorates with relatively large Jewish populations more likely to vote Liberal.
+
+# State effects
 
@@
 
-## Residuals by state
-
-It is often hypothesized that states have systematic differences that cause their electorates to vote differently.
-Boxplots of residuals grouped by state (Figure \@ref(fig:resstate)) reveal that the data reflects this -- there appears to be a state-specific effect not captured by the models.
-Tasmania and the Australian Capital Territory appear to have a bias towards Labor, whereas the Northern Territory tends towards voting Liberal.
-However, there are relatively few electorates in each of these states (five, two and two respectively), so this apparent result may be due to incumbent effects rather than an actual state-specific bias.
+It is often suggested that states have systematic differences that cause their electorates to vote differently.
+We can explore this effect by looking at the difference between the actual 2PP vote in each electorate and what our modelling would predict based only on electorate characteristics.
+We found a bias toward Labor in the ACT, and to a smaller extent in Tasmania, and a bias toward the Liberals in the Northern Territory.
+No other states showed a significantly different voting pattern from what you would expect using only socio-demographic information.
+That suggests that any differences between Queensland and Victoria (for example) are due to different voter socio-demographics.
 
@@
 States ordered by median residual.
-A state-specific bias not captured by the model is evident.", fig.asp=0.6, out.width="70%", fig.pos = 'h', out.extra = ''}
+A state-specific bias not captured by the model is evident.", fig.asp=0.6, out.width="70%", fig.pos = 'h', out.extra = '', eval=FALSE}
 all_resids %>%
@@
 
-## Outlier electorates
+# Against the tide
+
+As well as looking at whether some states have results different from what is predicted, we can also look at individual electorates.
+That is, when does an electorate vote very differently from what their socio-demographics would suggest?
+This suggests something is going on beyond the effect of socio-demographics.
+For example, a very good (or bad) local member could lead to voting patterns that are very different from what is predicted.
 
@@
 The authors would like to sincerely thank Anthony Ebert, Heike Hofmann, Thomas Lumley, Ben Marwick, Carson Sievert, Mingzhu Sun, Dilini Talagala, Nicholas Tierney, Nathaniel Tomasetti, Earo Wang and Fang Zhou, all of whom have contributed to the `eechidna` package.
-
-# From Emil
-
-In terms of structure, start with a brief paragraph or two on what the research is, and what the reader can expect to find below.
-
-Then, break the article into small sections for each of the factors that have an effect (a short paragraph or two for each factor).
-And it would be great to include examples of specific seats where the relationship was strong, or interesting for other reasons.
-
-So the sections will be:
-
-- Age
-- Income
-- Unemployment
-- Industry and type of work
-- Education
-- Diversity
-- Household mobility
-- Relationships
-
-An example of this structure can be seen here:
-https://theconversation.com/confused-about-aged-care-in-the-home-these-10-charts-explain-how-it-works-113923
-
-Questions on particular factors:
-
-Incomes:
-- Was this correlation stronger in 2016? (and increasingly influential?)
-
-Household mobility
-- In the report, you say: "“Having controlled for characteristics of house ownership and rental prices”, whis effect is somewhat surprising”.
-Why is this surprising to you?
-(I would have expected this to support Labor more)
-
-Age
-- Would it be possible to show how the correlation changes for different age brackets (ie. younger voters = more Labor, and this swings more to Liberal as the age brackets gets older)?
-- If not, can you go into more detail specific age brackets? Eg. Is it the over-65s that skew Liberal?
-
-Diversity
-- Migrant population: Would it possible to break this out into particular groups?
-eg.
-Seats with higher Chinese population correlates with Liberal, seats with higher SEAsia pop correlates with Labor, etc?
-- The same for languages?

### 2019-05-08 elections_conversation.tex
Coefficient analysis

@@
 
-{\centering \includegraphics[width=0.9\linewidth]{elections_conversation_files/figure-latex/timeline-1}
+{\centering \includegraphics{elections_conversation_files/figure-latex/timeline-1}
 
@@
 
-\caption{Timeline of Australian elections and Censuses. They do not always occur in the same year.}\label{fig:timeline}
+\caption{Timeline of Australian elections and Censuses. Censuses happen regularly, elections not quite regularly, which means data needs to be interpolated by time.}\label{fig:timeline}
 \end{figure}
@@
 
+First, we show the estimated two-party preferred vote for an ``average'' electorate.
+
 \begin{figure}[H]
@@
 
-\caption{Estimated intercept for each election, which represents the two-party preferred vote for an electorate with average characteristics.}\label{fig:plotintercept}
+\caption{Estimated two-party preferred vote for an electorate with average characteristics at each election.}\label{fig:plotintercept}
 \end{figure}
 
-First, we show the estimated two-party preferred vote for an ``average'' electorate.
-Figure \ref{fig:plotintercept} shows that the baseline of party preference has varied over the elections, with the biggest swing occurring in the 2007 election where the mean electorate shifted more than five percentage points in favour of the Labor party.
+This shows that the baseline of party preference has varied over the elections, with the biggest swing occurring in the 2007 election where the mean electorate shifted more than five percentage points in favour of the Labor party.
 The dots represent the estimated 2PP value, and the lines indicate a 95\% confidence interval providing a guide to the uncertainty in the estimate.
+Where the vertical lines corresponding to each election cross the 50\% horizontal line, the 2PP vote for an average electorate is not statistically distinguishable from 50\% for that election.
 
@@
 \addcontentsline{toc}{subsection}{Age}
+
+Regions comprising more older people are often believed to be more conservative, and indeed we found that electorates with a higher median age are more likely to support the Liberal party --- although this effect is significant only in 2004 and 2007.
 
@@
 
-\caption{Estimated age effect for each election, which represents the impact of electorate median age on the two-party preferred vote. Higher values indicate a }\label{fig:age}
+\caption{Estimated impact of electorate median age on the two-party preferred vote.}\label{fig:age}
 \end{figure}
 
-Regions comprising more older people are often believed to be more conservative, and indeed we found that electorates with a higher median age are more likely to support the Liberal party --- although this effect is significant only in 2007 and 2010.
+This graph shows the effect of median age on the 2PP vote.
+The vertical scale can be interpreted as a score measuring how much the 2PP vote differs between an average electorate, and an electorate with median age in the top 10\% of all electorates (with all other socio-demographic variables unchanged).
+The reverse is true for young electorates --- the score can be interpreted as how much the 2PP vote would be reduced if an average electorate kept all socio-demographic variables fixed but median age was reduced to be in the bottom 10\% of all eletorates.
 
-Figure \ref{fig:age} shows the effect of median age on the 2PP vote.
-The vertical scale can be interpreted as a score measuring how much the 2PP vote differs between an average electorate, and an electorate with median age in the top third of all electorates.
+\hypertarget{income}{%
+\subsection*{Income}\label{income}}
+\addcontentsline{toc}{subsection}{Income}
 
-\begin{itemize}
-\tightlist
-\item
-  Would it be possible to show how the correlation changes for different age brackets (ie. younger voters = more Labor, and this swings more to Liberal as the age brackets gets older)?
-\item
-  If not, can you go into more detail specific age brackets? Eg. Is it the over-65s that skew Liberal?
-\end{itemize}
-
-\hypertarget{income-and-unemployment}{%
-\subsection*{Income and unemployment}\label{income-and-unemployment}}
-\addcontentsline{toc}{subsection}{Income and unemployment}
+Typically the Labor party campaigns on more progressive policies, which often include tax reform that adversely affects higher income earners, and more generous social assistance programs.
+Perhaps it is due to these policies that higher income electorates appear more likely to support the Liberal party, as the \texttt{Incomes} factor has a positive effect on Liberal preference.
+The effect also seems to have been increasing over time.
 
@@
 
-\caption{Estimated age effect for each election, which represents the impact of electorate median age on the two-party preferred vote. Higher values indicate a }\label{fig:income}
+\caption{Estimated impact of electorate income on the two-party preferred vote.}\label{fig:income}
 \end{figure}
 
-Typically the Labor party campaigns on more progressive policies, which often include tax reform that adversely affects higher income earners, and more generous social assistance programs.
-Perhaps it is due to these policies that higher income electorates appear more likely to support the Liberal party, as the \texttt{Incomes} factor has a positive effect on Liberal preference (see row 1 in Figure \ref{fig:partresplot1}).
-This effect is significant in every election aside from 2004, where it is only marginally insignificant (\(p = 0.0613\)).
+As before, the vertical scale can be interpreted as a score measuring how much the 2PP vote differs between an average electorate, and an electorate with income in the top 10\% of all electorates (with all other socio-demographic variables unchanged).
+
+\hypertarget{unemployment}{%
+\subsection*{Unemployment}\label{unemployment}}
+\addcontentsline{toc}{subsection}{Unemployment}
+
 Unemployment however, is not as influential.
-In 2001 and 2004, electorates with higher unemployment align with Labor, but over time this shifts towards support for the Liberal party, culminating in a significantly positive effect in 2016.
+In 2001 and 2004, electorates with higher unemployment align with Labor, but over time this shifts towards support for the Liberal party, culminating in a significantly positive (but small) effect in 2016.
+
+\begin{figure}[H]
+
+{\centering \includegraphics{elections_conversation_files/figure-latex/unemployment-1}
+
+}
+
+\caption{Estimated impact of electorate unemployment on the two-party preferred vote.}\label{fig:unemployment}
+\end{figure}
+
+The vertical scale can be interpreted as a score measuring how much the 2PP vote differs between an average electorate, and an electorate with unemployment in the top 10\% of all electorates (with all other socio-demographic variables unchanged)
 
@@
 
-Electorates with higher proportions of workers in mining, gas, water, agriculture, waste and electricity (grouped as \texttt{Extractive} industries) are consistently linked with higher support for the Liberal party, with the magnitude of this effect slightly increasing over the years (see row 3 in Figure \ref{fig:partresplot1}).
+Electorates with higher proportions of workers in extractive industries (mining, gas, water, agriculture, waste and electricity) are consistently linked with higher support for the Liberal party, with the magnitude of this effect slightly increasing over the years.
+
+\begin{figure}[H]
+
+{\centering \includegraphics{elections_conversation_files/figure-latex/extractive-1}
+
+}
+
+\caption{Estimated impact of employment in the extractive industries on the two-party preferred vote.}\label{fig:extractive}
+\end{figure}
+
 This is unsurprising, as the Liberal party has close ties with these traditional energy industries, and typically present policies to reduce taxation on energy production.
-Furthermore, electorates with more workers in construction or manufacturing industries (\texttt{Transformative}) are also more likely to support the Liberal party (see row 4 in Figure \ref{fig:partresplot1}).
 
-Similarly, the proportion of workers in managerial, administrative, clerical and sales roles (\texttt{ManagerAdmin}) is also a significant predictor of two-party preference vote across all six elections, with a higher proportion of people working these jobs increasing Liberal support.
-The magnitude of this effect also seems to increase over the years.
+Electorates with more workers in transformative industries (construction or manufacturing) are also more likely to support the Liberal party.
+
+\begin{figure}[H]
+
+{\centering \includegraphics{elections_conversation_files/figure-latex/transformative-1}
+
+}
+
+\caption{Estimated impact of employment in the transformative industries on the two-party preferred vote.}\label{fig:transformative}
+\end{figure}
+
+The proportion of workers in managerial, administrative, clerical and sales roles is also a significant predictor of two-party preference vote across all six elections, with a higher proportion of people working these jobs increasing Liberal support.
+
+\begin{figure}[H]
+
+{\centering \includegraphics{elections_conversation_files/figure-latex/manager-1}
+
+}
+
+\caption{Estimated impact of employment in management or administration on the two-party preferred vote.}\label{fig:manager}
+\end{figure}
 
@@
 
-In each of the six elections, electorates with a higher proportion of people that have recently (in the past five years) moved house (\texttt{DiffAddress}) are more likely to support the Liberal party, although this effect was marginally insignificant in 2013 (see row 6 in Figure \ref{fig:partresplot1}.
-Having controlled for characteristics of house ownership and rental prices (via the factors \texttt{PropertyOwned} and \texttt{RentLoan} respectively), this effect is somewhat surprising.
+In each of the six elections, electorates with a higher proportion of people that have recently (in the past five years) moved house are more likely to support the Liberal party.
+
+\begin{figure}[H]
+
+{\centering \includegraphics{elections_conversation_files/figure-latex/mobility-1}
+
+}
+
+\caption{Estimated impact of household mobility on the two-party preferred vote.}\label{fig:mobility}
+\end{figure}
+
+Our analysis controls for characteristics of home ownership and rental prices, so this effect is not simply due to electorates having low rates of home ownership, or due to electorates having high rental prices.
+Instead, it suggests that people who are more transient are also more likely to be conservative voters, regardless of their home ownership or rental status.
+(This would need further study, as we do not have individual level voting data.)
 
@@
 
+\begin{figure}[H]
+
+{\centering \includegraphics{elections_conversation_files/figure-latex/defacto-1}
+
+}
+
+\caption{Estimated impact of de facto relationships on the two-party preferred vote.}\label{fig:defacto}
+\end{figure}
+
 \hypertarget{education}{%
@@
 
-Since 2007, electorates with higher education levels are associated with supporting the Labor party, although this effect is significant only in 2016.
-Before 2007, education has an almost zero effect (see row 3 in Figure \ref{fig:partresplot1}).
+Since 2007, electorates with higher education levels are associated with supporting the Labor party, although this effect is significant only in 2016. Before 2007, education had a negligible effect.
+
+\begin{figure}[H]
+
+{\centering \includegraphics{elections_conversation_files/figure-latex/education-1}
+
+}
+
+\caption{Estimated impact of education on the two-party preferred vote.}\label{fig:education}
+\end{figure}
 
@@
 Larger migrant populations from Asia, the Middle East, South-Eastern Europe, the United Kingdom and elsewhere, are either associated with Labor support, or have no effect.
-Of these areas, only South-Eastern European populations appear significant in each election, with the proportion of Asian migrants also being significant in 2010.
-Speaking other languages (aside from English) however, appears to have a far stronger effect, as observed through the \texttt{OtherLanguage} variable.
-Electorates with more diverse speech are associated with higher support for the Liberal party from 2004 onwards, with this effect being significant in 2007, 2010 and 2016.
-Furthermore, of the variables relating to religion, only Judaism shows a consistent effect, with electorates with relatively large Jewish populations more likely to vote Liberal.
+Of these areas, only differences in South-Eastern European populations appear to have had a significant impact in each election, with the proportion of Asian migrants also being significant in 2010.
 
-\hypertarget{a-closer-look-at-the-residuals}{%
-\section{A closer look at the residuals}\label{a-closer-look-at-the-residuals}}
+Speaking languages other than English however, appears to have a far stronger effect.
+Electorates with more diverse languages are associated with higher support for the Liberal party from 2004 onwards, with this effect being significant in 2007, 2010 and 2016.
 
-\hypertarget{residuals-by-state}{%
-\subsection{Residuals by state}\label{residuals-by-state}}
+\begin{figure}[H]
 
-It is often hypothesized that states have systematic differences that cause their electorates to vote differently.
-Boxplots of residuals grouped by state (Figure \ref{fig:resstate}) reveal that the data reflects this -- there appears to be a state-specific effect not captured by the models.
-Tasmania and the Australian Capital Territory appear to have a bias towards Labor, whereas the Northern Territory tends towards voting Liberal.
-However, there are relatively few electorates in each of these states (five, two and two respectively), so this apparent result may be due to incumbent effects rather than an actual state-specific bias.
-
-\begin{figure}[h]
-
-{\centering \includegraphics[width=0.7\linewidth,]{elections_conversation_files/figure-latex/resstate-1}
+{\centering \includegraphics{elections_conversation_files/figure-latex/otherlanguage-1}
 
@@
 
-\caption{Boxplot of residuals by state with jittered points. States ordered by median residual. A state-specific bias not captured by the model is evident.}\label{fig:resstate}
+\caption{Estimated impact of ethnic diversity on the two-party preferred vote.}\label{fig:otherlanguage}
 \end{figure}
 
-\hypertarget{outlier-electorates}{%
-\subsection{Outlier electorates}\label{outlier-electorates}}
+Of the variables relating to religion, only Judaism shows a consistent effect, with electorates with relatively large Jewish populations more likely to vote Liberal.
+
+\hypertarget{state-effects}{%
+\section{State effects}\label{state-effects}}
+
+It is often suggested that states have systematic differences that cause their electorates to vote differently.
+We can explore this effect by looking at the difference between the actual 2PP vote in each electorate and what our modelling would predict based only on electorate characteristics.
+We found a bias toward Labor in the ACT, and to a smaller extent in Tasmania, and a bias toward the Liberals in the Northern Territory.
+No other states showed a significantly different voting pattern from what you would expect using only socio-demographic information.
+That suggests that any differences between Queensland and Victoria (for example) are due to different voter socio-demographics.
+
+\hypertarget{against-the-tide}{%
+\section{Against the tide}\label{against-the-tide}}
+
+As well as looking at whether some states have results different from what is predicted, we can also look at individual electorates.
+That is, when does an electorate vote very differently from what their socio-demographics would suggest?
+This suggests something is going on beyond the effect of socio-demographics.
+For example, a very good (or bad) local member could lead to voting patterns that are very different from what is predicted.
 
@@
 
-\hypertarget{from-emil}{%
-\section{From Emil}\label{from-emil}}
-
-In terms of structure, start with a brief paragraph or two on what the research is, and what the reader can expect to find below.
-
-Then, break the article into small sections for each of the factors that have an effect (a short paragraph or two for each factor).
-And it would be great to include examples of specific seats where the relationship was strong, or interesting for other reasons.
-
-So the sections will be:
-
-\begin{itemize}
-\tightlist
-\item
-  Age
-\item
-  Income
-\item
-  Unemployment
-\item
-  Industry and type of work
-\item
-  Education
-\item
-  Diversity
-\item
-  Household mobility
-\item
-  Relationships
-\end{itemize}
-
-An example of this structure can be seen here:
-\url{https://theconversation.com/confused-about-aged-care-in-the-home-these-10-charts-explain-how-it-works-113923}
-
-Questions on particular factors:
-
-Incomes:
-- Was this correlation stronger in 2016? (and increasingly influential?)
-
-Household mobility
-- In the report, you say: ````Having controlled for characteristics of house ownership and rental prices'', whis effect is somewhat surprising''.
-Why is this surprising to you?
-(I would have expected this to support Labor more)
-
-Age
-- Would it be possible to show how the correlation changes for different age brackets (ie. younger voters = more Labor, and this swings more to Liberal as the age brackets gets older)?
-- If not, can you go into more detail specific age brackets? Eg. Is it the over-65s that skew Liberal?
-
-Diversity
-- Migrant population: Would it possible to break this out into particular groups?
-eg.
-Seats with higher Chinese population correlates with Liberal, seats with higher SEAsia pop correlates with Labor, etc?
-- The same for languages?
-
 \printbibliography

### 2019-05-08 wrangle.R

@@ -84,7 +84,7 @@ my_df <- bind_rows(
     LNP_Votes, ALP_Votes, ALP_Percent, TotalVotes, Swing,
     InternetUse, InternetAccess, EnglishOnly,
     Other_NonChrist, OtherChrist, Catholic, Anglican,
-    Volunteer, EmuneratedElsewhere
+    Volunteer, EmuneratedElsewhere, DivisionID
   ))
 
 # Create final df for modelling

### 2019-05-08 elections_conversation.tex
Finished for now

@@
 
-A simple way to measure voting patterns is to consider the two-party preferred (2PP) vote, which is based on the tally of preferences for the Labor and Liberal candidates, ignoring all other candidates.
-By convention, this is recorded as a percentage preference in favour of the Liberal party --- for example, a 2PP value of 45\% indicates that 45\% of voters ranked the Liberal candidate higher than the Labor candidate, while the remaining 55\% ranked them in the reverse order.
+A simple way to measure voting patterns is to consider the two-party preferred (2PP) vote, which is based on the tally of preferences for the Labor and Coalition candidates, ignoring all other candidates.
+We express this as a percentage preference in favour of the Coalition --- for example, a 2PP value of 45\% indicates that 45\% of voters ranked the Coalition candidate higher than the Labor candidate, while the remaining 55\% ranked them in the reverse order.
 
@@
 In this article, we highlight the variables with the strongest relationship to the two-party preferred vote, or which have had substantial changes over time.
-The full analysis is available at \url{https://robjhyndman.com/publications/elections/}.
 
@@
 
-This shows that the baseline of party preference has varied over the elections, with the biggest swing occurring in the 2007 election where the mean electorate shifted more than five percentage points in favour of the Labor party.
+This shows that the baseline of party preference has varied over the elections, with the biggest swing occurring in the 2007 election where the average electorate shifted more than five percentage points in favour of the Labor party.
 The dots represent the estimated 2PP value, and the lines indicate a 95\% confidence interval providing a guide to the uncertainty in the estimate.
@@
 
-Regions comprising more older people are often believed to be more conservative, and indeed we found that electorates with a higher median age are more likely to support the Liberal party --- although this effect is significant only in 2004 and 2007.
+Regions comprising more older people are often believed to be more conservative, and indeed we found that electorates with a higher median age are more likely to support the Coalition party.
 
@@
 The vertical scale can be interpreted as a score measuring how much the 2PP vote differs between an average electorate, and an electorate with median age in the top 10\% of all electorates (with all other socio-demographic variables unchanged).
-The reverse is true for young electorates --- the score can be interpreted as how much the 2PP vote would be reduced if an average electorate kept all socio-demographic variables fixed but median age was reduced to be in the bottom 10\% of all eletorates.
+The reverse is true for young electorates --- the score can be interpreted as how much the 2PP vote would be reduced if an average electorate kept all socio-demographic variables fixed but median age was reduced to be in the bottom 10\% of all electorates.
+
+When the uncertainty range covers the horizontal 0 line, the resulting impact can be considered statistically undetectable. Here age is a statistically significant effect only in 2004 and 2007.
 
@@
 Typically the Labor party campaigns on more progressive policies, which often include tax reform that adversely affects higher income earners, and more generous social assistance programs.
-Perhaps it is due to these policies that higher income electorates appear more likely to support the Liberal party, as the \texttt{Incomes} factor has a positive effect on Liberal preference.
+Perhaps it is due to these policies that higher income electorates appear more likely to support the Coalition, as the Incomes factor has a positive effect on Coalition preference.
 The effect also seems to have been increasing over time.
@@
 Unemployment however, is not as influential.
-In 2001 and 2004, electorates with higher unemployment align with Labor, but over time this shifts towards support for the Liberal party, culminating in a significantly positive (but small) effect in 2016.
+In 2001 and 2004, electorates with higher unemployment align with Labor, but over time this shifts towards support for the Coalition, culminating in a significantly positive (but small) effect in 2016.
 
@@
 
-Electorates with higher proportions of workers in extractive industries (mining, gas, water, agriculture, waste and electricity) are consistently linked with higher support for the Liberal party, with the magnitude of this effect slightly increasing over the years.
+Electorates with higher proportions of workers in extractive industries (mining, gas, water, agriculture, waste and electricity) are consistently linked with higher support for the Coalition, with the magnitude of this effect slightly increasing over the years.
 
@@
 
-This is unsurprising, as the Liberal party has close ties with these traditional energy industries, and typically present policies to reduce taxation on energy production.
+This is unsurprising, as the Coalition has close ties with these traditional energy industries, and typically present policies to reduce taxation on energy production.
 
-Electorates with more workers in transformative industries (construction or manufacturing) are also more likely to support the Liberal party.
+Electorates with more workers in transformative industries (construction or manufacturing) are also more likely to support the Coalition.
 
@@
 
-The proportion of workers in managerial, administrative, clerical and sales roles is also a significant predictor of two-party preference vote across all six elections, with a higher proportion of people working these jobs increasing Liberal support.
+The proportion of workers in managerial, administrative, clerical and sales roles is also a significant predictor of two-party preference vote across all six elections, with a higher proportion of people working these jobs increasing Coalition support.
 
@@
 
-In each of the six elections, electorates with a higher proportion of people that have recently (in the past five years) moved house are more likely to support the Liberal party.
+In each of the six elections, electorates with a higher proportion of people that have recently (in the past five years) moved house are more likely to support the Coalition.
 
@@
 Speaking languages other than English however, appears to have a far stronger effect.
-Electorates with more diverse languages are associated with higher support for the Liberal party from 2004 onwards, with this effect being significant in 2007, 2010 and 2016.
+Electorates with more diverse languages are associated with higher support for the Coalition from 2004 onwards, with this effect being significant in 2007, 2010 and 2016.
 
@@
 
-Of the variables relating to religion, only Judaism shows a consistent effect, with electorates with relatively large Jewish populations more likely to vote Liberal.
+Of the variables relating to religion, only Judaism shows a consistent effect, with electorates with relatively large Jewish populations more likely to vote for the Coalition.
 
@@
 We can explore this effect by looking at the difference between the actual 2PP vote in each electorate and what our modelling would predict based only on electorate characteristics.
-We found a bias toward Labor in the ACT, and to a smaller extent in Tasmania, and a bias toward the Liberals in the Northern Territory.
+We found a bias toward Labor in the ACT, and to a smaller extent in Tasmania, and a bias toward the Coalition in the Northern Territory.
 No other states showed a significantly different voting pattern from what you would expect using only socio-demographic information.
@@
 
-Based on the distribution of the Cook's distance values, a Cook's distance greater than \(0.1\) is considered to be influential and a potential outlier.
-The electorate of Sydney (NSW) has a large Cook's distance from 2001 to 2013, due to its diverse population (language, birthplace and religion), high number of defacto relationships, high income, high household mobility and small amount of workers in extractive and transformative jobs.
-It has remained a strong supporter of the Labor party and the Liberal vote is severely overpredicted by the model, making it an outlier.
-Nearby in metropolitan NSW, the electorate of Wentworth is found to be an outlier in all but the 2007 election.
-Although historically Liberal, its two-party vote jumped by over 10 percentage points in 2010 without experiencing any notable changes in its socio-demographic makeup --- implying that this may be the direct effect of its Liberal member, Malcolm Turnbull, becoming the leader of the Liberal party.
+By this measure, the top five most surprising results in the last six elections were the following:
+
+\begin{tabular}{llrr}
+\toprule
+Electorate & Election & Actual 2PP & Predicted 2PP\\
+\midrule
+Capricornia & 2004 & 44.9 & 60.6\\
+Capricornia & 2001 & 43.1 & 58.8\\
+Herbert & 2013 & 56.2 & 41.6\\
+Wentworth & 2013 & 67.7 & 53.1\\
+Herbert & 2016 & 50.0 & 35.8\\
+\bottomrule
+\end{tabular}
+
+Capricornia is on the Queensland coast, taking in Rockhampton, Yeppoon and parts of Mackay.
+It was won in 2001 and 2004 by Labour MP Kirsten Livermore, defying the socio-demographic prediction which suggested it would be a safe Coalition seat.
+She also won it in 2007 and 2010, but with a smaller number of votes, so the result was not so different from predicted.
+It is currently held by Michelle Landry for the Coalition.
+
+Herbert is another Queensland seat taking in Townsville.
+It was won in 2013 for the Coalition by Ewen Jones, and in 2016 by Cathy O'Toole for Labor by just 37 votes.
+The model predicts a much lower vote for the Coalition than tends to happen in practice.
+
+Wentworth probably needs no explanation --- the seat formerly held by Malcolm Turnbull which he won with a big majority in 2013.
+Its two-party vote jumped by over 10 percentage points in 2010 without experiencing any notable changes in its socio-demographic makeup --- implying that this may be the direct effect of its Liberal member, Malcolm Turnbull, becoming the leader of the Liberal party.
 Liberal support in Wentworth is underpredicted by the model in each year, and more so with Turnbull as Liberal leader.
 
-Lingiari, an electorate taking up almost all of the Northern Territory, is an outlier in the 2001--2007 elections due to its large Indigenous population, young age profile and low rates of property ownership.
-Fowler (NSW) has a diverse population with a high proportion of migrants, many Buddhists and Muslims, and has strong Labor support, making it influential in 2001, 2004 and 2010.
-Other electorates with large Cook's distance are Barton (NSW) and Leichhardt (QLD) in 2016, and Canberra (ACT) in 2007.
+\hypertarget{true-blue-seats}{%
+\section{True Blue Seats}\label{true-blue-seats}}
 
-\hypertarget{conclusion}{%
-\section{Conclusion}\label{conclusion}}
+Another way the model can be used is to identify the seats with the strongest propensity to vote one way or the other, based on their socio-demographics.
 
-This paper explores the effects of electoral socio-demographic characteristics on the two-party preferred vote in the 2001--2016 elections, using information from the corresponding Australian federal elections and Censuses.
-As a Census does not always occur in the same year as an election, Census data for the 2004--2013 elections are generated by employing a method of spatio-temporal imputation.
-This imputes electoral socio-demographics for the electoral boundaries in place at the time of the election --- an approach that is distinctly different from previous work on modelling election outcomes, where Census and election data are typically joined without addressing their temporal differences.
-Before estimating a model, these socio-demographic variables are standardized (to adjust for changing variable scales) and many variables (representing similar information) are combined into factors, resulting in a reduced predictor set.
-A spatial error model is then estimated for each election, accounting for the inherent spatial structure of the data.
+The top five most blue-ribbon Coalition seats in the last six elections are listed below.
+Mitchell takes in the north-western outer suburbs of Sydney, including Baulkham Hills, Kellyville, and parts of Castle Hill.
+Maranoa occupies a large part of south west Queensland including towns such as Charleville, Kingaroy and Winton.
+Mallee is, of course, the north west corner of Victoria including Mildura, Swan Hill and Stawell.
 
-Across the past six elections, most of the socio-demographics that drive the electoral two-party preferred vote are found to remain steady, whilst a few (typically weaker) effects vary over time.
-Industry and type of work are particularly influential, with energy-related and manufacturing/construction jobs, as well as administrative roles being strongly linked with the Liberal party in all elections.
-Incomes have a similarly consistent effect, with higher income areas supporting Liberal.
-Higher levels of unemployment shift from weak association with Labor to a significant Liberal effect over the years, and higher education levels are associated with Labor from 2007 (although significant only in 2016).
-It is also found that electorates with higher household mobility support Liberal, birthplace diversity favours Labor and more de facto relationships align with Labor preference --- although marriages, family and household sizes have no material influence.
-Furthermore, the neighbourhood (spatial) effects are found to be positive in all elections, although significant only in 2001 and 2016, meaning that in the 2004--2013 elections, electorates effectively voted independently.
+\begin{tabular}{llrr}
+\toprule
+Electorate & Election & Actual 2PP & Predicted 2PP\\
+\midrule
+Mitchell & 2001 & 71.3 & 76.4\\
+Mitchell & 2004 & 70.7 & 75.9\\
+Maranoa & 2013 & 72.3 & 75.6\\
+Maranoa & 2016 & 67.5 & 73.9\\
+Mallee & 2013 & 73.7 & 73.8\\
+\bottomrule
+\end{tabular}
 
-The findings in this paper complement the existing literature by modelling temporal trends, which as far as the authors are aware, has not been done previously for Australian elections using a regression framework.
-It is also the first study to model any Australian election since 2010 using Census information.
+The top five seats most likely to vote Labour, based on their socio-demographics, are listed below.
+Batman has since been abolished, but occupied the inner north of Melbourne.
+Gellibrand takes in the inner western suburbs of Melbourne including Altona and Williamstown, while Fowler is in the western suburbs of Sydney, including Cabramatta and Liverpool.
 
-Additionally, a key contribution of this research is the wrangling of the raw data and imputed data sets for the 2004, 2007, 2010 and 2013 elections, which have been contributed to the \texttt{eechidna} \texttt{R} package --- providing a rich, accessible data resource for future Australian electoral analysis.
+\begin{tabular}{llrr}
+\toprule
+Electorate & Election & Actual 2PP & Predicted 2PP\\
+\midrule
+Batman & 2007 & 24.1 & 22.3\\
+Batman & 2010 & 25.1 & 22.5\\
+Fowler & 2004 & 28.6 & 23.6\\
+Fowler & 2001 & 28.5 & 24.9\\
+Gellibrand & 2010 & 26.1 & 25.2\\
+\bottomrule
+\end{tabular}
 
@@
 
-This paper was produced using \texttt{RMarkdown} \autocite{rmarkdown} and \texttt{knitr} \autocite{knitr}.
-All corresponding code for this paper can be found in the github repository \href{https://github.com/jforbes14/eechidna-paper}{github.com/jforbes14/eechidna-paper}, and the data used is available in the \texttt{eechidna} package \autocite{eechidna}.
-All raw data was obtained from the Australian Electoral Commission, the Australian Bureau of Statistics and the Australian Government.
+This analysis is based on a research paper coauthored by Rob Hyndman and Di Cook from the Department of Econometrics and Business Statistics at the Monash Business School, and Jeremy Forbes, a former Monash University Honours student in econometrics.
+The full analysis is available at \url{https://robjhyndman.com/publications/elections/}.
 
-\hypertarget{software}{%
-\section{Software}\label{software}}
+The code used for the analysis can be found in the github repository \href{https://github.com/jforbes14/eechidna-paper}{github.com/jforbes14/eechidna-paper}.
 
-All election and Census datasets, along with electoral maps and more, are available in the \texttt{eechidna} (Exploring Election and Census Highly Informative Data Nationally for Australia) \texttt{R} package, which can be downloaded from CRAN.
-The \texttt{eechidna} package makes it easy to look at the data from the Australian Federal elections and Censuses that occurred between 2001 and 2016.
-This study contributed a large revision to the \texttt{eechidna} package, which included the addition of election and Census data for 2001--2010, voting outcomes for polling booths and imputed Census data for election years.
-For more details on using \texttt{eechidna}, please see the articles (vignettes) on the github page \href{https://ropenscilabs.github.io/eechidna/}{ropenscilabs.github.io/eechidna/}.
-
-The authors would like to sincerely thank Anthony Ebert, Heike Hofmann, Thomas Lumley, Ben Marwick, Carson Sievert, Mingzhu Sun, Dilini Talagala, Nicholas Tierney, Nathaniel Tomasetti, Earo Wang and Fang Zhou, all of whom have contributed to the \texttt{eechidna} package.
+All election and Census data since 2001, along with electoral maps and more, are available in the \href{https://cran.r-project.org/package=eechidna}{\texttt{eechidna}} (Exploring Election and Census Highly Informative Data Nationally for Australia) \texttt{R} package, which can be downloaded from CRAN.
+The eechidna package uses data obtained from the Australian Electoral Commission, the Australian Bureau of Statistics and the Australian Government.
 

### 2019-05-08 functions.R

@@ -132,6 +132,7 @@ my_fgls <- function(my_formula, my_data, sp_weights) {
   gls_model$gls_data <- gls_data
   gls_model$my_data <- my_data
   gls_model$actual_residuals <- solve(trans_mat) %*% gls_model$residuals
+  gls_model$predictions <- fitted(ols_model) # Use OLS predictions later
 
   return(gls_model)
 }

### 2019-06-08 submission_250419/anzjs_elections_250419.Rmd
Original submission files as submitted on 25 April 2019.

--- 2019-04-25 paper.Rmd
@@
 Based on the distribution of the Cook's distance values, a Cook's distance greater than $0.1$ is considered to be influential and a potential outlier.
-The electorate of Sydney (NSW) has a large Cook's distance from 2001 to 2013, due to its diverse population (language, birthplace and religion), high number of defacto relationships, high income, high household mobility and small amount of workers in extractive and transformative jobs.
+The electorate of Sydney (NSW) has a large Cook's distance from 2001 to 2013, due to its diverse population (language, birthplace and religion), high number of de facto relationships, high income, high household mobility and small amount of workers in extractive and transformative jobs.
 It has remained a strong supporter of the Labor party and the Liberal vote is severely overpredicted by the model, making it an outlier.

### 2019-06-08 response_to_review.Rmd
Comments to be addressed from the editor, associate editor and two reviewers.

---
title: "Response to Reviewer Feedback"
author: "Jeremy Forbes"
date: "08/06/2019"
output: html_document
---

## Editor's comments

*1. Address reviewer 1's question about why spatial imputation is required given that the census is reported at a low level of aggregation (lines 150-151) that could presumably be clumped to give a good approximation to electorate boundaries.*

...

*2. I note that you do mention the ecological fallacy (lines 285-286) but reviewer 1's reaction suggests that you should expand your remarks here.*

...

*3. It appears that your approach to variable selection has confused everybody. My understanding of what you have done is that:*
*- You begin with 65 variables (are they all the available census data or a subset?).*

...

*- Then you standardise the variables.*

*- Then you remove some variables to do with Age Bands and Christianity (how many variables out of the 65 does this leave? Were these variables totally removed or lumped in some way? Christianity still appears in table 2 so they obviously weren't all removed) *

...

*- Then you use the PCA to identify variables that covary. If variables appear together in a PC with loading higher than 0.15 they are combined together into a factor, e.g. a new variable Income replaces two previously existing variables. (In the description of this it is unclear how many variables are being combined in each factor,  the 'ands' and commas make things tricky, .e.g  is 'renting and government housing' one variable or two? Do the factors use the loading weights or just the signs?)*

...

*- The steps above get you from 65 variables down to 30 (I can count 10 variables that  get accounted for via the creation of factors, but that doesn't get from 65 to 30 so presumably some other variables have been removed? Or is this difference all in the age bands and variants of Christianity? More clarity is required)*

...

*4. As you note in the paper, it seems likely that there are still issues with multi-collinearity. Figs 6 and 7 show the variables with significant effects, do the results change much if you fit a model with just these 13 variables instead of the 30?*

...

## Associate Editor's comments
*Some conceptually simple imputation methods are used to align polling and census data to investigate associations between census variables and voting behaviour at electoral division level.*

...

*The main thing I found confusing about this paper was the variable selection procedure. It seems to be heading in the direction of four PCs, then it veers into six factors before ending with 30 variables. This all requires greater clarity and justification.*

...

*The authors opt for “visualization” of the fixed effects rather than the usual table of estimated parameters. Perhaps this is a good choice given that there are so many. But then spatial parameters are exhibited in a huge table. This seems inconsistent.*

...

*I found the writing unclear in a few places. What are the “time-varying restrictions” on a variable (line228)? And the description of a partial residual (line 276) doesn’t make sense to me.*

...

*Some of the interpretations of results seem a little naïve. For example, does Figure 4 really justify concluding that “electorates effectively voted independently” except for 2001 and 2016? (Is it a coincidence that the spatial autoregression is greatest in 2001 and 2016 where there is no need for imputation?)*

...

## Reviewer 1's comments

*Its goal is to account for – though the word ‘explain’ is used! – patterns of voting at Australian federal elections, using ecological data, but with no reference to the ecological fallacy.*

...

*The paper is written in virtual total ignorance of the large amount of work done on Australian voting patterns; apart from reference to a few papers by geographers there is not a single mention of work by political scientists – of which there is a great deal. As a consequence there is no theoretical background whatsoever, let alone hypotheses to be tested. It is banal empiricism: throw a lot of data (what are called ‘electoral socio-demographic variables’) at something and see what comes out.*

...

*The banality of this approach is illustrated by the finding that ‘more de facto relationships’ are associated with greater support for Labor. No rationale for this is offered – why should people in such relationships be more likely to vote Labor? And while there may be a statistically significant relationship (i.e. slope coefficient) it may be substantively trivial – a fact hidden because the data are transformed. If there is a small SD then the ‘real’ difference between a place with many de facto relationships and another with few will be small and will not account for much of the absolute variation in Labour support at all. Further, since all of the relationships with individual variables are partial, it may well be that when ‘all other variables are held constant’ the difference in Labour support is trivial.*

...

*Much is done – unnecessarily in my view – in preparing the data through imputation. This has two components. First there is the creation of census data for the 150 constituencies because boundaries of the large census reporting areas may not coincide with the constituencies’ at the time of an election. Imputation is deployed by assuming that if 50% of area A is in district X then 50% of A’s population is allocated to X without any recognition that area A may be internally heterogeneous. But that is in any case unnecessary. Australian census data are reported – as the authors note – for very small areas with populations of only a few hundred each, and there are shape files associated with these areas. It is thus a straightforward task to amalgamate the small areas into the electoral districts and get the ‘real’ data for each without any (or at best a very, very small amount of) imputation.*

...

*The second component involves imputing the values for each Census variable at the date of the election, which can never be more than three years – at most – distant from the nearest Census (either before or after the election). This seems totally unnecessary; the likelihood of any substantial change in the relative position of any district on any variable sufficient to impact on the regression outcomes is so minimal that a lot of effort could have been saved. It just wasn’t worth doing.*

...

*The authors recognise that there is likely to be collinearity among 65 Census variables – why those; we are never told?  And so, as others do, they did a pca of the data matrix, reducing it to four components – we are not told what rotation was employed to achieve simple structure. We are not given the results of this analysis, merely a naming of the six factors identified – how do you get six ‘factors’ from four components? That naming was based on the loadings and, we are told, that any loading greater than 0.15 (presumably either positive or negative) was ‘considered large’. Since loadings are the equivalent of correlations then a loading of 0.15 means that the component accounts for just 2.25 per cent of the variation in that variable. This is – quite frankly – ludicrous.*

...

*My expectation was that they would then fit a model using the four components as the independent variables, thus avoiding any collinearity and confounding. But they didn’t. Instead they selected 30 of the 65 variables – which must still be collinear; the resulting equation in Table 2 clearly implies this given the large number of (a) insignificant relations (normal with collinearity – see a recent paper in Quality and Quantity), (b) unreasonably large coefficients (-8.07 for de facto, for example), and (c) very large standard errors for many of the variables.*

...

*The authors fit a spatial autoregressive model, on the assumption that neighbouring places are more likely to be similar than distant places (the ‘first law of geography’). This is only a significant finding at two of the elections and is wrongly interpreted: it does not necessarily show that then ‘the political climate of an electorate appears to be affected by the attitudes of its neighbours’ but, much more likely, that neighbouring places share characteristics in common that affect voting behaviour but are not captured by the independent variables included in the model. Spatial autocorrelation may reflect spatial diffusion processes – but in which direction? – but is much more likely to reflect an under-specified model.*

...

*Having fitted the model we than get two pages of ‘interpretation’, which is inductivism of the worst kind – we’ve found some relationships, how do we account for them?*

...

## Reviewer 2's comments

*1. It seems very strange to assume that population is uniform within electorates (line 133) so that weighting is proportional to area, and not to the number of people in an area of intersection. An electorate on the urban fringe may be strongly affected by loss of a relatively small but highly populated area but little affected by incorporating a large rural area. A simple indirect way to get at this might be to address the question: Would the results change much if A_{s,t} was log area, for each s,t ?*

...

*2. It seems to me that the use of Principal Components analysis is unclear.*
*a. PC appears to have been used as a model selection criterion, with the variables related to the first four PCs included into the model in Table 2. However typically when principal components are computed they only involve the explanatory variables, not the response, and so there is no guarantee that all the first four PCs (and only them) will be related to the response. For example using the eechidna data for 2016 I found the fifth PC was extremely significant. But this is a minor point.*

...

*b. The variables which load highly on a PC are likely to be highly correlated, and so if these high-loading variables are used together in the regression model we may again have multicollinearity. I found this was the case when applying the author’s regression model to the 2016 data. This may be why several of the predictors in Table 2 are never related significantly related to the response in any of the years.*

...

*c. Conversely there could be a variable which is independent of the other variables (so representing 1 / 65th of the explanatory information) but by dint of that fact not be included in the first four PCs. In short, the narrative leaves me unconvinced that all the variables included are necessary or that all the relevant variables are included in the regression.*

...

*d. I presume the “factors” referred to lines 210-220 were created as the first principal component of the constituent raw variables?*

...

*e. The description of the variables included in “Education” is not clear.*

...

*f. I could not get the variable ‘OtherLanguageHome’ in 2016 to be significant, but it also had a high VIF with other explanatory variables. Please check this.*

...

*3. I am not convinced it is reasonable to summarise the effect of age by median age. Perhaps age effects are more at the edges of the age distribution than in the middle? I would have thought the percentage of the population in different age bands e.g. 20-34 years and Over 65years, might be significant. For 2016 the % aged 20-34 years was highly significant and removed median age.*

...

*4. Is there any tendency for the vote in an election to depend on how long the incumbent has been in power? In your cross-sectional models this would be confounded with any other effects in the intercept, but a comment would be nice to know. It would help in terms of interpreting the variations in intercept, and setting a reader’s mind to rest that the issue has been considered. Perhaps a something like a simple regression of average vote (or the intercept) on length of incumbency could address the issue.*

...

### 2019-06-08 anzjs_elections.Rmd

--- 2019-05-01 elections.Rmd
@@
 title: "Spatial modelling of the two-party preferred vote in Australian federal elections: 2001--2016"
-author:
-- familyname: Forbes
-  othernames: Jeremy
-  address: Monash University\newline Melbourne, Australia
-  email: jeremyforbes1995@gmail.com
-  correspondingauthor: true
-- familyname: Cook
-  othernames: Dianne
-  address: Monash University\newline Melbourne, Australia
-- familyname: Hyndman
-  othernames: Rob J
-  address: Monash University\newline Melbourne, Australia
-abstract: "We examine the relationships between electoral socio-demographic characteristics and two-party preference in the six Australian federal elections held between 2001 to 2016.
+author: Jeremy Forbes$^*$, Dianne Cook, Rob J Hyndman
+toc: false
+documentclass: anzsauth
+classoption: times, doublespace
+fontfamily: times
+bibliography: references.bib
+biblio-style: anzsj
+output: 
+  bookdown::pdf_document2:
+    fig_caption: yes
+    citation_package: natbib
+    keep_tex: yes
+    number_sections: yes
+    fig_height: 5
+    fig_width: 8
+    includes:
+      in_header: anzjs_preamble.tex
+---
+
+# Summary{-}
+
+We examine the relationships between electoral socio-demographic characteristics and two-party preference in the six Australian federal elections held between 2001 to 2016.
 Socio-demographic information is derived from the Australian Census, which occurs every five years.
@@
 It is also found that between 2004 and 2013, the spatial effect was insignificant, meaning that electorates voted effectively independently.
-All data featured in this study has been contributed to the `eechidna` `R` package (available on CRAN)."
-keywords: "federal election, Census, Australia, spatial modelling, imputation, data science, socio-demographics, electorates, R, eechidna"
-wpnumber:
-jelcodes: C31, C33, D72
-blind: false
-cover: false
-toc: false
-bibliography: references.bib
-biblio-style: authoryear-comp
-header-includes: \usepackage{float}
-output:
-  MonashEBSTemplates::workingpaper:
-    fig_caption: yes
-    fig_height: 5
-    fig_width: 8
-    includes:
-      in_header: preamble.tex
-    keep_tex: yes
-    number_sections: yes
-    citation_package: biblatex
----
+All data featured in this study has been contributed to the `eechidna` `R` package (available on CRAN).
+
+**Keywords:** federal election, Census, Australia, spatial modelling, imputation, data science, socio-demographics, electorates, R, eechidna
 

### 2019-06-08 elections.Rmd

@@
 keywords: "federal election, Census, Australia, spatial modelling, imputation, data science, socio-demographics, electorates, R, eechidna"
-wpnumber:
+wpnumber: 08/19
 jelcodes: C31, C33, D72
 blind: false
-cover: false
+cover: true
 toc: false

### 2019-06-08 elections_conversation.Rmd

@@
 
-A simple way to measure voting patterns is to consider the two-party preferred (2PP) vote, which is based on the tally of preferences for the Labor and Liberal candidates, ignoring all other candidates.
-By convention, this is recorded as a percentage preference in favour of the Liberal party --- for example, a 2PP value of 45% indicates that 45% of voters ranked the Liberal candidate higher than the Labor candidate, while the remaining 55% ranked them in the reverse order.
+A simple way to measure voting patterns is to consider the two-party preferred (2PP) vote, which is based on the tally of preferences for the Labor and Coalition candidates, ignoring all other candidates.
+We express this as a percentage preference in favour of the Coalition --- for example, a 2PP value of 45% indicates that 45% of voters ranked the Coalition candidate higher than the Labor candidate, while the remaining 55% ranked them in the reverse order.
 
@@
 In this article, we highlight the variables with the strongest relationship to the two-party preferred vote, or which have had substantial changes over time.
-The full analysis is available at https://robjhyndman.com/publications/elections/.
 
@@
 
-This shows that the baseline of party preference has varied over the elections, with the biggest swing occurring in the 2007 election where the mean electorate shifted more than five percentage points in favour of the Labor party.
+This shows that the baseline of party preference has varied over the elections, with the biggest swing occurring in the 2007 election where the average electorate shifted more than five percentage points in favour of the Labor party.
 The dots represent the estimated 2PP value, and the lines indicate a 95% confidence interval providing a guide to the uncertainty in the estimate.
@@
 The vertical scale can be interpreted as a score measuring how much the 2PP vote differs between an average electorate, and an electorate with median age in the top 10% of all electorates (with all other socio-demographic variables unchanged).
-The reverse is true for young electorates --- the score can be interpreted as how much the 2PP vote would be reduced if an average electorate kept all socio-demographic variables fixed but median age was reduced to be in the bottom 10% of all eletorates.
+The reverse is true for young electorates --- the score can be interpreted as how much the 2PP vote would be reduced if an average electorate kept all socio-demographic variables fixed but median age was reduced to be in the bottom 10% of all electorates.
+
+When the uncertainty range covers the horizontal 0 line, the resulting impact can be considered statistically undetectable. Here age is a statistically significant effect only in 2004 and 2007.
 
@@
 Typically the Labor party campaigns on more progressive policies, which often include tax reform that adversely affects higher income earners, and more generous social assistance programs.
-Perhaps it is due to these policies that higher income electorates appear more likely to support the Liberal party, as the `Incomes` factor has a positive effect on Liberal preference.
+Perhaps it is due to these policies that higher income electorates appear more likely to support the Coalition, as the Incomes factor has a positive effect on Coalition preference.
 The effect also seems to have been increasing over time.
@@
 
-```{r resids, include = FALSE}
-all_resids <- data.frame(
-  Residuals = c(glsmod16$actual_residuals, glsmod13$actual_residuals, glsmod10$actual_residuals, glsmod07$actual_residuals, glsmod04$actual_residuals, glsmod01$actual_residuals),
-  gls_residuals = c(glsmod16$residuals, glsmod13$residuals, glsmod10$residuals, glsmod07$residuals, glsmod04$residuals, glsmod01$residuals),
-  bind_rows(
-    glsmod16$my_data %>% left_join(tpp16 %>% dplyr::select(DivisionNm, DivisionID, StateAb), by = "DivisionNm"),
-    glsmod13$my_data %>% left_join(tpp13 %>% dplyr::select(DivisionNm, DivisionID, StateAb), by = "DivisionNm"),
-    glsmod10$my_data %>% left_join(tpp10 %>% dplyr::select(DivisionNm, DivisionID, StateAb), by = "DivisionNm"),
-    glsmod07$my_data %>% left_join(tpp07 %>% dplyr::select(DivisionNm, DivisionID, StateAb), by = "DivisionNm"),
-    glsmod04$my_data %>% left_join(tpp04 %>% dplyr::select(DivisionNm, DivisionID, StateAb), by = "DivisionNm"),
-    glsmod01$my_data %>% left_join(tpp01 %>% dplyr::select(DivisionNm, StateAb), by = "DivisionNm")
-  )
-)
-# %>% dplyr::select(Residuals, DivisionNm, DivisionID, StateAb, year, LNP_Percent, gls_residuals)
-```
-
 It is often suggested that states have systematic differences that cause their electorates to vote differently.
 We can explore this effect by looking at the difference between the actual 2PP vote in each electorate and what our modelling would predict based only on electorate characteristics.
-We found a bias toward Labor in the ACT, and to a smaller extent in Tasmania, and a bias toward the Liberals in the Northern Territory.
+We found a bias toward Labor in the ACT, and to a smaller extent in Tasmania, and a bias toward the Coalition in the Northern Territory.
 No other states showed a significantly different voting pattern from what you would expect using only socio-demographic information.
 That suggests that any differences between Queensland and Victoria (for example) are due to different voter socio-demographics.
-
-```{r resstate, fig.cap = "Boxplot of residuals by state with jittered points.
-States ordered by median residual.
-A state-specific bias not captured by the model is evident.", fig.asp=0.6, out.width="70%", fig.pos = 'h', out.extra = '', eval=FALSE}
-all_resids %>%
-  ggplot(aes(x = fct_reorder(StateAb, Residuals), y = Residuals)) +
-  geom_hline(aes(yintercept = 0), col = "grey50") +
-  geom_boxplot() +
-  geom_jitter(alpha = 0.5, shape = 1) +
-  guides(col = FALSE, shape = FALSE, DivisionNm = FALSE) +
-  theme_bw() +
-  labs(x = "State", y = "Residuals")
-```
 
@@
 
-```{r readcooksd, eval = FALSE}
-read_chunk('cooksd.R')
-```
-
-```{r getcooksd, eval = FALSE}
-<<GetCookDistance>>
-```
-
-```{r cooksd, eval = FALSE}
-cooksd_df <- data.frame(
-  Residuals = c(glsmod16$actual_residuals, glsmod13$actual_residuals, glsmod10$actual_residuals, glsmod07$actual_residuals, glsmod04$actual_residuals, glsmod01$actual_residuals),
-  gls_residuals = c(glsmod16$residuals, glsmod13$residuals, glsmod10$residuals, glsmod07$residuals, glsmod04$residuals, glsmod01$residuals),
-  cooksd = c(glsmod16$cooksd, glsmod13$cooksd, glsmod10$cooksd, glsmod07$cooksd, glsmod04$cooksd, glsmod01$cooksd),
-  bind_rows(
-    glsmod16$my_data %>% left_join(tpp16 %>% dplyr::select(DivisionNm, DivisionID, StateAb), by = "DivisionNm"),
-    glsmod13$my_data %>% left_join(tpp13 %>% dplyr::select(DivisionNm, DivisionID, StateAb), by = "DivisionNm"),
-    glsmod10$my_data %>% left_join(tpp10 %>% dplyr::select(DivisionNm, DivisionID, StateAb), by = "DivisionNm"),
-    glsmod07$my_data %>% left_join(tpp07 %>% dplyr::select(DivisionNm, DivisionID, StateAb), by = "DivisionNm"),
-    glsmod04$my_data %>% left_join(tpp04 %>% dplyr::select(DivisionNm, DivisionID, StateAb), by = "DivisionNm"),
-    glsmod01$my_data %>% left_join(tpp01 %>% dplyr::select(DivisionNm, DivisionID, StateAb), by = "DivisionNm"))
+```{r outliers}
+res <- bind_rows(
+  glsmod16$my_data %>% mutate(pred=glsmod16$predictions),
+  glsmod13$my_data %>% mutate(pred=glsmod13$predictions),
+  glsmod10$my_data %>% mutate(pred=glsmod10$predictions),
+  glsmod07$my_data %>% mutate(pred=glsmod07$predictions),
+  glsmod04$my_data %>% mutate(pred=glsmod04$predictions),
+  glsmod01$my_data %>% mutate(pred=glsmod01$predictions)
+) %>%
+  dplyr::select(DivisionNm, year, LNP_Percent, pred) %>%
+  mutate(residuals = LNP_Percent - pred) %>%
+  arrange(-abs(residuals)) %>%
+  mutate(
+    DivisionNm = str_to_title(DivisionNm)
   )
@@
 
-```{r plotcooksd, fig.cap = "Electorates' Cook's Distance values in each election.", out.height = "15%", eval = FALSE}
-cooksd_df %>%
-  ggplot(aes(x = cooksd)) +
-  geom_dotplot() +
-  facet_wrap(~year, nrow = 1) +
-  # coord_fixed(ratio = .5) +
-  labs(x = "Cook's Distance", y = "") +
-  theme(axis.text.x = element_text(angle = 60, hjust = 1, size = 8), axis.text.y = element_text(size = 8))
+By this measure, the top five most surprising results in the last six elections were the following:
+
+```{r surprises}
+res %>% dplyr::select(DivisionNm, year, LNP_Percent, pred) %>% head(5) %>%
+  kable(booktabs=TRUE, digits=c(0,0,1,1),
+        col.names=c("Electorate", "Election", "Actual 2PP", "Predicted 2PP"))
 ```
 
-```{r, eval = FALSE}
-cooksd_df %>%
-  filter(cooksd > 0.1) %>%
-  select(DivisionNm, StateAb, year, cooksd, Residuals, LNP_Percent, everything()) %>%
-  View()
+Capricornia is on the Queensland coast, taking in Rockhampton, Yeppoon and parts of Mackay.
+It was won in 2001 and 2004 by Labour MP Kirsten Livermore, defying the socio-demographic prediction which suggested it would be a safe Coalition seat.
+She also won it in 2007 and 2010, but with a smaller number of votes, so the result was not so different from predicted.
+It is currently held by Michelle Landry for the Coalition.
+
+Herbert is another Queensland seat taking in Townsville.
+It was won in 2013 for the Coalition by Ewen Jones, and in 2016 by Cathy O'Toole for Labor by just 37 votes.
+The model predicts a much lower vote for the Coalition than tends to happen in practice.
+
+Wentworth probably needs no explanation --- the seat formerly held by Malcolm Turnbull which he won with a big majority in 2013.
+Its two-party vote jumped by over 10 percentage points in 2010 without experiencing any notable changes in its socio-demographic makeup --- implying that this may be the direct effect of its Liberal member, Malcolm Turnbull, becoming the leader of the Liberal party.
+Liberal support in Wentworth is underpredicted by the model in each year, and more so with Turnbull as Liberal leader.
+
+# True Blue Seats
+
+Another way the model can be used is to identify the seats with the strongest propensity to vote one way or the other, based on their socio-demographics.
+
+The top five most blue-ribbon Coalition seats in the last six elections are listed below.
+Mitchell takes in the north-western outer suburbs of Sydney, including Baulkham Hills, Kellyville, and parts of Castle Hill.
+Maranoa occupies a large part of south west Queensland including towns such as Charleville, Kingaroy and Winton.
+Mallee is, of course, the north west corner of Victoria including Mildura, Swan Hill and Stawell.
+
+```{r trueblue}
+res %>% dplyr::select(DivisionNm, year, LNP_Percent, pred) %>% 
+  arrange(-pred) %>%
+  head(5) %>%
+  kable(booktabs=TRUE, digits=c(0,0,1,1),
+        col.names=c("Electorate", "Election", "Actual 2PP", "Predicted 2PP"))
 ```
 
-Based on the distribution of the Cook's distance values, a Cook's distance greater than $0.1$ is considered to be influential and a potential outlier.
-The electorate of Sydney (NSW) has a large Cook's distance from 2001 to 2013, due to its diverse population (language, birthplace and religion), high number of defacto relationships, high income, high household mobility and small amount of workers in extractive and transformative jobs.
-It has remained a strong supporter of the Labor party and the Liberal vote is severely overpredicted by the model, making it an outlier.
-Nearby in metropolitan NSW, the electorate of Wentworth is found to be an outlier in all but the 2007 election.
-Although historically Liberal, its two-party vote jumped by over 10 percentage points in 2010 without experiencing any notable changes in its socio-demographic makeup --- implying that this may be the direct effect of its Liberal member, Malcolm Turnbull, becoming the leader of the Liberal party.
-Liberal support in Wentworth is underpredicted by the model in each year, and more so with Turnbull as Liberal leader.
+The top five seats most likely to vote Labour, based on their socio-demographics, are listed below.
+Batman has since been abolished, but occupied the inner north of Melbourne.
+Gellibrand takes in the inner western suburbs of Melbourne including Altona and Williamstown, while Fowler is in the western suburbs of Sydney, including Cabramatta and Liverpool.
 
-Lingiari, an electorate taking up almost all of the Northern Territory, is an outlier in the 2001--2007 elections due to its large Indigenous population, young age profile and low rates of property ownership.
-Fowler (NSW) has a diverse population with a high proportion of migrants, many Buddhists and Muslims, and has strong Labor support, making it influential in 2001, 2004 and 2010.
-Other electorates with large Cook's distance are Barton (NSW) and Leichhardt (QLD) in 2016, and Canberra (ACT) in 2007.
-
-```{r, eval = FALSE}
-x <- model_df %>%
-  filter(year == "2016") %>%
-  select(-c(year, DivisionNm)) %>%
-  as.matrix()
-myhat <- hat(x)
-
-plot(
-  glsmod07$cooksd,
-  hat(
-    model_df %>%
-      filter(year == "2007") %>%
-      select(-c(year, DivisionNm)) %>%
-      as.matrix()
-  )
-)
+```{r truered}
+res %>% dplyr::select(DivisionNm, year, LNP_Percent, pred) %>% 
+  arrange(pred) %>%
+  head(5) %>%
+  kable(booktabs=TRUE, digits=c(0,0,1,1),
+        col.names=c("Electorate", "Election", "Actual 2PP", "Predicted 2PP"))
 ```
-
-# Conclusion
-
-This paper explores the effects of electoral socio-demographic characteristics on the two-party preferred vote in the 2001--2016 elections, using information from the corresponding Australian federal elections and Censuses.
-As a Census does not always occur in the same year as an election, Census data for the 2004--2013 elections are generated by employing a method of spatio-temporal imputation.
-This imputes electoral socio-demographics for the electoral boundaries in place at the time of the election --- an approach that is distinctly different from previous work on modelling election outcomes, where Census and election data are typically joined without addressing their temporal differences.
-Before estimating a model, these socio-demographic variables are standardized (to adjust for changing variable scales) and many variables (representing similar information) are combined into factors, resulting in a reduced predictor set.
-A spatial error model is then estimated for each election, accounting for the inherent spatial structure of the data.
-
-Across the past six elections, most of the socio-demographics that drive the electoral two-party preferred vote are found to remain steady, whilst a few (typically weaker) effects vary over time.
-Industry and type of work are particularly influential, with energy-related and manufacturing/construction jobs, as well as administrative roles being strongly linked with the Liberal party in all elections.
-Incomes have a similarly consistent effect, with higher income areas supporting Liberal.
-Higher levels of unemployment shift from weak association with Labor to a significant Liberal effect over the years, and higher education levels are associated with Labor from 2007 (although significant only in 2016).
-It is also found that electorates with higher household mobility support Liberal, birthplace diversity favours Labor and more de facto relationships align with Labor preference --- although marriages, family and household sizes have no material influence.
-Furthermore, the neighbourhood (spatial) effects are found to be positive in all elections, although significant only in 2001 and 2016, meaning that in the 2004--2013 elections, electorates effectively voted independently.
-
-The findings in this paper complement the existing literature by modelling temporal trends, which as far as the authors are aware, has not been done previously for Australian elections using a regression framework.
-It is also the first study to model any Australian election since 2010 using Census information.
-
-Additionally, a key contribution of this research is the wrangling of the raw data and imputed data sets for the 2004, 2007, 2010 and 2013 elections, which have been contributed to the `eechidna` `R` package --- providing a rich, accessible data resource for future Australian electoral analysis.
 
@@
 
-This paper was produced using `RMarkdown` [@rmarkdown] and `knitr` [@knitr].
-All corresponding code for this paper can be found in the github repository [github.com/jforbes14/eechidna-paper](https://github.com/jforbes14/eechidna-paper), and the data used is available in the `eechidna` package [@eechidna].
-All raw data was obtained from the Australian Electoral Commission, the Australian Bureau of Statistics and the Australian Government.
+This analysis is based on a research paper coauthored by Rob Hyndman and Di Cook from the Department of Econometrics and Business Statistics at the Monash Business School, and Jeremy Forbes, a former Monash University Honours student in econometrics.
+The full analysis is available at https://robjhyndman.com/publications/elections/.
 
-# Software
+The code used for the analysis can be found in the github repository [github.com/jforbes14/eechidna-paper](https://github.com/jforbes14/eechidna-paper).
 
-All election and Census datasets, along with electoral maps and more, are available in the `eechidna` (Exploring Election and Census Highly Informative Data Nationally for Australia) `R` package, which can be downloaded from CRAN.
-The `eechidna`  package makes it easy to look at the data from the Australian Federal elections and Censuses that occurred between 2001 and 2016.
-This study contributed a large revision to the `eechidna` package, which included the addition of election and Census data for 2001--2010, voting outcomes for polling booths and imputed Census data for election years.
-For more details on using `eechidna`, please see the articles (vignettes) on the github page [ropenscilabs.github.io/eechidna/](https://ropenscilabs.github.io/eechidna/).
-
-The authors would like to sincerely thank Anthony Ebert, Heike Hofmann, Thomas Lumley, Ben Marwick, Carson Sievert, Mingzhu Sun, Dilini Talagala, Nicholas Tierney, Nathaniel Tomasetti, Earo Wang and Fang Zhou, all of whom have contributed to the `eechidna` package.
+All election and Census data since 2001, along with electoral maps and more, are available in the [`eechidna`](https://cran.r-project.org/package=eechidna) (Exploring Election and Census Highly Informative Data Nationally for Australia) `R` package, which can be downloaded from CRAN.
+The eechidna package uses data obtained from the Australian Electoral Commission, the Australian Bureau of Statistics and the Australian Government.

### 2019-08-05 anzjs_elections.Rmd
Census data is now SA1 imputed

@@
 
+```{r temporary_load_local_data, include = FALSE}
+load("data/abs2004_cd.rda")
+load("data/abs2007_cd.rda")
+load("data/abs2010_cd.rda")
+load("data/abs2013_cd.rda")
+```
+
 ```{r wrangle, include = FALSE}
@@
   )
-abs2013 <- eechidna::abs2013 %>%
+abs2013 <- abs2013_cd %>% #eechidna::abs2013 %>%
   mutate(
@@
   )
-abs2010 <- eechidna::abs2010 %>%
+abs2010 <- abs2010_cd %>% #eechidna::abs2010 %>%
   mutate(
@@
   )
-abs2007 <- eechidna::abs2007 %>%
+abs2007 <- abs2007_cd %>% #eechidna::abs2007 %>%
   mutate(
@@
   )
-abs2004 <- eechidna::abs2004 %>%
+abs2004 <- abs2004_cd %>% #eechidna::abs2004 %>%
   mutate(
@@
   left_join(eechidna::tpp04,
-    standardise_vars(abs2004) %>%
-      dplyr::select(-UniqueID),
+    standardise_vars(abs2004),
     by = c("DivisionNm")
@@
   left_join(eechidna::tpp07,
-    standardise_vars(abs2007) %>%
-      dplyr::select(-UniqueID),
+    standardise_vars(abs2007),
     by = c("DivisionNm")
@@
   left_join(eechidna::tpp10,
-    standardise_vars(abs2010) %>%
-      dplyr::select(-UniqueID),
+    standardise_vars(abs2010),
     by = c("DivisionNm")
@@
   left_join(eechidna::tpp13,
-    standardise_vars(abs2013) %>%
-      dplyr::select(-UniqueID),
+    standardise_vars(abs2013),
     by = c("DivisionNm")
@@
 ) %>%
-  mutate(year = factor(year)) %>%
+  mutate(year = factor(year), 
+    Band_00_19 = Age00_04 + Age05_14 + Age15_19,
+    Band_20_34 = Age20_24 + Age25_34,
+    Band_35_54 = Age35_44 + Age45_54,
+    Band_55plus = Age55_64 + Age65_74 + Age75_84 + Age85plus) %>% 
   select(-c(
-    starts_with("Age"), StateAb,
+    starts_with("Age"), MedianAge, StateAb,
     LNP_Votes, ALP_Votes, ALP_Percent, TotalVotes, Swing,
@@
     Other_NonChrist, OtherChrist, Catholic, Anglican,
-    Volunteer, EmuneratedElsewhere, DivisionID
+    Volunteer, EmuneratedElsewhere, Population, UniqueID
   ))
@@
     BornAsia = Born_Asia,
-    OneParentHouse = OneParent_House
-  )
+    OneParentHouse = OneParent_House,
+    Pop_00_19 = Band_00_19,
+    Pop_20_34 = Band_20_34,
+    Pop_35_54 = Band_35_54
+  ) %>% 
+  select(-Band_55plus)
 ```

### 2019-08-05 anzjs_elections.Rmd
Spatial imputation method is now aggregation over SA1 centroids

@@
 
-To account for spatial differences, the piecewise approximation method in @Goodchild1993 is adopted.
-Consider a map of source zones $s = 1,\dots,S$, for which socio-demographic information is available, and a set of target zones $t = 1,\dots,T$ for which information is to be imputed.
+The finest level of disaggregation available for Census data is the region classification called Statistical Area 1 (SA1).
+In 2016, Australia was divded into over 55,000 SA1s.
+Consider each of these SA1 regions as a source zone, $s = 1,\dots,S$, for which socio-demographic information is available.
+For simplicity, let each source zone be wholly summarised by its centroid.
+A set of target zones, $t = 1,\dots,T$, are defined as regions for which information is to be imputed - these are the electoral boundaries for a particular election.
 This is described in the context of a single election, and a single neighbouring Census.
-
-Let the map of electoral boundaries at the time of a Census define the source zones, and let the boundaries at the time of the election be the target zones.
-Denote the area of intersection between a source zone $s$ and a target zone $t$ as $A_{s,t}$.
-Additionally, let the population of the source zone $s$ be $U_s$ and the population of intersection between source zone $s$ and target zone $t$ be $P_{s,t}$.
-The estimated population of intersection is given by
-$$
-  \hat{P}_{s,t} = \frac{U_s*A_{s,t}}{\sum_{t=1}^T A_{s,t}},
-  \quad\text{for all $s=1,\dots,S$ and $t=1,\dots,T$}.
-$$
-Note that this estimator implicitly assumes that populations are uniformly distributed within each source zone.
-
-In order to calculate socio-demographic information for each of the target zones, a weighted average is taken using the estimated intersection populations as weights.
+Let $I_{s,t}$ be an indicator variable, for which $I_{s,t} = 1$ if the centroid of source zone $s$ falls within target zone $t$, and $0$ otherwise.
+Additionally, let the population of the source zone $s$ be $U_s$ and the population of target zone $t$ be $P_{t}$.
+
+In order to calculate socio-demographic information for each of the target zones, a weighted average of source zones is taken using their populations as weights.
 Denote a given Census variable for the target zone by $C_t$, and the same Census variable for the source zone as $D_s$.
@@
 $$
-  \hat{C}_t = \frac{\sum_{s=1}^{S}{D_s*\hat{P}_{s,t}}}%
-  {\sum_{s=1}^{S}{\hat{P}_{s,t}}},
+  \hat{C}_t = \frac{\sum_{s=1}^{S}{I_{s,t}*D_s*U_s}}%
+  {\sum_{s=1}^{S}{I_{s,t}*U_s}},
   \quad\text{for each $t=1,\dots,T$}.
 $$
+
 This concludes the spatial imputation of the socio-demographic characteristics for one target zone (a single electoral boundary), at the time of only one of the neighbouring Censuses.

### 2019-08-05 anzjs_elections.Rmd
Data is imputed using SA1s. Revised spatial imputation section. Edited factor creation section.

@@
 This is described in the context of a single election, and a single neighbouring Census.
-Let $I_{s,t}$ be an indicator variable, for which $I_{s,t} = 1$ if the centroid of source zone $s$ falls within target zone $t$, and $0$ otherwise.
+Take the example of the Melbourne Ports electorate from the 2013 federal election, illustrated in Figure \@ref(fig:melbports13).
+The purple region in this figure represents the target zone and the source zones are the centroid locations from the 2016 Census SA1 areas.
+
+```{r melbports13, fig.cap = "The electoral boundaries for Melbourne Ports (shown in purple) and surrounding electorates, with centroids for Statistical Area 1 regions from the 2016 Census overlaid.
+The centroids falling within the purple region are attributed to Melbourne Ports.", out.height="28%"}
+# Get Vic shape
+load("data/vic_map.rda")
+load("data/MP_sa1_dots_2011.rda")
+
+p1 <- ggplot(data = vic_map) +
+  geom_polygon(aes(x = long, y = lat, group = group, fill = Elect_div == "Melbourne Ports"),
+    colour = "black", alpha = 0.4
+  ) +
+  scale_fill_manual(
+    name = "Electoral Boundaries",
+    values = c("white", "purple"),
+    labels = c("Other Electorates", "Melbourne Ports (VIC)")
+  ) +
+  theme_map() +
+  coord_equal(xlim = c(144.88, 145.07), ylim = c(-37.92, -37.78))
+
+p1 + geom_point(aes(x = long_c, y = lat_c, shape = shape), data = MP_sa1_dots_2011 %>% mutate(shape = "1"), colour = "grey25", size = 1.3) + scale_shape_manual(name = "Statistical Area 1", values = 3, labels = "Centroids")
+```
+
+Furthermore, let $I_{s,t}$ be an indicator variable, for which $I_{s,t} = 1$ if the centroid of source zone $s$ falls within target zone $t$, and $0$ otherwise.
 Additionally, let the population of the source zone $s$ be $U_s$ and the population of target zone $t$ be $P_{t}$.
@@
 
-### An illustration of the spatio-temporal imputation {-}
-
-Census data is publicly available at different levels of aggregation, ranging from SA1 (over 50,000 zones) to electoral divisions (150 zones).
-For this study, electoral divisions are used as source zones, and the imputation method is applied to produce socio-demographic variables for each of the 2004, 2007, 2010 and 2013 elections.
-As mentioned earlier, there is no need to impute socio-demographic variables for the 2001 and 2016 elections.
-To illustrate the method, consider the imputation of socio-demographic variables for the electorate of Hume in New South Wales (NSW) at the time of the 2013 federal election.
-The boundaries shown in Figure \@ref(fig:hume13) define all target zones in NSW for 2013, with the target zone of interest (Hume) shaded purple.
-
-```{r hume13, fig.cap = "Some of the electoral boundaries in NSW for 2013, with the electoral boundary for Hume shown in purple.", out.height="28%"}
-hume_area13 <- nat_map13 %>%
-  filter(state %in% c("ACT", "NSW"), long < 154)
-
-ggplot(data = hume_area13) +
-  geom_polygon(aes(x = long, y = lat, group = group, fill = elect_div == "HUME"),
-    colour = "grey50", alpha = 0.4
-  ) +
-  scale_fill_manual(
-    name = "Electoral boundaries",
-    values = c("white", "purple"),
-    labels = c("Other 2013 Electorates", "Hume 2013")
-  ) +
-  theme_map() +
-  coord_equal()
-```
-
-```{r hume16, fig.cap = "Census division boundaries in NSW for 2016, with the 2013 electoral boundary for Hume, shown in purple.
-The purple region is not contained within a single Census division.", out.height="28%"}
-hume_area16 <- nat_map16 %>%
-  filter(state %in% c("ACT", "NSW"), long < 154) %>%
-  mutate(year = "2016") %>%
-  bind_rows(
-    hume_area13 %>%
-      filter(elect_div == "HUME") %>%
-      mutate(year = "2013")
-  )
-
-ggplot(data = hume_area16) +
-  geom_polygon(aes(
-    x = long, y = lat, group = group, fill = year == "2013",
-    alpha = year == "2013", colour = year == "2013"
-  )) +
-  scale_fill_manual(
-    name = "Electoral boundaries",
-    values = c("white", "purple"),
-    labels = c("2016 Electorates", "Hume 2013")
-  ) +
-  scale_alpha_manual(values = c(0, 0.4)) +
-  scale_color_manual(values = c("grey50", NA)) +
-  theme_map() +
-  coord_equal() +
-  guides(alpha = FALSE, col = FALSE, 
-    fill = guide_legend(override.aes = list(col = c("grey50", NA), alpha = 0.4)))
-```
-
-The corresponding source zones from the 2016 Census are shown in Figure \@ref(fig:hume16).
-As can be seen, the Hume boundary from the 2013 election (shaded purple) does not perfectly match any of the source zones.
-
-There are many source zones from the 2016 Census that intersect with this purple region, including the divisions of Riverina, Eden-Monaro and Hume, along with smaller intersecting areas with Fenner, Calare, Gilmore and Whitlam.
-The proportion of each source zone that overlaps with the purple region is calculated, and used to obtain the intersecting populations $\hat{P}_{s,t}$.
-
-```{r hume1}
-hume <- abs2016 %>%
-  filter(DivisionNm %in%
-           c("HUME", "RIVERINA", "EDEN-MONARO", "CANBERRA",
-             "FENNER", "WHITLAM", "GILMORE", "CALARE")) %>%
-  mutate(
-    Pst = c(21, 548, 16358, 474, 86, 145427, 39117, 92),
-    Proportion = Pst/Population,
-    DivisionNm = tools::toTitleCase(tolower(DivisionNm))
-  ) %>%
-  select(DivisionNm, Proportion, Population, Pst, AusCitizen) %>%
-  arrange(desc(Proportion))
-
-hume %>%
-  kable(longtable=FALSE, booktabs=TRUE, digits=c(0,4,0,0,1),
-        format='latex',escape=FALSE,
-        col.names=c(
-          "Source zone (2016)",
-          "Proportion",
-          "Source zone population",
-          "Population allocated to purple region: $\\hat{P}_{s,t}$",
-          "AusCitizen (\\%): $D_s$"),
-    caption = "Population from each intersecting Census source zone (2016) that is allocated to the target zone (purple region - Hume electoral boundary in 2013), and the corresponding proportion of Australian citizens in each of these source zones.") %>%
-  kableExtra::kable_styling(font_size=12, latex_options='scale_down')
-```
-
-```{r wave, include=FALSE}
-wave2016 <- weighted.mean(x = hume$AusCitizen, w = hume$Pst)
-hume11 <- abs2011 %>%
-  filter(DivisionNm %in%
-           c("HUME", "RIVERINA", "EDEN-MONARO", "CANBERRA",
-             "FENNER", "WHITLAM", "GILMORE", "CALARE")) %>%
-  mutate(
-    Pst = c(21, 548, 16358, 474, 86, 145427), # Needs adjusting
-    Proportion = Pst/Population,
-    DivisionNm = tools::toTitleCase(tolower(DivisionNm))
-  ) %>%
-  select(DivisionNm, Proportion, Population, Pst, AusCitizen) %>%
-  arrange(desc(Proportion))
-wave2011 <- 91.00 #weighted.mean(x = hume11$AusCitizen, w = hume11$Pst)
-```
-
-Now consider the socio-demographic variable *AusCitizen*, the proportion of people in the region who are Australian citizens.
-A weighted average of *AusCitizen*, with the allocated population from each source zone as weights, yields $\hat{C}_{\text{Hume},2016} = `r sprintf("%4.2f",wave2016)`\%$.
-Repeating this process using the 2011 Census yields $\hat{C}_{\text{Hume},2011} = `r sprintf("%4.2f",wave2011)`\%$.
-Finally, linear interpolation between 2011 and 2016 yields the 2013 estimate:
+Continuing with the example of Melbourne Ports in the 2013 election, the estimate for a given Census variable in 2016, $\hat{C}_{MelbPorts, 2016}$ would be obtained by computing the weighted average of this variable amongst the SA1s within the purple region shown in Figure \@ref(fig:melbports13).
+This would be repeated with the 2011 Census SA1s to obtain $\hat{C}_{MelbPorts, 2011}$, from which the final estimate is given by
 $$
-  \hat{C}_{\text{Hume},2013}
-    = \frac{3}{5} \hat{C}_{\text{Hume},2011} + \frac{2}{5} \hat{C}_{\text{Hume},2016}
-    % = \frac{3}{5} \cdot 91.00 \% + \frac{2}{5} \cdot 89.65 \% \\
-    =  90.46\%.
+  \hat{C}_{\text{MelbPorts},2013}
+    = \frac{3}{5} \hat{C}_{\text{MelbPorts},2011} + \frac{2}{5} \hat{C}_{\text{MelbPorts},2016}.
 $$
-This is done for each of the 65 socio-demographic variables, and is repeated for each of the 149 remaining target zones corresponding with 2013 electorates.
+This is done for each of the socio-demographic variables, and is repeated for each of the 149 remaining target zones corresponding with 2013 electorates.
 
@@
 First, variable scales change over the years, making it important to standardize variables.
-Second, many variables represent similar information and where appropriate, should be combined in some way.
-Principal component analysis is used to identify variables covary, from which intuitive groupings are selected to be combined into a single variable.
-This also reduces the dimension of the data.
+Second, many variables represent similar information and where appropriate, should be grouped together.
+To determine which variables should be grouped, principal component analysis (PCA) is used.
+The intuition here is that PCA will identify which variables covary in the directions of maximal variance, from which intuitive groupings of variables can be chosen to combine into a single variable.
 After these steps, a model specification is chosen.
@@
 Many of the socio-demographic variables have changing scales over the years.
-For example, inflation-adjusted median rental prices increased across almost all electorates, with median rent of 200 dollars per week placing an electorate in the 90th percentile in 2001, but only the 30th percentile in 2016.
+For example, inflation-adjusted median rental prices increased across almost all electorates, with median rent of 225 dollars per week placing an electorate in the 90th percentile in 2001, but only the 45th percentile in 2016.
 In order for socio-demographic effects to be comparable across years, all explanatory variables are standardized to have mean zero and variance one within each election year.
@@
 
-There are only $N = 150$ observations (electorates) in each election and $p = 65$ socio-demographic variables in each cross-section, with many variables represent similar information about an electorate.
-Any model that uses all variables would face serious problems with multi-collinearity and over-fitting, which would likely lead to erroneous conclusions regarding variable significance.
-To address this, groups of variables that represent similar information are combined into "factors"[^2].
+There are only $N = 150$ observations (electorates) in each election and $p = 50$ socio-demographic variables in each cross-section, with many variables represent similar information about an electorate.
+Any model that uses all variables would face problems with multi-collinearity and over-fitting, which would likely lead to erroneous conclusions regarding variable significance.
+To address this, variables that represent similar information are combined into a single variable, which will be referred to as a "factor".
 
-[^2]: A preliminary step involved removing all age bands, because age is represented by median age, and to remove variables relating to particular denominations of Christianity.
+A factor is created from a group of variables if there is an intuitive reason as to why these variables should represent similar information and if there is evidence to suggest that they covary.
+For example, a potential group would be variables relating to electoral incomes --- median family, household and personal incomes.
+To determine which variables covary, principal component analysis is used on a combined dataset of socio-demographic variables from all six elections[^2].
+The only variables exempted from the principal component analysis are the four variables representing age brackets (the proportion of the population aged 0-19 years old, 20-34 years old, 45-54 years old and 55 years plus), which are to be included in the model without grouping.
 
-A factor is created from a group of variables if there is an intuitive reason as to why they should represent similar information and if there is evidence to suggest that they covary.
-For example, a potential group would be variables relating to electoral incomes --- median family, household and personal incomes.
-To determine which variables covary, principal component analysis is used on a combined dataset of socio-demographic variables from all six elections.
-It is appropriate to compute principal components in this way because when computed separately for each election, scree plots level off after four components and the loadings of the first four components are similar across the elections.
+[^2]: It is appropriate to compute principal components on a combined dataset of all six elections because when computed separately for each election, scree plots level off after four components and the loadings of the first four components are similar across the elections.
 
@@
 These are:
-`Incomes` (median personal, household and family incomes);
-`Unemployment` (unemployment and labor force participation rates);
-`ProportyOwned` (rates of housing ownership, mortgages, renting and government housing);
-`RentLoanPrice` (median rental and loan repayments);
-`FamHouseSize` (average household size, ratio of people to families and household makeup (single person, couple with kids and couple without kids); and
-`Education` (high school and university qualifications, jobs requiring higher levels of education as well as vocational course completions and jobs that do not require higher education levels, such as laborer or tradesperson).
-For each of these groups, variables with positive loadings are added and those with negative loadings are subtracted to create a factor.
-After computing these sums, each factor is standardized to have mean zero and variance one, within each election.
+`Incomes` (median personal income, median household income, median family income);
+`Unemployment` (unemployment rate, labor force participation rate);
+`PropertyOwned` (proportion of dwellings that are owned, proportion of dwellings that are mortgages, proportion of dwellings that are rented, proportion of dwellings that classified as government housing);
+`RentLoanPrice` (median rental payment amount, median loan repayment amount);
+`FamHouseSize` (average household size, ratio of people to families, incidence of single person households, incidence of households containing a couple with kids, incidence of households containing a couple without kids); and
+`Education` (high school completions, university qualifications, jobs requiring higher levels of education, vocational course completions, jobs that do not require higher education levels).
+For each of these groupings, a factor is created by taking a weighted sum of the variables.
+The weightings are allocated on the basis of whether the variable had a positive or negative loading in the principal component from which the grouping was identified.
+Variables with a positive loading are allocated a weight of $+1$ and those with negative loadings are allocated a weight of $-1$.
+After computing these weighted sums, the factor is standardized to have mean zero and variance one, within each election.
 
-There are $p=30$ variables in the resultant predictor set, with all of these used in the regression for each election.
+There are $p=32$ variables in the resultant predictor set, with all of these used in the regression for each election.
 

### 2019-08-05 response_to_review.Rmd
More annotations

@@
 
-## Editor's comments
+## Editor's comments (A)
 
@@
 
-...
+Impute using Census SA1
+
+-- Re-write the imputation section to use this method. See other paper for terminology.
 
@@
 
+-- Be more explicit in the introduction and/or the results section. "Study aims to see how the aggregate region of electorates behave, rather than make inferences about individual voters."
+
 *3. It appears that your approach to variable selection has confused everybody. My understanding of what you have done is that:*
 *- You begin with 65 variables (are they all the available census data or a subset?).*
-
-...
 
@@
 
+Don't even refer to any variables that aren't included.
+Use three age bands instead of MedianAge (effectively becomes two because of multicollinearity)
+
+-- Be concise about which variables are included to begin with.
+
 *- Then you use the PCA to identify variables that covary.
@@
 ...
+More of an explanation of why and how we are using PCA. Then more details of exactly what the resultant factors are.
+
+-- Provide the logic for why a procedure like PCA is used here.
+A large number of candidate covariates.
+Many of which represent similar information.
+Group similar variables.
+This will also help reduce the dimension of the data.
+PCA is used to inform the selection of these groups, as it will reveal which variables co-vary in the direction of maximal variance within the observed data.
 
@@
 ...
+With the final variables, use TOUR to see if there are associations.
 
-## Associate Editor's comments
+-- Fit model with significant effects for each year. Comment on sign of effects and if coefficients are within confidence bands.
+
+## Associate Editor's comments (B)
 *Some conceptually simple imputation methods are used to align polling and census data to investigate associations between census variables and voting behaviour at electoral division level.*
 
-...
+... A
 
@@
 
-...
+... A
 
@@
 ...
+Explain what the difference between the table and plots are. Keep both. 'Estimated model parameters' instead of 'spatial model'.
 
@@
 ...
+Refine description of partial residuals. Explain what a time-varying restriction means.
 
@@
 ...
+With SA1s we should be able to answer this. Less spatial blurring in imputation, so the spatial effect could become clearer.
 
-## Reviewer 1's comments
+-- Re-write spatial effect section.
+Effects are significant in 2001, 2004 and 2016.
+Do not claim that electorates voted independently, rather just say the effect is always positive, and significant in those years.
+
+## Reviewer 1's comments (C)
 
@@
 ...
+Be more explicit about what the conclusions will be - electorate level.
 
@@
 ...
+NA
 
@@
 ...
+Check the SD of the original variables and see how it compares to their estimated effects.
+If there are variables with small SDs but big effects, we should look into this further.
+This implies that we need to take the log of the variable.
+(question to answer: should logs be taken before PCA?)
 
@@
 ...
+A
 
@@
 
-...
+...A
 
@@
 
-...
+...Think about this - are there omitted variables? How might this be captured in the spatial effect?
 
@@
 
-## Reviewer 2's comments
+## Reviewer 2's comments (D)
 
@@
 
-...
+... ?
 
@@
 
-...
+... Are there more variables we should get? I think no.
 
@@
 
-...
+... Be more precise in explaning the factors created.
 
@@
 
-...
+... TOUR - ask Di for help.
 
@@
 
-...
+...  Don't include incumbent in model, but do plot residuals against incumbency to see if there is an effect. We want the model to only capture the socio-demographic effects.
+
+-- Section on incumbent effect (need to get winners each year from 1998, 2001 to do this)

### 2019-08-06 anzjs_elections.Rmd
added install instructions for patchwork

@@
 library(spdep)
+# remotes::install_github("thomasp85/patchwork")
 library(patchwork)

### 2019-08-23 edits.Rmd
edits to be included - this is just a space for workings

--- 2019-04-21 paper.Rmd
@@
 ---
-title: "Spatial modelling of the two-party preferred vote in Australian federal elections: 2001--2016"
-author: Jeremy Forbes$^*$, Dianne Cook, Rob J Hyndman
-toc: false
-documentclass: anzsauth
-classoption: times, doublespace
-fontfamily: times
-bibliography: references.bib
-biblio-style: anzsj
-output: 
-  bookdown::pdf_document2:
-    fig_caption: yes
-    citation_package: natbib
-    keep_tex: yes
-    number_sections: yes
-    fig_height: 5
-    fig_width: 8
-    includes:
-      in_header: anzjs_preamble.tex
+title: "edits"
+author: "Jeremy Forbes"
+date: "05/08/2019"
+output: html_document
 ---
-
-# Summary{-}
-
-We examine the relationships between electoral socio-demographic characteristics and two-party preference in the six Australian federal elections held between 2001 to 2016.
-Socio-demographic information is derived from the Australian Census, which occurs every five years.
-Since a Census is not directly available for each election, spatio-temporal imputation is employed to estimate Census data for the electorates at the time of each election.
-This accounts for both spatial and temporal changes in electoral characteristics between Censuses.
-To capture any spatial heterogeneity, a spatial error model is estimated for each election, which incorporates a spatially structured random effect vector that can be thought of as the unobserved political climate in each electorate.
-Over time, the impact of most socio-demographic characteristics that affect electoral two-party preference do not vary, with industry of work, incomes, household mobility and de facto relationships having strong effects in each of the six elections.
-Education and unemployment are amongst those that have varying effects.
-It is also found that between 2004 and 2013, the spatial effect was insignificant, meaning that electorates voted effectively independently.
-All data featured in this study has been contributed to the `eechidna` `R` package (available on CRAN).
-
-**Keywords:** federal election, Census, Australia, spatial modelling, imputation, data science, socio-demographics, electorates, R, eechidna
-
-# Introduction {#intro}
-
-```{r setup, include=FALSE, message=FALSE}
-knitr::opts_chunk$set(
-  echo = FALSE,
-  fig.pos = "h",
-  fig.align = "center",
-  cache = TRUE,
-  message = FALSE
-)
-library(tidyverse)
-library(eechidna)
-library(forcats)
-library(ggthemes)
-library(knitr)
-library(nlme)
-library(rgeos)
-library(spdep)
-library(patchwork)
-#library(gridExtra)
-#library(grid)
-library(predictmeans)
-source("functions.R")
-```
-
-```{r wrangle, include = FALSE}
-# Take log of indigenous, judaism, islam, buddhism
-abs2016 <- eechidna::abs2016 %>%
-  mutate(
-    Indigenous = log(Indigenous),
-    Judaism = log(Judaism),
-    Islam = log(Islam),
-    Buddhism = log(Buddhism)
-  )
-abs2013 <- eechidna::abs2013 %>%
-  mutate(
-    Indigenous = log(Indigenous),
-    Judaism = log(Judaism),
-    Islam = log(Islam),
-    Buddhism = log(Buddhism)
-  )
-abs2010 <- eechidna::abs2010 %>%
-  mutate(
-    Indigenous = log(Indigenous),
-    Judaism = log(Judaism),
-    Islam = log(Islam),
-    Buddhism = log(Buddhism)
-  )
-abs2007 <- eechidna::abs2007 %>%
-  mutate(
-    Indigenous = log(Indigenous),
-    Judaism = log(Judaism),
-    Islam = log(Islam),
-    Buddhism = log(Buddhism)
-  )
-abs2004 <- eechidna::abs2004 %>%
-  mutate(
-    Indigenous = log(Indigenous),
-    Judaism = log(Judaism),
-    Islam = log(Islam),
-    Buddhism = log(Buddhism)
-  )
-abs2001 <- eechidna::abs2001 %>%
-  mutate(
-    Indigenous = log(Indigenous),
-    Judaism = log(Judaism),
-    Islam = log(Islam),
-    Buddhism = log(Buddhism)
-  )
-
-# Combine and standardize
-my_df <- bind_rows(
-  left_join(eechidna::tpp01,
-    standardise_vars(abs2001) %>%
-      dplyr::select(-c(UniqueID, Area, ends_with("NS"), Population)),
-    by = c("DivisionNm", "StateAb" = "State")
-  ) %>% mutate(year = "2001"),
-  left_join(eechidna::tpp04,
-    standardise_vars(abs2004) %>%
-      dplyr::select(-UniqueID),
-    by = c("DivisionNm")
-  ) %>% mutate(year = "2004"),
-  left_join(eechidna::tpp07,
-    standardise_vars(abs2007) %>%
-      dplyr::select(-UniqueID),
-    by = c("DivisionNm")
-  ) %>% mutate(year = "2007"),
-  left_join(eechidna::tpp10,
-    standardise_vars(abs2010) %>%
-      dplyr::select(-UniqueID),
-    by = c("DivisionNm")
-  ) %>% mutate(year = "2010"),
-  left_join(eechidna::tpp13,
-    standardise_vars(abs2013) %>%
-      dplyr::select(-UniqueID),
-    by = c("DivisionNm")
-  ) %>% mutate(year = "2013"),
-  left_join(eechidna::tpp16,
-    standardise_vars(abs2016) %>%
-      dplyr::select(-c(UniqueID, Area, ends_with("NS"), Population)),
-    by = c("DivisionNm", "StateAb" = "State")
-  ) %>% mutate(year = "2016")
-) %>%
-  mutate(year = factor(year)) %>%
-  dplyr::select(-c(
-    starts_with("Age"), StateAb,
-    LNP_Votes, ALP_Votes, ALP_Percent, TotalVotes, Swing,
-    InternetUse, InternetAccess, EnglishOnly,
-    Other_NonChrist, OtherChrist, Catholic, Anglican,
-    Volunteer, EmuneratedElsewhere, UniqueID
-  ))
-
-# Create final df for modelling
-
-factors_df <- my_df %>%
-  mutate(
-    Education = BachelorAbv + HighSchool + Professional + Finance - Laborer - Tradesperson - DipCert,
-    FamHouseSize = FamilyRatio + AverageHouseholdSize + Couple_WChild_House - Couple_NoChild_House - SP_House,
-    PropertyOwned = Owned + Mortgage - Renting - PublicHousing,
-    RentLoanPrice = MedianRent + MedianLoanPay,
-    Incomes = MedianFamilyIncome + MedianHouseholdIncome + MedianPersonalIncome,
-    Unemployment = Unemployed - LFParticipation
-  ) %>%
-  dplyr::select(-c(
-    BachelorAbv, HighSchool, Professional, Finance, Laborer, Tradesperson, DipCert, FamilyRatio,
-    AverageHouseholdSize, Couple_WChild_House, Couple_NoChild_House, SP_House, Owned, Mortgage, Renting,
-    PublicHousing, MedianFamilyIncome, MedianHouseholdIncome, MedianPersonalIncome, MedianRent,
-    MedianLoanPay, Unemployed, LFParticipation
-  ))
-
-# Now standardize factors
-
-small_df <- bind_rows(
-  factors_df %>% filter(year == "2001") %>% standardise_vars(),
-  factors_df %>% filter(year == "2004") %>% standardise_vars(),
-  factors_df %>% filter(year == "2007") %>% standardise_vars(),
-  factors_df %>% filter(year == "2010") %>% standardise_vars(),
-  factors_df %>% filter(year == "2013") %>% standardise_vars(),
-  factors_df %>% filter(year == "2016") %>% standardise_vars()
-)
-
-# Order electorates in alphabetical order to match spatial matrix
-model_df <- small_df %>%
-  arrange(year, DivisionNm) %>%
-  dplyr::select(order(colnames(.))) %>%
-  rename(
-    ManagerAdmin = ManagerAdminClericalSales,
-    OtherLanguage = OtherLanguageHome,
-    BornSEEuro = Born_SE_Europe,
-    BornAsia = Born_Asia,
-    OneParentHouse = OneParent_House
-  )
-```
-
-```{r spatialweights, eval = FALSE}
-# Get spatial weights
-# This is not run, as weights are pulled from github, but this can be run to get them from scratch.
-sF_16 <- sF_download(2016)
-sF_13 <- sF_download(2013)
-sF_10 <- sF_download(2010)
-sF_07 <- sF_download(2007)
-sF_04 <- sF_download(2004)
-sF_01 <- sF_download(2001)
-
-sp_weights_16 <- sp_weights_matrix(sF_16)
-sp_weights_13 <- sp_weights_matrix(sF_13)
-sp_weights_10 <- sp_weights_matrix(sF_10)
-sp_weights_07 <- sp_weights_matrix(sF_07)
-sp_weights_04 <- sp_weights_matrix(sF_04)
-sp_weights_01 <- sp_weights_matrix(sF_01)
-```
-
-```{r FGLS, include = FALSE}
-# Instead of running spatial weights, load from github
-load("data/sp_weights_01.rda")
-load("data/sp_weights_04.rda")
-load("data/sp_weights_07.rda")
-load("data/sp_weights_10.rda")
-load("data/sp_weights_13.rda")
-load("data/sp_weights_16.rda")
-
-## Run full models for each year
-full_formula <- "LNP_Percent ~ ."
-
-# 2016
-glsmod16 <- my_fgls(full_formula,
-  my_data = model_df %>% filter(year == "2016"),
-  sp_weights = sp_weights_16
-)
-
-# 2013
-glsmod13 <- my_fgls(full_formula,
-  my_data = model_df %>% filter(year == "2013"),
-  sp_weights = sp_weights_13
-)
-
-# 2010
-glsmod10 <- my_fgls(full_formula,
-  my_data = model_df %>% filter(year == "2010"),
-  sp_weights = sp_weights_10
-)
-
-# 2007
-glsmod07 <- my_fgls(full_formula,
-  my_data = model_df %>% filter(year == "2007"),
-  sp_weights = sp_weights_07
-)
-
-# 2004
-glsmod04 <- my_fgls(full_formula,
-  my_data = model_df %>% filter(year == "2004"),
-  sp_weights = sp_weights_04
-)
-
-# 2001
-glsmod01 <- my_fgls(full_formula,
-  my_data = model_df %>% filter(year == "2001"),
-  sp_weights = sp_weights_01
-)
-
-## Visualise coefficients and significance
-
-coef_df <- bind_rows(
-  data.frame(
-    variable = glsmod16$coefficients %>% names(),
-    estimate = glsmod16$coefficients %>% unname(),
-    se = summary(glsmod16)$tTable[, "Std.Error"] %>% unname(),
-    p = summary(glsmod16)$tTable[, "p-value"] %>% unname(),
-    year = 2016),
-  data.frame(
-    variable = glsmod13$coefficients %>% names(),
-    estimate = glsmod13$coefficients %>% unname(),
-    se = summary(glsmod13)$tTable[, "Std.Error"] %>% unname(),
-    p = summary(glsmod13)$tTable[, "p-value"] %>% unname(),
-    year = 2013),
-  data.frame(
-    variable = glsmod10$coefficients %>% names(),
-    estimate = glsmod10$coefficients %>% unname(),
-    se = summary(glsmod10)$tTable[, "Std.Error"] %>% unname(),
-    p = summary(glsmod10)$tTable[, "p-value"] %>% unname(),
-    year = 2010),
-  data.frame(
-    variable = glsmod07$coefficients %>% names(),
-    estimate = glsmod07$coefficients %>% unname(),
-    se = summary(glsmod07)$tTable[, "Std.Error"] %>% unname(),
-    p = summary(glsmod07)$tTable[, "p-value"] %>% unname(),
-    year = 2007),
-  data.frame(
-    variable = glsmod04$coefficients %>% names(),
-    estimate = glsmod04$coefficients %>% unname(),
-    se = summary(glsmod04)$tTable[, "Std.Error"] %>% unname(),
-    p = summary(glsmod04)$tTable[, "p-value"] %>% unname(),
-    year = 2004),
-  data.frame(
-    variable = glsmod01$coefficients %>% names(),
-    estimate = glsmod01$coefficients %>% unname(),
-    se = summary(glsmod01)$tTable[, "Std.Error"] %>% unname(),
-    p = summary(glsmod01)$tTable[, "p-value"] %>% unname(),
-    year = 2001)
-)
-```
-
-Australia has changed in many ways over the last two decades.
-Rising house prices, country-wide improvements in education, an ageing population, and a decline in religious affiliation, are just a few facets of the country's evolving socio-demographic characteristics.
-At the same time, political power has moved back and forth between the two major parties.
-In the 2007 and 2010 federal elections, the Australian Labor Party (Labor) was victorious, whereas the 2001, 2004, 2013 and 2016 elections were won by the Liberal National coalition (Liberal).
-The two-party preferred vote, a measure of support between these two parties, fluctuated between 47.3% and 53.5% (in favour of the Liberal party) over this period.
-This study explores how electoral socio-demographic characteristics relate to two-party preference, and whether their effects have changed over time.
-
-Electoral socio-demographics are derived from the Australian Census, and vote counts are obtained from Australian federal elections.
-Joining these two data sources is problematic as there is an inherent asynchronicity in the two types of events.
-A Census is conducted by the Australian Bureau of Statistics (ABS) every five years, whereas federal elections, conducted by the Australian Electoral Commission (AEC), usually occur every three years or so.
-The first problem addressed is that of constructing appropriate Census data for the 2004, 2007, 2010 and 2013 elections --- election years in which a Census does not occur.
-The predominant approach in previous studies was to join voting outcomes to the nearest Census, without accounting for any temporal differences [see @DavisStimson98;@Stimson06;@Liao09;@Stimson09].
-Furthermore, electoral boundaries change regularly, so spatial discrepancies also arise when matching with electoral data.
-To obtain appropriate "Census-like" data for these four elections, electoral socio-demographics are constructed using a spatio-temporal imputation that combines areal interpolation [@Goodchild1993] and linear time-interpolation.
-Collecting and wrangling the raw data, along with the imputation process, are detailed in Section \@ref(data).
-All data and associated documentation relating to this procedure are available in the `eechidna` `R` package [@eechidna], providing a resource for future analysis.
-
-Previous work on modelling Australian federal elections has found that aggregate socio-demographics are relatively good predictors of voting outcomes.
-@Forrest01 used multiple regression to model the Liberal and Labor primary vote for polling booths in the Farrer electorate in 1998 as a function of Census variables from 1996.
-@Stimson06, @Stimson09 and @Stimson12 used principal component analysis of polling booths in the 2001, 2004 and 2007 elections respectively, also finding that socio-demographic characteristics of polling booths are linked to their two-party preferred vote.
-In contrast, @Stimson09 models the polling booth swing vote (change in the two-party preferred vote) in the 2007 election, finding that little of the swing vote can be explained by Census data.
-
-Instead of analyzing a single election in isolation, this paper employs a consistent model framework across six elections so that temporal changes in the effects of socio-demographics can be observed.
-Each federal election is modelled with a cross-sectional dataset.
-<!--The use of a regression framework to examine these socio-political relationships over time is seemingly absent from previous Australian studies.
-It also appears that no study has attempted any type of statistical analysis of socio-demographics in conjunction with voter behaviour in Australia since 2007, making this paper distinctly different from previous work.-->
-The cross-sectional dataset for each election used here consists of the two-party preferred vote (as the response variable), and a set of common socio-demographic variables (as the explanatory variables) that characterize each electorate.
-To prepare these datasets, socio-demographic variables are first standardized, and then a principal component analysis is used to group variables into "factors".
-To account for the inherent spatial structure of the data, a spatial error model is then estimated for each election.
-
-The paper is organised as follows.
-Section \@ref(data) describes the data collection, joining and cleaning.
-These pre-processing steps and model details are discussed in Section \@ref(modelling).
-Section \@ref(results) describes the inference conducted to determine signficance of effects and how these change over time.
-Section \@ref(conclusion) summarises the work.
-Two supplementary sections document the contributions of others to this work and the software.
-
-# Data collection, wrangling and imputation {#data}
-
-## Collecting the data
-
-The voting outcome of interest is the electoral two-party preferred vote, which is provided by the Australian Electoral Commission (AEC) for the 2001, 2004, 2007, 2010, 2013 and 2016 elections via the AEC Tally Room.
-The AEC divides Australia into 150 regions, called electorates, with each corresponding to a single seat in the House of Representatives.
-Voting is compulsory in Australia, and each voter assigns a numbered preference to each available candidate in their electorate.
-The two-party preferred vote is determined by a tally of these preferences where, by convention, only the ranks of the Labor and Liberal candidates are considered.
-This is recorded as a percentage preference in favour of the Liberal party.
-
-Socio-demographic variables were derived from the Australian Census of Population and Housing (Census), which is a survey of every household in Australia, recording information such as age, gender, ethnicity, education level and income.
-There have been four Censuses so far in the 21st century, conducted in 2001, 2006, 2011 and 2016.
-The Australian Bureau of Statistics (ABS) conducts the Census and publishes aggregated information.
-The ABS uses electoral boundaries as defined by the AEC at the time of each Census, which may not match those in place at the subsequent and previous elections.
-From the available Census information aggregated at the electorate level, 65 socio-demographic variables were defined for each of the electorates to be used in the analysis.
-
-Raw data was sourced online from the AEC and ABS websites in `.csv` and `.xlsx` files.
-The formats of these files differ over the years, making extracting the appropriate information a big task.
-The functions available in the `dplyr` [@dplyr] and `readxl` [@readxl] `R` packages are particularly useful, as they provide fast consistent tools for data manipulation and functions to import `.xlsx` files (respectively).
-The 2001 and 2006 Census data are published in a format where the information for each electorate is held in a separate document making it difficult to use the `dplyr` tools.
-Instead, cells have to be selected from each individual file to construct the desired variables.
-All scripts required for the data wrangling process can be found in the github repository for the `eechidna` `R` package [@eechidna], along with the raw data.
-The `eechidna` package makes this study entirely reproducible and provides a resource to help wrangle data for future Censuses and elections, when they become available.
-
-## Joining Census and election data
-
-### Differences between Census and election data {-}
-
-Between 2001 and 2016 there were six elections and four Censuses (see Figure \@ref(fig:timeline)).
-Electoral boundaries are redistributed regularly by the AEC, meaning that only in the years where both a Census and election occur are all boundaries likely to match --- the case for the 2001 and 2016 elections.
-Therefore, for the four elections between 2004 and 2013, both temporal and spatial differences in electorates need to be accounted for when joining the electoral two-party preferred vote with Census data.
-For these elections a spatio-temporal imputation method was employed to obtain electoral socio-demographics.
-This method uses Census information from both before and after the election of interest.
-
-```{r timeline, echo = FALSE, fig.cap="Timeline of Australian elections and Censuses. They do not always occur in the same year.", out.width = "90%", fig.align="center"}
-timeline <- data.frame(year = c(2001:2016)) %>%
-  mutate(
-    Election = ifelse(year %in% c(2001, 2004, 2007, 2010, 2013, 2016), 1, 0),
-    Census = ifelse(year %in% c(2001, 2006, 2011, 2016), 1.4, 0)
-  ) %>%
-  gather(c(Election, Census), key = event, value = measure) %>%
-  filter(measure != 0) %>%
-  mutate(year_vis = ifelse(event == "Census", year + 0.1, year))
-
-ggplot(timeline, aes(x = year_vis, y = measure)) +
-  geom_point(aes(color = as.factor(event), shape = as.factor(event)), size = 4) +
-  labs(color = "Event", shape = "Event", x = "Year") +
-  geom_hline(yintercept = 0, size = 1, color = "grey", show.legend = FALSE) +
-  geom_segment(x = 2001, xend = 2001, y = 1, yend = 0, color = "#00BFC4") +
-  geom_segment(x = 2004, xend = 2004, y = 1, yend = 0, color = "#00BFC4") +
-  geom_segment(x = 2007, xend = 2007, y = 1, yend = 0, color = "#00BFC4") +
-  geom_segment(x = 2010, xend = 2010, y = 1, yend = 0, color = "#00BFC4") +
-  geom_segment(x = 2013, xend = 2013, y = 1, yend = 0, color = "#00BFC4") +
-  geom_segment(x = 2016, xend = 2016, y = 1, yend = 0, color = "#00BFC4") +
-  geom_segment(x = 2001.1, xend = 2001.1, y = 1.4, yend = 0, color = "#F8766D") +
-  geom_segment(x = 2006.1, xend = 2006.1, y = 1.4, yend = 0, color = "#F8766D") +
-  geom_segment(x = 2011.1, xend = 2011.1, y = 1.4, yend = 0, color = "#F8766D") +
-  geom_segment(x = 2016.1, xend = 2016.1, y = 1.4, yend = 0, color = "#F8766D") +
-  theme(
-    text = element_text(size = 10),
-    axis.text.x = element_text(size = 6),
-    axis.text.y = element_blank(),
-    axis.title.y = element_blank(),
-    plot.background = element_blank(),
-    axis.ticks.y = element_blank(),
-    panel.background = element_blank()
-  ) +
-  scale_x_discrete(limits = c(2001:2016)) +
-  coord_fixed(ratio = 2)
-```
-
-### Spatio-temporal imputation {-}
-
-For each election, neighbouring Census information has to be combined in some way so that it represents the boundaries in place at the time of the election.
-This is done by taking the electoral boundaries and imputing the corresponding socio-demographic characteristics for each of neighbouring Censuses, thereby addressing the spatial aspect.
-Next, to deal with the temporal component, characteristics at the time of the election are constructed using linear interpolation between the spatially imputed neighbouring Census variables.
-
-To account for spatial differences, the piecewise approximation method in @Goodchild1993 is adopted.
-Consider a map of source zones $s = 1,\dots,S$, for which socio-demographic information is available, and a set of target zones $t = 1,\dots,T$ for which information is to be imputed.
-This is described in the context of a single election, and a single neighbouring Census.
-
-Let the map of electoral boundaries at the time of a Census define the source zones, and let the boundaries at the time of the election be the target zones.
-Denote the area of intersection between a source zone $s$ and a target zone $t$ as $A_{s,t}$.
-Additionally, let the population of the source zone $s$ be $U_s$ and the population of intersection between source zone $s$ and target zone $t$ be $P_{s,t}$.
-The estimated population of intersection is given by
-$$
-  \hat{P}_{s,t} = \frac{U_s*A_{s,t}}{\sum_{t=1}^T A_{s,t}},
-  \quad\text{for all $s=1,\dots,S$ and $t=1,\dots,T$}.
-$$
-Note that this estimator implicitly assumes that populations are uniformly distributed within each source zone.
-
-In order to calculate socio-demographic information for each of the target zones, a weighted average is taken using the estimated intersection populations as weights.
-Denote a given Census variable for the target zone by $C_t$, and the same Census variable for the source zone as $D_s$.
-Then, estimate $C_t$ using
-$$
-  \hat{C}_t = \frac{\sum_{s=1}^{S}{D_s*\hat{P}_{s,t}}}%
-  {\sum_{s=1}^{S}{\hat{P}_{s,t}}},
-  \quad\text{for each $t=1,\dots,T$}.
-$$
-This concludes the spatial imputation of the socio-demographic characteristics for one target zone (a single electoral boundary), at the time of only one of the neighbouring Censuses.
-This process is repeated for all of the target zones, and then for the other neighbouring Census.
-
-\enlargethispage*{0.5cm}
-
-To account for temporal changes, linear interpolation is used between Census years to get the final estimate of a Census variable for the target zone in the election year.
-Let $y_1$ be the year of the Census preceding an election, let $y_2$ be the year of the election, and $y_3$ be the year of the Census that follows.
-Add this year subscript to the Census variable estimate $\hat{C}_t$, resulting in $\hat{C}_{t,y}$.
-Linear interpolating between these Census years results an imputed value for the election year, given by
-$$
-  \hat{C}_{t,y_2} = \frac{y_3-y_2}{y_3-y_1} * \hat{C}_{t,y_1} +
-    \frac{y_2-y_1}{y_3-y_1} * \hat{C}_{t,y_3}.
-$$
-Implicitly this assumes that population characteristics change in a linear manner over time.
 
@@
 
-Census data is publicly available at different levels of aggregation, ranging from SA1 (over 50,000 zones) to electoral divisions (150 zones).
-For this study, electoral divisions are used as source zones, and the imputation method is applied to produce socio-demographic variables for each of the 2004, 2007, 2010 and 2013 elections.
-As mentioned earlier, there is no need to impute socio-demographic variables for the 2001 and 2016 elections.
-To illustrate the method, consider the imputation of socio-demographic variables for the electorate of Hume in New South Wales (NSW) at the time of the 2013 federal election.
-The boundaries shown in Figure \@ref(fig:hume13) define all target zones in NSW for 2013, with the target zone of interest (Hume) shaded purple.
+To illustrate the method, consider the imputation of socio-demographic variables for the electorate of Melbourne Ports in Victoria (VIC) at the time of the 2013 federal election.
+The boundaries shown in Figure \@ref(fig:melbports13) depict the target zone of Melbourne Ports for 2013 (shaded purple), along with its neighbouring electorates and the centroids of the SA1 regions from the 2016 Census.
 
-```{r hume13, fig.cap = "Some of the electoral boundaries in NSW for 2013, with the electoral boundary for Hume shown in purple.", out.height="28%"}
-hume_area13 <- nat_map13 %>%
-  filter(state %in% c("ACT", "NSW"), long < 154)
+```{r melbports13, fig.cap = "The electoral boundaries for Melbourne Ports (shown in purple) and surrounding electorates, with centroids for Statistical Area 1 regions from the 2016 Census overlaid.
+The centroids falling within the purple region are attributed to Melbourne Ports.", out.height="28%"}
+# Get Vic shape
+load("data/vic_map.rda")
+load("data/MP_sa1_dots_2011.rda")
 
-ggplot(data = hume_area13) +
-  geom_polygon(aes(x = long, y = lat, group = group, fill = elect_div == "HUME"),
-    colour = "grey50", alpha = 0.4
+p1 <- ggplot(data = vic_map) +
+  geom_polygon(aes(x = long, y = lat, group = group, fill = Elect_div == "Melbourne Ports"),
+    colour = "black", alpha = 0.4
   ) +
   scale_fill_manual(
-    name = "Electoral boundaries",
+    name = "Electoral Boundaries",
     values = c("white", "purple"),
-    labels = c("Other 2013 Electorates", "Hume 2013")
+    labels = c("Other Electorates", "Melbourne Ports (VIC)")
   ) +
   theme_map() +
-  coord_equal()
+  coord_equal(xlim = c(144.88, 145.07), ylim = c(-37.92, -37.78))
+
+p1 + geom_point(aes(x = long_c, y = lat_c, shape = shape), data = MP_sa1_dots_2011 %>% mutate(shape = "1"), colour = "grey25", size = 1.3) + scale_shape_manual(name = "Statistical Area 1", values = 3, labels = "Centroids")
 ```
-
-```{r hume16, fig.cap = "Census division boundaries in NSW for 2016, with the 2013 electoral boundary for Hume, shown in purple.
-The purple region is not contained within a single Census division.", out.height="28%"}
-hume_area16 <- nat_map16 %>%
-  filter(state %in% c("ACT", "NSW"), long < 154) %>%
-  mutate(year = "2016") %>%
-  bind_rows(
-    hume_area13 %>%
-      filter(elect_div == "HUME") %>%
-      mutate(year = "2013")
-  )
-
-ggplot(data = hume_area16) +
-  geom_polygon(aes(
-    x = long, y = lat, group = group, fill = year == "2013",
-    alpha = year == "2013", colour = year == "2013"
-  )) +
-  scale_fill_manual(
-    name = "Electoral boundaries",
-    values = c("white", "purple"),
-    labels = c("2016 Electorates", "Hume 2013")
-  ) +
-  scale_alpha_manual(values = c(0, 0.4)) +
-  scale_color_manual(values = c("grey50", NA)) +
-  theme_map() +
-  coord_equal() +
-  guides(alpha = FALSE, col = FALSE)
-```
-
-The corresponding source zones from the 2016 Census are shown in Figure \@ref(fig:hume16).
-As can be seen, the Hume boundary from the 2013 election (shaded purple) does not perfectly match any of the source zones.
 
@@
 
-# Modelling {#modelling}
+#### Robustness
 
-Following this process, electoral socio-demographic variables are available for each of the six elections and can be joined with their corresponding two-party preferred votes.
-Before choosing an appropriate model, two issues with the socio-demographic variables need to be addressed.
-First, variable scales change over the years, making it important to standardize variables.
-Second, many variables represent similar information and where appropriate, should be combined in some way.
-Principal component analysis is used to identify variables covary, from which intuitive groupings are selected to be combined into a single variable.
-This also reduces the dimension of the data.
-After these steps, a model specification is chosen.
-
-## Standardizing variables
-
-Many of the socio-demographic variables have changing scales over the years.
-For example, inflation-adjusted median rental prices increased across almost all electorates, with median rent of 200 dollars per week placing an electorate in the 90th percentile in 2001, but only the 30th percentile in 2016.
-In order for socio-demographic effects to be comparable across years, all explanatory variables are standardized to have mean zero and variance one within each election year.
-By standardizing, each variable is reported as a relative measure compared to all other electorates in the same year.
-
-## Creating factors
-
-There are only $N = 150$ observations (electorates) in each election and $p = 65$ socio-demographic variables in each cross-section, with many variables represent similar information about an electorate.
-Any model that uses all variables would face serious problems with multi-collinearity and over-fitting, which would likely lead to erroneous conclusions regarding variable significance.
-To address this, groups of variables that represent similar information are combined into "factors"[^2].
-
-[^2]: A preliminary step involved removing all age bands, because age is represented by median age, and to remove variables relating to particular denominations of Christianity.
-
-A factor is created from a group of variables if there is an intuitive reason as to why they should represent similar information and if there is evidence to suggest that they covary.
-For example, a potential group would be variables relating to electoral incomes --- median family, household and personal incomes.
-To determine which variables covary, principal component analysis is used on a combined dataset of socio-demographic variables from all six elections.
-It is appropriate to compute principal components in this way because when computed separately for each election, scree plots level off after four components and the loadings of the first four components are similar across the elections.
-
-Only the first four principal components from the combined dataset are considered, as the scree plot corresponding to the combined dataset levels off after the fourth component.
-Variables that have a large loading in a particular component are deemed to covary, with a loading with magnitude greater than 0.15 being considered large.
-Six factors are created using this criteria.
-These are:
-`Incomes` (median personal, household and family incomes);
-`Unemployment` (unemployment and labor force participation rates);
-`ProportyOwned` (rates of housing ownership, mortgages, renting and government housing);
-`RentLoanPrice` (median rental and loan repayments);
-`FamHouseSize` (average household size, ratio of people to families and household makeup (single person, couple with kids and couple without kids); and
-`Education` (high school and university qualifications, jobs requiring higher levels of education as well as vocational course completions and jobs that do not require higher education levels, such as laborer or tradesperson).
-For each of these groups, variables with positive loadings are added and those with negative loadings are subtracted to create a factor.
-After computing these sums, each factor is standardized to have mean zero and variance one, within each election.
-
-There are $p=30$ variables in the resultant predictor set, with all of these used in the regression for each election.
-
-## Regression incorporating spatially dependent errors
-
-An identical model specification is used for each of the six elections, with each election modelled separately.
-This allows the socio-demographic effects to be estimated separately for each election year, facilitating analysis of temporal changes in variable effects.
-This approach is preferable to using a single longitudinal model because it avoids any concerns of undue bias stemming from incorrectly imposed time-varying restrictions on any variable.
-Without such restrictions, a pooled cross-sectional model does not yield any distinct advantage over separate cross-sections.
-The panel approach is avoided because of how frequently electoral boundaries change, noting that electorates with the same name across elections are not guaranteed to represent the same geographical region.
-Therefore any fixed or random effects models would be difficult to estimate without implementing consistent boundaries, which would require further imputation.
-
-For each cross-section, let the response $\bm{y}$ be the vector two-party preferred vote in favour of the Liberal party; for example,  $y_i = 70$ represents a 70% preference for Liberal, 30% for Labor, in electorate $i$.
-Although $y_i$ lies in the interval $(0,100)$, observed values are never close to 0 or 100 (minimum $24.05 \%$ and maximum $74.90 \%$), so there is no need to formally impose the constraint of $y_i \in [0,100]$.
-Furthermore, the responses are found to be spatially correlated in each election (Moran's I test, $p \le 7\cdot10^{-15}$).
-This is not surprising as electorates are aggregate spatial units, and hence the spatial structure of the data must modelled appropriately.
-
-The spatial error model [@Anselin88] is chosen because it captures spatial heterogeneity by incorporating a spatially structured random effect vector [@LeSage2009].
-In this context, the random effect can be thought of as capturing the unobserved political climate in each electorate, where this climate is correlated with the climate in neighbouring electorates, under the assumption that the climate is independent of electoral socio-demographics.
-
-```{r printmodels, warning=FALSE, message=FALSE, include=FALSE}
-# Add rho and GLS residual standard error
-regtable <- stargazer::stargazer(
-  glsmod01, glsmod04, glsmod07, glsmod10, glsmod13, glsmod16,
-  header = FALSE,
-  column.sep.width = "2pt",
-  no.space=TRUE,
-  model.numbers=FALSE,
-  keep.stat = c("n","ser"),
-  dep.var.caption="",
-  dep.var.labels="\\vspace{-0.4cm}",
-  title = "Estimated spatial error model parameters (standard errors) for each of the six election years.\\vspace*{-0.4cm}",
-  #dep.var.labels = "Two-party preferred vote in favor of the Liberal party",
-  font.size = "scriptsize",
-  digits = 2,
-  column.labels = c("2001", "2004", "2007", "2010", "2013", "2016"),
-  label="tab:printmodels"
-  #notes = c("Estimated coefficients for variable named in column one shown for", "election year indicated by column heading, with estimated standard", "deviation for each coefficient shown below in parenthesis.
-Overall", "summary measures for each regression equation are provided in the", "bottom panel."),
-  #notes.align = "l"
-)
-# Add correlations
-getrho <- function(x, element) {x$rho_df[[element]]}
-rho <- map_dbl(list(glsmod01,glsmod04,glsmod07,glsmod10,glsmod13,glsmod16),
-           getrho, element="estimate")
-rhop <- map_dbl(list(glsmod01,glsmod04,glsmod07,glsmod10,glsmod13,glsmod16),
-           getrho, element="p")
-rhose <- map_dbl(list(glsmod01,glsmod04,glsmod07,glsmod10,glsmod13,glsmod16),
-           getrho, element="se")
-rholine1 <- paste("$\\rho$",
-            paste(" & ",
-                  sprintf("%3.2f",rho),
-                  ifelse(rhop < 0.01, "$^{***}$",
-                  ifelse(rhop < 0.05, "$^{**}$",
-                  ifelse(rhop < 0.10, "$^{*}$",""))),
-              sep="", collapse=""),
-            "\\\\", sep="", collapse="")
-rholine2 <- paste("",
-            paste(" & (",
-                  sprintf("%3.2f",rhose), ")",
-              sep="", collapse=""),
-            "\\\\", sep="", collapse="")
-regtable <- c(regtable[1:11], rholine1, rholine2, regtable[12:length(regtable)])
-# Add residual standard error (GLS)
-getsigma <- function(x, element) {x$sigma}
-sigma <- map_dbl(list(glsmod01,glsmod04,glsmod07,glsmod10,glsmod13,glsmod16),
-           getsigma)
-sigmaline <- paste("Residual Standard Error (GLS)",
-             paste(" & ",
-                   sprintf("%4.2f", sigma),
-                   sep="", collapse=""),
-             "\\\\", sep="", collapse="")
-regtable <- c(regtable[1:77],sigmaline, regtable[78:length(regtable)])
-# Squeeze some more space
-regtable[2] <- "\\begin{table}[!htbp]\\vspace*{-0.4cm}\\centering "     
-# Allow alignment on decimal point
-regtable[6] <- "\\begin{tabular}{@{\\extracolsep{1pt}}lrrrrrr}" 
-regtable[12:75] <- gsub(")","\\\\rlap{)}",regtable[12:75])
-regtable[12:75] <- gsub("$^{*}$","\\rlap{$^{*}$}", regtable[12:75], fixed=TRUE)
-regtable[12:75] <- gsub("$^{**}$","\\rlap{$^{**}$}", regtable[12:75], fixed=TRUE)
-regtable[12:75] <- gsub("$^{***}$","\\rlap{$^{***}$}", regtable[12:75], fixed=TRUE)
-# Column spacing
-regtable[5] <- "\\scriptsize\\tabcolsep=0.35cm"
-# Remove "Note"
-regtable[81] <- "& \\multicolumn{6}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\\\"
-```
-
-```{r printmodels2, results='asis', dependson='printmodels'}
-cat(regtable)
-```
-
-Spatial weights are calculated in accordance with the assumption that an electorate is equally correlated with any electorate that shares a part of its boundary.
-Let $\rho$ be the spatial autoregressive coefficient, $\bm{v}$ be a spherical error term, $\bm{W}$ be a matrix of spatial weights (containing information about the neighbouring regions), $\bm{X}$ be a matrix of socio-demographic covariates, $\bm{\beta}$ be a vector of regression coefficients and $\bm{a}$ be a spatially structured random effect vector.
-$$
-  \bm{y} = \bm{X} \bm{\beta} + \bm{a},
-$$
-and
-$$
-  \bm{a} = \rho \bm{W} \bm{a} + \bm{v},
-$$
-where $\bm{v} \sim N(\bm{0}, \sigma^2 \bm{I_n})$, and hence
-$$
-  \bm{y} = \bm{X} \bm{\beta} + (\bm{I}_n-\rho \bm{W})^{-1}\bm{v}.
-$$
-Estimation of the above spatial error model is undertaken using feasible generalized least squares.
-
-Table \@ref(tab:printmodels) details the estimated model coefficients and their estimated standard errors, for each of the six elections.
-An interpretation of these estimated values is provided in the next section.
-
-# Results
-
-## Spatial autoregressive parameter
-
-The spatial autoregressive coefficient $\rho$ is positive and significant in only the 2001 and 2016 elections (Figure \@ref(fig:rhovis)), meaning that in these elections, the political climate of an electorate appears to be affected by the attitudes of its neighbours.
-Conversely, in the other four elections, the spatial effect weakens to become insignificant.
-In these years, it appears that the spatial component does not explain anything not already explained by the electoral socio-demographics, meaning electorates effectively voted independently.
-
-```{r rhovis, fig.cap="Estimates of the spatial autoregressive parameter for each of the six elections, reported with their individual 95\\% confidence intervals.
-Only in 2001 and 2016 is there a significant spatial component.", out.height='20%', out.extra = ''}
-rho_df <- bind_rows(
-  glsmod16$rho_df %>% mutate(year = "2016"),
-  glsmod13$rho_df %>% mutate(year = "2013"),
-  glsmod10$rho_df %>% mutate(year = "2010"),
-  glsmod07$rho_df %>% mutate(year = "2007"),
-  glsmod04$rho_df %>% mutate(year = "2004"),
-  glsmod01$rho_df %>% mutate(year = "2001")
-) %>%
-  mutate(
-    upper95 = estimate + 1.96 * se,
-    lower95 = estimate - 1.96 * se
-  )
-
-rho_df %>%
-  ggplot() +
-  geom_point(aes(x = year, y = estimate), size = 3) +
-  geom_linerange(aes(x = year, ymin = lower95, ymax = upper95), size = 1.2) +
-  geom_hline(aes(yintercept = 0), alpha = 0.5, size = 1) +
-  scale_color_manual(values = c("grey50", "black")) +
-  labs(x = "Election year", y = "Estimate of spatial autoregressive parameter") +
-  lims(y = c(-1, 1)) +
-  guides(col = FALSE) +
-  coord_fixed(ratio = 1) +
-  theme_bw()
-```
-
-## Country-wide trend
-
-```{r plotintercept, fig.cap = "Estimated intercept for each election, which represents the two-party preferred vote for an electorate with mean characteristics.", out.height="20%", out.extra = ''}
-intercept_df <- bind_rows(
-  data.frame(summary(glsmod16)$tTable)[1, ] %>% mutate(year = "2016"),
-  data.frame(summary(glsmod13)$tTable)[1, ] %>% mutate(year = "2013"),
-  data.frame(summary(glsmod10)$tTable)[1, ] %>% mutate(year = "2010"),
-  data.frame(summary(glsmod07)$tTable)[1, ] %>% mutate(year = "2007"),
-  data.frame(summary(glsmod04)$tTable)[1, ] %>% mutate(year = "2004"),
-  data.frame(summary(glsmod01)$tTable)[1, ] %>% mutate(year = "2001")
-) %>%
-  rename(se = `Std.Error`, estimate = Value, p = `p.value`) %>%
-  mutate(upper95 = estimate + 1.96 * se, lower95 = estimate - 1.96 * se)
-
-intercept_df %>%
-  ggplot() +
-  geom_point(aes(x = year, y = estimate), size = 3) +
-  geom_linerange(aes(x = year, ymin = lower95, ymax = upper95), size = 1.2) +
-  geom_hline(aes(yintercept = 50), alpha = 0.5, size = 1) +
-  labs(x = "Election year", y = "Estimate of intercept") +
-  guides(col = FALSE) +
-  coord_fixed(ratio = 0.2) +
-  lims(y = c(45, 55)) +
-  theme_bw()
-```
-
-Since all socio-demographics have been standardized to have a mean of zero and a variance of one, the intercept in each model can be interpreted as the estimated two-party preferred vote for an electorate with mean characteristics[^3].
-Figure \@ref(fig:plotintercept) shows that the baseline of party preference has varied over the elections, with the biggest swing occurring in the 2007 election where the mean electorate shifted more than five percentage points in favour of the Labor party.
-
-[^3]: Mean of all variables aside from Judaism, Indigenous, Islam and Buddhism, where it assumes the mean of the log value.
-
-## Influential socio-demographics
-
-```{r prepplot, include = FALSE}
-p1 <- grid_visreg("Extractive", top = TRUE)
-p2 <- grid_visreg("DiffAddress")
-p3 <- grid_visreg("ManagerAdmin")
-p4 <- grid_visreg("Transformative")
-p5 <- grid_visreg("Incomes")
-p6 <- grid_visreg("Unemployment")
-
-p7 <- grid_visreg("MedianAge", top = TRUE)
-p8 <- grid_visreg("Judaism")
-p9 <- grid_visreg("OtherLanguage")
-temp <- grid_visreg("OtherLanguage")
-p10 <- grid_visreg("BornAsia")
-p11 <- grid_visreg("Education")
-p12 <- grid_visreg("BornSEEuro")
-p13 <- grid_visreg("DeFacto")
-```
-
-```{r partresplot1, fig.cap = "Partial residual plots by election year for a selection of predictors.
-Linear model with 95\\% confidence bands overlaid.
-Most predictors have a positive relationship: the larger the value the more likely the electorate preferences the Coalition.
-The relationship is relatively robust over time, with the exception of Unemployment.", fig.height = 8.5, fig.width = 6, warning = FALSE, fig.pos = 'h'}
-(p1 / p2 / (p3 + ylab("Two-party preferred vote (%)                     ")) / p4 / p5 / p6)
-```
-
-```{r partresplot2, fig.cap = "Partial residual plots by election year for a selection of predictors.
-Linear model with 95\\% confidence bands overlaid.
-Several predictors have a negative relationship: with larger values indicating the electorate more likely preferences Labor.
-Most relationships are relatively stable over elections, except OtherLanguage and Education.", fig.height = 8.5,  fig.width = 6, warning = FALSE, fig.pos = 'h'}
-(p7 / p8 / (p9 + ylab("Two-party preferred vote (%)                     ")) / p10 / p11 / p12 / p13)
-```
-
-To investigate the socio-demographics that have a strong effect on the two-party preferred vote, partial residual plots are used and shown in Figures \@ref(fig:partresplot1) and \@ref(fig:partresplot2).
-The partial residuals are the residuals from the fitted model with the estimated effect an individual variable added.
-These show the direction, size and significance of an estimated effect --- the slope of the prediction line matches the estimated coefficient, and the shaded region represents a 95% confidence band, computed using the method  @visreg.
-If a horizontal line can be drawn through the confidence band, then the effect is insignificant.
-The estimated intercept is also added to the partial residuals for interpretability.
-Plots for each election are faceted to compare the effects over time in Figures \@ref(fig:partresplot1) and \@ref(fig:partresplot2).
-Only socio-demographics that have a significant effect in at least one election are displayed in Figures \@ref(fig:partresplot1) and \@ref(fig:partresplot2).
-
-It is important here to note the ecological fallacy: insights are being drawn at the electorate level, and cannot be inferred for another disaggregate level (e.g. individual voters).
-
-### Income and unemployment {-}
-
-Typically the Labor party campaigns on more progressive policies, which often include tax reform that adversely affects higher income earners, and more generous social assistance programs.
-Perhaps it is due to these policies that higher income electorates appear more likely to support the Liberal party, as the `Incomes` factor has a positive effect on Liberal preference (see row 1 in Figure \@ref(fig:partresplot1)).
-This effect is significant in every election aside from 2004, where it is only marginally insignificant ($p = 0.0613$).
-Unemployment however, is not as influential.
-In 2001 and 2004, electorates with higher unemployment align with Labor, but over time this shifts towards support for the Liberal party, culminating in a significantly positive effect in 2016.
-
-### Industry and type of work {-}
-
-Electorates with higher proportions of workers in mining, gas, water, agriculture, waste and electricity (grouped as `Extractive` industries) are consistently linked with higher support for the Liberal party, with the magnitude of this effect slightly increasing over the years (see row 3 in Figure \@ref(fig:partresplot1)).
-This is unsurprising, as the Liberal party has close ties with these traditional energy industries, and typically present policies to reduce taxation on energy production.
-Furthermore, electorates with more workers in construction or manufacturing industries (`Transformative`) are also more likely to support the Liberal party (see row 4 in Figure \@ref(fig:partresplot1)).
-
-Similarly, the proportion of workers in managerial, administrative, clerical and sales roles (`ManagerAdmin`) is also a significant predictor of two-party preference vote across all six elections, with a higher proportion of people working these jobs increasing Liberal support.
-The magnitude of this effect also seems to increase over the years.
-
-### Household mobility {-}
-
-In each of the six elections, electorates with a higher proportion of people that have recently (in the past five years) moved house (`DiffAddress`) are more likely to support the Liberal party, although this effect was marginally insignificant in 2013 (see row 6 in Figure \@ref(fig:partresplot1).
-Having controlled for characteristics of house ownership and rental prices (via the factors `PropertyOwned` and `RentLoan` respectively), this effect is somewhat surprising.
-
-### Relationships {-}
-
-De facto relationships, but not marriages, are found to be an important (and significant) predictor of the two-party preferred vote in all six elections, with more de facto relationships associated with higher support for the Labor party.
-The proportion of individuals who are married however, is insignificant (not shown).
-
-### Age {-}
-
-Regions comprising more older people are often believed to be more conservative, and indeed it found that electorates with a higher median age are more likely to support the Liberal party --- although this effect is only significant in 2007 and 2010 (see row 2 in Figure \@ref(fig:partresplot2)).
-
-### Education {-}
-
-Since 2007, electorates with higher education levels are associated with supporting the Labor party, although this effect is only significant in 2016.
-Before 2007, education has an almost zero effect (see row 3 in Figure \@ref(fig:partresplot1)).
-
-### Diversity {-}
-
-Larger migrant populations from Asia, the Middle East, South-Eastern Europe, the United Kingdom and elsewhere, are either associated with Labor support, or have no effect.
-Of these areas, only South-Eastern European populations appear significant in each election, with the proportion of Asian migrants also being significant in 2010.
-Speaking other languages (aside from English) however, appears to have a far stronger effect, as observed through the `OtherLanguage` variable.
-Electorates with more diverse speech are associated with higher support for the Liberal party from 2004 onwards, with this effect being significant in 2007, 2010 and 2016.
-Furthermore, of the variables relating to religion, only Judaism shows a consistent effect, with electorates with relatively large Jewish populations more likely to vote Liberal.
-
-### A note on similar variables {-}
-
-Many of the Census variables represent similar information, which is why factors were created and some variables were removed.
-However, some variables remain which are closely related.
-For example, an electorate's income level (via `Incomes`) is likely to be related to electoral unemployment and labor force participation (via `Unemployment`).
-In 2001, the coefficient estimate for `Unemployment` is negative but not significant, whilst the `Incomes` variables is significant.
-If the `Incomes` variable is removed from the model in 2001, `Unemployment` absorbs the negative effect, becoming significant ($p = 0.0056$).
-
-```{r removeunemp, eval = FALSE}
-unemp01 <- my_fgls(full_formula,
-  my_data = model_df %>% filter(year == "2001") %>% dplyr::select(-Incomes),
-  sp_weights = sp_weights_01
-)
-
-summary(unemp01)$tTable["Unemployment",]
-```
-
-## A closer look at the residuals
-
-```{r resids, include = FALSE}
-all_resids <- data.frame(
-  Residuals = c(glsmod16$actual_residuals, glsmod13$actual_residuals, glsmod10$actual_residuals, glsmod07$actual_residuals, glsmod04$actual_residuals, glsmod01$actual_residuals),
-  gls_residuals = c(glsmod16$residuals, glsmod13$residuals, glsmod10$residuals, glsmod07$residuals, glsmod04$residuals, glsmod01$residuals),
-  bind_rows(
-    glsmod16$my_data %>% left_join(tpp16 %>% dplyr::select(DivisionNm, UniqueID, StateAb), by = "DivisionNm"),
-    glsmod13$my_data %>% left_join(tpp13 %>% dplyr::select(DivisionNm, UniqueID, StateAb), by = "DivisionNm"),
-    glsmod10$my_data %>% left_join(tpp10 %>% dplyr::select(DivisionNm, UniqueID, StateAb), by = "DivisionNm"),
-    glsmod07$my_data %>% left_join(tpp07 %>% dplyr::select(DivisionNm, UniqueID, StateAb), by = "DivisionNm"),
-    glsmod04$my_data %>% left_join(tpp04 %>% dplyr::select(DivisionNm, UniqueID, StateAb), by = "DivisionNm"),
-    glsmod01$my_data %>% left_join(tpp01 %>% dplyr::select(DivisionNm, UniqueID, StateAb), by = "DivisionNm")
-  )
-)
-# %>% dplyr::select(Residuals, DivisionNm, UniqueID, StateAb, year, LNP_Percent, gls_residuals)
-```
-
-### Residuals by state {-}
-
-It is often hypothesized that states have systematic differences that cause their electorates to vote differently.
-Boxplots of residuals grouped by state (Figure \@ref(fig:resstate)) reveal that the data reflects this -- there appears to be a state-specific effect not captured by the models.
-Tasmania and the Australian Capital Territory appear to have a bias towards Labor, whereas the Northern Territory tends towards voting Liberal.
-However, there are relatively few electorates in each of these states (five, two and two respectively), so this apparent result may be due to incumbent effects rather than an actual state-specific bias.
-
-```{r resstate, fig.cap = "Boxplot of residuals by state with jittered points.
-States ordered by median residual.
-A state-specific bias not captured by the model is evident.", fig.asp=0.6, out.width="70%", fig.pos = 'h', out.extra = ''}
-all_resids %>%
-  ggplot(aes(x = fct_reorder(StateAb, Residuals), y = Residuals)) +
-  geom_hline(aes(yintercept = 0), col = "grey50") +
-  geom_boxplot() +
-  geom_jitter(alpha = 0.5, shape = 1) +
-  guides(col = FALSE, shape = FALSE, DivisionNm = FALSE) +
-  theme_bw()
-```
-
-### Outlier electorates {-}
-
-```{r readcooksd, eval = FALSE}
-read_chunk('cooksd.R')
-```
-
-```{r getcooksd, eval = FALSE}
-<<GetCookDistance>>
-```
-
-```{r cooksd, eval = FALSE}
-cooksd_df <- data.frame(
-  Residuals = c(glsmod16$actual_residuals, glsmod13$actual_residuals, glsmod10$actual_residuals, glsmod07$actual_residuals, glsmod04$actual_residuals, glsmod01$actual_residuals),
-  gls_residuals = c(glsmod16$residuals, glsmod13$residuals, glsmod10$residuals, glsmod07$residuals, glsmod04$residuals, glsmod01$residuals),
-  cooksd = c(glsmod16$cooksd, glsmod13$cooksd, glsmod10$cooksd, glsmod07$cooksd, glsmod04$cooksd, glsmod01$cooksd),
-  bind_rows(
-    glsmod16$my_data %>% left_join(tpp16 %>% dplyr::select(DivisionNm, UniqueID, StateAb), by = "DivisionNm"),
-    glsmod13$my_data %>% left_join(tpp13 %>% dplyr::select(DivisionNm, UniqueID, StateAb), by = "DivisionNm"),
-    glsmod10$my_data %>% left_join(tpp10 %>% dplyr::select(DivisionNm, UniqueID, StateAb), by = "DivisionNm"),
-    glsmod07$my_data %>% left_join(tpp07 %>% dplyr::select(DivisionNm, UniqueID, StateAb), by = "DivisionNm"),
-    glsmod04$my_data %>% left_join(tpp04 %>% dplyr::select(DivisionNm, UniqueID, StateAb), by = "DivisionNm"),
-    glsmod01$my_data %>% left_join(tpp01 %>% dplyr::select(DivisionNm, UniqueID, StateAb), by = "DivisionNm"))
-)
-```
-
-```{r plotcooksd, fig.cap = "Electorates' Cook's Distance values in each election.", out.height = "15%", eval = FALSE}
-cooksd_df %>%
-  ggplot(aes(x = cooksd)) +
-  geom_dotplot() +
-  facet_wrap(~year, nrow = 1) +
-  # coord_fixed(ratio = .5) +
-  labs(x = "Cook's Distance", y = "") +
-  theme(axis.text.x = element_text(angle = 60, hjust = 1, size = 8), axis.text.y = element_text(size = 8))
-```
-
-```{r, eval = FALSE}
-cooksd_df %>%
-  filter(cooksd > 0.1) %>%
-  select(DivisionNm, StateAb, year, cooksd, Residuals, LNP_Percent, everything()) %>%
-  View()
-```
-
-Based on the distribution of the Cook's distance values, a Cook's distance greater than $0.1$ is considered to be influential and a potential outlier.
-The electorate of Sydney (NSW) has a large Cook's distance from 2001 to 2013, due to its diverse population (language, birthplace and religion), high number of defacto relationships, high income, high household mobility and small amount of workers in extractive and transformative jobs.
-It has remained a strong supporter of the Labor party and the Liberal vote is severely overpredicted by the model, making it an outlier.
-Nearby in metropolitan NSW, the electorate of Wentworth is found to be an outlier in all but the 2007 election.
-Although historically Liberal, its two-party vote jumped by over 10 percentage points in 2010 without experiencing any notable changes in its socio-demographic makeup --- implying that this may be the direct effect of its Liberal member, Malcolm Turnbull, becoming the leader of the Liberal party.
-Liberal support in Wentworth is underpredicted by the model in each year, and more so with Turnbull as Liberal leader.
-
-Lingiari, an electorate taking up almost all of the Northern Territory, is an outlier in the 2001--2007 elections due to its large Indigenous population, young age profile and low rates of property ownership.
-Fowler (NSW) has a diverse population with a high proportion of migrants, many Buddhists and Muslims, and has strong Labor support, making it influential in 2001, 2004 and 2010.
-Other electorates with large Cook's distance are Barton (NSW) and Leichhardt (QLD) in 2016, and Canberra (ACT) in 2007.
-
-```{r, eval = FALSE}
-x <- model_df %>%
-  filter(year == "2016") %>%
-  select(-c(year, DivisionNm)) %>%
-  as.matrix()
-myhat <- hat(x)
-
-plot(
-  glsmod07$cooksd,
-  hat(
-    model_df %>%
-      filter(year == "2007") %>%
-      select(-c(year, DivisionNm)) %>%
-      as.matrix()
-  )
-)
-```
-
-# Conclusion
-
-This paper explores the effects of electoral socio-demographic characteristics on the two-party preferred vote in the 2001--2016 elections, using information from the corresponding Australian federal elections and Censuses.
-As a Census does not always occur in the same year as an election, Census data for the 2004--2013 elections are generated by employing a method of spatio-temporal imputation.
-This imputes electoral socio-demographics for the electoral boundaries in place at the time of the election --- an approach that is distinctly different from previous work on modelling election outcomes, where Census and election data are typically joined without addressing their temporal differences.
-Before estimating a model, these socio-demographic variables are standardized (to adjust for changing variable scales) and many variables (representing similar information) are combined into factors, resulting in a reduced predictor set.
-A spatial error model is then estimated for each election, accounting for the inherent spatial structure of the data.
-
-Across the past six elections, most of the socio-demographics that drive the electoral two-party preferred vote are found to remain steady, whilst a few (typically weaker) effects vary over time.
-Industry and type of work are particularly influential, with energy-related and manufacturing/construction jobs, as well as administrative roles being strongly linked with the Liberal party in all elections.
-Incomes have a similarly consistent effect, with higher income areas supporting Liberal.
-Higher levels of unemployment shift from weak association with Labor to a significant Liberal effect over the years, and higher education levels are associated with Labor from 2007 (although only significant in 2016).
-It is also found that electorates with higher household mobility support Liberal, birthplace diversity favours Labor and more de facto relationships align with Labor preference --- although marriages, family and household sizes have no material influence.
-Furthermore, the neighbourhood (spatial) effects are found to be positive in all elections, although only significant in 2001 and 2016, meaning that in the 2004--2013 elections, electorates effectively voted independently.
-
-The findings in this paper complement the existing literature by modelling temporal trends, which as far as the authors are aware, has not been done previously for Australian elections using a regression framework.
-It is also the first study to model any Australian election since 2010 using Census information.
-
-Additionally, a key contribution of this research is the wrangling of the raw data and imputed data sets for the 2004, 2007, 2010 and 2013 elections, which have been contributed to the `eechidna` `R` package --- providing a rich, accessible data resource for future Australian electoral analysis.
-
-# Acknowledgements
-
-This paper was produced using `RMarkdown` [@rmarkdown] and `knitr` [@knitr].
-All corresponding code for this paper can be found in the github repository [github.com/jforbes14/eechidna-paper](https://github.com/jforbes14/eechidna-paper), and the data used is available in the `eechidna` package [@eechidna].
-All raw data was obtained from the Australian Electoral Commission, the Australian Bureau of Statistics and the Australian Government.
-
-# Software
-
-All election and Census datasets, along with electoral maps and more, are available in the `eechidna` (Exploring Election and Census Highly Informative Data Nationally for Australia) `R` package, which can be downloaded from CRAN.
-The `eechidna`  package makes it easy to look at the data from the Australian Federal elections and Censuses that occurred between 2001 and 2016.
-This study contributed a large revision to the `eechidna` package, which included the addition of election and Census data for 2001--2010, voting outcomes for polling booths and imputed Census data for election years.
-For more details on using `eechidna`, please see the articles (vignettes) on the github page [ropenscilabs.github.io/eechidna/](https://ropenscilabs.github.io/eechidna/).
-
-The authors would like to sincerely thank Anthony Ebert, Heike Hofmann, Thomas Lumley, Ben Marwick, Carson Sievert, Mingzhu Sun, Dilini Talagala, Nicholas Tierney, Nathaniel Tomasetti, Earo Wang and Fang Zhou, all of whom have contributed to the `eechidna` package.
-
-# References
+Two robustness checks are conducted to confirm model stability.
+First, a model for each election is fit using only the variables that were found to be significant in at least one of the six elections.
+The estimated coefficients of the variables in these models all fall within their respective $95\%$ confidence intervals from the full models, confirming that the significant effects are robust to variable selection.
+The second check involves computing pairwise correlations amongst the explanatory variables, and seeing if the removal of one variable (from a pair) affects the other's estimated effect.
+For example, the variables `Born_Asia` and `BornElsewhere` have a pairwise correlation of $0.93$ - the largest of any pair.
+A model is fit omitting `BornElsewhere` (the reduced model).
+The estimated coefficent of `Born_Asia` is then compared with the confidence interval for `Born_Asia` in the full model.
+It is found that for each of the top ten pairwise correlations, the estimated effect in the reduced model lies within the confidence interval from the full model.
+This second check indicates that variables with high pairwise correlation are not masking the effects of each other.

### 2019-08-25 anzjs_elections.Rmd
added section on robustness of model, minor edit to modelling section

@@
 The ABS uses electoral boundaries as defined by the AEC at the time of each Census, which may not match those in place at the subsequent and previous elections.
-From the available Census information aggregated at the electorate level, 65 socio-demographic variables were defined for each of the electorates to be used in the analysis.
+From the available Census information aggregated at the electorate level, 50 socio-demographic variables were defined for each of the electorates to be used in the analysis.
 
@@
 To determine which variables should be grouped, principal component analysis (PCA) is used.
-The intuition here is that PCA will identify which variables covary in the directions of maximal variance, from which intuitive groupings of variables can be chosen to combine into a single variable.
+The intuition here is that PCA will identify which variables covary, from which intuitive groupings of variables can be chosen to combine into a single variable.
 After these steps, a model specification is chosen.
@@
 To determine which variables covary, principal component analysis is used on a combined dataset of socio-demographic variables from all six elections[^2].
-The only variables exempted from the principal component analysis are the four variables representing age brackets (the proportion of the population aged 0-19 years old, 20-34 years old, 45-54 years old and 55 years plus), which are to be included in the model without grouping.
+The only variables exempted from the principal component analysis are the four variables representing age brackets (the proportion of the population aged 0-19 years old, 20-34 years old, 45-54 years old and 55 years plus), which are to be included in the model as separate variables by construction.
 
@@
 
+### Robustness
+
+Two robustness checks are conducted to confirm model stability.
+First, a model for each election is fit using only the variables that were found to be significant in at least one of the six elections.
+The estimated coefficients of the variables in these models all fall within their respective $95\%$ confidence intervals from the full models, confirming that the significant effects are robust to variable selection.
+The second check involves computing pairwise correlations amongst the explanatory variables, and seeing if the removal of one variable (from a pair) affects the other's estimated effect.
+For example, the variables `Born_Asia` and `BornElsewhere` have a pairwise correlation of $0.93$ - the largest of any pair.
+A model is fit omitting `BornElsewhere` (the reduced model).
+The estimated coefficent of `Born_Asia` is then compared with the confidence interval for `Born_Asia` in the full model.
+It is found that for each of the top ten pairwise correlations, the estimated effect in the reduced model lies within the confidence interval from the full model.
+This second check indicates that variables with high pairwise correlation are not masking the effects of each other.
+
 ### A note on similar variables {-}
+(Maybe remove this?)
 

### 2019-08-25 anzjs_elections.Rmd
clarification of partial residuals

@@
 To investigate the socio-demographics that have a strong effect on the two-party preferred vote, partial residual plots are used and shown in Figures \@ref(fig:partresplot1) and \@ref(fig:partresplot2).
-The partial residuals are the residuals from the fitted model with the estimated effect an individual variable added.
-These show the direction, size and significance of an estimated effect --- the slope of the prediction line matches the estimated coefficient, and the shaded region represents a 95% confidence band, computed using the method in @visreg.
+Partial residuals for a variable are the residuals from the fitted model with the estimated effect of that variable added to it.
+These plots show the direction, size and significance of an estimated effect, as well as any deviations from linearity.
+In each plot, the slope of the prediction line matches the estimated coefficient and the shaded region represents a 95% confidence band.
+Plots are computed using the method in @visreg.
 If a horizontal line can be drawn through the confidence band, then the effect is insignificant.

### 2019-08-25 anzjs_elections.Rmd
clarification of time-varying restrictions

@@
 This allows the socio-demographic effects to be estimated separately for each election year, facilitating analysis of temporal changes in variable effects.
-This approach is preferable to using a single longitudinal model because it avoids any concerns of undue bias stemming from incorrectly imposed time-varying restrictions on any variable.
-Without such restrictions, a pooled cross-sectional model does not yield any distinct advantage over separate cross-sections.
-The panel approach is avoided because of how frequently electoral boundaries change, noting that electorates with the same name across elections are not guaranteed to represent the same geographical region.
-Therefore any fixed or random effects models would be difficult to estimate without implementing consistent boundaries, which would require further imputation.
+This approach is preferable to using a single longitudinal model because it avoids imposing any time-varying restrictions on the effect of a variable.
+To gain any advantage over modelling each election as a separate cross-section, a pooled cross-sectional model would have to force the effects at least some variables to be invariant over time.
+This would not allow for conclusions to be drawn about how the effects of these variables change over time.
+Furthermore, the panel approach is avoided because of how frequently electoral boundaries change, noting that electorates with the same name across elections are not guaranteed to represent the same geographical region.
+Therefore any fixed or random effects models would be difficult to estimate without implementing consistent boundaries, which would require further imputation (of voting information).
 

### 2019-08-25 anzjs_elections.Rmd
add more emphasis on this being a study at the aggregate level and the ecological fallacy - moved to introduction.

@@
 The two-party preferred vote, a measure of support between these two parties, fluctuated between 47.3% and 53.5% (in favour of the Liberal party) over this period.
-This study explores how electoral socio-demographic characteristics relate to two-party preference, and whether their effects have changed over time.
+This study explores how electoral (aggregate) socio-demographic characteristics relate to two-party preference, and whether their effects have changed over time.
 
-Data on electoral socio-demographics are derived from the Australian Census, and vote counts are obtained from Australian federal elections.
+The Australian Electoral Commission (AEC) divides Australia into 150 regions, called electorates, with each corresponding to a single seat in the House of Representatives.
+If a party wins a majority of seats, they become the governing party.
+Data on the socio-demographics of these electorates are derived from the Australian Census, and vote counts are obtained from Australian federal elections.
 Joining these two data sources is problematic as there is an inherent asynchronicity in the two types of events.
-A Census is conducted by the Australian Bureau of Statistics (ABS) every five years, whereas federal elections, conducted by the Australian Electoral Commission (AEC), usually occur every three years or so.
+A Census is conducted by the Australian Bureau of Statistics (ABS) every five years, whereas federal elections, conducted by the AEC, usually occur every three years or so.
 The first problem addressed is that of constructing appropriate Census data for the 2004, 2007, 2010 and 2013 elections --- election years in which a Census does not occur.
@@
 Instead of analyzing a single election in isolation, this paper employs a consistent model framework across six elections so that temporal changes in the effects of socio-demographics can be observed.
-Each federal election is modelled with a cross-sectional dataset.
-<!--The use of a regression framework to examine these socio-political relationships over time is seemingly absent from previous Australian studies.
-It also appears that no study has attempted any type of statistical analysis of socio-demographics in conjunction with voter behaviour in Australia since 2007, making this paper distinctly different from previous work.-->
-The cross-sectional dataset for each election used here consists of the two-party preferred vote (as the response variable), and a set of common socio-demographic variables (as the explanatory variables) that characterize each electorate.
-To prepare these datasets, socio-demographic variables are first standardized, and then a principal component analysis is used to group variables into "factors".
+Each federal election is modelled with a cross-sectional dataset, where each observation is one of the 150 electorates.This dataset consists of the two-party preferred vote (as the response variable) and a set of common socio-demographic variables (as the explanatory variables).<!--The use of a regression framework to examine these socio-political relationships over time is seemingly absent from previous Australian studies.
+It also appears that no study has attempted any type of statistical analysis of socio-demographics in conjunction with voter behaviour in Australia since 2007, making this paper distinctly different from previous work.--> To prepare these datasets, socio-demographic variables are first standardized, and then a principal component analysis is used to group variables into "factors".
 To account for the inherent spatial structure of the data, a spatial error model is then estimated for each election.
+In interpreting these models, is important to be mindful of the ecological fallacy.
+Insights are being drawn at the electorate level and cannot be inferred for another disaggregate level (this means that drivers of individual voter behaviour may vary from what is observed at the electorate level).
 
@@
 
-It is important here to note the ecological fallacy: insights are being drawn at the electorate level, and cannot be inferred for another disaggregate level (e.g. individual voters).
-
 ### Income and unemployment {-}

### 2019-08-30 anzjs_elections.Rmd
type of work update

@@
     BornAsia = Born_Asia,
+    BornUK = Born_UK,
+    BornMidEast = Born_MidEast,
     OneParentHouse = OneParent_House,
@@
 p1 <- grid_visreg("Extractive", top = TRUE)
-p2 <- grid_visreg("DiffAddress")
-p3 <- grid_visreg("ManagerAdmin")
-p4 <- grid_visreg("Transformative")
-p5 <- grid_visreg("Incomes")
-p6 <- grid_visreg("Unemployment")
+p2 <- grid_visreg("ManagerAdmin")
+p3 <- grid_visreg("Transformative")
+p4 <- grid_visreg("SocialServ")
+p5 <- grid_visreg("DiffAddress")
+p6 <- grid_visreg("Incomes")
+p7 <- grid_visreg("OtherLanguage")
+p8 <- grid_visreg("Pop_00_19")
 
-p7 <- grid_visreg("MedianAge", top = TRUE)
-p8 <- grid_visreg("Judaism")
-p9 <- grid_visreg("OtherLanguage")
-temp <- grid_visreg("OtherLanguage")
-p10 <- grid_visreg("BornAsia")
+p9 <- grid_visreg("Unemployment", top = TRUE)
+p10 <- grid_visreg("BornSEEuro")
 p11 <- grid_visreg("Education")
-p12 <- grid_visreg("BornSEEuro")
+p12 <- grid_visreg("Married")
 p13 <- grid_visreg("DeFacto")
+p14 <- grid_visreg("Pop_35_54")
+p15 <- grid_visreg("Pop_20_34")
 ```
@@
 Most predictors have a positive relationship: the larger the value the more likely the electorate preferences the Coalition.
-The relationship is relatively robust over time, with the exception of Unemployment.", fig.height = 8.5, fig.width = 6, warning = FALSE, fig.pos = 'h'}
-(p1 / p2 / (p3 + ylab("Two-party preferred vote (%)                     ")) / p4 / p5 / p6)
+The relationship is relatively robust over time, with the exception of OtherLanguageHome and Pop_00_19.", fig.height = 8.5, fig.width = 6, warning = FALSE, fig.pos = 'h'}
+(p1 / p2 / p3 / (p4 + ylab("Two-party preferred vote (%)                     ")) / p5 / p6 / p7 / p8)
 ```
@@
 Several predictors have a negative relationship: with larger values indicating the electorate more likely preferences Labor.
-Most relationships are relatively stable over elections, except OtherLanguage and Education.", fig.height = 8.5,  fig.width = 6, warning = FALSE, fig.pos = 'h'}
-(p7 / p8 / (p9 + ylab("Two-party preferred vote (%)                     ")) / p10 / p11 / p12 / p13)
+Most relationships are relatively stable over elections, except Unemployment and Education.", fig.height = 8.5,  fig.width = 6, warning = FALSE, fig.pos = 'h'}
+(p9 / p10  / p11 / (p12 + ylab("Two-party preferred vote (%)")) / p13 / p14 / p15)
 ```
@@
 To investigate the socio-demographics that have a strong effect on the two-party preferred vote, partial residual plots are used and shown in Figures \@ref(fig:partresplot1) and \@ref(fig:partresplot2).
-Partial residuals for a variable are the residuals from the fitted model with the estimated effect of that variable added to it.
+Partial residuals, for a given variable, are the residuals from the fitted model with the estimated effect of that variable added to it.
 These plots show the direction, size and significance of an estimated effect, as well as any deviations from linearity.
@@
 Plots for each election are faceted to compare the effects over time in Figures \@ref(fig:partresplot1) and \@ref(fig:partresplot2).
-Only socio-demographics that have a significant effect in at least one election are displayed in Figures \@ref(fig:partresplot1) and \@ref(fig:partresplot2).
+Only socio-demographics that have a significant effect in at least two elections are displayed in Figures \@ref(fig:partresplot1) and \@ref(fig:partresplot2).
+
+### Industry and type of work {-}
+
+Electorates with higher proportions of workers in mining, gas, water, agriculture, waste and electricity (grouped as `Extractive` industries) are consistently linked with higher support for the Liberal party, with the magnitude of this effect slightly increasing over the years (see row 3 in Figure \@ref(fig:partresplot1)).
+This is unsurprising, as the Liberal party has close ties with these traditional energy industries, and typically present policies to reduce taxation on energy production.
+Furthermore, electorates with more workers in construction or manufacturing industries (`Transformative`) are also more likely to support the Liberal party (see row 4 in Figure \@ref(fig:partresplot1)), from 2004 onwards.
+
+Similarly, the proportion of workers in managerial, administrative, clerical and sales roles (`ManagerAdmin`), is also a significant predictor of two-party preference vote across all six elections, with a higher proportion of people working these jobs increasing Liberal support.
+
+Of these job related variables, the most surprising effect is that associated with the proportion of workers in education, healthcare, social work, community and arts (`SocialServ`).
+Typically the Labor party has more generous funding schemes affecting these areas of work, so one might expect `SocialServ` to have a negative effect on two-party preference.
+However, in every election this effect is found to be positive and significant.
 
@@
 In 2001 and 2004, electorates with higher unemployment align with Labor, but over time this shifts towards support for the Liberal party, culminating in a significantly positive effect in 2016.
-
-### Industry and type of work {-}
-
-Electorates with higher proportions of workers in mining, gas, water, agriculture, waste and electricity (grouped as `Extractive` industries) are consistently linked with higher support for the Liberal party, with the magnitude of this effect slightly increasing over the years (see row 3 in Figure \@ref(fig:partresplot1)).
-This is unsurprising, as the Liberal party has close ties with these traditional energy industries, and typically present policies to reduce taxation on energy production.
-Furthermore, electorates with more workers in construction or manufacturing industries (`Transformative`) are also more likely to support the Liberal party (see row 4 in Figure \@ref(fig:partresplot1)).
-
-Similarly, the proportion of workers in managerial, administrative, clerical and sales roles (`ManagerAdmin`) is also a significant predictor of two-party preference vote across all six elections, with a higher proportion of people working these jobs increasing Liberal support.
-The magnitude of this effect also seems to increase over the years.
 

### 2019-08-30 anzjs_elections.Rmd
Age, Relationships, Income & Unemployment updates

@@
 Perhaps it is due to these policies that higher income electorates appear more likely to support the Liberal party, as the `Incomes` factor has a positive effect on Liberal preference (see row 1 in Figure \@ref(fig:partresplot1)).
-This effect is significant in every election aside from 2004, where it is only marginally insignificant ($p = 0.0613$).
+This effect is significant in every election aside from 2004 and 2010.
 Unemployment however, is not as influential.
-In 2001 and 2004, electorates with higher unemployment align with Labor, but over time this shifts towards support for the Liberal party, culminating in a significantly positive effect in 2016.
+In 2001 and 2004, electorates with higher unemployment align with Labor, but over time this shifts towards support for the Liberal party, culminating in a positive (insignificant) effect in 2016.
 
@@
 
-De facto relationships, but not marriages, are found to be an important (and significant) predictor of the two-party preferred vote in all six elections, with more de facto relationships associated with higher support for the Labor party.
-The proportion of individuals who are married however, is insignificant (not shown).
+From 2007 onwards, both marriages (`Married`) and de facto relationships (`DeFacto`) are found to be strong predictors of the two-party preferred vote in opposition to the Liberal party.
+In 2001 and 2004 neither of these variables were significant (see rows 4 and 5 in Figure \@ref(fig:partresplot2)).
 
@@
 
-Regions comprising more older people are often believed to be more conservative, and indeed it found that electorates with a higher median age are more likely to support the Liberal party --- although this effect is significant only in 2007 and 2010 (see row 2 in Figure \@ref(fig:partresplot2)).
+The older Australian population is often believed to be more conservative, and the left leaning political parties (including Labor) typically have a stronger appeal to younger people.
+This effect is indeed observed across the six elections, with populations between 20 and 34 years of age (`Pop_20_34`) being very strongly aligned with Labor preference (bottom row in Figure \@ref(fig:partresplot2)).
+Larger populations of 35 to 54 year olds (`Pop_35_54`) are also associated with Labor, but the magnitude of this effect is far smaller.
+Populations under 20 years of age is only significant in 2007, where `Pop_00_19` increased Liberal support.
 

### 2019-08-30 anzjs_elections.Rmd
Diversity, education and households. also added OneParentHouse to plot.

@@
 p13 <- grid_visreg("DeFacto")
-p14 <- grid_visreg("Pop_35_54")
-p15 <- grid_visreg("Pop_20_34")
+p14 <- grid_visreg("OneParentHouse")
+p15 <- grid_visreg("Pop_35_54")
+p16 <- grid_visreg("Pop_20_34")
 ```
@@
 Most relationships are relatively stable over elections, except Unemployment and Education.", fig.height = 8.5,  fig.width = 6, warning = FALSE, fig.pos = 'h'}
-(p9 / p10  / p11 / (p12 + ylab("Two-party preferred vote (%)")) / p13 / p14 / p15)
+(p9 / p10  / p11 / (p12 + ylab("Two-party preferred vote (%)                     ")) / p13 / p14 / p15 / p16)
 ```
@@
 
-### Household mobility {-}
+### Households {-}
 
-In each of the six elections, electorates with a higher proportion of people that have recently (in the past five years) moved house (`DiffAddress`) are more likely to support the Liberal party, although this effect was marginally insignificant in 2013 (see row 6 in Figure \@ref(fig:partresplot1).
-Having controlled for characteristics of house ownership and rental prices (via the factors `PropertyOwned` and `RentLoan` respectively), this effect is somewhat surprising.
+In 2001, 2004 and 2016, higher proportions of people that have recently (in the past five years) moved house (`DiffAddress`) increased electoral support for the Liberal party (see row 5 in Figure \@ref(fig:partresplot1)).
+Having controlled for characteristics of house ownership and rental prices (`PropertyOwned` and `RentLoan` - both of which have no effect), this effect is somewhat surprising.
+
+Higher proportions of single parent households are associated with Labor support in all elections (albeit insignificant in 2004, see row 6 in Figure \@ref(fig:partresplot2)), whereas the electoral family and household sizes (via the `FamHouseSize` variable) do not appear to be associated with either party in any election.
 
@@
 
-Since 2007, electorates with higher education levels are associated with supporting the Labor party, although this effect is significant only in 2016.
-Before 2007, education has an almost zero effect (see row 3 in Figure \@ref(fig:partresplot1)).
+Since 2007, electorates with higher education levels are associated with supporting the Labor party, with this effect being significant in 2007, 2013 and 2016 and only marginally insignificant in 2010.
+In the elections before 2007, education has a negligible effect (see row 3 in Figure \@ref(fig:partresplot1)).
+Additionally, student populations (`CurrentlyStudying`) do not affect electoral party preference in any election (not shown).
 
@@
 Larger migrant populations from Asia, the Middle East, South-Eastern Europe, the United Kingdom and elsewhere, are either associated with Labor support, or have no effect.
-Of these areas, only South-Eastern European populations appear significant in each election, with the proportion of Asian migrants also being significant in 2010.
+Of these areas, only South-Eastern European populations significant affect party preference, with larger populations associating with Labor in 2013 and 2016 (row 2, Figure \@ref(fig:partresplot2)).
 Speaking other languages (aside from English) however, appears to have a far stronger effect, as observed through the `OtherLanguage` variable.
-Electorates with more diverse speech are associated with higher support for the Liberal party from 2004 onwards, with this effect being significant in 2007, 2010 and 2016.
-Furthermore, of the variables relating to religion, only Judaism shows a consistent effect, with electorates with relatively large Jewish populations more likely to vote Liberal.
+Electorates with more diverse speech are associated with higher support for the Liberal party from 2004 onwards, with this effect being significant in 2013 and 2016 row 7, Figure \@ref(fig:partresplot1)).
+Furthermore, none of the variables relating to religious beliefs aside from Christianity have a material effect in any election (this includes the Buddhist, Muslim, Jewish, areligious and Indigenous Australian populations).
+The association between Christian populations (`Christianity`) and the Liberal party steady increases over the years, becoming positive and significant in 2016.
 

### 2019-08-30 anzjs_elections.Rmd
removed 'a note on similar variables'

@@
 
-### A note on similar variables {-}
-(Maybe remove this?)
-
-Many of the Census variables represent similar information, which is why factors were created and some variables were removed.
-However, some variables remain which are closely related.
-For example, an electorate's income level (via `Incomes`) is likely to be related to electoral unemployment and labor force participation (via `Unemployment`).
-In 2001, the coefficient estimate for `Unemployment` is negative but not significant, whilst the `Incomes` variables is significant.
-If the `Incomes` variable is removed from the model in 2001, `Unemployment` absorbs the negative effect, becoming significant ($p = 0.0056$).
-
-```{r removeunemp, eval = FALSE}
-unemp01 <- my_fgls(full_formula,
-  my_data = model_df %>% filter(year == "2001") %>% dplyr::select(-Incomes),
-  sp_weights = sp_weights_01
-)
-
-summary(unemp01)$tTable["Unemployment",]
-```
-
 ## A closer look at the residuals

### 2019-08-30 anzjs_elections.Rmd
add incumbent effect and edit residuals by state

@@
 It is often hypothesized that states have systematic differences that cause their electorates to vote differently.
-Boxplots of residuals grouped by state (Figure \@ref(fig:resstate)) reveal that the data reflects this -- there appears to be a state-specific effect not captured by the models.
+Boxplots of residuals grouped by state (Figure \@ref(fig:resstate)) reveal that the data reflects this to a limited extent.
 Tasmania and the Australian Capital Territory appear to have a bias towards Labor, whereas the Northern Territory tends towards voting Liberal.
@@
   labs(x = "State", y = "Residuals")
+```
+
+### Residuals by party incumbency {-}
+
+The incumbent party appears to have a distinct advantage at the next election.
+The boxplots in Figure \@ref(fig:resincumbent) show that if either Labor or Liberal won the seat at the previous election, the electorate is likely to vote in their favour, over and above any socio-demographic effects.
+
+```{r resincumbent, fig.cap = "Boxplot of residuals by state with jittered points.
+States ordered by median residual.
+A state-specific bias not captured by the model is evident.", fig.asp=0.6, out.width="70%", fig.pos = 'h', out.extra = ''}
+# Incumbents
+incumbent <- function(fp) {
+  df <- fp %>% filter(HistoricElected == "Y", PartyAb == "LNP") %>% select(DivisionNm, HistoricElected)
+  return(df)
+}
+
+# 2004-2016
+
+incum_resids <- data.frame(
+  Residuals = c(glsmod16$actual_residuals, glsmod13$actual_residuals, glsmod10$actual_residuals, glsmod07$actual_residuals, glsmod04$actual_residuals), 
+  bind_rows(
+    glsmod16$my_data %>% select(DivisionNm, year) %>% left_join(incumbent(fp16), by = "DivisionNm"),
+    glsmod13$my_data %>% select(DivisionNm, year) %>% left_join(incumbent(fp13), by = "DivisionNm"),
+    glsmod10$my_data %>% select(DivisionNm, year) %>% left_join(incumbent(fp10), by = "DivisionNm"),
+    glsmod07$my_data %>% select(DivisionNm, year) %>% left_join(incumbent(fp07), by = "DivisionNm"),
+    glsmod04$my_data %>% select(DivisionNm, year) %>% left_join(
+      fp04 %>% filter(PartyAb == "LNP") %>% 
+        left_join(
+          fp01 %>% rename(HistoricElected = Elected) %>% select(UniqueID, HistoricElected, PartyAb), 
+          by = c("UniqueID", "PartyAb")
+        ) %>% incumbent() %>% unique(), by = "DivisionNm")) %>%
+    mutate(Incumbent = ifelse(is.na(HistoricElected), "N", "Y")))
+
+ggplot(aes(x = year, y = Residuals, fill = Incumbent), data = incum_resids) + 
+  geom_boxplot() +
+  geom_hline(aes(yintercept = 0), col = "grey50") +
+  scale_fill_manual(values = c('white', 'grey'), labels = c("No", "Yes")) +
+  theme_bw()
+
 ```

### 2019-08-30 anzjs_elections.Rmd
Added party incumbent, edited outlier electorates

@@
 States ordered by median residual.
-A state-specific bias not captured by the model is evident.", fig.asp=0.6, out.width="70%", fig.pos = 'h', out.extra = ''}
+A state-specific bias present only in the smaller states appears to have not been captured by the model.", fig.asp=0.6, out.width="70%", fig.pos = 'h', out.extra = ''}
 all_resids %>%
@@
 The incumbent party appears to have a distinct advantage at the next election.
-The boxplots in Figure \@ref(fig:resincumbent) show that if either Labor or Liberal won the seat at the previous election, the electorate is likely to vote in their favour, over and above any socio-demographic effects.
-
-```{r resincumbent, fig.cap = "Boxplot of residuals by state with jittered points.
-States ordered by median residual.
-A state-specific bias not captured by the model is evident.", fig.asp=0.6, out.width="70%", fig.pos = 'h', out.extra = ''}
+The boxplots in Figure \@ref(fig:resincumbent) show that if either of the Labor or Liberal parties won the seat at the previous election, the electorate is likely to vote in their favour, over and above any socio-demographic effects --- this effect has not been captured by the model.
+
+```{r resincumbent, fig.cap = "Boxplot of residuals for incumbent and non-incumbent parties each year.
+An incumbent advantage is evident and has not been captured by the model.", fig.asp=0.6, out.width="70%", fig.pos = 'h', out.extra = ''}
 # Incumbents
@@
   labs(x = "Cook's Distance", y = "") +
-  theme(axis.text.x = element_text(angle = 60, hjust = 1, size = 8), axis.text.y = element_text(size = 8))
+  theme(axis.text.x = element_text(angle = 60, hjust = 1, size = 8), axis.text.y = element_text(size = 8)) +
+  theme_bw()
 ```
@@
   View()
+
+cooksd_df %>%
+  filter(DivisionNm == "WENTWORTH") %>%
+  dplyr::select(DivisionNm, StateAb, year, cooksd, Residuals, LNP_Percent, everything()) %>%
+  View()
 ```
@@
 Based on the distribution of the Cook's distance values, a Cook's distance greater than $0.1$ is considered to be influential and a potential outlier.
-The electorate of Sydney (NSW) has a large Cook's distance from 2001 to 2013, due to its diverse population (language, birthplace and religion), high number of defacto relationships, high income, high household mobility and small amount of workers in extractive and transformative jobs.
+The electorate of Sydney (NSW) has a large Cook's distance from 2001 to 2010, due to its diverse population (language, birthplace and religion), high number of defacto relationships, high income, high household mobility and small amount of workers in extractive and transformative jobs.
 It has remained a strong supporter of the Labor party and the Liberal vote is severely overpredicted by the model, making it an outlier.
-Nearby in metropolitan NSW, the electorate of Wentworth is found to be an outlier in all but the 2007 election.
+Nearby in metropolitan NSW, the electorate of Wentworth is found to be an outlier in the 2013 and 2016 elections.
 Although historically Liberal, its two-party vote jumped by over 10 percentage points in 2010 without experiencing any notable changes in its socio-demographic makeup --- implying that this may be the direct effect of its Liberal member, Malcolm Turnbull, becoming the leader of the Liberal party.
-Liberal support in Wentworth is underpredicted by the model in each year, and more so with Turnbull as Liberal leader.
-
-Lingiari, an electorate taking up almost all of the Northern Territory, is an outlier in the 2001--2007 elections due to its large Indigenous population, young age profile and low rates of property ownership.
+In the elections since, the model underpredicts Wentworth's Liberal support.
+
+Lingiari, an electorate taking up almost all of the Northern Territory, is an outlier in the all but the 2013 election due to its large Indigenous population, young age profile and low rates of property ownership.
 Fowler (NSW) has a diverse population with a high proportion of migrants, many Buddhists and Muslims, and has strong Labor support, making it influential in 2001, 2004 and 2010.
-Other electorates with large Cook's distance are Barton (NSW) and Leichhardt (QLD) in 2016, and Canberra (ACT) in 2007.
+Other electorates with large Cook's distance are Canberra (ACT) and Durack (WA) in 2013, and Solomon (NT) in 2016.
 
@@
   filter(year == "2016") %>%
-  select(-c(year, DivisionNm)) %>%
+  dplyr::select(-c(year, DivisionNm)) %>%
   as.matrix()
@@
 plot(
-  glsmod07$cooksd,
+  glsmod16$cooksd,
   hat(
     model_df %>%
-      filter(year == "2007") %>%
-      select(-c(year, DivisionNm)) %>%
+      filter(year == "2016") %>%
+      dplyr::select(-c(year, DivisionNm)) %>%
       as.matrix()

### 2019-08-30 anzjs_elections.Rmd
updated spatial coefficient section - removed electorates voting independently. minor edit to robustness. shift around effects in Influential Socio-demographics. edited conclusion.

@@
 
-The spatial autoregressive coefficient $\rho$ is positive and significant in only the 2001 and 2016 elections (Figure \@ref(fig:rhovis)), meaning that in these elections, the political climate of an electorate appears to be affected by the attitudes of its neighbours.
-Conversely, in the other four elections, the spatial effect weakens to become insignificant.
-In these years, it appears that the spatial component does not explain anything not already explained by the electoral socio-demographics, meaning electorates effectively voted independently.
+The spatial autoregressive coefficient $\rho$ is positive and significant in the 2001, 2004 and 2016 elections (Figure \@ref(fig:rhovis)).
+In these three elections, there is evidence to suggest that the political climate of an electorate appears to be affected by the attitudes of its neighbours, affecting the electorate's two-party preference.
+Conversely, in the other three elections, the spatial effect weakens to become insignificant (although still positive).
 
@@
 
-### Households {-}
-
-In 2001, 2004 and 2016, higher proportions of people that have recently (in the past five years) moved house (`DiffAddress`) increased electoral support for the Liberal party (see row 5 in Figure \@ref(fig:partresplot1)).
-Having controlled for characteristics of house ownership and rental prices (`PropertyOwned` and `RentLoan` - both of which have no effect), this effect is somewhat surprising.
-
-Higher proportions of single parent households are associated with Labor support in all elections (albeit insignificant in 2004, see row 6 in Figure \@ref(fig:partresplot2)), whereas the electoral family and household sizes (via the `FamHouseSize` variable) do not appear to be associated with either party in any election.
-
-### Relationships {-}
-
-From 2007 onwards, both marriages (`Married`) and de facto relationships (`DeFacto`) are found to be strong predictors of the two-party preferred vote in opposition to the Liberal party.
-In 2001 and 2004 neither of these variables were significant (see rows 4 and 5 in Figure \@ref(fig:partresplot2)).
-
 ### Age {-}
@@
 
+### Households {-}
+
+In 2001, 2004 and 2016, higher proportions of people that have recently (in the past five years) moved house (`DiffAddress`) increased electoral support for the Liberal party (see row 5 in Figure \@ref(fig:partresplot1)).
+Having controlled for characteristics of house ownership and rental prices (`PropertyOwned` and `RentLoan` - both of which have no effect), this effect is somewhat surprising.
+
+Higher proportions of single parent households are associated with Labor support in all elections (albeit insignificant in 2004, see row 6 in Figure \@ref(fig:partresplot2)), whereas the electoral family and household sizes (via the `FamHouseSize` variable) do not appear to be associated with either party in any election.
+
+### Relationships {-}
+
+From 2007 onwards, both marriages (`Married`) and de facto relationships (`DeFacto`) are found to be strong predictors of the two-party preferred vote in opposition to the Liberal party.
+In 2001 and 2004 neither of these variables are significant (see rows 4 and 5 in Figure \@ref(fig:partresplot2)).
+
 ### Robustness
@@
 Two robustness checks are conducted to confirm model stability.
-First, a model for each election is fit using only the variables that were found to be significant in at least one of the six elections.
+First, a model for each election is fit using only the variables that are found to be significant in at least one of the six elections.
 The estimated coefficients of the variables in these models all fall within their respective $95\%$ confidence intervals from the full models, confirming that the significant effects are robust to variable selection.
@@
 It is found that for each of the top ten pairwise correlations, the estimated effect in the reduced model lies within the confidence interval from the full model.
-This second check indicates that variables with high pairwise correlation are not masking the effects of each other.
+This second check indicates that variables with the highest pairwise correlation are not masking the effects of each other.
 
@@
 Across the past six elections, most of the socio-demographics that drive the electoral two-party preferred vote are found to remain steady, whilst a few (typically weaker) effects vary over time.
-Industry and type of work are particularly influential, with energy-related and manufacturing/construction jobs, as well as administrative roles being strongly linked with the Liberal party in all elections.
+Industry and type of work are particularly influential.
+Energy-related and manufacturing/construction jobs, as well as administrative roles and jobs in education and social services are strongly linked with the Liberal party in all elections.
 Incomes have a similarly consistent effect, with higher income areas supporting Liberal.
-Higher levels of unemployment shift from weak association with Labor to a significant Liberal effect over the years, and higher education levels are associated with Labor from 2007 (although significant only in 2016).
-It is also found that electorates with higher household mobility support Liberal, birthplace diversity favours Labor and more de facto relationships align with Labor preference --- although marriages, family and household sizes have no material influence.
-Furthermore, the neighbourhood (spatial) effects are found to be positive in all elections, although significant only in 2001 and 2016, meaning that in the 2004--2013 elections, electorates effectively voted independently.
+Higher levels of unemployment shift from weak association with Labor to a significant Liberal effect over the years, and higher education levels are associated with Labor from 2007 (although marginally insignificant in 2010).
+Electorates with large populations 20 to 34 years are strongly associated with Labor, whilst the 35 to 54 year old bracket also increases Labor support, but to a lesser extent.
+It is also found that birthplace diversity slightly favours Labor, relationships (both marriages and de facto relationships) align with Labor preference from 2010 onwards, and the influence of Christian populations has trended towards Liberal support whilst other religions have negligible effects.
+Family and household sizes have minimal influence, although electorates with more single parent households are linked with Labor support.
+Furthermore, the neighbourhood (spatial) effects are found to be positive in all elections and significant in 2001, 2004 and 2016, meaning that neighbouring party preference appears to be an influential factor in those years.
 

### 2019-08-31 anzjs_elections.Rmd
Added a bit more detail to factor creation and variable selection. Split robustness checks into 2 paragraphs.

@@
 
-Only the first four principal components from the combined dataset are considered, as the scree plot corresponding to the combined dataset levels off after the fourth component.
+Only the first four principal components from the combined dataset are considered, as the scree plot levels off after the fourth component.
 Variables that have a large loading in a particular component are deemed to covary, with a loading with magnitude greater than 0.15 being considered large.
+Each principal component is considered seperately.
+If a subset of variables have a large loading in a given component, and there is an intuitive reason as to why they should be grouped together, then this subset will be combined to become a factor.
 Six factors are created using this criteria.
@@
 `FamHouseSize` (average household size, ratio of people to families, incidence of single person households, incidence of households containing a couple with kids, incidence of households containing a couple without kids); and
-`Education` (high school completions, university qualifications, jobs requiring higher levels of education, vocational course completions, jobs that do not require higher education levels).
+`Education` (high school completions, university qualifications, proportion of people working as professionals, proportion of jobs in finance,  proportion of workers who are laborers, proportion of workers who work as a tradesperson, diploma and certificate qualifications).
 For each of these groupings, a factor is created by taking a weighted sum of the variables.
@@
 
-There are $p=32$ variables in the resultant predictor set, with all of these used in the regression for each election.
+There are $p=33$ variables in the resultant set obtained from this procedure.
+One of the age brackets (`Pop_55_plus`) is removed from this set to avoid multicollinearity, because the other three age brackets are included.
+Hence, the final predictor set contains $p=32$ variables.
 
@@
 First, a model for each election is fit using only the variables that are found to be significant in at least one of the six elections.
-The estimated coefficients of the variables in these models all fall within their respective $95\%$ confidence intervals from the full models, confirming that the significant effects are robust to variable selection.
+Let this model be the reduced model, and the original model be the full model.
+The estimated coefficients of the variables in the reduced models all fall within their respective $95\%$ confidence intervals from the full models, confirming that the significant effects are robust to variable selection.
+
 The second check involves computing pairwise correlations amongst the explanatory variables, and seeing if the removal of one variable (from a pair) affects the other's estimated effect.

### 2019-08-31 response_for_ANZJS_editor.Rmd
Response to editors comments. Next to do associate editors comments.

---
title: "Response For The Editor"
author: "Jeremy Forbes"
date: "31/08/2019"
output: html_document
---

Dear XXXX,

Thank you for your feedback on our paper titled 'XXXX'. You raised many valid points in your response, and we have refined our work to address the queries raised.

Below we have detailed the changes made, or our response to, each point raised in your feedback. This is split into two sections - editor's comments and associate editor's comments.

If you have any further questions, please do not hestitate to reach out to the corresponding author, Jeremy Forbes (`jeremyforbes1995@gmail.com`) directly.

#### Editor's Comments

1. Address reviewer 1's question about why spatial imputation is required given that the census is reported at a low level of aggregation (lines 150-151) that could presumably be clumped to give a good approximation to electorate boundaries.

*This is a valid point, and we have revised the imputation method to adopt this suggestion. We now use Census data at the lowest level of aggregation available - the geographic region type of 'Statistical Area 1' - to obtain the approximation of electoral boundaries (see Section 2.2). Our original method was chosen because we believed that the imputation method would not materially affect results. We found that our results did not change much by adopting this new imputation method.*

2. I note that you do mention the ecological fallacy (lines 285-286) but reviewer 1's reaction suggests that you should expand your remarks here.

*We agree that more attention could have been given to the ecological fallacy in our original submission. In response, we have expanded our remarks and moved them to the introduction (Section 1). We are explicit in stating that "drivers of individual voter behaviour may vary from what is observed at the electorate level", which may not have been clear enough in the original submission.*

3. It appears that your approach to variable selection has confused everybody. My understanding of what you have done is that: You begin with 65 variables (are they all the available census data or a subset?). Then you standardise the variables. Then you remove some variables to do with Age Bands and Christianity (how many variables out of the 65 does this leave? Were these variables totally removed or lumped in some way? Christianity still appears in table 2 so they obviously weren't all removed)

*The variable selection method does have multiple steps and in retrospect, we agree that more clarity was needed here.*

*Firstly, we have refined the subset of variables considered for analysis in this paper to 50 Census variables. Census data is provided as a head count for categories for many different questions, and there are thousands of variables that could be constructed from the raw data. These 50 variables were chosen because they appear in previous studies of Australian elections (those mentioned in the Introduction). Previously we started with 65 variables, but many of these variables were removed in the variable selection process because they provided additional granularity that we did not want to pursue. These never should have been included to begin with - and we have removed them in our updated submission. For example, we had previously included variables relating to Christian denominations (Anglican, Catholic) in our 65 variables, but did not include them in the model.*

*The variable Christianity represents the proportion of people identifying with any Christian denomination, which we have kept in the model.*

4. ...Then you use the PCA to identify variables that covary. If variables appear together in a PC with loading higher than 0.15 they are combined together into a factor, e.g. a new variable Income replaces two previously existing variables. (In the description of this it is unclear how many variables are being combined in each factor, the ‘ands’ and commas make things tricky, .e.g is ‘renting and government housing’ one variable or two? Do the factors use the loading weights or just the signs?)

*We have made an effort to clarify this process in Section 3.2, which now details the variables being combined into each factor. For consistency, only commas are used to separate the variable descriptions. Only the signs are used, which is now made explicit.*

5. The steps above get you from 65 variables down to 30 (I can count 10 variables that  get accounted for via the creation of factors, but that doesn't get from 65 to 30 so presumably some other variables have been removed? Or is this difference all in the age bands and variants of Christianity? More clarity is required)

*This has been reconciled in Section 3.2. We begin with 50 variables. Six factors are created using a combined 23 variables. One of the age brackets (populated aged 55 and over) is removed, and the other three age brackets are kept (to avoid multicollinearity). This results in a final predictor set of 32 variables. Note that two variables no longer meet the criteria for being included in a factor, so they appear as separate variables - hence a final set of 32 variables.*

6. As you note in the paper, it seems likely that there are still issues with multi-collinearity. Figs 6 and 7 show the variables with significant effects, do the results change much if you fit a model with just these 13 variables instead of the 30?

*This was a very helpful suggestion to explore, and as such an additional section on robustness has been included in the revised paper (Section 4.4). When we fit a model using just the variables with significant effects, we find that the results do not change much. Each of the estimated coefficients from the reduced model (containing only the subset significant variables) lie within the 95% confidence interval from its corresponding full model (containing all 32 variables).*

*Furthermore, we have added another robustness check, where we consider the largest pairwise correlations amongst all variables (computed on a dataset of all elections combined). Taking each pair one by one, we remove one of the two variables, and find that the estimated coefficient for the remaining variable lies within the 95% confidence interval from the full model. This is true for all of the top 10 pairwise correlations.*

#### Associate Editor's Comments

1. Some conceptually simple imputation methods are used to align polling and census data to investigate associations between census variables and voting behaviour at electoral division level.

2. The main thing I found confusing about this paper was the variable selection procedure. It seems to be heading in the direction of four PCs, then it veers into six factors before ending with 30 variables. This all requires greater clarity and justification.

3. The authors opt for “visualization” of the fixed effects rather than the usual table of estimated parameters. Perhaps this is a good choice given that there are so many. But then spatial parameters are exhibited in a huge table. This seems inconsistent.

4. I found the writing unclear in a few places. What are the “time-varying restrictions” on a variable (line228)? And the description of a partial residual (line 276) doesn’t make sense to me.

5. Some of the interpretations of results seem a little naïve. For example, does Figure 4 really justify concluding that “electorates effectively voted independently” except for 2001 and 2016? (Is it a coincidence that the spatial autoregression is greatest in 2001 and 2016 where there is no need for imputation?)

### 2019-09-05 anzjs_elections.Rmd
added example in time-varying restriction

@@
 This approach is preferable to using a single longitudinal model because it avoids imposing any time-varying restrictions on the effect of a variable.
-To gain any advantage over modelling each election as a separate cross-section, a pooled cross-sectional model would have to force the effects at least some variables to be invariant over time.
+To gain any advantage over modelling each election as a separate cross-section, a pooled cross-sectional model would have to force the effect of at least one variable to be invariant over time (e.g.
+forcing the estimated coefficient of the variable `Unemployed` to be the same in 2013 and 2016 elections).
 This would not allow for conclusions to be drawn about how the effects of these variables change over time.

### 2019-09-05 response_for_ANZJS_editor.Rmd
added response to associate editor

@@
 
+*No response required.*
+
 2.
@@
 This all requires greater clarity and justification.
+
+*See items (2), (3) and (4) in section above.*
 
@@
 
+*Partial residual plots are presented to give the reader a visual feel for the direction, size and significance of an estimated effect, as well as the distribution of the data points (for that covariate).
+Whilst this is a valid point as there is duplication of information here, we think that the partial residual plots play an important role in illustrating the effect of a given variable across elections, and that the estimated model parameters should be included for completeness.*
+
 4. I found the writing unclear in a few places. What are the “time-varying restrictions” on a variable (line228)? And the description of a partial residual (line 276) doesn’t make sense to me.
+
+*We have expanded on descriptions of both in our revised version.
+Time-varying restrictions are when one forces the effect of a variable to be invariant over time, for example  - forcing the estimated coefficient of the variable `Unemployed` to be the same in 2013 and 2016 elections (section 3.2).*
+
+*Partial residuals, for a variable X1, are the residuals from the fitted model with the estimated effect of X1 added to it.
+Partial residual plots show the direction, size and significance of an estimated effect, as well as any deviations from linearity (section 4.3).*
 
@@
 (Is it a coincidence that the spatial autoregression is greatest in 2001 and 2016 where there is no need for imputation?)
+
+*In retrospect, we agree that this is perhaps a naïve conclusion.
+It is possible that unobserved variables are being captured in the spatial effect, and as such we have removed this interpretation from the revised paper.
+Additionally, on revision of the imputation method, we find that the spatial effect is significant in 2004 as well as 2001 and 2016.*

### 2019-09-11 anzjs_elections.tex
code to do multicollinearity check

@@
 The two-party preferred vote, a measure of support between these two parties, fluctuated between 47.3\% and 53.5\% (in favour of the Liberal party) over this period.
-This study explores how electoral socio-demographic characteristics relate to two-party preference, and whether their effects have changed over time.
+This study explores how electoral (aggregate) socio-demographic characteristics relate to two-party preference, and whether their effects have changed over time.
 
-Electoral socio-demographics are derived from the Australian Census, and vote counts are obtained from Australian federal elections.
+The Australian Electoral Commission (AEC) divides Australia into 150 regions, called electorates, with each corresponding to a single seat in the House of Representatives.
+If a party wins a majority of seats, they become the governing party.
+Data on the socio-demographics of these electorates are derived from the Australian Census, and vote counts are obtained from Australian federal elections.
 Joining these two data sources is problematic as there is an inherent asynchronicity in the two types of events.
-A Census is conducted by the Australian Bureau of Statistics (ABS) every five years, whereas federal elections, conducted by the Australian Electoral Commission (AEC), usually occur every three years or so.
+A Census is conducted by the Australian Bureau of Statistics (ABS) every five years, whereas federal elections, conducted by the AEC, usually occur every three years or so.
 The first problem addressed is that of constructing appropriate Census data for the 2004, 2007, 2010 and 2013 elections --- election years in which a Census does not occur.
@@
 Instead of analyzing a single election in isolation, this paper employs a consistent model framework across six elections so that temporal changes in the effects of socio-demographics can be observed.
-Each federal election is modelled with a cross-sectional dataset.
-The cross-sectional dataset for each election used here consists of the two-party preferred vote (as the response variable), and a set of common socio-demographic variables (as the explanatory variables) that characterize each electorate.
+Each federal election is modelled with a cross-sectional dataset, where each observation is one of the 150 electorates.This dataset consists of the two-party preferred vote (as the response variable) and a set of common socio-demographic variables (as the explanatory variables).
 To prepare these datasets, socio-demographic variables are first standardized, and then a principal component analysis is used to group variables into ``factors''.
 To account for the inherent spatial structure of the data, a spatial error model is then estimated for each election.
+In interpreting these models, is important to be mindful of the ecological fallacy.
+Insights are being drawn at the electorate level and cannot be inferred for another disaggregate level (this means that drivers of individual voter behaviour may vary from what is observed at the electorate level).
 
@@
 
-{\centering \includegraphics[width=0.9\linewidth]{paper_files/figure-latex/timeline-1}
+{\centering \includegraphics[width=0.9\linewidth]{anzjs_elections_files/figure-latex/timeline-1}
 
@@
 
-To account for spatial differences, the piecewise approximation method in \citet{Goodchild1993} is adopted.
-Consider a map of source zones \(s = 1,\dots,S\), for which socio-demographic information is available, and a set of target zones \(t = 1,\dots,T\) for which information is to be imputed.
+The finest level of disaggregation available for Census data is the region classification called Statistical Area 1 (SA1).
+In 2016, Australia was divded into over 55,000 SA1s.
+Consider each of these SA1 regions as a source zone, \(s = 1,\dots,S\), for which socio-demographic information is available.
+For simplicity, let each source zone be wholly summarised by its centroid.
+A set of target zones, \(t = 1,\dots,T\), are defined as regions for which information is to be imputed - these are the electoral boundaries for a particular election.
 This is described in the context of a single election, and a single neighbouring Census.
+Take the example of the Melbourne Ports electorate from the 2013 federal election, illustrated in Figure \ref{fig:melbports13}.
+The purple region in this figure represents the target zone and the source zones are the centroid locations from the 2016 Census SA1 areas.
 
-Let the map of electoral boundaries at the time of a Census define the source zones, and let the boundaries at the time of the election be the target zones.
-Denote the area of intersection between a source zone \(s\) and a target zone \(t\) as \(A_{s,t}\).
-Additionally, let the population of the source zone \(s\) be \(U_s\) and the population of intersection between source zone \(s\) and target zone \(t\) be \(P_{s,t}\).
-The estimated population of intersection is given by
-\[
-  \hat{P}_{s,t} = \frac{U_s*A_{s,t}}{\sum_{t=1}^T A_{s,t}},
-  \quad\text{for all $s=1,\dots,S$ and $t=1,\dots,T$}.
-\]
-Note that this estimator implicitly assumes that populations are uniformly distributed within each source zone.
+\begin{figure}[h]
 
-In order to calculate socio-demographic information for each of the target zones, a weighted average is taken using the estimated intersection populations as weights.
+{\centering \includegraphics[height=0.28\textheight]{anzjs_elections_files/figure-latex/melbports13-1}
+
+}
+
+\caption{The electoral boundaries for Melbourne Ports (shown in purple) and surrounding electorates, with centroids for Statistical Area 1 regions from the 2016 Census overlaid.
+The centroids falling within the purple region are attributed to Melbourne Ports.}\label{fig:melbports13}
+\end{figure}
+
+Furthermore, let \(I_{s,t}\) be an indicator variable, for which \(I_{s,t} = 1\) if the centroid of source zone \(s\) falls within target zone \(t\), and \(0\) otherwise.
+Additionally, let the population of the source zone \(s\) be \(U_s\) and the population of target zone \(t\) be \(P_{t}\).
+
+In order to calculate socio-demographic information for each of the target zones, a weighted average of source zones is taken using their populations as weights.
 Denote a given Census variable for the target zone by \(C_t\), and the same Census variable for the source zone as \(D_s\).
@@
 \[
-  \hat{C}_t = \frac{\sum_{s=1}^{S}{D_s*\hat{P}_{s,t}}}%
-  {\sum_{s=1}^{S}{\hat{P}_{s,t}}},
+  \hat{C}_t = \frac{\sum_{s=1}^{S}{I_{s,t}*D_s*U_s}}%
+  {\sum_{s=1}^{S}{I_{s,t}*U_s}},
   \quad\text{for each $t=1,\dots,T$}.
 \]
+
 This concludes the spatial imputation of the socio-demographic characteristics for one target zone (a single electoral boundary), at the time of only one of the neighbouring Censuses.
@@
 
-\hypertarget{an-illustration-of-the-spatio-temporal-imputation}{%
-\subsubsection*{An illustration of the spatio-temporal imputation}\label{an-illustration-of-the-spatio-temporal-imputation}}
-\addcontentsline{toc}{subsubsection}{An illustration of the spatio-temporal imputation}
-
-Census data is publicly available at different levels of aggregation, ranging from SA1 (over 50,000 zones) to electoral divisions (150 zones).
-For this study, electoral divisions are used as source zones, and the imputation method is applied to produce socio-demographic variables for each of the 2004, 2007, 2010 and 2013 elections.
-As mentioned earlier, there is no need to impute socio-demographic variables for the 2001 and 2016 elections.
-To illustrate the method, consider the imputation of socio-demographic variables for the electorate of Hume in New South Wales (NSW) at the time of the 2013 federal election.
-The boundaries shown in Figure \ref{fig:hume13} define all target zones in NSW for 2013, with the target zone of interest (Hume) shaded purple.
-
-\begin{figure}[h]
-
-{\centering \includegraphics[height=0.28\textheight]{paper_files/figure-latex/hume13-1}
-
-}
-
-\caption{Some of the electoral boundaries in NSW for 2013, with the electoral boundary for Hume shown in purple.}\label{fig:hume13}
-\end{figure}
-\begin{figure}[h]
-
-{\centering \includegraphics[height=0.28\textheight]{paper_files/figure-latex/hume16-1}
-
-}
-
-\caption{Census division boundaries in NSW for 2016, with the 2013 electoral boundary for Hume, shown in purple. The purple region is not contained within a single Census division.}\label{fig:hume16}
-\end{figure}
-
-The corresponding source zones from the 2016 Census are shown in Figure \ref{fig:hume16}.
-As can be seen, the Hume boundary from the 2013 election (shaded purple) does not perfectly match any of the source zones.
-
-There are many source zones from the 2016 Census that intersect with this purple region, including the divisions of Riverina, Eden-Monaro and Hume, along with smaller intersecting areas with Fenner, Calare, Gilmore and Whitlam.
-The proportion of each source zone that overlaps with the purple region is calculated, and used to obtain the intersecting populations \(\hat{P}_{s,t}\).
-
-\begin{table}[t]
-
-\caption{\label{tab:hume1}Population from each intersecting Census source zone (2016) that is allocated to the target zone (purple region - Hume electoral boundary in 2013), and the corresponding proportion of Australian citizens in each of these source zones.}
-\centering
-\resizebox{\linewidth}{!}{
-\fontsize{12}{14}\selectfont
-\begin{tabular}{lrrrr}
-\toprule
-Source zone (2016) & Proportion & Source zone population & Population allocated to purple region: $\hat{P}_{s,t}$ & AusCitizen (\%): $D_s$\\
-\midrule
-Hume & 0.9654 & 150643 & 145427 & 90.0\\
-Riverina & 0.2511 & 155793 & 39117 & 89.1\\
-Eden-Monaro & 0.1109 & 147532 & 16358 & 88.0\\
-Canberra & 0.0028 & 196037 & 548 & 85.5\\
-Fenner & 0.0023 & 202955 & 474 & 83.6\\
-\addlinespace
-Whitlam & 0.0006 & 152280 & 92 & 89.5\\
-Gilmore & 0.0006 & 150436 & 86 & 89.0\\
-Calare & 0.0001 & 161298 & 21 & 87.6\\
-\bottomrule
-\end{tabular}}
-\end{table}
-
-Now consider the socio-demographic variable \emph{AusCitizen}, the proportion of people in the region who are Australian citizens.
-A weighted average of \emph{AusCitizen}, with the allocated population from each source zone as weights, yields \(\hat{C}_{\text{Hume},2016} = 89.65\%\).
-Repeating this process using the 2011 Census yields \(\hat{C}_{\text{Hume},2011} = 91.00\%\).
-Finally, linear interpolation between 2011 and 2016 yields the 2013 estimate:
+Continuing with the example of Melbourne Ports in the 2013 election, the estimate for a given Census variable in 2016, \(\hat{C}_{MelbPorts, 2016}\) would be obtained by computing the weighted average of this variable amongst the SA1s within the purple region shown in Figure \ref{fig:melbports13}.
+This would be repeated with the 2011 Census SA1s to obtain \(\hat{C}_{MelbPorts, 2011}\), from which the final estimate is given by
 \[
-  \hat{C}_{\text{Hume},2013}
-    = \frac{3}{5} \hat{C}_{\text{Hume},2011} + \frac{2}{5} \hat{C}_{\text{Hume},2016}
-    % = \frac{3}{5} \cdot 91.00 \% + \frac{2}{5} \cdot 89.65 \% \\
-    =  90.46\%.
+  \hat{C}_{\text{MelbPorts},2013}
+    = \frac{3}{5} \hat{C}_{\text{MelbPorts},2011} + \frac{2}{5} \hat{C}_{\text{MelbPorts},2016}.
 \]
-This is done for each of the 65 socio-demographic variables, and is repeated for each of the 149 remaining target zones corresponding with 2013 electorates.
+This is done for each of the socio-demographic variables, and is repeated for each of the 149 remaining target zones corresponding with 2013 electorates.
 
@@
 First, variable scales change over the years, making it important to standardize variables.
-Second, many variables represent similar information and where appropriate, should be combined in some way.
-Principal component analysis is used to identify variables covary, from which intuitive groupings are selected to be combined into a single variable.
-This also reduces the dimension of the data.
+Second, many variables represent similar information and where appropriate, should be grouped together.
+To determine which variables should be grouped, principal component analysis (PCA) is used.
+The intuition here is that PCA will identify which variables covary, from which intuitive groupings of variables can be chosen to combine into a single variable.
 After these steps, a model specification is chosen.
@@
 
-There are only \(N = 150\) observations (electorates) in each election and \(p = 65\) socio-demographic variables in each cross-section, with many variables represent similar information about an electorate.
-Any model that uses all variables would face serious problems with multi-collinearity and over-fitting, which would likely lead to erroneous conclusions regarding variable significance.
-To address this, groups of variables that represent similar information are combined into ``factors''\footnote{A preliminary step involved removing all age bands, because age is represented by median age, and to remove variables relating to particular denominations of Christianity.}.
+There are only \(N = 150\) observations (electorates) in each election and \(p = 50\) socio-demographic variables in each cross-section, with many variables represent similar information about an electorate.
+Any model that uses all variables would face problems with multi-collinearity and over-fitting, which would likely lead to erroneous conclusions regarding variable significance.
+To address this, variables that represent similar information are combined into a single variable, which will be referred to as a ``factor''.
 
-A factor is created from a group of variables if there is an intuitive reason as to why they should represent similar information and if there is evidence to suggest that they covary.
+A factor is created from a group of variables if there is an intuitive reason as to why these variables should represent similar information and if there is evidence to suggest that they covary.
 For example, a potential group would be variables relating to electoral incomes --- median family, household and personal incomes.
-To determine which variables covary, principal component analysis is used on a combined dataset of socio-demographic variables from all six elections.
-It is appropriate to compute principal components in this way because when computed separately for each election, scree plots level off after four components and the loadings of the first four components are similar across the elections.
+To determine which variables covary, principal component analysis is used on a combined dataset of socio-demographic variables from all six elections\footnote{It is appropriate to compute principal components on a combined dataset of all six elections because when computed separately for each election, scree plots level off after four components and the loadings of the first four components are similar across the elections.}.
+The only variables exempted from the principal component analysis are the four variables representing age brackets (the proportion of the population aged 0-19 years old, 20-34 years old, 45-54 years old and 55 years plus), which are to be included in the model as separate variables by construction.
 
-Only the first four principal components from the combined dataset are considered, as the scree plot corresponding to the combined dataset levels off after the fourth component.
+Only the first four principal components from the combined dataset are considered, as the scree plot levels off after the fourth component.
 Variables that have a large loading in a particular component are deemed to covary, with a loading with magnitude greater than 0.15 being considered large.
+Each principal component is considered seperately.
+If a subset of variables have a large loading in a given component, and there is an intuitive reason as to why they should be grouped together, then this subset will be combined to become a factor.
 Six factors are created using this criteria.
 These are:
-\texttt{Incomes} (median personal, household and family incomes);
-\texttt{Unemployment} (unemployment and labor force participation rates);
-\texttt{ProportyOwned} (rates of housing ownership, mortgages, renting and government housing);
-\texttt{RentLoanPrice} (median rental and loan repayments);
-\texttt{FamHouseSize} (average household size, ratio of people to families and household makeup (single person, couple with kids and couple without kids); and
-\texttt{Education} (high school and university qualifications, jobs requiring higher levels of education as well as vocational course completions and jobs that do not require higher education levels, such as laborer or tradesperson).
-For each of these groups, variables with positive loadings are added and those with negative loadings are subtracted to create a factor.
-After computing these sums, each factor is standardized to have mean zero and variance one, within each election.
+\texttt{Incomes} (median personal income, median household income, median family income);
+\texttt{Unemployment} (unemployment rate, labor force participation rate);
+\texttt{PropertyOwned} (proportion of dwellings that are owned, proportion of dwellings that are mortgages, proportion of dwellings that are rented, proportion of dwellings that classified as government housing);
+\texttt{RentLoanPrice} (median rental payment amount, median loan repayment amount);
+\texttt{FamHouseSize} (average household size, ratio of people to families, incidence of single person households, incidence of households containing a couple with kids, incidence of households containing a couple without kids); and
+\texttt{Education} (high school completions, university qualifications, proportion of people working as professionals, proportion of jobs in finance, proportion of workers who are laborers, proportion of workers who work as a tradesperson, diploma and certificate qualifications).
+For each of these groupings, a factor is created by taking a weighted sum of the variables.
+The weightings are allocated on the basis of whether the variable had a positive or negative loading in the principal component from which the grouping was identified.
+Variables with a positive loading are allocated a weight of \(+1\) and those with negative loadings are allocated a weight of \(-1\).
+After computing these weighted sums, the factor is standardized to have mean zero and variance one, within each election.
 
-There are \(p=30\) variables in the resultant predictor set, with all of these used in the regression for each election.
+There are \(p=33\) variables in the resultant set obtained from this procedure.
+One of the age brackets (\texttt{Pop\_55\_plus}) is removed from this set to avoid multicollinearity, because the other three age brackets are included.
+Hence, the final predictor set contains \(p=32\) variables.
 
@@
 This allows the socio-demographic effects to be estimated separately for each election year, facilitating analysis of temporal changes in variable effects.
-This approach is preferable to using a single longitudinal model because it avoids any concerns of undue bias stemming from incorrectly imposed time-varying restrictions on any variable.
-Without such restrictions, a pooled cross-sectional model does not yield any distinct advantage over separate cross-sections.
-The panel approach is avoided because of how frequently electoral boundaries change, noting that electorates with the same name across elections are not guaranteed to represent the same geographical region.
-Therefore any fixed or random effects models would be difficult to estimate without implementing consistent boundaries, which would require further imputation.
+This approach is preferable to using a single longitudinal model because it avoids imposing any time-varying restrictions on the effect of a variable.
+To gain any advantage over modelling each election as a separate cross-section, a pooled cross-sectional model would have to force the effect of at least one variable to be invariant over time (e.g.~forcing the estimated coefficient of the variable \texttt{Unemployed} to be the same in 2013 and 2016 elections).
+This would not allow for conclusions to be drawn about how the effects of these variables change over time.
+Furthermore, the panel approach is avoided because of how frequently electoral boundaries change, noting that electorates with the same name across elections are not guaranteed to represent the same geographical region.
+Therefore any fixed or random effects models would be difficult to estimate without implementing consistent boundaries, which would require further imputation (of voting information).
 
@@
 
-\begin{table}[!htbp]\vspace*{-0.4cm}\centering    \caption{Estimated spatial error model parameters (standard errors) for each of the six election years.\vspace*{-0.4cm}}    \label{tab:printmodels}  \scriptsize\tabcolsep=0.35cm \begin{tabular}{@{\extracolsep{1pt}}lrrrrrr} \\[-1.8ex]\hline  \hline \\[-1.8ex]  \\[-1.8ex] & \multicolumn{6}{c}{\vspace{-0.4cm}} \\   & 2001 & 2004 & 2007 & 2010 & 2013 & 2016 \\  \hline \\[-1.8ex]  $\rho$ & 0.46\rlap{$^{***}$} & 0.29\rlap{$^{*}$} & 0.24 & 0.19 & 0.27\rlap{$^{*}$} & 0.50\rlap{$^{***}$}\\  & (0.15\rlap{)} & (0.17\rlap{)} & (0.17\rlap{)} & (0.16\rlap{)} & (0.16\rlap{)} & (0.17\rlap{)}\\  AusCitizen & $-$3.13 & $-$2.64 & $-$2.53 & $-$0.08 & $-$3.40 & $-$1.80 \\    & (2.26\rlap{)} & (2.43\rlap{)} & (2.34\rlap{)} & (2.79\rlap{)} & (2.76\rlap{)} & (2.71\rlap{)} \\    BornAsia & 2.22 & $-$0.95 & $-$1.60 & $-$6.83\rlap{$^{**}$} & $-$3.03 & $-$0.55 \\    & (2.18\rlap{)} & (2.44\rlap{)} & (2.19\rlap{)} & (2.73\rlap{)} & (2.71\rlap{)} & (2.17\rlap{)} \\    Born\_MidEast & $-$1.15 & $-$1.59 & $-$2.01\rlap{$^{*}$} & $-$2.03 & $-$0.92 & $-$1.44 \\    & (1.07\rlap{)} & (1.20\rlap{)} & (1.11\rlap{)} & (1.27\rlap{)} & (1.24\rlap{)} & (1.13\rlap{)} \\    BornSEEuro & $-$3.21\rlap{$^{**}$} & $-$4.24\rlap{$^{***}$} & $-$3.61\rlap{$^{***}$} & $-$4.14\rlap{$^{***}$} & $-$3.69\rlap{$^{***}$} & $-$2.72\rlap{$^{***}$} \\    & (1.42\rlap{)} & (1.46\rlap{)} & (1.02\rlap{)} & (1.19\rlap{)} & (1.07\rlap{)} & (0.97\rlap{)} \\    Born\_UK & 0.25 & $-$0.07 & 0.34 & 0.56 & $-$0.09 & $-$1.32 \\    & (1.00\rlap{)} & (0.98\rlap{)} & (0.90\rlap{)} & (1.07\rlap{)} & (1.04\rlap{)} & (1.04\rlap{)} \\    BornElsewhere & $-$5.04 & $-$4.91 & $-$4.13 & 2.35 & $-$5.23 & $-$4.14 \\    & (3.30\rlap{)} & (3.68\rlap{)} & (3.38\rlap{)} & (4.23\rlap{)} & (4.15\rlap{)} & (3.97\rlap{)} \\    Buddhism & $-$0.49 & $-$0.17 & $-$1.37 & $-$0.83 & $-$0.12 & $-$1.60 \\    & (1.39\rlap{)} & (1.61\rlap{)} & (1.61\rlap{)} & (1.80\rlap{)} & (1.68\rlap{)} & (1.56\rlap{)} \\    Christianity & $-$2.48 & $-$1.23 & 0.38 & 0.50 & 2.41 & 1.68 \\    & (1.73\rlap{)} & (1.85\rlap{)} & (1.83\rlap{)} & (1.99\rlap{)} & (1.85\rlap{)} & (1.78\rlap{)} \\    CurrentlyStudying & $-$2.19\rlap{$^{**}$} & $-$0.13 & 2.06\rlap{$^{*}$} & 2.12\rlap{$^{*}$} & 1.15 & $-$0.16 \\    & (0.99\rlap{)} & (1.13\rlap{)} & (1.17\rlap{)} & (1.25\rlap{)} & (1.26\rlap{)} & (1.18\rlap{)} \\    DeFacto & $-$6.44\rlap{$^{***}$} & $-$5.37\rlap{$^{**}$} & $-$6.43\rlap{$^{***}$} & $-$8.07\rlap{$^{***}$} & $-$6.56\rlap{$^{**}$} & $-$8.53\rlap{$^{***}$} \\    & (1.87\rlap{)} & (2.48\rlap{)} & (2.31\rlap{)} & (3.06\rlap{)} & (3.11\rlap{)} & (2.83\rlap{)} \\    DiffAddress & 3.88\rlap{$^{***}$} & 5.06\rlap{$^{***}$} & 4.22\rlap{$^{***}$} & 5.57\rlap{$^{***}$} & 3.53\rlap{$^{*}$} & 5.67\rlap{$^{***}$} \\    & (0.94\rlap{)} & (1.12\rlap{)} & (0.99\rlap{)} & (1.76\rlap{)} & (1.91\rlap{)} & (1.60\rlap{)} \\    Distributive & 1.27 & 2.01\rlap{$^{*}$} & 1.36 & 1.57 & 2.10\rlap{$^{*}$} & 1.20 \\    & (1.12\rlap{)} & (1.21\rlap{)} & (1.13\rlap{)} & (1.34\rlap{)} & (1.27\rlap{)} & (1.21\rlap{)} \\    Education & 1.08 & 0.52 & $-$5.52\rlap{$^{*}$} & $-$4.08 & $-$4.44 & $-$7.07\rlap{$^{**}$} \\    & (2.38\rlap{)} & (3.12\rlap{)} & (3.27\rlap{)} & (3.95\rlap{)} & (3.78\rlap{)} & (3.55\rlap{)} \\    Extractive & 4.83\rlap{$^{***}$} & 5.45\rlap{$^{***}$} & 5.37\rlap{$^{***}$} & 7.31\rlap{$^{***}$} & 6.71\rlap{$^{***}$} & 7.43\rlap{$^{***}$} \\    & (1.48\rlap{)} & (1.42\rlap{)} & (1.36\rlap{)} & (1.56\rlap{)} & (1.47\rlap{)} & (1.39\rlap{)} \\    FamHouseSize & $-$0.16 & 0.87 & $-$2.40 & $-$2.53 & $-$3.26 & $-$2.91 \\    & (2.19\rlap{)} & (2.72\rlap{)} & (2.69\rlap{)} & (3.25\rlap{)} & (3.28\rlap{)} & (2.90\rlap{)} \\    Incomes & 4.36\rlap{$^{**}$} & 5.03\rlap{$^{*}$} & 9.45\rlap{$^{***}$} & 7.09\rlap{$^{**}$} & 7.97\rlap{$^{***}$} & 12.20\rlap{$^{***}$} \\    & (1.77\rlap{)} & (2.66\rlap{)} & (2.75\rlap{)} & (3.25\rlap{)} & (2.92\rlap{)} & (2.75\rlap{)} \\    Indigenous & 2.91\rlap{$^{*}$} & 1.97 & 2.48 & 2.84 & 0.67 & $-$0.05 \\    & (1.68\rlap{)} & (1.95\rlap{)} & (1.75\rlap{)} & (2.16\rlap{)} & (2.14\rlap{)} & (2.00\rlap{)} \\    Islam & $-$0.92 & $-$0.97 & $-$0.54 & $-$2.50 & $-$0.82 & $-$0.95 \\    & (1.22\rlap{)} & (1.36\rlap{)} & (1.27\rlap{)} & (1.52\rlap{)} & (1.42\rlap{)} & (1.34\rlap{)} \\    Judaism & 1.88\rlap{$^{*}$} & 1.78 & 2.66\rlap{$^{***}$} & 1.97\rlap{$^{*}$} & 2.74\rlap{$^{**}$} & 1.65\rlap{$^{*}$} \\    & (1.05\rlap{)} & (1.13\rlap{)} & (1.01\rlap{)} & (1.15\rlap{)} & (1.10\rlap{)} & (1.00\rlap{)} \\    ManagerAdmin & 2.06\rlap{$^{***}$} & 3.32\rlap{$^{***}$} & 6.00\rlap{$^{***}$} & 5.47\rlap{$^{***}$} & 5.04\rlap{$^{***}$} & 5.78\rlap{$^{***}$} \\    & (0.71\rlap{)} & (0.93\rlap{)} & (0.90\rlap{)} & (1.08\rlap{)} & (1.03\rlap{)} & (1.06\rlap{)} \\    Married & 0.44 & 0.11 & $-$1.22 & $-$0.22 & 0.91 & $-$2.34 \\    & (2.31\rlap{)} & (2.96\rlap{)} & (2.83\rlap{)} & (3.15\rlap{)} & (3.03\rlap{)} & (2.81\rlap{)} \\    MedianAge & 2.32\rlap{$^{*}$} & 4.96\rlap{$^{***}$} & 3.66\rlap{$^{**}$} & 4.00\rlap{$^{*}$} & 2.30 & 2.87 \\    & (1.32\rlap{)} & (1.65\rlap{)} & (1.81\rlap{)} & (2.26\rlap{)} & (2.08\rlap{)} & (1.79\rlap{)} \\    NoReligion & $-$1.57 & $-$0.92 & 0.56 & $-$0.30 & 1.02 & 1.31 \\    & (1.59\rlap{)} & (1.71\rlap{)} & (1.73\rlap{)} & (1.92\rlap{)} & (1.94\rlap{)} & (2.04\rlap{)} \\    OneParentHouse & $-$1.73 & $-$0.45 & $-$0.75 & $-$1.46 & $-$0.77 & $-$0.74 \\    & (1.36\rlap{)} & (1.59\rlap{)} & (1.49\rlap{)} & (1.69\rlap{)} & (1.57\rlap{)} & (1.47\rlap{)} \\    OtherLanguage & $-$0.44 & 5.92 & 9.98\rlap{$^{**}$} & 11.24\rlap{$^{**}$} & 9.00\rlap{$^{*}$} & 9.84\rlap{$^{**}$} \\    & (3.22\rlap{)} & (4.16\rlap{)} & (3.91\rlap{)} & (4.76\rlap{)} & (4.66\rlap{)} & (4.44\rlap{)} \\    PropertyOwned & $-$0.46 & $-$0.53 & 0.67 & $-$0.94 & $-$0.48 & 1.41 \\    & (1.37\rlap{)} & (1.50\rlap{)} & (1.43\rlap{)} & (1.76\rlap{)} & (1.67\rlap{)} & (1.50\rlap{)} \\    RentLoanPrice & $-$1.57 & $-$3.32\rlap{$^{*}$} & $-$4.01\rlap{$^{**}$} & $-$0.97 & $-$0.70 & $-$0.89 \\    & (1.49\rlap{)} & (1.76\rlap{)} & (1.67\rlap{)} & (2.07\rlap{)} & (2.07\rlap{)} & (2.16\rlap{)} \\    SocialServ & 2.51\rlap{$^{*}$} & 1.65 & 2.47\rlap{$^{*}$} & 2.53\rlap{$^{*}$} & 2.35\rlap{$^{*}$} & 4.45\rlap{$^{***}$} \\    & (1.33\rlap{)} & (1.41\rlap{)} & (1.29\rlap{)} & (1.47\rlap{)} & (1.32\rlap{)} & (1.19\rlap{)} \\    Transformative & 3.24\rlap{$^{**}$} & 4.73\rlap{$^{***}$} & 4.84\rlap{$^{***}$} & 4.46\rlap{$^{**}$} & 3.56\rlap{$^{**}$} & 4.58\rlap{$^{***}$} \\    & (1.55\rlap{)} & (1.78\rlap{)} & (1.74\rlap{)} & (1.98\rlap{)} & (1.78\rlap{)} & (1.53\rlap{)} \\    Unemployment & $-$2.45\rlap{$^{*}$} & $-$3.07\rlap{$^{*}$} & 0.29 & 0.08 & 1.67 & 2.79\rlap{$^{**}$} \\    & (1.40\rlap{)} & (1.63\rlap{)} & (1.51\rlap{)} & (1.76\rlap{)} & (1.51\rlap{)} & (1.37\rlap{)} \\    Constant & 50.81\rlap{$^{***}$} & 52.60\rlap{$^{***}$} & 47.31\rlap{$^{***}$} & 49.93\rlap{$^{***}$} & 53.51\rlap{$^{***}$} & 50.49\rlap{$^{***}$} \\    & (0.71\rlap{)} & (0.58\rlap{)} & (0.52\rlap{)} & (0.57\rlap{)} & (0.58\rlap{)} & (0.80\rlap{)} \\   \hline \\[-1.8ex]  Observations & 150 & 150 & 150 & 150 & 150 & 150 \\  Residual Standard Error (GLS) & 4.69 & 5.04 & 4.79 & 5.63 & 5.18 & 4.88\\ \hline  \hline \\[-1.8ex]  & \multicolumn{6}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ \end{tabular}  \end{table}
+\begin{table}[!htbp]\vspace*{-0.4cm}\centering    \caption{Estimated spatial error model parameters (standard errors) for each of the six election years.\vspace*{-0.4cm}}    \label{tab:printmodels}  \scriptsize\tabcolsep=0.35cm \begin{tabular}{@{\extracolsep{1pt}}lrrrrrr} \\[-1.8ex]\hline  \hline \\[-1.8ex]  \\[-1.8ex] & \multicolumn{6}{c}{\vspace{-0.4cm}} \\   & 2001 & 2004 & 2007 & 2010 & 2013 & 2016 \\  \hline \\[-1.8ex]  $\rho$ & 0.53\rlap{$^{***}$} & 0.33\rlap{$^{**}$} & 0.21 & 0.17 & 0.27 & 0.39\rlap{$^{**}$}\\  & (0.15\rlap{)} & (0.16\rlap{)} & (0.18\rlap{)} & (0.17\rlap{)} & (0.17\rlap{)} & (0.17\rlap{)}\\  AusCitizen & $-$3.94\rlap{$^{*}$} & $-$1.39 & $-$2.18 & $-$1.28 & $-$3.89 & $-$2.66 \\    & (2.27\rlap{)} & (2.44\rlap{)} & (2.21\rlap{)} & (2.69\rlap{)} & (2.51\rlap{)} & (2.61\rlap{)} \\    Pop\_00\_19 & 0.49 & 2.66 & 9.39\rlap{$^{***}$} & 5.25 & 3.31 & 0.88 \\    & (2.54\rlap{)} & (3.91\rlap{)} & (3.63\rlap{)} & (3.64\rlap{)} & (2.91\rlap{)} & (2.62\rlap{)} \\    Pop\_20\_34 & $-$8.04\rlap{$^{***}$} & $-$7.72\rlap{$^{***}$} & $-$8.34\rlap{$^{***}$} & $-$11.68\rlap{$^{***}$} & $-$9.29\rlap{$^{***}$} & $-$9.21\rlap{$^{***}$} \\    & (1.80\rlap{)} & (2.21\rlap{)} & (2.18\rlap{)} & (2.90\rlap{)} & (2.62\rlap{)} & (2.37\rlap{)} \\    Pop\_35\_54 & $-$2.64\rlap{$^{***}$} & $-$2.78\rlap{$^{***}$} & $-$3.62\rlap{$^{***}$} & $-$3.13\rlap{$^{***}$} & $-$2.76\rlap{$^{**}$} & $-$2.13\rlap{$^{**}$} \\    & (0.84\rlap{)} & (0.89\rlap{)} & (0.83\rlap{)} & (1.10\rlap{)} & (1.11\rlap{)} & (1.06\rlap{)} \\    BornAsia & 3.58\rlap{$^{*}$} & $-$1.09 & 0.66 & $-$1.78 & $-$1.08 & $-$0.14 \\    & (2.09\rlap{)} & (2.52\rlap{)} & (1.99\rlap{)} & (2.74\rlap{)} & (2.54\rlap{)} & (2.17\rlap{)} \\    BornMidEast & $-$1.02 & $-$1.75 & $-$0.98 & $-$1.00 & $-$1.66 & $-$1.31 \\    & (1.00\rlap{)} & (1.17\rlap{)} & (1.09\rlap{)} & (1.33\rlap{)} & (1.23\rlap{)} & (1.11\rlap{)} \\    BornSEEuro & $-$1.63 & $-$3.17\rlap{$^{*}$} & $-$1.07 & $-$2.04 & $-$2.89\rlap{$^{***}$} & $-$2.53\rlap{$^{***}$} \\    & (1.37\rlap{)} & (1.68\rlap{)} & (1.06\rlap{)} & (1.29\rlap{)} & (1.11\rlap{)} & (0.97\rlap{)} \\    BornUK & 0.29 & 0.31 & 0.32 & 0.28 & $-$0.15 & $-$0.61 \\    & (1.02\rlap{)} & (1.04\rlap{)} & (0.87\rlap{)} & (1.06\rlap{)} & (0.99\rlap{)} & (0.99\rlap{)} \\    BornElsewhere & $-$4.13 & $-$1.51 & $-$1.03 & 2.45 & $-$4.21 & $-$2.17 \\    & (3.14\rlap{)} & (3.62\rlap{)} & (3.18\rlap{)} & (4.13\rlap{)} & (3.90\rlap{)} & (3.76\rlap{)} \\    Buddhism & $-$0.07 & 0.80 & 0.58 & $-$0.14 & $-$0.43 & $-$1.16 \\    & (1.31\rlap{)} & (1.54\rlap{)} & (1.39\rlap{)} & (1.66\rlap{)} & (1.60\rlap{)} & (1.58\rlap{)} \\    Christianity & $-$1.70 & $-$1.01 & $-$0.45 & 0.13 & 2.03 & 3.76\rlap{$^{**}$} \\    & (1.62\rlap{)} & (1.75\rlap{)} & (1.60\rlap{)} & (1.85\rlap{)} & (1.68\rlap{)} & (1.83\rlap{)} \\    CurrentlyStudying & $-$2.20\rlap{$^{*}$} & $-$0.01 & $-$0.14 & 1.35 & 0.32 & 0.22 \\    & (1.22\rlap{)} & (1.50\rlap{)} & (1.39\rlap{)} & (1.41\rlap{)} & (1.35\rlap{)} & (1.56\rlap{)} \\    DeFacto & $-$3.24 & $-$2.25 & $-$4.67\rlap{$^{**}$} & $-$7.75\rlap{$^{**}$} & $-$7.82\rlap{$^{**}$} & $-$10.39\rlap{$^{***}$} \\    & (2.07\rlap{)} & (2.62\rlap{)} & (2.27\rlap{)} & (3.09\rlap{)} & (3.08\rlap{)} & (3.15\rlap{)} \\    DiffAddress & 3.06\rlap{$^{***}$} & 2.75\rlap{$^{**}$} & 0.73 & 2.55 & 2.27 & 5.20\rlap{$^{***}$} \\    & (0.94\rlap{)} & (1.20\rlap{)} & (1.24\rlap{)} & (1.79\rlap{)} & (1.67\rlap{)} & (1.51\rlap{)} \\    Distributive & 1.60 & 1.89\rlap{$^{*}$} & 0.50 & 0.62 & 1.59 & 1.31 \\    & (1.06\rlap{)} & (1.14\rlap{)} & (0.99\rlap{)} & (1.27\rlap{)} & (1.20\rlap{)} & (1.18\rlap{)} \\    Education & $-$0.37 & $-$0.26 & $-$6.72\rlap{$^{**}$} & $-$7.31\rlap{$^{*}$} & $-$7.31\rlap{$^{**}$} & $-$8.55\rlap{$^{**}$} \\    & (2.35\rlap{)} & (3.34\rlap{)} & (3.00\rlap{)} & (3.90\rlap{)} & (3.63\rlap{)} & (3.37\rlap{)} \\    Extractive & 3.74\rlap{$^{***}$} & 4.96\rlap{$^{***}$} & 4.64\rlap{$^{***}$} & 6.46\rlap{$^{***}$} & 5.97\rlap{$^{***}$} & 6.38\rlap{$^{***}$} \\    & (1.43\rlap{)} & (1.47\rlap{)} & (1.20\rlap{)} & (1.45\rlap{)} & (1.35\rlap{)} & (1.38\rlap{)} \\    FamHouseSize & 1.94 & $-$2.55 & $-$6.47\rlap{$^{**}$} & $-$3.84 & $-$3.12 & $-$2.00 \\    & (2.61\rlap{)} & (3.66\rlap{)} & (3.28\rlap{)} & (3.87\rlap{)} & (3.52\rlap{)} & (3.06\rlap{)} \\    Incomes & 4.36\rlap{$^{***}$} & 2.42 & 5.52\rlap{$^{**}$} & 5.63\rlap{$^{*}$} & 8.02\rlap{$^{***}$} & 12.70\rlap{$^{***}$} \\    & (1.69\rlap{)} & (3.00\rlap{)} & (2.42\rlap{)} & (3.15\rlap{)} & (2.78\rlap{)} & (2.64\rlap{)} \\    Indigenous & 1.26 & 1.96 & 2.41 & 2.38 & 0.46 & $-$0.22 \\    & (1.61\rlap{)} & (1.89\rlap{)} & (1.59\rlap{)} & (2.00\rlap{)} & (1.88\rlap{)} & (1.90\rlap{)} \\    Islam & $-$0.75 & $-$0.91 & $-$0.60 & $-$2.01 & $-$0.88 & $-$1.09 \\    & (1.14\rlap{)} & (1.28\rlap{)} & (1.14\rlap{)} & (1.41\rlap{)} & (1.26\rlap{)} & (1.30\rlap{)} \\    Judaism & 1.32 & 0.93 & 1.47 & 0.28 & 1.35 & 1.15 \\    & (1.01\rlap{)} & (1.08\rlap{)} & (0.92\rlap{)} & (1.10\rlap{)} & (1.02\rlap{)} & (0.97\rlap{)} \\    ManagerAdmin & 2.62\rlap{$^{***}$} & 4.67\rlap{$^{***}$} & 7.47\rlap{$^{***}$} & 7.05\rlap{$^{***}$} & 5.93\rlap{$^{***}$} & 5.64\rlap{$^{***}$} \\    & (0.67\rlap{)} & (1.06\rlap{)} & (0.95\rlap{)} & (1.16\rlap{)} & (1.06\rlap{)} & (0.97\rlap{)} \\    Married & $-$3.93 & $-$2.72 & $-$9.35\rlap{$^{***}$} & $-$10.12\rlap{$^{***}$} & $-$7.91\rlap{$^{**}$} & $-$9.47\rlap{$^{**}$} \\    & (2.51\rlap{)} & (3.56\rlap{)} & (3.12\rlap{)} & (3.55\rlap{)} & (3.57\rlap{)} & (3.85\rlap{)} \\    NoReligion & $-$0.73 & 0.04 & 1.32 & 0.37 & 1.41 & 2.94 \\    & (1.50\rlap{)} & (1.65\rlap{)} & (1.51\rlap{)} & (1.75\rlap{)} & (1.74\rlap{)} & (2.03\rlap{)} \\    OneParentHouse & $-$4.77\rlap{$^{***}$} & $-$3.23 & $-$6.55\rlap{$^{***}$} & $-$7.03\rlap{$^{***}$} & $-$5.32\rlap{$^{***}$} & $-$4.94\rlap{$^{**}$} \\    & (1.49\rlap{)} & (1.99\rlap{)} & (1.81\rlap{)} & (2.04\rlap{)} & (1.97\rlap{)} & (2.03\rlap{)} \\    OtherLanguage & $-$1.02 & 6.88 & 6.21 & 7.80 & 10.13\rlap{$^{**}$} & 9.98\rlap{$^{**}$} \\    & (3.00\rlap{)} & (4.93\rlap{)} & (3.97\rlap{)} & (5.25\rlap{)} & (5.09\rlap{)} & (4.26\rlap{)} \\    PropertyOwned & $-$2.01 & $-$0.30 & 0.74 & $-$1.92 & $-$1.05 & 0.73 \\    & (1.35\rlap{)} & (1.49\rlap{)} & (1.36\rlap{)} & (1.74\rlap{)} & (1.67\rlap{)} & (1.48\rlap{)} \\    RentLoanPrice & $-$2.17 & 0.37 & 1.23 & 3.08 & 1.36 & $-$2.04 \\    & (1.46\rlap{)} & (1.93\rlap{)} & (1.76\rlap{)} & (2.23\rlap{)} & (2.20\rlap{)} & (2.07\rlap{)} \\    SocialServ & 3.31\rlap{$^{***}$} & 2.85\rlap{$^{**}$} & 3.46\rlap{$^{***}$} & 3.72\rlap{$^{**}$} & 2.98\rlap{$^{**}$} & 4.04\rlap{$^{***}$} \\    & (1.27\rlap{)} & (1.40\rlap{)} & (1.17\rlap{)} & (1.46\rlap{)} & (1.28\rlap{)} & (1.15\rlap{)} \\    Transformative & 2.30 & 4.71\rlap{$^{***}$} & 4.58\rlap{$^{***}$} & 4.55\rlap{$^{**}$} & 3.63\rlap{$^{**}$} & 4.05\rlap{$^{***}$} \\    & (1.48\rlap{)} & (1.77\rlap{)} & (1.51\rlap{)} & (1.87\rlap{)} & (1.67\rlap{)} & (1.47\rlap{)} \\    Unemployment & $-$3.39$^{**}$ & $-$3.47$^{**}$ & $-$0.40 & $-$0.68 & 0.81 & 1.93 \\    & (1.37) & (1.69) & (1.45) & (1.80) & (1.47) & (1.32) \\  Residual Standard Error (GLS) & 4.34 & 4.82 & 4.32 & 5.30 & 4.82 & 4.76\\   Constant & 50.80$^{***}$ & 52.63$^{***}$ & 47.31$^{***}$ & 49.92$^{***}$ & 53.52$^{***}$ & 50.46$^{***}$ \\    & (0.76) & (0.59) & (0.44) & (0.52) & (0.54) & (0.64) \\  & \multicolumn{6}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ Observations & 150 & 150 & 150 & 150 & 150 & 150 \\  \hline  \hline \\[-1.8ex]  \textit{Note:}  & \multicolumn{6}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\  \end{tabular}  \end{table}
 
@@
 
-The spatial autoregressive coefficient \(\rho\) is positive and significant in only the 2001 and 2016 elections (Figure \ref{fig:rhovis}), meaning that in these elections, the political climate of an electorate appears to be affected by the attitudes of its neighbours.
-Conversely, in the other four elections, the spatial effect weakens to become insignificant.
-In these years, it appears that the spatial component does not explain anything not already explained by the electoral socio-demographics, meaning electorates effectively voted independently.
+The spatial autoregressive coefficient \(\rho\) is positive and significant in the 2001, 2004 and 2016 elections (Figure \ref{fig:rhovis}).
+In these three elections, there is evidence to suggest that the political climate of an electorate appears to be affected by the attitudes of its neighbours, affecting the electorate's two-party preference.
+Conversely, in the other three elections, the spatial effect weakens to become insignificant (although still positive).
 
@@
 
-{\centering \includegraphics[height=0.2\textheight,]{paper_files/figure-latex/rhovis-1}
+{\centering \includegraphics[height=0.2\textheight,]{anzjs_elections_files/figure-latex/rhovis-1}
 
@@
 
-{\centering \includegraphics[height=0.2\textheight,]{paper_files/figure-latex/plotintercept-1}
+{\centering \includegraphics[height=0.2\textheight,]{anzjs_elections_files/figure-latex/plotintercept-1}
 
@@
 
-\begin{figure}[h]
+\textbackslash begin\{figure\}{[}h{]}
 
-{\centering \includegraphics{paper_files/figure-latex/partresplot1-1}
+\{\centering \includegraphics{anzjs_elections_files/figure-latex/partresplot1-1}
 
-}
+\}
 
-\caption{Partial residual plots by election year for a selection of predictors.
+\textbackslash caption\{Partial residual plots by election year for a selection of predictors.
 Linear model with 95\% confidence bands overlaid.
 Most predictors have a positive relationship: the larger the value the more likely the electorate preferences the Coalition.
-The relationship is relatively robust over time, with the exception of Unemployment.}\label{fig:partresplot1}
-\end{figure}
+The relationship is relatively robust over time, with the exception of OtherLanguageHome and Pop\_00\_19.\}\label{fig:partresplot1}
+\textbackslash end\{figure\}
 
@@
 
-{\centering \includegraphics{paper_files/figure-latex/partresplot2-1}
+{\centering \includegraphics{anzjs_elections_files/figure-latex/partresplot2-1}
 
@@
 Several predictors have a negative relationship: with larger values indicating the electorate more likely preferences Labor.
-Most relationships are relatively stable over elections, except OtherLanguage and Education.}\label{fig:partresplot2}
+Most relationships are relatively stable over elections, except Unemployment and Education.}\label{fig:partresplot2}
 \end{figure}
@@
 To investigate the socio-demographics that have a strong effect on the two-party preferred vote, partial residual plots are used and shown in Figures \ref{fig:partresplot1} and \ref{fig:partresplot2}.
-The partial residuals are the residuals from the fitted model with the estimated effect an individual variable added.
-These show the direction, size and significance of an estimated effect --- the slope of the prediction line matches the estimated coefficient, and the shaded region represents a 95\% confidence band, computed using the method \citet{visreg}.
+Partial residuals, for a given variable, are the residuals from the fitted model with the estimated effect of that variable added to it.
+These plots show the direction, size and significance of an estimated effect, as well as any deviations from linearity.
+In each plot, the slope of the prediction line matches the estimated coefficient and the shaded region represents a 95\% confidence band.
+Plots are computed using the method in \citet{visreg}.
 If a horizontal line can be drawn through the confidence band, then the effect is insignificant.
@@
 Plots for each election are faceted to compare the effects over time in Figures \ref{fig:partresplot1} and \ref{fig:partresplot2}.
-Only socio-demographics that have a significant effect in at least one election are displayed in Figures \ref{fig:partresplot1} and \ref{fig:partresplot2}.
+Only socio-demographics that have a significant effect in at least two elections are displayed in Figures \ref{fig:partresplot1} and \ref{fig:partresplot2}.
 
-It is important here to note the ecological fallacy: insights are being drawn at the electorate level, and cannot be inferred for another disaggregate level (e.g.~individual voters).
+\hypertarget{industry-and-type-of-work}{%
+\subsubsection*{Industry and type of work}\label{industry-and-type-of-work}}
+\addcontentsline{toc}{subsubsection}{Industry and type of work}
+
+Electorates with higher proportions of workers in mining, gas, water, agriculture, waste and electricity (grouped as \texttt{Extractive} industries) are consistently linked with higher support for the Liberal party, with the magnitude of this effect slightly increasing over the years (see row 3 in Figure \ref{fig:partresplot1}).
+This is unsurprising, as the Liberal party has close ties with these traditional energy industries, and typically present policies to reduce taxation on energy production.
+Furthermore, electorates with more workers in construction or manufacturing industries (\texttt{Transformative}) are also more likely to support the Liberal party (see row 4 in Figure \ref{fig:partresplot1}), from 2004 onwards.
+
+Similarly, the proportion of workers in managerial, administrative, clerical and sales roles (\texttt{ManagerAdmin}), is also a significant predictor of two-party preference vote across all six elections, with a higher proportion of people working these jobs increasing Liberal support.
+
+Of these job related variables, the most surprising effect is that associated with the proportion of workers in education, healthcare, social work, community and arts (\texttt{SocialServ}).
+Typically the Labor party has more generous funding schemes affecting these areas of work, so one might expect \texttt{SocialServ} to have a negative effect on two-party preference.
+However, in every election this effect is found to be positive and significant.
 
@@
 Perhaps it is due to these policies that higher income electorates appear more likely to support the Liberal party, as the \texttt{Incomes} factor has a positive effect on Liberal preference (see row 1 in Figure \ref{fig:partresplot1}).
-This effect is significant in every election aside from 2004, where it is only marginally insignificant (\(p = 0.0613\)).
+This effect is significant in every election aside from 2004 and 2010.
 Unemployment however, is not as influential.
-In 2001 and 2004, electorates with higher unemployment align with Labor, but over time this shifts towards support for the Liberal party, culminating in a significantly positive effect in 2016.
-
-\hypertarget{industry-and-type-of-work}{%
-\subsubsection*{Industry and type of work}\label{industry-and-type-of-work}}
-\addcontentsline{toc}{subsubsection}{Industry and type of work}
-
-Electorates with higher proportions of workers in mining, gas, water, agriculture, waste and electricity (grouped as \texttt{Extractive} industries) are consistently linked with higher support for the Liberal party, with the magnitude of this effect slightly increasing over the years (see row 3 in Figure \ref{fig:partresplot1}).
-This is unsurprising, as the Liberal party has close ties with these traditional energy industries, and typically present policies to reduce taxation on energy production.
-Furthermore, electorates with more workers in construction or manufacturing industries (\texttt{Transformative}) are also more likely to support the Liberal party (see row 4 in Figure \ref{fig:partresplot1}).
-
-Similarly, the proportion of workers in managerial, administrative, clerical and sales roles (\texttt{ManagerAdmin}) is also a significant predictor of two-party preference vote across all six elections, with a higher proportion of people working these jobs increasing Liberal support.
-The magnitude of this effect also seems to increase over the years.
-
-\hypertarget{household-mobility}{%
-\subsubsection*{Household mobility}\label{household-mobility}}
-\addcontentsline{toc}{subsubsection}{Household mobility}
-
-In each of the six elections, electorates with a higher proportion of people that have recently (in the past five years) moved house (\texttt{DiffAddress}) are more likely to support the Liberal party, although this effect was marginally insignificant in 2013 (see row 6 in Figure \ref{fig:partresplot1}.
-Having controlled for characteristics of house ownership and rental prices (via the factors \texttt{PropertyOwned} and \texttt{RentLoan} respectively), this effect is somewhat surprising.
-
-\hypertarget{relationships}{%
-\subsubsection*{Relationships}\label{relationships}}
-\addcontentsline{toc}{subsubsection}{Relationships}
-
-De facto relationships, but not marriages, are found to be an important (and significant) predictor of the two-party preferred vote in all six elections, with more de facto relationships associated with higher support for the Labor party.
-The proportion of individuals who are married however, is insignificant (not shown).
+In 2001 and 2004, electorates with higher unemployment align with Labor, but over time this shifts towards support for the Liberal party, culminating in a positive (insignificant) effect in 2016.
 
@@
 
-Regions comprising more older people are often believed to be more conservative, and indeed it found that electorates with a higher median age are more likely to support the Liberal party --- although this effect is only significant in 2007 and 2010 (see row 2 in Figure \ref{fig:partresplot2}).
+The older Australian population is often believed to be more conservative, and the left leaning political parties (including Labor) typically have a stronger appeal to younger people.
+This effect is indeed observed across the six elections, with populations between 20 and 34 years of age (\texttt{Pop\_20\_34}) being very strongly aligned with Labor preference (bottom row in Figure \ref{fig:partresplot2}).
+Larger populations of 35 to 54 year olds (\texttt{Pop\_35\_54}) are also associated with Labor, but the magnitude of this effect is far smaller.
+Populations under 20 years of age is only significant in 2007, where \texttt{Pop\_00\_19} increased Liberal support.
 
@@
 
-Since 2007, electorates with higher education levels are associated with supporting the Labor party, although this effect is only significant in 2016.
-Before 2007, education has an almost zero effect (see row 3 in Figure \ref{fig:partresplot1}).
+Since 2007, electorates with higher education levels are associated with supporting the Labor party, with this effect being significant in 2007, 2013 and 2016 and only marginally insignificant in 2010.
+In the elections before 2007, education has a negligible effect (see row 3 in Figure \ref{fig:partresplot1}).
+Additionally, student populations (\texttt{CurrentlyStudying}) do not affect electoral party preference in any election (not shown).
 
@@
 Larger migrant populations from Asia, the Middle East, South-Eastern Europe, the United Kingdom and elsewhere, are either associated with Labor support, or have no effect.
-Of these areas, only South-Eastern European populations appear significant in each election, with the proportion of Asian migrants also being significant in 2010.
+Of these areas, only South-Eastern European populations significant affect party preference, with larger populations associating with Labor in 2013 and 2016 (row 2, Figure \ref{fig:partresplot2}).
 Speaking other languages (aside from English) however, appears to have a far stronger effect, as observed through the \texttt{OtherLanguage} variable.
-Electorates with more diverse speech are associated with higher support for the Liberal party from 2004 onwards, with this effect being significant in 2007, 2010 and 2016.
-Furthermore, of the variables relating to religion, only Judaism shows a consistent effect, with electorates with relatively large Jewish populations more likely to vote Liberal.
+Electorates with more diverse speech are associated with higher support for the Liberal party from 2004 onwards, with this effect being significant in 2013 and 2016 row 7, Figure \ref{fig:partresplot1}).
+Furthermore, none of the variables relating to religious beliefs aside from Christianity have a material effect in any election (this includes the Buddhist, Muslim, Jewish, areligious and Indigenous Australian populations).
+The association between Christian populations (\texttt{Christianity}) and the Liberal party steady increases over the years, becoming positive and significant in 2016.
 
-\hypertarget{a-note-on-similar-variables}{%
-\subsubsection*{A note on similar variables}\label{a-note-on-similar-variables}}
-\addcontentsline{toc}{subsubsection}{A note on similar variables}
+\hypertarget{households}{%
+\subsubsection*{Households}\label{households}}
+\addcontentsline{toc}{subsubsection}{Households}
 
-Many of the Census variables represent similar information, which is why factors were created and some variables were removed.
-However, some variables remain which are closely related.
-For example, an electorate's income level (via \texttt{Incomes}) is likely to be related to electoral unemployment and labor force participation (via \texttt{Unemployment}).
-In 2001, the coefficient estimate for \texttt{Unemployment} is negative but not significant, whilst the \texttt{Incomes} variables is significant.
-If the \texttt{Incomes} variable is removed from the model in 2001, \texttt{Unemployment} absorbs the negative effect, becoming significant (\(p = 0.0056\)).
+In 2001, 2004 and 2016, higher proportions of people that have recently (in the past five years) moved house (\texttt{DiffAddress}) increased electoral support for the Liberal party (see row 5 in Figure \ref{fig:partresplot1}).
+Having controlled for characteristics of house ownership and rental prices (\texttt{PropertyOwned} and \texttt{RentLoan} - both of which have no effect), this effect is somewhat surprising.
+
+Higher proportions of single parent households are associated with Labor support in all elections (albeit insignificant in 2004, see row 6 in Figure \ref{fig:partresplot2}), whereas the electoral family and household sizes (via the \texttt{FamHouseSize} variable) do not appear to be associated with either party in any election.
+
+\hypertarget{relationships}{%
+\subsubsection*{Relationships}\label{relationships}}
+\addcontentsline{toc}{subsubsection}{Relationships}
+
+From 2007 onwards, both marriages (\texttt{Married}) and de facto relationships (\texttt{DeFacto}) are found to be strong predictors of the two-party preferred vote in opposition to the Liberal party.
+In 2001 and 2004 neither of these variables are significant (see rows 4 and 5 in Figure \ref{fig:partresplot2}).
+
+\hypertarget{robustness}{%
+\subsubsection{Robustness}\label{robustness}}
+
+Two robustness checks are conducted to confirm model stability.
+First, a model for each election is fit using only the variables that are found to be significant in at least one of the six elections.
+Let this model be the reduced model, and the original model be the full model.
+The estimated coefficients of the variables in the reduced models all fall within their respective \(95\%\) confidence intervals from the full models, confirming that the significant effects are robust to variable selection.
+
+The second check involves computing pairwise correlations amongst the explanatory variables, and seeing if the removal of one variable (from a pair) affects the other's estimated effect.
+For example, the variables \texttt{Born\_Asia} and \texttt{BornElsewhere} have a pairwise correlation of \(0.93\) - the largest of any pair.
+A model is fit omitting \texttt{BornElsewhere} (the reduced model).
+The estimated coefficent of \texttt{Born\_Asia} is then compared with the confidence interval for \texttt{Born\_Asia} in the full model.
+It is found that for each of the top ten pairwise correlations, the estimated effect in the reduced model lies within the confidence interval from the full model.
+This second check indicates that variables with the highest pairwise correlation are not masking the effects of each other.
 
@@
 It is often hypothesized that states have systematic differences that cause their electorates to vote differently.
-Boxplots of residuals grouped by state (Figure \ref{fig:resstate}) reveal that the data reflects this -- there appears to be a state-specific effect not captured by the models.
+Boxplots of residuals grouped by state (Figure \ref{fig:resstate}) reveal that the data reflects this to a limited extent.
 Tasmania and the Australian Capital Territory appear to have a bias towards Labor, whereas the Northern Territory tends towards voting Liberal.
@@
 
-{\centering \includegraphics[width=0.7\linewidth,]{paper_files/figure-latex/resstate-1}
+{\centering \includegraphics[width=0.7\linewidth,]{anzjs_elections_files/figure-latex/resstate-1}
 
@@
 
-\caption{Boxplot of residuals by state with jittered points. States ordered by median residual. A state-specific bias not captured by the model is evident.}\label{fig:resstate}
+\caption{Boxplot of residuals by state with jittered points.
+States ordered by median residual.
+A state-specific bias present only in the smaller states appears to have not been captured by the model.}\label{fig:resstate}
+\end{figure}
+
+\hypertarget{residuals-by-party-incumbency}{%
+\subsubsection*{Residuals by party incumbency}\label{residuals-by-party-incumbency}}
+\addcontentsline{toc}{subsubsection}{Residuals by party incumbency}
+
+The incumbent party appears to have a distinct advantage at the next election.
+The boxplots in Figure \ref{fig:resincumbent} show that if either of the Labor or Liberal parties won the seat at the previous election, the electorate is likely to vote in their favour, over and above any socio-demographic effects --- this effect has not been captured by the model.
+
+\begin{figure}[h]
+
+{\centering \includegraphics[width=0.7\linewidth,]{anzjs_elections_files/figure-latex/resincumbent-1}
+
+}
+
+\caption{Boxplot of residuals for incumbent and non-incumbent parties each year. An incumbent advantage is evident and has not been captured by the model.}\label{fig:resincumbent}
 \end{figure}
@@
 Based on the distribution of the Cook's distance values, a Cook's distance greater than \(0.1\) is considered to be influential and a potential outlier.
-The electorate of Sydney (NSW) has a large Cook's distance from 2001 to 2013, due to its diverse population (language, birthplace and religion), high number of de facto relationships, high income, high household mobility and small amount of workers in extractive and transformative jobs.
+The electorate of Sydney (NSW) has a large Cook's distance from 2001 to 2010, due to its diverse population (language, birthplace and religion), high number of defacto relationships, high income, high household mobility and small amount of workers in extractive and transformative jobs.
 It has remained a strong supporter of the Labor party and the Liberal vote is severely overpredicted by the model, making it an outlier.
-Nearby in metropolitan NSW, the electorate of Wentworth is found to be an outlier in all but the 2007 election.
+Nearby in metropolitan NSW, the electorate of Wentworth is found to be an outlier in the 2013 and 2016 elections.
 Although historically Liberal, its two-party vote jumped by over 10 percentage points in 2010 without experiencing any notable changes in its socio-demographic makeup --- implying that this may be the direct effect of its Liberal member, Malcolm Turnbull, becoming the leader of the Liberal party.
-Liberal support in Wentworth is underpredicted by the model in each year, and more so with Turnbull as Liberal leader.
+In the elections since, the model underpredicts Wentworth's Liberal support.
 
-Lingiari, an electorate taking up almost all of the Northern Territory, is an outlier in the 2001--2007 elections due to its large Indigenous population, young age profile and low rates of property ownership.
+Lingiari, an electorate taking up almost all of the Northern Territory, is an outlier in the all but the 2013 election due to its large Indigenous population, young age profile and low rates of property ownership.
 Fowler (NSW) has a diverse population with a high proportion of migrants, many Buddhists and Muslims, and has strong Labor support, making it influential in 2001, 2004 and 2010.
-Other electorates with large Cook's distance are Barton (NSW) and Leichhardt (QLD) in 2016, and Canberra (ACT) in 2007.
+Other electorates with large Cook's distance are Canberra (ACT) and Durack (WA) in 2013, and Solomon (NT) in 2016.
 
@@
 Across the past six elections, most of the socio-demographics that drive the electoral two-party preferred vote are found to remain steady, whilst a few (typically weaker) effects vary over time.
-Industry and type of work are particularly influential, with energy-related and manufacturing/construction jobs, as well as administrative roles being strongly linked with the Liberal party in all elections.
+Industry and type of work are particularly influential.
+Energy-related and manufacturing/construction jobs, as well as administrative roles and jobs in education and social services are strongly linked with the Liberal party in all elections.
 Incomes have a similarly consistent effect, with higher income areas supporting Liberal.
-Higher levels of unemployment shift from weak association with Labor to a significant Liberal effect over the years, and higher education levels are associated with Labor from 2007 (although only significant in 2016).
-It is also found that electorates with higher household mobility support Liberal, birthplace diversity favours Labor and more de facto relationships align with Labor preference --- although marriages, family and household sizes have no material influence.
-Furthermore, the neighbourhood (spatial) effects are found to be positive in all elections, although only significant in 2001 and 2016, meaning that in the 2004--2013 elections, electorates effectively voted independently.
+Higher levels of unemployment shift from weak association with Labor to a significant Liberal effect over the years, and higher education levels are associated with Labor from 2007 (although marginally insignificant in 2010).
+Electorates with large populations 20 to 34 years are strongly associated with Labor, whilst the 35 to 54 year old bracket also increases Labor support, but to a lesser extent.
+It is also found that birthplace diversity slightly favours Labor, relationships (both marriages and de facto relationships) align with Labor preference from 2010 onwards, and the influence of Christian populations has trended towards Liberal support whilst other religions have negligible effects.
+Family and household sizes have minimal influence, although electorates with more single parent households are linked with Labor support.
+Furthermore, the neighbourhood (spatial) effects are found to be positive in all elections and significant in 2001, 2004 and 2016, meaning that neighbouring party preference appears to be an influential factor in those years.
 

### 2019-09-18 anzjs_elections.Rmd
Robustness section within modelling - contains checks and influential & outlier sections

@@
 
-### Robustness
-
-Two robustness checks are conducted to confirm model stability.
-First, a model for each election is fit using only the variables that are found to be significant in at least one of the six elections.
-Let this model be the reduced model, and the original model be the full model.
-The estimated coefficients of the variables in the reduced models all fall within their respective $95\%$ confidence intervals from the full models, confirming that the significant effects are robust to variable selection.
-
-The second check involves computing pairwise correlations amongst the explanatory variables, and seeing if the removal of one variable (from a pair) affects the other's estimated effect.
-For example, the variables `Born_Asia` and `BornElsewhere` have a pairwise correlation of $0.93$ - the largest of any pair.
-A model is fit omitting `BornElsewhere` (the reduced model).
-The estimated coefficent of `Born_Asia` is then compared with the confidence interval for `Born_Asia` in the full model.
-It is found that for each of the top ten pairwise correlations, the estimated effect in the reduced model lies within the confidence interval from the full model.
-This second check indicates that variables with the highest pairwise correlation are not masking the effects of each other.
-
 ## A closer look at the residuals
@@
 
-### Outlier electorates {-}
+## Robustness
+
+### Checks
+Three robustness checks are conducted to confirm model stability.
+First, a model for each election is refit using only the variables that are found to be significant in at least one of the six elections.
+The estimated coefficients of the variables in the refitted models all fall within their respective $95\%$ confidence intervals from the full models.
+The second check involves the ten largest pairwise correlations.
+For each pair, a model for each election is refit omitting one of the two variables.
+It is found that for each of these pairs, the estimated effect of the remaining variable in the reduced model lies within the $95/%$ confidence interval from the full model.
+The final check is a visual exploration of different variable projections using a tour [@tourr], for each election.
+No definitive signs of multicollinearity are observed, and as expected (given the nature of spatial data), there is some clumping of electorates for certain projections.
+
+### Influential and outlier electorates {-}
 
@@
   filter(cooksd > 0.1) %>%
-  select(DivisionNm, StateAb, year, cooksd, Residuals, LNP_Percent, everything()) %>%
+  dplyr::select(DivisionNm, StateAb, year, cooksd, Residuals, LNP_Percent, everything()) %>%
   View()

### 2019-09-18 response_for_ANZJS_editor.Rmd
minor changes to robustness in response to editor. added response to reviewer 1.

@@
 
+$ DONE $
+
 *This is a valid point, and we have revised the imputation method to adopt this suggestion.
@@
 Our original method was chosen because we believed that the imputation method would not materially affect results.
-We found that our results did not change much by adopting this new imputation method.*
+With this change in imputation method, our results are effectively the same as the previous submission.*
+
+-- We made the change this way, and the results are effectively the same.
 
@@
 When we fit a model using just the variables with significant effects, we find that the results do not change much.
-Each of the estimated coefficients from the reduced model (containing only the subset significant variables) lie within the 95% confidence interval from its corresponding full model (containing all 32 variables).*
+Each of the estimated coefficients from the reduced model (containing only the subset significant variables) lie within the 95% confidence interval from its corresponding full model (containing all 32 variables).
+*
 
-*Furthermore, we have added another robustness check, where we consider the largest pairwise correlations amongst all variables (computed on a dataset of all elections combined).
-Taking each pair one by one, we remove one of the two variables, and find that the estimated coefficient for the remaining variable lies within the 95% confidence interval from the full model.
-This is true for all of the top 10 pairwise correlations.*
+*We also checked the effect of omitting a variable that is contained in the ten largest pairwise correlations.
+For each pair, a model for each election is refit omitting one of the two variables.
+It is found that for each of these pairs, the estimated effect of the remaining variable in the reduced model lies within the $95/%$ confidence interval from the full model.*
+
+*A third and final check is the visual exploration of different variable projections using a tour (for each election). No definitive signs of multicollinearity are observed.*
+
+$ DONE $
+-- Changes --
+
+Refitting without a variable. Doing this selectively on the ones that are more highly correlated. Re-fit with a different subset of variables.
+
+Another check is the TOUR.
+
+Bundle robustness and diagnostiscs - call it robustness:
+1) First para is on multicollinearity checks - don't need as much detail as I have
+2) Influential observations - couple paragraphs on cooks' dist and influence diagnostics
 
@@
 Additionally, on revision of the imputation method, we find that the spatial effect is significant in 2004 as well as 2001 and 2016.*
+
+#### Reviewer 1's comments
+
+Its goal is to account for – though the word ‘explain’ is used! – patterns of voting at Australian federal elections, using ecological data, but with no reference to the ecological fallacy.
+
+*We agree that more attention could have been given to the ecological fallacy in our original submission.
+In response, we have expanded our remarks and moved them to the introduction (Section 1).
+We are explicit in stating that "drivers of individual voter behaviour may vary from what is observed at the electorate level", which may not have been clear enough in the original submission.*
+
+The paper is written in virtual total ignorance of the large amount of work done on Australian voting patterns; apart from reference to a few papers by geographers there is not a single mention of work by political scientists – of which there is a great deal.
+As a consequence there is no theoretical background whatsoever, let alone hypotheses to be tested.
+It is banal empiricism: throw a lot of data (what are called ‘electoral socio-demographic variables’) at something and see what comes out.
+
+*We formed our approach to exploring this question by reviewing the Australian literature in this area, which has been referenced accordingly.
+As this is an exploratory piece, and we wanted to let the relationships reveal themselves, rather than expicitly stating hypotheses to be tested.*
+
+The banality of this approach is illustrated by the finding that ‘more de facto relationships’ are associated with greater support for Labor.
+No rationale for this is offered – why should people in such relationships be more likely to vote Labor?
+And while there may be a statistically significant relationship (i.e.
+slope coefficient) it may be substantively trivial – a fact hidden because the data are transformed.
+If there is a small SD then the ‘real’ difference between a place with many de facto relationships and another with few will be small and will not account for much of the absolute variation in Labour support at all.
+Further, since all of the relationships with individual variables are partial, it may well be that when ‘all other variables are held constant’ the difference in Labour support is trivial.
+
+*Since this study is of electorates, rather than individual voters, changing an electorate's demographics can have an effect on two-party preference beyond just the head count of population who's demographics are changed.
+All this means is that effect of de facto relationships on the electorate's vote is not necessarily just as a result of the votes from people who are in a de facto relationship.*
+
+Much is done – unnecessarily in my view – in preparing the data through imputation.
+This has two components.
+First there is the creation of census data for the 150 constituencies because boundaries of the large census reporting areas may not coincide with the constituencies’ at the time of an election.
+Imputation is deployed by assuming that if 50% of area A is in district X then 50% of A’s population is allocated to X without any recognition that area A may be internally heterogeneous.
+But that is in any case unnecessary.
+Australian census data are reported – as the authors note – for very small areas with populations of only a few hundred each, and there are shape files associated with these areas.
+It is thus a straightforward task to amalgamate the small areas into the electoral districts and get the ‘real’ data for each without any (or at best a very, very small amount of) imputation.
+
+*We have revised our imputation method to reflect this suggestion.
+We now use Census data at the lowest level of aggregation available - the geographic region type of 'Statistical Area 1' - to obtain the approximation of electoral boundaries (see Section 2.2).
+Our original method was chosen because we believed that the imputation method would not materially affect results.
+With this change in imputation method, our results are effectively the same as the previous submission.*
+
+The authors recognise that there is likely to be collinearity among 65 Census variables – why those; we are never told?
+
+*The 50 Census variables are chosen because they appear in previous studies of Australian elections (those mentioned in the Introduction), and we believe that they capture key characteristics of an electorate - age, income, employment, religion, birthplace and relationships.*
+
+*Note that we have refined the subset of variables considered for analysis in this paper to 50 Census variables.
+Previously we started with 65 variables, but many of these variables were removed in the variable selection process because they provided additional granularity that we did not want to pursue.
+These never should have been included to begin with - and we have removed them in our updated submission.
+For example, we had previously included variables relating to Christian denominations (Anglican, Catholic) in our 65 variables, but did not include them in the model.*
+
+And so, as others do, they did a pca of the data matrix, reducing it to four components – we are not told what rotation was employed to achieve simple structure.
+We are not given the results of this analysis, merely a naming of the six factors identified – how do you get six ‘factors’ from four components?
+
+*Our approach was to combine the election data together into a single dataset of 900 observations.
+We computed the principal components on this.
+The first four components are chosen because the amount of variance explained drops off after the fourth PC.
+The decision is made to omit these details of this type because they detract from the core steps of our variable selection procedure.*
+
+*The factors are not the principal components themselves.
+Rather, the loadings in the first four principal components are used to determine which variables should be grouped together to generate a new variable in their place, which we have called a factor.
+*
+
+My expectation was that they would then fit a model using the four components as the independent variables, thus avoiding any collinearity and confounding.
+But they didn’t.
+Instead they selected 30 of the 65 variables – which must still be collinear; the resulting equation in Table 2 clearly implies this given the large number of (a) insignificant relations (normal with collinearity – see a recent paper in Quality and Quantity), (b) unreasonably large coefficients (-8.07 for de facto, for example), and (c) very large standard errors for many of the variables.
+
+*This was a very helpful suggestion to explore further, and as such an additional section on robustness has been included in the revised paper (Section 4.4).
+When we fit a model using just the variables with significant effects, we find that the results do not change much.
+Each of the estimated coefficients from the reduced model (containing only the subset significant variables) lie within the 95% confidence interval from its corresponding full model (containing all 32 variables).
+*
+
+*We also checked the effect of omitting a variable that is contained in the ten largest pairwise correlations.
+For each pair, a model for each election is refit omitting one of the two variables.
+It is found that for each of these pairs, the estimated effect of the remaining variable in the reduced model lies within the $95/%$ confidence interval from the full model.*
+
+*A third and final check is the visual exploration of different variable projections using a tour (for each election). No definitive signs of multicollinearity are observed.*
+
+The authors fit a spatial autoregressive model, on the assumption that neighbouring places are more likely to be similar than distant places (the ‘first law of geography’).
+This is only a significant finding at two of the elections and is wrongly interpreted: it does not necessarily show that then ‘the political climate of an electorate appears to be affected by the attitudes of its neighbours’ but, much more likely, that neighbouring places share characteristics in common that affect voting behaviour but are not captured by the independent variables included in the model.
+Spatial autocorrelation may reflect spatial diffusion processes – but in which direction?
+– but is much more likely to reflect an under-specified model.
+
+*In retrospect, we agree that this is perhaps a naïve conclusion.
+It is possible that unobserved variables are being captured in the spatial effect, and as such we have removed this interpretation from the revised paper.
+Additionally, on revision of the imputation method, we find that the spatial effect is significant in 2004 as well as 2001 and 2016.*
+
+#### Reviewer 2's comments
+
+*1.
+It seems very strange to assume that population is uniform within electorates (line 133) so that weighting is proportional to area, and not to the number of people in an area of intersection.
+An electorate on the urban fringe may be strongly affected by loss of a relatively small but highly populated area but little affected by incorporating a large rural area.
+A simple indirect way to get at this might be to address the question: Would the results change much if A_{s,t} was log area, for each s,t ?*
+
+...A
+
+*2. It seems to me that the use of Principal Components analysis is unclear.*
+*a.
+PC appears to have been used as a model selection criterion, with the variables related to the first four PCs included into the model in Table 2.
+However typically when principal components are computed they only involve the explanatory variables, not the response, and so there is no guarantee that all the first four PCs (and only them) will be related to the response.
+For example using the eechidna data for 2016 I found the fifth PC was extremely significant.
+But this is a minor point.*
+
+...A
+
+*b.
+The variables which load highly on a PC are likely to be highly correlated, and so if these high-loading variables are used together in the regression model we may again have multicollinearity.
+I found this was the case when applying the author’s regression model to the 2016 data.
+This may be why several of the predictors in Table 2 are never related significantly related to the response in any of the years.*
+
+... ?
+
+*c.
+Conversely there could be a variable which is independent of the other variables (so representing 1 / 65th of the explanatory information) but by dint of that fact not be included in the first four PCs.
+In short, the narrative leaves me unconvinced that all the variables included are necessary or that all the relevant variables are included in the regression.*
+
+... Are there more variables we should get? I think no.
+
+*d. I presume the “factors” referred to lines 210-220 were created as the first principal component of the constituent raw variables?*
+
+... A
+
+*e. The description of the variables included in “Education” is not clear.*
+
+... Be more precise in explaning the factors created.
+
+*f. I could not get the variable ‘OtherLanguageHome’ in 2016 to be significant, but it also had a high VIF with other explanatory variables. Please check this.*
+
+... TOUR - ask Di for help.
+
+*3.
+I am not convinced it is reasonable to summarise the effect of age by median age.
+Perhaps age effects are more at the edges of the age distribution than in the middle?
+I would have thought the percentage of the population in different age bands e.g.
+20-34 years and Over 65years, might be significant.
+For 2016 the % aged 20-34 years was highly significant and removed median age.*
+
+... A
+
+*4.
+Is there any tendency for the vote in an election to depend on how long the incumbent has been in power?
+In your cross-sectional models this would be confounded with any other effects in the intercept, but a comment would be nice to know.
+It would help in terms of interpreting the variations in intercept, and setting a reader’s mind to rest that the issue has been considered.
+Perhaps a something like a simple regression of average vote (or the intercept) on length of incumbency could address the issue.*
+
+...  Don't include incumbent in model, but do plot residuals against incumbency to see if there is an effect. We want the model to only capture the socio-demographic effects.
+
+-- Section on incumbent effect (need to get winners each year from 1998, 2001 to do this)

### 2019-09-25 response_for_ANZJS_editor.Rmd
response to reviewers 1 and 2

@@
 
-*Our approach was to combine the election data together into a single dataset of 900 observations.
+*Our approach was to combine the six election datasets together into a single dataset of 900 observations.
 We computed the principal components on this.
@@
 
-*1.
 It seems very strange to assume that population is uniform within electorates (line 133) so that weighting is proportional to area, and not to the number of people in an area of intersection.
 An electorate on the urban fringe may be strongly affected by loss of a relatively small but highly populated area but little affected by incorporating a large rural area.
-A simple indirect way to get at this might be to address the question: Would the results change much if A_{s,t} was log area, for each s,t ?*
+A simple indirect way to get at this might be to address the question: Would the results change much if A_{s,t} was log area, for each s,t ?
 
-...A
+*We have revised our imputation method to avoid this assumption.
+We now use Census data at the lowest level of aggregation available - the geographic region type of 'Statistical Area 1' (SA1) - to obtain the approximation of electoral boundaries (see Section 2.2).
+There are on average approximately 400 people in each SA1.
+This approach will successfully capture any densely populated areas on the fringe of an electorate.
+Our original method was chosen because we believed that the imputation method would not materially affect results.
+With this change in imputation method, our results are effectively the same as the previous submission.*
 
-*2. It seems to me that the use of Principal Components analysis is unclear.*
-*a.
-PC appears to have been used as a model selection criterion, with the variables related to the first four PCs included into the model in Table 2.
+It seems to me that the use of Principal Components analysis is unclear.
+(a) PC appears to have been used as a model selection criterion, with the variables related to the first four PCs included into the model in Table 2.
 However typically when principal components are computed they only involve the explanatory variables, not the response, and so there is no guarantee that all the first four PCs (and only them) will be related to the response.
 For example using the eechidna data for 2016 I found the fifth PC was extremely significant.
-But this is a minor point.*
+But this is a minor point.
 
-...A
+*We have used principal components to inform which variables should be combined into factors, rather than use them directly as explanatory variables.
+This is done as a step to address any (near) multicollinearity in the explanatory variables.
+Any variables that are not used in creating a factor still remain in the model.
+The point raised is valid - there is no guarantee about any of these factors being related to the response.*
 
-*b.
-The variables which load highly on a PC are likely to be highly correlated, and so if these high-loading variables are used together in the regression model we may again have multicollinearity.
+(b) The variables which load highly on a PC are likely to be highly correlated, and so if these high-loading variables are used together in the regression model we may again have multicollinearity.
 I found this was the case when applying the author’s regression model to the 2016 data.
-This may be why several of the predictors in Table 2 are never related significantly related to the response in any of the years.*
+This may be why several of the predictors in Table 2 are never related significantly related to the response in any of the years.
 
-... ?
+*By taking a subset of variables with high loadings on a principal component and combining them into a single variable, we are trying to address potential multicollinearity.
+Two factors that have been created from the same PC will be correlated, which could mean that (near) multicollinearity persists.
+However, we have conducted three robustness tests for multicollinearity, and no conclusive signs of multicollinearity were observed.*
 
-*c.
-Conversely there could be a variable which is independent of the other variables (so representing 1 / 65th of the explanatory information) but by dint of that fact not be included in the first four PCs.
-In short, the narrative leaves me unconvinced that all the variables included are necessary or that all the relevant variables are included in the regression.*
+*The first check involved fitting a model using just the variables with significant effects.
+We find that the estimated coefficients in the reduced model (containing only the subset significant variables) all lie within the 95% confidence interval from its corresponding full model (containing all 32 variables).
+*
 
-... Are there more variables we should get? I think no.
+*We also checked the effect of omitting a variable that is contained in the ten largest pairwise correlations.
+For each pair, a model for each election is refit omitting one of the two variables.
+It is found that for each of these pairs, the estimated effect of the remaining variable in the reduced model lies within the $95/%$ confidence interval from the full model.*
 
-*d. I presume the “factors” referred to lines 210-220 were created as the first principal component of the constituent raw variables?*
+*A third and final check is the visual exploration of different variable projections using a tour (for each election). No definitive signs of multicollinearity are observed.*
 
-... A
+(c) Conversely there could be a variable which is independent of the other variables (so representing 1 / 65th of the explanatory information) but by dint of that fact not be included in the first four PCs.
+In short, the narrative leaves me unconvinced that all the variables included are necessary or that all the relevant variables are included in the regression.
 
-*e. The description of the variables included in “Education” is not clear.*
+*Any variables that are not used in creating a factor still remain in the model.
+The four PCs are used ONLY to inform which variables are correlated and should be combined into a factor.
+For example, median household, personal and family incomes all have large positive loadings in the first principal component, so we combine them into a factor called `Incomes`.
+Any variable that does not have a large loading in any of the four PCs are not discarded - they remain in the model.
+It is of course possible that there exist variables beyond what we have considered that could affect two party preference.
+However, we believe that the variables we have used cover the core demographic components of an electorate - age, income, employment, religion, birthplace and relationships.*
 
-... Be more precise in explaning the factors created.
+(d) I presume the “factors” referred to lines 210-220 were created as the first principal component of the constituent raw variables?*
 
-*f. I could not get the variable ‘OtherLanguageHome’ in 2016 to be significant, but it also had a high VIF with other explanatory variables. Please check this.*
+*In our revised submission, we have made an effort to clarify what we have defined a "factor" to be.
+A "factor" is created when there is evidence to suggest that a group of variables are explaning similar information, and if they were all included in the model as separate variables, there would be severe issues with multicollinearity.
+To inform which variables should be grouped, we use the first four PCs from PCA.
+In a single PC, if there are is a subset of variables with large (positive or negative) loadings that intuitively could represent similar information, we combine them into a factor using a weighted sum, where the weights are +1 or -1 depending on the sign of the loading in the that PC.
+This means that many factors can be created from a single PC, and that any variables that do not appear in a factor are still included in the model as stand-alone variables.*
 
-... TOUR - ask Di for help.
+(e) The description of the variables included in “Education” is not clear.
 
-*3.
+*We have provided more details on the variables that are combined into each of the factors, including `Education`.
+`Education` is a weighted sum of seven variables that represent: high school completions, undergraduate and postgraduate degrees, proportion of employed people working as professionals, proportion of jobs in finance, proportion of workers who are laborers, proportion of workers who work as a tradesperson, diploma and certificate qualifications.*
+
+(e) I could not get the variable ‘OtherLanguageHome’ in 2016 to be significant, but it also had a high VIF with other explanatory variables. Please check this.
+
+*Because of the revised imputation method, the data is now slightly different (however the conclusions do not change). We have checked that `OtherLanguageHome` is significant in 2016.*
+
 I am not convinced it is reasonable to summarise the effect of age by median age.
@@
 20-34 years and Over 65years, might be significant.
-For 2016 the % aged 20-34 years was highly significant and removed median age.*
+For 2016 the % aged 20-34 years was highly significant and removed median age.
 
-... A
+*This was a very useful suggestion.
+Instead of using median age, we summarise age of an electorate using its distribution amongst age brackets 0-19, 20-34, 35-54 and 55+ years.
+We find that higher populations in the 20-34 and 35-54 age brackets are associated with electoral support for Labor in all of the six elections, whilst populations 0-19 years have weakly positive associations with the Liberal party (aside from 2007, when this relationship was significantly positive).*
 
-*4.
 Is there any tendency for the vote in an election to depend on how long the incumbent has been in power?
@@
 It would help in terms of interpreting the variations in intercept, and setting a reader’s mind to rest that the issue has been considered.
-Perhaps a something like a simple regression of average vote (or the intercept) on length of incumbency could address the issue.*
+Perhaps a something like a simple regression of average vote (or the intercept) on length of incumbency could address the issue.
 
-...  Don't include incumbent in model, but do plot residuals against incumbency to see if there is an effect. We want the model to only capture the socio-demographic effects.
-
--- Section on incumbent effect (need to get winners each year from 1998, 2001 to do this)
+*This is an interesting question.
+We have included some boxplots of the residuals from each election, split by whether the party was the incumbent party.
+We find that the incumbent party has a clear advantage of around 3-5 percentage points.
+This can be found in Section 4.4.*

### 2019-09-25 anzjs_elections.Rmd
add axes labels to boxplot

@@
 `FamHouseSize` (average household size, ratio of people to families, incidence of single person households, incidence of households containing a couple with kids, incidence of households containing a couple without kids); and
-`Education` (high school completions, university qualifications, proportion of people working as professionals, proportion of jobs in finance,  proportion of workers who are laborers, proportion of workers who work as a tradesperson, diploma and certificate qualifications).
+`Education` (high school completions, undergraduate and postgraduate degrees, proportion of employed people working as professionals, proportion of jobs in finance, proportion of workers who are laborers, proportion of workers who work as a tradesperson, diploma and certificate qualifications).
 For each of these groupings, a factor is created by taking a weighted sum of the variables.
@@
   scale_fill_manual(values = c('white', 'grey'), labels = c("No", "Yes")) +
-  theme_bw()
+  theme_bw() +
+  labels(x = "Year", y = "Residuals")
 

### 2019-09-25 anzjs_elections.Rmd
Removed "electorates voted independently" and instead explain that this indicates an effect of other characteristics neighbours share. Added some detail to factor creation process. Added acknowledgements for reviewers.

@@
 Socio-demographic information is derived from the Australian Census, which occurs every five years.
-Since a Census is not directly available for each election, spatio-temporal imputation is employed to estimate Census data for the electorates at the time of each election.
+Since a Census is not directly available for each election, an imputation method is employed to estimate Census data for the electorates at the time of each election.
 This accounts for both spatial and temporal changes in electoral characteristics between Censuses.
-To capture any spatial heterogeneity, a spatial error model is estimated for each election, which incorporates a spatially structured random effect vector that can be thought of as the unobserved political climate in each electorate.
-Over time, the impact of most socio-demographic characteristics that affect electoral two-party preference do not vary, with industry of work, incomes, household mobility and de facto relationships having strong effects in each of the six elections.
+To capture any spatial heterogeneity, a spatial error model is estimated for each election, which incorporates a spatially structured random effect vector.
+Over time, the impact of most socio-demographic characteristics that affect electoral two-party preference do not vary, with age distribution, industry of work, incomes, household mobility and de facto relationships having strong effects in each of the six elections.
 Education and unemployment are amongst those that have varying effects.
-It is also found that between 2004 and 2013, the spatial effect was insignificant, meaning that electorates voted effectively independently.
 All data featured in this study has been contributed to the `eechidna` `R` package (available on CRAN).
@@
 From the available Census information aggregated at the electorate level, 50 socio-demographic variables were defined for each of the electorates to be used in the analysis.
+These variables include information relating to electoral age distributions, income, education qualifications, employment industries and job types, religion, birthplace, household characteristics and relationships.
 
@@
 
-Following this process, electoral socio-demographic variables are available for each of the six elections and can be joined with their corresponding two-party preferred votes.
+Following this imputation process, electoral socio-demographic variables are available for each of the six elections and can be joined with their corresponding two-party preferred votes.
 Before choosing an appropriate model, two issues with the socio-demographic variables need to be addressed.
@@
 Each principal component is considered seperately.
-If a subset of variables have a large loading in a given component, and there is an intuitive reason as to why they should be grouped together, then this subset will be combined to become a factor.
-Six factors are created using this criteria.
-These are:
+If a subset of variables have a large loading (positve or negative) in a given component, and there is an intuitive reason as to why they should be grouped together, then this subset of variables will be combined to become a factor.
+Note that more than one factor can be deduced from a principal component (i.e.
+two non-overlapping subsets of variables), and that any variables not included in a factor are not discarded.
+
+Six factors are created using this criteria. These are:
 `Incomes` (median personal income, median household income, median family income);
@@
 `Education` (high school completions, undergraduate and postgraduate degrees, proportion of employed people working as professionals, proportion of jobs in finance, proportion of workers who are laborers, proportion of workers who work as a tradesperson, diploma and certificate qualifications).
+
 For each of these groupings, a factor is created by taking a weighted sum of the variables.
@@
 
-There are $p=33$ variables in the resultant set obtained from this procedure.
-One of the age brackets (`Pop_55_plus`) is removed from this set to avoid multicollinearity, because the other three age brackets are included.
-Hence, the final predictor set contains $p=32$ variables.
+The final predictor set contains $p=32$ variables[^3].
+
+[^3]: The factor creation procedure reduce the variable set to $p=33$, however one of the age brackets (`Pop_55_plus`) is not included as a variable to avoid multicollinearity, because the other three age brackets are included.
 
@@
 The spatial error model [@Anselin88] is chosen because it captures spatial heterogeneity by incorporating a spatially structured random effect vector [@LeSage2009].
-In this context, the random effect can be thought of as capturing the unobserved political climate in each electorate, where this climate is correlated with the climate in neighbouring electorates, under the assumption that the climate is independent of electoral socio-demographics.
+In this context, the random effect can be thought of as capturing the effect of any characteristics that neighbourhoods share that have not been addressed by the independent variables included in the model.
 
@@
 The spatial autoregressive coefficient $\rho$ is positive and significant in the 2001, 2004 and 2016 elections (Figure \@ref(fig:rhovis)).
-In these three elections, there is evidence to suggest that the political climate of an electorate appears to be affected by the attitudes of its neighbours, affecting the electorate's two-party preference.
+In these three elections, there is evidence to suggest that neighbours share some influential characteristics outside the explanatory variables.
 Conversely, in the other three elections, the spatial effect weakens to become insignificant (although still positive).
@@
 This effect is significant in every election aside from 2004 and 2010.
-Unemployment however, is not as influential.
+`Unemployment` however, is not as influential.
 In 2001 and 2004, electorates with higher unemployment align with Labor, but over time this shifts towards support for the Liberal party, culminating in a positive (insignificant) effect in 2016.
@@
 Family and household sizes have minimal influence, although electorates with more single parent households are linked with Labor support.
-Furthermore, the neighbourhood (spatial) effects are found to be positive in all elections and significant in 2001, 2004 and 2016, meaning that neighbouring party preference appears to be an influential factor in those years.
+Furthermore, the spatial effects are found to be positive in all elections and significant in 2001, 2004 and 2016, meaning that other characteristics that neighbours have in common appear to be influential in those years.
 
@@
 All raw data was obtained from the Australian Electoral Commission, the Australian Bureau of Statistics and the Australian Government.
+
+The authors would like to sincerely thank the editor and associate editor of the Australian & New Zealand Journal of Statistics, and the two anonymous reviewers for providing helpful comments and suggestions on earlier drafts of the manuscript.
+Additionally, the authors would like to thank Anthony Ebert, Heike Hofmann, Thomas Lumley, Ben Marwick, Carson Sievert, Mingzhu Sun, Dilini Talagala, Nicholas Tierney, Nathaniel Tomasetti, Earo Wang and Fang Zhou, all of whom have contributed to the `eechidna` package.
 
@@
 
-The authors would like to sincerely thank Anthony Ebert, Heike Hofmann, Thomas Lumley, Ben Marwick, Carson Sievert, Mingzhu Sun, Dilini Talagala, Nicholas Tierney, Nathaniel Tomasetti, Earo Wang and Fang Zhou, all of whom have contributed to the `eechidna` package.
-
 # References

### 2019-09-25 anzjs_elections.Rmd
incomplete sentence

@@
 The spatial autoregressive coefficient $\rho$ is positive and significant in the 2001, 2004 and 2016 elections (Figure \@ref(fig:rhovis)).
-In these three elections, there is evidence to suggest that neighbours share some influential characteristics outside the explanatory variables.
+In these three elections, there is evidence to suggest that neighbours share some influential characteristics outside the explanatory variables, which affect the two-party preferred vote.
 Conversely, in the other three elections, the spatial effect weakens to become insignificant (although still positive).

### 2019-09-30 anzjs_elections.Rmd
Typos and small improvements in wording

@@
 
-# Summary{-}
-
-We examine the relationships between electoral socio-demographic characteristics and two-party preference in the six Australian federal elections held between 2001 to 2016.
-Socio-demographic information is derived from the Australian Census, which occurs every five years.
+# Summary {-}
+
+We examine the relationships between electoral socio-demographic characteristics and two-party preferences in the six Australian federal elections held between 2001 to 2016.
+Socio-demographic information is derived from the Australian Census which occurs every five years.
 Since a Census is not directly available for each election, an imputation method is employed to estimate Census data for the electorates at the time of each election.
@@
 library(spdep)
-# remotes::install_github("thomasp85/patchwork")
 library(patchwork)
-#library(gridExtra)
-#library(grid)
 library(predictmeans)
@@
 At the same time, political power has moved back and forth between the two major parties.
-In the 2007 and 2010 federal elections, the Australian Labor Party (Labor) was victorious, whereas the 2001, 2004, 2013 and 2016 elections were won by the Liberal National coalition (Liberal).
+In the 2007 and 2010 federal elections, the Australian Labor Party (hereafter Labor) was victorious, whereas the 2001, 2004, 2013 and 2016 elections were won by the Liberal National coalition (hereafter Liberal).
 The two-party preferred vote, a measure of support between these two parties, fluctuated between 47.3% and 53.5% (in favour of the Liberal party) over this period.
-This study explores how electoral (aggregate) socio-demographic characteristics relate to two-party preference, and whether their effects have changed over time.
+This study explores how electoral (aggregate) socio-demographic characteristics relate to two-party preferences, and whether their effects have changed over time.
 
@@
 A Census is conducted by the Australian Bureau of Statistics (ABS) every five years, whereas federal elections, conducted by the AEC, usually occur every three years or so.
-The first problem addressed is that of constructing appropriate Census data for the 2004, 2007, 2010 and 2013 elections --- election years in which a Census does not occur.
+The first problem addressed is that of constructing appropriate Census data for the 2004, 2007, 2010 and 2013 elections --- election years in which a Census did not occur.
 The predominant approach in previous studies was to join voting outcomes to the nearest Census, without accounting for any temporal differences [see @DavisStimson98;@Stimson06;@Liao09;@Stimson09].
@@
 Instead of analyzing a single election in isolation, this paper employs a consistent model framework across six elections so that temporal changes in the effects of socio-demographics can be observed.
-Each federal election is modelled with a cross-sectional dataset, where each observation is one of the 150 electorates.This dataset consists of the two-party preferred vote (as the response variable) and a set of common socio-demographic variables (as the explanatory variables).<!--The use of a regression framework to examine these socio-political relationships over time is seemingly absent from previous Australian studies.
+Each federal election is modelled with a cross-sectional dataset, where each observation is one of the 150 electorates.
+This dataset consists of the two-party preferred vote (as the response variable) and a set of common socio-demographic variables (as the explanatory variables).<!--The use of a regression framework to examine these socio-political relationships over time is seemingly absent from previous Australian studies.
 It also appears that no study has attempted any type of statistical analysis of socio-demographics in conjunction with voter behaviour in Australia since 2007, making this paper distinctly different from previous work.--> To prepare these datasets, socio-demographic variables are first standardized, and then a principal component analysis is used to group variables into "factors".
 To account for the inherent spatial structure of the data, a spatial error model is then estimated for each election.
-In interpreting these models, is important to be mindful of the ecological fallacy.
-Insights are being drawn at the electorate level and cannot be inferred for another disaggregate level (this means that drivers of individual voter behaviour may vary from what is observed at the electorate level).
+In interpreting these models, it is important to be mindful of the ecological fallacy.
+Insights are being drawn at the electorate level and cannot be inferred for another disaggregate level (in particular, drivers of individual voter behaviour may vary from what is observed at the electorate level).
 
 The paper is organised as follows.
-Section \@ref(data) describes the data collection, joining and cleaning.
-These pre-processing steps and model details are discussed in Section \@ref(modelling).
-Section \@ref(results) describes the inference conducted to determine signficance of effects and how these change over time.
+Section \@ref(data) describes the data collection, joining and cleaning, while model details are discussed in Section \@ref(modelling).
+Section \@ref(results) describes the inference conducted to determine significance of effects and how these change over time.
 Section \@ref(conclusion) summarises the work.
@@
 The formats of these files differ over the years, making extracting the appropriate information a big task.
-The functions available in the `dplyr` [@dplyr] and `readxl` [@readxl] `R` packages are particularly useful, as they provide fast consistent tools for data manipulation and functions to import `.xlsx` files (respectively).
+The functions available in the `dplyr` [@dplyr] and `readxl` [@readxl] `R` packages are particularly useful, as they provide fast consistent tools for data manipulation and functions to import `.xlsx` files.
 The 2001 and 2006 Census data are published in a format where the information for each electorate is held in a separate document making it difficult to use the `dplyr` tools.
@@
 Between 2001 and 2016 there were six elections and four Censuses (see Figure \@ref(fig:timeline)).
-Electoral boundaries are redistributed regularly by the AEC, meaning that only in the years where both a Census and election occur are all boundaries likely to match --- the case for the 2001 and 2016 elections.
+Electoral boundaries are redistributed regularly by the AEC, meaning that only in the years where both a Census and an election occur are all boundaries likely to match --- the case for the 2001 and 2016 elections.
 Therefore, for the four elections between 2004 and 2013, both temporal and spatial differences in electorates need to be accounted for when joining the electoral two-party preferred vote with Census data.
@@
 The finest level of disaggregation available for Census data is the region classification called Statistical Area 1 (SA1).
-In 2016, Australia was divded into over 55,000 SA1s.
+In 2016, Australia was divided into over 55,000 SA1s.
 Consider each of these SA1 regions as a source zone, $s = 1,\dots,S$, for which socio-demographic information is available.
 For simplicity, let each source zone be wholly summarised by its centroid.
-A set of target zones, $t = 1,\dots,T$, are defined as regions for which information is to be imputed - these are the electoral boundaries for a particular election.
-This is described in the context of a single election, and a single neighbouring Census.
+A set of target zones, $t = 1,\dots,T$, are defined as regions for which information is to be imputed --- these are the electoral boundaries for a particular election.
+
 Take the example of the Melbourne Ports electorate from the 2013 federal election, illustrated in Figure \@ref(fig:melbports13).
@@
 $$
-  \hat{C}_{t,y_2} = \frac{y_3-y_2}{y_3-y_1} * \hat{C}_{t,y_1} +
-    \frac{y_2-y_1}{y_3-y_1} * \hat{C}_{t,y_3}.
+  \hat{C}_{t,y_2} = \frac{y_3-y_2}{y_3-y_1} \hat{C}_{t,y_1} +
+    \frac{y_2-y_1}{y_3-y_1} \hat{C}_{t,y_3}.
 $$
@@
 
-Continuing with the example of Melbourne Ports in the 2013 election, the estimate for a given Census variable in 2016, $\hat{C}_{MelbPorts, 2016}$ would be obtained by computing the weighted average of this variable amongst the SA1s within the purple region shown in Figure \@ref(fig:melbports13).
-This would be repeated with the 2011 Census SA1s to obtain $\hat{C}_{MelbPorts, 2011}$, from which the final estimate is given by
+Continuing with the example of Melbourne Ports in the 2013 election, the estimate for a given Census variable in 2016, $\hat{C}_{\text{MelbPorts}, 2016}$ would be obtained by computing the weighted average of this variable amongst the SA1s within the purple region shown in Figure \@ref(fig:melbports13).
+This would be repeated with the 2011 Census SA1s to obtain $\hat{C}_{\text{MelbPorts}, 2011}$, from which the final estimate is given by
 $$
@@
 
-There are only $N = 150$ observations (electorates) in each election and $p = 50$ socio-demographic variables in each cross-section, with many variables represent similar information about an electorate.
+There are only $N = 150$ observations (electorates) in each election and $p = 50$ socio-demographic variables in each cross-section, with many variables representing similar information about an electorate.
 Any model that uses all variables would face problems with multi-collinearity and over-fitting, which would likely lead to erroneous conclusions regarding variable significance.
@@
 To determine which variables covary, principal component analysis is used on a combined dataset of socio-demographic variables from all six elections[^2].
-The only variables exempted from the principal component analysis are the four variables representing age brackets (the proportion of the population aged 0-19 years old, 20-34 years old, 45-54 years old and 55 years plus), which are to be included in the model as separate variables by construction.
+The only variables exempted from the principal component analysis are the four variables representing age brackets (the proportion of the population aged 0--19 years old, 20--34 years old, 45--54 years old and 55 years plus), which are included in the model as separate variables.
 
@@
 Variables that have a large loading in a particular component are deemed to covary, with a loading with magnitude greater than 0.15 being considered large.
-Each principal component is considered seperately.
-If a subset of variables have a large loading (positve or negative) in a given component, and there is an intuitive reason as to why they should be grouped together, then this subset of variables will be combined to become a factor.
-Note that more than one factor can be deduced from a principal component (i.e.
-two non-overlapping subsets of variables), and that any variables not included in a factor are not discarded.
-
-Six factors are created using this criteria. These are:
+Each principal component is considered separately.
+If a subset of variables have large loadings (positive or negative) in a given component, and there is an intuitive reason as to why they should be grouped together, then this subset of variables will be combined to become a factor.
+Note that more than one factor can be deduced from a principal component (i.e.\ two non-overlapping subsets of variables), and that any variables not included in a factor are not discarded.
+
+Six factors are created using this approach. These are:
 `Incomes` (median personal income, median household income, median family income);
-`Unemployment` (unemployment rate, labor force participation rate);
+`Unemployment` (unemployment rate, labour force participation rate);
 `PropertyOwned` (proportion of dwellings that are owned, proportion of dwellings that are mortgages, proportion of dwellings that are rented, proportion of dwellings that classified as government housing);
@@
 `FamHouseSize` (average household size, ratio of people to families, incidence of single person households, incidence of households containing a couple with kids, incidence of households containing a couple without kids); and
-`Education` (high school completions, undergraduate and postgraduate degrees, proportion of employed people working as professionals, proportion of jobs in finance, proportion of workers who are laborers, proportion of workers who work as a tradesperson, diploma and certificate qualifications).
+`Education` (high school completions, undergraduate and postgraduate degrees, proportion of employed people working as professionals, proportion of jobs in finance, proportion of workers who are labourers, proportion of workers who work as a tradesperson, diploma and certificate qualifications).
 
@@
 
-[^3]: The factor creation procedure reduce the variable set to $p=33$, however one of the age brackets (`Pop_55_plus`) is not included as a variable to avoid multicollinearity, because the other three age brackets are included.
+[^3]: The factor creation procedure reduced the variable set to $p=33$, however one of the age brackets (`Pop_55_plus`) is not included as a variable to avoid multicollinearity, because the other three age brackets are included.
 
@@
 This would not allow for conclusions to be drawn about how the effects of these variables change over time.
-Furthermore, the panel approach is avoided because of how frequently electoral boundaries change, noting that electorates with the same name across elections are not guaranteed to represent the same geographical region.
+Furthermore, the panel approach is avoided because of how frequently electoral boundaries change, noting that electorates with the same name across elections are not guaranteed to represent the same geographic region.
 Therefore any fixed or random effects models would be difficult to estimate without implementing consistent boundaries, which would require further imputation (of voting information).
@@
   title = "Estimated spatial error model parameters (standard errors) for each of the six election years.\\vspace*{-0.4cm}",
-  #dep.var.labels = "Two-party preferred vote in favor of the Liberal party",
+  #dep.var.labels = "Two-party preferred vote in favour of the Liberal party",
   font.size = "scriptsize",
@@
 Larger migrant populations from Asia, the Middle East, South-Eastern Europe, the United Kingdom and elsewhere, are either associated with Labor support, or have no effect.
-Of these areas, only South-Eastern European populations significant affect party preference, with larger populations associating with Labor in 2013 and 2016 (row 2, Figure \@ref(fig:partresplot2)).
+Of these areas, only South-Eastern European populations significantly affect party preference, with larger populations associating with Labor in 2013 and 2016 (row 2, Figure \@ref(fig:partresplot2)).
 Speaking other languages (aside from English) however, appears to have a far stronger effect, as observed through the `OtherLanguage` variable.
 Electorates with more diverse speech are associated with higher support for the Liberal party from 2004 onwards, with this effect being significant in 2013 and 2016 row 7, Figure \@ref(fig:partresplot1)).
-Furthermore, none of the variables relating to religious beliefs aside from Christianity have a material effect in any election (this includes the Buddhist, Muslim, Jewish, areligious and Indigenous Australian populations).
-The association between Christian populations (`Christianity`) and the Liberal party steady increases over the years, becoming positive and significant in 2016.
+Furthermore, none of the variables relating to religious beliefs aside from Christianity has a material effect in any election (this includes the Buddhist, Muslim, Jewish, non-religious and Indigenous Australian populations).
+The association between Christian populations (`Christianity`) and the Liberal party steadily increases over the years, becoming positive and significant in 2016.
 
@@
 In 2001, 2004 and 2016, higher proportions of people that have recently (in the past five years) moved house (`DiffAddress`) increased electoral support for the Liberal party (see row 5 in Figure \@ref(fig:partresplot1)).
-Having controlled for characteristics of house ownership and rental prices (`PropertyOwned` and `RentLoan` - both of which have no effect), this effect is somewhat surprising.
+Having controlled for characteristics of house ownership and rental prices (`PropertyOwned` and `RentLoan`,  both of which have no effect), this effect is somewhat surprising.
 
@@
 ### Checks
+
 Three robustness checks are conducted to confirm model stability.
-First, a model for each election is refit using only the variables that are found to be significant in at least one of the six elections.
-The estimated coefficients of the variables in the refitted models all fall within their respective $95\%$ confidence intervals from the full models.
+First, a model for each election is re-estimated using only the variables that are found to be significant in at least one of the six elections.
+The estimated coefficients of the variables in the re-estimated models all fall within their respective 95% confidence intervals from the full models.
 The second check involves the ten largest pairwise correlations.
-For each pair, a model for each election is refit omitting one of the two variables.
-It is found that for each of these pairs, the estimated effect of the remaining variable in the reduced model lies within the $95/%$ confidence interval from the full model.
-The final check is a visual exploration of different variable projections using a tour [@tourr], for each election.
+For each pair, a model for each election is re-estimated omitting one of the two variables.
+It is found that for each of these pairs, the estimated effect of the remaining variable in the reduced model lies within the 95% confidence interval from the full model.
+The final check is a visual exploration of different variable projections using a tour [@tourr] for each election.
 No definitive signs of multicollinearity are observed, and as expected (given the nature of spatial data), there is some clumping of electorates for certain projections.
@@
 
-Lingiari, an electorate taking up almost all of the Northern Territory, is an outlier in the all but the 2013 election due to its large Indigenous population, young age profile and low rates of property ownership.
+Lingiari, an electorate taking up almost all of the Northern Territory, is an outlier in all but the 2013 election due to its large Indigenous population, young age profile and low rates of property ownership.
 Fowler (NSW) has a diverse population with a high proportion of migrants, many Buddhists and Muslims, and has strong Labor support, making it influential in 2001, 2004 and 2010.

### 2019-10-01 anzjs_elections.tex
re-committing Rob's edits (without change) so that Jeremy can pull

@@
 
-We examine the relationships between electoral socio-demographic characteristics and two-party preference in the six Australian federal elections held between 2001 to 2016.
-Socio-demographic information is derived from the Australian Census, which occurs every five years.
-Since a Census is not directly available for each election, spatio-temporal imputation is employed to estimate Census data for the electorates at the time of each election.
+We examine the relationships between electoral socio-demographic characteristics and two-party preferences in the six Australian federal elections held between 2001 to 2016.
+Socio-demographic information is derived from the Australian Census which occurs every five years.
+Since a Census is not directly available for each election, an imputation method is employed to estimate Census data for the electorates at the time of each election.
 This accounts for both spatial and temporal changes in electoral characteristics between Censuses.
-To capture any spatial heterogeneity, a spatial error model is estimated for each election, which incorporates a spatially structured random effect vector that can be thought of as the unobserved political climate in each electorate.
-Over time, the impact of most socio-demographic characteristics that affect electoral two-party preference do not vary, with industry of work, incomes, household mobility and de facto relationships having strong effects in each of the six elections.
+To capture any spatial heterogeneity, a spatial error model is estimated for each election, which incorporates a spatially structured random effect vector.
+Over time, the impact of most socio-demographic characteristics that affect electoral two-party preference do not vary, with age distribution, industry of work, incomes, household mobility and de facto relationships having strong effects in each of the six elections.
 Education and unemployment are amongst those that have varying effects.
-It is also found that between 2004 and 2013, the spatial effect was insignificant, meaning that electorates voted effectively independently.
 All data featured in this study has been contributed to the \texttt{eechidna} \texttt{R} package (available on CRAN).
@@
 At the same time, political power has moved back and forth between the two major parties.
-In the 2007 and 2010 federal elections, the Australian Labor Party (Labor) was victorious, whereas the 2001, 2004, 2013 and 2016 elections were won by the Liberal National coalition (Liberal).
+In the 2007 and 2010 federal elections, the Australian Labor Party (hereafter Labor) was victorious, whereas the 2001, 2004, 2013 and 2016 elections were won by the Liberal National coalition (hereafter Liberal).
 The two-party preferred vote, a measure of support between these two parties, fluctuated between 47.3\% and 53.5\% (in favour of the Liberal party) over this period.
-This study explores how electoral (aggregate) socio-demographic characteristics relate to two-party preference, and whether their effects have changed over time.
+This study explores how electoral (aggregate) socio-demographic characteristics relate to two-party preferences, and whether their effects have changed over time.
 
@@
 A Census is conducted by the Australian Bureau of Statistics (ABS) every five years, whereas federal elections, conducted by the AEC, usually occur every three years or so.
-The first problem addressed is that of constructing appropriate Census data for the 2004, 2007, 2010 and 2013 elections --- election years in which a Census does not occur.
+The first problem addressed is that of constructing appropriate Census data for the 2004, 2007, 2010 and 2013 elections --- election years in which a Census did not occur.
 The predominant approach in previous studies was to join voting outcomes to the nearest Census, without accounting for any temporal differences \citep[see][]{DavisStimson98, Stimson06, Liao09, Stimson09}.
@@
 Instead of analyzing a single election in isolation, this paper employs a consistent model framework across six elections so that temporal changes in the effects of socio-demographics can be observed.
-Each federal election is modelled with a cross-sectional dataset, where each observation is one of the 150 electorates.This dataset consists of the two-party preferred vote (as the response variable) and a set of common socio-demographic variables (as the explanatory variables).
+Each federal election is modelled with a cross-sectional dataset, where each observation is one of the 150 electorates.
+This dataset consists of the two-party preferred vote (as the response variable) and a set of common socio-demographic variables (as the explanatory variables).
 To prepare these datasets, socio-demographic variables are first standardized, and then a principal component analysis is used to group variables into ``factors''.
 To account for the inherent spatial structure of the data, a spatial error model is then estimated for each election.
-In interpreting these models, is important to be mindful of the ecological fallacy.
-Insights are being drawn at the electorate level and cannot be inferred for another disaggregate level (this means that drivers of individual voter behaviour may vary from what is observed at the electorate level).
+In interpreting these models, it is important to be mindful of the ecological fallacy.
+Insights are being drawn at the electorate level and cannot be inferred for another disaggregate level (in particular, drivers of individual voter behaviour may vary from what is observed at the electorate level).
 
 The paper is organised as follows.
-Section \ref{data} describes the data collection, joining and cleaning.
-These pre-processing steps and model details are discussed in Section \ref{modelling}.
-Section \ref{results} describes the inference conducted to determine signficance of effects and how these change over time.
+Section \ref{data} describes the data collection, joining and cleaning, while model details are discussed in Section \ref{modelling}.
+Section \ref{results} describes the inference conducted to determine significance of effects and how these change over time.
 Section \ref{conclusion} summarises the work.
@@
 The formats of these files differ over the years, making extracting the appropriate information a big task.
-The functions available in the \texttt{dplyr} \citep{dplyr} and \texttt{readxl} \citep{readxl} \texttt{R} packages are particularly useful, as they provide fast consistent tools for data manipulation and functions to import \texttt{.xlsx} files (respectively).
+The functions available in the \texttt{dplyr} \citep{dplyr} and \texttt{readxl} \citep{readxl} \texttt{R} packages are particularly useful, as they provide fast consistent tools for data manipulation and functions to import \texttt{.xlsx} files.
 The 2001 and 2006 Census data are published in a format where the information for each electorate is held in a separate document making it difficult to use the \texttt{dplyr} tools.
@@
 Between 2001 and 2016 there were six elections and four Censuses (see Figure \ref{fig:timeline}).
-Electoral boundaries are redistributed regularly by the AEC, meaning that only in the years where both a Census and election occur are all boundaries likely to match --- the case for the 2001 and 2016 elections.
+Electoral boundaries are redistributed regularly by the AEC, meaning that only in the years where both a Census and an election occur are all boundaries likely to match --- the case for the 2001 and 2016 elections.
 Therefore, for the four elections between 2004 and 2013, both temporal and spatial differences in electorates need to be accounted for when joining the electoral two-party preferred vote with Census data.
@@
 The finest level of disaggregation available for Census data is the region classification called Statistical Area 1 (SA1).
-In 2016, Australia was divded into over 55,000 SA1s.
+In 2016, Australia was divided into over 55,000 SA1s.
 Consider each of these SA1 regions as a source zone, \(s = 1,\dots,S\), for which socio-demographic information is available.
 For simplicity, let each source zone be wholly summarised by its centroid.
-A set of target zones, \(t = 1,\dots,T\), are defined as regions for which information is to be imputed - these are the electoral boundaries for a particular election.
-This is described in the context of a single election, and a single neighbouring Census.
+A set of target zones, \(t = 1,\dots,T\), are defined as regions for which information is to be imputed --- these are the electoral boundaries for a particular election.
+
 Take the example of the Melbourne Ports electorate from the 2013 federal election, illustrated in Figure \ref{fig:melbports13}.
@@
 \[
-  \hat{C}_{t,y_2} = \frac{y_3-y_2}{y_3-y_1} * \hat{C}_{t,y_1} +
-    \frac{y_2-y_1}{y_3-y_1} * \hat{C}_{t,y_3}.
+  \hat{C}_{t,y_2} = \frac{y_3-y_2}{y_3-y_1} \hat{C}_{t,y_1} +
+    \frac{y_2-y_1}{y_3-y_1} \hat{C}_{t,y_3}.
 \]
@@
 
-Continuing with the example of Melbourne Ports in the 2013 election, the estimate for a given Census variable in 2016, \(\hat{C}_{MelbPorts, 2016}\) would be obtained by computing the weighted average of this variable amongst the SA1s within the purple region shown in Figure \ref{fig:melbports13}.
-This would be repeated with the 2011 Census SA1s to obtain \(\hat{C}_{MelbPorts, 2011}\), from which the final estimate is given by
+Continuing with the example of Melbourne Ports in the 2013 election, the estimate for a given Census variable in 2016, \(\hat{C}_{\text{MelbPorts}, 2016}\) would be obtained by computing the weighted average of this variable amongst the SA1s within the purple region shown in Figure \ref{fig:melbports13}.
+This would be repeated with the 2011 Census SA1s to obtain \(\hat{C}_{\text{MelbPorts}, 2011}\), from which the final estimate is given by
 \[
@@
 
-There are only \(N = 150\) observations (electorates) in each election and \(p = 50\) socio-demographic variables in each cross-section, with many variables represent similar information about an electorate.
+There are only \(N = 150\) observations (electorates) in each election and \(p = 50\) socio-demographic variables in each cross-section, with many variables representing similar information about an electorate.
 Any model that uses all variables would face problems with multi-collinearity and over-fitting, which would likely lead to erroneous conclusions regarding variable significance.
@@
 To determine which variables covary, principal component analysis is used on a combined dataset of socio-demographic variables from all six elections\footnote{It is appropriate to compute principal components on a combined dataset of all six elections because when computed separately for each election, scree plots level off after four components and the loadings of the first four components are similar across the elections.}.
-The only variables exempted from the principal component analysis are the four variables representing age brackets (the proportion of the population aged 0-19 years old, 20-34 years old, 45-54 years old and 55 years plus), which are to be included in the model as separate variables by construction.
+The only variables exempted from the principal component analysis are the four variables representing age brackets (the proportion of the population aged 0--19 years old, 20--34 years old, 45--54 years old and 55 years plus), which are included in the model as separate variables.
 
@@
 Variables that have a large loading in a particular component are deemed to covary, with a loading with magnitude greater than 0.15 being considered large.
-Each principal component is considered seperately.
-If a subset of variables have a large loading in a given component, and there is an intuitive reason as to why they should be grouped together, then this subset will be combined to become a factor.
-Six factors are created using this criteria.
-These are:
+Each principal component is considered separately.
+If a subset of variables have large loadings (positive or negative) in a given component, and there is an intuitive reason as to why they should be grouped together, then this subset of variables will be combined to become a factor.
+Note that more than one factor can be deduced from a principal component (i.e.~two non-overlapping subsets of variables), and that any variables not included in a factor are not discarded.
+
+Six factors are created using this approach. These are:
 \texttt{Incomes} (median personal income, median household income, median family income);
-\texttt{Unemployment} (unemployment rate, labor force participation rate);
+\texttt{Unemployment} (unemployment rate, labour force participation rate);
 \texttt{PropertyOwned} (proportion of dwellings that are owned, proportion of dwellings that are mortgages, proportion of dwellings that are rented, proportion of dwellings that classified as government housing);
@@
 \texttt{FamHouseSize} (average household size, ratio of people to families, incidence of single person households, incidence of households containing a couple with kids, incidence of households containing a couple without kids); and
-\texttt{Education} (high school completions, university qualifications, proportion of people working as professionals, proportion of jobs in finance, proportion of workers who are laborers, proportion of workers who work as a tradesperson, diploma and certificate qualifications).
+\texttt{Education} (high school completions, undergraduate and postgraduate degrees, proportion of employed people working as professionals, proportion of jobs in finance, proportion of workers who are labourers, proportion of workers who work as a tradesperson, diploma and certificate qualifications).
+
 For each of these groupings, a factor is created by taking a weighted sum of the variables.
@@
 
-There are \(p=33\) variables in the resultant set obtained from this procedure.
-One of the age brackets (\texttt{Pop\_55\_plus}) is removed from this set to avoid multicollinearity, because the other three age brackets are included.
-Hence, the final predictor set contains \(p=32\) variables.
+The final predictor set contains \(p=32\) variables\footnote{Mean of all variables aside from Judaism, Indigenous, Islam and Buddhism, where it assumes the mean of the log value.} which are listed in Table \ref{tab:printmodels}.
 
@@
 The spatial error model \citep{Anselin88} is chosen because it captures spatial heterogeneity by incorporating a spatially structured random effect vector \citep{LeSage2009}.
-In this context, the random effect can be thought of as capturing the unobserved political climate in each electorate, where this climate is correlated with the climate in neighbouring electorates, under the assumption that the climate is independent of electoral socio-demographics.
+In this context, the random effect can be thought of as capturing the effect of any characteristics that neighbourhoods share that have not been addressed by the independent variables included in the model.
 
@@
 The spatial autoregressive coefficient \(\rho\) is positive and significant in the 2001, 2004 and 2016 elections (Figure \ref{fig:rhovis}).
-In these three elections, there is evidence to suggest that the political climate of an electorate appears to be affected by the attitudes of its neighbours, affecting the electorate's two-party preference.
+In these three elections, there is evidence to suggest that neighbours share some influential characteristics outside the explanatory variables, which affect the two-party preferred vote.
 Conversely, in the other three elections, the spatial effect weakens to become insignificant (although still positive).
@@
 
-\textbackslash begin\{figure\}{[}h{]}
+\begin{figure}[h]
 
-\{\centering \includegraphics{anzjs_elections_files/figure-latex/partresplot1-1}
+{\centering \includegraphics{anzjs_elections_files/figure-latex/partresplot1-1}
 
-\}
+}
 
-\textbackslash caption\{Partial residual plots by election year for a selection of predictors.
+\caption{Partial residual plots by election year for a selection of predictors.
 Linear model with 95\% confidence bands overlaid.
 Most predictors have a positive relationship: the larger the value the more likely the electorate preferences the Coalition.
-The relationship is relatively robust over time, with the exception of OtherLanguageHome and Pop\_00\_19.\}\label{fig:partresplot1}
-\textbackslash end\{figure\}
+The relationship is relatively robust over time, with the exception of \texttt{OtherLanguageHome} and \texttt{Pop\_00\_19}.}\label{fig:partresplot1}
+\end{figure}
 
@@
 Several predictors have a negative relationship: with larger values indicating the electorate more likely preferences Labor.
-Most relationships are relatively stable over elections, except Unemployment and Education.}\label{fig:partresplot2}
+Most relationships are relatively stable over elections, except \texttt{Unemployment} and \texttt{Education}."}\label{fig:partresplot2}
 \end{figure}
@@
 This effect is significant in every election aside from 2004 and 2010.
-Unemployment however, is not as influential.
+\texttt{Unemployment} however, is not as influential.
 In 2001 and 2004, electorates with higher unemployment align with Labor, but over time this shifts towards support for the Liberal party, culminating in a positive (insignificant) effect in 2016.
@@
 Larger migrant populations from Asia, the Middle East, South-Eastern Europe, the United Kingdom and elsewhere, are either associated with Labor support, or have no effect.
-Of these areas, only South-Eastern European populations significant affect party preference, with larger populations associating with Labor in 2013 and 2016 (row 2, Figure \ref{fig:partresplot2}).
+Of these areas, only South-Eastern European populations significantly affect party preference, with larger populations associating with Labor in 2013 and 2016 (row 2, Figure \ref{fig:partresplot2}).
 Speaking other languages (aside from English) however, appears to have a far stronger effect, as observed through the \texttt{OtherLanguage} variable.
 Electorates with more diverse speech are associated with higher support for the Liberal party from 2004 onwards, with this effect being significant in 2013 and 2016 row 7, Figure \ref{fig:partresplot1}).
-Furthermore, none of the variables relating to religious beliefs aside from Christianity have a material effect in any election (this includes the Buddhist, Muslim, Jewish, areligious and Indigenous Australian populations).
-The association between Christian populations (\texttt{Christianity}) and the Liberal party steady increases over the years, becoming positive and significant in 2016.
+Furthermore, none of the variables relating to religious beliefs aside from Christianity has a material effect in any election (this includes the Buddhist, Muslim, Jewish, non-religious and Indigenous Australian populations).
+The association between Christian populations (\texttt{Christianity}) and the Liberal party steadily increases over the years, becoming positive and significant in 2016.
 
@@
 In 2001, 2004 and 2016, higher proportions of people that have recently (in the past five years) moved house (\texttt{DiffAddress}) increased electoral support for the Liberal party (see row 5 in Figure \ref{fig:partresplot1}).
-Having controlled for characteristics of house ownership and rental prices (\texttt{PropertyOwned} and \texttt{RentLoan} - both of which have no effect), this effect is somewhat surprising.
+Having controlled for characteristics of house ownership and rental prices (\texttt{PropertyOwned} and \texttt{RentLoan}, both of which have no effect), this effect is somewhat surprising.
 
@@
 In 2001 and 2004 neither of these variables are significant (see rows 4 and 5 in Figure \ref{fig:partresplot2}).
-
-\hypertarget{robustness}{%
-\subsubsection{Robustness}\label{robustness}}
-
-Two robustness checks are conducted to confirm model stability.
-First, a model for each election is fit using only the variables that are found to be significant in at least one of the six elections.
-Let this model be the reduced model, and the original model be the full model.
-The estimated coefficients of the variables in the reduced models all fall within their respective \(95\%\) confidence intervals from the full models, confirming that the significant effects are robust to variable selection.
-
-The second check involves computing pairwise correlations amongst the explanatory variables, and seeing if the removal of one variable (from a pair) affects the other's estimated effect.
-For example, the variables \texttt{Born\_Asia} and \texttt{BornElsewhere} have a pairwise correlation of \(0.93\) - the largest of any pair.
-A model is fit omitting \texttt{BornElsewhere} (the reduced model).
-The estimated coefficent of \texttt{Born\_Asia} is then compared with the confidence interval for \texttt{Born\_Asia} in the full model.
-It is found that for each of the top ten pairwise correlations, the estimated effect in the reduced model lies within the confidence interval from the full model.
-This second check indicates that variables with the highest pairwise correlation are not masking the effects of each other.
 
@@
 
-\hypertarget{outlier-electorates}{%
-\subsubsection*{Outlier electorates}\label{outlier-electorates}}
-\addcontentsline{toc}{subsubsection}{Outlier electorates}
+\hypertarget{robustness}{%
+\subsection{Robustness}\label{robustness}}
+
+\hypertarget{checks}{%
+\subsubsection{Checks}\label{checks}}
+
+Three robustness checks are conducted to confirm model stability.
+First, a model for each election is re-estimated using only the variables that are found to be significant in at least one of the six elections.
+The estimated coefficients of the variables in the re-estimated models all fall within their respective 95\% confidence intervals from the full models.
+The second check involves the ten largest pairwise correlations.
+For each pair, a model for each election is re-estimated omitting one of the two variables.
+It is found that for each of these pairs, the estimated effect of the remaining variable in the reduced model lies within the 95\% confidence interval from the full model.
+The final check is a visual exploration of different variable projections using a tour \citep{tourr} for each election.
+No definitive signs of multicollinearity are observed, and as expected (given the nature of spatial data), there is some clumping of electorates for certain projections.
+
+\hypertarget{influential-and-outlier-electorates}{%
+\subsubsection*{Influential and outlier electorates}\label{influential-and-outlier-electorates}}
+\addcontentsline{toc}{subsubsection}{Influential and outlier electorates}
 
@@
 
+The authors would like to sincerely thank the editor and associate editor of the Australian \& New Zealand Journal of Statistics, and the two anonymous reviewers for providing helpful comments and suggestions on earlier drafts of the manuscript.
+Additionally, the authors would like to thank Anthony Ebert, Heike Hofmann, Thomas Lumley, Ben Marwick, Carson Sievert, Mingzhu Sun, Dilini Talagala, Nicholas Tierney, Nathaniel Tomasetti, Earo Wang and Fang Zhou, all of whom have contributed to the \texttt{eechidna} package.
+
 \hypertarget{software}{%
@@
 
-The authors would like to sincerely thank Anthony Ebert, Heike Hofmann, Thomas Lumley, Ben Marwick, Carson Sievert, Mingzhu Sun, Dilini Talagala, Nicholas Tierney, Nathaniel Tomasetti, Earo Wang and Fang Zhou, all of whom have contributed to the \texttt{eechidna} package.
-
 \renewcommand\refname{References}

### 2019-10-01 response_for_ANZJS_editor.Rmd

@@
 
-Dear XXXX,
+Dear Dr Holland,
 
-Thank you for your feedback on our paper titled 'XXXX'. You raised many valid points in your response, and we have refined our work to address the queries raised.
+Thank you for your helpful comments on our paper titled 'Spatial modelling of the two-party preferred vote in Australian federal elections: 2001--2016'.
 
-Below we have detailed the changes made, or our response to, each point raised in your feedback. This is split into two sections - editor's comments and associate editor's comments.
-
-If you have any further questions, please do not hestitate to reach out to the corresponding author, Jeremy Forbes (`jeremyforbes1995@gmail.com`) directly.
+We have revised our paper to address the issues raised. Below we have detailed the changes made and our responses to the points raised in your email and by the reviewers.
 
@@
 
-$ DONE $
-
-*This is a valid point, and we have revised the imputation method to adopt this suggestion.
-We now use Census data at the lowest level of aggregation available - the geographic region type of 'Statistical Area 1' - to obtain the approximation of electoral boundaries (see Section 2.2).
-Our original method was chosen because we believed that the imputation method would not materially affect results.
+*We now use Census data at the lowest level of aggregation available --- the geographic region type of 'Statistical Area 1' --- to approximate the electoral boundaries (see Section 2.2).
+Some spatial imputation is still required as the SA1 boundaries do not exactly match the electorate boundaries.
 With this change in imputation method, our results are effectively the same as the previous submission.*
-
--- We made the change this way, and the results are effectively the same.
 
@@
 
-*We agree that more attention could have been given to the ecological fallacy in our original submission.
-In response, we have expanded our remarks and moved them to the introduction (Section 1).
-We are explicit in stating that "drivers of individual voter behaviour may vary from what is observed at the electorate level", which may not have been clear enough in the original submission.*
+*We have expanded our remarks and moved them to the introduction (Section 1) to give them more prominence.
+We are explicit in stating that "drivers of individual voter behaviour may vary from what is observed at the electorate level".*
 
@@
 
-*The variable selection method does have multiple steps and in retrospect, we agree that more clarity was needed here.*
-
-*Firstly, we have refined the subset of variables considered for analysis in this paper to 50 Census variables.
+*We have now explained our approach to variable selection in more detail and made some additional changes to to provide greater clarity.
+Previously we started with 65 variables, but many of these variables were removed in the variable selection process because they provided additional granularity that we did not want to pursue.
+We have now refined the subset of variables considered for analysis in this paper to 50 Census variables.
 Census data is provided as a head count for categories for many different questions, and there are thousands of variables that could be constructed from the raw data.
 These 50 variables were chosen because they appear in previous studies of Australian elections (those mentioned in the Introduction).
-Previously we started with 65 variables, but many of these variables were removed in the variable selection process because they provided additional granularity that we did not want to pursue.
-These never should have been included to begin with - and we have removed them in our updated submission.
-For example, we had previously included variables relating to Christian denominations (Anglican, Catholic) in our 65 variables, but did not include them in the model.*
-
-*The variable Christianity represents the proportion of people identifying with any Christian denomination, which we have kept in the model.*
+We hope this is now better explained.*
 
@@
 This results in a final predictor set of 32 variables.
-Note that two variables no longer meet the criteria for being included in a factor, so they appear as separate variables - hence a final set of 32 variables.*
+Note that two variables no longer meet the criteria for being included in a factor, so they appear as separate variables - hence a final set of 32 variables which are listed in Table 1.*
 
@@
 When we fit a model using just the variables with significant effects, we find that the results do not change much.
-Each of the estimated coefficients from the reduced model (containing only the subset significant variables) lie within the 95% confidence interval from its corresponding full model (containing all 32 variables).
-*
+Each of the estimated coefficients from the reduced model (containing only the subset significant variables) lie within the 95% confidence interval from its corresponding full model (containing all 32 variables).*
 
@@
 
-$ DONE $
--- Changes --
+7. Other minor points
 
-Refitting without a variable. Doing this selectively on the ones that are more highly correlated. Re-fit with a different subset of variables.
+  * line 120,  each of the
+  * line 129, is it really the boundaries that are the target /source zones or the areas defined by these boundaries?
+  * line 241, must be modelled
+  * line 261-266, is there anything in the fact that the spatial coefficient is significant only in the years where no imputation was done?
 
-Another check is the TOUR.
-
-Bundle robustness and diagnostiscs - call it robustness:
-1) First para is on multicollinearity checks - don't need as much detail as I have
-2) Influential observations - couple paragraphs on cooks' dist and influence diagnostics
+The first three minor issues have been corrected. For the last issue, the spatial AR parameter is now significant in 2001, 2004 and 2016, which do not correspond only to years with no imputation.
 
@@
 
-*See items (2), (3) and (4) in section above.*
+*See response to the editor.*
 
@@
 *Partial residual plots are presented to give the reader a visual feel for the direction, size and significance of an estimated effect, as well as the distribution of the data points (for that covariate).
-Whilst this is a valid point as there is duplication of information here, we think that the partial residual plots play an important role in illustrating the effect of a given variable across elections, and that the estimated model parameters should be included for completeness.*
+Whilst there is duplication of information here, we think that the partial residual plots play an important role in illustrating the effect of a given variable across elections, and that the estimated model parameters should be included for completeness.*
 
@@
 *We have expanded on descriptions of both in our revised version.
-Time-varying restrictions are when one forces the effect of a variable to be invariant over time, for example  - forcing the estimated coefficient of the variable `Unemployed` to be the same in 2013 and 2016 elections (section 3.2).*
+Time-varying restrictions are when one forces the effect of a variable to be invariant over time, for example, forcing the estimated coefficient of the variable `Unemployed` to be the same in 2013 and 2016 elections (section 3.2).*
 
-*Partial residuals, for a variable X1, are the residuals from the fitted model with the estimated effect of X1 added to it.
+*Partial residuals, for a variable X1, are the residuals from the fitted model with the estimated effect of X1 added to each.
 Partial residual plots show the direction, size and significance of an estimated effect, as well as any deviations from linearity (section 4.3).*
@@
 
-*We agree that more attention could have been given to the ecological fallacy in our original submission.
-In response, we have expanded our remarks and moved them to the introduction (Section 1).
+*We have expanded our remarks on the ecological fallacy (previously in section 4.3) and moved them to the introduction (Section 1) to give them more prominence.
 We are explicit in stating that "drivers of individual voter behaviour may vary from what is observed at the electorate level", which may not have been clear enough in the original submission.*
@@
 *We formed our approach to exploring this question by reviewing the Australian literature in this area, which has been referenced accordingly.
-As this is an exploratory piece, and we wanted to let the relationships reveal themselves, rather than expicitly stating hypotheses to be tested.*
+As this is an exploratory piece, we wanted to let the relationships reveal themselves, rather than expicitly stating hypotheses to be tested.*
 
@@
 
-*Since this study is of electorates, rather than individual voters, changing an electorate's demographics can have an effect on two-party preference beyond just the head count of population who's demographics are changed.
-All this means is that effect of de facto relationships on the electorate's vote is not necessarily just as a result of the votes from people who are in a de facto relationship.*
+*Since this study is of electorates, rather than individual voters, changing an electorate's demographics can have an effect on two-party preference beyond just the head count of population whose demographics are changed.
+In particular, the effect of de facto relationships on the electorate's vote is not necessarily just as a result of the votes from people who are in a de facto relationship.*
 
@@
 *We have revised our imputation method to reflect this suggestion.
-We now use Census data at the lowest level of aggregation available - the geographic region type of 'Statistical Area 1' - to obtain the approximation of electoral boundaries (see Section 2.2).
-Our original method was chosen because we believed that the imputation method would not materially affect results.
+We now use Census data at the lowest level of aggregation available --- the geographic region type of 'Statistical Area 1' --- to obtain the approximation of electoral boundaries (see Section 2.2).
+Some spatial imputation is still required as the SA1 boundaries do not exactly match the electorate boundaries.
 With this change in imputation method, our results are effectively the same as the previous submission.*
@@
 
-*The 50 Census variables are chosen because they appear in previous studies of Australian elections (those mentioned in the Introduction), and we believe that they capture key characteristics of an electorate - age, income, employment, religion, birthplace and relationships.*
+*The (now) 50 Census variables are chosen because they appear in previous studies of Australian elections (those mentioned in the Introduction), and we believe that they capture key characteristics of an electorate - age, income, employment, religion, birthplace and relationships.*
 
 *Note that we have refined the subset of variables considered for analysis in this paper to 50 Census variables.
-Previously we started with 65 variables, but many of these variables were removed in the variable selection process because they provided additional granularity that we did not want to pursue.
-These never should have been included to begin with - and we have removed them in our updated submission.
-For example, we had previously included variables relating to Christian denominations (Anglican, Catholic) in our 65 variables, but did not include them in the model.*
+Previously we started with 65 variables, but many of these variables were removed in the variable selection process because they provided additional granularity that we did not want to pursue.*
 
@@
 The first four components are chosen because the amount of variance explained drops off after the fourth PC.
-The decision is made to omit these details of this type because they detract from the core steps of our variable selection procedure.*
+We have now explained our variable selection process in more detail.*
 
 *The factors are not the principal components themselves.
-Rather, the loadings in the first four principal components are used to determine which variables should be grouped together to generate a new variable in their place, which we have called a factor.
-*
+Rather, the loadings in the first four principal components are used to determine which variables should be grouped together to generate a new variable in their place, which we have called a factor.*
 
@@
 
-*This was a very helpful suggestion to explore further, and as such an additional section on robustness has been included in the revised paper (Section 4.4).
+*In response to this suggestion, we have now added a section on robustness in the revised paper (Section 4.4).
 When we fit a model using just the variables with significant effects, we find that the results do not change much.
-Each of the estimated coefficients from the reduced model (containing only the subset significant variables) lie within the 95% confidence interval from its corresponding full model (containing all 32 variables).
-*
+Each of the estimated coefficients from the reduced model (containing only the subset significant variables) lie within the 95% confidence interval from its corresponding full model (containing all 32 variables).*
 
@@
 Any variables that are not used in creating a factor still remain in the model.
-The point raised is valid - there is no guarantee about any of these factors being related to the response.*
+Of course, there is no guarantee about any of these factors being related to the response.
+We have now explained our variable selection procedure in more detail in the revised paper.*
 
@@
 *The first check involved fitting a model using just the variables with significant effects.
-We find that the estimated coefficients in the reduced model (containing only the subset significant variables) all lie within the 95% confidence interval from its corresponding full model (containing all 32 variables).
-*
+We find that the estimated coefficients in the reduced model (containing only the subset significant variables) all lie within the 95% confidence interval from its corresponding full model (containing all 32 variables).*
 

### 2019-10-01 anzjs_elections.Rmd
Updated response

@@
 
-The final predictor set contains $p=32$ variables[^3].
+The final predictor set contains $p=32$ variables[^3] which are listed in Table \ref{tab:printmodels}.
 
@@
 
-```{r partresplot1, fig.cap = "Partial residual plots by election year for a selection of predictors.
-Linear model with 95\\% confidence bands overlaid.
+(ref:figcappartresplot1) Partial residual plots by election year for a selection of predictors.
+Linear model with 95% confidence bands overlaid.
 Most predictors have a positive relationship: the larger the value the more likely the electorate preferences the Coalition.
-The relationship is relatively robust over time, with the exception of OtherLanguageHome and Pop_00_19.", fig.height = 8.5, fig.width = 6, warning = FALSE, fig.pos = 'h'}
+The relationship is relatively robust over time, with the exception of `OtherLanguageHome` and `Pop_00_19`.
+
+```{r partresplot1, fig.cap = "(ref:figcappartresplot1)", fig.height = 8.5, fig.width = 6, warning = FALSE}
 (p1 / p2 / p3 / (p4 + ylab("Two-party preferred vote (%)                     ")) / p5 / p6 / p7 / p8)
@@
 
-```{r partresplot2, fig.cap = "Partial residual plots by election year for a selection of predictors.
-Linear model with 95\\% confidence bands overlaid.
+(ref:figcappartresplot2) Partial residual plots by election year for a selection of predictors.
+Linear model with 95% confidence bands overlaid.
 Several predictors have a negative relationship: with larger values indicating the electorate more likely preferences Labor.
-Most relationships are relatively stable over elections, except Unemployment and Education.", fig.height = 8.5,  fig.width = 6, warning = FALSE, fig.pos = 'h'}
+Most relationships are relatively stable over elections, except `Unemployment` and `Education`."
+
+```{r partresplot2, fig.cap = "(ref:figcappartresplot2)", fig.height = 8.5,  fig.width = 6, warning = FALSE}
 (p9 / p10  / p11 / (p12 + ylab("Two-party preferred vote (%)                     ")) / p13 / p14 / p15 / p16)
@@
 incumbent <- function(fp) {
-  df <- fp %>% filter(HistoricElected == "Y", PartyAb == "LNP") %>% select(DivisionNm, HistoricElected)
-  return(df)
+  fp %>% 
+    filter(HistoricElected == "Y", PartyAb == "LNP") %>% 
+    select(DivisionNm, HistoricElected)
 }
@@
   theme_bw() +
-  labels(x = "Year", y = "Residuals")
-
+  xlab("Year") + ylab("Residuals")
 ```

### 2019-10-02 response_for_ANZJS_editor.Rmd
tweaking response

@@
 ---
-title: "Response For The Editor"
+title: "Response to Review"
 author: "Jeremy Forbes"
@@
 
-We have revised our paper to address the issues raised. Below we have detailed the changes made and our responses to the points raised in your email and by the reviewers.
+We have revised our paper to address the issues raised. Below are detailed summaries of the changes made in response to the points raised by you, the Associate Editor and by the reviewers.
 

### 2019-10-02 multicollinearity.r

@@ -2,6 +2,7 @@
 library(tourr)
 library(tidyverse)
 library(eechidna)
+source("functions.R")
 
 load("data/abs2004_cd.rda")
 load("data/abs2007_cd.rda")
@@ -128,4 +129,16 @@ d <- small_df %>% filter(year == "2016") %>%
   select(-DivisionNm, -LNP_Percent, -year)
 quartz()
 animate_xy(d, axes="bottomleft", half_range=1)
+
+# For blog post
+d1 <- small_df %>% filter(year == "2001") %>%
+  select(-DivisionNm, -LNP_Percent, -year, -Band_55plus) %>%
+  rename_all(function(.) paste0("V", 1:32))
+
+d2 <- small_df %>% filter(year == "2016") %>%
+  select(-DivisionNm, -LNP_Percent, -year, -Band_55plus) %>%
+  rename_all(function(.) paste0("V", 1:32))
+
+save(d1, file="d1.rda") 
+save(d2, file="d2.rda") 
 

### 2019-10-02 anzjs_elections.Rmd
Updated table 1

@@
   header = FALSE,
-  column.sep.width = "2pt",
+  column.sep.width = "1pt",
   no.space=TRUE,
@@
              "\\\\", sep="", collapse="")
-regtable <- c(regtable[1:77],sigmaline, regtable[78:length(regtable)])
+regtable <- c(regtable[1:79],sigmaline, regtable[80:length(regtable)])
 # Squeeze some more space
@@
 regtable[6] <- "\\begin{tabular}{@{\\extracolsep{1pt}}lrrrrrr}"
-regtable[12:75] <- gsub(")","\\\\rlap{)}",regtable[12:75])
-regtable[12:75] <- gsub("$^{*}$","\\rlap{$^{*}$}", regtable[12:75], fixed=TRUE)
-regtable[12:75] <- gsub("$^{**}$","\\rlap{$^{**}$}", regtable[12:75], fixed=TRUE)
-regtable[12:75] <- gsub("$^{***}$","\\rlap{$^{***}$}", regtable[12:75], fixed=TRUE)
+regtable[12:80] <- gsub(")","\\\\rlap{)}",regtable[12:80])
+regtable[12:80] <- gsub("$^{*}$","\\rlap{$^{*}$}", regtable[12:80], fixed=TRUE)
+regtable[12:80] <- gsub("$^{**}$","\\rlap{$^{**}$}", regtable[12:80], fixed=TRUE)
+regtable[12:80] <- gsub("$^{***}$","\\rlap{$^{***}$}", regtable[12:80], fixed=TRUE)
 # Column spacing
-regtable[5] <- "\\scriptsize\\tabcolsep=0.35cm"
+regtable[5] <- "\\fontsize{7}{8}\\rm\\tabcolsep=0.42cm"
 # Remove "Note"
-regtable[81] <- "& \\multicolumn{6}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\\\"
+regtable[85] <- "& \\multicolumn{6}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\\\"
 ```

### 2019-10-02 anzjs_elections.tex
Updated table 1

@@
 
-\begin{table}[!htbp]\vspace*{-0.4cm}\centering    \caption{Estimated spatial error model parameters (standard errors) for each of the six election years.\vspace*{-0.4cm}}    \label{tab:printmodels}  \scriptsize\tabcolsep=0.35cm \begin{tabular}{@{\extracolsep{1pt}}lrrrrrr} \\[-1.8ex]\hline  \hline \\[-1.8ex]  \\[-1.8ex] & \multicolumn{6}{c}{\vspace{-0.4cm}} \\   & 2001 & 2004 & 2007 & 2010 & 2013 & 2016 \\  \hline \\[-1.8ex]  $\rho$ & 0.53\rlap{$^{***}$} & 0.33\rlap{$^{**}$} & 0.21 & 0.17 & 0.27 & 0.39\rlap{$^{**}$}\\  & (0.15\rlap{)} & (0.16\rlap{)} & (0.18\rlap{)} & (0.17\rlap{)} & (0.17\rlap{)} & (0.17\rlap{)}\\  AusCitizen & $-$3.94\rlap{$^{*}$} & $-$1.39 & $-$2.18 & $-$1.28 & $-$3.89 & $-$2.66 \\    & (2.27\rlap{)} & (2.44\rlap{)} & (2.21\rlap{)} & (2.69\rlap{)} & (2.51\rlap{)} & (2.61\rlap{)} \\    Pop\_00\_19 & 0.49 & 2.66 & 9.39\rlap{$^{***}$} & 5.25 & 3.31 & 0.88 \\    & (2.54\rlap{)} & (3.91\rlap{)} & (3.63\rlap{)} & (3.64\rlap{)} & (2.91\rlap{)} & (2.62\rlap{)} \\    Pop\_20\_34 & $-$8.04\rlap{$^{***}$} & $-$7.72\rlap{$^{***}$} & $-$8.34\rlap{$^{***}$} & $-$11.68\rlap{$^{***}$} & $-$9.29\rlap{$^{***}$} & $-$9.21\rlap{$^{***}$} \\    & (1.80\rlap{)} & (2.21\rlap{)} & (2.18\rlap{)} & (2.90\rlap{)} & (2.62\rlap{)} & (2.37\rlap{)} \\    Pop\_35\_54 & $-$2.64\rlap{$^{***}$} & $-$2.78\rlap{$^{***}$} & $-$3.62\rlap{$^{***}$} & $-$3.13\rlap{$^{***}$} & $-$2.76\rlap{$^{**}$} & $-$2.13\rlap{$^{**}$} \\    & (0.84\rlap{)} & (0.89\rlap{)} & (0.83\rlap{)} & (1.10\rlap{)} & (1.11\rlap{)} & (1.06\rlap{)} \\    BornAsia & 3.58\rlap{$^{*}$} & $-$1.09 & 0.66 & $-$1.78 & $-$1.08 & $-$0.14 \\    & (2.09\rlap{)} & (2.52\rlap{)} & (1.99\rlap{)} & (2.74\rlap{)} & (2.54\rlap{)} & (2.17\rlap{)} \\    BornMidEast & $-$1.02 & $-$1.75 & $-$0.98 & $-$1.00 & $-$1.66 & $-$1.31 \\    & (1.00\rlap{)} & (1.17\rlap{)} & (1.09\rlap{)} & (1.33\rlap{)} & (1.23\rlap{)} & (1.11\rlap{)} \\    BornSEEuro & $-$1.63 & $-$3.17\rlap{$^{*}$} & $-$1.07 & $-$2.04 & $-$2.89\rlap{$^{***}$} & $-$2.53\rlap{$^{***}$} \\    & (1.37\rlap{)} & (1.68\rlap{)} & (1.06\rlap{)} & (1.29\rlap{)} & (1.11\rlap{)} & (0.97\rlap{)} \\    BornUK & 0.29 & 0.31 & 0.32 & 0.28 & $-$0.15 & $-$0.61 \\    & (1.02\rlap{)} & (1.04\rlap{)} & (0.87\rlap{)} & (1.06\rlap{)} & (0.99\rlap{)} & (0.99\rlap{)} \\    BornElsewhere & $-$4.13 & $-$1.51 & $-$1.03 & 2.45 & $-$4.21 & $-$2.17 \\    & (3.14\rlap{)} & (3.62\rlap{)} & (3.18\rlap{)} & (4.13\rlap{)} & (3.90\rlap{)} & (3.76\rlap{)} \\    Buddhism & $-$0.07 & 0.80 & 0.58 & $-$0.14 & $-$0.43 & $-$1.16 \\    & (1.31\rlap{)} & (1.54\rlap{)} & (1.39\rlap{)} & (1.66\rlap{)} & (1.60\rlap{)} & (1.58\rlap{)} \\    Christianity & $-$1.70 & $-$1.01 & $-$0.45 & 0.13 & 2.03 & 3.76\rlap{$^{**}$} \\    & (1.62\rlap{)} & (1.75\rlap{)} & (1.60\rlap{)} & (1.85\rlap{)} & (1.68\rlap{)} & (1.83\rlap{)} \\    CurrentlyStudying & $-$2.20\rlap{$^{*}$} & $-$0.01 & $-$0.14 & 1.35 & 0.32 & 0.22 \\    & (1.22\rlap{)} & (1.50\rlap{)} & (1.39\rlap{)} & (1.41\rlap{)} & (1.35\rlap{)} & (1.56\rlap{)} \\    DeFacto & $-$3.24 & $-$2.25 & $-$4.67\rlap{$^{**}$} & $-$7.75\rlap{$^{**}$} & $-$7.82\rlap{$^{**}$} & $-$10.39\rlap{$^{***}$} \\    & (2.07\rlap{)} & (2.62\rlap{)} & (2.27\rlap{)} & (3.09\rlap{)} & (3.08\rlap{)} & (3.15\rlap{)} \\    DiffAddress & 3.06\rlap{$^{***}$} & 2.75\rlap{$^{**}$} & 0.73 & 2.55 & 2.27 & 5.20\rlap{$^{***}$} \\    & (0.94\rlap{)} & (1.20\rlap{)} & (1.24\rlap{)} & (1.79\rlap{)} & (1.67\rlap{)} & (1.51\rlap{)} \\    Distributive & 1.60 & 1.89\rlap{$^{*}$} & 0.50 & 0.62 & 1.59 & 1.31 \\    & (1.06\rlap{)} & (1.14\rlap{)} & (0.99\rlap{)} & (1.27\rlap{)} & (1.20\rlap{)} & (1.18\rlap{)} \\    Education & $-$0.37 & $-$0.26 & $-$6.72\rlap{$^{**}$} & $-$7.31\rlap{$^{*}$} & $-$7.31\rlap{$^{**}$} & $-$8.55\rlap{$^{**}$} \\    & (2.35\rlap{)} & (3.34\rlap{)} & (3.00\rlap{)} & (3.90\rlap{)} & (3.63\rlap{)} & (3.37\rlap{)} \\    Extractive & 3.74\rlap{$^{***}$} & 4.96\rlap{$^{***}$} & 4.64\rlap{$^{***}$} & 6.46\rlap{$^{***}$} & 5.97\rlap{$^{***}$} & 6.38\rlap{$^{***}$} \\    & (1.43\rlap{)} & (1.47\rlap{)} & (1.20\rlap{)} & (1.45\rlap{)} & (1.35\rlap{)} & (1.38\rlap{)} \\    FamHouseSize & 1.94 & $-$2.55 & $-$6.47\rlap{$^{**}$} & $-$3.84 & $-$3.12 & $-$2.00 \\    & (2.61\rlap{)} & (3.66\rlap{)} & (3.28\rlap{)} & (3.87\rlap{)} & (3.52\rlap{)} & (3.06\rlap{)} \\    Incomes & 4.36\rlap{$^{***}$} & 2.42 & 5.52\rlap{$^{**}$} & 5.63\rlap{$^{*}$} & 8.02\rlap{$^{***}$} & 12.70\rlap{$^{***}$} \\    & (1.69\rlap{)} & (3.00\rlap{)} & (2.42\rlap{)} & (3.15\rlap{)} & (2.78\rlap{)} & (2.64\rlap{)} \\    Indigenous & 1.26 & 1.96 & 2.41 & 2.38 & 0.46 & $-$0.22 \\    & (1.61\rlap{)} & (1.89\rlap{)} & (1.59\rlap{)} & (2.00\rlap{)} & (1.88\rlap{)} & (1.90\rlap{)} \\    Islam & $-$0.75 & $-$0.91 & $-$0.60 & $-$2.01 & $-$0.88 & $-$1.09 \\    & (1.14\rlap{)} & (1.28\rlap{)} & (1.14\rlap{)} & (1.41\rlap{)} & (1.26\rlap{)} & (1.30\rlap{)} \\    Judaism & 1.32 & 0.93 & 1.47 & 0.28 & 1.35 & 1.15 \\    & (1.01\rlap{)} & (1.08\rlap{)} & (0.92\rlap{)} & (1.10\rlap{)} & (1.02\rlap{)} & (0.97\rlap{)} \\    ManagerAdmin & 2.62\rlap{$^{***}$} & 4.67\rlap{$^{***}$} & 7.47\rlap{$^{***}$} & 7.05\rlap{$^{***}$} & 5.93\rlap{$^{***}$} & 5.64\rlap{$^{***}$} \\    & (0.67\rlap{)} & (1.06\rlap{)} & (0.95\rlap{)} & (1.16\rlap{)} & (1.06\rlap{)} & (0.97\rlap{)} \\    Married & $-$3.93 & $-$2.72 & $-$9.35\rlap{$^{***}$} & $-$10.12\rlap{$^{***}$} & $-$7.91\rlap{$^{**}$} & $-$9.47\rlap{$^{**}$} \\    & (2.51\rlap{)} & (3.56\rlap{)} & (3.12\rlap{)} & (3.55\rlap{)} & (3.57\rlap{)} & (3.85\rlap{)} \\    NoReligion & $-$0.73 & 0.04 & 1.32 & 0.37 & 1.41 & 2.94 \\    & (1.50\rlap{)} & (1.65\rlap{)} & (1.51\rlap{)} & (1.75\rlap{)} & (1.74\rlap{)} & (2.03\rlap{)} \\    OneParentHouse & $-$4.77\rlap{$^{***}$} & $-$3.23 & $-$6.55\rlap{$^{***}$} & $-$7.03\rlap{$^{***}$} & $-$5.32\rlap{$^{***}$} & $-$4.94\rlap{$^{**}$} \\    & (1.49\rlap{)} & (1.99\rlap{)} & (1.81\rlap{)} & (2.04\rlap{)} & (1.97\rlap{)} & (2.03\rlap{)} \\    OtherLanguage & $-$1.02 & 6.88 & 6.21 & 7.80 & 10.13\rlap{$^{**}$} & 9.98\rlap{$^{**}$} \\    & (3.00\rlap{)} & (4.93\rlap{)} & (3.97\rlap{)} & (5.25\rlap{)} & (5.09\rlap{)} & (4.26\rlap{)} \\    PropertyOwned & $-$2.01 & $-$0.30 & 0.74 & $-$1.92 & $-$1.05 & 0.73 \\    & (1.35\rlap{)} & (1.49\rlap{)} & (1.36\rlap{)} & (1.74\rlap{)} & (1.67\rlap{)} & (1.48\rlap{)} \\    RentLoanPrice & $-$2.17 & 0.37 & 1.23 & 3.08 & 1.36 & $-$2.04 \\    & (1.46\rlap{)} & (1.93\rlap{)} & (1.76\rlap{)} & (2.23\rlap{)} & (2.20\rlap{)} & (2.07\rlap{)} \\    SocialServ & 3.31\rlap{$^{***}$} & 2.85\rlap{$^{**}$} & 3.46\rlap{$^{***}$} & 3.72\rlap{$^{**}$} & 2.98\rlap{$^{**}$} & 4.04\rlap{$^{***}$} \\    & (1.27\rlap{)} & (1.40\rlap{)} & (1.17\rlap{)} & (1.46\rlap{)} & (1.28\rlap{)} & (1.15\rlap{)} \\    Transformative & 2.30 & 4.71\rlap{$^{***}$} & 4.58\rlap{$^{***}$} & 4.55\rlap{$^{**}$} & 3.63\rlap{$^{**}$} & 4.05\rlap{$^{***}$} \\    & (1.48\rlap{)} & (1.77\rlap{)} & (1.51\rlap{)} & (1.87\rlap{)} & (1.67\rlap{)} & (1.47\rlap{)} \\    Unemployment & $-$3.39$^{**}$ & $-$3.47$^{**}$ & $-$0.40 & $-$0.68 & 0.81 & 1.93 \\    & (1.37) & (1.69) & (1.45) & (1.80) & (1.47) & (1.32) \\  Residual Standard Error (GLS) & 4.34 & 4.82 & 4.32 & 5.30 & 4.82 & 4.76\\   Constant & 50.80$^{***}$ & 52.63$^{***}$ & 47.31$^{***}$ & 49.92$^{***}$ & 53.52$^{***}$ & 50.46$^{***}$ \\    & (0.76) & (0.59) & (0.44) & (0.52) & (0.54) & (0.64) \\  & \multicolumn{6}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ Observations & 150 & 150 & 150 & 150 & 150 & 150 \\  \hline  \hline \\[-1.8ex]  \textit{Note:}  & \multicolumn{6}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\  \end{tabular}  \end{table}
+\begin{table}[!htbp]\vspace*{-0.4cm}\centering    \caption{Estimated spatial error model parameters (standard errors) for each of the six election years.\vspace*{-0.4cm}}    \label{tab:printmodels}  \fontsize{7}{8}\rm\tabcolsep=0.42cm \begin{tabular}{@{\extracolsep{1pt}}lrrrrrr} \\[-1.8ex]\hline  \hline \\[-1.8ex]  \\[-1.8ex] & \multicolumn{6}{c}{\vspace{-0.4cm}} \\   & 2001 & 2004 & 2007 & 2010 & 2013 & 2016 \\  \hline \\[-1.8ex]  $\rho$ & 0.53\rlap{$^{***}$} & 0.33\rlap{$^{**}$} & 0.21 & 0.17 & 0.27 & 0.39\rlap{$^{**}$}\\  & (0.15\rlap{)} & (0.16\rlap{)} & (0.18\rlap{)} & (0.17\rlap{)} & (0.17\rlap{)} & (0.17\rlap{)}\\  AusCitizen & $-$3.94\rlap{$^{*}$} & $-$1.39 & $-$2.18 & $-$1.28 & $-$3.89 & $-$2.66 \\    & (2.27\rlap{)} & (2.44\rlap{)} & (2.21\rlap{)} & (2.69\rlap{)} & (2.51\rlap{)} & (2.61\rlap{)} \\    Pop\_00\_19 & 0.49 & 2.66 & 9.39\rlap{$^{***}$} & 5.25 & 3.31 & 0.88 \\    & (2.54\rlap{)} & (3.91\rlap{)} & (3.63\rlap{)} & (3.64\rlap{)} & (2.91\rlap{)} & (2.62\rlap{)} \\    Pop\_20\_34 & $-$8.04\rlap{$^{***}$} & $-$7.72\rlap{$^{***}$} & $-$8.34\rlap{$^{***}$} & $-$11.68\rlap{$^{***}$} & $-$9.29\rlap{$^{***}$} & $-$9.21\rlap{$^{***}$} \\    & (1.80\rlap{)} & (2.21\rlap{)} & (2.18\rlap{)} & (2.90\rlap{)} & (2.62\rlap{)} & (2.37\rlap{)} \\    Pop\_35\_54 & $-$2.64\rlap{$^{***}$} & $-$2.78\rlap{$^{***}$} & $-$3.62\rlap{$^{***}$} & $-$3.13\rlap{$^{***}$} & $-$2.76\rlap{$^{**}$} & $-$2.13\rlap{$^{**}$} \\    & (0.84\rlap{)} & (0.89\rlap{)} & (0.83\rlap{)} & (1.10\rlap{)} & (1.11\rlap{)} & (1.06\rlap{)} \\    BornAsia & 3.58\rlap{$^{*}$} & $-$1.09 & 0.66 & $-$1.78 & $-$1.08 & $-$0.14 \\    & (2.09\rlap{)} & (2.52\rlap{)} & (1.99\rlap{)} & (2.74\rlap{)} & (2.54\rlap{)} & (2.17\rlap{)} \\    BornMidEast & $-$1.02 & $-$1.75 & $-$0.98 & $-$1.00 & $-$1.66 & $-$1.31 \\    & (1.00\rlap{)} & (1.17\rlap{)} & (1.09\rlap{)} & (1.33\rlap{)} & (1.23\rlap{)} & (1.11\rlap{)} \\    BornSEEuro & $-$1.63 & $-$3.17\rlap{$^{*}$} & $-$1.07 & $-$2.04 & $-$2.89\rlap{$^{***}$} & $-$2.53\rlap{$^{***}$} \\    & (1.37\rlap{)} & (1.68\rlap{)} & (1.06\rlap{)} & (1.29\rlap{)} & (1.11\rlap{)} & (0.97\rlap{)} \\    BornUK & 0.29 & 0.31 & 0.32 & 0.28 & $-$0.15 & $-$0.61 \\    & (1.02\rlap{)} & (1.04\rlap{)} & (0.87\rlap{)} & (1.06\rlap{)} & (0.99\rlap{)} & (0.99\rlap{)} \\    BornElsewhere & $-$4.13 & $-$1.51 & $-$1.03 & 2.45 & $-$4.21 & $-$2.17 \\    & (3.14\rlap{)} & (3.62\rlap{)} & (3.18\rlap{)} & (4.13\rlap{)} & (3.90\rlap{)} & (3.76\rlap{)} \\    Buddhism & $-$0.07 & 0.80 & 0.58 & $-$0.14 & $-$0.43 & $-$1.16 \\    & (1.31\rlap{)} & (1.54\rlap{)} & (1.39\rlap{)} & (1.66\rlap{)} & (1.60\rlap{)} & (1.58\rlap{)} \\    Christianity & $-$1.70 & $-$1.01 & $-$0.45 & 0.13 & 2.03 & 3.76\rlap{$^{**}$} \\    & (1.62\rlap{)} & (1.75\rlap{)} & (1.60\rlap{)} & (1.85\rlap{)} & (1.68\rlap{)} & (1.83\rlap{)} \\    CurrentlyStudying & $-$2.20\rlap{$^{*}$} & $-$0.01 & $-$0.14 & 1.35 & 0.32 & 0.22 \\    & (1.22\rlap{)} & (1.50\rlap{)} & (1.39\rlap{)} & (1.41\rlap{)} & (1.35\rlap{)} & (1.56\rlap{)} \\    DeFacto & $-$3.24 & $-$2.25 & $-$4.67\rlap{$^{**}$} & $-$7.75\rlap{$^{**}$} & $-$7.82\rlap{$^{**}$} & $-$10.39\rlap{$^{***}$} \\    & (2.07\rlap{)} & (2.62\rlap{)} & (2.27\rlap{)} & (3.09\rlap{)} & (3.08\rlap{)} & (3.15\rlap{)} \\    DiffAddress & 3.06\rlap{$^{***}$} & 2.75\rlap{$^{**}$} & 0.73 & 2.55 & 2.27 & 5.20\rlap{$^{***}$} \\    & (0.94\rlap{)} & (1.20\rlap{)} & (1.24\rlap{)} & (1.79\rlap{)} & (1.67\rlap{)} & (1.51\rlap{)} \\    Distributive & 1.60 & 1.89\rlap{$^{*}$} & 0.50 & 0.62 & 1.59 & 1.31 \\    & (1.06\rlap{)} & (1.14\rlap{)} & (0.99\rlap{)} & (1.27\rlap{)} & (1.20\rlap{)} & (1.18\rlap{)} \\    Education & $-$0.37 & $-$0.26 & $-$6.72\rlap{$^{**}$} & $-$7.31\rlap{$^{*}$} & $-$7.31\rlap{$^{**}$} & $-$8.55\rlap{$^{**}$} \\    & (2.35\rlap{)} & (3.34\rlap{)} & (3.00\rlap{)} & (3.90\rlap{)} & (3.63\rlap{)} & (3.37\rlap{)} \\    Extractive & 3.74\rlap{$^{***}$} & 4.96\rlap{$^{***}$} & 4.64\rlap{$^{***}$} & 6.46\rlap{$^{***}$} & 5.97\rlap{$^{***}$} & 6.38\rlap{$^{***}$} \\    & (1.43\rlap{)} & (1.47\rlap{)} & (1.20\rlap{)} & (1.45\rlap{)} & (1.35\rlap{)} & (1.38\rlap{)} \\    FamHouseSize & 1.94 & $-$2.55 & $-$6.47\rlap{$^{**}$} & $-$3.84 & $-$3.12 & $-$2.00 \\    & (2.61\rlap{)} & (3.66\rlap{)} & (3.28\rlap{)} & (3.87\rlap{)} & (3.52\rlap{)} & (3.06\rlap{)} \\    Incomes & 4.36\rlap{$^{***}$} & 2.42 & 5.52\rlap{$^{**}$} & 5.63\rlap{$^{*}$} & 8.02\rlap{$^{***}$} & 12.70\rlap{$^{***}$} \\    & (1.69\rlap{)} & (3.00\rlap{)} & (2.42\rlap{)} & (3.15\rlap{)} & (2.78\rlap{)} & (2.64\rlap{)} \\    Indigenous & 1.26 & 1.96 & 2.41 & 2.38 & 0.46 & $-$0.22 \\    & (1.61\rlap{)} & (1.89\rlap{)} & (1.59\rlap{)} & (2.00\rlap{)} & (1.88\rlap{)} & (1.90\rlap{)} \\    Islam & $-$0.75 & $-$0.91 & $-$0.60 & $-$2.01 & $-$0.88 & $-$1.09 \\    & (1.14\rlap{)} & (1.28\rlap{)} & (1.14\rlap{)} & (1.41\rlap{)} & (1.26\rlap{)} & (1.30\rlap{)} \\    Judaism & 1.32 & 0.93 & 1.47 & 0.28 & 1.35 & 1.15 \\    & (1.01\rlap{)} & (1.08\rlap{)} & (0.92\rlap{)} & (1.10\rlap{)} & (1.02\rlap{)} & (0.97\rlap{)} \\    ManagerAdmin & 2.62\rlap{$^{***}$} & 4.67\rlap{$^{***}$} & 7.47\rlap{$^{***}$} & 7.05\rlap{$^{***}$} & 5.93\rlap{$^{***}$} & 5.64\rlap{$^{***}$} \\    & (0.67\rlap{)} & (1.06\rlap{)} & (0.95\rlap{)} & (1.16\rlap{)} & (1.06\rlap{)} & (0.97\rlap{)} \\    Married & $-$3.93 & $-$2.72 & $-$9.35\rlap{$^{***}$} & $-$10.12\rlap{$^{***}$} & $-$7.91\rlap{$^{**}$} & $-$9.47\rlap{$^{**}$} \\    & (2.51\rlap{)} & (3.56\rlap{)} & (3.12\rlap{)} & (3.55\rlap{)} & (3.57\rlap{)} & (3.85\rlap{)} \\    NoReligion & $-$0.73 & 0.04 & 1.32 & 0.37 & 1.41 & 2.94 \\    & (1.50\rlap{)} & (1.65\rlap{)} & (1.51\rlap{)} & (1.75\rlap{)} & (1.74\rlap{)} & (2.03\rlap{)} \\    OneParentHouse & $-$4.77\rlap{$^{***}$} & $-$3.23 & $-$6.55\rlap{$^{***}$} & $-$7.03\rlap{$^{***}$} & $-$5.32\rlap{$^{***}$} & $-$4.94\rlap{$^{**}$} \\    & (1.49\rlap{)} & (1.99\rlap{)} & (1.81\rlap{)} & (2.04\rlap{)} & (1.97\rlap{)} & (2.03\rlap{)} \\    OtherLanguage & $-$1.02 & 6.88 & 6.21 & 7.80 & 10.13\rlap{$^{**}$} & 9.98\rlap{$^{**}$} \\    & (3.00\rlap{)} & (4.93\rlap{)} & (3.97\rlap{)} & (5.25\rlap{)} & (5.09\rlap{)} & (4.26\rlap{)} \\    PropertyOwned & $-$2.01 & $-$0.30 & 0.74 & $-$1.92 & $-$1.05 & 0.73 \\    & (1.35\rlap{)} & (1.49\rlap{)} & (1.36\rlap{)} & (1.74\rlap{)} & (1.67\rlap{)} & (1.48\rlap{)} \\    RentLoanPrice & $-$2.17 & 0.37 & 1.23 & 3.08 & 1.36 & $-$2.04 \\    & (1.46\rlap{)} & (1.93\rlap{)} & (1.76\rlap{)} & (2.23\rlap{)} & (2.20\rlap{)} & (2.07\rlap{)} \\    SocialServ & 3.31\rlap{$^{***}$} & 2.85\rlap{$^{**}$} & 3.46\rlap{$^{***}$} & 3.72\rlap{$^{**}$} & 2.98\rlap{$^{**}$} & 4.04\rlap{$^{***}$} \\    & (1.27\rlap{)} & (1.40\rlap{)} & (1.17\rlap{)} & (1.46\rlap{)} & (1.28\rlap{)} & (1.15\rlap{)} \\    Transformative & 2.30 & 4.71\rlap{$^{***}$} & 4.58\rlap{$^{***}$} & 4.55\rlap{$^{**}$} & 3.63\rlap{$^{**}$} & 4.05\rlap{$^{***}$} \\    & (1.48\rlap{)} & (1.77\rlap{)} & (1.51\rlap{)} & (1.87\rlap{)} & (1.67\rlap{)} & (1.47\rlap{)} \\    Unemployment & $-$3.39\rlap{$^{**}$} & $-$3.47\rlap{$^{**}$} & $-$0.40 & $-$0.68 & 0.81 & 1.93 \\    & (1.37\rlap{)} & (1.69\rlap{)} & (1.45\rlap{)} & (1.80\rlap{)} & (1.47\rlap{)} & (1.32\rlap{)} \\    Constant & 50.80\rlap{$^{***}$} & 52.63\rlap{$^{***}$} & 47.31\rlap{$^{***}$} & 49.92\rlap{$^{***}$} & 53.52\rlap{$^{***}$} & 50.46\rlap{$^{***}$} \\    & (0.76\rlap{)} & (0.59\rlap{)} & (0.44\rlap{)} & (0.52\rlap{)} & (0.54\rlap{)} & (0.64\rlap{)} \\  Residual Standard Error (GLS\rlap{)} & 4.34 & 4.82 & 4.32 & 5.30 & 4.82 & 4.76\\  \hline \\[-1.8ex]  Observations & 150 & 150 & 150 & 150 & 150 & 150 \\  \hline  \hline \\[-1.8ex]  & \multicolumn{6}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ \end{tabular}  \end{table}
 

### 2019-10-08 cooksd.R
add hat values

@@ -34,6 +34,9 @@ gls_model <- gls(my_formula, gls_data)
 # Cooks distance
 glsmod16$cooksd <- unname(predictmeans::CookD(gls_model, plot = FALSE))
 
+# Hat values
+glsmod16$hatvalues <- hat(model_data %>% dplyr::select(-LNP_Percent))
+
 # REMOVE
 
 remove(gls_model)
@@ -84,6 +87,9 @@ gls_model <- gls(my_formula, gls_data)
 
 # Cooks distance
 glsmod13$cooksd <- unname(predictmeans::CookD(gls_model, plot = FALSE))
+
+# Hat values
+glsmod13$hatvalues <- hat(model_data %>% dplyr::select(-LNP_Percent))
 
 # REMOVE
 
@@ -136,6 +142,9 @@ gls_model <- gls(my_formula, gls_data)
 # Cooks distance
 glsmod10$cooksd <- unname(predictmeans::CookD(gls_model, plot = FALSE))
 
+# Hat values
+glsmod10$hatvalues <- hat(model_data %>% dplyr::select(-LNP_Percent))
+
 # REMOVE
 
 remove(gls_model)
@@ -186,6 +195,9 @@ gls_model <- gls(my_formula, gls_data)
 
 # Cooks distance
 glsmod07$cooksd <- unname(predictmeans::CookD(gls_model, plot = FALSE))
+
+# Hat values
+glsmod07$hatvalues <- hat(model_data %>% dplyr::select(-LNP_Percent))
 
 # REMOVE
 
@@ -238,6 +250,9 @@ gls_model <- gls(my_formula, gls_data)
 # Cooks distance
 glsmod04$cooksd <- unname(predictmeans::CookD(gls_model, plot = FALSE))
 
+# Hat values
+glsmod04$hatvalues <- hat(model_data %>% dplyr::select(-LNP_Percent))
+
 # REMOVE
 
 remove(gls_model)
@@ -288,6 +303,9 @@ gls_model <- gls(my_formula, gls_data)
 
 # Cooks distance
 glsmod01$cooksd <- unname(predictmeans::CookD(gls_model, plot = FALSE))
+
+# Hat values
+glsmod01$hatvalues <- hat(model_data %>% dplyr::select(-LNP_Percent))
 
 # REMOVE
 

### 2019-10-08 anzjs_elections.Rmd
added leverage to influence/outlier section

@@
     Band_55plus = Age55_64 + Age65_74 + Age75_84 + Age85plus) %>%
-  select(-c(
+  dplyr::select(-c(
     starts_with("Age"), MedianAge, StateAb,
@@
   ) %>%
-  select(-Band_55plus)
+  dplyr::select(-Band_55plus)
 ```
@@
   cooksd = c(glsmod16$cooksd, glsmod13$cooksd, glsmod10$cooksd, glsmod07$cooksd, glsmod04$cooksd, glsmod01$cooksd),
+  Hat = c(glsmod16$hatvalues, glsmod13$hatvalues, glsmod10$hatvalues, glsmod07$hatvalues, glsmod04$hatvalues, glsmod01$hatvalues),
   bind_rows(
@@
 )
+```
+
+```{r plothatandcooksd, eval = FALSE}
+cooksd_df %>%
+  ggplot(aes(x = cooksd, y = Hat)) +
+  geom_point(shape = 1) +
+  #geom_text(aes(label = DivisionNm), size = 3) +
+  facet_wrap(~year, nrow = 2) +
+  # coord_fixed(ratio = .5) +
+  labs(x = "Cook's Distance", y = "Hat values") +
+  theme(axis.text.x = element_text(angle = 60, hjust = 1, size = 8), axis.text.y = element_text(size = 8)) +
+  theme_bw()
 ```
@@
 
-Based on the distribution of the Cook's distance values, a Cook's distance greater than $0.1$ is considered to be influential and a potential outlier.
-The electorate of Sydney (NSW) has a large Cook's distance from 2001 to 2010, due to its diverse population (language, birthplace and religion), high number of defacto relationships, high income, high household mobility and small amount of workers in extractive and transformative jobs.
-It has remained a strong supporter of the Labor party and the Liberal vote is severely overpredicted by the model, making it an outlier.
+Based on the distribution of the Cook's distance values and the distribution of hat values, a Cook's distance greater than $0.1$ is considered to be influential, and a hat value greater than $0.5$ is considered to have high leverage.
+Electorates fitting these criteria are flagged and investigated to examine the characteristics driving these values.
+
+The electorate of Sydney (NSW) has a large Cook's distance and high leverage from 2001 to 2007, due to its diverse population (languages, birthplace and religion), high density of young adults (20 to 34 years old), high number of defacto relationships, high income, high household mobility and small amount of workers in extractive and transformative jobs.
+It has remained a strong supporter of the Labor party and the extent of this support is underpredicted by the model, making it an outlier.
 Nearby in metropolitan NSW, the electorate of Wentworth is found to be an outlier in the 2013 and 2016 elections.
@@
 
-Lingiari, an electorate taking up almost all of the Northern Territory, is an outlier in all but the 2013 election due to its large Indigenous population, young age profile and low rates of property ownership.
-Fowler (NSW) has a diverse population with a high proportion of migrants, many Buddhists and Muslims, and has strong Labor support, making it influential in 2001, 2004 and 2010.
+Lingiari, an electorate taking up almost all of the Northern Territory, has consistently high leverage (all years) and is an outlier in all but the 2013 election due to its large Indigenous population, low rates of property ownership and few workers in management or administrative jobs.
+Fowler (NSW) has a diverse population with a high proportion of migrants, many Buddhists and Muslims, as well as a high proportion of single parent households.
+These characteristics explain its high leverage in 2001, 2004, 2010 and 2013, and its strong Labor support makes it influential in 2001, 2004 and 2010.
 Other electorates with large Cook's distance are Canberra (ACT) and Durack (WA) in 2013, and Solomon (NT) in 2016.
+All of the electorates examined are not unduly influential in the model and therefore no action is required.
 

### 2019-10-08 functions.R
reduced font size of axis titles in partial residual plots

@@ -264,8 +264,8 @@ grid_visreg <- function(varname, plot = TRUE, myscale = "free_y", top = FALSE) {
         labs(x = "", y = "") +
         facet_grid(varname ~ year, scales = "free_y") +
         theme(plot.margin=unit(c(b = 0.05, l = 0, t = -0.2, r = 0),"cm"),
-          strip.text = element_text(size = 8),
-          axis.text = element_text(size = 8))
+          strip.text = element_text(size = 6),
+          axis.text = element_text(size = 6))
 
       if (top == FALSE) {
         return_object <- return_object + theme(strip.text.x = element_blank())

### 2019-10-08 anzjs_elections.Rmd
proofread - minor wording changes

@@
 This dataset consists of the two-party preferred vote (as the response variable) and a set of common socio-demographic variables (as the explanatory variables).<!--The use of a regression framework to examine these socio-political relationships over time is seemingly absent from previous Australian studies.
-It also appears that no study has attempted any type of statistical analysis of socio-demographics in conjunction with voter behaviour in Australia since 2007, making this paper distinctly different from previous work.--> To prepare these datasets, socio-demographic variables are first standardized, and then a principal component analysis is used to group variables into "factors".
+It also appears that no study has attempted any type of statistical analysis of socio-demographics in conjunction with voter behaviour in Australia since 2007, making this paper distinctly different from previous work.--> To prepare these datasets, socio-demographic variables are first standardized, and then a principal component analysis is used to group many of the variables into "factors".
 To account for the inherent spatial structure of the data, a spatial error model is then estimated for each election.
@@
 Section \@ref(data) describes the data collection, joining and cleaning, while model details are discussed in Section \@ref(modelling).
-Section \@ref(results) describes the inference conducted to determine significance of effects and how these change over time.
+Section \@ref(results) describes the inference conducted to determine significance of effects and how these change over time, as well as including details on model robustness.
 Section \@ref(conclusion) summarises the work.
@@
 
-Following this imputation process, electoral socio-demographic variables are available for each of the six elections and can be joined with their corresponding two-party preferred votes.
+From this imputation process, electoral socio-demographic variables are available for each of the six elections and can be joined with their corresponding two-party preferred votes.
 Before choosing an appropriate model, two issues with the socio-demographic variables need to be addressed.
@@
 Most predictors have a positive relationship: the larger the value the more likely the electorate preferences the Coalition.
-The relationship is relatively robust over time, with the exception of `OtherLanguageHome` and `Pop_00_19`.
+The relationship is relatively robust over time, with the exception of `Incomes`, `OtherLanguageHome`, `Pop_00_19` and `DiffAddress`.
 
@@
 Electorates with more diverse speech are associated with higher support for the Liberal party from 2004 onwards, with this effect being significant in 2013 and 2016 row 7, Figure \@ref(fig:partresplot1)).
-Furthermore, none of the variables relating to religious beliefs aside from Christianity has a material effect in any election (this includes the Buddhist, Muslim, Jewish, non-religious and Indigenous Australian populations).
+Furthermore, none of the variables relating to religious beliefs aside from Christianity have a material effect in any election (this includes the Buddhist, Muslim, Jewish, non-religious and Indigenous Australian populations).
 The association between Christian populations (`Christianity`) and the Liberal party steadily increases over the years, becoming positive and significant in 2016.
@@
 In 2001, 2004 and 2016, higher proportions of people that have recently (in the past five years) moved house (`DiffAddress`) increased electoral support for the Liberal party (see row 5 in Figure \@ref(fig:partresplot1)).
-Having controlled for characteristics of house ownership and rental prices (`PropertyOwned` and `RentLoan`,  both of which have no effect), this effect is somewhat surprising.
-
-Higher proportions of single parent households are associated with Labor support in all elections (albeit insignificant in 2004, see row 6 in Figure \@ref(fig:partresplot2)), whereas the electoral family and household sizes (via the `FamHouseSize` variable) do not appear to be associated with either party in any election.
+This is somewhat surprising as one might expect for house ownership and rental prices to be drivers of two-party preference, rather than household mobility (`PropertyOwned` and `RentLoan` are not significant in any election).
+
+Higher proportions of single parent households are associated with Labor support in all elections (albeit insignificant in 2004, see row 6 in Figure \@ref(fig:partresplot2)), whereas the electoral family and household sizes (via the `FamHouseSize` variable) do not appear to be associated with either party.
 
@@
 
-From 2007 onwards, both marriages (`Married`) and de facto relationships (`DeFacto`) are found to be strong predictors of the two-party preferred vote in opposition to the Liberal party.
+From 2007 onwards, both marriages (`Married`) and de facto relationships (`DeFacto`) are found to be strong predictors of the two-party preferred vote in favour of the Labor party.
 In 2001 and 2004 neither of these variables are significant (see rows 4 and 5 in Figure \@ref(fig:partresplot2)).
@@
 Family and household sizes have minimal influence, although electorates with more single parent households are linked with Labor support.
-Furthermore, the spatial effects are found to be positive in all elections and significant in 2001, 2004 and 2016, meaning that other characteristics that neighbours have in common appear to be influential in those years.
+Furthermore, the spatial effects are found to be positive in all elections and significant in 2001, 2004 and 2016, meaning that other characteristics that neighbours have in common  (outside of the variables in the model) appear to be influential in those years.
 
@@
 
-The authors would like to sincerely thank the editor and associate editor of the Australian & New Zealand Journal of Statistics, and the two anonymous reviewers for providing helpful comments and suggestions on earlier drafts of the manuscript.
+The authors would like to sincerely thank the editor and associate editor of the Australian & New Zealand Journal of Statistics and the two anonymous reviewers for providing helpful comments and suggestions on earlier drafts of the manuscript.
 Additionally, the authors would like to thank Anthony Ebert, Heike Hofmann, Thomas Lumley, Ben Marwick, Carson Sievert, Mingzhu Sun, Dilini Talagala, Nicholas Tierney, Nathaniel Tomasetti, Earo Wang and Fang Zhou, all of whom have contributed to the `eechidna` package.

### 2019-10-15 anzjs_elections.Rmd
rewording the point on why a sepearate model for each election is preferred

@@
 An identical model specification is used for each of the six elections, with each election modelled separately.
-This allows the socio-demographic effects to be estimated separately for each election year, facilitating analysis of temporal changes in variable effects.
-This approach is preferable to using a single longitudinal model because it avoids imposing any time-varying restrictions on the effect of a variable.
-To gain any advantage over modelling each election as a separate cross-section, a pooled cross-sectional model would have to force the effect of at least one variable to be invariant over time (e.g.
-forcing the estimated coefficient of the variable `Unemployed` to be the same in 2013 and 2016 elections).
-This would not allow for conclusions to be drawn about how the effects of these variables change over time.
-Furthermore, the panel approach is avoided because of how frequently electoral boundaries change, noting that electorates with the same name across elections are not guaranteed to represent the same geographic region.
+Separate models are preferred to a single model because of how frequently electoral boundaries change, noting that electorates with the same name across elections are not guaranteed to represent the same geographic region.
 Therefore any fixed or random effects models would be difficult to estimate without implementing consistent boundaries, which would require further imputation (of voting information).
+The separate models also allows the socio-demographic effects to be estimated separately for each election year, facilitating analysis of temporal changes in variable effects.
+This can be considered a special case of a longitudinal model where all coefficients are time-varying and heteroskedasticity is time-varying.
 

### 2019-10-25 anzjs_elections.Rmd
temporary geometry to knit on JF computer

--- 2019-10-08 anzjs_elections.Rmd
@@
       in_header: anzjs_preamble.tex
+geometry: "left=5cm, right=0cm, top=4cm, bottom=0cm"
 ---
@@
 
-We examine the relationships between electoral socio-demographic characteristics and two-party preferences in the six Australian federal elections held between 2001 to 2016.
+We examine the relationships between electoral socio-demographic characteristics and two-party preferences in the six Australian federal elections held between 2001 and 2016.
 Socio-demographic information is derived from the Australian Census which occurs every five years.
@@
 To capture any spatial heterogeneity, a spatial error model is estimated for each election, which incorporates a spatially structured random effect vector.
-Over time, the impact of most socio-demographic characteristics that affect electoral two-party preference do not vary, with age distribution, industry of work, incomes, household mobility and de facto relationships having strong effects in each of the six elections.
+Over time, the impact of most socio-demographic characteristics that affect electoral two-party preference do not vary, with age distribution, industry of work, incomes, household mobility and relationships having strong effects in each of the six elections.
 Education and unemployment are amongst those that have varying effects.
@@
 
-Raw data was sourced online from the AEC and ABS websites in `.csv` and `.xlsx` files.
+Raw data is sourced online from the AEC and ABS websites in `.csv` and `.xlsx` files.
 The formats of these files differ over the years, making extracting the appropriate information a big task.
@@
 Therefore, for the four elections between 2004 and 2013, both temporal and spatial differences in electorates need to be accounted for when joining the electoral two-party preferred vote with Census data.
-For these elections a spatio-temporal imputation method was employed to obtain electoral socio-demographics.
+For these elections a spatio-temporal imputation method is employed to obtain electoral socio-demographics.
 This method uses Census information from both before and after the election of interest.
@@
 
-Since all socio-demographics have been standardized to have a mean of zero and a variance of one, the intercept in each model can be interpreted as the estimated two-party preferred vote for an electorate with mean characteristics[^3].
+Since all socio-demographics have been standardized to have a mean of zero and a variance of one, the intercept in each model can be interpreted as the estimated two-party preferred vote for an electorate with mean characteristics[^4].
 Figure \@ref(fig:plotintercept) shows that the baseline of party preference has varied over the elections, with the biggest swing occurring in the 2007 election where the mean electorate shifted more than five percentage points in favour of the Labor party.
 
-[^3]: Mean of all variables aside from Judaism, Indigenous, Islam and Buddhism, where it assumes the mean of the log value.
+[^4]: Mean of all variables aside from Judaism, Indigenous, Islam and Buddhism, where it assumes the mean of the log value.
 
@@
 
-### Checks
+### Multicollinearity {-}
 

### 2019-10-26 anzjs_elections.Rmd
minor change to residuals by state

@@
 
-It is often hypothesized that states have systematic differences that cause their electorates to vote differently.
-Boxplots of residuals grouped by state (Figure \@ref(fig:resstate)) reveal that the data reflects this to a limited extent.
-Tasmania and the Australian Capital Territory appear to have a bias towards Labor, whereas the Northern Territory tends towards voting Liberal.
-However, there are relatively few electorates in each of these states (five, two and two respectively), so this apparent result may be due to incumbent effects rather than an actual state-specific bias.
+It is often hypothesized that states have a systematic bias towards one of the two major parties.
+Boxplots of residuals grouped by state (Figure \@ref(fig:resstate)) reveal that the data reflects this to only a limited extent.
+Tasmania and the Australian Capital Territory appear to have a bias towards Labor, whereas the South Australia and the Northern Territory tend towards voting Liberal.
+However, there are relatively few electorates in each of these states (five, two, eleven and two respectively), so this apparent result may be due to incumbent effects rather than an actual state-specific bias.
 

### 2019-10-26 response_for_ANZJS_editor.Rmd
version of response ready for submission

--- 2019-10-01 response_for_ANZJS_editor.Rmd
@@
 ---
-title: "Response For The Editor"
-author: "Jeremy Forbes"
-date: "31/08/2019"
-output: html_document
+title: "Response to the Referees"
+author: "Jeremy Forbes, Di Cook, Rob Hyndman"
+date: "`r format(Sys.time(), '%d %B, %Y')`"
+output: pdf_document
 ---
@@
 
-We have revised our paper to address the issues raised. Below we have detailed the changes made and our responses to the points raised in your email and by the reviewers.
+We have revised our paper to address the issues raised.
+Below are detailed summaries of the changes made in response to the points raised by yourself, the Associate Editor and by the anonymous reviewers.
 
-#### Editor's Comments
+## Editor's Comments
 
@@
 Some spatial imputation is still required as the SA1 boundaries do not exactly match the electorate boundaries.
-With this change in imputation method, our results are effectively the same as the previous submission.*
+Even with this change in imputation method, our results are effectively the same as the previous submission.*
 
@@
 *We have expanded our remarks and moved them to the introduction (Section 1) to give them more prominence.
-We are explicit in stating that "drivers of individual voter behaviour may vary from what is observed at the electorate level".*
+We are explicit in stating that "Insights are being drawn at the electorate level and cannot be inferred for another disaggregate level (in particular, drivers of individual voter behaviour may vary from what is observed at the electorate level)".*
 
@@
 
-*We have made an effort to clarify this process in Section 3.2, which now details the variables being combined into each factor.
-For consistency, only commas are used to separate the variable descriptions.
-Only the signs are used, which is now made explicit.*
+*We have made an effort to clarify this process in Section 3.2, which now details each of the variables being combined into each factor.
+For consistency, only commas are used to separate the variable descriptions.*
+
+*The factors are calculated using only the signs, not the loading weights. A total of 23 variables are combined into 6 different factors.*
 
@@
 We begin with 50 variables.
-Six factors are created using a combined 23 variables.
+Six factors are created using a total of 23 variables.
 One of the age brackets (populated aged 55 and over) is removed, and the other three age brackets are kept (to avoid multicollinearity).
 This results in a final predictor set of 32 variables.
-Note that two variables no longer meet the criteria for being included in a factor, so they appear as separate variables - hence a final set of 32 variables which are listed in Table 1.*
+Any variables that are not contained in a factor are not discarded, and are kept in the model as a explanatory variables.*
 
@@
 *This was a very helpful suggestion to explore, and as such an additional section on robustness has been included in the revised paper (Section 4.4).
-When we fit a model using just the variables with significant effects, we find that the results do not change much.
+When we fit a model using just the variables with significant effects, we find that the results are robust.
 Each of the estimated coefficients from the reduced model (containing only the subset significant variables) lie within the 95% confidence interval from its corresponding full model (containing all 32 variables).*
@@
 
-The first three minor issues have been corrected. For the last issue, the spatial AR parameter is now significant in 2001, 2004 and 2016, which do not correspond only to years with no imputation.
+*The first three minor issues have been corrected. For the last issue, the spatial coefficient is now significant in 2001, 2004 and 2016, which do not correspond only to years with no imputation.*
 
-#### Associate Editor's Comments
+## Associate Editor's Comments
 
-1. Some conceptually simple imputation methods are used to align polling and census data to investigate associations between census variables and voting behaviour at electoral division level.
-
-*No response required.*
-
-2.
+1.
 The main thing I found confusing about this paper was the variable selection procedure.
@@
 
-*See response to the editor.*
+*We have reworked Section 3.2 to better explain this process.*
 
-3.
+*We begin with 50 variables.
+Many of these variables represent similar information and where appropriate, should be grouped together.
+PCA is used to determine which variables covary, and hence decide which varibles should be combined into a factor.
+Each of the first four PCs are examined to determine which variables have large loadings.
+If there is an intuitive reason as to why a subset of variables with large loadings (in a given PC) should be grouped together, then they are combined into a factor using a weighted sum.
+The weights are simply +1 is the loading in that PC is positive, and -1 if it is negative.*
+
+*Six factors are created using a total of 23 variables.
+One of the age brackets (populated aged 55 and over) is removed, and the other three age brackets are kept (to avoid multicollinearity).
+This results in a final predictor set of 32 variables.
+Any variables that are not contained in a factor are not discarded, and are kept in the model as a explanatory variables.*
+
+2.
 The authors opt for “visualization” of the fixed effects rather than the usual table of estimated parameters.
@@
 
-4. I found the writing unclear in a few places. What are the “time-varying restrictions” on a variable (line228)? And the description of a partial residual (line 276) doesn’t make sense to me.
+3. I found the writing unclear in a few places. What are the “time-varying restrictions” on a variable (line228)?
 
-*We have expanded on descriptions of both in our revised version.
-Time-varying restrictions are when one forces the effect of a variable to be invariant over time, for example, forcing the estimated coefficient of the variable `Unemployed` to be the same in 2013 and 2016 elections (section 3.2).*
+*We have revised our explanation relating to the model specification (Section 3.3).
+Each election is modelled as a separate cross section, which can be considered a special case of a longitudinal model where all coefficients are time-varying and heteroskedasticity is time-varying.
+A time-varying coefficient is a coefficient that a different estimate for each time period $t=1,...,T$.
+We have removed the note about "time-varying restrictions".*
+
+4. And the description of a partial residual (line 276) doesn’t make sense to me.
 
 *Partial residuals, for a variable X1, are the residuals from the fitted model with the estimated effect of X1 added to each.
-Partial residual plots show the direction, size and significance of an estimated effect, as well as any deviations from linearity (section 4.3).*
+Partial residual plots show the direction, size and significance of an estimated effect, as well as any deviations from linearity (see Section 4.3).*
 
@@
 
-#### Reviewer 1's comments
+## Reviewer 1's comments
 
-Its goal is to account for – though the word ‘explain’ is used! – patterns of voting at Australian federal elections, using ecological data, but with no reference to the ecological fallacy.
+1. Its goal is to account for – though the word ‘explain’ is used! – patterns of voting at Australian federal elections, using ecological data, but with no reference to the ecological fallacy.
 
@@
 
+2.
 The paper is written in virtual total ignorance of the large amount of work done on Australian voting patterns; apart from reference to a few papers by geographers there is not a single mention of work by political scientists – of which there is a great deal.
@@
 
+3.
 The banality of this approach is illustrated by the finding that ‘more de facto relationships’ are associated with greater support for Labor.
@@
 
+4.
 Much is done – unnecessarily in my view – in preparing the data through imputation.
@@
 
-The authors recognise that there is likely to be collinearity among 65 Census variables – why those; we are never told?
+5. The authors recognise that there is likely to be collinearity among 65 Census variables – why those; we are never told?
 
@@
 
+6.
 And so, as others do, they did a pca of the data matrix, reducing it to four components – we are not told what rotation was employed to achieve simple structure.
@@
 
-*Our approach was to combine the six election datasets together into a single dataset of 900 observations.
+*We have reworked Section 3.2 to better explain this process.*
+
+*Our approach is to combine the six election datasets together into a single dataset of 900 observations.
 We computed the principal components on this.
 The first four components are chosen because the amount of variance explained drops off after the fourth PC.
-We have now explained our variable selection process in more detail.*
+We then use these four PCs to determine which variable covary, and hence decide which varibles should be combined into a factor.*
 
-*The factors are not the principal components themselves.
-Rather, the loadings in the first four principal components are used to determine which variables should be grouped together to generate a new variable in their place, which we have called a factor.*
+*Each of the first four PCs are examined to determine which variables have large loadings.
+If there is an intuitive reason as to why a subset of variables with large loadings (in a given PC) should be grouped together, then they are combined into a factor using a weighted sum.
+The weights are simply +1 is the loading in that PC is positive, and -1 if it is negative.
+Hence, the factors are not the principal components themselves.*
 
+7.
 My expectation was that they would then fit a model using the four components as the independent variables, thus avoiding any collinearity and confounding.
@@
 
-*In response to this suggestion, we have now added a section on robustness in the revised paper (Section 4.4).
-When we fit a model using just the variables with significant effects, we find that the results do not change much.
+*In response to this suggestion, we have now added an assessment of multicollinearity within the robustness section of the revised paper (Section 4.5).
+Three robustness checks are conducted as follows:*
+
+- *We fit a model using just the variables with significant effects, we find that the results do not change much.
 Each of the estimated coefficients from the reduced model (containing only the subset significant variables) lie within the 95% confidence interval from its corresponding full model (containing all 32 variables).*
 
-*We also checked the effect of omitting a variable that is contained in the ten largest pairwise correlations.
+- *We also checked the effect of omitting a variable that is contained in the ten largest pairwise correlations.
 For each pair, a model for each election is refit omitting one of the two variables.
@@
 
-*A third and final check is the visual exploration of different variable projections using a tour (for each election). No definitive signs of multicollinearity are observed.*
+- *A third and final check is the visual exploration of different variable projections using a tour (for each election). No definitive signs of multicollinearity are observed.*
 
+8.
 The authors fit a spatial autoregressive model, on the assumption that neighbouring places are more likely to be similar than distant places (the ‘first law of geography’).
@@
 *In retrospect, we agree that this is perhaps a naïve conclusion.
-It is possible that unobserved variables are being captured in the spatial effect, and as such we have removed this interpretation from the revised paper.
+It is possible that unobserved variables are being captured in the spatial effect, which is the interpretation that we have now included.
+We have removed the old interpretation from the revised paper.
 Additionally, on revision of the imputation method, we find that the spatial effect is significant in 2004 as well as 2001 and 2016.*
 
-#### Reviewer 2's comments
+## Reviewer 2's comments
 
+1.
 It seems very strange to assume that population is uniform within electorates (line 133) so that weighting is proportional to area, and not to the number of people in an area of intersection.
@@
 
+2.
 It seems to me that the use of Principal Components analysis is unclear.
@@
 Of course, there is no guarantee about any of these factors being related to the response.
-We have now explained our variable selection procedure in more detail in the revised paper.*
+We have now explained our variable selection procedure in more detail in the revised paper (Section 3.2).*
 
-(b) The variables which load highly on a PC are likely to be highly correlated, and so if these high-loading variables are used together in the regression model we may again have multicollinearity.
+3.
+The variables which load highly on a PC are likely to be highly correlated, and so if these high-loading variables are used together in the regression model we may again have multicollinearity.
 I found this was the case when applying the author’s regression model to the 2016 data.
@@
 Two factors that have been created from the same PC will be correlated, which could mean that (near) multicollinearity persists.
-However, we have conducted three robustness tests for multicollinearity, and no conclusive signs of multicollinearity were observed.*
+However, we have conducted three robustness tests for multicollinearity (see Section 4.5), and no conclusive signs of multicollinearity were observed.*
 
@@
 
-(c) Conversely there could be a variable which is independent of the other variables (so representing 1 / 65th of the explanatory information) but by dint of that fact not be included in the first four PCs.
+4.
+Conversely there could be a variable which is independent of the other variables (so representing 1 / 65th of the explanatory information) but by dint of that fact not be included in the first four PCs.
 In short, the narrative leaves me unconvinced that all the variables included are necessary or that all the relevant variables are included in the regression.
@@
 
-(d) I presume the “factors” referred to lines 210-220 were created as the first principal component of the constituent raw variables?*
+5. I presume the “factors” referred to lines 210-220 were created as the first principal component of the constituent raw variables?*
 
-*In our revised submission, we have made an effort to clarify what we have defined a "factor" to be.
+*In our revised submission, we have made an effort to clarify what we have defined a "factor" to be (Section 3.2).
 A "factor" is created when there is evidence to suggest that a group of variables are explaning similar information, and if they were all included in the model as separate variables, there would be severe issues with multicollinearity.
@@
 
-(e) The description of the variables included in “Education” is not clear.
+6. The description of the variables included in “Education” is not clear.
 
@@
 
-(e) I could not get the variable ‘OtherLanguageHome’ in 2016 to be significant, but it also had a high VIF with other explanatory variables. Please check this.
+7. I could not get the variable ‘OtherLanguageHome’ in 2016 to be significant, but it also had a high VIF with other explanatory variables. Please check this.
 
@@
 
+8.
 I am not convinced it is reasonable to summarise the effect of age by median age.
@@
 
+9.
 Is there any tendency for the vote in an election to depend on how long the incumbent has been in power?

### 2019-11-10 anzjs_elections.Rmd
minor typo fixes

@@
 
-Socio-demographic variables were derived from the Australian Census of Population and Housing (Census), which is a survey of every household in Australia, recording information such as age, gender, ethnicity, education level and income.
+Socio-demographic variables are derived from the Australian Census of Population and Housing (Census), which is a survey of every household in Australia, recording information such as age, gender, ethnicity, education level and income.
 There have been four Censuses so far in the 21st century, conducted in 2001, 2006, 2011 and 2016.
@@
 The ABS uses electoral boundaries as defined by the AEC at the time of each Census, which may not match those in place at the subsequent and previous elections.
-From the available Census information aggregated at the electorate level, 50 socio-demographic variables were defined for each of the electorates to be used in the analysis.
+From the available Census information aggregated at the electorate level, 50 socio-demographic variables are defined for each of the electorates to be used in the analysis.
 These variables include information relating to electoral age distributions, income, education qualifications, employment industries and job types, religion, birthplace, household characteristics and relationships.
@@
 For each election, neighbouring Census information has to be combined in some way so that it represents the boundaries in place at the time of the election.
-This is done by taking the electoral boundaries and imputing the corresponding socio-demographic characteristics for each of neighbouring Censuses, thereby addressing the spatial aspect.
+This is done by taking the electoral boundaries and imputing the corresponding socio-demographic characteristics for each of the neighbouring Censuses, thereby addressing the spatial aspect.
 Next, to deal with the temporal component, characteristics at the time of the election are constructed using linear interpolation between the spatially imputed neighbouring Census variables.
@@
 Therefore any fixed or random effects models would be difficult to estimate without implementing consistent boundaries, which would require further imputation (of voting information).
-The separate models also allows the socio-demographic effects to be estimated separately for each election year, facilitating analysis of temporal changes in variable effects.
+The separate models also allow the socio-demographic effects to be estimated separately for each election year, facilitating analysis of temporal changes in variable effects.
 This can be considered a special case of a longitudinal model where all coefficients are time-varying and heteroskedasticity is time-varying.
@@
 
-```{r prepplot, include = FALSE}
+```{r prepplot, include = FALSE, out.height="100%"}
 p1 <- grid_visreg("Extractive", top = TRUE)
@@
 Speaking other languages (aside from English) however, appears to have a far stronger effect, as observed through the `OtherLanguage` variable.
-Electorates with more diverse speech are associated with higher support for the Liberal party from 2004 onwards, with this effect being significant in 2013 and 2016 row 7, Figure \@ref(fig:partresplot1)).
+Electorates with more diverse speech are associated with higher support for the Liberal party from 2004 onwards, with this effect being significant in 2013 and 2016 (see row 7, Figure \@ref(fig:partresplot1)).
 Furthermore, none of the variables relating to religious beliefs aside from Christianity have a material effect in any election (this includes the Buddhist, Muslim, Jewish, non-religious and Indigenous Australian populations).
@@
 It is often hypothesized that states have a systematic bias towards one of the two major parties.
-Boxplots of residuals grouped by state (Figure \@ref(fig:resstate)) reveal that the data reflects this to only a limited extent.
+Boxplots of residuals grouped by state (Figure \@ref(fig:resstate)) show that the data reflects this to only a limited extent.
 Tasmania and the Australian Capital Territory appear to have a bias towards Labor, whereas the South Australia and the Northern Territory tend towards voting Liberal.
@@
 
-The authors would like to sincerely thank the editor and associate editor of the Australian & New Zealand Journal of Statistics and the two anonymous reviewers for providing helpful comments and suggestions on earlier drafts of the manuscript.
+The authors would like to sincerely thank the editor and associate editor of the Australian & New Zealand Journal of Statistics and the two anonymous reviewers for providing helpful comments and suggestions on earlier drafts of this manuscript.
 Additionally, the authors would like to thank Anthony Ebert, Heike Hofmann, Thomas Lumley, Ben Marwick, Carson Sievert, Mingzhu Sun, Dilini Talagala, Nicholas Tierney, Nathaniel Tomasetti, Earo Wang and Fang Zhou, all of whom have contributed to the `eechidna` package.

### 2019-11-10 anzjs_elections.tex
minor typo fixes

@@
 }{}
+\usepackage[left=5cm, right=0cm, top=4cm, bottom=0cm]{geometry}
 \usepackage{hyperref}
@@
 This dataset consists of the two-party preferred vote (as the response variable) and a set of common socio-demographic variables (as the explanatory variables).
-To prepare these datasets, socio-demographic variables are first standardized, and then a principal component analysis is used to group variables into ``factors''.
+To prepare these datasets, socio-demographic variables are first standardized, and then a principal component analysis is used to group many of the variables into ``factors''.
 To account for the inherent spatial structure of the data, a spatial error model is then estimated for each election.
@@
 Section \ref{data} describes the data collection, joining and cleaning, while model details are discussed in Section \ref{modelling}.
-Section \ref{results} describes the inference conducted to determine significance of effects and how these change over time.
+Section \ref{results} describes the inference conducted to determine significance of effects and how these change over time, as well as including details on model robustness.
 Section \ref{conclusion} summarises the work.
@@
 The ABS uses electoral boundaries as defined by the AEC at the time of each Census, which may not match those in place at the subsequent and previous elections.
-From the available Census information aggregated at the electorate level, 50 socio-demographic variables were defined for each of the electorates to be used in the analysis.
+From the available Census information aggregated at the electorate level, 50 socio-demographic variables are defined for each of the electorates to be used in the analysis.
 These variables include information relating to electoral age distributions, income, education qualifications, employment industries and job types, religion, birthplace, household characteristics and relationships.
 
-Raw data was sourced online from the AEC and ABS websites in \texttt{.csv} and \texttt{.xlsx} files.
+Raw data is sourced online from the AEC and ABS websites in \texttt{.csv} and \texttt{.xlsx} files.
 The formats of these files differ over the years, making extracting the appropriate information a big task.
@@
 
-The final predictor set contains \(p=32\) variables\footnote{Mean of all variables aside from Judaism, Indigenous, Islam and Buddhism, where it assumes the mean of the log value.} which are listed in Table \ref{tab:printmodels}.
+The final predictor set contains \(p=32\) variables\footnote{The factor creation procedure reduced the variable set to \(p=33\), however one of the age brackets (\texttt{Pop\_55\_plus}) is not included as a variable to avoid multicollinearity, because the other three age brackets are included.} which are listed in Table \ref{tab:printmodels}.
 
@@
 An identical model specification is used for each of the six elections, with each election modelled separately.
-This allows the socio-demographic effects to be estimated separately for each election year, facilitating analysis of temporal changes in variable effects.
-This approach is preferable to using a single longitudinal model because it avoids imposing any time-varying restrictions on the effect of a variable.
-To gain any advantage over modelling each election as a separate cross-section, a pooled cross-sectional model would have to force the effect of at least one variable to be invariant over time (e.g.~forcing the estimated coefficient of the variable \texttt{Unemployed} to be the same in 2013 and 2016 elections).
-This would not allow for conclusions to be drawn about how the effects of these variables change over time.
-Furthermore, the panel approach is avoided because of how frequently electoral boundaries change, noting that electorates with the same name across elections are not guaranteed to represent the same geographic region.
+Separate models are preferred to a single model because of how frequently electoral boundaries change, noting that electorates with the same name across elections are not guaranteed to represent the same geographic region.
 Therefore any fixed or random effects models would be difficult to estimate without implementing consistent boundaries, which would require further imputation (of voting information).
+The separate models also allow the socio-demographic effects to be estimated separately for each election year, facilitating analysis of temporal changes in variable effects.
+This can be considered a special case of a longitudinal model where all coefficients are time-varying and heteroskedasticity is time-varying.
 
@@
 Most predictors have a positive relationship: the larger the value the more likely the electorate preferences the Coalition.
-The relationship is relatively robust over time, with the exception of \texttt{OtherLanguageHome} and \texttt{Pop\_00\_19}.}\label{fig:partresplot1}
+The relationship is relatively robust over time, with the exception of \texttt{Incomes}, \texttt{OtherLanguageHome}, \texttt{Pop\_00\_19} and \texttt{DiffAddress}.}\label{fig:partresplot1}
 \end{figure}
@@
 Speaking other languages (aside from English) however, appears to have a far stronger effect, as observed through the \texttt{OtherLanguage} variable.
-Electorates with more diverse speech are associated with higher support for the Liberal party from 2004 onwards, with this effect being significant in 2013 and 2016 row 7, Figure \ref{fig:partresplot1}).
-Furthermore, none of the variables relating to religious beliefs aside from Christianity has a material effect in any election (this includes the Buddhist, Muslim, Jewish, non-religious and Indigenous Australian populations).
+Electorates with more diverse speech are associated with higher support for the Liberal party from 2004 onwards, with this effect being significant in 2013 and 2016 (see row 7, Figure \ref{fig:partresplot1}).
+Furthermore, none of the variables relating to religious beliefs aside from Christianity have a material effect in any election (this includes the Buddhist, Muslim, Jewish, non-religious and Indigenous Australian populations).
 The association between Christian populations (\texttt{Christianity}) and the Liberal party steadily increases over the years, becoming positive and significant in 2016.
@@
 In 2001, 2004 and 2016, higher proportions of people that have recently (in the past five years) moved house (\texttt{DiffAddress}) increased electoral support for the Liberal party (see row 5 in Figure \ref{fig:partresplot1}).
-Having controlled for characteristics of house ownership and rental prices (\texttt{PropertyOwned} and \texttt{RentLoan}, both of which have no effect), this effect is somewhat surprising.
+This is somewhat surprising as one might expect for house ownership and rental prices to be drivers of two-party preference, rather than household mobility (\texttt{PropertyOwned} and \texttt{RentLoan} are not significant in any election).
 
-Higher proportions of single parent households are associated with Labor support in all elections (albeit insignificant in 2004, see row 6 in Figure \ref{fig:partresplot2}), whereas the electoral family and household sizes (via the \texttt{FamHouseSize} variable) do not appear to be associated with either party in any election.
+Higher proportions of single parent households are associated with Labor support in all elections (albeit insignificant in 2004, see row 6 in Figure \ref{fig:partresplot2}), whereas the electoral family and household sizes (via the \texttt{FamHouseSize} variable) do not appear to be associated with either party.
 
@@
 
-From 2007 onwards, both marriages (\texttt{Married}) and de facto relationships (\texttt{DeFacto}) are found to be strong predictors of the two-party preferred vote in opposition to the Liberal party.
+From 2007 onwards, both marriages (\texttt{Married}) and de facto relationships (\texttt{DeFacto}) are found to be strong predictors of the two-party preferred vote in favour of the Labor party.
 In 2001 and 2004 neither of these variables are significant (see rows 4 and 5 in Figure \ref{fig:partresplot2}).
@@
 
-It is often hypothesized that states have systematic differences that cause their electorates to vote differently.
-Boxplots of residuals grouped by state (Figure \ref{fig:resstate}) reveal that the data reflects this to a limited extent.
-Tasmania and the Australian Capital Territory appear to have a bias towards Labor, whereas the Northern Territory tends towards voting Liberal.
-However, there are relatively few electorates in each of these states (five, two and two respectively), so this apparent result may be due to incumbent effects rather than an actual state-specific bias.
+It is often hypothesized that states have a systematic bias towards one of the two major parties.
+Boxplots of residuals grouped by state (Figure \ref{fig:resstate}) show that the data reflects this to only a limited extent.
+Tasmania and the Australian Capital Territory appear to have a bias towards Labor, whereas the South Australia and the Northern Territory tend towards voting Liberal.
+However, there are relatively few electorates in each of these states (five, two, eleven and two respectively), so this apparent result may be due to incumbent effects rather than an actual state-specific bias.
 
@@
 
-\hypertarget{checks}{%
-\subsubsection{Checks}\label{checks}}
+\hypertarget{multicollinearity}{%
+\subsubsection*{Multicollinearity}\label{multicollinearity}}
+\addcontentsline{toc}{subsubsection}{Multicollinearity}
 
@@
 
-Based on the distribution of the Cook's distance values, a Cook's distance greater than \(0.1\) is considered to be influential and a potential outlier.
-The electorate of Sydney (NSW) has a large Cook's distance from 2001 to 2010, due to its diverse population (language, birthplace and religion), high number of defacto relationships, high income, high household mobility and small amount of workers in extractive and transformative jobs.
-It has remained a strong supporter of the Labor party and the Liberal vote is severely overpredicted by the model, making it an outlier.
+Based on the distribution of the Cook's distance values and the distribution of hat values, a Cook's distance greater than \(0.1\) is considered to be influential, and a hat value greater than \(0.5\) is considered to have high leverage.
+Electorates fitting these criteria are flagged and investigated to examine the characteristics driving these values.
+
+The electorate of Sydney (NSW) has a large Cook's distance and high leverage from 2001 to 2007, due to its diverse population (languages, birthplace and religion), high density of young adults (20 to 34 years old), high number of defacto relationships, high income, high household mobility and small amount of workers in extractive and transformative jobs.
+It has remained a strong supporter of the Labor party and the extent of this support is underpredicted by the model, making it an outlier.
 Nearby in metropolitan NSW, the electorate of Wentworth is found to be an outlier in the 2013 and 2016 elections.
@@
 Family and household sizes have minimal influence, although electorates with more single parent households are linked with Labor support.
-Furthermore, the spatial effects are found to be positive in all elections and significant in 2001, 2004 and 2016, meaning that other characteristics that neighbours have in common appear to be influential in those years.
+Furthermore, the spatial effects are found to be positive in all elections and significant in 2001, 2004 and 2016, meaning that other characteristics that neighbours have in common (outside of the variables in the model) appear to be influential in those years.
 
@@
 
-The authors would like to sincerely thank the editor and associate editor of the Australian \& New Zealand Journal of Statistics, and the two anonymous reviewers for providing helpful comments and suggestions on earlier drafts of the manuscript.
+The authors would like to sincerely thank the editor and associate editor of the Australian \& New Zealand Journal of Statistics and the two anonymous reviewers for providing helpful comments and suggestions on earlier drafts of the manuscript.
 Additionally, the authors would like to thank Anthony Ebert, Heike Hofmann, Thomas Lumley, Ben Marwick, Carson Sievert, Mingzhu Sun, Dilini Talagala, Nicholas Tierney, Nathaniel Tomasetti, Earo Wang and Fang Zhou, all of whom have contributed to the \texttt{eechidna} package.

### 2019-11-10 response_for_ANZJS_editor.Rmd
minor edits

@@
 Some spatial imputation is still required as the SA1 boundaries do not exactly match the electorate boundaries.
-Even with this change in imputation method, our results are effectively the same as the previous submission.*
+Even with this change in imputation method, our results are effectively the same as the original submission.*
 
@@
 Census data is provided as a head count for categories for many different questions, and there are thousands of variables that could be constructed from the raw data.
-These 50 variables were chosen because they appear in previous studies of Australian elections (those mentioned in the Introduction).
+These 50 variables are chosen because they appear in previous studies of Australian elections (those mentioned in the Introduction).
 We hope this is now better explained.*
@@
 
-*This has been reconciled in Section 3.2.
+*This is reconciled in Section 3.2.
 We begin with 50 variables.
 Six factors are created using a total of 23 variables.
-One of the age brackets (populated aged 55 and over) is removed, and the other three age brackets are kept (to avoid multicollinearity).
+One of the age brackets (population aged 55 and over) is removed, and the other three age brackets are kept (to avoid multicollinearity).
 This results in a final predictor set of 32 variables.
@@
 *This was a very helpful suggestion to explore, and as such an additional section on robustness has been included in the revised paper (Section 4.4).
-When we fit a model using just the variables with significant effects, we find that the results are robust.
+When we fit a model using just the variables with significant effects, we find that the results are essentially unchanged.
 Each of the estimated coefficients from the reduced model (containing only the subset significant variables) lie within the 95% confidence interval from its corresponding full model (containing all 32 variables).*
@@
 For each pair, a model for each election is refit omitting one of the two variables.
-It is found that for each of these pairs, the estimated effect of the remaining variable in the reduced model lies within the $95/%$ confidence interval from the full model.*
+It is found that for each of these pairs, the estimated effect of the remaining variable in the reduced model lies within the 95% confidence interval from the full model.*
 
@@
 
-*The first three minor issues have been corrected. For the last issue, the spatial coefficient is now significant in 2001, 2004 and 2016, which do not correspond only to years with no imputation.*
+*The first three minor issues have been corrected. For the last issue, the spatial coefficient is now significant in 2001, 2004 and 2016, which no longer corresponds only to years without imputation.*
 
@@
 *We begin with 50 variables.
-Many of these variables represent similar information and where appropriate, should be grouped together.
-PCA is used to determine which variables covary, and hence decide which varibles should be combined into a factor.
+Many of these variables represent similar information and hence we argue, could be grouped together.
+PCA is used to determine which variables covary, and so we use it to decide variables should be combined into individual factors.
 Each of the first four PCs are examined to determine which variables have large loadings.
@@
 *We formed our approach to exploring this question by reviewing the Australian literature in this area, which has been referenced accordingly.
-As this is an exploratory piece, we wanted to let the relationships reveal themselves, rather than expicitly stating hypotheses to be tested.*
+As this is an exploratory piece, we wanted to let the data reveal the relationships themselves, rather than imposing and testing explicit hypotheses.*
 
@@
 Some spatial imputation is still required as the SA1 boundaries do not exactly match the electorate boundaries.
-With this change in imputation method, our results are effectively the same as the previous submission.*
+Despite these changes, our results are effectively the same as the initial submission.*
 
@@
 
-*Our approach is to combine the six election datasets together into a single dataset of 900 observations.
-We computed the principal components on this.
-The first four components are chosen because the amount of variance explained drops off after the fourth PC.
-We then use these four PCs to determine which variable covary, and hence decide which varibles should be combined into a factor.*
+*Our approach is to combine the six election datasets together into a single dataset of 900 observations, containing 50 socio-demographic variables.
+Many of these variables represent similar information and hence we argue, could be grouped together.
+PCA is used to determine which variables covary, and so we use it to guide the selection of variables to be combined into individual factors.
+Each of the first four PCs are examined to see which variables have large loadings.
+If there is an intuitive reason as to why a subset of these variables with large loadings (in a given PC) should be grouped together, then they are combined into a single factor using a weighted sum.
+The weights are simply +1 if the loading in that PC is positive, and -1 if it is negative.
+Hence, the factors are not the principal components themselves, rather they are each a weighted sum of a different subset of variables.*
 
-*Each of the first four PCs are examined to determine which variables have large loadings.
-If there is an intuitive reason as to why a subset of variables with large loadings (in a given PC) should be grouped together, then they are combined into a factor using a weighted sum.
-The weights are simply +1 is the loading in that PC is positive, and -1 if it is negative.
-Hence, the factors are not the principal components themselves.*
+*The subsets that comprise each of the six factors are explicitly detailed in paragraph four of Section 3.2.*
 
@@
 Our original method was chosen because we believed that the imputation method would not materially affect results.
-With this change in imputation method, our results are effectively the same as the previous submission.*
+Despite these changes, our results are effectively the same as the previous submission.*
 
@@
 Two factors that have been created from the same PC will be correlated, which could mean that (near) multicollinearity persists.
-However, we have conducted three robustness tests for multicollinearity (see Section 4.5), and no conclusive signs of multicollinearity were observed.*
+However, we have conducted three robustness tests for multicollinearity (see Section 4.5), and no conclusive signs of multicollinearity are observed.*
 

### 2019-11-10 anzjs_elections.Rmd
minor edits

@@
 Collecting and wrangling the raw data, along with the imputation process, are detailed in Section \@ref(data).
-All data and associated documentation relating to this procedure are available in the `eechidna` `R` package [@eechidna], providing a resource for future analysis.
+All data and associated documentation relating to this procedure are available in the `eechidna` `R` package [@eechidna], providing a resource for any future analysis.
 
@@
 First, variable scales change over the years, making it important to standardize variables.
-Second, many variables represent similar information and where appropriate, should be grouped together.
-To determine which variables should be grouped, principal component analysis (PCA) is used.
-The intuition here is that PCA will identify which variables covary, from which intuitive groupings of variables can be chosen to combine into a single variable.
+Second, many variables represent similar information and where appropriate, will be grouped together.
+To determine which variables should be grouped, principal component analysis (PCA) is used to guide the construction of specific factors.
+The intuition here is that PCA will identify which variables covary, from which intuitive groupings of variables can be chosen to combine into individual variables.
+Details are given in Section 3.2.
 After these steps, a model specification is chosen.
@@
 Most predictors have a positive relationship: the larger the value the more likely the electorate preferences the Coalition.
-The relationship is relatively robust over time, with the exception of `Incomes`, `OtherLanguageHome`, `Pop_00_19` and `DiffAddress`.
+The relationship is relatively robust over time, with the exception of `DiffAddress`, `Incomes`, `OtherLanguageHome` and `Pop_00_19`.
 
@@
 In 2001, 2004 and 2016, higher proportions of people that have recently (in the past five years) moved house (`DiffAddress`) increased electoral support for the Liberal party (see row 5 in Figure \@ref(fig:partresplot1)).
-This is somewhat surprising as one might expect for house ownership and rental prices to be drivers of two-party preference, rather than household mobility (`PropertyOwned` and `RentLoan` are not significant in any election).
+This is somewhat surprising as one might expect house ownership and rental prices to be linked to two-party preference, rather than household mobility (`PropertyOwned` and `RentLoan` are not significant in any election).
 
@@
 An incumbent advantage is evident and has not been captured by the model.", fig.asp=0.6, out.width="70%", fig.pos = 'h', out.extra = ''}
+# Load first preferences
+fp16 <- eechidna::fp16 %>% 
+  mutate(PartyAb = ifelse(PartyAb %in% c("LP", "NP"), "LNP", PartyAb))
+fp13 <- eechidna::fp13 %>% 
+  mutate(PartyAb = ifelse(PartyAb %in% c("LP", "NP"), "LNP", PartyAb))
+fp10 <- eechidna::fp10 %>% 
+  mutate(PartyAb = ifelse(PartyAb %in% c("LP", "NP"), "LNP", PartyAb))
+fp07 <- eechidna::fp07 %>% 
+  mutate(PartyAb = ifelse(PartyAb %in% c("LP", "NP"), "LNP", PartyAb))
+fp04 <- eechidna::fp04 %>% 
+  mutate(PartyAb = ifelse(PartyAb %in% c("LP", "NP"), "LNP", PartyAb))
+fp01 <- eechidna::fp01 %>% 
+  mutate(PartyAb = ifelse(PartyAb %in% c("LP", "NP"), "LNP", PartyAb))
+
 # Incumbents
@@
     filter(HistoricElected == "Y", PartyAb == "LNP") %>%
-    select(DivisionNm, HistoricElected)
+    dplyr::select(DivisionNm, HistoricElected)
 }
@@
   bind_rows(
-    glsmod16$my_data %>% select(DivisionNm, year) %>% left_join(incumbent(fp16), by = "DivisionNm"),
-    glsmod13$my_data %>% select(DivisionNm, year) %>% left_join(incumbent(fp13), by = "DivisionNm"),
-    glsmod10$my_data %>% select(DivisionNm, year) %>% left_join(incumbent(fp10), by = "DivisionNm"),
-    glsmod07$my_data %>% select(DivisionNm, year) %>% left_join(incumbent(fp07), by = "DivisionNm"),
-    glsmod04$my_data %>% select(DivisionNm, year) %>% left_join(
-      fp04 %>% filter(PartyAb == "LNP") %>%
+    glsmod16$my_data %>% dplyr::select(DivisionNm, year) %>% left_join(incumbent(fp16), by = "DivisionNm"),
+    glsmod13$my_data %>% dplyr::select(DivisionNm, year) %>% left_join(incumbent(fp13), by = "DivisionNm"),
+    glsmod10$my_data %>% dplyr::select(DivisionNm, year) %>% left_join(incumbent(fp10), by = "DivisionNm"),
+    glsmod07$my_data %>% dplyr::select(DivisionNm, year) %>% left_join(incumbent(fp07), by = "DivisionNm"),
+    glsmod04$my_data %>% dplyr::select(DivisionNm, year) %>% left_join(
+      fp04 %>% filter(PartyAb %in% c("LNP", "NP", "LP")) %>%
         left_join(
-          fp01 %>% rename(HistoricElected = Elected) %>% select(UniqueID, HistoricElected, PartyAb),
+          fp01 %>% rename(HistoricElected = Elected) %>% dplyr::select(UniqueID, HistoricElected, PartyAb),
           by = c("UniqueID", "PartyAb")
@@
 Other electorates with large Cook's distance are Canberra (ACT) and Durack (WA) in 2013, and Solomon (NT) in 2016.
+
 All of the electorates examined are not unduly influential in the model and therefore no action is required.
@@
 
-Additionally, a key contribution of this research is the wrangling of the raw data and imputed data sets for the 2004, 2007, 2010 and 2013 elections, which have been contributed to the `eechidna` `R` package --- providing a rich, accessible data resource for future Australian electoral analysis.
+Additionally, a key contribution of this research is the wrangling of the raw data and imputed data sets for the 2004, 2007, 2010 and 2013 elections, which have been contributed to the `eechidna` `R` package --- providing a rich, accessible data resource for any future Australian electoral analysis.
 

### 2019-11-10 anzjs_elections.tex
minor edits

@@
 Collecting and wrangling the raw data, along with the imputation process, are detailed in Section \ref{data}.
-All data and associated documentation relating to this procedure are available in the \texttt{eechidna} \texttt{R} package \citep{eechidna}, providing a resource for future analysis.
+All data and associated documentation relating to this procedure are available in the \texttt{eechidna} \texttt{R} package \citep{eechidna}, providing a resource for any future analysis.
 
@@
 Most predictors have a positive relationship: the larger the value the more likely the electorate preferences the Coalition.
-The relationship is relatively robust over time, with the exception of \texttt{Incomes}, \texttt{OtherLanguageHome}, \texttt{Pop\_00\_19} and \texttt{DiffAddress}.}\label{fig:partresplot1}
+The relationship is relatively robust over time, with the exception of \texttt{DiffAddress}, \texttt{Incomes}, \texttt{OtherLanguageHome} and \texttt{Pop\_00\_19}.}\label{fig:partresplot1}
 \end{figure}
@@
 In 2001, 2004 and 2016, higher proportions of people that have recently (in the past five years) moved house (\texttt{DiffAddress}) increased electoral support for the Liberal party (see row 5 in Figure \ref{fig:partresplot1}).
-This is somewhat surprising as one might expect for house ownership and rental prices to be drivers of two-party preference, rather than household mobility (\texttt{PropertyOwned} and \texttt{RentLoan} are not significant in any election).
+This is somewhat surprising as one might expect house ownership and rental prices to be linked to two-party preference, rather than household mobility (\texttt{PropertyOwned} and \texttt{RentLoan} are not significant in any election).
 
@@
 
-Additionally, a key contribution of this research is the wrangling of the raw data and imputed data sets for the 2004, 2007, 2010 and 2013 elections, which have been contributed to the \texttt{eechidna} \texttt{R} package --- providing a rich, accessible data resource for future Australian electoral analysis.
+Additionally, a key contribution of this research is the wrangling of the raw data and imputed data sets for the 2004, 2007, 2010 and 2013 elections, which have been contributed to the \texttt{eechidna} \texttt{R} package --- providing a rich, accessible data resource for any future Australian electoral analysis.
 
@@
 
-The authors would like to sincerely thank the editor and associate editor of the Australian \& New Zealand Journal of Statistics and the two anonymous reviewers for providing helpful comments and suggestions on earlier drafts of the manuscript.
+The authors would like to sincerely thank the editor and associate editor of the Australian \& New Zealand Journal of Statistics and the two anonymous reviewers for providing helpful comments and suggestions on earlier drafts of this manuscript.
 Additionally, the authors would like to thank Anthony Ebert, Heike Hofmann, Thomas Lumley, Ben Marwick, Carson Sievert, Mingzhu Sun, Dilini Talagala, Nicholas Tierney, Nathaniel Tomasetti, Earo Wang and Fang Zhou, all of whom have contributed to the \texttt{eechidna} package.

### 2019-11-14 anzjs_elections.Rmd
minor edits - removed margins for Rob to knit

@@
       in_header: anzjs_preamble.tex
-geometry: "left=5cm, right=0cm, top=4cm, bottom=0cm"
 ---
@@
 Furthermore, let $I_{s,t}$ be an indicator variable, for which $I_{s,t} = 1$ if the centroid of source zone $s$ falls within target zone $t$, and $0$ otherwise.
-Additionally, let the population of the source zone $s$ be $U_s$ and the population of target zone $t$ be $P_{t}$.
+Additionally, let the population of the source zone $s$ be $P_{s}$.
 
@@
 $$
-  \hat{C}_t = \frac{\sum_{s=1}^{S}{I_{s,t}*D_s*U_s}}%
-  {\sum_{s=1}^{S}{I_{s,t}*U_s}},
+  \hat{C}_t = \frac{\sum_{s=1}^{S}{I_{s,t}*D_s*P_s}}%
+  {\sum_{s=1}^{S}{I_{s,t}*P_s}},
   \quad\text{for each $t=1,\dots,T$}.
@@
 Any model that uses all variables would face problems with multi-collinearity and over-fitting, which would likely lead to erroneous conclusions regarding variable significance.
-To address this, variables that represent similar information are combined into a single variable, which will be referred to as a "factor".
+To address this, a subset of variables that represent similar information are combined into a single variable, which will be referred to as a "factor".
 
@@
 If a subset of variables have large loadings (positive or negative) in a given component, and there is an intuitive reason as to why they should be grouped together, then this subset of variables will be combined to become a factor.
-Note that more than one factor can be deduced from a principal component (i.e.\ two non-overlapping subsets of variables), and that any variables not included in a factor are not discarded.
+Note that more than one factor can be deduced from a principal component (i.e.\ multiple non-overlapping subsets of variables), and that any variables not included in a factor are not discarded.
 
@@
 
-[^3]: The factor creation procedure reduced the variable set to $p=33$, however one of the age brackets (`Pop_55_plus`) is not included as a variable to avoid multicollinearity, because the other three age brackets are included.
+[^3]: The factor creation procedure reduces the variable set to $p=33$, however one of the age brackets (`Pop_55_plus`) is not included as a variable to avoid multicollinearity, because the other three age brackets are included.
 
@@
 ```{r rhovis, fig.cap="Estimates of the spatial autoregressive parameter for each of the six elections, reported with their individual 95\\% confidence intervals.
-Only in 2001 and 2016 is there a significant spatial component.", out.height='20%', out.extra = ''}
+In 2001, 2004 and 2016 there is a significant spatial component.", out.height='20%', out.extra = ''}
 rho_df <- bind_rows(
@@
 Linear model with 95% confidence bands overlaid.
-Most predictors have a positive relationship: the larger the value the more likely the electorate preferences the Coalition.
+Most predictors have a positive relationship: the larger the value the more likely the electorate preferences Liberal.
 The relationship is relatively robust over time, with the exception of `DiffAddress`, `Incomes`, `OtherLanguageHome` and `Pop_00_19`.
@@
 
-Electorates with higher proportions of workers in mining, gas, water, agriculture, waste and electricity (grouped as `Extractive` industries) are consistently linked with higher support for the Liberal party, with the magnitude of this effect slightly increasing over the years (see row 3 in Figure \@ref(fig:partresplot1)).
+Electorates with higher proportions of workers in mining, gas, water, agriculture, waste and electricity (grouped as `Extractive` industries) are consistently linked with higher support for the Liberal party, with the magnitude of this effect slightly increasing over the years (see row 1 in Figure \@ref(fig:partresplot1)).
 This is unsurprising, as the Liberal party has close ties with these traditional energy industries, and typically present policies to reduce taxation on energy production.
-Furthermore, electorates with more workers in construction or manufacturing industries (`Transformative`) are also more likely to support the Liberal party (see row 4 in Figure \@ref(fig:partresplot1)), from 2004 onwards.
+Furthermore, electorates with more workers in construction or manufacturing industries (`Transformative`) are also more likely to support the Liberal party (see row 3 in Figure \@ref(fig:partresplot1)), from 2004 onwards.
 
@@
 Typically the Labor party campaigns on more progressive policies, which often include tax reform that adversely affects higher income earners, and more generous social assistance programs.
-Perhaps it is due to these policies that higher income electorates appear more likely to support the Liberal party, as the `Incomes` factor has a positive effect on Liberal preference (see row 1 in Figure \@ref(fig:partresplot1)).
+Perhaps it is due to these policies that higher income electorates appear more likely to support the Liberal party, as the `Incomes` factor has a positive effect on Liberal preference (see row 6 in Figure \@ref(fig:partresplot1)).
 This effect is significant in every election aside from 2004 and 2010.
@@
 Since 2007, electorates with higher education levels are associated with supporting the Labor party, with this effect being significant in 2007, 2013 and 2016 and only marginally insignificant in 2010.
-In the elections before 2007, education has a negligible effect (see row 3 in Figure \@ref(fig:partresplot1)).
+In the elections before 2007, education has a negligible effect (see row 3 in Figure \@ref(fig:partresplot2)).
 Additionally, student populations (`CurrentlyStudying`) do not affect electoral party preference in any election (not shown).
@@
 This imputes electoral socio-demographics for the electoral boundaries in place at the time of the election --- an approach that is distinctly different from previous work on modelling election outcomes, where Census and election data are typically joined without addressing their temporal differences.
-Before estimating a model, these socio-demographic variables are standardized (to adjust for changing variable scales) and many variables (representing similar information) are combined into factors, resulting in a reduced predictor set.
+Before estimating a model, these socio-demographic variables are standardized (to adjust for changing variable scales) and subsets of variables (representing similar information) are combined into factors, resulting in a reduced predictor set.
 A spatial error model is then estimated for each election, accounting for the inherent spatial structure of the data.

### 2019-11-14 anzjs_elections.tex
minor edits - removed margins for Rob to knit

@@
 Furthermore, let \(I_{s,t}\) be an indicator variable, for which \(I_{s,t} = 1\) if the centroid of source zone \(s\) falls within target zone \(t\), and \(0\) otherwise.
-Additionally, let the population of the source zone \(s\) be \(U_s\) and the population of target zone \(t\) be \(P_{t}\).
+Additionally, let the population of the source zone \(s\) be \(P_{s}\).
 
@@
 \[
-  \hat{C}_t = \frac{\sum_{s=1}^{S}{I_{s,t}*D_s*U_s}}%
-  {\sum_{s=1}^{S}{I_{s,t}*U_s}},
+  \hat{C}_t = \frac{\sum_{s=1}^{S}{I_{s,t}*D_s*P_s}}%
+  {\sum_{s=1}^{S}{I_{s,t}*P_s}},
   \quad\text{for each $t=1,\dots,T$}.
@@
 Any model that uses all variables would face problems with multi-collinearity and over-fitting, which would likely lead to erroneous conclusions regarding variable significance.
-To address this, variables that represent similar information are combined into a single variable, which will be referred to as a ``factor''.
+To address this, a subset of variables that represent similar information are combined into a single variable, which will be referred to as a ``factor''.
 
@@
 If a subset of variables have large loadings (positive or negative) in a given component, and there is an intuitive reason as to why they should be grouped together, then this subset of variables will be combined to become a factor.
-Note that more than one factor can be deduced from a principal component (i.e.~two non-overlapping subsets of variables), and that any variables not included in a factor are not discarded.
+Note that more than one factor can be deduced from a principal component (i.e.~multiple non-overlapping subsets of variables), and that any variables not included in a factor are not discarded.
 
@@
 
-The final predictor set contains \(p=32\) variables\footnote{The factor creation procedure reduced the variable set to \(p=33\), however one of the age brackets (\texttt{Pop\_55\_plus}) is not included as a variable to avoid multicollinearity, because the other three age brackets are included.} which are listed in Table \ref{tab:printmodels}.
+The final predictor set contains \(p=32\) variables\footnote{The factor creation procedure reduces the variable set to \(p=33\), however one of the age brackets (\texttt{Pop\_55\_plus}) is not included as a variable to avoid multicollinearity, because the other three age brackets are included.} which are listed in Table \ref{tab:printmodels}.
 
@@
 \caption{Estimates of the spatial autoregressive parameter for each of the six elections, reported with their individual 95\% confidence intervals.
-Only in 2001 and 2016 is there a significant spatial component.}\label{fig:rhovis}
+In 2001, 2004 and 2016 there is a significant spatial component.}\label{fig:rhovis}
 \end{figure}
@@
 Linear model with 95\% confidence bands overlaid.
-Most predictors have a positive relationship: the larger the value the more likely the electorate preferences the Coalition.
+Most predictors have a positive relationship: the larger the value the more likely the electorate preferences Liberal.
 The relationship is relatively robust over time, with the exception of \texttt{DiffAddress}, \texttt{Incomes}, \texttt{OtherLanguageHome} and \texttt{Pop\_00\_19}.}\label{fig:partresplot1}
@@
 
-Electorates with higher proportions of workers in mining, gas, water, agriculture, waste and electricity (grouped as \texttt{Extractive} industries) are consistently linked with higher support for the Liberal party, with the magnitude of this effect slightly increasing over the years (see row 3 in Figure \ref{fig:partresplot1}).
+Electorates with higher proportions of workers in mining, gas, water, agriculture, waste and electricity (grouped as \texttt{Extractive} industries) are consistently linked with higher support for the Liberal party, with the magnitude of this effect slightly increasing over the years (see row 1 in Figure \ref{fig:partresplot1}).
 This is unsurprising, as the Liberal party has close ties with these traditional energy industries, and typically present policies to reduce taxation on energy production.
-Furthermore, electorates with more workers in construction or manufacturing industries (\texttt{Transformative}) are also more likely to support the Liberal party (see row 4 in Figure \ref{fig:partresplot1}), from 2004 onwards.
+Furthermore, electorates with more workers in construction or manufacturing industries (\texttt{Transformative}) are also more likely to support the Liberal party (see row 3 in Figure \ref{fig:partresplot1}), from 2004 onwards.
 
@@
 Typically the Labor party campaigns on more progressive policies, which often include tax reform that adversely affects higher income earners, and more generous social assistance programs.
-Perhaps it is due to these policies that higher income electorates appear more likely to support the Liberal party, as the \texttt{Incomes} factor has a positive effect on Liberal preference (see row 1 in Figure \ref{fig:partresplot1}).
+Perhaps it is due to these policies that higher income electorates appear more likely to support the Liberal party, as the \texttt{Incomes} factor has a positive effect on Liberal preference (see row 6 in Figure \ref{fig:partresplot1}).
 This effect is significant in every election aside from 2004 and 2010.
@@
 Since 2007, electorates with higher education levels are associated with supporting the Labor party, with this effect being significant in 2007, 2013 and 2016 and only marginally insignificant in 2010.
-In the elections before 2007, education has a negligible effect (see row 3 in Figure \ref{fig:partresplot1}).
+In the elections before 2007, education has a negligible effect (see row 3 in Figure \ref{fig:partresplot2}).
 Additionally, student populations (\texttt{CurrentlyStudying}) do not affect electoral party preference in any election (not shown).

### 2020-04-01 anzjs_elections.Rmd
Formatting changes made for ANZJS print. Includes: using \textsf \url, increase plot size for rho and intercept (and increase font size), embed footnotes in text, fix "standardise",

@@
 Socio-demographic information is derived from the Australian Census which occurs every five years.
-Since a Census is not directly available for each election, an imputation method is employed to estimate Census data for the electorates at the time of each election.
-This accounts for both spatial and temporal changes in electoral characteristics between Censuses.
+Since a census is not directly available for each election, an imputation method is employed to estimate census data for the electorates at the time of each election.
+This accounts for both spatial and temporal changes in electoral characteristics between censuses.
 To capture any spatial heterogeneity, a spatial error model is estimated for each election, which incorporates a spatially structured random effect vector.
@@
 Education and unemployment are amongst those that have varying effects.
-All data featured in this study has been contributed to the `eechidna` `R` package (available on CRAN).
-
-**Keywords:** federal election, Census, Australia, spatial modelling, imputation, data science, socio-demographics, electorates, R, eechidna
+All data featured in this study has been contributed to the \textsf{eechidna} \textsf{R} package (available on CRAN).
+
+**Keywords:** federal election, census, Australia, spatial modelling, imputation, data science, socio-demographics, electorates, R, eechidna
 
@@
 
-# Combine and standardize
+# Combine and standardise
 my_df <- bind_rows(
@@
 
-# Now standardize factors
+# Now standardise factors
 
@@
 Joining these two data sources is problematic as there is an inherent asynchronicity in the two types of events.
-A Census is conducted by the Australian Bureau of Statistics (ABS) every five years, whereas federal elections, conducted by the AEC, usually occur every three years or so.
-The first problem addressed is that of constructing appropriate Census data for the 2004, 2007, 2010 and 2013 elections --- election years in which a Census did not occur.
-The predominant approach in previous studies was to join voting outcomes to the nearest Census, without accounting for any temporal differences [see @DavisStimson98;@Stimson06;@Liao09;@Stimson09].
+A census is conducted by the Australian Bureau of Statistics (ABS) every five years, whereas federal elections, conducted by the AEC, usually occur every three years or so.
+The first problem addressed is that of constructing appropriate census data for the 2004, 2007, 2010 and 2013 elections---election years in which a census did not occur.
+The predominant approach in previous studies was to join voting outcomes to the nearest census, without accounting for any temporal differences [see @DavisStimson98;@Stimson06;@Liao09;@Stimson09].
 Furthermore, electoral boundaries change regularly, so spatial discrepancies also arise when matching with electoral data.
-To obtain appropriate "Census-like" data for these four elections, electoral socio-demographics are constructed using a spatio-temporal imputation that combines areal interpolation [@Goodchild1993] and linear time-interpolation.
+To obtain appropriate 'census-like' data for these four elections, electoral socio-demographics are constructed using a spatio-temporal imputation that combines areal interpolation [@Goodchild1993] and linear time-interpolation.
 Collecting and wrangling the raw data, along with the imputation process, are detailed in Section \@ref(data).
-All data and associated documentation relating to this procedure are available in the `eechidna` `R` package [@eechidna], providing a resource for any future analysis.
+All data and associated documentation relating to this procedure are available in the \textsf{eechidna} \textsf{R} package [@eechidna], providing a resource for any future analysis.
 
 Previous work on modelling Australian federal elections has found that aggregate socio-demographics are relatively good predictors of voting outcomes.
-@Forrest01 used multiple regression to model the Liberal and Labor primary vote for polling booths in the Farrer electorate in 1998 as a function of Census variables from 1996.
+@Forrest01 used multiple regression to model the Liberal and Labor primary vote for polling booths in the Farrer electorate in 1998 as a function of census variables from 1996.
 @Stimson06, @Stimson09 and @Stimson12 used principal component analysis of polling booths in the 2001, 2004 and 2007 elections respectively, also finding that socio-demographic characteristics of polling booths are linked to their two-party preferred vote.
-In contrast, @Stimson09 models the polling booth swing vote (change in the two-party preferred vote) in the 2007 election, finding that little of the swing vote can be explained by Census data.
+In contrast, @Stimson09 models the polling booth swing vote (change in the two-party preferred vote) in the 2007 election, finding that little of the swing vote can be explained by census data.
 
@@
 This dataset consists of the two-party preferred vote (as the response variable) and a set of common socio-demographic variables (as the explanatory variables).<!--The use of a regression framework to examine these socio-political relationships over time is seemingly absent from previous Australian studies.
-It also appears that no study has attempted any type of statistical analysis of socio-demographics in conjunction with voter behaviour in Australia since 2007, making this paper distinctly different from previous work.--> To prepare these datasets, socio-demographic variables are first standardized, and then a principal component analysis is used to group many of the variables into "factors".
+It also appears that no study has attempted any type of statistical analysis of socio-demographics in conjunction with voter behaviour in Australia since 2007, making this paper distinctly different from previous work.--> To prepare these datasets, socio-demographic variables are first standardised, and then a principal component analysis is used to group many of the variables into 'factors'.
 To account for the inherent spatial structure of the data, a spatial error model is then estimated for each election.
@@
 
-Socio-demographic variables are derived from the Australian Census of Population and Housing (Census), which is a survey of every household in Australia, recording information such as age, gender, ethnicity, education level and income.
-There have been four Censuses so far in the 21st century, conducted in 2001, 2006, 2011 and 2016.
-The Australian Bureau of Statistics (ABS) conducts the Census and publishes aggregated information.
-The ABS uses electoral boundaries as defined by the AEC at the time of each Census, which may not match those in place at the subsequent and previous elections.
-From the available Census information aggregated at the electorate level, 50 socio-demographic variables are defined for each of the electorates to be used in the analysis.
+Socio-demographic variables are derived from the Australian Census of Population and Housing (census), which is a survey of every household in Australia, recording information such as age, gender, ethnicity, education level and income.
+There have been four censuses so far in the 21st century, conducted in 2001, 2006, 2011 and 2016.
+The Australian Bureau of Statistics (ABS) conducts the census and publishes aggregated information.
+The ABS uses electoral boundaries as defined by the AEC at the time of each census, which may not match those in place at the subsequent and previous elections.
+From the available census information aggregated at the electorate level, 50 socio-demographic variables are defined for each of the electorates to be used in the analysis.
 These variables include information relating to electoral age distributions, income, education qualifications, employment industries and job types, religion, birthplace, household characteristics and relationships.
 
-Raw data is sourced online from the AEC and ABS websites in `.csv` and `.xlsx` files.
+Raw data is sourced online from the AEC and ABS websites in \textsf{.csv} and \textsf{.xlsx} files.
 The formats of these files differ over the years, making extracting the appropriate information a big task.
-The functions available in the `dplyr` [@dplyr] and `readxl` [@readxl] `R` packages are particularly useful, as they provide fast consistent tools for data manipulation and functions to import `.xlsx` files.
-The 2001 and 2006 Census data are published in a format where the information for each electorate is held in a separate document making it difficult to use the `dplyr` tools.
+The functions available in the \textsf{dplyr} [@dplyr] and \textsf{readxl} [@readxl] \textsf{R} packages are particularly useful, as they provide fast consistent tools for data manipulation and functions to import \textsf{.xlsx} files.
+The 2001 and 2006 census data are published in a format where the information for each electorate is held in a separate document making it difficult to use the \textsf{dplyr} tools.
 Instead, cells have to be selected from each individual file to construct the desired variables.
-All scripts required for the data wrangling process can be found in the github repository for the `eechidna` `R` package [@eechidna], along with the raw data.
-The `eechidna` package makes this study entirely reproducible and provides a resource to help wrangle data for future Censuses and elections, when they become available.
-
-## Joining Census and election data
-
-### Differences between Census and election data {-}
-
-Between 2001 and 2016 there were six elections and four Censuses (see Figure \@ref(fig:timeline)).
-Electoral boundaries are redistributed regularly by the AEC, meaning that only in the years where both a Census and an election occur are all boundaries likely to match --- the case for the 2001 and 2016 elections.
-Therefore, for the four elections between 2004 and 2013, both temporal and spatial differences in electorates need to be accounted for when joining the electoral two-party preferred vote with Census data.
+All scripts required for the data wrangling process can be found in the github repository for the \textsf{eechidna} \textsf{R} package [@eechidna], along with the raw data.
+The \textsf{eechidna} package makes this study entirely reproducible and provides a resource to help wrangle data for future censuses and elections, when they become available.
+
+## Joining census and election data
+
+### Differences between census and election data {-}
+
+Between 2001 and 2016 there were six elections and four censuses (see Figure \@ref(fig:timeline)).
+Electoral boundaries are redistributed regularly by the AEC, meaning that only in the years where both a census and an election occur are all boundaries likely to match---the case for the 2001 and 2016 elections.
+Therefore, for the four elections between 2004 and 2013, both temporal and spatial differences in electorates need to be accounted for when joining the electoral two-party preferred vote with census data.
 For these elections a spatio-temporal imputation method is employed to obtain electoral socio-demographics.
-This method uses Census information from both before and after the election of interest.
-
-```{r timeline, echo = FALSE, fig.cap="Timeline of Australian elections and Censuses. They do not always occur in the same year.", out.width = "90%", fig.align="center"}
+This method uses census information from both before and after the election of interest.
+
+```{r timeline, echo = FALSE, fig.cap="Timeline of Australian elections and censuses. They do not always occur in the same year.", out.width = "90%", fig.align="center"}
 timeline <- data.frame(year = c(2001:2016)) %>%
@@
 
-For each election, neighbouring Census information has to be combined in some way so that it represents the boundaries in place at the time of the election.
-This is done by taking the electoral boundaries and imputing the corresponding socio-demographic characteristics for each of the neighbouring Censuses, thereby addressing the spatial aspect.
-Next, to deal with the temporal component, characteristics at the time of the election are constructed using linear interpolation between the spatially imputed neighbouring Census variables.
-
-The finest level of disaggregation available for Census data is the region classification called Statistical Area 1 (SA1).
+For each election, neighbouring census information has to be combined in some way so that it represents the boundaries in place at the time of the election.
+This is done by taking the electoral boundaries and imputing the corresponding socio-demographic characteristics for each of the neighbouring censuses, thereby addressing the spatial aspect.
+Next, to deal with the temporal component, characteristics at the time of the election are constructed using linear interpolation between the spatially imputed neighbouring census variables.
+
+The finest level of disaggregation available for census data is the region classification called Statistical Area 1 (SA1).
 In 2016, Australia was divided into over 55,000 SA1s.
@@
 For simplicity, let each source zone be wholly summarised by its centroid.
-A set of target zones, $t = 1,\dots,T$, are defined as regions for which information is to be imputed --- these are the electoral boundaries for a particular election.
+A set of target zones, $t = 1,\dots,T$, are defined as regions for which information is to be imputed---these are the electoral boundaries for a particular election.
 
 Take the example of the Melbourne Ports electorate from the 2013 federal election, illustrated in Figure \@ref(fig:melbports13).
-The purple region in this figure represents the target zone and the source zones are the centroid locations from the 2016 Census SA1 areas.
-
-```{r melbports13, fig.cap = "The electoral boundaries for Melbourne Ports (shown in purple) and surrounding electorates, with centroids for Statistical Area 1 regions from the 2016 Census overlaid.
+The purple region in this figure represents the target zone and the source zones are the centroid locations from the 2016 census SA1 areas.
+
+```{r melbports13, fig.cap = "The electoral boundaries for Melbourne Ports (shown in purple) and surrounding electorates, with centroids for Statistical Area 1 regions from the 2016 census overlaid.
 The centroids falling within the purple region are attributed to Melbourne Ports.", out.height="28%"}
@@
 In order to calculate socio-demographic information for each of the target zones, a weighted average of source zones is taken using their populations as weights.
-Denote a given Census variable for the target zone by $C_t$, and the same Census variable for the source zone as $D_s$.
+Denote a given census variable for the target zone by $C_t$, and the same census variable for the source zone as $D_s$.
 Then, estimate $C_t$ using
 $$
-  \hat{C}_t = \frac{\sum_{s=1}^{S}{I_{s,t}*D_s*P_s}}%
-  {\sum_{s=1}^{S}{I_{s,t}*P_s}},
+  \hat{C}_t = \frac{\sum_{s=1}^{S}{I_{s,t} \times D_s \times P_s}}%
+  {\sum_{s=1}^{S}{I_{s,t} \times P_s}},
   \quad\text{for each $t=1,\dots,T$}.
@@
 
-This concludes the spatial imputation of the socio-demographic characteristics for one target zone (a single electoral boundary), at the time of only one of the neighbouring Censuses.
-This process is repeated for all of the target zones, and then for the other neighbouring Census.
+This concludes the spatial imputation of the socio-demographic characteristics for one target zone (a single electoral boundary), at the time of only one of the neighbouring censuses.
+This process is repeated for all of the target zones, and then for the other neighbouring census.
 
@@
 
-To account for temporal changes, linear interpolation is used between Census years to get the final estimate of a Census variable for the target zone in the election year.
-Let $y_1$ be the year of the Census preceding an election, let $y_2$ be the year of the election, and $y_3$ be the year of the Census that follows.
-Add this year subscript to the Census variable estimate $\hat{C}_t$, resulting in $\hat{C}_{t,y}$.
-Linear interpolating between these Census years results an imputed value for the election year, given by
+To account for temporal changes, linear interpolation is used between census years to get the final estimate of a census variable for the target zone in the election year.
+Let $y_1$ be the year of the census preceding an election, let $y_2$ be the year of the election, and $y_3$ be the year of the census that follows.
+Add this year subscript to the census variable estimate $\hat{C}_t$, resulting in $\hat{C}_{t,y}$.
+Linear interpolating between these census years results an imputed value for the election year, given by
 $$
@@
 
-Continuing with the example of Melbourne Ports in the 2013 election, the estimate for a given Census variable in 2016, $\hat{C}_{\text{MelbPorts}, 2016}$ would be obtained by computing the weighted average of this variable amongst the SA1s within the purple region shown in Figure \@ref(fig:melbports13).
-This would be repeated with the 2011 Census SA1s to obtain $\hat{C}_{\text{MelbPorts}, 2011}$, from which the final estimate is given by
+Continuing with the example of Melbourne Ports in the 2013 election, the estimate for a given census variable in 2016, $\hat{C}_{\text{MelbPorts}, 2016}$ would be obtained by computing the weighted average of this variable amongst the SA1s within the purple region shown in Figure \@ref(fig:melbports13).
+This would be repeated with the 2011 census SA1s to obtain $\hat{C}_{\text{MelbPorts}, 2011}$, from which the final estimate is given by
 $$
@@
 Before choosing an appropriate model, two issues with the socio-demographic variables need to be addressed.
-First, variable scales change over the years, making it important to standardize variables.
+First, variable scales change over the years, making it important to standardise variables.
 Second, many variables represent similar information and where appropriate, will be grouped together.
@@
 For example, inflation-adjusted median rental prices increased across almost all electorates, with median rent of 225 dollars per week placing an electorate in the 90th percentile in 2001, but only the 45th percentile in 2016.
-In order for socio-demographic effects to be comparable across years, all explanatory variables are standardized to have mean zero and variance one within each election year.
+In order for socio-demographic effects to be comparable across years, all explanatory variables are standardised to have mean zero and variance one within each election year.
 By standardizing, each variable is reported as a relative measure compared to all other electorates in the same year.
+(Note that the log values were standardised for the variables `Judaism`, `Indigenous`, `Islam` and `Buddhism`.)
 
@@
 Any model that uses all variables would face problems with multi-collinearity and over-fitting, which would likely lead to erroneous conclusions regarding variable significance.
-To address this, a subset of variables that represent similar information are combined into a single variable, which will be referred to as a "factor".
+To address this, a subset of variables that represent similar information are combined into a single variable, which will be referred to as a 'factor'.
 
 A factor is created from a group of variables if there is an intuitive reason as to why these variables should represent similar information and if there is evidence to suggest that they covary.
-For example, a potential group would be variables relating to electoral incomes --- median family, household and personal incomes.
-To determine which variables covary, principal component analysis is used on a combined dataset of socio-demographic variables from all six elections[^2].
+For example, a potential group would be variables relating to electoral incomes---median family, household and personal incomes.
+To determine which variables covary, principal component analysis is used on a combined dataset of socio-demographic variables from all six elections.
 The only variables exempted from the principal component analysis are the four variables representing age brackets (the proportion of the population aged 0--19 years old, 20--34 years old, 45--54 years old and 55 years plus), which are included in the model as separate variables.
-
-[^2]: It is appropriate to compute principal components on a combined dataset of all six elections because when computed separately for each election, scree plots level off after four components and the loadings of the first four components are similar across the elections.
 
@@
 Variables with a positive loading are allocated a weight of $+1$ and those with negative loadings are allocated a weight of $-1$.
-After computing these weighted sums, the factor is standardized to have mean zero and variance one, within each election.
-
-The final predictor set contains $p=32$ variables[^3] which are listed in Table \ref{tab:printmodels}.
-
-[^3]: The factor creation procedure reduces the variable set to $p=33$, however one of the age brackets (`Pop_55_plus`) is not included as a variable to avoid multicollinearity, because the other three age brackets are included.
+After computing these weighted sums, the factor is standardised to have mean zero and variance one, within each election.
+
+The final predictor set contains $p=32$ variables which are listed in Table \ref{tab:printmodels}.
+(Note that the factor creation procedure reduces the variable set to $p=33$, however the `Pop_55_plus` age bracket is not included as a variable to avoid multicollinearity, because the other three age brackets are included.)
 
@@
 Let $\rho$ be the spatial autoregressive coefficient, $\bm{v}$ be a spherical error term, $\bm{W}$ be a matrix of spatial weights (containing information about the neighbouring regions), $\bm{X}$ be a matrix of socio-demographic covariates, $\bm{\beta}$ be a vector of regression coefficients and $\bm{a}$ be a spatially structured random effect vector.
+
+Let
 $$
@@
 $$
-Estimation of the above spatial error model is undertaken using feasible generalized least squares.
+Estimation of the above spatial error model is undertaken using feasible generalised least squares.
 
@@
 ```{r rhovis, fig.cap="Estimates of the spatial autoregressive parameter for each of the six elections, reported with their individual 95\\% confidence intervals.
-In 2001, 2004 and 2016 there is a significant spatial component.", out.height='20%', out.extra = ''}
+In 2001, 2004 and 2016 there is a significant spatial component.", out.height='25%', out.extra = ''}
 rho_df <- bind_rows(
@@
   ggplot() +
-  geom_point(aes(x = year, y = estimate), size = 3) +
-  geom_linerange(aes(x = year, ymin = lower95, ymax = upper95), size = 1.2) +
+  geom_point(aes(x = year, y = estimate), size = 4) +
+  geom_linerange(aes(x = year, ymin = lower95, ymax = upper95), size = 1.5) +
   geom_hline(aes(yintercept = 0), alpha = 0.5, size = 1) +
   scale_color_manual(values = c("grey50", "black")) +
-  labs(x = "Election year", y = "Estimate of spatial autoregressive parameter") +
-  lims(y = c(-1, 1)) +
+  labs(x = "Election year", y = "Estimate of spatial \nautoregressive parameter") +
+  lims(y = c(-0.2, 1)) +
   guides(col = FALSE) +
-  coord_fixed(ratio = 1) +
-  theme_bw()
+  coord_fixed(ratio = 1.5) +
+  theme_bw() +
+  theme(text = element_text(size = 14))
 ```
@@
 
-```{r plotintercept, fig.cap = "Estimated intercept for each election, which represents the two-party preferred vote for an electorate with mean characteristics.", out.height="20%", out.extra = ''}
+```{r plotintercept, fig.cap = "Estimated intercept for each election, which represents the two-party preferred vote for an electorate with mean characteristics.", out.height="25%", out.extra = ''}
 intercept_df <- bind_rows(
@@
   ggplot() +
-  geom_point(aes(x = year, y = estimate), size = 3) +
-  geom_linerange(aes(x = year, ymin = lower95, ymax = upper95), size = 1.2) +
+  geom_point(aes(x = year, y = estimate), size = 4) +
+  geom_linerange(aes(x = year, ymin = lower95, ymax = upper95), size = 1.5) +
   geom_hline(aes(yintercept = 50), alpha = 0.5, size = 1) +
@@
   lims(y = c(45, 55)) +
-  theme_bw()
-```
-
-Since all socio-demographics have been standardized to have a mean of zero and a variance of one, the intercept in each model can be interpreted as the estimated two-party preferred vote for an electorate with mean characteristics[^4].
+  theme_bw() +
+  theme(text = element_text(size = 14))
+```
+
+Since all socio-demographics have been standardised to have a mean of zero and a variance of one, the intercept in each model can be interpreted as the estimated two-party preferred vote for an electorate with mean characteristics (aside from `Judaism`, `Indigenous`, `Islam` and `Buddhism`, where it assumes the mean of the log value).
 Figure \@ref(fig:plotintercept) shows that the baseline of party preference has varied over the elections, with the biggest swing occurring in the 2007 election where the mean electorate shifted more than five percentage points in favour of the Labor party.
-
-[^4]: Mean of all variables aside from Judaism, Indigenous, Islam and Buddhism, where it assumes the mean of the log value.
 
@@
 
-It is often hypothesized that states have a systematic bias towards one of the two major parties.
+It is often hypothesised that states have a systematic bias towards one of the two major parties.
 Boxplots of residuals grouped by state (Figure \@ref(fig:resstate)) show that the data reflects this to only a limited extent.
@@
 The incumbent party appears to have a distinct advantage at the next election.
-The boxplots in Figure \@ref(fig:resincumbent) show that if either of the Labor or Liberal parties won the seat at the previous election, the electorate is likely to vote in their favour, over and above any socio-demographic effects --- this effect has not been captured by the model.
+The boxplots in Figure \@ref(fig:resincumbent) show that if either of the Labor or Liberal parties won the seat at the previous election, the electorate is likely to vote in their favour, over and above any socio-demographic effects---this effect has not been captured by the model.
 
@@
 Nearby in metropolitan NSW, the electorate of Wentworth is found to be an outlier in the 2013 and 2016 elections.
-Although historically Liberal, its two-party vote jumped by over 10 percentage points in 2010 without experiencing any notable changes in its socio-demographic makeup --- implying that this may be the direct effect of its Liberal member, Malcolm Turnbull, becoming the leader of the Liberal party.
+Although historically Liberal, its two-party vote jumped by over 10 percentage points in 2010 without experiencing any notable changes in its socio-demographic makeup---implying that this may be the direct effect of its Liberal member, Malcolm Turnbull, becoming the leader of the Liberal party.
 In the elections since, the model underpredicts Wentworth's Liberal support.
@@
 
-This paper explores the effects of electoral socio-demographic characteristics on the two-party preferred vote in the 2001--2016 elections, using information from the corresponding Australian federal elections and Censuses.
-As a Census does not always occur in the same year as an election, Census data for the 2004--2013 elections are generated by employing a method of spatio-temporal imputation.
-This imputes electoral socio-demographics for the electoral boundaries in place at the time of the election --- an approach that is distinctly different from previous work on modelling election outcomes, where Census and election data are typically joined without addressing their temporal differences.
-Before estimating a model, these socio-demographic variables are standardized (to adjust for changing variable scales) and subsets of variables (representing similar information) are combined into factors, resulting in a reduced predictor set.
+This paper explores the effects of electoral socio-demographic characteristics on the two-party preferred vote in the 2001--2016 elections, using information from the corresponding Australian federal elections and censuses.
+As a census does not always occur in the same year as an election, census data for the 2004--2013 elections are generated by employing a method of spatio-temporal imputation.
+This imputes electoral socio-demographics for the electoral boundaries in place at the time of the election---an approach that is distinctly different from previous work on modelling election outcomes, where census and election data are typically joined without addressing their temporal differences.
+Before estimating a model, these socio-demographic variables are standardised (to adjust for changing variable scales) and subsets of variables (representing similar information) are combined into factors, resulting in a reduced predictor set.
 A spatial error model is then estimated for each election, accounting for the inherent spatial structure of the data.
@@
 The findings in this paper complement the existing literature by modelling temporal trends, which as far as the authors are aware, has not been done previously for Australian elections using a regression framework.
-It is also the first study to model any Australian election since 2010 using Census information.
-
-Additionally, a key contribution of this research is the wrangling of the raw data and imputed data sets for the 2004, 2007, 2010 and 2013 elections, which have been contributed to the `eechidna` `R` package --- providing a rich, accessible data resource for any future Australian electoral analysis.
+It is also the first study to model any Australian election since 2010 using census information.
+
+Additionally, a key contribution of this research is the wrangling of the raw data and imputed data sets for the 2004, 2007, 2010 and 2013 elections, which have been contributed to the \textsf{eechidna} \textsf{R} package---providing a rich, accessible data resource for any future Australian electoral analysis.
 
@@
 
-This paper was produced using `RMarkdown` [@rmarkdown] and `knitr` [@knitr].
-All corresponding code for this paper can be found in the github repository [github.com/jforbes14/eechidna-paper](https://github.com/jforbes14/eechidna-paper), and the data used is available in the `eechidna` package [@eechidna].
+This paper was produced using \textsf{RMarkdown} [@rmarkdown] and \textsf{knitr} [@knitr].
+All corresponding code for this paper can be found in the github repository \url{github.com/jforbes14/eechidna-paper}, and the data used is available in the \textsf{eechidna} package [@eechidna].
 All raw data was obtained from the Australian Electoral Commission, the Australian Bureau of Statistics and the Australian Government.
@@
 The authors would like to sincerely thank the editor and associate editor of the Australian & New Zealand Journal of Statistics and the two anonymous reviewers for providing helpful comments and suggestions on earlier drafts of this manuscript.
-Additionally, the authors would like to thank Anthony Ebert, Heike Hofmann, Thomas Lumley, Ben Marwick, Carson Sievert, Mingzhu Sun, Dilini Talagala, Nicholas Tierney, Nathaniel Tomasetti, Earo Wang and Fang Zhou, all of whom have contributed to the `eechidna` package.
+Additionally, the authors would like to thank Anthony Ebert, Heike Hofmann, Thomas Lumley, Ben Marwick, Carson Sievert, Mingzhu Sun, Dilini Talagala, Nicholas Tierney, Nathaniel Tomasetti, Earo Wang and Fang Zhou, all of whom have contributed to the \textsf{eechidna} package.
 
@@
 
-All election and Census datasets, along with electoral maps and more, are available in the `eechidna` (Exploring Election and Census Highly Informative Data Nationally for Australia) `R` package, which can be downloaded from CRAN.
-The `eechidna` package makes it easy to look at the data from the Australian Federal elections and Censuses that occurred between 2001 and 2016.
-This study contributed a large revision to the `eechidna` package, which included the addition of election and Census data for 2001--2010, voting outcomes for polling booths and imputed Census data for election years.
-For more details on using `eechidna`, please see the articles (vignettes) on the github page [ropenscilabs.github.io/eechidna/](https://ropenscilabs.github.io/eechidna/).
+All election and census datasets, along with electoral maps and more, are available in the \textsf{eechidna} (Exploring Election and Census Highly Informative Data Nationally for Australia) \textsf{R} package, which can be downloaded from CRAN.
+The \textsf{eechidna} package makes it easy to look at the data from the Australian Federal elections and censuses that occurred between 2001 and 2016.
+This study contributed a large revision to the \textsf{eechidna} package, which included the addition of election and census data for 2001--2010, voting outcomes for polling booths and imputed census data for election years.
+For more details on using \textsf{eechidna}, please see the articles (vignettes) on the github page \url{ropenscilabs.github.io/eechidna}.
 

### 2020-04-01 anzjs_elections.Rmd
Results and Conclusion changed to past tense - at the request of the formatting/technical editor

@@
 
-The spatial autoregressive coefficient $\rho$ is positive and significant in the 2001, 2004 and 2016 elections (Figure \@ref(fig:rhovis)).
-In these three elections, there is evidence to suggest that neighbours share some influential characteristics outside the explanatory variables, which affect the two-party preferred vote.
-Conversely, in the other three elections, the spatial effect weakens to become insignificant (although still positive).
+The spatial autoregressive coefficient $\rho$ was positive and significant in the 2001, 2004 and 2016 elections (Figure \@ref(fig:rhovis)).
+In these three elections, there is evidence to suggest that neighbours shared some influential characteristics outside the explanatory variables, which affected the two-party preferred vote.
+Conversely, in the other three elections, the spatial effect is weaker and insignificant (although still positive).
 
@@
 
-Since all socio-demographics have been standardised to have a mean of zero and a variance of one, the intercept in each model can be interpreted as the estimated two-party preferred vote for an electorate with mean characteristics (aside from `Judaism`, `Indigenous`, `Islam` and `Buddhism`, where it assumes the mean of the log value).
-Figure \@ref(fig:plotintercept) shows that the baseline of party preference has varied over the elections, with the biggest swing occurring in the 2007 election where the mean electorate shifted more than five percentage points in favour of the Labor party.
+Since all socio-demographics were standardised to have a mean of zero and a variance of one, the intercept in each model can be interpreted as the estimated two-party preferred vote for an electorate with mean characteristics (aside from `Judaism`, `Indigenous`, `Islam` and `Buddhism`, where it assumes the mean of the log value).
+Figure \@ref(fig:plotintercept) shows that the baseline of party preference varied over the elections, with the biggest swing occurring in the 2007 election where the mean electorate shifted more than five percentage points in favour of the Labor party.
 
@@
 Linear model with 95% confidence bands overlaid.
-Most predictors have a positive relationship: the larger the value the more likely the electorate preferences Liberal.
-The relationship is relatively robust over time, with the exception of `DiffAddress`, `Incomes`, `OtherLanguageHome` and `Pop_00_19`.
+Most predictors had a positive relationship: the larger the value the more likely the electorate preferenced Liberal.
+The relationships were relatively robust over time, with the exception of `DiffAddress`, `Incomes`, `OtherLanguageHome` and `Pop_00_19`.
 
@@
 Linear model with 95% confidence bands overlaid.
-Several predictors have a negative relationship: with larger values indicating the electorate more likely preferences Labor.
-Most relationships are relatively stable over elections, except `Unemployment` and `Education`."
+Several predictors had a negative relationship: with larger values indicating the electorate more likely preferenced Labor.
+Most relationships were relatively stable over elections, except `Unemployment` and `Education`."
 
@@
 
-To investigate the socio-demographics that have a strong effect on the two-party preferred vote, partial residual plots are used and shown in Figures \@ref(fig:partresplot1) and \@ref(fig:partresplot2).
+To investigate the socio-demographics that had a strong effect on the two-party preferred vote, partial residual plots are used and shown in Figures \@ref(fig:partresplot1) and \@ref(fig:partresplot2).
 Partial residuals, for a given variable, are the residuals from the fitted model with the estimated effect of that variable added to it.
@@
 In each plot, the slope of the prediction line matches the estimated coefficient and the shaded region represents a 95% confidence band.
-Plots are computed using the method in @visreg.
+Plots were computed using the method in @visreg.
 If a horizontal line can be drawn through the confidence band, then the effect is insignificant.
@@
 Plots for each election are faceted to compare the effects over time in Figures \@ref(fig:partresplot1) and \@ref(fig:partresplot2).
-Only socio-demographics that have a significant effect in at least two elections are displayed in Figures \@ref(fig:partresplot1) and \@ref(fig:partresplot2).
+Only socio-demographics that had a significant effect in at least two elections are displayed in Figures \@ref(fig:partresplot1) and \@ref(fig:partresplot2).
 
@@
 
-Electorates with higher proportions of workers in mining, gas, water, agriculture, waste and electricity (grouped as `Extractive` industries) are consistently linked with higher support for the Liberal party, with the magnitude of this effect slightly increasing over the years (see row 1 in Figure \@ref(fig:partresplot1)).
-This is unsurprising, as the Liberal party has close ties with these traditional energy industries, and typically present policies to reduce taxation on energy production.
-Furthermore, electorates with more workers in construction or manufacturing industries (`Transformative`) are also more likely to support the Liberal party (see row 3 in Figure \@ref(fig:partresplot1)), from 2004 onwards.
-
-Similarly, the proportion of workers in managerial, administrative, clerical and sales roles (`ManagerAdmin`), is also a significant predictor of two-party preference vote across all six elections, with a higher proportion of people working these jobs increasing Liberal support.
+Electorates with higher proportions of workers in mining, gas, water, agriculture, waste and electricity (grouped as `Extractive` industries) were consistently linked with higher support for the Liberal party, with the magnitude of this effect slightly increasing over the years (see row 1 in Figure \@ref(fig:partresplot1)).
+This is unsurprising, as the Liberal party has maintained close ties with these traditional energy industries, and typically present policies to reduce taxation on energy production.
+Furthermore, electorates with more workers in construction or manufacturing industries (`Transformative`) were also more likely to support the Liberal party (see row 3 in Figure \@ref(fig:partresplot1)), from 2004 onwards.
+
+Similarly, the proportion of workers in managerial, administrative, clerical and sales roles (`ManagerAdmin`), was also a significant predictor of two-party preference vote across all six elections, where higher proportions of people working these jobs increased Liberal support.
 
 Of these job related variables, the most surprising effect is that associated with the proportion of workers in education, healthcare, social work, community and arts (`SocialServ`).
-Typically the Labor party has more generous funding schemes affecting these areas of work, so one might expect `SocialServ` to have a negative effect on two-party preference.
-However, in every election this effect is found to be positive and significant.
+Typically the Labor party has more generous funding schemes affecting these areas of work, so one might expect `SocialServ` to have had a negative effect on two-party preference.
+However, in every election this effect was found to be positive and significant.
 
@@
 
-Typically the Labor party campaigns on more progressive policies, which often include tax reform that adversely affects higher income earners, and more generous social assistance programs.
-Perhaps it is due to these policies that higher income electorates appear more likely to support the Liberal party, as the `Incomes` factor has a positive effect on Liberal preference (see row 6 in Figure \@ref(fig:partresplot1)).
-This effect is significant in every election aside from 2004 and 2010.
-`Unemployment` however, is not as influential.
-In 2001 and 2004, electorates with higher unemployment align with Labor, but over time this shifts towards support for the Liberal party, culminating in a positive (insignificant) effect in 2016.
+Typically the Labor party campaigns on more progressive policies, often including tax reform that adversely affects higher income earners, and more generous social assistance programs.
+Perhaps it is due to these policies that higher income electorates appeared more likely to support the Liberal party, as the `Incomes` factor had a positive effect on Liberal preference (see row 6 in Figure \@ref(fig:partresplot1)).
+This effect was significant in every election aside from 2004 and 2010.
+`Unemployment` however, was not as influential.
+In 2001 and 2004, electorates with higher unemployment aligned with Labor, but over time this shifted towards support for the Liberal party, culminating in a positive (insignificant) effect in 2016.
 
@@
 The older Australian population is often believed to be more conservative, and the left leaning political parties (including Labor) typically have a stronger appeal to younger people.
-This effect is indeed observed across the six elections, with populations between 20 and 34 years of age (`Pop_20_34`) being very strongly aligned with Labor preference (bottom row in Figure \@ref(fig:partresplot2)).
-Larger populations of 35 to 54 year olds (`Pop_35_54`) are also associated with Labor, but the magnitude of this effect is far smaller.
-Populations under 20 years of age is only significant in 2007, where `Pop_00_19` increased Liberal support.
+This effect is indeed observed across the six elections, where populations between 20 and 34 years of age (`Pop_20_34`) aligned strongly with Labor preference (bottom row in Figure \@ref(fig:partresplot2)).
+Larger populations of 35 to 54 year olds (`Pop_35_54`) were also associated with Labor, but the magnitude of this effect was far smaller.
+Populations under 20 years of age was only significant in 2007, where `Pop_00_19` increased Liberal support.
 
@@
 
-Since 2007, electorates with higher education levels are associated with supporting the Labor party, with this effect being significant in 2007, 2013 and 2016 and only marginally insignificant in 2010.
-In the elections before 2007, education has a negligible effect (see row 3 in Figure \@ref(fig:partresplot2)).
-Additionally, student populations (`CurrentlyStudying`) do not affect electoral party preference in any election (not shown).
+From 2007, electorates with higher education levels have associated with supporting the Labor party, with this effect being significant in 2007, 2013 and 2016 and only marginally insignificant in 2010.
+In the elections before 2007, education had a negligible effect (see row 3 in Figure \@ref(fig:partresplot2)).
+Additionally, student populations (`CurrentlyStudying`) did not affect electoral party preference in any election (not shown).
 
@@
 
-Larger migrant populations from Asia, the Middle East, South-Eastern Europe, the United Kingdom and elsewhere, are either associated with Labor support, or have no effect.
-Of these areas, only South-Eastern European populations significantly affect party preference, with larger populations associating with Labor in 2013 and 2016 (row 2, Figure \@ref(fig:partresplot2)).
-Speaking other languages (aside from English) however, appears to have a far stronger effect, as observed through the `OtherLanguage` variable.
-Electorates with more diverse speech are associated with higher support for the Liberal party from 2004 onwards, with this effect being significant in 2013 and 2016 (see row 7, Figure \@ref(fig:partresplot1)).
-Furthermore, none of the variables relating to religious beliefs aside from Christianity have a material effect in any election (this includes the Buddhist, Muslim, Jewish, non-religious and Indigenous Australian populations).
-The association between Christian populations (`Christianity`) and the Liberal party steadily increases over the years, becoming positive and significant in 2016.
+Larger migrant populations from Asia, the Middle East, South-Eastern Europe, the United Kingdom and elsewhere, were either associated with Labor support, or have no effect.
+Of these areas, only South-Eastern European populations significantly affected party preference, with larger populations associating with Labor in 2013 and 2016 (row 2, Figure \@ref(fig:partresplot2)).
+Speaking other languages (aside from English) however, appears to have had a far stronger effect, as observed through the `OtherLanguage` variable.
+Electorates with more diverse speech were associated with higher support for the Liberal party from 2004 onwards, with this effect being significant in 2013 and 2016 (see row 7, Figure \@ref(fig:partresplot1)).
+Furthermore, none of the variables relating to religious beliefs aside from Christianity had a material effect in any election (this includes the Buddhist, Muslim, Jewish, non-religious and Indigenous Australian populations).
+The association between Christian populations (`Christianity`) and the Liberal party steadily increased over the years, becoming positive and significant in 2016.
 
@@
 
-In 2001, 2004 and 2016, higher proportions of people that have recently (in the past five years) moved house (`DiffAddress`) increased electoral support for the Liberal party (see row 5 in Figure \@ref(fig:partresplot1)).
-This is somewhat surprising as one might expect house ownership and rental prices to be linked to two-party preference, rather than household mobility (`PropertyOwned` and `RentLoan` are not significant in any election).
-
-Higher proportions of single parent households are associated with Labor support in all elections (albeit insignificant in 2004, see row 6 in Figure \@ref(fig:partresplot2)), whereas the electoral family and household sizes (via the `FamHouseSize` variable) do not appear to be associated with either party.
+In 2001, 2004 and 2016, higher proportions of people that recently (in the past five years) moved house (`DiffAddress`) increased electoral support for the Liberal party (see row 5 in Figure \@ref(fig:partresplot1)).
+This is somewhat surprising as one might expect house ownership and rental prices to be linked to two-party preference, rather than household mobility (`PropertyOwned` and `RentLoan` were not significant in any election).
+
+Higher proportions of single parent households were associated with Labor support in all elections (albeit insignificant in 2004, see row 6 in Figure \@ref(fig:partresplot2)), whereas the electoral family and household sizes (via the `FamHouseSize` variable) did not appear to be associated with either party.
 
@@
 
-From 2007 onwards, both marriages (`Married`) and de facto relationships (`DeFacto`) are found to be strong predictors of the two-party preferred vote in favour of the Labor party.
-In 2001 and 2004 neither of these variables are significant (see rows 4 and 5 in Figure \@ref(fig:partresplot2)).
+From 2007 onwards, both marriages (`Married`) and de facto relationships (`DeFacto`) were found to be strong predictors of the two-party preferred vote in favour of the Labor party.
+In 2001 and 2004 neither of these variables were significant (see rows 4 and 5 in Figure \@ref(fig:partresplot2)).
 
@@
 Boxplots of residuals grouped by state (Figure \@ref(fig:resstate)) show that the data reflects this to only a limited extent.
-Tasmania and the Australian Capital Territory appear to have a bias towards Labor, whereas the South Australia and the Northern Territory tend towards voting Liberal.
-However, there are relatively few electorates in each of these states (five, two, eleven and two respectively), so this apparent result may be due to incumbent effects rather than an actual state-specific bias.
+Tasmania and the Australian Capital Territory appeared to have a bias towards Labor, whereas the South Australia and the Northern Territory tend towards voting Liberal.
+However, there were relatively few electorates in each of these states (five, two, eleven and two respectively), so this apparent result may be due to incumbent effects rather than an actual state-specific bias.
 
@@
 
-The incumbent party appears to have a distinct advantage at the next election.
+The incumbent party appeared to have a distinct advantage at the next election.
 The boxplots in Figure \@ref(fig:resincumbent) show that if either of the Labor or Liberal parties won the seat at the previous election, the electorate is likely to vote in their favour, over and above any socio-demographic effects---this effect has not been captured by the model.
@@
 
-Three robustness checks are conducted to confirm model stability.
-First, a model for each election is re-estimated using only the variables that are found to be significant in at least one of the six elections.
-The estimated coefficients of the variables in the re-estimated models all fall within their respective 95% confidence intervals from the full models.
-The second check involves the ten largest pairwise correlations.
-For each pair, a model for each election is re-estimated omitting one of the two variables.
-It is found that for each of these pairs, the estimated effect of the remaining variable in the reduced model lies within the 95% confidence interval from the full model.
-The final check is a visual exploration of different variable projections using a tour [@tourr] for each election.
-No definitive signs of multicollinearity are observed, and as expected (given the nature of spatial data), there is some clumping of electorates for certain projections.
+Three robustness checks were conducted to confirm model stability.
+First, a model for each election was re-estimated using only the variables that were found to be significant in at least one of the six elections.
+The estimated coefficients of the variables in the re-estimated models all fell within their respective 95% confidence intervals from the full models.
+The second check involved the ten largest pairwise correlations.
+For each pair, a model for each election was re-estimated omitting one of the two variables.
+It was found that for each of these pairs, the estimated effect of the remaining variable in the reduced model was within the 95% confidence interval from the full model.
+The final check was a visual exploration of different variable projections using a tour [@tourr] for each election.
+No definitive signs of multicollinearity were observed, and as expected (given the nature of spatial data), there was some clumping of electorates for certain projections.
 
@@
 
-Based on the distribution of the Cook's distance values and the distribution of hat values, a Cook's distance greater than $0.1$ is considered to be influential, and a hat value greater than $0.5$ is considered to have high leverage.
-Electorates fitting these criteria are flagged and investigated to examine the characteristics driving these values.
-
-The electorate of Sydney (NSW) has a large Cook's distance and high leverage from 2001 to 2007, due to its diverse population (languages, birthplace and religion), high density of young adults (20 to 34 years old), high number of defacto relationships, high income, high household mobility and small amount of workers in extractive and transformative jobs.
-It has remained a strong supporter of the Labor party and the extent of this support is underpredicted by the model, making it an outlier.
-Nearby in metropolitan NSW, the electorate of Wentworth is found to be an outlier in the 2013 and 2016 elections.
+Based on the distribution of the Cook's distance values and the distribution of hat values, a Cook's distance greater than $0.1$ was considered to be influential, and a hat value greater than $0.5$ was considered to have high leverage.
+Electorates fitting these criteria were flagged and investigated to examine the characteristics driving these values.
+
+The electorate of Sydney (NSW) had a large Cook's distance and high leverage from 2001 to 2007, due to its diverse population (languages, birthplace and religion), high density of young adults (20 to 34 years old), high number of defacto relationships, high income, high household mobility and small amount of workers in extractive and transformative jobs.
+It remained a strong supporter of the Labor party and the extent of this support is underpredicted by the model, making it an outlier.
+Nearby in metropolitan NSW, the electorate of Wentworth was found to be an outlier in the 2013 and 2016 elections.
 Although historically Liberal, its two-party vote jumped by over 10 percentage points in 2010 without experiencing any notable changes in its socio-demographic makeup---implying that this may be the direct effect of its Liberal member, Malcolm Turnbull, becoming the leader of the Liberal party.
@@
 
-Lingiari, an electorate taking up almost all of the Northern Territory, has consistently high leverage (all years) and is an outlier in all but the 2013 election due to its large Indigenous population, low rates of property ownership and few workers in management or administrative jobs.
-Fowler (NSW) has a diverse population with a high proportion of migrants, many Buddhists and Muslims, as well as a high proportion of single parent households.
-These characteristics explain its high leverage in 2001, 2004, 2010 and 2013, and its strong Labor support makes it influential in 2001, 2004 and 2010.
-Other electorates with large Cook's distance are Canberra (ACT) and Durack (WA) in 2013, and Solomon (NT) in 2016.
-
-All of the electorates examined are not unduly influential in the model and therefore no action is required.
+Lingiari, an electorate making up almost all of the Northern Territory, had consistently high leverage (all years) and was an outlier in all but the 2013 election due to its large Indigenous population, low rates of property ownership and few workers in management or administrative jobs.
+Fowler (NSW) had a diverse population with a high proportion of migrants, many Buddhists and Muslims, as well as a high proportion of single parent households.
+These characteristics explain its high leverage in 2001, 2004, 2010 and 2013, and its strong Labor support made it influential in 2001, 2004 and 2010.
+Other electorates with large Cook's distance were Canberra (ACT) and Durack (WA) in 2013, and Solomon (NT) in 2016.
+
+All of the electorates examined were not unduly influential in the model and therefore no action was required.
 
@@
 
-This paper explores the effects of electoral socio-demographic characteristics on the two-party preferred vote in the 2001--2016 elections, using information from the corresponding Australian federal elections and censuses.
-As a census does not always occur in the same year as an election, census data for the 2004--2013 elections are generated by employing a method of spatio-temporal imputation.
-This imputes electoral socio-demographics for the electoral boundaries in place at the time of the election---an approach that is distinctly different from previous work on modelling election outcomes, where census and election data are typically joined without addressing their temporal differences.
-Before estimating a model, these socio-demographic variables are standardised (to adjust for changing variable scales) and subsets of variables (representing similar information) are combined into factors, resulting in a reduced predictor set.
-A spatial error model is then estimated for each election, accounting for the inherent spatial structure of the data.
-
-Across the past six elections, most of the socio-demographics that drive the electoral two-party preferred vote are found to remain steady, whilst a few (typically weaker) effects vary over time.
-Industry and type of work are particularly influential.
-Energy-related and manufacturing/construction jobs, as well as administrative roles and jobs in education and social services are strongly linked with the Liberal party in all elections.
-Incomes have a similarly consistent effect, with higher income areas supporting Liberal.
-Higher levels of unemployment shift from weak association with Labor to a significant Liberal effect over the years, and higher education levels are associated with Labor from 2007 (although marginally insignificant in 2010).
-Electorates with large populations 20 to 34 years are strongly associated with Labor, whilst the 35 to 54 year old bracket also increases Labor support, but to a lesser extent.
-It is also found that birthplace diversity slightly favours Labor, relationships (both marriages and de facto relationships) align with Labor preference from 2010 onwards, and the influence of Christian populations has trended towards Liberal support whilst other religions have negligible effects.
-Family and household sizes have minimal influence, although electorates with more single parent households are linked with Labor support.
-Furthermore, the spatial effects are found to be positive in all elections and significant in 2001, 2004 and 2016, meaning that other characteristics that neighbours have in common  (outside of the variables in the model) appear to be influential in those years.
+This paper explored the effects of electoral socio-demographic characteristics on the two-party preferred vote in the 2001--2016 elections, using information from the corresponding Australian federal elections and censuses.
+As a census did not always occur in the same year as an election, census data for the 2004--2013 elections was generated by employing a method of spatio-temporal imputation.
+This imputed electoral socio-demographics for the electoral boundaries in place at the time of the election---an approach that is distinctly different from previous work on modelling election outcomes, where census and election data is typically joined without addressing their temporal differences.
+Before estimating a model, these socio-demographic variables were standardised (to adjust for changing variable scales) and subsets of variables (representing similar information) were combined into factors, resulting in a reduced predictor set.
+A spatial error model was then estimated for each election, accounting for the inherent spatial structure of the data.
+
+Across the past six elections, most of the socio-demographics driving the electoral two-party preferred vote were found to remain steady, whilst a few (typically weaker) effects varied over time.
+Industry and type of work were particularly influential.
+Energy-related and manufacturing/construction jobs, as well as administrative roles and jobs in education and social services were strongly linked with the Liberal party in all elections.
+Incomes had a similarly consistent effect, with higher income areas supporting Liberal.
+Higher levels of unemployment shifted from a weak association with Labor to a significant Liberal effect over the years, and higher education levels were associated with Labor from 2007 (although marginally insignificant in 2010).
+Electorates with large populations 20 to 34 years were strongly associated with Labor, whilst the 35 to 54 year old bracket also increased Labor support, but to a lesser extent.
+It was also found that birthplace diversity slightly favoured Labor, relationships (both marriages and de facto relationships) aligned with Labor preference from 2010 onwards, and the influence of Christian populations trended towards Liberal support whilst other religions had negligible effects.
+Family and household sizes had minimal influence, although electorates with more single parent households were linked with Labor support.
+Furthermore, the spatial effects were found to be positive in all elections and significant in 2001, 2004 and 2016, meaning that other characteristics that neighbours had in common  (outside of the variables in the model) appeared to be influential in those years.
 

### 2020-04-05 anzjs_elections.Rmd
Small text and graph fixes

@@
 Several predictors had a negative relationship: with larger values indicating the electorate more likely preferenced Labor.
-Most relationships were relatively stable over elections, except `Unemployment` and `Education`."
+Most relationships were relatively stable over elections, except `Unemployment` and `Education`.
 
@@
 Electorates with higher proportions of workers in mining, gas, water, agriculture, waste and electricity (grouped as `Extractive` industries) were consistently linked with higher support for the Liberal party, with the magnitude of this effect slightly increasing over the years (see row 1 in Figure \@ref(fig:partresplot1)).
-This is unsurprising, as the Liberal party has maintained close ties with these traditional energy industries, and typically present policies to reduce taxation on energy production.
+This is unsurprising, as the Liberal party had maintained close ties with these traditional energy industries, and typically present policies to reduce taxation on energy production.
 Furthermore, electorates with more workers in construction or manufacturing industries (`Transformative`) were also more likely to support the Liberal party (see row 3 in Figure \@ref(fig:partresplot1)), from 2004 onwards.
@@
 Typically the Labor party campaigns on more progressive policies, often including tax reform that adversely affects higher income earners, and more generous social assistance programs.
-Perhaps it is due to these policies that higher income electorates appeared more likely to support the Liberal party, as the `Incomes` factor had a positive effect on Liberal preference (see row 6 in Figure \@ref(fig:partresplot1)).
+Perhaps it is due to these policies that higher income electorates were more likely to support the Liberal party, as the `Incomes` factor had a positive effect on Liberal preference (see row 6 in Figure \@ref(fig:partresplot1)).
 This effect was significant in every election aside from 2004 and 2010.
@@
 
-From 2007, electorates with higher education levels have associated with supporting the Labor party, with this effect being significant in 2007, 2013 and 2016 and only marginally insignificant in 2010.
+From 2007, electorates with higher education levels were associated with support for the Labor party, with this effect being significant in 2007, 2013 and 2016 and only marginally insignificant in 2010.
 In the elections before 2007, education had a negligible effect (see row 3 in Figure \@ref(fig:partresplot2)).
@@
 
-Larger migrant populations from Asia, the Middle East, South-Eastern Europe, the United Kingdom and elsewhere, were either associated with Labor support, or have no effect.
+Larger migrant populations from Asia, the Middle East, South-Eastern Europe, the United Kingdom and elsewhere, were either associated with Labor support, or had no effect.
 Of these areas, only South-Eastern European populations significantly affected party preference, with larger populations associating with Labor in 2013 and 2016 (row 2, Figure \@ref(fig:partresplot2)).
@@
 Boxplots of residuals grouped by state (Figure \@ref(fig:resstate)) show that the data reflects this to only a limited extent.
-Tasmania and the Australian Capital Territory appeared to have a bias towards Labor, whereas the South Australia and the Northern Territory tend towards voting Liberal.
+Tasmania and the Australian Capital Territory appeared to have a bias towards Labor, whereas the South Australia and the Northern Territory tended towards voting Liberal.
 However, there were relatively few electorates in each of these states (five, two, eleven and two respectively), so this apparent result may be due to incumbent effects rather than an actual state-specific bias.
@@
   theme_bw() +
-  labs(x = "State", y = "Residuals")
+  labs(x = "State", y = "Residuals") +
+  theme(text = element_text(size = 16))
 ```
@@
   theme_bw() +
-  xlab("Year") + ylab("Residuals")
+  xlab("Year") + ylab("Residuals") +
+  theme(text = element_text(size = 16))
 ```
@@
 This paper explored the effects of electoral socio-demographic characteristics on the two-party preferred vote in the 2001--2016 elections, using information from the corresponding Australian federal elections and censuses.
-As a census did not always occur in the same year as an election, census data for the 2004--2013 elections was generated by employing a method of spatio-temporal imputation.
-This imputed electoral socio-demographics for the electoral boundaries in place at the time of the election---an approach that is distinctly different from previous work on modelling election outcomes, where census and election data is typically joined without addressing their temporal differences.
+As a census did not always occur in the same year as an election, census data for each of the 2004--2013 elections were generated by employing a method of spatio-temporal imputation.
+This method imputes electoral socio-demographics for the electoral boundaries in place at the time of the corresponding election---an approach that is distinctly different from previous work on modelling election outcomes, where census and election data is typically joined without addressing their temporal differences.
 Before estimating a model, these socio-demographic variables were standardised (to adjust for changing variable scales) and subsets of variables (representing similar information) were combined into factors, resulting in a reduced predictor set.
@@
 
-Additionally, a key contribution of this research is the wrangling of the raw data and imputed data sets for the 2004, 2007, 2010 and 2013 elections, which have been contributed to the \textsf{eechidna} \textsf{R} package---providing a rich, accessible data resource for any future Australian electoral analysis.
+Additionally, a key contribution of this research is the wrangling of raw data and imputation of data sets for the 2004, 2007, 2010 and 2013 elections, which have been contributed to the \textsf{eechidna} \textsf{R} package---providing a rich, accessible data resource for any future Australian electoral analysis.
 

### 2020-04-05 anzjs_elections.Rmd
Final tense changes

@@
 In these three elections, there is evidence to suggest that neighbours shared some influential characteristics outside the explanatory variables, which affected the two-party preferred vote.
-Conversely, in the other three elections, the spatial effect is weaker and insignificant (although still positive).
+Conversely, in the other three elections, the spatial effect was weaker and insignificant (although still positive).
 
 ```{r rhovis, fig.cap="Estimates of the spatial autoregressive parameter for each of the six elections, reported with their individual 95\\% confidence intervals.
-In 2001, 2004 and 2016 there is a significant spatial component.", out.height='25%', out.extra = ''}
+In 2001, 2004 and 2016 there was a significant spatial component.", out.height='25%', out.extra = ''}
 rho_df <- bind_rows(
@@
 
-To investigate the socio-demographics that had a strong effect on the two-party preferred vote, partial residual plots are used and shown in Figures \@ref(fig:partresplot1) and \@ref(fig:partresplot2).
+To investigate the socio-demographics that had a strong effect on the two-party preferred vote, partial residual plots were used and shown in Figures \@ref(fig:partresplot1) and \@ref(fig:partresplot2).
 Partial residuals, for a given variable, are the residuals from the fitted model with the estimated effect of that variable added to it.
@@
 Plots were computed using the method in @visreg.
-If a horizontal line can be drawn through the confidence band, then the effect is insignificant.
-The estimated intercept is also added to the partial residuals for interpretability.
-Plots for each election are faceted to compare the effects over time in Figures \@ref(fig:partresplot1) and \@ref(fig:partresplot2).
-Only socio-demographics that had a significant effect in at least two elections are displayed in Figures \@ref(fig:partresplot1) and \@ref(fig:partresplot2).
+If a horizontal line can be drawn through the confidence band, then the effect was insignificant.
+The estimated intercept was also added to the partial residuals for interpretability.
+Plots for each election are faceted in Figures \@ref(fig:partresplot1) and \@ref(fig:partresplot2) to compare the effects over time.
+Only socio-demographics that had a significant effect in at least two elections are displayed.
 
@@
 Electorates with higher proportions of workers in mining, gas, water, agriculture, waste and electricity (grouped as `Extractive` industries) were consistently linked with higher support for the Liberal party, with the magnitude of this effect slightly increasing over the years (see row 1 in Figure \@ref(fig:partresplot1)).
-This is unsurprising, as the Liberal party had maintained close ties with these traditional energy industries, and typically present policies to reduce taxation on energy production.
+This is unsurprising, as the Liberal party maintained close ties with these traditional energy industries, and typically presented policies to reduce taxation on energy production.
 Furthermore, electorates with more workers in construction or manufacturing industries (`Transformative`) were also more likely to support the Liberal party (see row 3 in Figure \@ref(fig:partresplot1)), from 2004 onwards.
@@
 Of these job related variables, the most surprising effect is that associated with the proportion of workers in education, healthcare, social work, community and arts (`SocialServ`).
-Typically the Labor party has more generous funding schemes affecting these areas of work, so one might expect `SocialServ` to have had a negative effect on two-party preference.
+Typically the Labor party has had more generous funding schemes affecting these areas of work, so one might expect `SocialServ` to have had a negative effect on two-party preference.
 However, in every election this effect was found to be positive and significant.
@@
 
-Typically the Labor party campaigns on more progressive policies, often including tax reform that adversely affects higher income earners, and more generous social assistance programs.
+Typically the Labor party has campaigned on more progressive policies, often including tax reform that adversely affects higher income earners, and more generous social assistance programs.
 Perhaps it is due to these policies that higher income electorates were more likely to support the Liberal party, as the `Incomes` factor had a positive effect on Liberal preference (see row 6 in Figure \@ref(fig:partresplot1)).
@@
 
-The older Australian population is often believed to be more conservative, and the left leaning political parties (including Labor) typically have a stronger appeal to younger people.
-This effect is indeed observed across the six elections, where populations between 20 and 34 years of age (`Pop_20_34`) aligned strongly with Labor preference (bottom row in Figure \@ref(fig:partresplot2)).
+The older Australian population has often been considered to be more conservative, and the left leaning political parties (including Labor) have typically had a stronger appeal to younger people.
+This effect was indeed observed across the six elections, where populations between 20 and 34 years of age (`Pop_20_34`) aligned strongly with Labor preference (bottom row in Figure \@ref(fig:partresplot2)).
 Larger populations of 35 to 54 year olds (`Pop_35_54`) were also associated with Labor, but the magnitude of this effect was far smaller.
@@
 Speaking other languages (aside from English) however, appears to have had a far stronger effect, as observed through the `OtherLanguage` variable.
-Electorates with more diverse speech were associated with higher support for the Liberal party from 2004 onwards, with this effect being significant in 2013 and 2016 (see row 7, Figure \@ref(fig:partresplot1)).
+Electorates with more diverse speech were linked with higher support for the Liberal party from 2004 onwards, with this effect being significant in 2013 and 2016 (see row 7, Figure \@ref(fig:partresplot1)).
 Furthermore, none of the variables relating to religious beliefs aside from Christianity had a material effect in any election (this includes the Buddhist, Muslim, Jewish, non-religious and Indigenous Australian populations).
-The association between Christian populations (`Christianity`) and the Liberal party steadily increased over the years, becoming positive and significant in 2016.
+The relationship between Christian populations (`Christianity`) and the Liberal party strengthened over the years, becoming positive and significant in 2016.
 
@@
 In 2001, 2004 and 2016, higher proportions of people that recently (in the past five years) moved house (`DiffAddress`) increased electoral support for the Liberal party (see row 5 in Figure \@ref(fig:partresplot1)).
-This is somewhat surprising as one might expect house ownership and rental prices to be linked to two-party preference, rather than household mobility (`PropertyOwned` and `RentLoan` were not significant in any election).
+This was somewhat surprising as one might expect house ownership and rental prices to be linked to two-party preference, rather than household mobility (`PropertyOwned` and `RentLoan` were not significant in any election).
 
@@
 It is often hypothesised that states have a systematic bias towards one of the two major parties.
-Boxplots of residuals grouped by state (Figure \@ref(fig:resstate)) show that the data reflects this to only a limited extent.
+Boxplots of residuals grouped by state (Figure \@ref(fig:resstate)) showed that the data reflects this to only a limited extent.
 Tasmania and the Australian Capital Territory appeared to have a bias towards Labor, whereas the South Australia and the Northern Territory tended towards voting Liberal.
@@
 States ordered by median residual.
-A state-specific bias present only in the smaller states appears to have not been captured by the model.", fig.asp=0.6, out.width="70%", fig.pos = 'h', out.extra = ''}
+A state-specific bias present only in the smaller states appeared to have not been captured by the model.", fig.asp=0.6, out.width="70%", fig.pos = 'h', out.extra = ''}
 all_resids %>%
@@
 The incumbent party appeared to have a distinct advantage at the next election.
-The boxplots in Figure \@ref(fig:resincumbent) show that if either of the Labor or Liberal parties won the seat at the previous election, the electorate is likely to vote in their favour, over and above any socio-demographic effects---this effect has not been captured by the model.
+The boxplots in Figure \@ref(fig:resincumbent) show that if either of the Labor or Liberal parties won the seat at the previous election, the electorate was likely to vote in their favour at the subsequent election, over and above any socio-demographic effects---this effect has not been captured by the model.
 
 ```{r resincumbent, fig.cap = "Boxplot of residuals for incumbent and non-incumbent parties each year.
-An incumbent advantage is evident and has not been captured by the model.", fig.asp=0.6, out.width="70%", fig.pos = 'h', out.extra = ''}
+An incumbent advantage was evident and had not been captured by the model.", fig.asp=0.6, out.width="70%", fig.pos = 'h', out.extra = ''}
 # Load first preferences
@@
 The electorate of Sydney (NSW) had a large Cook's distance and high leverage from 2001 to 2007, due to its diverse population (languages, birthplace and religion), high density of young adults (20 to 34 years old), high number of defacto relationships, high income, high household mobility and small amount of workers in extractive and transformative jobs.
-It remained a strong supporter of the Labor party and the extent of this support is underpredicted by the model, making it an outlier.
+It remained a strong supporter of the Labor party and the extent of this support was underpredicted by the model, making it an outlier.
 Nearby in metropolitan NSW, the electorate of Wentworth was found to be an outlier in the 2013 and 2016 elections.
 Although historically Liberal, its two-party vote jumped by over 10 percentage points in 2010 without experiencing any notable changes in its socio-demographic makeup---implying that this may be the direct effect of its Liberal member, Malcolm Turnbull, becoming the leader of the Liberal party.
-In the elections since, the model underpredicts Wentworth's Liberal support.
+In the elections that followed, the model underpredicted Wentworth's Liberal support.
 
@@
 As a census did not always occur in the same year as an election, census data for each of the 2004--2013 elections were generated by employing a method of spatio-temporal imputation.
-This method imputes electoral socio-demographics for the electoral boundaries in place at the time of the corresponding election---an approach that is distinctly different from previous work on modelling election outcomes, where census and election data is typically joined without addressing their temporal differences.
+This method imputes electoral socio-demographics for the electoral boundaries in place at the time of the corresponding election---an approach that is distinctly different from previous work on modelling election outcomes, where census and election data has typically been joined without addressing their temporal differences.
 Before estimating a model, these socio-demographic variables were standardised (to adjust for changing variable scales) and subsets of variables (representing similar information) were combined into factors, resulting in a reduced predictor set.
@@
 This study contributed a large revision to the \textsf{eechidna} package, which included the addition of election and census data for 2001--2010, voting outcomes for polling booths and imputed census data for election years.
-For more details on using \textsf{eechidna}, please see the articles (vignettes) on the github page \url{ropenscilabs.github.io/eechidna}.
+For more details on using \textsf{eechidna}, please see the articles (vignettes) on the github page: \url{ropenscilabs.github.io/eechidna}.
 

### 2020-04-05 anzjs_elections.tex
Recompiled
Compiled pdf

@@
-\documentclass[times, doublespace]{anzsauth}
+% Options for packages loaded elsewhere
+\PassOptionsToPackage{unicode}{hyperref}
+\PassOptionsToPackage{hyphens}{url}
+%
+\documentclass[
+  times, doublespace]{anzsauth}
 \usepackage[]{times}
@@
 \usepackage{ifxetex,ifluatex}
-\usepackage{fixltx2e} % provides \textsubscript
 \ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
@@
   \usepackage[utf8]{inputenc}
-\else % if luatex or xelatex
-  \ifxetex
-    \usepackage{mathspec}
-  \else
-    \usepackage{fontspec}
+  \usepackage{textcomp} % provide euro and other symbols
+\else % if luatex or xetex
+  \usepackage{unicode-math}
+  \defaultfontfeatures{Scale=MatchLowercase}
+  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
 \fi
-  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
-\fi
-% use upquote if available, for straight quotes in verbatim environments
+% Use upquote if available, for straight quotes in verbatim environments
 \IfFileExists{upquote.sty}{\usepackage{upquote}}{}
-% use microtype if available
-\IfFileExists{microtype.sty}{%
-\usepackage{microtype}
+\IfFileExists{microtype.sty}{% use microtype if available
+  \usepackage[]{microtype}
   \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
 }{}
-\usepackage[left=5cm, right=0cm, top=4cm, bottom=0cm]{geometry}
-\usepackage{hyperref}
-\hypersetup{unicode=true,
+\makeatletter
+\@ifundefined{KOMAClassName}{% if non-KOMA class
+  \IfFileExists{parskip.sty}{%
+    \usepackage{parskip}
+  }{% else
+    \setlength{\parindent}{0pt}
+    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
+}{% if KOMA class
+  \KOMAoptions{parskip=half}}
+\makeatother
+\usepackage{xcolor}
+\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
+\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
+\hypersetup{
   pdftitle={Spatial modelling of the two-party preferred vote in Australian federal elections: 2001--2016},
   pdfauthor={Jeremy Forbes\^{}*, Dianne Cook, Rob J Hyndman},
-            pdfborder={0 0 0},
-            breaklinks=true}
-\urlstyle{same}  % don't use monospace font for urls
-\usepackage{natbib}
-\bibliographystyle{anzsj}
+  hidelinks,
+  pdfcreator={LaTeX via pandoc}}
+\urlstyle{same} % disable monospaced font for URLs
 \usepackage{longtable,booktabs}
+% Correct order of tables after \paragraph or \subparagraph
+\usepackage{etoolbox}
+\makeatletter
+\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
+\makeatother
+% Allow footnotes in longtable head/foot
+\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
+\makesavenoteenv{longtable}
 \usepackage{graphicx,grffile}
@@
 \setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
-\IfFileExists{parskip.sty}{%
-\usepackage{parskip}
-}{% else
-\setlength{\parindent}{0pt}
-\setlength{\parskip}{6pt plus 2pt minus 1pt}
-}
+% Set default figure placement to htbp
+\makeatletter
+\def\fps@figure{htbp}
+\makeatother
 \setlength{\emergencystretch}{3em} % prevent overfull lines
@@
 \setcounter{secnumdepth}{5}
-% Redefines (sub)paragraphs to behave more like sections
-\ifx\paragraph\undefined\else
-\let\oldparagraph\paragraph
-\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
-\fi
-\ifx\subparagraph\undefined\else
-\let\oldsubparagraph\subparagraph
-\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
-\fi
-
-%%% Use protect on footnotes to avoid problems with footnotes in titles
-\let\rmarkdownfootnote\footnote%
-\def\footnote{\protect\rmarkdownfootnote}
-
-%%% Change title format to be more compact
-\usepackage{titling}
-
-% Create subtitle command for use in maketitle
-\providecommand{\subtitle}[1]{
-  \posttitle{
-    \begin{center}\large#1\end{center}
-    }
-}
-
-\setlength{\droptitle}{-2em}
-
-  \title{Spatial modelling of the two-party preferred vote in Australian federal elections: 2001--2016}
-    \pretitle{\vspace{\droptitle}\centering\huge}
-  \posttitle{\par}
-    \author{Jeremy Forbes\(^*\), Dianne Cook, Rob J Hyndman}
-    \preauthor{\centering\large\emph}
-  \postauthor{\par}
-    \date{}
-    \predate{}\postdate{}
-
 %% Any special functions or other packages can be loaded here.
@@
 \usepackage[UKenglish]{isodate}
+\usepackage{hyperref}
 
@@
 
+\usepackage{titling}
 \pretitle{\begin{center}\Large\bfseries}
@@
 Email: \texttt{jeremyforbes1995@gmail.com}}\end{center}}
+\usepackage[]{natbib}
+\bibliographystyle{anzsj}
+
+\title{Spatial modelling of the two-party preferred vote in Australian federal elections: 2001--2016}
+\author{Jeremy Forbes\(^*\), Dianne Cook, Rob J Hyndman}
+\date{}
 
@@
 Education and unemployment are amongst those that have varying effects.
-All data featured in this study has been contributed to the \texttt{eechidna} \texttt{R} package (available on CRAN).
+All data featured in this study has been contributed to the \textsf{eechidna} \textsf{R} package (available on CRAN).
 
-\textbf{Keywords:} federal election, Census, Australia, spatial modelling, imputation, data science, socio-demographics, electorates, R, eechidna
+\textbf{Keywords:} federal election, census, Australia, spatial modelling, imputation, data science, socio-demographics, electorates, R, eechidna
 
@@
 Joining these two data sources is problematic as there is an inherent asynchronicity in the two types of events.
-A Census is conducted by the Australian Bureau of Statistics (ABS) every five years, whereas federal elections, conducted by the AEC, usually occur every three years or so.
-The first problem addressed is that of constructing appropriate Census data for the 2004, 2007, 2010 and 2013 elections --- election years in which a Census did not occur.
-The predominant approach in previous studies was to join voting outcomes to the nearest Census, without accounting for any temporal differences \citep[see][]{DavisStimson98, Stimson06, Liao09, Stimson09}.
+A census is conducted by the Australian Bureau of Statistics (ABS) every five years, whereas federal elections, conducted by the AEC, usually occur every three years or so.
+The first problem addressed is that of constructing appropriate census data for the 2004, 2007, 2010 and 2013 elections---election years in which a census did not occur.
+The predominant approach in previous studies was to join voting outcomes to the nearest census, without accounting for any temporal differences \citep[see][]{DavisStimson98, Stimson06, Liao09, Stimson09}.
 Furthermore, electoral boundaries change regularly, so spatial discrepancies also arise when matching with electoral data.
-To obtain appropriate ``Census-like'' data for these four elections, electoral socio-demographics are constructed using a spatio-temporal imputation that combines areal interpolation \citep{Goodchild1993} and linear time-interpolation.
+To obtain appropriate `census-like' data for these four elections, electoral socio-demographics are constructed using a spatio-temporal imputation that combines areal interpolation \citep{Goodchild1993} and linear time-interpolation.
 Collecting and wrangling the raw data, along with the imputation process, are detailed in Section \ref{data}.
-All data and associated documentation relating to this procedure are available in the \texttt{eechidna} \texttt{R} package \citep{eechidna}, providing a resource for any future analysis.
+All data and associated documentation relating to this procedure are available in the \textsf{eechidna} \textsf{R} package \citep{eechidna}, providing a resource for any future analysis.
 
 Previous work on modelling Australian federal elections has found that aggregate socio-demographics are relatively good predictors of voting outcomes.
-\citet{Forrest01} used multiple regression to model the Liberal and Labor primary vote for polling booths in the Farrer electorate in 1998 as a function of Census variables from 1996.
+\citet{Forrest01} used multiple regression to model the Liberal and Labor primary vote for polling booths in the Farrer electorate in 1998 as a function of census variables from 1996.
 \citet{Stimson06}, \citet{Stimson09} and \citet{Stimson12} used principal component analysis of polling booths in the 2001, 2004 and 2007 elections respectively, also finding that socio-demographic characteristics of polling booths are linked to their two-party preferred vote.
-In contrast, \citet{Stimson09} models the polling booth swing vote (change in the two-party preferred vote) in the 2007 election, finding that little of the swing vote can be explained by Census data.
+In contrast, \citet{Stimson09} models the polling booth swing vote (change in the two-party preferred vote) in the 2007 election, finding that little of the swing vote can be explained by census data.
 
@@
 This dataset consists of the two-party preferred vote (as the response variable) and a set of common socio-demographic variables (as the explanatory variables).
-To prepare these datasets, socio-demographic variables are first standardized, and then a principal component analysis is used to group many of the variables into ``factors''.
+To prepare these datasets, socio-demographic variables are first standardised, and then a principal component analysis is used to group many of the variables into `factors'.
 To account for the inherent spatial structure of the data, a spatial error model is then estimated for each election.
@@
 
-Socio-demographic variables are derived from the Australian Census of Population and Housing (Census), which is a survey of every household in Australia, recording information such as age, gender, ethnicity, education level and income.
-There have been four Censuses so far in the 21st century, conducted in 2001, 2006, 2011 and 2016.
-The Australian Bureau of Statistics (ABS) conducts the Census and publishes aggregated information.
-The ABS uses electoral boundaries as defined by the AEC at the time of each Census, which may not match those in place at the subsequent and previous elections.
-From the available Census information aggregated at the electorate level, 50 socio-demographic variables are defined for each of the electorates to be used in the analysis.
+Socio-demographic variables are derived from the Australian Census of Population and Housing (census), which is a survey of every household in Australia, recording information such as age, gender, ethnicity, education level and income.
+There have been four censuses so far in the 21st century, conducted in 2001, 2006, 2011 and 2016.
+The Australian Bureau of Statistics (ABS) conducts the census and publishes aggregated information.
+The ABS uses electoral boundaries as defined by the AEC at the time of each census, which may not match those in place at the subsequent and previous elections.
+From the available census information aggregated at the electorate level, 50 socio-demographic variables are defined for each of the electorates to be used in the analysis.
 These variables include information relating to electoral age distributions, income, education qualifications, employment industries and job types, religion, birthplace, household characteristics and relationships.
 
-Raw data is sourced online from the AEC and ABS websites in \texttt{.csv} and \texttt{.xlsx} files.
+Raw data is sourced online from the AEC and ABS websites in \textsf{.csv} and \textsf{.xlsx} files.
 The formats of these files differ over the years, making extracting the appropriate information a big task.
-The functions available in the \texttt{dplyr} \citep{dplyr} and \texttt{readxl} \citep{readxl} \texttt{R} packages are particularly useful, as they provide fast consistent tools for data manipulation and functions to import \texttt{.xlsx} files.
-The 2001 and 2006 Census data are published in a format where the information for each electorate is held in a separate document making it difficult to use the \texttt{dplyr} tools.
+The functions available in the \textsf{dplyr} \citep{dplyr} and \textsf{readxl} \citep{readxl} \textsf{R} packages are particularly useful, as they provide fast consistent tools for data manipulation and functions to import \textsf{.xlsx} files.
+The 2001 and 2006 census data are published in a format where the information for each electorate is held in a separate document making it difficult to use the \textsf{dplyr} tools.
 Instead, cells have to be selected from each individual file to construct the desired variables.
-All scripts required for the data wrangling process can be found in the github repository for the \texttt{eechidna} \texttt{R} package \citep{eechidna}, along with the raw data.
-The \texttt{eechidna} package makes this study entirely reproducible and provides a resource to help wrangle data for future Censuses and elections, when they become available.
+All scripts required for the data wrangling process can be found in the github repository for the \textsf{eechidna} \textsf{R} package \citep{eechidna}, along with the raw data.
+The \textsf{eechidna} package makes this study entirely reproducible and provides a resource to help wrangle data for future censuses and elections, when they become available.
 
 \hypertarget{joining-census-and-election-data}{%
-\subsection{Joining Census and election data}\label{joining-census-and-election-data}}
+\subsection{Joining census and election data}\label{joining-census-and-election-data}}
 
 \hypertarget{differences-between-census-and-election-data}{%
-\subsubsection*{Differences between Census and election data}\label{differences-between-census-and-election-data}}
-\addcontentsline{toc}{subsubsection}{Differences between Census and election data}
+\subsubsection*{Differences between census and election data}\label{differences-between-census-and-election-data}}
+\addcontentsline{toc}{subsubsection}{Differences between census and election data}
 
-Between 2001 and 2016 there were six elections and four Censuses (see Figure \ref{fig:timeline}).
-Electoral boundaries are redistributed regularly by the AEC, meaning that only in the years where both a Census and an election occur are all boundaries likely to match --- the case for the 2001 and 2016 elections.
-Therefore, for the four elections between 2004 and 2013, both temporal and spatial differences in electorates need to be accounted for when joining the electoral two-party preferred vote with Census data.
+Between 2001 and 2016 there were six elections and four censuses (see Figure \ref{fig:timeline}).
+Electoral boundaries are redistributed regularly by the AEC, meaning that only in the years where both a census and an election occur are all boundaries likely to match---the case for the 2001 and 2016 elections.
+Therefore, for the four elections between 2004 and 2013, both temporal and spatial differences in electorates need to be accounted for when joining the electoral two-party preferred vote with census data.
 For these elections a spatio-temporal imputation method is employed to obtain electoral socio-demographics.
-This method uses Census information from both before and after the election of interest.
+This method uses census information from both before and after the election of interest.
 
@@
 
-\caption{Timeline of Australian elections and Censuses. They do not always occur in the same year.}\label{fig:timeline}
+\caption{Timeline of Australian elections and censuses. They do not always occur in the same year.}\label{fig:timeline}
 \end{figure}
@@
 
-For each election, neighbouring Census information has to be combined in some way so that it represents the boundaries in place at the time of the election.
-This is done by taking the electoral boundaries and imputing the corresponding socio-demographic characteristics for each of the neighbouring Censuses, thereby addressing the spatial aspect.
-Next, to deal with the temporal component, characteristics at the time of the election are constructed using linear interpolation between the spatially imputed neighbouring Census variables.
+For each election, neighbouring census information has to be combined in some way so that it represents the boundaries in place at the time of the election.
+This is done by taking the electoral boundaries and imputing the corresponding socio-demographic characteristics for each of the neighbouring censuses, thereby addressing the spatial aspect.
+Next, to deal with the temporal component, characteristics at the time of the election are constructed using linear interpolation between the spatially imputed neighbouring census variables.
 
-The finest level of disaggregation available for Census data is the region classification called Statistical Area 1 (SA1).
+The finest level of disaggregation available for census data is the region classification called Statistical Area 1 (SA1).
 In 2016, Australia was divided into over 55,000 SA1s.
@@
 For simplicity, let each source zone be wholly summarised by its centroid.
-A set of target zones, \(t = 1,\dots,T\), are defined as regions for which information is to be imputed --- these are the electoral boundaries for a particular election.
+A set of target zones, \(t = 1,\dots,T\), are defined as regions for which information is to be imputed---these are the electoral boundaries for a particular election.
 
 Take the example of the Melbourne Ports electorate from the 2013 federal election, illustrated in Figure \ref{fig:melbports13}.
-The purple region in this figure represents the target zone and the source zones are the centroid locations from the 2016 Census SA1 areas.
+The purple region in this figure represents the target zone and the source zones are the centroid locations from the 2016 census SA1 areas.
 
@@
 
-\caption{The electoral boundaries for Melbourne Ports (shown in purple) and surrounding electorates, with centroids for Statistical Area 1 regions from the 2016 Census overlaid.
+\caption{The electoral boundaries for Melbourne Ports (shown in purple) and surrounding electorates, with centroids for Statistical Area 1 regions from the 2016 census overlaid.
 The centroids falling within the purple region are attributed to Melbourne Ports.}\label{fig:melbports13}
@@
 In order to calculate socio-demographic information for each of the target zones, a weighted average of source zones is taken using their populations as weights.
-Denote a given Census variable for the target zone by \(C_t\), and the same Census variable for the source zone as \(D_s\).
+Denote a given census variable for the target zone by \(C_t\), and the same census variable for the source zone as \(D_s\).
 Then, estimate \(C_t\) using
 \[
-  \hat{C}_t = \frac{\sum_{s=1}^{S}{I_{s,t}*D_s*P_s}}%
-  {\sum_{s=1}^{S}{I_{s,t}*P_s}},
+  \hat{C}_t = \frac{\sum_{s=1}^{S}{I_{s,t} \times D_s \times P_s}}%
+  {\sum_{s=1}^{S}{I_{s,t} \times P_s}},
   \quad\text{for each $t=1,\dots,T$}.
@@
 
-To account for temporal changes, linear interpolation is used between Census years to get the final estimate of a Census variable for the target zone in the election year.
-Let \(y_1\) be the year of the Census preceding an election, let \(y_2\) be the year of the election, and \(y_3\) be the year of the Census that follows.
-Add this year subscript to the Census variable estimate \(\hat{C}_t\), resulting in \(\hat{C}_{t,y}\).
-Linear interpolating between these Census years results an imputed value for the election year, given by
+To account for temporal changes, linear interpolation is used between census years to get the final estimate of a census variable for the target zone in the election year.
+Let \(y_1\) be the year of the census preceding an election, let \(y_2\) be the year of the election, and \(y_3\) be the year of the census that follows.
+Add this year subscript to the census variable estimate \(\hat{C}_t\), resulting in \(\hat{C}_{t,y}\).
+Linear interpolating between these census years results an imputed value for the election year, given by
 \[
@@
 
-Continuing with the example of Melbourne Ports in the 2013 election, the estimate for a given Census variable in 2016, \(\hat{C}_{\text{MelbPorts}, 2016}\) would be obtained by computing the weighted average of this variable amongst the SA1s within the purple region shown in Figure \ref{fig:melbports13}.
-This would be repeated with the 2011 Census SA1s to obtain \(\hat{C}_{\text{MelbPorts}, 2011}\), from which the final estimate is given by
+Continuing with the example of Melbourne Ports in the 2013 election, the estimate for a given census variable in 2016, \(\hat{C}_{\text{MelbPorts}, 2016}\) would be obtained by computing the weighted average of this variable amongst the SA1s within the purple region shown in Figure \ref{fig:melbports13}.
+This would be repeated with the 2011 census SA1s to obtain \(\hat{C}_{\text{MelbPorts}, 2011}\), from which the final estimate is given by
 \[
@@
 For example, inflation-adjusted median rental prices increased across almost all electorates, with median rent of 225 dollars per week placing an electorate in the 90th percentile in 2001, but only the 45th percentile in 2016.
-In order for socio-demographic effects to be comparable across years, all explanatory variables are standardized to have mean zero and variance one within each election year.
+In order for socio-demographic effects to be comparable across years, all explanatory variables are standardised to have mean zero and variance one within each election year.
 By standardizing, each variable is reported as a relative measure compared to all other electorates in the same year.
+(Note that the log values were standardised for the variables \texttt{Judaism}, \texttt{Indigenous}, \texttt{Islam} and \texttt{Buddhism}.)
 
@@
 Any model that uses all variables would face problems with multi-collinearity and over-fitting, which would likely lead to erroneous conclusions regarding variable significance.
-To address this, a subset of variables that represent similar information are combined into a single variable, which will be referred to as a ``factor''.
+To address this, a subset of variables that represent similar information are combined into a single variable, which will be referred to as a `factor'.
 
 A factor is created from a group of variables if there is an intuitive reason as to why these variables should represent similar information and if there is evidence to suggest that they covary.
-For example, a potential group would be variables relating to electoral incomes --- median family, household and personal incomes.
-To determine which variables covary, principal component analysis is used on a combined dataset of socio-demographic variables from all six elections\footnote{It is appropriate to compute principal components on a combined dataset of all six elections because when computed separately for each election, scree plots level off after four components and the loadings of the first four components are similar across the elections.}.
+For example, a potential group would be variables relating to electoral incomes---median family, household and personal incomes.
+To determine which variables covary, principal component analysis is used on a combined dataset of socio-demographic variables from all six elections.
 The only variables exempted from the principal component analysis are the four variables representing age brackets (the proportion of the population aged 0--19 years old, 20--34 years old, 45--54 years old and 55 years plus), which are included in the model as separate variables.
@@
 Variables with a positive loading are allocated a weight of \(+1\) and those with negative loadings are allocated a weight of \(-1\).
-After computing these weighted sums, the factor is standardized to have mean zero and variance one, within each election.
+After computing these weighted sums, the factor is standardised to have mean zero and variance one, within each election.
 
-The final predictor set contains \(p=32\) variables\footnote{The factor creation procedure reduces the variable set to \(p=33\), however one of the age brackets (\texttt{Pop\_55\_plus}) is not included as a variable to avoid multicollinearity, because the other three age brackets are included.} which are listed in Table \ref{tab:printmodels}.
+The final predictor set contains \(p=32\) variables which are listed in Table \ref{tab:printmodels}.
+(Note that the factor creation procedure reduces the variable set to \(p=33\), however the \texttt{Pop\_55\_plus} age bracket is not included as a variable to avoid multicollinearity, because the other three age brackets are included.)
 
@@
 Let \(\rho\) be the spatial autoregressive coefficient, \(\bm{v}\) be a spherical error term, \(\bm{W}\) be a matrix of spatial weights (containing information about the neighbouring regions), \(\bm{X}\) be a matrix of socio-demographic covariates, \(\bm{\beta}\) be a vector of regression coefficients and \(\bm{a}\) be a spatially structured random effect vector.
+
+Let
 \[
@@
 \]
-Estimation of the above spatial error model is undertaken using feasible generalized least squares.
+Estimation of the above spatial error model is undertaken using feasible generalised least squares.
 
@@
 
-The spatial autoregressive coefficient \(\rho\) is positive and significant in the 2001, 2004 and 2016 elections (Figure \ref{fig:rhovis}).
-In these three elections, there is evidence to suggest that neighbours share some influential characteristics outside the explanatory variables, which affect the two-party preferred vote.
-Conversely, in the other three elections, the spatial effect weakens to become insignificant (although still positive).
+The spatial autoregressive coefficient \(\rho\) was positive and significant in the 2001, 2004 and 2016 elections (Figure \ref{fig:rhovis}).
+In these three elections, there is evidence to suggest that neighbours shared some influential characteristics outside the explanatory variables, which affected the two-party preferred vote.
+Conversely, in the other three elections, the spatial effect was weaker and insignificant (although still positive).
 
@@
 
-{\centering \includegraphics[height=0.2\textheight,]{anzjs_elections_files/figure-latex/rhovis-1}
+{\centering \includegraphics[height=0.25\textheight,]{anzjs_elections_files/figure-latex/rhovis-1}
 
@@
 \caption{Estimates of the spatial autoregressive parameter for each of the six elections, reported with their individual 95\% confidence intervals.
-In 2001, 2004 and 2016 there is a significant spatial component.}\label{fig:rhovis}
+In 2001, 2004 and 2016 there was a significant spatial component.}\label{fig:rhovis}
 \end{figure}
@@
 
-{\centering \includegraphics[height=0.2\textheight,]{anzjs_elections_files/figure-latex/plotintercept-1}
+{\centering \includegraphics[height=0.25\textheight,]{anzjs_elections_files/figure-latex/plotintercept-1}
 
@@
 
-Since all socio-demographics have been standardized to have a mean of zero and a variance of one, the intercept in each model can be interpreted as the estimated two-party preferred vote for an electorate with mean characteristics\footnote{Mean of all variables aside from Judaism, Indigenous, Islam and Buddhism, where it assumes the mean of the log value.}.
-Figure \ref{fig:plotintercept} shows that the baseline of party preference has varied over the elections, with the biggest swing occurring in the 2007 election where the mean electorate shifted more than five percentage points in favour of the Labor party.
+Since all socio-demographics were standardised to have a mean of zero and a variance of one, the intercept in each model can be interpreted as the estimated two-party preferred vote for an electorate with mean characteristics (aside from \texttt{Judaism}, \texttt{Indigenous}, \texttt{Islam} and \texttt{Buddhism}, where it assumes the mean of the log value).
+Figure \ref{fig:plotintercept} shows that the baseline of party preference varied over the elections, with the biggest swing occurring in the 2007 election where the mean electorate shifted more than five percentage points in favour of the Labor party.
 
@@
 Linear model with 95\% confidence bands overlaid.
-Most predictors have a positive relationship: the larger the value the more likely the electorate preferences Liberal.
-The relationship is relatively robust over time, with the exception of \texttt{DiffAddress}, \texttt{Incomes}, \texttt{OtherLanguageHome} and \texttt{Pop\_00\_19}.}\label{fig:partresplot1}
+Most predictors had a positive relationship: the larger the value the more likely the electorate preferenced Liberal.
+The relationships were relatively robust over time, with the exception of \texttt{DiffAddress}, \texttt{Incomes}, \texttt{OtherLanguageHome} and \texttt{Pop\_00\_19}.}\label{fig:partresplot1}
 \end{figure}
@@
 Linear model with 95\% confidence bands overlaid.
-Several predictors have a negative relationship: with larger values indicating the electorate more likely preferences Labor.
-Most relationships are relatively stable over elections, except \texttt{Unemployment} and \texttt{Education}."}\label{fig:partresplot2}
+Several predictors had a negative relationship: with larger values indicating the electorate more likely preferenced Labor.
+Most relationships were relatively stable over elections, except \texttt{Unemployment} and \texttt{Education}.}\label{fig:partresplot2}
 \end{figure}
 
-To investigate the socio-demographics that have a strong effect on the two-party preferred vote, partial residual plots are used and shown in Figures \ref{fig:partresplot1} and \ref{fig:partresplot2}.
+To investigate the socio-demographics that had a strong effect on the two-party preferred vote, partial residual plots were used and shown in Figures \ref{fig:partresplot1} and \ref{fig:partresplot2}.
 Partial residuals, for a given variable, are the residuals from the fitted model with the estimated effect of that variable added to it.
@@
 In each plot, the slope of the prediction line matches the estimated coefficient and the shaded region represents a 95\% confidence band.
-Plots are computed using the method in \citet{visreg}.
-If a horizontal line can be drawn through the confidence band, then the effect is insignificant.
-The estimated intercept is also added to the partial residuals for interpretability.
-Plots for each election are faceted to compare the effects over time in Figures \ref{fig:partresplot1} and \ref{fig:partresplot2}.
-Only socio-demographics that have a significant effect in at least two elections are displayed in Figures \ref{fig:partresplot1} and \ref{fig:partresplot2}.
+Plots were computed using the method in \citet{visreg}.
+If a horizontal line can be drawn through the confidence band, then the effect was insignificant.
+The estimated intercept was also added to the partial residuals for interpretability.
+Plots for each election are faceted in Figures \ref{fig:partresplot1} and \ref{fig:partresplot2} to compare the effects over time.
+Only socio-demographics that had a significant effect in at least two elections are displayed.
 
@@
 
-Electorates with higher proportions of workers in mining, gas, water, agriculture, waste and electricity (grouped as \texttt{Extractive} industries) are consistently linked with higher support for the Liberal party, with the magnitude of this effect slightly increasing over the years (see row 1 in Figure \ref{fig:partresplot1}).
-This is unsurprising, as the Liberal party has close ties with these traditional energy industries, and typically present policies to reduce taxation on energy production.
-Furthermore, electorates with more workers in construction or manufacturing industries (\texttt{Transformative}) are also more likely to support the Liberal party (see row 3 in Figure \ref{fig:partresplot1}), from 2004 onwards.
+Electorates with higher proportions of workers in mining, gas, water, agriculture, waste and electricity (grouped as \texttt{Extractive} industries) were consistently linked with higher support for the Liberal party, with the magnitude of this effect slightly increasing over the years (see row 1 in Figure \ref{fig:partresplot1}).
+This is unsurprising, as the Liberal party maintained close ties with these traditional energy industries, and typically presented policies to reduce taxation on energy production.
+Furthermore, electorates with more workers in construction or manufacturing industries (\texttt{Transformative}) were also more likely to support the Liberal party (see row 3 in Figure \ref{fig:partresplot1}), from 2004 onwards.
 
-Similarly, the proportion of workers in managerial, administrative, clerical and sales roles (\texttt{ManagerAdmin}), is also a significant predictor of two-party preference vote across all six elections, with a higher proportion of people working these jobs increasing Liberal support.
+Similarly, the proportion of workers in managerial, administrative, clerical and sales roles (\texttt{ManagerAdmin}), was also a significant predictor of two-party preference vote across all six elections, where higher proportions of people working these jobs increased Liberal support.
 
 Of these job related variables, the most surprising effect is that associated with the proportion of workers in education, healthcare, social work, community and arts (\texttt{SocialServ}).
-Typically the Labor party has more generous funding schemes affecting these areas of work, so one might expect \texttt{SocialServ} to have a negative effect on two-party preference.
-However, in every election this effect is found to be positive and significant.
+Typically the Labor party has had more generous funding schemes affecting these areas of work, so one might expect \texttt{SocialServ} to have had a negative effect on two-party preference.
+However, in every election this effect was found to be positive and significant.
 
@@
 
-Typically the Labor party campaigns on more progressive policies, which often include tax reform that adversely affects higher income earners, and more generous social assistance programs.
-Perhaps it is due to these policies that higher income electorates appear more likely to support the Liberal party, as the \texttt{Incomes} factor has a positive effect on Liberal preference (see row 6 in Figure \ref{fig:partresplot1}).
-This effect is significant in every election aside from 2004 and 2010.
-\texttt{Unemployment} however, is not as influential.
-In 2001 and 2004, electorates with higher unemployment align with Labor, but over time this shifts towards support for the Liberal party, culminating in a positive (insignificant) effect in 2016.
+Typically the Labor party has campaigned on more progressive policies, often including tax reform that adversely affects higher income earners, and more generous social assistance programs.
+Perhaps it is due to these policies that higher income electorates were more likely to support the Liberal party, as the \texttt{Incomes} factor had a positive effect on Liberal preference (see row 6 in Figure \ref{fig:partresplot1}).
+This effect was significant in every election aside from 2004 and 2010.
+\texttt{Unemployment} however, was not as influential.
+In 2001 and 2004, electorates with higher unemployment aligned with Labor, but over time this shifted towards support for the Liberal party, culminating in a positive (insignificant) effect in 2016.
 
@@
 
-The older Australian population is often believed to be more conservative, and the left leaning political parties (including Labor) typically have a stronger appeal to younger people.
-This effect is indeed observed across the six elections, with populations between 20 and 34 years of age (\texttt{Pop\_20\_34}) being very strongly aligned with Labor preference (bottom row in Figure \ref{fig:partresplot2}).
-Larger populations of 35 to 54 year olds (\texttt{Pop\_35\_54}) are also associated with Labor, but the magnitude of this effect is far smaller.
-Populations under 20 years of age is only significant in 2007, where \texttt{Pop\_00\_19} increased Liberal support.
+The older Australian population has often been considered to be more conservative, and the left leaning political parties (including Labor) have typically had a stronger appeal to younger people.
+This effect was indeed observed across the six elections, where populations between 20 and 34 years of age (\texttt{Pop\_20\_34}) aligned strongly with Labor preference (bottom row in Figure \ref{fig:partresplot2}).
+Larger populations of 35 to 54 year olds (\texttt{Pop\_35\_54}) were also associated with Labor, but the magnitude of this effect was far smaller.
+Populations under 20 years of age was only significant in 2007, where \texttt{Pop\_00\_19} increased Liberal support.
 
@@
 
-Since 2007, electorates with higher education levels are associated with supporting the Labor party, with this effect being significant in 2007, 2013 and 2016 and only marginally insignificant in 2010.
-In the elections before 2007, education has a negligible effect (see row 3 in Figure \ref{fig:partresplot2}).
-Additionally, student populations (\texttt{CurrentlyStudying}) do not affect electoral party preference in any election (not shown).
+From 2007, electorates with higher education levels were associated with support for the Labor party, with this effect being significant in 2007, 2013 and 2016 and only marginally insignificant in 2010.
+In the elections before 2007, education had a negligible effect (see row 3 in Figure \ref{fig:partresplot2}).
+Additionally, student populations (\texttt{CurrentlyStudying}) did not affect electoral party preference in any election (not shown).
 
@@
 
-Larger migrant populations from Asia, the Middle East, South-Eastern Europe, the United Kingdom and elsewhere, are either associated with Labor support, or have no effect.
-Of these areas, only South-Eastern European populations significantly affect party preference, with larger populations associating with Labor in 2013 and 2016 (row 2, Figure \ref{fig:partresplot2}).
-Speaking other languages (aside from English) however, appears to have a far stronger effect, as observed through the \texttt{OtherLanguage} variable.
-Electorates with more diverse speech are associated with higher support for the Liberal party from 2004 onwards, with this effect being significant in 2013 and 2016 (see row 7, Figure \ref{fig:partresplot1}).
-Furthermore, none of the variables relating to religious beliefs aside from Christianity have a material effect in any election (this includes the Buddhist, Muslim, Jewish, non-religious and Indigenous Australian populations).
-The association between Christian populations (\texttt{Christianity}) and the Liberal party steadily increases over the years, becoming positive and significant in 2016.
+Larger migrant populations from Asia, the Middle East, South-Eastern Europe, the United Kingdom and elsewhere, were either associated with Labor support, or had no effect.
+Of these areas, only South-Eastern European populations significantly affected party preference, with larger populations associating with Labor in 2013 and 2016 (row 2, Figure \ref{fig:partresplot2}).
+Speaking other languages (aside from English) however, appears to have had a far stronger effect, as observed through the \texttt{OtherLanguage} variable.
+Electorates with more diverse speech were linked with higher support for the Liberal party from 2004 onwards, with this effect being significant in 2013 and 2016 (see row 7, Figure \ref{fig:partresplot1}).
+Furthermore, none of the variables relating to religious beliefs aside from Christianity had a material effect in any election (this includes the Buddhist, Muslim, Jewish, non-religious and Indigenous Australian populations).
+The relationship between Christian populations (\texttt{Christianity}) and the Liberal party strengthened over the years, becoming positive and significant in 2016.
 
@@
 
-In 2001, 2004 and 2016, higher proportions of people that have recently (in the past five years) moved house (\texttt{DiffAddress}) increased electoral support for the Liberal party (see row 5 in Figure \ref{fig:partresplot1}).
-This is somewhat surprising as one might expect house ownership and rental prices to be linked to two-party preference, rather than household mobility (\texttt{PropertyOwned} and \texttt{RentLoan} are not significant in any election).
+In 2001, 2004 and 2016, higher proportions of people that recently (in the past five years) moved house (\texttt{DiffAddress}) increased electoral support for the Liberal party (see row 5 in Figure \ref{fig:partresplot1}).
+This was somewhat surprising as one might expect house ownership and rental prices to be linked to two-party preference, rather than household mobility (\texttt{PropertyOwned} and \texttt{RentLoan} were not significant in any election).
 
-Higher proportions of single parent households are associated with Labor support in all elections (albeit insignificant in 2004, see row 6 in Figure \ref{fig:partresplot2}), whereas the electoral family and household sizes (via the \texttt{FamHouseSize} variable) do not appear to be associated with either party.
+Higher proportions of single parent households were associated with Labor support in all elections (albeit insignificant in 2004, see row 6 in Figure \ref{fig:partresplot2}), whereas the electoral family and household sizes (via the \texttt{FamHouseSize} variable) did not appear to be associated with either party.
 
@@
 
-From 2007 onwards, both marriages (\texttt{Married}) and de facto relationships (\texttt{DeFacto}) are found to be strong predictors of the two-party preferred vote in favour of the Labor party.
-In 2001 and 2004 neither of these variables are significant (see rows 4 and 5 in Figure \ref{fig:partresplot2}).
+From 2007 onwards, both marriages (\texttt{Married}) and de facto relationships (\texttt{DeFacto}) were found to be strong predictors of the two-party preferred vote in favour of the Labor party.
+In 2001 and 2004 neither of these variables were significant (see rows 4 and 5 in Figure \ref{fig:partresplot2}).
 
@@
 
-It is often hypothesized that states have a systematic bias towards one of the two major parties.
-Boxplots of residuals grouped by state (Figure \ref{fig:resstate}) show that the data reflects this to only a limited extent.
-Tasmania and the Australian Capital Territory appear to have a bias towards Labor, whereas the South Australia and the Northern Territory tend towards voting Liberal.
-However, there are relatively few electorates in each of these states (five, two, eleven and two respectively), so this apparent result may be due to incumbent effects rather than an actual state-specific bias.
+It is often hypothesised that states have a systematic bias towards one of the two major parties.
+Boxplots of residuals grouped by state (Figure \ref{fig:resstate}) showed that the data reflects this to only a limited extent.
+Tasmania and the Australian Capital Territory appeared to have a bias towards Labor, whereas the South Australia and the Northern Territory tended towards voting Liberal.
+However, there were relatively few electorates in each of these states (five, two, eleven and two respectively), so this apparent result may be due to incumbent effects rather than an actual state-specific bias.
 
@@
 States ordered by median residual.
-A state-specific bias present only in the smaller states appears to have not been captured by the model.}\label{fig:resstate}
+A state-specific bias present only in the smaller states appeared to have not been captured by the model.}\label{fig:resstate}
 \end{figure}
@@
 
-The incumbent party appears to have a distinct advantage at the next election.
-The boxplots in Figure \ref{fig:resincumbent} show that if either of the Labor or Liberal parties won the seat at the previous election, the electorate is likely to vote in their favour, over and above any socio-demographic effects --- this effect has not been captured by the model.
+The incumbent party appeared to have a distinct advantage at the next election.
+The boxplots in Figure \ref{fig:resincumbent} show that if either of the Labor or Liberal parties won the seat at the previous election, the electorate was likely to vote in their favour at the subsequent election, over and above any socio-demographic effects---this effect has not been captured by the model.
 
@@
 
-\caption{Boxplot of residuals for incumbent and non-incumbent parties each year. An incumbent advantage is evident and has not been captured by the model.}\label{fig:resincumbent}
+\caption{Boxplot of residuals for incumbent and non-incumbent parties each year. An incumbent advantage was evident and had not been captured by the model.}\label{fig:resincumbent}
 \end{figure}
@@
 
-Three robustness checks are conducted to confirm model stability.
-First, a model for each election is re-estimated using only the variables that are found to be significant in at least one of the six elections.
-The estimated coefficients of the variables in the re-estimated models all fall within their respective 95\% confidence intervals from the full models.
-The second check involves the ten largest pairwise correlations.
-For each pair, a model for each election is re-estimated omitting one of the two variables.
-It is found that for each of these pairs, the estimated effect of the remaining variable in the reduced model lies within the 95\% confidence interval from the full model.
-The final check is a visual exploration of different variable projections using a tour \citep{tourr} for each election.
-No definitive signs of multicollinearity are observed, and as expected (given the nature of spatial data), there is some clumping of electorates for certain projections.
+Three robustness checks were conducted to confirm model stability.
+First, a model for each election was re-estimated using only the variables that were found to be significant in at least one of the six elections.
+The estimated coefficients of the variables in the re-estimated models all fell within their respective 95\% confidence intervals from the full models.
+The second check involved the ten largest pairwise correlations.
+For each pair, a model for each election was re-estimated omitting one of the two variables.
+It was found that for each of these pairs, the estimated effect of the remaining variable in the reduced model was within the 95\% confidence interval from the full model.
+The final check was a visual exploration of different variable projections using a tour \citep{tourr} for each election.
+No definitive signs of multicollinearity were observed, and as expected (given the nature of spatial data), there was some clumping of electorates for certain projections.
 
@@
 
-Based on the distribution of the Cook's distance values and the distribution of hat values, a Cook's distance greater than \(0.1\) is considered to be influential, and a hat value greater than \(0.5\) is considered to have high leverage.
-Electorates fitting these criteria are flagged and investigated to examine the characteristics driving these values.
+Based on the distribution of the Cook's distance values and the distribution of hat values, a Cook's distance greater than \(0.1\) was considered to be influential, and a hat value greater than \(0.5\) was considered to have high leverage.
+Electorates fitting these criteria were flagged and investigated to examine the characteristics driving these values.
 
-The electorate of Sydney (NSW) has a large Cook's distance and high leverage from 2001 to 2007, due to its diverse population (languages, birthplace and religion), high density of young adults (20 to 34 years old), high number of defacto relationships, high income, high household mobility and small amount of workers in extractive and transformative jobs.
-It has remained a strong supporter of the Labor party and the extent of this support is underpredicted by the model, making it an outlier.
-Nearby in metropolitan NSW, the electorate of Wentworth is found to be an outlier in the 2013 and 2016 elections.
-Although historically Liberal, its two-party vote jumped by over 10 percentage points in 2010 without experiencing any notable changes in its socio-demographic makeup --- implying that this may be the direct effect of its Liberal member, Malcolm Turnbull, becoming the leader of the Liberal party.
-In the elections since, the model underpredicts Wentworth's Liberal support.
+The electorate of Sydney (NSW) had a large Cook's distance and high leverage from 2001 to 2007, due to its diverse population (languages, birthplace and religion), high density of young adults (20 to 34 years old), high number of defacto relationships, high income, high household mobility and small amount of workers in extractive and transformative jobs.
+It remained a strong supporter of the Labor party and the extent of this support was underpredicted by the model, making it an outlier.
+Nearby in metropolitan NSW, the electorate of Wentworth was found to be an outlier in the 2013 and 2016 elections.
+Although historically Liberal, its two-party vote jumped by over 10 percentage points in 2010 without experiencing any notable changes in its socio-demographic makeup---implying that this may be the direct effect of its Liberal member, Malcolm Turnbull, becoming the leader of the Liberal party.
+In the elections that followed, the model underpredicted Wentworth's Liberal support.
 
-Lingiari, an electorate taking up almost all of the Northern Territory, has consistently high leverage (all years) and is an outlier in all but the 2013 election due to its large Indigenous population, low rates of property ownership and few workers in management or administrative jobs.
-Fowler (NSW) has a diverse population with a high proportion of migrants, many Buddhists and Muslims, as well as a high proportion of single parent households.
-These characteristics explain its high leverage in 2001, 2004, 2010 and 2013, and its strong Labor support makes it influential in 2001, 2004 and 2010.
-Other electorates with large Cook's distance are Canberra (ACT) and Durack (WA) in 2013, and Solomon (NT) in 2016.
+Lingiari, an electorate making up almost all of the Northern Territory, had consistently high leverage (all years) and was an outlier in all but the 2013 election due to its large Indigenous population, low rates of property ownership and few workers in management or administrative jobs.
+Fowler (NSW) had a diverse population with a high proportion of migrants, many Buddhists and Muslims, as well as a high proportion of single parent households.
+These characteristics explain its high leverage in 2001, 2004, 2010 and 2013, and its strong Labor support made it influential in 2001, 2004 and 2010.
+Other electorates with large Cook's distance were Canberra (ACT) and Durack (WA) in 2013, and Solomon (NT) in 2016.
 
-All of the electorates examined are not unduly influential in the model and therefore no action is required.
+All of the electorates examined were not unduly influential in the model and therefore no action was required.
 
@@
 
-This paper explores the effects of electoral socio-demographic characteristics on the two-party preferred vote in the 2001--2016 elections, using information from the corresponding Australian federal elections and Censuses.
-As a Census does not always occur in the same year as an election, Census data for the 2004--2013 elections are generated by employing a method of spatio-temporal imputation.
-This imputes electoral socio-demographics for the electoral boundaries in place at the time of the election --- an approach that is distinctly different from previous work on modelling election outcomes, where Census and election data are typically joined without addressing their temporal differences.
-Before estimating a model, these socio-demographic variables are standardized (to adjust for changing variable scales) and subsets of variables (representing similar information) are combined into factors, resulting in a reduced predictor set.
-A spatial error model is then estimated for each election, accounting for the inherent spatial structure of the data.
+This paper explored the effects of electoral socio-demographic characteristics on the two-party preferred vote in the 2001--2016 elections, using information from the corresponding Australian federal elections and censuses.
+As a census did not always occur in the same year as an election, census data for each of the 2004--2013 elections were generated by employing a method of spatio-temporal imputation.
+This method imputes electoral socio-demographics for the electoral boundaries in place at the time of the corresponding election---an approach that is distinctly different from previous work on modelling election outcomes, where census and election data has typically been joined without addressing their temporal differences.
+Before estimating a model, these socio-demographic variables were standardised (to adjust for changing variable scales) and subsets of variables (representing similar information) were combined into factors, resulting in a reduced predictor set.
+A spatial error model was then estimated for each election, accounting for the inherent spatial structure of the data.
 
-Across the past six elections, most of the socio-demographics that drive the electoral two-party preferred vote are found to remain steady, whilst a few (typically weaker) effects vary over time.
-Industry and type of work are particularly influential.
-Energy-related and manufacturing/construction jobs, as well as administrative roles and jobs in education and social services are strongly linked with the Liberal party in all elections.
-Incomes have a similarly consistent effect, with higher income areas supporting Liberal.
-Higher levels of unemployment shift from weak association with Labor to a significant Liberal effect over the years, and higher education levels are associated with Labor from 2007 (although marginally insignificant in 2010).
-Electorates with large populations 20 to 34 years are strongly associated with Labor, whilst the 35 to 54 year old bracket also increases Labor support, but to a lesser extent.
-It is also found that birthplace diversity slightly favours Labor, relationships (both marriages and de facto relationships) align with Labor preference from 2010 onwards, and the influence of Christian populations has trended towards Liberal support whilst other religions have negligible effects.
-Family and household sizes have minimal influence, although electorates with more single parent households are linked with Labor support.
-Furthermore, the spatial effects are found to be positive in all elections and significant in 2001, 2004 and 2016, meaning that other characteristics that neighbours have in common (outside of the variables in the model) appear to be influential in those years.
+Across the past six elections, most of the socio-demographics driving the electoral two-party preferred vote were found to remain steady, whilst a few (typically weaker) effects varied over time.
+Industry and type of work were particularly influential.
+Energy-related and manufacturing/construction jobs, as well as administrative roles and jobs in education and social services were strongly linked with the Liberal party in all elections.
+Incomes had a similarly consistent effect, with higher income areas supporting Liberal.
+Higher levels of unemployment shifted from a weak association with Labor to a significant Liberal effect over the years, and higher education levels were associated with Labor from 2007 (although marginally insignificant in 2010).
+Electorates with large populations 20 to 34 years were strongly associated with Labor, whilst the 35 to 54 year old bracket also increased Labor support, but to a lesser extent.
+It was also found that birthplace diversity slightly favoured Labor, relationships (both marriages and de facto relationships) aligned with Labor preference from 2010 onwards, and the influence of Christian populations trended towards Liberal support whilst other religions had negligible effects.
+Family and household sizes had minimal influence, although electorates with more single parent households were linked with Labor support.
+Furthermore, the spatial effects were found to be positive in all elections and significant in 2001, 2004 and 2016, meaning that other characteristics that neighbours had in common (outside of the variables in the model) appeared to be influential in those years.
 
 The findings in this paper complement the existing literature by modelling temporal trends, which as far as the authors are aware, has not been done previously for Australian elections using a regression framework.
-It is also the first study to model any Australian election since 2010 using Census information.
+It is also the first study to model any Australian election since 2010 using census information.
 
-Additionally, a key contribution of this research is the wrangling of the raw data and imputed data sets for the 2004, 2007, 2010 and 2013 elections, which have been contributed to the \texttt{eechidna} \texttt{R} package --- providing a rich, accessible data resource for any future Australian electoral analysis.
+Additionally, a key contribution of this research is the wrangling of raw data and imputation of data sets for the 2004, 2007, 2010 and 2013 elections, which have been contributed to the \textsf{eechidna} \textsf{R} package---providing a rich, accessible data resource for any future Australian electoral analysis.
 
@@
 
-This paper was produced using \texttt{RMarkdown} \citep{rmarkdown} and \texttt{knitr} \citep{knitr}.
-All corresponding code for this paper can be found in the github repository \href{https://github.com/jforbes14/eechidna-paper}{github.com/jforbes14/eechidna-paper}, and the data used is available in the \texttt{eechidna} package \citep{eechidna}.
+This paper was produced using \textsf{RMarkdown} \citep{rmarkdown} and \textsf{knitr} \citep{knitr}.
+All corresponding code for this paper can be found in the github repository \url{github.com/jforbes14/eechidna-paper}, and the data used is available in the \textsf{eechidna} package \citep{eechidna}.
 All raw data was obtained from the Australian Electoral Commission, the Australian Bureau of Statistics and the Australian Government.
@@
 The authors would like to sincerely thank the editor and associate editor of the Australian \& New Zealand Journal of Statistics and the two anonymous reviewers for providing helpful comments and suggestions on earlier drafts of this manuscript.
-Additionally, the authors would like to thank Anthony Ebert, Heike Hofmann, Thomas Lumley, Ben Marwick, Carson Sievert, Mingzhu Sun, Dilini Talagala, Nicholas Tierney, Nathaniel Tomasetti, Earo Wang and Fang Zhou, all of whom have contributed to the \texttt{eechidna} package.
+Additionally, the authors would like to thank Anthony Ebert, Heike Hofmann, Thomas Lumley, Ben Marwick, Carson Sievert, Mingzhu Sun, Dilini Talagala, Nicholas Tierney, Nathaniel Tomasetti, Earo Wang and Fang Zhou, all of whom have contributed to the \textsf{eechidna} package.
 
@@
 
-All election and Census datasets, along with electoral maps and more, are available in the \texttt{eechidna} (Exploring Election and Census Highly Informative Data Nationally for Australia) \texttt{R} package, which can be downloaded from CRAN.
-The \texttt{eechidna} package makes it easy to look at the data from the Australian Federal elections and Censuses that occurred between 2001 and 2016.
-This study contributed a large revision to the \texttt{eechidna} package, which included the addition of election and Census data for 2001--2010, voting outcomes for polling booths and imputed Census data for election years.
-For more details on using \texttt{eechidna}, please see the articles (vignettes) on the github page \href{https://ropenscilabs.github.io/eechidna/}{ropenscilabs.github.io/eechidna/}.
+All election and census datasets, along with electoral maps and more, are available in the \textsf{eechidna} (Exploring Election and Census Highly Informative Data Nationally for Australia) \textsf{R} package, which can be downloaded from CRAN.
+The \textsf{eechidna} package makes it easy to look at the data from the Australian Federal elections and censuses that occurred between 2001 and 2016.
+This study contributed a large revision to the \textsf{eechidna} package, which included the addition of election and census data for 2001--2010, voting outcomes for polling booths and imputed census data for election years.
+For more details on using \textsf{eechidna}, please see the articles (vignettes) on the github page: \url{ropenscilabs.github.io/eechidna}.
 

### 2020-04-05 anzjs_elections.Rmd
Increase font size in plot

@@
   theme_map() +
-  coord_equal(xlim = c(144.88, 145.07), ylim = c(-37.92, -37.78))
+  coord_equal(xlim = c(144.88, 145.07), ylim = c(-37.92, -37.78)) +
+  theme(text = element_text(size= 12))
 

### 2020-04-06 anzjs_elections.Rmd
increase text size in plot

@@
     text = element_text(size = 10),
-    axis.text.x = element_text(size = 6),
+    axis.text.x = element_text(size = 9),
     axis.text.y = element_blank(),

### 2020-04-06 anzjs_elections.Rmd
Small changes based on reviewer 2's second feedback

@@
 The older Australian population has often been considered to be more conservative, and the left leaning political parties (including Labor) have typically had a stronger appeal to younger people.
-This effect was indeed observed across the six elections, where populations between 20 and 34 years of age (`Pop_20_34`) aligned strongly with Labor preference (bottom row in Figure \@ref(fig:partresplot2)).
-Larger populations of 35 to 54 year olds (`Pop_35_54`) were also associated with Labor, but the magnitude of this effect was far smaller.
+This effect was indeed observed across all six elections, as electorates with higher proportions of people aged between 20 and 34 (`Pop_20_34`) aligned strongly with Labor preference (bottom row in Figure \@ref(fig:partresplot2)).
+Larger proportions of 35 to 54 year olds (`Pop_35_54`) were also associated with Labor, but the magnitude of this effect was far smaller.
 Populations under 20 years of age was only significant in 2007, where `Pop_00_19` increased Liberal support.
@@
 
-Higher proportions of single parent households were associated with Labor support in all elections (albeit insignificant in 2004, see row 6 in Figure \@ref(fig:partresplot2)), whereas the electoral family and household sizes (via the `FamHouseSize` variable) did not appear to be associated with either party.
+Higher proportions of single parent households were associated with Labor support in all elections (albeit insignificant in 2004, see row 6 in Figure \@ref(fig:partresplot2)), whereas family and household sizes (via the `FamHouseSize` variable) did not appear to be associated with either party.
 
@@
 
-From 2007 onwards, both marriages (`Married`) and de facto relationships (`DeFacto`) were found to be strong predictors of the two-party preferred vote in favour of the Labor party.
+From 2007 onwards, proportions of both marriages (`Married`) and de facto relationships (`DeFacto`) were found to be strong predictors of the two-party preferred vote in favour of the Labor party.
 In 2001 and 2004 neither of these variables were significant (see rows 4 and 5 in Figure \@ref(fig:partresplot2)).
@@
 
-This paper was produced using \textsf{RMarkdown} [@rmarkdown] and \textsf{knitr} [@knitr].
+This paper was produced using \textsf{R Markdown} [@rmarkdown] and \textsf{knitr} [@knitr].
 All corresponding code for this paper can be found in the github repository \url{github.com/jforbes14/eechidna-paper}, and the data used is available in the \textsf{eechidna} package [@eechidna].

### 2020-04-06 anzjs_elections.tex
Small changes based on reviewer 2's second feedback

@@
 The older Australian population has often been considered to be more conservative, and the left leaning political parties (including Labor) have typically had a stronger appeal to younger people.
-This effect was indeed observed across the six elections, where populations between 20 and 34 years of age (\texttt{Pop\_20\_34}) aligned strongly with Labor preference (bottom row in Figure \ref{fig:partresplot2}).
+This effect was indeed observed across all six elections, as electorates with higher proportions of people aged between 20 and 34 (\texttt{Pop\_20\_34}) aligned strongly with Labor preference (bottom row in Figure \ref{fig:partresplot2}).
 Larger populations of 35 to 54 year olds (\texttt{Pop\_35\_54}) were also associated with Labor, but the magnitude of this effect was far smaller.
@@
 
-Higher proportions of single parent households were associated with Labor support in all elections (albeit insignificant in 2004, see row 6 in Figure \ref{fig:partresplot2}), whereas the electoral family and household sizes (via the \texttt{FamHouseSize} variable) did not appear to be associated with either party.
+Higher proportions of single parent households were associated with Labor support in all elections (albeit insignificant in 2004, see row 6 in Figure \ref{fig:partresplot2}), whereas family and household sizes (via the \texttt{FamHouseSize} variable) did not appear to be associated with either party.
 
@@
 
-From 2007 onwards, both marriages (\texttt{Married}) and de facto relationships (\texttt{DeFacto}) were found to be strong predictors of the two-party preferred vote in favour of the Labor party.
+From 2007 onwards, proportions of both marriages (\texttt{Married}) and de facto relationships (\texttt{DeFacto}) were found to be strong predictors of the two-party preferred vote in favour of the Labor party.
 In 2001 and 2004 neither of these variables were significant (see rows 4 and 5 in Figure \ref{fig:partresplot2}).
@@
 
-This paper was produced using \textsf{RMarkdown} \citep{rmarkdown} and \textsf{knitr} \citep{knitr}.
+This paper was produced using \textsf{R Markdown} \citep{rmarkdown} and \textsf{knitr} \citep{knitr}.
 All corresponding code for this paper can be found in the github repository \url{github.com/jforbes14/eechidna-paper}, and the data used is available in the \textsf{eechidna} package \citep{eechidna}.

### 2020-04-07 anzjs_elections.Rmd
Final minor changes for submission

@@
 Linear model with 95% confidence bands overlaid.
-Several predictors had a negative relationship: with larger values indicating the electorate more likely preferenced Labor.
+Several predictors had a negative relationship: the larger the value the more likely the electorate preferenced Labor.
 Most relationships were relatively stable over elections, except `Unemployment` and `Education`.
@@
 
-From 2007 onwards, proportions of both marriages (`Married`) and de facto relationships (`DeFacto`) were found to be strong predictors of the two-party preferred vote in favour of the Labor party.
+From 2007 onwards, the percentage of people in both marriages (`Married`) and de facto relationships (`DeFacto`) were found to be strong predictors of the two-party preferred vote in favour of the Labor party.
 In 2001 and 2004 neither of these variables were significant (see rows 4 and 5 in Figure \@ref(fig:partresplot2)).
@@
 
-It is often hypothesised that states have a systematic bias towards one of the two major parties.
+It is often hypothesised that states have had a systematic bias towards one of the two major parties.
 Boxplots of residuals grouped by state (Figure \@ref(fig:resstate)) showed that the data reflects this to only a limited extent.
@@
 All election and census datasets, along with electoral maps and more, are available in the \textsf{eechidna} (Exploring Election and Census Highly Informative Data Nationally for Australia) \textsf{R} package, which can be downloaded from CRAN.
-The \textsf{eechidna} package makes it easy to look at the data from the Australian Federal elections and censuses that occurred between 2001 and 2016.
+The \textsf{eechidna} package makes it easy to look at the data from the Australian Federal elections and censuses that occurred between 2001 and 2019.
 This study contributed a large revision to the \textsf{eechidna} package, which included the addition of election and census data for 2001--2010, voting outcomes for polling booths and imputed census data for election years.

### 2020-04-07 submission_070420/anzjs_elections.tex
Files as submitted in final submission

--- 2020-04-06 anzjs_elections.tex
@@
 Linear model with 95\% confidence bands overlaid.
-Several predictors had a negative relationship: with larger values indicating the electorate more likely preferenced Labor.
+Several predictors had a negative relationship: the larger the value the more likely the electorate preferenced Labor.
 Most relationships were relatively stable over elections, except \texttt{Unemployment} and \texttt{Education}.}\label{fig:partresplot2}
@@
 
-From 2007 onwards, proportions of both marriages (\texttt{Married}) and de facto relationships (\texttt{DeFacto}) were found to be strong predictors of the two-party preferred vote in favour of the Labor party.
+From 2007 onwards, the percentage of people in both marriages (\texttt{Married}) and de facto relationships (\texttt{DeFacto}) were found to be strong predictors of the two-party preferred vote in favour of the Labor party.
 In 2001 and 2004 neither of these variables were significant (see rows 4 and 5 in Figure \ref{fig:partresplot2}).
@@
 
-It is often hypothesised that states have a systematic bias towards one of the two major parties.
+It is often hypothesised that states have had a systematic bias towards one of the two major parties.
 Boxplots of residuals grouped by state (Figure \ref{fig:resstate}) showed that the data reflects this to only a limited extent.
